### 2018-07-06 spherical-harmonics-paper/presentation-spherical-harmonics.tex
First draft at presentation slides

\documentclass[10pt]{beamer}

% \usepackage{beamerthemesplit} // Activate for custom appearance

\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent

\usepackage{natbib}
\usepackage{url}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plain}

\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex	
\usepackage{media9}

\usepackage{caption}
\usepackage{subcaption}
% \captionsetup[figure]{font=scriptsize, labelfont=scriptsize}
\usepackage{float}
\graphicspath{ {../sphere/plots} }
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{bm}

\usetheme{Madrid}

\usepackage{mleftright}
\usepackage{todonotes}
\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}

\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\No}{\mathbb{N}_0}
\newcommand{\Ylm}{Y^m_l}
\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
\newcommand{\Plm}{P^m_l}
\newcommand{\costheta}{\cos\theta}
\newcommand{\sintheta}{\sin\theta}
\newcommand{\cosphi}{\cos\varphi}
\newcommand{\sinphi}{\sin\varphi}
\newcommand{\eimphi}{e^{im\varphi}}
\newcommand{\alphalm}{\alpha^m_l}
\newcommand{\clm}{c^m_l}
\newcommand{\ctilde}{\tilde{c}^m_l}
\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
\newcommand{\chat}{\hat{c}^m_l}
\newcommand{\chatmod}{\hat{c}^{|m|}_l}
\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}

\newcommand{\Atilde}{\tilde{A}_{l,m}}
\newcommand{\Btilde}{\tilde{B}_{l,m}}
\newcommand{\Dtilde}{\tilde{D}_{l,m}}
\newcommand{\Etilde}{\tilde{E}_{l,m}}
\newcommand{\Ftilde}{\tilde{F}_{l,m}}
\newcommand{\Gtilde}{\tilde{G}_{l,m}}
\newcommand{\Alm}{A_{l,m}}
\newcommand{\Blm}{B_{l,m}}
\newcommand{\Dlm}{D_{l,m}}
\newcommand{\Elm}{E_{l,m}}
\newcommand{\Flm}{F_{l,m}}
\newcommand{\Glm}{G_{l,m}}

\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}

\newcommand{\bigP}{\mathbb{P}}
\newcommand{\Pl}{\mathbb{P}_l}
\newcommand{\gradP}{T\mathbb{P}}
\newcommand{\gradPl}{T\mathbb{P}_l}
\newcommand{\gradY}{\nabla Y}
\newcommand{\gradYlm}{\nabla Y^m_l}
\newcommand{\gradpY}{\nabla^\perp Y}
\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}

\newcommand{\Dlt}{D^T_l}

\newcommand{\curlyy}{\bm{\mathcal{Y}}}
\newcommand{\blone}{\beta_{l, 1}}
\newcommand{\blzero}{\beta_{l, 0}}
\newcommand{\blmone}{\beta_{l, -1}}
\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
\newcommand{\cgcoeff}{\mathcal{C}}

\newcommand{\alm}{a_{l,m}}
\newcommand{\blm}{b_{l,m}}
\newcommand{\dlm}{d_{l,m}}
\newcommand{\elm}{e_{l,m}}
\newcommand{\flm}{f_{l,m}}
\newcommand{\glm}{g_{l,m}}
\newcommand{\hlm}{h_{l,m}}
\newcommand{\jlm}{j_{l,m}}
\newcommand{\klm}{k_{l,m}}
\newcommand{\almperp}{a_{l,m}^\perp}
\newcommand{\blmperp}{b_{l,m}^\perp}
\newcommand{\dlmperp}{d_{l,m}^\perp}
\newcommand{\elmperp}{e_{l,m}^\perp}
\newcommand{\flmperp}{f_{l,m}^\perp}
\newcommand{\glmperp}{g_{l,m}^\perp}
\newcommand{\hlmperp}{h_{l,m}^\perp}
\newcommand{\jlmperp}{j_{l,m}^\perp}
\newcommand{\klmperp}{k_{l,m}^\perp}

\newcommand{\unitvec}{\hat{\bm{k}}}

\title{Spherical Harmonics and Vector Spherical Harmonics as Orthogonal Polynomials}
\author{Ben Snowball}
\date{\today}

\begin{document}

\frame{\titlepage}

\section[Outline]{}
\frame{\tableofcontents}
\section{Introduction}
\subsection{Spherical Harmonics (SHs) as orthogonal polynomials on the sphere in Cartesian coordinates}
\section{Surface of the sphere}
\subsection{Jacobi operators for multiplication of SHs by x, y, z}
\subsection{Recurrence relation and function evaluation}
\section{Tangent space}
\subsection{Spin-1 tensor spherical harmonics}
\subsection{Jacobi operators for multiplication of vector SHs by x, y, z}
\subsection{Recurrence relation and function evaluation}
\section{Example}
\subsection{Linearised SWE}
\section{Summary and Outlook}
\subsection{The circle arc}

\frame
{
  \frametitle{On the Unit Interval (1D)}

On the unit interval, \([-1,1]\), we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that
\begin{align}
& \ddx P^{(a,b)}_l (x) = \text{const.} \times P^{(a+1,b+1)}_{l-1}(x) \\
\implies & \dmdxm P_l(x) = \text{const.} \times P^{(m,m)}_{l-m}(x)
\end{align}
where \(P^{(a,b)}_l (x)\) is the degree \(l\) Jacobi polynomial, orthogonal with weight \(w(x) = (1-x)^a(1+x)^b\), and \(P_l(x) := P^{(0,0)}_l (x)\) is the Legendre polynomial of degree \(l\).

}

\frame
{
  \frametitle{On the Unit Interval (1D)}

Further, we can define associated Legendre polynomials that are also orthogonal:
\begin{align}
\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm P_l(x) = \chat (1-x^2)^\frac{m}{2} P^{(m,m)}_{l-m}(x) \\
P^{-m}_l(x) &:= \ctilde \Plm(x),
\end{align}
for \(m = 0,1,2,\ldots\) where
\begin{align}
\chat &:= \frac{\Gamma(l+m+1)}{(-2)^m\Gamma(l+1)} \\
\ctilde &:= \frac{(-1)^m\Gamma(l-m+1)}{\Gamma(l+m+1)}
\end{align}
using the gamma function \(\Gamma(n) := (n-1)!\) for \(n \in \N\).

}

\frame
{
    \frametitle{Spherical Harmonics as OPs of three variables}

Let \((x,y,z)\) be a point on the unit sphere such that \(x^2+y^2+z^2 = 1\). Let \(\theta, \varphi\) denote the angles such that
\begin{align}
x &= \sintheta \cosphi \\
y &= \sintheta \sinphi \\
z &= \costheta.
\end{align}
Further, define
\begin{align}
\clm := \Bigg(\frac{(2l+1)\Gamma(l-m+1)}{4\pi\Gamma(l+m+1)}\Bigg)^\frac{1}{2},
\end{align}
and
\begin{align}
\alphalm := \begin{cases} 
		\clm \chatmod \quad \quad \quad \text{if } m \ge 0 \\
		\clm \chatmod \ctildemod \quad \text{if } m < 0
	   \end{cases}.
\end{align}

}

\frame
{
    \frametitle{Spherical Harmonics as OPs of three variables}

\begin{itemize}

\item We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{DLMFSphericalandSpheroidalHarmonics, DLMFFerrers, DLMFAssociatedLegendre}:
\begin{align}
\Ylmfull &:= \clm \eimphi \Plm (\costheta) \nonumber \\
&= \alphalm (1 - (\costheta)^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(\costheta), 
\end{align}
where  $0 \le |m| \le l, \, l \in \No$.

\item The SHs are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$:
\begin{align}
\Ylm(x,y,z) &= \alphalm (1 - z^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(z), \quad \text{where } 0 \le |m| \le l, \, l \in \No.
\end{align}

\item Span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

\end{itemize}

}

\frame
{
    \frametitle{Jacobi Operators}

\begin{block}{Goal}
To use SHs as a basis for scalar valued functions on the unit sphere $S := \{ \bold{x} := (x,y,z) \quad | \quad ||\bold{x}|| = 1\} \subset \R^3$.
\end{block}

\begin{itemize}

\item Computationally work with coefficients of the expansion of a function.

\item Among the basic operators we require are those for multiplication by $x, y$ and $z$.
\begin{align}
x\,\Ylm(x,y,z) &= \Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z), \\ \nonumber \\
y\,\Ylm(x,y,z) &= - i \, \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \nonumber \\
		     &\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\ \nonumber \\
z\,\Ylm(x,y,z) &= \Flm Y^{m}_{l+1}(x,y,z) + \Glm Y^{m}_{l-1}(x,y,z) ,
\end{align}

\end{itemize}

}

\frame
{
    \frametitle{Jacobi Operators}

Define, for \(l \in \No\), \(\Pl\) as the column vector of the degree \(l\) spherical harmonic polynomials, and \(\bigP\) as the stacked block vector of the \(\Pl\)'s; that is
\begin{align}
\Pl := \begin{bmatrix}
		Y^{-l}_l \\
		\vdots \\
		Y^l_l
	\end{bmatrix} \in \C^{2l+1}, 
\quad \quad 
\bigP := \begin{bmatrix}
		\bigP_0 \\
		\hline
		\bigP_1 \\
		\hline
		\bigP_2 \\
		\vdots \\
	\end{bmatrix}.
\end{align}

Define the (Jacobi) matrices \(J^x, J^y, J^z\) by 
\begin{align}
J^x \bigP = x \bigP, \quad J^y \bigP = y \bigP, \quad J^z \bigP = z \bigP.
\end{align}

}

\frame
{
    \frametitle{Jacobi Operators}
Using eqns (14--16), the Jacobi matrices take the following Banded-Block-Banded form:
\begin{align}
J^x &= \begin{bmatrix}
		B^x_0 & A^x_0 & & & & \\
		C^x_1 & B^x_1 & A^x_1 & & & \\
		& C^x_2 & B^x_2 & A^x_2  & & & \\
		& & C^x_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix},
J^y = \begin{bmatrix}
		B^y_0 & A^y_0 & & & & \\
		C^y_1 & B^y_1 & A^y_1 & & & \\
		& C^y_2 & B^y_2 & A^y_2  & & & \\
		& & C^y_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}, \nonumber \\
&\quad \quad \quad \quad \quad \quad \quad J^z = \begin{bmatrix}
		B^z_0 & A^z_0 & & & & \\
		C^z_1 & B^z_1 & A^z_1 & & & \\
		& C^z_2 & B^z_2 & A^z_2  & & & \\
		& & C^z_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}.
\end{align}

}

\frame
{
    \frametitle{Three-term recurrence relation}

Define the joint matrices for each \(l \in \No\):
\begin{align}
A_l &:= \begin{bmatrix}
		A^x_l \\
		A^y_l \\
		A^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+3)}, \quad
C_l := \begin{bmatrix}
		C^x_l \\
		C^y_l \\
		C^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l-1)} \quad (l \ne 0), \\
B_l &:= \begin{bmatrix}
		B^x_l \\
		B^y_l \\
		B^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}, \quad
G_l(x,y,z) := \begin{bmatrix}
		xI_{2l+1} \\
		yI_{2l+1} \\
		zI_{2l+1}
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}.
\end{align}
Then our system (62) simply becomes
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1  & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigP
=
\begin{bmatrix}
	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
\end{bmatrix}.
\end{align}

}

\frame
{
    \frametitle{Three-term recurrence relation}

For each \(l \in \No\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e. such that \(\Dlt A_l = I_{2l+3}\). Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain the lower triangular system \citep{dunkl2014orthogonal}
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1  & & & \\
		D^T_0(B_0-G_0) & I_3 & & \\
		D^T_1 C_1 & D^T_1(B_1-G_1) & \quad I_5 \quad & \\
		& D^T_2 C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigP
=
\begin{bmatrix}
	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
\end{bmatrix}.
\end{align}

Expanding this we can solve the system (66) to find a three-term recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\):
\begin{align}
\begin{cases}
\bigP_{-1} := 0 \\
\bigP_{0} := \alpha^0_0 \\
\bigP_{l+1} = -\Dlt (B_l-G_l) \Pl - \Dlt C_l  \,\bigP_{l-1}, \quad l \in \N.
\end{cases}
\end{align}

}

\frame
{
    \frametitle{Function Evaluation}

We can use the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e. suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
\begin{align}
f(x,y,z) \approx \sum^N_{l=0} \bold{f}_l^T \, \Pl (x,y,z), \quad \text{where } \Pl (x,y,z), \bold{f}_l \in \R^{2l+1} \text{ for each } l \in \{0,\dots,N\}.
\end{align}

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}

}

\frame
{
    \frametitle{Tangent Space}

\begin{itemize}

\item The gradients and perpendicular gradients of the spherical harmonics \(\gradYlm\) and\(\gradpYlm\) are a basis for the tangent space. Note that the perpendicular gradient is related to the regular surface gradient by
\begin{align}
\gradpYlm(x,y,z) = \unitvec \times \gradYlm(x,y,z),
\end{align}
where \(\unitvec\) is the unit vector normal to the surface of the sphere at the point \((x,y,z)\).

\item \(\gradYlm\), \(\gradpYlm\) are vector-valued orthogonal polynomials, that span all vector-valued polynomials modulo the vanishing ideal $\{ \bm{p} \in (\Pi_3)^3  : \unitvec \cdot \bm{p} = 0 \}$.

\item A simple calculation shows that such orthogonal polynomials must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal.

\end{itemize}

}

\frame
{
    \frametitle{Vector Spherical Harmonics}

\begin{itemize}

\item In general, the tensor spherical harmonic is given by, for \(2l, \, j, \, 2s \in \No\),
\begin{align}
\curlyy^{j,s}_{l,m}(x,y,z) &= \sum_{m_s=-s}^s \langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle \: \: Y^{m-m_s}_j (x,y,z) \: \: \bm{\chi}_{s, m_s},
\end{align}

\item \(\bm{\chi}_{s, m_s}\) are the simultaneous eigenstates of the spin operators \(\bold{S}^2\) and \(S_z\).

\item \(\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle\) is a Clebsch-Gordan coefficient. 
\begin{itemize}
	\item C-G coefficients vanish unless \(|j - s| \le l \le j + s\). 
	\item There are simple calculable expressions for the C-G coefficients when the spin \(s=1\).
\end{itemize}

\item Both \(\gradYlm\), \(\gradpYlm\) can be expressed in terms of spin-1 tensor spherical harmonics (vector SHs), which in turn can each be expressed as a vector-weighted sum of spherical harmonics.

\end{itemize}

}

\frame
{
    \frametitle{Vector Spherical Harmonics}

To simplify the notation, for spin $s=1$, we define
\begin{align}
\cgcoeff^{L, m_s}_{l, m} := \, \langle L \quad m-m_s \; ; \; 1 \quad m_s \; | \; l \quad m \rangle.
\end{align}

Thus we have that the three spin-1 tensor spherical harmonics are given by
\begin{align}
\curlyy^{l \pm1,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
\curlyy^{l,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
\end{align}
where
\begin{align}
\bm{\chi}_{1, \pm 1} = \frac{1}{\sqrt{2}} \begin{bmatrix} \mp 1 \\ -i \\ 0 \end{bmatrix}, \quad \bm{\chi}_{1,0} = \begin{bmatrix} 0 \\ 0 \\ 1 \end{bmatrix}.
\end{align}

}

\frame
{
    \frametitle{Orthogonal Polynomials on the Tangent Space}

Then, for any \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) we have that
\begin{align}
\nabla \Ylm &= \blmone \, \curlyy^{l-1,1}_{l,m} + \blone \curlyy^{l+1,1}_{l,m}, \\
\nabla^\perp \Ylm &= \blzero \, \curlyy^{l,1}_{l,m},
\end{align}
where
\begin{align}
\blmone &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
\blzero := i \, \big(l(l+1)\big)^\half, \quad
\blone := l \, \Big(\frac{l+1}{2l+1}\Big)^\half.
\end{align}

}

\frame
{
    \frametitle{Jacobi Operators}

\begin{block}{Goal}
To use the gradient and perpendicular gradients of the SHs as a basis for vector valued functions in the tangent space on the unit sphere.
\end{block}

\begin{align}
x \,\gradYlm &= \alm \gradY^{m+1}_{l+1} + \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad + \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l}, \\
y \,\gradYlm &= i \, \Big[ -\alm \gradY^{m+1}_{l+1} - \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad - \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l} \Big], \\
z \,\gradYlm &= \flm \gradY^{m}_{l+1} + \glm \gradY^{m}_{l-1} + \klm \gradpY^{m}_{l},
\end{align}

}

\frame
{
    \frametitle{Jacobi Operators}

\begin{align}
x \,\gradpYlm &= \almperp \gradpY^{m+1}_{l+1} + \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad + \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
y \,\gradpYlm &= i \, \Big[ -\almperp \gradpY^{m+1}_{l+1} - \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad - \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
z \,\gradpYlm &= \flmperp \gradpY^{m}_{l+1} + \glmperp \gradpY^{m}_{l-1} + \klmperp \gradY^{m}_{l},
\end{align}
where it can be shown that
\begin{align}
\almperp &= \alm^*, \quad \blmperp = \blm^*, \quad \dlmperp = \dlm^*, \quad \elmperp = \elm^*, \nonumber \\
\flmperp = \flm^*,& \quad \glmperp = \glm^*, \quad \hlmperp = \hlm^*, \quad \jlmperp = \jlm^*, \quad \klmperp = \klm^*,
\end{align}
where \(^*\) denotes the complex conjugate.

}

\frame
{
    \frametitle{Jacobi Operators}

\begin{itemize}

\item Define \(\gradP\) as the block-column vector
\begin{align}
\gradP = \begin{bmatrix} \gradP_0 \\ \gradP_1 \\ \vdots \end{bmatrix}, \quad \text{where } \quad \gradP_l = \begin{bmatrix} \gradY^{-l}_l \\ \gradpY^{-l}_l \\ \vdots \\ \gradY^{l}_l \\ \gradpY^{l}_l  \end{bmatrix} \quad \forall \: l \in \No.
\end{align}

\item Let $S := \{ \bold{x} := (x,y,z) \quad | \quad ||\bold{x}|| = 1\}$ be the unit sphere in $\R^3$.

\item Define
\begin{align}
\bold{u}^c = \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
		    \end{bmatrix},		  
\quad
\text{where}
\quad
\bold{u}^c_l &= \begin{bmatrix}
				u_{l,-l} \\
				u^{\perp}_{l,-l} \\
				\vdots \\
				u_{l,l} \\
				u^{\perp}_{l,l}
		        \end{bmatrix} \in \R^{2(2l+1)} \quad \forall l = 0,1,2,\dots
\end{align}

\end{itemize}

}

\frame
{
    \frametitle{Jacobi Operators}

\begin{itemize}

\item Let $\mathbf{u}(x,y,z)$ be a function on $S$ with vector values in the tangent space, i.e. there exist coefficients $\{u_{l,m}, u^\perp_{l,m}\}$ such that, $\forall (x,y,z) \in S$:
\begin{align}
\mathbf{u}(x,y,z) &= \sum_{l=0}^\infty \sum_{m=-l}^l \big( u_{l,m} \gradYlm(x,y,z) + u^\perp_{l,m} \gradpYlm(x,y,z) \big) \\
&= (\mathbf{u}^c)^T \gradP(x,y,z)
\end{align}

\item \(J_\nabla^x\) is defined as the matrix such that 
\begin{align}
x \: \mathbf{u}(x,y,z) = \sum_{l=0}^\infty \sum_{m=-l}^l \big( \tilde{u}_{l,m} \gradYlm(x,y,z) + \tilde{u}^\perp_{l,m} \gradpYlm(x,y,z) \big),
\end{align}
$\forall (x,y,z) \in S$, where
\begin{align}
(J_\nabla^x)^T \mathbf{u}^c = \mathbf{\tilde{u}}^c,
\quad
\bold{\tilde{u}}^c = \begin{bmatrix}
			\bold{\tilde{u}}_0^c \\
			\bold{\tilde{u}}_1^c \\
			\vdots \\
		    \end{bmatrix},		  
\quad
\text{and}
\quad
\bold{\tilde{u}}^c_l &= \begin{bmatrix}
				\tilde{u}_{l,-l} \\
				\tilde{u}^{\perp}_{l,-l} \\
				\vdots \\
				\tilde{u}^{\perp}_{l,l}
		        \end{bmatrix} \quad \forall l = 0,1,2,\dots
\end{align}

\end{itemize}

}

\frame
{
    \frametitle{Jacobi Operators}

\begin{itemize}

\item In other words, abusing notation, we can write that
\begin{align}
J_\nabla^x \gradP = x \gradP
\end{align}
where each sub-vector entry \(\gradYlm\) and \(\gradpYlm\) of \(\gradP\) is treated as a single element of the vector \(\gradP\) in the matrix-vector calculation.

\item \(J_\nabla^y, J_\nabla^z\) are defined similarly.

\end{itemize}

}

\frame
{
    \frametitle{Jacobi Operators}

Using eqns (36--41), the Jacobi matrices take the similar Banded-Block-Banded form:
\begin{align}
J_\nabla^x &= \begin{bmatrix}
		B^x_0 & A^x_0 & & & & \\
		C^x_1 & B^x_1 & A^x_1 & & & \\
		& C^x_2 & B^x_2 & A^x_2  & & & \\
		& & C^x_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix},
J_\nabla^y = \begin{bmatrix}
		B^y_0 & A^y_0 & & & & \\
		C^y_1 & B^y_1 & A^y_1 & & & \\
		& C^y_2 & B^y_2 & A^y_2  & & & \\
		& & C^y_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}, \nonumber \\
&\quad \quad \quad \quad \quad \quad \quad J_\nabla^z = \begin{bmatrix}
		B^z_0 & A^z_0 & & & & \\
		C^z_1 & B^z_1 & A^z_1 & & & \\
		& C^z_2 & B^z_2 & A^z_2  & & & \\
		& & C^z_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}.
\end{align}

}

\frame
{
    \frametitle{Three-term recurrence relation}

For each \(l \in \N\) define the joint matrices:
\begin{align}
A_l &:= \begin{bmatrix}
		A^x_l \\
		A^y_l \\
		A^z_l
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l+3)}, \quad
C_l := \begin{bmatrix}
		C^x_l \\
		C^y_l \\
		C^z_l
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l-1)} \quad (l \ne 1), \\
B_l &:= \begin{bmatrix}
		B^x_l \\
		B^y_l \\
		B^z_l
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l+1)}, \quad
G_l(x,y,z) := \begin{bmatrix}
		xI_{2(2l+1)} \\
		yI_{2(2l+1)} \\
		zI_{2(2l+1)}
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l+1)}.
\end{align}
Then our system becomes
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		I_2 \\
		0 & I_6 & & \\
		C_1 & B_1-G_1 & A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\gradP
=
\begin{bmatrix}
	\underline{0}_3 \\ \underline{0}_3 \\ \gradP_1 \\ \underline{0}_3 \\ \\ \vdots \\
\end{bmatrix}.
\end{align}

}

\frame
{
    \frametitle{Three-term recurrence relation}

In a similar way as for the scalar case, for each \(l \in \N\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e. such that \(\Dlt A_l = I_{2(2l+3)}\). We can then obtain the lower triangular system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		I_2 \\
		0 & I_6 & & \\
		D^T_1 C_1 & D^T_1(B_1-G_1) & \quad I_{10} \quad & \\
		& D^T_2 C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\gradP
=
\begin{bmatrix}
	\underline{0}_3 \\ \underline{0}_3 \\ \gradP_1 \\ \underline{0}_3 \\ \\ \vdots \\
\end{bmatrix}.
\end{align}

}

\frame
{
    \frametitle{Three-term recurrence relation}

Expanding this we can find a three-term recurrence relation for each \(\gradP_{l+1}\) in terms of the previous two sub-vectors \(\gradPl\) and \(\gradP_{l-1}\):
\begin{align}
\begin{cases}
\gradP_{0}(x,y,z) := \underline{0}_6 \\
\\
\gradP_{1}(x,y,z) := \begin{bmatrix}
				\gradY_1^{-1}(x,y,z) \\ \gradpY_1^{-1}(x,y,z) \\ \gradY_1^{0}(x,y,z) \\ \gradpY_1^{0}(x,y,z) \\ \gradY_1^{1}(x,y,z) \\ \gradpY_1^{1}(x,y,z) \\
			\end{bmatrix} \\
\\
\gradP_{l+1}(x,y,z) = -\Dlt [B_l-G_l(x,y,z)] \gradPl(x,y,z) - \Dlt C_l  \,\gradP_{l-1}(x,y,z).
\end{cases}
\end{align}
}

\frame
{
    \frametitle{Function Evaluation}

We can use the Clenshaw algorithm again to evaluate a vector valued function in the tangent space at a point $(x,y,z) \in S$. Suppose
\begin{align}
\mathbf{u}(x,y,z) \approx \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big].
\end{align}

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = (\mathbf{u}^c_n)^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } \mathbf{u}(x,y,z) \approx \bm{\gamma}_{1}^T \gradP_1(x,y,z)
\end{align*}

}

\frame
{
    \frametitle{How does this work practically?}

\begin{itemize}

\item \(S := \{ \bold{x} := (x,y,z) \quad | \quad ||\bold{x}|| = 1\}\) be the unit sphere in \(\R^3\)

\item \(\bold{u}(x,y,z)\), \(\bold{v}(x,y,z)\) vector valued functions for \((x,y,z)\) on the unit sphere with values in the tangent space

\item \(h(x,y,z)\) scalar function on the sphere.

\end{itemize}

\begin{align}
\implies \bold{u} &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ], \\
\bold{v} &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ v_{l,m} \gradYlm + v^{\perp}_{l,m} \gradpYlm ], \\
h &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} h_{l,m} \Ylm,
\end{align}
for some real coefficients \(u_{l,m}, u^{\perp}_{l,m}, v_{l,m}, v^{\perp}_{l,m}, h_{l,m}\).

}

\frame
{
    \frametitle{How does this work practically?}

Define the vectors of these coefficients as 
\begin{align}
\bold{u}^c &= \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
		    \end{bmatrix},
\quad
\bold{v}^c = \begin{bmatrix}
			\bold{v}_0^c \\
			\bold{v}_1^c \\
			\vdots \\
		    \end{bmatrix},
\quad
\bold{h}^c = \begin{bmatrix}
			\bold{h}_0^c \\
			\bold{h}_1^c \\
			\vdots \\
		    \end{bmatrix},		  
\end{align}
where
\begin{align}
\bold{u}^c_l &= \begin{bmatrix}
				u_{l,-l} \\
				u^{\perp}_{l,-l} \\
				\vdots \\
				u_{l,l} \\
				u^{\perp}_{l,l}
		        \end{bmatrix},
\quad
\bold{v}^c_l = \begin{bmatrix}
				v_{l,-l} \\
				v^{\perp}_{l,-l} \\
				\vdots \\
				v_{l,l} \\
				v^{\perp}_{l,l}
		        \end{bmatrix},
\quad
\bold{h}^c_l = \begin{bmatrix}
				h_{l,-l} \\
				\vdots \\
				h_{l,l} \\
		        \end{bmatrix},
\quad \forall l \in \No	  
\end{align}.

}

\frame
{
    \frametitle{How does this work practically?}

For large enough \(N \in \N\) we have that
\begin{align}
\bold{u} &\approx \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ], \\
\bold{v} &\approx \sum_{l=0}^{N} \sum_{m=-l}^{l} [ v_{l,m} \gradYlm + v^{\perp}_{l,m} \gradpYlm ], \\
h &\approx \sum_{l=0}^{N} \sum_{m=-l}^{l} h_{l,m} \Ylm,
\end{align}
and so we define the truncated coefficient vectors for some \(N \in \N\) as
\begin{align}
\bold{u}^c &= \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
			\bold{u}_N^c
		    \end{bmatrix},
\quad
\bold{v}^c = \begin{bmatrix}
			\bold{v}_0^c \\
			\bold{v}_1^c \\
			\vdots \\
			\bold{v}_N^c
		    \end{bmatrix},
\quad
\bold{h}^c = \begin{bmatrix}
			\bold{h}_0^c \\
			\bold{h}_1^c \\
			\vdots \\
			\bold{h}_N^c
		    \end{bmatrix}.		  
\end{align}

}

\frame
{
    \frametitle{How does this work practically?}

Operators act on the coefficients vectors.

Easily derive operators for Grad, Div, curl with the unit vector, etc. as well as the multiplications by $x, y, z$.

- BandedBlockBanded structures

Can calculate dot products simply.

}

\frame
{
    \frametitle{Example: Linearised SWE}

Operators:

$K$ - cross product with $\unitvec$

$F$ - Coriolis parameter

$D$ - Divergence operator (of a function in the tangent space)

$G$ - Gradient operator (of a scalar function)

Let $\mathbf{u}(x,y,z)$ be the tangential velocity of the flow and $h(x,y,z)$ be the height of the water from some constant reference height $\mathcal{H}$.

The linear SWEs are
\begin{align}
\begin{cases}
\frac{\partial \mathbf{u}}{\partial t} + f \unitvec \times \mathbf{u} - \nabla h = \mathbf{0} \\
\frac{\partial h}{\partial t} + \mathcal{H} \nabla \cdot \mathbf{u} = 0
\end{cases}
\end{align}
where $f = 2 \Omega \cos(\theta) =  2 \Omega z$ is the Coriolis parameter.

}

\frame
{
    \frametitle{Example: Linearised SWE}

Using backward Euler:
\begin{align}
\mathbf{u}_{n+1} &= \mathbf{u}_{n} + \Delta t \: (\nabla h_{n+1} - f \unitvec \times \mathbf{u}_{n+1}) \\
h_{n+1} &= h_n - \Delta t \: \mathcal{H} \nabla \cdot \mathbf{u}_{n+1}
\end{align}

Then using the operators on the coefficient vectors, the timestepping method is: 
\begin{align}
\mathbf{u}^c_{n+1} &= \mathbf{u}^c_{n} + \Delta t \: (G \mathbf{h}^c_{n+1} - F K \mathbf{u}^c_{n+1}) \\
\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
\end{align}
$\iff$
\begin{align}
\mathbf{u}^c_{n+1} &= (I + \Delta t F K + \Delta t^2 \mathcal{H} G D)^{-1} \: (\mathbf{u}^c_{n} + \Delta t \: G \mathbf{h}^c_n) \\
\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
\end{align}

}

\frame
{
    \frametitle{Example: Linearised SWE}

\includemedia[
     width=6cm,height=6cm,
     activate=pageopen,
     addresource=swe.avi,
     flashvars={
         source=swe.avi
        &autoPlay=false
     }
]{}{VPlayer.swf}

}

\frame
{
    \frametitle{Where next?}

\begin{itemize}

\item Future work will focus on using similar techniques for a suitable basis of OPs on spherical triangles.

\item First, we are looking at the 2D analog - the circle arc

\begin{itemize}
	\item Used a version of the Lanczos algorithm to determine OPs
	\item Implemented function evaluation
	\item Operator for derivative along the arc segment
\end{itemize}

\end{itemize}

}

\frame
{
    \frametitle{Summary}

  \begin{itemize}

  \item The motivation was to look at creating a library for working with SHs as OPs in three variables on the sphere.

  \item Derived operators for multiplication by $x, y, z$ for functions expanded in either the scalar or tangent basis.

  \item Clenshaw's algorithm can be used to evaluate functions.

  \item Example to demonstrate how the operator matrices for common operations can be implemented in a simple timestepping method.

  \item Operators are BandedBlockBanded.

  \item Look ahead to using similar ideas for a suitable basis of OPs on spherical triangles (3D) by first working on the circle arc (2D).

  \end{itemize}

}

\frame
{
  \frametitle{References}
\bibliography{spherical-harmonics-bib}
}

\end{document}

### 2018-07-06 spherical-harmonics-paper/presentation-spherical-harmonics.tex
Second draft

@@
 
-Let \((x,y,z)\) be a point on the unit sphere such that \(x^2+y^2+z^2 = 1\). Let \(\theta, \varphi\) denote the angles such that
+\begin{block}{Motivation}
+To look at creating a library for working with spherical harmonics (SHs) as orthogonal polynomials (OPs) in three variables on the unit sphere.
+\end{block}
+
+Let \((x,y,z)\) be a point on the unit sphere, \(\theta, \varphi\) denote the angles such that
 \begin{align}
@@
 \quad &\text{2) } \text{For } n = N:-1:1 \\
-\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
+\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
 \quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
@@
 {
-    \frametitle{Vector Spherical Harmonics}
-
-\begin{itemize}
-
-\item In general, the tensor spherical harmonic is given by, for \(2l, \, j, \, 2s \in \No\),
-\begin{align}
-\curlyy^{j,s}_{l,m}(x,y,z) &= \sum_{m_s=-s}^s \langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle \: \: Y^{m-m_s}_j (x,y,z) \: \: \bm{\chi}_{s, m_s},
-\end{align}
-
-\item \(\bm{\chi}_{s, m_s}\) are the simultaneous eigenstates of the spin operators \(\bold{S}^2\) and \(S_z\).
-
-\item \(\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle\) is a Clebsch-Gordan coefficient. 
-\begin{itemize}
-	\item C-G coefficients vanish unless \(|j - s| \le l \le j + s\). 
-	\item There are simple calculable expressions for the C-G coefficients when the spin \(s=1\).
-\end{itemize}
-
-\item Both \(\gradYlm\), \(\gradpYlm\) can be expressed in terms of spin-1 tensor spherical harmonics (vector SHs), which in turn can each be expressed as a vector-weighted sum of spherical harmonics.
-
-\end{itemize}
-
-}
-
-\frame
-{
-    \frametitle{Vector Spherical Harmonics}
-
-To simplify the notation, for spin $s=1$, we define
-\begin{align}
-\cgcoeff^{L, m_s}_{l, m} := \, \langle L \quad m-m_s \; ; \; 1 \quad m_s \; | \; l \quad m \rangle.
-\end{align}
-
-Thus we have that the three spin-1 tensor spherical harmonics are given by
-\begin{align}
-\curlyy^{l \pm1,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
-\curlyy^{l,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
-\end{align}
-where
-\begin{align}
-\bm{\chi}_{1, \pm 1} = \frac{1}{\sqrt{2}} \begin{bmatrix} \mp 1 \\ -i \\ 0 \end{bmatrix}, \quad \bm{\chi}_{1,0} = \begin{bmatrix} 0 \\ 0 \\ 1 \end{bmatrix}.
-\end{align}
-
-}
-
-\frame
-{
     \frametitle{Orthogonal Polynomials on the Tangent Space}
@@
 \end{align}
+
+\begin{itemize}
+\item $\curlyy^{l-1,1}_{l,m}, \curlyy^{l,1}_{l,m}, \curlyy^{l+1,1}_{l,m}$ are spin-1 tensor spherical harmonics (vector SHs).
+\item VSHs can be expressed as a vector-weighted sum of spherical harmonics.
+\end{itemize}
 
@@
 
-Operators act on the coefficients vectors.
+\begin{itemize}
 
-Easily derive operators for Grad, Div, curl with the unit vector, etc. as well as the multiplications by $x, y, z$.
+\item Operators act on the coefficients vectors.
 
-- BandedBlockBanded structures
+\item Easily derive operators for Grad, Div, curl with the unit vector, etc. as well as the multiplications by $x, y, z$.
+\begin{itemize}
+	\item BandedBlockBanded structures
+\end{itemize}
 
-Can calculate dot products simply.
+\item Can calculate dot products simply.
+
+\end{itemize}
 

### 2018-11-20 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
laplace operator, and some working examples to test. transform operator not working for N>3

@@ -774,20 +774,42 @@ differentiatey(f::Fun, S::HalfDiskSpace) =
 
 
 function evalweightedderivativex(S::HalfDiskSpace, n, k, x, y)
-
+    j = Int((n+1)*n/2 + k)
+    p = Fun(S, [zeros(j); 1])(x,y)
+    x^(S.a-1) * (1-x^2-y^2)^(S.b-1) * (S.a * (1-x^2-y^2) * p
+                                       -2 * S.b * x^2 * p
+                                       + evalderivativex(S, n, k, x, y) * x * (1-x^2-y^2))
 end
 function evalweightedderivativey(S::HalfDiskSpace, n, k, x, y)
     j = Int((n+1)*n/2 + k)
     x^(S.a) * (1-x^2-y^2)^(S.b-1) * (-2 * S.b * y * Fun(S, [zeros(j); 1])(x,y)
                                         + evalderivativey(S, n, k, x, y) * (1-x^2-y^2))
 end
-function getweightedpartialoperatory(S::HalfDiskSpace, N)
-
+function getweightedpartialoperatorx(S::HalfDiskSpace, N)
+    W = BandedBlockBandedMatrix(Zeros{Float64}(sum(1:(N+3)),sum(1:(N+1))), (1:N+3, 1:N+1), (2,-1), (2,0))
+    for n = 0:N
+        for k = 0:n-1
+            p = (x,y) -> (OrthogonalPolynomialFamilies.evalweightedderivativex(S, n, k, x, y) * x^(1-S.a) * (1-x^2-y^2)^(1-S.b))
+            cfs = pad(Fun(p, (S.family)(S.a-1, S.b-1)).coefficients, sum(1:(n+3)))
+            cfs = PseudoBlockArray(cfs, 1:(n+3))
+            inds = k+1:2:k+3
+            view(W, Block(n+2, n+1))[inds, k+1] = cfs[Block(n+2)][inds]
+            view(W, Block(n+3, n+1))[inds, k+1] = cfs[Block(n+3)][inds]
+        end
+        k = n
+        p = (x,y) -> (OrthogonalPolynomialFamilies.evalweightedderivativex(S, n, k, x, y) * x^(1-S.a) * (1-x^2-y^2)^(1-S.b))
+        cfs = pad(Fun(p, (S.family)(S.a-1, S.b-1)).coefficients, sum(1:(n+3)))
+        cfs = PseudoBlockArray(cfs, 1:(n+3))
[…]

### 2018-11-20 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -1,7 +1,7 @@
 using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature, SingularIntegralEquations, Test
 import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule, gethalfdiskOP,
                                         jacobix, jacobiy, evalderivativex, evalderivativey,
-                                        differentiatex, differentiatey
+                                        differentiatex, differentiatey, resizecoeffs!, laplace
 
 
 
@@ -169,6 +169,30 @@ end
     h = 1e-5; @test (f(x,y+h)-f(x,y))/h ≈ differentiatey(f, f.space)(x,y) atol=100h
 end
 
+@test "Poisson" begin
+    # Model Problem: Δ(u*w)(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
+    #   where w(x,y) = x*(1-x^2-y^2) is the weight of the D(1.0,1.0) basis.
+    a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
+    x, y = 0.4, -0.2; z = [x; y] # Test point
+
+    # 1) f(x,y) = -8x => u(x,y) ≡ 1
+    N = 1 # degree of f
+    c = rand(1)[1]; f = Fun((x,y)->-c*8x, S)
+    Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+    u = Fun(S, Δ \ resizecoeffs!(f, N))
+    @test u(z) ≈ c # Result u(x,y) where Δ(u*w)(x,y) = f(x,y)
+
+    # 2) f(x,y) = 2 - 12xy - 14x^2 - 2y^2 => u(x,y) = x + y
+    U = Fun((x,y)->x+y, S)
+    N = 2 # degree of f
+    f = Fun((x,y)->(2 - 12x*y - 14x^2 - 2y^2), S)
+    Δ = laplace(D, N-1)
+    u = Fun(S, Δ \ resizecoeffs!(f, N))
+    @test u(z) ≈ U(z)
+end
+
+
+#=====#
 
 
 

### 2018-11-20 experiments/halfdisk.jl

@@ -17,32 +17,39 @@ using StaticArrays, LinearAlgebra, Test, Profile, BlockArrays,
 # differentiatespacey2(S) =
 #     OrthogonalPolynomialFamilies.differentiatespacey(
 #             OrthogonalPolynomialFamilies.differentiatespacey(S))
-# function resizecoeffs!(cfs::AbstractVector, N)
-#     m̃ = length(cfs)
-#     m = Int((N+1)*(N+2)/2)
-#     resize!(cfs, m)
-#     if m̃ < m
-#         cfs[m̃+1:end] .= 0.0
-#     end
-#     cfs
-# end
 
 
-function evalweightedderivativex(S::HalfDiskSpace, n, k, x, y)
+# Model Problem: Δ(u*w)(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
+#   where w(x,y) = x*(1-x^2-y^2) is the weight of the D(1.0,1.0) basis.
+a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
+x, y = 0.4, -0.2; z = [x; y] # Test point
 
-end
-function getweightedpartialoperatory(S::HalfDiskSpace, N)
+# 1) f(x,y) = -8x => u(x,y) ≡ 1
+N = 1 # degree of f
+c = rand(1)[1]; f = Fun((x,y)->-c*8x, S)
+resizecoeffs!(f, N)
+Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+Δ \ f.coefficients
+u = Fun(S, Δ \ f.coefficients)
+@test u(z) ≈ c # Result u(x,y) where Δ(u*w)(x,y) = f(x,y)
+# plot(x->u(x,y))
+# plot(y->u(x,y))
 
-end
+# 2) f(x,y) = 2 - 12xy - 14x^2 - 2y^2 => u(x,y) = x + y
+U = Fun((x,y)->x+y, S)
+N = 2 # degree of f
+f = Fun((x,y)->(2 - 12x*y - 14x^2 - 2y^2), S)
+Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+u = Fun(S, Δ \ resizecoeffs!(f, N))
+@test u(z) ≈ U(z)
 
-
-
-x, y = 0.2, 0.3; x^2 + y^2 < 1
-z = [x;y]
[…]

### 2018-11-22 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
operators work for higher N. Is slow.

@@ -707,7 +707,9 @@ function getpartialoperatorx(S::HalfDiskSpace, N)
                 differentiatespacex(S)).coefficients, sum(1:n))
     view(A, Block(n, n+1))[1, 1] = cfs[1]
     for n = 2:N, k = 0:n
-        cfs = pad(Fun((x,y) -> evalderivativex(S, n, k, x, y), differentiatespacex(S)).coefficients, sum(1:n))
+        m = Int((n+3)*(n+4))
+        cfs = pad(Fun((x,y) -> evalderivativex(S, n, k, x, y),
+                        differentiatespacex(S), m).coefficients, sum(1:n))
         cfs = PseudoBlockArray(cfs, 1:n)
         if k == 0
             inds1 = 1
@@ -735,7 +737,9 @@ end
 function getpartialoperatory(S::HalfDiskSpace, N)
     A = BandedBlockBandedMatrix(Zeros{Float64}(sum(1:N),sum(1:(N+1))), (1:N, 1:N+1), (-1,1), (-1,1))
     for n = 1:N, k = 1:n
-        cfs = pad(Fun((x,y) -> evalderivativey(S, n, k, x, y), differentiatespacey(S)).coefficients, sum(1:n))
+        m = Int((n+2)*(n+3))
+        cfs = pad(Fun((x,y) -> evalderivativey(S, n, k, x, y),
+                        differentiatespacey(S), m).coefficients, sum(1:n))
         cfs = PseudoBlockArray(cfs, 1:n)
         view(A, Block(n, n+1))[k, k+1] = cfs[Block(n)][k]
     end
@@ -788,17 +792,19 @@ end
 function getweightedpartialoperatorx(S::HalfDiskSpace, N)
     W = BandedBlockBandedMatrix(Zeros{Float64}(sum(1:(N+3)),sum(1:(N+1))), (1:N+3, 1:N+1), (2,-1), (2,0))
     for n = 0:N
+        m = Int((n+2+(2-S.a-S.b))*(n+3+(2-S.a-S.b)))
         for k = 0:n-1
             p = (x,y) -> (OrthogonalPolynomialFamilies.evalweightedderivativex(S, n, k, x, y) * x^(1-S.a) * (1-x^2-y^2)^(1-S.b))
[…]

### 2018-11-22 experiments/halfdisk.jl

@@ -27,7 +27,7 @@ x, y = 0.4, -0.2; z = [x; y] # Test point
 # 1) f(x,y) = -8x => u(x,y) ≡ 1
 N = 1 # degree of f
 c = rand(1)[1]; f = Fun((x,y)->-c*8x, S)
-resizecoeffs!(f, N)
+OrthogonalPolynomialFamilies.resizecoeffs!(f, N)
 Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
 Δ \ f.coefficients
 u = Fun(S, Δ \ f.coefficients)
@@ -40,16 +40,14 @@ U = Fun((x,y)->x+y, S)
 N = 2 # degree of f
 f = Fun((x,y)->(2 - 12x*y - 14x^2 - 2y^2), S)
 Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
-u = Fun(S, Δ \ resizecoeffs!(f, N))
+u = Fun(S, Δ \ OrthogonalPolynomialFamilies.resizecoeffs!(f, N))
 @test u(z) ≈ U(z)
 
 # 3) f(x,y) = y*exp(x)*[2-11x-6x^2-x^3-2y^2-xy^2] => u(x,y) = y*exp(x)
-# TODO: laplace(D, n) hangs/gives error for n >= 3. Think problem is in
-#       obtaining T^{(1,0)->(0,0)} operator.
 U = Fun((x,y)->y*exp(x), S)
-N = 4 # degree of f
-m = Int((N+1)*(N+2)/2)
+N = 7 # degree of f
+m = Int((N+1)*(N+2))
 f = Fun((x,y)->y*exp(x)*(2-11x-6x^2-x^3-2y^2-x*y^2), S, m)
-# Δ = OrthogonalPolynomialFamilies.laplace(D, N-1) # NOTE: hangs/error
-# u = Fun(S, Δ \ resizecoeffs!(f, N))
-# @test u(z) ≈ U(z)
+Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+u = Fun(S, Δ \ f.coefficients[1:size(Δ)[1]])
+@test u(z) ≈ U(z) atol=10.0^(-N)

### 2018-11-23 papers/esa/esa.tex
Current write up draft for ESA. Contains almost complete write up of Half Disk stuff

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mleftright}
\usepackage{todonotes}

\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

%SetFonts

%SetFonts

\usepackage{natbib}
\usepackage{url}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plain}

\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\No}{\mathbb{N}_0}
\newcommand{\Ylm}{Y^m_l}
\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
\newcommand{\Plm}{P^m_l}
\newcommand{\costheta}{\cos\theta}
\newcommand{\sintheta}{\sin\theta}
\newcommand{\cosphi}{\cos\varphi}
\newcommand{\sinphi}{\sin\varphi}
\newcommand{\eimphi}{e^{im\varphi}}
\newcommand{\alphalm}{\alpha^m_l}
\newcommand{\clm}{c^m_l}
\newcommand{\ctilde}{\tilde{c}^m_l}
\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
\newcommand{\chat}{\hat{c}^m_l}
\newcommand{\chatmod}{\hat{c}^{|m|}_l}
\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}

\newcommand{\Atilde}{\tilde{A}_{l,m}}
\newcommand{\Btilde}{\tilde{B}_{l,m}}
\newcommand{\Dtilde}{\tilde{D}_{l,m}}
\newcommand{\Etilde}{\tilde{E}_{l,m}}
\newcommand{\Ftilde}{\tilde{F}_{l,m}}
\newcommand{\Gtilde}{\tilde{G}_{l,m}}
\newcommand{\Alm}{A_{l,m}}
\newcommand{\Blm}{B_{l,m}}
\newcommand{\Dlm}{D_{l,m}}
\newcommand{\Elm}{E_{l,m}}
\newcommand{\Flm}{F_{l,m}}
\newcommand{\Glm}{G_{l,m}}

\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}

\newcommand{\bigP}{\mathbb{P}}
\newcommand{\Pl}{\mathbb{P}_l}
\newcommand{\gradP}{T\mathbb{P}}
\newcommand{\gradPl}{T\mathbb{P}_l}
\newcommand{\gradY}{\nabla Y}
\newcommand{\gradYlm}{\nabla Y^m_l}
\newcommand{\gradpY}{\nabla^\perp Y}
\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}

\newcommand{\Dlt}{D^T_l}

\newcommand{\curlyy}{\bm{\mathcal{Y}}}
\newcommand{\blone}{\beta_{l, 1}}
\newcommand{\blzero}{\beta_{l, 0}}
\newcommand{\blmone}{\beta_{l, -1}}
\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
\newcommand{\cgcoeff}{\mathcal{C}}

\newcommand{\alm}{a_{l,m}}
\newcommand{\blm}{b_{l,m}}
\newcommand{\dlm}{d_{l,m}}
\newcommand{\elm}{e_{l,m}}
\newcommand{\flm}{f_{l,m}}
\newcommand{\glm}{g_{l,m}}
\newcommand{\hlm}{h_{l,m}}
\newcommand{\jlm}{j_{l,m}}
\newcommand{\klm}{k_{l,m}}
\newcommand{\almperp}{a_{l,m}^\perp}
\newcommand{\blmperp}{b_{l,m}^\perp}
\newcommand{\dlmperp}{d_{l,m}^\perp}
\newcommand{\elmperp}{e_{l,m}^\perp}
\newcommand{\flmperp}{f_{l,m}^\perp}
\newcommand{\glmperp}{g_{l,m}^\perp}
\newcommand{\hlmperp}{h_{l,m}^\perp}
\newcommand{\jlmperp}{j_{l,m}^\perp}
\newcommand{\klmperp}{k_{l,m}^\perp}

\newcommand{\unitvec}{\hat{\bm{k}}}

\newcommand{\Pnk}{P_{n,k}}
\newcommand{\Pnkab}{P_{n,k}^{(a,b)}}
\newcommand{\Wab}{{W^{(a,b)}}}
\newcommand{\Pmjab}{P_{m,j}^{(a,b)}}
\newcommand{\alphaab}{\alpha^{(a,b)}}
\newcommand{\betaab}{\beta^{(a,b)}}
\newcommand{\bigPab}{\bigP^{(a,b)}}
\newcommand{\Dnt}{D^T_n}
\newcommand{\Wii}{W^{(1,1)}}
\newcommand{\Pii}{P^{(1,1)}}
\newcommand{\Poo}{P^{(0,0)}}
\newcommand{\Poi}{P^{(0,1)}}
\newcommand{\Pio}{P^{(1,0)}}
\newcommand{\dx}{\frac{\partial}{\partial x}}
\newcommand{\dy}{\frac{\partial}{\partial y}}

%\usepackage{media9}
%\graphicspath{ {../sphere/plots} }

\title{Early Stage Assessment}
\author{Ben Snowball (Principle Supervisor: Sheehan Olver)}
\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\section{Introduction}

This document is to provide a brief description of my research area which will be the topic of my thesis. All code used thus far and from hence forth will mainly be written in the Julia programming language.

\section{Professional Skills Development Requirement}
I have met the initial part of the professional skills development requirement which is to attend a minimum of 2 professional development workshops plus the online plagiarism awareness course by the Early Stage Assessment (ESA).  This is by virtue of the fact that I attended the Communication Skills Training in the MPE CDT on 03/11/2017.  This course has been given accreditation with the Graduate School programme and is worth 2 credits towards the requirement.

\section{Orthogonal polynomials on the sphere (spherical harmonics)}

The first step was to develop a small code library so that we can use spherical harmonics as orthogonal polynomials in three variables $x, y, z$  to evaluate functions on the sphere.

We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{DLMFSphericalandSpheroidalHarmonics, DLMFFerrers, DLMFAssociatedLegendre}:
\begin{align}
\Ylmfull &:= \clm \eimphi \Plm (\costheta) \\
&= \alphalm (1 - (\costheta)^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(\costheta), \quad \text{where } 0 \le |m| \le l, \, l \in \No.
\end{align} 
Note that we can then express \(\Ylm\) in terms of \(x,y,z\) instead of \(\theta, \varphi\) by noting that \(\costheta = z\) and that \( \eimphi\) can be expressed in terms of \(x,y,z\) for any \(m\in\mathbb{Z}\). Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

\citep{dunkl2014orthogonal} provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$. Clenshaw's algorithm then gives us a way of evaluating a function given its expansion in the spherical harmonic basis (up to a finite order).

\subsection{Deriving expressions for the multiplication by \(x, y, z\) of the spherical harmonics}

We start by expressing \(x\,\Ylm(x,y,z)\), \(y\,\Ylm(x,y,z)\) , and \(z\,\Ylm(x,y,z)\)  in terms of \(Y^{m'}_{l'}(x,y,z)\) for any point \((x,y,z)\) on the unit circle.

For \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\):
\begin{align}
x\,\Ylm(x,y,z) &= \Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z), \\
y\,\Ylm(x,y,z) &= - i \, \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \nonumber \\
		     &\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
z\,\Ylm(x,y,z) &= \Flm Y^{m}_{l+1}(x,y,z) + \Glm Y^{m}_{l-1}(x,y,z) ,
\end{align}
where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have derivable and calculable formulae.

\subsection{Jacobi matrices}

Define, for \(l \in \No\), \(\Pl\) as the column vector of the degree \(l\) spherical harmonic polynomials, and \(\bigP\) as the stacked block vector of the \(\Pl\)'s; that is
\begin{align}
\Pl := \begin{bmatrix}
		Y^{-l}_l \\
		\vdots \\
		Y^l_l
	\end{bmatrix} \in \C^{2l+1}, 
\quad \quad 
\bigP := \begin{bmatrix}
		\bigP_0 \\
		\hline
		\bigP_1 \\
		\hline
		\bigP_2 \\
		\vdots \\
	\end{bmatrix}.
\end{align}

Define the (Jacobi) matrices \(J^x, J^y, J^z\) by 
\begin{align}
J^x \bigP = x \bigP, \quad J^y \bigP = y \bigP, \quad J^z \bigP = z \bigP.
\end{align}
Then, using equations (3--5), we have that the Jacobi matrices each have the following block-tridiagonal form, allowing for easy application computationally:
\begin{align}
J^{x/y/z} &= \begin{bmatrix}
		B^{x/y/z}_0 & A^{x/y/z}_0 & & & & \\
		C^{x/y/z}_1 & B^{x/y/z}_1 & A^{x/y/z}_1 & & & \\
		& C^{x/y/z}_2 & B^{x/y/z}_2 & A^{x/y/z}_2  & & & \\
		& & C^{x/y/z}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}
\end{align}

\subsection{Three-term recurrence relation for \(\bigP\)}

Combining each system in (7) we can write the tridiagonal-block system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1  & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigP
=
\begin{bmatrix}
	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
\end{bmatrix}.
\end{align}
where for each \(l \in \No\):
\begin{align}
A_l &:= \begin{bmatrix}
		A^x_l \\
		A^y_l \\
		A^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+3)}, \quad
C_l := \begin{bmatrix}
		C^x_l \\
		C^y_l \\
		C^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l-1)} \quad (l \ne 0), \\
B_l &:= \begin{bmatrix}
		B^x_l \\
		B^y_l \\
		B^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}, \quad
G_l(x,y,z) := \begin{bmatrix}
		xI_{2l+1} \\
		yI_{2l+1} \\
		zI_{2l+1}
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}.
\end{align}

For each \(l \in \No\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e. such that \(\Dlt A_l = I_{2l+3}\). Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align}
\begin{cases}
\bigP_{-1} := 0 \\
\bigP_{0} := \alpha^0_0 \\
\bigP_{l+1} = -\Dlt (B_l-G_l) \Pl - \Dlt C_l  \,\bigP_{l-1}, \quad l \in \N.
\end{cases}
\end{align}

\subsection{Evaluation of a scalar function on the sphere}
We can use the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e. suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
\begin{align}
f(x,y,z) \approx \sum^N_{l=0} \bold{f}_l^T \, \Pl (x,y,z), \quad \text{where } \Pl (x,y,z), \bold{f}_l \in \R^{2l+1} \text{ for each } l \in \{0,\dots,N\}.
\end{align}

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}
where $\bigP_0 \equiv \alpha^0_0$.

We can use the Clenshaw algorithm to output a matrix operator for a function, given we know the coefficients of its expansion in the SH basis. We simply make the adjustment that
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = (\bold{f}_n^T)I - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(J^x,J^y,J^z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(J^x,J^y,J^z) \approx F := \alpha^0_0 f_0 + \bm{\gamma}_{1}^T \bigP_1(J^x,J^y,J^z) - \alpha^0_0 \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}
where $\bigP_1(J^x,J^y,J^z)$ is a vector of matrices, $G_n(J^x,J^y,J^z)$ is a matrix of matrices, and so each $\bm{\gamma}_n$ is a vector of matrices.

\subsection{Tangent space}

By noting that the tangent space is then spanned by $\{\gradYlm, \gradpYlm\}$, and that these gradients and perpendicular gradients are orthogonal polynomials themselves, we can use the same ideas to create a framework for functions in the tangent space as expansions in this OP basis.

Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient  and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics are a basis for the tangent space, namely \(\gradYlm\), \(\gradpYlm\). Note that the perpendicular gradient is related to the regular surface gradient by
\begin{align}
\gradpYlm(x,y,z) = \unitvec \times \gradYlm(x,y,z),
\end{align}
where \(\unitvec\) is the unit vector normal to the surface of the sphere at the point \((x,y,z)\), i.e. as we are looking at the unit sphere, \(\unitvec\) is simply given by
\begin{align}
\unitvec = \begin{bmatrix} x \\ y \\ z \end{bmatrix}.
\end{align}

Importantly,  \(\gradYlm\), \(\gradpYlm\) are in fact vector-valued orthogonal polynomials, that span all vector-valued polynomials modulo the vanishing ideal $\{ \bm{p} \in (\Pi_3)^3  : \unitvec \cdot \bm{p} = 0 \}$. A simple calculation shows that such orthogonal polynomials must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal. In the following, we determine the coefficients of these Jacobi operators, as well as other important operators needed.

\subsubsection{Spin-1 tensor spherical harmonics}

The gradient and perpendicular gradient of a spherical harmonic \(\Ylm(x,y,z)\) can be expressed in terms of spin-1 tensor spherical harmonics, which in turn can each be expressed as a vector-weighted sum of spherical harmonics. For the derivation and more on vector spherical harmonics, see the appendix.

For any \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) we have that
\begin{align}
\nabla \Ylm &= \blmone \, \curlyy^{l-1,1}_{l,m} + \blone \curlyy^{l+1,1}_{l,m}, \\
\nabla^\perp \Ylm &= \blzero \, \curlyy^{l,1}_{l,m},
\end{align}
where
\begin{align}
\blmone &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
\blzero := i \, \big(l(l+1)\big)^\half, \quad
\blone := l \, \Big(\frac{l+1}{2l+1}\Big)^\half.
\end{align}

We can then gain similar (vector analogue) derivations for recurrences, Jacobi matrices and the Clenshaw algorithm implementation for the tangent space vector-valued orthogonal polynomial basis.

The aim is to, given a PDE say, work with the coefficients of the functions involved (coefficients of their expansion in the OP bases), reframe the PDE in terms of the OP coefficients utilising a suitable timestepping method. For derivative, gradient, divergence etc. operators, we can formulate matrix operators to act on the coefficients, thus making our system to simulate a simple matrix vector system.

\subsection{Example: linear shallow water equations}
Let $\mathbf{u}(x,y,z)$ be the tangential velocity of the flow and $h(x,y,z)$ be the height of the water from some constant reference height $\mathcal{H}$. The linear SWEs are
\begin{align}
\begin{cases}
\frac{\partial \mathbf{u}}{\partial t} + f \unitvec \times \mathbf{u} - \nabla h = \mathbf{0} \\
\frac{\partial h}{\partial t} + \mathcal{H} \nabla \cdot \mathbf{u} = 0
\end{cases}
\end{align}
where $f = 2 \Omega \cos(\theta) =  2 \Omega z$ is the Coriolis parameter (note that the operator on the coefficients vector for multiplication by this parameter would be $F = 2 \Omega (J^z_\nabla)^T$).

Using backward Euler:
\begin{align}
\mathbf{u}_{n+1} &= \mathbf{u}_{n} + \Delta t \: (\nabla h_{n+1} - f \unitvec \times \mathbf{u}_{n+1}) \\
h_{n+1} &= h_n - \Delta t \: \mathcal{H} \nabla \cdot \mathbf{u}_{n+1}
\end{align}

Then using the operators on the coefficient vectors, the timestepping method is: 
\begin{align}
\mathbf{u}^c_{n+1} &= \mathbf{u}^c_{n} + \Delta t \: (G \mathbf{h}^c_{n+1} - F K \mathbf{u}^c_{n+1}) \\
\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
\end{align}
$\iff$
\begin{align}
\mathbf{u}^c_{n+1} &= (I + \Delta t F K + \Delta t^2 \mathcal{H} G D)^{-1} \: (\mathbf{u}^c_{n} + \Delta t \: G \mathbf{h}^c_n) \\
\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
\end{align}

These ideas will also be used to gain similar relationships for the circle arc, half disk, and spherical triangle orthogonal polynomials.

\section{Orthogonal polynomials on the circle arc}

As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$. We will use the lanczos algorithm from ApproxFun \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:
\begin{align}
\bigP_0(x,y) := T^h_0(x) \equiv T_0,
\quad \quad 
\bigP_n(x,y) := \begin{bmatrix}
		T^h_n(x) \\
		y \: U^h_{n-1}(x)
	\end{bmatrix} \in \R^2 \quad \forall n \in \N,
\quad \quad 
\bigP := \begin{bmatrix}
		\bigP_0 \\
		\hline
		\bigP_1 \\
		\hline
		\bigP_2 \\
		\vdots \\
	\end{bmatrix}.
\end{align}

\section{Orthogonal polynomials on the half disk (2D)}

By using a similar process to \cite{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\). We denote the half disk by \(\Omega\) where
\begin{align}
\Omega := \{(x,y) \in \R^2 \quad | \quad 0 < x < 1, \: -(1-x^2)^{\half} < y < (1-x^2)^{\half}\}.
\end{align}

Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:

1) \(\rho\) is a degree 1 polynomial, or

2) \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.

Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\).

We define \(\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)})\) for \( 0 \le k \le n, \: n = 0,1,2,\dots\) and note that the functions \(\{\Pnk\}\) are in fact polynomials in either cases above. Furthermore, they are orthogonal with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on \(R := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}\).

Let \(a, b \in \R\). For the half disk, if we set:

\begin{align}
\alpha &:= 0 \\
\beta &:= 1 \\
\gamma &:= -1 \\
\delta &:= -c = 1 \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= x^a \: (1-x^2)^b \\
w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{align}

we can obtain the 1D orthogonal polynomials \(\{H_n^{(a, b)}\}\) orthogonal with respect to the weight \(x^a \: (1-x^2)^b\), and \(\{P_n^{(b, b)}\}\) orthogonal with respect to the weight \((1-x^2)^b\) (so \(\{P_n^{(b, b)}\}\) are in fact Jacobi polynomials), leading to the two parameter 2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}
 \Pnk(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight 
\begin{align}
W^{(a,b)}(x,y) := x^a \: (1-x^2-y^2)^b, \quad (x,y) \in \Omega.
\end{align}

This provides us with a family of orthogonal polynomial spaces on the half disk.

\subsection{Obtaining the coefficients for expansion of a function}

Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by
\begin{align}
f(x,y) + \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
\end{align}
where
\begin{align}
\bigP(x,y) := \begin{bmatrix}
		\bigP_0(x,y) \\
		\hline
		\bigP_1(x,y) \\
		\hline
		\bigP_2(x,y) \\
		\hline
		\vdots \\
		\hline
		\bigP_N(x,y)
	\end{bmatrix},
\quad
\bigP_n(x,y) := \begin{bmatrix}
		P_{n,0}(x,y) \\
		\vdots \\
		P_{n,n}(x,y)
	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
\end{align}
and where
\begin{align}
\mathbf{f} &:= \begin{bmatrix}
		\mathbf{f}_0 \\
		\hline
		\mathbf{f}_1 \\
		\hline
		\mathbf{f}_2\\
		\hline
		\vdots \\
		\hline
		\mathbf{f}_N
	\end{bmatrix},
\quad
\mathbf{f}_n := \begin{bmatrix}
		f_{n,0} \\
		\vdots \\
		f_{n,n}
	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \\
f_{n,k} &:= \frac{< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
\end{align}

We can develop a quadrature rule on \(\Omega\) in order to evaluate (40) and obtain the coefficients vector \(\mathbf{f}\) as follows.

We will use the substitution that
\begin{align}
x &= s, \quad y = (1-s^2)^\half t.
\end{align}
First, note that
\begin{align}
W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
\end{align}
Using the Golub-Welsch algorithm, we can obtain nodes and weights for 1D quadrature rules on \([0,1]\) with weight \(s^a \: (1-s^2)^{b+\half}\), and on \([-1,1]\) with weight \((1-t^2)^b\).

Let \(f(x,y)\) be a degree \(d\) polynomial that is even in \(y\). Then there exists a polynomial in \(x,y\) \(\tilde{f}(x,y)\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\). Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree \(\tilde{d}\) \bstodo{What is the degree of g?} \ in \(s\) and \(t\). Thus our quadrature rule will be exact:
\begin{align}
& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
&= \sum_j \sum_k w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
&= \sum_j \sum_k w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
\end{align} \bstodo{What is the degree of g?}
where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\).

We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
\begin{align}
\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
\end{align}

Relabelling the weights and nodes as 
\begin{align}
x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
y_{l+(i-1)N} &:= (1-s_l^2)^\half \: t_l, \quad i,l = 1,\dots,N, \\
w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
\end{align}

Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
\begin{align}
\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
\end{align}
exact for any polynomial where \(deg(f(x,\cdot)) \le 2N-1\) and \(deg(f(\cdot,y)) \le 2N-1\). Thus,
\begin{align}
< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
\end{align}

\subsection{Jacobi matrices}

We obtain the 1D OPs via the Lanczos algorithm, which also gives us the coefficients of the relations
\begin{align}
x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) \\
y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
\end{align}
where we note that \(\gamma_n^{(b)} = 0\) \(\forall n = 0,1,2,\dots\).

Next, note that \(<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
\begin{align}
<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &= \int \int_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
&= \begin{cases}
    	\delta_k^{(b)} \: <H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
	\delta_{k-1}^{(b)} \: <H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
	0 & \text{o/w}
      \end{cases}.
\end{align}

We then have that multiplication by \(x\) and \(y\) yields the 3-term recurrences as follows:
\begin{align}
x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
\end{align}
for \((x,y) \in \Omega\), where
\begin{align}
\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
\end{align}

Now, set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
\begin{align}
J_x^{(a,b)} \bigPab(x,y) = x \bigPab(x,y), \quad J_y^{(a,b)} \bigPab(x,y) = y \bigPab(x,y).
\end{align}
Then  \(J_x^{(a,b)}, J_y^{(a,b)}\) have the Banded-Block-Banded structures
\begin{align}
J_{x/y}^{(a,b)} &= \begin{bmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}
\end{align}
where
\begin{align}
A^x_n &:= \begin{bmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
B^x_n &:= \begin{bmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{bmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots,N \\
C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots,N, \\ 
\nonumber \\
A^y_n &:= \begin{bmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
B^y_n &:= \begin{bmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{bmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots,N \\
C^y_n &:= \begin{bmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{bmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots,N.
\end{align}

\subsection{Clenshaw on the half disk}

Combining each system in (71) we can write the block-tridiagonal system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigPab(x,y)
=
\begin{bmatrix}
	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
\end{bmatrix}.
\end{align}
where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each \(n = 0,1,\dots,N\):
\begin{align}
A_n &:= \begin{bmatrix}
		A^x_n \\
		A^y_n \\
		A^z_n
	    \end{bmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
C_n := \begin{bmatrix}
		C^x_n \\
		C^y_n
	    \end{bmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{bmatrix}
		B^x_n \\
		B^y_n
	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y,z) := \begin{bmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}.
\end{align}

For each \(n = 0,1,\dots,N-1\) let \(\Dnt\) be any matrix that is a left inverse of \(A_n\), i.e. such that \(\Dnt A_n = I_{n+3}\). Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align}
\begin{cases}
\bigPab_{-1}(x,y) := 0 \\
\bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots,N.
\end{cases}
\end{align}

Note that we can define \(\Dnt\) as follows. For \(n\) even:
\begin{align}
\Dnt := \begin{bmatrix}
		\frac{1}{\alphaab_{n+1,0,1}} & & &  \\
		& \ddots & & & \\
		& & \frac{1}{\alphaab_{n+1,n,1}} & \\
		0 & \hdots & 0 & \eta_{m-1} & \hdots & 0 & \eta_1 & 0 & \eta_0
	    \end{bmatrix},
\end{align}
where
\begin{align}
m &= \frac{n}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-1.
\end{align}

For \(n\) odd:

\begin{align}
\Dnt := \begin{bmatrix}
		\frac{1}{\alphaab_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaab_{n+1,n,1}} & \\
		\xi & 0 & \hdots & 0 & 0 & \eta_{m-2} & \hdots & 0 & \eta_0
	    \end{bmatrix},
\end{align}
 where
\begin{align}
m &= \frac{n+1}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-2, \\
\xi &= -\frac{\betaab_{n,1,5} \: \eta_{m-2}}{\alphaab_{n+1,0,1}}. \nonumber
\end{align}

\subsection{Example: Laplace operator}

We can obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space
\[
\Wii(x,y) \: \Pii(x,y)
\]
to coefficients in the non-weighted space
\[
\Pii(x,y).
\]
This operator will have Banded-Block-Banded structure. We first note that, for general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
\begin{align}
\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^T \: D_x^{(a,b)} \: \mathbf{f}, \\
\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^T \: D_y^{(a,b)} \: \mathbf{f}, \\
\dx[\Wab(x,y) \: f(x,y)] &= W^{(a-1,b-1)}(x,y) \: \bigP^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f},
\end{align}
Further, we note that there exist Banded-Block-Banded matrix operators \(T^{(1,0)\to(0,0)}\) and \(T^{(0,1)\to(1,1)}\) such that:
\begin{align}
x \bigP^{(1,0)}(x,y) &= \Big(T^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y)
\end{align}
Then, the operator \(\Delta\) is given by
\begin{align}
    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)}
\end{align}

An example is as follows. Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
\begin{align}
    \Delta(\Wii \: u)(x,y) = f(x,y) \quad \text{in } \Omega
    \end{align}
    where we recall
    \begin{align}
    \Wii(x,y) = x \: (1-x^2-y^2)
\end{align}
which ensures a boundary condition that \((\Wii \: u)(x,y) = 0\) on \(\partial \Omega\).

The first step is to obtain the expansion of \(f\) in the \(\Pii\) basis, i.e. obtain the vector \(\mathbf{f}\) such that \(f(x,y) = \mathbf{f}^T \: \bigP^{(1,1)}(x,y) = \sum_{n=0}^N \mathbf{f}_n^T \: \bigP_n^{(1,1)}(x,y)\) for some \(N \in \N\). The coefficients of \(u\) in the \(\Pii\) basis are then given by 
\begin{align}
    \mathbf{u} = \Delta \text{\textbackslash} \mathbf{f}.
\end{align}

\section{Future work - orthogonal polynomials on the hemisphere and spherical triangles}

We will use a similar idea of projecting up from 1D OPs to the interval, to 2D OPs on the disk, to 3D OPs on the sphere. Initially, we will add to the ApproxFun library by adding capability to work with OPs on the half disk (2D) with weights governed by parameters that will be incremented/decremented after an application of an operator (e.g. partial derivative). With this, we can obtain a family of OPs on the hemisphere, with sparse relations.

The goal here is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model. The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.

\bibliography{esa-bib}

\end{document}

### 2018-11-23 papers/spherical-harmonics-paper/spherical-harmonics.tex
Moved tex files into new directory

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mleftright}
\usepackage{todonotes}

\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

%SetFonts

%SetFonts

\usepackage{natbib}
\usepackage{url}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plain}

\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\No}{\mathbb{N}_0}
\newcommand{\Ylm}{Y^m_l}
\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
\newcommand{\Plm}{P^m_l}
\newcommand{\costheta}{\cos\theta}
\newcommand{\sintheta}{\sin\theta}
\newcommand{\cosphi}{\cos\varphi}
\newcommand{\sinphi}{\sin\varphi}
\newcommand{\eimphi}{e^{im\varphi}}
\newcommand{\alphalm}{\alpha^m_l}
\newcommand{\clm}{c^m_l}
\newcommand{\ctilde}{\tilde{c}^m_l}
\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
\newcommand{\chat}{\hat{c}^m_l}
\newcommand{\chatmod}{\hat{c}^{|m|}_l}
\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}

\newcommand{\Atilde}{\tilde{A}_{l,m}}
\newcommand{\Btilde}{\tilde{B}_{l,m}}
\newcommand{\Dtilde}{\tilde{D}_{l,m}}
\newcommand{\Etilde}{\tilde{E}_{l,m}}
\newcommand{\Ftilde}{\tilde{F}_{l,m}}
\newcommand{\Gtilde}{\tilde{G}_{l,m}}
\newcommand{\Alm}{A_{l,m}}
\newcommand{\Blm}{B_{l,m}}
\newcommand{\Dlm}{D_{l,m}}
\newcommand{\Elm}{E_{l,m}}
\newcommand{\Flm}{F_{l,m}}
\newcommand{\Glm}{G_{l,m}}

\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}

\newcommand{\bigP}{\mathbb{P}}
\newcommand{\Pl}{\mathbb{P}_l}
\newcommand{\gradP}{T\mathbb{P}}
\newcommand{\gradPl}{T\mathbb{P}_l}
\newcommand{\gradY}{\nabla Y}
\newcommand{\gradYlm}{\nabla Y^m_l}
\newcommand{\gradpY}{\nabla^\perp Y}
\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}

\newcommand{\Dlt}{D^T_l}

\newcommand{\curlyy}{\bm{\mathcal{Y}}}
\newcommand{\blone}{\beta_{l, 1}}
\newcommand{\blzero}{\beta_{l, 0}}
\newcommand{\blmone}{\beta_{l, -1}}
\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
\newcommand{\cgcoeff}{\mathcal{C}}

\newcommand{\alm}{a_{l,m}}
\newcommand{\blm}{b_{l,m}}
\newcommand{\dlm}{d_{l,m}}
\newcommand{\elm}{e_{l,m}}
\newcommand{\flm}{f_{l,m}}
\newcommand{\glm}{g_{l,m}}
\newcommand{\hlm}{h_{l,m}}
\newcommand{\jlm}{j_{l,m}}
\newcommand{\klm}{k_{l,m}}
\newcommand{\almperp}{a_{l,m}^\perp}
\newcommand{\blmperp}{b_{l,m}^\perp}
\newcommand{\dlmperp}{d_{l,m}^\perp}
\newcommand{\elmperp}{e_{l,m}^\perp}
\newcommand{\flmperp}{f_{l,m}^\perp}
\newcommand{\glmperp}{g_{l,m}^\perp}
\newcommand{\hlmperp}{h_{l,m}^\perp}
\newcommand{\jlmperp}{j_{l,m}^\perp}
\newcommand{\klmperp}{k_{l,m}^\perp}

\newcommand{\unitvec}{\hat{\bm{k}}}

\usepackage{media9}
\graphicspath{ {../sphere/plots} }

\title{Spherical harmonics as orthogonal polynomials in three variables}
\author{Ben Snowball}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\section{Introduction}

\sotodoinline{Give outline}
\sotodoinline{Mention shallow water equation. Include plot.}

We use spherical harmonics as orthogonal polynomials in three variables $x, y, z$  to evaluate functions on the sphere. We lay out our derivations behind how we can evaluate a function on the sphere given coefficients of its expansion in the spherical harmonic basis, as well as demonstrate similar work with vector valued functions in the tangent space of the sphere. Finally, we provide details on how these techniques are used in an example of the linear shallow water equations (SWE).

\subsection{On the Unit Interval (1D)}

On the unit interval, \([-1,1]\), we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite{DLMFJacobiDerivative}:
\begin{align}
& \ddx P^{(a,b)}_l (x) = \text{const.} \times P^{(a+1,b+1)}_{l-1}(x) \\
\implies & \dmdxm P_l(x) = \text{const.} \times P^{(m,m)}_{l-m}(x)
\end{align}
where \(P^{(a,b)}_l (x)\) is the \(l\) degree Jacobi polynomial, orthogonal with weight \(w(x) = (1-x)^a(1+x)^b\), and \(P_l(x) := P^{(0,0)}_l (x)\) is the Legendre polynomial of degree \(l\).

Further, we can define associated Legendre polynomials that are also orthogonal \bstodo{Add reference}:
\begin{align}
\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm P_l(x) = \chat (1-x^2)^\frac{m}{2} P^{(m,m)}_{l-m}(x) \\
P^{-m}_l(x) &:= \ctilde \Plm(x),
\end{align}
for \( m \in \No\) \sotodo{I prefer to write $m = 0,1,2,\ldots$ to avoid any ahance of confusion} where
\begin{align}
\chat &:= \frac{\Gamma(l+m+1)}{(-2)^m\Gamma(l+1)} \\
\ctilde &:= \frac{(-1)^m\Gamma(l-m+1)}{\Gamma(l+m+1)}
\end{align}
using the gamma function \(\Gamma(n) := (n-1)!\) for \(n \in \N\).

\subsection{On the Unit Sphere (3D)}

Let \((x,y,z)\) be a point on the unit sphere such that \(x^2+y^2+z^2 = 1\). Let \(\theta, \varphi\) denote the angles such that
\begin{align}
x &= \sintheta \cosphi \\
y &= \sintheta \sinphi \\
z &= \costheta.
\end{align}
Further, define
\begin{align}
\clm := \Bigg(\frac{(2l+1)\Gamma(l-m+1)}{4\pi\Gamma(l+m+1)}\Bigg)^\frac{1}{2},
\end{align}
and
\begin{align}
\alphalm := \begin{cases} 
		\clm \chatmod \quad \quad \quad \text{if } m \ge 0 \\
		\clm \chatmod \ctildemod \quad \text{if } m < 0
	   \end{cases}.
\end{align}

We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{DLMFSphericalandSpheroidalHarmonics, DLMFFerrers, DLMFAssociatedLegendre}:
\begin{align}
\Ylmfull &:= \clm \eimphi \Plm (\costheta) \\
&= \alphalm (1 - (\costheta)^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(\costheta), \quad \text{where } 0 \le |m| \le l, \, l \in \No.
\end{align} 
Note that we can then express \(\Ylm\) in terms of \(x,y,z\) instead of \(\theta, \varphi\) by noting that \(\costheta = z\) and that \( \eimphi\) can be expressed in terms of \(x,y,z\) for any \(m\in\mathbb{Z}\). Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

\section{Surface of the sphere}

\subsection{Deriving expressions for the multiplication by \(x, y, z\) of the spherical harmonics}

We start by expressing \(x\,\Ylm(x,y,z)\), \(y\,\Ylm(x,y,z)\) , and \(z\,\Ylm(x,y,z)\)  in terms of \(Y^{m'}_{l'}(x,y,z)\) for any point \((x,y,z)\) on the unit circle. For the derivation, see the appendix.

\sotodoinline{Move the derivations to an appendix to make the text flow better}

For \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\):
\begin{align}
x\,\Ylm(x,y,z) &= \Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z), \\
y\,\Ylm(x,y,z) &= - i \, \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \nonumber \\
		     &\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
z\,\Ylm(x,y,z) &= \Flm Y^{m}_{l+1}(x,y,z) + \Glm Y^{m}_{l-1}(x,y,z) ,
\end{align}
where
\begin{align}
\Alm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \Atilde \quad \text{if } m\ge0 \\
			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \tilde{D}_{l,|m|} \quad \text{if } m<0 
	       \end{cases} \\
\Blm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \Btilde \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \tilde{E}_{l,|m|} \quad \text{if } m<0 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise.} 
	        \end{cases} \\
\Dlm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \Dtilde \quad \text{if } m>0 \\
			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \tilde{A}_{l,|m|} \quad \text{if } m\le0 
	       \end{cases} \\
\Elm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \Etilde \quad \text{if } m>0 \\
			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \tilde{B}_{l,|m|} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise.} 
	        \end{cases} \\ 
\Flm &:= \frac{\alphalm}{\alpha^{m}_{l+1}} \Ftilde \\
\Glm &:= \frac{\alphalm}{\alpha^{m}_{l-1}} \Gtilde.
\end{align}

\subsection{Jacobi matrices}

Define, for \(l \in \No\), \(\Pl\) as the column vector of the degree \(l\) spherical harmonic polynomials, and \(\bigP\) as the stacked block vector of the \(\Pl\)'s; that is
\begin{align}
\Pl := \begin{bmatrix}
		Y^{-l}_l \\
		\vdots \\
		Y^l_l
	\end{bmatrix} \in \C^{2l+1}, 
\quad \quad 
\bigP := \begin{bmatrix}
		\bigP_0 \\
		\hline
		\bigP_1 \\
		\hline
		\bigP_2 \\
		\vdots \\
	\end{bmatrix}.
\end{align}

Define the (Jacobi) matrices \(J^x, J^y, J^z\) by 
\begin{align}
J^x \bigP = x \bigP, \quad J^y \bigP = y \bigP, \quad J^z \bigP = z \bigP.
\end{align}
Then, using equations (38--40), we have that the Jacobi matrices take the following block-tridiagonal form:
\begin{align}
J^x &= \begin{bmatrix}
		B^x_0 & A^x_0 & & & & \\
		C^x_1 & B^x_1 & A^x_1 & & & \\
		& C^x_2 & B^x_2 & A^x_2  & & & \\
		& & C^x_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}, \\
J^y &= \begin{bmatrix}
		B^y_0 & A^y_0 & & & & \\
		C^y_1 & B^y_1 & A^y_1 & & & \\
		& C^y_2 & B^y_2 & A^y_2  & & & \\
		& & C^y_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}, \\
J^z &= \begin{bmatrix}
		B^z_0 & A^z_0 & & & & \\
		C^z_1 & B^z_1 & A^z_1 & & & \\
		& C^z_2 & B^z_2 & A^z_2  & & & \\
		& & C^z_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix},
\end{align}
where for \(l \in \No\),
\begin{align}
A^x_l &:= \begin{bmatrix}
		D_{l,-l} & 0 & A_{l,-l} & & \\
		& \ddots & \ddots & \ddots & \\
		& & D_{l,l} & 0 & A_{l,l} \\
	    \end{bmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
B^x_l &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
C^x_l &:= \begin{bmatrix}
		B_{l,-l} & & \\
		0 & \ddots & \\
		E_{l,-l+2} & \ddots & B_{l,l-2} \\
		& \ddots & 0 \\
		& & E_{l,l}
	    \end{bmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0), \\
A^y_l &:= -i \begin{bmatrix}
		-D_{l,-l} & 0 & A_{l,-l} & & \\
		& \ddots & \ddots & \ddots & \\
		& & -D_{l,l} & 0 & A_{l,l} \\
	    \end{bmatrix} \in \C^{(2l+1)\times(2l+3)}, \\
B^y_l &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
C^y_l &:= -i \begin{bmatrix}
		B_{l,-l} & & \\
		0 & \ddots & \\
		-E_{l,-l+2} & \ddots & B_{l,l-2} \\
		& \ddots & 0 \\
		& & -E_{l,l}
	    \end{bmatrix} \in \C^{(2l+1)\times(2l-1)} \quad (l \ne 0), \\
A^z_l &:= \begin{bmatrix}
		0 & F_{l,-l} & 0 & & \\
		& \ddots & \ddots & \ddots & \\
		& & 0 & F_{l,l} & 0 \\
	    \end{bmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
B^z_l &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
C^z_l &:= \begin{bmatrix}
		0 & & \\
		G_{l,-l+1} & \ddots & \\
		0 & \ddots & 0 \\
		& \ddots & G_{l,l-1} \\
		& & 0
	    \end{bmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0). \\
\end{align}

\subsection{Three-term recurrence relation for \(\bigP\)}

Combining each system in (48) we can write the tridiagonal-block system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{c|c|cc}
		1  & & & \\
		\hline 
		B^x_0 - xI_1 & A^x_0 & & \\
		B^y_0 - yI_1 & A^y_0 & & \\
		B^z_0 - zI_1 & A^z_0 & & \\
		\hline 
		C^x_1 & B^x_1 - xI_{3} & \quad A^x_1 \quad & \\
		C^y_1 & B^y_1 - yI_{3} & \quad A^y_1 \quad & \\
		C^z_1 & B^z_1 - zI_{3} & \quad A^z_1 \quad & \\
		\hline
		& C^x_2 & \ddots & \ddots \\
		& C^y_2 & & \\
		& C^z_2 & & \\
		\hline
		& & \ddots &
\end{array}
\mright]
\bigP
=
\begin{bmatrix}
	\alpha^0_0 \\ 0 \\ 0 \\ 0 \\ \\ \vdots \\ \\ \vdots \\ \\ \vdots \\ \\
\end{bmatrix},
\end{align}
where \(I_{2l+1}\) is the \((2l+1)\times(2l+1)\) identity matrix.

Let us define the joint matrices that comprise each block. For each \(l \in \No\):
\begin{align}
A_l &:= \begin{bmatrix}
		A^x_l \\
		A^y_l \\
		A^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+3)}, \quad
C_l := \begin{bmatrix}
		C^x_l \\
		C^y_l \\
		C^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l-1)} \quad (l \ne 0), \\
B_l &:= \begin{bmatrix}
		B^x_l \\
		B^y_l \\
		B^z_l
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}, \quad
G_l(x,y,z) := \begin{bmatrix}
		xI_{2l+1} \\
		yI_{2l+1} \\
		zI_{2l+1}
	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}.
\end{align}
Then our system (62) simply becomes
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1  & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigP
=
\begin{bmatrix}
	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
\end{bmatrix}.
\end{align}

\sotodoinline{Cite Dunkl and Xu here.}

For each \(l \in \No\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e. such that \(\Dlt A_l = I_{2l+3}\). Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain the lower triangular system \citep{dunkl2014orthogonal}
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1  & & & \\
		D^T_0(B_0-G_0) & I_3 & & \\
		D^T_1 C_1 & D^T_1(B_1-G_1) & \quad I_5 \quad & \\
		& D^T_2 C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigP
=
\begin{bmatrix}
	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
\end{bmatrix}.
\end{align}

Expanding this we can solve the system (66) to find a three-term recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\):
\begin{align}
\begin{cases}
\bigP_{-1} := 0 \\
\bigP_{0} := \alpha^0_0 \\
\bigP_{l+1} = -\Dlt (B_l-G_l) \Pl - \Dlt C_l  \,\bigP_{l-1}, \quad l \in \N.
\end{cases}
\end{align}

We note that we can choose the matrices \(\Dlt\) in the following way. For \(l \in \N\), we set
\begin{align}
\Dlt = \begin{bmatrix}
		\hat{A}^{x,y}_l & 0_{(2l+3)\times(2l+1)}
	  \end{bmatrix} \in \R^{(2l+3)\times3(2l+1)}
\end{align}
where \(0_{(2l+3)\times(2l+1)}\) the zero matrix in \(\R^{(2l+3)\times(2l+1)}\), and \(\hat{A}^{x,y}_l \in \R^{(2l+3)\times2(2l+1)}\) is the left inverse of the matrix \(\begin{bmatrix} A^x_l \\ A^y_l \end{bmatrix}\), given by
\begin{align}
\hat{A}^{x,y}_l = \begin{bmatrix}
		\frac{1}{2D_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2D_{l,-l}} & 0 & \hdots & 0 \\
		& \ddots & & & & \ddots & & \\
		& & \ddots & & & & \ddots & \\
		& & & \frac{1}{2D_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2D_{l,l}} \\
		0 & \hdots & \frac{1}{2A_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2A_{l,l-1}} & 0 \\
		0 & \hdots & 0 & \frac{1}{2A_{l,l}} & 0 & \hdots & 0 & \frac{i}{2A_{l,l}}
	  \end{bmatrix}.
\end{align}
For \(l = 0\) we set
\begin{align}
D^T_0 = \begin{bmatrix}
		\frac{1}{2D_{0,0}}&  -\frac{i}{2D_{0,0}} & 0 \\
		0 & 0 & \frac{1}{F_{0,0}} \\
		\frac{1}{2A_{0,0}}&  \frac{i}{2A_{0,0}} & 0 
	     \end{bmatrix}.
\end{align}

\subsection{Evaluation of a scalar function on the sphere}
We can use the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e. suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
\begin{align}
f(x,y,z) \approx \sum^N_{l=0} \bold{f}_l^T \, \Pl (x,y,z), \quad \text{where } \Pl (x,y,z), \bold{f}_l \in \R^{2l+1} \text{ for each } l \in \{0,\dots,N\}.
\end{align}

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}
where $\bigP_0 \equiv \alpha^0_0$.

\sotodoinline{Mention $a(J^x, J^y, J^z)$ for variable coefficients here. Mention it is banded-block-banded.}
We can use the same algorithm to output a matrix operator for a function, given we know the coefficients of its expansion in the SH basis. We simply make the adjustment that
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = (\bold{f}_n^T)I - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(J^x,J^y,J^z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } f(x,y,z) \approx \alpha^0_0 f_0 + \bm{\gamma}_{1}^T \bigP_1(J^x,J^y,J^z) - \alpha^0_0 \bm{\gamma}_{2}^T D^T_1 C_1.
\end{align*}
where $\bigP_1(J^x,J^y,J^z)$ is a vector of matrices, $G_n(J^x,J^y,J^z)$ is a matrix of matrices, and so each $\bm{\gamma}_n$ is a vector of matrices.

\section{Tangent space}

Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient  and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics are a basis for the tangent space, namely \(\gradYlm\), \(\gradpYlm\). Note that the perpendicular gradient is related to the regular surface gradient by
\begin{align}
\gradpYlm(x,y,z) = \unitvec \times \gradYlm(x,y,z),
\end{align}
where \(\unitvec\) is the unit vector normal to the surface of the sphere at the point \((x,y,z)\), i.e. as we are looking at the unit sphere, \(\unitvec\) is simply given by
\begin{align}
\unitvec = \begin{bmatrix} x \\ y \\ z \end{bmatrix}.
\end{align}

Importantly,  \(\gradYlm\), \(\gradpYlm\) are in fact vector-valued orthogonal polynomials, that span all vector-valued polynomials modulo the vanishing ideal $\{ \bm{p} \in (\Pi_3)^3  : \unitvec \cdot \bm{p} = 0 \}$. A simple calculation shows that such orthogonal polynomials must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal. In the following, we determine the coefficients of these Jacobi operators, as well as other important operators needed.

\subsection{Spin-1 tensor spherical harmonics}

The gradient and perpendicular gradient of a spherical harmonic \(\Ylm(x,y,z)\) can be expressed in terms of spin-1 tensor spherical harmonics, which in turn can each be expressed as a vector-weighted sum of spherical harmonics. For the derivation and more on vector spherical harmonics, see the appendix.

For any \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) we have that
\begin{align}
\nabla \Ylm &= \blmone \, \curlyy^{l-1,1}_{l,m} + \blone \curlyy^{l+1,1}_{l,m}, \\
\nabla^\perp \Ylm &= \blzero \, \curlyy^{l,1}_{l,m},
\end{align}
where
\begin{align}
\blmone &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
\blzero := i \, \big(l(l+1)\big)^\half, \quad
\blone := l \, \Big(\frac{l+1}{2l+1}\Big)^\half.
\end{align}

\sotodoinline{I think the following should also be moved to appendix because it's technical details}

\subsection{Deriving expressions for the multiplication by \(x, y, z\) of \(\nabla \Ylm\), \(\nabla^\perp \Ylm\) }

We start by establishing some equations to be used to find \(x \,\gradYlm (x,y,z)\), \(x \,\gradpYlm (x,y,z)\) etc. in terms of \(\gradY^{m'}_{l'} (x,y,z)\), \(\gradpY^{m'}_{l'} (x,y,z)\). First, we have that
\begin{align}
x \,\gradYlm 
&= x \, \blmone \, \curlyy^{l-1,1}_{l,m} + x \, \blone \curlyy^{l+1,1}_{l,m} \nonumber
\\
&= \sum_{m_s=-1}^{1} \chivec \Big[ \blmone \, \cgcoeff^{l-1,m_s}_{l,m} \big\{ A_{l-1,m-m_s} \, Y^{m-m_s+1}_{l} + B_{l-1,m-m_s} \, Y^{m-m_s+1}_{l-2} \nonumber \\ 
& \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + D_{l-1,m-m_s} \, Y^{m-m_s-1}_{l} + E_{l-1,m-m_s} \, Y^{m-m_s-1}_{l-2} \big\} \nonumber \\
& \quad \quad \quad \quad \quad \quad \quad + \blone \, \cgcoeff^{l+1,m_s}_{l,m} \big\{ A_{l+1,m-m_s} \, Y^{m-m_s+1}_{l+2} + B_{l+1,m-m_s} \, Y^{m-m_s+1}_{l} \nonumber \\ 
& \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + D_{l+1,m-m_s} \, Y^{m-m_s-1}_{l+2} + E_{l+1,m-m_s} \, Y^{m-m_s-1}_{l} \big\} \Big]
\end{align}
\begin{align}
\implies x \,\gradYlm &= \alm \gradY^{m+1}_{l+1} + \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad + \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l},
\end{align}
where, for any valid \(m_s\) value,
\begin{align}
\alm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} A_{l+1,m-m_s}, \\
\blm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} B_{l-1,m-m_s}, \\
\dlm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} D_{l+1,m-m_s}, \\
\elm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m-1}} E_{l-1,m-m_s}, \\
\hlm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, A_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, B_{l+1,m-m_s} \nonumber \\
& \quad \quad \quad \quad \quad \quad \quad - \alm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - \blm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
\jlm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, D_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, E_{l+1,m-m_s} \nonumber \\
& \quad \quad \quad \quad \quad \quad \quad - \dlm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - \elm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big].
\end{align}

Note that (it can be shown) these are constants for each \(l,m\) pair despite appearing to depend on the value of \(m_s\); we need only use any valid \(m_s\) value for each expression. By ``valid'', we mean the Clebsch--Gordan coefficients do not vanish for that \(m_s\) value when used.

Similarly, we have that
\begin{align}
y \,\gradYlm &= i \, \Big[ -\alm \gradY^{m+1}_{l+1} - \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad - \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l} \Big].
\end{align}

Further, we also have that
\begin{align}
z \,\gradYlm 
&= z \, \blzero \, \curlyy^{l,1}_{l,m} \nonumber
\\
&= \sum_{m_s=-1}^{1} \chivec \Big[ \blmone \, \cgcoeff^{l-1,m_s}_{l,m} \big\{ F_{l-1,m-m_s} \, Y^{m-m_s}_{l} + G_{l-1,m-m_s} \, Y^{m-m_s}_{l-2} \big\} \nonumber \\ 
& \quad \quad \quad \quad \quad \quad \quad + \blone \, \cgcoeff^{l+1,m_s}_{l,m} \big\{ F_{l+1,m-m_s} \, Y^{m-m_s}_{l+2} + G_{l+1,m-m_s} \, Y^{m-m_s}_{l} \big\} \Big]
\end{align}
\begin{align}
\implies z \,\gradYlm &= \flm \gradY^{m}_{l+1} + \glm \gradY^{m}_{l-1} + \klm \gradpY^{m}_{l},
\end{align}
where, for any valid \(m_s\) value,
\begin{align}
\flm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} F_{l+1,m-m_s}, \\
\glm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} G_{l-1,m-m_s}, \\
\klm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, F_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, G_{l+1,m-m_s}  \nonumber \\
& \quad \quad \quad \quad \quad \quad \quad - \flm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \glm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big].
\end{align}

We can similarly yield the relations for the perpendicular gradients:
\begin{align}
x \,\gradpYlm &= \almperp \gradpY^{m+1}_{l+1} + \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad + \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
y \,\gradpYlm &= i \, \Big[ -\almperp \gradpY^{m+1}_{l+1} - \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
& \quad \quad \quad - \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
z \,\gradpYlm &= \flmperp \gradpY^{m}_{l+1} + \glmperp \gradpY^{m}_{l-1} + \klmperp \gradY^{m}_{l},
\end{align}
where it can be shown that
\begin{align}
\almperp &= \alm^*, \quad \blmperp = \blm^*, \quad \dlmperp = \dlm^*, \quad \elmperp = \elm^*, \nonumber \\
\flmperp = \flm^*,& \quad \glmperp = \glm^*, \quad \hlmperp = \hlm^*, \quad \jlmperp = \jlm^*, \quad \klmperp = \klm^*,
\end{align}
where \(^*\) denotes the complex conjugate.

\subsection{Jacobi matrices}

Define \(\gradP\) \sotodo{This notation could be confusing. I'd say use either $T{\mathbb P}$ or ${\mathbb P}^T$} as the column vector
\begin{align}
\gradP = \begin{bmatrix} \gradP_0 \\ \gradP_1 \\ \vdots \end{bmatrix}, \quad \text{where } \quad \gradP_l = \begin{bmatrix} \gradY^{-l}_l \\ \gradpY^{-l}_l \\ \vdots \\ \gradY^{l}_l \\ \gradpY^{l}_l  \end{bmatrix} \quad \forall \: l \in \No.
\end{align}

Further, let
\begin{align}
\bold{u}^c = \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
		    \end{bmatrix},		  
\quad
\text{where}
\quad
\bold{u}^c_l &= \begin{bmatrix}
				u_{l,-l} \\
				u^{\perp}_{l,-l} \\
				\vdots \\
				u_{l,l} \\
				u^{\perp}_{l,l}
		        \end{bmatrix} \in \R^{2(2l+1)} \quad \forall l = 0,1,2,\dots
\end{align}

We can define the Jacobi operators \(J_\nabla^x, J_\nabla^y, J_\nabla^z\) as follows. Let $S := \{ \bold{x} := (x,y,z) \quad | \quad ||\bold{x}|| = 1\}$ be the unit sphere in $\R^3$ and $\mathbf{u}(x,y,z)$ be a function on $S$ with vector values in the tangent space. That is, there exist coefficients $\{u_{l,m}, u^\perp_{l,m}\}$ such that, $\forall (x,y,z) \in S$:
\begin{align}
\mathbf{u}(x,y,z) &= \sum_{l=0}^\infty \sum_{m=-l}^l \big( u_{l,m} \gradYlm(x,y,z) + u^\perp_{l,m} \gradpYlm(x,y,z) \big) \\
&= (\mathbf{u}^c)^T \gradP(x,y,z)
\end{align}

Then \(J_\nabla^x\) is defined as the matrix such that 
\begin{align}
x \mathbf{u}(x,y,z) = \sum_{l=0}^\infty \sum_{m=-l}^l \big( \tilde{u}_{l,m} \gradYlm(x,y,z) + \tilde{u}^\perp_{l,m} \gradpYlm(x,y,z) \big), \quad \forall (x,y,z) \in S,
\end{align}
where
\begin{align}
(J_\nabla^x)^T \mathbf{u}^c = \mathbf{\tilde{u}}^c,
\quad
\bold{\tilde{u}}^c = \begin{bmatrix}
			\bold{\tilde{u}}_0^c \\
			\bold{\tilde{u}}_1^c \\
			\vdots \\
		    \end{bmatrix},		  
\quad
\text{and}
\quad
\bold{\tilde{u}}^c_l &= \begin{bmatrix}
				\tilde{u}_{l,-l} \\
				\tilde{u}^{\perp}_{l,-l} \\
				\vdots \\
				\tilde{u}_{l,l} \\
				\tilde{u}^{\perp}_{l,l}
		        \end{bmatrix} \quad \forall l = 0,1,2,\dots,
\end{align}
and \(J_\nabla^y, J_\nabla^z\) are defined similarly.

In other words, abusing notation, we can write that
\begin{align}
J_\nabla^x \gradP = x \gradP, \quad J_\nabla^y \gradP = y \gradP, \quad J_\nabla^z \gradP = z \gradP,
\end{align}
where each sub-vector entry \(\gradYlm\) and \(\gradpYlm\) of \(\gradP\) is treated as a single element of the vector \(\gradP\) in the matrix-vector calculation. \sotodo{I don't know what this means}

The Jacobi matrices have the following form:
\begin{align}
J_\nabla^x &= \begin{bmatrix}
		B^x_0 & A^x_0 & & & & \\
		C^x_1 & B^x_1 & A^x_1 & & & \\
		& C^x_2 & B^x_2 & A^x_2  & & & \\
		& & C^x_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}, \\
J_\nabla^y &= \begin{bmatrix}
		B^y_0 & A^y_0 & & & & \\
		C^y_1 & B^y_1 & A^y_1 & & & \\
		& C^y_2 & B^y_2 & A^y_2  & & & \\
		& & C^y_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}, \\
J_\nabla^z &= \begin{bmatrix}
		B^z_0 & A^z_0 & & & & \\
		C^z_1 & B^z_1 & A^z_1 & & & \\
		& C^z_2 & B^z_2 & A^z_2  & & & \\
		& & C^z_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix},
\end{align}
where for \(l \in \No\),

\begin{align}
A^x_l &:= \begin{bmatrix}
		d_{l,-l} & 0 & 0 & 0 & a_{l,-l} & & \\
		& d^\perp_{l,-l} & 0 & 0 & 0 & a_{l,-l} & & \\
		& & \ddots & \ddots & \ddots & \ddots & \ddots & \\
		& & & d_{l,l} & 0 & 0 & 0 & a_{l,l} \\
		& & & & d^\perp_{l,l} & 0 & 0 & 0 & a^\perp_{l,-l}  \\
	    \end{bmatrix} \in \R^{2(2l+1)\times2(2l+3)}, 
\end{align}
\begin{align}
B^x_l &:= \begin{bmatrix}
		0 & 0 & 0 & h_{l,-l} & & \\
		0 & 0 & h^\perp_{l,-l} & 0 & & \\
		0 & j_{l,-l+1} & 0 & 0 & 0 & h_{l,-l+1} \\
		j^\perp_{l,-l+1} & 0 & 0 & 0 & h^\perp_{l,-l+1} & 0 \\
		& \ddots & & \ddots & & \ddots \\
		& & \ddots & & \ddots & & \ddots \\
		& & 0 & j_{l,l-1} & 0 & 0 & 0 & h_{l,l-1} \\
		& & j^\perp_{l,l-1} & 0 & 0 & 0 & h^\perp_{l,l-1} & 0 \\
		& & & & 0 & j_{l,l} & 0 & 0 \\
		& & & & j^\perp_{l,l} & 0 & 0 & 0 \\
	\end{bmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, 
\end{align}
\begin{align}
C^x_l &:= \begin{bmatrix}
		b_{l,-l} & & \\
		0 & b^\perp_{l,-l} \\
		0 & 0 & \ddots & \\
		0 & 0 & \ddots & \ddots \\
		e_{l,-l+2} & 0 & \ddots & \ddots & b_{l,l-2} \\
		& e^\perp_{l,-l+2} & \ddots & \ddots & 0 & b^\perp_{l,l-2} \\
		& & \ddots & \ddots & 0 & 0 \\
		& & & \ddots & 0 & 0 \\
		& & & & e_{l,l} & 0 \\
		& & & & & e^\perp_{l,l} \\
	    \end{bmatrix} \in \R^{2(2l+1)\times2(2l-1)} \quad (l \ne 0),
\end{align}
\begin{align}
A^y_l &:= \begin{bmatrix}
		-d_{l,-l} & 0 & 0 & 0 & a_{l,-l} & & \\
		& -d^\perp_{l,-l} & 0 & 0 & 0 & a_{l,-l} & & \\
		& & \ddots & \ddots & \ddots & \ddots & \ddots & \\
		& & & -d_{l,l} & 0 & 0 & 0 & a_{l,l} \\
		& & & & -d^\perp_{l,l} & 0 & 0 & 0 & a^\perp_{l,-l}  \\
	    \end{bmatrix} \in \R^{2(2l+1)\times2(2l+3)},
\end{align}
\begin{align}
B^y_l &:= \begin{bmatrix}
		0 & 0 & 0 & h_{l,-l} & & \\
		0 & 0 & h^\perp_{l,-l} & 0 & & \\
		0 & -j_{l,-l+1} & 0 & 0 & 0 & h_{l,-l+1} \\
		-j^\perp_{l,-l+1} & 0 & 0 & 0 & h^\perp_{l,-l+1} & 0 \\
		& \ddots & & \ddots & & \ddots \\
		& & \ddots & & \ddots & & \ddots \\
		& & 0 & -j_{l,l-1} & 0 & 0 & 0 & h_{l,l-1} \\
		& & -j^\perp_{l,l-1} & 0 & 0 & 0 & h^\perp_{l,l-1} & 0 \\
		& & & & 0 & -j_{l,l} & 0 & 0 \\
		& & & & -j^\perp_{l,l} & 0 & 0 & 0 \\
		\end{bmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, 
\end{align}
\begin{align}
C^y_l &:= \begin{bmatrix}
		b_{l,-l} & & \\
		0 & b^\perp_{l,-l} \\
		0 & 0 & \ddots & \\
		0 & 0 & \ddots & \ddots \\
		-e_{l,-l+2} & 0 & \ddots & \ddots & b_{l,l-2} \\
		& -e^\perp_{l,-l+2} & \ddots & \ddots & 0 & b^\perp_{l,l-2} \\
		& & \ddots & \ddots & 0 & 0 \\
		& & & \ddots & 0 & 0 \\
		& & & & -e_{l,l} & 0 \\
		& & & & & -e^\perp_{l,l} \\
	    \end{bmatrix} \in \R^{2(2l+1)\times2(2l-1)} \quad (l \ne 0),
\end{align}
\begin{align}
A^z_l &:= \begin{bmatrix}
		0 & 0 & f_{l,-l} \\
		& & & f^\perp_{l,-l} \\
		& & & & \ddots \\
		& & & & & f_{l,l} \\
		& & & & & & f^\perp_{l,l} & 0 & 0 \\
	    \end{bmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
\end{align}
\begin{align}
B^z_l &:= \begin{bmatrix}
		0 & k_{l,-l} \\
		k^\perp_{l,-l} & 0 \\
		& & \ddots \\
		& & & 0 & k_{l,l} \\
		& & & k^\perp_{l,l} & 0 \\
	\end{bmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, 
\end{align}
\begin{align}
C^z_l &:= \begin{bmatrix}
		0 \\
		0 \\
		g_{l,-l+1} \\
		& g^\perp_{l,-l+1} \\
		& & \ddots \\
		& & & g_{l,l-1} \\
		& & & & g^\perp_{l,l-1} \\
		& & & & 0 \\
		& & & & 0 \\
	    \end{bmatrix} \in \R^{2(2l+1)\times2(2l-1)} \quad (l \ne 0). \\
\end{align}

\subsection{Three-term recurrence relation for \(\gradP\)}

Combining each system in (102) we can write, for \((x,y,z)\) on the unit sphere, the tridiagonal-block system\sotodo{Is $0_3$ wrong below?}
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{c|c|c|cc}
		I_{2}  & & & \\
		\hline
		& I_{6}  & & & \\
		\hline 
		& B^x_1 - xI_6 & A^x_1 & & \\
		& B^y_1 - yI_6 & A^y_1 & & \\
		& B^z_1 - zI_6 & A^z_1 & & \\
		\hline 
		& C^x_2 & B^x_2 - xI_{10} & \quad A^x_2 \quad & \\
		& C^y_2 & B^y_2 - yI_{10} & \quad A^y_2 \quad & \\
		& C^z_2 & B^z_2 - zI_{10} & \quad A^z_2 \quad & \\
		\hline
		& & C^x_3 & \ddots & \ddots \\
		& & C^y_3 & & \\
		& & C^z_3 & & \\
		\hline
		& & & \ddots &
\end{array}
\mright]
\gradP
=
\begin{bmatrix}
	\underline{0}_3 \\ \underline{0}_3 \\ \gradP_1 \\ \underline{0}_3 \\ \\ \vdots \\ \\ \vdots \\ \\ \vdots \\ \\
\end{bmatrix},
\end{align}
where \(I_{2l+1}\) is the \((2l+1)\times(2l+1)\) identity matrix. For clarity we again treat each sub-vector \(\gradYlm, \gradpYlm\) of \(\gradP\) in the matrix-vector multiplication as a single element.

Let us define the joint matrices that comprise each block. For each \(l \in \N\):
\begin{align}
A_l &:= \begin{bmatrix}
		A^x_l \\
		A^y_l \\
		A^z_l
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l+3)}, \quad
C_l := \begin{bmatrix}
		C^x_l \\
		C^y_l \\
		C^z_l
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l-1)} \quad (l \ne 1), \\
B_l &:= \begin{bmatrix}
		B^x_l \\
		B^y_l \\
		B^z_l
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l+1)}, \quad
G_l(x,y,z) := \begin{bmatrix}
		xI_{2(2l+1)} \\
		yI_{2(2l+1)} \\
		zI_{2(2l+1)}
	    \end{bmatrix} \in \R^{6(2l+1)\times2(2l+1)}.
\end{align}
Then our system simply becomes
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		I_2 \\
		0 & I_6 & & \\
		C_1 & B_1-G_1 & A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\gradP
=
\begin{bmatrix}
	\underline{0}_3 \\ \underline{0}_3 \\ \gradP_1 \\ \underline{0}_3 \\ \\ \vdots \\
\end{bmatrix}.
\end{align}

For each \(l \in \N\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e. such that \(\Dlt A_l = I_{2(2l+3)}\). Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain the lower triangular system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		I_2 \\
		0 & I_6 & & \\
		D^T_1 C_1 & D^T_1(B_1-G_1) & \quad I_{10} \quad & \\
		& D^T_2 C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\gradP
=
\begin{bmatrix}
	\underline{0}_3 \\ \underline{0}_3 \\ \gradP_1 \\ \underline{0}_3 \\ \\ \vdots \\
\end{bmatrix}.
\end{align}

Expanding this we can find a three-term recurrence relation for each \(\gradP_{l+1}\) in terms of the previous two sub-vectors \(\gradPl\) and \(\gradP_{l-1}\):
\begin{align}
\begin{cases}
\gradP_{0}(x,y,z) := \underline{0}_6 \\
\\
\gradP_{1}(x,y,z) := \begin{bmatrix}
				\gradY_1^{-1}(x,y,z) \\ \gradpY_1^{-1}(x,y,z) \\ \gradY_1^{0}(x,y,z) \\ \gradpY_1^{0}(x,y,z) \\ \gradY_1^{1}(x,y,z) \\ \gradpY_1^{1}(x,y,z) \\
			\end{bmatrix} \\
\\
\gradP_{l+1}(x,y,z) = -\Dlt [B_l-G_l(x,y,z)] \gradPl(x,y,z) - \Dlt C_l  \,\gradP_{l-1}(x,y,z), \quad l \in \N.
\end{cases}
\end{align}

We note that we can choose the matrices \(\Dlt\) in the following way. For \(l \in \N\), we set
\begin{align}
\Dlt = \begin{bmatrix}
		\hat{A}^{x,y}_l & 0_{2(2l+3)\times2(2l+1)}
	  \end{bmatrix} \in \R^{2(2l+3)\times6(2l+1)}
\end{align}
where \(0_{2(2l+3)\times2(2l+1)}\) the zero matrix in \(\R^{2(2l+3)\times2(2l+1)}\), and \(\hat{A}^{x,y}_l \in \R^{2(2l+3)\times4(2l+1)}\) is the left inverse of the matrix \(\begin{bmatrix} A^x_l \\ A^y_l \end{bmatrix}\), given by
\begin{align}
\hat{A}^{x,y}_l = \begin{bmatrix}
		\frac{1}{2d_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2d_{l,-l}} & 0 & \hdots &  \\
		0 & \frac{1}{2d^\perp_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2d^\perp_{l,-l}} & 0 & \hdots &  \\
		& & \ddots & & & & \ddots & & \\
		\vdots & & & \ddots & & & & \ddots & \\
		& & & & \frac{1}{2d_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2d_{l,l}} & \\
		& & & & & \frac{1}{2d^\perp_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2d^\perp_{l,l}} \\
		0 & \hdots & \frac{1}{2a_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2a_{l,l-1}} & 0 & 0 & 0 \\
		0 & \hdots & 0 & \frac{1}{2a^\perp_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2a^\perp_{l,l-1}} & 0 & 0\\
		0 & \hdots & \hdots & 0 & \frac{1}{2a_{l,l}} & 0 & \hdots & 0 & \frac{i}{2a_{l,l}} & 0 \\
		0 & \hdots & \hdots & & 0 & \frac{1}{2a^\perp_{l,l}} & 0 & \hdots & 0 & \frac{i}{2a^\perp_{l,l}} \\
	  \end{bmatrix}.
\end{align}

\subsection{Evaluation of a vector valued function in the tangent space on the sphere}

We can use the Clenshaw algorithm again to evaluate a vector valued function in the tangent space at a point $(x,y,z) \in S$. Suppose
\begin{align}
\mathbf{u}(x,y,z) \approx \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big].
\end{align}

The Clenshaw algorithm is then as follows:
\begin{align*}
\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
\quad &\text{2) } \text{For } n = N:-1:1 \\
\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = (\mathbf{u}^c_n)^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
\quad &\text{3) } \text{Output: } \mathbf{u}(x,y,z) \approx \bm{\gamma}_{1}^T \gradP_1(x,y,z) \\
&\quad \quad \quad \quad \Big( = (\bm{\gamma}_{1})_1 \gradY_1^{-1} + (\bm{\gamma}_{1})_2 \gradpY_1^{-1} + (\bm{\gamma}_{1})_3 \gradY_1^{0} + (\bm{\gamma}_{1})_4 \gradpY_1^{0} + (\bm{\gamma}_{1})_5 \gradY_1^{1} + (\bm{\gamma}_{1})_6 \gradpY_1^{1} \Big)
\end{align*}

\subsection{Deriving matrices and calculations for certain operations}

Define \(\unitvec\) as the unit outward normal vector at the point on the sphere \((x,y,z)\), so that
\begin{align}
\unitvec = \begin{bmatrix} x \\ y \\ z \end{bmatrix}.
\end{align}

Let \(S := \{ \bold{x} := (x,y,z) \quad | \quad ||\bold{x}|| = 1\}\) be the unit sphere in \(\R^3\) and let \(T_x^S\) denote the tangent space at the point \(\bold{x} \in S\). Further, let \(\bold{u}(x,y,z)\), \(\bold{v}(x,y,z)\) be two vector valued functions for \((x,y,z)\) on the unit sphere with values in the tangent space (representing the tangential velocity of a flow for example) and let \(h(x,y,z)\) be a scalar function on the sphere.

Then, 
\begin{align}
\bold{u} &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ], \\
\bold{v} &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ v_{l,m} \gradYlm + v^{\perp}_{l,m} \gradpYlm ], \\
h &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} h_{l,m} \Ylm,
\end{align}
for some real coefficients \(u_{l,m}, u^{\perp}_{l,m}, v_{l,m}, v^{\perp}_{l,m}, h_{l,m}\). Define the vectors of these coefficients as 
\begin{align}
\bold{u}^c &= \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
		    \end{bmatrix},
\quad
\bold{v}^c = \begin{bmatrix}
			\bold{v}_0^c \\
			\bold{v}_1^c \\
			\vdots \\
		    \end{bmatrix},
\quad
\bold{h}^c = \begin{bmatrix}
			\bold{h}_0^c \\
			\bold{h}_1^c \\
			\vdots \\
		    \end{bmatrix},		  
\end{align}
where
\begin{align}
\bold{u}^c_l &= \begin{bmatrix}
				u_{l,-l} \\
				u^{\perp}_{l,-l} \\
				\vdots \\
				u_{l,l} \\
				u^{\perp}_{l,l}
		        \end{bmatrix},
\quad
\bold{v}^c_l = \begin{bmatrix}
				v_{l,-l} \\
				v^{\perp}_{l,-l} \\
				\vdots \\
				v_{l,l} \\
				v^{\perp}_{l,l}
		        \end{bmatrix},
\quad
\bold{h}^c_l = \begin{bmatrix}
				h_{l,-l} \\
				\vdots \\
				h_{l,l} \\
		        \end{bmatrix},
\quad \forall l \in \No	  
\end{align}.

Then, for large enough \(N \in \N\) we have that
\begin{align}
\bold{u} &\approx \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ], \\
\bold{v} &\approx \sum_{l=0}^{N} \sum_{m=-l}^{l} [ v_{l,m} \gradYlm + v^{\perp}_{l,m} \gradpYlm ], \\
h &\approx \sum_{l=0}^{N} \sum_{m=-l}^{l} h_{l,m} \Ylm,
\end{align}
and so we define the truncated coefficient vectors for some \(N \in \N\) as
\begin{align}
\bold{u}^c &= \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
			\bold{u}_N^c
		    \end{bmatrix},
\quad
\bold{v}^c = \begin{bmatrix}
			\bold{v}_0^c \\
			\bold{v}_1^c \\
			\vdots \\
			\bold{v}_N^c
		    \end{bmatrix},
\quad
\bold{h}^c = \begin{bmatrix}
			\bold{h}_0^c \\
			\bold{h}_1^c \\
			\vdots \\
			\bold{h}_N^c
		    \end{bmatrix}.		  
\end{align}

We note that
\begin{align}
\gradpYlm = \unitvec \times \gradYlm, \quad \forall \; l \in \No, \; m \in \Z \; \text{s.t.} \; |m| \le l.
\end{align}

\sotodo{Isn't there a name for this Operator? Maybe "Unit Curl Operator"?}
\subsubsection{Unit Curl Operator}
\begin{align}
\unitvec \times \bold{u} &= \unitvec \times \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ] \\
&= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \unitvec \times \gradYlm + u^{\perp}_{l,m} \unitvec \times \gradpYlm ] \\
&= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \unitvec \times \gradYlm + u^{\perp}_{l,m} \unitvec \times ( \unitvec \times \gradYlm ) ] \\
&= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \gradpYlm + u^{\perp}_{l,m} (( \unitvec \cdot \gradpYlm) \unitvec - ( \unitvec \cdot \unitvec ) \gradYlm ) ] \\
&= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} [ u_{l,m} \gradpYlm - u^{\perp}_{l,m} \gradYlm ) ]. \\
\end{align}

Thus the operator matrix for the cross product from the left by the normal unit vector at the point \((x,y,z)\) is given by
\begin{align}
K =  \begin{bmatrix}
		0 & -1 & & & & \\
		1 & 0 & & & & \\
		& & 0 & -1 & & \\
		& & 1 & 0 & & \\
		& & & & \ddots \\
		& & & & & \ddots 
	  \end{bmatrix} \quad \in \quad \R^{2(N+1)^2 \times 2(N+1)^2}.
\end{align}
so that \(K\bold{u}^c\) gives the vector of coefficients for for expansion of \(\unitvec \times \bold{u}\) in the vector spherical harmonic basis (up to degree N).

\subsubsection{Operator for Div}
\begin{align}
\nabla \cdot \bold{u} &= \nabla \cdot \Big( \sum_l \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ] \Big) \\
&= \sum_l \sum_{m=-l}^{l} [ u_{l,m} \: \Delta \Ylm + u^{\perp}_{l,m} \nabla \cdot (\unitvec \times \gradYlm) ] \\
&= \sum_l \sum_{m=-l}^{l} u_{l,m} \: \Delta \Ylm \\
&= \sum_l \sum_{m=-l}^{l} - u_{l,m} \: l(l+1) \Ylm, \\
\end{align}
using the fact that 
\begin{align}
\Delta \Ylm(x,y,z) = -l(l+1)\Ylm(x,y,z), \quad \forall \; l \in \No, \; m \in \Z \; \text{s.t.} \; |m| \le l.
\end{align}

Thus, the operator matrix for the divergence of a vector in the tangent space is
\begin{align}
D &=  \begin{bmatrix}
		-l(l+1)|_{l=0} & 0 & & & & & & & \\
		& & -l(l+1)|_{l=1} & 0 & & & & & \\
		& & & & -l(l+1)|_{l=1} & 0 & & & \\
		& & & & & & -l(l+1)|_{l=1} & 0 & \\
		& & & & & & & \ddots 
	  \end{bmatrix} \\
	  &=  \begin{bmatrix}
		0 & 0 & & & & & & & \\
		& & -2 & 0 & & & & & \\
		& & & & -2 & 0 & & & \\
		& & & & & & -2 & 0 & \\
		& & & & & & & \ddots 
	  \end{bmatrix} \quad \in \quad \R^{(N+1)^2 \times 2(N+1)^2},
\end{align}
so that \(D\bold{u}^c\) gives the vector of coefficients of the expansion for \(\nabla \cdot u\) in the spherical harmonic basis.

\subsubsection{Operator for Grad}
\begin{align}
\nabla h &= \nabla \Big( \sum_l \sum_{m=-l}^{l} h_{l,m} \Ylm \Big) \\
&= \sum_l \sum_{m=-l}^{l} h_{l,m} \gradYlm
\end{align}

Thus, the operator matrix for the gradient of a scalar function on the sphere is
\begin{align}
G &=  \begin{bmatrix}
		\tilde{G_0} & & & \\
		& \tilde{G_1} & & \\
		& & \tilde{G_2} & \\
		& & & \ddots
	  \end{bmatrix}  \quad \in \quad \R^{2(N+1)^2 \times (N+1)^2},
\end{align}
where
\begin{align}
\tilde{G_l} &=  \begin{bmatrix}
		1 & & & \\
		0 & & & \\
		& 1 & & \\
		& 0 & & \\
		& & 1 & \\
		& & 0 & \\
		& & & \ddots \\
	  \end{bmatrix} \quad \in \quad \R^{2(2l+1) \times (2l+1)},
\end{align}
so that \(G\bold{h}^c\) gives the vector of coefficients of the expansion for \(\nabla h\) in the tangent space (vector spherical harmonic) basis.

\subsubsection{Dot products}

We consider the general case that \(\bold{u}\) is expanded up to order \(N_1 \in \N\) and \(\bold{v}\) is expanded up to order \(N_2 \in \N\), i.e. 
\begin{align}
\bold{u}^c &= \begin{bmatrix}
			\bold{u}_0^c \\
			\bold{u}_1^c \\
			\vdots \\
			\bold{u}_{N_1}^c
		    \end{bmatrix},
\quad
\bold{v}^c = \begin{bmatrix}
			\bold{v}_0^c \\
			\bold{v}_1^c \\
			\vdots \\
			\bold{v}_{N_2}^c
		    \end{bmatrix}.
\end{align}

Let \((x,y,z)\) lie on the unit sphere and let \(l = 1\), \(m \in {-1, 0, 1}\) and define \(\tilde{N} := N_1 + N_2\). Then
\begin{align}
\gradYlm(x,y,z) \cdot \bold{v}(x,y,z) = (\bold{b}^c)^T  \: \bigP(x,y,z)
\end{align}
where
\begin{align}
\bigP &:= \begin{bmatrix}
		\bigP_0 \\
		\bigP_1 \\
		\bigP_2 \\
		\vdots \\
		\bigP_{N_2}
	  \end{bmatrix}, \quad 
\Pl := \begin{bmatrix}
		Y_l^{-l} \\
		\vdots \\	
		Y_l^l
	  \end{bmatrix},
\end{align}
for some coefficient vector \(\bold{b}^c \in \R^{(N_2+1)^2}\). In other words, the dot product of one of the \(l=1\) tangent space's orthogonal polynomials with a vector valued function on the unit sphere can naturally be written as an expansion in the scalar spherical harmonic OP basis. Similarly,
\begin{align}
\gradpYlm(x,y,z) \cdot \bold{v}(x,y,z) = (\bold{b^\perp}^c)^T  \: \bigP(x,y,z)
\end{align}
for some coefficient vector \(\bold{b^\perp}^c \in \R^{(N_2+2)^2}\).

Then, 
\begin{align}
\bold{b}^c &= J_{l,m} \: \bold{v}^c, \quad 
\bold{b^\perp}^c = J^\perp_{l,m} \: \bold{v}^c
\end{align}
where \(J_{l,m}\), \(J^\perp_{l,m}\) are (operator) matrices for the dot product with \(\gradYlm(x,y,z)\) and \(\gradpYlm(x,y,z)\) respectively.
Each \(J_{l,m}\), \(J^\perp_{l,m}\) will have the \((\tilde{N} \times N_2)\) block structure:
\begin{align}
	\begin{array}{c|c|c|c}
		1 \times 6 & 1 \times 10 & \\
		\hline
		3 \times 6 & 3 \times 10 & 3 \times 14 & \\
		\hline
		5 \times 6 & 5 \times 10 & 5 \times 14 & \ddots \\
		\hline
		&  7 \times 10 & 7 \times 14 & \ddots \\
		\hline
		& & \ddots & \ddots \\
	  \end{array}
\end{align}
i.e. a {\tt BandedBlockBandedMatrix(J, (rows, cols), (2,1), (?,?))} [note: I am not sure what the sub-block bands should be] where 
\begin{verbatim}
    rows = 1:2:2N+1 
    cols = 6:4:2(2N_2+1).
\end{verbatim}

Define
\begin{align}
\mathbb{T}^\bigP_0 &:= 0, \quad
\mathbb{T}^\bigP_1 := \begin{bmatrix}
			J_{1,-1} \\
			J^\perp_{1,-1} \\
			J_{1,0} \\
			J^\perp_{1,0} \\
			J_{1,1} \\
			J^\perp_{1,1}
	  	\end{bmatrix}.
\end{align}
Then by linearity of the dot product, we can use the recurrence relation (122) for \(\gradPl\) to gain a recurrence for \(\mathbb{T}^\bigP_l\):
\begin{align}
\mathbb{T}^\bigP_{l+1} = -\Dlt [B_l-G_l(J^x, J^y, J^z)] \: \mathbb{T}^\bigP_l - \Dlt C_l  \, \mathbb{T}^\bigP_{l-1}, \quad l \in \{1,\dots,N_1-1\},
\end{align}
where \(J^x, J^y, J^z\) are the \(\tilde{N} \times \tilde{N}\) Jacobi operator matrices for multiplication of the scalar spherical harmonic basis by \(x, y, z\) respectively. We then have that for each \(l \in {1,\dots,N_1}\),
\begin{align}
\mathbb{T}^\bigP_l := \begin{bmatrix}
			J_{l,-l} \\
			J^\perp_{l,-l} \\
			\vdots \\
			J_{l,l} \\
			J^\perp_{l,l}
	  	\end{bmatrix}.
\end{align}

Then, 
\begin{align}
\bold{u}(x,y,z) \cdot \bold{v}(x,y,z) = (\bold{b}^c)^T  \: \bigP(x,y,z)
\end{align}
where \(\bold{b}^c \in \R^{(\tilde{N}+1)^2}\) and is given by
\begin{align}
\bold{b}^c = \Big(\sum_{l=0}^{N_1} \sum_{m=-l}^{l} [ u_{l,m} J_{l,m} + u^{\perp}_{l,m} J^\perp_{l,m} ] \Big) \: \bold{v}^c.
\end{align}

\subsection{Example}

Let $\mathbf{u}(x,y,z)$ be the tangential velocity of the flow and $h(x,y,z)$ be the height of the water from some constant reference height $\mathcal{H}$. The linear SWEs are
\begin{align}
\begin{cases}
\frac{\partial \mathbf{u}}{\partial t} + f \unitvec \times \mathbf{u} - \nabla h = \mathbf{0} \\
\frac{\partial h}{\partial t} + \mathcal{H} \nabla \cdot \mathbf{u} = 0
\end{cases}
\end{align}
where $f = 2 \Omega \cos(\theta) =  2 \Omega z$ is the Coriolis parameter (note that the operator on the coefficients vector for multiplication by this parameter would be $F = 2 \Omega (J^z_\nabla)^T$).

Using backward Euler:
\begin{align}
\mathbf{u}_{n+1} &= \mathbf{u}_{n} + \Delta t \: (\nabla h_{n+1} - f \unitvec \times \mathbf{u}_{n+1}) \\
h_{n+1} &= h_n - \Delta t \: \mathcal{H} \nabla \cdot \mathbf{u}_{n+1}
\end{align}

Then using the operators on the coefficient vectors, the timestepping method is: 
\begin{align}
\mathbf{u}^c_{n+1} &= \mathbf{u}^c_{n} + \Delta t \: (G \mathbf{h}^c_{n+1} - F K \mathbf{u}^c_{n+1}) \\
\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
\end{align}
$\iff$
\begin{align}
\mathbf{u}^c_{n+1} &= (I + \Delta t F K + \Delta t^2 \mathcal{H} G D)^{-1} \: (\mathbf{u}^c_{n} + \Delta t \: G \mathbf{h}^c_n) \\
\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
\end{align}

\bstodo{Image}

\section{Appendix}

\subsection{Derivation of multiplication by $x,y,z$ relationships for the spherical harmonics}

Using (7)--(9), we can write \sotodo{Add references to the DLMF. Be precise about which equation as it makes it easier for yourself to look up.}
\begin{align}
\Ylm(x,y,z) &= \alphalm (1 - z^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(z), \quad \text{where } 0 \le |m| \le l, \, l \in \No.
\end{align}

Note that the recurrence relationship for the Jacobi polynomials satisfies \cite{DLMFJacobiRecurrence}
\begin{align}
z P^{(m,m)}_k(z) = \frac{1}{\kappa_{k,m}} \big[ P^{(m,m)}_{k+1}(z) - \lambda_{k,m} P^{(m,m)}_k(z) + \mu_{k,m} P^{(m,m)}_{k-1}(z) \big],
\end{align}
for \(k\ge0, \, m\in\Z\), where
\begin{align}
\kappa_{k,m} &:= \frac{(2k+2m+1)(k+m+1)}{(k+1)(k+2m+1)}, \\
\lambda_{k,m} &:= 0\\
\mu_{k,m} &:= \frac{k+m+1}{(k+1)(k+2m+1)}.
\end{align}
Thus,
\begin{align}
z P^{(m,m)}_{l-m}(z) = \Ftilde P^{(m,m)}_{l-m+1}(z) + \Gtilde P^{(m,m)}_{l-m-1}(z),
\end{align}
for \(k\ge0, \, m\in\Z\), where
\begin{align}
\Ftilde &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}, \\
\Gtilde &:= \begin{cases}
			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
			\quad 0 \quad \quad \text{otherwise.}
		\end{cases} .
\end{align}

Further, note that
\begin{align}
\cosphi \, \eimphi &= \frac{1}{2} (e^{i\varphi} + e^{-i\varphi}) \eimphi =  \frac{1}{2} (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \\
\sinphi \, \eimphi &= \frac{1}{2i} (e^{i\varphi} - e^{-i\varphi}) \eimphi =  \frac{-i}{2} (e^{i(m+1)\varphi} - e^{i(m-1)\varphi})
\end{align}

Finally, note the relationship between the Jacobi polynomials \(P^{(a,b)}_{l}(z)\) and the ultraspherical polynomials \(C^{(\lambda)}_{n}(z)\) \cite{DLMFUltraJacobi}, as well as the three-term recurrence relation for the ultraspherical polynomials \cite{DLMFUltraRecurrence, DLMFUltraRecurrence2}, i.e. for \(m \in \No\):
\begin{align}
\nu_{l,m} C^{(m+1/2)}_{l-m}(z) = P^{(m,m)}_{l-m}(z),
\end{align}
and
\begin{align}
C^{(\lambda)}_n(z) &= \xi^{(1)}_{n, \lambda} \big[ C^{(\lambda+1)}_n(z) - C^{(\lambda+1)}_{n-2}(z) \big]  \\
(1-z^2) \, C^{(\lambda)}_n(z) &= \frac{1}{\xi^{(2)}_{n, \lambda}} \big[ \xi^{(3)}_{n, \lambda} C^{(\lambda-1)}_n(z) - \xi^{(4)}_{n, \lambda}C^{(\lambda-1)}_{n+2}(z) \big] ,
\end{align}
where
\begin{align}
\nu_{l,m} &:= \frac{\Gamma(l+1)\Gamma(2m+1)}{\Gamma(l+m+1)\Gamma(m+1)} \\
\xione&:= \frac{\lambda}{n+\lambda} \\
\xitwo &:= 4(\lambda-1)(n+\lambda) \\
\xithree &:= (n+2\lambda-2)(n+2\lambda-1) \\
\xifour &:= (n+1)(n+2).
\end{align}

Thus, we can write three-term recurrences for the Jacobi polynomials as:
\begin{align}
P^{(m,m)}_{l-m}(z) &= \Atilde P^{(m+1,m+1)}_{l-m}(z) + \Btilde P^{(m+1,m+1)}_{l-m-2}(z) \\
(1-z^2) \, P^{(m,m)}_{l-m}(z) &= \Dtilde P^{(m-1,m-1)}_{l-m+2}(z) + \Etilde P^{(m-1,m-1)}_{l-m}(z),
\end{align}
for \(l, m \in \No\) s.t. \(0 \le m \le l\), where
\begin{align}
\Atilde &:= \frac{(l+m+2)(l+m+1)}{2(2l+1)(l+1)} \\
\Btilde &:= \begin{cases}
			- \frac{l}{2(2l+1)} \quad \text{if } l - m \ge 2 \\
			\quad \quad 0 \quad \quad \text{otherwise.}
		\end{cases} \\
\Dtilde &:= - \frac{2(l-m+2)(l-m+1)}{(2l+1)(l+1)} \\
\Etilde &:= \frac{2l}{2l+1} .
\end{align}

We can now write down expressions for the multiplication of \(\Ylm(x,y,z)\) by \(x, y \) or \(z\) for some point \((x,y,z)\) on the unit sphere for \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) as follows:
\begin{align}
x\,\Ylm(x,y,z) &= \alphalm \cosphi \eimphi \sintheta (1-z^2)^{|m|/2} P^{(|m|,|m|)}_{l-|m|}(z) \nonumber \\
		     &= \half \alphalm (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) (1-z^2)^{\frac{|m|+1}{2}} P^{(|m|,|m|)}_{l-|m|}(z) \nonumber \\
		     &= \half \alphalm e^{i(m+1)\varphi} (1-z^2)^{\frac{|m|+1}{2}} \big[ \Atilde P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \Btilde P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \nonumber \\
		     &\quad \quad +  \half \alphalm e^{i(m-1)\varphi} (1-z^2)^{\frac{|m|-1}{2}} \big[ \Dtilde P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \Etilde P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \nonumber \\
		     &= \Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z),
\end{align}
\begin{align}
y\,\Ylm(x,y,z) &= \alphalm \sinphi \eimphi \sintheta (1-z^2)^{|m|/2} P^{(|m|,|m|)}_{l-|m|}(z) \nonumber \\
		     &= -\half i \alphalm (e^{i(m+1)\varphi} - e^{i(m-1)\varphi}) (1-z^2)^{\frac{|m|+1}{2}} P^{(|m|,|m|)}_{l-|m|}(z) \nonumber \\
		     &= - i \, \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \nonumber \\
		     &\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big],
\end{align}
\begin{align}
z\,\Ylm(x,y,z) &= \alphalm \eimphi (1-z^2)^{|m|/2} z P^{(|m|,|m|)}_{l-|m|}(z) \nonumber \\
		     &= \alphalm \eimphi (1-z^2)^{|m|/2} \big[ \Ftilde P^{(|m|,|m|)}_{l-|m|+1}(z) + \Gtilde P^{(|m|,|m|)}_{l-|m|-1}(z) \big]\nonumber \\
		     &= \Flm Y^{m}_{l+1}(x,y,z) + \Glm Y^{m}_{l-1}(x,y,z) ,
\end{align}
where
\begin{align}
\Alm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \Atilde \quad \text{if } m\ge0 \\
			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \tilde{D}_{l,|m|} \quad \text{if } m<0 
	       \end{cases} \\
\Blm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \Btilde \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \tilde{E}_{l,|m|} \quad \text{if } m<0 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise.} 
	        \end{cases} \\
\Dlm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \Dtilde \quad \text{if } m>0 \\
			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \tilde{A}_{l,|m|} \quad \text{if } m\le0 
	       \end{cases} \\
\Elm &:= \begin{cases}
			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \Etilde \quad \text{if } m>0 \\
			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \tilde{B}_{l,|m|} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise.} 
	        \end{cases} \\ 
\Flm &:= \frac{\alphalm}{\alpha^{m}_{l+1}} \Ftilde \\
\Glm &:= \frac{\alphalm}{\alpha^{m}_{l-1}} \Gtilde.
\end{align}

\subsection{Vector spherical harmonics}

We start by defining what we mean by a spin-1 tensor spherical harmonic. In general, the tensor spherical harmonic is given by, for \(2l, \, j, \, 2s \in \No\),
\begin{align}
\curlyy^{j,s}_{l,m}(x,y,z) = \sum_{m_s=-s}^s \langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle \, Y^{m-m_s}_j (x,y,z) \, \bm{\chi}_{s, m_s},
\end{align}
where \(\bm{\chi}_{s, m_s}\) are the simultaneous eigenstates of the spin operators \(\bold{S}^2\) and \(S_z\), and where \(\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle\) \sotodo{This notation is hard to read.} is a Clebsch-Gordan coefficient. We note that a property of the Clebsch--Gordan coefficients means that they vanish unless \(|j - s| \le l \le j + s\). We further note that there are simple calculable expressions for the Clebsch--Gordan coefficients when the spin \(s=1\).

To simplify the notation, for spin $s=1$, we define
\begin{align}
\cgcoeff^{L, m_s}_{l, m} := \, \langle L \quad m-m_s \; ; \; 1 \quad m_s \; | \; l \quad m \rangle.
\end{align}

Thus we have that the three spin-1 tensor spherical harmonics are given by
\begin{align}
\curlyy^{l \pm1,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
\curlyy^{l,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
\end{align}

Here, the vectors \(\chivec\) are the orthonormal eigenvectors of the matrix  \sotodo{Fix}
\begin{align}
S_3 = \begin{bmatrix} 0 & -i & 0 \\ i & 0 & 0 \\ 0 & 0 & 0 \end{bmatrix},
\end{align}
and so are given as
\begin{align}
\bm{\chi}_{1, \pm 1} = \frac{1}{\sqrt{2}} \begin{bmatrix} \mp 1 \\ -i \\ 0 \end{bmatrix}, \quad \bm{\chi}_{1,0} = \begin{bmatrix} 0 \\ 0 \\ 1 \end{bmatrix}.
\end{align}

Then, for any \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) we have that
\begin{align}
\nabla \Ylm &= \blmone \, \curlyy^{l-1,1}_{l,m} + \blone \curlyy^{l+1,1}_{l,m}, \\
\nabla^\perp \Ylm &= \blzero \, \curlyy^{l,1}_{l,m},
\end{align}
where
\begin{align}
\blmone &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
\blzero := i \, \big(l(l+1)\big)^\half, \quad
\blone := l \, \Big(\frac{l+1}{2l+1}\Big)^\half.
\end{align}

\bibliography{spherical-harmonics-bib}

\end{document}

### 2018-11-27 papers/esa/esa.tex
Add comments to ESA

@@
 
+\sotodoinline{Outline the "big picture" here. Say that the ultimate goal is to solve PDEs on spherical triangles, and what is listed below are steps on the way to that goal.}
+
 This document is to provide a brief description of my research area which will be the topic of my thesis.
@@
 They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.
+\sotodo{Mention something like: this is equivalent to standard techniques in the literature on spherical harmonics, but the goal here is to put it in a language that will translate to spherical triangles}
 
@@
 Clenshaw's algorithm then gives us a way of evaluating a function given its expansion in the spherical harmonic basis (up to a finite order).
+\sotodo{Mention that "this provides an optimal complexity method for evaluating spherical harmonic expansions.
+This is an alternative to using the explicit expressions that may have benefits, though further investigation is required."}
 
@@
 \end{align}
-where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have derivable and calculable formulae.
+where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have derivable and calculable formulae.\sotodo{Mention that the formulae are explicit}
 
@@
 \subsection{Evaluation of a scalar function on the sphere}
-We can use the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e.
+We can use the Clenshaw algorithm \sotodo{I would say "a multivariate analogue of Clenshaw's algorithm}  to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e.
 suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
@@
 
-By noting that the tangent space is then spanned by $\{\gradYlm, \gradpYlm\}$, and that these gradients and perpendicular gradients are orthogonal polynomials themselves, we can use the same ideas to create a framework for functions in the tangent space as expansions in this OP basis.
+By noting that the tangent space is then spanned by\sotodo{Connect these to "vector spherical harmonics"} $\{\gradYlm, \gradpYlm\}$, and that these gradients and perpendicular gradients are orthogonal polynomials themselves, we can use the same ideas to create a framework for functions in the tangent space as expansions in this OP basis.
 
@@
 
-Then using the operators on the coefficient vectors, the timestepping method is: 
+Then using the operators on the coefficient vectors, the timestepping method is: \sotodo{Explain what $F$, $K$ , and $G$ are and emphasise they can all be constructed}
 \begin{align}
@@
 As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
-We will use the lanczos algorithm from ApproxFun \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:
+We will use the lanczos algorithm from\sotodo{"as implemented", not "from"} ApproxFun \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:\sotodo{Cite Huybrechs reference for connection in literature}
 \begin{align}
@@
 
-By using a similar process to \cite{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
+By using a similar process to\sotodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is} \cite{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
 We denote the half disk by \(\Omega\) where
@@
 
-Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by
+Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by\sotodo{The $+$ should be $\approx$}
 \begin{align}
@@
 	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \\
-f_{n,k} &:= \frac{< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
+f_{n,k} &:= \frac{\langle f, \: \Pnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
 \end{align}
@@
 
-We can obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space
+We can obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space \sotodo{Mention this vanishes on the boundary so imposes zero Dirichlet}
 \[
@@
 \]
-to coefficients in the non-weighted space
+to coefficients in the non-weighted space\sotodo{The notation $\Pii$ is not defined I think}
 \[
@@
 This operator will have Banded-Block-Banded structure.
-We first note that, for general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
+We first note that, for general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that: \sotodo{Mention this is analoguous to Jacobi polynomials, DLMF ...}
 \begin{align}
@@
 \end{align}
-Then, the operator \(\Delta\) is given by
+Then, the operator \(\Delta\) is given by \sotodo{Mention it is also Banded-Block-Banded, hence sparse}
 \begin{align}
@@
 The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.
+\sotodo{This last paragraph could be in the intro}
+
+\sotodoinline{Maybe mention that isosceles triangles should follow by the same procedure as the half sphere, but general spherical triangles will require further ideas.}
+
+\sotodoinline{Expand bibliography: cite spherical harmonics books, Huybrechs, ECMWF paper, etc.}
 

### 2018-11-27 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -169,7 +169,7 @@ end
     h = 1e-5; @test (f(x,y+h)-f(x,y))/h ≈ differentiatey(f, f.space)(x,y) atol=100h
 end
 
-@test "Poisson" begin
+@testset "Poisson" begin
     # Model Problem: Δ(u*w)(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
     #   where w(x,y) = x*(1-x^2-y^2) is the weight of the D(1.0,1.0) basis.
     a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
@@ -178,7 +178,7 @@ end
     # 1) f(x,y) = -8x => u(x,y) ≡ 1
     N = 1 # degree of f
     c = rand(1)[1]; f = Fun((x,y)->-c*8x, S)
-    Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+    Δ = laplace(D, N-1)
     u = Fun(S, Δ \ resizecoeffs!(f, N))
     @test u(z) ≈ c # Result u(x,y) where Δ(u*w)(x,y) = f(x,y)
 

### 2018-12-03 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Sped up laplace operator construction

@@ -346,25 +346,27 @@ end
 # NOTE these are squared norms
 function getopnorms(S::HalfDiskSpace, m)
     m̃ = length(S.opnorms)
-    n = m̃ == 0 ? -1 : -1 + Int(round(sqrt(1 + 2(m̃ - 1))))
-    k = m̃ - Int((n + 1)n / 2)
-    resize!(S.opnorms, m)
-    P = S.P(S.b, S.b)
-    p = Fun(P, [1])
-    ptsp, wp = pointswithweights(P, 1)
-    normP = inner(P, p, p, ptsp, wp)
-    for j = m̃+1:m
-        if k > n
-            n += 1
-            k = 0
+    if m > m̃
+        n = m̃ == 0 ? -1 : -1 + Int(round(sqrt(1 + 2(m̃ - 1))))
+        k = m̃ - Int((n + 1)n / 2)
+        resize!(S.opnorms, m)
+        P = S.P(S.b, S.b)
+        p = Fun(P, [1])
+        ptsp, wp = pointswithweights(P, 1)
+        normP = inner(P, p, p, ptsp, wp)
+        for j = m̃+1:m
+            if k > n
+                n += 1
+                k = 0
+            end
+            # Pnk = Fun(S, [zeros(j-1); 1])
+            # S.opnorms[j] = inner(S, Pnk, Pnk, pts, w)
+            H = S.H(S.a, S.b + k + 0.5)
+            h = Fun(H, [zeros(n-k); 1])
+            ptsh, wh = pointswithweights(H, Int(ceil(n-k+0.5)))
+            S.opnorms[j] = inner(H, h, h, ptsh, wh) * normP
+            k += 1
         end
-        # Pnk = Fun(S, [zeros(j-1); 1])
-        # S.opnorms[j] = inner(S, Pnk, Pnk, pts, w)
-        H = S.H(S.a, S.b + k + 0.5)
-        h = Fun(H, [zeros(n-k); 1])
-        ptsh, wh = pointswithweights(H, Int(ceil(n-k+0.5)))
-        S.opnorms[j] = inner(H, h, h, ptsh, wh) * normP
-        k += 1
     end
     S
 end
@@ -697,51 +699,69 @@ function evalderivativey(S::HalfDiskSpace, n, k, x, y)
     H(x) * S.ρ(x)^(k-1) * differentiateop(S.P(S.b, S.b), k)(y/S.ρ(x))
 end
 function getpartialoperatorx(S::HalfDiskSpace, N)
[…]

### 2018-12-03 papers/esa/esa.tex
Edits to ESA

--- 2018-11-23 papers/esa/esa.tex
@@
 
-This document is to provide a brief description of my research area which will be the topic of my thesis.
+The goal of my research is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model.
+The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.
+
+This document is to provide a brief outline of my work to date, with a closer look at our most recent area of research (solving PDEs on the half disk using sparse spectral methods).
+The work outlined in this report serves as initial steps to pave the way to solving PDEs on spherical triangles.
 All code used thus far and from hence forth will mainly be written in the Julia programming language.
@@
 They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.
+While this is equivalent to standard techniques in the litrature concerning spherical harmonics, we approach this with our goal in mind of developing sparse spectral methods on spherical triangles.
+Thus, we try to use language and formulations that will translate as we move through our steps towards that goal.
 
-\citep{dunkl2014orthogonal} provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$.
+\citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is} provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$.
 Clenshaw's algorithm then gives us a way of evaluating a function given its expansion in the spherical harmonic basis (up to a finite order).
+This provides an optimal complexity method for evaluating spherical harmonic expansions.
+This is an alternative to using the explicit expressions that may have benefits, though further investigation is required.
+\bstodoinline{Reword above}
 
@@
 \end{align}
-where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have derivable and calculable formulae.
+where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have explicit derivable and calculable formulae.
 
@@
 such that \(\Dlt A_l = I_{2l+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is}, which can be expanded to obtain the recurrence:
 \begin{align}
@@
 \subsection{Evaluation of a scalar function on the sphere}
-We can use the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e.
+We can use a multivariate analogue of the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e.
 suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
@@
 By noting that the tangent space is then spanned by $\{\gradYlm, \gradpYlm\}$, and that these gradients and perpendicular gradients are orthogonal polynomials themselves, we can use the same ideas to create a framework for functions in the tangent space as expansions in this OP basis.
-
-Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient  and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics are a basis for the tangent space, namely \(\gradYlm\), \(\gradpYlm\).
 Note that the perpendicular gradient is related to the regular surface gradient by
@@
 \end{align}
+For any scalar spherical harmonic (SSH) $\Ylm$, the triad of $\gradYlm, \: \unitvec \times \gradYlm, \: \unitvec \Ylm$ can be referred to as vector spherical harmonics \cite{BarreraEtAl}.
+
+Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient  and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics are a basis for the tangent space, namely \(\gradYlm\), \(\gradpYlm\).
 
@@
 The gradient and perpendicular gradient of a spherical harmonic \(\Ylm(x,y,z)\) can be expressed in terms of spin-1 tensor spherical harmonics, which in turn can each be expressed as a vector-weighted sum of spherical harmonics.
-For the derivation and more on vector spherical harmonics, see the appendix.
 
@@
 \end{align}
+where $F, K, G, D$ are all operators corresponding to multiplying by the Coriolis parameter, outward normal vector cross product, gradient and divergence operations respectively.
+It is important to note that these can be easily constructed due to the relationships between the vector spherical harmonics and scalar spherical harmonics we have outlined, and have \bstodo{Banded-Block-Banded?} a sparse Block-Banded structure.
 
@@
 As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
-We will use the lanczos algorithm from ApproxFun \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:
+We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:\sotodo{Cite Huybrechs reference for connection in literature}
 \begin{align}
@@
 
-By using a similar process to \cite{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
+By using a similar process to \cite{dunkl2014orthogonal} (p??) \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is} we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
 We denote the half disk by \(\Omega\) where
@@
 \begin{align}
-f(x,y) + \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
+f(x,y) \approx \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
 \end{align}
@@
 such that \(\Dnt A_n = I_{n+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is}, which can be expanded to obtain the recurrence:
 \begin{align}
@@
 \[
-\Wii(x,y) \: \Pii(x,y)
+\Wii(x,y) \: \bigP^{(1,1)}(x,y)
 \]
@@
 \[
-\Pii(x,y).
+\bigP^{(1,1)}(x,y).
 \]
-This operator will have Banded-Block-Banded structure.
-We first note that, for general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
+Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse.
+Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
+
+For general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
 \begin{align}
@@
 \dx[\Wab(x,y) \: f(x,y)] &= W^{(a-1,b-1)}(x,y) \: \bigP^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
-\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f},
+\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
 \end{align}
+The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials, as seen in the DLMF \cite{DLMFDerivatives}.
 Further, we note that there exist Banded-Block-Banded matrix operators \(T^{(1,0)\to(0,0)}\) and \(T^{(0,1)\to(1,1)}\) such that:
@@
 
-The goal here is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model.
-The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.
+Beyond this, solving PDEs on isosceles triangles should follow using the same procedure.
+However, we are expecting that to solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas beyond the procedures given in this document.
+
+\sotodoinline{Expand bibliography: cite spherical harmonics books, Huybrechs, ECMWF paper, etc.}
 

### 2018-12-03 papers/halfdisk/halfdisk.tex
Halfdisk paper first draft

--- 2018-12-03 papers/esa/esa.tex
@@
 
-\title{Early Stage Assessment}
-\author{Ben Snowball (Principle Supervisor: Sheehan Olver)}
-\date{}							% Activate to display a given date or no date
+\title{Sparse Spectral Methods and P-Finite Element Methods on Disk-Splices and Rhombi - Part I: Zero Boundary Conditions}
+\author{Sheehan Olver, Ben Snowball}
+%\date{}							% Activate to display a given date or no date
 
@@
 
+%
 \section{Introduction}
 
-The goal of my research is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model.
-The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.
+This document is to derive a framework for solving PDEs on the half disk using sparse spectral methods.
+We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk.
+We will derive a quadrature rule on the half disk that can be used to expand a function in the OP basis up to a given order.
+We will demonstrate that these families will have sparse operators (including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators), and how we can implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
+We can extend this work directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc.
+can be found similarly.
 
-This document is to provide a brief outline of my work to date, with a closer look at our most recent area of research (solving PDEs on the half disk using sparse spectral methods).
-The work outlined in this report serves as initial steps to pave the way to solving PDEs on spherical triangles.
-All code used thus far and from hence forth will mainly be written in the Julia programming language.
-
-\section{Professional Skills Development Requirement}
-I have met the initial part of the professional skills development requirement which is to attend a minimum of 2 professional development workshops plus the online plagiarism awareness course by the Early Stage Assessment (ESA).
-This is by virtue of the fact that I attended the Communication Skills Training in the MPE CDT on 03/11/2017.
-This course has been given accreditation with the Graduate School programme and is worth 2 credits towards the requirement.
-
-\section{Orthogonal polynomials on the sphere (spherical harmonics)}
-
-The first step was to develop a small code library so that we can use spherical harmonics as orthogonal polynomials in three variables $x, y, z$  to evaluate functions on the sphere.
-
-We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{DLMFSphericalandSpheroidalHarmonics, DLMFFerrers, DLMFAssociatedLegendre}:
-\begin{align}
-\Ylmfull &:= \clm \eimphi \Plm (\costheta) \\
-&= \alphalm (1 - (\costheta)^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(\costheta), \quad \text{where } 0 \le |m| \le l, \, l \in \No.
-\end{align} 
-Note that we can then express \(\Ylm\) in terms of \(x,y,z\) instead of \(\theta, \varphi\) by noting that \(\costheta = z\) and that \( \eimphi\) can be expressed in terms of \(x,y,z\) for any \(m\in\mathbb{Z}\).
-Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$.
-They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.
-While this is equivalent to standard techniques in the litrature concerning spherical harmonics, we approach this with our goal in mind of developing sparse spectral methods on spherical triangles.
-Thus, we try to use language and formulations that will translate as we move through our steps towards that goal.
-
-\citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is} provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$.
-Clenshaw's algorithm then gives us a way of evaluating a function given its expansion in the spherical harmonic basis (up to a finite order).
-This provides an optimal complexity method for evaluating spherical harmonic expansions.
-This is an alternative to using the explicit expressions that may have benefits, though further investigation is required.
-\bstodoinline{Reword above}
-
-\subsection{Deriving expressions for the multiplication by \(x, y, z\) of the spherical harmonics}
-
-We start by expressing \(x\,\Ylm(x,y,z)\), \(y\,\Ylm(x,y,z)\) , and \(z\,\Ylm(x,y,z)\)  in terms of \(Y^{m'}_{l'}(x,y,z)\) for any point \((x,y,z)\) on the unit circle.
-
-For \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\):
-\begin{align}
-x\,\Ylm(x,y,z) &= \Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \nonumber \\
-		     & \quad \quad \quad + \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z), \\
-y\,\Ylm(x,y,z) &= - i \, \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \nonumber \\
-		     &\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
-z\,\Ylm(x,y,z) &= \Flm Y^{m}_{l+1}(x,y,z) + \Glm Y^{m}_{l-1}(x,y,z) ,
-\end{align}
-where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have explicit derivable and calculable formulae.
-
-\subsection{Jacobi matrices}
-
-Define, for \(l \in \No\), \(\Pl\) as the column vector of the degree \(l\) spherical harmonic polynomials, and \(\bigP\) as the stacked block vector of the \(\Pl\)'s; that is
-\begin{align}
-\Pl := \begin{bmatrix}
-		Y^{-l}_l \\
-		\vdots \\
-		Y^l_l
-	\end{bmatrix} \in \C^{2l+1}, 
-\quad \quad 
-\bigP := \begin{bmatrix}
-		\bigP_0 \\
-		\hline
-		\bigP_1 \\
-		\hline
-		\bigP_2 \\
-		\vdots \\
-	\end{bmatrix}.
-\end{align}
-
-Define the (Jacobi) matrices \(J^x, J^y, J^z\) by 
-\begin{align}
-J^x \bigP = x \bigP, \quad J^y \bigP = y \bigP, \quad J^z \bigP = z \bigP.
-\end{align}
-Then, using equations (3--5), we have that the Jacobi matrices each have the following block-tridiagonal form, allowing for easy application computationally:
-\begin{align}
-J^{x/y/z} &= \begin{bmatrix}
-		B^{x/y/z}_0 & A^{x/y/z}_0 & & & & \\
-		C^{x/y/z}_1 & B^{x/y/z}_1 & A^{x/y/z}_1 & & & \\
-		& C^{x/y/z}_2 & B^{x/y/z}_2 & A^{x/y/z}_2  & & & \\
-		& & C^{x/y/z}_3 & \ddots & \ddots & \\
-		& & & \ddots & \ddots & \ddots \\
-	\end{bmatrix}
-\end{align}
-
-\subsection{Three-term recurrence relation for \(\bigP\)}
-
-Combining each system in (7) we can write the tridiagonal-block system
-\begin{align}
-\renewcommand\arraystretch{1.3}
-\mleft[
-\begin{array}{cccc}
-		1  & & & \\
-		B_0-G_0(x,y,z) & A_0 & & \\
-		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
-		& C_2 & \ddots & \ddots \\
-		& & \ddots &
-\end{array}
-\mright]
-\bigP
-=
-\begin{bmatrix}
-	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
-\end{bmatrix}.
-\end{align}
-where for each \(l \in \No\):
-\begin{align}
-A_l &:= \begin{bmatrix}
-		A^x_l \\
-		A^y_l \\
-		A^z_l
-	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+3)}, \quad
-C_l := \begin{bmatrix}
-		C^x_l \\
-		C^y_l \\
-		C^z_l
-	    \end{bmatrix} \in \R^{3(2l+1)\times(2l-1)} \quad (l \ne 0), \\
-B_l &:= \begin{bmatrix}
-		B^x_l \\
-		B^y_l \\
-		B^z_l
-	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}, \quad
-G_l(x,y,z) := \begin{bmatrix}
-		xI_{2l+1} \\
-		yI_{2l+1} \\
-		zI_{2l+1}
-	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}.
-\end{align}
-
-For each \(l \in \No\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e.
-such that \(\Dlt A_l = I_{2l+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is}, which can be expanded to obtain the recurrence:
-\begin{align}
-\begin{cases}
-\bigP_{-1} := 0 \\
-\bigP_{0} := \alpha^0_0 \\
-\bigP_{l+1} = -\Dlt (B_l-G_l) \Pl - \Dlt C_l  \,\bigP_{l-1}, \quad l \in \N.
-\end{cases}
-\end{align}
-
-\subsection{Evaluation of a scalar function on the sphere}
-We can use a multivariate analogue of the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e.
-suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
-\begin{align}
-f(x,y,z) \approx \sum^N_{l=0} \bold{f}_l^T \, \Pl (x,y,z), \quad \text{where } \Pl (x,y,z), \bold{f}_l \in \R^{2l+1} \text{ for each } l \in \{0,\dots,N\}.
-\end{align}
-
-The Clenshaw algorithm is then as follows:
-\begin{align*}
-\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
-\quad &\text{2) } \text{For } n = N:-1:1 \\
-\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
-\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
-\end{align*}
-where $\bigP_0 \equiv \alpha^0_0$.
-
-We can use the Clenshaw algorithm to output a matrix operator for a function, given we know the coefficients of its expansion in the SH basis. We simply make the adjustment that
-\begin{align*}
-\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
-\quad &\text{2) } \text{For } n = N:-1:1 \\
-\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = (\bold{f}_n^T)I - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(J^x,J^y,J^z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
-\quad &\text{3) } \text{Output: } f(J^x,J^y,J^z) \approx F := \alpha^0_0 f_0 + \bm{\gamma}_{1}^T \bigP_1(J^x,J^y,J^z) - \alpha^0_0 \bm{\gamma}_{2}^T D^T_1 C_1.
-\end{align*}
-where $\bigP_1(J^x,J^y,J^z)$ is a vector of matrices, $G_n(J^x,J^y,J^z)$ is a matrix of matrices, and so each $\bm{\gamma}_n$ is a vector of matrices.
-
-\subsection{Tangent space}
-
-By noting that the tangent space is then spanned by $\{\gradYlm, \gradpYlm\}$, and that these gradients and perpendicular gradients are orthogonal polynomials themselves, we can use the same ideas to create a framework for functions in the tangent space as expansions in this OP basis.
-Note that the perpendicular gradient is related to the regular surface gradient by
-\begin{align}
-\gradpYlm(x,y,z) = \unitvec \times \gradYlm(x,y,z),
-\end{align}
-where \(\unitvec\) is the unit vector normal to the surface of the sphere at the point \((x,y,z)\), i.e. as we are looking at the unit sphere, \(\unitvec\) is simply given by
-\begin{align}
-\unitvec = \begin{bmatrix} x \\ y \\ z \end{bmatrix}.
-\end{align}
-For any scalar spherical harmonic (SSH) $\Ylm$, the triad of $\gradYlm, \: \unitvec \times \gradYlm, \: \unitvec \Ylm$ can be referred to as vector spherical harmonics \cite{BarreraEtAl}.
-
-Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient  and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics are a basis for the tangent space, namely \(\gradYlm\), \(\gradpYlm\).
-
-Importantly,  \(\gradYlm\), \(\gradpYlm\) are in fact vector-valued orthogonal polynomials, that span all vector-valued polynomials modulo the vanishing ideal $\{ \bm{p} \in (\Pi_3)^3  : \unitvec \cdot \bm{p} = 0 \}$.
-A simple calculation shows that such orthogonal polynomials must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal.
-In the following, we determine the coefficients of these Jacobi operators, as well as other important operators needed.
-
-\subsubsection{Spin-1 tensor spherical harmonics}
-
-The gradient and perpendicular gradient of a spherical harmonic \(\Ylm(x,y,z)\) can be expressed in terms of spin-1 tensor spherical harmonics, which in turn can each be expressed as a vector-weighted sum of spherical harmonics.
-
-For any \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) we have that
-\begin{align}
-\nabla \Ylm &= \blmone \, \curlyy^{l-1,1}_{l,m} + \blone \curlyy^{l+1,1}_{l,m}, \\
-\nabla^\perp \Ylm &= \blzero \, \curlyy^{l,1}_{l,m},
-\end{align}
-where
-\begin{align}
-\blmone &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
-\blzero := i \, \big(l(l+1)\big)^\half, \quad
-\blone := l \, \Big(\frac{l+1}{2l+1}\Big)^\half.
-\end{align}
-
-We can then gain similar (vector analogue) derivations for recurrences, Jacobi matrices and the Clenshaw algorithm implementation for the tangent space vector-valued orthogonal polynomial basis.
-
-The aim is to, given a PDE say, work with the coefficients of the functions involved (coefficients of their expansion in the OP bases), reframe the PDE in terms of the OP coefficients utilising a suitable timestepping method.
-For derivative, gradient, divergence etc.
-operators, we can formulate matrix operators to act on the coefficients, thus making our system to simulate a simple matrix vector system.
-
-\subsection{Example: linear shallow water equations}
-Let $\mathbf{u}(x,y,z)$ be the tangential velocity of the flow and $h(x,y,z)$ be the height of the water from some constant reference height $\mathcal{H}$. The linear SWEs are
-\begin{align}
-\begin{cases}
-\frac{\partial \mathbf{u}}{\partial t} + f \unitvec \times \mathbf{u} - \nabla h = \mathbf{0} \\
-\frac{\partial h}{\partial t} + \mathcal{H} \nabla \cdot \mathbf{u} = 0
-\end{cases}
-\end{align}
-where $f = 2 \Omega \cos(\theta) =  2 \Omega z$ is the Coriolis parameter (note that the operator on the coefficients vector for multiplication by this parameter would be $F = 2 \Omega (J^z_\nabla)^T$).
-
-Using backward Euler:
-\begin{align}
-\mathbf{u}_{n+1} &= \mathbf{u}_{n} + \Delta t \: (\nabla h_{n+1} - f \unitvec \times \mathbf{u}_{n+1}) \\
-h_{n+1} &= h_n - \Delta t \: \mathcal{H} \nabla \cdot \mathbf{u}_{n+1}
-\end{align}
-
-Then using the operators on the coefficient vectors, the timestepping method is: 
-\begin{align}
-\mathbf{u}^c_{n+1} &= \mathbf{u}^c_{n} + \Delta t \: (G \mathbf{h}^c_{n+1} - F K \mathbf{u}^c_{n+1}) \\
-\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
-\end{align}
-$\iff$
-\begin{align}
-\mathbf{u}^c_{n+1} &= (I + \Delta t F K + \Delta t^2 \mathcal{H} G D)^{-1} \: (\mathbf{u}^c_{n} + \Delta t \: G \mathbf{h}^c_n) \\
-\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
-\end{align}
-where $F, K, G, D$ are all operators corresponding to multiplying by the Coriolis parameter, outward normal vector cross product, gradient and divergence operations respectively.
-It is important to note that these can be easily constructed due to the relationships between the vector spherical harmonics and scalar spherical harmonics we have outlined, and have \bstodo{Banded-Block-Banded?} a sparse Block-Banded structure.
-
-These ideas will also be used to gain similar relationships for the circle arc, half disk, and spherical triangle orthogonal polynomials.
-
-\section{Orthogonal polynomials on the circle arc}
-
-As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
-We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:\sotodo{Cite Huybrechs reference for connection in literature}
-\begin{align}
-\bigP_0(x,y) := T^h_0(x) \equiv T_0,
-\quad \quad 
-\bigP_n(x,y) := \begin{bmatrix}
-		T^h_n(x) \\
-		y \: U^h_{n-1}(x)
-	\end{bmatrix} \in \R^2 \quad \forall n \in \N,
-\quad \quad 
-\bigP := \begin{bmatrix}
-		\bigP_0 \\
-		\hline
-		\bigP_1 \\
-		\hline
-		\bigP_2 \\
-		\vdots \\
-	\end{bmatrix}.
-\end{align}
-
-\section{Orthogonal polynomials on the half disk (2D)}
+\section{Orthogonal polynomials  and function approximation on the half disk (2D)}
 
@@
 
-\subsection{Example: Laplace operator}
+%
+\section{Sparse partial differential operators}
 
@@
 
-An example is as follows. Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+%
+\section{P-finite element methods using sparse operators}
+Consider the 2D Dirichlet problem on a disk-splice or rhobus $\Omega$:
+\begin{align}
+	\begin{cases}
+                \Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
+                u(x,y) = 0 \quad \text{on } \partial \Omega
+         \end{cases}
+\end{align}
+This has the weak formulation for any test function $v$
+\begin{align}
+	L(v) := \int \int_\Omega f \: v \: dy \: dx = - \int \int_\Omega \nabla u \cdot \nabla v \: dy \: dx =: a(u,v).
+\end{align}
+
+We can choose our finite dimensional space $V$ to
+
+%
+\section{Examples on the half disk with zero BCs}
+
+\subsection{Poisson}
+
+Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
 \begin{align}
@@
 
+\subsection{Helmholtz}
+
+\subsection{Advection-Diffusion}
+
+%
 \section{Future work - orthogonal polynomials on the hemisphere and spherical triangles}
@@
 
-\bibliography{esa-bib}
+\bibliography{halfdisk-bib}
 

### 2018-12-05 papers/esa/esa.tex
Further edits

@@
 
-The goal of my research is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model.
+The goal of my research is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model \cite{wedi2013fast}.
 The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.
@@
 
-We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{DLMFSphericalandSpheroidalHarmonics, DLMFFerrers, DLMFAssociatedLegendre}:
+We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{DLMFSphericalandSpheroidalHarmonics, DLMFFerrers, DLMFAssociatedLegendre, efthimiou2014spherical, dai2013approximation}:
 \begin{align}
@@
 
-\citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is} provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$.
+\citep{dunkl2014orthogonal} (p75-85) provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$.
 Clenshaw's algorithm then gives us a way of evaluating a function given its expansion in the spherical harmonic basis (up to a finite order).
-This provides an optimal complexity method for evaluating spherical harmonic expansions.
-This is an alternative to using the explicit expressions that may have benefits, though further investigation is required.
-\bstodoinline{Reword above}
+This provides an optimal complexity method for evaluating spherical harmonic expansions; an alternative to using the explicit expressions that may have benefits, though further investigation is required.
 
@@
 such that \(\Dlt A_l = I_{2l+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} (p78), which can be expanded to obtain the recurrence:
 \begin{align}
@@
 where $F, K, G, D$ are all operators corresponding to multiplying by the Coriolis parameter, outward normal vector cross product, gradient and divergence operations respectively.
-It is important to note that these can be easily constructed due to the relationships between the vector spherical harmonics and scalar spherical harmonics we have outlined, and have \bstodo{Banded-Block-Banded?} a sparse Block-Banded structure.
+It is important to note that these can be easily constructed due to the relationships between the vector spherical harmonics and scalar spherical harmonics we have outlined, and have a sparse Block-Banded structure.
 
@@
 As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
-We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to:\sotodo{Cite Huybrechs reference for connection in literature}
+We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to \cite{huybrechs2010fourier}:
 \begin{align}
@@
 
-By using a similar process to \cite{dunkl2014orthogonal} (p??) \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is} we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
+By using a similar process to \cite{dunkl2014orthogonal} (p55-56) we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
 We denote the half disk by \(\Omega\) where
@@
 such that \(\Dnt A_n = I_{n+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} \bstodo{Try to always include page numbers in citations to books, as otherwise you'll forget where it is}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} (p78), which can be expanded to obtain the recurrence:
 \begin{align}
@@
 
-\sotodoinline{Expand bibliography: cite spherical harmonics books, Huybrechs, ECMWF paper, etc.}
-
 \bibliography{esa-bib}

### 2018-12-05 papers/halfdisk/halfdisk.tex
Further edits

@@
 
-\sotodoinline{Expand bibliography: cite spherical harmonics books, Huybrechs, ECMWF paper, etc.}
-
 \bibliography{halfdisk-bib}

### 2018-12-06 papers/esa/esa.tex
ESA 2nd draft done. Further edits to paper.

@@
 
-Let \(f(x,y)\) be a degree \(d\) polynomial that is even in \(y\).
-Then there exists a polynomial in \(x,y\) \(\tilde{f}(x,y)\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\).
-Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree \(\tilde{d}\) \bstodo{What is the degree of g?} \ in \(s\) and \(t\).
-Thus our quadrature rule will be exact:
+Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$.
+Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e.
+that $d_2$ is an even integer).
+Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$.
+Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
 \begin{align}
@@
 &= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
-&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
-&= \sum_j \sum_k w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
-&= \sum_j \sum_k w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
-\end{align} \bstodo{What is the degree of g?}
-where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\).
+&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
+&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
+&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
+\end{align}
+where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
 
@@
 \end{align}
-exact for any polynomial where \(deg(f(x,\cdot)) \le 2N-1\) and \(deg(f(\cdot,y)) \le 2N-1\). Thus,
+and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$. Thus,
 \begin{align}
 < f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
-|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
+|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
 \end{align}
+for suitably large $N$.
 

### 2018-12-06 papers/halfdisk/halfdisk.tex
ESA 2nd draft done. Further edits to paper.

@@
 
-\subsection{Helmholtz}
+\subsection{Inhomogenous Helmholtz}
 
-\subsection{Advection-Diffusion}
+Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+\begin{align}
+    \Delta(\Wii u)(x,y) + k^2 u(x,y) = - f(x,y) \quad \text{in } \Omega
+\end{align}
+noting that this ensures a boundary condition that \((\Wii \: u)(x,y) = 0\) on \(\partial \Omega\).
+
+Again, the first step is to obtain the expansion of \(f\) in the \(\Pii\) basis. The coefficients of \(u\) in the \(\Pii\) basis are then given by 
+\begin{align}
+    \mathbf{u} = - (k^2I + \Delta) \text{\textbackslash} \mathbf{f}.
+\end{align}
+
+\subsection{Heat Equation}
+
+Fix \(a, b = 1,1\), and consider
+\begin{align}
+\frac{\partial u}{\partial t}(x,y) = \Delta(\Wii \: u)(x,y) \quad \text{in } \Omega
+\end{align}
+
+Using a backward Euler time-stepping method, we can formulate this for \(t = 0,1,2,\dots\) and step length $h$:
+\begin{align}
+u_{t+1}(x,y) - u_t (x,y) = h \Delta(\Wii \: u_{t+1})(x,y) \quad \text{for } (x,y) \in \Omega.
+\end{align}
+
+There exist $N \in \N$ such that, for each $t = 0,1,2,\dots$, $u_t (x,y) = \sum_{n=0}^N \mathbf{u_t}_n^T \: \bigP_n^{(1,1)}(x,y) = \mathbf{u}_t^T \; \bigP^{(1,1)}(x,y)$.
+Then, using the operator $\Delta$, we can solve this system for a given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
+\begin{align}
+    \mathbf{u_{t+1}} &= (I - h \Delta) \: \text{\textbackslash} \: \mathbf{u_t}, \\
+    \big( u_{t+1}(x,y) &= \mathbf{u}_{t+1}^T \; \bigP^{(1,1)}(x,y) \big) \nonumber.
+\end{align}
 

### 2018-12-11 papers/halfdisk/halfdisk.tex
Further edits to paper

@@
 \newcommand{\dy}{\frac{\partial}{\partial y}}
+\newcommand{\element}{\tau}
+\newcommand{\refelement}{\hat{\tau}}
+\newcommand{\FEset}{\mathcal{T}}
 
@@
 \begin{align}
+\begin{cases}
 \alpha &:= 0 \\
@@
 \gamma &:= -1 \\
-\delta &:= -c = 1 \\
+\delta &:= -\gamma = 1 \\
 \rho(x) &:= (1-x^2)^{\half} \\
@@
 w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b,
+\end{cases}
 \end{align}
@@
 This provides us with a family of orthogonal polynomial spaces on the half disk.
+Note that we can obtain OP families on other disk-splices and rhombi by appropriately choosing the functions and constants in (2).
+\bstodo{Should I add in the rhombi / other disk splices formulations here?}
 
@@
 We can develop a quadrature rule on \(\Omega\) in order to evaluate (40) and obtain the coefficients vector \(\mathbf{f}\) as follows.
+\bstodo{Should I add in the rhombi / other disk splices formulations here?}
 
@@
 
+Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs.
+In other words, sparsity is not dependent on the domain.
+
 \subsection{Clenshaw on the half disk}
@@
 \begin{align}
-    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)}
+    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
 \end{align}
+The sparsity of the differential operators can be shown using a simply integration by parts argument, again showing that the sparsity for differential operators is not dependent on the domain.
 
@@
 \section{P-finite element methods using sparse operators}
-Consider the 2D Dirichlet problem on a disk-splice or rhobus $\Omega$:
+\bstodoinline{Need to discuss this}
+We follow the method of \cite{beuchler2006new}. Consider the 2D Dirichlet problem on a domain $\Omega$:
 \begin{align}
 	\begin{cases}
-                \Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
-                u(x,y) = 0 \quad \text{on } \partial \Omega
+         - \Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
+         u = 0 \quad \text{on } \partial \Omega
          \end{cases}
 \end{align}
-This has the weak formulation for any test function $v$
+This has the weak formulation for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$,
 \begin{align}
-	L(v) := \int \int_\Omega f \: v \: dy \: dx = - \int \int_\Omega \nabla u \cdot \nabla v \: dy \: dx =: a(u,v).
+	L(v) := \int_\Omega f \: v \: d\mathbf{x} = \int_\Omega \nabla u \cdot \nabla v \: d\mathbf{x} =: a(u,v).
 \end{align}
 
-We can choose our finite dimensional space $V$ to
+Let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a rhombus for example.
+We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h)|_\element \le p \: \forall \element \in \FEset\}$.
+
+We seek $u_h \in V_h$ s.t.
+\begin{align}
+	L(v_h) = a(u_h,v_h) \quad \forall \: v_h \in V_h.
+\end{align}
+
+We can choose a basis for $V_h$ by using the orthogonal polynomials 
+\begin{align}
+\bigP(x,y) := \begin{bmatrix}
+		\bigP_0(x,y) \\
+		\hline
+		\bigP_1(x,y) \\
+		\hline
+		\bigP_2(x,y) \\
+		\hline
+		\vdots \\
+		\hline
+		\bigP_p(x,y)
+	\end{bmatrix},
+\quad
+\bigP_n(x,y) := \begin{bmatrix}
+		P_{n,0}(x,y) \\
+		\vdots \\
+		P_{n,n}(x,y)
+	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
+\end{align}
+Appropriately relabelling these OPs as the set $\{\phi_1,\dots,\phi_N\}$, we can define the global stiffness matrix $A$ by $A_{ij} =  a(\phi_i, \phi_j)$.
+The global stiffness matrix can be computed using the element stiffness matrices $A^\element = \int_\element \nabla \bigP \cdot \nabla \bigP \: d\mathbf{x}, \: \element \in \FEset$.
+Then we have the linear system for our finite element problem:
+\begin{align}
+A\mathbf{u} = \mathbf{f}
+\end{align}
+where the vectors $\mathbf{u}, \mathbf{f}$ are the coefficient vectors for expansion of the solution $u_h$ and for the function $f$ respectively.
+
+Using the quadrature rule we have formulated for integrals on a rhombus element $\element$, and the sparse operator matrices for divergence (partial derivatives), we obtain a sparse global stiffness matrix.
 
@@
 \subsection{Heat Equation}
-
+\bstodo{Struggling to get plotting working}
 Fix \(a, b = 1,1\), and consider

### 2018-12-11 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Examples for paper. Confused about plotting

@@ -880,29 +880,34 @@ function getweightedpartialoperatory(S::HalfDiskSpace, N)
     end
     W
 end
-function gettransformoperator(S, N)
+function gettransformoperator(S::HalfDiskSpace, N)
     if Int(S.a) == 1 && Int(S.b) == 0
+        St = (S.family)(0.0, 0.0)
         T = BandedBlockBandedMatrix(Zeros{Float64}(sum(1:(N+2)),sum(1:(N+1))), (1:N+2, 1:N+1), (1,0), (0,0))
-        for n = 0:N, k = 0:n
-            j = Int(n*(n+1)/2 + k)
-            p = (x,y) -> (x * Fun(S, [zeros(j); 1])(x,y))
-            m = Int((n+3)*(n+4))
-            cfs = pad(Fun(p, (S.family)(0.0, 0.0), m).coefficients, sum(1:(n+2)))
-            cfs = PseudoBlockArray(cfs, 1:(n+2))
-            view(T, Block(n+1, n+1))[k+1, k+1] = cfs[Block(n+1)][k+1]
-            view(T, Block(n+2, n+1))[k+1, k+1] = cfs[Block(n+2)][k+1]
+        getopnorms(St, sum(1:N+2))
+        pts, w = pointswithweights(St, Int(ceil(N+1.5)^2))
+        for n=0:N, k=0:n
+            j = getindex!(n, k)
+            p = (x,y) -> (x * Fun(S, [zeros(j-1); 1])(x,y))
+            for m = n:n+1
+                i = getindex!(m, k)
+                view(T, Block(m+1, n+1))[k+1, k+1] = inner(St, p, Fun(St, [zeros(i-1); 1]), pts, w) / St.opnorms[i]
+            end
         end
         T
     elseif Int(S.a) == 0 && Int(S.b) == 1
+        St = (S.family)(1.0, 1.0)
+        getopnorms(St, sum(1:N+2))
+        pts, w = pointswithweights(St, Int(ceil(N+1.5)^2))
         T = BandedBlockBandedMatrix(Zeros{Float64}(sum(1:(N+1)),sum(1:(N+1))), (1:N+1, 1:N+1), (0,1), (0,0))
         for n = 0:N
             for k = 0:n-1
-                j = Int(n*(n+1)/2 + k)
-                m = Int((n+2)*(n+3))
-                cfs = pad(Fun(Fun(S, [zeros(j); 1]), (S.family)(1.0, 1.0), m).coefficients, sum(1:(n+1)))
[…]

### 2018-12-11 experiments/halfdisk.jl

@@ -1,23 +1,7 @@
 using OrthogonalPolynomialFamilies, ApproxFun
 using StaticArrays, LinearAlgebra, Test, Profile, BlockArrays,
-        BlockBandedMatrices, SparseArrays, Plots
-
-# # Model Problem: Δu(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
-# function getpartialoperatorx2(S, N)
-#     (OrthogonalPolynomialFamilies.getpartialoperatorx(OrthogonalPolynomialFamilies.differentiatespacex(S), N-1)
-#         * OrthogonalPolynomialFamilies.getpartialoperatorx(S, N))
-# end
-# function getpartialoperatory2(S, N)
-#     (OrthogonalPolynomialFamilies.getpartialoperatory(OrthogonalPolynomialFamilies.differentiatespacey(S), N-1)
-#         * OrthogonalPolynomialFamilies.getpartialoperatory(S, N))
-# end
-# differentiatespacex2(S) =
-#     OrthogonalPolynomialFamilies.differentiatespacex(
-#             OrthogonalPolynomialFamilies.differentiatespacex(S))
-# differentiatespacey2(S) =
-#     OrthogonalPolynomialFamilies.differentiatespacey(
-#             OrthogonalPolynomialFamilies.differentiatespacey(S))
-
+        BlockBandedMatrices, SparseArrays
+using Makie
 
 # Model Problem: Δ(u*w)(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
 #   where w(x,y) = x*(1-x^2-y^2) is the weight of the D(1.0,1.0) basis.
@@ -51,3 +35,79 @@ f = Fun((x,y)->y*exp(x)*(2-11x-6x^2-x^3-2y^2-x*y^2), S, m)
 Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
 u = Fun(S, Δ \ f.coefficients[1:size(Δ)[1]])
 @test u(z) ≈ U(z) atol=10.0^(-N)
+
+
+
+#====#
+#=
+Model problem: Helmholtz eqn:
+    Δ(W*u)(x,y) + k²u(x,y) = -f(x,y)
[…]

### 2018-12-13 papers/halfdisk/halfdisk.tex
Further edits to FE stuff mainly

@@
 \newcommand{\Pii}{P^{(1,1)}}
+\newcommand{\bigPii}{{\mathbb{P}^{(1,1)}}}
 \newcommand{\Poo}{P^{(0,0)}}
+\newcommand{\bigPoo}{{\mathbb{P}^{(0,0)}}}
 \newcommand{\Poi}{P^{(0,1)}}
@@
 \newcommand{\FEset}{\mathcal{T}}
+\newcommand{\bigW}{\mathbb{W}}
+\newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
 
@@
 Note that we can obtain OP families on other disk-splices and rhombi by appropriately choosing the functions and constants in (2).
-\bstodo{Should I add in the rhombi / other disk splices formulations here?}
 
@@
 
-We can develop a quadrature rule on \(\Omega\) in order to evaluate (40) and obtain the coefficients vector \(\mathbf{f}\) as follows.
-\bstodo{Should I add in the rhombi / other disk splices formulations here?}
+We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficients vector \(\mathbf{f}\) as follows.
 
@@
 
-We can obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space
-\[
-\Wii(x,y) \: \bigP^{(1,1)}(x,y)
-\]
-to coefficients in the non-weighted space
-\[
-\bigP^{(1,1)}(x,y).
-\]
-Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse.
-Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
-
 For general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
@@
 \end{align}
-Then, the operator \(\Delta\) is given by
+
+We can then obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space
+\[
+\Wii(x,y) \: \bigPii(x,y)
+\]
+to coefficients in the non-weighted space
+\[
+\bigPii(x,y).
+\]
+Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse.
+Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
+The matrix operator \(\Delta\) acting on the coefficients vector is then given by
 \begin{align}
@@
 \end{align}
-The sparsity of the differential operators can be shown using a simply integration by parts argument, again showing that the sparsity for differential operators is not dependent on the domain.
+The sparsity of the these operators can be shown using a simply integration by parts argument, again showing that the sparsity for differential operators is not dependent on the domain.
 
@@
 \section{P-finite element methods using sparse operators}
-\bstodoinline{Need to discuss this}
+
 We follow the method of \cite{beuchler2006new}. Consider the 2D Dirichlet problem on a domain $\Omega$:
@@
 
-Let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a rhombus for example.
-We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h)|_\element \le p \: \forall \element \in \FEset\}$.
+In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a rhombus or disk slice for example.
+For the purposes of this paper, we limit this to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/rhombus domain.
+We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h|_\element) \le p\}$ for some $p \in \N$.
 
@@
 
-We can choose a basis for $V_h$ by using the orthogonal polynomials 
+Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^T \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to.
+Note that due to orthogonality this is a diagonal matrix.
+We can choose a basis for $V_h$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
 \begin{align}
-\bigP(x,y) := \begin{bmatrix}
-		\bigP_0(x,y) \\
+\bigWii(x,y) &:= \begin{bmatrix}
+		\bigWii_0(x,y) \\
 		\hline
-		\bigP_1(x,y) \\
+		\bigWii_1(x,y) \\
 		\hline
-		\bigP_2(x,y) \\
+		\bigWii_2(x,y) \\
 		\hline
@@
 		\hline
-		\bigP_p(x,y)
-	\end{bmatrix},
-\quad
-\bigP_n(x,y) := \begin{bmatrix}
-		P_{n,0}(x,y) \\
+		\bigWii_p(x,y)
+	\end{bmatrix}, \\
+\bigWii_n(x,y) &:= \begin{bmatrix}
+		\Wii(x,y) \: \Pii_{n,0}(x,y) \\
 		\vdots \\
-		P_{n,n}(x,y)
+		\Wii(a,y) \: \Pii_{n,n}(x,y)
 	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
 \end{align}
-Appropriately relabelling these OPs as the set $\{\phi_1,\dots,\phi_N\}$, we can define the global stiffness matrix $A$ by $A_{ij} =  a(\phi_i, \phi_j)$.
-The global stiffness matrix can be computed using the element stiffness matrices $A^\element = \int_\element \nabla \bigP \cdot \nabla \bigP \: d\mathbf{x}, \: \element \in \FEset$.
-Then we have the linear system for our finite element problem:
+and rewrite (63) in matrix form:
 \begin{align}
-A\mathbf{u} = \mathbf{f}
+	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
+	&= \int_\element \begin{bmatrix}
+					\partial_x v_h \\
+					\partial_y v_h
+				\end{bmatrix}^T 
+				\begin{bmatrix}
+					\partial_x u_h \\
+					\partial_y u_h
+				\end{bmatrix}
+				\\
+	&= \int_\element \begin{bmatrix}
+					\bigPoo^T \Wii_x \mathbf{v} \\
+					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
+				\end{bmatrix}^T 
+				\begin{bmatrix}
+					\bigPoo^T \Wii_x \mathbf{u} \\
+					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
+				\end{bmatrix}
+				\\
+	&= \int_\element \Big( \mathbf{v}^T {\Wii_x}^T \bigPoo \bigPoo^T \Wii_x \mathbf{u} \nonumber \\
+					& \quad \quad \quad + \mathbf{v}^T ({T^{(1,0)\to(0,0)} \Wii_y})^T \bigPoo \bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
+	&= \mathbf{v}^T \: \Big( {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
 \end{align}
-where the vectors $\mathbf{u}, \mathbf{f}$ are the coefficient vectors for expansion of the solution $u_h$ and for the function $f$ respectively.
+where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_h, v_h \in V_h$ respectively in the $V_h$ basis ($\bigWii$ OPs), and
+\begin{align}
+	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
+	&= \int_\element \: \mathbf{v}^T \: \bigWii \: \bigPii^T \: \mathbf{f} \: d\mathbf{x} \\
+	&= \mathbf{v}^T \Lambda^{(1,1)} \: \mathbf{f},
+\end{align}
+where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bigPii$ OP basis.
+We can also not that naturally $\langle \bigPii, {\bigPii}^T \rangle_{\Wii}$ is just a diagonal matrix.
 
-Using the quadrature rule we have formulated for integrals on a rhombus element $\element$, and the sparse operator matrices for divergence (partial derivatives), we obtain a sparse global stiffness matrix.
+Since (63) is equivalent to stating that
+\begin{align}
+	L(\Wii \Pii_{n,k}) = a(u_h,\Wii \Pii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
+\end{align}
+(i.e. holds for all basis functions of $V_h$) by choosing $v_h$ as each basis function, we can equivalently write the linear system for our finite element problem as:
+\begin{align}
+A\mathbf{u} = \tilde{\mathbf{f}}.
+\end{align}
+where the (element) stiffness matrix $A$ is defined by 
+\begin{align}
+A = {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
+\end{align}
+and the load vector $\tilde{\mathbf{f}}$ is given by 
+\begin{align}
+\tilde{\mathbf{f}} = \Lambda^{(1,1)} \: \mathbf{f}.
+\end{align}
+
+Note the since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
 

### 2018-12-13 experiments/halfdisk.jl
Convergence pictures for some examples. Trying to get an operator version of clenshaw in 2d working. Start at general OPFamily2D types and code.

@@ -36,6 +36,29 @@ f = Fun((x,y)->y*exp(x)*(2-11x-6x^2-x^3-2y^2-x*y^2), S, m)
 u = Fun(S, Δ \ f.coefficients[1:size(Δ)[1]])
 @test u(z) ≈ U(z) atol=10.0^(-N)
 
+# 4) f(x,y) = 2zcos(x) - 8xsin(x) - xzsin(x) - 4x^2cos(x) - 12xy => u(x,y) = sin(x) + y
+# (where z = 1-x^2-y^2)
+using Plots
+Uc = Fun((x,y)->sin(x), S).coefficients
+m = 200
+f = Fun((x,y)->(2*(1-x^2-y^2)*cos(x) - 8*x*sin(x) - x*(1-x^2-y^2)*sin(x) - 4*x^2*cos(x) - 12*x*y),
+        S, m)
+N = 2
+û = zeros(1)
+Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+u = Δ \ f.coefficients[1:size(Δ)[1]]
+res = abs.(u - append!(û, zeros(size(u)[1]-size(û)[1])))
+plt = plot(res, yaxis=:log)
+û = copy(u)
+for N = 3:10 # degree of estimation
+    Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
+    u = Δ \ f.coefficients[1:size(Δ)[1]]
+    res = abs.(u - append!(û, zeros(size(u)[1]-size(û)[1])))
+    plot!(plt, res)
+    global û = copy(u)
+end
+plt
+savefig("example4coeffs")
 
 
 #====#
@@ -79,35 +102,158 @@ Model problem: Heat eqn
 a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
 x, y = 0.4, -0.2; z = [x; y] # Test point
 h = 1e-3
-N = 3
-uc = OrthogonalPolynomialFamilies.resizecoeffs!(Fun((x,y)->x, S), N)
+N = 5
 Δ = OrthogonalPolynomialFamilies.laplacesquare(D, N-1)
-sparse(I - h * Δ)
-for it = 1:10
-    global uc = sparse(I - h * Δ) \ uc
+û = OrthogonalPolynomialFamilies.resizecoeffs!(Fun((x,y)->x, S), N)
+plt = plot()
+maxits = 1000
+nplots = 10
+for it = 1:1000
[…]

### 2018-12-13 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Parameter conversion operator added

@@ -912,6 +912,31 @@ function gettransformoperator(S::HalfDiskSpace, N)
             view(T, Block(n+1, n+1))[n+1, n+1] = 1.0
         end
         T
+    elseif Int(S.a) == 0 && Int(S.b) == 2
+        St = (S.family)(2.0, 2.0)
+        getopnorms(St, sum(1:N+2))
+        pts, w = pointswithweights(St, Int(ceil(N+1.5)^2))
+        T = BandedBlockBandedMatrix(Zeros{Float64}(sum(1:(N+1)),sum(1:(N+1))), (1:N+1, 1:N+1), (0,2), (0,0))
+        n = 0; view(T, Block(n+1, n+1))[n+1, n+1] = 1.0
+        for n = 1:N
+            for k = 0:n-2
+                j = getindex!(n, k)
+                p = Fun(S, [zeros(j-1); 1])
+                for m = n-2:n
+                    i = getindex!(m, k)
+                    view(T, Block(m+1, n+1))[k+1, k+1] = inner(St, p, Fun(St, [zeros(i-1); 1]), pts, w) / St.opnorms[i]
+                end
+            end
+            k = n-1
+            j = getindex!(n, k)
+            p = Fun(S, [zeros(j-1); 1])
+            for m = n-1:n
+                i = getindex!(m, k)
+                view(T, Block(m+1, n+1))[k+1, k+1] = inner(St, p, Fun(St, [zeros(i-1); 1]), pts, w) / St.opnorms[i]
+            end
+            view(T, Block(n+1, n+1))[n+1, n+1] = 1.0
+        end
+        T
     else
         error("Invalid HalfDiskSpace")
     end

### 2018-12-13 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -1,7 +1,8 @@
 using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature, SingularIntegralEquations, Test
 import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule, gethalfdiskOP,
                                         jacobix, jacobiy, evalderivativex, evalderivativey,
-                                        differentiatex, differentiatey, resizecoeffs!, laplace
+                                        differentiatex, differentiatey, resizecoeffs!, laplace,
+                                        gettransformoperator
 
 
 
@@ -189,6 +190,24 @@ end
     Δ = laplace(D, N-1)
     u = Fun(S, Δ \ resizecoeffs!(f, N))
     @test u(z) ≈ U(z)
+end
+
+@testset "Conversion operators" begin
+    a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
+    x, y = 0.457, -0.209; z = [x; y] # Test point
+    N = 5
+    f1 = Fun(D(0.0, 2.0), rand(sum(1:N+1)))
+    T = gettransformoperator(D(0.0, 2.0), N)
+    f2 = Fun(D(2.0, 2.0), T * f1.coefficients)
+    @test f2(z) ≈ f1(z)
+    f1 = Fun(D(0.0, 1.0), rand(sum(1:N+1)))
+    T = gettransformoperator(D(0.0, 1.0), N)
+    f2 = Fun(D(1.0, 1.0), T * f1.coefficients)
+    @test f2(z) ≈ f1(z)
+    f1 = Fun(D(1.0, 0.0), rand(sum(1:N+1)))
+    T = gettransformoperator(D(1.0, 0.0), N)
+    f2 = Fun(D(0.0, 0.0), T * f1.coefficients)
+    @test f2(z) ≈ x*f1(z)
 end
 
 

### 2018-12-14 experiments/halfdisk.jl
Minor changes

@@ -48,17 +48,17 @@ û = zeros(1)
 Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
 u = Δ \ f.coefficients[1:size(Δ)[1]]
 res = abs.(u - append!(û, zeros(size(u)[1]-size(û)[1])))
-plt = plot(res, yaxis=:log)
+plt = plot(res, yaxis=:log, label="N=$N")
 û = copy(u)
 for N = 3:10 # degree of estimation
     Δ = OrthogonalPolynomialFamilies.laplace(D, N-1)
     u = Δ \ f.coefficients[1:size(Δ)[1]]
     res = abs.(u - append!(û, zeros(size(u)[1]-size(û)[1])))
-    plot!(plt, res)
+    plot!(plt, res, label="N=$N")
     global û = copy(u)
 end
 plt
-savefig("example4coeffs")
+savefig("experiments/example4coeffs")
 
 
 #====#
@@ -99,32 +99,36 @@ Model problem: Heat eqn
     u1 - u0 = h * Δ(W*u1)
 =>  (I - hΔ)\u0 = u1
 =#
+using Plots
 a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
 x, y = 0.4, -0.2; z = [x; y] # Test point
-h = 1e-3
+h = 1e-2
 N = 5
 Δ = OrthogonalPolynomialFamilies.laplacesquare(D, N-1)
-û = OrthogonalPolynomialFamilies.resizecoeffs!(Fun((x,y)->x, S), N)
+û = OrthogonalPolynomialFamilies.resizecoeffs!(Fun((x,y)->x*(1-x^2)*y, S), N)
 plt = plot()
-maxits = 1000
-nplots = 10
-for it = 1:1000
+maxits = 100
+nplots = 5
+for it = 1:99
     u = sparse(I - h * Δ) \ û
     if round(nplots*it/maxits) == nplots*it/maxits
         # plot!(plt, abs.(u-û))
     end
     global û = copy(u)
 end
-for it = 1001:maxits+1001
+for it = 100:maxits+100
     u = sparse(I - h * Δ) \ û
     if round(nplots*it/maxits) == nplots*it/maxits
-        plot!(plt, abs.(u-û))
+        plot!(plt, abs.(u), label="t=$it")
     end
     global û = copy(u)
 end
 plt
-u = Fun(S, uc)
-u(z)
[…]

### 2018-12-14 papers/halfdisk/halfdisk.tex
Further edits. Added a couple of plots.

@@
 \usepackage{todonotes}
+
+\graphicspath{ {images/} }
 
@@
 
+\begin{figure} 
+\center
+\begin{tikzpicture} 
+% Box/square (4x4)
+\draw[black,solid,ultra thin] (0,0)--(0,4);
+\draw[black,solid,ultra thin] (0,0)--(4,0);
+\draw[black,solid,ultra thin] (4,0)--(4,4);
+\draw[black,solid,ultra thin] (0,4)--(4,4);
+% Dots
+\draw[black,fill=black] (0,0) circle (.5ex);
+\draw[black,fill=black] (2,2) circle (.5ex);
+\draw[black,fill=black] (4,4) circle (.5ex);
+\draw[black,fill=black] (0,2) circle (.5ex);
+\draw[black,fill=black] (0,4) circle (.5ex);
+% Arrows
+\draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
+\draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
+\draw[black,thick,->] (0,0.2)--(0,1.8);
+\draw[black,thick,->] (0,2.2)--(0,3.8);
+\draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
+% Node (parameter) labels
+\draw[] (0,0) node[anchor=east] {$(0,0)$};
+\draw[] (0,4) node[anchor=east] {$(0,2)$};
+\draw[] (4,0) node[anchor=west] {$(2,0)$};
+\draw[] (4,4) node[anchor=west] {$(2,2)$};
+% Arrow labels
+\draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
+\draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
+\draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
+\draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
+% The key and headings
+\draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
+\draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
+\draw[] (-2.3,3.2) node[anchor=west] {$a$};
+\draw[] (-2.7,3.6) node[anchor=south] {$b$};
+\draw[] (2,4.6) node[anchor=south] {basis conversion};
+\end{tikzpicture} 
+\caption{The Laplace operator acting on vectors of $P_{n,k}=P_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{P^{(2,2)}_{n,k}}$ coefficients.
+Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{P_{n,k}^{(a,b)}}$ coefficients, where $(a,b)$ is at the tail of the arrow, and the range is $\smash{P_{n,k}^{(\tilde{a},\tilde{b})}}$ coefficients, where $(\tilde{a},\tilde{b})$ is at the head of the arrow.
+%For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{P_{n,k}^{(0,0,0)}}$ (resp.~$\smash{P_{n,k}^{(1,0,1)}}$) and the range in $\smash{P_{n,k}^{(1,0,1)}}$ (resp.~$\smash{P_{n,k}^{(2,0,2)}}$).
+}
+\label{fig:Laplace} 
+\end{figure}
+
 For general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
@@
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials, as seen in the DLMF \cite{DLMFDerivatives}.
-Further, we note that there exist Banded-Block-Banded matrix operators \(T^{(1,0)\to(0,0)}\) and \(T^{(0,1)\to(1,1)}\) such that:
+Further, we note that there exist Banded-Block-Banded matrix operators \(T^{(1,0)\to(0,0)}\), \(T^{(0,1)\to(1,1)}\) and $T^{(0,2)\to(2,2)}$ such that:
 \begin{align}
 x \bigP^{(1,0)}(x,y) &= \Big(T^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
-\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y)
+\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y) \\
+\bigP^{(0,2)}(x,y) &= \Big(T^{(0,2)\to(2,2)} \Big)^T \: \bigP^{(2,2)}(x,y)
 \end{align}
@@
 \end{align}
+
+Further, we can also obtain the matrix operator for \(\Delta\) that will take us from coefficients for expansion in the space
+\[
+\bigPoo(x,y)
+\]
+to coefficients in the space
+\[
+\bigP^{(2,2)}(x,y).
+\]
+Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse. The matrix operator \(\Delta\) acting on the coefficients vector is then given by
+\begin{align}
+    D := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(0,2)\to(2,2)} \: D_y^{(0,1)} \: D_y^{(0,0)}.
+\end{align}
+
 The sparsity of the these operators can be shown using a simply integration by parts argument, again showing that the sparsity for differential operators is not dependent on the domain.
@@
 
+\begin{figure}
+	\includegraphics[scale=0.4]{example4coeffs}
+\centering
+\caption{Plot of the difference between the previous solution's coefficients and the current solution's coefficients under increasing degree of approximation (number of coefficients used) for $N = 2,\dots,10$, using RHS of $f(x,y) = 2(1-x^2-y^2)\cos(x) - 8x\sin(x) - x(1-x^2-y^2)\sin(x) - 4x^2 \cos(x) - 12xy$ for the Poisson equation.
+The solution is converging as seen by the decreasing (absolute value) difference of coefficient values.}
+\centering
+\label{fig:poisson}
+\end{figure}
+
 \subsection{Inhomogenous Helmholtz}
@@
 
+\begin{figure}
+	\includegraphics[scale=0.4]{example2timesteppingcovnvergence}
+\centering
+\caption{Plot of the current iteration's coefficients for the heat equation example, using degree of approximation $N = 5$ with initial condition of $u_0(x,y) = \Wii(x,y)$ and step size $h=0.01$, plotted for iterations $t=100, 120, \dots, 200$.
+The solution tends to zero, as seen by the decreasing (absolute value) coefficient values.}
+\centering
+\label{fig:heateqn}
+\end{figure}
+
 %

### 2019-01-02 papers/halfdisk/halfdisk.tex
Work on paper

@@
 
+\usepackage{amsthm}
+
+\newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma} 
+\newtheorem{theorem}{Theorem} 
+\newtheorem{definition}{Definition}
+
+\include somacros
+
 %\usepackage{media9}
@@
 
-\title{Sparse Spectral Methods and P-Finite Element Methods on Disk-Splices and Rhombi - Part I: Zero Boundary Conditions}
+\title{Sparse spectral and $p$-finite element methods on disk-slices and rhombi I: zero boundary conditions}
 \author{Sheehan Olver, Ben Snowball}
@@
 
+\begin{abstract}
+In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
+a half disk) and rhombi.
+This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
+\end{abstract}
+
 %
@@
 
-This document is to derive a framework for solving PDEs on the half disk using sparse spectral methods.
+This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and rhombi.  
+More precisely, let  \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
+
+\begin{enumerate}
+\item  \(\rho\) is a degree 1 polynomial, or 
+\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
+\end{enumerate}
+
+We consider the solution of partial differential equations on the domain
+\begin{align}
+\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+\end{align}
+For simplicity of discussion we focus on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.
+
+We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{SOATUltraspherical} and its analogue on the triangle \cite{SOATGVTriangle}.
+On the half-disk the family of weights we consider are of the form
+$$
+W^{(a,b)}(x,y) = x^a (1-x-y)^b
+$$
+with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$.
+Sparsity comes from expanding the domain and range of an operator  using different polynomials.
+Where the sparsity pattern and entries derived in \cite{SOATGVTriangleRecurrence,SOATGVTriangle}  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
+Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{TriangleFEM}.
+
+The motivation for this work is solving partial differential equations on sub-domains of the sphere.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk.
+Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
+Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids.
+
  We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk.
@@
 
-\section{Orthogonal polynomials  and function approximation on the half disk (2D)}
+\section{Orthogonal polynomials  on the half disk}
 
-By using a similar process to \cite{dunkl2014orthogonal} (p55-56) we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
-We denote the half disk by \(\Omega\) where
-\begin{align}
-\Omega := \{(x,y) \in \R^2 \quad | \quad 0 < x < 1, \: -(1-x^2)^{\half} < y < (1-x^2)^{\half}\}.
-\end{align}
+In this section we outline some basic properties of $P_{n,k}^{(a,b)}(x,y)$.
+The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.
 
-Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
+\subsection{Explicit construction}
 
-1) \(\rho\) is a degree 1 polynomial, or
+By using a similar process to \cite[p55--56]{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
 
-2) \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
+\begin{proposition}
+Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\),
+Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\).
+Then
+$$
+\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
+$$
+are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on $\Omega$. 
+\end{proposition}
 
-Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\).
-
-We define \(\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)})\) for \( 0 \le k \le n, \: n = 0,1,2,\dots\) and note that the functions \(\{\Pnk\}\) are in fact polynomials in either cases above.
-Furthermore, they are orthogonal with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on \(R := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}\).
-
-Let \(a, b \in \R\). For the half disk, if we set:
+For the half disk, the weight $W^{(a,b)}(x,y)$ results from setting:
 
@@
 \begin{cases}
-\alpha &:= 0 \\
-\beta &:= 1 \\
-\gamma &:= -1 \\
-\delta &:= -\gamma = 1 \\
+(\alpha,\beta) &:= (0.1) \\
+(\gamma,\delta) &:= (-1,1) \\
 \rho(x) &:= (1-x^2)^{\half} \\
@@
 
-we can obtain the 1D orthogonal polynomials \(\{H_n^{(a, b)}\}\) orthogonal with respect to the weight \(x^a \: (1-x^2)^b\), and \(\{P_n^{(b, b)}\}\) orthogonal with respect to the weight \((1-x^2)^b\) (so \(\{P_n^{(b, b)}\}\) are in fact Jacobi polynomials), leading to the two parameter 2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\).
+The weight $w_1(x)$ is non-classical and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
+Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
- \Pnk(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+ \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
-orthogonal on \(\Omega\) with respect to the weight 
-\begin{align}
-W^{(a,b)}(x,y) := x^a \: (1-x^2-y^2)^b, \quad (x,y) \in \Omega.
-\end{align}
-
-This provides us with a family of orthogonal polynomial spaces on the half disk.
-Note that we can obtain OP families on other disk-splices and rhombi by appropriately choosing the functions and constants in (2).
-
-\subsection{Obtaining the coefficients for expansion of a function}
-
-Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by
-\begin{align}
-f(x,y) \approx \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
-\end{align}
-where
-\begin{align}
-\bigP(x,y) := \begin{bmatrix}
-		\bigP_0(x,y) \\
-		\hline
-		\bigP_1(x,y) \\
-		\hline
-		\bigP_2(x,y) \\
-		\hline
-		\vdots \\
-		\hline
-		\bigP_N(x,y)
-	\end{bmatrix},
-\quad
-\bigP_n(x,y) := \begin{bmatrix}
-		P_{n,0}(x,y) \\
-		\vdots \\
-		P_{n,n}(x,y)
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
-\end{align}
-and where
-\begin{align}
-\mathbf{f} &:= \begin{bmatrix}
-		\mathbf{f}_0 \\
-		\hline
-		\mathbf{f}_1 \\
-		\hline
-		\mathbf{f}_2\\
-		\hline
-		\vdots \\
-		\hline
-		\mathbf{f}_N
-	\end{bmatrix},
-\quad
-\mathbf{f}_n := \begin{bmatrix}
-		f_{n,0} \\
-		\vdots \\
-		f_{n,n}
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \\
-f_{n,k} &:= \frac{\langle f, \: \Pnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
-\end{align}
-
-We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficients vector \(\mathbf{f}\) as follows.
-
-We will use the substitution that
-\begin{align}
-x &= s, \quad y = (1-s^2)^\half t.
-\end{align}
-First, note that
-\begin{align}
-W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
-		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
-\end{align}
-Using the Golub-Welsch algorithm, we can obtain nodes and weights for 1D quadrature rules on \([0,1]\) with weight \(s^a \: (1-s^2)^{b+\half}\), and on \([-1,1]\) with weight \((1-t^2)^b\).
-
-Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$.
-Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e.
-that $d_2$ is an even integer).
-Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$.
-Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
-\begin{align}
-& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
-&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
-&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
-&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
-&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
-\end{align}
-where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
-
-We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
-\begin{align}
-\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
-\end{align}
-
-Relabelling the weights and nodes as 
-\begin{align}
-x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
-y_{l+(i-1)N} &:= (1-s_l^2)^\half \: t_l, \quad i,l = 1,\dots,N, \\
-w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
-\end{align}
-
-Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
-\begin{align}
-\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
-\end{align}
-and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$. Thus,
-\begin{align}
-< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
-|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
-\end{align}
-for suitably large $N$.
+orthogonal on \(\Omega\) with respect to the weight $W^{(a,b)}(x,y)$.
 
@@
 
-We obtain the 1D OPs via the Lanczos algorithm, which also gives us the coefficients of the relations
+Express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
 \begin{align}
@@
 \end{align}
-where we note that \(\gamma_n^{(b)} = 0\) \(\forall n = 0,1,2,\dots\).
+where we note that \(\gamma_n^{(b)} = 0\) \(\forall n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:
 
-Next, note that \(<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
-\begin{align}
-<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &= \int \int_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
-&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
-& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
-&= \begin{cases}
-    	\delta_k^{(b)} \: <H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
-	\delta_{k-1}^{(b)} \: <H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
-	0 & \text{o/w}
-      \end{cases}.
-\end{align}
+\sotodoinline{Define norms and inner products below}
 
-We then have that multiplication by \(x\) and \(y\) yields the 3-term recurrences as follows:
+\begin{lemma}
+$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
 \begin{align}
@@
 \alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
+\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
+\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
+\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
+\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
+\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
+\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
 \end{align}
 
-Now, set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
+\end{lemma}
+
+\begin{proof}
+Note that \(\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
+\begin{align}
+\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &= \int \int_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
+&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
+& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
+&= \begin{cases}
+    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
+	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
+	0 & \text{otherwise}
+      \end{cases}.
+\end{align}
+
+\end{proof}
+
+Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
+Set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
 \begin{align}
@@
 
-\subsection{Clenshaw on the half disk}
+\subsection{Building the OPs}
 
@@
 
-For general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
+For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
 \begin{align}
@@
 
+\section{Computational aspects}
+
+\subsection{Constructing $H_n^{(a,b)}(x)$}
+
+\subsection{Quadrature rules}
+
+In this section we construct a quadrature rule exact for polynomials in $\Omega$ that can be used to expand functions in $P_{n,k}^{(a,b)}(x,y)$.
+
+\begin{theorem}
+
+Denote the  Gauss quadrature nodes and weight on \([0,1]\) with weight \(s^a \: (1-s^2)^{b+\half}\) as $(s_k,w_k^{(s)})$ , and
+ on \([-1,1]\) with weight \((1-t^2)^b\) as $(t_k,w_k^{(t)})$. Define
+\begin{align}
+x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
+y_{l+(i-1)N} &:= (1-s_l^2)^\half \: t_l, \quad i,l = 1,\dots,N, \\
+w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
+\end{align}
+If $f(x,y)$ is an at most degree $N$ polynomial in $x$ and at most degree $2N-1$ in $y$ then
+$$
+\iint_\Omega f(x,y) \D A =\sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
+$$
+
+\end{theorem}
+\begin{proof}
+
+We will use the substitution that
+\begin{align}
+x &= s, \quad y = (1-s^2)^\half t.
+\end{align}
+First, note that
+\begin{align}
+W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
+		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
+\end{align}
+
+\end{proof}
+
+\subsection{Obtaining the coefficients for expansion of a function}
+
+Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by
+\begin{align*}
+f(x,y) \approx \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
+\end{align*}
+where
+\begin{align*}
+\bigP_n(x,y) &:= \begin{bmatrix}
+		P_{n,0}(x,y) \\
+		\vdots \\
+		P_{n,n}(x,y)
+	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
+\end{align*}
+and where
+\begin{align*}
+\mathbf{f}_n &:= \begin{bmatrix}
+		f_{n,0} \\
+		\vdots \\
+		f_{n,n}
+	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
+f_{n,k} := \frac{\langle f, \: \Pnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
+\end{align*}
+
+We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.
+
+Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$\sotodo{Clarify this}.
+Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e.
+that $d_2$ is an even integer).
+Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$.
+Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
+\begin{align}
+& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
+&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
+&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
+&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
+&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
+\end{align}
+where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
+
+We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
+\begin{align}
+\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
+\end{align}
+
+Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
+\begin{align}
+\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
+\end{align}
+and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$. Thus,
+\begin{align}
+\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
+|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
+\end{align}
+for suitably large $N$.
+
 %
@@
 		\bigWii_0(x,y) \\
-		\hline
 		\bigWii_1(x,y) \\
-		\hline
 		\bigWii_2(x,y) \\
-		\hline
 		\vdots \\
-		\hline
 		\bigWii_p(x,y)

### 2019-01-02 arcexperiments.jl

@@ -1,42 +1,169 @@
-using CircleArc, ApproxFun, SO, BlockArrays
+include("/Users/sheehanolver/Documents/SUpervision/project-phd/circle-arc/CircleArc.jl")
 
-N = 40; h = 0
-    D = arc_derivative_operator_in_Q(N, h)
+using CircleArc, ApproxFun, SO, BlockArrays, Plots
+
+N = 100; h = 0
 C = arc2Q(N, h)
 
 
-n = 30; C[Block.(n:n+1), Block.(n:n+1)]
+n = 30; inv(C[Block.(n:n+1), Block.(n:n+1)])
 
 
 
 P = initialise_arc_ops(N, h)
 
+θ = 0.1
+
 j = 11;c = [zeros(j); 1; zeros(2N-j)]
-    arc_func_eval(c, h, P..., cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 4E-15
+    arc_func_eval(c, h, P, cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 4E-15
 j = 12;c = [zeros(j); 1; zeros(2N-j)]
-    arc_func_eval(c, h, P..., cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 4E-8
+    arc_func_eval(c, h, P, cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 4E-8
 j = 21;c = [zeros(j); 1; zeros(2N-j)]
-    arc_func_eval(c, h, P..., cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 1E-14
+    arc_func_eval(c, h, P, cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 1E-14
 j = 22;c = [zeros(j); 1; zeros(2N-j)]
-    arc_func_eval(c, h, P..., cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 0.7
+    arc_func_eval(c, h, P, cos(θ), sin(θ)) - Q_func_eval(C*c, h, cos(θ), sin(θ)) # 0.7
 
 
-norm(C*c)
+using SO
 
-0.001*C
[…]

### 2019-01-03 papers/halfdisk/halfdisk.tex
Edits to paper

@@
 
-\include somacros
+\include{somacros}
 
@@
 
-\sotodoinline{Define norms and inner products below}
-
 \begin{lemma}
@@
 \end{align}
+where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight.
 
@@
 
-Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$\sotodo{Clarify this}.
+Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$ (i.e.
+$f(\cdot, y)$ with $y$ fixed is a degree $d_1$ polynomial, and $f(x, \cdot)$ with $x$ fixed is an even degree $d_2$ polynomial).
 Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e.
@@
 \subsection{Heat Equation}
-\bstodo{Struggling to get plotting working}
 Fix \(a, b = 1,1\), and consider

### 2019-02-08 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Methods to plot figures in the experiments/halfdisk

@@ -64,38 +64,6 @@ function lanczos!(w, P, β, γ, N₀=0)
 
     P, β, γ
 end
-function lanczos2!(S, pts, w, F, N₀=0)
-    # F is the added factor as a Fun
-    P, β, γ = S.ops, S.a, S.b
-
-    # x * P[n](x) == (γ[n] * P[n+1](x) + β[n] * P[n](x) + γ[n-1] * P[n-1](x))
-
-    # We use our own quadrature rule here, as given by pts and w
-
-    N = length(β)
-    X = Fun(identity, domain(S.weight))
-
-    if N₀ <= 0
-        N₀ = 1
-        f1 = Fun(1 / sqrt(sum(S.weight)), space(X))
-        P[1] = f1
-        v = X * P[1]
-        β[1] = inner(S, v * F, P[1], pts, w)
-        v = v - β[1] * P[1]
-        γ[1] = sqrt(inner(S, v * F, v, pts, w))
-        P[2] = v / γ[1]
-    end
-
-    for k = N₀+1:N
-        v = X * P[k] - γ[k-1] * P[k-1]
-        β[k] = inner(S, v * F, P[k], pts, w)
-        v = v - β[k] * P[k]
-        γ[k] = sqrt(inner(S, v * F, v, pts, w))
-        P[k+1] = v / γ[k]
-    end
-
-    P, β, γ
-end
 
 # Finds the OPs and recurrence for weight w
 function lanczos(w, N)
@@ -105,41 +73,6 @@ function lanczos(w, N)
     γ = Array{eltype(w)}(undef, N)
     lanczos!(w, P, β, γ)
 end
-
-# Decides which lanczos! to call depending on if we would run into
-# ill-conditioned errors etc., and then calls it FA,WW,F,D,B,R,N
-function lanczosdecider(S::OrthogonalPolynomialSpace{<:Any,<:Any,<:Any,<:Any,<:Any,T,<:Any}, N₀) where T
-    # TODO: Make this cleaner...
-    ϕ = Vector{T}(undef, length(S.params))
-    ψ = []
-    for i = 1:length(S.params)
-        α = S.params[i]
-        if floor(α) == α || α < 1
-            ϕ[i] = α
-        else # We have non-integer parameter
-            ϕ[i] = α - floor(α)
-            append!(ψ, i)
-        end
-    end
-    if length(ψ) == 0
[…]

### 2019-02-08 experiments/halfdisk.jl

@@ -14,20 +14,132 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
 # points(), and use in a new basiseval method (using 3-term recurrence)
 
 a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
-N = 15
+x, y = 0.4, -0.2; z = [x; y]
+N = 10
+laplace(D, N)
+
+N = 30
+S = D(a, b)
+D
 @time W = OrthogonalPolynomialFamilies.weightedpartialoperatorx(S, N)
 @time W = OrthogonalPolynomialFamilies.weightedpartialoperatory(S, N)
 @time A = OrthogonalPolynomialFamilies.partialoperatorx(S, N+2)
 @time A = OrthogonalPolynomialFamilies.partialoperatory(S, N+2)
 S = D(1, 0)
-@time T = OrthogonalPolynomialFamilies.transformoperator(S, N+1)
+@time T = transformoperator(S, N+1)
 S = D(0, 1)
-@time T = OrthogonalPolynomialFamilies.transformoperator(S, N+1)
+@time T = transformoperator(S, N+1)
 S = D(0, 2)
-@time T = OrthogonalPolynomialFamilies.transformoperator(S, N+1)
+@time T = transformoperator(S, N+1)
 
 D = HalfDiskFamily(); N = 15
 @time laplace(D, N)
+
+#====#
+#=
+Sparsity of laplace operator
+=#
+using PyPlot
+a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
+N = 30
+Δ = laplacesquare(D, N)
+DDD = copy(Δ)
+PyPlot.clf()
+PyPlot.spy(Array(sparse(Δ)))
+PyPlot.axis(xmin=0, xmax=500, ymin=500, ymax=0)
+PyPlot.title("Sparsity of Δ operator")
+savefig("experiments/sparsityoflaplacian")
+
+#=
+Solution of Δu = f, where f(x,y) = 1 + erf(5(1 - 10((x - 0.5)^2 + (y)^2)))
+                        # f = y*exp(x)*(2-11x-6x^2-x^3-2y^2-x*y^2)
+=#
+using SpecialFunctions
+using Makie, FileIO
[…]

### 2019-02-08 posters/jamboree2019.tex
First draft of poster for HD. Minor edits to paper.

--- 2018-12-14 papers/halfdisk/halfdisk.tex
@@
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
+\documentclass{imposter}
+%\documentclass[landscape]{imposter}
+
+%\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
 %\geometry{landscape}                		% Activate for rotated page geometry
 \usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
+%\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
 								% TeX will automatically convert eps --> pdf in pdflatex
-\usepackage{caption}
-\usepackage{subcaption}
+
+\usepackage{wrapfig}								
+\usepackage[font=scriptsize,labelfont=bf]{caption}
+\usepackage[font=scriptsize]{subcaption}
 \usepackage{float}
+% \graphicspath{ {images/poster-eps/} }
 \usepackage{amssymb}
 \usepackage{amsmath}
-\usepackage{bm}
 \usepackage{bbm}
-\usepackage{mleftright}
+
 \usepackage{todonotes}
 
+\graphicspath{ {../papers/halfdisk/images/} }
 \graphicspath{ {images/} }
@@
 \usepackage{url}
-%\bibliographystyle{elsarticle-harv}
-\bibliographystyle{plain}
+\bibliographystyle{abbrv}
 
@@
 
-%\usepackage{media9}
-%\graphicspath{ {../sphere/plots} }
+\usepackage{amsthm}
 
-\title{Sparse Spectral Methods and P-Finite Element Methods on Disk-Splices and Rhombi - Part I: Zero Boundary Conditions}
-\author{Sheehan Olver, Ben Snowball}
-%\date{}							% Activate to display a given date or no date
+\newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma} 
+\newtheorem{theorem}{Theorem} 
+\newtheorem{definition}{Definition}
+
+\include{somacros}
+
+\usepackage{listings}
+
+\title{Sparse Spectral Methods on Disk Slices}
+\author{Ben Snowball (Supervisors: Sheehan Olver (Main), Colin Cotter)}
+\department{Department of Mathematics, MPE CDT}
+\institution{Imperial College London}
+
+%% Primary university logo.
+%\logo{imperial.eps}
+% Secondary logos (not supported in all poster styles)
+%\logoa{AMCG.eps}
+
+%% Options to set colors of multiple poster features.
+\titlecolor{named}{Blue}
+\postercolor{named}{Blue}
+%\pagecolor{green}
+
+\posterstyle{mpecdtstyle}
+
+% This sets up boxes to have nice rounded corners and somewhat thicker lines
+% than the default.
+\psset{linecolor=postercolor,linewidth=.2,cornersize=absolute,linearc=0.5,framesep=0.5}
 
 \begin{document}
+%\showgrid
 
-\maketitle
+% LHS
 
-%
-\section{Introduction}
+  \leftbox{96}{
+        \section{Abstract}
 
-This document is to derive a framework for solving PDEs on the half disk using sparse spectral methods.
-We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk.
-We will derive a quadrature rule on the half disk that can be used to expand a function in the OP basis up to a given order.
-We will demonstrate that these families will have sparse operators (including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators), and how we can implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
-We can extend this work directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc.
-can be found similarly.
+In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials (OPs) on intervals, disks, and triangles (e.g.
+\cite{olver2018recurrence}).
+Here, we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
+a half disk), building on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate OPs.
 
-\section{Orthogonal polynomials  and function approximation on the half disk (2D)}
+    }
 
-By using a similar process to \cite{dunkl2014orthogonal} (p55-56) we can construct 2D orthogonal polynomials on the half disk from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
-We denote the half disk by \(\Omega\) where
-\begin{align}
-\Omega := \{(x,y) \in \R^2 \quad | \quad 0 < x < 1, \: -(1-x^2)^{\half} < y < (1-x^2)^{\half}\}.
-\end{align}
+  \leftbox{81}{
+      \section{Orthogonal polynomials on the half disk}
 
-Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
+By using a similar process to \cite[p55--56]{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
 
-1) \(\rho\) is a degree 1 polynomial, or
+\begin{proposition}
+Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\),
+Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\).
+Then
+$$
+\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
+$$
+are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on $\Omega$. 
+\end{proposition}
 
-2) \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
-
-Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\).
-
-We define \(\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)})\) for \( 0 \le k \le n, \: n = 0,1,2,\dots\) and note that the functions \(\{\Pnk\}\) are in fact polynomials in either cases above.
-Furthermore, they are orthogonal with respect to the weight \(W(x,y) := w_1(x) \: w_2(\frac{y}{\rho(x)}) \) on \(R := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}\).
-
-Let \(a, b \in \R\). For the half disk, if we set:
-
+Taking $\Omega$ as the half disk, the weight $W^{(a,b)}(x,y)$ results from setting:
 \begin{align}
 \begin{cases}
-\alpha &:= 0 \\
-\beta &:= 1 \\
-\gamma &:= -1 \\
-\delta &:= -\gamma = 1 \\
+(\alpha,\beta, \gamma,\delta) &:= (0.1, -1, 1) \\
 \rho(x) &:= (1-x^2)^{\half} \\
 w_1(x) &:= x^a \: (1-x^2)^b \\
-w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b,
+w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b.
 \end{cases}
@@
 
-we can obtain the 1D orthogonal polynomials \(\{H_n^{(a, b)}\}\) orthogonal with respect to the weight \(x^a \: (1-x^2)^b\), and \(\{P_n^{(b, b)}\}\) orthogonal with respect to the weight \((1-x^2)^b\) (so \(\{P_n^{(b, b)}\}\) are in fact Jacobi polynomials), leading to the two parameter 2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+The weight $w_2(x)$ corresponds to the weight for the Jacobi polynomials \(\{P_n^{(b, b)}\}\).
+The weight $w_1(x)$ is non-classical and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
+Thus we arrive at a two parameter family of 2D OPs \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
- \Pnk(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+ \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
-orthogonal on \(\Omega\) with respect to the weight 
+orthogonal on \(\Omega\) with respect to the weight $W^{(a,b)}(x,y)$.
+
+    }
+
+    \leftbox{37}{
+    	\section{Jacobi matrices and building the OPs}
+Define, for some $N \in \No$:
 \begin{align}
-W^{(a,b)}(x,y) := x^a \: (1-x^2-y^2)^b, \quad (x,y) \in \Omega.
+\bigPab_n := \begin{bmatrix}
+		P^{(a,b)}_{n,-n}(x,y) \\
+		\vdots \\
+		P^{(a,b)}_{n,n}(x,y)
+	\end{bmatrix} \in \R^{n+1}, 
+\quad \quad 
+\bigPab := \begin{bmatrix}
+		\bigPab_0 \\
+		\bigPab_1 \\
+		\vdots \\
+		\bigPab_N
+	\end{bmatrix}.
 \end{align}
 
-This provides us with a family of orthogonal polynomial spaces on the half disk.
-Note that we can obtain OP families on other disk-splices and rhombi by appropriately choosing the functions and constants in (2).
-
-\subsection{Obtaining the coefficients for expansion of a function}
-
-Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by
-\begin{align}
-f(x,y) \approx \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
-\end{align}
-where
-\begin{align}
-\bigP(x,y) := \begin{bmatrix}
-		\bigP_0(x,y) \\
-		\hline
-		\bigP_1(x,y) \\
-		\hline
-		\bigP_2(x,y) \\
-		\hline
-		\vdots \\
-		\hline
-		\bigP_N(x,y)
-	\end{bmatrix},
-\quad
-\bigP_n(x,y) := \begin{bmatrix}
-		P_{n,0}(x,y) \\
-		\vdots \\
-		P_{n,n}(x,y)
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
-\end{align}
-and where
-\begin{align}
-\mathbf{f} &:= \begin{bmatrix}
-		\mathbf{f}_0 \\
-		\hline
-		\mathbf{f}_1 \\
-		\hline
-		\mathbf{f}_2\\
-		\hline
-		\vdots \\
-		\hline
-		\mathbf{f}_N
-	\end{bmatrix},
-\quad
-\mathbf{f}_n := \begin{bmatrix}
-		f_{n,0} \\
-		\vdots \\
-		f_{n,n}
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \\
-f_{n,k} &:= \frac{\langle f, \: \Pnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
-\end{align}
-
-We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficients vector \(\mathbf{f}\) as follows.
-
-We will use the substitution that
-\begin{align}
-x &= s, \quad y = (1-s^2)^\half t.
-\end{align}
-First, note that
-\begin{align}
-W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
-		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
-\end{align}
-Using the Golub-Welsch algorithm, we can obtain nodes and weights for 1D quadrature rules on \([0,1]\) with weight \(s^a \: (1-s^2)^{b+\half}\), and on \([-1,1]\) with weight \((1-t^2)^b\).
-
-Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$.
-Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e.
-that $d_2$ is an even integer).
-Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$.
-Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
-\begin{align}
-& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
-&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
-&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
-&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
-&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
-\end{align}
-where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
-
-We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
-\begin{align}
-\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
-\end{align}
-
-Relabelling the weights and nodes as 
-\begin{align}
-x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
-y_{l+(i-1)N} &:= (1-s_l^2)^\half \: t_l, \quad i,l = 1,\dots,N, \\
-w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
-\end{align}
-
-Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
-\begin{align}
-\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
-\end{align}
-and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$. Thus,
-\begin{align}
-< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
-|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
-\end{align}
-for suitably large $N$.
-
-\subsection{Jacobi matrices}
-
-We obtain the 1D OPs via the Lanczos algorithm, which also gives us the coefficients of the relations
-\begin{align}
-x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) \\
-y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
-\end{align}
-where we note that \(\gamma_n^{(b)} = 0\) \(\forall n = 0,1,2,\dots\).
-
-Next, note that \(<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
-\begin{align}
-<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &= \int \int_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
-&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
-& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
-&= \begin{cases}
-    	\delta_k^{(b)} \: <H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
-	\delta_{k-1}^{(b)} \: <H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
-	0 & \text{o/w}
-      \end{cases}.
-\end{align}
-
-We then have that multiplication by \(x\) and \(y\) yields the 3-term recurrences as follows:
-\begin{align}
-x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
-y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
-		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
-		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
-\end{align}
-for \((x,y) \in \Omega\), where
-\begin{align}
-\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
-\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: <H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
-\end{align}
-
-Now, set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
+\begin{lemma}
+The three-term recurrences that $\Pnk^{(a,b)}(x,y)$ satisfies lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
+Set $J_x^{(a,b)}, J_y^{(a,b)} \in \R^{\half (N+1)(N+2) \times \half (N+1)(N+2)}$ as the Jacobi matrices corresponding to
 \begin{align}
@@
 \end{align}
-Then  \(J_x^{(a,b)}, J_y^{(a,b)}\) have the Banded-Block-Banded structures
+Then  \(J_x^{(a,b)}, J_y^{(a,b)}\) have similar sparse (block-tridiagonal) Banded-Block-Banded structures.
+\end{lemma}
+Combining each system in (4) we can write this as a block-tridiagonal system of equations.
+Applying a preconditioner matrix to the system \citep{dunkl2014orthogonal} (p78), we can expand the resulting lower-triangular system to obtain the recurrence:
     \begin{align}
-J_{x/y}^{(a,b)} &= \begin{bmatrix}
-		B^{x/y}_0 & A^{x/y}_0 & & & & \\
-		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
-		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
-		& & C^{x/y}_3 & \ddots & \ddots & \\
-		& & & \ddots & \ddots & \ddots \\
-	\end{bmatrix}
+\begin{cases}
+\bigPab_{-1}(x,y) := 0, \quad \bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
+\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n(x,y) - \Dnt C_n  \,\bigPab_{n-1}(x,y).
+\end{cases}
 \end{align}
-where
+where the matrix coefficients $\Dnt, B_n, C_n, G_n(x,y)$ can be explicitly found.
+    }
+
+    \rightbox{96}{
+    	\section{Sparse partial differential operators}
+
+For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that (as seen in the DLMF \cite{DLMFDerivatives} for other OP families):
+\begin{align*}
+\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^T \: D_x^{(a,b)} \: \mathbf{f}, \\
+\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^T \: D_y^{(a,b)} \: \mathbf{f}, \\
+\dx[\Wab(x,y) \: f(x,y)] &= W^{(a-1,b-1)}(x,y) \: \bigP^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
+\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
+\end{align*}
+Further, we note that there exist Banded-Block-Banded matrix operators $T^{(1,0)\to(0,0)}\), \(T^{(0,1)\to(1,1)}$ such that:
+\begin{align*}
+x \bigP^{(1,0)}(x,y) &= \Big(T^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
+\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y)
+\end{align*}
 \begin{align}
-A^x_n &:= \begin{bmatrix}
-		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
-		& \ddots & & \vdots & \\
-		& & \alphaab_{n+1,n,1} & 0 \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
-B^x_n &:= \begin{bmatrix}
-		\alphaab_{n,0,2} & & \\
-		& \ddots & \\
-		& & \alphaab_{n,n,2} \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots,N \\
-C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots,N, \\ 
-\nonumber \\
-A^y_n &:= \begin{bmatrix}
-		0 & \betaab_{n,0,6} & & & \\
-		\betaab_{n,1,5} & 0 & & & \\
-		& \ddots & \ddots & \ddots & \\
-		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
-B^y_n &:= \begin{bmatrix}
-		0 & \betaab_{n,0,4} & & \\
-		\betaab_{n,1,3} & 0 & \ddots & \\
-		& \ddots & \ddots & \betaab_{n,n-1,4} \\
-		& & \betaab_{n,n,3} & 0
-	    \end{bmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots,N \\
-C^y_n &:= \begin{bmatrix}
-		0 & \betaab_{n,0,2} & & \\
-		\betaab_{n,1,1} & 0 & \ddots & \\
-		& \ddots & \ddots & \betaab_{n,n-2,2} \\
-		& & \ddots & 0 \\
-		& & & \betaab_{n,n,1} \\
-	    \end{bmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots,N.
+    \implies \Delta^{(1,1)} := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)},
 \end{align}
 
-Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs.
-In other words, sparsity is not dependent on the domain.
-
-\subsection{Clenshaw on the half disk}
-
-Combining each system in (71) we can write the block-tridiagonal system
-\begin{align}
-\renewcommand\arraystretch{1.3}
-\mleft[
-\begin{array}{cccc}
-		1 & & & \\
-		B_0-G_0(x,y) & A_0 & & \\
-		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
-		& C_2 & \ddots & \ddots \\
-		& & \ddots &
-\end{array}
-\mright]
-\bigPab(x,y)
-=
-\begin{bmatrix}
-	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
-\end{bmatrix}.
-\end{align}
-where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each \(n = 0,1,\dots,N\):
-\begin{align}
-A_n &:= \begin{bmatrix}
-		A^x_n \\
-		A^y_n \\
-		A^z_n
-	    \end{bmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
-C_n := \begin{bmatrix}
-		C^x_n \\
-		C^y_n
-	    \end{bmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
-B_n &:= \begin{bmatrix}
-		B^x_n \\
-		B^y_n
-	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
-G_n(x,y,z) := \begin{bmatrix}
-		xI_{n+1} \\
-		yI_{n+1}
-	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}.
-\end{align}
-
-For each \(n = 0,1,\dots,N-1\) let \(\Dnt\) be any matrix that is a left inverse of \(A_n\), i.e.
-such that \(\Dnt A_n = I_{n+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} (p78), which can be expanded to obtain the recurrence:
-\begin{align}
-\begin{cases}
-\bigPab_{-1}(x,y) := 0 \\
-\bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
-\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots,N.
-\end{cases}
-\end{align}
-
-Note that we can define \(\Dnt\) as follows. For \(n\) even:
-\begin{align}
-\Dnt := \begin{bmatrix}
-		\frac{1}{\alphaab_{n+1,0,1}} & & &  \\
-		& \ddots & & & \\
-		& & \frac{1}{\alphaab_{n+1,n,1}} & \\
-		0 & \hdots & 0 & \eta_{m-1} & \hdots & 0 & \eta_1 & 0 & \eta_0
-	    \end{bmatrix},
-\end{align}
-where
-\begin{align}
-m &= \frac{n}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
-\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-1.
-\end{align}
-
-For \(n\) odd:
-
-\begin{align}
-\Dnt := \begin{bmatrix}
-		\frac{1}{\alphaab_{n+1,0,1}} & &  \\
-		& \ddots & & &  \\
-		& & \ddots & & \\
-		& & & \frac{1}{\alphaab_{n+1,n,1}} & \\
-		\xi & 0 & \hdots & 0 & 0 & \eta_{m-2} & \hdots & 0 & \eta_0
-	    \end{bmatrix},
-\end{align}
- where
-\begin{align}
-m &= \frac{n+1}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
-\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-2, \\
-\xi &= -\frac{\betaab_{n,1,5} \: \eta_{m-2}}{\alphaab_{n+1,0,1}}. \nonumber
-\end{align}
-
-%
-\section{Sparse partial differential operators}
-
-\begin{figure} 
+\begin{wrapfigure}{R}{0.7\textwidth}
+        \center 
+   \begin{subfigure}[t]{0.35\textwidth}
+         \centering
+        \includegraphics[scale=1.0]{sparsityoflaplacian}
+        \label{fig:sparsity}
+    \end{subfigure}
+   \begin{subfigure}[t]{0.3\textwidth}
+         \centering
+        \includegraphics[scale=0.6]{solution}
+        \label{fig:solution}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.3\textwidth}
                 \center
@@
         \end{tikzpicture} 
-\caption{The Laplace operator acting on vectors of $P_{n,k}=P_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{P^{(2,2)}_{n,k}}$ coefficients.
-Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{P_{n,k}^{(a,b)}}$ coefficients, where $(a,b)$ is at the tail of the arrow, and the range is $\smash{P_{n,k}^{(\tilde{a},\tilde{b})}}$ coefficients, where $(\tilde{a},\tilde{b})$ is at the head of the arrow.
-%For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{P_{n,k}^{(0,0,0)}}$ (resp.~$\smash{P_{n,k}^{(1,0,1)}}$) and the range in $\smash{P_{n,k}^{(1,0,1)}}$ (resp.~$\smash{P_{n,k}^{(2,0,2)}}$).
+	\label{fig:Laplace}
+   \end{subfigure}
+    \centering
+    \captionsetup{width=.7\linewidth}
+    \caption{Left: A "spy" plot of the $\Delta$ operator for $\Wii(x,y) \: \bigPii(x,y) \to \bigPii(x,y)$ showing its sparsity.
+Right: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 ?
+10((x ?
+0.5)^2 + y^2)))$.
+Bottom: Diagram illustrating how the parameters change under partial differentiation.
 }
-\label{fig:Laplace} 
-\end{figure}
+\end{wrapfigure}
 
-For general \(a,b\) and general function \(f(x,y) = \mathbf{f}^T \: \bigPab(x,y)\), there exist Banded-Block-Banded matrix operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
-\begin{align}
-\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^T \: D_x^{(a,b)} \: \mathbf{f}, \\
-\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^T \: D_y^{(a,b)} \: \mathbf{f}, \\
-\dx[\Wab(x,y) \: f(x,y)] &= W^{(a-1,b-1)}(x,y) \: \bigP^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
-\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
-\end{align}
-The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials, as seen in the DLMF \cite{DLMFDerivatives}.
-Further, we note that there exist Banded-Block-Banded matrix operators \(T^{(1,0)\to(0,0)}\), \(T^{(0,1)\to(1,1)}\) and $T^{(0,2)\to(2,2)}$ such that:
-\begin{align}
-x \bigP^{(1,0)}(x,y) &= \Big(T^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
-\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y) \\
-\bigP^{(0,2)}(x,y) &= \Big(T^{(0,2)\to(2,2)} \Big)^T \: \bigP^{(2,2)}(x,y)
-\end{align}
-
-We can then obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space
-\[
-\Wii(x,y) \: \bigPii(x,y)
-\]
-to coefficients in the non-weighted space
-\[
-\bigPii(x,y).
-\]
-Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse.
-Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
-The matrix operator \(\Delta\) acting on the coefficients vector is then given by
-\begin{align}
-    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
-\end{align}
-
-Further, we can also obtain the matrix operator for \(\Delta\) that will take us from coefficients for expansion in the space
-\[
-\bigPoo(x,y)
-\]
-to coefficients in the space
-\[
-\bigP^{(2,2)}(x,y).
-\]
-Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse. The matrix operator \(\Delta\) acting on the coefficients vector is then given by
-\begin{align}
-    D := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(0,2)\to(2,2)} \: D_y^{(0,1)} \: D_y^{(0,0)}.
-\end{align}
+where $\Delta^{(1,1)}$ is the matrix operator for the Laplacian, that will take us from coefficients for expansion in the weighted space $\Wii(x,y) \: \bigPii(x,y)$ to coefficients in the non-weighted space $\bigPii(x,y)$.
+This operator will have Banded-Block-Banded structure, and hence sparse.
+Note that this construction ensures imposition of Dirichlet zero boundary conditions.
 
@@
 
-%
-\section{P-finite element methods using sparse operators}
+    }
 
-We follow the method of \cite{beuchler2006new}. Consider the 2D Dirichlet problem on a domain $\Omega$:
-\begin{align}
-	\begin{cases}
-         - \Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
-         u = 0 \quad \text{on } \partial \Omega
-         \end{cases}
-\end{align}
-This has the weak formulation for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$,
-\begin{align}
-	L(v) := \int_\Omega f \: v \: d\mathbf{x} = \int_\Omega \nabla u \cdot \nabla v \: d\mathbf{x} =: a(u,v).
-\end{align}
+     \rightbox{30.5}{
+      \section{Summary and Future Work}
 
-In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a rhombus or disk slice for example.
-For the purposes of this paper, we limit this to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/rhombus domain.
-We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h|_\element) \le p\}$ for some $p \in \N$.
+\begin{itemize}
+  \item We have detailed a method for obtaining families of 2D OPs on certain domains.
+Using the three-term recurrence relationships the OPs have, we can calculate sparse operator matrices for multiplication by $x, y$ and in turn sparse differential operators, allowing us to solve PDEs on the half disk.
+  \item The same (or similar) relations hold for other suitable domains (namely, disk slices). 
+  \item This 2D half disk framework can be extended for a family of OPs on the hemisphere, with sparse relations.
+   \item To solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas.
+\end{itemize}
 
-We seek $u_h \in V_h$ s.t.
-\begin{align}
-	L(v_h) = a(u_h,v_h) \quad \forall \: v_h \in V_h.
-\end{align}
+    }
 
-Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^T \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to.
-Note that due to orthogonality this is a diagonal matrix.
-We can choose a basis for $V_h$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
-\begin{align}
-\bigWii(x,y) &:= \begin{bmatrix}
-		\bigWii_0(x,y) \\
-		\hline
-		\bigWii_1(x,y) \\
-		\hline
-		\bigWii_2(x,y) \\
-		\hline
-		\vdots \\
-		\hline
-		\bigWii_p(x,y)
-	\end{bmatrix}, \\
-\bigWii_n(x,y) &:= \begin{bmatrix}
-		\Wii(x,y) \: \Pii_{n,0}(x,y) \\
-		\vdots \\
-		\Wii(a,y) \: \Pii_{n,n}(x,y)
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
-\end{align}
-and rewrite (63) in matrix form:
-\begin{align}
-	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
-	&= \int_\element \begin{bmatrix}
-					\partial_x v_h \\
-					\partial_y v_h
-				\end{bmatrix}^T 
-				\begin{bmatrix}
-					\partial_x u_h \\
-					\partial_y u_h
-				\end{bmatrix}
-				\\
-	&= \int_\element \begin{bmatrix}
-					\bigPoo^T \Wii_x \mathbf{v} \\
-					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
-				\end{bmatrix}^T 
-				\begin{bmatrix}
-					\bigPoo^T \Wii_x \mathbf{u} \\
-					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
-				\end{bmatrix}
-				\\
-	&= \int_\element \Big( \mathbf{v}^T {\Wii_x}^T \bigPoo \bigPoo^T \Wii_x \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^T ({T^{(1,0)\to(0,0)} \Wii_y})^T \bigPoo \bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
-	&= \mathbf{v}^T \: \Big( {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
-\end{align}
-where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_h, v_h \in V_h$ respectively in the $V_h$ basis ($\bigWii$ OPs), and
-\begin{align}
-	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
-	&= \int_\element \: \mathbf{v}^T \: \bigWii \: \bigPii^T \: \mathbf{f} \: d\mathbf{x} \\
-	&= \mathbf{v}^T \Lambda^{(1,1)} \: \mathbf{f},
-\end{align}
-where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bigPii$ OP basis.
-We can also not that naturally $\langle \bigPii, {\bigPii}^T \rangle_{\Wii}$ is just a diagonal matrix.
+%Also do graphics.
+    \rightbox{11}{
+        {\footnotesize\bibliography{../papers/halfdisk/halfdisk-bib}}
 
-Since (63) is equivalent to stating that
-\begin{align}
-	L(\Wii \Pii_{n,k}) = a(u_h,\Wii \Pii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
-\end{align}
-(i.e. holds for all basis functions of $V_h$) by choosing $v_h$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-\begin{align}
-A\mathbf{u} = \tilde{\mathbf{f}}.
-\end{align}
-where the (element) stiffness matrix $A$ is defined by 
-\begin{align}
-A = {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
-\end{align}
-and the load vector $\tilde{\mathbf{f}}$ is given by 
-\begin{align}
-\tilde{\mathbf{f}} = \Lambda^{(1,1)} \: \mathbf{f}.
-\end{align}
-
-Note the since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-
-%
-\section{Examples on the half disk with zero BCs}
-
-\subsection{Poisson}
-
-Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
-\begin{align}
-    \Delta(\Wii \: u)(x,y) = f(x,y) \quad \text{in } \Omega
-\end{align}
-    where we recall
-\begin{align}
-    \Wii(x,y) = x \: (1-x^2-y^2)
-\end{align}
-which ensures a boundary condition that \((\Wii \: u)(x,y) = 0\) on \(\partial \Omega\).
-
-The first step is to obtain the expansion of \(f\) in the \(\Pii\) basis, i.e.
-obtain the vector \(\mathbf{f}\) such that \(f(x,y) = \mathbf{f}^T \: \bigP^{(1,1)}(x,y) = \sum_{n=0}^N \mathbf{f}_n^T \: \bigP_n^{(1,1)}(x,y)\) for some \(N \in \N\).
-The coefficients of \(u\) in the \(\Pii\) basis are then given by
-\begin{align}
-    \mathbf{u} = \Delta \text{\textbackslash} \mathbf{f}.
-\end{align}
-
-\begin{figure}
-	\includegraphics[scale=0.4]{example4coeffs}
-\centering
-\caption{Plot of the difference between the previous solution's coefficients and the current solution's coefficients under increasing degree of approximation (number of coefficients used) for $N = 2,\dots,10$, using RHS of $f(x,y) = 2(1-x^2-y^2)\cos(x) - 8x\sin(x) - x(1-x^2-y^2)\sin(x) - 4x^2 \cos(x) - 12xy$ for the Poisson equation.
-The solution is converging as seen by the decreasing (absolute value) difference of coefficient values.}
-\centering
-\label{fig:poisson}
-\end{figure}
-
-\subsection{Inhomogenous Helmholtz}
-
-Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
-\begin{align}
-    \Delta(\Wii u)(x,y) + k^2 u(x,y) = - f(x,y) \quad \text{in } \Omega
-\end{align}
-noting that this ensures a boundary condition that \((\Wii \: u)(x,y) = 0\) on \(\partial \Omega\).
-
-Again, the first step is to obtain the expansion of \(f\) in the \(\Pii\) basis. The coefficients of \(u\) in the \(\Pii\) basis are then given by 
-\begin{align}
-    \mathbf{u} = - (k^2I + \Delta) \text{\textbackslash} \mathbf{f}.
-\end{align}
-
-\subsection{Heat Equation}
-\bstodo{Struggling to get plotting working}
-Fix \(a, b = 1,1\), and consider
-\begin{align}
-\frac{\partial u}{\partial t}(x,y) = \Delta(\Wii \: u)(x,y) \quad \text{in } \Omega
-\end{align}
-
-Using a backward Euler time-stepping method, we can formulate this for \(t = 0,1,2,\dots\) and step length $h$:
-\begin{align}
-u_{t+1}(x,y) - u_t (x,y) = h \Delta(\Wii \: u_{t+1})(x,y) \quad \text{for } (x,y) \in \Omega.
-\end{align}
-
-There exist $N \in \N$ such that, for each $t = 0,1,2,\dots$, $u_t (x,y) = \sum_{n=0}^N \mathbf{u_t}_n^T \: \bigP_n^{(1,1)}(x,y) = \mathbf{u}_t^T \; \bigP^{(1,1)}(x,y)$.
-Then, using the operator $\Delta$, we can solve this system for a given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
-\begin{align}
-    \mathbf{u_{t+1}} &= (I - h \Delta) \: \text{\textbackslash} \: \mathbf{u_t}, \\
-    \big( u_{t+1}(x,y) &= \mathbf{u}_{t+1}^T \; \bigP^{(1,1)}(x,y) \big) \nonumber.
-\end{align}
-
-\begin{figure}
-	\includegraphics[scale=0.4]{example2timesteppingcovnvergence}
-\centering
-\caption{Plot of the current iteration's coefficients for the heat equation example, using degree of approximation $N = 5$ with initial condition of $u_0(x,y) = \Wii(x,y)$ and step size $h=0.01$, plotted for iterations $t=100, 120, \dots, 200$.
-The solution tends to zero, as seen by the decreasing (absolute value) coefficient values.}
-\centering
-\label{fig:heateqn}
-\end{figure}
-
-%
-\section{Future work - orthogonal polynomials on the hemisphere and spherical triangles}
-
-We will use a similar idea of projecting up from 1D OPs to the interval, to 2D OPs on the disk, to 3D OPs on the sphere.
-Initially, we will add to the ApproxFun library by adding capability to work with OPs on the half disk (2D) with weights governed by parameters that will be incremented/decremented after an application of an operator (e.g.
-partial derivative).
-With this, we can obtain a family of OPs on the hemisphere, with sparse relations.
-
-Beyond this, solving PDEs on isosceles triangles should follow using the same procedure.
-However, we are expecting that to solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas beyond the procedures given in this document.
-
-\bibliography{halfdisk-bib}
+    }
 

### 2019-02-11 experiments/halfdisk.jl
Plots

@@ -104,39 +104,47 @@ FileIO.save("experiments/solution.png", s)
 
 
 #=
-Plotting the norms of each block of coeffs for various function approximations
+Plotting the norms of each block of coeffs for solutions of Poisson for
+different RHSs
 =#
-m = 400
-f1 = Fun((x,y)->1.0, S, m)
-N = getnk(length(f1.coefficients))[1]
-f2 = Fun((x,y)->x^2 + y^2, S, m)
-f3 = Fun((x,y)->x^2*y^2*(1-x^2-y^2)^2, S, m)
+m = 800
 f4 = Fun((x,y)->exp(-1000((x-0.2)^2+(y-0.2)^2)), S, m)
-f1coeffs = PseudoBlockArray(f1.coefficients, [i+1 for i=0:N])
-f2coeffs = PseudoBlockArray(f2.coefficients, [i+1 for i=0:N])
-f3coeffs = PseudoBlockArray(f3.coefficients, [i+1 for i=0:N])
-f4coeffs = PseudoBlockArray(f4.coefficients, [i+1 for i=0:N])
-f1norms = zeros(N+1)
-f2norms = zeros(N+1)
-f3norms = zeros(N+1)
-f4norms = zeros(N+1)
+N = getnk(length(f4.coefficients))[1]
+Δc = Δ[1:getopindex(N, N), 1:getopindex(N, N)] # laplacesquare(D, N)
+f1 = Fun((x,y)->1.0, S)
+f2 = Fun((x,y)->x^2 + y^2 - 1, S)
+f3 = Fun((x,y)->x^2*y^2*(1-x^2-y^2)^2, S)
+u1 = Fun(S, sparse(Δc) \ resizecoeffs!(f1, N))
+u2 = Fun(S, sparse(Δc) \ resizecoeffs!(f2, N))
+u3 = Fun(S, sparse(Δc) \ resizecoeffs!(f3, N))
+u4 = Fun(S, sparse(Δc) \ f4.coefficients)
+u1coeffs = PseudoBlockArray(u1.coefficients, [i+1 for i=0:N])
+u2coeffs = PseudoBlockArray(u2.coefficients, [i+1 for i=0:N])
+u3coeffs = PseudoBlockArray(u3.coefficients, [i+1 for i=0:N])
+u4coeffs = PseudoBlockArray(u4.coefficients, [i+1 for i=0:N])
+u1norms = zeros(N+1)
[…]

### 2019-02-19 OrthogonalPolynomialFamilies/test/runtests.jl
Reframing or renaming of the transform operators

@@ -2,8 +2,8 @@ using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature,
         SingularIntegralEquations, Test
 import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule,
         gethalfdiskOP, jacobix, jacobiy, differentiatex, differentiatey,
-        resizecoeffs!, laplace, transformoperator, operatorclenshaw, getnk,
-        convertweightedtononoperator, increaseparamsoperator, getopindex
+        resizecoeffs!, laplace, transformparamsoperator, operatorclenshaw, getnk,
+        getopindex
 
 
 
@@ -171,17 +171,20 @@ end
     a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
     x, y = 0.457, -0.209; z = [x; y] # Test point
     N = 5
-    f1 = Fun(D(0.0, 2.0), rand(sum(1:N+1)))
-    T = transformoperator(D(0.0, 2.0), N)
-    f2 = Fun(D(2.0, 2.0), T * f1.coefficients)
+    Sfrom = D(0.0, 2.0); Sto = D(2.0, 2.0)
+    f1 = Fun(Sfrom, rand(sum(1:N+1)))
+    T = transformparamsoperator(Sfrom, Sto, N)
+    f2 = Fun(D(Sto, T * f1.coefficients)
     @test f2(z) ≈ f1(z)
-    f1 = Fun(D(0.0, 1.0), rand(sum(1:N+1)))
-    T = transformoperator(D(0.0, 1.0), N)
-    f2 = Fun(D(1.0, 1.0), T * f1.coefficients)
+    Sfrom = D(0.0, 1.0); Sto = D(1.0, 1.0)
+    f1 = Fun(Sfrom, rand(sum(1:N+1)))
+    T = transformparamsoperator(Sfrom, Sto, N)
+    f2 = Fun(D(Sto, T * f1.coefficients)
     @test f2(z) ≈ f1(z)
-    f1 = Fun(D(1.0, 0.0), rand(sum(1:N+1)))
-    T = transformoperator(D(1.0, 0.0), N)
-    f2 = Fun(D(0.0, 0.0), T * f1.coefficients)
+    Sfrom = D(1.0, 0.0); Sto = D(0.0, 0.0)
+    f1 = Fun(Sfrom, rand(sum(1:N+1)))
[…]

### 2019-02-19 papers/halfdisk/halfdisk.tex
First draft of poster for HD. Minor edits to paper.
Further edits to paper

@@
 \newcommand{\bigW}{\mathbb{W}}
+\newcommand{\bigWab}{\mathbb{W}^{(a,b)}}
 \newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
@@
 
-For the half disk, the weight $W^{(a,b)}(x,y)$ results from setting:
+For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
 
@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
-Set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
+Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for some $N \in \No$:
+\begin{align}
+\bigPab_n := \begin{bmatrix}
+		P^{(a,b)}_{n,-n}(x,y) \\
+		\vdots \\
+		P^{(a,b)}_{n,n}(x,y)
+	\end{bmatrix} \in \R^{n+1}, 
+\quad \quad 
+\bigPab := \begin{bmatrix}
+		\bigPab_0 \\
+		\bigPab_1 \\
+		\vdots \\
+		\bigPab_N
+	\end{bmatrix}
+\end{align}
+and set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
 \begin{align}
@@
 
+Firstly, we introduce further notation that
+\begin{align}
+\bigWab(x,y) := \Wab(x,y) \bigPab(x,y).
+\end{align}
+
 For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
@@
 \dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^T \: D_y^{(a,b)} \: \mathbf{f}, \\
-\dx[\Wab(x,y) \: f(x,y)] &= W^{(a-1,b-1)}(x,y) \: \bigP^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
-\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
+\dx[\Wab(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
+\dy[\Wab(x,y) \: f(x,y)] &= \bigW^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
 \end{align}
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials, as seen in the DLMF \cite{DLMFDerivatives}.
-Further, we note that there exist Banded-Block-Banded matrix operators \(T^{(1,0)\to(0,0)}\), \(T^{(0,1)\to(1,1)}\) and $T^{(0,2)\to(2,2)}$ such that:
+
+Further, we note that there exist "parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+We denote these operators as $T_W^{(1,0)\to(0,0)}$, $T_W^{(1,1)\to(0,0)}$, $T^{(0,1)\to(1,1)}$, $T^{(0,2)\to(2,2)}$ and the general $T^{(a,b)\to(a+1,b+1)}$ where:
 \begin{align}
-x \bigP^{(1,0)}(x,y) &= \Big(T^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
+\bigW^{(1,0)}(x,y) &= \Big(T_W^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
+\bigW^{(1,1)}(x,y) &= \Big(T_W^{(1,1)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
 \bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y) \\
-\bigP^{(0,2)}(x,y) &= \Big(T^{(0,2)\to(2,2)} \Big)^T \: \bigP^{(2,2)}(x,y)
+\bigP^{(0,2)}(x,y) &= \Big(T^{(0,2)\to(2,2)} \Big)^T \: \bigP^{(2,2)}(x,y) \\
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigP^{(a+1,b+1)}(x,y)
 \end{align}
@@
 \[
-\Wii(x,y) \: \bigPii(x,y)
+\bigWii(x,y) = \Wii(x,y) \: \bigPii(x,y)
 \]
@@
 \begin{align}
-    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
+    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
 \end{align}
@@
 \subsection{Constructing $H_n^{(a,b)}(x)$}
+
+????
 
@@
 
+%
+\section{Examples on the half disk with zero BCs}
+
+\subsection{Poisson}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.4]{solutionblocknorms}
+	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.}
+	\centering
+	\label{fig:solutionblocknorms}
+\end{figure}
+
+Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+\begin{align}
+	\begin{cases}
+    		\Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
+		u(x,y) = 0 \quad \text{on } \partial \Omega
+	\end{cases}.
+\end{align}
+noting the imposition of zero Dirichlet boundary conditions on $u$.
+
+We can tackle the problem as follows.
+Fix \(a, b = 1,1\).
+Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$.
+Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
+In matrix-vector notation, our system hence becomes:
+\begin{align}
+    \Delta^{(1,1)} \mathbf{u} = \mathbf{f}
+\end{align}
+which can be solved to find $\mathbf{u}$.
+
+%\begin{figure}
+%	\includegraphics[scale=0.4]{example4coeffs}
+%\centering
+%\caption{Plot of the difference between the previous solution's coefficients and the current solution's coefficients under increasing degree of approximation (number of coefficients used) for $N = 2,\dots,10$, using RHS of $f(x,y) = 2(1-x^2-y^2)\cos(x) - 8x\sin(x) - x(1-x^2-y^2)\sin(x) - 4x^2 \cos(x) - 12xy$ for the Poisson equation.
+The solution is converging as seen by the decreasing (absolute value) difference of coefficient values.}
+%\centering
+%\label{fig:poisson}
+%\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.4]{sparsityoflaplacian}
+        \label{fig:sparsity}
+    	\caption{A "spy" plot of the $\Delta$ operator for $\bigWii(x,y) \to \bigPii(x,y)$ showing its sparsity.}
+	\centering
+\end{figure}
+
+\subsection{Inhomogenous Helmholtz}
+
+Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+\begin{align}
+	\begin{cases}
+    		\Delta u(x,y) + k^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
+		u(x,y) = 0 \quad \text{on } \partial \Omega
+	\end{cases}.
+\end{align}
+noting the imposition of zero Dirichlet boundary conditions on $u$.
+
+We can tackle the problem as follows.
+Fix \(a, b = 1,1\).
+Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$.
+Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
+In matrix-vector notation, our system hence becomes:
+\begin{align}
+    (\Delta^{(1,1)} + k^2 T^{(0,0)\to(1,1)} T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
+\end{align}
+which can be solved to find $\mathbf{u}$.
+
+We can extend this to constant non-zero boundary conditions by simply noting that the problem 
+\begin{align}
+	\begin{cases}
+    		\Delta u(x,y) + k^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
+		u(x,y) = c \in \R \quad \text{on } \partial \Omega
+	\end{cases}
+\end{align}
+is equivalent to letting $u = v + c$ and solving
+\begin{align}
+	\begin{cases}
+    		\Delta v(x,y) + k^2 v(x,y) = f(x,y) - k^2 c =: g(x,y)  \quad \text{in } \Omega \\
+		v(x,y) = 0 \quad \text{on } \partial \Omega
+	\end{cases}.
+\end{align}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.3]{solution}
+        	\label{fig:solution}
+    	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
+	\centering
+\end{figure}
+
+\subsection{Heat Equation}
+Consider
+\begin{align}
+	\begin{cases}
+		\frac{\partial u}{\partial t}(t,x,y) = \Delta u(t,x,y) \quad \text{in } \Omega \\
+		u(t, x, y) = 0 \quad \text{on } \partial \Omega \\
+		u(0, x, y) = u_0(x,y) := \Wii(x,y) \: f(x,y)
+	\end{cases}
+\end{align}
+noting the imposition of zero Dirichlet boundary conditions.
+
+Using a backward Euler time-stepping method, we can formulate this for \(t = 0,1,2,\dots\) and step length $h$:
+\begin{align}
+u_{t+1}(x,y) - u_t (x,y) = h \: \Delta u_{t+1}(x,y) \quad \text{for } (x,y) \in \Omega.
+\end{align}
+
+We can tackle the problem as follows.
+Fix \(a, b = 1,1\).
+Denote the coefficient vector for expansion of $u_0$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ we denote $\mathbf{u}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bigWii$ OP basis up to degree $N$.
+Further, denote the coefficient vector for expansion of $u_0$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{\tilde{u}}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ denote $\mathbf{\tilde{u}}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bigPii$ OP basis up to degree $N$.
+
+Then, using the operator $\Delta$, we can solve this system for the given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
+\begin{align}
+	&T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (\mathbf{u}_{t+1} - \mathbf{u}_t) = h \: \Delta^{(1,1)} \mathbf{u}_{t+1} \nonumber \\
+	\iff& (T^{(0,0)\to(1,1)} \:T_W^{(1,1)\to(0,0)} - h \: \Delta^{(1,1)}) \: \mathbf{u}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
+	\iff& \mathbf{u}_{t+1} = (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \Delta^{(1,1)})^{-1} T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
+	\iff& \mathbf{\tilde{u}}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \Delta^{(1,1)})^{-1} \: \mathbf{\tilde{u}}_t.
+\end{align}
+
+%\begin{figure}
+%	\includegraphics[scale=0.4]{example2timesteppingcovnvergence}
+%\centering
+%\caption{Plot of the current iteration's coefficients for the heat equation example, using degree of approximation $N = 5$ with initial condition of $u_0(x,y) = \Wii(x,y)$ and step size $h=0.01$, plotted for iterations $t=100, 120, \dots, 200$.
+The solution tends to zero, as seen by the decreasing (absolute value) coefficient values.}
+%\centering
+%\label{fig:heateqn}
+%\end{figure}
+
+%
+\section{Future work - orthogonal polynomials on the hemisphere and spherical triangles}
+
+We will use a similar idea of projecting up from 1D OPs to the interval, to 2D OPs on the disk, to 3D OPs on the sphere.
+Initially, we will add to the ApproxFun library by adding capability to work with OPs on the half disk (2D) with weights governed by parameters that will be incremented/decremented after an application of an operator (e.g.
+partial derivative).
+With this, we can obtain a family of OPs on the hemisphere, with sparse relations.
+
+Beyond this, solving PDEs on isosceles triangles should follow using the same procedure.
+However, we are expecting that to solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas beyond the procedures given in this document.
+
+%
+\appendix
 %
@@
 
-%
-\section{Examples on the half disk with zero BCs}
+\section{Disk Slices}
 
-\subsection{Poisson}
-
-Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+The work in this paper on the half disk can be easily transferred to the domain of a disk slice by which we mean
 \begin{align}
-    \Delta(\Wii \: u)(x,y) = f(x,y) \quad \text{in } \Omega
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
 \end{align}
-    where we recall
-\begin{align}
-    \Wii(x,y) = x \: (1-x^2-y^2)
-\end{align}
-which ensures a boundary condition that \((\Wii \: u)(x,y) = 0\) on \(\partial \Omega\).
-
-The first step is to obtain the expansion of \(f\) in the \(\Pii\) basis, i.e.
-obtain the vector \(\mathbf{f}\) such that \(f(x,y) = \mathbf{f}^T \: \bigP^{(1,1)}(x,y) = \sum_{n=0}^N \mathbf{f}_n^T \: \bigP_n^{(1,1)}(x,y)\) for some \(N \in \N\).
-The coefficients of \(u\) in the \(\Pii\) basis are then given by
-\begin{align}
-    \mathbf{u} = \Delta \text{\textbackslash} \mathbf{f}.
-\end{align}
-
-\begin{figure}
-	\includegraphics[scale=0.4]{example4coeffs}
-\centering
-\caption{Plot of the difference between the previous solution's coefficients and the current solution's coefficients under increasing degree of approximation (number of coefficients used) for $N = 2,\dots,10$, using RHS of $f(x,y) = 2(1-x^2-y^2)\cos(x) - 8x\sin(x) - x(1-x^2-y^2)\sin(x) - 4x^2 \cos(x) - 12xy$ for the Poisson equation.
-The solution is converging as seen by the decreasing (absolute value) difference of coefficient values.}
-\centering
-\label{fig:poisson}
-\end{figure}
-
-\subsection{Inhomogenous Helmholtz}
-
-Fix \(a, b = 1,1\). Find \(u(x,y)\) given a function \(f(x,y)\) such that:
-\begin{align}
-    \Delta(\Wii u)(x,y) + k^2 u(x,y) = - f(x,y) \quad \text{in } \Omega
-\end{align}
-noting that this ensures a boundary condition that \((\Wii \: u)(x,y) = 0\) on \(\partial \Omega\).
-
-Again, the first step is to obtain the expansion of \(f\) in the \(\Pii\) basis. The coefficients of \(u\) in the \(\Pii\) basis are then given by 
-\begin{align}
-    \mathbf{u} = - (k^2I + \Delta) \text{\textbackslash} \mathbf{f}.
-\end{align}
-
-\subsection{Heat Equation}
-Fix \(a, b = 1,1\), and consider
-\begin{align}
-\frac{\partial u}{\partial t}(x,y) = \Delta(\Wii \: u)(x,y) \quad \text{in } \Omega
-\end{align}
-
-Using a backward Euler time-stepping method, we can formulate this for \(t = 0,1,2,\dots\) and step length $h$:
-\begin{align}
-u_{t+1}(x,y) - u_t (x,y) = h \Delta(\Wii \: u_{t+1})(x,y) \quad \text{for } (x,y) \in \Omega.
-\end{align}
-
-There exist $N \in \N$ such that, for each $t = 0,1,2,\dots$, $u_t (x,y) = \sum_{n=0}^N \mathbf{u_t}_n^T \: \bigP_n^{(1,1)}(x,y) = \mathbf{u}_t^T \; \bigP^{(1,1)}(x,y)$.
-Then, using the operator $\Delta$, we can solve this system for a given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
-\begin{align}
-    \mathbf{u_{t+1}} &= (I - h \Delta) \: \text{\textbackslash} \: \mathbf{u_t}, \\
-    \big( u_{t+1}(x,y) &= \mathbf{u}_{t+1}^T \; \bigP^{(1,1)}(x,y) \big) \nonumber.
-\end{align}
-
-\begin{figure}
-	\includegraphics[scale=0.4]{example2timesteppingcovnvergence}
-\centering
-\caption{Plot of the current iteration's coefficients for the heat equation example, using degree of approximation $N = 5$ with initial condition of $u_0(x,y) = \Wii(x,y)$ and step size $h=0.01$, plotted for iterations $t=100, 120, \dots, 200$.
-The solution tends to zero, as seen by the decreasing (absolute value) coefficient values.}
-\centering
-\label{fig:heateqn}
-\end{figure}
-
-%
-\section{Future work - orthogonal polynomials on the hemisphere and spherical triangles}
-
-We will use a similar idea of projecting up from 1D OPs to the interval, to 2D OPs on the disk, to 3D OPs on the sphere.
-Initially, we will add to the ApproxFun library by adding capability to work with OPs on the half disk (2D) with weights governed by parameters that will be incremented/decremented after an application of an operator (e.g.
-partial derivative).
-With this, we can obtain a family of OPs on the hemisphere, with sparse relations.
-
-Beyond this, solving PDEs on isosceles triangles should follow using the same procedure.
-However, we are expecting that to solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas beyond the procedures given in this document.
+with $\alpha, \beta \in [0,1]$, $\rho(x) = \sqrt{1-x^2}$, and  $-\gamma = \delta  = $
 

### 2019-02-22 papers/halfdisk/halfdisk.tex
Further edits to paper

@@
 \end{align}
-and rewrite (63) in matrix form:
+and rewrite (73) in matrix form:
 \begin{align}
@@
 	&= \int_\element \: \mathbf{v}^T \: \bigWii \: \bigPii^T \: \mathbf{f} \: d\mathbf{x} \\
+	&= \mathbf{v}^T \: \langle \bigPii, {\bigPii}^T \rangle_{\Wii} \: d\mathbf{x} \\
 	&= \mathbf{v}^T \Lambda^{(1,1)} \: \mathbf{f},
@@
 where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bigPii$ OP basis.
-We can also not that naturally $\langle \bigPii, {\bigPii}^T \rangle_{\Wii}$ is just a diagonal matrix.
 
-Since (63) is equivalent to stating that
+Since (73) is equivalent to stating that
 \begin{align}
@@
 
-\section{Disk Slices}
+\section{Disk slices}
 
@@
 \end{align}
-with $\alpha, \beta \in [0,1]$, $\rho(x) = \sqrt{1-x^2}$, and  $-\gamma = \delta  = $
+with 
+\begin{align}
+\begin{cases}
+-1 \le \alpha < &\beta \le 1 \\
+(\gamma, \delta) &:= (-1,1) \\
+\rho(x) &:= (1-x^2)^{\half} \\
+w^{(a,b,c)}_1(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c} = (\beta - x)^a \: (x - \alpha)^{b} \: (1-x^2)^{c} \\
+w^{(b)}_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b.
+\end{cases}
+\end{align}
+The weight $w^{(b)}_2(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(b, b)}\}$.
+$w^{(a,b,c)}_1(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a, b,c)}\}\).
+Thus we arrive at the two parameter family of 2D orthogonal polynomials $\{\Pnk\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\begin{align}
+	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(c, a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+\end{align}
+orthogonal with respect to the weight
+\begin{align}
+	W^{(a,b,c)}(x,y) &:= w^{(c,a,b)}_1(x) w^{(b)}_2(\frac{y}{\rho(x)}) \nonumber \\
+	&= (\beta - x)^c \: (x - \alpha)^{a} \: (\rho(x)^2 -y^2)^b \nonumber \\
+	&= (\beta - x)^c \: (x - \alpha)^{a} \: (1 - x^2 -y^2)^b , \quad (x,y) \in \Omega.
+\end{align}
+
+Note that by setting $c = 0$, $\alpha = 0$ and $\beta = 1$ we obtain the 2D orthogonal polynomials on the half disk.
+
+\section{Hemisphere}
+
+We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the hemisphere.
 

### 2019-02-27 papers/halfdisk/halfdisk.tex
Further edits to paper

@@
 \newcommand{\Pnk}{P_{n,k}}
+\newcommand{\Pab}{P^{(a,b)}}
 \newcommand{\Pnkab}{P_{n,k}^{(a,b)}}
@@
 \newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
+\newcommand{\bigQ}{\mathbb{Q}}
+\newcommand{\bigQab}{\bigQ^{(a,b)}}
 
@@
 In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a rhombus or disk slice for example.
-For the purposes of this paper, we limit this to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/rhombus domain.
+
+\subsection{Single element p-FEM}
+In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/rhombus domain.
 We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h|_\element) \le p\}$ for some $p \in \N$.
@@
 
+\subsection{Two-element p-FEM on the disk (without BCs)}
+We now consider the case of using two elements to discretise a unit disk domain $\Omega$, that is the each element $\element$ is a half-disk, with common edge defined by the line $x=0$ for $y \in [-1,1]$.
+To ensure continuity across the edge interface, we need to expand our basis functions to be used.
+We require basis functions that correspond to each edge (in this case there is only one edge), that is basis functions with support as the two elements connected by said edge (for more than two elements - and hence more than one edge - we would also stipulate that each edge's corresponding basis functions would vanish on any other edge).
+We then also require the remaining (interior) basis functions be such that they have support of one element only, and vanish on (each) edge.
+Using this, our basis functions for $\le N$-degree polynomials can then be written as
+\begin{align}
+	\bigQab_n(x,y) := \begin{bmatrix}
+		P^{(a,b)}_{n,n}(x,y) \\
+		x \: \bigPab_{n-1}(x,y)
+	\end{bmatrix} \in \R^{n+1}, 
+	\quad \quad 
+	\bigQab := \begin{bmatrix}
+		\bigQab_0 \\
+		\bigQab_1 \\
+		\vdots \\
+		\bigQab_N
+	\end{bmatrix}.
+\end{align}
+We note that for each $n = 0,\dots,N$, $P^{(a,b)}_{n,n}(x,y) = \rho(x)^n \: P^{(a,b)}_{n}(\frac{y}{\rho(x)})$ and so at $x=0$ these become just the Jacobi polynomials in $y$, thus span the edge interface.
+The polynomials that make up the $\bigQab$ basis can also be shown to have similar sparse relations and yield similar sparse operator matrices.
+We can then proceed to use a similar approach to section A.1 on a reference element.
+
 \section{Disk slices}
@@
 We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the hemisphere.
+Define $\rho(x) := (1-x^2)^\half$ for $x \in [0,1]$ and let the half disk be denoted by $\omega$, that is let
+\begin{align}
+	\omega := \{(x,y) \in \R^2 \quad | \quad 1 < x < 0, \: \rho(x) < y < \rho(x)\}.
+\end{align}
+Further, define the hemisphere by
+\begin{align}
+	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
+\end{align}
+We can define the 3D OP family on the hemisphere as the set of polynomials $\bigQab(x,y,z)$ where
+\begin{align}
+	\bigQab_n(x,y,z) := \begin{bmatrix}
+		\bigPab_n(x,y) \\
+		z \: \bigP^{(a,b+1)}_{n-1}(x,y)
+	\end{bmatrix} \in \R^{2n+1}, 
+	\quad \quad 
+	\bigQab := \begin{bmatrix}
+		\bigQab_0 \\
+		\bigQab_1 \\
+		\vdots \\
+		\bigQab_N
+	\end{bmatrix}.
+\end{align}
+The relations for multiplication by $x, y$ are then trivial.
+For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
+\begin{align}
+	z \: \Pnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: P^{(a,b+1)}_{m,j}(x,y), \\
+	z^2 \: \Pnkab(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b)\to(a,b-1)} \: z \: P^{(a,b-1)}_{m,j}(x,y).
+\end{align}
+A simple calculation shows that the expansions are sparse, that is that most coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are $0$, and are also given by
+\begin{align}
+	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \Pnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
+	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \\
+	&= \frac{\norm{\Pnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)} .
+\end{align}
+The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are just the entries to the parameter tranformation operator matrices
+\begin{align}
+	\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
+	W^{(0,1)}(x,y) \bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)}_ \Big)^T \: \bigP^{(a,b-1)}(x,y)
+\end{align}
+respectively.
 

### 2019-02-27 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
New parameter transformation operators for half disk (useful for hemisphere jacobi operators entries)

@@ -950,7 +950,8 @@ function transformparamsoperator(S::HalfDiskSpace{<:Any, <:Any, T},
         end
         C
     # Case 2
-    elseif Int(S.a) == 1 && Int(S.b) == 1 && weightedfrom == true && weightedto == false
+elseif (Int(St.a) == 0 && Int(St.b) == 0 && Int(S.a) == 1 && Int(S.b) == 1
+            && weightedfrom == true && weightedto == false)
         # NOTE: Only works for W11 -> P00
         # Outputs the sum(1:N+4) × sum(1:N+1) matrix operator
         Sinner = (S.family)(2S.a-1, 2S.b-1)
@@ -1035,6 +1036,52 @@ function transformparamsoperator(S::HalfDiskSpace{<:Any, <:Any, T},
             j = getopindex(n, k)
             for nn = max(0, n-3):n
                 for kk = (k < 2 ? k : k-2):2:k
+                    kk > nn && continue
+                    jj = getopindex(nn, kk)
+                    val = (inner2(S, opevalatpts(S, j, pts), opevalatpts(St, jj, pts), w)
+                            / St.opnorms[jj])
+                    view(C, Block(nn+1, n+1))[kk+1, k+1] = val
+                end
+            end
+        end
+        C
+    elseif St.a == S.a && St.b == S.b+1 && weightedfrom == false && weightedto == false
+        # Takes the space P^{a,b} -> P^{a,b+1}
+        # Outputs the sum(1:N+1) × sum(1:N+1) matrix operator
+        m = 2N^2 - 1 # TODO: check how many points are required for all points() calls
+        pts, w = pointswithweights(St, m)
+        getopptseval(St, N-2, pts)
+        getopptseval(S, N+1, pts)
+        C = BandedBlockBandedMatrix(
+                Zeros{T}(sum(1:(N+1)),sum(1:(N+1))), (1:N+1, 1:N+1), (0,2), (0,2))
+        for n = 0:N, k = 0:n
+            j = getopindex(n, k)
+            for nn = max(0, n-2):n
+                for kk = (k < 2 ? k : k-2):2:k
+                    kk > nn && continue
+                    jj = getopindex(nn, kk)
+                    val = (inner2(S, opevalatpts(S, j, pts), opevalatpts(St, jj, pts), w)
+                            / St.opnorms[jj])
+                    view(C, Block(nn+1, n+1))[kk+1, k+1] = val
+                end
[…]

### 2019-02-27 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -222,8 +222,9 @@ end
     end
 end
 
-@testset "Increment parameters operator" begin
+@testset "Increment/decrement parameters operators" begin
     tol = 1e-8
+
     a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a-1, b-1)
     x, y = 0.4, -0.2; z = [x; y] # Test point
     St = (S.family)(S.a+1, S.b+1)
@@ -250,6 +251,37 @@ end
         resizecoeffs!(p, N)
         q = Fun(St, C * p.coefficients)
         res = abs(q(z) - p(z))
+        res > tol && @show getnk(j), j, res
+        @test res < tol
+    end
+
+    tol = 1e-13
+    
+    a, b = 2.0, 4.0; D = HalfDiskFamily(); S = D(a, b)
+    x, y = 0.4, -0.2; z = [x; y] # Test point
+    St = (S.family)(S.a, S.b+1)
+    maxop = 150
+    N = getnk(maxop)[1] + 1
+    C = transformparamsoperator(S, St, N)
+    for j = 1:maxop
+        p = Fun(S, [zeros(j-1); 1])
+        resizecoeffs!(p, N)
+        q = Fun(St, C * p.coefficients)
+        res = abs(q(z) - p(z))
+        res > tol && @show getnk(j), j, res
+        @test res < tol
+    end
+
+    a, b = 2.0, 4.0; D = HalfDiskFamily(); S = D(a, b)
+    x, y = 0.4, -0.2; z = [x; y] # Test point
+    St = (S.family)(S.a, S.b-1)
+    maxop = 150
+    N = getnk(maxop)[1] + 1
+    C = transformparamsoperator(S, St, N, weightedfrom=true)
+    for j = 1:maxop
+        p = Fun(S, [zeros(j-1); 1])
+        q = Fun(St, C * pad(p.coefficients, size(C)[1]))
+        res = abs(q(z) - (1-z[1]^2-z[2]^2) * p(z))
         res > tol && @show getnk(j), j, res
         @test res < tol
     end

### 2019-02-27 papers/halfdisk/halfdisk.tex
Paper additions
Added reference, and some edits

@@
 $$
-\iint_\Omega f(x,y) \D A =\sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
+\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A = \half \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
 $$
@@
 
-Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) we can approximate (or exactly express) \(f\) by
+Fix \(a,b \in \R\). Then for any function \(f : \Omega \to \R\) of degree $N$ we can express \(f\) by
 \begin{align*}
-f(x,y) \approx \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
+f(x,y) = \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
 \end{align*}
@@
 \end{align}
-and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$. Thus,
+and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$.
+Note that for a general polynomial where the overall degree is $N$, we still would hence need $N^2$ points and weights to make the rule exact.
+Thus,
 \begin{align}
@@
 We now consider the case of using two elements to discretise a unit disk domain $\Omega$, that is the each element $\element$ is a half-disk, with common edge defined by the line $x=0$ for $y \in [-1,1]$.
+We take inspiration from \cite{karniadakis2013spectral, beuchler2006new}.
 To ensure continuity across the edge interface, we need to expand our basis functions to be used.
@@
 	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \Pnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
-	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \\
-	&= \frac{\norm{\Pnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)} .
+	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
+	&= \frac{\ip< \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
+	&= \frac{\norm{\Pnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)}.
 \end{align}
@@
 	\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
-	W^{(0,1)}(x,y) \bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)}_ \Big)^T \: \bigP^{(a,b-1)}(x,y)
+	W^{(0,1)}(x,y) \bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)} \Big)^T \: \bigP^{(a,b-1)}(x,y)
 \end{align}

### 2019-03-04 papers/halfdisk/halfdisk.tex
Some revisions

@@
 \begin{abstract}
-In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
@@
 
-We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{SOATUltraspherical} and its analogue on the triangle \cite{SOATGVTriangle}.
+We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.
 On the half-disk the family of weights we consider are of the form
@@
 $$
-with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$.
-Sparsity comes from expanding the domain and range of an operator  using different polynomials.
-Where the sparsity pattern and entries derived in \cite{SOATGVTriangleRecurrence,SOATGVTriangle}  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
-Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{TriangleFEM}.
+with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$. We define these to be orthogonalised lexicographically, that is,
+$$
+P_{n,k}^{(a,b)}(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
+$$
+where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.
+
+ Sparsity comes from expanding the domain and range of an operator  using different choices of $a$ and $b$.
+Where the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
+Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}.
 
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk\sotodo{Explain this}.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
 
- We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk.
-We will derive a quadrature rule on the half disk that can be used to expand a function in the OP basis up to a given order.
-We will demonstrate that these families will have sparse operators (including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators), and how we can implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
-We can extend this work directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc.
-can be found similarly.
+Here is an overview of the paper:
 
-\section{Orthogonal polynomials on the half disk}
+\noindent \secref{OPs}: We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk.
+
+\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
+
+\noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice.
+We  derive a quadrature rule on the half disk that can be used to expand a function in the OP basis up to a given order.
+We implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
+
+\noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the half-disk.
+
+The motivation for this work is to solve surface PDEs on the sphere.
+In subsequent work we will extend  directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc. can be found similarly.
+
+\section{Orthogonal polynomials on the half disk}\label{Section:OPs}
 
@@
 The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\).
-The weight $w_1(x)$ is non-classical and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
+The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
 Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
- \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+ \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \end{align}
-where we note that \(\gamma_n^{(b)} = 0\) \(\forall n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:
+where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:
 
@@
 \begin{align}
-\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &= \int \int_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
+\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &=  \iint_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
 &= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for some $N \in \No$:
+Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for some $N \in \No$: \sotodo{Why bmatrix, not pmatrix?}
 \begin{align}
@@
 \end{align}
-and set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
+\sotodo{Why not just make $\bigPab$ infinite for now?}and set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
 \begin{align}
@@
 Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs.
-In other words, sparsity is not dependent on the domain.
+In other words, sparsity is not dependent on the domain\sotodo{Huh?}.
 
@@
 	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
-G_n(x,y,z) := \begin{bmatrix}
+G_n(x,y) := \begin{bmatrix}
 		xI_{n+1} \\
@@
 such that \(\Dnt A_n = I_{n+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} (p78), which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
 \begin{align}
@@
 
-Note that we can define \(\Dnt\) as follows. For \(n\) even:
+Note that we can define an explicit \(\Dnt\) as follows. For \(n\) even:
 \begin{align}
@@
 \begin{align}
-m &= \frac{n}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
+m &= \frac{n}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}},\qqand \nonumber \\
 \eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-1.
@@
 \begin{align}
-m &= \frac{n+1}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
-\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-2, \\
-\xi &= -\frac{\betaab_{n,1,5} \: \eta_{m-2}}{\alphaab_{n+1,0,1}}. \nonumber
+m &= \frac{n+1}{2} + 1,\quad  \eta_0 = \frac{1}{\betaab_{n,n,6}}, \quad \xi = -\frac{\betaab_{n,1,5} \: \eta_{m-2}}{\alphaab_{n+1,0,1}},\qqand \nonumber \\
+\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-2.
 \end{align}
+It follows that we can apply $D_n^T$ in $O(n)$ complexity, and thereby construct $\bigPab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
 
 %
-\section{Sparse partial differential operators}
+\section{Sparse partial differential operators}\label{Section:PDOs}
 
@@
 
-\section{Computational aspects}
+\section{Computational aspects}\label{Section:Computation}
 
@@
 %
-\section{Examples on the half disk with zero BCs}
+\section{Examples on the half disk with zero BCs}\label{Section:Examples}
 

### 2019-03-05 papers/halfdisk/halfdisk.tex
Address the points Sheehan made.

@@
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk\sotodo{Explain this}.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family\bstodo{Is this OK?}.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for some $N \in \No$: \sotodo{Why bmatrix, not pmatrix?}
+Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
+Define, for $n=0,1,2,\dots$: \sotodo{Why bmatrix, not pmatrix?}\bstodo{No reason, I am not familiar with pmatrix}
 \begin{align}
@@
 		\bigPab_1 \\
+		\bigPab_2 \\
 		\vdots \\
-		\bigPab_N
 	\end{bmatrix}
 \end{align}
-\sotodo{Why not just make $\bigPab$ infinite for now?}and set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
+\sotodo{Why not just make $\bigPab$ infinite for now?}and set $J_x^{(a,b)}, J_y^{(a,b)}$ as the Jacobi matrices corresponding to
 \begin{align}
-J_x^{(a,b)} \bigPab(x,y) = x \bigPab(x,y), \quad J_y^{(a,b)} \bigPab(x,y) = y \bigPab(x,y).
+J_x^{(a,b)} \: \bigPab(x,y) = x \: \bigPab(x,y), \quad J_y^{(a,b)} \: \bigPab(x,y) = y \: \bigPab(x,y).
 \end{align}
-Then  \(J_x^{(a,b)}, J_y^{(a,b)}\) have the Banded-Block-Banded structures
+Then $J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded structures
 \begin{align}
@@
 		& & \alphaab_{n+1,n,1} & 0 \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
+	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
 B^x_n &:= \begin{bmatrix}
@@
 		& & \alphaab_{n,n,2} \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots,N \\
-C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots,N, \\ 
+	    \end{bmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
+C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
 \nonumber \\
@@
 		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
+	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
 B^y_n &:= \begin{bmatrix}
@@
 		& & \betaab_{n,n,3} & 0
-	    \end{bmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots,N \\
+	    \end{bmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
 C^y_n &:= \begin{bmatrix}
@@
 		& & & \betaab_{n,n,1} \\
-	    \end{bmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots,N.
+	    \end{bmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
 \end{align}
 
-Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs.
-In other words, sparsity is not dependent on the domain\sotodo{Huh?}.
+Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs, meaning that the sparsity is not limited to the specific half-disk case.
 
@@
 \end{align}
-where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each \(n = 0,1,\dots,N\):
+where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
 \begin{align}
@@
 
-For each \(n = 0,1,\dots,N-1\) let \(\Dnt\) be any matrix that is a left inverse of \(A_n\), i.e.
-such that \(\Dnt A_n = I_{n+3}\).
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e.
+such that $\Dnt A_n = I_{n+3}$.
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
 \begin{align}
@@
 \bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
-\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots,N.
+\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots.
 \end{cases}
@@
 
-????
+To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (4), we use the Lanczos algorithm. \bstodo{Not sure if this is required / what else to say here}
 
@@
 
-We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.
+\bstodo{This section contains repetition / expansion on the quadrature rule section above - not sure if we should remove it or move it?} We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.
 

### 2019-03-07 papers/halfdisk/halfdisk.tex
bmatrix -> pmatrix

@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
-Define, for $n=0,1,2,\dots$: \sotodo{Why bmatrix, not pmatrix?}\bstodo{No reason, I am not familiar with pmatrix}
+Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
 \begin{align}
-\bigPab_n := \begin{bmatrix}
+\bigPab_n := \begin{pmatrix}
 		P^{(a,b)}_{n,-n}(x,y) \\
@@
 		P^{(a,b)}_{n,n}(x,y)
-	\end{bmatrix} \in \R^{n+1}, 
+	\end{pmatrix} \in \R^{n+1}, 
 \quad \quad 
-\bigPab := \begin{bmatrix}
+\bigPab := \begin{pmatrix}
 		\bigPab_0 \\
@@
 		\vdots \\
-	\end{bmatrix}
+	\end{pmatrix}
 \end{align}
-\sotodo{Why not just make $\bigPab$ infinite for now?}and set $J_x^{(a,b)}, J_y^{(a,b)}$ as the Jacobi matrices corresponding to
+and set $J_x^{(a,b)}, J_y^{(a,b)}$ as the Jacobi matrices corresponding to
 \begin{align}
@@
 \begin{align}
-J_{x/y}^{(a,b)} &= \begin{bmatrix}
+J_{x/y}^{(a,b)} &= \begin{pmatrix}
 		B^{x/y}_0 & A^{x/y}_0 & & & & \\
@@
 		& & & \ddots & \ddots & \ddots \\
-	\end{bmatrix}
+	\end{pmatrix}
 \end{align}
@@
 \begin{align}
-A^x_n &:= \begin{bmatrix}
+A^x_n &:= \begin{pmatrix}
 		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
@@
 		& & \alphaab_{n+1,n,1} & 0 \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
-B^x_n &:= \begin{bmatrix}
+	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
+B^x_n &:= \begin{pmatrix}
 		\alphaab_{n,0,2} & & \\
@@
 		& & \alphaab_{n,n,2} \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
+	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
 C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
 \nonumber \\
-A^y_n &:= \begin{bmatrix}
+A^y_n &:= \begin{pmatrix}
 		0 & \betaab_{n,0,6} & & & \\
@@
 		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
-	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
-B^y_n &:= \begin{bmatrix}
+	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
+B^y_n &:= \begin{pmatrix}
 		0 & \betaab_{n,0,4} & & \\
@@
 		& & \betaab_{n,n,3} & 0
-	    \end{bmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
-C^y_n &:= \begin{bmatrix}
+	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
+C^y_n &:= \begin{pmatrix}
 		0 & \betaab_{n,0,2} & & \\
@@
 		& & & \betaab_{n,n,1} \\
-	    \end{bmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
+	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
 \end{align}
@@
 =
-\begin{bmatrix}
+\begin{pmatrix}
 	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
-\end{bmatrix}.
+\end{pmatrix}.
 \end{align}
@@
 \begin{align}
-A_n &:= \begin{bmatrix}
+A_n &:= \begin{pmatrix}
 		A^x_n \\
@@
 		A^z_n
-	    \end{bmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
-C_n := \begin{bmatrix}
+	    \end{pmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
+C_n := \begin{pmatrix}
 		C^x_n \\
 		C^y_n
-	    \end{bmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
-B_n &:= \begin{bmatrix}
+	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
+B_n &:= \begin{pmatrix}
 		B^x_n \\
 		B^y_n
-	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
-G_n(x,y) := \begin{bmatrix}
+	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
+G_n(x,y) := \begin{pmatrix}
 		xI_{n+1} \\
 		yI_{n+1}
-	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}.
+	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
 \end{align}
@@
 \begin{align}
-\Dnt := \begin{bmatrix}
+\Dnt := \begin{pmatrix}
 		\frac{1}{\alphaab_{n+1,0,1}} & & &  \\
@@
 		0 & \hdots & 0 & \eta_{m-1} & \hdots & 0 & \eta_1 & 0 & \eta_0
-	    \end{bmatrix},
+	    \end{pmatrix},
 \end{align}
@@
 \begin{align}
-\Dnt := \begin{bmatrix}
+\Dnt := \begin{pmatrix}
 		\frac{1}{\alphaab_{n+1,0,1}} & &  \\
@@
 		\xi & 0 & \hdots & 0 & 0 & \eta_{m-2} & \hdots & 0 & \eta_0
-	    \end{bmatrix},
+	    \end{pmatrix},
 \end{align}
@@
 \begin{align*}
-\bigP_n(x,y) &:= \begin{bmatrix}
+\bigP_n(x,y) &:= \begin{pmatrix}
 		P_{n,0}(x,y) \\
@@
 		P_{n,n}(x,y)
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
+	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
 \end{align*}
@@
 \begin{align*}
-\mathbf{f}_n &:= \begin{bmatrix}
+\mathbf{f}_n &:= \begin{pmatrix}
 		f_{n,0} \\
@@
 		f_{n,n}
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
+	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
 f_{n,k} := \frac{\langle f, \: \Pnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
@@
 \begin{align}
-\bigWii(x,y) &:= \begin{bmatrix}
+\bigWii(x,y) &:= \begin{pmatrix}
 		\bigWii_0(x,y) \\
@@
 		\bigWii_p(x,y)
-	\end{bmatrix}, \\
-\bigWii_n(x,y) &:= \begin{bmatrix}
+	\end{pmatrix}, \\
+\bigWii_n(x,y) &:= \begin{pmatrix}
 		\Wii(x,y) \: \Pii_{n,0}(x,y) \\
@@
 		\Wii(a,y) \: \Pii_{n,n}(x,y)
-	\end{bmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
+	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
 \end{align}
@@
 	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
-	&= \int_\element \begin{bmatrix}
+	&= \int_\element \begin{pmatrix}
 					\partial_x v_h \\
 					\partial_y v_h
-				\end{bmatrix}^T 
-				\begin{bmatrix}
+				\end{pmatrix}^T 
+				\begin{pmatrix}
 					\partial_x u_h \\
 					\partial_y u_h
-				\end{bmatrix}
+				\end{pmatrix}
 				\\
-	&= \int_\element \begin{bmatrix}
+	&= \int_\element \begin{pmatrix}
 					\bigPoo^T \Wii_x \mathbf{v} \\
 					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
-				\end{bmatrix}^T 
-				\begin{bmatrix}
+				\end{pmatrix}^T 
+				\begin{pmatrix}
 					\bigPoo^T \Wii_x \mathbf{u} \\
 					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
-				\end{bmatrix}
+				\end{pmatrix}
 				\\
@@
 \begin{align}
-	\bigQab_n(x,y) := \begin{bmatrix}
+	\bigQab_n(x,y) := \begin{pmatrix}
 		P^{(a,b)}_{n,n}(x,y) \\
 		x \: \bigPab_{n-1}(x,y)
-	\end{bmatrix} \in \R^{n+1}, 
+	\end{pmatrix} \in \R^{n+1}, 
 	\quad \quad 
-	\bigQab := \begin{bmatrix}
+	\bigQab := \begin{pmatrix}
 		\bigQab_0 \\
@@
 		\bigQab_N
-	\end{bmatrix}.
+	\end{pmatrix}.
 \end{align}
@@
 \begin{align}
-	\bigQab_n(x,y,z) := \begin{bmatrix}
+	\bigQab_n(x,y,z) := \begin{pmatrix}
 		\bigPab_n(x,y) \\
 		z \: \bigP^{(a,b+1)}_{n-1}(x,y)
-	\end{bmatrix} \in \R^{2n+1}, 
+	\end{pmatrix} \in \R^{2n+1}, 
 	\quad \quad 
-	\bigQab := \begin{bmatrix}
+	\bigQab := \begin{pmatrix}
 		\bigQab_0 \\

### 2019-03-07 posters/jamboree2019.tex
edits to poster (final)

@@
 
-\graphicspath{ {../papers/halfdisk/images/} }
+% \graphicspath{ {../papers/halfdisk/images/} }
 \graphicspath{ {images/} }
@@
 In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials (OPs) on intervals, disks, and triangles (e.g.
-\cite{olver2018recurrence}).
+\cite{olver2019triangle}).
 Here, we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
-a half disk), building on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate OPs.
+a half disk), building on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical or semi-classical) univariate OPs.
 
@@
 Combining each system in (4) we can write this as a block-tridiagonal system of equations.
-Applying a preconditioner matrix to the system \citep{dunkl2014orthogonal} (p78), we can expand the resulting lower-triangular system to obtain the recurrence:
+Applying a preconditioner matrix to the system \cite[p78]{dunkl2014orthogonal}, we can expand the resulting lower-triangular system to obtain the recurrence:
     \begin{align}
@@
          \centering
-        \includegraphics[scale=0.6]{solution}
+        \includegraphics[scale=0.65]{solution}
         \label{fig:solution}
     \end{subfigure}
-    \begin{subfigure}[t]{0.3\textwidth}
+    \begin{subfigure}[t]{0.6\textwidth}
                 \center
@@
         % Box/square (4x4)
-        \draw[black,solid,ultra thin] (0,0)--(0,4);
-        \draw[black,solid,ultra thin] (0,0)--(4,0);
-        \draw[black,solid,ultra thin] (4,0)--(4,4);
-        \draw[black,solid,ultra thin] (0,4)--(4,4);
+        \draw[black,solid,ultra thin] (0,0)--(0,6);
+        \draw[black,solid,ultra thin] (0,0)--(6,0);
+        \draw[black,solid,ultra thin] (6,0)--(6,6);
+        \draw[black,solid,ultra thin] (0,6)--(6,6);
         % Dots
         \draw[black,fill=black] (0,0) circle (.5ex);
-        \draw[black,fill=black] (2,2) circle (.5ex);
-        \draw[black,fill=black] (4,4) circle (.5ex);
-        \draw[black,fill=black] (0,2) circle (.5ex);
-        \draw[black,fill=black] (0,4) circle (.5ex);
+        \draw[black,fill=black] (3,3) circle (.5ex);
+        \draw[black,fill=black] (6,6) circle (.5ex);
+        \draw[black,fill=black] (0,3) circle (.5ex);
+        \draw[black,fill=black] (0,6) circle (.5ex);
         % Arrows
-        \draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
-        \draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
-        \draw[black,thick,->] (0,0.2)--(0,1.8);
-        \draw[black,thick,->] (0,2.2)--(0,3.8);
-        \draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
+        \draw[black,thick,->] (0.3,0.3)--(2.7,2.7);
+        \draw[black,thick,->] (3.3,3.3)--(5.7,5.7);
+        \draw[black,thick,->] (0,0.5)--(0,2.5);
+        \draw[black,thick,->] (0,3.5)--(0,5.5);
+        \draw [black,thick,->] (0.3,6.3) to [out=20,in=160] (5.7,6.3);
         % Node (parameter) labels
         \draw[] (0,0) node[anchor=east] {$(0,0)$};
-        \draw[] (0,4) node[anchor=east] {$(0,2)$};
-        \draw[] (4,0) node[anchor=west] {$(2,0)$};
-        \draw[] (4,4) node[anchor=west] {$(2,2)$};
+        \draw[] (0,6) node[anchor=east] {$(0,2)$};
+        \draw[] (6,0) node[anchor=west] {$(2,0)$};
+        \draw[] (6,6) node[anchor=west] {$(2,2)$};
         % Arrow labels
-        \draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
-        \draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
-        \draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
-        \draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
+        \draw[] (2.2,2.2) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
+        \draw[] (5,5) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
+        \draw[] (-1.8,1.5) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
+        \draw[] (-1.8,4.5) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
         % The key and headings
-        \draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
-        \draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
-        \draw[] (-2.3,3.2) node[anchor=west] {$a$};
-        \draw[] (-2.7,3.6) node[anchor=south] {$b$};
-        \draw[] (2,4.6) node[anchor=south] {basis conversion};
+        \draw[black,thick,->] (8.5,2.5)--(8.5,3.5);
+        \draw[black,thick,->] (8.5,2.5)--(9.5,2.5);
+        \draw[] (9.5,2.5) node[anchor=west] {$a$};
+        \draw[] (8.5,3.5) node[anchor=south] {$b$};
+        \draw[] (3,6.9) node[anchor=south] {basis conversion};
         \end{tikzpicture} 
@@
 Note that this construction ensures imposition of Dirichlet zero boundary conditions.
-
-The sparsity of the these operators can be shown using a simply integration by parts argument, again showing that the sparsity for differential operators is not dependent on the domain.
+The sparsity of the these operators can be shown using a simply integration by parts argument.
 
@@
 
-     \rightbox{30.5}{
+     \rightbox{32}{
       \section{Summary and Future Work}
@@
 Using the three-term recurrence relationships the OPs have, we can calculate sparse operator matrices for multiplication by $x, y$ and in turn sparse differential operators, allowing us to solve PDEs on the half disk.
+We can extend this to p-FEM on the disk using half-disk elements.
   \item The same (or similar) relations hold for other suitable domains (namely, disk slices). 

### 2019-03-08 papers/halfdisk/halfdisk.tex
Added algorithm

@@
 \newtheorem{definition}{Definition}
+\newtheorem{algorithm}{Algorithm}
 
@@
 
-To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (4), we use the Lanczos algorithm. \bstodo{Not sure if this is required / what else to say here}
+To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (4), we use a version of the Stieltjes algorithm. Recall that
+\begin{align}
+x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
+\end{align}
+where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$.
+
+\begin{algorithm}
+Inputs: max degree $N$, weight $w(x)$
+\begin{flalign*}
+	&\text{Define } H_0(x) \equiv \frac{1}{\norm{1}_w} \\
+	&\text{Define } v(x) = x  H_0(x) \\
+	&\text{Set } \alpha_0 = \ip<v, H_0>_w \\
+	&\text{Define } \tilde{v}(x) = v(x) - \alpha_0 H_0(x) \\
+	&\text{Set } \beta_0 = \norm{\tilde{v}}_w \\
+	&\text{Define } H_1(x) = \frac{\tilde{v}(x)}{\beta_0} \\
+	&\text{FOR } n = 1:N \\
+		& \quad \quad \text{Define } v(x) = x H_n - \beta_{n-1} H_{n-1} \\
+		& \quad \quad \text{Set } \alpha_n = \ip<v, H_n>_w \\
+		& \quad \quad \text{Define } \tilde{v}(x) = v(x) - \alpha_n H_n(x) \\
+		& \quad \quad \text{Set } \beta_n = \norm{\tilde{v}}_w \\
+		& \quad \quad \text{Define } H_{n+1}(x) = \frac{\tilde{v}(x)}{\beta_n} \\
+	&\text{END }
+\end{flalign*}
+Outputs: OP set $(H_0, \dots, H_N)$, coefficients $(\alpha_0, \dots, \alpha_N)$ and $(\beta_0, \dots, \beta_N)$
+\end{algorithm}
+We use the ApproxFun \cite{ApproxFun} package to calculate the inner products and norms in this algorithm to obtain the coefficients.
+
+%    # x * P[n](x) == (?[n] * P[n+1](x) + ?[n] * P[n](x) + ?[n-1] * P[n-1](x))
+%
+%    N = length(?)
+%    # x = Fun(identity, space(w)) # NOTE: space(w) does not "work" sometimes
+%    x = Fun(identity, domain(w))
+%
+%    if N? <= 0
+%        N? = 1
+%        f1 = Fun(1/sqrt(sum(w)),space(x))
+%        P[1] = f1
+%        v = x*P[1]
+%        ?[1] = sum(w*v*P[1])
+%        v = v - ?[1]*P[1]
+%        ?[1] = sqrt(sum(w*v^2))
+%        P[2] = v/?[1]
+%    end
+%
+%    for k = N?+1:N
+%        v = x*P[k] - ?[k-1]*P[k-1]
+%        ?[k] = sum(w*v*P[k])
+%        v = v - ?[k]*P[k]
+%        ?[k] = sqrt(sum(w*v^2))
+%        P[k+1] = v/?[k]
+%    end
+%
+%    P, ?, ?
 

### 2019-03-21 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Changed 3 of 4 of the partial derivative operators to use 1D inner products. This speeds them up. weighted dx to still do.

@@ -31,6 +31,7 @@ struct OrthogonalPolynomialSpace{FA,WW,F,D,B,R,N} <: PolynomialSpace{D,R}
     opnorm::Vector{R} # The norm of the OPs (all OPs of a OPSpace have the same norm).
                         # NOTE this is the value of the norm squared
     opptseval::Vector{Vector{R}}
+    derivopptseval::Vector{Vector{R}}
 end
 
 
@@ -81,7 +82,7 @@ tocanonical(S::OrthogonalPolynomialSpace, x) = x
 OrthogonalPolynomialSpace(fam::SpaceFamily{D,R}, w::Fun, α::NTuple{N,B}) where {D,R,B,N} =
     OrthogonalPolynomialSpace{typeof(fam),typeof(w),Fun,D,B,R,N}(
         fam, w, α, Vector{Fun}(), Vector{R}(), Vector{R}(), Vector{R}(),
-        Vector{Vector{R}}())
+        Vector{Vector{R}}(), Vector{Vector{R}}())
 
 function resizedata!(S::OrthogonalPolynomialSpace, n)
     N₀ = length(S.a) - 1
@@ -275,15 +276,68 @@ function opevalatpts(S::OrthogonalPolynomialSpace{<:Any,<:Any,<:Any,<:Any,<:Any,
 end
 resetopptseval(S::OrthogonalPolynomialSpace) = resize!(S.opptseval, 0)
 
+# Method to gather and evaluate the op derivatives of space S at the transform pts given
+# NOTE: getopptseval() has to be called first
+function getderivopptseval(S::OrthogonalPolynomialSpace, N, pts)
+    resetderivopptseval(S)
+    for n = 0:N
+        derivopevalatpts(S, n+1, pts)
+    end
+    S.derivopptseval
+end
+function derivopevalatpts(S::OrthogonalPolynomialSpace{<:Any,<:Any,<:Any,<:Any,<:Any,T,<:Any}, j, pts) where T
+    # Here, j refers to the index (i.e. deg(p) - 1)
+    N = length(S.derivopptseval) - 1
+    if N ≥ j - 1
+        return S.derivopptseval[j]
+    end
+
+    # We iterate up from the last obtained pts eval
+    if  N != j - 2
+        error("Invalid index")
+    end
+
+    resizedata!(S, j)
[…]

### 2019-03-21 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
changed weighted partial dx operator. Doesnt seem to speed it up. Looking at more efficient ways of doing laplace()

@@ -958,32 +958,60 @@ function partialoperatory(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
     end
     A
 end
+function getweightedpartialoperatorxval(S::HalfDiskSpace{<:Any, <:Any, T},
+                                    ptsp, wp, ptsh, wh, n, k, m, j) where T
+    # We should have already called getopptseval etc for
+    Sx = differentiateweightedspacex(S)
+    P = S.family.P(S.b, S.b)
+    Px = Sx.family.P(Sx.b, Sx.b)
+    H = S.family.H(S.a, S.b+k+0.5)
+    Hx = S.family.H(Sx.a, Sx.b+j+0.5)
+    valp = inner2(Px, (-ptsp.^2 .+ 1) .* opevalatpts(P, k+1, ptsp),
+                    opevalatpts(Px, j+1, ptsp), wp)
+    valh = inner2(H, (ptsh.^2) .* opevalatpts(H, n-k+1, ptsh),
+                    (T.(S.family.ρ.(ptsh)).^(k+j)) .* opevalatpts(Hx, m-j+1, ptsh), wh)
+
+    val = S.a * valp * inner2(H, (T.(S.family.ρ.(ptsh)).^(k+j+2)) .* opevalatpts(H, n-k+1, ptsh),
+                     opevalatpts(Hx, m-j+1, ptsh), wh)
+    val += - 2S.b * valh * inner2(Px, opevalatpts(P, k+1, ptsp),
+                                    opevalatpts(Px, j+1, ptsp), wp)
+    val += valp * inner2(H, ptsh .* derivopevalatpts(H, n-k+1, ptsh),
+                            (T.(S.family.ρ.(ptsh)).^(k+j+2)) .* opevalatpts(Hx, m-j+1, ptsh), wh)
+    val -= k * valh * valp
+    val += valh * inner2(Px, ptsp .* derivopevalatpts(P, k+1, ptsp),
+                            (-ptsp.^2 .+ 1) .* opevalatpts(Px, j+1, ptsp), wp)
+    val /= Sx.opnorms[getopindex(m, j)]
+    val
+end
 function weightedpartialoperatorx(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
     # Takes weighted space ∂/∂x(W^{a,b}) -> W^{a-1,b-1}
     W = BandedBlockBandedMatrix(
[…]

### 2019-03-22 posters/jamboree2019.tex
Added reference to ApproxFun
Edits to halfdisk poster

@@
 \usepackage{natbib}
+\setlength{\bibsep}{0.0pt}
 \usepackage{url}
@@
 \title{Sparse Spectral Methods on Disk Slices}
-\author{Ben Snowball (Supervisors: Sheehan Olver (Main), Colin Cotter)}
+\author{Ben Snowball, Sheehan Olver}
 \department{Department of Mathematics, MPE CDT}
@@
   \leftbox{96}{
-        \section{Abstract}
+        \section{Introduction}
 
-In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials (OPs) on intervals, disks, and triangles (e.g.
+In recent years, \textbf{sparse spectral methods} have been derived using hierarchies of classical \textbf{orthogonal polynomials} (OPs) on intervals, disks, and triangles (e.g.
 \cite{olver2019triangle}).
-Here, we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
-a half disk), building on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical or semi-classical) univariate OPs.
+We extend this methodology to a hierarchy of non-classical orthogonal polynomials on \textbf{disk slices} (e.g.
+a half disk).
+Sparsity is guaranteed due to the boundary being defined by an algebraic curve, and the entries of partial differential operators can be determined using formulae in terms of (non/semi-classical) univariate OPs.
+This work will lead to investigating p-FEM using disk-slice elements, and further to solving PDEs on the 3D hemisphere.
 
@@
 
-  \leftbox{81}{
+  \leftbox{79.5}{
       \section{Orthogonal polynomials on the half disk}
@@
 The weight $w_1(x)$ is non-classical and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
-Thus we arrive at a two parameter family of 2D OPs \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+Thus we arrive at a \textbf{two parameter family of 2D OPs} \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
@@
 
-    \leftbox{37}{
+    \leftbox{35.5}{
     	\section{Jacobi matrices and building the OPs}
@@
 		\bigPab_0 \\
-		\bigPab_1 \\
 		\vdots \\
@@
 \begin{lemma}
-The three-term recurrences that $\Pnk^{(a,b)}(x,y)$ satisfies lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
-Set $J_x^{(a,b)}, J_y^{(a,b)} \in \R^{\half (N+1)(N+2) \times \half (N+1)(N+2)}$ as the Jacobi matrices corresponding to
+The three-term recurrences that $\Pnk^{(a,b)}(x,y)$ satisfies lead to Jacobi operators $J_x^{(a,b)}, J_y^{(a,b)} \in \R^{\half (N+1)(N+2) \times \half (N+1)(N+2)}$ that correspond to multiplication by $x$ and $y$, i.e.
 \begin{align}
@@
 
-For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that (as seen in the DLMF \cite{DLMFDerivatives} for other OP families):
+For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist \textbf{sparse operators} \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that (as seen in the DLMF \cite{DLMFDerivatives} for other OP families):
 \begin{align*}
@@
     \centering
-    \captionsetup{width=.7\linewidth}
+    \captionsetup{width=.8\linewidth}
     \caption{Left: A "spy" plot of the $\Delta$ operator for $\Wii(x,y) \: \bigPii(x,y) \to \bigPii(x,y)$ showing its sparsity.
-Right: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 ?
-10((x ?
-0.5)^2 + y^2)))$.
+Right: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+Code using ApproxFun \cite{ApproxFun}.
 Bottom: Diagram illustrating how the parameters change under partial differentiation.
@@
 
-where $\Delta^{(1,1)}$ is the matrix operator for the Laplacian, that will take us from coefficients for expansion in the weighted space $\Wii(x,y) \: \bigPii(x,y)$ to coefficients in the non-weighted space $\bigPii(x,y)$.
-This operator will have Banded-Block-Banded structure, and hence sparse.
+where $\Delta^{(1,1)}$ is the matrix operator for the \textbf{Laplacian}, that will take us from coefficients for expansion in the weighted space $\Wii(x,y) \: \bigPii(x,y)$ to coefficients in the non-weighted space $\bigPii(x,y)$.
+This operator will have \textbf{Banded-Block-Banded} structure, and hence is \textbf{sparse}.
 Note that this construction ensures imposition of Dirichlet zero boundary conditions.
-The sparsity of the these operators can be shown using a simply integration by parts argument.
+The sparsity of the these operators can be shown using a simple integration by parts argument.
 
@@
 
-     \rightbox{32}{
+     \rightbox{31.5}{
       \section{Summary and Future Work}
@@
   \item This 2D half disk framework can be extended for a family of OPs on the hemisphere, with sparse relations.
-   \item To solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas.
+   \item The aim of solving PDEs on spherical triangles will require further ideas.
 \end{itemize}
@@
 %Also do graphics.
-    \rightbox{11}{
+    \rightbox{11.2}{
         {\footnotesize\bibliography{../papers/halfdisk/halfdisk-bib}}

### 2019-03-22 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Further edits to dx operators

@@ -864,19 +864,19 @@ end
 
 function getpartialoperatorxval(S::HalfDiskSpace{<:Any, <:Any, T},
                                     ptsp, wp, ptsh, wh, n, k, m, j) where T
+    # NOTE: ρ(x) is explicitly assumed to be sqrt(1-x^2), as calling ρ(x) is too
+    # expensive
+    # We should have already called getopptseval etc
     Sx = differentiatespacex(S)
     P = S.family.P(S.b, S.b)
     Px = Sx.family.P(Sx.b, Sx.b)
     H = S.family.H(S.a, S.b+k+0.5)
     Hx = S.family.H(Sx.a, Sx.b+j+0.5)
-    getopptseval(H, n-k, ptsh)
-    getderivopptseval(H, n-k, ptsh)
-    getopptseval(Hx, m-j, ptsh)
     valp = inner2(Px, opevalatpts(P, k+1, ptsp), opevalatpts(Px, j+1, ptsp), wp)
     valh = inner2(H, (ptsh.^2) .* opevalatpts(H, n-k+1, ptsh),
-                    (T.(S.family.ρ.(ptsh)).^(k+j)) .* opevalatpts(Hx, m-j+1, ptsh), wh)
+                    (-ptsh.^2 .+ 1).^(0.5k + 0.5j) .* opevalatpts(Hx, m-j+1, ptsh), wh)
     val = valp * inner2(H, ptsh .* derivopevalatpts(H, n-k+1, ptsh),
-                        (T.(S.family.ρ.(ptsh)).^(k+j+2)) .* opevalatpts(Hx, m-j+1, ptsh), wh)
+                        (-ptsh.^2 .+ 1).^(0.5k + 0.5j + 1) .* opevalatpts(Hx, m-j+1, ptsh), wh)
     val -= k * valh * valp
     val += valh * inner2(Px, ptsp .* derivopevalatpts(P, k+1, ptsp),
                             opevalatpts(Px, j+1, ptsp), wp)
@@ -899,16 +899,25 @@ function partialoperatorx(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
     A = BandedBlockBandedMatrix(
         Zeros{T}(sum(1:N),sum(1:(N+1))), (1:N, 1:N+1), (-1,2), (0, 2))
 
+    # Get pt evals for the H OPs
+    for k = 0:N
+        H = S.family.H(S.a, S.b+k+0.5)
+        getopptseval(H, N-k, ptsh)
[…]

### 2019-03-22 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Converted 2d inner products to 1d for the transform params operators. need to finish the missing operators. made them work now for general params a, b.

@@ -1056,189 +1056,118 @@ function weightedpartialoperatory(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
 end
 
 function transformparamsoperator(S::HalfDiskSpace{<:Any, <:Any, T},
-            St::HalfDiskSpace{<:Any, <:Any, T}, N;
-            weightedfrom=false, weightedto=false
-            ) where T
-
+            St::HalfDiskSpace{<:Any, <:Any, T}, N; weighted=false) where T
     # Cases we can handle:
-    # Case 1
-    if Int(S.a) == 1 && Int(S.b) == 0 && weightedfrom == true && weightedto == false
+    # Case 1: Takes the space P^{a,b} -> P^{a+1,b}
+    if weighted == false && St.a == S.a+1 && St.b == S.b
+        # Outputs the relevant sum(1:N+1) × sum(1:N+1) matrix operator
+        C = BandedBlockBandedMatrix(Zeros{T}(sum(1:(N+1)),sum(1:(N+1))),
+                                    (1:N+1, 1:N+1), (0,1), (0,0))
+
+        P = S.family.P(S.b, S.b)
+        ptsh, wh = pointswithweights(St.family.H(St.a, St.b + 0.5), N+1)
+        getopnorms(St, sum(1:N+1))
+
+        # Get pt evals for H OPs
+        for k = 0:N
+            H = S.family.H(S.a, S.b+k+0.5)
+            Ht = St.family.H(St.a, St.b+k+0.5)
+            getopptseval(H, N-k, ptsh)
+            getopptseval(Ht, N-k, ptsh)
+        end
+
+        n, k = 0, 0; m = n
+        view(C, Block(m+1, n+1))[k+1, k+1] = sum(wh) * getopnorm(P) / St.opnorms[getopindex(m, k)]
+        for n=1:N, k=0:n
+            H = S.family.H(S.a, S.b+k+0.5)
+            Ht = St.family.H(St.a, St.b+k+0.5)
+            M = (k == n ? [n] : [n-1, n])
+            for m in M
+                val = inner2(H, opevalatpts(H, n-k+1, ptsh),
+                                (-ptsh.^2 .+ 1).^k .* opevalatpts(Ht, m-k+1, ptsh), wh)
+                val *= getopnorm(P)
+                view(C, Block(m+1, n+1))[k+1, k+1] = val / St.opnorms[getopindex(m, k)]
+            end
+        end
+        C
+    # Case 2: Takes the space W^{a,b} -> W^{a-1,b}
[…]

### 2019-03-22 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -209,7 +209,7 @@ end
     St = (S.family)(S.a-1, S.b-1)
     maxop = 100
     N = getnk(maxop)[1] + 3
-    W = transformparamsoperator(S, St, N, weightedfrom=true)
+    W = transformparamsoperator(S, St, N, weighted=true)
     for j = 1:maxop
         p = Fun(S, [zeros(j-1); 1])
         resizecoeffs!(p, N)
@@ -256,7 +256,7 @@ end
     end
 
     tol = 1e-13
-    
+
     a, b = 2.0, 4.0; D = HalfDiskFamily(); S = D(a, b)
     x, y = 0.4, -0.2; z = [x; y] # Test point
     St = (S.family)(S.a, S.b+1)
@@ -277,7 +277,7 @@ end
     St = (S.family)(S.a, S.b-1)
     maxop = 150
     N = getnk(maxop)[1] + 1
-    C = transformparamsoperator(S, St, N, weightedfrom=true)
+    C = transformparamsoperator(S, St, N, weighted=true)
     for j = 1:maxop
         p = Fun(S, [zeros(j-1); 1])
         q = Fun(St, C * pad(p.coefficients, size(C)[1]))

### 2019-03-25 papers/halfdisk/halfdisk.tex
Edits to paper, notably the operators section

@@
 \newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
+\newcommand{\pddx}{\frac{\partial}{\partial x}}
+\newcommand{\pddy}{\frac{\partial}{\partial y}}
 \newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}
@@
 
-\title{Sparse spectral and $p$-finite element methods on disk-slices and rhombi I: zero boundary conditions}
+\title{Sparse spectral and $p$-finite element methods on disk-slices and trapeziums I: zero boundary conditions}
 \author{Sheehan Olver, Ben Snowball}
@@
 In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
-a half disk) and rhombi.
+a half disk) and trapeziums.
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
@@
 
-This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and rhombi.  
+This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
 More precisely, let  \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
@@
 
-For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
+Let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
+Recall that $\Wab(x,y) := w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $w_H^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
+
+There exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
 \begin{align}
@@
 Further, we note that there exist "parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
-We denote these operators as $T_W^{(1,0)\to(0,0)}$, $T_W^{(1,1)\to(0,0)}$, $T^{(0,1)\to(1,1)}$, $T^{(0,2)\to(2,2)}$ and the general $T^{(a,b)\to(a+1,b+1)}$ where:
-\begin{align}
-\bigW^{(1,0)}(x,y) &= \Big(T_W^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
-\bigW^{(1,1)}(x,y) &= \Big(T_W^{(1,1)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
-\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y) \\
-\bigP^{(0,2)}(x,y) &= \Big(T^{(0,2)\to(2,2)} \Big)^T \: \bigP^{(2,2)}(x,y) \\
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigP^{(a+1,b+1)}(x,y)
+We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T^{(a,b)\to(a-1,b)}$, $T^{(a,b)\to(a,b-1)}$ and $T^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
+\begin{align}
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^T \: \bigP^{(a+1,b)}(x,y) \\
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigP^{(a+1,b+1)}(x,y) \\
+\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^T \: \bigW^{(a-1,b)}(x,y) \\
+\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^T \: \bigW^{(a,b-1)}(x,y) \\
+\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigW^{(a-1,b-1)}(x,y) \\
 \end{align}
@@
 \begin{align}
-    D := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(0,2)\to(2,2)} \: D_y^{(0,1)} \: D_y^{(0,0)}.
-\end{align}
-
-The sparsity of the these operators can be shown using a simply integration by parts argument, again showing that the sparsity for differential operators is not dependent on the domain.
+    \Delta := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(1,2)\to(2,2)} \: T^{(0,2)\to(1,2)}  \: D_y^{(0,1)} \: D_y^{(0,0)}.
+\end{align}
+
+The sparsity of the these operators can be shown using a simple integration by parts argument. We note that the non-zero entries of the differential operators are calculated as follows.
+
+Note that we can expand $\pddx \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
+The coefficients of these expansion are then the entries of the relevant operator matrix.
+By a simple integration by parts argument, we have that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
+These are then calculated by taking inner products, i.e.
+\begin{align*}
+c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}
+\end{align*}
+where 
+\begin{align}
+	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
+	&= \ip< x \: H_{n-k}^{(a, b+k+\half)} \prime (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
+	& \quad - k \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
+	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+		\: \ip< y \: P_k^{(b,b)} \prime (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
+\end{align}
+We can then use quadrature to calculate the 1D inner products. This is much cheaper than using the 2D quadrature rule.
+
+Further, we can expand $\pddy \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+The coefficients of these expansion are then the entries of the relevant operator matrix.
+By a simple integration by parts argument, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
+This is then calculated by taking inner products, i.e.
+\begin{align*}
+c_{n-1,k-1} = \frac{\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
+\end{align*}
+where 
+\begin{align}
+	\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
+		\: \ip< P_k^{(b,b)} \prime, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
+\end{align}
+where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ $\forall m, n \in \N_0$.
+
+We can gain the non-zero entries of the weighted differential operators similarly, by noting that
+\begin{align}
+	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \\
+	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y)
+\end{align}
+and also that
+\begin{align}
+	\ip< \Wab \Pnkab, \Wab \Pab_{m,j} >_{W^{(-a, -b)}} = \ip< \Pnkab, \Pab_{m,j} >_\Wab = \delta_{n,m} \: \delta_{k,j}.
+\end{align}
+
+Calculating the non-zero entries of the parameter transformation operators is similarly done, i.e. use a substitution in the 2D inner product integral to yield a product of 1D inner product integrals.
 
@@
 We use the ApproxFun \cite{ApproxFun} package to calculate the inner products and norms in this algorithm to obtain the coefficients.
-
-%    # x * P[n](x) == (?[n] * P[n+1](x) + ?[n] * P[n](x) + ?[n-1] * P[n-1](x))
-%
-%    N = length(?)
-%    # x = Fun(identity, space(w)) # NOTE: space(w) does not "work" sometimes
-%    x = Fun(identity, domain(w))
-%
-%    if N? <= 0
-%        N? = 1
-%        f1 = Fun(1/sqrt(sum(w)),space(x))
-%        P[1] = f1
-%        v = x*P[1]
-%        ?[1] = sum(w*v*P[1])
-%        v = v - ?[1]*P[1]
-%        ?[1] = sqrt(sum(w*v^2))
-%        P[2] = v/?[1]
-%    end
-%
-%    for k = N?+1:N
-%        v = x*P[k] - ?[k-1]*P[k-1]
-%        ?[k] = sum(w*v*P[k])
-%        v = v - ?[k]*P[k]
-%        ?[k] = sqrt(sum(w*v^2))
-%        P[k+1] = v/?[k]
-%    end
-%
-%    P, ?, ?
 
@@
 
-In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a rhombus or disk slice for example.
+In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a trapezium or disk slice for example.
 
 \subsection{Single element p-FEM}
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/rhombus domain.
+In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/trapezium domain.
 We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h|_\element) \le p\}$ for some $p \in \N$.

### 2019-03-26 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Fix in HalfDiskTransformPlan

@@ -520,10 +520,10 @@ function HalfDiskTransformPlan(S::HalfDiskSpace{<:Any, <:Any, T}, vals) where T
     # Vandermonde matrices transposed, for each set of pts (x, ±y)
     VTp = Array{Float64}(undef, m, n)
     VTm = copy(VTp)
+    getopptseval(S, N, pts)
     for k = 1:m
-        Pnk = Fun(S, [zeros(k-1); 1])
-        VTp[k, :] = Pnk.(pts[1:n])
-        VTm[k, :] = Pnk.(pts[n+1:end])
+        VTp[k, :] = opevalatpts(S, k, pts)[1:n]
+        VTm[k, :] = opevalatpts(S, k, pts)[n+1:end]
     end
     W = Diagonal(w)
     U = Diagonal(1 ./ S.opnorms[1:m])
@@ -567,7 +567,8 @@ function recα(S::HalfDiskSpace{<:Any, <:Any, T}, n, k, j) where T
     end
 end
 
-# TODO
+# TODO: Speed up clenshaw! Maybe call pointswithweights(H) outide of this once,
+# and just adjust the inner products...
 function recβ(S::HalfDiskSpace{<:Any, <:Any, T}, n, k, j) where T
     # We get the norms of the 2D OPs
     m = Int((n+2)*(n+3) / 2)
@@ -580,7 +581,6 @@ function recβ(S::HalfDiskSpace{<:Any, <:Any, T}, n, k, j) where T
     H3 = (S.family.H)(S.a, S.b + k + 1.5)
     P = (S.family.P)(S.b, S.b)
     getopnorm(P)
-    δ1 = recβ(T, P, k+1) * P.opnorm[1]
 
     if isodd(j)
         pts, w = pointswithweights(H2, Int(ceil(n-k+1.5)))

### 2019-03-27 papers/halfdisk/halfdisk.tex
Further edits to paper. Included updated figures

@@
 \begin{align}
-\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}.
 \end{align}
@@
 $$
-with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$. We define these to be orthogonalised lexicographically, that is,
+with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
 $$
@@
 
-The motivation for this work is to solve surface PDEs on the sphere.
-In subsequent work we will extend  directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc. can be found similarly.
-
 \section{Orthogonal polynomials on the half disk}\label{Section:OPs}
 
-In this section we outline some basic properties of $P_{n,k}^{(a,b)}(x,y)$.
+In this section we outline the construction and some basic properties of $P_{n,k}^{(a,b)}(x,y)$.
 The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.
@@
 \begin{proposition}
-Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\),
-Then, \(\forall n,k = 0,1,2,\dots\) let \(\{p_{n,k}\}\) be polynomials orthogonal with respect to the weight \(\rho(x)^{2k+1} w_1(x)\), and \(\{q_{n}\}\) be polynomials orthogonal with respect to the weight \(w_2(x)\).
-Then
+Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
+\begin{enumerate}
+\item  \(\rho\) is a degree 1 polynomial, or 
+\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
+\end{enumerate}
+$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$.
+Then the 2D polynomials defined on $\Omega$
 $$
@@
 \end{proposition}
+\begin{proof}
+See \cite[p55--56]{dunkl2014orthogonal}.
+\end{proof}
 
 For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
-
 \begin{align}
@@
 \rho(x) &:= (1-x^2)^{\half} \\
-w_1(x) &:= x^a \: (1-x^2)^b \\
-w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b,
+w_1(x) &:= w_H^{(a,b)}(x) := x^a \: (1-x^2)^b \\
+w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
 \end{cases}
@@
 The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
-Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
@@
 \end{align}
-orthogonal on \(\Omega\) with respect to the weight $W^{(a,b)}(x,y)$.
+orthogonal on \(\Omega\) with respect to the weight
+\begin{align}
+W^{(a,b)}(x,y) = w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
+\end{align}
 
@@
 
-Express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
+We can express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
 \begin{align}
@@
 \alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
-\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
-\end{align}
-where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight.
+\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
+\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \\
+\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
+\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \\
+\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
+\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}}. 
+\end{align}
+where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{w_H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight of $w_H^{(a, b)}$.
 
@@
 &= \begin{cases}
-    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
-	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
+    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
+	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
 	0 & \text{otherwise}
@@
 Further, we note that there exist "parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
-We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T^{(a,b)\to(a-1,b)}$, $T^{(a,b)\to(a,b-1)}$ and $T^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
+We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
 \begin{align}
@@
 \end{align}
-If $f(x,y)$ is an at most degree $N$ polynomial in $x$ and at most degree $2N-1$ in $y$ then
+Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
 $$
-\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A = \half \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
+\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A \approx \half \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
 $$
-
+where if $f(\cdot, y)$ for fixed $y$ is an at most degree $N$ polynomial and $f(x, \cdot)$ for fixed $x$  is an at most degree $2N-1$ polynomial, the quadrature rule is exact.
 \end{theorem}
+
 \begin{proof}
-
 We will use the substitution that
@@
 \end{align}
-
 \end{proof}
@@
 
-\bstodo{This section contains repetition / expansion on the quadrature rule section above - not sure if we should remove it or move it?} We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.
-
-Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$ (i.e.
+Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$:
+\begin{align}
+\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
+|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
+\end{align}
+
+%\bstodo{This section contains repetition / expansion on the quadrature rule section above - not sure if we should remove it or move it?} We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.
+%
+%Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$ (i.e.
 $f(\cdot, y)$ with $y$ fixed is a degree $d_1$ polynomial, and $f(x, \cdot)$ with $x$ fixed is an even degree $d_2$ polynomial).
@@
 Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
-\begin{align}
-& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
-&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
-&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
-&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
-&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
-\end{align}
-where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
-
-We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
-\begin{align}
-\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
-\end{align}
-
-Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
-\begin{align}
-\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
-\end{align}
-and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$.
+%\begin{align}
+%& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
+%&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
+%&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
+%&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
+%&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
+%\end{align}
+%where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
+%
+%We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
+%\begin{align}
+%\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
+%\end{align}
+%
+%
+%Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
+%\begin{align}
+%\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
+%\end{align}
+%and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$.
 Note that for a general polynomial where the overall degree is $N$, we still would hence need $N^2$ points and weights to make the rule exact.
 Thus,
-\begin{align}
-\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
-|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
-\end{align}
-for suitably large $N$.
+%\begin{align}
+%\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
+%|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
+%\end{align}
+%for suitably large $N$.
 
@@
 %
-\section{Future work - orthogonal polynomials on the hemisphere and spherical triangles}
+\section{Future work - orthogonal polynomials on the half-spheres and spherical triangles}
 
@@
 partial derivative).
-With this, we can obtain a family of OPs on the hemisphere, with sparse relations.
-
-Beyond this, solving PDEs on isosceles triangles should follow using the same procedure.
-However, we are expecting that to solve PDEs on spherical triangles (as is our ultimate goal) will require further ideas beyond the procedures given in this document.
+With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
+An extension to this will be to look at a $p$-finite element method on the sphere using two half-sphere elements.
+We plan to extend this work to construct similar methods on other 3D spherical domains, such as spherical caps and spherical triangles, which will have applications in numerical weather prediction.
+
+Other extensions include a full $hp$-finite element method on sections of a disk which has applications in fluids.
 
@@
 
-\section{Hemisphere}
-
-We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the hemisphere.
+%
+%\section{Hemisphere}
+%
+%We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the hemisphere.
 Define $\rho(x) := (1-x^2)^\half$ for $x \in [0,1]$ and let the half disk be denoted by $\omega$, that is let
-\begin{align}
-	\omega := \{(x,y) \in \R^2 \quad | \quad 1 < x < 0, \: \rho(x) < y < \rho(x)\}.
-\end{align}
-Further, define the hemisphere by
-\begin{align}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
-\end{align}
-We can define the 3D OP family on the hemisphere as the set of polynomials $\bigQab(x,y,z)$ where
-\begin{align}
-	\bigQab_n(x,y,z) := \begin{pmatrix}
-		\bigPab_n(x,y) \\
-		z \: \bigP^{(a,b+1)}_{n-1}(x,y)
-	\end{pmatrix} \in \R^{2n+1}, 
-	\quad \quad 
-	\bigQab := \begin{pmatrix}
-		\bigQab_0 \\
-		\bigQab_1 \\
-		\vdots \\
-		\bigQab_N
-	\end{pmatrix}.
-\end{align}
-The relations for multiplication by $x, y$ are then trivial.
+%\begin{align}
+%	\omega := \{(x,y) \in \R^2 \quad | \quad 1 < x < 0, \: \rho(x) < y < \rho(x)\}.
+%\end{align}
+%Further, define the hemisphere by
+%\begin{align}
+%	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
+%\end{align}
+%We can define the 3D OP family on the hemisphere as the set of polynomials $\bigQab(x,y,z)$ where
+%\begin{align}
+%	\bigQab_n(x,y,z) := \begin{pmatrix}
+%		\bigPab_n(x,y) \\
+%		z \: \bigP^{(a,b+1)}_{n-1}(x,y)
+%	\end{pmatrix} \in \R^{2n+1}, 
+%	\quad \quad 
+%	\bigQab := \begin{pmatrix}
+%		\bigQab_0 \\
+%		\bigQab_1 \\
+%		\vdots \\
+%		\bigQab_N
+%	\end{pmatrix}.
+%\end{align}
+%The relations for multiplication by $x, y$ are then trivial.
 For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
-\begin{align}
-	z \: \Pnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: P^{(a,b+1)}_{m,j}(x,y), \\
-	z^2 \: \Pnkab(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b)\to(a,b-1)} \: z \: P^{(a,b-1)}_{m,j}(x,y).
-\end{align}
-A simple calculation shows that the expansions are sparse, that is that most coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are $0$, and are also given by
-\begin{align}
-	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \Pnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
-	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
-	&= \frac{\ip< \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
-	&= \frac{\norm{\Pnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)}.
-\end{align}
-The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are just the entries to the parameter tranformation operator matrices
-\begin{align}
-	\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
-	W^{(0,1)}(x,y) \bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)} \Big)^T \: \bigP^{(a,b-1)}(x,y)
-\end{align}
-respectively.
+%\begin{align}
+%	z \: \Pnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: P^{(a,b+1)}_{m,j}(x,y), \\
+%	z^2 \: \Pnkab(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b)\to(a,b-1)} \: z \: P^{(a,b-1)}_{m,j}(x,y).
+%\end{align}
+%A simple calculation shows that the expansions are sparse, that is that most coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are $0$, and are also given by
+%\begin{align}
+%	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \Pnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
+%	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
+%	&= \frac{\ip< \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
+%	&= \frac{\norm{\Pnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)}.
+%\end{align}
+%The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are just the entries to the parameter tranformation operator matrices
+%\begin{align}
+%	\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
+%	W^{(0,1)}(x,y) \bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)} \Big)^T \: \bigP^{(a,b-1)}(x,y)
+%\end{align}
+%respectively.
 

### 2019-03-28 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
resizedata() for OPSpaces N0 changed to stop repeating lanczos iterations. New operatorclenshaw() overload.

@@ -56,6 +56,7 @@ function lanczos!(w, P, β, γ, N₀=0)
     end
 
     for k = N₀+1:N
+        @show k
         v = x*P[k] - γ[k-1]*P[k-1]
         β[k] = sum(w*v*P[k])
         v = v - β[k]*P[k]
@@ -85,8 +86,8 @@ OrthogonalPolynomialSpace(fam::SpaceFamily{D,R}, w::Fun, α::NTuple{N,B}) where 
         Vector{Vector{R}}(), Vector{Vector{R}}())
 
 function resizedata!(S::OrthogonalPolynomialSpace, n)
-    N₀ = length(S.a) - 1
-    n ≤ N₀ + 1 && return S
+    N₀ = length(S.a)
+    n ≤ N₀ && return S
     resize!(S.a, n)
     resize!(S.b, n)
     resize!(S.ops, n + 1)
@@ -1327,6 +1328,7 @@ function operatorclenshaw(cfs, S::HalfDiskSpace)
     (γ1 * P0)[1]
 end
 operatorclenshaw(f::Fun, S::HalfDiskSpace) = operatorclenshaw(f.coefficients, S)
+operatorclenshaw(f::Fun, S::HalfDiskSpace, N) = operatorclenshaw(resizecoeffs!(f, N), S)
 
 
 # Method to gather and evaluate the ops of space S at the transform pts given

### 2019-03-29 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Implemented all parameter transfrom operators we require (increment for non-weighted, decrement for weighted spaces)

@@ -954,6 +954,7 @@ function partialoperatory(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
     P = S.family.P(S.b, S.b)
     Py = Sy.family.P(Sy.b, Sy.b)
     pts, w = pointswithweights(Py, N)
+    getopptseval(P, N, pts)
     getderivopptseval(P, N, pts)
     getopptseval(Py, N, pts)
     getopnorms(Sy, getopindex(N-1, N-1))
@@ -1064,9 +1065,9 @@ function transformparamsoperator(S::HalfDiskSpace{<:Any, <:Any, T},
         # Outputs the relevant sum(1:N+1) × sum(1:N+1) matrix operator
         C = BandedBlockBandedMatrix(Zeros{T}(sum(1:(N+1)),sum(1:(N+1))),
                                     (1:N+1, 1:N+1), (0,1), (0,0))
-
         P = S.family.P(S.b, S.b)
         ptsh, wh = pointswithweights(St.family.H(St.a, St.b + 0.5), N+1)
+        rho2ptsh = -ptsh.^2 .+ 1
         getopnorms(St, sum(1:N+1))
 
         # Get pt evals for H OPs
@@ -1085,18 +1086,52 @@ function transformparamsoperator(S::HalfDiskSpace{<:Any, <:Any, T},
             M = (k == n ? [n] : [n-1, n])
             for m in M
                 val = inner2(H, opevalatpts(H, n-k+1, ptsh),
-                                (-ptsh.^2 .+ 1).^k .* opevalatpts(Ht, m-k+1, ptsh), wh)
+                                rho2ptsh.^k .* opevalatpts(Ht, m-k+1, ptsh), wh)
                 val *= getopnorm(P)
                 view(C, Block(m+1, n+1))[k+1, k+1] = val / St.opnorms[getopindex(m, k)]
             end
         end
         C
-    # Case 2: Takes the space W^{a,b} -> W^{a-1,b}
+    # Case 2: Takes the space P^{a,b} -> P^{a,b+1}
+    # and Case 3: Takes the space P^{a,b} -> P^{a+1,b+1}
+    elseif ((weighted == false && St.a == S.a && St.b == S.b+1)
+            || (weighted == false && St.a == S.a+1 && St.b == S.b+1))
+        # Outputs the relevant sum(1:N+1) × sum(1:N+1) matrix operator
[…]

### 2019-03-29 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -3,7 +3,7 @@ using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature,
 import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule,
         gethalfdiskOP, jacobix, jacobiy, differentiatex, differentiatey,
         resizecoeffs!, laplace, transformparamsoperator, operatorclenshaw, getnk,
-        getopindex
+        getopindex, weight
 
 
 
@@ -154,38 +154,17 @@ end
     # 1) f(x,y) = -8x => u(x,y) ≡ 1
     N = 1 # degree of f
     c = rand(1)[1]; f = Fun((x,y)->-c*8x, S)
-    Δ = laplace(D, N-1)
-    u = Fun(S, Δ \ resizecoeffs!(f, N))
+    Δ = laplacesquare(D, N)
+    u = Fun(S, sparse(Δ) \ resizecoeffs!(f, N))
     @test u(z) ≈ c # Result u(x,y) where Δ(u*w)(x,y) = f(x,y)
 
     # 2) f(x,y) = 2 - 12xy - 14x^2 - 2y^2 => u(x,y) = x + y
     U = Fun((x,y)->x+y, S)
     N = 2 # degree of f
     f = Fun((x,y)->(2 - 12x*y - 14x^2 - 2y^2), S)
-    Δ = laplace(D, N-1)
-    u = Fun(S, Δ \ f.coefficients[1:getopindex(N,N)])
+    Δ = laplacesquare(D, N)
+    u = Fun(S, sparse(Δ) \ f.coefficients[1:getopindex(N,N)])
     @test u(z) ≈ U(z)
-end
-
-@testset "Conversion operators" begin
-    a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
-    x, y = 0.457, -0.209; z = [x; y] # Test point
-    N = 5
-    Sfrom = D(0.0, 2.0); Sto = D(2.0, 2.0)
-    f1 = Fun(Sfrom, rand(sum(1:N+1)))
-    T = transformparamsoperator(Sfrom, Sto, N)
-    f2 = Fun(D(Sto, T * f1.coefficients)
-    @test f2(z) ≈ f1(z)
-    Sfrom = D(0.0, 1.0); Sto = D(1.0, 1.0)
-    f1 = Fun(Sfrom, rand(sum(1:N+1)))
[…]

### 2019-04-01 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Laplacian operators and biharmonic operator

@@ -895,7 +895,9 @@ function partialoperatorx(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
     getopptseval(Px, N, ptsp)
     H = S.family.H(S.a, S.b+0.5)
     ptsh, wh = pointswithweights(H, 2N+2)
+    @show N
     getopnorms(Sx, getopindex(N-1, N-1))
+    @show "done"
 
     A = BandedBlockBandedMatrix(
         Zeros{T}(sum(1:N),sum(1:(N+1))), (1:N, 1:N+1), (-1,2), (0, 2))
@@ -1213,25 +1215,94 @@ function laplace(D::HalfDiskFamily, N)
     A * B + C * E * F * G
 end
 
-function laplacesquare(D::HalfDiskFamily{<:Any, T, <:Any, <:Any}, N) where T
-    # sparse([laplace(D, N) zeros(sum(1:N+2), sum(1:N+2)-sum(1:N+1))])
-
-    # Outputs the sum(1:N+1) × sum(1:N+1) matrix operator
-    Δ = BandedBlockBandedMatrix(
-        Zeros{T}(sum(1:(N+1)),sum(1:(N+1))), (1:N+1, 1:N+1), (1,1), (2,5))
-    L = laplace(D, N+1)
-    i = 1
-    view(Δ, Block(i, i)) .= view(L, Block(i, i))
-    view(Δ, Block(i, i+1)) .= view(L, Block(i, i+1))
-    for i = 2:N
-        view(Δ, Block(i, i-1)) .= view(L, Block(i, i-1))
-        view(Δ, Block(i, i)) .= view(L, Block(i, i))
-        view(Δ, Block(i, i+1)) .= view(L, Block(i, i+1))
+function laplaceoperator(S::HalfDiskSpace{<:Any, <:Any, T},
+            St::HalfDiskSpace{<:Any, <:Any, T}, N;
+            weighted=false, square=true) where T
+    # Outputs the sum(1:N+1) × sum(1:N+1) matrix operator if square=true
+    # TODO: Is it more efficient to allocate blocks manually?
+    D = S.family
+    if (weighted == true && Int(S.a) == 1 && Int(S.b) == 1
+            && Int(St.a) == 1 && Int(St.b) == 1)
+        A = partialoperatorx(D(S.a-1,S.b-1), N+2)
+        @show "A done"
+        B = weightedpartialoperatorx(D(S.a,S.b), N)
[…]

### 2019-04-01 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -145,28 +145,6 @@ end
     h = 1e-5; @test (f(x,y+h)-f(x,y))/h ≈ differentiatey(f, f.space)(x,y) atol=100h
 end
 
-@testset "Poisson" begin
-    # Model Problem: Δ(u*w)(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
-    #   where w(x,y) = x*(1-x^2-y^2) is the weight of the D(1.0,1.0) basis.
-    a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
-    x, y = 0.4, -0.2; z = [x; y] # Test point
-
-    # 1) f(x,y) = -8x => u(x,y) ≡ 1
-    N = 1 # degree of f
-    c = rand(1)[1]; f = Fun((x,y)->-c*8x, S)
-    Δ = laplacesquare(D, N)
-    u = Fun(S, sparse(Δ) \ resizecoeffs!(f, N))
-    @test u(z) ≈ c # Result u(x,y) where Δ(u*w)(x,y) = f(x,y)
-
-    # 2) f(x,y) = 2 - 12xy - 14x^2 - 2y^2 => u(x,y) = x + y
-    U = Fun((x,y)->x+y, S)
-    N = 2 # degree of f
-    f = Fun((x,y)->(2 - 12x*y - 14x^2 - 2y^2), S)
-    Δ = laplacesquare(D, N)
-    u = Fun(S, sparse(Δ) \ f.coefficients[1:getopindex(N,N)])
-    @test u(z) ≈ U(z)
-end
-
 @testset "Operator Clenshaw" begin
     a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
     x, y = 0.4, -0.2; z = [x; y] # Test point
@@ -177,28 +155,6 @@ end
     resizecoeffs!(f, N)
     foper = Fun(S, A*f.coefficients)
     @test foper(z) ≈ f(z) * oper(z)
-end
-
-@testset "Weighted to non-weighted space conversion" begin
-    # NOTE: Operator takes W11*P11 -> P00 only
-    tol = 1e-8
-    a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
-    x, y = 0.4, -0.2; z = [x; y] # Test point
[…]

### 2019-04-01 papers/halfdisk/halfdisk.tex
Additions to paper

@@
 \newcommand{\pddy}{\frac{\partial}{\partial y}}
+\newcommand{\pddn}{\frac{\partial}{\partial n}}
 \newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}
@@
 \newcommand{\dy}{\frac{\partial}{\partial y}}
+\newcommand{\laplacewii}{\Delta_W^{(1,1)\to(1,1)}}
+\newcommand{\laplacewtt}{\Delta_W^{(2,2)\to(0,0)}}
+\newcommand{\laplaceoo}{\Delta^{(0,0)\to(2,2)}}
+\newcommand{\biharmonic}{_2\Delta_W^{(2,2)\to(2,2)}}
+
 \newcommand{\element}{\tau}
@@
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family\bstodo{Is this OK?}.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
 
-We can then obtain the matrix operator for \(\Delta\), that will take us from coefficients for expansion in the weighted space
+We can then obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
 \[
@@
 \]
-Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse.
+Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in FIGURE.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
-The matrix operator \(\Delta\) acting on the coefficients vector is then given by
-\begin{align}
-    \Delta := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
-\end{align}
-
-Further, we can also obtain the matrix operator for \(\Delta\) that will take us from coefficients for expansion in the space
-\[
+The matrix operator for the Laplacian we denote $\laplacewii$ acting on the coefficients vector is then given by
+\begin{align}
+    \laplacewii := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
+\end{align}
+
+Further, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space
+$$
 \bigPoo(x,y)
-\]
+$$
 to coefficients in the space
-\[
+$$
 \bigP^{(2,2)}(x,y).
-\]
-Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse. The matrix operator \(\Delta\) acting on the coefficients vector is then given by
-\begin{align}
-    \Delta := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(1,2)\to(2,2)} \: T^{(0,2)\to(1,2)}  \: D_y^{(0,1)} \: D_y^{(0,0)}.
-\end{align}
-
-The sparsity of the these operators can be shown using a simple integration by parts argument. We note that the non-zero entries of the differential operators are calculated as follows.
-
-Note that we can expand $\pddx \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
-The coefficients of these expansion are then the entries of the relevant operator matrix.
-By a simple integration by parts argument, we have that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
-These are then calculated by taking inner products, i.e.
-\begin{align*}
-c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}
-\end{align*}
-where 
-\begin{align}
-	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \ip< x \: H_{n-k}^{(a, b+k+\half)} \prime (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
-		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad - k \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
-		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
-		\: \ip< y \: P_k^{(b,b)} \prime (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
-\end{align}
-We can then use quadrature to calculate the 1D inner products. This is much cheaper than using the 2D quadrature rule.
-
-Further, we can expand $\pddy \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
-The coefficients of these expansion are then the entries of the relevant operator matrix.
-By a simple integration by parts argument, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
-This is then calculated by taking inner products, i.e.
-\begin{align*}
-c_{n-1,k-1} = \frac{\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
-\end{align*}
-where 
-\begin{align}
-	\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
-		\: \ip< P_k^{(b,b)} \prime, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
-\end{align}
-where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ $\forall m, n \in \N_0$.
-
-We can gain the non-zero entries of the weighted differential operators similarly, by noting that
-\begin{align}
-	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \\
-	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y)
-\end{align}
-and also that
-\begin{align}
-	\ip< \Wab \Pnkab, \Wab \Pab_{m,j} >_{W^{(-a, -b)}} = \ip< \Pnkab, \Pab_{m,j} >_\Wab = \delta_{n,m} \: \delta_{k,j}.
-\end{align}
-
-Calculating the non-zero entries of the parameter transformation operators is similarly done, i.e. use a substitution in the 2D inner product integral to yield a product of 1D inner product integrals.
+$$
+Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in FIGURE.
+We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
+\begin{align}
+    \laplaceoo := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(1,2)\to(2,2)} \: D_y^{(1,1)} \: T^{(0,1)\to(1,1)} \: D_y^{(0,0)}.
+\end{align}
+
+We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions.
+On this note, we can further represent the Laplacian as a map from coefficients in the space
+$$
+\bigW^{(2,2)}
+$$
+to coefficients in the space
+$$
+\bigPoo.
+$$
+Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
+We denote this matrix operator as $\laplacewtt$, and is given by
+\begin{align}
+	\laplacewtt := W_x^{(1,1)} \: W_x^{(2,2)} + T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)} \: T_W^{(2,1)\to(1,1)} \: W_y^{(2,2)}.
+\end{align}
+The sparsity and structure of this matrix can be seen in FIGURE.
+
+The sparsity of the these operators can be shown using a simple integration by parts argument, and can be seen in FIGURE.
 
@@
 
+\subsection{Calculating non-zero entries of the operator matrices}
+
+First, note that we can expand $\pddx \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
+The coefficients of the expansion are then the entries of the relevant operator matrix.
+By a simple integration by parts argument, we have that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
+These are then calculated by taking inner products, i.e.
+\begin{align*}
+c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}
+\end{align*}
+where 
+\begin{align}
+	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
+	&= \ip< x \: H_{n-k}^{(a, b+k+\half)} \prime (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
+	& \quad - k \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
+	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+		\: \ip< y \: P_k^{(b,b)} \prime (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
+\end{align}
+We can then use quadrature to calculate the 1D inner products. This is much cheaper than using the 2D quadrature rule.
+
+Further, we can expand $\pddy \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+The coefficients of the expansion are then the entries of the relevant operator matrix.
+By a simple integration by parts argument, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
+This is then calculated by taking inner products, i.e.
+\begin{align*}
+c_{n-1,k-1} = \frac{\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
+\end{align*}
+where 
+\begin{align}
+	\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
+		\: \ip< P_k^{(b,b)} \prime, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
+\end{align}
+where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ $\forall m, n = 0, 1, 2, \dots$.
+
+We can gain the non-zero entries of the weighted differential operators similarly, by noting that
+\begin{align}
+	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \\
+	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y)
+\end{align}
+and also that
+\begin{align}
+	\ip< \Wab \Pnkab, \Wab \Pab_{m,j} >_{W^{(-a, -b)}} = \ip< \Pnkab, \Pab_{m,j} >_\Wab = \delta_{n,m} \: \delta_{k,j}.
+\end{align}
+
+Calculating the non-zero entries of the parameter transformation operators is similarly done, i.e. use a substitution in the 2D inner product integral to yield a product of 1D inner product integrals.
+
 %
@@
 \begin{align}
-    \Delta^{(1,1)} \mathbf{u} = \mathbf{f}
+    \laplacewii \mathbf{u} = \mathbf{f}
 \end{align}
 which can be solved to find $\mathbf{u}$.
+In FIGURE we see the solution to the Poisson equation with zero boundary conditions given in EQN in the half-disk $\Omega$.
+In FIGURE we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 999????.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
+\bstodo{Describe what plot tells us}
 
@@
         \label{fig:sparsity}
-    	\caption{A "spy" plot of the $\Delta$ operator for $\bigWii(x,y) \to \bigPii(x,y)$ showing its sparsity.}
+    	\caption{A "spy" plot of the $\laplacewii$ matrix operator for the Laplacian showing its sparsity.}
 	\centering
@@
 
-\subsection{Inhomogenous Helmholtz}
-
-Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+\subsection{Inhomogenous variable-coefficient Helmholtz}
+
+Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
 \begin{align}
 	\begin{cases}
-    		\Delta u(x,y) + k^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
+    		\Delta u(x,y) + k^2 v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
 		u(x,y) = 0 \quad \text{on } \partial \Omega
@@
 \end{align}
-noting the imposition of zero Dirichlet boundary conditions on $u$.
+where $k \in \R$, noting the imposition of zero Dirichlet boundary conditions on $u$.
 
@@
 Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^T, {J_y^{(0,0)}}^T$, yielding an operator matrix of the same dimension as the input Jacobi matries.
+We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^T, {J_y^{(0,0)}}^T)$.
 In matrix-vector notation, our system hence becomes:
 \begin{align}
-    (\Delta^{(1,1)} + k^2 T^{(0,0)\to(1,1)} T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
+    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^T, {J_y^{(0,0)}}^T) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
 \end{align}
 which can be solved to find $\mathbf{u}$.
+We can see the sparsity and structure of this matrix system in FIGURE with $v(x,y) = xy^2$ as an example.
+In FIGURE we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in EQN in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.
 
@@
 	\begin{cases}
-    		\Delta u(x,y) + k^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
+    		\Delta u(x,y) + k^2 v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
 		u(x,y) = c \in \R \quad \text{on } \partial \Omega
@@
 \end{align}
-is equivalent to letting $u = v + c$ and solving
+is equivalent to letting $u = \tilde{u} + c$ and solving
 \begin{align}
 	\begin{cases}
-    		\Delta v(x,y) + k^2 v(x,y) = f(x,y) - k^2 c =: g(x,y)  \quad \text{in } \Omega \\
-		v(x,y) = 0 \quad \text{on } \partial \Omega
+    		\Delta \tilde{u}(x,y) + k^2 v(x,y) \; \tilde{u}(x,y) = f(x,y) - k^2 v(x,y) \; =: g(x,y)  \quad \text{in } \Omega \\
+		\tilde{u}(x,y) = 0 \quad \text{on } \partial \Omega
 	\end{cases}.
@@
 
-\subsection{Heat Equation}
+\subsection{Biharmonic equation}
+
+Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+\begin{align}
+	\begin{cases}
+    		\Delta^2 u(x,y) = 0 \quad \text{in } \Omega \\
+		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
+	\end{cases}.
+\end{align}
+where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
+We can construct a matrix operator for $\Delta^2$ that will take coefficients in the space
+$$
+\bigW^{(2,2)}
+$$
+to coefficients in the space
+$$
+\bigP^{(2,2)}.
+$$
+Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$.
+The matrix operator for the Biharmonic operator is then given by
+\begin{align}
+	\biharmonic = \laplaceoo \: \laplacewtt.
+\end{align}
+In FIGURE we see the solution to the Biharmonic equation EQN in the half-disk $\Omega$.
+
+\subsection{Heat equation}
 Consider
@@
 
-Then, using the operator $\Delta$, we can solve this system for the given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
-\begin{align}
-	&T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (\mathbf{u}_{t+1} - \mathbf{u}_t) = h \: \Delta^{(1,1)} \mathbf{u}_{t+1} \nonumber \\
-	\iff& (T^{(0,0)\to(1,1)} \:T_W^{(1,1)\to(0,0)} - h \: \Delta^{(1,1)}) \: \mathbf{u}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
-	\iff& \mathbf{u}_{t+1} = (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \Delta^{(1,1)})^{-1} T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
-	\iff& \mathbf{\tilde{u}}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \Delta^{(1,1)})^{-1} \: \mathbf{\tilde{u}}_t.
+We can solve this system for the given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
+\begin{align}
+	&T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (\mathbf{u}_{t+1} - \mathbf{u}_t) = h \: \laplacewii \mathbf{u}_{t+1} \nonumber \\
+	\iff& (T^{(0,0)\to(1,1)} \:T_W^{(1,1)\to(0,0)} - h \: \laplacewii) \: \mathbf{u}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
+	\iff& \mathbf{u}_{t+1} = (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \laplacewii)^{-1} T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
+	\iff& \mathbf{\tilde{u}}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \laplacewii)^{-1} \: \mathbf{\tilde{u}}_t.
 \end{align}
@@
 %
-\section{Future work - orthogonal polynomials on the half-spheres and spherical triangles}
+\section{Future work - orthogonal polynomials on half-spheres and spherical triangles}
 

### 2019-04-02 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Changed how 2D norms are calculated and stored

@@ -28,7 +28,7 @@ struct OrthogonalPolynomialSpace{FA,WW,F,D,B,R,N} <: PolynomialSpace{D,R}
     ops::Vector{F}  # Cache the ops we get for free from lanczos
     a::Vector{R} # Diagonal recurrence coefficients
     b::Vector{R} # Off diagonal recurrence coefficients
-    opnorm::Vector{R} # The norm of the OPs (all OPs of a OPSpace have the same norm).
+    opnorm::Vector{R} # The norm of the OPs (all OPs of an OPSpace have the same norm).
                         # NOTE this is the value of the norm squared
     opptseval::Vector{Vector{R}}
     derivopptseval::Vector{Vector{R}}
@@ -477,24 +477,26 @@ function inner2(S::HalfDiskSpace, fpts, gpts, w)
             + [fpts[pt] * gpts[pt] for pt = n+1:2n]) .* w) / 2
 end
 
-function getopnorms(S::HalfDiskSpace, m)
+# TODO only store...
+function getopnorms(S::HalfDiskSpace{<:Any, <:Any, T}, k) where T
     # NOTE these are squared norms
-    m̃ = length(S.opnorms)
-    if m > m̃
-        n = m̃ == 0 ? -1 : -1 + Int(round(sqrt(1 + 2(m̃ - 1))))
-        k = m̃ - Int((n + 1)n / 2)
-        resize!(S.opnorms, m)
+    m = length(S.opnorms)
+    if k + 1 > m
+        resize!(S.opnorms, k+1)
         P = S.family.P(S.b, S.b)
         getopnorm(P)
-        for j = m̃+1:m
-            if k > n
-                n += 1
-                k = 0
-            end
-            H = S.family.H(S.a, S.b + k + 0.5)
-            getopnorm(H)
-            S.opnorms[j] = H.opnorm[1] * P.opnorm[1]
-            k += 1
+        # TODO: This is hardcoded, need to find a better way
+        # Doing H(S.a, S.b + k + 0.5) for large k leads to H.weight
+        # reaching maximum number of coefficients when calculating due to
+        # BigFloats
+        X = Fun(T(0)..1)
+        W1 = X^T(S.a) * (1-X^2)^(T(S.b)+0.5)
[…]

### 2019-04-02 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -1,8 +1,8 @@
 using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature,
-        SingularIntegralEquations, Test
+        SingularIntegralEquations, Test, SparseArrays
 import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule,
         gethalfdiskOP, jacobix, jacobiy, differentiatex, differentiatey,
-        resizecoeffs!, laplace, transformparamsoperator, operatorclenshaw, getnk,
+        resizecoeffs!, laplaceoperator, transformparamsoperator, operatorclenshaw, getnk,
         getopindex, weight
 
 
@@ -235,6 +235,7 @@ end
     end
     d2yweight(S::HalfDiskSpace, z) = d2yweight(S, z[1], z[2])
 
+    x, y = 0.34, -0.29; z = [x; y]
     # W11->P11
     D = HalfDiskFamily()
     a, b = 1.0, 1.0; S = D(a, b)
@@ -265,8 +266,3 @@ end
 end
 
 #=====#
-
-a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
-N = 5
-S.family.H(S.a,S.b)
-laplace(D, N)

### 2019-04-02 papers/halfdisk/halfdisk.tex
References to eqns and figs

@@
 
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.4]{sparsityoflaplacian}
+    	\caption{A "spy" plot of the $\laplacewii$ matrix operator for the Laplacian showing its sparsity.}
+        \label{fig:sparsity}
+	\centering
+\end{figure}
+
 We can then obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
@@
 \]
-Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in FIGURE.
+Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
@@
 $$
-Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in FIGURE.
+Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
 We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
@@
 \end{align}
-The sparsity and structure of this matrix can be seen in FIGURE.
-
-The sparsity of the these operators can be shown using a simple integration by parts argument, and can be seen in FIGURE.
+The sparsity and structure of this matrix can be seen in Figure \ref{fig:sparsity}.
+
+The sparsity of the these operators can be shown using a simple integration by parts argument.
 
@@
 \begin{figure}
+	\begin{subfigure}[t]{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution}
 	\centering
+	\caption{}
+        	\label{fig:solution-poisson}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.4\textwidth}
         	\includegraphics[scale=0.4]{solutionblocknorms}
-	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.}
 	\centering
+	\caption{}
         	\label{fig:solutionblocknorms}
+	\end{subfigure}
+	\caption{(a) The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+(b) The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.}
+	\centering
 \end{figure}
@@
 	\end{cases}.
+	\label{eqn:poisson}
 \end{align}
@@
 which can be solved to find $\mathbf{u}$.
-In FIGURE we see the solution to the Poisson equation with zero boundary conditions given in EQN in the half-disk $\Omega$.
-In FIGURE we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 999????.
+In Figure \ref{fig:solution-poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the half-disk $\Omega$.
+In Figure \ref{fig:solutionblocknorms} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
 \bstodo{Describe what plot tells us}
-
-%\begin{figure}
-%	\includegraphics[scale=0.4]{example4coeffs}
-%\centering
-%\caption{Plot of the difference between the previous solution's coefficients and the current solution's coefficients under increasing degree of approximation (number of coefficients used) for $N = 2,\dots,10$, using RHS of $f(x,y) = 2(1-x^2-y^2)\cos(x) - 8x\sin(x) - x(1-x^2-y^2)\sin(x) - 4x^2 \cos(x) - 12xy$ for the Poisson equation.
-The solution is converging as seen by the decreasing (absolute value) difference of coefficient values.}
-%\centering
-%\label{fig:poisson}
-%\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.4]{sparsityoflaplacian}
-        \label{fig:sparsity}
-    	\caption{A "spy" plot of the $\laplacewii$ matrix operator for the Laplacian showing its sparsity.}
-	\centering
-\end{figure}
 
@@
 	\end{cases}.
+	\label{eqn:helmholtz}
 \end{align}
@@
 We can see the sparsity and structure of this matrix system in FIGURE with $v(x,y) = xy^2$ as an example.
-In FIGURE we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in EQN in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.
+In FIGURE we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.3]{solution}
-        	\label{fig:solution}
-    	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
-	\centering
-\end{figure}
-
 \subsection{Biharmonic equation}
@@
 	\end{cases}.
+	\label{eqn:biharmonic}
 \end{align}
@@
 \end{align}
-In FIGURE we see the solution to the Biharmonic equation EQN in the half-disk $\Omega$.
+In FIGURE we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
 

### 2019-04-02 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
print statements added

@@ -56,7 +56,6 @@ function lanczos!(w, P, β, γ, N₀=0)
     end
 
     for k = N₀+1:N
-        @show k
         v = x*P[k] - γ[k-1]*P[k-1]
         β[k] = sum(w*v*P[k])
         v = v - β[k]*P[k]
@@ -114,13 +113,24 @@ function OrthogonalPolynomialFamily(::Type{R}, w::Vararg{Fun{<:Space{D,B}},N}) w
 end
 
 function (P::OrthogonalPolynomialFamily{<:Any,<:Any,<:Any,R,B,N})(α::Vararg{B,N}) where {R,B,N}
-    if length(α) == 1
-        haskey(P.spaces,α) && return P.spaces[α]
-        P.spaces[α] = OrthogonalPolynomialSpace(P, (P.factors.^α)[1], α)
+    haskey(P.spaces,α) && return P.spaces[α]
+    # NOTE: This ensures we dont hang on calculating the weight for large
+    # parameter values
+    maxp = 100
+    if any(α .> maxp)
+        β = []
+        f = []
+        for j = 1:length(α)
+            r = α[j] % maxp
+            t = Int(div(α[j], maxp))
+            β = append!(β, [tt == 0 ? r : maxp for tt=0:t])
+            f = append!(f, [P.factors[j] for tt=0:t])
+        end
+        W = length(α) == 1 ? (f.^β)[1] : prod(f.^β)
     else
-        haskey(P.spaces,α) && return P.spaces[α]
-        P.spaces[α] = OrthogonalPolynomialSpace(P, prod(P.factors.^α), α)
+        W = length(α) == 1 ? (P.factors.^α)[1] : prod(P.factors.^α)
     end
+    P.spaces[α] = OrthogonalPolynomialSpace(P, W, α)
 end
 
 #####
@@ -890,30 +900,28 @@ function partialoperatorx(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
     Px = Sx.family.P(Sx.b, Sx.b)
     ptsp, wp = pointswithweights(Px, N+2) # TODO
     getopptseval(P, N, ptsp)
+    @show "here1"
     getderivopptseval(P, N, ptsp)
+    @show "here2"
     getopptseval(Px, N, ptsp)
+    @show "here3"
     H = S.family.H(S.a, S.b+0.5)
     ptsh, wh = pointswithweights(H, 2N+2)
[…]

### 2019-04-03 papers/halfdisk/halfdisk.tex
Trapezium appendix

@@
 
-Note that by setting $c = 0$, $\alpha = 0$ and $\beta = 1$ we obtain the 2D orthogonal polynomials on the half disk.
+Note that by setting $c = 0$, $\alpha = 0$ and $\beta = 1$ we obtain the 2D orthogonal polynomials on the half-disk.
+
+\section{Trapeziums}
+
+We can further extend this work to trapezium shaped domains. Note that for any trapezium there exists an affine map to the canonical trapezium domain that we consider here, given by
+\begin{align}
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+\end{align}
+with 
+\begin{align}
+\begin{cases}
+(\alpha, \beta) &:= (0,1) \\
+(\gamma, \delta) &:= (0,1) \\
+\rho(x) &:= 1-\half x \\
+w^{(a,b,c)}_H(x) &:= (x - \alpha)^{a} \: (\beta - x)^b \: \rho(x)^{c} = x^a \: (1-x)^{b} \: (1-\half x)^{c} \\
+w^{(a,b)}_P(x) &:= (x - \gamma)^a \: (\delta-x)^b.
+\end{cases}
+\end{align}
+The weight $w^{(b)}_2(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(b, b)}\}$.
+$w^{(a,b,c)}_1(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a, b,c)}\}\).
+Thus we arrive at the two parameter family of 2D orthogonal polynomials $\{\Pnk\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\begin{align}
+	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(c, a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+\end{align}
+orthogonal with respect to the weight
+\begin{align}
+	W^{(a,b,c)}(x,y) &:= w^{(a,b,c)}_H(x) w^{(c,d)}_P(\frac{y}{\rho(x)}) \nonumber \\
+	&= x^a \: (1-x)^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega.
+\end{align}
 

### 2019-04-03 papers/halfdisk/halfdisk.tex
Edits to trapezium section

@@
 \rho(x) &:= 1-\half x \\
-w^{(a,b,c)}_H(x) &:= (x - \alpha)^{a} \: (\beta - x)^b \: \rho(x)^{c} = x^a \: (1-x)^{b} \: (1-\half x)^{c} \\
-w^{(a,b)}_P(x) &:= (x - \gamma)^a \: (\delta-x)^b.
+w^{(a,b,c)}_H(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\half x)^{c} \\
+w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
 \end{cases}
 \end{align}
-The weight $w^{(b)}_2(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(b, b)}\}$.
-$w^{(a,b,c)}_1(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a, b,c)}\}\).
+The weight $w^{(a,b)}_P(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
+We note that the shifted Jacobi polynomials relate the the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = P_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
+$w^{(a,b,c)}_H(x)$ is the (non-classical) weight for the OPs we dentote $\{H_n^{(a,b,c)}\}\).
 Thus we arrive at the two parameter family of 2D orthogonal polynomials $\{\Pnk\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
-	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(c, a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\Pnk^{(a,b,c,d)}(x,y) := H_{n-k}^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \begin{align}
-	W^{(a,b,c)}(x,y) &:= w^{(a,b,c)}_H(x) w^{(c,d)}_P(\frac{y}{\rho(x)}) \nonumber \\
-	&= x^a \: (1-x)^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega.
-\end{align}
+	W^{(a,b,c,d)}(x,y) &:= w^{(a,b,c+d)}_H(x) \: w^{(d,c)}_P(\frac{y}{\rho(x)}) \nonumber \\
+	&= (1-x)^a \: x^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega,
+\end{align}
+i.e. that
+\begin{align}
+	\int \int_\Omega \Pnk^{(a,b,c,d)} \: P_{m,j}^{(a,b,c,d)} \: W^{(a,b,c,d)} \: \dx \dy &:= C_{n,k}^{(a,b,c,d)} \: \delta_{n,m}
+ \: \delta_{j,k},
+\end{align}
+for some constant $C_{n,k}^{(a,b,c,d)}$.
 

### 2019-04-03 papers/halfdisk/halfdisk.tex
Added explanation for general case of constructing the OPs to get 4 param family of 2D OPs

@@
 This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
-More precisely, let  \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
-
+More precisely, we consider the solution of partial differential equations on the domain
+\begin{align}
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+\end{align}
+where we require that either of the following scenarios hold:
 \begin{enumerate}
 \item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
+\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and the weight function \(w_2\) on the interval $(\gamma, \delta)$ is an even function.
 \end{enumerate}
-
-We consider the solution of partial differential equations on the domain
-\begin{align}
-\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}.
-\end{align}
-For simplicity of discussion we focus on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.
+We will explain where $w_2$ comes into play in Section \ref{Section:OPs}.
+
+For simplicity of discussion we focus in this paper on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.
 
@@
 
-By using a similar process to \cite[p55--56]{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([0,1]\) and \([-1,1]\).
+By using a similar process to \cite[p55--56]{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).
+We generalise this process to demonstrate a method of constructing multi-parameter families of 2D orthogonal polynomials on disk-slice and trapezium shaped domains $\Omega$ from multi-parameter families of 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).
 
@@
 
+We can generalise this for our purposes in the following proposition.
+\begin{proposition}
+Let
+\begin{align*}
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+\end{align*}
+with $\rho : (\alpha, \beta) \to \R$. Further, let $w^{(a,b,c)}_H(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
+\begin{align}
+\begin{cases}
+w^{(a,b,c)}_H(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
+w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b
+\end{cases}
+\end{align}
+where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{H_n^{(a,b,c)}\}$ are orthogonal with respect to $w^{(a,b,c)}_H(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
+
+Assume that either of the following hold:
+\begin{enumerate}
+\item  \(\rho\) is a degree 1 polynomial, or 
+\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w^{(a,b)}_P\) is an even function (i.e.
+$a = b$, and we can hence denote the weight as $w^{(a)}_P(x) = (\delta-x^2)^a$).
+\end{enumerate}
+Then the four-parameter 2D polynomials defined on $\Omega$ given by
+\begin{align}
+	\Pnk^{(a,b,c,d)}(x,y) := H_{n-k}^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+\end{align}
+are orthogonal polynomials with respect to the weight 
+\begin{align}
+	W^{(a,b,c,d)}(x,y) := w^{(a,b,c+d)}_H(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega. 
+\end{align}
+\end{proposition}
+\begin{proof}
+See \cite[p55--56]{dunkl2014orthogonal}.
+\end{proof}
+
 For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
@@
 \end{align}
+Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
+This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.
 
@@
 \begin{cases}
--1 \le \alpha < &\beta \le 1 \\
+(\alpha, \beta) &\subset (-1,1) \\
 (\gamma, \delta) &:= (-1,1) \\
-\rho(x) &:= (1-x^2)^{\half} \\
-w^{(a,b,c)}_1(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c} = (\beta - x)^a \: (x - \alpha)^{b} \: (1-x^2)^{c} \\
-w^{(b)}_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b.
+\rho(x) &:= (1-x^2)^{\half}.
 \end{cases}
 \end{align}
-The weight $w^{(b)}_2(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(b, b)}\}$.
-$w^{(a,b,c)}_1(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a, b,c)}\}\).
-Thus we arrive at the two parameter family of 2D orthogonal polynomials $\{\Pnk\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align}
-	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(c, a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+Our 1D weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively are then given by:
+\begin{align}
+\begin{cases}
+w^{(a,b,c)}_H(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
+w^{(a)}_P(x) &:= (1-x^2)^a.
+\end{cases}
+\end{align}
+
+The weight $w^{(a)}_P(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(a, a)}\}$.
+$w^{(a,b,c)}_H(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a,b,c)}\}\).
+Thus we arrive at the three-parameter family of 2D orthogonal polynomials $\{\Pnk^{(a,b,c)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\begin{align}
+	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: P_k^{(c,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \begin{align}
-	W^{(a,b,c)}(x,y) &:= w^{(c,a,b)}_1(x) w^{(b)}_2(\frac{y}{\rho(x)}) \nonumber \\
-	&= (\beta - x)^c \: (x - \alpha)^{a} \: (\rho(x)^2 -y^2)^b \nonumber \\
-	&= (\beta - x)^c \: (x - \alpha)^{a} \: (1 - x^2 -y^2)^b , \quad (x,y) \in \Omega.
-\end{align}
-
-Note that by setting $c = 0$, $\alpha = 0$ and $\beta = 1$ we obtain the 2D orthogonal polynomials on the half-disk.
+	W^{(a,b,c)}(x,y) &:= w^{(a,b,2c)}_1(x) w^{(c)}_2(\frac{y}{\rho(x)}) \nonumber \\
+	&= (\beta - x)^a \: (x - \alpha)^{b} \: (\rho(x)^2 -y^2)^c \nonumber \\
+	&= (\beta - x)^a \: (x - \alpha)^{b} \: (1 - x^2 -y^2)^c , \quad (x,y) \in \Omega.
+\end{align}
+
+We can note that by setting the parameter $c = 0$, and $\alpha = 0$, $\beta = 1$ we obtain the 2D orthogonal polynomials on the half-disk (note again also that we made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ for the half-disk case, for ease of writing and computation).
 
@@
 $w^{(a,b,c)}_H(x)$ is the (non-classical) weight for the OPs we dentote $\{H_n^{(a,b,c)}\}\).
-Thus we arrive at the two parameter family of 2D orthogonal polynomials $\{\Pnk\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\Pnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}
@@
 	W^{(a,b,c,d)}(x,y) &:= w^{(a,b,c+d)}_H(x) \: w^{(d,c)}_P(\frac{y}{\rho(x)}) \nonumber \\
-	&= (1-x)^a \: x^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega,
-\end{align}
-i.e. that
-\begin{align}
-	\int \int_\Omega \Pnk^{(a,b,c,d)} \: P_{m,j}^{(a,b,c,d)} \: W^{(a,b,c,d)} \: \dx \dy &:= C_{n,k}^{(a,b,c,d)} \: \delta_{n,m}
- \: \delta_{j,k},
-\end{align}
-for some constant $C_{n,k}^{(a,b,c,d)}$.
+	&= (1-x)^a \: x^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega.
+\end{align}
 

### 2019-04-03 papers/halfdisk/halfdisk.tex
Minor edit

@@
 \begin{figure}
-	\begin{subfigure}[t]{0.3\textwidth}
 	\includegraphics[scale=0.3]{solution}
 	\centering
-	\caption{}
+	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
         	\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.4\textwidth}
-        	\includegraphics[scale=0.4]{solutionblocknorms}
-	\centering
-	\caption{}
-        	\label{fig:solutionblocknorms}
-	\end{subfigure}
-	\caption{(a) The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-(b) The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.}
 	\centering
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-\bstodo{Describe what plot tells us}
+We see that\bstodo{Describe what plot tells us}
+
+\begin{figure}
+        	\includegraphics[scale=0.5]{solutionblocknorms}
+	\centering
+	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.}
+        	\label{fig:solutionblocknorms}
+\end{figure}
 

### 2019-04-03 papers/halfdisk/halfdisk.tex
Added more refs to eqns

@@
 We can generalise this for our purposes in the following proposition.
+\bstodoinline{I have added in this proposition to explain how I think this generalises to the main domain shapes we are thinking about, including your triangle case.}
 \begin{proposition}
@@
 \end{align}
+\bstodoinline{Hope the following makes sense.
+The generalisation I have, and how I originally wrote/coded up the half-disk case use slightly different conventions regarding the parameters - would it be worth reframing the paper to match this generalisation, or is this fine?}
 Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
@@
 \begin{align}
-x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) \\
+x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) 
+\label{eqn:Hrecurrence} \\
 y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
@@
 J_x^{(a,b)} \: \bigPab(x,y) = x \: \bigPab(x,y), \quad J_y^{(a,b)} \: \bigPab(x,y) = y \: \bigPab(x,y).
+\label{eqn:jacobimatricesdefinition}
 \end{align}
@@
 
-Combining each system in (71) we can write the block-tridiagonal system
+Combining each system in (\ref{eqn:jacobimatricesdefinition}) we can write the block-tridiagonal system
 \begin{align}
@@
 
-To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (4), we use a version of the Stieltjes algorithm. Recall that
+To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
 \begin{align}
@@
 	L(v_h) = a(u_h,v_h) \quad \forall \: v_h \in V_h.
+	\label{eqn:FEMweakform}
 \end{align}
@@
 \end{align}
-and rewrite (73) in matrix form:
+and rewrite (\ref{eqn:FEMweakform}) in matrix form:
 \begin{align}
@@
 
-Since (73) is equivalent to stating that
+Since (\ref{eqn:FEMweakform}) is equivalent to stating that
 \begin{align}

### 2019-04-03 papers/halfdisk/halfdisk.tex
Further edits

@@
 \end{align}
-Then $J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded structures
+The matrices $J_x^{(a,b)}, J_y^{(a,b)}$ act on the coefficients vector of a functions expansion in the $\{\Pnkab\}$ basis.
+For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
+Then $x \: f(x,y)$ is approximated by $\bigPab(x,y)^\top {J_x^{(a,b)}}^\top \mathbf{f}$.
+In other words, ${J_x^{(a,b)}}^\top \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\Pnkab\}$ basis.
+
+$J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded structures
 \begin{align}
@@
 
-Let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
-Recall that $\Wab(x,y) := w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $w_H^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
+Recall that, for $a, b$ be general parameters, a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
+Recall also that $\Wab(x,y) := w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $w_H^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
 
@@
 \begin{align}
-	\ip< \Wab \Pnkab, \Wab \Pab_{m,j} >_{W^{(-a, -b)}} = \ip< \Pnkab, \Pab_{m,j} >_\Wab = \delta_{n,m} \: \delta_{k,j}.
+	\ip< \Wab \Pnkab, W^{(\tilde{a},\tilde{b})} P^{(\tilde{a},\tilde{b})}_{m,j} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \Pnkab, P^{(\tilde{a},\tilde{b})}_{m,j} >_\Wab.
 \end{align}

### 2019-04-04 papers/halfdisk/halfdisk.tex
Added new figures. Replaced transposes using letter T with top

@@
 
-\newcommand{\Dlt}{D^T_l}
+\newcommand{\Dlt}{D^\top_l}
 
@@
 \newcommand{\bigPab}{\bigP^{(a,b)}}
-\newcommand{\Dnt}{D^T_n}
+\newcommand{\Dnt}{D^\top_n}
 \newcommand{\Wii}{W^{(1,1)}}
@@
 	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
-C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
+C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
 \nonumber \\
@@
 \end{align}
-It follows that we can apply $D_n^T$ in $O(n)$ complexity, and thereby construct $\bigPab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
+It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby construct $\bigPab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
 
@@
 \begin{align}
-\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^T \: D_x^{(a,b)} \: \mathbf{f}, \\
-\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^T \: D_y^{(a,b)} \: \mathbf{f}, \\
-\dx[\Wab(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
-\dy[\Wab(x,y) \: f(x,y)] &= \bigW^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
+\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^\top \: D_x^{(a,b)} \: \mathbf{f}, \\
+\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^\top \: D_y^{(a,b)} \: \mathbf{f}, \\
+\dx[\Wab(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1)}(x,y)^\top \: W_x^{(a,b)} \: \mathbf{f}, \\
+\dy[\Wab(x,y) \: f(x,y)] &= \bigW^{(a,b-1)}(x,y)^\top \: W_y^{(a,b)} \: \mathbf{f}.
 \end{align}
@@
 \begin{align}
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^T \: \bigP^{(a+1,b)}(x,y) \\
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigP^{(a+1,b+1)}(x,y) \\
-\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^T \: \bigW^{(a-1,b)}(x,y) \\
-\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^T \: \bigW^{(a,b-1)}(x,y) \\
-\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^T \: \bigW^{(a-1,b-1)}(x,y) \\
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^\top \: \bigP^{(a+1,b)}(x,y) \\
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^\top \: \bigP^{(a,b+1)}(x,y) \\
+\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigP^{(a+1,b+1)}(x,y) \\
+\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^\top \: \bigW^{(a-1,b)}(x,y) \\
+\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^\top \: \bigW^{(a,b-1)}(x,y) \\
+\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigW^{(a-1,b-1)}(x,y) \\
 \end{align}
@@
 \begin{figure}
+	\begin{subfigure}[t]{0.45\textwidth}
+	\includegraphics[scale=0.4]{sparsityoflaplacian-w11}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.45\textwidth}
+	\includegraphics[scale=0.4]{sparsityoflaplacian-w22}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.45\textwidth}
+	\includegraphics[scale=0.4]{sparsityoflaplacian-00}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.45\textwidth}
+	\includegraphics[scale=0.4]{sparsityofhelmholtz}
+	\end{subfigure}
 	\centering
-	\includegraphics[scale=0.4]{sparsityoflaplacian}
-    	\caption{A "spy" plot of the $\laplacewii$ matrix operator for the Laplacian showing its sparsity.}
+    	\caption{"Spy" plots of the different operator matrices for the Laplacian operator $\Delta$, as well as the operator matrix for the Helmholtz equation, showing their sparsity.
+Top Left: $\laplacewii$.
+Top Right: $\laplacewtt$.
+Bottom Left: $\laplaceoo$.
+Bottom Right: the weighted variable coefficient Helmholtz operator $\laplacewii + k^2 \: T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}$ for $v(x,y) = xy^2$ and $k = 0.25$.}
         \label{fig:sparsity}
-	\centering
 \end{figure}
@@
 \begin{align*}
-f(x,y) = \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^T \: \mathbf{f}_n
+f(x,y) = \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^\top \: \mathbf{f}_n
 \end{align*}
@@
 Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^T, {J_y^{(0,0)}}^T$, yielding an operator matrix of the same dimension as the input Jacobi matries.
-We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^T, {J_y^{(0,0)}}^T)$.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matries.
+We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$.
 In matrix-vector notation, our system hence becomes:
 \begin{align}
-    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^T, {J_y^{(0,0)}}^T) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
+    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
 \end{align}
@@
 
-Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^T \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to.
+Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^\top \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to.
 Note that due to orthogonality this is a diagonal matrix.
@@
 					\partial_y v_h
-				\end{pmatrix}^T 
+				\end{pmatrix}^\top 
 				\begin{pmatrix}
@@
 	&= \int_\element \begin{pmatrix}
-					\bigPoo^T \Wii_x \mathbf{v} \\
-					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
-				\end{pmatrix}^T 
+					\bigPoo^\top \Wii_x \mathbf{v} \\
+					\bigPoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
+				\end{pmatrix}^\top 
 				\begin{pmatrix}
-					\bigPoo^T \Wii_x \mathbf{u} \\
-					\bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
+					\bigPoo^\top \Wii_x \mathbf{u} \\
+					\bigPoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
 				\end{pmatrix}
 				\\
-	&= \int_\element \Big( \mathbf{v}^T {\Wii_x}^T \bigPoo \bigPoo^T \Wii_x \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^T ({T^{(1,0)\to(0,0)} \Wii_y})^T \bigPoo \bigPoo^T T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
-	&= \mathbf{v}^T \: \Big( {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
+	&= \int_\element \Big( \mathbf{v}^\top {\Wii_x}^\top \bigPoo \bigPoo^\top \Wii_x \mathbf{u} \nonumber \\
+					& \quad \quad \quad + \mathbf{v}^\top ({T^{(1,0)\to(0,0)} \Wii_y})^\top \bigPoo \bigPoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
+	&= \mathbf{v}^\top \: \Big( {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
 \end{align}
@@
 	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
-	&= \int_\element \: \mathbf{v}^T \: \bigWii \: \bigPii^T \: \mathbf{f} \: d\mathbf{x} \\
-	&= \mathbf{v}^T \: \langle \bigPii, {\bigPii}^T \rangle_{\Wii} \: d\mathbf{x} \\
-	&= \mathbf{v}^T \Lambda^{(1,1)} \: \mathbf{f},
+	&= \int_\element \: \mathbf{v}^\top \: \bigWii \: \bigPii^\top \: \mathbf{f} \: d\mathbf{x} \\
+	&= \mathbf{v}^\top \: \langle \bigPii, {\bigPii}^\top \rangle_{\Wii} \: d\mathbf{x} \\
+	&= \mathbf{v}^\top \Lambda^{(1,1)} \: \mathbf{f},
 \end{align}
@@
 \begin{align}
-A = {\Wii_x}^T \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^T \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
+A = {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
 \end{align}

### 2019-04-08 papers/halfdisk/halfdisk.tex
New figures for paper

@@
 \begin{figure}
-	\includegraphics[scale=0.3]{solution}
+	\begin{subfigure}[t]{0.3\textwidth}
 	\centering
-	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
+	\includegraphics[scale=0.3]{solution-poisson}
+	\caption{}
 	\label{fig:solution-poisson}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.3\textwidth}
 	\centering
+	\includegraphics[scale=0.3]{solution-helmholtz}
+	\caption{}
+	\label{fig:solution-helmholtz}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.3]{solution-biharmonic}
+	\caption{}
+	\label{fig:solution-biharmonic}
+	\end{subfigure}
+	\centering
+    	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 0.25$.
+(c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
+        \label{fig:solution}
 \end{figure}
@@
 which can be solved to find $\mathbf{u}$.
-We can see the sparsity and structure of this matrix system in FIGURE with $v(x,y) = xy^2$ as an example.
-In FIGURE we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.
+We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
+In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.
 
@@
 \end{align}
-In FIGURE we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
+In Figure \ref{fig:solution-biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
 

### 2019-04-08 papers/halfdisk/halfdisk.tex
readthrough

@@
 
-\title{Sparse spectral and $p$-finite element methods on disk-slices and trapeziums I: zero boundary conditions}
+\title{Sparse spectral and $p$-finite element methods for partial differential equations on disk-slices and trapeziums}
 \author{Sheehan Olver, Ben Snowball}
@@
 \item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and the weight function \(w_2\) on the interval $(\gamma, \delta)$ is an even function.
+\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
 \end{enumerate}
-We will explain where $w_2$ comes into play in Section \ref{Section:OPs}.
 
@@
 Sparsity comes from expanding the domain and range of an operator  using different choices of $a$ and $b$.
-Where the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
-Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}.
+Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
+Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
 
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \sotodo{Add citation to this}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
 $$
-\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
+\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
 $$
@@
 We can generalise this for our purposes in the following proposition.
-\bstodoinline{I have added in this proposition to explain how I think this generalises to the main domain shapes we are thinking about, including your triangle case.}
 \begin{proposition}
@@
 The generalisation I have, and how I originally wrote/coded up the half-disk case use slightly different conventions regarding the parameters - would it be worth reframing the paper to match this generalisation, or is this fine?}
+\sotodoinline{Yes I think its fine} 
 Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
@@
 The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
-Thus we arrive at  a two parameter family of 2D orthogonal polynomials \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align}
+Thus we arrive at  a two parameter family of 2D orthogonal polynomials\sotodo{We should probably use a different letter than $P$ to avoid confusion with Jacobi polynomials}
+\ \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\begin{align}\label{eq:diskpolys}
  \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
@@
 \alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
-\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
-\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \\
-\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \\
-\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}}. 
+\betaab_{n,k,1} &:= {\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k}^{(a, b+k-1/2)}>_{w_H^{(a, b+k+1/2)}} \\
+\betaab_{n,k,2} &:= {\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k-2}^{(a, b+k+{3/2})}>_{w_H^{(a, b+k+{3/2})}} \\
+\betaab_{n,k,3} &:= {\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k+1}^{(a, b+k-1/2)}>_{w_H^{(a, b+k+1/2)}} \\
+\betaab_{n,k,4} &:= {\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k-1}^{(a, b+k+{3/2})}>_{w_H^{(a, b+k+{3/2})}} \\
+\betaab_{n,k,5} &:= {\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k+2}^{(a, b+k-1/2)}>_{w_H^{(a, b+k+1/2)}} \\
+\betaab_{n,k,6} &:= {\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k}^{(a, b+k+{3/2})}>_{w_H^{(a, b+k+{3/2})}}. 
 \end{align}
@@
 For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
-Then $x \: f(x,y)$ is approximated by $\bigPab(x,y)^\top {J_x^{(a,b)}}^\top \mathbf{f}$.
-In other words, ${J_x^{(a,b)}}^\top \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\Pnkab\}$ basis.
-
-$J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded structures
+Then $x \: f(x,y)$ is approximated by $\bigPab(x,y)^\top {J_x^{(a,b)\top}} \mathbf{f}$.
+In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\Pnkab\}$ basis.
+
+$J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded\sotodo{Define Banded-Block-sBanded} structures
 \begin{align}
@@
 \end{align}
-where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
+where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:\sotodo{Remove $A^z_n$?}
 \begin{align}
@@
 For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e.
-such that $\Dnt A_n = I_{n+3}$.
+such that $\Dnt A_n = I_{n+3}$\sotodo{n+1?}.
 Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
@@
 \begin{align}
-\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^\top \: D_x^{(a,b)} \: \mathbf{f}, \\
-\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^\top \: D_y^{(a,b)} \: \mathbf{f}, \\
-\dx[\Wab(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1)}(x,y)^\top \: W_x^{(a,b)} \: \mathbf{f}, \\
-\dy[\Wab(x,y) \: f(x,y)] &= \bigW^{(a,b-1)}(x,y)^\top \: W_y^{(a,b)} \: \mathbf{f}.
-\end{align}
-The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials, as seen in the DLMF \cite{DLMFDerivatives}.
-
-Further, we note that there exist "parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+{\partial f \over \partial x}&= \bigP^{(a+1,b+1)}(x,y)^\top \: D_x^{(a,b)} \: \mathbf{f}, \\
+{\partial f \over \partial y} &= \bigP^{(a,b+1)}(x,y)^\top \: D_y^{(a,b)} \: \mathbf{f}, \\
+{\partial \over \partial x}[\Wab(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1)}(x,y)^\top \: W_x^{(a,b)} \: \mathbf{f}, \\
+{\partial \over \partial y}[\Wab(x,y) \: f(x,y)] &= \bigW^{(a,b-1)}(x,y)^\top \: W_y^{(a,b)} \: \mathbf{f}.
+\end{align}
+The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite{DLMFDerivatives}, and on the triangle \cite{olver2018recurrence}.
+
+Further, we note that there exist ``parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
@@
 	\centering
-    	\caption{"Spy" plots of the different operator matrices for the Laplacian operator $\Delta$, as well as the operator matrix for the Helmholtz equation, showing their sparsity.
+    	\caption{\sotodoinline{Fix titles} "Spy" plots of the different operator matrices for the Laplacian operator $\Delta$, as well as the operator matrix for the Helmholtz equation, showing their sparsity.
 Top Left: $\laplacewii$.
@@
 \]
-to coefficients in the non-weighted space
-\[
-\bigPii(x,y).
-\]
+to coefficients in the non-weighted space $\bigPii(x,y)$.
+
 Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
@@
 Further, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space
-$$
+$
 \bigPoo(x,y)
-$$
+$
 to coefficients in the space
-$$
+$
 \bigP^{(2,2)}(x,y).
-$$
+$
 Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
@@
 On this note, we can further represent the Laplacian as a map from coefficients in the space
-$$
+$
 \bigW^{(2,2)}
-$$
+$
 to coefficients in the space
-$$
+$
 \bigPoo.
-$$
+$
 Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
@@
 
-The sparsity of the these operators can be shown using a simple integration by parts argument.
+The sparsity of the these operators can be shown using a simple integration by parts argument. \sotodoinline{Add argument}
 
@@
 where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$.
+We use a variant of the Stieltjes algorithm where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
+This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $b$ is large, as required for large $n$ in \eqref{eq:diskpolys}.
+The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm.
 
 \begin{algorithm}
-Inputs: max degree $N$, weight $w(x)$
+\sotodo{This is just Stieltjes right? We can probably delete it and reference Gautschi.}Inputs: max degree $N$, weight $w(x)$
 \begin{flalign*}
@@
 \end{algorithm}
-We use the ApproxFun \cite{ApproxFun} package to calculate the inner products and norms in this algorithm to obtain the coefficients.
 
@@
 
-Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$:
-\begin{align}
-\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
+Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: \sotodo{Add sentence explaining how symmetry is imprtant here}
+\begin{align}
+\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j) \big], \\
 || \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
@@
 \end{align*}
-where 
+where \sotodo{Why giant $\prime$?}
 \begin{align}
@@
 \end{align}
-where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ $\forall m, n = 0, 1, 2, \dots$.
+where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ \sotodo{Aren't these $= 1$ since we orthonormalize?} $\forall m, n = 0, 1, 2, \dots$.
 
@@
     	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 0.25$.
+(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 0.25$ \sotodoinline{Make a lot higher}.
 (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
@@
 In Figure \ref{fig:solution-poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the half-disk $\Omega$.
-In Figure \ref{fig:solutionblocknorms} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200.
+In Figure \ref{fig:solutionblocknorms} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
 	\centering
-	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.}
+	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
+This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.
+\sotodoinline{More blocks}}
         	\label{fig:solutionblocknorms}
@@
 We can tackle the problem as follows.
-Fix \(a, b = 1,1\).
 Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matries.
+Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule above.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
 We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$.
@@
 
-Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+Find \(u(x,y)\) given a function \(f(x,y)\) such that:\sotodo{I assume there should be an $f$ in the right-hand side?}
 \begin{align}
@@
 We can construct a matrix operator for $\Delta^2$ that will take coefficients in the space
-$$
+$
 \bigW^{(2,2)}
-$$
+$
 to coefficients in the space
-$$
+$
 \bigP^{(2,2)}.
-$$
+$
 Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$.
-The matrix operator for the Biharmonic operator is then given by
+The matrix operator for the Biharmonic operator is then given by \sotodo{Move up to other discussion on constructing Biharmonic}
 \begin{align}
@@
 
-Using a backward Euler time-stepping method, we can formulate this for \(t = 0,1,2,\dots\) and step length $h$:
+For simplicity of exposition, we use a backward Euler time-stepping method to discretise in time, which can be generalised to more sophisticated time-stepping schemes.
+That is, formulate this for \(t = 0,1,2,\dots\) and step length $h$:
 \begin{align}
@@
 \section{Future work - orthogonal polynomials on half-spheres and spherical triangles}
+
+\sotodoinline{Rewrite in a more "paper" like tone}
 
@@
 \subsection{Two-element p-FEM on the disk (without BCs)}
+
+\sotodoinline{Delete section?}
 We now consider the case of using two elements to discretise a unit disk domain $\Omega$, that is the each element $\element$ is a half-disk, with common edge defined by the line $x=0$ for $y \in [-1,1]$.
@@
 
+\sotodoinline{Fix DLMF References.}
+\sotodoinline{Abbreviate Journal titles.} 
 \bibliography{halfdisk-bib}

### 2019-04-11 papers/halfdisk/halfdisk.tex
Addressed some of SO's points

@@
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \sotodo{Add citation to this}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \sotodo{Add citation to this} \bstodo{should we instead take the Xu approach?}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
 \end{align}
-\bstodoinline{Hope the following makes sense.
-The generalisation I have, and how I originally wrote/coded up the half-disk case use slightly different conventions regarding the parameters - would it be worth reframing the paper to match this generalisation, or is this fine?}
-\sotodoinline{Yes I think its fine} 
 Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
@@
 The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\).
-The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
+The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
 Thus we arrive at  a two parameter family of 2D orthogonal polynomials\sotodo{We should probably use a different letter than $P$ to avoid confusion with Jacobi polynomials}
@@
 
-$J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded\sotodo{Define Banded-Block-sBanded} structures
+$J_x^{(a,b)}, J_y^{(a,b)}$ have Banded-Block-Banded structures.
+A block-banded matrix is a block $n$-diagonal matrix where $n$ is reasonably small for the size of the full matrix.
+A banded-block-banded matrix is one where the blocks themselves are banded matrices.
+$J_x^{(a,b)}, J_y^{(a,b)}$ are given by:
 \begin{align}
@@
 \end{align}
-where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:\sotodo{Remove $A^z_n$?}
+where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
 \begin{align}
@@
 		A^x_n \\
-		A^y_n \\
-		A^z_n
-	    \end{pmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
+		A^y_n
+	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)} \quad (n \ne N), \quad
 C_n := \begin{pmatrix}
@@
 For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e.
-such that $\Dnt A_n = I_{n+3}$\sotodo{n+1?}.
+such that $\Dnt A_n = I_{n+2}$.
 Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
@@
 \bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
-\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots.
+\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n(x,y) - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots.
 \end{cases}
@@
 We can then obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
-\[
+$$
 \bigWii(x,y) = \Wii(x,y) \: \bigPii(x,y)
-\]
+$$
 to coefficients in the non-weighted space $\bigPii(x,y)$.
-
-Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
@@
 \end{align}
-
-Further, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space
-$
-\bigPoo(x,y)
-$
-to coefficients in the space
-$
-\bigP^{(2,2)}(x,y).
-$
-Importantly, again, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+
+We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions.
+On this note, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bigPoo(x,y)$ to coefficients in the space $\bigP^{(2,2)}(x,y)$.
 We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
@@
 \end{align}
-
-We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions.
-On this note, we can further represent the Laplacian as a map from coefficients in the space
-$
-\bigW^{(2,2)}
-$
-to coefficients in the space
-$
-\bigPoo.
-$
+Further represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bigPoo$.
 Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
@@
 \end{align}
-The sparsity and structure of this matrix can be seen in Figure \ref{fig:sparsity}.
-
-The sparsity of the these operators can be shown using a simple integration by parts argument. \sotodoinline{Add argument}
+We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bigP^{(2,2)}$.
+Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$.
+The matrix operator for the Biharmonic operator is then given by
+\begin{align}
+	\biharmonic = \laplaceoo \: \laplacewtt.
+\end{align}
+The sparsity and structure of this biharmonic operator is seen in Figure \ref{fig:sparsity}.
+
+The sparsity of the these operators can be shown using a simple integration by parts argument, see Section \ref{subsection:Computation-operatormatrices}.
 
@@
 where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$.
-We use a variant of the Stieltjes algorithm where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
+We use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
 This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $b$ is large, as required for large $n$ in \eqref{eq:diskpolys}.
 The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm.
-
-\begin{algorithm}
-\sotodo{This is just Stieltjes right? We can probably delete it and reference Gautschi.}Inputs: max degree $N$, weight $w(x)$
-\begin{flalign*}
-	&\text{Define } H_0(x) \equiv \frac{1}{\norm{1}_w} \\
-	&\text{Define } v(x) = x  H_0(x) \\
-	&\text{Set } \alpha_0 = \ip<v, H_0>_w \\
-	&\text{Define } \tilde{v}(x) = v(x) - \alpha_0 H_0(x) \\
-	&\text{Set } \beta_0 = \norm{\tilde{v}}_w \\
-	&\text{Define } H_1(x) = \frac{\tilde{v}(x)}{\beta_0} \\
-	&\text{FOR } n = 1:N \\
-		& \quad \quad \text{Define } v(x) = x H_n - \beta_{n-1} H_{n-1} \\
-		& \quad \quad \text{Set } \alpha_n = \ip<v, H_n>_w \\
-		& \quad \quad \text{Define } \tilde{v}(x) = v(x) - \alpha_n H_n(x) \\
-		& \quad \quad \text{Set } \beta_n = \norm{\tilde{v}}_w \\
-		& \quad \quad \text{Define } H_{n+1}(x) = \frac{\tilde{v}(x)}{\beta_n} \\
-	&\text{END }
-\end{flalign*}
-Outputs: OP set $(H_0, \dots, H_N)$, coefficients $(\alpha_0, \dots, \alpha_N)$ and $(\beta_0, \dots, \beta_N)$
-\end{algorithm}
 
@@
 \end{proof}
+
+Note here that due to the symmetry in the domain, the integral of an odd function in the second variable (i.e.
+if $f(x, \dot)$ is odd for fixed $x$) over $\Omega$ is zero.
+Thus, the quadrature rule is such that we only consider the even part of a function.
 
@@
 
-Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: \sotodo{Add sentence explaining how symmetry is imprtant here}
+Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: 
 \begin{align}
@@
 
-\subsection{Calculating non-zero entries of the operator matrices}
-
-First, note that we can expand $\pddx \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
+\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
+
+Since $\{P^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \Pnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
-By a simple integration by parts argument, we have that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
-These are then calculated by taking inner products, i.e.
+We can use an integration by parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
+First, note that
 \begin{align*}
-c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}
+c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
 \end{align*}
-where \sotodo{Why giant $\prime$?}
+Second, note that
+\begin{align}
+	\rho'(x) = -x \: \rho(x)^{-2}. \label{eqn:rhoderivative}
+\end{align}
+Now, using a change of variables and (\ref{eqn:rhoderivative}), we have that
 \begin{align}
 	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \ip< x \: H_{n-k}^{(a, b+k+\half)} \prime (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+	&= \ip< x \: H_{n-k}^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
@@
 	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
-		\: \ip< y \: P_k^{(b,b)} \prime (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
-\end{align}
-We can then use quadrature to calculate the 1D inner products. This is much cheaper than using the 2D quadrature rule.
-
-Further, we can expand $\pddy \Pnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
+\end{align}
+We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
+Now, using integration by parts and noting that the weight $w_H^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
+\begin{align*}
+	&\ip< x \: H_{n-k}^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}}  \\
+	&= \ip<H_{n-k}^{(a, b+k+\half) \: \prime}, \: w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half)} >_1 \\
+	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \ddx \Big[w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half)} \Big] >_1 \\
+	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \Big[(a+1) \: w_H^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: w_H^{(a+2, b+\half(k+j+1))} \Big] \: H_{m-j}^{(a, b+1+j+\half)} >_1 \\
+		&\quad \quad \quad + \ip<H_{n-k}^{(a, b+k+\half)}, \: w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half) \: \prime}>_1 \\
+	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \Big[(a+1) \: w_H^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: w_H^{(2, \half(j-k))} \Big] \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+k+\half)}} \\
+		&\quad \quad \quad + \ip<H_{n-k}^{(a, b+k+\half)}, \: w_H^{(1, 1+\half(j-k))} \: H_{m-j}^{(a, b+1+j+\half) \: \prime}>_{w_H^{(a, b+k+\half)}}
+\end{align*}
+which is zero for $m < n - 2$, and further that
+\begin{align*}
+	&\ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} \\
+	&= \ip< H_{n-k}^{(a, b+k+\half)}(x), \: w_H^{(2, \half(j-k))} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+k+\half)}}
+\end{align*}
+which is also zero for $m < n - 2$.
+%, and since the weight $w_H^{(a+1, b+1)}$ will vanish at $x=0,1$ for non-negative parameters $a,b$, we have that for a polynomial $g(x,y)$ of degree $m$ on $\Omega$: 
+%\begin{align*}
+%	&\ip<\pddx \Pnkab, g >_{W^{(a+1, b+1)}} \\
+%	&= \int_0^1 \int_{-\rho(x)}^{\rho(x)} \pddx f \: g \: W^{(a+1,b+1)} \: \D y \: \D x \\
+%	&= \int_{-1}^1 \Big( \int_0^1 \pddx \Pnkab \: g \: w_H^{(a+1,b+1)} \: \rho(x) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
+%	&= - \int_{-1}^1 \Big( \int_0^1 \Pnkab \: \pddx \big(g \: w_H^{(a+1,b+1)}(x) \: \rho(x)\big) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
+%	&= - \int_{-1}^1 \Big( \int_0^1 \Pnkab \: \pddx \big(g \: \rho(x) \: w_H^{(a+1,b+1)}(x) \: w_P^{(b+1)}(y)\big) \: \D x  \Big) \: \D y \\
+%	&= - \int_0^1 \int_{-\rho(x)}^{\rho(x)} \Pnkab \: \pddx \big(g \: \rho(x) \: W^{(a+1,b+1)} \big) \: \rho(x)^{-1} \: \D y \: \D x \\
+%	&= - \iint_\Omega \Pnkab \: \big[\pddx g \: W^{(a+1,b+1)} + g \: \pddx W^{(a+1,b+1)} + \: g \rho'(x) \rho(x)^{-1} \: W^{(a+1, b+1)} \big] \: \D y \: \D x \\
+%	&= - \ip<\Pnkab,\: \rho'(x) \rho(x)^{-1} \: W^{(1, 1)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:weightderivativex}) \\
+%	&= - \ip<\Pnkab,\: W^{(2, 0)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:rhoderivative})
+%\end{align*}
+%Now, as the polynomial on the RHS of the inner product is of degree $m+2$ and $\Pnkab$ are orthogonal, we have that the above is zero for $m+2 < n$, i.e $c_{m,j}$ is non-zero for $m = n-1, n-2$ only.
+
+We can then use quadrature to calculate the 1D inner products and find the non-zero $c_{m,j}$ entries to the partial differential operator matrix.
+This is much cheaper than using the 2D quadrature rule.
+
+For partial differentiation by $y$, we can also expand $\pddy \Pnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
-By a simple integration by parts argument, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
+By a similar argument to before, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
 This is then calculated by taking inner products, i.e.
 \begin{align*}
-c_{n-1,k-1} = \frac{\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
+c_{n-1,k-1} = \frac{\ip<\pddy \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
 \end{align*}
-where \sotodo{Why giant $\prime$?}
-\begin{align}
-	\ip<\pddx \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
-		\: \ip< P_k^{(b,b)} \prime, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
-\end{align}
-where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ \sotodo{Aren't these $= 1$ since we orthonormalize?} $\forall m, n = 0, 1, 2, \dots$.
+where
+\begin{align}
+	\ip<\pddy \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
+		\: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
+\end{align}
+where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ $\forall m, n = 0, 1, 2, \dots$, and thus is in fact given by the integral of the weight function $w_H^{(a,b)}$ over $(0,1)$.
 
@@
 \begin{align}
-	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \\
-	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y)
+	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \label{eqn:weightderivativex} \\
+	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y) \label{eqn:weightderivativey}
 \end{align}
@@
 
-Find \(u(x,y)\) given a function \(f(x,y)\) such that:\sotodo{I assume there should be an $f$ in the right-hand side?}
+Find \(u(x,y)\) given a function \(f(x,y)\) such that:
 \begin{align}
 	\begin{cases}
-    		\Delta^2 u(x,y) = 0 \quad \text{in } \Omega \\
+    		\Delta^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
 		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
@@
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-We can construct a matrix operator for $\Delta^2$ that will take coefficients in the space
-$
-\bigW^{(2,2)}
-$
-to coefficients in the space
-$
-\bigP^{(2,2)}.
-$
-Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$.
-The matrix operator for the Biharmonic operator is then given by \sotodo{Move up to other discussion on constructing Biharmonic}
-\begin{align}
-	\biharmonic = \laplaceoo \: \laplacewtt.
-\end{align}
 In Figure \ref{fig:solution-biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
@@
 Note the since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-
-\subsection{Two-element p-FEM on the disk (without BCs)}
-
-\sotodoinline{Delete section?}
-We now consider the case of using two elements to discretise a unit disk domain $\Omega$, that is the each element $\element$ is a half-disk, with common edge defined by the line $x=0$ for $y \in [-1,1]$.
-We take inspiration from \cite{karniadakis2013spectral, beuchler2006new}.
-To ensure continuity across the edge interface, we need to expand our basis functions to be used.
-We require basis functions that correspond to each edge (in this case there is only one edge), that is basis functions with support as the two elements connected by said edge (for more than two elements - and hence more than one edge - we would also stipulate that each edge's corresponding basis functions would vanish on any other edge).
-We then also require the remaining (interior) basis functions be such that they have support of one element only, and vanish on (each) edge.
-Using this, our basis functions for $\le N$-degree polynomials can then be written as
-\begin{align}
-	\bigQab_n(x,y) := \begin{pmatrix}
-		P^{(a,b)}_{n,n}(x,y) \\
-		x \: \bigPab_{n-1}(x,y)
-	\end{pmatrix} \in \R^{n+1}, 
-	\quad \quad 
-	\bigQab := \begin{pmatrix}
-		\bigQab_0 \\
-		\bigQab_1 \\
-		\vdots \\
-		\bigQab_N
-	\end{pmatrix}.
-\end{align}
-We note that for each $n = 0,\dots,N$, $P^{(a,b)}_{n,n}(x,y) = \rho(x)^n \: P^{(a,b)}_{n}(\frac{y}{\rho(x)})$ and so at $x=0$ these become just the Jacobi polynomials in $y$, thus span the edge interface.
-The polynomials that make up the $\bigQab$ basis can also be shown to have similar sparse relations and yield similar sparse operator matrices.
-We can then proceed to use a similar approach to section A.1 on a reference element.
 

### 2019-04-11 papers/halfdisk/halfdisk.tex
New figures

@@
 \begin{figure}
-	\begin{subfigure}[t]{0.45\textwidth}
-	\includegraphics[scale=0.4]{sparsityoflaplacian-w11}
+	\centering
+	\begin{subfigure}[t]{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.35]{sparsityoflaplacian-w11}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.45\textwidth}
-	\includegraphics[scale=0.4]{sparsityoflaplacian-w22}
+	\begin{subfigure}[t]{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.35]{sparsityofhelmholtz}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.45\textwidth}
-	\includegraphics[scale=0.4]{sparsityoflaplacian-00}
+	\begin{subfigure}[t]{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.35]{sparsityofbiharmonic}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.45\textwidth}
-	\includegraphics[scale=0.4]{sparsityofhelmholtz}
-	\end{subfigure}
-	\centering
-    	\caption{\sotodoinline{Fix titles} "Spy" plots of the different operator matrices for the Laplacian operator $\Delta$, as well as the operator matrix for the Helmholtz equation, showing their sparsity.
-Top Left: $\laplacewii$.
-Top Right: $\laplacewtt$.
-Bottom Left: $\laplaceoo$.
-Bottom Right: the weighted variable coefficient Helmholtz operator $\laplacewii + k^2 \: T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}$ for $v(x,y) = xy^2$ and $k = 0.25$.}
+    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
+Left: the Laplace operator $\laplacewii$.
+Centre: the weighted variable coefficient Helmholtz operator $\laplacewii + k^2 \: T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
+Right: the biharmonic operator $\biharmonic$.}
         \label{fig:sparsity}

### 2019-04-11 papers/halfdisk/halfdisk.tex
Changed naming convention for the OPs

@@
 
-\newcommand{\Pnk}{P_{n,k}}
-\newcommand{\Pab}{P^{(a,b)}}
-\newcommand{\Pnkab}{P_{n,k}^{(a,b)}}
+\newcommand{\hdop}{H}
+\newcommand{\bighdop}{\mathbb{\hdop}}
+\newcommand{\hdopnk}{\hdop_{n,k}}
+\newcommand{\hdopab}{\hdop^{(a,b)}}
+\newcommand{\hdopnkab}{\hdop_{n,k}^{(a,b)}}
 \newcommand{\Wab}{{W^{(a,b)}}}
-\newcommand{\Pmjab}{P_{m,j}^{(a,b)}}
+\newcommand{\hdopmj}{\hdop_{m,j}}
+\newcommand{\hdopmjab}{\hdop_{m,j}^{(a,b)}}
 \newcommand{\alphaab}{\alpha^{(a,b)}}
 \newcommand{\betaab}{\beta^{(a,b)}}
-\newcommand{\bigPab}{\bigP^{(a,b)}}
+\newcommand{\bighdopab}{\bighdop^{(a,b)}}
 \newcommand{\Dnt}{D^\top_n}
 \newcommand{\Wii}{W^{(1,1)}}
-\newcommand{\Pii}{P^{(1,1)}}
-\newcommand{\bigPii}{{\mathbb{P}^{(1,1)}}}
-\newcommand{\Poo}{P^{(0,0)}}
-\newcommand{\bigPoo}{{\mathbb{P}^{(0,0)}}}
-\newcommand{\Poi}{P^{(0,1)}}
-\newcommand{\Pio}{P^{(1,0)}}
+\newcommand{\hdopii}{\hdop^{(1,1)}}
+\newcommand{\bighdopii}{{\mathbb{\hdop}^{(1,1)}}}
+\newcommand{\hdopoo}{\hdop^{(0,0)}}
+\newcommand{\bighdopoo}{{\mathbb{\hdop}^{(0,0)}}}
+\newcommand{\hdopoi}{\hdop^{(0,1)}}
+\newcommand{\hdopio}{\hdop^{(1,0)}}
+\newcommand{\genjac}{R}
+\newcommand{\genjacnmk}{\genjac_{n-k}}
+\newcommand{\genjacmmj}{\genjac_{m-j}}
+\newcommand{\genjacw}{w_\genjac}
+
 \newcommand{\dx}{\frac{\partial}{\partial x}}
@@
 $$
-with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
+with corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
 $$
-P_{n,k}^{(a,b)}(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
+\hdopnkab(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
 $$
@@
 
-In this section we outline the construction and some basic properties of $P_{n,k}^{(a,b)}(x,y)$.
+In this section we outline the construction and some basic properties of $\hdopnkab(x,y)$.
 The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.
@@
 $$
-\Pnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
+\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
 $$
@@
 \end{align*}
-with $\rho : (\alpha, \beta) \to \R$. Further, let $w^{(a,b,c)}_H(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
+with $\rho : (\alpha, \beta) \to \R$.
+Further, let $\genjacw^{(a,b,c)}(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
 \begin{align}
 \begin{cases}
-w^{(a,b,c)}_H(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
+\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
 w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b
@@
 \end{align}
-where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{H_n^{(a,b,c)}\}$ are orthogonal with respect to $w^{(a,b,c)}_H(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
+where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{\genjac_n^{(a,b,c)}\}$ are orthogonal with respect to $\genjacw^{(a,b,c)}(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
 
@@
 \begin{align}
-	\Pnk^{(a,b,c,d)}(x,y) := H_{n-k}^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \begin{align}
-	W^{(a,b,c,d)}(x,y) := w^{(a,b,c+d)}_H(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega. 
+	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega. 
 \end{align}
@@
 \rho(x) &:= (1-x^2)^{\half} \\
-w_1(x) &:= w_H^{(a,b)}(x) := x^a \: (1-x^2)^b \\
+w_1(x) &:= \genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b \\
 w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
@@
 \end{align}
-Note here the we have made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
+Note here the we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
 This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.
@@
 The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\).
-The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials.
+The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation \(\{\genjac_n^{(a, b)}\}\) for the orthonormal polynomials.
 Thus we arrive at  a two parameter family of 2D orthogonal polynomials\sotodo{We should probably use a different letter than $P$ to avoid confusion with Jacobi polynomials}
-\ \(\{\Pnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\ \(\{\hdopnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
 \begin{align}\label{eq:diskpolys}
- \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+ \hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \begin{align}
-W^{(a,b)}(x,y) = w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
+W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
 \end{align}
@@
 
-We can express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
-\begin{align}
-x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) 
+We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as
+\begin{align}
+x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) 
 \label{eqn:Hrecurrence} \\
@@
 \end{align}
-where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:
+where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $\hdopnkab(x,y)$:
 
 \begin{lemma}
-$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
-\begin{align}
-x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
-y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
-		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
-		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
+$\hdopnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
+\begin{align}
+x \hdopnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: \hdop_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: \hdop_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b)}(x, y), \\
+y \hdopnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: \hdop_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
+		& \quad \quad + \betaab_{n,k,3} \: \hdop_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: \hdop_{n, k+1}^{(a,b)}(x, y) \nonumber \\
+		& \quad \quad + \betaab_{n,k,5} \: \hdop_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: \hdop_{n+1, k+1}^{(a,b)}(x, y),
 \end{align}
@@
 \alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-\betaab_{n,k,1} &:= {\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k}^{(a, b+k-1/2)}>_{w_H^{(a, b+k+1/2)}} \\
-\betaab_{n,k,2} &:= {\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k-2}^{(a, b+k+{3/2})}>_{w_H^{(a, b+k+{3/2})}} \\
-\betaab_{n,k,3} &:= {\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k+1}^{(a, b+k-1/2)}>_{w_H^{(a, b+k+1/2)}} \\
-\betaab_{n,k,4} &:= {\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k-1}^{(a, b+k+{3/2})}>_{w_H^{(a, b+k+{3/2})}} \\
-\betaab_{n,k,5} &:= {\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k+2}^{(a, b+k-1/2)}>_{w_H^{(a, b+k+1/2)}} \\
-\betaab_{n,k,6} &:= {\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<H_{n-k}^{(a, b+k+1/2)}, H_{n-k}^{(a, b+k+{3/2})}>_{w_H^{(a, b+k+{3/2})}}. 
-\end{align}
-where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{w_H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight of $w_H^{(a, b)}$.
+\betaab_{n,k,1} &:= {\delta_{k-1}^{(b)}}{|| \hdop_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
+\betaab_{n,k,2} &:= {\delta_{k}^{(b)}}{|| \hdop_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
+\betaab_{n,k,3} &:= {\delta_{k-1}^{(b)}}{|| \hdop_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
+\betaab_{n,k,4} &:= {\delta_{k}^{(b)}}{|| \hdop_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
+\betaab_{n,k,5} &:= {\delta_{k-1}^{(b)}}{|| \hdop_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
+\betaab_{n,k,6} &:= {\delta_{k}^{(b)}}{|| \hdop_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
+\end{align}
+where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{\genjacw^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight of $\genjacw^{(a, b)}$.
 
@@
 \begin{proof}
-Note that \(\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
-\begin{align}
-\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &=  \iint_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
-&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
+Note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
+\begin{align}
+\ip<y \hdopnkab, \hdopmjab>_\Wab &=  \iint_\Omega \hdopnkab(x,y) \: \hdopmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
+&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(s) \: \genjacmmj^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
 & \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
 &= \begin{cases}
-    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{w_H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
-	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{w_H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
+    	\delta_k^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
+	\delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
 	0 & \text{otherwise}
@@
 \begin{align}
-\bigPab_n := \begin{pmatrix}
-		P^{(a,b)}_{n,-n}(x,y) \\
+\bighdopab_n := \begin{pmatrix}
+		\hdopab_{n,0}(x,y) \\
 		\vdots \\
-		P^{(a,b)}_{n,n}(x,y)
+		\hdopab_{n,n}(x,y)
 	\end{pmatrix} \in \R^{n+1}, 
 \quad \quad 
-\bigPab := \begin{pmatrix}
-		\bigPab_0 \\
-		\bigPab_1 \\
-		\bigPab_2 \\
+\bighdopab := \begin{pmatrix}
+		\bighdopab_0 \\
+		\bighdopab_1 \\
+		\bighdopab_2 \\
 		\vdots \\
@@
 \begin{align}
-J_x^{(a,b)} \: \bigPab(x,y) = x \: \bigPab(x,y), \quad J_y^{(a,b)} \: \bigPab(x,y) = y \: \bigPab(x,y).
+J_x^{(a,b)} \: \bighdopab(x,y) = x \: \bighdopab(x,y), \quad J_y^{(a,b)} \: \bighdopab(x,y) = y \: \bighdopab(x,y).
 \label{eqn:jacobimatricesdefinition}
 \end{align}
-The matrices $J_x^{(a,b)}, J_y^{(a,b)}$ act on the coefficients vector of a functions expansion in the $\{\Pnkab\}$ basis.
-For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
-Then $x \: f(x,y)$ is approximated by $\bigPab(x,y)^\top {J_x^{(a,b)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\Pnkab\}$ basis.
+The matrices $J_x^{(a,b)}, J_y^{(a,b)}$ act on the coefficients vector of a functions expansion in the $\{\hdopnkab\}$ basis.
+For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
+Then $x \: f(x,y)$ is approximated by $\bighdopab(x,y)^\top {J_x^{(a,b)\top}} \mathbf{f}$.
+In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\hdopnkab\}$ basis.
 
@@
 \mright]
-\bigPab(x,y)
+\bighdopab(x,y)
 =
@@
 \end{align}
-where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
+where we note \(P^{(a,b)}_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
 \begin{align}
@@
 \begin{cases}
-\bigPab_{-1}(x,y) := 0 \\
-\bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
-\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n(x,y) - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots.
+\bighdopab_{-1}(x,y) := 0 \\
+\bighdopab_{0}(x,y) := P^{(a,b)}_{0,0} \\
+\bighdopab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopab_n(x,y) - \Dnt C_n  \,\bighdopab_{n-1}(x,y), \quad n = 1,2,3,\dots.
 \end{cases}
@@
 \end{align}
-It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby construct $\bigPab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
+It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby construct $\bighdopab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
 
@@
 \end{tikzpicture} 
-\caption{The Laplace operator acting on vectors of $P_{n,k}=P_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{P^{(2,2)}_{n,k}}$ coefficients.
-Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{P_{n,k}^{(a,b)}}$ coefficients, where $(a,b)$ is at the tail of the arrow, and the range is $\smash{P_{n,k}^{(\tilde{a},\tilde{b})}}$ coefficients, where $(\tilde{a},\tilde{b})$ is at the head of the arrow.
-%For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{P_{n,k}^{(0,0,0)}}$ (resp.~$\smash{P_{n,k}^{(1,0,1)}}$) and the range in $\smash{P_{n,k}^{(1,0,1)}}$ (resp.~$\smash{P_{n,k}^{(2,0,2)}}$).
+\caption{The Laplace operator acting on vectors of $\hdop_{n,k}=\hdop_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{P^{(2,2)}_{n,k}}$ coefficients.
+Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{\hdopnkab}$ coefficients, where $(a,b)$ is at the tail of the arrow, and the range is $\smash{\hdop_{n,k}^{(\tilde{a},\tilde{b})}}$ coefficients, where $(\tilde{a},\tilde{b})$ is at the head of the arrow.
+%For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{\hdop_{n,k}^{(0,0,0)}}$ (resp.~$\smash{\hdop_{n,k}^{(1,0,1)}}$) and the range in $\smash{\hdop_{n,k}^{(1,0,1)}}$ (resp.~$\smash{\hdop_{n,k}^{(2,0,2)}}$).
 }
@@
 \begin{align}
-\bigWab(x,y) := \Wab(x,y) \bigPab(x,y).
-\end{align}
-
-Recall that, for $a, b$ be general parameters, a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$.
-Recall also that $\Wab(x,y) := w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $w_H^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
+\bigWab(x,y) := \Wab(x,y) \bighdopab(x,y).
+\end{align}
+
+Recall that, for $a, b$ be general parameters, a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
+Recall also that $\Wab(x,y) := \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $\genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
 
@@
 \begin{align}
-{\partial f \over \partial x}&= \bigP^{(a+1,b+1)}(x,y)^\top \: D_x^{(a,b)} \: \mathbf{f}, \\
-{\partial f \over \partial y} &= \bigP^{(a,b+1)}(x,y)^\top \: D_y^{(a,b)} \: \mathbf{f}, \\
+{\partial f \over \partial x}&= \bighdop^{(a+1,b+1)}(x,y)^\top \: D_x^{(a,b)} \: \mathbf{f}, \\
+{\partial f \over \partial y} &= \bighdop^{(a,b+1)}(x,y)^\top \: D_y^{(a,b)} \: \mathbf{f}, \\
 {\partial \over \partial x}[\Wab(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1)}(x,y)^\top \: W_x^{(a,b)} \: \mathbf{f}, \\
@@
 \begin{align}
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^\top \: \bigP^{(a+1,b)}(x,y) \\
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^\top \: \bigP^{(a,b+1)}(x,y) \\
-\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigP^{(a+1,b+1)}(x,y) \\
+\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^\top \: \bighdop^{(a+1,b)}(x,y) \\
+\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^\top \: \bighdop^{(a,b+1)}(x,y) \\
+\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bighdop^{(a+1,b+1)}(x,y) \\
 \bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^\top \: \bigW^{(a-1,b)}(x,y) \\
@@
 $$
-\bigWii(x,y) = \Wii(x,y) \: \bigPii(x,y)
+\bigWii(x,y) = \Wii(x,y) \: \bighdopii(x,y)
 $$
-to coefficients in the non-weighted space $\bigPii(x,y)$.
+to coefficients in the non-weighted space $\bighdopii(x,y)$.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$.
@@
 We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions.
-On this note, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bigPoo(x,y)$ to coefficients in the space $\bigP^{(2,2)}(x,y)$.
+On this note, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopoo(x,y)$ to coefficients in the space $\bighdop^{(2,2)}(x,y)$.
 We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
@@
 \end{align}
-Further represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bigPoo$.
+Further represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopoo$.
 Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
@@
 \end{align}
-We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bigP^{(2,2)}$.
+We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdop^{(2,2)}$.
 Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$.
@@
 
-\subsection{Constructing $H_n^{(a,b)}(x)$}
-
-To obtain the recurrence coefficients for the $\{H_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
-\begin{align}
-x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
+\subsection{Constructing $\genjac_n^{(a,b)}(x)$}
+
+To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
+\begin{align}
+x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
 \end{align}
@@
 
-In this section we construct a quadrature rule exact for polynomials in $\Omega$ that can be used to expand functions in $P_{n,k}^{(a,b)}(x,y)$.
+In this section we construct a quadrature rule exact for polynomials in $\Omega$ that can be used to expand functions in $\hdopnkab(x,y)$.
 
@@
 \begin{align*}
-f(x,y) = \sum_{n=0}^N \bigP_n^{(a,b)}(x,y)^\top \: \mathbf{f}_n
+f(x,y) = \sum_{n=0}^N \bighdop_n^{(a,b)}(x,y)^\top \: \mathbf{f}_n
 \end{align*}
@@
 \begin{align*}
-\bigP_n(x,y) &:= \begin{pmatrix}
-		P_{n,0}(x,y) \\
+\bighdop_n(x,y) &:= \begin{pmatrix}
+		\hdop_{n,0}(x,y) \\
 		\vdots \\
-		P_{n,n}(x,y)
+		\hdop_{n,n}(x,y)
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
@@
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
-f_{n,k} := \frac{\langle f, \: \Pnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \Pnk^{(a,b)} ||_{W^{(a,b)}}}
+f_{n,k} := \frac{\langle f, \: \hdopnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}}
 \end{align*}
@@
 \begin{align}
-	\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j) \big], \\
-	|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big].
+	\ip< f, \: \hdopnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j) \big], \\
+	|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \hdopnk^{(a,b)}(x_j, y_j)^2 + \hdopnk^{(a,b)}(x_j, -y_j)^2 \big].
 \end{align}
@@
 %\begin{align}
-%\ip< f, \: \Pnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \Pnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \Pnk^{(a,b)}(x_j, -y_j)\big) \big], \\
-%|| \Pnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \Pnk^{(a,b)}(x_j, y_j)^2 + \Pnk^{(a,b)}(x_j, -y_j)^2 \big],
+%\ip< f, \: \hdopnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j)\big) \big], \\
+%|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \hdopnk^{(a,b)}(x_j, y_j)^2 + \hdopnk^{(a,b)}(x_j, -y_j)^2 \big],
 %\end{align}
@@
 
-Since $\{P^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \Pnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
+Since $\{P^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 \begin{align*}
-c_{m,j} = \frac{\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
+c_{m,j} = \frac{\ip<\pddx \hdopnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
 \end{align*}
@@
 \begin{align}
-	&\ip<\pddx \Pnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \ip< x \: H_{n-k}^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+	&\ip<\pddx \hdopnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
+	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad - k \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad + \: \ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} 
+	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
@@
 We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
-Now, using integration by parts and noting that the weight $w_H^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
+Now, using integration by parts and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
 \begin{align*}
-	&\ip< x \: H_{n-k}^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}}  \\
-	&= \ip<H_{n-k}^{(a, b+k+\half) \: \prime}, \: w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half)} >_1 \\
-	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \ddx \Big[w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half)} \Big] >_1 \\
-	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \Big[(a+1) \: w_H^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: w_H^{(a+2, b+\half(k+j+1))} \Big] \: H_{m-j}^{(a, b+1+j+\half)} >_1 \\
-		&\quad \quad \quad + \ip<H_{n-k}^{(a, b+k+\half)}, \: w_H^{(a+1, b+1+\half(k+j+1))} \: H_{m-j}^{(a, b+1+j+\half) \: \prime}>_1 \\
-	&= \ip<H_{n-k}^{(a, b+k+\half)}, \: \Big[(a+1) \: w_H^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: w_H^{(2, \half(j-k))} \Big] \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+k+\half)}} \\
-		&\quad \quad \quad + \ip<H_{n-k}^{(a, b+k+\half)}, \: w_H^{(1, 1+\half(j-k))} \: H_{m-j}^{(a, b+1+j+\half) \: \prime}>_{w_H^{(a, b+k+\half)}}
+	&\ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
+	&= \ip<\genjacnmk^{(a, b+k+\half) \: \prime}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \ddx \Big[\genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} \Big] >_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \\
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}}
 \end{align*}
@@
 \begin{align*}
-	&\ip< x^2 \: H_{n-k}^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+\half)}} \\
-	&= \ip< H_{n-k}^{(a, b+k+\half)}(x), \: w_H^{(2, \half(j-k))} \: H_{m-j}^{(a, b+1+j+\half)} >_{w_H^{(a, b+k+\half)}}
+	&\ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
+	&= \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
 \end{align*}
 which is also zero for $m < n - 2$.
-%, and since the weight $w_H^{(a+1, b+1)}$ will vanish at $x=0,1$ for non-negative parameters $a,b$, we have that for a polynomial $g(x,y)$ of degree $m$ on $\Omega$: 
+%, and since the weight $\genjacw^{(a+1, b+1)}$ will vanish at $x=0,1$ for non-negative parameters $a,b$, we have that for a polynomial $g(x,y)$ of degree $m$ on $\Omega$: 
 %\begin{align*}
-%	&\ip<\pddx \Pnkab, g >_{W^{(a+1, b+1)}} \\
+%	&\ip<\pddx \hdopnkab, g >_{W^{(a+1, b+1)}} \\
 %	&= \int_0^1 \int_{-\rho(x)}^{\rho(x)} \pddx f \: g \: W^{(a+1,b+1)} \: \D y \: \D x \\
-%	&= \int_{-1}^1 \Big( \int_0^1 \pddx \Pnkab \: g \: w_H^{(a+1,b+1)} \: \rho(x) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
-%	&= - \int_{-1}^1 \Big( \int_0^1 \Pnkab \: \pddx \big(g \: w_H^{(a+1,b+1)}(x) \: \rho(x)\big) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
-%	&= - \int_{-1}^1 \Big( \int_0^1 \Pnkab \: \pddx \big(g \: \rho(x) \: w_H^{(a+1,b+1)}(x) \: w_P^{(b+1)}(y)\big) \: \D x  \Big) \: \D y \\
-%	&= - \int_0^1 \int_{-\rho(x)}^{\rho(x)} \Pnkab \: \pddx \big(g \: \rho(x) \: W^{(a+1,b+1)} \big) \: \rho(x)^{-1} \: \D y \: \D x \\
-%	&= - \iint_\Omega \Pnkab \: \big[\pddx g \: W^{(a+1,b+1)} + g \: \pddx W^{(a+1,b+1)} + \: g \rho'(x) \rho(x)^{-1} \: W^{(a+1, b+1)} \big] \: \D y \: \D x \\
-%	&= - \ip<\Pnkab,\: \rho'(x) \rho(x)^{-1} \: W^{(1, 1)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:weightderivativex}) \\
-%	&= - \ip<\Pnkab,\: W^{(2, 0)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:rhoderivative})
+%	&= \int_{-1}^1 \Big( \int_0^1 \pddx \hdopnkab \: g \: \genjacw^{(a+1,b+1)} \: \rho(x) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
+%	&= - \int_{-1}^1 \Big( \int_0^1 \hdopnkab \: \pddx \big(g \: \genjacw^{(a+1,b+1)}(x) \: \rho(x)\big) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
+%	&= - \int_{-1}^1 \Big( \int_0^1 \hdopnkab \: \pddx \big(g \: \rho(x) \: \genjacw^{(a+1,b+1)}(x) \: w_P^{(b+1)}(y)\big) \: \D x  \Big) \: \D y \\
+%	&= - \int_0^1 \int_{-\rho(x)}^{\rho(x)} \hdopnkab \: \pddx \big(g \: \rho(x) \: W^{(a+1,b+1)} \big) \: \rho(x)^{-1} \: \D y \: \D x \\
+%	&= - \iint_\Omega \hdopnkab \: \big[\pddx g \: W^{(a+1,b+1)} + g \: \pddx W^{(a+1,b+1)} + \: g \rho'(x) \rho(x)^{-1} \: W^{(a+1, b+1)} \big] \: \D y \: \D x \\
+%	&= - \ip<\hdopnkab,\: \rho'(x) \rho(x)^{-1} \: W^{(1, 1)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:weightderivativex}) \\
+%	&= - \ip<\hdopnkab,\: W^{(2, 0)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:rhoderivative})
 %\end{align*}
-%Now, as the polynomial on the RHS of the inner product is of degree $m+2$ and $\Pnkab$ are orthogonal, we have that the above is zero for $m+2 < n$, i.e $c_{m,j}$ is non-zero for $m = n-1, n-2$ only.
+%Now, as the polynomial on the RHS of the inner product is of degree $m+2$ and $\hdopnkab$ are orthogonal, we have that the above is zero for $m+2 < n$, i.e $c_{m,j}$ is non-zero for $m = n-1, n-2$ only.
 
@@
 
-For partial differentiation by $y$, we can also expand $\pddy \Pnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+For partial differentiation by $y$, we can also expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 \begin{align*}
-c_{n-1,k-1} = \frac{\ip<\pddy \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
+c_{n-1,k-1} = \frac{\ip<\pddy \hdopnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
 \end{align*}
@@
 \begin{align}
-	\ip<\pddy \Pnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{H_0^{(a,b+k+\half)}}^2_{w_H^{(a, b+k+\half)}} 
+	\ip<\pddy \hdopnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{\genjac_0^{(a,b+k+\half)}}^2_{\genjacw^{(a, b+k+\half)}} 
 		\: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
 \end{align}
-where we note that $\norm{H_n^{(a,b)}}_{w_H^{(a,b)}} = \norm{H_m^{(a,b)}}_{w_H^{(a,b)}}$ $\forall m, n = 0, 1, 2, \dots$, and thus is in fact given by the integral of the weight function $w_H^{(a,b)}$ over $(0,1)$.
+where we note that $\norm{\genjac_n^{(a,b)}}_{\genjacw^{(a,b)}} = \norm{\genjac_m^{(a,b)}}_{\genjacw^{(a,b)}}$ $\forall m, n = 0, 1, 2, \dots$, and thus is in fact given by the integral of the weight function $\genjacw^{(a,b)}$ over $(0,1)$.
 
@@
 \begin{align}
-	\ip< \Wab \Pnkab, W^{(\tilde{a},\tilde{b})} P^{(\tilde{a},\tilde{b})}_{m,j} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \Pnkab, P^{(\tilde{a},\tilde{b})}_{m,j} >_\Wab.
+	\ip< \Wab \hdopnkab, W^{(\tilde{a},\tilde{b})} \hdopmj^{(\tilde{a},\tilde{b})} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wab.
 \end{align}
@@
 Fix \(a, b = 1,1\).
-Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$.
+Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{f}$.
 Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
@@
 We can tackle the problem as follows.
-Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$.
+Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{f}$.
 Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule above.
 We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
-We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$.
+We can denote the resulting operator acting on coefficients in the $\bighdopoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$.
 In matrix-vector notation, our system hence becomes:
@@
 Denote the coefficient vector for expansion of $u_0$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ we denote $\mathbf{u}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bigWii$ OP basis up to degree $N$.
-Further, denote the coefficient vector for expansion of $u_0$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{\tilde{u}}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ denote $\mathbf{\tilde{u}}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bigPii$ OP basis up to degree $N$.
+Further, denote the coefficient vector for expansion of $u_0$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{\tilde{u}}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ denote $\mathbf{\tilde{u}}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bighdopii$ OP basis up to degree $N$.
 
@@
 
-Define $\Lambda^{(a,b)} := \langle \bigP^{(a,b)}, \: {\bigP^{(a,b)}}^\top \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bigP^{(a,b)}$ are orthogonal with respect to.
+Define $\Lambda^{(a,b)} := \langle \bighdop^{(a,b)}, \: {\bighdop^{(a,b)}}^\top \rangle_{W^{(a,b)}}$ where $W^{(a,b)}$ is the weight with which the OPs in $\bighdop^{(a,b)}$ are orthogonal with respect to.
 Note that due to orthogonality this is a diagonal matrix.
@@
 \bigWii_n(x,y) &:= \begin{pmatrix}
-		\Wii(x,y) \: \Pii_{n,0}(x,y) \\
+		\Wii(x,y) \: \hdopii_{n,0}(x,y) \\
 		\vdots \\
-		\Wii(a,y) \: \Pii_{n,n}(x,y)
+		\Wii(a,y) \: \hdopii_{n,n}(x,y)
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
@@
 	&= \int_\element \begin{pmatrix}
-					\bigPoo^\top \Wii_x \mathbf{v} \\
-					\bigPoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
+					\bighdopoo^\top \Wii_x \mathbf{v} \\
+					\bighdopoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
 				\end{pmatrix}^\top 
 				\begin{pmatrix}
-					\bigPoo^\top \Wii_x \mathbf{u} \\
-					\bigPoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
+					\bighdopoo^\top \Wii_x \mathbf{u} \\
+					\bighdopoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
 				\end{pmatrix}
 				\\
-	&= \int_\element \Big( \mathbf{v}^\top {\Wii_x}^\top \bigPoo \bigPoo^\top \Wii_x \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^\top ({T^{(1,0)\to(0,0)} \Wii_y})^\top \bigPoo \bigPoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
+	&= \int_\element \Big( \mathbf{v}^\top {\Wii_x}^\top \bighdopoo \bighdopoo^\top \Wii_x \mathbf{u} \nonumber \\
+					& \quad \quad \quad + \mathbf{v}^\top ({T^{(1,0)\to(0,0)} \Wii_y})^\top \bighdopoo \bighdopoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
 	&= \mathbf{v}^\top \: \Big( {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
@@
 	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
-	&= \int_\element \: \mathbf{v}^\top \: \bigWii \: \bigPii^\top \: \mathbf{f} \: d\mathbf{x} \\
-	&= \mathbf{v}^\top \: \langle \bigPii, {\bigPii}^\top \rangle_{\Wii} \: d\mathbf{x} \\
+	&= \int_\element \: \mathbf{v}^\top \: \bigWii \: \bighdopii^\top \: \mathbf{f} \: d\mathbf{x} \\
+	&= \mathbf{v}^\top \: \langle \bighdopii, {\bighdopii}^\top \rangle_{\Wii} \: d\mathbf{x} \\
 	&= \mathbf{v}^\top \Lambda^{(1,1)} \: \mathbf{f},
 \end{align}
-where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bigPii$ OP basis.
+where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bighdopii$ OP basis.
 
@@
 \begin{align}
-	L(\Wii \Pii_{n,k}) = a(u_h,\Wii \Pii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
+	L(\Wii \hdopii_{n,k}) = a(u_h,\Wii \hdopii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
 \end{align}
@@
 \begin{cases}
-w^{(a,b,c)}_H(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
+\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
 w^{(a)}_P(x) &:= (1-x^2)^a.
@@
 The weight $w^{(a)}_P(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(a, a)}\}$.
-$w^{(a,b,c)}_H(x)$ is the (non-classical) weight for the OPs denoted $\{H_n^{(a,b,c)}\}\).
-Thus we arrive at the three-parameter family of 2D orthogonal polynomials $\{\Pnk^{(a,b,c)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align}
-	\Pnk^{(a,b,c)}(x,y) := H_{n-k}^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: P_k^{(c,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+$\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs denoted $\{\genjac_n^{(a,b,c)}\}\).
+Thus we arrive at the three-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\begin{align}
+	\hdopnk^{(a,b,c)}(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: P_k^{(c,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 
-We can note that by setting the parameter $c = 0$, and $\alpha = 0$, $\beta = 1$ we obtain the 2D orthogonal polynomials on the half-disk (note again also that we made the adjustment that $w^{(a,b,c)}_H(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ for the half-disk case, for ease of writing and computation).
+We can note that by setting the parameter $c = 0$, and $\alpha = 0$, $\beta = 1$ we obtain the 2D orthogonal polynomials on the half-disk (note again also that we made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ for the half-disk case, for ease of writing and computation).
 
@@
 \rho(x) &:= 1-\half x \\
-w^{(a,b,c)}_H(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\half x)^{c} \\
+\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\half x)^{c} \\
 w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
@@
 We note that the shifted Jacobi polynomials relate the the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = P_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
-$w^{(a,b,c)}_H(x)$ is the (non-classical) weight for the OPs we dentote $\{H_n^{(a,b,c)}\}\).
-Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\Pnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align}
-	\Pnk^{(a,b,c,d)}(x,y) := H_{n-k}^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+$\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\).
+Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\begin{align}
+	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \begin{align}
-	W^{(a,b,c,d)}(x,y) &:= w^{(a,b,c+d)}_H(x) \: w^{(d,c)}_P(\frac{y}{\rho(x)}) \nonumber \\
+	W^{(a,b,c,d)}(x,y) &:= \genjacw^{(a,b,c+d)}(x) \: w^{(d,c)}_P(\frac{y}{\rho(x)}) \nonumber \\
 	&= (1-x)^a \: x^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega.
@@
 %	\bigQab_n(x,y,z) := \begin{pmatrix}
-%		\bigPab_n(x,y) \\
-%		z \: \bigP^{(a,b+1)}_{n-1}(x,y)
+%		\bighdopab_n(x,y) \\
+%		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
 %	\end{pmatrix} \in \R^{2n+1}, 
@@
 %\begin{align}
-%	z \: \Pnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: P^{(a,b+1)}_{m,j}(x,y), \\
-%	z^2 \: \Pnkab(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b)\to(a,b-1)} \: z \: P^{(a,b-1)}_{m,j}(x,y).
+%	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: P^{(a,b+1)}_{m,j}(x,y), \\
+%	z^2 \: \hdopnkab(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b)\to(a,b-1)} \: z \: P^{(a,b-1)}_{m,j}(x,y).
 %\end{align}
@@
 %\begin{align}
-%	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \Pnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
-%	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
-%	&= \frac{\ip< \Pnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
-%	&= \frac{\norm{\Pnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)}.
+%	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
+%	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \hdopnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
+%	&= \frac{\ip< \hdopnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
+%	&= \frac{\norm{\hdopnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)}.
 %\end{align}
@@
 %\begin{align}
-%	\bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bigP^{(a,b+1)}(x,y) \\
-%	W^{(0,1)}(x,y) \bigP^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)} \Big)^T \: \bigP^{(a,b-1)}(x,y)
+%	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y) \\
+%	W^{(0,1)}(x,y) \bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)} \Big)^T \: \bighdop^{(a,b-1)}(x,y)
 %\end{align}

### 2019-04-15 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Changed how HalfDiskTransform plan and associated multiplication is done - this means we dont crash for large N

@@ -36,7 +36,7 @@ end
 
 
 # Finds the OPs and recurrence for weight w, having already found N₀ OPs
-function lanczos!(w, P, β, γ, N₀=0)
+function lanczos!(w, P, β, γ; N₀=0)
 
     # x * P[n](x) == (γ[n] * P[n+1](x) + β[n] * P[n](x) + γ[n-1] * P[n-1](x))
 
@@ -56,11 +56,13 @@ function lanczos!(w, P, β, γ, N₀=0)
     end
 
     for k = N₀+1:N
-        v = x*P[k] - γ[k-1]*P[k-1]
-        β[k] = sum(w*v*P[k])
-        v = v - β[k]*P[k]
-        γ[k] = sqrt(sum(w*v^2))
-        P[k+1] = v/γ[k]
+        @show k
+        v = x*P[2] - γ[k-1]*P[1]
+        β[k] = sum(P[2]*w*v)
+        v = v - β[k]*P[2]
+        γ[k] = sqrt(sum(w*v*v))
+        P[1] = P[2]
+        P[2] = v/γ[k]
     end
 
     P, β, γ
@@ -89,8 +91,8 @@ function resizedata!(S::OrthogonalPolynomialSpace, n)
     n ≤ N₀ && return S
     resize!(S.a, n)
     resize!(S.b, n)
-    resize!(S.ops, n + 1)
-    lanczos!(S.weight, S.ops, S.a, S.b, N₀)
+    resize!(S.ops, 2)
+    lanczos!(S.weight, S.ops, S.a, S.b, N₀=N₀)
     S
 end
 
@@ -513,46 +515,54 @@ function getopnorms(S::HalfDiskSpace{<:Any, <:Any, T}, k) where T
 end
 
 struct HalfDiskTransformPlan{T}
-    U::Diagonal{T,Vector{T}}
-    VTp::Matrix{T}
-    W::Diagonal{T,Vector{T}}
-    VTm::Matrix{T}
+    w::Vector{T}
+    pts::Vector{SArray{Tuple{2},T,1,2}}
+    S::HalfDiskSpace{<:Any, <:Any, T}
 end
 
 function HalfDiskTransformPlan(S::HalfDiskSpace{<:Any, <:Any, T}, vals) where T
-    m2 = Int(length(vals) / 2)
-    N = Int(sqrt(m2)) - 1
-    m1 = Int((N+1)*(N+2) / 2)
-    pts, w = pointswithweights(S, m2)
-
[…]

### 2019-04-15 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Tried to speed up OperatorClenshaw()

@@ -1368,7 +1368,7 @@ end
 #===#
 
 # Operator Clenshaw
-function operatorclenshawG(S::HalfDiskSpace{<:Any, <:Any, T}, n, Jx, Jy) where T
+function operatorclenshawG(S::HalfDiskSpace{<:Any, <:Any, T}, n, Jx, Jy, zeromat) where T
     @show "G", n
     G = Matrix{SparseMatrixCSC{T}}(undef, 2(n+1), n+1)
     for i = 1:n+1
@@ -1377,27 +1377,39 @@ function operatorclenshawG(S::HalfDiskSpace{<:Any, <:Any, T}, n, Jx, Jy) where T
                 G[i,j] = Jx
                 G[i+n+1,j] = Jy
             else
-                G[i,j] = zeros(size(Jx))
-                G[i+n+1,j] = zeros(size(Jy))
+                G[i,j] = zeromat
+                G[i+n+1,j] = zeromat
             end
         end
     end
     G
 end
-function converttooperatorclenshawvector(S::HalfDiskSpace{<:Any, <:Any, T}, v, Jx) where T
-    nn = size(Jx)
+function operatorclenshawvector(S::HalfDiskSpace{<:Any, <:Any, T}, v, id) where T
     s = size(v)[1]
     B = Array{SparseMatrixCSC{T}}(undef, (1, s))
     for i = 1:s
-        B[1,i] = sparse(I, nn) * v[i]
+        B[1,i] = id * v[i]
     end
     B
 end
-function converttooperatorclenshawmatrix(S::HalfDiskSpace{<:Any, <:Any, T}, A, Jx) where T
-    nn = size(Jx)
+function operatorclenshawmatrixDT(S::HalfDiskSpace{<:Any, <:Any, T}, A, id) where T
     B = Array{SparseMatrixCSC{T}}(undef, size(A))
     for ij = 1:length(A)
-        B[ij] = sparse(I, nn) * A[ij]
+        B[ij] = id * A[ij]
+    end
+    B
+end
+function operatorclenshawmatrixBmG(S::HalfDiskSpace{<:Any, <:Any, T}, A, id, Jx, Jy) where T
+    ii, jj = size(A)
+    B = Array{SparseMatrixCSC{T}}(undef, (ii, jj))
+    for i = 1:jj, j = 1:jj
+        if i == j
+            B[i, j] = (id * A[i, j]) - Jx
[…]

### 2019-04-15 papers/halfdisk/halfdisk.tex
New figures. Comments added.

@@
 The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation \(\{\genjac_n^{(a, b)}\}\) for the orthonormal polynomials.
-Thus we arrive at  a two parameter family of 2D orthogonal polynomials\sotodo{We should probably use a different letter than $P$ to avoid confusion with Jacobi polynomials}
-\ \(\{\hdopnkab\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+Thus we arrive at  a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
 \begin{align}\label{eq:diskpolys}
@@
 \end{align}
-where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $\hdopnkab(x,y)$:
+where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$:
 
@@
     	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 0.25$ \sotodoinline{Make a lot higher}.
+(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 200$.
 (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that\bstodo{Describe what plot tells us}
+We see that...
+\bstodoinline{Describe what plot tells us - I dont want to copy and paste your Triangle paper here but not sure what else to write}
 
@@
 This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.
-\sotodoinline{More blocks}}
+\bstodoinline{N=200 blocks here.
+When I have the time to run N=1000 or so I will do}}
         	\label{fig:solutionblocknorms}

### 2019-04-15 experiments/halfdisk.jl
jl file that i use to get the figs etc.

@@ -1,123 +1,240 @@
-using OrthogonalPolynomialFamilies, ApproxFun
+using Revise
+using ApproxFun
 using Test, Profile
 using StaticArrays, LinearAlgebra, BlockArrays, BlockBandedMatrices,
         SparseArrays
 # using Makie
-
+using OrthogonalPolynomialFamilies
 import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
-                    opevalatpts, inner2, laplace, getopindex, getnk, resizecoeffs!,
-                    convertweightedtononoperator, increaseparamsoperator,
-                    transformoperator, laplacesquare, convertweightedtononoperatorsquare
+                    opevalatpts, inner2, getopindex, getnk, resizecoeffs!,
+                    transformparamsoperator, weightedpartialoperatorx,
+                    weightedpartialoperatory, partialoperatorx, partialoperatory,
+                    laplaceoperator, derivopevalatpts, getderivopptseval, getopnorms,
+                    getopnorm, operatorclenshaw, weight, biharmonicoperator
+using SpecialFunctions
 
 #===#
-# Speed up the code! Add storage of the basis evaluated at the points from
-# points(), and use in a new basiseval method (using 3-term recurrence)
 
-a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
-x, y = 0.4, -0.2; z = [x; y]
-N = 10
-laplace(D, N)
+# Set up for experiments
+precision(BigFloat)
+setprecision(800)
+D = HalfDiskFamily()
+a, b = 1.0, 1.0; S = D(a, b)
+x, y = 0.42, -0.2; z = [x; y]
+N = 25 # Sparsity plots
+@time bihar = biharmonicoperator(D(2.0, 2.0), N)
+N = 100 # Solution plots
+@time bihar = biharmonicoperator(D(2.0, 2.0), N)
[…]

### 2019-04-17 papers/halfdisk/halfdisk.tex
Further edits

@@
 %
-\section{Future work - orthogonal polynomials on half-spheres and spherical triangles}
+\section{Conclusions}
 
@@
 
-We will use a similar idea of projecting up from 1D OPs to the interval, to 2D OPs on the disk, to 3D OPs on the sphere.
-Initially, we will add to the ApproxFun library by adding capability to work with OPs on the half disk (2D) with weights governed by parameters that will be incremented/decremented after an application of an operator (e.g.
-partial derivative).
+We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on certain domains given by an algebraic curve - notably here the half-disk - with Dirichlet boundary conditions.
+We can develop $hp-$finite element methods methods to solve PDEs on other polygonal domains by using suitable shaped elements (for example, by using disk-slice elements on the disk, which has applications in fluids).
+
+A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
 With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
-An extension to this will be to look at a $p$-finite element method on the sphere using two half-sphere elements.
-We plan to extend this work to construct similar methods on other 3D spherical domains, such as spherical caps and spherical triangles, which will have applications in numerical weather prediction.
-
-Other extensions include a full $hp$-finite element method on sections of a disk which has applications in fluids.
+
+This work serves as a stepping stone to constructing similar methods on other 3D spherical domains, such as spherical caps and spherical triangles, which will have applications in numerical weather prediction.
 

### 2019-04-17 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Changed some show statements

@@ -56,7 +56,7 @@ function lanczos!(w, P, β, γ; N₀=0)
     end
 
     for k = N₀+1:N
-        @show k
+        @show "lanczos", N, k
         v = x*P[2] - γ[k-1]*P[1]
         β[k] = sum(P[2]*w*v)
         v = v - β[k]*P[2]
@@ -542,7 +542,7 @@ function *(P::HalfDiskTransformPlan{T}, vals) where T
     getopnorms(P.S, N)
     for i = 1:m2
         if i % 100 == 0
-            @show i, m2
+            @show m2, i
         end
         pt = [P.pts[i]]
         getopptseval(P.S, N, pt)
@@ -732,13 +732,13 @@ function resizedata!(S::HalfDiskSpace, N)
     resize!(S.C, N + 2)
     resize!(S.DT, N + 2)
     getBs!(S, N, N₀)
-    @show "hereB"
+    @show "resizedata!", "done getBs!"
     getCs!(S, N, N₀)
-    @show "hereC"
+    @show "resizedata!", "done getCs!"
     getAs!(S, N, N₀)
-    @show "hereA"
+    @show "resizedata!", "done getAs!"
     getDTs!(S, N, N₀)
-    @show "hereDT"
+    @show "resizedata!", "done getDTs!"
     S
 end
 
@@ -932,7 +932,7 @@ function partialoperatorx(S::HalfDiskSpace{<:Any, <:Any, T}, N) where T
 
     # Get pt evals for the H OPs
     for k = 0:N
-        @show k
+        @show "dxoperator", N, k
         H = S.family.H(S.a, S.b+k+0.5)
         getopptseval(H, N-k, ptsh)
         getderivopptseval(H, N-k, ptsh)
@@ -1221,22 +1221,6 @@ function transformparamsoperator(S::HalfDiskSpace{<:Any, <:Any, T},
     end
 end
 
-function laplace(D::HalfDiskFamily, N)
-    A = partialoperatorx(D(0.0,0.0), N+2)
-    @show "A done"
-    B = weightedpartialoperatorx(D(1.0,1.0), N)
-    @show "B done"
-    C = transformparamsoperator(D(0.0,1.0), D(1.0,1.0), N+1)
-    @show "C done"
-    E = partialoperatory(D(0.0,0.0), N+2)
-    @show "E done"
-    F = transformparamsoperator(D(1.0,0.0), D(0.0,0.0), N+1, weighted=true)
[…]

### 2019-04-17 papers/halfdisk/halfdisk.tex
New figs. Some comments. New ref.

@@
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \sotodo{Add citation to this} \bstodo{should we instead take the Xu approach?}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
+In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \cite[Theorem 3.1]{olver2018orthogonal}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
     	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 200$.
+(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
 (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that...
-\bstodoinline{Describe what plot tells us - I dont want to copy and paste your Triangle paper here but not sure what else to write}
+We see that we achieve algebraic convergence for the first three examples, with the final Gaussian bump example achieving spectral convergence.
+We note that for right hand sides that vanish at the corners of our half disk ($y = \pm 1$) we see faster convergence.
+\bstodoinline{Improve above}
 
@@
 This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.
-\bstodoinline{N=200 blocks here.
-When I have the time to run N=1000 or so I will do}}
+\bstodoinline{N=200 blocks here}}
         	\label{fig:solutionblocknorms}
@@
 
-\sotodoinline{Rewrite in a more "paper" like tone}
+\bstodoinline{Rewritten}
 

### 2019-04-24 papers/halfdisk/halfdisk.tex
Some edits to correct how Jacobi operator entries are calculated

@@
 \newcommand{\genjacw}{w_\genjac}
+\newcommand{\normgenjac}{\omega_\genjac}
+\newcommand{\normjac}{\omega_P}
 
@@
 
+The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are not orthonormal, but have (squared) norms given by the integral of their respective weight functions we denote as
+\begin{align}
+	\normgenjac^{(a,b)} := \int_0^1 \: \genjacw^{(a,b)}(x) \: \D x, \quad \normjac^{(b)} := \int_{-1}^1 \: w_P^{(b)}(y) \: \D y,
+\end{align}
+i.e. $\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}}^2 = \normgenjac^{(a,b)}$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}}^2 = \normjac^{(b)}$ $\forall n = 0,1,2,\dots$.
+
+For clarity, we use the notation that 
+\begin{align*}
+&\ip<p, q>_{\genjacw^{(a,b)}} := \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \quad
+\ip<p, q>_{w_P^{(b)}} := \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y, \\
+&\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}, \quad
+\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}.
+\end{align*}
+
 \subsection{Jacobi matrices}
@@
 \alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-\betaab_{n,k,1} &:= {\delta_{k-1}^{(b)}}{|| \hdop_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
-\betaab_{n,k,2} &:= {\delta_{k}^{(b)}}{|| \hdop_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
-\betaab_{n,k,3} &:= {\delta_{k-1}^{(b)}}{|| \hdop_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
-\betaab_{n,k,4} &:= {\delta_{k}^{(b)}}{|| \hdop_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
-\betaab_{n,k,5} &:= {\delta_{k-1}^{(b)}}{|| \hdop_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
-\betaab_{n,k,6} &:= {\delta_{k}^{(b)}}{|| \hdop_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^{-2}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
-\end{align}
-where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{\genjacw^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight of $\genjacw^{(a, b)}$.
+\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
+\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
+\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
+\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
+\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
+\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
+\end{align}
 
@@
 \begin{proof}
-Note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
+First, note that, using a change of variable,
+\begin{align}
+	\norm{\hdopnkab}_\Wab^2 &= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}(\frac{y}{\rho(x)}}^2 \: \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
+	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(y)^2 \: w_P^{(b)}(y) \: \D y \Big) \nonumber \\
+	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}.
+\end{align}
+where we use the notation that $\ip<p, q>_{\Wab} := \iint_\Omega p(x,y) \: q(x,y) \: \Wab(x,y) \: \D y \: \D x$ and $\norm{p}_{W^{(a,b)}}^2 := \ip<p, p>_{\Wab}$ .
+
+Finally, note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
 \begin{align}
@@
 &= \begin{cases}
-    	\delta_k^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
-	\delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
+    	\delta_k^{(b)} \: \normjac^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
+	\delta_{k-1}^{(b)} \: \normjac^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
 	0 & \text{otherwise}
@@
 \begin{align*}
-c_{m,j} = \frac{\ip<\pddx \hdopnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{P^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
+c_{m,j} = \frac{\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
 \end{align*}
@@
 \begin{align}
-	&\ip<\pddx \hdopnkab, P^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
+	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
 	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
@@
 \begin{align*}
-c_{n-1,k-1} = \frac{\ip<\pddy \hdopnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{P^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
+c_{n-1,k-1} = \frac{\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
 \end{align*}
@@
 \begin{align}
-	\ip<\pddy \hdopnkab, P^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{\genjac_0^{(a,b+k+\half)}}^2_{\genjacw^{(a, b+k+\half)}} 
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{\genjac_0^{(a,b+k+\half)}}^2_{\genjacw^{(a, b+k+\half)}} 
 		\: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}

### 2019-04-25 papers/halfdisk/halfdisk.tex
Minor edit
Readthrough

@@
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
+We apply the framework to solving the Poisson, variable coefficient Helmholtz, and Biharmonic equations.
  \end{abstract}
@@
 More precisely, we consider the solution of partial differential equations on the domain
-\begin{align}
+\begin{align*}
 	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align}
-where we require that either of the following scenarios hold:
+\end{align*}
+where  either of the following scenarios hold:
 \begin{enumerate}
@@
 
-For simplicity of discussion we focus in this paper on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.
+For simplicity of presentation we focus on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, and discuss extension to other geometries in the appendix.
 
@@
 $$
-W^{(a,b)}(x,y) = x^a (1-x-y)^b
+W^{(a,b)}(x,y) = x^a (1-x-y)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
 $$
@@
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk \cite[Theorem 3.1]{olver2018orthogonal}, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
+In particular, OPs in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere  can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D.
+%, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
@@
 \noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the half-disk.
+
+\noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods.
+
+\noindent\appref{DiskSlices}: We discuss extension to disk slices.
+
+\noindent\appref{Trapezium}: We discuss extension to trapezia.
 
@@
 
-We can generalise this for our purposes in the following proposition.
+We can generalise\sotodo{specialise?} this for our purposes in the following proposition.
+\sotodoinline{Can this be simplified to just the definition?}
 \begin{proposition}
@@
 Further, let $\genjacw^{(a,b,c)}(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
-\begin{align}
+\begin{align*}
 \begin{cases}
@@
 \end{cases}
-\end{align}
+\end{align*}
 where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{\genjac_n^{(a,b,c)}\}$ are orthogonal with respect to $\genjacw^{(a,b,c)}(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
@@
 Then the four-parameter 2D polynomials defined on $\Omega$ given by
-\begin{align}
+\begin{align*}
 	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align}
+\end{align*}
 are orthogonal polynomials with respect to the weight 
-\begin{align}
+\begin{align*}
 	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega. 
-\end{align}
+\end{align*}
 \end{proposition}
@@
 For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
-\begin{align}
+\begin{align*}
 \begin{cases}
@@
 \end{cases}
-\end{align}
+\end{align*}
 Note here the we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
@@
 orthogonal on \(\Omega\) with respect to the weight
-\begin{align}
+\begin{align*}
 W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
-\end{align}
+\end{align*}
 
 The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are not orthonormal, but have (squared) norms given by the integral of their respective weight functions we denote as
-\begin{align}
+\begin{align*}
 	\normgenjac^{(a,b)} := \int_0^1 \: \genjacw^{(a,b)}(x) \: \D x, \quad \normjac^{(b)} := \int_{-1}^1 \: w_P^{(b)}(y) \: \D y,
-\end{align}
+\end{align*}
 i.e. $\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}}^2 = \normgenjac^{(a,b)}$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}}^2 = \normjac^{(b)}$ $\forall n = 0,1,2,\dots$.
@@
 $\hdopnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
-\begin{align}
+\begin{align*}
 x \hdopnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: \hdop_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: \hdop_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b)}(x, y), \\
@@
 		& \quad \quad + \betaab_{n,k,5} \: \hdop_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: \hdop_{n+1, k+1}^{(a,b)}(x, y),
-\end{align}
+\end{align*}
 for \((x,y) \in \Omega\), where
-\begin{align}
-\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
-\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
+\begin{align*}
+\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)}, \qquad \alphaab_{n,k,2} := \alpha_{n-k}^{(a, b+k+\half)} \\
 \betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
@@
 \betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
-\end{align}
+\end{align*}
 
@@
 First, note that, using a change of variable,
-\begin{align}
-	\norm{\hdopnkab}_\Wab^2 &= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}(\frac{y}{\rho(x)}}^2 \: \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
+\begin{align*}
+	\norm{\hdopnkab}_\Wab^2 &= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
 	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(y)^2 \: w_P^{(b)}(y) \: \D y \Big) \nonumber \\
 	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}.
-\end{align}
+\end{align*}
 where we use the notation that $\ip<p, q>_{\Wab} := \iint_\Omega p(x,y) \: q(x,y) \: \Wab(x,y) \: \D y \: \D x$ and $\norm{p}_{W^{(a,b)}}^2 := \ip<p, p>_{\Wab}$ .
@@
 Finally, note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
-\begin{align}
+\begin{align*}
 \ip<y \hdopnkab, \hdopmjab>_\Wab &=  \iint_\Omega \hdopnkab(x,y) \: \hdopmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
@@
       \end{cases}.
-\end{align}
+\end{align*}
 
@@
 Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
-\begin{align}
+\begin{align*}
 \bighdopab_n := \begin{pmatrix}
@@
 	\end{pmatrix}
-\end{align}
+\end{align*}
 and set $J_x^{(a,b)}, J_y^{(a,b)}$ as the Jacobi matrices corresponding to
@@
 
-$J_x^{(a,b)}, J_y^{(a,b)}$ have Banded-Block-Banded structures.
+$J_x^{(a,b)}, J_y^{(a,b)}$ have banded-block-banded structures.
 A block-banded matrix is a block $n$-diagonal matrix where $n$ is reasonably small for the size of the full matrix.
@@
 $J_x^{(a,b)}, J_y^{(a,b)}$ are given by:
-\begin{align}
+\begin{align*}
 J_{x/y}^{(a,b)} &= \begin{pmatrix}
@@
 	\end{pmatrix}
-\end{align}
+\end{align*}
 where
-\begin{align}
+\begin{align*}
 A^x_n &:= \begin{pmatrix}
@@
 	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
-\end{align}
+\end{align*}
 
@@
 Combining each system in (\ref{eqn:jacobimatricesdefinition}) we can write the block-tridiagonal system
-\begin{align}
+\begin{align*}
 \renewcommand\arraystretch{1.3}
@@
 \end{pmatrix}.
-\end{align}
+\end{align*}
 where we note \(P^{(a,b)}_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
-\begin{align}
+\begin{align*}
 A_n &:= \begin{pmatrix}
@@
 	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
-\end{align}
+\end{align*}
 
@@
 Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
-\begin{align}
+\begin{align*}
 \begin{cases}
@@
 \end{cases}
-\end{align}
+\end{align*}
 
 Note that we can define an explicit \(\Dnt\) as follows. For \(n\) even:
-\begin{align}
+\begin{align*}
 \Dnt := \begin{pmatrix}
@@
 	    \end{pmatrix},
-\end{align}
+\end{align*}
 where
-\begin{align}
+\begin{align*}
 m &= \frac{n}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}},\qqand \nonumber \\
 \eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-1.
-\end{align}
+\end{align*}
 
@@
 
-\begin{align}
+\begin{align*}
 \Dnt := \begin{pmatrix}
@@
 	    \end{pmatrix},
-\end{align}
+\end{align*}
  where
-\begin{align}
+\begin{align*}
 m &= \frac{n+1}{2} + 1,\quad  \eta_0 = \frac{1}{\betaab_{n,n,6}}, \quad \xi = -\frac{\betaab_{n,1,5} \: \eta_{m-2}}{\alphaab_{n+1,0,1}},\qqand \nonumber \\
 \eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-2.
-\end{align}
-It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby construct $\bighdopab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
+\end{align*}
+It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopab_{0}(x,y)$  through $\bighdopab_{n}(x,y)$ in optimal $O(n^2)$ complexity.
 
@@
 
-Firstly, we introduce further notation that
-\begin{align}
+Denote the weighted OPs by
+\begin{align*}
 \bigWab(x,y) := \Wab(x,y) \bighdopab(x,y).
-\end{align}
+\end{align*}
 
@@
 
-There exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
-\begin{align}
+There exist sparse \sotodo{Explain why these are sparse} operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
+\begin{align*}
 {\partial f \over \partial x}&= \bighdop^{(a+1,b+1)}(x,y)^\top \: D_x^{(a,b)} \: \mathbf{f}, \\
@@
 {\partial \over \partial y}[\Wab(x,y) \: f(x,y)] &= \bigW^{(a,b-1)}(x,y)^\top \: W_y^{(a,b)} \: \mathbf{f}.
-\end{align}
-The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite{DLMFDerivatives}, and on the triangle \cite{olver2018recurrence}.
-
-Further, we note that there exist ``parameter transformation" Banded-Block-Banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+\end{align*}
+The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
+
+Further, we note that there exist ``parameter transformation" banded-block-banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
-\begin{align}
+\begin{align*}
 \bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b)} \Big)^\top \: \bighdop^{(a+1,b)}(x,y) \\
@@
 \bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigW^{(a-1,b-1)}(x,y) \\
-\end{align}
+\end{align*}
 
@@
 Centre: the weighted variable coefficient Helmholtz operator $\laplacewii + k^2 \: T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
-Right: the biharmonic operator $\biharmonic$.}
+Right: the biharmonic operator $\biharmonic$.
+	\sotodoinline{Use PDFs for pictures to avoid resolution problems}
+	}
         \label{fig:sparsity}
@@
 The matrix operator for the Laplacian we denote $\laplacewii$ acting on the coefficients vector is then given by
-\begin{align}
+\begin{align*}
     \laplacewii := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)}.
-\end{align}
-Importantly, this operator will have Banded-Block-Banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+\end{align*}
+Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
 
@@
 We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
-\begin{align}
+\begin{align*}
     \laplaceoo := D_x^{(1,1)} \: D_x^{(0,0)} + T^{(1,2)\to(2,2)} \: D_y^{(1,1)} \: T^{(0,1)\to(1,1)} \: D_y^{(0,0)}.
-\end{align}
+\end{align*}
 Further represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopoo$.
@@
 We denote this matrix operator as $\laplacewtt$, and is given by
-\begin{align}
+\begin{align*}
 	\laplacewtt := W_x^{(1,1)} \: W_x^{(2,2)} + T_W^{(1,0)\to(0,0)} \: W_y^{(1,1)} \: T_W^{(2,1)\to(1,1)} \: W_y^{(2,2)}.
-\end{align}
+\end{align*}
 We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdop^{(2,2)}$.
@@
 The matrix operator for the Biharmonic operator is then given by
-\begin{align}
+\begin{align*}
 	\biharmonic = \laplaceoo \: \laplacewtt.
-\end{align}
+\end{align*}
 The sparsity and structure of this biharmonic operator is seen in Figure \ref{fig:sparsity}.
 
-The sparsity of the these operators can be shown using a simple integration by parts argument, see Section \ref{subsection:Computation-operatormatrices}.
+The sparsity of the these operators can be shown using a simple integration-by-parts argument, see Section \ref{subsection:Computation-operatormatrices}.
 
@@
 To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
-\begin{align}
+\begin{align*}
 x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
-\end{align}
+\end{align*}
 where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$.
@@
  on \([-1,1]\) with weight \((1-t^2)^b\) as $(t_k,w_k^{(t)})$. Define
-\begin{align}
+\begin{align*}
 x_{i+(k-1)N} &:= s_k, \quad i,k = 1,\dots,N, \\
@@
 w_{l+(k-1)N} &:= w_k^{(s)} w_l^{(t)}, \quad k,l = 1,\dots,N.
-\end{align}
+\end{align*}
 Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
@@
 We will use the substitution that
-\begin{align}
+\begin{align*}
 x &= s, \quad y = (1-s^2)^\half t.
-\end{align}
+\end{align*}
 First, note that
-\begin{align}
+\begin{align*}
 W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
 		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
-\end{align}
+\end{align*}
+
+\sotodoinline{Rest of proof?}
 \end{proof}
@@
 Note here that due to the symmetry in the domain, the integral of an odd function in the second variable (i.e.
-if $f(x, \dot)$ is odd for fixed $x$) over $\Omega$ is zero.
+if $f(x, \cdot)$ is odd for fixed $x$) over $\Omega$ is zero.
 Thus, the quadrature rule is such that we only consider the even part of a function.
@@
 Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: 
-\begin{align}
+\begin{align*}
 	\ip< f, \: \hdopnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j) \big], \\
 	|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \hdopnk^{(a,b)}(x_j, y_j)^2 + \hdopnk^{(a,b)}(x_j, -y_j)^2 \big].
-\end{align}
+\end{align*}
 
@@
 Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
-%\begin{align}
+%\begin{align*}
 %& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
@@
 %&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
-%\end{align}
+%\end{align*}
 %where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
@@
 %We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
-%\begin{align}
+%\begin{align*}
 %\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
-%\end{align}
+%\end{align*}
 %
@@
 %Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
-%\begin{align}
+%\begin{align*}
 %\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
-%\end{align}
+%\end{align*}
 %and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$.
@@
 Thus,
-%\begin{align}
+%\begin{align*}
 %\ip< f, \: \hdopnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j)\big) \big], \\
 %|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \hdopnk^{(a,b)}(x_j, y_j)^2 + \hdopnk^{(a,b)}(x_j, -y_j)^2 \big],
-%\end{align}
+%\end{align*}
 %for suitably large $N$.
@@
 \subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
+
+\sotodoinline{The first part of this section should be moved up to Section 3, and stated as a "Lemma". }
 
@@
 The coefficients of the expansion are then the entries of the relevant operator matrix.
-We can use an integration by parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
+We can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
 First, note that
@@
 Now, using a change of variables and (\ref{eqn:rhoderivative}), we have that
-\begin{align}
+\begin{align*}
 	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
@@
 		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
-\end{align}
+\end{align*}
 We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
-Now, using integration by parts and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
+Now, using integration-by-parts and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
 \begin{align*}
@@
 %, and since the weight $\genjacw^{(a+1, b+1)}$ will vanish at $x=0,1$ for non-negative parameters $a,b$, we have that for a polynomial $g(x,y)$ of degree $m$ on $\Omega$: 
-%\begin{align*}
+%\begin{align}
 %	&\ip<\pddx \hdopnkab, g >_{W^{(a+1, b+1)}} \\
@@
 %	&= - \ip<\hdopnkab,\: W^{(2, 0)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:rhoderivative})
-%\end{align*}
+%\end{align}
 %Now, as the polynomial on the RHS of the inner product is of degree $m+2$ and $\hdopnkab$ are orthogonal, we have that the above is zero for $m+2 < n$, i.e $c_{m,j}$ is non-zero for $m = n-1, n-2$ only.
@@
 where
-\begin{align}
-	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \norm{\genjac_0^{(a,b+k+\half)}}^2_{\genjacw^{(a, b+k+\half)}} 
-		\: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}
-\end{align}
-where we note that $\norm{\genjac_n^{(a,b)}}_{\genjacw^{(a,b)}} = \norm{\genjac_m^{(a,b)}}_{\genjacw^{(a,b)}}$ $\forall m, n = 0, 1, 2, \dots$, and thus is in fact given by the integral of the weight function $\genjacw^{(a,b)}$ over $(0,1)$.
+\begin{align*}
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \normgenjac^{(a, b+k+\half)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+\end{align*}
 
@@
 and also that
-\begin{align}
+\begin{align*}
 	\ip< \Wab \hdopnkab, W^{(\tilde{a},\tilde{b})} \hdopmj^{(\tilde{a},\tilde{b})} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wab.
-\end{align}
+\end{align*}
 
@@
 In matrix-vector notation, our system hence becomes:
-\begin{align}
+\begin{align*}
     \laplacewii \mathbf{u} = \mathbf{f}
-\end{align}
+\end{align*}
 which can be solved to find $\mathbf{u}$.
@@
 In matrix-vector notation, our system hence becomes:
-\begin{align}
+\begin{align*}
     (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
-\end{align}
+\end{align*}
 which can be solved to find $\mathbf{u}$.
@@
 We can extend this to constant non-zero boundary conditions by simply noting that the problem 
-\begin{align}
+\begin{align*}
 	\begin{cases}
@@
 	\end{cases}
-\end{align}
+\end{align*}
 is equivalent to letting $u = \tilde{u} + c$ and solving
-\begin{align}
+\begin{align*}
 	\begin{cases}
@@
 	\end{cases}.
-\end{align}
+\end{align*}
 
@@
 \subsection{Heat equation}
+
+\sotodoinline{Skip example?}
 Consider
-\begin{align}
+\begin{align*}
 	\begin{cases}
@@
 	\end{cases}
-\end{align}
+\end{align*}
 noting the imposition of zero Dirichlet boundary conditions.
@@
 That is, formulate this for \(t = 0,1,2,\dots\) and step length $h$:
-\begin{align}
+\begin{align*}
 u_{t+1}(x,y) - u_t (x,y) = h \: \Delta u_{t+1}(x,y) \quad \text{for } (x,y) \in \Omega.
-\end{align}
+\end{align*}
 
@@
 We can solve this system for the given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
-\begin{align}
+\begin{align*}
 	&T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (\mathbf{u}_{t+1} - \mathbf{u}_t) = h \: \laplacewii \mathbf{u}_{t+1} \nonumber \\
@@
 	\iff& \mathbf{\tilde{u}}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \laplacewii)^{-1} \: \mathbf{\tilde{u}}_t.
-\end{align}
+\end{align*}
 
@@
 %
-\section{P-finite element methods using sparse operators}
-
-We follow the method of \cite{beuchler2006new}. Consider the 2D Dirichlet problem on a domain $\Omega$:
-\begin{align}
+\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
+
+We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
+Consider the 2D Dirichlet problem on a domain $\Omega$:
+\begin{align*}
 	\begin{cases}
@@
          \end{cases}
-\end{align}
+\end{align*}
 This has the weak formulation for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$,
-\begin{align}
+\begin{align*}
 	L(v) := \int_\Omega f \: v \: d\mathbf{x} = \int_\Omega \nabla u \cdot \nabla v \: d\mathbf{x} =: a(u,v).
-\end{align}
+\end{align*}
 
@@
 
-\subsection{Single element p-FEM}
 In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/trapezium domain.
-We can choose our finite dimensional space $V_h = \{v_h \in V \quad | \quad deg(v_h|_\element) \le p\}$ for some $p \in \N$.
+We can choose our finite dimensional space\sotodo{Use of $h$ is mis-leading here.
+Maybe $V_p$?} $V_h = \{v_h \in V \quad | \quad {\rm deg}\,(v_h|_\element) \le p\}$ for some $p \in \N$.
 
@@
 We can choose a basis for $V_h$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
-\begin{align}
+\begin{align*}
 \bigWii(x,y) &:= \begin{pmatrix}
@@
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
-\end{align}
+\end{align*}
 and rewrite (\ref{eqn:FEMweakform}) in matrix form:
-\begin{align}
+\begin{align*}
 	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
@@
 	&= \mathbf{v}^\top \: \Big( {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
-\end{align}
+\end{align*}
 where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_h, v_h \in V_h$ respectively in the $V_h$ basis ($\bigWii$ OPs), and
-\begin{align}
+\begin{align*}
 	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
@@
 	&= \mathbf{v}^\top \Lambda^{(1,1)} \: \mathbf{f},
-\end{align}
+\end{align*}
 where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bighdopii$ OP basis.
@@
 Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-\begin{align}
+\begin{align*}
 	L(\Wii \hdopii_{n,k}) = a(u_h,\Wii \hdopii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
-\end{align}
+\end{align*}
 (i.e. holds for all basis functions of $V_h$) by choosing $v_h$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-\begin{align}
+\begin{align*}
 A\mathbf{u} = \tilde{\mathbf{f}}.
-\end{align}
+\end{align*}
 where the (element) stiffness matrix $A$ is defined by 
-\begin{align}
+\begin{align*}
 A = {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
-\end{align}
+\end{align*}
 and the load vector $\tilde{\mathbf{f}}$ is given by 
-\begin{align}
+\begin{align*}
 \tilde{\mathbf{f}} = \Lambda^{(1,1)} \: \mathbf{f}.
-\end{align}
+\end{align*}
 
@@
 
-\section{Disk slices}
+\section{Disk slices}\label{Appendix:DiskSlices}
 
 The work in this paper on the half disk can be easily transferred to the domain of a disk slice by which we mean
-\begin{align}
+\begin{align*}
 	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align}
+\end{align*}
 with 
-\begin{align}
+\begin{align*}
 \begin{cases}
@@
 \end{cases}
-\end{align}
+\end{align*}
 Our 1D weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively are then given by:
-\begin{align}
+\begin{align*}
 \begin{cases}
@@
 \end{cases}
-\end{align}
+\end{align*}
 
@@
 Thus we arrive at the three-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align}
-	\hdopnk^{(a,b,c)}(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: P_k^{(c,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align}
+\begin{align*}
+	\hdopnk^{(a,b,c)}(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: P_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+\end{align*}
 orthogonal with respect to the weight
-\begin{align}
-	W^{(a,b,c)}(x,y) &:= w^{(a,b,2c)}_1(x) w^{(c)}_2(\frac{y}{\rho(x)}) \nonumber \\
+\begin{align*}
+	W^{(a,b,c)}(x,y) &:= w^{(a,b,2c)}_1(x) w^{(c)}_2\fpr(\frac{y}{\rho(x)}) \nonumber \\
 	&= (\beta - x)^a \: (x - \alpha)^{b} \: (\rho(x)^2 -y^2)^c \nonumber \\
 	&= (\beta - x)^a \: (x - \alpha)^{b} \: (1 - x^2 -y^2)^c , \quad (x,y) \in \Omega.
-\end{align}
+\end{align*}
 
@@
 
-\section{Trapeziums}
+\sotodoinline{Comment on how the sparse operators generalise to such domains}
+
+\section{Trapeziums}\label{Appendix:Trapezium}
 
 We can further extend this work to trapezium shaped domains. Note that for any trapezium there exists an affine map to the canonical trapezium domain that we consider here, given by
-\begin{align}
+\begin{align*}
 	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align}
+\end{align*}
 with 
-\begin{align}
+\begin{align*}
 \begin{cases}
@@
 \end{cases}
-\end{align}
+\end{align*}
 The weight $w^{(a,b)}_P(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
@@
 Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align}
+\begin{align*}
 	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align}
+\end{align*}
 orthogonal with respect to the weight
-\begin{align}
+\begin{align*}
 	W^{(a,b,c,d)}(x,y) &:= \genjacw^{(a,b,c+d)}(x) \: w^{(d,c)}_P(\frac{y}{\rho(x)}) \nonumber \\
 	&= (1-x)^a \: x^b \: y^c \: (1-\half x - y)^d, \quad (x,y) \in \Omega.
-\end{align}
+\end{align*}
 
@@
 
-\sotodoinline{Fix DLMF References.}
-\sotodoinline{Abbreviate Journal titles.} 
+\sotodoinline{More Xu references.}
+\sotodoinline{Add missing references to Bib file}
 \bibliography{halfdisk-bib}

### 2019-04-25 papers/halfdisk/halfdisk.tex
addressed some of SO comments

@@
 \begin{align*}
-x &= s, \quad y = (1-s^2)^\half t.
-\end{align*}
-First, note that
-\begin{align*}
-W^{(a,b)}(x,y) &= x^a \: (1-x^2-y^2)^b, \quad \text{for } (x,y) \in \Omega, \\
-		      &= s^a \: (1-s^2)^{b} \: (1-t^2)^b =: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
-\end{align*}
-
-\sotodoinline{Rest of proof?}
+	x &= s, \quad y = \rho(s) \: t.
+\end{align*}
+First, note that, for $(x,y) \in \Omega$,
+\begin{align*}
+	W^{(a,b)}(x,y) &= \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \\
+	&= \genjacw^{(a,b)}(s) \: w_P^{(b)}(t) \\
+	&= s^a \: \rho(s)^{2b} \: (1-t^2)^b \\
+	&=: V^{(a,b)}(s,t), \quad \text{for } (s,t) \in [0,1] \times [-1,1].
+\end{align*}
+Let $f : \Omega \to \R$ be such that $y \mapsto f(x,y)$ for fixed $x$ is an even function. Then,
+\begin{align*}
+	\iint_\Omega f(x,y) \: \Wab(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_\gamma^\delta f\big(s,\rho(s) t\big) \: V^{(a,b)}(s,t) \: \rho(s) \: \D t \: \D s \\
+	&= \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \Big( \int_\gamma^\delta f\big(s,\rho(s) t\big) \: w_P^{(b)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
+	&\approx \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \sum_{k=1}^N \Big( w_k^{(t)} f\big(s,\rho(s) t_k\big) \Big) \: \D s \quad (\star) \\
+	&\approx \sum_{k=1}^N \Bigg( w_k^{(s)} \: \sum_{l=1}^N \Big( w_l^{(t)} f\big(s_k,\rho(s_k) t_l\big) \Big) \Bigg) \quad (\star \star) \\
+	&= \sum_{j=1}^{N^2}  w_j \: f(x_j, y_j),
+\end{align*}
+where we achieve equality at $(\star)$ if $y \mapsto f(x,y)$ for fixed $x$ is a polynomial of degree at most $2N-1$ and we achieve equality at $(\star \star)$ if $x \mapsto f(x,y)$ for fixed $y$ is a polynomial of degree at most $N$.
+Note also that the inner integral at $(\dagger)$ the inner integral over $t$ would be zero if $y \mapsto f(x,y)$ for fixed $x$ was an odd function, due to the symmetry over the domain (i.e.
+since $\gamma = -\delta$ in the this case).
+
+\bstodoinline{I cannot remember/work out why it was necessary to split f into odd and even parts?}
 \end{proof}
-
-Note here that due to the symmetry in the domain, the integral of an odd function in the second variable (i.e.
-if $f(x, \cdot)$ is odd for fixed $x$) over $\Omega$ is zero.
-Thus, the quadrature rule is such that we only consider the even part of a function.
 
@@
 
-\bstodoinline{Rewritten}
-
 We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on certain domains given by an algebraic curve - notably here the half-disk - with Dirichlet boundary conditions.
@@
 In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/trapezium domain.
-We can choose our finite dimensional space\sotodo{Use of $h$ is mis-leading here.
-Maybe $V_p$?} $V_h = \{v_h \in V \quad | \quad {\rm deg}\,(v_h|_\element) \le p\}$ for some $p \in \N$.
-
-We seek $u_h \in V_h$ s.t.
+We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.
+
+We seek $u_p \in V_p$ s.t.
 \begin{align}
-	L(v_h) = a(u_h,v_h) \quad \forall \: v_h \in V_h.
+	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
 	\label{eqn:FEMweakform}
@@
 Note that due to orthogonality this is a diagonal matrix.
-We can choose a basis for $V_h$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
+We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
 \begin{align*}
@@
 \begin{align*}
-	a(u_h,v_h) &= \int_\element \nabla u_h \cdot \nabla v_h \: d\mathbf{x} \\
+	a(u_p,v_p) &= \int_\element \nabla u_p \cdot \nabla v_p \: d\mathbf{x} \\
 	&= \int_\element \begin{pmatrix}
-					\partial_x v_h \\
-					\partial_y v_h
+					\partial_x v_p \\
+					\partial_y v_p
 				\end{pmatrix}^\top 
 				\begin{pmatrix}
-					\partial_x u_h \\
-					\partial_y u_h
+					\partial_x u_p \\
+					\partial_y u_p
 				\end{pmatrix}
@@
 \end{align*}
-where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_h, v_h \in V_h$ respectively in the $V_h$ basis ($\bigWii$ OPs), and
-\begin{align*}
-	L(v_h) &= \int_\element \: v_h \: f \: d\mathbf{x} \\
+where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWii$ OPs), and
+\begin{align*}
+	L(v_p) &= \int_\element \: v_p \: f \: d\mathbf{x} \\
 	&= \int_\element \: \mathbf{v}^\top \: \bigWii \: \bighdopii^\top \: \mathbf{f} \: d\mathbf{x} \\
@@
 \begin{align*}
-	L(\Wii \hdopii_{n,k}) = a(u_h,\Wii \hdopii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
-\end{align*}
-(i.e. holds for all basis functions of $V_h$) by choosing $v_h$ as each basis function, we can equivalently write the linear system for our finite element problem as:
+	L(\Wii \hdopii_{n,k}) = a(u_p,\Wii \hdopii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
+\end{align*}
+(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
 \begin{align*}
@@
 \begin{align*}
-	W^{(a,b,c)}(x,y) &:= w^{(a,b,2c)}_1(x) w^{(c)}_2\fpr(\frac{y}{\rho(x)}) \nonumber \\
+	W^{(a,b,c)}(x,y) &:= \genjacw^{(a,b,2c)}(x) w^{(c)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
 	&= (\beta - x)^a \: (x - \alpha)^{b} \: (\rho(x)^2 -y^2)^c \nonumber \\
@@
 
-We can note that by setting the parameter $c = 0$, and $\alpha = 0$, $\beta = 1$ we obtain the 2D orthogonal polynomials on the half-disk (note again also that we made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ for the half-disk case, for ease of writing and computation).
+We note again that by making the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$, and setting the first parameter $a$ to zero and removing it for the family and weight for $\{\genjac_n^{(a,b,c)}\}$, and taking and $\alpha = 0$, $\beta = 1$, we recover the half-disk case.
 

### 2019-04-29 papers/halfdisk/halfdisk.tex
Further edits

@@
 
-\begin{proposition}
+\begin{proposition}\label{prop:construction}
 Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
@@
 
-We can generalise\sotodo{specialise?} this for our purposes in the following proposition.
-\sotodoinline{Can this be simplified to just the definition?}
-\begin{proposition}
+For disk slices and trapeziums, we can specialise the framework from Proposition \ref{prop:construction} in the following definition.
+\begin{definition}\label{def:construction}
 Let
@@
 \end{enumerate}
-Then the four-parameter 2D polynomials defined on $\Omega$ given by
+We define the four-parameter 2D orthogonal polynomials on $\Omega$ by
 \begin{align*}
@@
 \end{align*}
-are orthogonal polynomials with respect to the weight 
+which are orthogonal with respect to the weight
 \begin{align*}
@@
 \end{align*}
-\end{proposition}
-\begin{proof}
-See \cite[p55--56]{dunkl2014orthogonal}.
-\end{proof}
+\end{definition}
 
@@
 
-There exist sparse \sotodo{Explain why these are sparse} operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that:
+\begin{definition}\label{def:sparsityofdifferentialoperators}
+Define the operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ according to:
 \begin{align*}
@@
 \end{align*}
+\end{definition}
+
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
 
-Further, we note that there exist ``parameter transformation" banded-block-banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+\begin{lemma}\label{lemma:sparsityofdifferentialoperators}
+The operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ from Definition \ref{def:sparsityofdifferentialoperators} are sparse, with banded-block-banded structure.
+\end{lemma}
+
+\begin{proof}
+We proceed with the case for the operator $D_x^{(a,b)}$ for partial differentiation by $x$.
+Since $\{\hdop^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+1, b+1)}_{m,j}$.
+The coefficients of the expansion are then the entries of the relevant operator matrix.
+We can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
+First, note that
+\begin{align*}
+c_{m,j} = \frac{\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
+\end{align*}
+Second, note that
+\begin{align}
+	\rho'(x) = -x \: \rho(x)^{-2}. \label{eqn:rhoderivative}
+\end{align}
+Now, using a change of variables and (\ref{eqn:rhoderivative}), we have that
+\begin{align*}
+	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
+	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
+	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
+	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
+\end{align*}
+We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
+Now, using integration-by-parts and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
+\begin{align*}
+	&\ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
+	&= \ip<\genjacnmk^{(a, b+k+\half) \: \prime}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \ddx \Big[\genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} \Big] >_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \\
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}}
+\end{align*}
+which is zero for $m < n - 2$, and further that
+\begin{align*}
+	&\ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
+	&= \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
+\end{align*}
+which is also zero for $m < n - 2$.
+
+\bstodoinline{Finish this proof!!!}
+
+For the operator for partial differentiation by $y$, we can also expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+The coefficients of the expansion are then the entries of the relevant operator matrix.
+By a similar argument to before, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
+This is then calculated by taking inner products, i.e.
+\begin{align*}
+c_{n-1,k-1} = \frac{\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
+\end{align*}
+where
+\begin{align*}
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \normgenjac^{(a, b+k+\half)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+\end{align*}
+
+We can gain the non-zero entries of the weighted differential operators similarly, by noting that
+\begin{align}
+	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \label{eqn:weightderivativex} \\
+	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y) \label{eqn:weightderivativey}
+\end{align}
+and also that
+\begin{align*}
+	\ip< \Wab \hdopnkab, W^{(\tilde{a},\tilde{b})} \hdopmj^{(\tilde{a},\tilde{b})} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wab.
+\end{align*}
+\end{proof}
+
+Further, we note that there exist "parameter transformation" banded-block-banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
@@
 $$
-where if $f(\cdot, y)$ for fixed $y$ is an at most degree $N$ polynomial and $f(x, \cdot)$ for fixed $x$  is an at most degree $2N-1$ polynomial, the quadrature rule is exact.
+and the quadrature rule is exact if $x \mapsto f(x,y)$ for fixed $y$ is an at most degree $N$ polynomial and $y \mapsto f(x,y)$ for fixed $x$ is an at most degree $2N-1$ polynomial.
 \end{theorem}
@@
 \end{align*}
-Let $f : \Omega \to \R$ be such that $y \mapsto f(x,y)$ for fixed $x$ is an even function. Then,
+Let $f : \Omega \to \R$ be such that $y \mapsto f(x,y)$ for fixed $x$ is an even function.
+Note that if $f$ is a polynomial, then $f(s, \rho(s)t)$ is a polynomial in $s \in [\alpha,\beta]$ for fixed $t$.
+Hence,
 \begin{align*}
@@
 \end{align*}
-where we achieve equality at $(\star)$ if $y \mapsto f(x,y)$ for fixed $x$ is a polynomial of degree at most $2N-1$ and we achieve equality at $(\star \star)$ if $x \mapsto f(x,y)$ for fixed $y$ is a polynomial of degree at most $N$.
-Note also that the inner integral at $(\dagger)$ the inner integral over $t$ would be zero if $y \mapsto f(x,y)$ for fixed $x$ was an odd function, due to the symmetry over the domain (i.e.
+where we achieve equality at $(\star)$ if $y \mapsto f(x,y)$ for fixed $x$ is a polynomial of degree at most $2N-1$ and we achieve equality at $(\star \star)$ if also $x \mapsto f(x,y)$ for fixed $y$ is a polynomial of degree at most $N$.
+
+Finally, note that the inner integral at $(\dagger)$ over $t$ would be zero if $y \mapsto f(x,y)$ for fixed $x$ was an odd function, due to the symmetry over the domain (i.e.
 since $\gamma = -\delta$ in the this case).
-
-\bstodoinline{I cannot remember/work out why it was necessary to split f into odd and even parts?}
 \end{proof}
@@
 
-%\bstodo{This section contains repetition / expansion on the quadrature rule section above - not sure if we should remove it or move it?} We can develop a quadrature rule on the half disk \(\Omega\) in order to evaluate (40) and obtain the coefficient vectors \(\mathbf{f}_n\) as follows.
-%
-%Let \(f(x,y)\) be polynomial that is even in \(y\) of degree $d_1$ in $x$ and degree $d_2$ in $y$ (i.e.
-$f(\cdot, y)$ with $y$ fixed is a degree $d_1$ polynomial, and $f(x, \cdot)$ with $x$ fixed is an even degree $d_2$ polynomial).
-Then there exists a polynomial \(\tilde{f}(x,y)\) in \(x,y\) such that \(g(s,t) := f(x, y) = f(s, (1-s^2)^\half t) = \tilde{f}(s, (1-s^2) t^2)\) (i.e.
-that $d_2$ is an even integer).
-Now, as \(\tilde{f}\) is a polynomial, so must \(g\) be a polynomial of degree $d_1+\frac{d_2}{2}$ in $s$ and degree $d_2$ in $t$.
-Thus a quadrature rule of the following will be exact for the polynomial $f(x,y)$ (even in \(y\) of degree $\le d_1$ in $x$ and degree $\le d_2$ in $y$):
-%\begin{align*}
-%& \quad \int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \\
-%&= \int_0^1 \int_{-1}^1 V^{(a,b)}(s,t) \: g(s,t) \: (1-s^2)^\half \: dt \: ds \\
-%&= \int_0^1 s^a \: (1-s^2)^{b+\half} \: \Big( \sum_k^{\half(d_2 + 1)} w_k^{(t)} \: g(s, t_k) \Big) \: ds \\
-%&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: g(s_j, t_k) \\
-%&= \sum_j^{\half(d_1 + \frac{d_2}{2} + 1)} \sum_k^{\half(d_2 + 1)} w_j^{(s)} w_k^{(t)} \: f(x_j, y_{j,k}),
-%\end{align*}
-%where \(x_j := s_j,\) \(y_{j,k} := \rho(s_j) \: t_k\) are the quadrature points.
-%
-%We further note that for any polynomial \(f(x,y)\) that is odd in \(y\),
-%\begin{align*}
-%\int_0^1 \int_{-\rho(x)}^{\rho(x)} W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx = 0.
-%\end{align*}
-%
-%
-%Then our quadrature rule for any polynomial \(f(x,y)\) on \(\Omega\) is:
-%\begin{align*}
-%\int \int_\Omega W^{(a,b)}(x,y) \: f(x,y) \: dy \: dx \approx \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
-%\end{align*}
-%and is exact for any polynomial $f(x,y)$ where the degree in $x$ is $\le N$ and the degree in $y$ is $\le 2N-1$.
-Note that for a general polynomial where the overall degree is $N$, we still would hence need $N^2$ points and weights to make the rule exact.
-Thus,
-%\begin{align*}
-%\ip< f, \: \hdopnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \big(f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j)\big) +\big(f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j)\big) \big], \\
-%|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \hdopnk^{(a,b)}(x_j, y_j)^2 + \hdopnk^{(a,b)}(x_j, -y_j)^2 \big],
-%\end{align*}
-%for suitably large $N$.
-
 \subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
-
-\sotodoinline{The first part of this section should be moved up to Section 3, and stated as a "Lemma". }
-
-Since $\{P^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a+1, b+1)}_{m,j}$.
-The coefficients of the expansion are then the entries of the relevant operator matrix.
-We can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
-First, note that
-\begin{align*}
-c_{m,j} = \frac{\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
-\end{align*}
-Second, note that
-\begin{align}
-	\rho'(x) = -x \: \rho(x)^{-2}. \label{eqn:rhoderivative}
-\end{align}
-Now, using a change of variables and (\ref{eqn:rhoderivative}), we have that
-\begin{align*}
-	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
-\end{align*}
-We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
-Now, using integration-by-parts and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
-\begin{align*}
-	&\ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
-	&= \ip<\genjacnmk^{(a, b+k+\half) \: \prime}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \ddx \Big[\genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} \Big] >_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}}
-\end{align*}
-which is zero for $m < n - 2$, and further that
-\begin{align*}
-	&\ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
-	&= \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
-\end{align*}
-which is also zero for $m < n - 2$.
-%, and since the weight $\genjacw^{(a+1, b+1)}$ will vanish at $x=0,1$ for non-negative parameters $a,b$, we have that for a polynomial $g(x,y)$ of degree $m$ on $\Omega$: 
-%\begin{align}
-%	&\ip<\pddx \hdopnkab, g >_{W^{(a+1, b+1)}} \\
-%	&= \int_0^1 \int_{-\rho(x)}^{\rho(x)} \pddx f \: g \: W^{(a+1,b+1)} \: \D y \: \D x \\
-%	&= \int_{-1}^1 \Big( \int_0^1 \pddx \hdopnkab \: g \: \genjacw^{(a+1,b+1)} \: \rho(x) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
-%	&= - \int_{-1}^1 \Big( \int_0^1 \hdopnkab \: \pddx \big(g \: \genjacw^{(a+1,b+1)}(x) \: \rho(x)\big) \: \D x  \Big)  \: w_P^{(b+1)}(y) \: \D y \\
-%	&= - \int_{-1}^1 \Big( \int_0^1 \hdopnkab \: \pddx \big(g \: \rho(x) \: \genjacw^{(a+1,b+1)}(x) \: w_P^{(b+1)}(y)\big) \: \D x  \Big) \: \D y \\
-%	&= - \int_0^1 \int_{-\rho(x)}^{\rho(x)} \hdopnkab \: \pddx \big(g \: \rho(x) \: W^{(a+1,b+1)} \big) \: \rho(x)^{-1} \: \D y \: \D x \\
-%	&= - \iint_\Omega \hdopnkab \: \big[\pddx g \: W^{(a+1,b+1)} + g \: \pddx W^{(a+1,b+1)} + \: g \rho'(x) \rho(x)^{-1} \: W^{(a+1, b+1)} \big] \: \D y \: \D x \\
-%	&= - \ip<\hdopnkab,\: \rho'(x) \rho(x)^{-1} \: W^{(1, 1)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:weightderivativex}) \\
-%	&= - \ip<\hdopnkab,\: W^{(2, 0)} \: g + [(a+1)W^{(0, 1)} - 2(b+1)W^{(2, 0)}] \: g + W^{(1, 1)} \: \pddx g>_{W^{(a, b)}} \quad \text{by } (\ref{eqn:rhoderivative})
-%\end{align}
-%Now, as the polynomial on the RHS of the inner product is of degree $m+2$ and $\hdopnkab$ are orthogonal, we have that the above is zero for $m+2 < n$, i.e $c_{m,j}$ is non-zero for $m = n-1, n-2$ only.
 
@@
 This is much cheaper than using the 2D quadrature rule.
-
-For partial differentiation by $y$, we can also expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
-The coefficients of the expansion are then the entries of the relevant operator matrix.
-By a similar argument to before, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
-This is then calculated by taking inner products, i.e.
-\begin{align*}
-c_{n-1,k-1} = \frac{\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
-\end{align*}
-where
-\begin{align*}
-	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \normgenjac^{(a, b+k+\half)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
-\end{align*}
-
-We can gain the non-zero entries of the weighted differential operators similarly, by noting that
-\begin{align}
-	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \label{eqn:weightderivativex} \\
-	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y) \label{eqn:weightderivativey}
-\end{align}
-and also that
-\begin{align*}
-	\ip< \Wab \hdopnkab, W^{(\tilde{a},\tilde{b})} \hdopmj^{(\tilde{a},\tilde{b})} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wab.
-\end{align*}
 
@@
 
-\subsection{Heat equation}
-
-\sotodoinline{Skip example?}
-Consider
-\begin{align*}
-	\begin{cases}
-		\frac{\partial u}{\partial t}(t,x,y) = \Delta u(t,x,y) \quad \text{in } \Omega \\
-		u(t, x, y) = 0 \quad \text{on } \partial \Omega \\
-		u(0, x, y) = u_0(x,y) := \Wii(x,y) \: f(x,y)
-	\end{cases}
-\end{align*}
-noting the imposition of zero Dirichlet boundary conditions.
-
-For simplicity of exposition, we use a backward Euler time-stepping method to discretise in time, which can be generalised to more sophisticated time-stepping schemes.
-That is, formulate this for \(t = 0,1,2,\dots\) and step length $h$:
-\begin{align*}
-u_{t+1}(x,y) - u_t (x,y) = h \: \Delta u_{t+1}(x,y) \quad \text{for } (x,y) \in \Omega.
-\end{align*}
-
-We can tackle the problem as follows.
-Fix \(a, b = 1,1\).
-Denote the coefficient vector for expansion of $u_0$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ we denote $\mathbf{u}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bigWii$ OP basis up to degree $N$.
-Further, denote the coefficient vector for expansion of $u_0$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{\tilde{u}}_0$, and similarly for each $t = 1,2,3,\dots,t_{max}$ denote $\mathbf{\tilde{u}}_t$ as the coefficient vector of the expansion of $u_t$ in the $\bighdopii$ OP basis up to degree $N$.
-
-We can solve this system for the given initial condition $\mathbf{u_0}$ by iterating for $t = 0,1,2,\dots,t_{max}$:
-\begin{align*}
-	&T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (\mathbf{u}_{t+1} - \mathbf{u}_t) = h \: \laplacewii \mathbf{u}_{t+1} \nonumber \\
-	\iff& (T^{(0,0)\to(1,1)} \:T_W^{(1,1)\to(0,0)} - h \: \laplacewii) \: \mathbf{u}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
-	\iff& \mathbf{u}_{t+1} = (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \laplacewii)^{-1} T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: \mathbf{u}_t \nonumber \\
-	\iff& \mathbf{\tilde{u}}_{t+1} = T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} \: (T^{(0,0)\to(1,1)} \: T_W^{(1,1)\to(0,0)} - h \: \laplacewii)^{-1} \: \mathbf{\tilde{u}}_t.
-\end{align*}
-
-%\begin{figure}
-%	\includegraphics[scale=0.4]{example2timesteppingcovnvergence}
-%\centering
-%\caption{Plot of the current iteration's coefficients for the heat equation example, using degree of approximation $N = 5$ with initial condition of $u_0(x,y) = \Wii(x,y)$ and step size $h=0.01$, plotted for iterations $t=100, 120, \dots, 200$.
-The solution tends to zero, as seen by the decreasing (absolute value) coefficient values.}
-%\centering
-%\label{fig:heateqn}
-%\end{figure}
-
 %
@@
 
-\sotodoinline{Comment on how the sparse operators generalise to such domains}
+The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such disk-slice domains.
+For instance, if we inspect the proof of Lemma \ref{lemma:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for a disk-slice.
+\bstodoinline{Does this require more?}
 

### 2019-05-01 papers/halfdisk/halfdisk.tex
Further edits

@@
 \newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
+\newcommand{\ddy}{\frac{\mathrm{d}}{\mathrm{d}y}}
 \newcommand{\pddx}{\frac{\partial}{\partial x}}
@@
 \begin{proof}
-We proceed with the case for the operator $D_x^{(a,b)}$ for partial differentiation by $x$.
+We proceed with the case for the operator $D_y^{(a,b)}$ for partial differentiation by $y$, we can also expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+The coefficients of the expansion are then the entries of the relevant operator matrix.
+We can use an integration-by-parts argument to show that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
+First, note that
+\begin{align*}
+	c_{m,j} = \frac{\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{m,j}}^2_{W^{(a, b+1)}}}.
+\end{align*}
+Then, using the change of variable $\tilde{y} = \frac{y}{\rho(x)}$, we have that
+\begin{align*}
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} &= \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} \\
+	&= \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+\end{align*}
+Now, integrating by parts, noting that the weight $w_P^{(b)}$ is a polynomial degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
+\begin{align*}
+	\ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} &= \ip< P_k^{(b,b) \: \prime}, \: w_P^{(b+1)} \: P_{j}^{(b+1,b+1)} >_1 \\
+	&= \ip< P_k^{(b,b)}, \: \ddy [w_P^{(b+1)} \: P_{j}^{(b+1,b+1)}] >_1 \\
+	&= \ip< P_k^{(b,b)}, \: P_{j}^{(b+1,b+1) \: \prime} >_{w_P^{(b+1)}} - 2 b \ip< P_k^{(b,b)}, \: y \: P_{j}^{(b+1,b+1)} >_{w_P^{(b)}} \\
+	&= \ip< P_k^{(b,b)}, \: w_P^{(1)} \: P_{j}^{(b+1,b+1) \: \prime} -2b \:y \: P_{j}^{(b+1,b+1)} >_{w_P^{(b)}} 
+\end{align*}
+which is zero for $j < k-1$. Further,
+\begin{align*}
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,k-1} >_{W^{(a, b+1)}} &= \ip< \genjacnmk^{(a,b+k+\half)}, \genjac_{m+1-k}^{(a,b+k+\half)} >_{\genjacw^{(a,b+k+\half)}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}},
+\end{align*}
+which is zero for $m < n-1$. Finally, we simply have that
+\begin{align*}
+	c_{n-1,k-1} &= \frac{1}{\normjac^{(b)}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+\end{align*}
+
+We next consider the case for the operator $D_x^{(a,b)}$ for partial differentiation by $x$.
 Since $\{\hdop^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+1, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
-We can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-2$ and $0 \le j \le m$.
+As before, we can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-1, k-2$ and $0 \le j \le m$.
 First, note that
@@
 \end{align}
-Now, using a change of variables and (\ref{eqn:rhoderivative}), we have that
+Now, using the change of variable $\tilde{y} = \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that
 \begin{align*}
 	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
@@
 \begin{align*}
-	&\ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
-	&= \ip<\genjacnmk^{(a, b+k+\half) \: \prime}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \ddx \Big[\genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} \Big] >_1 \\
+	&\ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
+	&= \ip<\genjacnmk^{(a, b+k+\half) \: \prime}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half)} >_1 \\
+	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \ddx \Big[\genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half)} \Big] >_1 \\
 	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_1 \\
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime}>_1 \\
 	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}}
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}}
 \end{align*}
@@
 \begin{align*}
-	&\ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
-	&= \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
+	&\ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
+	&= \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
 \end{align*}
 which is also zero for $m < n - 2$.
-
-\bstodoinline{Finish this proof!!!}
-
-For the operator for partial differentiation by $y$, we can also expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
-The coefficients of the expansion are then the entries of the relevant operator matrix.
-By a similar argument to before, we have that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
-This is then calculated by taking inner products, i.e.
-\begin{align*}
-c_{n-1,k-1} = \frac{\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{n-1,k-1}}^2_{W^{(a, b+1)}}}
-\end{align*}
-where
-\begin{align*}
-	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{n-1,k-1} >_{W^{(a, b+1)}} = \normgenjac^{(a, b+k+\half)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
-\end{align*}
 
@@
 \end{align*}
+
 \end{proof}

### 2019-05-01 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
updates for ApproxFun v0.11.1

@@ -1,11 +1,13 @@
 module OrthogonalPolynomialFamilies
 
 using ApproxFun
-    import ApproxFun: evaluate, PolynomialSpace, recα, recβ, recγ, recA, recB, recC, domain,
-                        domainspace, rangespace, bandwidths, prectype, canonicaldomain, tocanonical,
-                        spacescompatible, points, transform, itransform, AbstractProductSpace, tensorizer,
-                        columnspace, checkpoints, plan_transform, clenshaw
-    import Base: in, *
+import ApproxFun: evaluate, domain,
+                    domainspace, rangespace, bandwidths, prectype, canonicaldomain, tocanonical,
+                    spacescompatible, points, transform, itransform, AbstractProductSpace, 
+                    checkpoints, plan_transform, clenshaw
+import ApproxFunOrthogonalPolynomials: PolynomialSpace, recα, recβ, recγ, recA, recB, recC
+import ApproxFunBase: tensorizer, columnspace            
+import Base: in, *
 using StaticArrays
 using FastGaussQuadrature
 using LinearAlgebra
@@ -56,7 +58,7 @@ function lanczos!(w, P, β, γ; N₀=0)
     end
 
     for k = N₀+1:N
-        @show "lanczos", N, k
+        # @show "lanczos", N, k
         v = x*P[2] - γ[k-1]*P[1]
         β[k] = sum(P[2]*w*v)
         v = v - β[k]*P[2]

### 2019-05-01 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -3,7 +3,7 @@ using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature,
 import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule,
         gethalfdiskOP, jacobix, jacobiy, differentiatex, differentiatey,
         resizecoeffs!, laplaceoperator, transformparamsoperator, operatorclenshaw, getnk,
-        getopindex, weight
+        getopindex, weight, resizedata!
 
 
 
@@ -266,3 +266,52 @@ end
 end
 
 #=====#
+
+# using Profile
+# D = HalfDiskFamily()
+# S = H = D.H(BigFloat(1),BigFloat(2.5));
+# @time resizedata!(S,300);
+# @time resizedata!(S,1000);
+# Profile.print()
+
+
+# w = S.weight
+# x = Fun(identity, domain(w))
+# g = cos.(1000x);
+
+# g = Fun(Chebyshev(domain(x)), BigFloat.(randn(1000)))
+# # @time sum(w*g^2)
+# @time sum(g^2)
+
+# g = Fun(Legendre(domain(x)), BigFloat.(randn(1000)))
+
+
+# g = Fun(NormalizedLegendre(domain(x)), BigFloat.(randn(1000)))
+# @time norm(g.coefficients)
+
+# x = Fun(BigFloat(-1)..1)
+# g = cos.(x)
+# @time norm(g)
+# Fun(g, NormalizedLegendre(domain(x)))(0.1)
+# g(0.1)
+# Fun(x -> cos(x), Legendre(domain(x)))(0.1)
+# @time norm(Fun(g, NormalizedLegendre(domain(x))).coefficients)
+
+# x = Fun()
+# g = cos.(x)
+# sqrt(sum(g^2))
+# norm(Fun(g, NormalizedLegendre(domain(x))).coefficients)
+# g^2
+
+
+# norm(Fun(g, Legendre(domain(x))).coefficients)
+
+
+# @time norm(g)
+
+
+
+# , S.ops, S.a, S.b, N₀=N₀)
+
+
+# Profile.print()    

### 2019-05-02 OrthogonalPolynomialFamilies/test/runtests.jl
Slight correction to jacobi matrices test

@@ -131,8 +131,8 @@ end
     N = Int(ceil(0.5 * (-1 + sqrt(1 + 8m)))) - 1
     Jx = jacobix(S, N+1)
     Jy = jacobiy(S, N+1)
-    @test evaluate(Jx'[1:m, 1:m] * cfs, S, z) ≈ z[1] * f(z...)
-    @test evaluate(Jy'[1:m, 1:m] * cfs, S, z) ≈ z[2] * f(z...)
+    @test evaluate(Jx[1:m, 1:m] * cfs, S, z) ≈ z[1] * f(z...)
+    @test evaluate(Jy[1:m, 1:m] * cfs, S, z) ≈ z[2] * f(z...)
 end
 
 @testset "Evaluate partial derivative of (random) function" begin
@@ -314,4 +314,4 @@ end
 # , S.ops, S.a, S.b, N₀=N₀)
 
 
-# Profile.print()    
+# Profile.print()

### 2019-05-02 papers/halfdisk/halfdisk.tex
Edits to paper

@@
 								% TeX will automatically convert eps --> pdf in pdflatex		
-\usepackage{caption}
+\usepackage[margin=10pt,font=footnotesize,labelfont=bf]{caption}
 \usepackage{subcaption}
@@
 
-The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\).
-The weight $w_1(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation \(\{\genjac_n^{(a, b)}\}\) for the orthonormal polynomials.
+The weight $w_P(x)$ is an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials  $\{P_n^{(b, b)}\}$.
+The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials.
 Thus we arrive at a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
@@
 \begin{proof}
-First, note that, using a change of variable,
-\begin{align*}
-	\norm{\hdopnkab}_\Wab^2 &= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
-	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(y)^2 \: w_P^{(b)}(y) \: \D y \Big) \nonumber \\
+The 3-term recurrence for multiplication by $x$ follows from equation (\ref{eqn:Hrecurrence}).
+For the recurrence for multiplication by $y$, first note that, using a change of variable $t = \frac{y}{\rho(x)}$,
+\begin{align*}
+	\norm{\hdopnkab}_\Wab^2 &=  \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \Wab(x,y) \: \D y \: \D x \nonumber \\
+	&= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b+k)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
+	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(t)^2 \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
 	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}.
@@
 
-Finally, note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
+Finally, note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\) using the same change of variable:
 \begin{align*}
 	\ip<y \hdopnkab, \hdopmjab>_\Wab &=  \iint_\Omega \hdopnkab(x,y) \: \hdopmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
-&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(s) \: \genjacmmj^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
-& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
-&= \begin{cases}
+	&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(x) \: \genjacmmj^{(a, b+j+\half)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b)}(x) \: \D x \Big) \nonumber \\
+	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: w_P^{(b)}(t) \: \D t \Big) \\
+	&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(x) \: \genjacmmj^{(a, b+j+\half)}(x) \: \genjacw^{(a, b+1+\half(k+j))}(x) \: \D x \Big) \nonumber \\
+	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: w_P^{(b)}(t) \: \D t \Big) \\
+	&= 
+	\begin{cases}
     		\delta_k^{(b)} \: \normjac^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
@@
 \end{align}
-The matrices $J_x^{(a,b)}, J_y^{(a,b)}$ act on the coefficients vector of a functions expansion in the $\{\hdopnkab\}$ basis.
+The matrices $J_x^{(a,b)}, J_y^{(a,b)}$ act on the coefficients vector of a function's expansion in the $\{\hdopnkab\}$ basis.
 For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
 Then $x \: f(x,y)$ is approximated by $\bighdopab(x,y)^\top {J_x^{(a,b)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\hdopnkab\}$ basis.
+In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkab\}$ basis.
 
@@
 C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
-\nonumber \\
+\nonumber
+\end{align*}
+\begin{align*}
 A^y_n &:= \begin{pmatrix}
@@
 \renewcommand\arraystretch{1.3}
-\mleft[
-\begin{array}{cccc}
+\begin{pmatrix}
 		1 & & & \\
@@
 		& & \ddots &
-\end{array}
-\mright]
+\end{pmatrix}
 \bighdopab(x,y)
@@
 \begin{pmatrix}
-	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
+	\hdopab_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
 \end{pmatrix}.
 \end{align*}
-where we note \(P^{(a,b)}_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}\), and for each $n = 0,1,2\dots$:
+where we note $\hdopab_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}$, and for each $n = 0,1,2\dots$:
 \begin{align*}
@@
 \bighdopab_{0}(x,y) := P^{(a,b)}_{0,0} \\
-\bighdopab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopab_n(x,y) - \Dnt C_n  \,\bighdopab_{n-1}(x,y), \quad n = 1,2,3,\dots.
+\bighdopab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopab_n(x,y) - \Dnt C_n  \,\bighdopab_{n-1}(x,y), \quad n = 0,1,2,\dots.
 \end{cases}
@@
 
-\begin{definition}\label{def:sparsityofdifferentialoperators}
+\begin{definition}\label{def:differentialoperators}
 Define the operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ according to:
@@
 \begin{lemma}\label{lemma:sparsityofdifferentialoperators}
-The operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ from Definition \ref{def:sparsityofdifferentialoperators} are sparse, with banded-block-banded structure.
+The operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
 \end{lemma}
@@
 \begin{proof}
-We proceed with the case for the operator $D_y^{(a,b)}$ for partial differentiation by $y$, we can also expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: P^{(a, b+1)}_{m,j}$.
+We proceed with the case for the operator $D_y^{(a,b)}$ for partial differentiation by $y$.
+Since $\{\hdop^{(a, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: \hdop^{(a, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 \end{align*}
-Then, using the change of variable $\tilde{y} = \frac{y}{\rho(x)}$, we have that
+Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
 \begin{align*}
@@
 \end{align}
-Now, using the change of variable $\tilde{y} = \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that
+Now, using the change of variable $t= \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that
 \begin{align*}
@@
 
-Further, we note that there exist "parameter transformation" banded-block-banded matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
-We denote these operators as $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and similarly $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, where:
+Further, we note that there exist "parameter transformation" matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+\begin{definition}\label{def:parametertransformationoperators}
+Define the operator matrices $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$ and $T^{(a,b)\to(a+1,b+1)}$ for conversion between non-weighted spaces, and $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ for conversion between weighted spaces, according to:
 \begin{align*}
@@
 	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^\top \: \bigW^{(a,b-1)}(x,y) \\
-\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigW^{(a-1,b-1)}(x,y) \\
-\end{align*}
+	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigW^{(a-1,b-1)}(x,y).
+\end{align*}
+\end{definition}
+
+\begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
+The operator matrices $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$, $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ from Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure.
+\end{lemma}
+
+\begin{proof}
+We proceed with the case for the non-weighted operators $T^{(a,b)\to(a+\lambda,b+\mu)}$, where $\lambda, \mu \in \{0,1\}$.
+Since $\{\hdop^{(a+\lambda, b+\mu)}_{m,j}\}$ for $m = 0,\dots,n$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n$ polynomial, we can expand $\hdopnkab = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+\lambda, b+\mu)}_{m,j}$.
+The coefficients of the expansion are then the entries of the relevant operator matrix.
+We will see that the only non-zero coefficients are for $m \ge n - \lambda - 2\mu $, $j \ge k-2\mu$ and $0 \le j \le m$.
+First, note that
+\begin{align*}
+	c_{m,j} = \frac{\ip< \hdopnkab, \hdop^{(a+\lambda, b+\mu)}_{m,j} >_{W^{(a+\lambda, b+\mu)}}}{\norm{\hdop^{(a+\lambda, b+\mu)}_{m,j}}^2_{W^{(a+\lambda, b+\mu)}}}.
+\end{align*}
+Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
+\begin{align*}
+	\ip< \hdopnkab, \hdop^{(a+\lambda, b+\mu)}_{m,j} >_{W^{(a+\lambda, b+\mu)}} &= \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a+\lambda,b+\mu)}} \\
+	&\quad \quad \quad \cdot \ip< P_k^{(b,b)}, \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b+\mu)}} \\
+	&= \ip< \genjacnmk^{(a,b+k+\half)}, \genjacw^{(\lambda, \: \mu+\half(j-k))} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a,b+k+\half)}} \\
+	&\quad \quad \quad \cdot \ip< P_k^{(b,b)}, \: w_P^{(\mu)} \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b)}}
+\end{align*}
+Since $w_P^{(\mu)}$ is a polynomial degree $2\mu$, we have that the above is then zero for $j < k-2\mu$.
+Further, since $\genjacw^{(\lambda, \: \mu+\half(j-k))}$ is a polynomial of degree $\lambda + 2\mu + j - k$, we have that the above is zero for $m-j < n-k - (\lambda + 2\mu + j - k) \iff m < n - \lambda - 2\mu$.
+
+The sparsity argument for the weighted parameter transformation operators follows similarly.
+\end{proof}
 
@@
 	\centering
-	\begin{subfigure}[t]{0.3\textwidth}
+	\begin{subfigure}[t]{0.32\textwidth}
 	\centering
@@
 	\end{subfigure}
-	\begin{subfigure}[t]{0.3\textwidth}
+	\begin{subfigure}[t]{0.32\textwidth}
 	\centering
@@
 	\end{subfigure}
+	\centering
     	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
@@
 Centre: the weighted variable coefficient Helmholtz operator $\laplacewii + k^2 \: T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
-Right: the biharmonic operator $\biharmonic$.
-	\sotodoinline{Use PDFs for pictures to avoid resolution problems}
-	}
+Right: the biharmonic operator $\biharmonic$.}
+	\centering
         \label{fig:sparsity}
@@
 
-The sparsity of the these operators can be shown using a simple integration-by-parts argument, see Section \ref{subsection:Computation-operatormatrices}.
-
 \section{Computational aspects}\label{Section:Computation}
@@
 
-We can then use quadrature to calculate the 1D inner products and find the non-zero $c_{m,j}$ entries to the partial differential operator matrix.
-This is much cheaper than using the 2D quadrature rule.
-
-Calculating the non-zero entries of the parameter transformation operators is similarly done, i.e. use a substitution in the 2D inner product integral to yield a product of 1D inner product integrals.
+The proofs of Lemma \ref{lemma:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
+We can simply use quadrature to calculate the 1D inner products.
+This proves much cheaper computationally than using the 2D quadrature rule to calculate the 2D inner products.
 
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that we achieve algebraic convergence for the first three examples, with the final Gaussian bump example achieving spectral convergence.
-We note that for right hand sides that vanish at the corners of our half disk ($y = \pm 1$) we see faster convergence.
-\bstodoinline{Improve above}
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+For the final Gaussian bump example, we see that we achieve spectral convergence.
 
@@
 	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
-This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.
-\bstodoinline{N=200 blocks here}}
+This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
         	\label{fig:solutionblocknorms}
@@
 For instance, if we inspect the proof of Lemma \ref{lemma:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for a disk-slice.
-\bstodoinline{Does this require more?}
 

### 2019-05-03 papers/halfdisk/halfdisk.tex
Further edits

@@
 \end{align*}
-i.e. $\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}}^2 = \normgenjac^{(a,b)}$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}}^2 = \normjac^{(b)}$ $\forall n = 0,1,2,\dots$.
-
+i.e.
+$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}}^2 = \normgenjac^{(a,b)}$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}}^2 = \normjac^{(b)}$ $\forall n = 0,1,2,\dots$.
 For clarity, we use the notation that
@@
 \end{align*}
+Further, we have that the norms of the 2D half-disk OPs for a particular parameter family $(a,b)$ are then given by
+\begin{align}
+	\norm{\hdopnkab}_\Wab^2 &=  \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \Wab(x,y) \: \D y \: \D x \nonumber \\
+	&= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b+k)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
+	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(t)^2 \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
+	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)} \label{eqn:normhdop}
+\end{align}
+where we have used the change of variable $t = \frac{y}{\rho(x)}$.
+Again, for clairty we use the notation that $\ip<p, q>_{\Wab} := \iint_\Omega p(x,y) \: q(x,y) \: \Wab(x,y) \: \D y \: \D x$ and $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$ .
 
@@
 \begin{align*}
-\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)}, \qquad \alphaab_{n,k,2} := \alpha_{n-k}^{(a, b+k+\half)} \\
+	\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
+	\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
 	\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
@@
 The 3-term recurrence for multiplication by $x$ follows from equation (\ref{eqn:Hrecurrence}).
-For the recurrence for multiplication by $y$, first note that, using a change of variable $t = \frac{y}{\rho(x)}$,
-\begin{align*}
-	\norm{\hdopnkab}_\Wab^2 &=  \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \Wab(x,y) \: \D y \: \D x \nonumber \\
-	&= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b+k)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
-	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(t)^2 \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
-	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}.
-\end{align*}
-where we use the notation that $\ip<p, q>_{\Wab} := \iint_\Omega p(x,y) \: q(x,y) \: \Wab(x,y) \: \D y \: \D x$ and $\norm{p}_{W^{(a,b)}}^2 := \ip<p, p>_{\Wab}$ .
-
-Finally, note that \(\ip<y \hdopnkab, \hdopmjab>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\) using the same change of variable:
+For the recurrence for multiplication by $y$, since $\{\hdopmjab\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkab(x,) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjab(x,y)$.
+These coefficients are given by
+\begin{align*}
+	c_{m,j} = \frac{\ip< y \: \hdopnkab, \hdopmjab >_{\Wab}}{\norm{\hdopmjab}^2_{\Wab}}.
+\end{align*}
+Recall from equation (\ref{eqn:normhdop}) that $\norm{\hdopnkab}_{\Wab}^2 = \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}$.
+Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
 \begin{align*}
@@
 \end{align*}
+where $\ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}}$ and $\ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k+\half)}}$ are zero for $m < n-1$.
 
@@
 Finally, note that the inner integral at $(\dagger)$ over $t$ would be zero if $y \mapsto f(x,y)$ for fixed $x$ was an odd function, due to the symmetry over the domain (i.e.
-since $\gamma = -\delta$ in the this case).
+since $\gamma = -\delta$).
 \end{proof}
@@
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
-f_{n,k} := \frac{\langle f, \: \hdopnk^{(a,b)} \rangle_{W^{(a,b)}}}{|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}}
-\end{align*}
-
-Using the quadrature rule detailed in Section 4.2, we can calculate exactly the coefficients $f_{n,k}$: 
-\begin{align*}
-	\ip< f, \: \hdopnk^{(a,b)} >_{W^{(a,b)}} &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j) \big], \\
-	|| \hdopnk^{(a,b)} ||_{W^{(a,b)}}^2 &= \half \: \sum_{j=1}^{N^2} w_j \: \big[ \hdopnk^{(a,b)}(x_j, y_j)^2 + \hdopnk^{(a,b)}(x_j, -y_j)^2 \big].
+f_{n,k} := \frac{\langle f, \: \hdopnk^{(a,b)} \rangle_{W^{(a,b)}}}{\norm{\hdopnk^{(a,b)}}_{W^{(a,b)}}}
+\end{align*}
+
+Recall from (\ref{eqn:normhdop}) that $\norm{\hdopnkab}^2_\Wab = \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}$.
+Using the quadrature rule detailed in Section 4.2 for the inner product, we can calculate exactly the coefficients $f_{n,k}$:
+\begin{align*}
+	f_{n,k} &= \frac{1}{2 \: \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}} \: \sum_{j=1}^{N^2} w_j \: \big[ f(x_j, y_j) \: \hdopnk^{(a,b)}(x_j, y_j) +f(x_j, -y_j) \: \hdopnk^{(a,b)}(x_j, -y_j) \big].
 \end{align*}

### 2019-05-09 papers/halfdisk/halfdisk.tex
Removed SO notes and commented out stuff
comments from Xu

@@
 $$
-W^{(a,b)}(x,y) = x^a (1-x-y)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
+W^{(a,b)}(x,y) = x^a (1-x^2-y^2)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
 $$
-with corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
+with corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$.
+We define these to be orthogonalised lexicographically, that is, \sotodo{Are these meant to be orthonormal?}
 $$
@@
 Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
+
+\sotodoinline{Mention that this is a computational challenge}
+
+\sotodoinline{Make clear this is more than plugging in the basis into a Petrov--Galerkin method.}
 
@@
 \end{align*}
-where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{\genjac_n^{(a,b,c)}\}$ are orthogonal with respect to $\genjacw^{(a,b,c)}(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
+where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{\genjac_n^{(a,b,c)}\}$ are \sotodo{Make clear whether they are orthonormal or not} orthogonal with respect to $\genjacw^{(a,b,c)}(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
 
@@
 \begin{cases}
-(\alpha,\beta) &:= (0.1) \\
+(\alpha,\beta) &:= (0,1) \\
 (\gamma,\delta) &:= (-1,1) \\
@@
 
-We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as
+We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as \sotodo{These hold only for orthonormal, which isn't the case for $P_n$}
 \begin{align}
@@
 \end{align*}
-where
+where \sotodo{$C^x_n := \big( A^x_n \big)^\top$ is only true if orthonormal}
 \begin{align*}
@@
 
-To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
+To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that\sotodo{check this}
 \begin{align*}
@@
 
-%
-%\section{Hemisphere}
-%
-%We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the hemisphere.
-Define $\rho(x) := (1-x^2)^\half$ for $x \in [0,1]$ and let the half disk be denoted by $\omega$, that is let
-%\begin{align}
-%	\omega := \{(x,y) \in \R^2 \quad | \quad 1 < x < 0, \: \rho(x) < y < \rho(x)\}.
-%\end{align}
-%Further, define the hemisphere by
-%\begin{align}
-%	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
-%\end{align}
-%We can define the 3D OP family on the hemisphere as the set of polynomials $\bigQab(x,y,z)$ where
-%\begin{align}
-%	\bigQab_n(x,y,z) := \begin{pmatrix}
-%		\bighdopab_n(x,y) \\
-%		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
-%	\end{pmatrix} \in \R^{2n+1}, 
-%	\quad \quad 
-%	\bigQab := \begin{pmatrix}
-%		\bigQab_0 \\
-%		\bigQab_1 \\
-%		\vdots \\
-%		\bigQab_N
-%	\end{pmatrix}.
-%\end{align}
-%The relations for multiplication by $x, y$ are then trivial.
-For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
-%\begin{align}
-%	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: P^{(a,b+1)}_{m,j}(x,y), \\
-%	z^2 \: \hdopnkab(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b)\to(a,b-1)} \: z \: P^{(a,b-1)}_{m,j}(x,y).
-%\end{align}
-%A simple calculation shows that the expansions are sparse, that is that most coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are $0$, and are also given by
-%\begin{align}
-%	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: P^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{P^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
-%	d_{n,k,m,j}^{(a,b)\to(a,b-1)} &= \frac{\ip< W^{(0,1)} \hdopnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b-1)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
-%	&= \frac{\ip< \hdopnkab, \: P^{(a,b-1)}_{m,j} >_{W^{(a,b)}}}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \nonumber \\
-%	&= \frac{\norm{\hdopnkab}_\Wab^2}{\norm{P^{(a,b-1)}_{m,j}}^2_{W^{(a,b-1)}}} \: c_{m,j,n,k}^{(a,b-1)\to(a,b)}.
-%\end{align}
-%The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b)\to(a,b-1)}$ are just the entries to the parameter tranformation operator matrices
-%\begin{align}
-%	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y) \\
-%	W^{(0,1)}(x,y) \bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b-1)} \Big)^T \: \bighdop^{(a,b-1)}(x,y)
-%\end{align}
-%respectively.
-
-\sotodoinline{More Xu references.}
-\sotodoinline{Add missing references to Bib file}
 \bibliography{halfdisk-bib}

### 2019-05-10 papers/halfdisk/halfdisk.tex
Fixed some typos. Added reference.

@@
 %, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
-Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction.
+Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
 Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids.
@@
 		A^y_n
-	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)} \quad (n \ne N), \quad
+	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)}, \quad
 C_n := \begin{pmatrix}
@@
 \end{align*}
-Further represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopoo$.
+Further, we can represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopoo$.
 Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
@@
 \end{align*}
-where we define $\beta_{-1}^{(a,b)} = 0$ and $H_{-1}^{(a,b)}(x) \equiv 0$.
+where we define $\beta_{-1}^{(a,b)} = 0$ and $\genjac_{-1}^{(a,b)}(x) \equiv 0$.
 We use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
@@
 The weight $w^{(a,b)}_P(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
-We note that the shifted Jacobi polynomials relate the the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = P_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
+We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = P_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
 $\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\).

### 2019-05-13 papers/halfdisk/halfdisk.tex
Further edits

@@
 \end{align*}
-where the three-parameter family of orthogonal polynomials on $[\alpha,\beta]$ given by $\{\genjac_n^{(a,b,c)}\}$ are \sotodo{Make clear whether they are orthonormal or not} orthogonal with respect to $\genjacw^{(a,b,c)}(x)$, and the two-parameter family of orthogonal polynomials on $[\gamma,\delta]$ given by $\{P_n^{(a,b)}\}$ are orthogonal with respect to $w^{(a,b)}_P(x)$.
+and define the associated inner products by: \bstodo{Is it best to define these ip s with the normalisation?}
+\begin{align}
+	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x, \label{eqn:ipgenjac} \\
+	\ip<p, q>_{w_P^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \:\int_\gamma^\delta p(y) \: q(y) \: w_P^{(a,b)}(y)\: \D y. \label{eqn:ipjac}
+\end{align}
+where
+\begin{align*}
+	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: w_P^{(a,b)}(y) \: \D y.
+\end{align*}
+Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{P_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
 
@@
 \begin{align*}
-	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega. 
-\end{align*}
+	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
+\end{align*}
+that is,
+\begin{align*}
+	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b+c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
+\end{align*}
+where
+$\ip< f, \: g >_W^{(a,b,c,d)} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x$.
 \end{definition}
@@
 
-The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are not orthonormal, but have (squared) norms given by the integral of their respective weight functions we denote as
+The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
+\begin{align*}
+	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
+	\ip<p, q>_{w_P^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y,
+\end{align*}
+where 
 \begin{align*}
@@
 i.e.
-$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}}^2 = \normgenjac^{(a,b)}$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}}^2 = \normjac^{(b)}$ $\forall n = 0,1,2,\dots$.
-For clarity, we use the notation that
-\begin{align*}
-	&\ip<p, q>_{\genjacw^{(a,b)}} := \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \quad
-		\ip<p, q>_{w_P^{(b)}} := \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y, \\
-	&\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}, \quad
-		\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}.
-\end{align*}
+$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}} = 1$ $\forall n = 0,1,2,\dots$.
+For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}$.
+
 Further, we have that the norms of the 2D half-disk OPs for a particular parameter family $(a,b)$ are then given by
@@
 
-The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomials families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
+The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
 
@@
 \end{align*}
-which is zero for $m < n-1$. Finally, we simply have that
+which is zero for $m < n-1$. Finally, we simply have that \bstodo{Make sure correct ip definition is being used here}
 \begin{align*}

### 2019-05-13 experiments/halfdisk.jl
Experiments and examples script

@@ -12,35 +12,35 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
                     laplaceoperator, derivopevalatpts, getderivopptseval, getopnorms,
                     getopnorm, operatorclenshaw, weight, biharmonicoperator
 using SpecialFunctions
+using JLD
+
+#===#
+
+# Investigations on computation time/lanczos
+setprecision(800)
+B = BigFloat; a = 1.0; b = 1.0
+D = HalfDiskFamily(); S = D(a, b)
+N = 100
+@time laplaceoperator(S, S, N, weighted=true) # Compile
+A = laplaceoperator(S, S, N, weighted=true)
+k = 400; H = D.H(S.a, S.b+k+0.5)
+OrthogonalPolynomialFamilies.resizedata!(H, 200)
+
+
+
 
 #===#
 
 # Set up for experiments
-precision(BigFloat)
-setprecision(800)
-D = HalfDiskFamily()
-a, b = 1.0, 1.0; S = D(a, b)
-x, y = 0.42, -0.2; z = [x; y]
-N = 25 # Sparsity plots
-@time bihar = biharmonicoperator(D(2.0, 2.0), N)
-N = 100 # Solution plots
-@time bihar = biharmonicoperator(D(2.0, 2.0), N)
-@time errfuncfs = Fun((x,y)->(1 + erf(5(1 - 10((x - 0.5)^2 + (y)^2)))), S, N*(N+1)).coefficients
-N = 200 # solutionblocknorms plot
-ff = (x,y)->exp(-1000((x-0.2)^2 + (y-0.2)^2))
-pts = pointswithweights(S, N*(N+1))[1]
-vals = [ff(pt...) for pt in pts]
-HDTP = OrthogonalPolynomialFamilies.HalfDiskTransformPlan(S, vals)
-@time f4cfs = HDTP * vals
-S
-
-
-
-using JLD
 B = BigFloat
+precision(B)
+setprecision(800)
+a, b = 1.0, 1.0
+x, y = 0.42, -0.2; z = [x; y]
 # save("experiments/saved/laplacian-w11.jld", "Lw11", Δw)
 # save("experiments/saved/biharmonic.jld", "bihar", bihar)
 # save("experiments/saved/halfdiskspace11.jld", "opnorms", S.opnorms, "A", S.A, "B", S.B, "C", S.C, "DT", S.DT)
[…]

### 2019-05-14 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
improved 1D transform()

@@ -3,10 +3,10 @@ module OrthogonalPolynomialFamilies
 using ApproxFun
 import ApproxFun: evaluate, domain,
                     domainspace, rangespace, bandwidths, prectype, canonicaldomain, tocanonical,
-                    spacescompatible, points, transform, itransform, AbstractProductSpace, 
+                    spacescompatible, points, transform, itransform, AbstractProductSpace,
                     checkpoints, plan_transform, clenshaw
 import ApproxFunOrthogonalPolynomials: PolynomialSpace, recα, recβ, recγ, recA, recB, recC
-import ApproxFunBase: tensorizer, columnspace            
+import ApproxFunBase: tensorizer, columnspace
 import Base: in, *
 using StaticArrays
 using FastGaussQuadrature
@@ -187,14 +187,20 @@ spacescompatible(A::OrthogonalPolynomialSpace, B::OrthogonalPolynomialSpace) =
 function transform(S::OrthogonalPolynomialSpace, vals::Vector{T}) where T
     n = length(vals)
     pts, w = pointswithweights(S, n)
-    # Vandermonde matrix transposed, including weights and normalisations
-    Ṽ = Array{T}(undef, n, n)
+    getopptseval(S, n-1, pts)
+    cfs = zeros(n)
     for k = 0:n-1
-        pk = Fun(S, [zeros(k); 1])
-        nrm = sum([pk(pts[j])^2 * w[j] for j = 1:n])
-        Ṽ[k+1, :] = pk.(pts) .* w / nrm
+        cfs[k+1] = inner2(S, opevalatpts(S, k+1, pts), vals, w) / getopnorm(S)
     end
-    Ṽ * vals
+    cfs
+    # # Vandermonde matrix transposed, including weights and normalisations
+    # Ṽ = Array{T}(undef, n, n)
+    # for k = 0:n-1
+    #     pk = Fun(S, [zeros(k); 1])
+    #     nrm = sum([pk(pts[j])^2 * w[j] for j = 1:n])
+    #     Ṽ[k+1, :] = pk.(pts) .* w / nrm
+    # end
+    # Ṽ * vals
 end
 
 # Inputs: OP space, coeffs of a function f for its expansion in the OPSpace OPs
@@ -202,13 +208,19 @@ end
 function itransform(S::OrthogonalPolynomialSpace, cfs::Vector{T}) where T
[…]

### 2019-05-14 papers/halfdisk/halfdisk.tex
Added comments

@@
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
-Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids.
+Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids \bstodo{Reference?}.
 
@@
 
+\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig. Add in solution block norms fig  - what should RHSs be?}
+
 Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
@@
 \subsection{Biharmonic equation}
+
+\bstodoinline{Add in solution block norms fig - what should RHSs be?}
 

### 2019-05-16 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Fixed operator clenshaw method

@@ -552,8 +552,8 @@ function *(P::HalfDiskTransformPlan{T}, vals) where T
     @show m1, m2
 
     ret = zeros(m1)
-    # We store the norms of the OPs
-    getopnorms(P.S, N)
+    resizedata!(P.S, N)
+    getopnorms(P.S, N) # We store the norms of the OPs
     for i = 1:m2
         if i % 100 == 0
             @show m2, i
@@ -1410,34 +1410,38 @@ function operatorclenshawmatrixBmG(S::HalfDiskSpace{<:Any, <:Any, T}, A, id, Jx,
     end
     B
 end
-function operatorclenshaw(cfs, S::HalfDiskSpace)
+function operatorclenshaw(cfs, S::HalfDiskSpace, N)
+    # Outputs the operator NxN-blocked matrix operator corresponding to the
+    # function f given by its coefficients of its expansion in the space S
     m̃ = length(cfs)
-    N = -1 + Int(round(sqrt(1+2(m̃-1))))
-    resizedata!(S, N+1)
-    m = Int((N+1)*(N+2)/2)
-    Jx = sparse(jacobix(S, N))
-    Jy = sparse(jacobiy(S, N))
-    if m̃ < m
+    M = getnk(m̃)[1] # Degree of f
+    m = getopindex(M, M)
+    if m̃ < getopindex(M, M)
+        # Pad cfs to correct size
         resize!(cfs, m)
         cfs[m̃+1:end] .= 0.0
     end
+    resizedata!(S, M)
+    Jx = sparse(jacobix(S, M))
+    Jy = sparse(jacobiy(S, M))
+
     @show "Operator Clenshaw"
     P0 = 1.0
-    if N == 0
+    if M == 0
         return cfs[1] * id * P0
     end
     id = sparse(I, size(Jx))
-    if N == 1
+    if M == 1
         P0 * (operatorclenshawvector(S, cfs[1], id)[1] - (operatorclenshawmatrixDT(S, S.DT[1], id) * operatorclenshawmatrixBmG(S, S.B[1], id, Jx, Jy))[1])
     end
-    n = N; @show "Operator Clenshaw", N, n
-    inds2 = m-N:m
+    n = M; @show "Operator Clenshaw", N, n
+    inds2 = m-M:m
     γ2 = operatorclenshawvector(S, view(cfs, inds2), id)
[…]

### 2019-05-16 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -99,10 +99,10 @@ end
 end # NOTE: Fun(f, S) is a λ function, and not a Fun
 
 @testset "Evaluation for HalfDiskSpace (transform())" begin
-    n = 10; a = 0.5; b = 1.5
+    n = 10; a = 1.0; b = 2.0
     f = (x,y) -> y*x^2 + x
-    F = HalfDiskFamily()
-    S = F(a, b)
+    D = HalfDiskFamily()
+    S = D(a, b)
     pts = points(S, n)
     vals = [f(pt...) for pt in pts]
     cfs = transform(S, vals)

### 2019-05-17 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Change to opnorms() method

@@ -511,18 +511,18 @@ function getopnorms(S::HalfDiskSpace{<:Any, <:Any, T}, k) where T
         resize!(S.opnorms, k+1)
         P = S.family.P(S.b, S.b)
         getopnorm(P)
-        # TODO: This is hardcoded, need to find a better way
-        # Doing H(S.a, S.b + k + 0.5) for large k leads to H.weight
-        # reaching maximum number of coefficients when calculating due to
-        # BigFloats
-        X = Fun(T(0)..1)
-        W1 = X^T(S.a) * (1-X^2)^(T(S.b)+0.5)
-        W = W1 * (1-X^2)^m
+        # # TODO: This is hardcoded, need to find a better way
+        # # Doing H(S.a, S.b + k + 0.5) for large k leads to H.weight
+        # # reaching maximum number of coefficients when calculating due to
+        # # BigFloats
+        # X = Fun(T(0)..1)
+        # W1 = X^T(S.a) * (1-X^2)^(T(S.b)+0.5)
+        # W = W1 * (1-X^2)^m
         for j = m+1:k+1
-            # H = S.family.H(S.a, S.b + k + 0.5)
-            # S.opnorms[j] = getopnorm(H) * P.opnorm[1]
-            S.opnorms[j] = sum(W) * P.opnorm[1]
-            W = W * (1-X^2)
+            H = S.family.H(S.a, S.b + j - 1 + 0.5)
+            S.opnorms[j] = getopnorm(H) * P.opnorm[1]
+            # S.opnorms[j] = sum(W) * P.opnorm[1]
+            # W = W * (1-X^2)
         end
     end
     S

### 2019-05-17 papers/halfdisk/halfdisk.tex
Some edits. Added figs.

@@
 $$
-W^{(a,b)}(x,y) = x^a (1-x^2-y^2)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
+W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b, \qfor 0 \leq x \leq 1,\quad -\sqrt{1-x^2} \leq y \leq \sqrt{1-x^2},
 $$
-with corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$.
-We define these to be orthogonalised lexicographically, that is, \sotodo{Are these meant to be orthonormal?}
+The corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
 $$
@@
 \end{align*}
-and define the associated inner products by: \bstodo{Is it best to define these ip s with the normalisation?}
+and define the associated inner products by:
 \begin{align}
-	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x, \label{eqn:ipgenjac} \\
-	\ip<p, q>_{w_P^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \:\int_\gamma^\delta p(y) \: q(y) \: w_P^{(a,b)}(y)\: \D y. \label{eqn:ipjac}
+	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ipgenjac} \\
+	\ip<p, q>_{w_P^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: w_P^{(a,b)}(y)\: \D y \label{eqn:ipjac}
 \end{align}
 where
-\begin{align*}
-	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: w_P^{(a,b)}(y) \: \D y.
-\end{align*}
+\begin{align}
+	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: w_P^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
+\end{align}
 Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{P_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
@@
 \begin{align*}
-	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b+c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
-\end{align*}
-where
-$\ip< f, \: g >_W^{(a,b,c,d)} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x$.
+	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
+\end{align*}
+where for $f, g : \Omega \to \R$ the inner product is defined as 
+\begin{align*}
+	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
+\end{align*}
 \end{definition}
@@
 \end{align*}
-Note here the we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
+Note here we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
 This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.
@@
 The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials.
+The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
+\begin{align*}
+	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
+	\ip<p, q>_{w_P^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y,
+\end{align*}
+i.e.
+$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}} = 1$ $\forall n = 0,1,2,\dots$.
+For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}$.
+
 Thus we arrive at a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
@@
 \end{align}
-orthogonal on \(\Omega\) with respect to the weight
-\begin{align*}
-W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)}) = x^a \: (1-x^2-y^2)^b.
-\end{align*}
-
-The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
-\begin{align*}
-	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
-	\ip<p, q>_{w_P^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y,
-\end{align*}
-where 
-\begin{align*}
-	\normgenjac^{(a,b)} := \int_0^1 \: \genjacw^{(a,b)}(x) \: \D x, \quad \normjac^{(b)} := \int_{-1}^1 \: w_P^{(b)}(y) \: \D y,
-\end{align*}
-i.e.
-$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}} = 1$ $\forall n = 0,1,2,\dots$.
-For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}$.
-
-Further, we have that the norms of the 2D half-disk OPs for a particular parameter family $(a,b)$ are then given by
+orthogonal on $\Omega$ with respect to the weight
+\begin{align*}
+	W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) =  \frac{x^a \: (1-x^2-y^2)^b}{\normgenjac^{(a,b)} \normjac^{(b)}}.
+\end{align*}
+That is, they are orthogonal (for given parameters $a,b$) with respect to the inner product given by
+\begin{align*}
+	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \hdopnkab(x,y) \: \hdopmjab(x,y) \: \Wab(x,y) \: \D y \: \D x.
+\end{align*}
+We can see that they are indeed orthogonal using the change of variable $t = \frac{y}{\rho(x)}$:
 \begin{align}
-	\norm{\hdopnkab}_\Wab^2 &=  \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \rho(x)^{2k} \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \Wab(x,y) \: \D y \: \D x \nonumber \\
-	&= \iint_\Omega \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: {P_k^{(b,b)}\fpr(\frac{y}{\rho(x)})^2} \: \genjacw^{(a,b+k)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \: \D y \: \D x \nonumber \\
-	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x)^2 \: \genjacw^{(a,b+k+\half)}(x) \: \D x \Big) \: \Big( \int_\gamma^\delta \: {P^{(b,b)}}(t)^2 \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
-	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)} \label{eqn:normhdop}
+	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \Big[ \genjacnmk^{(a,b+k+\half)}(x) \: \genjacmmj^{(a,b+j+\half)}(x) \: \rho(x)^{k+j}  \nonumber \\
+		&\quad \quad \quad \quad \quad \quad \cdot \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}) \: P_j^{(b,b)}\fpr(\frac{y}{\rho(x)})  \: \Wab(x,y) \Big] \: \D y \: \D x \nonumber \\
+	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x) \: {\genjacmmj^{(a,b+j+\half)}}(x) \: \genjacw^{(a,b+\half(k+j+1))}(x) \: \D x \Big) \nonumber \\
+	&\quad \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \: {P_k^{(b,b)}}(t) \:  {P_j^{(b,b)}}(t) \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
+	&= \normjac^{(b)} \: \delta_{k,j} \: \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x) \: {\genjac_{m-k}^{(a,b+k+\half)}}(x) \: \genjacw^{(a,b+k+\half)}(x) \: \D x \nonumber \\ 
+	&= \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)} \: \delta_{n,m} \: \delta_{k,j}. \label{eqn:normhdop}
 \end{align}
-where we have used the change of variable $t = \frac{y}{\rho(x)}$.
-Again, for clairty we use the notation that $\ip<p, q>_{\Wab} := \iint_\Omega p(x,y) \: q(x,y) \: \Wab(x,y) \: \D y \: \D x$ and $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$ .
+Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.
 
@@
 
-We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as \sotodo{These hold only for orthonormal, which isn't the case for $P_n$}
+We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as 
 \begin{align}
-x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) 
-\label{eqn:Hrecurrence} \\
+	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Hrecurrence} \\
 	y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
@@
 	\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
-	\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
-	\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
-	\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
-	\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
-	\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{\normgenjac^{(a, b + k - \half)}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k+1/2)}} \\
-	\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{\normgenjac^{(a, b + k + \frac{3}{2})}} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
+	\betaab_{n,k,1} &:= \delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \: \rho(x)^2 \genjacnmk^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k-1/2)}} \\
+	\betaab_{n,k,2} &:= \delta_{k}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-2}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
+	\betaab_{n,k,3} &:= \delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \: \rho(x)^2 \genjac_{n-k+1}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k-1/2)}} \\
+	\betaab_{n,k,4} &:= \delta_{k}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjac_{n-k-1}^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}} \\
+	\betaab_{n,k,5} &:= \delta_{k-1}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \: \rho(x)^2 \genjac_{n-k+2}^{(a, b+k-1/2)}>_{\genjacw^{(a, b+k-1/2)}} \\
+	\betaab_{n,k,6} &:= \delta_{k}^{(b)} \: \ip<\genjacnmk^{(a, b+k+1/2)}, \genjacnmk^{(a, b+k+{3/2})}>_{\genjacw^{(a, b+k+{3/2})}}. 
 \end{align*}
@@
 \end{align*}
-Recall from equation (\ref{eqn:normhdop}) that $\norm{\hdopnkab}_{\Wab}^2 = \normgenjac^{(a,b+k+\half)} \: \normjac^{(b)}$.
+Recall from equation (\ref{eqn:normhdop}) that $\norm{\hdopmjab}_{\Wab}^2 = \normgenjac^{(a,b+j+\half)} \: \normjac^{(b)}$.
 Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
@@
 	&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(x) \: \genjacmmj^{(a, b+j+\half)}(x) \: \genjacw^{(a, b+1+\half(k+j))}(x) \: \D x \Big) \nonumber \\
-	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: w_P^{(b)}(t) \: \D t \Big) \\
-	&= 
+	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: w_P^{(b)}(t) \: \D t \Big)
+\end{align*}
+\begin{align*}
+	=
 	\begin{cases}
-    		\delta_k^{(b)} \: \normjac^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
-		\delta_{k-1}^{(b)} \: \normjac^{(b)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
+    		\delta_k^{(b)} \: \normjac^{(b)} \: \normgenjac^{(a,b+k+\frac{3}{2})} \: \ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
+		\delta_{k-1}^{(b)} \: \normjac^{(b)} \: \normgenjac^{(a,b+k-\half)} \: \ip<\genjacnmk^{(a, b+k+\half)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k-\half)}} \quad& \text{if } j = k-1 \\
 		0 & \text{otherwise}
-      	\end{cases}.
-\end{align*}
-where $\ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}}$ and $\ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k+\half)}}$ are zero for $m < n-1$.
+      	\end{cases}
+\end{align*}
+where $\ip<\genjacnmk^{(a, b+k+\half)}, \genjac_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{\genjacw^{(a, b+k+\frac{3}{2})}}$ and $\ip<\genjacnmk^{(a, b+k+\half)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b+k-\half)}>_{\genjacw^{(a, b+k-\half)}}$ are zero for $m < n-1$.
 
@@
 \end{align*}
-where \sotodo{$C^x_n := \big( A^x_n \big)^\top$ is only true if orthonormal}
+where
 \begin{align*}
@@
 \begin{proof}
+First, note that:
+\begin{align}
+	\genjacw^{(a,b) \: \prime}(x) &= a \: \genjacw^{(a-1,b)}(x) - 2b \: \genjacw^{(a+1,b-1)}(x), \label{eqn:derivativeofweightgenjac} \\
+	w_P^{(b) \: \prime}(y) &= - 2b \: y \: w_P^{(b-1)}(y), \label{eqn:derivativeofweightjac}
+\end{align}
+
 We proceed with the case for the operator $D_y^{(a,b)}$ for partial differentiation by $y$.
@@
 \begin{align*}
-	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} &= \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} \\
-	&= \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
-\end{align*}
-Now, integrating by parts, noting that the weight $w_P^{(b)}$ is a polynomial degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
-\begin{align*}
-	\ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} &= \ip< P_k^{(b,b) \: \prime}, \: w_P^{(b+1)} \: P_{j}^{(b+1,b+1)} >_1 \\
-	&= \ip< P_k^{(b,b)}, \: \ddy [w_P^{(b+1)} \: P_{j}^{(b+1,b+1)}] >_1 \\
-	&= \ip< P_k^{(b,b)}, \: P_{j}^{(b+1,b+1) \: \prime} >_{w_P^{(b+1)}} - 2 b \ip< P_k^{(b,b)}, \: y \: P_{j}^{(b+1,b+1)} >_{w_P^{(b)}} \\
-	&= \ip< P_k^{(b,b)}, \: w_P^{(1)} \: P_{j}^{(b+1,b+1) \: \prime} -2b \:y \: P_{j}^{(b+1,b+1)} >_{w_P^{(b)}} 
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} &= \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \\
+	&\quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} \\
+	&= \normgenjac^{(a, b+1+\half(k+j))} \: \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \\
+		&\quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+\end{align*}
+Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $w_P^{(b)}$ is a polynomial degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
+\begin{align*}
+	\normjac^{(b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} &= \int_{-1}^1 \: P_k^{(b,b) \: \prime}(y) \: P_{j}^{(b+1,b+1)}(y) \: w_P^{(b+1)}(y) \: \D y \\
+	&= - \int_{-1}^1 \: P_k^{(b,b)}(y) \: \ddy [w_P^{(b+1)}(y) \: P_{j}^{(b+1,b+1)}(y)] \: \D y \\
+	&= - \int_{-1}^1 \: P_k^{(b,b)} \: [ P_{j}^{(b+1,b+1) \: \prime} \: w_P^{(b+1)} - 2 b \: y \: P_{j}^{(b+1,b+1)} \:w_P^{(b)} ] \: \D y \\
+	&= - \: \normjac^{(b)} \: \ip< P_k^{(b,b)}, \: w_P^{(1)} \: P_{j}^{(b+1,b+1) \: \prime} -2b \:y \: P_{j}^{(b+1,b+1)} >_{w_P^{(b)}}
 \end{align*}
@@
 \begin{align*}
-	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,k-1} >_{W^{(a, b+1)}} &= \ip< \genjacnmk^{(a,b+k+\half)}, \genjac_{m+1-k}^{(a,b+k+\half)} >_{\genjacw^{(a,b+k+\half)}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}},
-\end{align*}
-which is zero for $m < n-1$. Finally, we simply have that \bstodo{Make sure correct ip definition is being used here}
-\begin{align*}
-	c_{n-1,k-1} &= \frac{1}{\normjac^{(b)}} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+	\ip< \genjacnmk^{(a,b+k+\half)}, \genjac_{m+1-k}^{(a,b+k+\half)} >_{\genjacw^{(a,b+k+\half)}} = \delta_{n,m+1}
+\end{align*}
+showing the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally, we simply have that 
+\begin{align*}
+	c_{n-1,k-1} &= \normgenjac^{(a, b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
 \end{align*}
@@
 	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+	&= \normgenjac^{(a, b+\half)} \: \normjac^{(b+1)} \: \Big\{ \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+	& \quad \quad \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
 		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  
-\end{align*}
+	& \quad \quad \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}} \Big\}
+\end{align*}
+
 We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
-Now, using integration-by-parts and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
-\begin{align*}
-	&\ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
-	&= \ip<\genjacnmk^{(a, b+k+\half) \: \prime}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half)} >_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \ddx \Big[\genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half)} \Big] >_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_1 \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime}>_1 \\
-	&= \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}}
+Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
+\begin{align*}
+	&\normgenjac^{(a, b+\half)} \: \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime}, \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}}  \\
+	&= \int_0^1 \: \genjacnmk^{(a, b+k+\half) \: \prime} \: \genjacmmj^{(a+1, b+1+j+\half)} \: \genjacw^{(a+1, b+1+\half(k+j+1))} \: \D x \\
+	&= - \int_0^1 \: \genjacnmk^{(a, b+k+\half)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1+j+\half)} \: \genjacw^{(a+1, b+1+\half(k+j+1))} \Big] \: \D x \\
+	&= - \int_0^1 \: \genjacnmk^{(a, b+k+\half)} \: \Big\{ \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime} \\
+	&\quad \quad \quad \quad + \Big[ (a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} \Big\} \: \D x \\
+	&= - \: \normgenjac^{(a,b+k+\half)} \:  \Big\{ \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}} \\
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \Big\}
 \end{align*}
@@
 \begin{align*}
-	&\ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
-	&= \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
+	&\normgenjac^{(a, b+\half)} \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\
+	&= \normgenjac^{(a, b+k+\half)} \: \ip< \genjacnmk^{(a, b+k+\half)}(x), \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}}
 \end{align*}
@@
 
-Further, we note that there exist "parameter transformation" matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+There exist "parameter transformation" matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+
 \begin{definition}\label{def:parametertransformationoperators}
@@
 \begin{align*}
-	\ip< \hdopnkab, \hdop^{(a+\lambda, b+\mu)}_{m,j} >_{W^{(a+\lambda, b+\mu)}} &= \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a+\lambda,b+\mu)}} \\
-	&\quad \quad \quad \cdot \ip< P_k^{(b,b)}, \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b+\mu)}} \\
-	&= \ip< \genjacnmk^{(a,b+k+\half)}, \genjacw^{(\lambda, \: \mu+\half(j-k))} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a,b+k+\half)}} \\
-	&\quad \quad \quad \cdot \ip< P_k^{(b,b)}, \: w_P^{(\mu)} \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b)}}
+	&\ip< \hdopnkab, \hdop^{(a+\lambda, b+\mu)}_{m,j} >_{W^{(a+\lambda, b+\mu)}} \\
+	&\quad = \normgenjac^{(a+\lambda, b+\mu)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a+\lambda,b+\mu)}} \\ 
+	&\quad \quad \quad \cdot \: \normjac^{(b+\mu)} \: \ip< P_k^{(b,b)}, \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b+\mu)}} \\
+	&\quad= \normgenjac^{(a, b+k+\half)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacw^{(\lambda, \: \mu+\half(j-k))} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a,b+k+\half)}} \\
+	&\quad \quad \quad \cdot \: \normjac^{(b)} \: \ip< P_k^{(b,b)}, \: w_P^{(\mu)} \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b)}}
 \end{align*}
@@
 
-To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that\sotodo{check this}
+To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
 \begin{align*}
@@
 
+% FIGURES
 \begin{figure}
@@
 
+\begin{figure}[t]
+        	\includegraphics[scale=0.4]{solutionblocknorms-poisson}
+	\centering
+	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
+This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
+        	\label{fig:solutionblocknorms-poisson}
+\end{figure}
+
+\begin{figure}[t]
+        	\includegraphics[scale=0.4]{solutionblocknorms-helmholtz-k=20}
+	\centering
+	\caption{The norms of each block of the computed solution of the helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
+        	\label{fig:solutionblocknorms-helmoltz}
+\end{figure}
+
+\begin{figure}[t]
+        	\includegraphics[scale=0.4]{solutionblocknorms-biharmonic}
+	\centering
+	\caption{The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
+        	\label{fig:solutionblocknorms-biharmonic}
+\end{figure}
+
 Find \(u(x,y)\) given a function \(f(x,y)\) such that:
@@
 In Figure \ref{fig:solution-poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the half-disk $\Omega$.
-In Figure \ref{fig:solutionblocknorms} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:solutionblocknorms-poisson} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
 
-\begin{figure}
-        	\includegraphics[scale=0.5]{solutionblocknorms}
-	\centering
-	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
-This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
-        	\label{fig:solutionblocknorms}
-\end{figure}
-
 \subsection{Inhomogenous variable-coefficient Helmholtz}
 
-\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig. Add in solution block norms fig  - what should RHSs be?}
+\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig (do N=200, k=20). Add in solution block norms fig  (for N=200, k = 20).}
 
@@
 	\begin{cases}
-    		\Delta u(x,y) + k^2 v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
+    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
 		u(x,y) = 0 \quad \text{on } \partial \Omega
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.
+In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
+In Figure \ref{fig:solutionblocknorms-helmholtz} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
+\bstodo{Correct the explanation of what we observe} We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+For the final Gaussian bump example, we see that we achieve spectral convergence.
 
@@
 	\begin{cases}
-    		\Delta u(x,y) + k^2 v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
+    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
 		u(x,y) = c \in \R \quad \text{on } \partial \Omega
@@
 	\begin{cases}
-    		\Delta \tilde{u}(x,y) + k^2 v(x,y) \; \tilde{u}(x,y) = f(x,y) - k^2 v(x,y) \; =: g(x,y)  \quad \text{in } \Omega \\
+    		\Delta \tilde{u}(x,y) + k^2 \: v(x,y) \; \tilde{u}(x,y) = f(x,y) - c \: k^2 \: v(x,y) \; =: g(x,y)  \quad \text{in } \Omega \\
 		\tilde{u}(x,y) = 0 \quad \text{on } \partial \Omega
@@
 \subsection{Biharmonic equation}
-
-\bstodoinline{Add in solution block norms fig - what should RHSs be?}
 
@@
 In Figure \ref{fig:solution-biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
+In Figure \ref{fig:solutionblocknorms-biharmonic} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
+\bstodo{Correct the explanation of what we observe} We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+For the final Gaussian bump example, we see that we achieve spectral convergence.
 

### 2019-05-17 papers/halfdisk/halfdisk.tex
Include new image. Update text.

@@
 	\centering
-	\includegraphics[scale=0.3]{solution-helmholtz}
+	\includegraphics[scale=0.3]{solution-helmholtz-k=100}
 	\caption{}
@@
     	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
+(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
 (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
@@
 
-\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig (do N=200, k=20). Add in solution block norms fig  (for N=200, k = 20).}
-
 Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
-In Figure \ref{fig:solutionblocknorms-helmholtz} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
+In Figure \ref{fig:solutionblocknorms-helmoltz} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.

### 2019-05-21 papers/halfdisk/halfdisk.tex
Fixed typo
readthrough

@@
 \begin{abstract}
-In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
@@
 where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.
+The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:construction}.
 
@@
 Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
+In particular, by exploiting the connection with one-dimensional orthogonal polynomials we are able to construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
+This compares favourably to $O(p^6)$ operations if one proceeds na\"ively.
 Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
 
-\sotodoinline{Mention that this is a computational challenge}
-
-\sotodoinline{Make clear this is more than plugging in the basis into a Petrov--Galerkin method.}
+An alternative approach is to use \sotodoinline{Compare with "Discontinuous Galerkin scheme for the spherical shallow water equations with applications to tsunami modeling and prediction"}
 
@@
 Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
-Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids \bstodo{Reference?}.
+Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
 
@@
 
-\noindent \secref{OPs}: We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk.
+\noindent \secref{OPs}: We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.
 
@@
 \noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods.
+This lays the groundwork for a future $hp$-finite element method in a disk, where the elements capture the circular geometry precisely.
 
@@
 
-By using a similar process to \cite[p55--56]{dunkl2014orthogonal} we can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).
-We generalise this process to demonstrate a method of constructing multi-parameter families of 2D orthogonal polynomials on disk-slice and trapezium shaped domains $\Omega$ from multi-parameter families of 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).
-
-\begin{proposition}\label{prop:construction}
+We can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).
+
+\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
 Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
@@
 \end{proposition}
-\begin{proof}
-See \cite[p55--56]{dunkl2014orthogonal}.
-\end{proof}
-
-For disk slices and trapeziums, we can specialise the framework from Proposition \ref{prop:construction} in the following definition.
+
+For disk slices and trapeziums, we  specialise  Proposition \ref{prop:construction} in the following definition.
 \begin{definition}\label{def:construction}
@@
 \item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w^{(a,b)}_P\) is an even function (i.e.
-$a = b$, and we can hence denote the weight as $w^{(a)}_P(x) = (\delta-x^2)^a$).
+$a = b$, and we can hence denote the weight as $w^{(a)}_P(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$).
 \end{enumerate}
@@
 Note here we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
-This is for ease of writing and computation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.
-
-The weight $w_P(x)$ is an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials  $\{P_n^{(b, b)}\}$.
+This is for ease of presentiation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.
+
+The weight $w_P(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials  $\{\tilde P_n^{(b, b)}\}$ \sotodo{Change $P_n -> \tilde P_n$ everywhere}.
 The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials.
@@
 \end{align}
-where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\). We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$:
+where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\).
+We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$.
+Importantly, we can deduce sparsity in the recurrence relationships:
 
@@
 \begin{align*}
-	c_{m,j} = \frac{\ip< y \: \hdopnkab, \hdopmjab >_{\Wab}}{\norm{\hdopmjab}^2_{\Wab}}.
+	c_{m,j} = {\ip< y \: \hdopnkab, \hdopmjab >_{\Wab}}{\norm{\hdopmjab}^{-2}_{\Wab}}.
 \end{align*}
@@
 $J_x^{(a,b)}, J_y^{(a,b)}$ have banded-block-banded structures.
-A block-banded matrix is a block $n$-diagonal matrix where $n$ is reasonably small for the size of the full matrix.
+A block-banded matrix is a block  matrix where only a finite number of block-bands are nonzero.
 A banded-block-banded matrix is one where the blocks themselves are banded matrices.
-$J_x^{(a,b)}, J_y^{(a,b)}$ are given by:
+For example, $J_x^{(a,b)}, J_y^{(a,b)}$ are block-tridiagonal:
 \begin{align*}
@@
 \end{align*}
-where
+where the blocks themselves are diagonal for $J_x^{(a,b)}$,
 \begin{align*}
@@
 \end{align*}
+and tridiagonal for $J_u^{(a,b)}$,
 \begin{align*}
@@
 
-Combining each system in (\ref{eqn:jacobimatricesdefinition}) we can write the block-tridiagonal system
+We can combine each system in (\ref{eqn:jacobimatricesdefinition}) into a block-tridiagonal system:
 \begin{align*}
@@
 	\hdopab_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
-\end{pmatrix}.
-\end{align*}
-where we note $\hdopab_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}$, and for each $n = 0,1,2\dots$:
+\end{pmatrix},
+\end{align*}
+where we note $\hdopab_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}$, and for each $n = 0,1,2\dots$,
 \begin{align*}
@@
 
-Recall that, for $a, b$ be general parameters, a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
+Recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
 Recall also that $\Wab(x,y) := \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $\genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
@@
 
-\begin{lemma}\label{lemma:sparsityofdifferentialoperators}
+\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
 The operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
-\end{lemma}
+\sotodoinline{Give block-bandwidths}
+\end{theorem}
 
@@
 We can use an integration-by-parts argument to show that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
-First, note that
-\begin{align*}
-	c_{m,j} = \frac{\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{m,j}}^2_{W^{(a, b+1)}}}.
+First, note that\sotodo{Maybe $c_{m,j}^ys$ to avoid confusion}
+\begin{align*}
+	c_{m,j} = {\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
 \end{align*}
@@
 As before, we can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-1, k-2$ and $0 \le j \le m$.
-First, note that
-\begin{align*}
-c_{m,j} = \frac{\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^2_{W^{(a+1, b+1)}}}.
+First, note that\sotodo{Maybe $c_{m,j}^x$ to avoid confusion}
+\begin{align*}
+c_{m,j} = {\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^{-2}_{W^{(a+1, b+1)}}}.
 \end{align*}
@@
 \end{align}
-Now, using the change of variable $t= \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that
+Now, using the change of variable $t= \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that\sotodo{Try to avoid overflow}
 \begin{align*}
@@
 
-There exist "parameter transformation" matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+There exist ``parameter transformation'' matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 
@@
 	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bighdop^{(a+1,b+1)}(x,y) \\
-	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b)} \Big)^\top \: \bigW^{(a-1,b)}(x,y) \\
-	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b+1)} \Big)^\top \: \bigW^{(a,b-1)}(x,y) \\
-	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a+1,b+1)} \Big)^\top \: \bigW^{(a-1,b-1)}(x,y).
+	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a-1,b)} \Big)^\top \: \bigW^{(a-1,b)}(x,y) \\
+	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a,b-1)} \Big)^\top \: \bigW^{(a,b-1)}(x,y) \\
+	\bigW^{(a,b)}(x,y) &= \Big(T_W^{(a,b)\to(a-1,b-1)} \Big)^\top \: \bigW^{(a-1,b-1)}(x,y).
 \end{align*}
@@
 The operator matrices $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$, $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ from Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure.
+\sotodoinline{State block-bandwidths}
 \end{lemma}
@@
 \begin{align*}
-	c_{m,j} = \frac{\ip< \hdopnkab, \hdop^{(a+\lambda, b+\mu)}_{m,j} >_{W^{(a+\lambda, b+\mu)}}}{\norm{\hdop^{(a+\lambda, b+\mu)}_{m,j}}^2_{W^{(a+\lambda, b+\mu)}}}.
+	c_{m,j} = {\ip< \hdopnkab, \hdop^{(a+\lambda, b+\mu)}_{m,j} >_{W^{(a+\lambda, b+\mu)}}}{\norm{\hdop^{(a+\lambda, b+\mu)}_{m,j}}^{-2}_{W^{(a+\lambda, b+\mu)}}}.
 \end{align*}
@@
 
-We can then obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
+General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
+As a canonical example, we can obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
 $$
@@
 
-We can use the same technique for constructing a sparse Laplacian operator to construct the Biharmonic operator $\Delta^2$, that ensures we use a basis that satisfies both zero Dirichlet and Neumann conditions.
-On this note, we can also obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopoo(x,y)$ to coefficients in the space $\bighdop^{(2,2)}(x,y)$.
+Another important example is the Biharmonic operator $\Delta^2$, where we assume zero Dirichlet and Neumann conditions.
+To construct this operator, we first note that we can obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopoo(x,y)$ to coefficients in the space $\bighdop^{(2,2)}(x,y)$.
 We denote this matrix operator that acts on the coefficients vector as $\laplaceoo$, and is given by
@@
 
-To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a version of the Stieltjes algorithm. Recall that
-\begin{align*}
-	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x), \quad n = 0,1,2,\dots
-\end{align*}
-where we define $\beta_{-1}^{(a,b)} = 0$ and $\genjac_{-1}^{(a,b)}(x) \equiv 0$.
-We use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
+To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
 This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $b$ is large, as required for large $n$ in \eqref{eq:diskpolys}.
-The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm.
+The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm, utilising the {\tt BigFloat} type to handle high-precision calculations.
+
+\noindent{\bf Remark}: 
+This is the most expensive part of the current calculation, but note that we can reuse this computation for multiple partial differential equations.
+Efficient construction of OPs with large parameters in the weights is an important topic, but one tangential to the proposed scheme.
 
@@
 \end{align*}
+\sotodoinline{Explain that even and odd polynomials are treated separately}
 Let $f : \Omega \to \R$ be such that $y \mapsto f(x,y)$ for fixed $x$ is an even function.
@@
 
-The proofs of Lemma \ref{lemma:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
+The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
 We can simply use quadrature to calculate the 1D inner products.
 This proves much cheaper computationally than using the 2D quadrature rule to calculate the 2D inner products.
+
+\sotodoinline{Give complexity}
 
@@
 (b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-(c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
+(c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+\sotodoinline{Combine this figure with subsequent ones}}
         \label{fig:solution}
@@
 We can tackle the problem as follows.
-Fix \(a, b = 1,1\).
 Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{f}$.
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-\bstodo{Correct the explanation of what we observe} We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
 For the final Gaussian bump example, we see that we achieve spectral convergence.
@@
 In Figure \ref{fig:solutionblocknorms-biharmonic} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-\bstodo{Correct the explanation of what we observe} We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
 For the final Gaussian bump example, we see that we achieve spectral convergence.
@@
 
-We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on certain domains given by an algebraic curve - notably here the half-disk - with Dirichlet boundary conditions.
-We can develop $hp-$finite element methods methods to solve PDEs on other polygonal domains by using suitable shaped elements (for example, by using disk-slice elements on the disk, which has applications in fluids).
-
-A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
+We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains given by an algebraic curve---notably here the half-disk---with Dirichlet boundary conditions.
+This forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitable shaped elements, for example, by dividing the disk int disk-slice elements.
+
+%A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
 With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
@@
 
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk slice/trapezium domain.
+In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk-slice/trapezium domain.
 We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.
@@
 \end{align*}
-with 
+with\sotodo{Why $\subset$?} 
 \begin{align*}
@@
 The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such disk-slice domains.
-For instance, if we inspect the proof of Lemma \ref{lemma:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for a disk-slice.
+For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for a disk-slice.
 

### 2019-05-23 papers/halfdisk/halfdisk.tex
Addressed SO comments - reformatted figs, changed labelling of P 1D OPs and added to proofs/lemmas

@@
 
+\newcommand{\bigO}{\mathcal{O}}
 \newcommand{\half}{\frac{1}{2}}
@@
 \newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
-\newcommand{\Plm}{P^m_l}
+\newcommand{\Plm}{\jac^m_l}
 \newcommand{\costheta}{\cos\theta}
@@
 \newcommand{\hdopio}{\hdop^{(1,0)}}
+\newcommand{\jac}{{\tilde P}}
 \newcommand{\genjac}{R}
@@
 \newcommand{\genjacw}{w_\genjac}
+\newcommand{\jacw}{w_\jac}
 \newcommand{\normgenjac}{\omega_\genjac}
-\newcommand{\normjac}{\omega_P}
+\newcommand{\normjac}{\omega_\jac}
 
@@
 with $\rho : (\alpha, \beta) \to \R$.
-Further, let $\genjacw^{(a,b,c)}(x)$ and $w^{(a,b)}_P(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
+Further, let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
 \begin{align*}
@@
 \genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
-w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b
+\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b
 \end{cases}
@@
 	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ipgenjac} \\
-	\ip<p, q>_{w_P^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: w_P^{(a,b)}(y)\: \D y \label{eqn:ipjac}
+	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ipjac}
 \end{align}
@@
 \begin{align}
-	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: w_P^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
+	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
 \end{align}
-Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{P_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
+Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
 
@@
 \item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w^{(a,b)}_P\) is an even function (i.e.
-$a = b$, and we can hence denote the weight as $w^{(a)}_P(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$).
+\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(\jacw^{(a,b)}\) is an even function (i.e.
+$a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$).
 \end{enumerate}
@@
 \begin{align*}
-	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: P_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align*}
@@
 \begin{align*}
-	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: w_P^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
+	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
 \end{align*}
@@
 w_1(x) &:= \genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b \\
-w_2(x) &:= w_P^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
+w_2(x) &:= \jacw^{(b)}(x) := (1-x^2)^b = (1-x)^b \: (1+x)^b,
 \end{cases}
@@
 
-The weight $w_P(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials  $\{\tilde P_n^{(b, b)}\}$ \sotodo{Change $P_n -> \tilde P_n$ everywhere}.
+The weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\jac_n^{(b, b)}\}$.
 The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials.
-The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{P_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
+The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{\jac_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
 \begin{align*}
 	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
-	\ip<p, q>_{w_P^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: w_P^{(b)}(y)\: \D y,
+	\ip<p, q>_{\jacw^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: \jacw^{(b)}(y)\: \D y,
 \end{align*}
 i.e.
-$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{P_{n}^{(b,b)}}_{w_P^{(b)}} = 1$ $\forall n = 0,1,2,\dots$.
-For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{w_P^{(b)}}^2 := \ip<p, p>_{w_P^{(b)}}$.
+$\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{\jac_{n}^{(b,b)}}_{\jacw^{(b)}} = 1$ $\forall n = 0,1,2,\dots$.
+For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{\jacw^{(b)}}^2 := \ip<p, p>_{\jacw^{(b)}}$.
 
@@
 \begin{align}\label{eq:diskpolys}
-	\hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align}
@@
 \begin{align*}
-	W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) =  \frac{x^a \: (1-x^2-y^2)^b}{\normgenjac^{(a,b)} \normjac^{(b)}}.
+	W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: \jacw^{(b)}\fpr(\frac{y}{\rho(x)}) =  \frac{x^a \: (1-x^2-y^2)^b}{\normgenjac^{(a,b)} \normjac^{(b)}}.
 \end{align*}
@@
 	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \Big[ \genjacnmk^{(a,b+k+\half)}(x) \: \genjacmmj^{(a,b+j+\half)}(x) \: \rho(x)^{k+j}  \nonumber \\
-		&\quad \quad \quad \quad \quad \quad \cdot \: P_k^{(b,b)}\fpr(\frac{y}{\rho(x)}) \: P_j^{(b,b)}\fpr(\frac{y}{\rho(x)})  \: \Wab(x,y) \Big] \: \D y \: \D x \nonumber \\
+		&\quad \quad \quad \quad \quad \quad \cdot \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}) \: \jac_j^{(b,b)}\fpr(\frac{y}{\rho(x)})  \: \Wab(x,y) \Big] \: \D y \: \D x \nonumber \\
 	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x) \: {\genjacmmj^{(a,b+j+\half)}}(x) \: \genjacw^{(a,b+\half(k+j+1))}(x) \: \D x \Big) \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \: {P_k^{(b,b)}}(t) \:  {P_j^{(b,b)}}(t) \: w_P^{(b)}(t) \: \D t \Big) \nonumber \\
+	&\quad \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \: {\jac_k^{(b,b)}}(t) \:  {\jac_j^{(b,b)}}(t) \: \jacw^{(b)}(t) \: \D t \Big) \nonumber \\
 	&= \normjac^{(b)} \: \delta_{k,j} \: \int_\alpha^\beta \: {\genjacnmk^{(a,b+k+\half)}}(x) \: {\genjac_{m-k}^{(a,b+k+\half)}}(x) \: \genjacw^{(a,b+k+\half)}(x) \: \D x \nonumber \\ 
@@
 
-We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $P_n^{(b,b)}$ as 
+We can express the three-term recurrences associated to $\genjac_n^{(a,b)}$ and $\jac_n^{(b,b)}$ as 
 \begin{align}
 	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Hrecurrence} \\
-	y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
+	y \jac_n^{(b,b)}(y) &= \delta_n^{(b)} \jac_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} \jac_n^{(b,b)}(y) + \delta_{n-1}^{(b)} \jac_{n-1}^{(b,b)}(y),
 \end{align}
@@
 	&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(x) \: \genjacmmj^{(a, b+j+\half)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b)}(x) \: \D x \Big) \nonumber \\
-	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: w_P^{(b)}(t) \: \D t \Big) \\
+	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} \jac_k^{(b,b)}(t) \: \jac_j^{(b,b)}(t) \: t \: \jacw^{(b)}(t) \: \D t \Big) \\
 	&= \Big( \int^1_0 \genjacnmk^{(a, b+k+\half)}(x) \: \genjacmmj^{(a, b+j+\half)}(x) \: \genjacw^{(a, b+1+\half(k+j))}(x) \: \D x \Big) \nonumber \\
-	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: w_P^{(b)}(t) \: \D t \Big)
+	& \quad \quad \quad\cdot \: \Big( \int^1_{-1} \jac_k^{(b,b)}(t) \: \jac_j^{(b,b)}(t) \: t \: \jacw^{(b)}(t) \: \D t \Big)
 \end{align*}
@@
 \end{align*}
-where we note $\hdopab_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: P_0^{(b,b)}$, and for each $n = 0,1,2\dots$,
+where we note $\hdopab_{0,0}(x,y) \equiv \genjac_0^{(a,b)} \: \jac_0^{(b,b)}$, and for each $n = 0,1,2\dots$,
 \begin{align*}
@@
 \bighdopab_{-1}(x,y) := 0 \\
-\bighdopab_{0}(x,y) := P^{(a,b)}_{0,0} \\
+\bighdopab_{0}(x,y) := \jac^{(a,b)}_{0,0} \\
 \bighdopab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopab_n(x,y) - \Dnt C_n  \,\bighdopab_{n-1}(x,y), \quad n = 0,1,2,\dots.
@@
 \end{tikzpicture} 
-\caption{The Laplace operator acting on vectors of $\hdop_{n,k}=\hdop_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{P^{(2,2)}_{n,k}}$ coefficients.
+\caption{The Laplace operator acting on vectors of $\hdop_{n,k}=\hdop_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{\jac^{(2,2)}_{n,k}}$ coefficients.
 Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{\hdopnkab}$ coefficients, where $(a,b)$ is at the tail of the arrow, and the range is $\smash{\hdop_{n,k}^{(\tilde{a},\tilde{b})}}$ coefficients, where $(\tilde{a},\tilde{b})$ is at the head of the arrow.
@@
 Recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
-Recall also that $\Wab(x,y) := \genjacw^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $\genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.
+Recall also that $\Wab(x,y) := \genjacw^{(a,b)}(x) \: \jacw^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $\genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $\jacw^{(b)}(y) := (1-y^2)^b$ for the half disk case.
 
@@
 \begin{theorem}\label{theorem:sparsityofdifferentialoperators}
-The operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
-\sotodoinline{Give block-bandwidths}
+The operator matrices $D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure. More specifically:
+\begin{itemize}
+	\item $D_x^{(a,b)}$ has a block-bandwidth of $(-1,2)$, with the blocks themselves having a bandwidth of $(0, 2)$.
+  	\item $D_y^{(a,b)}$ has a block-bandwidth of $(-1,1)$, with the blocks themselves having a bandwidth of $(-1,1)$.
+	\item $W_x^{(a,b)}$ has a block-bandwidth of $(2,-1)$, with the blocks themselves having a bandwidth of $(2, 0)$.
+  	\item $W_y^{(a,b)}$ has a block-bandwidth of $(1,-1)$, with the blocks themselves having a bandwidth of $(1,-1)$.
+\end{itemize}
 \end{theorem}
@@
 	\genjacw^{(a,b) \: \prime}(x) &= a \: \genjacw^{(a-1,b)}(x) - 2b \: \genjacw^{(a+1,b-1)}(x), \label{eqn:derivativeofweightgenjac} \\
-	w_P^{(b) \: \prime}(y) &= - 2b \: y \: w_P^{(b-1)}(y), \label{eqn:derivativeofweightjac}
+	\jacw^{(b) \: \prime}(y) &= - 2b \: y \: \jacw^{(b-1)}(y), \label{eqn:derivativeofweightjac}
 \end{align}
@@
 We can use an integration-by-parts argument to show that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
-First, note that\sotodo{Maybe $c_{m,j}^ys$ to avoid confusion}
-\begin{align*}
-	c_{m,j} = {\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
+First, note that
+\begin{align*}
+	c_{m,j}^y = {\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}}}{\norm{\hdop^{(a, b+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
 \end{align*}
@@
 \begin{align*}
-	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} &= \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \\
-	&\quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} \\
-	&= \normgenjac^{(a, b+1+\half(k+j))} \: \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \\
-		&\quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
-\end{align*}
-Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $w_P^{(b)}$ is a polynomial degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
-\begin{align*}
-	\normjac^{(b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{j}^{(b+1,b+1)} >_{w_P^{(b+1)}} &= \int_{-1}^1 \: P_k^{(b,b) \: \prime}(y) \: P_{j}^{(b+1,b+1)}(y) \: w_P^{(b+1)}(y) \: \D y \\
-	&= - \int_{-1}^1 \: P_k^{(b,b)}(y) \: \ddy [w_P^{(b+1)}(y) \: P_{j}^{(b+1,b+1)}(y)] \: \D y \\
-	&= - \int_{-1}^1 \: P_k^{(b,b)} \: [ P_{j}^{(b+1,b+1) \: \prime} \: w_P^{(b+1)} - 2 b \: y \: P_{j}^{(b+1,b+1)} \:w_P^{(b)} ] \: \D y \\
-	&= - \: \normjac^{(b)} \: \ip< P_k^{(b,b)}, \: w_P^{(1)} \: P_{j}^{(b+1,b+1) \: \prime} -2b \:y \: P_{j}^{(b+1,b+1)} >_{w_P^{(b)}}
+	&\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} \\
+	&\quad \quad = \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \\
+	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{j}^{(b+1,b+1)} >_{\jacw^{(b+1)}} \\
+	&\quad \quad = \normgenjac^{(a, b+1+\half(k+j))} \: \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \\
+		&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{j}^{(b+1,b+1)} >_{\jacw^{(b+1)}}.
+\end{align*}
+Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(b)}$ is a polynomial degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
+\begin{align*}
+	\normjac^{(b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{j}^{(b+1,b+1)} >_{\jacw^{(b+1)}} &= \int_{-1}^1 \: \jac_k^{(b,b) \: \prime}(y) \: \jac_{j}^{(b+1,b+1)}(y) \: \jacw^{(b+1)}(y) \: \D y \\
+	&= - \int_{-1}^1 \: \jac_k^{(b,b)}(y) \: \ddy [\jacw^{(b+1)}(y) \: \jac_{j}^{(b+1,b+1)}(y)] \: \D y \\
+	&= - \int_{-1}^1 \: \jac_k^{(b,b)} \: [ \jac_{j}^{(b+1,b+1) \: \prime} \: \jacw^{(b+1)} - 2 b \: y \: \jac_{j}^{(b+1,b+1)} \:\jacw^{(b)} ] \: \D y \\
+	&= - \: \normjac^{(b)} \: \ip< \jac_k^{(b,b)}, \: \jacw^{(1)} \: \jac_{j}^{(b+1,b+1) \: \prime} -2b \:y \: \jac_{j}^{(b+1,b+1)} >_{\jacw^{(b)}}
 \end{align*}
@@
 \begin{align*}
-	c_{n-1,k-1} &= \normgenjac^{(a, b+1)} \: \ip< P_k^{(b,b) \: \prime}, \: P_{k-1}^{(b+1,b+1)} >_{w_P^{(b+1)}}.
+	c_{n-1,k-1}^y &= \normgenjac^{(a, b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{k-1}^{(b+1,b+1)} >_{\jacw^{(b+1)}}.
 \end{align*}
@@
 As before, we can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2$, $j = k, k-1, k-2$ and $0 \le j \le m$.
-First, note that\sotodo{Maybe $c_{m,j}^x$ to avoid confusion}
-\begin{align*}
-c_{m,j} = {\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^{-2}_{W^{(a+1, b+1)}}}.
+First, note that
+\begin{align*}
+	c_{m,j}^x = {\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}}}{\norm{\hdop^{(a+1, b+1)}_{m,j}}^{-2}_{W^{(a+1, b+1)}}}.
 \end{align*}
@@
 \end{align}
-Now, using the change of variable $t= \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that\sotodo{Try to avoid overflow}
+Now, using the change of variable $t= \frac{y}{\rho(x)}$ and (\ref{eqn:rhoderivative}), we have that
 \begin{align*}
 	&\ip<\pddx \hdopnkab, \hdop^{(a+1, b+1)}_{m,j} >_{W^{(a+1, b+1)}} \nonumber \\ 
-	&= \normgenjac^{(a, b+\half)} \: \normjac^{(b+1)} \: \Big\{ \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad \quad \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< P_k^{(b,b)}, \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}}  \nonumber \\ 
-	& \quad \quad \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-		\: \ip< y \: P_k^{(b,b) \: \prime} (y), \: P_j^{(b+1,b+1)} >_{w_P^{(b+1)}} \Big\}
+	&= \normgenjac^{(a, b+\half)} \: \normjac^{(b+1)} \: \Big\{ \ip< x \: \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\ 
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \cdot \: \ip< \jac_k^{(b,b)}, \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}}  \nonumber \\ 
+	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+		\: \ip< \jac_k^{(b,b)}, \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}}  \nonumber \\ 
+	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
+		\: \ip< y \: \jac_k^{(b,b) \: \prime} (y), \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}} \Big\}
 \end{align*}
@@
 	&= - \int_0^1 \: \genjacnmk^{(a, b+k+\half)} \: \Big\{ \genjacw^{(a+1, b+1+\half(k+j+1))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime} \\
-	&\quad \quad \quad \quad + \Big[ (a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} \Big\} \: \D x \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (a+1) \: \genjacw^{(a, b+1+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad -2(b+1+\half(k+j+1)) \: \genjacw^{(a+2, b+\half(k+j+1))} \: \genjacmmj^{(a, b+1+j+\half)} \Big\} \: \D x \\
 	&= - \: \normgenjac^{(a,b+k+\half)} \:  \Big\{ \ip<\genjacnmk^{(a, b+k+\half)}, \: \genjacw^{(1, 1+\half(j-k))} \: \genjacmmj^{(a+1, b+1+j+\half) \: \prime}>_{\genjacw^{(a, b+k+\half)}} \\
-		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: \Big[(a+1) \: \genjacw^{(0,1+\half(j-k))} -2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \Big] \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \Big\}
+		&\quad \quad \quad + \ip<\genjacnmk^{(a, b+k+\half)}, \: (a+1) \: \genjacw^{(0,1+\half(j-k))} \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \\
+		&\quad \quad \quad - \ip<\genjacnmk^{(a, b+k+\half)}, \: 2(b+1+\half(k+j+1)) \: \genjacw^{(2, \half(j-k))} \: \genjacmmj^{(a, b+1+j+\half)} >_{\genjacw^{(a, b+k+\half)}} \Big\}
 \end{align*}
@@
 The operator matrices $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$, $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ from Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure.
-\sotodoinline{State block-bandwidths}
+More specifically:
+\begin{itemize}
+	\item $T^{(a,b)\to(a+1,b)}$ has a block-bandwidth of $(0,1)$, with the blocks themselves being diagonal.
+  	\item $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$ have a block-bandwidth of $(0,3)$, with the blocks themselves having a bandwidth of $(0,2)$.
+	\item $T_W^{(a,b)\to(a-1,b)}$ has a block-bandwidth of $(1,0)$, with the blocks themselves being diagonal.
+  	\item $T_W^{(a,b)\to(a,b-1)}$, $T_W^{(a,b)\to(a-1,b-1)}$ have a block-bandwidth of $(3,0)$, with the blocks themselves having a bandwidth of $(2,0)$.
+\end{itemize}
 \end{lemma}
@@
 	&\quad = \normgenjac^{(a+\lambda, b+\mu)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a+\lambda,b+\mu)}} \\ 
-	&\quad \quad \quad \cdot \: \normjac^{(b+\mu)} \: \ip< P_k^{(b,b)}, \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b+\mu)}} \\
+	&\quad \quad \quad \cdot \: \normjac^{(b+\mu)} \: \ip< \jac_k^{(b,b)}, \: \jac_{j}^{(b+\mu,b+\mu)} >_{\jacw^{(b+\mu)}} \\
 	&\quad= \normgenjac^{(a, b+k+\half)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacw^{(\lambda, \: \mu+\half(j-k))} \: \genjacmmj^{(a+\lambda,b+\mu+j+\half)} >_{\genjacw^{(a,b+k+\half)}} \\
-	&\quad \quad \quad \cdot \: \normjac^{(b)} \: \ip< P_k^{(b,b)}, \: w_P^{(\mu)} \: P_{j}^{(b+\mu,b+\mu)} >_{w_P^{(b)}}
-\end{align*}
-Since $w_P^{(\mu)}$ is a polynomial degree $2\mu$, we have that the above is then zero for $j < k-2\mu$.
+	&\quad \quad \quad \cdot \: \normjac^{(b)} \: \ip< \jac_k^{(b,b)}, \: \jacw^{(\mu)} \: \jac_{j}^{(b+\mu,b+\mu)} >_{\jacw^{(b)}}
+\end{align*}
+Since $\jacw^{(\mu)}$ is a polynomial degree $2\mu$, we have that the above is then zero for $j < k-2\mu$.
 Further, since $\genjacw^{(\lambda, \: \mu+\half(j-k))}$ is a polynomial of degree $\lambda + 2\mu + j - k$, we have that the above is zero for $m-j < n-k - (\lambda + 2\mu + j - k) \iff m < n - \lambda - 2\mu$.
@@
 
-\begin{figure}
-	\centering
-	\begin{subfigure}[t]{0.32\textwidth}
-	\centering
+\begin{figure}[t]
+	\begin{subfigure}{0.32\textwidth}
 	\includegraphics[scale=0.35]{sparsityoflaplacian-w11}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.32\textwidth}
-	\centering
+	\begin{subfigure}{0.32\textwidth}
 	\includegraphics[scale=0.35]{sparsityofhelmholtz}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.32\textwidth}
-	\centering
+	\begin{subfigure}{0.32\textwidth}
 	\includegraphics[scale=0.35]{sparsityofbiharmonic}
 	\end{subfigure}
-	\centering
     	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
@@
 Right: the biharmonic operator $\biharmonic$.}
+        \label{fig:sparsity}
         \centering
-        \label{fig:sparsity}
 \end{figure}
@@
 \begin{align*}
-	W^{(a,b)}(x,y) &= \genjacw^{(a,b)}(x) \: w_P^{(b)}\fpr(\frac{y}{\rho(x)}) \\
-	&= \genjacw^{(a,b)}(s) \: w_P^{(b)}(t) \\
+	W^{(a,b)}(x,y) &= \genjacw^{(a,b)}(x) \: \jacw^{(b)}\fpr(\frac{y}{\rho(x)}) \\
+	&= \genjacw^{(a,b)}(s) \: \jacw^{(b)}(t) \\
 	&= s^a \: \rho(s)^{2b} \: (1-t^2)^b \\
@@
 \end{align*}
-\sotodoinline{Explain that even and odd polynomials are treated separately}
-Let $f : \Omega \to \R$ be such that $y \mapsto f(x,y)$ for fixed $x$ is an even function.
-Note that if $f$ is a polynomial, then $f(s, \rho(s)t)$ is a polynomial in $s \in [\alpha,\beta]$ for fixed $t$.
+
+Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
+\begin{align*}
+	f_e(x,y) &:= \half \Big(f(x, y) + f(x, -y)\Big), \quad \forall (x,y) \in \Omega\\
+	f_o(x,y) &:= \half \Big(f(x, y) - f(x, -y)\Big), \quad \forall (x,y) \in \Omega
+\end{align*}
+so that $y \mapsto f_e(x,y)$ for fixed $x$ is an even function, and $y \mapsto f_o(x,y)$ for fixed $x$ is an odd function.
+Note that if $f$ is a polynomial, then $f_e(s, \rho(s)t)$ is a polynomial in $s \in [\alpha,\beta]$ for fixed $t$.
+
+Now, we have that
+\begin{align*}
+	\iint_\Omega f_e(x,y) \: \Wab(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_\gamma^\delta f_e\big(s,\rho(s) t\big) \: V^{(a,b)}(s,t) \: \rho(s) \: \D t \: \D s \\
+	&= \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \Big( \int_\gamma^\delta f_e\big(s,\rho(s) t\big) \: \jacw^{(b)}(t) \: \D t \Big) \: \D s \\
+	&\approx \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \sum_{k=1}^N \Big( w_k^{(t)} f_e\big(s,\rho(s) t_k\big) \Big) \: \D s \quad (\star) \\
+	&\approx \sum_{k=1}^N \Bigg( w_k^{(s)} \: \sum_{l=1}^N \Big( w_l^{(t)} f_e\big(s_k,\rho(s_k) t_l\big) \Big) \Bigg) \quad (\star \star) \\
+	&= \sum_{j=1}^{N^2}  w_j \: f_e(x_j, y_j),
+\end{align*}
+where we achieve equality at $(\star)$ if $y \mapsto f_e(x,y)$ for fixed $x$ is a polynomial of degree at most $2N-1$ and we achieve equality at $(\star \star)$ if also $x \mapsto f_e(x,y)$ for fixed $y$ is a polynomial of degree at most $N$.
+
+Next, note that 
+\begin{align*}
+	\iint_\Omega f_o(x,y) \: \Wab(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_\gamma^\delta f_o\big(s,\rho(s) t\big) \: V^{(a,b)}(s,t) \: \rho(s) \: \D t \: \D s \\
+	&= \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \Big( \int_\gamma^\delta f_o\big(s,\rho(s) t\big) \: \jacw^{(b)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
+	&= 0
+\end{align*}
+since the inner integral at $(\dagger)$ over $t$ is zero, due to the symmetry over the domain (i.e. since $\gamma = -\delta$).
+
 Hence,
 \begin{align*}
-	\iint_\Omega f(x,y) \: \Wab(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_\gamma^\delta f\big(s,\rho(s) t\big) \: V^{(a,b)}(s,t) \: \rho(s) \: \D t \: \D s \\
-	&= \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \Big( \int_\gamma^\delta f\big(s,\rho(s) t\big) \: w_P^{(b)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
-	&\approx \int_\alpha^\beta  \genjacw^{(a,b+\half)}(s) \: \sum_{k=1}^N \Big( w_k^{(t)} f\big(s,\rho(s) t_k\big) \Big) \: \D s \quad (\star) \\
-	&\approx \sum_{k=1}^N \Bigg( w_k^{(s)} \: \sum_{l=1}^N \Big( w_l^{(t)} f\big(s_k,\rho(s_k) t_l\big) \Big) \Bigg) \quad (\star \star) \\
-	&= \sum_{j=1}^{N^2}  w_j \: f(x_j, y_j),
-\end{align*}
-where we achieve equality at $(\star)$ if $y \mapsto f(x,y)$ for fixed $x$ is a polynomial of degree at most $2N-1$ and we achieve equality at $(\star \star)$ if also $x \mapsto f(x,y)$ for fixed $y$ is a polynomial of degree at most $N$.
-
-Finally, note that the inner integral at $(\dagger)$ over $t$ would be zero if $y \mapsto f(x,y)$ for fixed $x$ was an odd function, due to the symmetry over the domain (i.e.
-since $\gamma = -\delta$).
+	\iint_\Omega f(x,y) \: \Wab(x,y) \: \D y \: \D x &= \iint_\Omega \Big(f_e(x,y) + f_o(x,y)\Big) \: \Wab(x,y) \: \D y \: \D x \\
+	&= \iint_\Omega f_e(x,y) \: \Wab(x,y) \: \D y \: \D x \\
+	&\approx \sum_{j=1}^{N^2}  w_j \: f_e(x_j, y_j) \quad (\star),
+\end{align*}
+where we achieve equality at $(\star)$ if $y \mapsto f(x,y)$ for fixed $x$ is a polynomial of degree at most $2N-1$ and also if $x \mapsto f(x,y)$ for fixed $y$ is a polynomial of degree at most $N$.
 \end{proof}
@@
 The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
-We can simply use quadrature to calculate the 1D inner products.
-This proves much cheaper computationally than using the 2D quadrature rule to calculate the 2D inner products.
-
-\sotodoinline{Give complexity}
+We can simply use quadrature to calculate the 1D inner products -- this has a complexity of $\bigO(N^3)$.
+This proves much cheaper computationally than using the 2D quadrature rule to calculate the 2D inner products, which has a complexity of $\bigO(N^4)$.
 
@@
 % FIGURES
-\begin{figure}
-	\begin{subfigure}[t]{0.3\textwidth}
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-poisson}
 	\centering
-	\includegraphics[scale=0.3]{solution-poisson}
-	\caption{}
-	\label{fig:solution-poisson}
+	%\label{fig:solution-poisson}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.3\textwidth}
+	\begin{subfigure}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.5]{solutionblocknorms-poisson}
+        	%\label{fig:solutionblocknorms-poisson}
+	\end{subfigure}
+	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
+This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
+	\centering
+	\label{fig:poisson}
+\end{figure}
+
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
 	\centering
 	\includegraphics[scale=0.3]{solution-helmholtz-k=100}
-	\caption{}
-	\label{fig:solution-helmholtz}
+	%\label{fig:solution-helmholtz}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.3\textwidth}
+	\begin{subfigure}{0.5\textwidth}
+	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-k=20}
+	\centering
+        	%\label{fig:solutionblocknorms-helmholtz}
+	\end{subfigure}
+	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
+Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
+	\centering
+	\label{fig:helmholtz}
+\end{figure}
+
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
 	\centering
 	\includegraphics[scale=0.3]{solution-biharmonic}
-	\caption{}
-	\label{fig:solution-biharmonic}
+	%\label{fig:solution-biharmonic}
 	\end{subfigure}
+	\begin{subfigure}{0.5\textwidth}
+	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic}
 	\centering
-    	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-(b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-(c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-\sotodoinline{Combine this figure with subsequent ones}}
-        \label{fig:solution}
+        	%\label{fig:solutionblocknorms-biharmonic}
+	\end{subfigure}
+	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
+	\centering
+	\label{fig:biharmonic}
 \end{figure}
-
-\begin{figure}[t]
-        	\includegraphics[scale=0.4]{solutionblocknorms-poisson}
-	\centering
-	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
-This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
-        	\label{fig:solutionblocknorms-poisson}
-\end{figure}
-
-\begin{figure}[t]
-        	\includegraphics[scale=0.4]{solutionblocknorms-helmholtz-k=20}
-	\centering
-	\caption{The norms of each block of the computed solution of the helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
-        	\label{fig:solutionblocknorms-helmoltz}
-\end{figure}
-
-\begin{figure}[t]
-        	\includegraphics[scale=0.4]{solutionblocknorms-biharmonic}
-	\centering
-	\caption{The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
-        	\label{fig:solutionblocknorms-biharmonic}
-\end{figure}
+% END FIGURES
 
@@
 which can be solved to find $\mathbf{u}$.
-In Figure \ref{fig:solution-poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the half-disk $\Omega$.
-In Figure \ref{fig:solutionblocknorms-poisson} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the half-disk $\Omega$.
+In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
-In Figure \ref{fig:solutionblocknorms-helmoltz} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
+In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-In Figure \ref{fig:solution-biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
-In Figure \ref{fig:solutionblocknorms-biharmonic} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
+In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
@@
 \end{align*}
-with\sotodo{Why $\subset$?} 
+with\sotodoinline{Why $\subset$?} \bstodoinline{I mean to say that $-1 \le \alpha < \beta \le 1$, otherwise this is the half disk case/full disk}
 \begin{align*}
@@
 \genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
-w^{(a)}_P(x) &:= (1-x^2)^a.
+\jacw^{(a)}(x) &:= (1-x^2)^a.
 \end{cases}
@@
 
-The weight $w^{(a)}_P(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{P_n^{(a, a)}\}$.
+The weight $\jacw^{(a)}(x)$ is a still an ultraspherical weight, and the corresponding OPs are the Jacobi polynomials $\{\jac_n^{(a, a)}\}$.
 $\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs denoted $\{\genjac_n^{(a,b,c)}\}\).
@@
 \begin{align*}
-	\hdopnk^{(a,b,c)}(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: P_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\hdopnk^{(a,b,c)}(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
 \end{align*}
@@
 \genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\half x)^{c} \\
-w^{(a,b)}_P(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
+\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
 \end{cases}
 \end{align*}
-The weight $w^{(a,b)}_P(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
-We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = P_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
+The weight $\jacw^{(a,b)}(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
+We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = \jac_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
 $\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\).

### 2019-05-29 papers/halfdisk/halfdisk.tex
draft

@@
 \newcommand{\genjacw}{w_\genjac}
-\newcommand{\jacw}{w_\jac}
+\newcommand{\jacw}{w_P}
 \newcommand{\normgenjac}{\omega_\genjac}
-\newcommand{\normjac}{\omega_\jac}
+\newcommand{\normjac}{\omega_P}
 
@@
 
-\title{Sparse spectral and $p$-finite element methods for partial differential equations on disk-slices and trapeziums}
+\title{Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums}
 \author{Sheehan Olver, Ben Snowball}
@@
 In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g.
-a half disk) and trapeziums.
+a half-disk) and trapeziums.
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
@@
 
-An alternative approach is to use \sotodoinline{Compare with "Discontinuous Galerkin scheme for the spherical shallow water equations with applications to tsunami modeling and prediction"}
+%An alternative approach is to use \sotodoinline{Compare with "Discontinuous Galerkin scheme for the spherical shallow water equations with applications to tsunami modeling and prediction"}
 
 The motivation for this work is solving partial differential equations on sub-domains of the sphere.
-In particular, OPs in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere  can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D.
+In particular, OPs in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere  can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere.
+The resulting sparsity in high-polynomial degree discretisations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater}).
+%
 %, where we would require $2n+1$ 3D OPs for each degree $n$, of which $n+1$ would be the $n+1$ degree $n$ OPs on the half-disk of the first family (polynomials in $x,y$) while the remaining $n$ OPs would be given by $z$ multiplied by the $n$ degree $n-1$ OPs on the half-disk of the second family.
@@
 
-\noindent \secref{OPs}: We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.
+\noindent \secref{OPs}: We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half-disk domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.
 
@@
 \noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice.
-We  derive a quadrature rule on the half disk that can be used to expand a function in the OP basis up to a given order.
+We  derive a quadrature rule on the half-disk that can be used to expand a function in the OP basis up to a given order.
 We implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
@@
 
-\section{Orthogonal polynomials on the half disk}\label{Section:OPs}
+\section{Orthogonal polynomials on the half-disk}\label{Section:OPs}
 
@@
 
-For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definition.
-\begin{definition}\label{def:construction}
-Let
-\begin{align*}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
-with $\rho : (\alpha, \beta) \to \R$.
-Further, let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
+For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definition. First we introduce notation for two families of univariate OPs
+\begin{definition}\label{def:OPconstruction}
+Let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
 \begin{align*}
@@
 \end{align}
-Further, denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
-
-Assume that either of the following hold:
+Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
+\end{definition}
+
+\begin{definition}\label{def:constuction}
+Define the four-parameter 2D orthogonal polynomials via:
+\begin{align*}
+	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+\end{align*}
+\end{definition}
+
+ $\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
+\begin{align*}
+	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
+\end{align*}
+assuming that either of the following hold:
 \begin{enumerate}
@@
 \end{enumerate}
-We define the four-parameter 2D orthogonal polynomials on $\Omega$ by
-\begin{align*}
-	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align*}
-which are orthogonal with respect to the weight
-\begin{align*}
-	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
-\end{align*}
-that is,
+That is,
 \begin{align*}
@@
 \end{align*}
-\end{definition}
-
-For the half disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
+
+For the half-disk, the weight $W^{(a,b)}(x,y) = x^a \: (1-x^2-y^2)^b$ results from setting:
 \begin{align*}
@@
 \end{align*}
-Note here we have made the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and removed it for the specific half-disk case.
-This is for ease of presentiation, by ensuring we reduce the need to incorporate unnecessary parameters and instead yield a two-parameter family of 2D orthogonal polynomials.
-
-The weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\jac_n^{(b, b)}\}$.
-The weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}) and we introduce the notation $\{\genjac_n^{(a, b)}\}$ for the orthogonal polynomials.
-The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{\jac_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
-\begin{align*}
-	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
-	\ip<p, q>_{\jacw^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: \jacw^{(b)}(y)\: \D y,
-\end{align*}
-i.e.
+Note here we make the adjustment that $\genjacw^{(a,b,c)}(x) = (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{2c}$ and simply set the first parameter to zero and remove it:
+\begin{align}\label{eq:diskpolys}
+	\hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
+\end{align}
+where $\genjac_n^{(a, b)}  := \genjac_n^{(0,a, b)}$, and we use $W^{(a,b)}(x,y) := W^{(0,a,b,b)}$.
+In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
+%
+%The two 1D families of orthogonal polynomials we have denoted $\{\genjac_{n}^{(a,b)}\}$ and $\{\jac_n^{(b,b)}\}$ are orthonormal, with respect to the inner products respectively given by:
+%\begin{align*}
+%	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_0^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \\
+%	\ip<p, q>_{\jacw^{(b)}} &:= \frac{1}{\normjac^{(b)}} \: \int_{-1}^1 p(y) \: q(y) \: \jacw^{(b)}(y)\: \D y,
+%\end{align*}
+%i.e.
 $\norm{\genjac_{n}^{(a,b)}}_{\genjacw^{(a,b)}} = 1$ $\forall n = 0,1,2,\dots$ and $\norm{\jac_{n}^{(b,b)}}_{\jacw^{(b)}} = 1$ $\forall n = 0,1,2,\dots$.
 For clarity, we use the notation that $\norm{p}_{\genjacw^{(a,b)}}^2 := \ip<p, p>_{\genjacw^{(a,b)}}$ and that $\norm{p}_{\jacw^{(b)}}^2 := \ip<p, p>_{\jacw^{(b)}}$.
-
-Thus we arrive at a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
-\begin{align}\label{eq:diskpolys}
-	\hdopnk^{(a,b)}(x,y) := \genjacnmk^{(a, b+k+\half)}(x) \: \rho(x)^k \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align}
-orthogonal on $\Omega$ with respect to the weight
-\begin{align*}
-	W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: \jacw^{(b)}\fpr(\frac{y}{\rho(x)}) =  \frac{x^a \: (1-x^2-y^2)^b}{\normgenjac^{(a,b)} \normjac^{(b)}}.
-\end{align*}
-That is, they are orthogonal (for given parameters $a,b$) with respect to the inner product given by
-\begin{align*}
-	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \hdopnkab(x,y) \: \hdopmjab(x,y) \: \Wab(x,y) \: \D y \: \D x.
-\end{align*}
-We can see that they are indeed orthogonal using the change of variable $t = \frac{y}{\rho(x)}$:
+%
+%Thus we arrive at a two parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ given by, for $0 \le k \le n, \: n = 0,1,2,\dots,$
+
+%orthogonal on $\Omega$ with respect to the weight
+%\begin{align*}
+%	W^{(a,b)}(x,y) = \genjacw^{(a,b)}(x) \: \jacw^{(b)}\fpr(\frac{y}{\rho(x)}) =  \frac{x^a \: (1-x^2-y^2)^b}{\normgenjac^{(a,b)} \normjac^{(b)}}.
+%\end{align*}
+%That is, they are orthogonal (for given parameters $a,b$) with respect to the inner product given by
+%\begin{align*}
+%	\ip< \hdopnkab, \: \hdopmjab >_\Wab &=  \iint_\Omega \: \hdopnkab(x,y) \: \hdopmjab(x,y) \: \Wab(x,y) \: \D y \: \D x.
+%\end{align*}
+We can see that they are indeed orthogonal using the change of variable $t = \frac{y}{\rho(x)}$, for the following normalisation:
 \begin{align}
@@
 \end{align}
-Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.
+%Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.
 
@@
 The 3-term recurrence for multiplication by $x$ follows from equation (\ref{eqn:Hrecurrence}).
-For the recurrence for multiplication by $y$, since $\{\hdopmjab\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkab(x,) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjab(x,y)$.
+For the recurrence for multiplication by $y$, since $\{\hdopmjab\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkab(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjab(x,y)$.
 These coefficients are given by
@@
 
-$J_x^{(a,b)}, J_y^{(a,b)}$ have banded-block-banded structures.
-A block-banded matrix is a block  matrix where only a finite number of block-bands are nonzero.
-A banded-block-banded matrix is one where the blocks themselves are banded matrices.
-For example, $J_x^{(a,b)}, J_y^{(a,b)}$ are block-tridiagonal:
+$J_x^{(a,b)}, J_y^{(a,b)}$ are banded-block-banded matrices:
+
+\begin{definition}
+A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
+A matrix where the block-bandwidths and sub-blockbandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
+\end{definition}
+
+For example, $J_x^{(a,b)}, J_y^{(a,b)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
 \begin{align*}
@@
 \end{align*}
-where the blocks themselves are diagonal for $J_x^{(a,b)}$,
+where the blocks themselves are diagonal for $J_x^{(a,b)}$ (sub-block-bandwidths $(0,0)$),
 \begin{align*}
@@
 \end{align*}
-and tridiagonal for $J_u^{(a,b)}$,
+and tridiagonal for $J_y^{(a,b)}$ (sub-block-bandwidths $(1,1)$),
 \begin{align*}
@@
 Recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopab(x,y)^\top \mathbf{f}$.
-Recall also that $\Wab(x,y) := \genjacw^{(a,b)}(x) \: \jacw^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $\genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $\jacw^{(b)}(y) := (1-y^2)^b$ for the half disk case.
+Recall also that $\Wab(x,y) := \genjacw^{(a,b)}(x) \: \jacw^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $\genjacw^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $\jacw^{(b)}(y) := (1-y^2)^b$ for the half-disk case.
 
@@
 \begin{itemize}
-	\item $D_x^{(a,b)}$ has a block-bandwidth of $(-1,2)$, with the blocks themselves having a bandwidth of $(0, 2)$.
-  	\item $D_y^{(a,b)}$ has a block-bandwidth of $(-1,1)$, with the blocks themselves having a bandwidth of $(-1,1)$.
-	\item $W_x^{(a,b)}$ has a block-bandwidth of $(2,-1)$, with the blocks themselves having a bandwidth of $(2, 0)$.
-  	\item $W_y^{(a,b)}$ has a block-bandwidth of $(1,-1)$, with the blocks themselves having a bandwidth of $(1,-1)$.
+	\item $D_x^{(a,b)}$ has  block-bandwidths $(-1,2)$, and sub-block-bandwidths $(0, 2)$.
+  	\item $D_y^{(a,b)}$ has  block-bandwidths $(-1,1)$, and sub-block-bandwidths $(-1,1)$.
+	\item $W_x^{(a,b)}$ has  block-bandwidths $(2,-1)$, and sub-block-bandwidths $(2, 0)$.
+  	\item $W_y^{(a,b)}$ has  block-bandwidths $(1,-1)$, and sub-block-bandwidths $(1,-1)$.
 \end{itemize}
@@
 
-There exist ``parameter transformation'' matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+There exist conversion matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 
@@
 The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
-We can simply use quadrature to calculate the 1D inner products -- this has a complexity of $\bigO(N^3)$.
+We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$.
 This proves much cheaper computationally than using the 2D quadrature rule to calculate the 2D inner products, which has a complexity of $\bigO(N^4)$.
@@
 %
-\section{Examples on the half disk with zero BCs}\label{Section:Examples}
+\section{Examples on the half-disk with zero BCs}\label{Section:Examples}
 
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half-disk ($x=0, \: y = \pm 1$) we see faster convergence.
 For the final Gaussian bump example, we see that we achieve spectral convergence.
@@
 In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half-disk ($x=0, \: y = \pm 1$) we see faster convergence.
 For the final Gaussian bump example, we see that we achieve spectral convergence.
@@
 We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains given by an algebraic curve---notably here the half-disk---with Dirichlet boundary conditions.
-This forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitable shaped elements, for example, by dividing the disk int disk-slice elements.
-
+This forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitable shaped elements, for example, by dividing the disk into disk slice elements.
 %A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
 With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
-
-This work serves as a stepping stone to constructing similar methods on other 3D spherical domains, such as spherical caps and spherical triangles, which will have applications in numerical weather prediction.
+This work serves as a stepping stone to constructing similar methods on other 3D spherical domains, such as spherical caps and spherical triangles.
 
@@
 
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half disk/disk-slice/trapezium domain.
+In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a half-disk/disk slice/trapezium domain.
 We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.
@@
 
-The work in this paper on the half disk can be easily transferred to the domain of a disk slice by which we mean
+The work in this paper on the half-disk can be easily transferred to the domain of a disk slice by which we mean
 \begin{align*}
@@
 \end{align*}
-with\sotodoinline{Why $\subset$?} \bstodoinline{I mean to say that $-1 \le \alpha < \beta \le 1$, otherwise this is the half disk case/full disk}
+with
 \begin{align*}
@@
 
-The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such disk-slice domains.
-For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for a disk-slice.
+The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such disk slice domains.
+For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for a disk slice.
 

### 2019-06-04 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
New struct and framework to 2D general disk slice as given in paper. Needs finishing. (To take over from HalfDiskSpace etc.). To extend to spherical caps framework.

@@ -1,0 +1,975 @@
+# module DiskSliceFamilies
+
+# using ApproxFun
+#     import ApproxFun: evaluate, PolynomialSpace, recα, recβ, recγ, recA, recB, recC, domain,
+#                         domainspace, rangespace, bandwidths, prectype, canonicaldomain, tocanonical,
+#                         spacescompatible, points, transform, itransform, AbstractProductSpace, tensorizer,
+#                         columnspace, checkpoints, plan_transform, clenshaw
+#     import Base: in, *
+# using OrthogonalPolynomialFamilies
+# using StaticArrays
+# using FastGaussQuadrature
+# using LinearAlgebra
+# using SparseArrays
+# using BlockBandedMatrices
+# using BlockArrays
+# # using SingularIntegralEquations
+# using Test
+
+export DiskSliceFamily, DiskSliceSpace
+
+# R should be Float64, B should be BigFloat
+abstract type DiskFamily{B,R,N} end
+struct DiskSlice{B,T} <: Domain{SVector{2,T}} end
+DiskSlice() = DiskSlice{BigFloat, Float64}()
+checkpoints(::DiskSlice) = [SVector(0.1,0.23), SVector(0.3,0.12)]
+
+struct DiskSliceSpace{DF, B, T, N} <: Space{DiskSlice{B,T}, T}
+    family::DF # Pointer back to the family
+    params::NTuple{N,B} # Parameters
+    opnorms::Vector{T} # squared norms
+    opptseval::Vector{Vector{T}} # Store the ops evaluated at the transform pts
+    xderivopptseval::Vector{Vector{T}} # Store the x deriv of the ops evaluated
+                                    # at the transform pts
+    yderivopptseval::Vector{Vector{T}} # Store the y deriv of the ops evaluated
+                                    # at the transform pts
+    A::Vector{SparseMatrixCSC{T}}
+    B::Vector{SparseMatrixCSC{T}}
+    C::Vector{SparseMatrixCSC{T}}
+    DT::Vector{SparseMatrixCSC{T}}
+end
+
[…]

### 2019-06-04 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl

@@ -36,7 +36,6 @@ struct OrthogonalPolynomialSpace{FA,WW,F,D,B,R,N} <: PolynomialSpace{D,R}
     derivopptseval::Vector{Vector{R}}
 end
 
-
 # Finds the OPs and recurrence for weight w, having already found N₀ OPs
 function lanczos!(w, P, β, γ; N₀=0)
 
@@ -756,7 +755,7 @@ function resizedata!(S::HalfDiskSpace, N)
     S
 end
 
-function jacobix(S, N)
+function jacobix(S::HalfDiskSpace, N)
     resizedata!(S, N)
     rows = cols = 1:N+1
     l, u = 1, 1
@@ -774,7 +773,7 @@ function jacobix(S, N)
     J
 end
 
-function jacobiy(S, N)
+function jacobiy(S::HalfDiskSpace, N)
     # Transposed operator, so acts directly on coeffs vec
     resizedata!(S, N)
     rows = cols = 1:N+1
@@ -794,7 +793,7 @@ function jacobiy(S, N)
     J
 end
 
-function clenshawG(n, z)
+function clenshawG(::HalfDiskSpace, n, z)
     sp = sparse(I, n+1, n+1)
     [z[1] * sp; z[2] * sp]
 end
@@ -815,11 +814,11 @@ function clenshaw(cfs::AbstractVector, S::HalfDiskSpace, z)
     inds2 = m-N:m
     inds1 = (m-2N):(m-N-1)
     γ2 = view(cfs, inds2)'
-    γ1 = view(cfs, inds1)' - γ2 * S.DT[N] * (S.B[N] - clenshawG(N-1, z))
+    γ1 = view(cfs, inds1)' - γ2 * S.DT[N] * (S.B[N] - clenshawG(S, N-1, z))
     for n = N-2:-1:0
         ind = sum(1:n)
         γ = (view(cfs, ind+1:ind+n+1)'
-             - γ1 * S.DT[n+1] * (S.B[n+1] - clenshawG(n, z))
+             - γ1 * S.DT[n+1] * (S.B[n+1] - clenshawG(S, n, z))
              - γ2 * S.DT[n+2] * S.C[n+2])
         γ2 = copy(γ1)
         γ1 = copy(γ)
@@ -1501,7 +1500,7 @@ function opevalatpts(S::HalfDiskSpace{<:Any, <:Any, T}, j, pts) where T
         nm1 = getopindex(n-1, 0)
         for r = 1:length(pts)
[…]

### 2019-06-04 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -1,6 +1,6 @@
 using ApproxFun, OrthogonalPolynomialFamilies, FastGaussQuadrature,
         SingularIntegralEquations, Test, SparseArrays
-import OrthogonalPolynomialFamilies: golubwelsch, lanczos, halfdiskquadrule,
+import OrthogonalPolynomialFamilies: golubwelsch, halfdiskquadrule,
         gethalfdiskOP, jacobix, jacobiy, differentiatex, differentiatey,
         resizecoeffs!, laplaceoperator, transformparamsoperator, operatorclenshaw, getnk,
         getopindex, weight, resizedata!
@@ -266,52 +266,78 @@ end
 end
 
 #=====#
+# DiskSlice
 
-# using Profile
-# D = HalfDiskFamily()
-# S = H = D.H(BigFloat(1),BigFloat(2.5));
-# @time resizedata!(S,300);
-# @time resizedata!(S,1000);
-# Profile.print()
+@testset "Evaluation for DiskSliceSpace (transform())" begin
+    n = 10; a = 1.0; b = 2.0; c = 1.0
+    f = (x,y) -> y*x^2 + x
+    α, β = 0.4, 0.8
+    D = DiskSliceFamily(α, β)
+    S = D(a, b, c)
+    pts = points(S, n)
+    vals = [f(pt...) for pt in pts]
+    cfs = transform(S, vals)
+    x, y = 0.5673, -0.2786; z = [x; y]
+    @test (x^2 + y^2 < 1 && D.α ≤ z[1] ≤ D.β && D.γ*D.ρ(z[1]) ≤ z[2] ≤ D.δ*D.ρ(z[1]))
+    F = Fun(S, cfs)
+    @test F(z) ≈ f(z...)
+    @test itransform(S, cfs) ≈ vals
+    F = Fun(f, S, 10)
+    @test F(z) ≈ f(z...)
+    F = Fun(f, S)
+    @test F(z) ≈ f(z...)
+end
 
+@testset "Jacobi matrices" begin
+    n = 10; a, b, c = 1.0, 1.0, 2.0
+    α, β = 0.2, 0.9
+    D = DiskSliceFamily(α, β)
+    S = D(a, b, c)
+    x, y = α + 0.1673, -0.2786; z = [x; y]
[…]

### 2019-06-08 papers/halfdisk/halfdisk.tex
Fixed a few typos
readthrough

@@
 \usepackage{todonotes}
+
+%\bibliographystyle{apalike}
 
@@
 where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.
-The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:construction}.
+The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
 
 Sparsity comes from expanding the domain and range of an operator  using different choices of $a$ and $b$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
+Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials we are able to construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 In particular, OPs in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere  can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere.
-The resulting sparsity in high-polynomial degree discretisations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater}).
+The resulting sparsity in high-polynomial degree discretisations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
 %
@@
 
-For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definition. First we introduce notation for two families of univariate OPs
+For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definition. First we introduce notation for two families of univariate OPs.
 \begin{definition}\label{def:OPconstruction}
@@
 	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Hrecurrence} \\
-	y \jac_n^{(b,b)}(y) &= \delta_n^{(b)} \jac_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} \jac_n^{(b,b)}(y) + \delta_{n-1}^{(b)} \jac_{n-1}^{(b,b)}(y),
+	y \jac_n^{(b,b)}(y) &= \delta_n^{(b)} \jac_{n+1}^{(b,b)}(y) + \delta_{n-1}^{(b)} \jac_{n-1}^{(b,b)}(y).
 \end{align}
-where we note that \(\gamma_n^{(b)} = 0\), for \( n = 0,1,2,\dots\).
-We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$.
-Importantly, we can deduce sparsity in the recurrence relationships:
+ We can use these to determine the 2D recurrences for $\hdopnkab(x,y)$. Importantly, we can deduce sparsity in the recurrence relationships:
 
@@
 In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkab\}$ basis.
-
-$J_x^{(a,b)}, J_y^{(a,b)}$ are banded-block-banded matrices:
+Further, note that $J_x^{(a,b)}, J_y^{(a,b)}$ are banded-block-banded matrices:
 
@@
 A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
-A matrix where the block-bandwidths and sub-blockbandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
+A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
 \end{definition}
@@
 We proceed with the case for the operator $D_y^{(a,b)}$ for partial differentiation by $y$.
-Since $\{\hdop^{(a, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: \hdop^{(a, b+1)}_{m,j}$.
+Since $\{\hdop^{(a, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^y \: \hdop^{(a, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 \begin{align*}
-	&\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} \\
-	&\quad \quad = \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \\
+	\ip<\pddy \hdopnkab, \hdop^{(a, b+1)}_{m,j} >_{W^{(a, b+1)}} &= \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \rho(x)^{k+j} \: \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1)}} \\
 	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{j}^{(b+1,b+1)} >_{\jacw^{(b+1)}} \\
-	&\quad \quad = \normgenjac^{(a, b+1+\half(k+j))} \: \normgenjac^{(a, b+1)} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \\
+	&= \normgenjac^{(a, b+1+\half(k+j))} \: \ip< \genjacnmk^{(a,b+k+\half)}, \genjacmmj^{(a,b+1+j+\half)} >_{\genjacw^{(a,b+1+\half(k+j))}} \\
 		&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{j}^{(b+1,b+1)} >_{\jacw^{(b+1)}}.
 \end{align*}
-Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(b)}$ is a polynomial degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
+Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(b)}$ is a polynomial of degree $2b$ and vanishes at the limits of the integral for positive parameter $b$, we have that
 \begin{align*}
@@
 \begin{align*}
-	\ip< \genjacnmk^{(a,b+k+\half)}, \genjac_{m+1-k}^{(a,b+k+\half)} >_{\genjacw^{(a,b+k+\half)}} = \delta_{n,m+1}
-\end{align*}
-showing the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally, we simply have that 
-\begin{align*}
-	c_{n-1,k-1}^y &= \normgenjac^{(a, b+1)} \: \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{k-1}^{(b+1,b+1)} >_{\jacw^{(b+1)}}.
+	\ip< \genjacnmk^{(a,b+k+\half)}, \genjac_{m+1-k}^{(a,b+k+\half)} >_{\genjacw^{(a,b+k+\half)}} = \delta_{n,m+1},
+\end{align*}
+showing that the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally,
+\begin{align*}
+	c_{n-1,k-1}^y &= \ip< \jac_k^{(b,b) \: \prime}, \: \jac_{k-1}^{(b+1,b+1)} >_{\jacw^{(b+1)}}.
 \end{align*}
@@
 We next consider the case for the operator $D_x^{(a,b)}$ for partial differentiation by $x$.
-Since $\{\hdop^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+1, b+1)}_{m,j}$.
+Since $\{\hdop^{(a+1, b+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkab = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^x \: \hdop^{(a+1, b+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 \begin{align}
-	\rho'(x) = -x \: \rho(x)^{-2}. \label{eqn:rhoderivative}
+	\rho'(x) = -x \: \rho(x)^{-1}. \label{eqn:rhoderivative}
 \end{align}
@@
 
-We can see that, from each term's second factor as written, the above is zero for $j < k-2$.
+We can see  from each term's second factor that the above is zero for $j < k-2$.
 Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b)}$ is a polynomial degree $a+2b$ and vanishes at the limits of the integral for positive parameters $a,b$, we have that
@@
 \begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
-The operator matrices $T^{(a,b)\to(a+1,b)}$, $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$, $T_W^{(a,b)\to(a-1,b)}$, $T_W^{(a,b)\to(a,b-1)}$ and $T_W^{(a,b)\to(a-1,b-1)}$ from Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure.
-More specifically:
+The operator matrices in Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
 \begin{itemize}
-	\item $T^{(a,b)\to(a+1,b)}$ has a block-bandwidth of $(0,1)$, with the blocks themselves being diagonal.
-  	\item $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$ have a block-bandwidth of $(0,3)$, with the blocks themselves having a bandwidth of $(0,2)$.
-	\item $T_W^{(a,b)\to(a-1,b)}$ has a block-bandwidth of $(1,0)$, with the blocks themselves being diagonal.
-  	\item $T_W^{(a,b)\to(a,b-1)}$, $T_W^{(a,b)\to(a-1,b-1)}$ have a block-bandwidth of $(3,0)$, with the blocks themselves having a bandwidth of $(2,0)$.
+	\item $T^{(a,b)\to(a+1,b)}$ has block-bandwidth $(0,1)$, with diagonal blocks.
+  	\item $T^{(a,b)\to(a,b+1)}$, $T^{(a,b)\to(a+1,b+1)}$ have block-bandwidth $(0,3)$ and sub-block-bandwidth $(0,2)$.
+	\item $T_W^{(a,b)\to(a-1,b)}$ has block-bandwidth  $(1,0)$ with diagonal blocks.
+  	\item $T_W^{(a,b)\to(a,b-1)}$, $T_W^{(a,b)\to(a-1,b-1)}$ have block-bandwidth $(3,0)$ and sub-block-bandwidth $(2,0)$.
 \end{itemize}
@@
 
+In this section we discuss how to take advantage of the proposed basis and sparsity structure in partial differential operators in practical computational applications.
+
 \subsection{Constructing $\genjac_n^{(a,b)}(x)$}
@@
 %
-\section{Examples on the half-disk with zero BCs}\label{Section:Examples}
+\section{Examples on the half-disk with zero Dirichlet conditions}\label{Section:Examples}
+
+We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions.
+We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, demonstrating the versatility of the approach.
 
@@
 
-Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+The Poisson equation is the classic problem of finding \(u(x,y)\) given a function \(f(x,y)\) such that:
 \begin{align}
@@
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half-disk ($x=0, \: y = \pm 1$) we see faster convergence.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our half-disk ($x=0, \: y = \pm 1$) we observe faster convergence.
 For the final Gaussian bump example, we see that we achieve spectral convergence.
 
-\subsection{Inhomogenous variable-coefficient Helmholtz}
+\subsection{Inhomogeneous variable-coefficient Helmholtz}
 
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
+In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
     		\Delta \tilde{u}(x,y) + k^2 \: v(x,y) \; \tilde{u}(x,y) = f(x,y) - c \: k^2 \: v(x,y) \; =: g(x,y)  \quad \text{in } \Omega \\
-		\tilde{u}(x,y) = 0 \quad \text{on } \partial \Omega
+		\tilde{u}(x,y) = 0 \quad \text{on } \partial \Omega.
 	\end{cases}.
@@
 
-We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains given by an algebraic curve---notably here the half-disk---with Dirichlet boundary conditions.
+We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the half-disk---with Dirichlet boundary conditions.
 This forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitable shaped elements, for example, by dividing the disk into disk slice elements.
@@
 					\bighdopoo^\top \Wii_x \mathbf{v} \\
-					\bighdopoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
+					\bighdopoo^\top T_W^{(1,0)\to(0,0)} \Wii_y \mathbf{v}
 				\end{pmatrix}^\top 
@@
 					\bighdopoo^\top \Wii_x \mathbf{u} \\
-					\bighdopoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
+					\bighdopoo^\top T_W^{(1,0)\to(0,0)} \Wii_y \mathbf{u}
 				\end{pmatrix}
@@
 	&= \int_\element \Big( \mathbf{v}^\top {\Wii_x}^\top \bighdopoo \bighdopoo^\top \Wii_x \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^\top ({T^{(1,0)\to(0,0)} \Wii_y})^\top \bighdopoo \bighdopoo^\top T^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
-	&= \mathbf{v}^\top \: \Big( {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
+					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1,0)\to(0,0)} \Wii_y})^\top \bighdopoo \bighdopoo^\top T_W^{(1,0)\to(0,0)} \Wii_y \mathbf{u}  \Big) \: d\mathbf{x} \\
+	&= \mathbf{v}^\top \: \Big( {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T_W^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T_W^{(1,0)\to(0,0)} \Wii_y \Big) \: \mathbf{u}
 \end{align*}
@@
 \begin{align*}
-A = {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T^{(1,0)\to(0,0)} \Wii_y, 
+A = {\Wii_x}^\top \Lambda^{(0,0)} \Wii_x + ({T_W^{(1,0)\to(0,0)} \Wii_y})^\top \Lambda^{(0,0)} T_W^{(1,0)\to(0,0)} \Wii_y, 
 \end{align*}

### 2019-06-11 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
Fixed operator clenshaw. Edits to tests.

@@ -744,6 +744,7 @@ function resizedata!(S::HalfDiskSpace, N)
     resize!(S.B, N + 2)
     resize!(S.C, N + 2)
     resize!(S.DT, N + 2)
+    @show "resizedata!"
     getBs!(S, N, N₀)
     @show "resizedata!", "done getBs!"
     getCs!(S, N, N₀)
@@ -762,6 +763,9 @@ function jacobix(S::HalfDiskSpace, N)
     λ, μ = 0, 0
     J = BandedBlockBandedMatrix(0.0I, (rows, cols), (l, u), (λ, μ))
     J[1, 1] = S.B[1][1, 1]
+    if N == 0
+        return J
+    end
     view(J, Block(1, 2)) .= S.A[1][1, :]'
     for n = 2:N
         view(J, Block(n, n-1)) .= S.C[n][1:Int(end/2), :]
@@ -780,6 +784,9 @@ function jacobiy(S::HalfDiskSpace, N)
     l, u = 1, 1
     λ, μ = 1, 1
     J = BandedBlockBandedMatrix(0.0I, (rows, cols), (l, u), (λ, μ))
+    if N == 0
+        return J
+    end
     n = 1
     J[1, 1] = S.B[1][2, 1]
     view(J, Block(n, n+1)) .= S.C[n+1][Int(end/2)+1:end, :]'
@@ -1414,47 +1421,45 @@ function operatorclenshaw(cfs, S::HalfDiskSpace, N)
     # function f given by its coefficients of its expansion in the space S
     m̃ = length(cfs)
     M = getnk(m̃)[1] # Degree of f
+
+    # Pad cfs to correct size
     m = getopindex(M, M)
-    if m̃ < getopindex(M, M)
-        # Pad cfs to correct size
+    if N < M
+        error("Size requested has lower degree than function for the operator")
+    end
+    if m̃ < m
         resize!(cfs, m)
         cfs[m̃+1:end] .= 0.0
     end
+
     resizedata!(S, M)
-    Jx = sparse(jacobix(S, M))
-    Jy = sparse(jacobiy(S, M))
+    Jx = sparse(jacobix(S, N))
+    Jy = sparse(jacobiy(S, N))
+    id = sparse(I, size(Jx))
 
     @show "Operator Clenshaw"
     P0 = 1.0
     if M == 0
-        return cfs[1] * id * P0
-    end
-    id = sparse(I, size(Jx))
[…]

### 2019-06-11 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -149,10 +149,22 @@ end
     a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
     x, y = 0.4, -0.2; z = [x; y] # Test point
     N = 14
-    oper = Fun((x,y)->cos(x)*y^2, S, 200)
-    A = operatorclenshaw(oper, S)
+    oper = Fun((x,y)->cos(x)*y^2, D(a-1, b-1), 200)
     f = Fun((x,y)->x*y, S)
     resizecoeffs!(f, N)
+    A = operatorclenshaw(oper, D(a-1, b-1), N+3)
+    T = transformparamsoperator(D(a-1, b-1), S, N+3)
+    Tw = transformparamsoperator(S, D(a-1, b-1), N, weighted=true)
+    foper = Fun(S, T*A*Tw*f.coefficients)
+    @test foper(z) ≈ OrthogonalPolynomialFamilies.weight(S, z) * f(z) * oper(z)
+
+    a, b = 1.0, 1.0; D = HalfDiskFamily(); S = D(a, b)
+    x, y = 0.4, -0.2; z = [x; y] # Test point
+    N = 14
+    oper = Fun((x,y)->cos(x)*y^2, S, 200)
+    f = Fun((x,y)->x*y, S)
+    resizecoeffs!(f, N)
+    A = operatorclenshaw(oper, S, N)
     foper = Fun(S, A*f.coefficients)
     @test foper(z) ≈ f(z) * oper(z)
 end

### 2019-06-11 papers/halfdisk/halfdisk.tex
New figures

@@
 	\includegraphics[scale=0.35]{sparsityoflaplacian-w11}
+        \centering
 	\end{subfigure}
@@
 	\includegraphics[scale=0.35]{sparsityofhelmholtz}
+        \centering
 	\end{subfigure}
@@
 	\includegraphics[scale=0.35]{sparsityofbiharmonic}
+        \centering
 	\end{subfigure}
@@
 	\centering
-	\includegraphics[scale=0.3]{solution-helmholtz-k=100}
+	\includegraphics[scale=0.3]{solution-helmholtz-k=100-n=1500}
 	%\label{fig:solution-helmholtz}
@@
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
+	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
 Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$.
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = x(1-x^2-y^2)e^x$.
 In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns.

### 2019-06-18 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
DiskSLiceFamily code (all but biharmonic) done. tests added. Can replace HalfDisk code.

@@ -118,7 +118,7 @@ getPspace(S::DiskSliceSpace) = (S.family.P)(S.params[end], S.params[end])
 #===#
 # Weight eval functions
 weight(S::DiskSliceSpace{<:Any,<:Any,T,<:Any}, x, y) where T =
-    T(getRspace(S).weight(x) * getPspace(S).weight(y/ρ(x)))
+    T(getRspace(S).weight(x) * getPspace(S).weight(y/S.family.ρ(x)))
 weight(S::DiskSliceSpace, z) = weight(S, z[1], z[2])
 
 
@@ -583,47 +583,45 @@ function operatorclenshaw(cfs, S::DiskSliceSpace, N)
     # function f given by its coefficients of its expansion in the space S
     m̃ = length(cfs)
     M = getnk(m̃)[1] # Degree of f
+
+    # Pad cfs to correct size
     m = getopindex(M, M)
-    if m̃ < getopindex(M, M)
-        # Pad cfs to correct size
+    if N < M
+        error("Size requested has lower degree than function for the operator")
+    end
+    if m̃ < m
         resize!(cfs, m)
         cfs[m̃+1:end] .= 0.0
     end
+
     resizedata!(S, M)
-    Jx = sparse(jacobix(S, M))
-    Jy = sparse(jacobiy(S, M))
+    Jx = sparse(jacobix(S, N))
+    Jy = sparse(jacobiy(S, N))
+    id = sparse(I, size(Jx))
 
     @show "Operator Clenshaw"
     P0 = 1.0
     if M == 0
-        return cfs[1] * id * P0
-    end
-    id = sparse(I, size(Jx))
-    if M == 1
-        P0 * (operatorclenshawvector(S, cfs[1], id)[1] - (operatorclenshawmatrixDT(S, S.DT[1], id) * operatorclenshawmatrixBmG(S, S.B[1], id, Jx, Jy))[1])
-    end
-    n = M; @show "Operator Clenshaw", N, n
-    inds2 = m-M:m
-    γ2 = operatorclenshawvector(S, view(cfs, inds2), id)
-    n = M - 1; @show "Operator Clenshaw", M, n
-    inds1 = (m-2M):(m-M-1)
-    γ1 = (operatorclenshawvector(S, view(cfs, inds1), id)
[…]

### 2019-06-18 OrthogonalPolynomialFamilies/test/runtests.jl

@@ -243,7 +243,7 @@ end
     d2xweight(S::HalfDiskSpace, z) = d2xweight(S, z[1], z[2])
     function d2yweight(S::HalfDiskSpace{<:Any, <:Any, T}, x, y) where T
         D = S.family
-        - 2 * T(S.b) * (weight(D(S.a,S.b-1), x, y) - 2 * T(S.b) * (T(S.b) - 1) * y^2 * weight(D(S.a,S.b-2), x, y))
+        - 2 * T(S.b) * (weight(D(S.a,S.b-1), x, y) + 4 * T(S.b) * (T(S.b) - 1) * y^2 * weight(D(S.a,S.b-2), x, y))
     end
     d2yweight(S::HalfDiskSpace, z) = d2yweight(S, z[1], z[2])
 
@@ -298,6 +298,24 @@ end
     @test F(z) ≈ f(z...)
     F = Fun(f, S)
     @test F(z) ≈ f(z...)
+
+    n = 10; a = 1.0; b = 2.0; c = 1.0
+    f = (x,y) -> y*x^2 + x
+    α, β = 0.4, 0.8
+    D = DiskSliceFamily(α)
+    S = D(b, c)
+    pts = points(S, n)
+    vals = [f(pt...) for pt in pts]
+    cfs = transform(S, vals)
+    x, y = 0.5673, -0.2786; z = [x; y]
+    @test (x^2 + y^2 < 1 && D.α ≤ z[1] ≤ D.β && D.γ*D.ρ(z[1]) ≤ z[2] ≤ D.δ*D.ρ(z[1]))
+    F = Fun(S, cfs)
+    @test F(z) ≈ f(z...)
+    @test itransform(S, cfs) ≈ vals
+    F = Fun(f, S, 10)
+    @test F(z) ≈ f(z...)
+    F = Fun(f, S)
+    @test F(z) ≈ f(z...)
 end
 
 @testset "Jacobi matrices" begin
@@ -305,6 +323,23 @@ end
     α, β = 0.2, 0.9
     D = DiskSliceFamily(α, β)
     S = D(a, b, c)
+    x, y = α + 0.1673, -0.2786; z = [x; y]
+    @test (x^2 + y^2 < 1 && D.α ≤ z[1] ≤ D.β && D.γ*D.ρ(z[1]) ≤ z[2] ≤ D.δ*D.ρ(z[1]))
+    f = (x,y)->x*y + x
+    pts = points(S, n)
[…]

### 2019-06-18 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
biharmonic operator done

@@ -1195,13 +1195,13 @@ function laplaceoperator(S::DiskSliceSpace{<:Any, <:Any, T, <:Any},
         end
     elseif (weighted == true && S.params == ntuple(x->2, D.nparams)
             && St.params == ntuple(x->0, D.nparams))
-        A = weightedpartialoperatorx(differentiateweightedspacex(S), N+2)
+        A = weightedpartialoperatorx(differentiateweightedspacex(S), N+D.nparams)
         @show "laplaceoperator", "1 of 6 done"
         B = weightedpartialoperatorx(S, N)
         @show "laplaceoperator", "2 of 6 done"
-        C = transformparamsoperator(differentiateweightedspacey(D(S.params .- 1)), D(S.params .- 2), N+3, weighted=true)
+        C = transformparamsoperator(differentiateweightedspacey(D(S.params .- 1)), D(S.params .- 2), N+D.nparams+1, weighted=true)
         @show "laplaceoperator", "3 of 6 done"
-        E = weightedpartialoperatory(D(S.params .- 1), N+2)
+        E = weightedpartialoperatory(D(S.params .- 1), N+D.nparams)
         @show "laplaceoperator", "4 of 6 done"
         F = transformparamsoperator(differentiateweightedspacey(S), D(S.params .- 1), N+1, weighted=true)
         @show "laplaceoperator", "5 of 6 done"
@@ -1210,7 +1210,7 @@ function laplaceoperator(S::DiskSliceSpace{<:Any, <:Any, T, <:Any},
         L = A * B + C * E * F * G
         if square
             m = sum(1:(N+1))
-            Δ = BandedBlockBandedMatrix(L[1:m, 1:m], (1:N+1, 1:N+1), (4,-2), (4,0))
+            Δ = BandedBlockBandedMatrix(L[1:m, 1:m], (1:N+1, 1:N+1), (L.l, L.u), (L.λ, L.μ))
         else
             L
         end
@@ -1246,13 +1246,14 @@ end
 function biharmonicoperator(S::DiskSliceSpace{<:Any, <:Any, T, <:Any}, N; square=true) where T
     D = S.family
     if S.params == ntuple(x->2, D.nparams)
-        B = (laplaceoperator(D(S.a-2, S.b-2), S, N+2; square=false)
-                * laplaceoperator(S, D(S.a-2, S.b-2), N; weighted=true, square=false))
[…]

### 2019-06-18 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
First framework for spherical caps family code

@@ -1,0 +1,65 @@
+# Spherical/Polar Caps
+
+
+export SphericalCapFamily, SphericalCapSpace
+
+# R should be Float64, B should be BigFloat
+abstract type SphericalFamily{B,R,N} end
+struct SphericalCap{B,T} <: Domain{SVector{2,T}} end
+SphericalCap() = SphericalCap{BigFloat, Float64}()
+checkpoints(::SphericalCap) = [SVector(0.1,0.23), SVector(0.3,0.12)]
+
+struct SphericalCapSpace{DF, B, T, N} <: Space{SphericalCap{B,T}, T}
+    family::DF # Pointer back to the family
+    params::NTuple{N,B} # Parameters
+    opnorms::Vector{T} # squared norms
+    opptseval::Vector{Vector{T}} # Store the ops evaluated at the transform pts
+    xderivopptseval::Vector{Vector{T}} # Store the x deriv of the ops evaluated
+                                    # at the transform pts
+    yderivopptseval::Vector{Vector{T}} # Store the y deriv of the ops evaluated
+                                    # at the transform pts
+    A::Vector{SparseMatrixCSC{T}}
+    B::Vector{SparseMatrixCSC{T}}
+    C::Vector{SparseMatrixCSC{T}}
+    DT::Vector{SparseMatrixCSC{T}}
+end
+
+function SphericalCapSpace(fam::SphericalFamily{B,T,N}, params::NTuple{N,B}) where {B,T,N}
+    SphericalCapSpace{typeof(fam), B, T, N}(
+        fam, params, Vector{T}(), Vector{Vector{T}}(),
+        Vector{Vector{T}}(), Vector{Vector{T}}(), Vector{SparseMatrixCSC{T}}(),
+        Vector{SparseMatrixCSC{T}}(), Vector{SparseMatrixCSC{T}}(),
+        Vector{SparseMatrixCSC{T}}())
+end
+
+# TODO
+in(x::SVector{2}, D::SphericalCap) = D.α ≤ x[1] ≤ D.β && D.γ*D.ρ(x[1]) ≤ x[2] ≤ D.δ*D.ρ(x[1])
+
+spacescompatible(A::SphericalCapSpace, B::SphericalCapSpace) = (A.params == B.params)
+
+domain(::SphericalCapSpace) = SphericalCap()
+
+# R should be Float64, B BigFloat
[…]

### 2019-11-13 experiments/diskslice.jl
more experients on Rn coeffs

@@ -567,22 +567,26 @@ R̃ = OrthogonalPolynomialFamily(β-X, X-α, 1-X, 1+X)
 
 let
     N = 300
-    for k = 0:100
+    for k = 0:99
         @show k, N
         R00 = R(S.params[1], S.params[2], B(0.5)+k)
         if k == 0
-            @show "begin initialisation"
-            OrthogonalPolynomialFamilies.resizedata!(R00, N+1)
-            @show "end initialisation"
+            # @show "begin initialisation"
+            # OrthogonalPolynomialFamilies.resizedata!(R00, N+1)
+            # @show "end initialisation"
+
+            resize!(R00.a, N+1); resize!(R00.b, N+1)
+            R00.a[:] = load("experiments/saved/R110p5-rec-coeffs-a.jld", "a")
+            R00.b[:] = load("experiments/saved/R110p5-rec-coeffs-b.jld", "b")
         end
         getopptseval(R00, N, [1.0])
         # interim coeffs
         R10 = R̃(S.params[1], S.params[2], B(0.5)+k+1, B(0.5)+k)
         getopnorm(R10)
         resize!(R10.a, N-1); resize!(R10.b, N-1)
-        C00 = zeros(N)
+        C00 = zeros(B, N)
         for n = 0:N-1
-            C00[n+1] = sqrt(getopnorm(R10) / (getopnorm(R00)
+            C00[n+1] = sqrt(B(getopnorm(R10)) / B(getopnorm(R00)
                                                 * R00.b[n+1]
                                                 * R00.opptseval[n+1][1]
                                                 * R00.opptseval[n+2][1]))
@@ -600,10 +604,10 @@ let
         R11 = R(S.params[1], S.params[2], B(0.5)+k+1)
         getopnorm(R11)
         resize!(R11.a, N-2); resize!(R11.b, N-2)
-        D10 = zeros(N-1)
+        D10 = zeros(B, N-1)
         getopptseval(R10, N, [-1.0])
         for n = 0:N-2
-            D10[n+1] = (-1)^(n) * sqrt(- getopnorm(R11) / (getopnorm(R10)
+            D10[n+1] = (-1)^(n) * sqrt(- B(getopnorm(R11)) / B(getopnorm(R10)
[…]

### 2019-11-15 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
revised paper on disk slices first draft (no images)

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage[margin=10pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mleftright}
\usepackage{todonotes}

%\bibliographystyle{apalike}

\graphicspath{ {images/} }

\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

%SetFonts

%SetFonts

\usepackage[square,numbers]{natbib}
\usepackage{url}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plain}

\newcommand{\bigO}{\mathcal{O}}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\No}{\mathbb{N}_0}
\newcommand{\Ylm}{Y^m_l}
\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
\newcommand{\Plm}{\jac^m_l}
\newcommand{\costheta}{\cos\theta}
\newcommand{\sintheta}{\sin\theta}
\newcommand{\cosphi}{\cos\varphi}
\newcommand{\sinphi}{\sin\varphi}
\newcommand{\eimphi}{e^{im\varphi}}
\newcommand{\alphalm}{\alpha^m_l}
\newcommand{\clm}{c^m_l}
\newcommand{\ctilde}{\tilde{c}^m_l}
\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
\newcommand{\chat}{\hat{c}^m_l}
\newcommand{\chatmod}{\hat{c}^{|m|}_l}
\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
\newcommand{\ddy}{\frac{\mathrm{d}}{\mathrm{d}y}}
\newcommand{\pddx}{\frac{\partial}{\partial x}}
\newcommand{\pddy}{\frac{\partial}{\partial y}}
\newcommand{\pddn}{\frac{\partial}{\partial n}}
\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}

\newcommand{\Atilde}{\tilde{A}_{l,m}}
\newcommand{\Btilde}{\tilde{B}_{l,m}}
\newcommand{\Dtilde}{\tilde{D}_{l,m}}
\newcommand{\Etilde}{\tilde{E}_{l,m}}
\newcommand{\Ftilde}{\tilde{F}_{l,m}}
\newcommand{\Gtilde}{\tilde{G}_{l,m}}
\newcommand{\Alm}{A_{l,m}}
\newcommand{\Blm}{B_{l,m}}
\newcommand{\Dlm}{D_{l,m}}
\newcommand{\Elm}{E_{l,m}}
\newcommand{\Flm}{F_{l,m}}
\newcommand{\Glm}{G_{l,m}}

\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}

\newcommand{\bigP}{\mathbb{P}}
\newcommand{\Pl}{\mathbb{P}_l}
\newcommand{\gradP}{T\mathbb{P}}
\newcommand{\gradPl}{T\mathbb{P}_l}
\newcommand{\gradY}{\nabla Y}
\newcommand{\gradYlm}{\nabla Y^m_l}
\newcommand{\gradpY}{\nabla^\perp Y}
\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}

\newcommand{\Dlt}{D^\top_l}

\newcommand{\curlyy}{\bm{\mathcal{Y}}}
\newcommand{\blone}{\beta_{l, 1}}
\newcommand{\blzero}{\beta_{l, 0}}
\newcommand{\blmone}{\beta_{l, -1}}
\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
\newcommand{\cgcoeff}{\mathcal{C}}

\newcommand{\alm}{a_{l,m}}
\newcommand{\blm}{b_{l,m}}
\newcommand{\dlm}{d_{l,m}}
\newcommand{\elm}{e_{l,m}}
\newcommand{\flm}{f_{l,m}}
\newcommand{\glm}{g_{l,m}}
\newcommand{\hlm}{h_{l,m}}
\newcommand{\jlm}{j_{l,m}}
\newcommand{\klm}{k_{l,m}}
\newcommand{\almperp}{a_{l,m}^\perp}
\newcommand{\blmperp}{b_{l,m}^\perp}
\newcommand{\dlmperp}{d_{l,m}^\perp}
\newcommand{\elmperp}{e_{l,m}^\perp}
\newcommand{\flmperp}{f_{l,m}^\perp}
\newcommand{\glmperp}{g_{l,m}^\perp}
\newcommand{\hlmperp}{h_{l,m}^\perp}
\newcommand{\jlmperp}{j_{l,m}^\perp}
\newcommand{\klmperp}{k_{l,m}^\perp}

\newcommand{\unitvec}{\hat{\bm{k}}}

\newcommand{\hdop}{H}
\newcommand{\bighdop}{\mathbb{\hdop}}
\newcommand{\hdopnk}{\hdop_{n,k}}
\newcommand{\hdopnkab}{\hdop_{n,k}^{(a,b)}}
\newcommand{\hdopab}{\hdop^{(a,b)}}
\newcommand{\Wab}{{W^{(a,b)}}}
\newcommand{\hdopmj}{\hdop_{m,j}}
\newcommand{\hdopmjab}{\hdop_{m,j}^{(a,b)}}
\newcommand{\alphaab}{\alpha^{(a,b)}}
\newcommand{\betaab}{\beta^{(a,b)}}
\newcommand{\bighdopab}{\bighdop^{(a,b)}}
\newcommand{\Dnt}{D^\top_n}
\newcommand{\Wii}{W^{(1,1)}}
\newcommand{\hdopii}{\hdop^{(1,1)}}
\newcommand{\bighdopii}{{\mathbb{\hdop}^{(1,1)}}}
\newcommand{\hdopoo}{\hdop^{(0,0)}}
\newcommand{\bighdopoo}{{\mathbb{\hdop}^{(0,0)}}}
\newcommand{\hdopoi}{\hdop^{(0,1)}}
\newcommand{\hdopio}{\hdop^{(1,0)}}
%\newcommand{\jac}{{\tilde P}}
\newcommand{\jac}{{P}}
\newcommand{\genjac}{R}
\newcommand{\genjacnmk}{\genjac_{n-k}}
\newcommand{\genjacmmj}{\genjac_{m-j}}
\newcommand{\genjacw}{w_\genjac}
\newcommand{\jacw}{w_P}
\newcommand{\normgenjac}{\omega_\genjac}
\newcommand{\normjac}{\omega_P}

\newcommand{\dx}{\frac{\partial}{\partial x}}
\newcommand{\dy}{\frac{\partial}{\partial y}}
\newcommand{\laplacewii}{\Delta_W^{(1,1)\to(1,1)}}
\newcommand{\laplacewtt}{\Delta_W^{(2,2)\to(0,0)}}
\newcommand{\laplaceoo}{\Delta^{(0,0)\to(2,2)}}
\newcommand{\biharmonic}{_2\Delta_W^{(2,2)\to(2,2)}}

\newcommand{\element}{\tau}
\newcommand{\refelement}{\hat{\tau}}
\newcommand{\FEset}{\mathcal{T}}
\newcommand{\bigW}{\mathbb{W}}
\newcommand{\bigWab}{\mathbb{W}^{(a,b)}}
\newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
\newcommand{\bigQ}{\mathbb{Q}}
\newcommand{\bigQab}{\bigQ^{(a,b)}}

\newcommand{\hdopnkabc}{\hdop_{n,k}^{(a,b,c)}}
\newcommand{\hdopabc}{\hdop^{(a,b,c)}}
\newcommand{\Wabc}{{W^{(a,b,c)}}}
\newcommand{\hdopmjabc}{\hdop_{m,j}^{(a,b,c)}}
\newcommand{\alphaabc}{\alpha^{(a,b,c)}}
\newcommand{\betaabc}{\beta^{(a,b,c)}}
\newcommand{\bighdopabc}{\bighdop^{(a,b,c)}}
\newcommand{\Wiii}{W^{(1,1,1)}}
\newcommand{\hdopiii}{\hdop^{(1,1,1)}}
\newcommand{\bighdopiii}{{\mathbb{\hdop}^{(1,1,1)}}}
\newcommand{\hdopooo}{\hdop^{(0,0,0)}}
\newcommand{\bighdopooo}{{\mathbb{\hdop}^{(0,0,0)}}}
\newcommand{\hdopooi}{\hdop^{(0,0,1)}}
\newcommand{\hdopiio}{\hdop^{(1,1,0)}}
\newcommand{\laplacewiii}{\Delta_W^{(1,1,1)\to(1,1,1)}}
\newcommand{\laplacewttt}{\Delta_W^{(2,2,2)\to(0,0,0)}}
\newcommand{\laplaceooo}{\Delta^{(0,0,0)\to(2,2,2)}}
\newcommand{\biharmonictwo}{_2\Delta_W^{(2,2,2)\to(2,2,2)}}
\newcommand{\bigWiii}{{\mathbb{W}^{(1,1,1)}}}
\newcommand{\bigWabc}{\mathbb{W}^{(a,b,c)}}

\newcommand{\alphaabcd}{\alpha^{(a,b,c,d)}}
\newcommand{\betaabcd}{\beta^{(a,b,c,d)}}
\newcommand{\hdopnkabcd}{\hdop_{n,k}^{(a,b,c,d)}}
\newcommand{\hdopabcd}{\hdop^{(a,b,c,d)}}
\newcommand{\Wabcd}{{W^{(a,b,c,d)}}}
\newcommand{\hdopmjabcd}{\hdop_{m,j}^{(a,b,c,d)}}
\newcommand{\bighdopabcd}{\bighdop^{(a,b,c,d)}}
\newcommand{\bigWabcd}{\mathbb{W}^{(a,b,c,d)}}

\newcommand{\genjact}{\tilde{\genjac}}
\newcommand{\genjactnmk}{\genjact_{n-k}}
\newcommand{\genjactmmj}{\genjact_{m-j}}
\newcommand{\genjactw}{w_{\genjact}}
\newcommand{\normgenjact}{\omega_{\genjact}}

\usepackage{amsthm}

\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma} 
\newtheorem{theorem}{Theorem} 
\newtheorem{definition}{Definition}
\newtheorem{algorithm}{Algorithm}

\include{somacros}

%\usepackage{media9}
%\graphicspath{ {../sphere/plots} }

\title{Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums}
\author{Sheehan Olver, Ben Snowball}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\begin{abstract}
Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles. In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on \bstodo{Reviewer 1, comment 3}disk slices and trapeziums, with a focus here on the half-disk case. This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials. We apply the framework to solving the Poisson, variable coefficient Helmholtz, and Biharmonic equations. \bstodo{Reviewer 1, comment 3}In this paper we focus on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
 \end{abstract}

%
\section{Introduction}

This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
More precisely, we consider the solution of partial differential equations on the domain
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
where  either of the following scenarios hold:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
\end{enumerate}

For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss \bstodo{typo 1}an extension to other geometries in the appendix (including the half-disk and trapeziums).

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.  On the disk-slice the family of weights we consider are of the form
$$
\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \qfor \alpha \leq x \leq \beta,\quad -\rho(x) \leq y \leq \rho(x).
$$
The corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
$$
\hdopnkab(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
$$
where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms. The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.

Sparsity comes from expanding the domain and range of an operator  using different choices of $a$ and $b$. Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.  In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This compares favourably to $O(p^6)$ operations if one proceeds na\"ively. Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.

%An alternative approach is to use \sotodoinline{Compare with "Discontinuous Galerkin scheme for the spherical shallow water equations with applications to tsunami modeling and prediction"}

Here is an overview of the paper:

\noindent \secref{OPs}: We present our procedure to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the disk-slice domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.

\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.

\noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice. We present a method for explicitly deriving the recurrence coefficients of the non-classical 1D OPs we detail in \secref{OPs}. We derive a quadrature rule on the disk-slice that can be used to expand a function in the OP basis up to a given order. Further, we implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.

\noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the disk-slice.

\noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods. This lays the groundwork for a future $hp$-finite element method in a disk, where the elements capture the circular geometry precisely.

\noindent\appref{HalfDisk}: We discuss extension to the special case of the end-disk-slice.

\noindent\appref{Trapezium}: We discuss extension to trapezia.

\section{Orthogonal polynomials on the disk-slice and the trapezium}\label{Section:OPs}

In this section we outline the construction and some basic properties of $\hdopnkabcd(x,y)$. The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.

\subsection{Explicit construction}

We can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).

\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
\end{enumerate}
$\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$. Then the 2D polynomials defined on $\Omega$
$$
\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
$$
are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)}) \) on $\Omega$. 
\end{proposition}

For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definition. First we introduce notation for two families of univariate OPs.
\begin{definition}\label{def:OPconstruction}
Let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
\begin{align*}
\begin{cases}
\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b
\end{cases}
\end{align*}
and define the associated inner products by:
\begin{align}
	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ipgenjac} \\
	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ipjac}
\end{align}
where
\begin{align}
	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
\end{align}
Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
\end{definition}

\begin{definition}\label{def:constuction}
Define the four-parameter 2D orthogonal polynomials via:
\begin{align*}
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align*}
\end{definition}

$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
\begin{align*}
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
\end{align*}
assuming that either of the following hold:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(\jacw^{(a,b)}\) is an even function (i.e. $a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$).
\end{enumerate}
That is,
\begin{align*}
	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
\end{align*}
where for $f, g : \Omega \to \R$ the inner product is defined as 
\begin{align*}
	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
\end{align*}

We can see that they are indeed orthogonal using the change of variable $t = \frac{y}{\rho(x)}$, for the following normalisation:
\begin{align}
	&\ip< \hdopnkabcd, \: \hdopmjabcd >_\Wabcd \\
	&\quad \quad \quad =  \iint_\Omega \: \Big[ \genjacnmk^{(a,b,c+d+2k+1)}(x) \: \genjacmmj^{(a,b,c+d+2j+1)}(x) \: \rho(x)^{k+j}  \nonumber \\
		&\quad \quad \quad \quad \quad \quad \cdot \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}) \: \jac_j^{(d,c)}\fpr(\frac{y}{\rho(x)})  \: \Wabcd(x,y) \Big] \: \D y \: \D x \nonumber \\
	&\quad \quad \quad = \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b,c+d+2k+1)}}(x) \: {\genjacmmj^{(a,b,c+d+2j+1)}}(x) \: \genjacw^{(a,b,c+d+k+j+1)}(x) \: \D x \Big) \nonumber \\
	&\quad \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \: {\jac_k^{(d,c)}}(t) \:  {\jac_j^{(d,c)}}(t) \: \jacw^{(d,)}(t) \: \D t \Big) \nonumber \\
	&\quad \quad \quad = \normjac^{(d,c)} \: \delta_{k,j} \: \int_\alpha^\beta \: {\genjacnmk^{(a,b,c+d+2k+1)}}(x) \: {\genjac_{m-k}^{(a,b,c+d+2k+1)}}(x) \: \genjacw^{(a,b,c+d+2k+1)}(x) \: \D x \nonumber \\ 
	&\quad \quad \quad = \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}. \label{eqn:normhdop}
\end{align}

For the disk-slice, the weight $\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c$ results from setting:
\begin{align*}
\begin{cases}
	(\alpha,\beta) &\subset (0,1) \\
	(\gamma,\delta) &:= (-1,1) \\
	\rho(x) &:= (1-x^2)^{\half}
\end{cases}
\end{align*}
so that
\begin{align*}
\begin{cases}
	\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c} \\
	\jacw^{(c)}(x) := (1-x)^c \: (1+x)^c = (1-x^2)^c.
\end{cases}
\end{align*}
Note here we can simply remove the need for including a fourth parameter $d$. The 2D OPs orthogonal with respect to the weight above on the disk-slice $\Omega$ are then given by:
\begin{align}\label{eq:diskpolys}
	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
\end{align}
In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).

%Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.

\subsection{Jacobi matrices}

We can express the three-term recurrences associated \bstodo{typo 3}with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$ as 
\begin{align}
	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:Hrecurrence} \\
	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y).
\end{align}
Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0 \quad \forall n=0,1,2,\dots$. We can use these to determine the 2D recurrences for $\hdopnkabcd(x,y)$. Importantly, we can deduce sparsity in the recurrence relationships:

\begin{lemma}
$\hdopnkabc(x,y)$ satisfy the following 3-term recurrences:
\begin{align*}
x \hdopnkabcd(x,y) &= \alphaabc_{n,k,1} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabc_{n,k,2} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabc_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
y \hdopnkabcd(x,y) &= \betaab_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaab_{n,k,2} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \betaab_{n,k,3} \: \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,4} \: \hdop_{n, k-1}^{(a,b,c,d)}(x, y) + \betaab_{n,k,5} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaab_{n,k,6} \: \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,7} \: \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaab_{n,k,8} \: \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaab_{n,k,9} \: \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
\end{align*}
for \((x,y) \in \Omega\), where
\begin{align*}
	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
	\alphaabcd_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,2} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k-1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,3} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-2}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
	\betaabcd_{n,k,4} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,5} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,6} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
	\betaabcd_{n,k,7} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+2}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,8} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k+1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
	\betaabcd_{n,k,9} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjacnmk^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}. 
\end{align*}

\end{lemma}

\begin{proof}
The 3-term recurrence for multiplication by $x$ follows from equation (\ref{eqn:Hrecurrence}). For the recurrence for multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$. These coefficients are given by
\begin{align*}
	c_{m,j} = {\ip< y \: \hdopnkabcd, \hdopmjabcd >_{\Wabcd}}{\norm{\hdopmjabcd}^{-2}_{\Wabcd}}.
\end{align*}
Recall from equation (\ref{eqn:normhdop}) that $\norm{\hdopmjabcd}_{\Wabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$. Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
\begin{align*}
	&\ip<y \hdopnkabcd, \hdopmjabcd>_\Wabcd \\
	&\quad \quad =  \iint_\Omega \hdopnkabcd(x,y) \: \hdopmjabcd(x,y) \: y \: \Wabcd(x,y) \: dy \: dx \\
	&\quad \quad = \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b,c+d)}(x) \: \D x \Big) \nonumber \\
	&\quad \quad \quad \quad \quad\cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
	&\quad \quad = \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \genjacw^{(a, b,c+d+k+j+2)}(x) \: \D x \Big) \nonumber \\
	&\quad \quad \quad \quad \quad\cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big)
\end{align*}
\begin{align*}
	=
	\begin{cases}
    		\delta_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+3)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \quad& \text{if } j = k+1 \\
		\gamma_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+1)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \quad& \text{if } j = k \\
		\delta_{k-1}^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k-1)} \: \ip<\genjacnmk^{(a, b, c+d+2k-1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} \quad& \text{if } j = k-1 \\
		0 & \text{otherwise}
      	\end{cases}
\end{align*}
where, by orthogonality,
\begin{align*}
	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} &= 0 \qfor m < n-1, \\
	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} &= 0 \qfor m < n-1.
\end{align*}

Finally, recall that either:
\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(\jacw^{(d,c)}\) is an even function.
\end{enumerate}
For case 1, we have that 
$$
\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} = 0 \qfor m < n-1.
$$
For case 2 we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.

\end{proof}

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
\begin{align*}
\bighdopabcd_n := \begin{pmatrix}
		\hdopabcd_{n,0}(x,y) \\
		\vdots \\
		\hdopabcd_{n,n}(x,y)
	\end{pmatrix} \in \R^{n+1}, 
\quad \quad 
\bighdopabcd := \begin{pmatrix}
		\bighdopabcd_0 \\
		\bighdopabcd_1 \\
		\bighdopabcd_2 \\
		\vdots \\
	\end{pmatrix}
\end{align*}
and set $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ as the Jacobi matrices corresponding to
\begin{align}
J_x^{(a,b,c,d)} \: \bighdopabcd(x,y) = x \: \bighdopabcd(x,y), \quad J_y^{(a,b,c,d)} \: \bighdopabcd(x,y) = y \: \bighdopabcd(x,y).
\label{eqn:jacobimatricesdefinition}
\end{align}
The matrices $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ act on the coefficients vector of a function's expansion in the $\{\hdopnkabcd\}$ basis. For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopabcd(x,y)^\top \mathbf{f}$. Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b)\top}} \mathbf{f}$. In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkabcd\}$ basis. Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices:

\begin{definition}
A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$. A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix. 
\end{definition}

For example, $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
\begin{align*}
J_{x/y}^{(a,b,c,d)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align*}
where the blocks themselves are diagonal for $J_x^{(a,b,c,d)}$ (sub-block-bandwidths $(0,0)$),
\begin{align*}
A^x_n &:= \begin{pmatrix}
		\alphaabcd_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaabcd_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaabcd_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaabcd_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber
\end{align*}
and tridiagonal for $J_y^{(a,b,c,d)}$ (sub-block-bandwidths $(1,1)$),
\begin{align*}
A^y_n &:= \begin{pmatrix}
		\betaabcd_{n,0,8} & \betaabcd_{n,0,9} & & & \\
		\betaabcd_{n,1,7}& \ddots & \ddots & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaabcd_{n,n,7} & \betaabcd_{n,n,8} & \betaabcd_{n,n,9} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		\betaabcd_{n,0,5} & \betaabcd_{n,0,6} & & \\
		\betaabcd_{n,1,4} & \ddots & \ddots & \\
		& \ddots & \ddots & \betaabcd_{n,n-1,6} \\
		& & \betaabcd_{n,n,4} & \betaabcd_{n,n,5}
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		\betaabcd_{n,0,2} & \betaabcd_{n,0,3} & & \\
		\betaabcd_{n,1,1} & \ddots & \ddots & \\
		& \ddots & \ddots & \betaabcd_{n,n-2,3} \\
		& & \ddots & \betaabcd_{n,n-1,2} \\
		& & & \betaabcd_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align*}

Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs, meaning that the sparsity is not limited to the specific disk-slice case.

\subsection{Building the OPs}

We can combine each system in (\ref{eqn:jacobimatricesdefinition}) into a block-tridiagonal system:
\begin{align*}
\renewcommand\arraystretch{1.3}
\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{pmatrix}
\bighdopabcd(x,y)
=
\begin{pmatrix}
	\hdopabcd_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
\end{pmatrix},
\end{align*}
where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
\begin{align*}
A_n &:= \begin{pmatrix}
		A^x_n \\
		A^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)}, \quad
C_n := \begin{pmatrix}
		C^x_n \\
		C^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{pmatrix}
		B^x_n \\
		B^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y) := \begin{pmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
\end{align*}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+2}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align*}
\begin{cases}
\bighdopabcd_{-1}(x,y) := 0 \\
\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y), \quad n = 0,1,2,\dots.
\end{cases}
\end{align*}

Note that we can define an explicit \(\Dnt\) as follows:
\begin{align*}
\Dnt := \begin{pmatrix}
		\frac{1}{\alphaabcd_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaabcd_{n+1,n,1}} & \\
		\eta_{0} & 0 & \hdots & 0 & \eta_{1} & \hdots & \hdots & \eta_{n+1}
	    \end{pmatrix},
\end{align*}
 where
 \begin{align*}
	\eta_{n+1} &= \frac{1}{\betaabcd_{n,n,9}}, \\ 
	\eta_{n} &= -\frac{1}{\betaabcd_{n,n-1,9}} \big( \betaabcd_{n,n,8} \: \eta_{n+1} \big), \\
	\eta_j &= \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
	\eta_0 &= \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
\end{align*}
It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopabcd_{0}(x,y)$  through $\bighdopabcd_{n}(x,y)$ in optimal $O(n^2)$ complexity.

For the disk-slice, we further note that $\betaabc_{n,k,2} = \betaabc_{n,k,5} = \betaabc_{n,k,8} \equiv 0$ for any $n, k$.

%
\section{Sparse partial differential operators}\label{Section:PDOs}

\begin{figure} 
\center
\begin{tikzpicture} 
% Box/square (4x4)
\draw[black,solid,ultra thin] (0,0)--(0,4);
\draw[black,solid,ultra thin] (0,0)--(4,0);
\draw[black,solid,ultra thin] (4,0)--(4,4);
\draw[black,solid,ultra thin] (0,4)--(4,4);
% Dots
\draw[black,fill=black] (0,0) circle (.5ex);
\draw[black,fill=black] (2,2) circle (.5ex);
\draw[black,fill=black] (4,4) circle (.5ex);
\draw[black,fill=black] (0,2) circle (.5ex);
\draw[black,fill=black] (0,4) circle (.5ex);
% Arrows
\draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
\draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
\draw[black,thick,->] (0,0.2)--(0,1.8);
\draw[black,thick,->] (0,2.2)--(0,3.8);
\draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
% Node (parameter) labels
\draw[] (0,0) node[anchor=east] {$(0,0,0)$};
\draw[] (0,4) node[anchor=east] {$(0,0,2)$};
\draw[] (4,0) node[anchor=west] {$(2,2,0)$};
\draw[] (4,4) node[anchor=west] {$(2,2,2)$};
% Arrow labels
\draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
\draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
% The key and headings
\draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
\draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
\draw[] (-2.3,3.2) node[anchor=west] {$(a,b)$};
\draw[] (-2.7,3.6) node[anchor=south] {$c$};
\draw[] (2,4.6) node[anchor=south] {basis conversion};
\end{tikzpicture} 
\caption{The Laplace operator acting on vectors of $\smash{\hdopnk^{(0,0,0)}}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{\hdopnk^{(2,2,2)}}$ coefficients. Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{\hdopnkabc}$ coefficients, where $(a,b,c)$ is at the tail of the arrow, and the range is $\smash{\hdopnk^{(\tilde{a},\tilde{b},\tilde{c})}}$ coefficients, where $(\tilde{a},\tilde{b},\tilde{c})$ is at the head of the arrow. %For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{\hdop_{n,k}^{(0,0,0)}}$ (resp.~$\smash{\hdop_{n,k}^{(1,0,1)}}$) and the range in $\smash{\hdop_{n,k}^{(1,0,1)}}$ (resp.~$\smash{\hdop_{n,k}^{(2,0,2)}}$).
}
\label{fig:Laplace} 
\end{figure}

In this section, we concentrate on the disk-slice case, and simply note that similar arguments apply for the trapezium case. Recall that, for the disk-slice,
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
where
\begin{align*}
\begin{cases}
	(\alpha,\beta) &\subset (0,1) \\
	(\gamma,\delta) &:= (-1,1) \\
	\rho(x) &:= (1-x^2)^{\half}
\end{cases}.
\end{align*}
The 2D OPs on the disk-slice $\Omega$, orthogonal with respect to the weight
\begin{align*}
	\Wabc(x,y) &:= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
	&= (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \quad (x,y) \in \Omega,
\end{align*}
are then given by:
\begin{align*}
	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
\end{align*}
where the 1D OPs $\{\genjac^{(a, b, c)}_n\}$ are orthogonal on the interval $(\alpha, \beta)$ with respect to the weight
$$
\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c}
$$
and the 1D OPs $\{\jac^{(c, c)}_n\}$ are orthogonal on the interval $(\gamma, \delta) = (-1, 1)$ with respect to the weight
$$
\jacw^{(c)}(x) := (1-x)^c \: (1+x)^c = (1-x^2)^c.
$$

Denote the weighted OPs by
\begin{align*}
\bigWabc(x,y) := \Wabc(x,y) \: \bighdopabc(x,y),
\end{align*}
and recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopabc(x,y)^\top \mathbf{f}$.

\begin{definition}\label{def:differentialoperators}
Define the operator matrices $D_x^{(a,b,c)}, \: D_y^{(a,b,c)}, \: W_x^{(a,b,c)}, \: W_y^{(a,b,c)}$ according to:
\begin{align*}
{\partial f \over \partial x}&= \bighdop^{(a+1,b+1,c+1)}(x,y)^\top \: D_x^{(a,b,c)} \: \mathbf{f}, \\
{\partial f \over \partial y} &= \bighdop^{(a,b,c+1)}(x,y)^\top \: D_y^{(a,b,c)} \: \mathbf{f}, \\
{\partial \over \partial x}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1,c-1)}(x,y)^\top \: W_x^{(a,b,c)} \: \mathbf{f}, \\
{\partial \over \partial y}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a,b,c-1)}(x,y)^\top \: W_y^{(a,b,c)} \: \mathbf{f}.
\end{align*}
\end{definition}

The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.

\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
The operator matrices $D_x^{(a,b,c)}, \: D_y^{(a,b,c)}, \: W_x^{(a,b,c)}, \: W_y^{(a,b,c)}$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure. More specifically:
\begin{itemize}
	\item $D_x^{(a,b,c)}$ has  block-bandwidths $(-1,3)$, and sub-block-bandwidths $(0, 2)$.
  	\item $D_y^{(a,b,c)}$ has  block-bandwidths $(-1,1)$, and sub-block-bandwidths $(-1,1)$.
	\item $W_x^{(a,b,c)}$ has  block-bandwidths $(3,-1)$, and sub-block-bandwidths $(2, 0)$.
  	\item $W_y^{(a,b,c)}$ has  block-bandwidths $(1,-1)$, and sub-block-bandwidths $(1,-1)$.
\end{itemize}
\end{theorem}

\begin{proof}
First, note that:
\begin{align}
	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
	\jacw^{(b) \: \prime}(y) &= - 2b \: y \: \jacw^{(b-1)}(y), \label{eqn:derivativeofweightjac} \\
	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
\end{align}

We proceed with the case for the operator $D_y^{(a,b,c)}$ for partial differentiation by $y$. Since $\{\hdop^{(a,b,c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^y \: \hdop^{(a, b, c+1)}_{m,j}$. The coefficients of the expansion are then the entries of the relevant operator matrix. We can use an integration-by-parts argument to show that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$. First, note that
\begin{align*}
	c_{m,j}^y = {\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
\end{align*}
Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
\begin{align*}
	&\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
	&= \iint_\Omega \: \Big[ \genjacnmk^{(a,b,2c+2k+1)}(x) \: \rho(x)^{k-1} \: \jac_k^{(c,c) \: \prime}\fpr(\frac{y}{\rho(x)}) \\
	& \quad \quad \quad \quad \quad \quad \cdot \: \genjacmmj^{(a,b,2c+2j+3)}(x) \: \rho(x)^{j} \: \jac_j^{(c+1,c+1)}\fpr(\frac{y}{\rho(x)}) \Big] \: \D y \: \D x \\
	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}
\end{align*}
Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
\begin{align*}
	\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} &= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddy [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
	&= - \int_{-1}^1 \: \jac_k^{(c,c)} \: [ \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: y \: \jac_{j}^{(c+1,c+1)} \:\jacw^{(c)} ] \: \D y \\
	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}}
\end{align*}
which is zero for $j < k-1$ by orthogonality. Further, when $j = k-1$, we have that
\begin{align*}
	&\normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacmmj^{(a,b,2c+2k+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
	&= \normgenjac^{(a, b, 2c+2k+1)} \: \delta_{n,m+1},
\end{align*}
showing that the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally,
\begin{align*}
	c_{n-1,k-1}^y &= \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{k-1}^{(c+1,c+1)} >_{\jacw^{(c+1)}}.
\end{align*}

We next consider the case for the operator $D_x^{(a,b,c)}$ for partial differentiation by $x$. Since $\{\hdop^{(a+1, b+1, c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^x \: \hdop^{(a+1, b+1, c+1)}_{m,j}$. The coefficients of the expansion are then the entries of the relevant operator matrix. As before, we can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2, n-3$, $j = k, k-1, k-2$ and $0 \le j \le m$. First, note that
\begin{align*}
	c_{m,j}^x = {\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
\end{align*}
Now, again using the change of variable $t= \frac{y}{\rho(x)}$, and (\ref{eqn:rhoderivative}), we have that
\begin{align*}
	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \\
	&\quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \\ 
	& \quad + k \: \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \\
	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \\
	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \\
	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \\
\end{align*}
%\begin{align*}
%	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
%	&= \normgenjac^{(a, b+\half)} \: \normjac^{(c+1)} \: \Big\{ \ip< \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\ 
%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \cdot \: \ip< \jac_k^{(b,b)}, \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}}  \nonumber \\ 
%	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
%		\: \ip< \jac_k^{(b,b)}, \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}}  \nonumber \\ 
%	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
%		\: \ip< y \: \jac_k^{(b,b) \: \prime} (y), \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}} \Big\}
%\end{align*}

We can see from each term's second factor that the above is only possibly non-zero for $j = k-2, k$. \bstodo{Need to give the working here!!}. Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
\begin{align*}
	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x  \\
	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \\
	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \\
	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \\
	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \\
		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \\
		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \\
		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}
\end{align*}
By orthogonality, each term is is zero for $m-j+3+j-k < n-k \iff m < n - 3$ (since $j-k$ is even by the earlier argument and so $\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)})$,  $\text{deg} (\genjacw^{(1,0, j-k+2)})$, $\text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k$).

Finally,
\begin{align*}
	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \\
	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}}
\end{align*}
which is also zero for $m < n - 3$.

We can gain the non-zero entries of the weighted differential operators similarly, by noting that for the disk-slice
\begin{align}
	\pddx \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
	\pddy \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
\end{align}
and also that
\begin{align*}
	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{W^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wabc.
\end{align*}

\end{proof}

There exist conversion matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.

\begin{definition}\label{def:parametertransformationoperators}
Define the operator matrices $T^{(a,b,c)\to(a+1,b+1,c)}$, $T^{(a,b,c)\to(a,b,c+1)}$ and $T^{(a,b,c)\to(a+1,b+1,c+1)}$ for conversion between non-weighted spaces, and $T_W^{(a,b,c)\to(a-1,b-1,c)}$, $T_W^{(a,b,c)\to(a,b,c-1)}$ and $T_W^{(a,b,c)\to(a-1,b-1,c-1)}$ for conversion between weighted spaces, according to:
\begin{align*}
	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c)} \Big)^\top \: \bighdop^{(a+1,b+1,c)}(x,y) \\
	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a,b,c+1)} \Big)^\top \: \bighdop^{(a,b,c+1)}(x,y) \\
	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c+1)} \Big)^\top \: \bighdop^{(a+1,b+1,c+1)}(x,y) \\
	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a-1,b-1,c)} \Big)^\top \: \bigW^{(a-1,b-1,c)}(x,y) \\
	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a,b,c-1)} \Big)^\top \: \bigW^{(a,b,c-1)}(x,y) \\
	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a-1,b-1,c-1)} \Big)^\top \: \bigW^{(a-1,b-1,c-1)}(x,y).
\end{align*}
\end{definition}

\begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
The operator matrices in Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
\begin{itemize}
	\item $T^{(a,b,c)\to(a+1,b+1,c)}$ has block-bandwidth $(0,2)$, with diagonal blocks.
  	\item $T^{(a,b,c)\to(a,b,c+1)}$ has block-bandwidth $(0,2)$ and sub-block-bandwidth $(0,2)$.
	\item $T^{(a,b,c)\to(a+1,b+1,c+1)}$ has block-bandwidth $(0,4)$ and sub-block-bandwidth $(0,2)$.
	\item $T_W^{(a,b,c)\to(a-1,b-1,c)}$ has block-bandwidth $(2,0)$ with diagonal blocks.
  	\item $T_W^{(a,b,c)\to(a,b,c-1)}$ has block-bandwidth $(2,0)$ and sub-block-bandwidth $(2,0)$.
	\item $T_W^{(a,b,c)\to(a-1,b-1,c-1)}$ has block-bandwidth $(4,0)$ and sub-block-bandwidth $(2,0)$.
\end{itemize}
\end{lemma}

\begin{proof}
We proceed with the case for the non-weighted operators $T^{(a,b)\to(a+\tilde{a},b+\tilde{b},c+\tilde{c})}$, where $\tilde{a}, \tilde{b}, \tilde{c} \in \{0,1\}$. Since $\{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}\}$ for $m = 0,\dots,n$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n$ polynomial, we can expand $\hdopnkabc = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}$. The coefficients of the expansion are then the entries of the relevant operator matrix. We will show that the only non-zero coefficients are for $m \ge n - \tilde{a} - \tilde{b} - 2\tilde{c} $, $j \ge k-2\tilde{c}$ and $0 \le j \le m$. First, note that
\begin{align*}
	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
\end{align*}
Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
\begin{align*}
	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
	&\quad = \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\ 
	&\quad \quad \quad \cdot \: \normjac^{(c+\tilde{c})} \: \ip< \jac_k^{(c,c)}, \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c+\tilde{c})}} \\
	&\quad= \normgenjac^{(a,b,2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
	&\quad \quad \quad \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
\end{align*}
Since $\jacw^{(\tilde{c})}$ is a polynomial degree $2\tilde{c}$, we have that the above is then zero for $j < k-2\tilde{c}$. Further, since $\genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)}$ is a polynomial of degree $\tilde{a}+\tilde{b}+2\tilde{c}+j-k$, we have that the above is zero for $m-j+\tilde{a}+\tilde{b}+2\tilde{c}+j-k < n-k \iff m < n - \tilde{a}-\tilde{b}-2\tilde{c}$.

The sparsity argument for the weighted parameter transformation operators follows similarly.
\end{proof}

\begin{figure}[t]
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityoflaplacian-w11}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofhelmholtz}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofbiharmonic}
        \centering
	\end{subfigure}
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. Left: the Laplace operator $\laplacewiii$. Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. Right: the biharmonic operator $\biharmonictwo$.}
        \label{fig:sparsity}
        \centering
\end{figure}

General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators. As a canonical example, we can obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
$$
\bigWiii(x,y) = \Wiii(x,y) \: \bighdopiii(x,y)
$$
to coefficients in the non-weighted space $\bighdopiii(x,y)$. Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$. The matrix operator for the Laplacian we denote $\laplacewiii$ acting on the coefficients vector is then given by
\begin{align*}
    \laplacewiii := D_x^{(0,0,0)} \: W_x^{(1,1,1)} + T^{(0,0,1)\to(1,1)} \: D_y^{(0,0,0)} \: T_W^{(1,1,0)\to(0,0,0)} \: W_y^{(1,1,1)}.
\end{align*}
Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.

Another important example is the Biharmonic operator $\Delta^2$, where we assume zero Dirichlet and Neumann conditions. To construct this operator, we first note that we can obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopooo(x,y)$ to coefficients in the space $\bighdop^{(2,2)}(x,y)$. We denote this matrix operator that acts on the coefficients vector as $\laplaceooo$, and is given by
\begin{align*}
    \laplaceooo := D_x^{(1,1,1)} \: D_x^{(0,0,0)} + T^{(1,1,2)\to(2,2,2)} \: D_y^{(1,1,1)} \: T^{(0,0,1)\to(1,1,1)} \: D_y^{(0,0,0)}.
\end{align*}
Further, we can represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopooo$. Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$. We denote this matrix operator as $\laplacewttt$, and is given by
\begin{align*}
	\laplacewttt := W_x^{(1,1,1)} \: W_x^{(2,2,2)} + T_W^{(1,1,0)\to(0,0,0)} \: W_y^{(1,1,1)} \: T_W^{(2,2,1)\to(1,1,1)} \: W_y^{(2,2,2)}.
\end{align*}
We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2,2)}$ to coefficients in the space $\bighdop^{(2,2,2)}$. Note that any function expanded in the $\bigW^{(2,2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$. The matrix operator for the Biharmonic operator is then given by
\begin{align*}
	\biharmonictwo = \laplaceooo \: \laplacewttt.
\end{align*}
The sparsity and structure of this biharmonic operator \bstodo{typo 4}are seen in Figure \ref{fig:sparsity}.

\section{Computational aspects}\label{Section:Computation}

In this section we discuss how to take advantage of the proposed basis and sparsity structure in partial differential operators in practical computational applications.

\subsection{Constructing $\genjac_n^{(a,b,c)}(x)$}

%To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature. This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $c$ is large, as required for large $n$ in \eqref{eq:diskpolys}. The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm, utilising the {\tt BigFloat} type to handle high-precision calculations.

It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \bstodo{Link!!!}. We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case. We thus first need to define a new set of `interim' 1D OPs.
\begin{definition}\label{def:InterimOPconstruction}
Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
\begin{align}
	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ipgenjact}
\end{align}
where
\begin{align}
	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
\end{align}
Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjact}).
\end{definition}
Note that the OPs $\{\genjac_n^{(a,b,2c)}\}$ are then equivalent to the OPs $\{\genjact_n^{(a,b,c,c)}\}$. Let the recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d)}\}$ be given by:
\begin{align}
	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
\end{align}
\begin{proposition}
There exist constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ such that
\begin{align}
	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
\end{align}
\end{proposition}

\begin{proof}
Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
\begin{align*}
	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
\end{align*}
and define \bstodo{check signs}
\begin{align}
	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half. \label{eqn:genjactnormalisationii}
\end{align}
Now, by the Christoffel-Darboux formula \bstodo{cite}, we have that for any $x, y \in \R$,
\begin{align}
	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}. \label{eqn:christoffeldarboux}
\end{align}
Then,
\begin{align*}
	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
\end{align*}
using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
\begin{align*}
	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
\end{align*}
using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
\end{proof}

\begin{proposition}
The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c+1,d)}\}$ are given by:
\begin{align}
	\tilde{\alpha}_n^{(a,b,c+1,d)} &= \frac{\genjact_{n+2}^{(a,b,c,d)}(1)}{\genjact_{n+1}^{(a,b,c,d)}(1)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} - \frac{\genjact_{n+1}^{(a,b,c,d)}(1)}{\genjact_{n}^{(a,b,c,d)}(1)} \: \tilde{\beta}_{n}^{(a,b,c,d)} + \tilde{\alpha}_{n+1}^{(a,b,c,d)}, \\
	\tilde{\beta}_n^{(a,b,c+1,d)} &= \frac{\mathcal{C}_n^{(a,b,c,d)}}{\mathcal{C}_{n+1}^{(a,b,c,d)}} \: \frac{\genjact_{n}^{(a,b,c,d)}(1)}{\genjact_{n+1}^{(a,b,c,d)}(1)} \: \tilde{\beta}_n^{(a,b,c,d)}.
\end{align}
The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d+1)}\}$ are given by:
\begin{align}
	\tilde{\alpha}_n^{(a,b,c,d+1)} &= \frac{\genjact_{n+2}^{(a,b,c,d)}(-1)}{\genjact_{n+1}^{(a,b,c,d)}(-1)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} - \frac{\genjact_{n+1}^{(a,b,c,d)}(-1)}{\genjact_{n}^{(a,b,c,d)}(-1)} \: \tilde{\beta}_{n}^{(a,b,c,d)} + \tilde{\alpha}_{n+1}^{(a,b,c,d)}, \\
	\tilde{\beta}_n^{(a,b,c,d+1)} &= \frac{\mathcal{D}_n^{(a,b,c,d)}}{\mathcal{D}_{n+1}^{(a,b,c,d)}} \: \frac{\genjact_{n}^{(a,b,c,d)}(-1)}{\genjact_{n+1}^{(a,b,c,d)}(-1)} \: \tilde{\beta}_n^{(a,b,c,d)}.
\end{align}
\end{proposition}

\begin{proof}
First, using (\ref{eqn:genjacexpansioni}) and (\ref{eqn:christoffeldarboux}) we have that
\begin{align}
	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: x \: \Big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\Big] \nonumber \\
	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \nonumber \\
	& \quad \quad \cdot \Big[\Big(\tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)\Big) \: \genjact_{n+1}^{(a,b,c,d)}(1) \nonumber \\
	& \quad \quad \quad - \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \: \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x)\Big) \: \genjact_{n}^{(a,b,c,d)}(1) \Big] \label{eqn:genjactproofa}
\end{align}
Next, note that the recurrence coefficients for $\genjact_{n}^{(a,b,c+1,d)}(x)$ satisfy
\begin{align}
	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
	&= (1-x) \: \Big[\tilde{\beta}_{n}^{(a,b,c+1,d)} \: \genjact_{n+1}^{(a,b,c+1,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \genjact_{n}^{(a,b,c+1,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \genjact_{n-1}^{(a,b,c+1,d)}(x)\Big] \nonumber \\
	&= \mathcal{C}_{n+1}^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} \: \Big( \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n+2}^{(a,b,c,d)}(1) - \genjact_{n+2}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) \Big) \nonumber \\
	& \quad \quad + \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
	& \quad \quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \: \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:genjactproofb}
\end{align}
We can set $\tilde{\beta}_{-1}^{(a,b,c+1,d)} = 0$. By comparing coefficients of $\genjact_{n+2}^{(a,b,c,d)}(x)$ and $\genjact_{n+1}^{(a,b,c,d)}(x)$ in both (\ref{eqn:genjactproofa}) and (\ref{eqn:genjactproofb}) we obtain the desired recurrence coefficients for the OP $\genjact_{n}^{(a,b,c+1,d)}(x)$. The recurrence coefficients for the OPs $\genjact_{n}^{(a,b,c,d+1)}(x)$ are found similarly.
\end{proof}

These two propositions allow us to recursively obtain the recurrence coefficients for the OPs $\{\genjac_{n-k}^{(a,b,2c+2k + 1)}\}$ as $k$ increases to be large.

\noindent{\bf Remark}:  \bstodo{improve this remark}
This is the most expensive part of the current calculation, in particular the initialisation required (i.e. using the Lanczos algorithm to compute the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+1)}\}$, $\{\beta_{n}^{(a,b,2c+1)}\}$ -- however, we only need to compute these once, and can store and save this initialisation to disk once computed, for the given values of $a, b, c$). Note also that in order to obtain the recurrence coefficients $\{\alpha_{n-k}^{(a,b,2c+2k+1)}\}$, $\{\beta_{n-k}^{(a,b,2c+2k+1)}\}$ for $n = 0,\dots,N$ for some $N$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{n}^{(a,b,2c+2(k-1)+1)}\}$ for $n = 0,\dots,N+3$. Thus, for large $k$, this recursive method of obtaining the recurrence coefficients requires a large initialisation.
%This is the most expensive part of the current calculation, but note that we can reuse this computation for multiple partial differential equations. Efficient construction of OPs with large parameters in the weights is an important topic, but one tangential to the proposed scheme.

\subsection{Quadrature rule on the disk-slice}

In this section we construct a quadrature rule exact for polynomials in the disk-slice $\Omega$ that can be used to expand functions in $\hdopnkabc(x,y)$ when $\Omega$ is a disk-slice.

\begin{theorem}
Denote the  Gauss quadrature nodes and weight on $[\alpha,\beta]$ with weight $(\beta - s)^a \: (s - \alpha)^b \: \rho(s)^{2c+1}$ as $(s_k,w_k^{(s)})$ , and
 on \([-1,1]\) with weight \((1-t^2)^c\) as $(t_k,w_k^{(t)})$. Define
\begin{align*}
	x_{i+(j-1)N} &:= s_j, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
	y_{i+(j-1)N} &:= \rho(s_j) \: t_i, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
	w_{i+(j-1)N} &:= w_j^{(s)} w_i^{(t)}, \quad  i,j = 1,\dots, \ceil{\frac{N+1}{2}}.
\end{align*}
Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
$$
\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
$$
where $M = \ceil{\half(N+1)}^2$, and the quadrature rule is exact if $f(x,y)$ is a polynomial of degree $\le N$.
\end{theorem}

\begin{proof}
We will use the substitution that
\begin{align*}
	x &= s, \quad y = \rho(s) \: t.
\end{align*}
First, note that, for $(x,y) \in \Omega$,
\begin{align*}
	\Wabc(x,y) &= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
	&= \genjacw^{(a,b,c2)}(s) \: \jacw^{(c)}(t) \\
	&=: V^{(a,b,c)}(s,t), \quad \text{for } (s,t) \in [\alpha,\beta] \times [-1,1].
\end{align*}

Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
\begin{align*}
	f_e(x,y) &:= \half \Big(f(x, y) + f(x, -y)\Big), \quad \forall (x,y) \in \Omega\\
	f_o(x,y) &:= \half \Big(f(x, y) - f(x, -y)\Big), \quad \forall (x,y) \in \Omega
\end{align*}
so that $y \mapsto f_e(x,y)$ for fixed $x$ is an even function, and $y \mapsto f_o(x,y)$ for fixed $x$ is an odd function. Note that if $f$ is a polynomial, then $f_e(s, \rho(s)t)$ is a polynomial in $s \in [\alpha,\beta]$ for fixed $t$.

Now, we have that
\begin{align*}
	\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
	&= \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \\
	&\approx \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
	&= \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
\end{align*}
Suppose $f$ is a polynomial in $x$ and $y$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial. First, note that the degree of the polynomial given by $x \mapsto f_e(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f_e(x,y)$ for fixed $x$ is $\le N$. Also note that $s \mapsto f_e\big(s,\rho(s) t\big)$ for fixed $t$ is then a degree $N$ polynomial (since $\rho$ is a degree $1$ polynomial). Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.

Next, note that 
\begin{align*}
	\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
	&= \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
	&= 0
\end{align*}
since the inner integral at $(\dagger)$ over $t$ is zero, due to the symmetry over the domain.

Hence, for a polynomial $f$ in $x$ and $y$ of degree $N$,
\begin{align*}
	\iint_\Omega f(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \iint_\Omega \Big(f_e(x,y) + f_o(x,y)\Big) \: \Wabc(x,y) \: \D y \: \D x \\
	&= \iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x \\
	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j),
\end{align*}
where $M = \ceil{\half(N+1)}^2$.
\end{proof}

\subsection{Obtaining the coefficients for expansion of a function on the disk-slice}

Fix \(a,b,c \in \R\). Then for any function \(f : \Omega \to \R\) we can express \(f\) by
\begin{align*}
	f(x,y) \approx \sum_{n=0}^N \bighdop_n^{(a,b,c)}(x,y)^\top \: \mathbf{f}_n
\end{align*}
for N sufficiently large, where
\begin{align*}
	\bighdop^{(a,b,c)}_n(x,y) &:= \begin{pmatrix}
		\hdop^{(a,b,c)}_{n,0}(x,y) \\
		\vdots \\
		\hdop^{(a,b,c)}_{n,n}(x,y)
	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
\end{align*}
and where
\begin{align*}
	\mathbf{f}_n &:= \begin{pmatrix}
		f_{n,0} \\
		\vdots \\
		f_{n,n}
	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
	f_{n,k} := \frac{\ip< f, \: \hdopnk^{(a,b,c)}>_{\Wabc}}{\norm{\hdopnk^{(a,b,c)}}_{\Wabc}}
\end{align*}
Recall from (\ref{eqn:normhdop}) that $\norm{\hdopnkabc}^2_\Wabc = \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}$. Using the quadrature rule detailed in Section 4.2 for the inner product, we can calculate the coefficients $f_{n,k}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$: 
\begin{align*}
	f_{n,k} &= \frac{1}{2 \: \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}} \: \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) \: \hdopnkabc(x_j, y_j) +f(x_j, -y_j) \: \hdopnkabc(x_j, -y_j) \big]
\end{align*}
where $M = \ceil{\half(N+1)}^2$.

\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}

The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}. We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$. This proves much cheaper computationally than using the 2D quadrature rule to calculate the 2D inner products, which has a complexity of $\bigO(N^4)$.

%
\section{Examples on the disk-slice with zero Dirichlet conditions}\label{Section:Examples}

We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions. We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, demonstrating the versatility of the approach.

\subsection{Poisson}

% FIGURES
\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-poisson}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.5]{solutionblocknorms-poisson}
        	%\label{fig:solutionblocknorms-poisson}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
	\centering
	\label{fig:poisson}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz-k=100-n=1500}
	%\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-k=20}
	\centering
        	%\label{fig:solutionblocknorms-helmholtz}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
	\centering
	\label{fig:helmholtz}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic}
	%\label{fig:solution-biharmonic}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic}
	\centering
        	%\label{fig:solutionblocknorms-biharmonic}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
	\centering
	\label{fig:biharmonic}
\end{figure}
% END FIGURES

The Poisson equation is the classic problem of finding \(u(x,y)\) given a function \(f(x,y)\) such that:
\begin{align}
	\begin{cases}
    		\Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
		u(x,y) = 0 \quad \text{on } \partial \Omega
	\end{cases}.
	\label{eqn:poisson}
\end{align}
noting the imposition of zero Dirichlet boundary conditions on $u$.

We can tackle the problem as follows. Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above. In matrix-vector notation, our system hence becomes:
\begin{align*}
    \laplacewiii \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$.
In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$. In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns. The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution. We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence. For the final Gaussian bump example, we see that we achieve spectral convergence.

\subsection{Inhomogeneous variable-coefficient Helmholtz}

Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
\begin{align}
	\begin{cases}
    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
		u(x,y) = 0 \quad \text{on } \partial \Omega
	\end{cases}.
	\label{eqn:helmholtz}
\end{align}
where $k \in \R$, noting the imposition of zero Dirichlet boundary conditions on $u$.

We can tackle the problem as follows. Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule above. We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}. We can denote the resulting operator acting on coefficients in the $\bighdopooo$ space by $V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top)$. In matrix-vector notation, our system hence becomes:
\begin{align*}
    (\laplacewiii + k^2 T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}) \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example. In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$. In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using N = 200, that is, 20,301 unknowns. The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution. We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence. For the final Gaussian bump example, we see that we achieve spectral convergence.

We can extend this to constant non-zero boundary conditions by simply noting that the problem 
\begin{align*}
	\begin{cases}
    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
		u(x,y) = c \in \R \quad \text{on } \partial \Omega
	\end{cases}
\end{align*}
is equivalent to letting $u = \tilde{u} + c$ and solving
\begin{align*}
	\begin{cases}
    		\Delta \tilde{u}(x,y) + k^2 \: v(x,y) \; \tilde{u}(x,y) = f(x,y) - c \: k^2 \: v(x,y) \; =: g(x,y)  \quad \text{in } \Omega \\
		\tilde{u}(x,y) = 0 \quad \text{on } \partial \Omega.
	\end{cases}.
\end{align*}

\subsection{Biharmonic equation}

Find \(u(x,y)\) given a function \(f(x,y)\) such that:
\begin{align}
	\begin{cases}
    		\Delta^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
	\end{cases}.
	\label{eqn:biharmonic}
\end{align}
where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$. In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$. In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.  We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence. For the final Gaussian bump example, we see that we achieve spectral convergence.

%
\section{Conclusions}

\bstodo{Reviewer 2, comments 2, 3}
We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions. \bstodo{Reviewer 1, comment 3} This work is related to the triangle case \cite{olver2019triangle}, where ????. This work forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements. 
%A further extension is to develop similar relationships for 3D domains such as the half-sphere surface. With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres. 
This work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles. In particular, orthogonal polynomials (OPs) in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere. The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}). Constructing these sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}. Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.

%
\appendix
%
\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}

We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric. Consider the 2D Dirichlet problem on a domain $\Omega$:
\begin{align*}
	\begin{cases}
         - \Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
         u = 0 \quad \text{on } \partial \Omega
         \end{cases}
\end{align*}
This has the weak formulation for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$,
\begin{align*}
	L(v) := \int_\Omega f \: v \: d\mathbf{x} = \int_\Omega \nabla u \cdot \nabla v \: d\mathbf{x} =: a(u,v).
\end{align*}

In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a trapezium or disk slice for example.

In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a disk-slice domain. We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.

We seek $u_p \in V_p$ s.t.
\begin{align}
	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
	\label{eqn:FEMweakform}
\end{align}

Define $\Lambda^{(a,b,c)} :=  \ip< \bighdop^{(a,b,c)}, \: {\bighdop^{(a,b,c)}}^\top >_{\Wabc}$ where $\Wabc$ is the weight with which the OPs in $\bighdop^{(a,b,c)}$ are orthogonal with respect to. Note that due to orthogonality this is a diagonal matrix. We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
\begin{align*}
\bigWiii(x,y) &:= \begin{pmatrix}
		\bigWiii_0(x,y) \\
		\bigWiii_1(x,y) \\
		\bigWiii_2(x,y) \\
		\vdots \\
		\bigWiii_p(x,y)
	\end{pmatrix}, \\
\bigWiii_n(x,y) &:=  \Wiii(x,y) \begin{pmatrix}
						\hdopiii_{n,0}(x,y) \\
						\vdots \\
						\hdopiii_{n,n}(x,y)
					\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
\end{align*}
and rewrite (\ref{eqn:FEMweakform}) in matrix form:
\begin{align*}
	a(u_p,v_p) &= \int_\element \nabla u_p \cdot \nabla v_p \: \D \mathbf{x} \\
	&= \int_\element \begin{pmatrix}
					\partial_x v_p \\
					\partial_y v_p
				\end{pmatrix}^\top 
				\begin{pmatrix}
					\partial_x u_p \\
					\partial_y u_p
				\end{pmatrix} \: \D \mathbf{x}
				\\
	&= \int_\element \begin{pmatrix}
					\bighdopooo^\top \Wiii_x \mathbf{v} \\
					\bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \mathbf{v}
				\end{pmatrix}^\top 
				\begin{pmatrix}
					\bighdopooo^\top \Wiii_x \mathbf{u} \\
					\bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \mathbf{u}
				\end{pmatrix} \: \D \mathbf{x}
				\\
	&= \int_\element \Big( \mathbf{v}^\top {\Wiii_x}^\top \bighdopooo \bighdopooo^\top \Wiii_x \mathbf{u} \nonumber \\
					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \bighdopooo \bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \mathbf{u}  \Big) \: \D \mathbf{x} \\
	&= \mathbf{v}^\top \: \Big( {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x \\
	& \quad \quad \quad \quad + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \Big) \: \mathbf{u}
\end{align*}
where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs), and
\begin{align*}
	L(v_p) &= \int_\element \: v_p \: f \: \D \mathbf{x}\\
	&= \int_\element \: \mathbf{v}^\top \: \bigWiii \: \bighdopiii^\top \: \mathbf{f} \: \D \mathbf{x} \\
	&= \mathbf{v}^\top \: \ip< \bighdopiii, {\bighdopiii}^\top >_{\Wiii} \: \D \mathbf{x} \\
	&= \mathbf{v}^\top \Lambda^{(1,1,1)} \: \mathbf{f},
\end{align*}
where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bighdopiii$ OP basis.

Since (\ref{eqn:FEMweakform}) is equivalent to stating that
\begin{align*}
	L(\Wiii \hdopiii_{n,k}) = a(u_p,\Wiii \hdopiii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
\end{align*}
(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
\begin{align*}
	A\mathbf{u} = \tilde{\mathbf{f}}.
\end{align*}
where the (element) stiffness matrix $A$ is defined by 
\begin{align*}
	A = {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y
\end{align*}
and the load vector $\tilde{\mathbf{f}}$ is given by 
\begin{align*}
	\tilde{\mathbf{f}} = \Lambda^{(1,1,1)} \: \mathbf{f}.
\end{align*}

Note \bstodo{typo 6}that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).

\section{End-Disk-Slice}\label{Appendix:HalfDisk}

The work in this paper on the disk-slice can be easily transferred to the special-case domain of the end-disk-slice by which we mean
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
with
\begin{align*}
\begin{cases}
\alpha &\in (0,1) \\
\beta &:= 1 \\
(\gamma, \delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half}.
\end{cases}
\end{align*}
Our 1D weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively are then given by:
\begin{align*}
\begin{cases}
\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b} \\
\jacw^{(a)}(x) &:= (1-x^2)^b.
\end{cases}
\end{align*}

Note here how we can remove the need for third parameter, which is why we consider this a special case. This will make some calculations easier, and the operator matrices more sparse. The weight $\jacw^{(b)}(x)$ is a still the same ultraspherical weight (and the corresponding OPs are the Jacobi polynomials $\{\jac_n^{(b, b)}\}$). $\genjacw^{(a,b)}(x)$ is the (non-classical) weight for the OPs denoted $\{\genjac_n^{(a,b)}\}$. Thus we arrive at the two-parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align*}
	\hdopnkab(x,y) := \genjacnmk^{(a, 2b+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align*}
orthogonal with respect to the weight
\begin{align*}
	\Wab(x,y) &:= \genjacw^{(a,2b)}(x) w^{(b)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
	&= (x - \alpha)^{a} \: (\rho(x)^2 -y^2)^b \nonumber \\
	&= (x - \alpha)^{a} \: (1 - x^2 -y^2)^b , \quad (x,y) \in \Omega.
\end{align*}

The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such end-disk-slice domains. For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for an end-disk-slice.

\section{Trapeziums}\label{Appendix:Trapezium}

We can further extend this work to trapezium shaped domains. Note that for any trapezium there exists an affine map to the canonical trapezium domain that we consider here, given by
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
with 
\begin{align*}
\begin{cases}
(\alpha, \beta) &:= (0,1) \\
(\gamma, \delta) &:= (0,1) \\
\rho(x) &:= 1- \xi x, \quad \xi \in (0,1) \\
\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\xi x)^{c} \\
\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
\end{cases}
\end{align*}
The weight $\jacw^{(a,b)}(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$. We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = \jac_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$. $\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\). Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align*}
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align*}
orthogonal with respect to the weight
\begin{align*}
	W^{(a,b,c,d)}(x,y) &:= \genjacw^{(a,b,c+d)}(x) \: w^{(d,c)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
	&= (1-x)^a \: x^b \: y^c \: (1- \xi x - y)^d, \quad (x,y) \in \Omega.
\end{align*}

%\subsection{Quadrature rule on the trapezium}
%
%In this section we construct a quadrature rule exact for polynomials in the trapezium $\Omega$ that can be used to expand functions in $\hdopnkabcd(x,y)$.
%
%\begin{theorem}
%Denote the Gauss quadrature nodes and weight on $[\alpha,\beta]$ with weight $\genjacw^{(a,b,c+d)}(s) = (\beta - s)^a \: (s - \alpha)^b \: \rho(s)^{c+d+1}$ as $(s_k,w_k^{(s)})$ , and
% on $[0,1]$ with weight $\jacw^{(d,c)}(t) = t^c \: (1-t)^d$ as $(t_k,w_k^{(t)})$. Define
%\begin{align*}
%	x_{i+(j-1)N} &:= s_j, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
%	y_{i+(j-1)N} &:= \rho(s_j) \: t_i, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
%	w_{i+(j-1)N} &:= w_j^{(s)} w_i^{(t)}, \quad  i,j = 1,\dots, \ceil{\frac{N+1}{2}}.
%\end{align*}
%Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
%$$
%\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D A \approx \sum_{j=1}^{M} w_j \: f(x_j, y_j)
%$$
%where $M = \ceil{\half(N+1)}^2$, and the quadrature rule is exact if $f(x,y)$ is a polynomial of degree $\le N$.
%\end{theorem}
%
%\begin{proof}
%\begin{align*}
%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
%	&\approx \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
%	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
%	&= \sum_{k=1}^{M_1 M_2}  w_k \: f(x_k, y_k).
%\end{align*}
%Suppose f is a polynomial in $x$ and $y$ of degree $N$. First, note that the degree of the polynomial given by $x \mapsto f(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f(x,y)$ for fixed $x$ is $\le N$. Also note that $s \mapsto f\big(s,\rho(s) t\big)$ for fixed $t$ is then a degree $N$ polynomial (since $\rho$ is a degree $1$ polynomial). Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
%
%Thus, for a polynomial $f$ in $x$ and $y$ of degree $N$, 
%\begin{align*}
%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
%	&= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad \\
%	&= \sum_{j=1}^{\ceil{\half(N+1)}} \Bigg( w_j^{(s)} \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \\
%	&= \sum_{k=1}^{M}  w_k \: f(x_k, y_k)
%\end{align*}
%where $M = \ceil{\half(N+1)}^2$.
%
%\end{proof}

\bstodo{typo 8 (capitalization of names in some reference titles}
\bibliography{halfdisk}

\end{document}

### 2019-11-19 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
further edits to proof of lemma.

@@
 	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
-	\jacw^{(b) \: \prime}(y) &= - 2b \: y \: \jacw^{(b-1)}(y), \label{eqn:derivativeofweightjac} \\
+	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(b-1)}(y), \label{eqn:derivativeofweightjac} \\
 	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
@@
 \end{align*}
-Now, again using the change of variable $t= \frac{y}{\rho(x)}$, and (\ref{eqn:rhoderivative}), we have that
-\begin{align*}
+Now, again using the change of variable $t= \frac{y}{\rho(x)}$, we have that
+\begin{align}
 	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
-	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \\
-	&\quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \\ 
-	& \quad + k \: \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \\
-	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \\
-	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \\
-	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \\
-\end{align*}
-%\begin{align*}
-%	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
-%	&= \normgenjac^{(a, b+\half)} \: \normjac^{(c+1)} \: \Big\{ \ip< \genjacnmk^{(a, b+k+\half) \: \prime} (x), \: \rho(x)^{k+j+2} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} \\ 
-%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \cdot \: \ip< \jac_k^{(b,b)}, \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}}  \nonumber \\ 
-%	& \quad - k \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-%		\: \ip< \jac_k^{(b,b)}, \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}}  \nonumber \\ 
-%	& \quad + \: \ip< x^2 \: \genjacnmk^{(a, b+k+\half)}(x), \: \rho(x)^{k+j} \: \genjacmmj^{(a+1, b+1+j+\half)} >_{\genjacw^{(a, b+\half)}} 
-%		\: \ip< y \: \jac_k^{(b,b) \: \prime} (y), \: \jac_j^{(b+1,b+1)} >_{\jacw^{(b+1)}} \Big\}
-%\end{align*}
-
-We can see from each term's second factor that the above is only possibly non-zero for $j = k-2, k$.
-\bstodo{Need to give the working here!!}.
+	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
+	&\quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\ 
+	& \quad + k \: \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
+	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\
+	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
+	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:dxsparsity}
+\end{align}
+We will first show that the second factor of each term in (\ref{eqn:dxsparsity}) are zero for $j < k - 2$ and also for $j = k - 1$.
+To this end, observe that, for any integer $c$, $\jac^{(c,c)}(-t) = (-1)^k \: \jac^{(c,c)}(t)$ and so $\jac_k^{(c,c)}$ is an even polynomial for even $k$, and an odd polynomial for odd $k$.
+Thus, $\jac_k^{(c,c)} \: \jac_{k-1}^{(c+1,c+1)}$ is an odd polynomial for any $k$.
+Hence
+\begin{align*}
+	\int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t = \int_{-\delta}^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} \: \jacw^{(c)} \: \D t 
+\end{align*}
+is zero for $j < k - 2$ by orthogonality, and is zero for $j = k-1$ due to symmetry over the domain.
+Moreover, $t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t)$ is also an odd polynomial for any $k$ and so
+\begin{align*}
+	\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t) \: \jacw^{(c+1)}(t) \: \D t
+\end{align*}
+is zero for $j = k-1$ due to symmetry over the domain, and
+\begin{align*}
+	&\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \\
+	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \frac{\D}{\D t} \big[ t \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \big] \: \D t \\
+	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \big[ \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} + t \: \jac_j^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: t^2 \: \jac_j^{(c+1,c+1)} \: \jacw^{(c)} \big] \: \D t \\
+	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}}
+\end{align*}
+which is zero for $j < k - 2$ by orthogonality. Thus, (\ref{eqn:dxsparsity}) is zero for $j \notin \{k-2, k\}$.
+
 Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
-\begin{align*}
-	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x  \\
-	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \\
-	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \\
-	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \\
-	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \\
-		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \\
-		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \\
-		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}
-\end{align*}
-By orthogonality, each term is is zero for $m-j+3+j-k < n-k \iff m < n - 3$ (since $j-k$ is even by the earlier argument and so $\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)})$,  $\text{deg} (\genjacw^{(1,0, j-k+2)})$, $\text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k$).
+\begin{align}
+	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \nonumber  \\
+	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \nonumber \\
+	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \nonumber \\
+	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \nonumber \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \nonumber \\
+	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:dxsparsityR}
+\end{align}
+By recalling (\ref{eqn:rhoderivative}) and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
+$$\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)}) = \text{deg} (\genjacw^{(1,0, j-k+2)}) = \text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k.$$ 
+Hence, by orthogonality, each term in (\ref{eqn:dxsparsityR}) is is zero for $m-j+3+j-k < n-k \iff m < n - 3$.
 
@@
 \end{align*}
-which is also zero for $m < n - 3$.
+which is also zero for $m < n - 3$. Thus
+\begin{align*}
+	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
+\end{align*}
+showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,\dots,n$ and $j = k-2,k$ $(j \le m)$.
 
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityoflaplacian-w11}
+	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofhelmholtz}
+	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofbiharmonic}
+	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\centering
-	\includegraphics[scale=0.5]{solutionblocknorms-poisson}
+	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-k=20}
+	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
 	\centering
@@
 \section{End-Disk-Slice}\label{Appendix:HalfDisk}
+
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-poisson}
+	\centering
+	%\label{fig:solution-poisson}
+	\end{subfigure}
+	\begin{subfigure}{0.5\textwidth}
+	\centering
+	\includegraphics[scale=0.5]{solutionblocknorms-poisson}
+        	%\label{fig:solutionblocknorms-poisson}
+	\end{subfigure}
+	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
+Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
+This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
+	\centering
+	\label{fig:poisson}
+\end{figure}
 

### 2019-11-19 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
resolving merge issues

@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{sparsityoflaplacian-w11} %-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{sparsityofhelmholtz} %-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{sparsityofbiharmonic} %-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\centering
-	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.5]{solutionblocknorms-poisson} %-diskslice-alpha=0p2-beta=0p8}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
+	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-k=20} %-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
 	\centering

### 2019-11-20 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
R rec coeffs methods

@@ -57,7 +57,7 @@ function lanczos!(w, P, β, γ; N₀=0)
     end
 
     for k = N₀+1:N
-        # @show "lanczos", N, k
+        @show "lanczos", N, k
         v = x*P[2] - γ[k-1]*P[1]
         β[k] = sum(P[2]*w*v)
         v = v - β[k]*P[2]

### 2019-11-20 experiments/diskslice.jl

@@ -231,8 +231,8 @@ a, b, c = 1.0, 1.0, 1.0
 S = D(a, b, c)
 x, y = 0.4, -0.765; z = [x; y]
 @test isindomain(z, D)
-# Δw = laplaceoperator(S, S, 200; weighted=true)
-# save("experiments/saved/laplacian-w11-array-diskslice-alpha=$α-beta=$β.jld", "Lw11", sparse(Δw))
+Δw = laplaceoperator(S, S, 990; weighted=true)
+# save("experiments/saved/laplacian-w11-array-diskslice-alpha=$α-beta=$β-N=310.jld", "Lw11", sparse(Δw))
 # save("experiments/saved/biharmonic-array-diskslice-alpha=$α-beta=$β.jld", "bihar", sparse(bihar))
 bihar = biharmonicoperator(D(2.0, 2.0, 2.0), 200)
 S
@@ -402,10 +402,11 @@ S
 f1 = Fun((x,y)->1.0, S); 1.0
 f2 = Fun((x,y)->y^2 - 1, S); 1.0
 f3 = Fun((x,y)->weight(S, x, y)^3, S, 300); f3.coefficients
-# f4 = Fun((x,y)->exp(-1000((x-0.5)^2+(y-0.5)^2)), S, 2*getopindex(N,N)); f4.coefficients
+f4 = Fun((x,y)->exp(-1000((x-0.5)^2+(y-0.5)^2)), S, 2*getopindex(N,N)); f4.coefficients
 # f4cfs = f4.coefficients[1:getopindex(N,N)]
 # save("experiments/saved/solutionblocknorms-diskslice-alpha=$α-beta=$β-f4cfs.jld", "f4cfs", f4cfs)
 f4cfs = load("experiments/saved/solutionblocknorms-diskslice-alpha=$α-beta=$β-f4cfs.jld", "f4cfs")
+mm = length(f4cfs); resize!(f4cfs, sum(1:N+1)); f4cfs[mm+1:end] .= 0.0
 u1norms, u2norms, u3norms, u4norms = getsolutionblocknorms(N, S, Δw, resizecoeffs!(f1, N),
                 resizecoeffs!(f2, N), resizecoeffs!(f3, N), f4cfs)
 using Plots
@@ -415,20 +416,23 @@ Plots.plot!(u3norms, line=(3, :dashdot), label="f(x,y) = W{(1,1,1)}^3")
[…]

### 2019-11-20 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
fixed some typos

@@
 	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:Hrecurrence} \\
-	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y).
+	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:Precurrence}
 \end{align}
 Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0 \quad \forall n=0,1,2,\dots$.
-We can use these to determine the 2D recurrences for $\hdopnkabcd(x,y)$.
+We can use (\ref{eqn:Hrecurrence}) and (\ref{eqn:Precurrence}) to determine the 2D recurrences for $\hdopnkabcd(x,y)$.
 Importantly, we can deduce sparsity in the recurrence relationships:
@@
 \begin{lemma}
-$\hdopnkabc(x,y)$ satisfy the following 3-term recurrences:
-\begin{align*}
-x \hdopnkabcd(x,y) &= \alphaabc_{n,k,1} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabc_{n,k,2} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabc_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
-y \hdopnkabcd(x,y) &= \betaab_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaab_{n,k,2} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \betaab_{n,k,3} \: \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
-		& \quad \quad + \betaab_{n,k,4} \: \hdop_{n, k-1}^{(a,b,c,d)}(x, y) + \betaab_{n,k,5} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaab_{n,k,6} \: \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
-		& \quad \quad + \betaab_{n,k,7} \: \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaab_{n,k,8} \: \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaab_{n,k,9} \: \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
+$\hdopnkabcd(x,y)$ satisfy the following 3-term recurrences:
+\begin{align*}
+x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
+y \hdopnkabcd(x,y) &= \betaabcd_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,3} \: \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
+		& \quad \quad + \betaabcd_{n,k,4} \: \hdop_{n, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,5} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \: \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
+		& \quad \quad + \betaabcd_{n,k,7} \: \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \: \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,9} \: \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
 \end{align*}
@@
 For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopabcd(x,y)^\top \mathbf{f}$.
-Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a,b)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkabcd\}$ basis.
+Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b,c,d)\top}} \mathbf{f}$.
+In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkabcd\}$ basis.
 Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices:
@@
 
-For the disk-slice, we further note that $\betaabc_{n,k,2} = \betaabc_{n,k,5} = \betaabc_{n,k,8} \equiv 0$ for any $n, k$.
+For the disk-slice, $\betaabc_{n,k,2} = \betaabc_{n,k,5} = \betaabc_{n,k,8} \equiv 0$ for any $n, k$.
 
@@
 	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
-	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(b-1)}(y), \label{eqn:derivativeofweightjac} \\
+	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(c-1)}(y), \label{eqn:derivativeofweightjac} \\
 	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}

### 2019-11-21 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Start readthrough

@@
 Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
-In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on \bstodo{Reviewer 1, comment 3}disk slices and trapeziums, with a focus here on the half-disk case.
+In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on \bstodo{Reviewer 1, comment 3}disk slices and trapeziums.
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
@@
 Sparsity comes from expanding the domain and range of an operator  using different choices of $a$ and $b$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.
+Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules\sotodo{Mention Christoffel--Darboux here?} to determine the entries.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 \noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
+
+\sotodoinline{Mention Christoffell--Darboux here}
 

### 2019-11-21 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Make conditions references.

@@
 \newtheorem{definition}{Definition}
+\newtheorem{condition}{Condition}
 \newtheorem{algorithm}{Algorithm}
+
+\def\condref#1{Condition~\ref{cond:#1}}
 
@@
 \end{align*}
-where  either of the following scenarios hold:
-\begin{enumerate}
-\item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
-\end{enumerate}
+where  either of the following conditions hold:
+
+\begin{condition}\label{cond:trap}
+ \(\rho\) is a degree 1 polynomial.
+ \end{condition}
+
+ \begin{condition}\label{cond:disk}
+\(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
+\end{condition}
 
@@
 \begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
-Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:
-\begin{enumerate}
-\item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
-\end{enumerate}
+Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either \condref{trap} or \condref{disk} with \(w_2\) being an even function hold.
 $\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$.
@@
 \end{align*}
-assuming that either of the following hold:
-\begin{enumerate}
-\item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(\jacw^{(a,b)}\) is an even function (i.e.
-$a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$).
-\end{enumerate}
+assuming that either \condref{trap} or \condref{disk} with  \(\jacw^{(a,b)}\) being an even function (i.e.
+$a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$) hold.
 That is,
@@
 
-Finally, recall that either:
-\begin{enumerate}
-\item  \(\rho\) is a degree 1 polynomial, or 
-\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(\jacw^{(d,c)}\) is an even function.
-\end{enumerate}
-For case 1, we have that 
+Finally, if \condref{trap} holds we have that 
 $$
@@
 $$
-For case 2 we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.
+If \condref{disk} holds we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.
 

### 2020-01-06 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Further edits to paper, inc figs. Made some comments.

@@
 \newtheorem{algorithm}{Algorithm}
+\newtheorem{corollary}{Corollary}
 
@@
 Sparsity comes from expanding the domain and range of an operator using different choices of $a$ and $b$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle  and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules\sotodo{Mention Christoffel--Darboux here?} to determine the entries.
+Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 \noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
-
-\sotodoinline{Mention Christoffell--Darboux here}
+We present a method involving use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.
 
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityoflaplacian-w11} %-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofhelmholtz} %-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofbiharmonic} %-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
         \centering
@@
 
-It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \bstodo{Link!!!}.
+It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \cite[18.2.12]{DLMF}.
 We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case.
@@
 \end{align*}
-and define \bstodo{check signs}
+and define
 \begin{align}
@@
 \end{align}
-Now, by the Christoffel-Darboux formula \bstodo{cite}, we have that for any $x, y \in \R$,
+Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
 \begin{align}
@@
 
+\begin{corollary}
+The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c+1,d)}\}$ can be written as:
+\begin{align}
+	\tilde{\alpha}_n^{(a,b,c+1,d)} &= \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(1)} - \frac{\tilde{\beta}_{n}^{(a,b,c,d)}}{\chi_{n}^{(a,b,c,d)}(1)} + \tilde{\alpha}_{n}^{(a,b,c,d)}, \\
+	\tilde{\beta}_n^{(a,b,c+1,d)} &= \fpr(\frac{1 - \tilde{\alpha}_{n+1}^{(a,b,c,d)} - \frac{\tilde{\beta}_n^{(a,b,c,d)}}{\chi_n^{(a,b,c,d)}(1)}}{1 - \tilde{\alpha}_{n}^{(a,b,c,d)} - \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(1)}})^\half \tilde{\beta}_n^{(a,b,c,d)}.
+\end{align}
+The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d+1)}\}$ can be written as:
+\begin{align}
+	\tilde{\alpha}_n^{(a,b,c,d+1)} &= \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(-1)} - \frac{\tilde{\beta}_{n}^{(a,b,c,d)}}{\chi_{n}^{(a,b,c,d)}(-1)} + \tilde{\alpha}_{n}^{(a,b,c,d)}, \\
+	\tilde{\beta}_n^{(a,b,c,d+1)} &= \fpr(\frac{- 1 + \tilde{\alpha}_{n+1}^{(a,b,c,d)} + \frac{\tilde{\beta}_n^{(a,b,c,d)}}{\chi_n^{(a,b,c,d)}(-1)}}{- 1 + \tilde{\alpha}_{n}^{(a,b,c,d)} + \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(-1)}})^\half \tilde{\beta}_n^{(a,b,c,d)}.
+\end{align}
+where 
+\begin{align}
+	\chi_{n}^{(a,b,c,d)}(y) &:= \frac{\genjact_{n+1}^{(a,b,c,d)}(y)}{\genjact_{n}^{(a,b,c,d)}(y)} \\
+	&= \frac{1}{\tilde{\beta}_{n}^{(a,b,c,d)}} \: \fpr(y - \tilde{\alpha}_{n}^{(a,b,c,d)} - \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(y)}), \quad y \in \{-1,1\}.
+\end{align}
+\end{corollary}
+
 These two propositions allow us to recursively obtain the recurrence coefficients for the OPs $\{\genjac_{n-k}^{(a,b,2c+2k + 1)}\}$ as $k$ increases to be large.
 
-\noindent{\bf Remark}:  \bstodo{improve this remark}
-This is the most expensive part of the current calculation, in particular the initialisation required (i.e.
+\noindent{\bf Remark}:
+The Corollary demonstrates that in order to obtain the recurrence coefficients $\{\alpha_{m}^{(a,b,2c+2k+1)}\}$, $\{\beta_{m}^{(a,b,2c+2k+1)}\}$ for some $m$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{m+2}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{m+2}^{(a,b,2c+2(k-1)+1)}\}$.
+Thus, for large $N$, this recursive method of obtaining the recurrence coefficients requires a large initialisation (i.e.
 using the Lanczos algorithm to compute the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+1)}\}$, $\{\beta_{n}^{(a,b,2c+1)}\}$ -- however, we only need to compute these once, and can store and save this initialisation to disk once computed, for the given values of $a, b, c$).
-Note also that in order to obtain the recurrence coefficients $\{\alpha_{n-k}^{(a,b,2c+2k+1)}\}$, $\{\beta_{n-k}^{(a,b,2c+2k+1)}\}$ for $n = 0,\dots,N$ for some $N$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{n}^{(a,b,2c+2(k-1)+1)}\}$ for $n = 0,\dots,N+3$.
-Thus, for large $k$, this recursive method of obtaining the recurrence coefficients requires a large initialisation.
 %This is the most expensive part of the current calculation, but note that we can reuse this computation for multiple partial differential equations.
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson}
+	\includegraphics[scale=0.3]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.5]{solutionblocknorms-poisson} %-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\centering
-	\includegraphics[scale=0.3]{solution-helmholtz-k=100-n=1500}
+	\includegraphics[scale=0.3]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
 	%\label{fig:solution-helmholtz}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-k=20} %-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
+	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.3]{solution-biharmonic}
+	\includegraphics[scale=0.3]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
 	%\label{fig:solution-biharmonic}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic}
+	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic-diskslice-alpha=0p2-beta=0p8-N=200}
 	\centering
@@
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
-For the final Gaussian bump example, we see that we achieve spectral convergence.
+For the final Gaussian bump example, we see that we also achieve algebraic convergence.
+\bstodoinline{Note that it is no longer correct to say that we achieve spectral convergence for ex 4 for the disk slice.
+Is there something more we can say?
+Same goes for the other examples.}
 
@@
 We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
-\bstodo{Reviewer 1, comment 3} This work is related to the triangle case \cite{olver2019triangle}, where ????.
-This work forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements.
+\bstodo{Reviewer 1, comment 3} This work is related to the triangle case \cite{olver2019triangle} \bstodoinline{add more here?}, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements.
 %A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
@@
 \section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
+
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
+        	%\label{fig:solutionblocknorms-poisson}
+	\centering
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-poisson}
+	\centering
+	%\label{fig:solution-poisson}
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
+	%\label{fig:solution-poisson}
+	\centering
+	\end{subfigure}
+	\centering
+	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
+Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
+Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
+in the trapezium.}
+	\centering
+	\label{fig:appendixfigs}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.5]{solutionblocknorms-poisson}
-        	%\label{fig:solutionblocknorms-poisson}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
-This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
-	\centering
-	\label{fig:poisson}
-\end{figure}
-
 The work in this paper on the disk-slice can be easily transferred to the special-case domain of the end-disk-slice by which we mean
@@
 For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for an end-disk-slice.
+
+In Figure \ref{fig:appendixfigs} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
 
@@
 \end{align*}
+
+In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
 

### 2020-01-06 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
fixes - for TrapFam, BigFloats used now. For DSFam, BBM mult wasnt working now, so using sparse when multiplying in laplaceoperator methods.

@@ -438,9 +438,10 @@ function getBs!(S::DiskSliceSpace{<:Any, <:Any, T, <:Any}, N, N₀) where T
         m += 1
     end
     for n = N+1:-1:m
-        if n > 200
-            @show "getsBs!", N, n
-        end
+        # if n > 200
+        #     @show "getsBs!", N, n
+        # end
+        @show "getsBs!", N, n
         v1 = [recα(T, S, n, k, 2) for k = 0:n]
         v2 = [recβ(T, S, n, k, 4) for k = 0:n-1]
         v3 = [recβ(T, S, n, k, 3) for k = 1:n]
@@ -1365,7 +1366,16 @@ function laplaceoperator(S::DiskSliceSpace{<:Any, <:Any, T, <:Any},
         @show "laplaceoperator", "5 of 6 done"
         G = weightedpartialoperatory(S, N)
         @show "laplaceoperator", "6 of 6 done"
-        L = A * B + C * E * F * G
+        # NOTE: Multiplying the BlockBandedMatrices fails with an error, so
+        #       convert to sparse
+        AAl, AAu = A.l + B.l, A.u + B.u
+        BBl, BBu = C.l + E.l + F.l + G.l, C.u + E.u + F.u + G.u
+        AAλ, AAμ = A.λ + B.λ, A.μ + B.μ
+        BBλ, BBμ = C.λ + E.λ + F.λ + G.λ, C.μ + E.μ + F.μ + G.μ
+        AA = sparse(A) * sparse(B)
+        BB = sparse(C) * sparse(E) * sparse(F) * sparse(G)
+        L = BandedBlockBandedMatrix(AA + BB, (1:nblocks(A)[1], 1:nblocks(B)[2]),
+                                    (max(AAl,BBl),max(AAu,BBu)), (max(AAλ,BBλ),max(AAμ,BBμ)))
         if square
             m = sum(1:(N+1))
             Δ = BandedBlockBandedMatrix(L[1:m, 1:m], (1:N+1, 1:N+1), (L.l, L.u), (L.λ, L.μ))
@@ -1386,10 +1396,16 @@ function laplaceoperator(S::DiskSliceSpace{<:Any, <:Any, T, <:Any},
         @show "laplaceoperator", "5 of 6 done"
         G = partialoperatory(S, N+2)
         @show "laplaceoperator", "6 of 6 done"
-        AA = A * B
[…]

### 2020-01-06 OrthogonalPolynomialFamilies/src/TrapeziumFamilies.jl

@@ -270,7 +270,7 @@ end
 #===#
 # Jacobi operator entries
 
-function recα(S::TrapeziumSpace{<:Any, <:Any, T, <:Any}, n, k, j) where T
+function recα(::Type{T}, S::TrapeziumSpace, n, k, j) where T
     R = getRspace(S, k)
     if j == 1
         recγ(T, R, n-k+1)
@@ -280,7 +280,7 @@ function recα(S::TrapeziumSpace{<:Any, <:Any, T, <:Any}, n, k, j) where T
         error("Invalid entry to function")
     end
 end
-function recβ(S::TrapeziumSpace{<:Any, <:Any, T, <:Any}, n, k, j) where T
+function recβ(::Type{T}, S::TrapeziumSpace, n, k, j) where T
     # j ∈ {1,...,9}
 
     # We get the norms of the 2D OPs
@@ -314,90 +314,91 @@ function recβ(S::TrapeziumSpace{<:Any, <:Any, T, <:Any}, n, k, j) where T
 
     if j == 1
         getopptseval(R1, n-k+1, pts)
-        (inner2(R2, opevalatpts(R2, n-k+1, pts), opevalatpts(R1, n-k+1, pts), w)
+        T(inner2(R2, opevalatpts(R2, n-k+1, pts), opevalatpts(R1, n-k+1, pts), w)
             * δ / S.opnorms[k])
     elseif j == 3
         getopptseval(R3, n-k-1, pts)
-        (inner2(R3, opevalatpts(R2, n-k+1, pts), opevalatpts(R3, n-k-1, pts), w)
+        T(inner2(R3, opevalatpts(R2, n-k+1, pts), opevalatpts(R3, n-k-1, pts), w)
             * δ / S.opnorms[k+2])
     elseif j == 4
         getopptseval(R1, n-k+2, pts)
-        (inner2(R2, opevalatpts(R2, n-k+1, pts), opevalatpts(R1, n-k+2, pts), w)
+        T(inner2(R2, opevalatpts(R2, n-k+1, pts), opevalatpts(R1, n-k+2, pts), w)
             * δ / S.opnorms[k])
     elseif j == 6
         getopptseval(R3, n-k, pts)
-        (inner2(R3, opevalatpts(R2, n-k+1, pts), opevalatpts(R3, n-k, pts), w)
+        T(inner2(R3, opevalatpts(R2, n-k+1, pts), opevalatpts(R3, n-k, pts), w)
             * δ / S.opnorms[k+2])
[…]

### 2020-01-07 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Added review comments and some other changes. New example figs.

@@
 \end{figure}
+solution-exact-poisson-diskslice-alpha=0.2-beta=0.8-u=y3expx-actual
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
+	\centering
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
+	\centering
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
+	\centering
+	\end{subfigure}
+	\centering
+	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$.
+Left: Computed.
+Centre: Exact.
+Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
+	\centering
+	\label{fig:poissonexact}
+\end{figure}
 
@@
 Same goes for the other examples.}
+
+In Figure \ref{fig:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error.
+The example was chosen so that the exact solution was $u(x,y) = \Wiii (x,y) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y) = \Delta \big[\Wiii (x,y) y^3 \exp(x)\big]$.
+We see that the computed solution is almost exact.
 

### 2020-01-07 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Added some more notes to refer to the refs comments

@@
 
+\bstodo{Reviewer 1, comments 1,2}
 It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \cite[18.2.12]{DLMF}.
@@
 We see that the computed solution is almost exact.
+\bstodo{Reviewer 2, comment 5}
 
@@
 
-In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
+In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$. \bstodo{Reviewer 2, comment 6}
 

### 2020-01-08 papers/lsr/lsr.tex
LSR draft (should be about 30 pages, hence ive removed the proofs and other bits)

--- 2020-01-06 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
@@
 
-%\bibliographystyle{apalike}
-
-\graphicspath{ {images/} }
-
 \newcommand{\sotodo}{\todo[color=green]}
@@
 
+\graphicspath{ {images/} }
+
 %SetFonts
@@
 
-\usepackage[square,numbers]{natbib}
+\usepackage{natbib}
 \usepackage{url}
@@
 
-%\usepackage{media9}
-%\graphicspath{ {../sphere/plots} }
-
-\title{Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums}
-\author{Sheehan Olver, Ben Snowball}
-%\date{}							% Activate to display a given date or no date
+\title{Late Stage Review}
+\author{Ben Snowball (Principle Supervisor: Sheehan Olver)}
+\date{}							% Activate to display a given date or no date
 
@@
 
-\begin{abstract}
+\section{Introduction}
+
+The goal of my research is to eventually investigate sparse spectral methods on triangulations of the sphere as an alternative to the spherical harmonics approach currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) in their weather and climate model \cite{wedi2013fast}.
+The aim would be to preserve the predictive skill of the current approach whilst avoiding the parallel scalability bottleneck from the global spectral transform, which is expected to inhibit the future performance of the ECMWF model.
+
+This document is to provide a brief outline of my work to date, with a closer look at our most recent area of research (solving PDEs on the half disk using sparse spectral methods).
+The work outlined in this report serves as initial steps to pave the way to solving PDEs on spherical triangles.
+All code used thus far and from hence forth will mainly be written in the Julia programming language.
+
+\section{Professional Skills Development Requirement}
+I have met the initial part of the professional skills development requirement which is to attend a minimum of 2 further professional development workshops by the Late Stage Review (LSR).
+The 3 further courses taken are:
+\begin{itemize}
+	\item Maximising Management Skills 1: Becoming an Effective Researcher
+	\item Maximising Management Skills 2: Time Management Strategies for your PhD
+	\item Maximising Management Skills 4: Put Project Management into Action
+\end{itemize}
+The 100 hours of PG courses/activities have been met by virtue of the courses taken in the first year of my studies with the Mathematics of Planet Earth CDT.
+
+\section{Orthogonal polynomials on the sphere (spherical harmonics)}
+
+The first step was to develop a small code library so that we can use spherical harmonics as orthogonal polynomials in three variables $x, y, z$  to evaluate functions on the sphere.
+
+We can the define the spherical harmonics, orthogonal on the unit sphere as \cite{efthimiou2014spherical, dai2013approximation}, \cite[14.30.1, 14.3.1, 14.3.6]{DLMF}:
+\begin{align}
+\Ylmfull &:= \clm \eimphi \Plm (\costheta) \\
+&= \alphalm (1 - (\costheta)^2)^\frac{|m|}{2} \eimphi P^{(|m|,|m|)}_{l-|m|}(\costheta), \quad \text{where } 0 \le |m| \le l, \, l \in \No.
+\end{align} 
+Note that we can then express \(\Ylm\) in terms of \(x,y,z\) instead of \(\theta, \varphi\) by noting that \(\costheta = z\) and that \( \eimphi\) can be expressed in terms of \(x,y,z\) for any \(m\in\mathbb{Z}\).
+Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$.
+They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.
+While this is equivalent to standard techniques in the litrature concerning spherical harmonics, we approach this with our goal in mind of developing sparse spectral methods on spherical triangles.
+Thus, we try to use language and formulations that will translate as we move through our steps towards that goal.
+
+\citep{dunkl2014orthogonal} (p75-85) provided a framework for gaining a recurrence relation for each \(\bigP_{l+1}\) in terms of the previous two sub-vectors \(\Pl\) and \(\bigP_{l-1}\), utilising the sub matrices that make up the block-tridiagnonal Jacobi operator matrices for multiplication by $x, y$ and $z$.
+Clenshaw's algorithm then gives us a way of evaluating a function given its expansion in the spherical harmonic basis (up to a finite order).
+This provides an optimal complexity method for evaluating spherical harmonic expansions; an alternative to using the explicit expressions that may have benefits, though further investigation is required.
+
+%\subsection{Deriving expressions for the multiplication by \(x, y, z\) of the spherical harmonics}
+%
+%We start by expressing \(x\,\Ylm(x,y,z)\), \(y\,\Ylm(x,y,z)\) , and \(z\,\Ylm(x,y,z)\)  in terms of \(Y^{m'}_{l'}(x,y,z)\) for any point \((x,y,z)\) on the unit circle.
+%
+%For \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\):
+%\begin{align}
+%x\,\Ylm(x,y,z) &= \Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \nonumber \\
+%		     & \quad \quad \quad + \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z), \\
+%y\,\Ylm(x,y,z) &= - i \, \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \nonumber \\
+%		     &\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
+%z\,\Ylm(x,y,z) &= \Flm Y^{m}_{l+1}(x,y,z) + \Glm Y^{m}_{l-1}(x,y,z) ,
+%\end{align}
+%where \(\Alm, \Blm, \Dlm, \Elm, \Flm, \Glm \in \R\) have explicit derivable and calculable formulae.
+%
+%\subsection{Jacobi matrices}
+%
+%Define, for \(l \in \No\), \(\Pl\) as the column vector of the degree \(l\) spherical harmonic polynomials, and \(\bigP\) as the stacked block vector of the \(\Pl\)'s; that is
+%\begin{align}
+%\Pl := \begin{bmatrix}
+%		Y^{-l}_l \\
+%		\vdots \\
+%		Y^l_l
+%	\end{bmatrix} \in \C^{2l+1}, 
+%\quad \quad 
+%\bigP := \begin{bmatrix}
+%		\bigP_0 \\
+%		\hline
+%		\bigP_1 \\
+%		\hline
+%		\bigP_2 \\
+%		\vdots \\
+%	\end{bmatrix}.
+%\end{align}
+%
+%Define the (Jacobi) matrices \(J^x, J^y, J^z\) by 
+%\begin{align}
+%J^x \bigP = x \bigP, \quad J^y \bigP = y \bigP, \quad J^z \bigP = z \bigP.
+%\end{align}
+%Then, using equations (3--5), we have that the Jacobi matrices each have the following block-tridiagonal form, allowing for easy application computationally:
+%\begin{align}
+%J^{x/y/z} &= \begin{bmatrix}
+%		B^{x/y/z}_0 & A^{x/y/z}_0 & & & & \\
+%		C^{x/y/z}_1 & B^{x/y/z}_1 & A^{x/y/z}_1 & & & \\
+%		& C^{x/y/z}_2 & B^{x/y/z}_2 & A^{x/y/z}_2  & & & \\
+%		& & C^{x/y/z}_3 & \ddots & \ddots & \\
+%		& & & \ddots & \ddots & \ddots \\
+%	\end{bmatrix}
+%\end{align}
+%
+%\subsection{Three-term recurrence relation for \(\bigP\)}
+%
+%Combining each system in (7) we can write the tridiagonal-block system
+%\begin{align}
+%\renewcommand\arraystretch{1.3}
+%\mleft[
+%\begin{array}{cccc}
+%		1  & & & \\
+%		B_0-G_0(x,y,z) & A_0 & & \\
+%		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
+%		& C_2 & \ddots & \ddots \\
+%		& & \ddots &
+%\end{array}
+%\mright]
+%\bigP
+%=
+%\begin{bmatrix}
+%	\alpha^0_0 \\ 0 \\ 0 \\ \vdots \\ \\
+%\end{bmatrix}.
+%\end{align}
+%where for each \(l \in \No\):
+%\begin{align}
+%A_l &:= \begin{bmatrix}
+%		A^x_l \\
+%		A^y_l \\
+%		A^z_l
+%	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+3)}, \quad
+%C_l := \begin{bmatrix}
+%		C^x_l \\
+%		C^y_l \\
+%		C^z_l
+%	    \end{bmatrix} \in \R^{3(2l+1)\times(2l-1)} \quad (l \ne 0), \\
+%B_l &:= \begin{bmatrix}
+%		B^x_l \\
+%		B^y_l \\
+%		B^z_l
+%	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}, \quad
+%G_l(x,y,z) := \begin{bmatrix}
+%		xI_{2l+1} \\
+%		yI_{2l+1} \\
+%		zI_{2l+1}
+%	    \end{bmatrix} \in \R^{3(2l+1)\times(2l+1)}.
+%\end{align}
+%
+%For each \(l \in \No\) let \(\Dlt\) be any matrix that is a left inverse of \(A_l\), i.e.
+such that \(\Dlt A_l = I_{2l+3}\).
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dlt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} (p78), which can be expanded to obtain the recurrence:
+%\begin{align}
+%\begin{cases}
+%\bigP_{-1} := 0 \\
+%\bigP_{0} := \alpha^0_0 \\
+%\bigP_{l+1} = -\Dlt (B_l-G_l) \Pl - \Dlt C_l  \,\bigP_{l-1}, \quad l \in \N.
+%\end{cases}
+%\end{align}
+%
+%\subsection{Evaluation of a scalar function on the sphere}
+%We can use a multivariate analogue of the Clenshaw algorithm to evaluate a function at a given point \((x,y,z)\) on the unit sphere provided we know the coefficients of the function when expanded in the spherical harmonic basis, i.e.
+suppose \(f(x,y,z)\) is a function and we know the set \(\{\bold{f}_l\}\) s.t.
+%\begin{align}
+%f(x,y,z) \approx \sum^N_{l=0} \bold{f}_l^T \, \Pl (x,y,z), \quad \text{where } \Pl (x,y,z), \bold{f}_l \in \R^{2l+1} \text{ for each } l \in \{0,\dots,N\}.
+%\end{align}
+%
+%The Clenshaw algorithm is then as follows:
+%\begin{align*}
+%\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
+%\quad &\text{2) } \text{For } n = N:-1:1 \\
+%\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = \bold{f}_n^T - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
+%\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bigP_0(x,y,z) f_0 + \bm{\gamma}_{1}^T \bigP_1(x,y,z) - \bigP_0(x,y,z) \bm{\gamma}_{2}^T D^T_1 C_1.
+%\end{align*}
+%where $\bigP_0 \equiv \alpha^0_0$.
+%
+%We can use the Clenshaw algorithm to output a matrix operator for a function, given we know the coefficients of its expansion in the SH basis. We simply make the adjustment that
+%\begin{align*}
+%\quad &\text{1) } \text{Set } \bm{\gamma}_{N+2} = \bold{0}, \: \bm{\gamma}_{N+2} = \bold{0}. \\
+%\quad &\text{2) } \text{For } n = N:-1:1 \\
+%\quad & \quad \quad \quad \text{set } \bm{\gamma}_{n}^T = (\bold{f}_n^T)I - \bm{\gamma}_{n+1}^T D^T_n (B_n - G_n(J^x,J^y,J^z)) -  \bm{\gamma}_{n+2}^T D^T_{n+1}C_{n+1} \\
+%\quad &\text{3) } \text{Output: } f(J^x,J^y,J^z) \approx F := \alpha^0_0 f_0 + \bm{\gamma}_{1}^T \bigP_1(J^x,J^y,J^z) - \alpha^0_0 \bm{\gamma}_{2}^T D^T_1 C_1.
+%\end{align*}
+%where $\bigP_1(J^x,J^y,J^z)$ is a vector of matrices, $G_n(J^x,J^y,J^z)$ is a matrix of matrices, and so each $\bm{\gamma}_n$ is a vector of matrices.
+
+\subsection{Tangent space}
+
+By noting that the tangent space is then spanned by $\{\gradYlm, \gradpYlm\}$, and that these gradients and perpendicular gradients are orthogonal polynomials themselves, we can use the same ideas to create a framework for functions in the tangent space as expansions in this OP basis.
+Note that the perpendicular gradient is related to the regular surface gradient by
+\begin{align}
+\gradpYlm(x,y,z) = \unitvec \times \gradYlm(x,y,z),
+\end{align}
+where \(\unitvec\) is the unit vector normal to the surface of the sphere at the point \((x,y,z)\), i.e. as we are looking at the unit sphere, \(\unitvec\) is simply given by
+\begin{align}
+\unitvec = \begin{bmatrix} x \\ y \\ z \end{bmatrix}.
+\end{align}
+For any scalar spherical harmonic (SSH) $\Ylm$, the triad of $\gradYlm, \: \unitvec \times \gradYlm, \: \unitvec \Ylm$ can be referred to as vector spherical harmonics \cite{barrera1985vector}.
+
+Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient  and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics are a basis for the tangent space, namely \(\gradYlm\), \(\gradpYlm\).
+
+Importantly,  \(\gradYlm\), \(\gradpYlm\) are in fact vector-valued orthogonal polynomials, that span all vector-valued polynomials modulo the vanishing ideal $\{ \bm{p} \in (\Pi_3)^3  : \unitvec \cdot \bm{p} = 0 \}$.
+A simple calculation shows that such orthogonal polynomials must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal.
+In the following, we determine the coefficients of these Jacobi operators, as well as other important operators needed.
+
+\subsubsection{Spin-1 tensor spherical harmonics}
+
+The gradient and perpendicular gradient of a spherical harmonic \(\Ylm(x,y,z)\) can be expressed in terms of spin-1 tensor spherical harmonics, which in turn can each be expressed as a vector-weighted sum of spherical harmonics.
+
+For any \(l \in \No\), \(m \in \Z\) s.t. \(0 \le |m| \le l\) we have that
+\begin{align}
+\nabla \Ylm &= \blmone \, \curlyy^{l-1,1}_{l,m} + \blone \curlyy^{l+1,1}_{l,m}, \\
+\nabla^\perp \Ylm &= \blzero \, \curlyy^{l,1}_{l,m},
+\end{align}
+where
+\begin{align}
+\blmone &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
+\blzero := i \, \big(l(l+1)\big)^\half, \quad
+\blone := l \, \Big(\frac{l+1}{2l+1}\Big)^\half.
+\end{align}
+
+We can then gain similar (vector analogue) derivations for recurrences, Jacobi matrices and the Clenshaw algorithm implementation for the tangent space vector-valued orthogonal polynomial basis.
+
+The aim is to, given a PDE say, work with the coefficients of the functions involved (coefficients of their expansion in the OP bases), reframe the PDE in terms of the OP coefficients utilising a suitable timestepping method.
+For derivative, gradient, divergence etc.
+operators, we can formulate matrix operators to act on the coefficients, thus making our system to simulate a simple matrix vector system.
+
+\subsection{Example: linear shallow water equations}
+Let $\mathbf{u}(x,y,z)$ be the tangential velocity of the flow and $h(x,y,z)$ be the height of the water from some constant reference height $\mathcal{H}$. The linear SWEs are
+\begin{align}
+\begin{cases}
+\frac{\partial \mathbf{u}}{\partial t} + f \unitvec \times \mathbf{u} - \nabla h = \mathbf{0} \\
+\frac{\partial h}{\partial t} + \mathcal{H} \nabla \cdot \mathbf{u} = 0
+\end{cases}
+\end{align}
+where $f = 2 \Omega \cos(\theta) =  2 \Omega z$ is the Coriolis parameter (note that the operator on the coefficients vector for multiplication by this parameter would be $F = 2 \Omega (J^z_\nabla)^T$).
+
+Using backward Euler:
+\begin{align}
+\mathbf{u}_{n+1} &= \mathbf{u}_{n} + \Delta t \: (\nabla h_{n+1} - f \unitvec \times \mathbf{u}_{n+1}) \\
+h_{n+1} &= h_n - \Delta t \: \mathcal{H} \nabla \cdot \mathbf{u}_{n+1}
+\end{align}
+
+Then using the operators on the coefficient vectors, the timestepping method is: 
+\begin{align}
+\mathbf{u}^c_{n+1} &= \mathbf{u}^c_{n} + \Delta t \: (G \mathbf{h}^c_{n+1} - F K \mathbf{u}^c_{n+1}) \\
+\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
+\end{align}
+$\iff$
+\begin{align}
+\mathbf{u}^c_{n+1} &= (I + \Delta t F K + \Delta t^2 \mathcal{H} G D)^{-1} \: (\mathbf{u}^c_{n} + \Delta t \: G \mathbf{h}^c_n) \\
+\mathbf{h}^c_{n+1} &= \mathbf{h}^c_n - \Delta t \: \mathcal{H} D \mathbf{u}^c_{n+1}
+\end{align}
+where $F, K, G, D$ are all operators corresponding to multiplying by the Coriolis parameter, outward normal vector cross product, gradient and divergence operations respectively.
+It is important to note that these can be easily constructed due to the relationships between the vector spherical harmonics and scalar spherical harmonics we have outlined, and have a sparse Block-Banded structure.
+
+These ideas will also be used to gain similar relationships for the circle arc, half disk, and spherical triangle orthogonal polynomials.
+
+\section{Orthogonal polynomials on the circle arc}
+
+As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
+We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to \cite{huybrechs2010fourier}:
+\begin{align}
+\bigP_0(x,y) := T^h_0(x) \equiv T_0,
+\quad \quad 
+\bigP_n(x,y) := \begin{bmatrix}
+		T^h_n(x) \\
+		y \: U^h_{n-1}(x)
+	\end{bmatrix} \in \R^2 \quad \forall n \in \N,
+\quad \quad 
+\bigP := \begin{bmatrix}
+		\bigP_0 \\
+		\hline
+		\bigP_1 \\
+		\hline
+		\bigP_2 \\
+		\vdots \\
+	\end{bmatrix}.
+\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Orthogonal Polynomials on the disk-slice (2D)}
+
 Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
-In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on \bstodo{Reviewer 1, comment 3}disk slices and trapeziums.
+In the work in this section we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices and trapeziums.
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
 We apply the framework to solving the Poisson, variable coefficient Helmholtz, and Biharmonic equations.
-\bstodo{Reviewer 1, comment 3}In this paper we focus on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
- \end{abstract}
-
-%
-\section{Introduction}
-
-This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
+For now, we have focused on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
+
+\subsection{Explicit construction}
+
+Here, we have developed sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.
 More precisely, we consider the solution of partial differential equations on the domain
-\begin{align*}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
+\begin{align}
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\} \label{eqn:domaindef}
+\end{align}
 where either of the following conditions hold:
@@
 
-For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss \bstodo{typo 1}an extension to other geometries in the appendix (including the half-disk and trapeziums).
-
-We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.
-On the disk-slice the family of weights we consider are of the form
-$$
-\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \qfor \alpha \leq x \leq \beta,\quad -\rho(x) \leq y \leq \rho(x).
-$$
-The corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
-$$
-\hdopnkab(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
-$$
-where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.
-The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
-
-Sparsity comes from expanding the domain and range of an operator using different choices of $a$ and $b$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
-In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
-This compares favourably to $O(p^6)$ operations if one proceeds na\"ively.
-Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
-
-%An alternative approach is to use \sotodoinline{Compare with "Discontinuous Galerkin scheme for the spherical shallow water equations with applications to tsunami modeling and prediction"}
-
-Here is an overview of the paper:
-
-\noindent \secref{OPs}: We present our procedure to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the disk-slice domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.
-
-\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
-We present a method involving use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.
-
-\noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice.
-We present a method for explicitly deriving the recurrence coefficients of the non-classical 1D OPs we detail in \secref{OPs}.
-We derive a quadrature rule on the disk-slice that can be used to expand a function in the OP basis up to a given order.
-Further, we implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
-
-\noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the disk-slice.
-
-\noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods.
-This lays the groundwork for a future $hp$-finite element method in a disk, where the elements capture the circular geometry precisely.
-
-\noindent\appref{HalfDisk}: We discuss extension to the special case of the end-disk-slice.
-
-\noindent\appref{Trapezium}: We discuss extension to trapezia.
-
-\section{Orthogonal polynomials on the disk-slice and the trapezium}\label{Section:OPs}
-
-In this section we outline the construction and some basic properties of $\hdopnkabcd(x,y)$.
-The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.
-
-\subsection{Explicit construction}
+For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss an extension to other the end-disk-slice in the appendix (note we omit detailed discussion  on the extension to trapezium in this document).
 
@@
 
-We can express the three-term recurrences associated \bstodo{typo 3}with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$ as 
+We can express the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$ as 
 \begin{align}
@@
 %
-\section{Sparse partial differential operators}\label{Section:PDOs}
+\subsection{Sparse partial differential operators}\label{Section:PDOs}
 
@@
 
-\begin{proof}
-First, note that:
-\begin{align}
-	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
-	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(c-1)}(y), \label{eqn:derivativeofweightjac} \\
-	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
-\end{align}
-
-We proceed with the case for the operator $D_y^{(a,b,c)}$ for partial differentiation by $y$.
+%\begin{proof}
+%First, note that:
+%\begin{align}
+%	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
+%	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(c-1)}(y), \label{eqn:derivativeofweightjac} \\
+%	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
+%\end{align}
+%
+%We proceed with the case for the operator $D_y^{(a,b,c)}$ for partial differentiation by $y$.
 Since $\{\hdop^{(a,b,c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^y \: \hdop^{(a, b, c+1)}_{m,j}$.
@@
 First, note that
-\begin{align*}
-	c_{m,j}^y = {\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
-\end{align*}
-Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
-\begin{align*}
-	&\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
-	&= \iint_\Omega \: \Big[ \genjacnmk^{(a,b,2c+2k+1)}(x) \: \rho(x)^{k-1} \: \jac_k^{(c,c) \: \prime}\fpr(\frac{y}{\rho(x)}) \\
-	& \quad \quad \quad \quad \quad \quad \cdot \: \genjacmmj^{(a,b,2c+2j+3)}(x) \: \rho(x)^{j} \: \jac_j^{(c+1,c+1)}\fpr(\frac{y}{\rho(x)}) \Big] \: \D y \: \D x \\
-	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}
-\end{align*}
-Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
-\begin{align*}
-	\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} &= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
-	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddy [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
-	&= - \int_{-1}^1 \: \jac_k^{(c,c)} \: [ \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: y \: \jac_{j}^{(c+1,c+1)} \:\jacw^{(c)} ] \: \D y \\
-	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}}
-\end{align*}
-which is zero for $j < k-1$ by orthogonality. Further, when $j = k-1$, we have that
-\begin{align*}
-	&\normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacmmj^{(a,b,2c+2k+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-	&= \normgenjac^{(a, b, 2c+2k+1)} \: \delta_{n,m+1},
-\end{align*}
-showing that the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally,
-\begin{align*}
-	c_{n-1,k-1}^y &= \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{k-1}^{(c+1,c+1)} >_{\jacw^{(c+1)}}.
-\end{align*}
-
-We next consider the case for the operator $D_x^{(a,b,c)}$ for partial differentiation by $x$.
+%\begin{align*}
+%	c_{m,j}^y = {\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
+%\end{align*}
+%Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
+%\begin{align*}
+%	&\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
+%	&= \iint_\Omega \: \Big[ \genjacnmk^{(a,b,2c+2k+1)}(x) \: \rho(x)^{k-1} \: \jac_k^{(c,c) \: \prime}\fpr(\frac{y}{\rho(x)}) \\
+%	& \quad \quad \quad \quad \quad \quad \cdot \: \genjacmmj^{(a,b,2c+2j+3)}(x) \: \rho(x)^{j} \: \jac_j^{(c+1,c+1)}\fpr(\frac{y}{\rho(x)}) \Big] \: \D y \: \D x \\
+%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
+%	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}
+%\end{align*}
+%Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
+%\begin{align*}
+%	\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} &= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
+%	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddy [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
+%	&= - \int_{-1}^1 \: \jac_k^{(c,c)} \: [ \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: y \: \jac_{j}^{(c+1,c+1)} \:\jacw^{(c)} ] \: \D y \\
+%	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}}
+%\end{align*}
+%which is zero for $j < k-1$ by orthogonality. Further, when $j = k-1$, we have that
+%\begin{align*}
+%	&\normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
+%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacmmj^{(a,b,2c+2k+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
+%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \delta_{n,m+1},
+%\end{align*}
+%showing that the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally,
+%\begin{align*}
+%	c_{n-1,k-1}^y &= \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{k-1}^{(c+1,c+1)} >_{\jacw^{(c+1)}}.
+%\end{align*}
+%
+%We next consider the case for the operator $D_x^{(a,b,c)}$ for partial differentiation by $x$.
 Since $\{\hdop^{(a+1, b+1, c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^x \: \hdop^{(a+1, b+1, c+1)}_{m,j}$.
@@
 First, note that
-\begin{align*}
-	c_{m,j}^x = {\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
-\end{align*}
-Now, again using the change of variable $t= \frac{y}{\rho(x)}$, we have that
-\begin{align}
-	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
-	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-	&\quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\ 
-	& \quad + k \: \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\
-	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:dxsparsity}
-\end{align}
-We will first show that the second factor of each term in (\ref{eqn:dxsparsity}) are zero for $j < k - 2$ and also for $j = k - 1$.
+%\begin{align*}
+%	c_{m,j}^x = {\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
+%\end{align*}
+%Now, again using the change of variable $t= \frac{y}{\rho(x)}$, we have that
+%\begin{align}
+%	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
+%	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
+%	&\quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\ 
+%	& \quad + k \: \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
+%	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\
+%	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
+%	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:dxsparsity}
+%\end{align}
+%We will first show that the second factor of each term in (\ref{eqn:dxsparsity}) are zero for $j < k - 2$ and also for $j = k - 1$.
 To this end, observe that, for any integer $c$, $\jac^{(c,c)}(-t) = (-1)^k \: \jac^{(c,c)}(t)$ and so $\jac_k^{(c,c)}$ is an even polynomial for even $k$, and an odd polynomial for odd $k$.
@@
 Hence
-\begin{align*}
-	\int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t = \int_{-\delta}^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} \: \jacw^{(c)} \: \D t 
-\end{align*}
-is zero for $j < k - 2$ by orthogonality, and is zero for $j = k-1$ due to symmetry over the domain.
+%\begin{align*}
+%	\int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t = \int_{-\delta}^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} \: \jacw^{(c)} \: \D t 
+%\end{align*}
+%is zero for $j < k - 2$ by orthogonality, and is zero for $j = k-1$ due to symmetry over the domain.
 Moreover, $t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t)$ is also an odd polynomial for any $k$ and so
-\begin{align*}
-	\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t) \: \jacw^{(c+1)}(t) \: \D t
-\end{align*}
-is zero for $j = k-1$ due to symmetry over the domain, and
-\begin{align*}
-	&\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \\
-	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \frac{\D}{\D t} \big[ t \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \big] \: \D t \\
-	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \big[ \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} + t \: \jac_j^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: t^2 \: \jac_j^{(c+1,c+1)} \: \jacw^{(c)} \big] \: \D t \\
-	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}}
-\end{align*}
-which is zero for $j < k - 2$ by orthogonality. Thus, (\ref{eqn:dxsparsity}) is zero for $j \notin \{k-2, k\}$.
-
-Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
-\begin{align}
-	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \nonumber  \\
-	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \nonumber \\
-	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \nonumber \\
-	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \nonumber \\
-	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:dxsparsityR}
-\end{align}
-By recalling (\ref{eqn:rhoderivative}) and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
-$$\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)}) = \text{deg} (\genjacw^{(1,0, j-k+2)}) = \text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k.$$ 
-Hence, by orthogonality, each term in (\ref{eqn:dxsparsityR}) is is zero for $m-j+3+j-k < n-k \iff m < n - 3$.
-
-Finally,
-\begin{align*}
-	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \\
-	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}}
-\end{align*}
-which is also zero for $m < n - 3$. Thus
-\begin{align*}
-	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
-\end{align*}
-showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,\dots,n$ and $j = k-2,k$ $(j \le m)$.
-
-We can gain the non-zero entries of the weighted differential operators similarly, by noting that for the disk-slice
-\begin{align}
-	\pddx \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
-	\pddy \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
-\end{align}
-and also that
-\begin{align*}
-	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{W^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wabc.
-\end{align*}
-
-\end{proof}
+%\begin{align*}
+%	\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t) \: \jacw^{(c+1)}(t) \: \D t
+%\end{align*}
+%is zero for $j = k-1$ due to symmetry over the domain, and
+%\begin{align*}
+%	&\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \\
+%	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \frac{\D}{\D t} \big[ t \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \big] \: \D t \\
+%	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \big[ \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} + t \: \jac_j^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: t^2 \: \jac_j^{(c+1,c+1)} \: \jacw^{(c)} \big] \: \D t \\
+%	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}}
+%\end{align*}
+%which is zero for $j < k - 2$ by orthogonality. Thus, (\ref{eqn:dxsparsity}) is zero for $j \notin \{k-2, k\}$.
+%
+%Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
+%\begin{align}
+%	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \nonumber  \\
+%	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \nonumber \\
+%	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \nonumber \\
+%	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \nonumber \\
+%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
+%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
+%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \nonumber \\
+%	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+%		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+%		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+%		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:dxsparsityR}
+%\end{align}
+%By recalling (\ref{eqn:rhoderivative}) and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
+%$$\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)}) = \text{deg} (\genjacw^{(1,0, j-k+2)}) = \text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k.$$ 
+%Hence, by orthogonality, each term in (\ref{eqn:dxsparsityR}) is is zero for $m-j+3+j-k < n-k \iff m < n - 3$.
+%
+%Finally,
+%\begin{align*}
+%	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \\
+%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}}
+%\end{align*}
+%which is also zero for $m < n - 3$. Thus
+%\begin{align*}
+%	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
+%\end{align*}
+%showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,\dots,n$ and $j = k-2,k$ $(j \le m)$.
+%
+%We can gain the non-zero entries of the weighted differential operators similarly, by noting that for the disk-slice
+%\begin{align}
+%	\pddx \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
+%	\pddy \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
+%\end{align}
+%and also that
+%\begin{align*}
+%	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{W^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wabc.
+%\end{align*}
+%
+%\end{proof}
 
@@
 
-\begin{proof}
-We proceed with the case for the non-weighted operators $T^{(a,b)\to(a+\tilde{a},b+\tilde{b},c+\tilde{c})}$, where $\tilde{a}, \tilde{b}, \tilde{c} \in \{0,1\}$.
+%\begin{proof}
+%We proceed with the case for the non-weighted operators $T^{(a,b)\to(a+\tilde{a},b+\tilde{b},c+\tilde{c})}$, where $\tilde{a}, \tilde{b}, \tilde{c} \in \{0,1\}$.
 Since $\{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}\}$ for $m = 0,\dots,n$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n$ polynomial, we can expand $\hdopnkabc = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}$.
@@
 First, note that
-\begin{align*}
-	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
-\end{align*}
-Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
-\begin{align*}
-	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
-	&\quad = \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\
-	&\quad \quad \quad \cdot \: \normjac^{(c+\tilde{c})} \: \ip< \jac_k^{(c,c)}, \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c+\tilde{c})}} \\
-	&\quad= \normgenjac^{(a,b,2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-	&\quad \quad \quad \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
-\end{align*}
-Since $\jacw^{(\tilde{c})}$ is a polynomial degree $2\tilde{c}$, we have that the above is then zero for $j < k-2\tilde{c}$.
+%\begin{align*}
+%	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
+%\end{align*}
+%Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
+%\begin{align*}
+%	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
+%	&\quad = \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\
+%	&\quad \quad \quad \cdot \: \normjac^{(c+\tilde{c})} \: \ip< \jac_k^{(c,c)}, \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c+\tilde{c})}} \\
+%	&\quad= \normgenjac^{(a,b,2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
+%	&\quad \quad \quad \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
+%\end{align*}
+%Since $\jacw^{(\tilde{c})}$ is a polynomial degree $2\tilde{c}$, we have that the above is then zero for $j < k-2\tilde{c}$.
 Further, since $\genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)}$ is a polynomial of degree $\tilde{a}+\tilde{b}+2\tilde{c}+j-k$, we have that the above is zero for $m-j+\tilde{a}+\tilde{b}+2\tilde{c}+j-k < n-k \iff m < n - \tilde{a}-\tilde{b}-2\tilde{c}$.
-
-The sparsity argument for the weighted parameter transformation operators follows similarly.
-\end{proof}
+%
+%The sparsity argument for the weighted parameter transformation operators follows similarly.
+%\end{proof}
 
@@
 \end{align*}
-The sparsity and structure of this biharmonic operator \bstodo{typo 4}are seen in Figure \ref{fig:sparsity}.
-
-\section{Computational aspects}\label{Section:Computation}
+The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
+
+\subsection{Computational aspects}\label{Section:Computation}
 
@@
 
-\subsection{Constructing $\genjac_n^{(a,b,c)}(x)$}
+\subsubsection{Constructing $\genjac_n^{(a,b,c)}(x)$}
 
@@
 	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
-	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
-\end{align}
-\end{proposition}
-
-\begin{proof}
-Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
-\begin{align*}
-	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
-\end{align*}
-and define
+	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x). \label{eqn:genjacexpansionii}
+\end{align}
+The constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ are given by
 \begin{align}
@@
 \end{align}
-Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
-\begin{align}
-	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}.
+\end{proposition}
+
+%\begin{proof}
+%Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
+%\begin{align*}
+%	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
+%\end{align*}
+%and define
+%\begin{align}
+%	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
+%	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half.
+\label{eqn:genjactnormalisationii}
+%\end{align}
+%Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
+%\begin{align}
+%	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}.
 \label{eqn:christoffeldarboux}
-\end{align}
-Then,
-\begin{align*}
-	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
-	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
-	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
-	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
-	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
-	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
-	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
-\end{align*}
-using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
-\begin{align*}
-	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
-	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
-	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
-	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
-	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
-	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
-	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
-\end{align*}
-using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
-\end{proof}
+%\end{align}
+%Then,
+%\begin{align*}
+%	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
+%	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
+%	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
+%	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
+%	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
+%	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
+%	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
+%\end{align*}
+%using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
+%\begin{align*}
+%	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
+%	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
+%	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
+%	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
+%	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
+%	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
+%	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
+%\end{align*}
+%using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
+%\end{proof}
 
@@
 
-\begin{proof}
-First, using (\ref{eqn:genjacexpansioni}) and (\ref{eqn:christoffeldarboux}) we have that
-\begin{align}
-	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
-	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: x \: \Big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\Big] \nonumber \\
-	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \nonumber \\
-	& \quad \quad \cdot \Big[\Big(\tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)\Big) \: \genjact_{n+1}^{(a,b,c,d)}(1) \nonumber \\
-	& \quad \quad \quad - \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \: \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x)\Big) \: \genjact_{n}^{(a,b,c,d)}(1) \Big] \label{eqn:genjactproofa}
-\end{align}
-Next, note that the recurrence coefficients for $\genjact_{n}^{(a,b,c+1,d)}(x)$ satisfy
-\begin{align}
-	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
-	&= (1-x) \: \Big[\tilde{\beta}_{n}^{(a,b,c+1,d)} \: \genjact_{n+1}^{(a,b,c+1,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \genjact_{n}^{(a,b,c+1,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \genjact_{n-1}^{(a,b,c+1,d)}(x)\Big] \nonumber \\
-	&= \mathcal{C}_{n+1}^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} \: \Big( \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n+2}^{(a,b,c,d)}(1) - \genjact_{n+2}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) \Big) \nonumber \\
-	& \quad \quad + \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
-	& \quad \quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \: \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:genjactproofb}
-\end{align}
-We can set $\tilde{\beta}_{-1}^{(a,b,c+1,d)} = 0$.
+%\begin{proof}
+%First, using (\ref{eqn:genjacexpansioni}) and (\ref{eqn:christoffeldarboux}) we have that
+%\begin{align}
+%	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
+%	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: x \: \Big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\Big] \nonumber \\
+%	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \nonumber \\
+%	& \quad \quad \cdot \Big[\Big(\tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)\Big) \: \genjact_{n+1}^{(a,b,c,d)}(1) \nonumber \\
+%	& \quad \quad \quad - \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \: \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x)\Big) \: \genjact_{n}^{(a,b,c,d)}(1) \Big] \label{eqn:genjactproofa}
+%\end{align}
+%Next, note that the recurrence coefficients for $\genjact_{n}^{(a,b,c+1,d)}(x)$ satisfy
+%\begin{align}
+%	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
+%	&= (1-x) \: \Big[\tilde{\beta}_{n}^{(a,b,c+1,d)} \: \genjact_{n+1}^{(a,b,c+1,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \genjact_{n}^{(a,b,c+1,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \genjact_{n-1}^{(a,b,c+1,d)}(x)\Big] \nonumber \\
+%	&= \mathcal{C}_{n+1}^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} \: \Big( \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n+2}^{(a,b,c,d)}(1) - \genjact_{n+2}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) \Big) \nonumber \\
+%	& \quad \quad + \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
+%	& \quad \quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \: \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:genjactproofb}
+%\end{align}
+%We can set $\tilde{\beta}_{-1}^{(a,b,c+1,d)} = 0$.
 By comparing coefficients of $\genjact_{n+2}^{(a,b,c,d)}(x)$ and $\genjact_{n+1}^{(a,b,c,d)}(x)$ in both (\ref{eqn:genjactproofa}) and (\ref{eqn:genjactproofb}) we obtain the desired recurrence coefficients for the OP $\genjact_{n}^{(a,b,c+1,d)}(x)$.
 The recurrence coefficients for the OPs $\genjact_{n}^{(a,b,c,d+1)}(x)$ are found similarly.
-\end{proof}
+%\end{proof}
 
@@
 
-\subsection{Quadrature rule on the disk-slice}
+\subsubsection{Quadrature rule on the disk-slice}
 
@@
 
-\begin{proof}
-We will use the substitution that
-\begin{align*}
-	x &= s, \quad y = \rho(s) \: t.
-\end{align*}
-First, note that, for $(x,y) \in \Omega$,
-\begin{align*}
-	\Wabc(x,y) &= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
-	&= \genjacw^{(a,b,c2)}(s) \: \jacw^{(c)}(t) \\
-	&=: V^{(a,b,c)}(s,t), \quad \text{for } (s,t) \in [\alpha,\beta] \times [-1,1].
-\end{align*}
-
-Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-\begin{align*}
-	f_e(x,y) &:= \half \Big(f(x, y) + f(x, -y)\Big), \quad \forall (x,y) \in \Omega\\
-	f_o(x,y) &:= \half \Big(f(x, y) - f(x, -y)\Big), \quad \forall (x,y) \in \Omega
-\end{align*}
-so that $y \mapsto f_e(x,y)$ for fixed $x$ is an even function, and $y \mapsto f_o(x,y)$ for fixed $x$ is an odd function.
+%\begin{proof}
+%We will use the substitution that
+%\begin{align*}
+%	x &= s, \quad y = \rho(s) \: t.
+%\end{align*}
+%First, note that, for $(x,y) \in \Omega$,
+%\begin{align*}
+%	\Wabc(x,y) &= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
+%	&= \genjacw^{(a,b,c2)}(s) \: \jacw^{(c)}(t) \\
+%	&=: V^{(a,b,c)}(s,t), \quad \text{for } (s,t) \in [\alpha,\beta] \times [-1,1].
+%\end{align*}
+%
+%Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
+%\begin{align*}
+%	f_e(x,y) &:= \half \Big(f(x, y) + f(x, -y)\Big), \quad \forall (x,y) \in \Omega\\
+%	f_o(x,y) &:= \half \Big(f(x, y) - f(x, -y)\Big), \quad \forall (x,y) \in \Omega
+%\end{align*}
+%so that $y \mapsto f_e(x,y)$ for fixed $x$ is an even function, and $y \mapsto f_o(x,y)$ for fixed $x$ is an odd function.
 Note that if $f$ is a polynomial, then $f_e(s, \rho(s)t)$ is a polynomial in $s \in [\alpha,\beta]$ for fixed $t$.
-
-Now, we have that
-\begin{align*}
-	\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
-	&= \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \\
-	&\approx \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
-	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
-	&= \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
-\end{align*}
-Suppose $f$ is a polynomial in $x$ and $y$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
+%
+%Now, we have that
+%\begin{align*}
+%	\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
+%	&= \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \\
+%	&\approx \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
+%	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
+%	&= \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
+%\end{align*}
+%Suppose $f$ is a polynomial in $x$ and $y$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
 First, note that the degree of the polynomial given by $x \mapsto f_e(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f_e(x,y)$ for fixed $x$ is $\le N$.
@@
 Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-
-Next, note that 
-\begin{align*}
-	\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
-	&= \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
-	&= 0
-\end{align*}
-since the inner integral at $(\dagger)$ over $t$ is zero, due to the symmetry over the domain.
-
-Hence, for a polynomial $f$ in $x$ and $y$ of degree $N$,
-\begin{align*}
-	\iint_\Omega f(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \iint_\Omega \Big(f_e(x,y) + f_o(x,y)\Big) \: \Wabc(x,y) \: \D y \: \D x \\
-	&= \iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x \\
-	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j),
-\end{align*}
-where $M = \ceil{\half(N+1)}^2$.
-\end{proof}
+%
+%Next, note that 
+%\begin{align*}
+%	\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
+%	&= \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
+%	&= 0
+%\end{align*}
+%since the inner integral at $(\dagger)$ over $t$ is zero, due to the symmetry over the domain.
+%
+%Hence, for a polynomial $f$ in $x$ and $y$ of degree $N$,
+%\begin{align*}
+%	\iint_\Omega f(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \iint_\Omega \Big(f_e(x,y) + f_o(x,y)\Big) \: \Wabc(x,y) \: \D y \: \D x \\
+%	&= \iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x \\
+%	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j),
+%\end{align*}
+%where $M = \ceil{\half(N+1)}^2$.
+%\end{proof}
 
@@
 
-\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
+\subsubsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
 
@@
 %
-\section{Examples on the disk-slice with zero Dirichlet conditions}\label{Section:Examples}
+\subsection{Examples on the disk-slice with zero Dirichlet conditions}\label{Section:Examples}
 
@@
 
-\subsection{Poisson}
+\subsubsection{Poisson}
 
@@
 which can be solved to find $\mathbf{u}$.
-In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
-For the final Gaussian bump example, we see that we also achieve algebraic convergence.
-\bstodoinline{Note that it is no longer correct to say that we achieve spectral convergence for ex 4 for the disk slice.
-Is there something more we can say?
-Same goes for the other examples.}
-
-\subsection{Inhomogeneous variable-coefficient Helmholtz}
+
+\subsubsection{Inhomogeneous variable-coefficient Helmholtz}
 
@@
 which can be solved to find $\mathbf{u}$.
-We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
-For the final Gaussian bump example, we see that we achieve spectral convergence.
 
@@
 
-\subsection{Biharmonic equation}
-
-Find \(u(x,y)\) given a function \(f(x,y)\) such that:
-\begin{align}
-	\begin{cases}
-    		\Delta^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
-		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
-	\end{cases}.
-	\label{eqn:biharmonic}
-\end{align}
-where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
-For the final Gaussian bump example, we see that we achieve spectral convergence.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Orthogonal Polynomials on spherical caps, spherical strips and spherical triangles (3D)}
+
+Moving forward, our aim is to try and develop a similar framework of using 3D OPs to expand functions and obtaining sparse differential operators on these spherical sub-domains, in order to solve partial differential equations and even lead to finite element methods on the sphere using these spherical sub-domains as elements.
+Developing these methods on the spherical cap is the natural first step from the work completed on the disk-slice to 3D domains, and where our work will head moving forward.
+
+One approach we will look at will be to build the 3D OPs on the spherical cap by in a similar vein as to how we could build the circle arc OPs, by projecting up from the 2D orthogonal polynomials on the disk to 3D OPs on the sphere.
+
+Define the end-disk-slice, denoted by $\omega$, by
+\begin{align}
+	\omega := \{(x,y) \in \R^2 \quad | \quad \alpha \le x \le 1, \: - \rho(x) \le y \le \rho(x)\}.
+\end{align}
+where $\alpha \ge 0$ and $\rho(x) := (1-x^2)^{\half}$, $x \in (\alpha,1)$. Further, define the spherical cap, denoted by $\Omega$, by
+\begin{align}
+	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
+\end{align}
+We can define the 3D OP family on $\Omega$ as the set of polynomials $\bigQab(x,y,z)$ where
+\begin{align}
+	\bigQab_n(x,y,z) := \begin{pmatrix}
+		\bighdopab_n(x,y) \\
+		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
+	\end{pmatrix} \in \R^{2n+1}, 
+	\quad \quad 
+	\bigQab := \begin{pmatrix}
+		\bigQab_0 \\
+		\bigQab_1 \\
+		\vdots \\
+		\bigQab_N
+	\end{pmatrix}.
+\end{align}
+The relations for multiplication by $x, y$ are then trivial.
+For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
+\begin{align}
+	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: \hdop^{(a,b+1)}_{m,j}(x,y), \\
+	z^2 \: \hdopnk^{(a,b+1)}(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b+1)\to(a,b)} \: \hdopmjab(x,y).
+\end{align}
+The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b+1)\to(a,b)}$ are given by
+\begin{align}
+	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: \hdop^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{\hdop^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
+	d_{n,k,m,j}^{(a,b+1)\to(a,b)} &= \frac{\ip< W^{(0,1)} \hdopnk^{(a,b+1)}, \: \hdopmjab >_\Wab}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
+        	&= \frac{\ip< \hdopnk^{(a,b+1)}, \: \hdopmjab >_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
+        	&= \frac{\norm{\hdopnk^{(a,b+1)}}^2_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \: c_{m,j,n,k}^{(a,b)\to(a,b+1)}.
+\end{align}
+The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ are just the entries to the parameter transformation operator matrix $T^{(a,b)\to(a,b+1)}$, defined by
+\begin{align}
+	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y),
+\end{align}
+which we have shown to be sparse with banded-block-banded structure, meaning the resulting operator matrix for the Jacobi operator for multiplication by $z$ here is sparse.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\bstodo{Reviewer 2, comments 2, 3}
+In this document I have presented the two main components of my PhD thus far, namely the introductory work I did using spherical harmonics in cartesian coordinates to solve multivariate (3D) PDEs on the sphere.
+From here, this allowed me to use my newly gained knowledge to begin developing similar sparse spectral methods on the 2D domains as defined by equation (\ref{eqn:domaindef}) with Conditions (\ref{cond:trap}, \ref{cond:disk}).
+
 We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
-\bstodo{Reviewer 1, comment 3} This work is related to the triangle case \cite{olver2019triangle} \bstodoinline{add more here?}, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements.
-%A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
-With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
+This work is related to the triangle case \cite{olver2019triangle}, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements, for example, by dividing the disk into disk slice elements.
+
 This work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles.
@@
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
-Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
-
-%
+Constructing these sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is the next steps of my PhD, which has applications in weather prediction \cite{staniforth2012horizontal}.
+Other extensions to the disk-slice work could include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \appendix
@@
 \section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-        	%\label{fig:solutionblocknorms-poisson}
-	\centering
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-	%\label{fig:solution-poisson}
-	\centering
-	\end{subfigure}
-	\centering
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
-Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
-Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-in the trapezium.}
-	\centering
-	\label{fig:appendixfigs}
-\end{figure}
 
@@
 
-Note \bstodo{typo 6}that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-
-\section{End-Disk-Slice}\label{Appendix:HalfDisk}
+Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+
+\section{End-Disk-Slice (e.g. the Half-Disk)}\label{Appendix:HalfDisk}
 
@@
 
-In Figure \ref{fig:appendixfigs} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
-
-\section{Trapeziums}\label{Appendix:Trapezium}
-
-We can further extend this work to trapezium shaped domains. Note that for any trapezium there exists an affine map to the canonical trapezium domain that we consider here, given by
-\begin{align*}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
-with 
-\begin{align*}
-\begin{cases}
-(\alpha, \beta) &:= (0,1) \\
-(\gamma, \delta) &:= (0,1) \\
-\rho(x) &:= 1- \xi x, \quad \xi \in (0,1) \\
-\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\xi x)^{c} \\
-\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
-\end{cases}
-\end{align*}
-The weight $\jacw^{(a,b)}(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
-We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = \jac_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
-$\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\).
-Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align*}
-	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align*}
-orthogonal with respect to the weight
-\begin{align*}
-	W^{(a,b,c,d)}(x,y) &:= \genjacw^{(a,b,c+d)}(x) \: w^{(d,c)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
-	&= (1-x)^a \: x^b \: y^c \: (1- \xi x - y)^d, \quad (x,y) \in \Omega.
-\end{align*}
-
-In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
-
-%\subsection{Quadrature rule on the trapezium}
-%
-%In this section we construct a quadrature rule exact for polynomials in the trapezium $\Omega$ that can be used to expand functions in $\hdopnkabcd(x,y)$.
-%
-%\begin{theorem}
-%Denote the Gauss quadrature nodes and weight on $[\alpha,\beta]$ with weight $\genjacw^{(a,b,c+d)}(s) = (\beta - s)^a \: (s - \alpha)^b \: \rho(s)^{c+d+1}$ as $(s_k,w_k^{(s)})$ , and
-% on $[0,1]$ with weight $\jacw^{(d,c)}(t) = t^c \: (1-t)^d$ as $(t_k,w_k^{(t)})$. Define
-%\begin{align*}
-%	x_{i+(j-1)N} &:= s_j, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
-%	y_{i+(j-1)N} &:= \rho(s_j) \: t_i, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
-%	w_{i+(j-1)N} &:= w_j^{(s)} w_i^{(t)}, \quad  i,j = 1,\dots, \ceil{\frac{N+1}{2}}.
-%\end{align*}
-%Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
-%$$
-%\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D A \approx \sum_{j=1}^{M} w_j \: f(x_j, y_j)
-%$$
-%where $M = \ceil{\half(N+1)}^2$, and the quadrature rule is exact if $f(x,y)$ is a polynomial of degree $\le N$.
-%\end{theorem}
-%
-%\begin{proof}
-%\begin{align*}
-%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
-%	&\approx \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
-%	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
-%	&= \sum_{k=1}^{M_1 M_2}  w_k \: f(x_k, y_k).
-%\end{align*}
-%Suppose f is a polynomial in $x$ and $y$ of degree $N$.
-First, note that the degree of the polynomial given by $x \mapsto f(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f(x,y)$ for fixed $x$ is $\le N$.
-Also note that $s \mapsto f\big(s,\rho(s) t\big)$ for fixed $t$ is then a degree $N$ polynomial (since $\rho$ is a degree $1$ polynomial).
-Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-%
-%Thus, for a polynomial $f$ in $x$ and $y$ of degree $N$, 
-%\begin{align*}
-%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
-%	&= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad \\
-%	&= \sum_{j=1}^{\ceil{\half(N+1)}} \Bigg( w_j^{(s)} \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \\
-%	&= \sum_{k=1}^{M}  w_k \: f(x_k, y_k)
-%\end{align*}
-%where $M = \ceil{\half(N+1)}^2$.
-%
-%\end{proof}
-
-\bstodo{typo 8 (capitalization of names in some reference titles}
-\bibliography{halfdisk}
+\bibliography{lsr}
 

### 2020-01-10 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
fixed typo
updates to paper

@@
 
-We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the triangle \cite{olver2018recurrence,olver2019triangle}.
+We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
 On the disk-slice the family of weights we consider are of the form
@@
 $$
-where $C_{n,k} \neq 0$ and ``lower order terms'' can  include degree $n$ polynomials with fewer $y$ terms.
+where $C_{n,k} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
 The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
@@
 Sparsity comes from expanding the domain and range of an operator using different choices of $a$ and $b$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
+Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 \noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
-We present a method involving use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.
+We present a method involving use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.
 
@@
 
-\noindent\appref{HalfDisk}: We discuss extension to the special case of the end-disk-slice.
+\noindent\appref{HalfDisk}: We discuss extension to the special case of end-disk-slices (e.g., half disks).
 
@@
 \begin{align*}
-	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
-	\alphaabcd_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
+	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b+k+\half)} \\
 	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
@@
 		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
-		& C_2 & \ddots & \ddots \\
-		& & \ddots &
+		& C_2 & B_2 - G_2(x,y)  & \ddots \\
+		& & \ddots &\ddots
 \end{pmatrix}
@@
 \begin{pmatrix}
-	\hdopabcd_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
+	\hdopabcd_{0,0} \\ 0 \\ 0 \\ 0 \\ \vdots  \\
 \end{pmatrix},
@@
 \begin{definition}\label{def:parametertransformationoperators}
-Define the operator matrices $T^{(a,b,c)\to(a+1,b+1,c)}$, $T^{(a,b,c)\to(a,b,c+1)}$ and $T^{(a,b,c)\to(a+1,b+1,c+1)}$ for conversion between non-weighted spaces, and $T_W^{(a,b,c)\to(a-1,b-1,c)}$, $T_W^{(a,b,c)\to(a,b,c-1)}$ and $T_W^{(a,b,c)\to(a-1,b-1,c-1)}$ for conversion between weighted spaces, according to:
+Define the operator matrices 
+$$
+T^{(a,b,c)\to(a+1,b+1,c)}, \quad T^{(a,b,c)\to(a,b,c+1)}\qqand T^{(a,b,c)\to(a+1,b+1,c+1)}
+$$ 
+for conversion between non-weighted spaces, and 
+$$
+T_W^{(a,b,c)\to(a-1,b-1,c)}, \quad T_W^{(a,b,c)\to(a,b,c-1)} \qqand T_W^{(a,b,c)\to(a-1,b-1,c-1)}
+$$ 
+for conversion between weighted spaces, according to:
 \begin{align*}
@@
 \bstodo{Reviewer 1, comments 1,2}
-It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \cite[18.2.12]{DLMF}.
+It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
 We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case.
@@
 \end{figure}
-solution-exact-poisson-diskslice-alpha=0.2-beta=0.8-u=y3expx-actual
+
 \begin{figure}[t]
@@
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
-For the final Gaussian bump example, we see that we also achieve algebraic convergence.
+
+%For the final Gaussian bump example, we see that we also achieve algebraic convergence.
 \bstodoinline{Note that it is no longer correct to say that we achieve spectral convergence for ex 4 for the disk slice.
@@
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
-For the final Gaussian bump example, we see that we achieve spectral convergence.
+
+% For the final Gaussian bump example, we see that we achieve spectral convergence.
 
@@
 We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
-\bstodo{Reviewer 1, comment 3} This work is related to the triangle case \cite{olver2019triangle} \bstodoinline{add more here?}, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements.
+\bstodo{Reviewer 1, comment 3} This work extends  the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements.
 %A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
@@
 
-The work in this paper on the disk-slice can be easily transferred to the special-case domain of the end-disk-slice by which we mean
+The work in this paper on the disk-slice can be easily transferred to the special-case domain of the end-disk-slice , such as half disks, by which we mean
 \begin{align*}

### 2020-01-10 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
Fixed typos

@@
 $$
-The corresponding OPs denoted $\hdopnkab(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
-$$
-\hdopnkab(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
+The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
+$$
+\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
 $$
@@
 
-Sparsity comes from expanding the domain and range of an operator using different choices of $a$ and $b$.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$.
 Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.

### 2020-01-10 papers/disk-slices-and-trapeziums/revised/review-comments.tex
review comments

\documentclass[10pt]{letter}

\usepackage[papersize={8.5in,11in},margin=1in]{geometry}
\usepackage[final,stretch=10,shrink=10]{microtype} % default of stretch=shrink=20 is a little too much
\usepackage{amsmath,amsfonts,amsxtra,amssymb}
\usepackage{url}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}

\definecolor{bluey}{rgb}{0.0,0.2,0.5}
\newcommand{\comment}[1]{\textit{\color{bluey}#1}}

%\definecolor{rev1}{HTML}{cb270f}
%\definecolor{rev2}{HTML}{1c8235}

\usepackage{todonotes}
\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

\newcommand{\half}{\frac{1}{2}}
\newcommand{\genjac}{R}

\begin{document}

\thispagestyle{empty}

\hfill
\begin{flushright}
Correspondence to: \\
Sheehan Olver \\
s.olver@imperial.ac.uk \\
\end{flushright}

\vspace{1em}
\hfill\today

Dear Jose Bautista Francisco III

Thank you for forwarding the reviews of our article. We greatly appreciate your time as well as that of the referees.  Below we detail our response to the referees' reports.

Ben Snowball and Sheehan Olver

\bigskip

\centerline{\textbf{Response to Referee 1}}

We have fixed the typos and minor suggestions. The other comments are addressed below.

\begin{itemize}[parsep=1em,leftmargin=1em]

\item \comment{Conditioning for large b: In Section 4.1, the authors mention that there is an ill-conditioning issue when b is large. First, does this ill-conditioning grow algebraically with b (I suspect so)? Second, I do not understand where this ill-conditioning is coming from. Can the authors explain?}

Firstly, we note that that with the changes we have made in focusing on the disk-slice case rather than the half-disk, the parameter $b$ being refered to here is now $c$. We have altered our approach for finding the recurrence coefficients of the (non-classical) 1D OPs $\{\genjac_n^{(a,b,c)}\}$, from using the Lanczos algorithm (that lead to the ill-conditioning for large $c$) to using an alternative recursive approach involving the Christoffel-Darboux formula.

\item \comment{Explicit recurrence relations: The manuscript demonstrates that sparse differentiation and conversion recurrence relations exist between $H_{n,k}$ but the recurrences are not given explicitly (as they are defined via integrals). This means that one must numerically generate the entries of the differentiation matrices as well as the conversion operators. I am strongly encouraging the authors to derive the explicit recurrence relationship for this paper: both for conversion and differentiation.}

To the first comment, we have now implemented a way to exactly calculate the recurrence coefficients for the non-classical 1D OPs, allowing one to exactly evaluate these integrals using quadrature.

\item \comment{Boundary conditions: I think a few more details are needed on how the authors are imposing boundary conditions. I understand for zero Dirichlet conditions and zero Dirichlet+Neumann, but how does one impose general boundary conditions in this spectral method? For example, can one impose zero Neumann condition with a zero mean condition? Or, non-homogeneous Dirichlet conditions. The authors mention that they can now use this to construct an element method. The authors may like to add a comment on how one might patch together domains (and impose any interface constraints on the solution)}

We have addressed point in the abstract that we focus on the BCs mentioned. We have also added a point in the conclusion about how this work relates to the triangle, and the work done there on implementing other BCs.

\item \comment{p-finite element method with sparse operator: Preserving symmetry and sparsity in the discretization sounds like an all-round win. Why are the authors not advocating this approach in the body of the manuscript? Are these symmetric FEM discretizations as sparse?}

\bstodoinline{something about not wanting to make this a FEM paper, when we are looking at a single element?}

\end{itemize}

\bigskip

\centerline{\textbf{Response to Referee 2}}

We have fixed the typos and minor suggestions. The other comments are addressed below.

\begin{itemize}[parsep=1em,leftmargin=1em]

\item \comment{The title says "disk slices, and trapezium", but then the definition of $\Omega$ gives a definition in terms of the function $\rho(x)$. It's not obvious (without digging out paper) that these are the same. Please make the connection for the reader. Next it says that you specialise to $\rho(x) = (1-x^2)^\half$, which is the half disk. Later in the paper it's not obvious if your more technical results apply only to the half disk, or if the half disk is simply used to get the conversation started. It seems to me that the paper is not about "disk slices, and trapezium", but really the "half disk". If so, then I think the authors should do one of two things:}

A) \comment{Explain why, in detail, why the half disk implies the same results automatically for disk slices, and trapezium. Or,}

B) \comment{Change the title to the paper to reflect the focus on the half disk. }

\comment{If the choice is option B, then I strongly suggest rewriting the introduction to reflect the more limited scope. This means talking about the half disk right away, rather than to defining $\Omega$ so generally.}

We have altered the paper so that we now focus on the general disk-slice case, rather than the specific half-disk. In the main body of the paper, we present how to build the OPs for a general case, and then focus on how to build the operators for the disk-slice case. In the appendix we include an explanation for how the framework we present in the paper can also be applied to the trapezium and "end-disk-slice" (of which the half-disk is a specific case).

\item \comment{The final full paragraph before the overview is highly misleading. This paper is not about spheres, weather, or turbulent pipe flow. This seems a bit like pandering. I don't think the authors need such tactics to justify the importance of their work. The mathematics are enough justification. I would rather see a short summary of where this work sits in comparison to some of the other similar past work. As I said, this work is part of a larger program to understand orthogonal polynomials on interesting domains.}

Removed and added to conclusions.

\item \comment{The conclusion is very short. If the authors want to talk about spheres, or "disk slices, and trapezium", then this might be a good place to put it. There could be a short discussion about how the methods and relations might be extended to the more general cases. I have no doubt this is possible.}

Added the motivation/future work of spherical sub-domains to the conclusions.

\item \comment{Looking at the past work, it seem Olver was involved in a similar effort for the full disk.}

\comment{I would like to see an explanation of how the half disk is more than just half the full disk? It seems in the full disk that Fourier series was used in the angular direction. But the full 2D functions are Cartesian polynomials. How do these problems map into each other? }

\comment{It seem that one could obtain a basis on the half disk by restricting to even and odd functions on the full disk. But half-period Fourier series have a hard time representing all functions on the 1D interval. This is one of the motivations behind Chebyshev polynomials, and the authors surely know this.}

\comment{My guess is that the half-disk basis contains something different than the even/odd restricted full-disk basis. It would be good to know what.}

\bstodoinline{I am not to sure how to address this comment.}

\item \comment{The examples could be improved somewhat.}

\comment{The Norm plots on the right-hand side of the figures needs more explanation. Is this roughly the norm of the coefficients on the solution? It seems this is a proxy for the error of the solution. But it's not exactly that. }

\comment{I would like to see at least one example where the authors compare the numerical solution to a known exact solution. This could happen by putting an answer into an operator and generating a right-hand side. Or is this what was done?}

Added in example showing comparison to an exact known solution.

\item \comment{It would be useful to see one example with the action near the straight boundary. }

\comment{Relating to the earlier discussion about the difference between even/odd full disk, and the half disk, I believe the benefit to this basis is the ability to capture things near the outer edge. }

\comment{Even better would be a demonstration with non-zero boundary conditions at the straight edge. I don't think the even/odd full disk basis would easily (or at all) be able to handle arbitrary boundary conditions at $x=0$.}

We have added an example demonstrating the Helmholtz equation with the action in one corner of the trapezium.

\bstodoinline{Should I put more here?}

\end{itemize}

\end{document}

### 2020-01-10 papers/disk-slices-and-trapeziums/revised/review-comments.tex
modify comments

@@
 
-Thank you for forwarding the reviews of our article. We greatly appreciate your time as well as that of the referees.  Below we detail our response to the referees' reports.
+Thank you for forwarding the reviews of our article.
+We greatly appreciate your time as well as that of the referees.
+We have made substantial changes in this revision, having changed the exposition to focus on general disk slices and to use the Christofel-Darboux formula in generating the necessary families of orthogonal polynomials.
+Below we detail our response to the referees' reports.
+
+Note that we are  hoping to improve the figures in the final version to go beyond degree $N = 200$ polynomials.
+We have decided to resubmit the draft now, as improving the figures will be a minor change.
 
@@
 
-\bstodoinline{something about not wanting to make this a FEM paper, when we are looking at a single element?}
+The $p$-finite element method only applies to more specific classes of PDEs so we have chosen to focus on the more general strong formulation version.
+The extension to $p$-FEM then follows as a specialisation of the results, which does indeed have substantial benefits in terms of symmetry.
 
@@
 
-\bstodoinline{I am not to sure how to address this comment.}
+We don't believe the suggested restriction of the full disk polynomials  spans all polynomials in cartesian coordinates.
+In any case, the changed emphasis to general disk-slices makes these comments less relevant.
 
@@
 
-Added in example showing comparison to an exact known solution.
+We added an example showing comparison to an exact known solution.
 
@@
 
-\bstodoinline{Should I put more here?}
-
 \end{itemize}

### 2020-01-11 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
final comments

@@
 In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay\bstodo{Reviewer 2, comment 5}.
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
@@
 Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
+
+\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.  The second author was supported in part by a Leverhulme Trust Research Grant.
 

### 2020-01-23 papers/lsr/lsr.tex
LSR

@@
 \newcommand{\bigQab}{\bigQ^{(a,b)}}
+\newcommand{\Qab}{Q^{(a,b)}}
 
@@
 \newcommand{\normgenjact}{\omega_{\genjact}}
+
+\newcommand{\xvec}{\mathbf{x}}
 
@@
 
-\section{Orthogonal polynomials on the circle arc}
-
-As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
-We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to \cite{huybrechs2010fourier}:
-\begin{align}
-\bigP_0(x,y) := T^h_0(x) \equiv T_0,
-\quad \quad 
-\bigP_n(x,y) := \begin{bmatrix}
-		T^h_n(x) \\
-		y \: U^h_{n-1}(x)
-	\end{bmatrix} \in \R^2 \quad \forall n \in \N,
-\quad \quad 
-\bigP := \begin{bmatrix}
-		\bigP_0 \\
-		\hline
-		\bigP_1 \\
-		\hline
-		\bigP_2 \\
-		\vdots \\
-	\end{bmatrix}.
-\end{align}
-
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
-In the work in this section we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices and trapeziums.
+In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices and trapeziums.
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
 We apply the framework to solving the Poisson, variable coefficient Helmholtz, and Biharmonic equations.
-For now, we have focused on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
+In this paper we focus on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
 
 \subsection{Explicit construction}
-
 Here, we have developed sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.
 More precisely, we consider the solution of partial differential equations on the domain
-\begin{align}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\} \label{eqn:domaindef}
+\begin{align}\label{eqn:domaindef}
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
 \end{align}
@@
 
-For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss an extension to other the end-disk-slice in the appendix (note we omit detailed discussion  on the extension to trapezium in this document).
+For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss an extension to the end-disk-slice in the appendix (note we omit detailed discussion on the extension to the trapezium in this document).
 
@@
 
-$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
+The OPs $\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
 \begin{align*}
@@
 
-%Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.
-
 \subsection{Jacobi matrices}
@@
 
-%\begin{proof}
-%First, note that:
-%\begin{align}
-%	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
-%	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(c-1)}(y), \label{eqn:derivativeofweightjac} \\
-%	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
-%\end{align}
-%
-%We proceed with the case for the operator $D_y^{(a,b,c)}$ for partial differentiation by $y$.
-Since $\{\hdop^{(a,b,c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^y \: \hdop^{(a, b, c+1)}_{m,j}$.
-The coefficients of the expansion are then the entries of the relevant operator matrix.
-We can use an integration-by-parts argument to show that the only non-zero coefficient of this expansion is when $m = n-1$, $j = k-1$.
-First, note that
-%\begin{align*}
-%	c_{m,j}^y = {\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
-%\end{align*}
-%Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
-%\begin{align*}
-%	&\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
-%	&= \iint_\Omega \: \Big[ \genjacnmk^{(a,b,2c+2k+1)}(x) \: \rho(x)^{k-1} \: \jac_k^{(c,c) \: \prime}\fpr(\frac{y}{\rho(x)}) \\
-%	& \quad \quad \quad \quad \quad \quad \cdot \: \genjacmmj^{(a,b,2c+2j+3)}(x) \: \rho(x)^{j} \: \jac_j^{(c+1,c+1)}\fpr(\frac{y}{\rho(x)}) \Big] \: \D y \: \D x \\
-%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-%	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}
-%\end{align*}
-%Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
-%\begin{align*}
-%	\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} &= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
-%	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddy [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
-%	&= - \int_{-1}^1 \: \jac_k^{(c,c)} \: [ \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: y \: \jac_{j}^{(c+1,c+1)} \:\jacw^{(c)} ] \: \D y \\
-%	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}}
-%\end{align*}
-%which is zero for $j < k-1$ by orthogonality. Further, when $j = k-1$, we have that
-%\begin{align*}
-%	&\normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacmmj^{(a,b,2c+2k+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \delta_{n,m+1},
-%\end{align*}
-%showing that the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally,
-%\begin{align*}
-%	c_{n-1,k-1}^y &= \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{k-1}^{(c+1,c+1)} >_{\jacw^{(c+1)}}.
-%\end{align*}
-%
-%We next consider the case for the operator $D_x^{(a,b,c)}$ for partial differentiation by $x$.
-Since $\{\hdop^{(a+1, b+1, c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^x \: \hdop^{(a+1, b+1, c+1)}_{m,j}$.
-The coefficients of the expansion are then the entries of the relevant operator matrix.
-As before, we can use an integration-by-parts argument to show that the only non-zero coefficients of this expansion are when $m = n-1, n-2, n-3$, $j = k, k-1, k-2$ and $0 \le j \le m$.
-First, note that
-%\begin{align*}
-%	c_{m,j}^x = {\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
-%\end{align*}
-%Now, again using the change of variable $t= \frac{y}{\rho(x)}$, we have that
-%\begin{align}
-%	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
-%	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-%	&\quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\ 
-%	& \quad + k \: \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-%	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big) \nonumber \\
-%	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-%	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:dxsparsity}
-%\end{align}
-%We will first show that the second factor of each term in (\ref{eqn:dxsparsity}) are zero for $j < k - 2$ and also for $j = k - 1$.
-To this end, observe that, for any integer $c$, $\jac^{(c,c)}(-t) = (-1)^k \: \jac^{(c,c)}(t)$ and so $\jac_k^{(c,c)}$ is an even polynomial for even $k$, and an odd polynomial for odd $k$.
-Thus, $\jac_k^{(c,c)} \: \jac_{k-1}^{(c+1,c+1)}$ is an odd polynomial for any $k$.
-Hence
-%\begin{align*}
-%	\int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t = \int_{-\delta}^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} \: \jacw^{(c)} \: \D t 
-%\end{align*}
-%is zero for $j < k - 2$ by orthogonality, and is zero for $j = k-1$ due to symmetry over the domain.
-Moreover, $t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t)$ is also an odd polynomial for any $k$ and so
-%\begin{align*}
-%	\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t) \: \jacw^{(c+1)}(t) \: \D t
-%\end{align*}
-%is zero for $j = k-1$ due to symmetry over the domain, and
-%\begin{align*}
-%	&\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \\
-%	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \frac{\D}{\D t} \big[ t \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \big] \: \D t \\
-%	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \big[ \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} + t \: \jac_j^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: t^2 \: \jac_j^{(c+1,c+1)} \: \jacw^{(c)} \big] \: \D t \\
-%	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}}
-%\end{align*}
-%which is zero for $j < k - 2$ by orthogonality. Thus, (\ref{eqn:dxsparsity}) is zero for $j \notin \{k-2, k\}$.
-%
-%Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
-%\begin{align}
-%	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \nonumber  \\
-%	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \nonumber \\
-%	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \nonumber \\
-%	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \nonumber \\
-%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
-%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
-%	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \nonumber \\
-%	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-%		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-%		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-%		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:dxsparsityR}
-%\end{align}
-%By recalling (\ref{eqn:rhoderivative}) and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
-%$$\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)}) = \text{deg} (\genjacw^{(1,0, j-k+2)}) = \text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k.$$ 
-%Hence, by orthogonality, each term in (\ref{eqn:dxsparsityR}) is is zero for $m-j+3+j-k < n-k \iff m < n - 3$.
-%
-%Finally,
-%\begin{align*}
-%	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \\
-%	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}}
-%\end{align*}
-%which is also zero for $m < n - 3$. Thus
-%\begin{align*}
-%	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
-%\end{align*}
-%showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,\dots,n$ and $j = k-2,k$ $(j \le m)$.
-%
-%We can gain the non-zero entries of the weighted differential operators similarly, by noting that for the disk-slice
-%\begin{align}
-%	\pddx \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
-%	\pddy \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
-%\end{align}
-%and also that
-%\begin{align*}
-%	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{W^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wabc.
-%\end{align*}
-%
-%\end{proof}
-
 There exist conversion matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
@@
 
-%\begin{proof}
-%We proceed with the case for the non-weighted operators $T^{(a,b)\to(a+\tilde{a},b+\tilde{b},c+\tilde{c})}$, where $\tilde{a}, \tilde{b}, \tilde{c} \in \{0,1\}$.
-Since $\{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}\}$ for $m = 0,\dots,n$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n$ polynomial, we can expand $\hdopnkabc = \sum_{m=0}^{n} \sum_{j=0}^m c_{m,j} \: \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}$.
-The coefficients of the expansion are then the entries of the relevant operator matrix.
-We will show that the only non-zero coefficients are for $m \ge n - \tilde{a} - \tilde{b} - 2\tilde{c} $, $j \ge k-2\tilde{c}$ and $0 \le j \le m$.
-First, note that
-%\begin{align*}
-%	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
-%\end{align*}
-%Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
-%\begin{align*}
-%	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
-%	&\quad = \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\
-%	&\quad \quad \quad \cdot \: \normjac^{(c+\tilde{c})} \: \ip< \jac_k^{(c,c)}, \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c+\tilde{c})}} \\
-%	&\quad= \normgenjac^{(a,b,2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-%	&\quad \quad \quad \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
-%\end{align*}
-%Since $\jacw^{(\tilde{c})}$ is a polynomial degree $2\tilde{c}$, we have that the above is then zero for $j < k-2\tilde{c}$.
-Further, since $\genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)}$ is a polynomial of degree $\tilde{a}+\tilde{b}+2\tilde{c}+j-k$, we have that the above is zero for $m-j+\tilde{a}+\tilde{b}+2\tilde{c}+j-k < n-k \iff m < n - \tilde{a}-\tilde{b}-2\tilde{c}$.
-%
-%The sparsity argument for the weighted parameter transformation operators follows similarly.
-%\end{proof}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
-        \centering
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
-        \centering
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
-        \centering
-	\end{subfigure}
-    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
-Left: the Laplace operator $\laplacewiii$.
-Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
-Right: the biharmonic operator $\biharmonictwo$.}
-        \label{fig:sparsity}
-        \centering
-\end{figure}
-
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
@@
 
+\begin{figure}[t]
+	\begin{subfigure}{0.32\textwidth}
+	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
+        \centering
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
+        \centering
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
+        \centering
+	\end{subfigure}
+    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
+Left: the Laplace operator $\laplacewiii$.
+Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
+Right: the biharmonic operator $\biharmonictwo$.}
+        \label{fig:sparsity}
+        \centering
+\end{figure}
+
 \subsection{Computational aspects}\label{Section:Computation}
@@
 
-%To obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), we use a variant of the Stieltjes procedure \cite{gautschi1982generating} where the polynomials are expressed as Chebyshev polynomial expansions and the inner products are calculated via Clenshaw--Curtis quadrature.
-This has the benefit that it is easier to incorporate high-precision arithmetic, which we use to overcome ill-conditioning present when $c$ is large, as required for large $n$ in \eqref{eq:diskpolys}.
-The ApproxFun.jl \cite{ApproxFun} package gives a convenient way to manipulate Chebyshev and Jacobi expansions, which we use to calculate the inner products and norms in this algorithm, utilising the {\tt BigFloat} type to handle high-precision calculations.
-
-It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel-Darboux formula \cite[18.2.12]{DLMF}.
+It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
 We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case.
@@
 	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
-	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x). \label{eqn:genjacexpansionii}
-\end{align}
-The constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ are given by
-\begin{align}
-	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
-	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half.
-\label{eqn:genjactnormalisationii}
+	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
 \end{align}
 \end{proposition}
-
-%\begin{proof}
-%Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
-%\begin{align*}
-%	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
-%\end{align*}
-%and define
-%\begin{align}
-%	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
-%	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half.
-\label{eqn:genjactnormalisationii}
-%\end{align}
-%Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
-%\begin{align}
-%	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}.
-\label{eqn:christoffeldarboux}
-%\end{align}
-%Then,
-%\begin{align*}
-%	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
-%	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
-%	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
-%	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
-%	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
-%	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
-%	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
-%\end{align*}
-%using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
-%\begin{align*}
-%	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
-%	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
-%	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
-%	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
-%	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
-%	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
-%	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
-%\end{align*}
-%using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
-%\end{proof}
 
@@
 \end{proposition}
-
-%\begin{proof}
-%First, using (\ref{eqn:genjacexpansioni}) and (\ref{eqn:christoffeldarboux}) we have that
-%\begin{align}
-%	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
-%	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: x \: \Big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\Big] \nonumber \\
-%	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \nonumber \\
-%	& \quad \quad \cdot \Big[\Big(\tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)\Big) \: \genjact_{n+1}^{(a,b,c,d)}(1) \nonumber \\
-%	& \quad \quad \quad - \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \: \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x)\Big) \: \genjact_{n}^{(a,b,c,d)}(1) \Big] \label{eqn:genjactproofa}
-%\end{align}
-%Next, note that the recurrence coefficients for $\genjact_{n}^{(a,b,c+1,d)}(x)$ satisfy
-%\begin{align}
-%	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
-%	&= (1-x) \: \Big[\tilde{\beta}_{n}^{(a,b,c+1,d)} \: \genjact_{n+1}^{(a,b,c+1,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \genjact_{n}^{(a,b,c+1,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \genjact_{n-1}^{(a,b,c+1,d)}(x)\Big] \nonumber \\
-%	&= \mathcal{C}_{n+1}^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} \: \Big( \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n+2}^{(a,b,c,d)}(1) - \genjact_{n+2}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) \Big) \nonumber \\
-%	& \quad \quad + \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
-%	& \quad \quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \: \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:genjactproofb}
-%\end{align}
-%We can set $\tilde{\beta}_{-1}^{(a,b,c+1,d)} = 0$.
-By comparing coefficients of $\genjact_{n+2}^{(a,b,c,d)}(x)$ and $\genjact_{n+1}^{(a,b,c,d)}(x)$ in both (\ref{eqn:genjactproofa}) and (\ref{eqn:genjactproofb}) we obtain the desired recurrence coefficients for the OP $\genjact_{n}^{(a,b,c+1,d)}(x)$.
-The recurrence coefficients for the OPs $\genjact_{n}^{(a,b,c,d+1)}(x)$ are found similarly.
-%\end{proof}
 
@@
 using the Lanczos algorithm to compute the recurrence coefficients $\{\alpha_{n}^{(a,b,2c+1)}\}$, $\{\beta_{n}^{(a,b,2c+1)}\}$ -- however, we only need to compute these once, and can store and save this initialisation to disk once computed, for the given values of $a, b, c$).
-%This is the most expensive part of the current calculation, but note that we can reuse this computation for multiple partial differential equations.
-Efficient construction of OPs with large parameters in the weights is an important topic, but one tangential to the proposed scheme.
 
@@
 
-%\begin{proof}
-%We will use the substitution that
-%\begin{align*}
-%	x &= s, \quad y = \rho(s) \: t.
-%\end{align*}
-%First, note that, for $(x,y) \in \Omega$,
-%\begin{align*}
-%	\Wabc(x,y) &= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
-%	&= \genjacw^{(a,b,c2)}(s) \: \jacw^{(c)}(t) \\
-%	&=: V^{(a,b,c)}(s,t), \quad \text{for } (s,t) \in [\alpha,\beta] \times [-1,1].
-%\end{align*}
-%
-%Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-%\begin{align*}
-%	f_e(x,y) &:= \half \Big(f(x, y) + f(x, -y)\Big), \quad \forall (x,y) \in \Omega\\
-%	f_o(x,y) &:= \half \Big(f(x, y) - f(x, -y)\Big), \quad \forall (x,y) \in \Omega
-%\end{align*}
-%so that $y \mapsto f_e(x,y)$ for fixed $x$ is an even function, and $y \mapsto f_o(x,y)$ for fixed $x$ is an odd function.
-Note that if $f$ is a polynomial, then $f_e(s, \rho(s)t)$ is a polynomial in $s \in [\alpha,\beta]$ for fixed $t$.
-%
-%Now, we have that
-%\begin{align*}
-%	\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
-%	&= \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \\
-%	&\approx \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
-%	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
-%	&= \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
-%\end{align*}
-%Suppose $f$ is a polynomial in $x$ and $y$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
-First, note that the degree of the polynomial given by $x \mapsto f_e(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f_e(x,y)$ for fixed $x$ is $\le N$.
-Also note that $s \mapsto f_e\big(s,\rho(s) t\big)$ for fixed $t$ is then a degree $N$ polynomial (since $\rho$ is a degree $1$ polynomial).
-Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-%
-%Next, note that 
-%\begin{align*}
-%	\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
-%	&= \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
-%	&= 0
-%\end{align*}
-%since the inner integral at $(\dagger)$ over $t$ is zero, due to the symmetry over the domain.
-%
-%Hence, for a polynomial $f$ in $x$ and $y$ of degree $N$,
-%\begin{align*}
-%	\iint_\Omega f(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \iint_\Omega \Big(f_e(x,y) + f_o(x,y)\Big) \: \Wabc(x,y) \: \D y \: \D x \\
-%	&= \iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x \\
-%	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j),
-%\end{align*}
-%where $M = \ceil{\half(N+1)}^2$.
-%\end{proof}
-
-\subsection{Obtaining the coefficients for expansion of a function on the disk-slice}
+\subsubsection{Obtaining the coefficients for expansion of a function on the disk-slice}
 
@@
 	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
+	\centering
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\centering
+	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
+	\centering
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
+	\centering
+	\end{subfigure}
+	\centering
+	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$.
+Left: Computed.
+Centre: Exact.
+Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
+	\centering
+	\label{fig:poissonexact}
+\end{figure}
+
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
 	\centering
@@
 which can be solved to find $\mathbf{u}$.
+In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
+In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
+
+In Figure \ref{fig:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error.
+The example was chosen so that the exact solution was $u(x,y) = \Wiii (x,y) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y) = \Delta \big[\Wiii (x,y) y^3 \exp(x)\big]$.
+We see that the computed solution is almost exact.
 
@@
 which can be solved to find $\mathbf{u}$.
+We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
+In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using N = 200, that is, 20,301 unknowns.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
+
+% For the final Gaussian bump example, we see that we achieve spectral convergence.
 
@@
 
+\subsubsection{Biharmonic equation}
+
+Find \(u(x,y)\) given a function \(f(x,y)\) such that:
+\begin{align}
+	\begin{cases}
+    		\Delta^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
+		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
+	\end{cases}.
+	\label{eqn:biharmonic}
+\end{align}
+where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
+In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
+In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
+We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Orthogonal Polynomials on spherical caps, spherical strips and spherical triangles (3D)}
+%
+%\section{Orthogonal polynomials on the circle arc}
+%
+%As a stepping stone to working on spherical triangles, we will try to construct orthogonal polynomials on a circle arc, given by $x \in [h, 1]$, $y = (1-x^2)^\frac{1}{2}$.
+We will use the Lanczos algorithm as implemented in the ApproxFun package \cite{ApproxFun} to obtain 1D OPs $\{T^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{-\frac{1}{2}}$) and $\{U^h_n\}$ (orthogonal wrt $w(x) = (1-x^2)^{\frac{1}{2}}$) on the interval $[h,1]$, and create our 2D OPs according to \cite{huybrechs2010fourier}:
+%\begin{align}
+%\bigP_0(x,y) := T^h_0(x) \equiv T_0,
+%\quad \quad 
+%\bigP_n(x,y) := \begin{bmatrix}
+%		T^h_n(x) \\
+%		y \: U^h_{n-1}(x)
+%	\end{bmatrix} \in \R^2 \quad \forall n \in \N,
+%\quad \quad 
+%\bigP := \begin{bmatrix}
+%		\bigP_0 \\
+%		\hline
+%		\bigP_1 \\
+%		\hline
+%		\bigP_2 \\
+%		\vdots \\
+%	\end{bmatrix}.
+%\end{align}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Orthogonal Polynomials on spherical caps and spherical triangles (3D)}
 
@@
 
-One approach we will look at will be to build the 3D OPs on the spherical cap by in a similar vein as to how we could build the circle arc OPs, by projecting up from the 2D orthogonal polynomials on the disk to 3D OPs on the sphere.
-
-Define the end-disk-slice, denoted by $\omega$, by
-\begin{align}
-	\omega := \{(x,y) \in \R^2 \quad | \quad \alpha \le x \le 1, \: - \rho(x) \le y \le \rho(x)\}.
-\end{align}
-where $\alpha \ge 0$ and $\rho(x) := (1-x^2)^{\half}$, $x \in (\alpha,1)$. Further, define the spherical cap, denoted by $\Omega$, by
-\begin{align}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
-\end{align}
-We can define the 3D OP family on $\Omega$ as the set of polynomials $\bigQab(x,y,z)$ where
-\begin{align}
-	\bigQab_n(x,y,z) := \begin{pmatrix}
-		\bighdopab_n(x,y) \\
-		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
-	\end{pmatrix} \in \R^{2n+1}, 
-	\quad \quad 
-	\bigQab := \begin{pmatrix}
-		\bigQab_0 \\
-		\bigQab_1 \\
-		\vdots \\
-		\bigQab_N
-	\end{pmatrix}.
-\end{align}
-The relations for multiplication by $x, y$ are then trivial.
+One approach would be to build the 3D OPs on the spherical cap by in a similar vein as to how we could build the circle arc OPs, by projecting up from the 2D orthogonal polynomials on the disk to 3D OPs on the sphere.
+The approach that we will first consider however will involve developing the 3D OPs on the spherical cap in a similar way to how we developed the 2D OPs on the disk-slice, similar to the approach of \cite{olver2019orthogonal}.
+
+Define the spherical cap, denoted by $\Omega$, by
+\begin{align}
+	\Omega := \{(\xvec, t) \in \R^3 \quad | \quad t \in (\alpha, 1), \: \xvec \in \R^2, \: ||\xvec|| = \rho(t)\}.
+\end{align}
+where $0 \le \alpha < 1$ and $\rho(t) := (1 - t^2)^\half$.
+
+Let $\genjacw^{(a,b)}(t)$ a weight function on the interval $(\alpha, 1)$ given by:
+\begin{align*}
+	\genjacw^{(a,b)}(t) &:= (t - \alpha)^{a} \: \rho(t)^{b}
+\end{align*}
+and define the associated inner product by:
+\begin{align}
+	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^\beta p(t) \: q(t) \: \genjacw^{(a,b)}(t) \: \D t \label{eqn:ipgenjactwoparam}
+\end{align}
+where
+\begin{align}
+	\normgenjac^{(a,b)} := \int_\alpha^\beta \: \genjacw^{(a,b)}(t) \: \D t.
+\end{align}
+Denote the two-parameter family of orthonormal polynomials on $[\alpha,1]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjactwoparam}).
+
+Let $\{Y_{m,k}\}$ for $m = 0, 1, \dots$, $k = 0, 1$ be the 2D OPs known as the circular harmonics on the unit circle; that is, if we define $\theta$ by $x = \cos(\theta)$, $y = \sin(\theta)$ for $\xvec := (x, y)$ where $||\xvec|| = 1$,
+\begin{align}
+	Y_{m,0}(\xvec) = Y_{m,0}(x, y) := \cos(m \theta), \quad Y_{m,1}(\xvec) = Y_{m,1}(x, y) := \sin(m \theta)
+\end{align}
+orthonormal with respect to the inner product
+\begin{align}
+	\ip< p, \: q >_{Y} &:= \frac{1}{\sqrt{\pi}} \: \int_{||\xvec|| = 1} p(\xvec) \: q(\xvec) \: \D \sigma(\xvec) = \frac{1}{\sqrt{\pi}} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
+\label{eqn:ipcircularharmonics}
+\end{align}
+We can then define the 3D family of OPs on $\Omega$ as the set of polynomials $\{\Qab_{n,m,k}\}$ where
+\begin{align}
+	\Qab_{n,m,k}(\xvec, t) := \genjac_{n-m}^{(a,2b+2m+1)}(t) \: \rho(t)^m \: Y_{m,0}\fpr(\frac{\xvec}{\rho(t)})
+\end{align}
+orthogonal on $\Omega$ with respect to
+\begin{align}
+	\ip< p, \: q >_{\Qab} &:= \int_\Omega p(\xvec, t) \: q(\xvec, t) \: \genjacw^{(a,2b)}(t) \: \D \sigma(\xvec, t) \nonumber \\ 
+	&= \int_\alpha^1 \int_{||\xvec|| = \rho(t)} p(\xvec, t) \: q(\xvec, t) \: \D \sigma(\xvec) \: \genjacw^{(a,2b)}(t) \: \D t \nonumber \\
+	&= \int_\alpha^1 \rho(t) \int_{||\xvec|| = 1} p(\rho(t) \xvec, t) \: q(\rho(t) \xvec, t) \: \D \sigma(\xvec) \: \genjacw^{(a,2b)}(t) \: \D t. \label{eqn:ipsphericalcaps}
+\end{align}
+
+%Define the end-disk-slice, denoted by $\omega$, by
+%\begin{align}
+%	\omega := \{(x,y) \in \R^2 \quad | \quad \alpha \le x \le 1, \: - \rho(x) \le y \le \rho(x)\}.
+%\end{align}
+%where $\alpha \ge 0$ and $\rho(x) := (1-x^2)^{\half}$, $x \in (\alpha,1)$. Further, define the spherical cap, denoted by $\Omega$, by
+%\begin{align}
+%	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
+%\end{align}
+%We can define the 3D OP family on $\Omega$ as the set of polynomials $\bigQab(x,y,z)$ where
+%\begin{align}
+%	\bigQab_n(x,y,z) := \begin{pmatrix}
+%		\bighdopab_n(x,y) \\
+%		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
+%	\end{pmatrix} \in \R^{2n+1}, 
+%	\quad \quad 
+%	\bigQab := \begin{pmatrix}
+%		\bigQab_0 \\
+%		\bigQab_1 \\
+%		\vdots \\
+%		\bigQab_N
+%	\end{pmatrix}.
+%\end{align}
+%The relations for multiplication by $x, y$ are then trivial.
 For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
-\begin{align}
-	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: \hdop^{(a,b+1)}_{m,j}(x,y), \\
-	z^2 \: \hdopnk^{(a,b+1)}(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b+1)\to(a,b)} \: \hdopmjab(x,y).
-\end{align}
-The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b+1)\to(a,b)}$ are given by
-\begin{align}
-	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: \hdop^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{\hdop^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
-	d_{n,k,m,j}^{(a,b+1)\to(a,b)} &= \frac{\ip< W^{(0,1)} \hdopnk^{(a,b+1)}, \: \hdopmjab >_\Wab}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
-        	&= \frac{\ip< \hdopnk^{(a,b+1)}, \: \hdopmjab >_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
-        	&= \frac{\norm{\hdopnk^{(a,b+1)}}^2_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \: c_{m,j,n,k}^{(a,b)\to(a,b+1)}.
-\end{align}
-The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ are just the entries to the parameter transformation operator matrix $T^{(a,b)\to(a,b+1)}$, defined by
-\begin{align}
-	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y),
-\end{align}
-which we have shown to be sparse with banded-block-banded structure, meaning the resulting operator matrix for the Jacobi operator for multiplication by $z$ here is sparse.
+%\begin{align}
+%	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: \hdop^{(a,b+1)}_{m,j}(x,y), \\
+%	z^2 \: \hdopnk^{(a,b+1)}(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b+1)\to(a,b)} \: \hdopmjab(x,y).
+%\end{align}
+%The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b+1)\to(a,b)}$ are given by
+%\begin{align}
+%	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: \hdop^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{\hdop^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
+%	d_{n,k,m,j}^{(a,b+1)\to(a,b)} &= \frac{\ip< W^{(0,1)} \hdopnk^{(a,b+1)}, \: \hdopmjab >_\Wab}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
+%        	&= \frac{\ip< \hdopnk^{(a,b+1)}, \: \hdopmjab >_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
+%        	&= \frac{\norm{\hdopnk^{(a,b+1)}}^2_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \: c_{m,j,n,k}^{(a,b)\to(a,b+1)}.
+%\end{align}
+%The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ are just the entries to the parameter transformation operator matrix $T^{(a,b)\to(a,b+1)}$, defined by
+%\begin{align}
+%	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y),
+%\end{align}
+%which we have shown to be sparse with banded-block-banded structure, meaning the resulting operator matrix for the Jacobi operator for multiplication by $z$ here is sparse.
 
@@
 This work is related to the triangle case \cite{olver2019triangle}, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements, for example, by dividing the disk into disk slice elements.
-
-This work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles.
-In particular, orthogonal polynomials (OPs) in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere.
+Other extensions to the disk-slice work could include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
+
+Moving forward, this work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles.
+In particular, orthogonal polynomials (OPs) in cartesian coordinates ($x$, $y$, and $z$) on a spherical cap can be represented using two families of OPs on the end-disk-slice (see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D), or from a rotation framework involving the non-classical 1D OPs defined in this document and circular harmonics (see \cite{olver2019orthogonal} for a similar construction of OPs in other quadratic surfaces of revolution).
+Further, we aim to show that the latter's construction leads to discretizations of spherical gradients and Laplacian's that are sparse on half-spheres and other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
 Constructing these sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is the next steps of my PhD, which has applications in weather prediction \cite{staniforth2012horizontal}.
-Other extensions to the disk-slice work could include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
 
@@
 %
-\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
+\section{$p$--finite element methods using sparse operators}\label{Appendix:PFEM}
 

### 2020-01-23 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
remove comments

--- 2020-01-10 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
@@
 \graphicspath{ {images/} }
-
-\newcommand{\sotodo}{\todo[color=green]}
-\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
-\newcommand{\bstodo}{\todo[color=pink]}
-\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}
 
@@
 Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
-In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on \bstodo{Reviewer 1, comment 3}disk slices and trapeziums.
+In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices and trapeziums.
 This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
 We apply the framework to solving the Poisson, variable coefficient Helmholtz, and Biharmonic equations.
-\bstodo{Reviewer 1, comment 3}In this paper we focus on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
+In this paper we focus on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
  \end{abstract}
@@
 
-For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss \bstodo{typo 1}an extension to other geometries in the appendix (including the half-disk and trapeziums).
+For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss an extension to other geometries in the appendix (including the half-disk and trapeziums).
 
@@
 Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
-In particular, by exploiting the connection with one-dimensional orthogonal polynomials \bstodo{typo 2}we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
+In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
 This compares favourably to $O(p^6)$ operations if one proceeds na\"ively.
 Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
-
-%An alternative approach is to use \sotodoinline{Compare with "Discontinuous Galerkin scheme for the spherical shallow water equations with applications to tsunami modeling and prediction"}
 
@@
 \end{align*}
-The sparsity and structure of this biharmonic operator \bstodo{typo 4}are seen in Figure \ref{fig:sparsity}.
+The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
 
@@
 
-\bstodo{Reviewer 1, comments 1,2}
 It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
@@
 In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
-
-%For the final Gaussian bump example, we see that we also achieve algebraic convergence.
-\bstodoinline{Note that it is no longer correct to say that we achieve spectral convergence for ex 4 for the disk slice.
-Is there something more we can say?
-Same goes for the other examples.}
 
@@
 We see that the computed solution is almost exact.
-\bstodo{Reviewer 2, comment 5}
 
@@
 
-\bstodo{Reviewer 2, comments 2, 3}
 We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
-\bstodo{Reviewer 1, comment 3} This work extends  the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using \bstodo{typo 5}suitably shaped elements, for example, by dividing the disk into disk slice elements.
+This work extends  the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements, for example, by dividing the disk into disk slice elements.
 %A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
@@
 
-Note \bstodo{typo 6}that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
 
@@
 
-In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$. \bstodo{Reviewer 2, comment 6}
+In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
 
@@
 
-\bstodo{typo 8 (capitalization of names in some reference titles}
 \bibliography{halfdisk}

### 2020-01-29 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
add acknowledgment

@@
 \title{Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums}
-\author{Sheehan Olver, Ben Snowball}
+\author{Sheehan Olver, Ben Snowball\thanks{Department of Mathematics, Imperial College, London, England}}
 %\date{}							% Activate to display a given date or no date
@@
 
-\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.  The second author was supported in part by a Leverhulme Trust Research Grant.
+\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.
+The second author was supported in part by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces.
 

### 2020-01-31 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
BigFloats produce BigFloats (full BF arithmitic used, and output for operators, coeffs, etc.)

@@ -20,9 +20,9 @@ export DiskSliceFamily, DiskSliceSpace
 
 # R should be Float64, B should be BigFloat
 abstract type DiskFamily{B,R,N} end
-struct DiskSlice{B,T} <: Domain{SVector{2,T}} end
-DiskSlice() = DiskSlice{BigFloat, Float64}()
-checkpoints(::DiskSlice) = [SVector(0.1,0.23), SVector(0.3,0.12)]
+struct DiskSlice{B,T} <: Domain{SVector{2,B}} end
+DiskSlice(::Type{B}, ::Type{T}) where {B,T} = DiskSlice{B,T}()
+checkpoints(::DiskSlice) = [SVector(0.23,-0.78), SVector(0.76,0.56)]
 
 struct DiskSliceSpace{DF, B, T, N} <: Space{DiskSlice{B,T}, T}
     family::DF # Pointer back to the family
@@ -33,10 +33,10 @@ struct DiskSliceSpace{DF, B, T, N} <: Space{DiskSlice{B,T}, T}
                                     # at the transform pts
     yderivopptseval::Vector{Vector{B}} # Store the y deriv of the ops evaluated
                                     # at the transform pts
-    A::Vector{SparseMatrixCSC{T}}
-    B::Vector{SparseMatrixCSC{T}}
-    C::Vector{SparseMatrixCSC{T}}
-    DT::Vector{SparseMatrixCSC{T}}
+    A::Vector{SparseMatrixCSC{B}}
+    B::Vector{SparseMatrixCSC{B}}
+    C::Vector{SparseMatrixCSC{B}}
+    DT::Vector{SparseMatrixCSC{B}}
 end
 
 function DiskSliceSpace(fam::DiskFamily{B,T,N}, params::NTuple{N,B}) where {B,T,N}
@@ -52,7 +52,7 @@ in(x::SVector{2}, D::DiskSlice) = D.α ≤ x[1] ≤ D.β && D.γ*D.ρ(x[1]) ≤ 
 
 spacescompatible(A::DiskSliceSpace, B::DiskSliceSpace) = (A.params == B.params)
 
-domain(::DiskSliceSpace) = DiskSlice()
+domain(::DiskSliceSpace{DF,B,T,N}) where {DF,B,T,N} = DiskSlice(B,T)
 
 # R should be Float64, B BigFloat
 struct DiskSliceFamily{B,T,N,FAR,FAP,F,I} <: DiskFamily{B,T,N}
[…]

### 2020-01-31 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl

@@ -117,6 +117,7 @@ function resizedata!(S::OrthogonalPolynomialSpace, n)
     resize!(S.b, n)
     resize!(S.ops, 2)
     # We set the weight here when this is called
+    @show "resizedata! for OPSpace", length(S.params)
     lanczos!(getweightfun(S), S.ops, S.a, S.b, N₀=N₀)
     S
 end

### 2020-01-31 experiments/diskslice.jl

@@ -43,6 +43,101 @@ F = SphericalCapFamily(α)
 a, b = 1.0, 1.0
 S = SphericalCapSpace(a, b)
 
+#====#
+
+# Model Problem: Δ(u*w)(x,y) = f(x,y) in Ω=halfdisk; u(x,y) ≡ 0 on ∂Ω.
+#   where w(x,y) = x*(1-x^2-y^2) is the weight of the D(1.0,1.0) basis.
+D
+S = D(a, b, c)
+
+evalu(u, z) = isindomain(z, u.space.family) ? u(z) : NaN
+isindomain(x, D::DiskSliceFamily) = D.α ≤ x[1] ≤ D.β && D.γ*D.ρ(x[1]) ≤ x[2] ≤ D.δ*D.ρ(x[1])
+x, y = 0.4, -0.765; z = [x; y]
+@test isindomain(z, D)
+
+# 1) u(x,y) ≡ c
+N = 20
+c = rand(1)[1]; f = Fun((x,y)->c * (d2xweight(S, x, y) + d2yweight(S, x, y)), S)
+U = Fun(S, sparse(Δw)[1:getopindex(N,N), 1:getopindex(N,N)] \ resizecoeffs!(f, N))
+U(z) - c
+@test U(z) ≈ c # Result u(x,y) where Δ(u*w)(x,y) = f(x,y)
+
+# 2) u(x,y) = x + y
+u = (x,y)->x + y
+ux = (x,y)->1.0
+uxx = (x,y)->0.0
+uy = (x,y)->1.0
+uyy = (x,y)->0.0
+N = 20
+f = Fun((x,y)->(d2xweight(S,x,y) * u(x,y) + 2 * dxweight(S,x,y) * ux(x,y) + weight(S,x,y) * uxx(x,y)
+                + d2yweight(S,x,y) * u(x,y) + 2 * dyweight(S,x,y) * uy(x,y) + weight(S,x,y) * uyy(x,y))
+        , S)
+f.coefficients
+U = Fun(S, sparse(Δw)[1:getopindex(N,N), 1:getopindex(N,N)] \ resizecoeffs!(f, N))
+U(z) - u(z...)
+@test U(z) ≈ u(z...)
+
+# 3) u(x,y) = y^3 * exp(x)
+u = (x,y)->y^3 * exp(x)
+ux = (x,y)->y^3 * exp(x)
+uxx = (x,y)->y^3 * exp(x)
[…]

### 2020-01-31 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
Initial attempt to tidy code/clarify some comments. Need to do same for TrapeziumFamily.

@@ -1,24 +1,7 @@
-# module DiskSliceFamilies
-
-# using ApproxFun
-#     import ApproxFun: evaluate, PolynomialSpace, recα, recβ, recγ, recA, recB, recC, domain,
-#                         domainspace, rangespace, bandwidths, prectype, canonicaldomain, tocanonical,
-#                         spacescompatible, points, transform, itransform, AbstractProductSpace, tensorizer,
-#                         columnspace, checkpoints, plan_transform, clenshaw
-#     import Base: in, *
-# using OrthogonalPolynomialFamilies
-# using StaticArrays
-# using FastGaussQuadrature
-# using LinearAlgebra
-# using SparseArrays
-# using BlockBandedMatrices
-# using BlockArrays
-# # using SingularIntegralEquations
-# using Test
 
 export DiskSliceFamily, DiskSliceSpace
 
-# R should be Float64, B should be BigFloat
+# R should be Float64, B should be BigFloat for best accuracy
 abstract type DiskFamily{B,R,N} end
 struct DiskSlice{B,T} <: Domain{SVector{2,B}} end
 DiskSlice(::Type{B}, ::Type{T}) where {B,T} = DiskSlice{B,T}()
@@ -33,7 +16,7 @@ struct DiskSliceSpace{DF, B, T, N} <: Space{DiskSlice{B,T}, T}
                                     # at the transform pts
     yderivopptseval::Vector{Vector{B}} # Store the y deriv of the ops evaluated
                                     # at the transform pts
-    A::Vector{SparseMatrixCSC{B}}
+    A::Vector{SparseMatrixCSC{B}} # A, B, C, DT store the clenshaw matrices (for function evaluation)
     B::Vector{SparseMatrixCSC{B}}
     C::Vector{SparseMatrixCSC{B}}
     DT::Vector{SparseMatrixCSC{B}}
@@ -54,7 +37,6 @@ spacescompatible(A::DiskSliceSpace, B::DiskSliceSpace) = (A.params == B.params)
 
 domain(::DiskSliceSpace{DF,B,T,N}) where {DF,B,T,N} = DiskSlice(B,T)
 
-# R should be Float64, B BigFloat
 struct DiskSliceFamily{B,T,N,FAR,FAP,F,I} <: DiskFamily{B,T,N}
[…]

### 2020-01-31 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl

@@ -15,7 +15,6 @@ using SparseArrays
 using BlockBandedMatrices
 using BlockArrays
 using GenericLinearAlgebra
-# using SingularIntegralEquations
 using Test
 
 export OrthogonalPolynomialFamily, HalfDisk, HalfDiskSpace, HalfDiskFamily
@@ -32,7 +31,7 @@ struct OrthogonalPolynomialSpace{FA,WW,F,D,B,R,N} <: PolynomialSpace{D,R}
     a::Vector{B} # Diagonal recurrence coefficients
     b::Vector{B} # Off diagonal recurrence coefficients
     opnorm::Vector{B} # The norm of the OPs (all OPs of an OPSpace have the same norm).
-                        # NOTE this is the value of the norm squared
+                      # NOTE this is the value of the norm squared
     opptseval::Vector{Vector{B}}
     derivopptseval::Vector{Vector{B}}
 end
@@ -97,6 +96,7 @@ OrthogonalPolynomialSpace(fam::SpaceFamily{D,R}, α::NTuple{N,B}) where {D,R,B,N
         fam, Vector{Fun}(), α, Vector{Fun}(), Vector{B}(), Vector{B}(),
         Vector{B}(), Vector{Vector{B}}(), Vector{Vector{B}}())
 
+# Creates and returns the Fun() representing the weight function for the OPSpace
 function getweightfun(S::OrthogonalPolynomialSpace)
     if length(S.weight) == 0
         # @show "getweightfun() for OPSpace", S.params
@@ -110,6 +110,7 @@ function getweightfun(S::OrthogonalPolynomialSpace)
     S.weight[1]
 end
 
+# Calls lanczos!() to get the recurrence coeffs for the OPSpace up to deg n
 function resizedata!(S::OrthogonalPolynomialSpace, n)
     N₀ = length(S.a)
     n ≤ N₀ && return S
@@ -146,6 +147,9 @@ function (P::OrthogonalPolynomialFamily{<:Any,<:Any,<:Any,R,B,N})(α::Vararg{B,N
     P.spaces[α] = OrthogonalPolynomialSpace(P, α)
 end
 
+#======#
+# Methods to return the recurrence coeffs
+
 #####
 # recα/β/γ are given by
 #       x p_{n-1} = γ_n p_{n-2} + α_n p_{n-1} +  p_n β_n
@@ -168,6 +172,8 @@ recB(::Type{T}, S::OrthogonalPolynomialSpace, n) where T =
[…]

### 2020-03-20 posters/jamboree2020/jamboree2020.tex
poster for jamboree on disk slices work

\documentclass{imposter}
%\documentclass[landscape]{imposter}

%\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
%\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex

\usepackage{wrapfig}								
\usepackage[font=scriptsize,labelfont=bf]{caption}
\usepackage[font=scriptsize]{subcaption}
\usepackage{float}
% \graphicspath{ {images/poster-eps/} }
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}

\usepackage{todonotes}

% \graphicspath{ {../papers/halfdisk/images/} }
\graphicspath{ {images/} }

\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

\newcommand{\rightboxnoborder}[2]{{\rput[tr](82.1,#1){\psframebox[linecolor=white]{\begin{minipage}{38.9cm}#2\end{minipage}}}}}
\newcommand{\lefttboxnoborder}[2]{{\rput[tl](0,#1){\psframebox[linecolor=white]{\begin{minipage}{38.9cm}#2\end{minipage}}}}}
\newcommand{\rightboxreferences}[2]{{\rput[tr](82.1,#1){\psframebox[linecolor=titleframecolor]{\begin{minipage}{38.9cm}#2\end{minipage}}}}}
\newcommand{\leftboxreferences}[2]{{\rput[tl](0,#1){\psframebox[linecolor=titleframecolor]{\begin{minipage}{38.9cm}#2\end{minipage}}}}}

%SetFonts

%SetFonts

\usepackage{natbib}
\setlength{\bibsep}{0.0pt}
\usepackage{url}
\bibliographystyle{abbrv}

\newcommand{\bigO}{\mathcal{O}}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\No}{\mathbb{N}_0}
\newcommand{\Ylm}{Y^m_l}
\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
\newcommand{\Plm}{\jac^m_l}
\newcommand{\costheta}{\cos\theta}
\newcommand{\sintheta}{\sin\theta}
\newcommand{\cosphi}{\cos\varphi}
\newcommand{\sinphi}{\sin\varphi}
\newcommand{\eimphi}{e^{im\varphi}}
\newcommand{\alphalm}{\alpha^m_l}
\newcommand{\clm}{c^m_l}
\newcommand{\ctilde}{\tilde{c}^m_l}
\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
\newcommand{\chat}{\hat{c}^m_l}
\newcommand{\chatmod}{\hat{c}^{|m|}_l}
\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
\newcommand{\ddy}{\frac{\mathrm{d}}{\mathrm{d}y}}
\newcommand{\pddx}{\frac{\partial}{\partial x}}
\newcommand{\pddy}{\frac{\partial}{\partial y}}
\newcommand{\pddn}{\frac{\partial}{\partial n}}
\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}

\newcommand{\Atilde}{\tilde{A}_{l,m}}
\newcommand{\Btilde}{\tilde{B}_{l,m}}
\newcommand{\Dtilde}{\tilde{D}_{l,m}}
\newcommand{\Etilde}{\tilde{E}_{l,m}}
\newcommand{\Ftilde}{\tilde{F}_{l,m}}
\newcommand{\Gtilde}{\tilde{G}_{l,m}}
\newcommand{\Alm}{A_{l,m}}
\newcommand{\Blm}{B_{l,m}}
\newcommand{\Dlm}{D_{l,m}}
\newcommand{\Elm}{E_{l,m}}
\newcommand{\Flm}{F_{l,m}}
\newcommand{\Glm}{G_{l,m}}

\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}

\newcommand{\bigP}{\mathbb{P}}
\newcommand{\Pl}{\mathbb{P}_l}
\newcommand{\gradP}{T\mathbb{P}}
\newcommand{\gradPl}{T\mathbb{P}_l}
\newcommand{\gradY}{\nabla Y}
\newcommand{\gradYlm}{\nabla Y^m_l}
\newcommand{\gradpY}{\nabla^\perp Y}
\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}

\newcommand{\Dlt}{D^\top_l}

\newcommand{\curlyy}{\bm{\mathcal{Y}}}
\newcommand{\blone}{\beta_{l, 1}}
\newcommand{\blzero}{\beta_{l, 0}}
\newcommand{\blmone}{\beta_{l, -1}}
\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
\newcommand{\cgcoeff}{\mathcal{C}}

\newcommand{\alm}{a_{l,m}}
\newcommand{\blm}{b_{l,m}}
\newcommand{\dlm}{d_{l,m}}
\newcommand{\elm}{e_{l,m}}
\newcommand{\flm}{f_{l,m}}
\newcommand{\glm}{g_{l,m}}
\newcommand{\hlm}{h_{l,m}}
\newcommand{\jlm}{j_{l,m}}
\newcommand{\klm}{k_{l,m}}
\newcommand{\almperp}{a_{l,m}^\perp}
\newcommand{\blmperp}{b_{l,m}^\perp}
\newcommand{\dlmperp}{d_{l,m}^\perp}
\newcommand{\elmperp}{e_{l,m}^\perp}
\newcommand{\flmperp}{f_{l,m}^\perp}
\newcommand{\glmperp}{g_{l,m}^\perp}
\newcommand{\hlmperp}{h_{l,m}^\perp}
\newcommand{\jlmperp}{j_{l,m}^\perp}
\newcommand{\klmperp}{k_{l,m}^\perp}

\newcommand{\unitvec}{\hat{\bm{k}}}

\newcommand{\hdop}{H}
\newcommand{\bighdop}{\mathbb{\hdop}}
\newcommand{\hdopnk}{\hdop_{n,k}}
\newcommand{\hdopnkab}{\hdop_{n,k}^{(a,b)}}
\newcommand{\hdopab}{\hdop^{(a,b)}}
\newcommand{\Wab}{{W^{(a,b)}}}
\newcommand{\hdopmj}{\hdop_{m,j}}
\newcommand{\hdopmjab}{\hdop_{m,j}^{(a,b)}}
\newcommand{\alphaab}{\alpha^{(a,b)}}
\newcommand{\betaab}{\beta^{(a,b)}}
\newcommand{\bighdopab}{\bighdop^{(a,b)}}
\newcommand{\Dnt}{D^\top_n}
\newcommand{\Wii}{W^{(1,1)}}
\newcommand{\hdopii}{\hdop^{(1,1)}}
\newcommand{\bighdopii}{{\mathbb{\hdop}^{(1,1)}}}
\newcommand{\hdopoo}{\hdop^{(0,0)}}
\newcommand{\bighdopoo}{{\mathbb{\hdop}^{(0,0)}}}
\newcommand{\hdopoi}{\hdop^{(0,1)}}
\newcommand{\hdopio}{\hdop^{(1,0)}}
%\newcommand{\jac}{{\tilde P}}
\newcommand{\jac}{{P}}
\newcommand{\genjac}{R}
\newcommand{\genjacnmk}{\genjac_{n-k}}
\newcommand{\genjacmmj}{\genjac_{m-j}}
\newcommand{\genjacw}{w_\genjac}
\newcommand{\jacw}{w_P}
\newcommand{\normgenjac}{\omega_\genjac}
\newcommand{\normjac}{\omega_P}

\newcommand{\dx}{\frac{\partial}{\partial x}}
\newcommand{\dy}{\frac{\partial}{\partial y}}
\newcommand{\laplacewii}{\Delta_W^{(1,1)\to(1,1)}}
\newcommand{\laplacewtt}{\Delta_W^{(2,2)\to(0,0)}}
\newcommand{\laplaceoo}{\Delta^{(0,0)\to(2,2)}}
\newcommand{\biharmonic}{_2\Delta_W^{(2,2)\to(2,2)}}

\newcommand{\element}{\tau}
\newcommand{\refelement}{\hat{\tau}}
\newcommand{\FEset}{\mathcal{T}}
\newcommand{\bigW}{\mathbb{W}}
\newcommand{\bigWab}{\mathbb{W}^{(a,b)}}
\newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
\newcommand{\bigQ}{\mathbb{Q}}
\newcommand{\bigQab}{\bigQ^{(a,b)}}
\newcommand{\Qab}{Q^{(a,b)}}

\newcommand{\hdopnkabc}{\hdop_{n,k}^{(a,b,c)}}
\newcommand{\hdopabc}{\hdop^{(a,b,c)}}
\newcommand{\Wabc}{{W^{(a,b,c)}}}
\newcommand{\hdopmjabc}{\hdop_{m,j}^{(a,b,c)}}
\newcommand{\alphaabc}{\alpha^{(a,b,c)}}
\newcommand{\betaabc}{\beta^{(a,b,c)}}
\newcommand{\bighdopabc}{\bighdop^{(a,b,c)}}
\newcommand{\Wiii}{W^{(1,1,1)}}
\newcommand{\hdopiii}{\hdop^{(1,1,1)}}
\newcommand{\bighdopiii}{{\mathbb{\hdop}^{(1,1,1)}}}
\newcommand{\hdopooo}{\hdop^{(0,0,0)}}
\newcommand{\bighdopooo}{{\mathbb{\hdop}^{(0,0,0)}}}
\newcommand{\hdopooi}{\hdop^{(0,0,1)}}
\newcommand{\hdopiio}{\hdop^{(1,1,0)}}
\newcommand{\laplacewiii}{\Delta_W^{(1,1,1)\to(1,1,1)}}
\newcommand{\laplacewttt}{\Delta_W^{(2,2,2)\to(0,0,0)}}
\newcommand{\laplaceooo}{\Delta^{(0,0,0)\to(2,2,2)}}
\newcommand{\biharmonictwo}{_2\Delta_W^{(2,2,2)\to(2,2,2)}}
\newcommand{\bigWiii}{{\mathbb{W}^{(1,1,1)}}}
\newcommand{\bigWabc}{\mathbb{W}^{(a,b,c)}}

\newcommand{\alphaabcd}{\alpha^{(a,b,c,d)}}
\newcommand{\betaabcd}{\beta^{(a,b,c,d)}}
\newcommand{\hdopnkabcd}{\hdop_{n,k}^{(a,b,c,d)}}
\newcommand{\hdopabcd}{\hdop^{(a,b,c,d)}}
\newcommand{\Wabcd}{{W^{(a,b,c,d)}}}
\newcommand{\hdopmjabcd}{\hdop_{m,j}^{(a,b,c,d)}}
\newcommand{\bighdopabcd}{\bighdop^{(a,b,c,d)}}
\newcommand{\bigWabcd}{\mathbb{W}^{(a,b,c,d)}}

\newcommand{\genjact}{\tilde{\genjac}}
\newcommand{\genjactnmk}{\genjact_{n-k}}
\newcommand{\genjactmmj}{\genjact_{m-j}}
\newcommand{\genjactw}{w_{\genjact}}
\newcommand{\normgenjact}{\omega_{\genjact}}

\newcommand{\xvec}{\mathbf{x}}

\usepackage{amsthm}

\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma} 
\newtheorem{theorem}{Theorem} 
\newtheorem{definition}{Definition}
\newtheorem{condition}{Condition}
\newtheorem{algorithm}{Algorithm}
\newtheorem{corollary}{Corollary}

\def\condref#1{Condition~\ref{cond:#1}}

\include{somacros}

\usepackage{listings}

\title{Sparse Spectral Methods for PDEs on Disk-Slices, Trapeziums and Spherical Bands}
\author{Ben Snowball, Sheehan Olver}
\department{Department of Mathematics, Imperial College London with the MPE CDT}

%% Primary university logo.
%\logo{imperial.eps}
% Secondary logos (not supported in all poster styles)
%\logoa{AMCG.eps}

%% Options to set colors of multiple poster features.
\titlecolor{named}{Blue}
\postercolor{named}{Blue}
%\pagecolor{green}

\posterstyle{mpecdtstyle}

% This sets up boxes to have nice rounded corners and somewhat thicker lines
% than the default.
\psset{linecolor=postercolor,linewidth=.2,cornersize=absolute,linearc=0.5,framesep=0.5}

\begin{document}
%\showgrid

% LHS

  \leftbox{96}{
        \section{Introduction}

In recent years, \textbf{sparse spectral methods} (SSMs) have been derived using hierarchies of classical \textbf{orthogonal polynomials} (OPs) on intervals, disks, and triangles (e.g. \cite{olver2019triangle}). We extend this methodology to a hierarchy of non-classical orthogonal polynomials on \textbf{disk slices} (e.g. a half-disk) and \textbf{trapeziums}. Sparsity is guaranteed due to the boundary being defined by an algebraic curve, and the entries of partial differential operators can be determined using formulae in terms of (non/semi-classical) univariate OPs. This work will lead to investigating \textbf{p-FEM} using disk-slice or trapezium shaped elements, and further to solving PDEs on the 3D \textbf{spherical band} or \textbf{polar cap}.
    }

  \leftbox{77.5}{
      \section{Construction of OPs on the disk-slice and trapezium}

Here, we develop SSMs for solving linear PDEs on a special class of geometries that includes disk-slices and trapeziums:
\begin{align*}
	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align*}
where either of the following conditions hold:
\begin{condition}\label{cond:trap}
	$\rho$ is a degree 1 polynomial.
\end{condition}
\begin{condition}\label{cond:disk}
	$\rho$ is the square root of a non-negative degree $\le$ 2 polynomial, and $-\gamma = \delta > 0$.
\end{condition}

We can construct 2D orthogonal polynomials on $\Omega$ by specialising the Proposition in \cite[p55--56]{dunkl2014orthogonal} in the following.

\begin{definition}
Let $\{\genjac_n^{(a,b,c)}\}$ be a three-parameter family of 1D OPs orthonormal w.r.t. the weight $\genjacw^{(a,b,c)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c}$ on $(\alpha,\beta)$, and let $\{\jac_n^{(a,b,c)}\}$ be a two-parameter family of 1D OPs orthonormal w.r.t. the weight $\jacw^{(a,b)}(x) := (\delta-x)^a \: (x - \gamma)^b$ on $(\gamma,\delta)$.

Define the \textbf{four-parameter 2D orthogonal polynomial family on $\Omega$} via:
\begin{align*}
	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align*}
which are orthogonal w.r.t. the weight
\begin{align*}
	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
\end{align*}
\end{definition}

A function $f: \Omega \to \R$ can then be expressed as a vector $\textbf{f} = \{f_{n,k}\}$ of coefficients in our OP basis: $f(x,y) \approx \sum_{n=0}^N \sum_{k=0}^n f_{n,k} \; \hdopnkabcd(x,y)$, for sufficiently large $N$ (with equality if $f$ is a degree $N$ polynomial).

\subsection{Example: Disk-Slice}

For a disk-slice, we have $(\gamma,\delta) := (-1,1)$ and $\rho(x) := (1-x^2)^{\half}$. Thus we can simply remove the need for including a fourth parameter $d$, since we are in Condition \ref{cond:disk}:
\begin{align*}\label{eq:diskpolys}
	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
\end{align*}

\noindent{\bf Remark}: In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while $\genjacw(x)$ is non-classical (it is in fact semi-classical, see \cite[\S5]{magnus1995painleve}).

    }

    \rightbox{96}{
    	\section{Sparse Operators}

\begin{wrapfigure}{L}{0.4\textwidth}
\center
\begin{tikzpicture} 
% Box/square (4x4)
\draw[black,solid,ultra thin] (0,0)--(0,4);
\draw[black,solid,ultra thin] (0,0)--(4,0);
\draw[black,solid,ultra thin] (4,0)--(4,4);
\draw[black,solid,ultra thin] (0,4)--(4,4);
% Dots
\draw[black,fill=black] (0,0) circle (.5ex);
\draw[black,fill=black] (2,2) circle (.5ex);
\draw[black,fill=black] (4,4) circle (.5ex);
\draw[black,fill=black] (0,2) circle (.5ex);
\draw[black,fill=black] (0,4) circle (.5ex);
% Arrows
\draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
\draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
\draw[black,thick,->] (0,0.2)--(0,1.8);
\draw[black,thick,->] (0,2.2)--(0,3.8);
\draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
% Node (parameter) labels
\draw[] (0,-0.5) node[anchor=east] {$(0,0,0)$};
\draw[] (0,4.5) node[anchor=east] {$(0,0,2)$};
\draw[] (4,-0.5) node[anchor=west] {$(2,2,0)$};
\draw[] (4,4.5) node[anchor=west] {$(2,2,2)$};
% Arrow labels
\draw[] (1.5,1.5) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (3.5,3.5) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (-2,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
\draw[] (-2,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
% The key and headings
\draw[] (2,5.3) node[anchor=south] {Parameter Conversion};
\end{tikzpicture} 
\label{fig:Laplace} 
\end{wrapfigure}

The construction leads to \textbf{sparse Jacobi operator matrices} for multiplication by $x$, $y$. We can hence derive a recurrence relation and construct methods to expand functions in our OP basis, evaluate functions using the \textbf{Clenshaw algorithm}, derive sparse relations for partial derivatives of the OPs. This is implemented in a code library that utilises ApproxFun \cite{ApproxFun}. \\

We are also guaranteed sparsity of the partial differential relationships, and hence obtain \textbf{sparse differential operator matrices} with \textbf{banded-block-banded} structure. \\

\begin{wrapfigure}{R}{0.7\textwidth}
	\centering
	\begin{subfigure}{0.2\textwidth}
	\centering
	\includegraphics[scale=0.6]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.2\textwidth}
	\includegraphics[scale=0.6]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.2\textwidth}
	\includegraphics[scale=0.6]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\captionsetup{width=.8\linewidth}
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. Left: the Laplace operator. Centre: the weighted variable coefficient Helmholtz operator $\Delta + k^2 \: v$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. Right: the biharmonic operator.}
        \label{fig:sparsity}
        \centering
\end{wrapfigure}

The \textbf{Laplacian operator} for example can be constructed as a combination of the sparse partial differential operators and \textbf{sparse parameter increment/decrement operators}. The incrementing and decrementing of parameters here is analogous to other well known OP families' derivatives (see \cite[(18.9.3)]{DLMF}, \cite{olver2018recurrence}).

}

	\rightboxnoborder{51}{

%$ $
%
%\begin{wrapfigure}{L}{\textwidth}
%	\centering
%        \begin{subfigure}{0.35\linewidth}
%        \centering
%        \includegraphics[scale=0.6]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
%        \end{subfigure}
%        \centering
%        \begin{subfigure}{.15\linewidth}
%        \centering
%        \includegraphics[scale=0.55]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
%        \end{subfigure}
%        \centering
%        \begin{subfigure}{.15\linewidth}
%        \centering
%        \includegraphics[scale=0.55]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
%        \end{subfigure}
%        \centering
%        \begin{subfigure}{.15\linewidth}
%        \centering
%        \includegraphics[scale=0.58]{solution-trapezium-helmholtz-k=100-n=1500}
%        \end{subfigure}
%        \centering
%        \captionsetup{width=.8\linewidth}
%        \caption{Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders. Right 3: The computed solutions $u$ with zero BCs to: Left: $\Delta u = f$, $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Centre: $\Delta u + k^2 \: v \: u = f$, $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$, $k = 100$. Right: $\Delta u + k^2 \: v \: u = f$, $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$, $k = 100$.}
%        \centering
%\end{wrapfigure}
%
%$$
%$$

\begin{center}
	\includegraphics[scale=0.73]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=950-f64}
	\quad
	\includegraphics[scale=0.65]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
	\includegraphics[scale=0.65]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
	\includegraphics[scale=0.69]{solution-trapezium-helmholtz-k=100-n=1500}
\end{center}
$ $\\
\scriptsize{\textbf{Figure 2: } Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates \textbf{algebraic convergence} with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders. Right 3: The computed solutions $u$ with zero BCs to: Left: $\Delta u = f$, $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Centre: $\Delta u + k^2 \: v \: u = f$, $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$, $k = 100$. Right: $\Delta u + k^2 \: v \: u = f$, $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$, $k = 100$.}

}

      \rightbox{28}{
      \section{OPs on the spherical cap and spherical band}

We can develop a similar framework on the spherical cap and spherical band. The 3D OPs will be simply the two-parameter family given by
\begin{align*}
	\tilde{Q}^{(a,b)}_{n,k}(x, y, z) \equiv \Qab_{n,k}(\theta, z) &:= \genjac_{n-k}^{(a,b,2k+1)}(z) \: \rho(z)^k \: e^{i k \theta}, \quad (x,y,z) \in \Omega
\end{align*}
where $\Omega := \{(x,y,z) \in \R^3 \quad | \quad z \in (\alpha, \beta), \: ||(x,y)|| = \rho(z)\}$, and $\theta, \phi$ are defined according to $x = \cos \theta \sin \phi$, $y = \sin \theta \sin \phi$, $z = \cos \phi$. Note the comparison to the \textbf{spherical harmonics} defined on the whole sphere.

    }

    \leftboxreferences{16}{
        {\footnotesize\bibliography{halfdisk}}

    }

     \rightbox{12.5}{
      \subsection{Summary and Future Work}

\begin{itemize}
	\item Presented a method for obtaining families of 2D OPs on certain domains. 
	\item Using the three-term recurrence relationships, we can calculate sparse operator matrices, allowing us to efficiently solve PDEs on these domains. 
	\item We can extend this to p-FEM on the disk using disk-slice elements for example.
	\item Next goal is to solve PDEs on the spherical cap and spherical band.
\end{itemize}

    }

%    
%        \rightbox{50}{
%
%\begin{wrapfigure}{R}{1.0\textwidth}
%	\centering
%	\begin{subfigure}{0.3\textwidth}
%	\includegraphics[scale=0.7]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
%	\centering
%	%\label{fig:solution-poisson}
%	\end{subfigure}
%	\centering
%	\begin{subfigure}{0.3\textwidth}
%	\centering
%	\includegraphics[scale=0.7]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
%	%\label{fig:solution-helmholtz}
%	\end{subfigure}
%	\begin{subfigure}{0.3\textwidth}
%	\includegraphics[scale=0.8]{solution-trapezium-helmholtz-k=100-n=1500}
%	%\label{fig:solution-poisson}
%	\centering
%	\end{subfigure}
%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Centre: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$ in the trapezium. Code utilises ApproxFun \cite{ApproxFun}}
%	\centering
%	\label{fig:solutions}
%\end{wrapfigure}
%
%\begin{wrapfigure}{R}{0.7\textwidth}
%	\centering
%	\begin{subfigure}{0.2\textwidth}
%	\centering
%	\includegraphics[scale=0.6]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
%        \centering
%	\end{subfigure}
%	\begin{subfigure}{0.2\textwidth}
%	\includegraphics[scale=0.6]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
%        \centering
%	\end{subfigure}
%	\begin{subfigure}{0.2\textwidth}
%	\includegraphics[scale=0.6]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
%        \centering
%	\end{subfigure}
%    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. Left: the Laplace operator. Centre: the weighted variable coefficient Helmholtz operator $\Delta + k^2 \: v$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. Right: the biharmonic operator.}
%        \label{fig:sparsity}
%        \centering
%\end{wrapfigure}
%
%\begin{wrapfigure}{R}{0.6\textwidth}
%	\centering
%	\begin{subfigure}{0.5\textwidth}
%	\centering
%	\includegraphics[scale=0.7]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
%        	%\label{fig:solutionblocknorms-poisson}
%	\end{subfigure}
%	\caption{The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders. Code utilises ApproxFun \cite{ApproxFun}}
%	\centering
%	\label{fig:solutionblocknorms}
%\end{wrapfigure}
%
%    }

\end{document}

### 2020-03-20 OrthogonalPolynomialFamilies/src/OrthogonalPolynomialFamilies.jl
More work on Sph Caps code

@@ -337,10 +337,10 @@ resetopptseval(S::OrthogonalPolynomialSpace) = resize!(S.opptseval, 0)
 # This func returns the S.opptseval for the degree n OP
 # We assume that we have called getopptseval(S, N, pts)
 function getptsevalforop(S::OrthogonalPolynomialSpace, n)
-    if length(S.opevalatpts) > n+1
-        error('Invalid OP requested in getptsevalforop - getopptseval(S,N,pts) may not have been correctly called')
+    if length(S.opptseval) < n+1
+        error("Invalid OP requested in getptsevalforop - getopptseval(S,N,pts) may not have been correctly called")
     else
-        S.opevalatpts[n+1]
+        S.opptseval[n+1]
     end
 end
 
@@ -402,7 +402,7 @@ resetderivopptseval(S::OrthogonalPolynomialSpace) = resize!(S.derivopptseval, 0)
 # We assume that we have called getderivopptseval(S, N, pts)
 function getderivptsevalforop(S::OrthogonalPolynomialSpace, n)
     if length(S.derivopptseval) > n+1
-        error('Invalid OP requested in getderivptsevalforop - getderivopptseval(S,N,pts) may not have been correctly called')
+        error("Invalid OP requested in getderivptsevalforop - getderivopptseval(S,N,pts) may not have been correctly called")
     else
         S.derivopptseval[n+1]
     end
@@ -1423,7 +1423,7 @@ function resizecoeffs!(f::Fun, N)
         cfs[m̃+1:end] .= 0.0
     elseif m̃ > m
         for j = m+1:m̃
-            if cfs[j] > 1e-15
+            if cfs[j] > 1e-16
                 error("Trying to decrease degree of f")
             end
         end
@@ -1742,6 +1742,6 @@ weightderivativey(S::HalfDiskSpace, z) = weightderivativey(S, z[1], z[2])
 # Include other OP Families
 include("DiskSliceFamilies.jl")
 include("TrapeziumFamilies.jl")
-# include("SphericalCaps.jl")
+include("SphericalCaps.jl")
 
 end # module

### 2020-03-20 OrthogonalPolynomialFamilies/src/SphericalCaps.jl

@@ -12,51 +12,58 @@ mode (not by n but by k)
 
 i.e. for a given max degree N:
 
-    ℚ^{(a,b)}_N := [ℚ^{(a,b)}_{N,N};...;ℚ^{(a,b)}_{N,0}] ∈ ℂ^{(N+1)(N+2)/2}
-    ℚ^{(a,b)}_{N,k} := [Q^{(a,b)}_{N,k};...;Q^{(a,b)}_{k,k}] ∈ ℂ^{N-k+1}
+    ℚ^{(a,b)}_N := [ℚ^{(a,b)}_{N,0};...;ℚ^{(a,b)}_{N,N}] ∈ ℂ^{(N+1)(N+2)/2}
+    ℚ^{(a,b)}_{N,k} := [Q^{(a,b)}_{k,k};...;Q^{(a,b)}_{N,k}] ∈ ℂ^{N-k+1}
                                                             for k = 0,...,N
 
 =#
 
 export SphericalCapFamily, SphericalCapSpace
 
-# R should be Float64, B should be BigFloat
-abstract type SphericalFamily{B,R,N} end
+# T should be Float64, B should be BigFloat
+
+# TODO Sort out checkpoints, in, and domain - should
+abstract type SphericalFamily{B,T,N} end
 struct SphericalCap{B,T} <: Domain{SVector{2,T}} end
 SphericalCap() = SphericalCap{BigFloat, Float64}()
-checkpoints(::SphericalCap) = [SVector(0.1, 0.23), SVector(0.3, 0.12)]
+function checkpoints(::SphericalCap)
+    # Return 2 3D points that will lie on the domain TODO
+    y, z = -0.3, 0.8; x = sqrt(1 - z^2 - y^2); p1 = SVector(x, y, z)
+    y, z = 0.4, 0.8; x = sqrt(1 - z^2 - y^2); p2 = SVector(x, y, z)
+    [p1, p2]
+end
+in(x::SVector{3}, D::SphericalCap) = D.α ≤ x[3] ≤ D.β && sqrt(x[1]^2 + x[2]^2) == D.ρ(x[3])
 
-struct SphericalCapSpace{DF, B, T, N} <: Space{SphericalCap{B,T}, T}
+struct SphericalCapSpace{DF, B, T, CB, N} <: Space{SphericalCap{B,T}, T}
     family::DF # Pointer back to the family
     params::NTuple{N,B} # Parameters
-    opnorms::Vector{T} # squared norms
[…]

### 2020-05-26 papers/disk-slices-and-trapeziums/proof/finalcomments.tex
final comments

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage[margin=10pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mleftright}
\usepackage{todonotes}

\include{somacros}

\begin{document}

Queries:

1) Yes

2) Yes

3) Reference Figure 1 after Definition 4.
	"... and on the triangle. An illustration of how the non-weighted differential operators increment the parameters (a,b,c) is seen in Figure 1"

4) OK

5) Should be ``2020, IMA. J. Numer. Anal., https://doi.org/10.1093/imanum/draa001''

6) Replace the problematic sentence with the following:
	"Recall that the OPs $\mathbb{H}^{(a,b,c)}$ are orthogonal with respect to the weight $W^{(a,b,c)}$ on $\Omega$, and define the matrix $\Lambda^{(a,b,c)} := \ip< \mathbb{H}^{(a,b,c)}, \: {\mathbb{H}^{(a,b,c)}}^\top >_{W^{(a,b,c)}}$"

Typos:

1) In Lemma 1, it should say 
$$
\alpha^{(a,b,c,d)}_{n,k,1} := \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alpha^{(a,b,c,d)}_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)}
$$

2) At the end of Section 3, we are missing a couple of minus signs. We should have:
\begin{align*}
	\eta_j &= - \frac{1}{\beta^{(a,b,c,d)}_{n,j-1,9}} \big( \beta^{(a,b,c,d)}_{n,n+j+1,7} \: \eta_{j+2} + \beta^{(a,b,c,d)}_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
	\eta_0 &= - \frac{1}{\alpha^{(a,b,c,d)}_{n+1,0,1}} \big( \beta^{(a,b,c,d)}_{n,1,7} \: \eta_{2} + \beta^{(a,b,c,d)}_{n,0,8} \: \eta_{1} \big).
\end{align*}

3) Near the end of the proof for Theorem 1, it should read:
	"... nonzero coefficients $c^x_{m,j}$ are when $m = n-3, n-2, n-1$ and $j = k-2, k$"
(and not $m = n-3,...,n$)

4) Near the end of Section 3, it should read:
	"... for the Laplacian $\Delta$ that will take us from the coefficients for expansion in the space $\mathbb{H}^{(0,0,0)}$ to coefficients in the space $\mathbb{H}^{(2,2,2)}$"
(and not $\mathbb{H}^{(2,2)}$)

5) Similarly, it should read:
	"... Laplacian as a map from coefficients in the space $\mathbb{W}^{(2,2,2)}$ to coefficients in the space $\mathbb{H}^{(0,0,0)}$. Note a function expanded in the $\mathbb{W}^{(2,2,2)}$ basis will satisfy ..."
(and not $\mathbb{W}^{(2,2)}$)

Additional changes for Figures:

1) Figure 3b will be a new figure, that increases the N value used in the original. We would also then need to update the text in Section 5.1 that reads "N = 200, that is, 20,301 unknowns" to instead be "N = 990, that is, 491,536 unknowns"

2) Figure 5b will be a new figure, that increases the N value used in the original. We would also then need to update the text in Section 5.1 that reads "N = 200, that is, 20,301 unknowns" to instead be "N = 500, that is, 125,751 unknowns" (Note: Depends how high I can go before I run out of memory with the factorization/time)

3) Figure 6b will be a new figure, that increases the N value used in the original. We would also then need to update the text in Section 5.1 that reads "N = 200, that is, 20,301 unknowns" to instead be "N = 500, that is, 125,751 unknowns" (Note: Depends how high I can go before I run out of memory with the factorization/time)

\end{document}

### 2020-05-27 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
Laplacian (without rho2 factor) now implemented for special case (weighted, param a=1)

@@ -180,12 +180,30 @@ function getopnorms(S::DiskSliceSpace{<:Any, B, T, <:Any}, k) where {B,T}
         P = getPspace(S)
         getopnorm(P)
 
-        # for j = m+1:k+1
-        #     S.opnorms[j] = getopnorm(getRspace(S, j-1)) * P.opnorm[1]
-        # end
-
         # We get the weights for the R integral
         ptsr, wr = pointswithweights(B, getRspace(S, 0), k+1)
+        rhoptsr = S.family.ρ.(ptsr)
+        for j = m:k
+            R = getRspace(S, j)
+            if !isnormset(R)
+                setopnorm(R, sum(rhoptsr.^(2j) .* wr))
+            end
+            S.opnorms[j+1] = R.opnorm[1] * P.opnorm[1]
+        end
+    end
+    S
+end
+function getopnorms(S::DiskSliceSpace{<:Any, B, T, <:Any}, k::Int,
+                    ptsr::Vector{B}, wr::Vector{B}) where {B,T}
+    # NOTE Here ptsr, wr = pointswithweights(B, getRspace(S, 0), k+1)
+    # NOTE these are squared norms
+    m = length(S.opnorms)
+    if k + 1 > m
+        resize!(S.opnorms, k+1)
+        P = getPspace(S)
+        getopnorm(P)
+
+        # We get the weights for the R integral
         rhoptsr = S.family.ρ.(ptsr)
         for j = m:k
             R = getRspace(S, j)
@@ -351,8 +369,57 @@ function recα(::Type{T}, S::DiskSliceSpace, n, k, j) where T
     end
 end
 
+function recβ(::Type{T}, S::DiskSliceSpace, n, k, j, pts, w) where T
+    # NOTE Here we expect that pts, w are the pts, weights wrt weight of getRspace(S, 0)
+
+    getopnorms(S, k+1, pts, w)
+
+    R1 = getRspace(S, k-1)
+    R2 = getRspace(S, k)
+    R3 = getRspace(S, k+1)
+    P = getPspace(S)
+    getopnorm(P)
+
+    if isodd(j)
+        # pts, w = pointswithweights(T, R2, Int(ceil(n-k+1.5)))
[…]

### 2020-05-27 OrthogonalPolynomialFamilies/src/SphericalCaps.jl

@@ -25,6 +25,8 @@ i.e. for a given max degree N:
 
 =#
 
+# TODO there should only be one parameter, not two (as b is not used)
+
 export SphericalCapFamily, SphericalCapSpace
 
 # T should be Float64, B should be BigFloat
@@ -1200,7 +1202,6 @@ end
 function transformparamsoperator(S::SphericalCapSpace{<:Any, B, T, <:Any},
             St::SphericalCapSpace{<:Any, B, T, <:Any}, N;
             weighted=false) where {B,T}
-    # TODO do the other parameter conversions
 
     # The St space is the target space. Applying this operator to coeffs in the
     # (weighted) S space will result in coeffs in the (weighted) St space.
@@ -1242,7 +1243,25 @@ function transformparamsoperator(S::SphericalCapSpace{<:Any, B, T, <:Any},
     end
     C
 end
-
+function convertweightedtononweightedoperator(S::SphericalCapSpace, N; Nout=-1)
+    # Converts coeffs in W^{a} to coeffs in Q^{a}
+    # If Nout is set, then we increase the resulting dimension so that the
+    # operator outputs coeffs of degree Nout (probably to be used for Helholtz
+    # type equations, to match dimension of Laplacian, in which case we should
+    # have Nout=N+3)
+    S0 = S.family(0.0, 0.0)
+    T = transformparamsoperator(S0, S, N+1; weighted=false)
+    Tw = transformparamsoperator(S, S0, N; weighted=true)
+    if Nout > 0 # ie is set
+        if Nout <= N
+            error("invalid Nout value given - must be bigger than N")
+        else
+            increasedegreeoperator(S, N+1, Nout) * T * Tw
+        end
+    else
+        T * Tw
+    end
+end
 
 #===#
 # "Coeffs degree increaser" operator matrix
@@ -1270,31 +1289,91 @@ end
 #===#
 # Spherical Laplacian operator matrix
 
-# TODO this is currently the operator for ρ²Δ
-function laplaceoperator(S::SphericalCapSpace, St::SphericalCapSpace, N;
-                            weighted=true, square=false)
+function getlaplacianval(S::SphericalCapSpace{<:Any, B, T, <:Any},
[…]

### 2020-05-27 experiments/sphericalcaptests.jl

@@ -7,72 +7,89 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
                     opevalatpts, inner2, getopindex, getnk, resizecoeffs!,
                     transformparamsoperator, weightedpartialoperatorx,
                     weightedpartialoperatory, partialoperatorx, partialoperatory,
-                    laplaceoperator, derivopevalatpts, getderivopptseval, getopnorms,
+                    derivopevalatpts, getderivopptseval, getopnorms,
                     getopnorm, operatorclenshaw, weight, biharmonicoperator,
                     getPspace, getRspace, differentiatespacex, differentiatespacey,
                     differentiateweightedspacex, differentiateweightedspacey,
                     resizedata!,
                     resizedataonedimops!, getnki, convertcoeffsvecorder,
                     diffoperatorphi, diffoperatortheta, diffoperatortheta2,
-                    differentiatespacephi, increasedegreeoperator
+                    differentiatespacephi, increasedegreeoperator,
+                    getptsevalforop, getderivptsevalforop, laplacianoperator,
+                    rholaplacianoperator
 using JLD
 
 
-# Useful functions for testing
-function converttopseudo(S::SphericalCapSpace, cfs; converttobydegree=true)
-    N = getnki(S, length(cfs))[1]
-    if converttobydegree
-        PseudoBlockArray(convertcoeffsvecorder(S, cfs), [2n+1 for n=0:N])
-    else
-        PseudoBlockArray(cfs, [2n+1 for n=0:N])
-    end
-end
 isindomain(pt, D::SphericalCapFamily) = D.α ≤ pt[3] ≤ D.β && norm(pt) == 1.0
 isindomain(pt, S::SphericalCapSpace) = isindomain(pt, S.family)
-"""
[…]

### 2020-07-30 papers/spherical-caps/spherical-caps.tex
First draft at notes on spherical caps (basis of a paper). Extra section 7 on tangent space notes.

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage[margin=10pt,font=footnotesize,labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mleftright}
\usepackage{todonotes}

%\bibliographystyle{apalike}

\graphicspath{ {images/} }

\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

%SetFonts

%SetFonts

\usepackage[square,numbers]{natbib}
\usepackage{url}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plain}

\newcommand{\bigO}{\mathcal{O}}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\No}{\mathbb{N}_0}
\newcommand{\Ylm}{Y^m_l}
\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
\newcommand{\Plm}{\jac^m_l}
\newcommand{\costheta}{\cos\theta}
\newcommand{\sintheta}{\sin\theta}
\newcommand{\cosphi}{\cos\varphi}
\newcommand{\sinphi}{\sin\varphi}
\newcommand{\eimphi}{e^{im\varphi}}
\newcommand{\alphalm}{\alpha^m_l}
\newcommand{\clm}{c^m_l}
\newcommand{\ctilde}{\tilde{c}^m_l}
\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
\newcommand{\chat}{\hat{c}^m_l}
\newcommand{\chatmod}{\hat{c}^{|m|}_l}
\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
\newcommand{\ddy}{\frac{\mathrm{d}}{\mathrm{d}y}}
\newcommand{\pddx}{\frac{\partial}{\partial x}}
\newcommand{\pddy}{\frac{\partial}{\partial y}}
\newcommand{\pddn}{\frac{\partial}{\partial n}}
\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}

\newcommand{\Atilde}{\tilde{A}_{l,m}}
\newcommand{\Btilde}{\tilde{B}_{l,m}}
\newcommand{\Dtilde}{\tilde{D}_{l,m}}
\newcommand{\Etilde}{\tilde{E}_{l,m}}
\newcommand{\Ftilde}{\tilde{F}_{l,m}}
\newcommand{\Gtilde}{\tilde{G}_{l,m}}
\newcommand{\Alm}{A_{l,m}}
\newcommand{\Blm}{B_{l,m}}
\newcommand{\Dlm}{D_{l,m}}
\newcommand{\Elm}{E_{l,m}}
\newcommand{\Flm}{F_{l,m}}
\newcommand{\Glm}{G_{l,m}}

\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}

\newcommand{\bigP}{\mathbb{P}}
\newcommand{\Pl}{\mathbb{P}_l}
\newcommand{\gradP}{T\mathbb{P}}
\newcommand{\gradPl}{T\mathbb{P}_l}
\newcommand{\gradY}{\nabla Y}
\newcommand{\gradYlm}{\nabla Y^m_l}
\newcommand{\gradpY}{\nabla^\perp Y}
\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}

\newcommand{\Dlt}{D^\top_l}

\newcommand{\curlyy}{\bm{\mathcal{Y}}}
\newcommand{\blone}{\beta_{l, 1}}
\newcommand{\blzero}{\beta_{l, 0}}
\newcommand{\blmone}{\beta_{l, -1}}
\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
\newcommand{\cgcoeff}{\mathcal{C}}

\newcommand{\alm}{a_{l,m}}
\newcommand{\blm}{b_{l,m}}
\newcommand{\dlm}{d_{l,m}}
\newcommand{\elm}{e_{l,m}}
\newcommand{\flm}{f_{l,m}}
\newcommand{\glm}{g_{l,m}}
\newcommand{\hlm}{h_{l,m}}
\newcommand{\jlm}{j_{l,m}}
\newcommand{\klm}{k_{l,m}}
\newcommand{\almperp}{a_{l,m}^\perp}
\newcommand{\blmperp}{b_{l,m}^\perp}
\newcommand{\dlmperp}{d_{l,m}^\perp}
\newcommand{\elmperp}{e_{l,m}^\perp}
\newcommand{\flmperp}{f_{l,m}^\perp}
\newcommand{\glmperp}{g_{l,m}^\perp}
\newcommand{\hlmperp}{h_{l,m}^\perp}
\newcommand{\jlmperp}{j_{l,m}^\perp}
\newcommand{\klmperp}{k_{l,m}^\perp}

\newcommand{\unitvec}{\hat{\bm{k}}}

\newcommand{\hdop}{H}
\newcommand{\bighdop}{\mathbb{\hdop}}
\newcommand{\hdopnk}{\hdop_{n,k}}
\newcommand{\hdopnkab}{\hdop_{n,k}^{(a,b)}}
\newcommand{\hdopab}{\hdop^{(a,b)}}
\newcommand{\Wab}{{W^{(a,b)}}}
\newcommand{\hdopmj}{\hdop_{m,j}}
\newcommand{\hdopmjab}{\hdop_{m,j}^{(a,b)}}
\newcommand{\alphaab}{\alpha^{(a,b)}}
\newcommand{\betaab}{\beta^{(a,b)}}
\newcommand{\bighdopab}{\bighdop^{(a,b)}}
\newcommand{\Dnt}{D^\top_n}
\newcommand{\Wii}{W^{(1,1)}}
\newcommand{\hdopii}{\hdop^{(1,1)}}
\newcommand{\bighdopii}{{\mathbb{\hdop}^{(1,1)}}}
\newcommand{\hdopoo}{\hdop^{(0,0)}}
\newcommand{\bighdopoo}{{\mathbb{\hdop}^{(0,0)}}}
\newcommand{\hdopoi}{\hdop^{(0,1)}}
\newcommand{\hdopio}{\hdop^{(1,0)}}
%\newcommand{\jac}{{\tilde P}}
\newcommand{\jac}{{P}}
\newcommand{\genjac}{R}
\newcommand{\genjacnmk}{\genjac_{n-k}}
\newcommand{\genjacmmj}{\genjac_{m-j}}
\newcommand{\genjacw}{w_\genjac}
\newcommand{\jacw}{w_P}
\newcommand{\normgenjac}{\omega_\genjac}
\newcommand{\normjac}{\omega_P}

\newcommand{\dx}{\frac{\partial}{\partial x}}
\newcommand{\dy}{\frac{\partial}{\partial y}}
\newcommand{\laplacewii}{\Delta_W^{(1,1)\to(1,1)}}
\newcommand{\laplacewtt}{\Delta_W^{(2,2)\to(0,0)}}
\newcommand{\laplaceoo}{\Delta^{(0,0)\to(2,2)}}
\newcommand{\biharmonic}{_2\Delta_W^{(2,2)\to(2,2)}}

\newcommand{\element}{\tau}
\newcommand{\refelement}{\hat{\tau}}
\newcommand{\FEset}{\mathcal{T}}
\newcommand{\bigW}{\mathbb{W}}
\newcommand{\bigWab}{\mathbb{W}^{(a,b)}}
\newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
\newcommand{\bigQ}{\mathbb{Q}}
\newcommand{\bigQab}{\bigQ^{(a,b)}}

\newcommand{\hdopnkabc}{\hdop_{n,k}^{(a,b,c)}}
\newcommand{\hdopabc}{\hdop^{(a,b,c)}}
\newcommand{\Wabc}{{W^{(a,b,c)}}}
\newcommand{\hdopmjabc}{\hdop_{m,j}^{(a,b,c)}}
\newcommand{\alphaabc}{\alpha^{(a,b,c)}}
\newcommand{\betaabc}{\beta^{(a,b,c)}}
\newcommand{\bighdopabc}{\bighdop^{(a,b,c)}}
\newcommand{\Wiii}{W^{(1,1,1)}}
\newcommand{\hdopiii}{\hdop^{(1,1,1)}}
\newcommand{\bighdopiii}{{\mathbb{\hdop}^{(1,1,1)}}}
\newcommand{\hdopooo}{\hdop^{(0,0,0)}}
\newcommand{\bighdopooo}{{\mathbb{\hdop}^{(0,0,0)}}}
\newcommand{\hdopooi}{\hdop^{(0,0,1)}}
\newcommand{\hdopiio}{\hdop^{(1,1,0)}}
\newcommand{\laplacewiii}{\Delta_W^{(1,1,1)\to(1,1,1)}}
\newcommand{\laplacewttt}{\Delta_W^{(2,2,2)\to(0,0,0)}}
\newcommand{\laplaceooo}{\Delta^{(0,0,0)\to(2,2,2)}}
\newcommand{\biharmonictwo}{_2\Delta_W^{(2,2,2)\to(2,2,2)}}
\newcommand{\bigWiii}{{\mathbb{W}^{(1,1,1)}}}
\newcommand{\bigWabc}{\mathbb{W}^{(a,b,c)}}

\newcommand{\alphaabcd}{\alpha^{(a,b,c,d)}}
\newcommand{\betaabcd}{\beta^{(a,b,c,d)}}
\newcommand{\hdopnkabcd}{\hdop_{n,k}^{(a,b,c,d)}}
\newcommand{\hdopabcd}{\hdop^{(a,b,c,d)}}
\newcommand{\Wabcd}{{W^{(a,b,c,d)}}}
\newcommand{\hdopmjabcd}{\hdop_{m,j}^{(a,b,c,d)}}
\newcommand{\bighdopabcd}{\bighdop^{(a,b,c,d)}}
\newcommand{\bigWabcd}{\mathbb{W}^{(a,b,c,d)}}

\newcommand{\genjact}{\tilde{\genjac}}
\newcommand{\genjactnmk}{\genjact_{n-k}}
\newcommand{\genjactmmj}{\genjact_{m-j}}
\newcommand{\genjactw}{w_{\genjact}}
\newcommand{\normgenjact}{\omega_{\genjact}}

% Spherical Cap commands
\newcommand{\scop}{Q}
\newcommand{\scopnki}{\scop_{n,k,i}}
\newcommand{\scopmjh}{\scop_{m,j,h}}
\newcommand{\scopa}{\scop^{(a)}}
\newcommand{\scopnkia}{\scopnki^{(a)}}
\newcommand{\scopmjha}{\scopmjh^{(a)}}
\newcommand{\bigscop}{{\mathbb{Q}}}
\newcommand{\bigscopa}{\bigscop^{(a)}}
\newcommand{\bigscopN}{\bigscop_{N}}
\newcommand{\bigscopNa}{\bigscopa_{N}}
\newcommand{\bigscopNka}{\bigscopa_{N,k}}
\newcommand{\bigscopt}{\mathbb{\tilde{Q}}}
\newcommand{\bigscopta}{\bigscopt^{(a)}}
\newcommand{\bigscoptn}{\bigscopt_{n}}
\newcommand{\bigscoptna}{\bigscopta_{n}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\ch}{Y}
\newcommand{\chki}{\ch_{k,i}}
\newcommand{\chjh}{\ch_{j,h}}
\newcommand{\chkitheta}{\ch_{k,i}(\theta)}
\newcommand{\alphaa}{\alpha^{(a)}}
\newcommand{\betaa}{\beta^{(a)}}
\newcommand{\gammaa}{\gamma^{(a)}}
\newcommand{\Wa}{W^{(a)}}
\newcommand{\bigWNa}{\mathbb{W}_N^{(a)}}
\newcommand{\ppphi}{\frac{\partial}{\partial \phi}}
\newcommand{\ppz}{\frac{\partial}{\partial z}}
\newcommand{\rhoppphi}{\rho \ppphi}
\newcommand{\pptheta}{\frac{\partial}{\partial \theta}}
\newcommand{\bigWNi}{\mathbb{W}_N^{(1)}}
\newcommand{\scopi}{\scop^{(1)}}
\newcommand{\scopnkii}{\scopnki^{(1)}}
\newcommand{\bigWNo}{\mathbb{W}_N^{(0)}}
\newcommand{\scopo}{\scop^{(0)}}
\newcommand{\scopnkio}{\scopnki^{(0)}}
\newcommand{\bigscopi}{\bigscop^{(1)}}
\newcommand{\bigscopNi}{\bigscopi_{N}}
\newcommand{\bigscopNki}{\bigscopi_{N,k}}
\newcommand{\bigscopo}{\bigscop^{(0)}}
\newcommand{\bigscopNo}{\bigscopo_{N}}
\newcommand{\bigscopNko}{\bigscopo_{N,k}}

% Tangent space
\newcommand{\tangentspace}{{\Omega_T}}
\newcommand{\phivec}{\hat{\bm{\phi}}}
\newcommand{\thetavec}{\hat{\bm{\theta}}}
\newcommand{\bigtsop}{\mathbb{T}}
\newcommand{\bigtsopt}{\tilde{\bigtsop}}
\newcommand{\bigtsoptn}{\bigtsopt_n}
\newcommand{\bigtsopN}{\bigtsop_N}
\newcommand{\bigtsopNk}{\bigtsop_{N,k}}
\newcommand{\tsopi}{\bm{\Phi}}
\newcommand{\tsopii}{\bm{\Psi}}
\newcommand{\tsopiN}{\tsopi_N}
\newcommand{\tsopiiN}{\tsopii_N}
\newcommand{\tsopinki}{\tsopi_{n,k,i}}
\newcommand{\tsopiinki}{\tsopii_{n,k,i}}
\newcommand{\tsopinkit}{\tsopi_{n,k,i}^\top}
\newcommand{\tsopiinkit}{\tsopii_{n,k,i}\top}
\newcommand{\tsopit}{\tsopi^\top}
\newcommand{\tsopiit}{\tsopii^\top}
\newcommand{\rvec}{\hat{\bm{r}}}
\newcommand{\jacobimattangent}{{}^T\!J}
\newcommand{\jacobimattangentx}{\jacobimattangent_x}
\newcommand{\jacobimattangenty}{\jacobimattangent_y}
\newcommand{\jacobimattangentz}{\jacobimattangent_z}
\newcommand{\alphao}{\alpha^{(0)}}
\newcommand{\alphaonkj}{\alphao_{n,k,j}}
\newcommand{\betao}{\beta^{(0)}}
\newcommand{\betaonkj}{\betao_{n,k,j}}
\newcommand{\gammao}{\gamma^{(0)}}
\newcommand{\gammaonkj}{\gammao_{n,k,j}}

\usepackage{amsthm}

\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma} 
\newtheorem{theorem}{Theorem} 
\newtheorem{definition}{Definition}
\newtheorem{condition}{Condition}
\newtheorem{algorithm}{Algorithm}
\newtheorem{corollary}{Corollary}

\def\condref#1{Condition~\ref{cond:#1}}

\include{somacros}

%\usepackage{media9}
%\graphicspath{ {../sphere/plots} }

\title{Sparse spectral and $p$-finite element methods for partial differential equations on spherical caps}
\author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\begin{abstract}
In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles. In this work we extend the methodology on disk-slices developed in \cite{snowball2019sparse} to a hierarchy of non-classical orthogonal polynomials on spherical caps. This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials. 
\end{abstract}

\section{Introduction}

We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the sphere. Define the disk-slice, denoted by $\omega$, by
\begin{align}
	\omega := \{(x,y) \in \R^2 \quad | \quad \alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)\}.
\end{align}
where:
\begin{align*}
	\begin{cases}
		(\alpha,\beta) &:= (\alpha,1), \quad \alpha \ge 0 \\
		(\gamma,\delta) &:= (-1,1) \\
		\rho(x) &:= (1-x^2)^{\half}, \quad x \in (\alpha,\beta) \\
	\end{cases}
\end{align*}

Further, define the spherical cap, denoted by $\Omega$, by
\begin{align}
	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
\end{align}
We can define the 3D OP family on $\Omega$ as the set of polynomials $\bigQab(x,y,z)$ where
\begin{align}
	\bigQab_n(x,y,z) := \begin{pmatrix}
		\bighdopab_n(x,y) \\
		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
	\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigQab := \begin{pmatrix}
		\bigQab_0 \\
		\bigQab_1 \\
		\vdots \\
		\bigQab_N
	\end{pmatrix}.
\end{align}
The relations for multiplication by $x, y$ are then trivial. For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
\begin{align}
	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: \hdop^{(a,b+1)}_{m,j}(x,y), \\
	z^2 \: \hdopnk^{(a,b+1)}(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b+1)\to(a,b)} \: \hdopmjab(x,y).
\end{align}
The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b+1)\to(a,b)}$ are given by
\begin{align}
	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: \hdop^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{\hdop^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
	d_{n,k,m,j}^{(a,b+1)\to(a,b)} &= \frac{\ip< W^{(0,1)} \hdopnk^{(a,b+1)}, \: \hdopmjab >_\Wab}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
        	&= \frac{\ip< \hdopnk^{(a,b+1)}, \: \hdopmjab >_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
        	&= \frac{\norm{\hdopnk^{(a,b+1)}}^2_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \: c_{m,j,n,k}^{(a,b)\to(a,b+1)}.
\end{align}
The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ are just the entries to the parameter transformation operator matrix $T^{(a,b)\to(a,b+1)}$, defined by
\begin{align}
	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y),
\end{align}
which can be shown to be sparse with banded-block-banded structure, meaning the resulting operator matrix for the Jacobi operator for multiplication by $z$ is sparse.

%
\section{Introduction}

This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
\begin{align*}
	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| = \gamma \rho(z)\}
\end{align*}
and consider the case where $\alpha \in (0,1)$ and where
\begin{align*}
	 \rho(z) &:= \sqrt{1-z^2} \\
	 \beta &:= 1 \\
	 \gamma &:= 1.
\end{align*}
This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
\begin{align*}
	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < 1, \: x^2 + y^2 + z^2 = 1\}
\end{align*}
Of course, there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$, however for simplicity we focus on the spherical cap.

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.  On the disk-slice the family of weights we consider are of the form
$$
\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \qfor \alpha \leq x \leq \beta,\quad -\rho(x) \leq y \leq \rho(x).
$$
The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
$$
\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
$$
where $C_{n,k} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.

Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$. Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries. In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree. This compares favourably to $O(p^6)$ operations if one proceeds na\"ively. Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.

Here is an overview of the paper:

\noindent \secref{OPs}: We present our procedure to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the disk-slice domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.

\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators. We present a method involving use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.

\noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice. We present a method for explicitly deriving the recurrence coefficients of the non-classical 1D OPs we detail in \secref{OPs}. We derive a quadrature rule on the disk-slice that can be used to expand a function in the OP basis up to a given order. Further, we implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.

\noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the disk-slice.

\noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods. This lays the groundwork for a future $hp$-finite element method in a disk, where the elements capture the circular geometry precisely.

\noindent\appref{HalfDisk}: We discuss extension to the special case of end-disk-slices (e.g., half disks).

\noindent\appref{Trapezium}: We discuss extension to trapezia.

\section{Orthogonal polynomials on the disk-slice and the trapezium}\label{Section:OPs}

In this section we outline the construction and some basic properties of $\scopnkia(x,y)$.

\subsection{Explicit construction}

We can construct the 3D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from the 2D circular harmonics on the unit circle.

\begin{definition}
Define $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ as the 2D OPs on $(x,y)\in\R$ s.t. $x^2 + y^2 = 1$ (known as the circular harmonics on the unit circle); that is, if we define $\theta$ by $x = \cos(\theta)$, $y = \sin(\theta)$ for $\xvec := (x, y)$ where $||\xvec|| = 1$,
\begin{align}
	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) \equiv \ch_{0,0}(\theta) &:= \frac{\sqrt{2}}{2} =: \ch_{0} \nonumber \\
	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \equiv \ch_{k,0}(\theta) &:= \cos(k \theta), \quad k = 1,2,3,\dots\label{eqn:circularharmonics} \\
	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) \equiv \ch_{k,0}(\theta)& := \sin(k \theta), \quad k = 1,2,3,\dots \nonumber
\end{align}
orthonormal with respect to the inner product
\begin{align}
	\ip< p, \: q >_{\ch} &:= \frac{1}{\sqrt{\pi}} \: \int_{||\xvec|| = 1} p(\xvec) \: q(\xvec) \: \D \sigma(\xvec) = \frac{1}{\sqrt{\pi}} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta. \label{eqn:ipcircularharmonics}
\end{align}
\end{definition}

\begin{proposition}\label{prop:construction}
Let $w : (\alpha,\beta) \: \to \R$ be a weight function. $\forall$, $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$
\begin{align*}
	\scopnki(x,y,z) \equiv \scopnki(\xvec,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr(\frac{\xvec}{\rho(z)})
\end{align*}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\begin{align*}
	\ip< p, \: q > &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: w(z) \: \D \sigma(\xvec, z) \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: w(z) \: \D z.
\end{align*}
on $\Omega$. 
\end{proposition}

\begin{proof}
\begin{align*}
	\ip< \scopnki, \: \scopmjh > &= \int_\alpha^1 \int_0^{2\pi} r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}(\theta) \: r_{m-j,j}(z) \: \rho(z)^j \: \ch_{j,h}(\theta) \: \D \theta \: w(z) \: \D z \\
	&= \Big(\int_0^{2\pi}  \: \ch_{k,i}(\theta) \: \: \ch_{j,h}(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: r_{n-k,k}(z) \: \rho(z)^k \: r_{m-j,j}(z) \: \rho(z)^j \: w(z) \: \D z \Big) \\
	&= \pi \: \delta_{k,j} \: \delta_{i,h} \:  \int_\alpha^1 \: r_{n-k,k}(z) \: r_{m-k,k}(z) \: \rho(z)^{2k} \: w(z) \: \D z \\
	&= C \: \pi \: \delta_{k,j} \: \delta_{i,h}
\end{align*}
for some constant $C$.
\end{proof}

For the spherical cap, we formalise Proposition \ref{prop:construction} in the following definitions. First we introduce notation for two families of univariate OPs.
\begin{definition}\label{def:OPconstruction}
Let $\genjacw^{(a,b)}(x)$ be a weight function on the interval $(\alpha, \beta)$ given by:
\begin{align*}
	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{b} \: \rho(x)^{b}
\end{align*}
and define the associated inner product by:
\begin{align}
	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipgenjac}
\end{align}
where
\begin{align}
	\normgenjac^{(a,b)} := \int_\alpha^\beta \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipnormalisation}
\end{align}
is a normalising constant.
Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}).
\end{definition}

\begin{definition}\label{def:constuction}
Define the one-parameter 3D orthogonal polynomials via:
\begin{align}
	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(x) \: \rho(x)^k \: \chki\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
\end{definition}

$\{\scopnkia\}$ are orthogonal with respect to the inner product
\begin{align}
	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D \sigma(\xvec, z) \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z.
\end{align}

We note that the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}), and has been used in the construction of 2D orthogonal polynomials on disk-slices and trapeziums \cite{snowball2019sparse}.

\subsection{Jacobi matrices}

We can express the three-term recurrences associated with $\genjac_n^{(a,b)}$ as
\begin{align}
	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Rrecurrence}
\end{align}
where the coefficients are calculatable (see \cite{snowball2019sparse}). We can use (\ref{eqn:Rrecurrence}) to determine the 3D recurrences for $\scopnkia(x,y,z)$. Importantly, we can deduce sparsity in the recurrence relationships.  We first require the following Lemma.

\begin{lemma}\label{lemma:Yrecurrence} 
The following identities hold for $k = 2,3,\dots$:
\begin{align*}
	&1) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \cos(\theta) \: \D \theta = \ch_0 \: \pi \: \delta_{0,h} \: \delta_{1,j} \\
	&2) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \sin(\theta) \: \D \theta = \ch_0 \: \pi \: \delta_{1,h} \: \delta_{1,j} \\
	&3) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta = \pi \: \delta_{i,h} \: (\ch_0 \: \delta_{0,j} + \half \delta_{2,j}) \\
	&4) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta = \pi \:  \delta_{|i-1|,h} \: ((-1)^{i+1} \: \ch_0 \: \delta_{0,j} + (-1)^i \: \half \: \delta_{2,j}) \\
	&5) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta = \half \: \pi \: \delta_{i,h} \: (\delta_{k-1,j} + \delta_{k+1,j}) \\
	&6) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta = \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \delta_{k-1,j} + (-1)^i \: \delta_{k+1,j}).
\end{align*}
\end{lemma}
\begin{proof}
Each follows from the definitions of $\ch_{k,i}$ and $\ch_0$, as well as the relationships:
\begin{align*}
	2 \: \cos(k \theta) \cos(\theta) &= \cos((k-1)\theta) + \cos((k+1)\theta) \\
	2 \: \sin(k \theta) \cos(\theta) &= \sin((k-1)\theta) + \sin((k+1)\theta) \\
	2 \: \cos(k \theta) \sin(\theta) &= - \sin((k-1)\theta) + \sin((k+1)\theta) \\
	2 \: \sin(k \theta) \sin(\theta) &= \cos((k-1)\theta) - \cos((k+1)\theta).
\end{align*}
\end{proof}

\begin{lemma}\label{lemma:Qrecurrence} 
Define
\begin{align}
	\gamma_{k} :=
		\begin{cases}
			0 &\text{if } k < 0 \\
			\ch_0 &\text{if } k = 0 \\
			\half &\text{otherwise}
		\end{cases}
\end{align}
$\scopnkia(x,y,z)$ satisfy the following 3-term recurrences:
\begin{align*}
x \: \scopnkia(x,y,z) &= \alphaa_{n,k,1} \:  \scopa_{n-1, k-1, i}(x, y, z) + \alphaa_{n,k,2} \:  \scopa_{n-1, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\
y \: \scopnkia(x,y,z) &= \betaa_{n,k,i,1} \:  \scopa_{n-1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,2} \:  \scopa_{n-1, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z), \\
\end{align*}
for $(x,y,z) \in \Omega$, where
\begin{align*}
	\alphaa_{n,k,1} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,2} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\alphaa_{n,k,3} &:=\gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,4} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\alphaa_{n,k,5} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
	\alphaa_{n,k,6} &:=\gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
	\betaa_{n,k,i,j} &:= 
		\begin{cases}
			- \alphaa_{n,k,j} \quad &\text{if }(i = 0 \text{ and } j \text{ is odd}) \text{ or }(i = 1 \text{ and } j \text{ is even}) \\
			\alphaa_{n,k,j} \quad &\text{otherwise}
		\end{cases} \\	
	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)} \\
\end{align*}
\end{lemma}

\begin{proof}
The 3-term recurrence for multiplication by $z$ follows from (\ref{eqn:Rrecurrence}). For the recurrence for multiplication by $x$, since $\{\scopmjha\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n+1$ polynomial on $\Omega$, we can expand 
$$
x \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j} \: \scopmjha(x,y,z).
$$
These coefficients are given by
\begin{align*}
	c_{m,j} = {\ip< x \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
\end{align*}
where we show the non-zero coefficients that result are the $\alphaa_{n,k,1},\dots,\alphaa_{n,k,6}$ in the Lemma.
Recall from equation (\ref{eqn:normscop}) that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$. Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos(\theta) \sin(\phi), \: \sin(\theta)\sin(\phi), \: \cos(\phi)) = (x, y, z)$:
\begin{align*}
	&\ip<x \: \scopnkia, \scopmjha>_{\scopa} \\
	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: x \: \genjacw^{(a,0)}(z) \: \D\sigma(\xvec, z) \\
	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta \Big) \\
	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta \Big) \\
	&= \half \: \pi \: \delta_{i,h} \: (\gamma_{k-1} \: \delta_{k-1, j} + \gamma_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
\end{align*}
where $\delta_{k, j}$ is the standard Kronecker delta function, using Lemma \ref{lemma:Yrecurrence}. Similarly, for the recurrence for multiplication by $y$, we can expand 
$$
y \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 d_{m,j} \: \scopmjha(x,y,z).
$$ 
These coefficients are given by
\begin{align*}
	d_{m,j} = {\ip< y \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
\end{align*}
where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the Lemma:
\begin{align*}
	&\ip<y \: \scopnkia, \scopmjha>_{\scopa} \\
	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: y \: \genjacw^{(a,0)}(z) \: \D\sigma(\xvec, z) \\
	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta \Big) \\
	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta \Big) \\
	&= \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \gamma_{k-1} \: \delta_{k-1, j} + (-1)^i \: \gamma_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
\end{align*}
where again $\delta_{k, j}$ is the standard Kronecker delta function, and we have used Lemma \ref{lemma:Yrecurrence}.

\end{proof}

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Define, for $n=0,1,2,\dots$: 
\begin{align*}
\bigscoptna := \begin{pmatrix}
		\scopa_{n,0,0}(x,y,z) \\
		\scopa_{n,1,0}(x,y,z) \\
		\scopa_{n,1,1}(x,y,z) \\
		\vdots \\
		\scopa_{n,n,0}(x,y,z) \\
		\scopa_{n,n,1}(x,y,z)
	\end{pmatrix} \in \R^{2n+1}, 
\quad \quad 
\bigscopta := \begin{pmatrix}
		\bigscopta_0 \\
		\bigscopta_1 \\
		\bigscopta_2 \\
		\vdots \\
	\end{pmatrix}
\end{align*}
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
\begin{align}
	J_x^{(a,b,c,d)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z), \nonumber
\end{align}
The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated by its expansion $f(x,y,z) = \bigscopta(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopta(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. Further, note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:

\begin{definition}
A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$. A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix. 
\end{definition}

For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
\begin{align*}
J_{x/y/z}^{(a)} &= \begin{pmatrix}
		B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
		C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
		& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
		& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align*}
For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
\begin{align*}
A^{(a)}_{x,n} &:= \begin{pmatrix}
		0 & A^{(a)}_{n,0,6} & 0 & & \\
		A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
		& \ddots & \ddots & \ddots & \\
		& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
	    \end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
B^{(a)}_{x,n} &:= \begin{pmatrix}
		0 & A^{(a)}_{n,0,4} & & \\
		A^{(a)}_{n,1,3} & \ddots & \ddots & \\
		& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
		& & A^{(a)}_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
C^{(a)}_{x,n} &:= \begin{pmatrix}
		0 & A^{(a)}_{n,0,2} & & \\
		A^{(a)}_{n,1,1} & \ddots & \ddots & \\
		& \ddots & \ddots &A^{(a)}_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & A^{(a)}_{n,n,1} \\
	    \end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
\end{align*}
where
\begin{align*}
A^{(a)}_{n,k,j} &:= 
	\begin{pmatrix}
		\alphaa_{n,k,j} & 0 \\
		0 & \alphaa_{n,k,j}
	\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
A^{(a)}_{n,0,j} &:=
	\begin{pmatrix}
		\alphaa_{n,0,j} & 0
	\end{pmatrix} \in \R^{1\times2}, \quad j \text{ even} \\
A^{(a)}_{n,1,j} &:=
	\begin{pmatrix}
		\alphaa_{n,1,j} \\
		0
	\end{pmatrix} \in \R^{2\times1}, \quad j \text{ odd} \\
\end{align*}
For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
\begin{align*}
A^{(a)}_{y,n} &:= \begin{pmatrix}
		0 & B^{(a)}_{n,0,6} & 0 & & \\
		B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
		& \ddots & \ddots & \ddots & \\
		& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
	    \end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
B^{(a)}_{y,n} &:= \begin{pmatrix}
		0 & B^{(a)}_{n,0,4} & & \\
		B^{(a)}_{n,1,3} & \ddots & \ddots & \\
		& \ddots & \ddots & B^{(a)}_{n,n-1,4} \\
		& & B^{(a)}_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
C^{(a)}_{y,n} &:= \begin{pmatrix}
		0 & B^{(a)}_{n,0,2} & & \\
		B^{(a)}_{n,1,1} & \ddots & \ddots & \\
		& \ddots & \ddots & B^{(a)}_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & B^{(a)}_{n,n,1} \\
	    \end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
\end{align*}
where
\begin{align*}
B^{(a)}_{n,k,j} &:= 
	\begin{pmatrix}
		0 & \betaa_{n,k,j} \\
		\betaa_{n,k,j} & 0
	\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
B^{(a)}_{n,0,j} &:=
	\begin{pmatrix}
		0 & \betaa_{n,0,j}
	\end{pmatrix} \in \R^{1\times2}, \quad j \text{ even} \\
B^{(a)}_{n,1,j} &:=
	\begin{pmatrix}
		0 \\
		\betaa_{n,1,j}
	\end{pmatrix} \in \R^{2\times1}, \quad j \text{ odd} \\
\end{align*}
For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
\begin{align*}
A^{(a)}_{z,n} &:= \begin{pmatrix}
		\Gamma^{(a)}_{n,0,3} & 0 & \\
		0 & \ddots & \ddots & & \\
		& \ddots & \ddots & \ddots & \\
		& & 0 & \Gamma^{(a)}_{n,0,3} & 0 \\
	    \end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
B^{(a)}_{z,n} &:= \begin{pmatrix}
		\Gamma^{(a)}_{n,0,2} & \\
		& \ddots & & \\
		& & \ddots & \\
		& & & \Gamma^{(a)}_{n,0,2}
	    \end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
C^{(a)}_{z,n} &:= \begin{pmatrix}
		\Gamma^{(a)}_{n,0,1} & 0 & & \\
		0 & \ddots & \ddots & \\
		& \ddots & \ddots & 0 \\
		& & \ddots & \Gamma^{(a)}_{n,n,1} \\
		& & & 0 \\
	    \end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
\end{align*}
where
\begin{align*}
\Gamma^{(a)}_{n,k,j} &:= 
	\begin{pmatrix}
		\gamma_{n,k,j} & 0 \\
		0 & \gamma_{n,k,j}
	\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \\
\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}
\end{align*}

Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.

\subsection{Building the OPs}

We can combine each system in (\ref{eqn:jacobimatricesdefinition}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:

\begin{align*}
\renewcommand\arraystretch{1.3}
\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
		& & \ddots &\ddots
\end{pmatrix}
\bigscopta(x,y,z)
=
\begin{pmatrix}
	 \scopa_0 \\ 0 \\ 0 \\ 0 \\ \vdots  \\
\end{pmatrix},
\end{align*}
where we note $\scopa_{0,0,0}(x,y,z) =: \scopa_0 \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
\begin{align*}
A_n &:= \begin{pmatrix}
		A^{(a)}_{x,n} \\
		A^{(a)}_{y,n} \\
		A^{(a)}_{z,n}
	    \end{pmatrix} \in \R^{3(2n+1)\times(2n+3)}, \quad
C_n := \begin{pmatrix}
		C^{(a)}_{x,n} \\
		C^{(a)}_{y,n} \\
		C^{(a)}_{z,n}
	    \end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{pmatrix}
		B^{(a)}_{x,n} \\
		B^{(a)}_{y,n} \\
		B^{(a)}_{z,n}
	    \end{pmatrix} \in \R^{3(2n+1)\times(2n+1)}, \quad
G_n(x,y,z) := \begin{pmatrix}
		xI_{2n+1} \\
		yI_{2n+1} \\
		zI_{2n+1}
	    \end{pmatrix} \in \R^{3(2n+1)\times(n+1)}.
\end{align*}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{2n+3}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align*}
	\begin{cases}
		\bigscopta_{-1}(x,y,z) := 0 \\
		\bigscopta_{0}(x,y) := \scopa_0 \\
		\bigscopta_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
	\end{cases}
\end{align*}

Note that we can define an explicit \(\Dnt\) as follows:
\begin{align*}
\Dnt := \begin{pmatrix}
		0 & & & 0 & & & 1 / \gammaa_{n,0,3} & & \\
		& \ddots & & & \ddots & & & \ddots \\
		& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
		& & & & & \underline{\eta}^\top_{1} & & & \\
		& & & & & \underline{\eta}^\top_{0} & & &
	    \end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
\end{align*}
for $n = 2, 3, \dots$ where $\underline{\eta}_{0}, \underline{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
\begin{align*}
	\big(\underline{\eta}_{0}\big)_j &= 
		\begin{cases}
			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2 \\
			0 & o/w
		\end{cases} \\
	\big(\underline{\eta}_{1}\big)_j &= 
		\begin{cases}
			\big(\underline{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
			0 & o/w
		\end{cases}
\end{align*}
For $n=0, 1$, we can simply take
\begin{align*}
D^\top_0 &:= \begin{pmatrix}
		0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
		\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
		0 & \frac{1}{\betaa_{0,0,6}} & 0
	    \end{pmatrix} \in \R^{3\times3}, \\
D^\top_1 &:= \begin{pmatrix}
		0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,0,3} & & \\
		0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3} \\
		0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3}  \\
		0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & \eta & 0 & 0 \\
		0 & 0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & 0 & 0
	    \end{pmatrix} \in \R^{5\times9},
\end{align*}
where $\eta = \frac{- \: \alphaa_{1,1,5}}{\alphaa_{1,1,6} \: \gammaa_{1, 0, 3}}$.

It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigscopta_{0}(x,y)$  through $\bigscopta_{n}(x,y)$ in optimal $O(n^2)$ complexity. \bstodo{correct for this case}

%
\section{Sparse partial differential operators}\label{Section:PDOs}

\begin{figure} 
\center
\begin{tikzpicture} 
% Box/square (4x4)
\draw[black,solid,ultra thin] (0,0)--(0,4);
\draw[black,solid,ultra thin] (0,0)--(4,0);
\draw[black,solid,ultra thin] (4,0)--(4,4);
\draw[black,solid,ultra thin] (0,4)--(4,4);
% Dots
\draw[black,fill=black] (0,0) circle (.5ex);
\draw[black,fill=black] (2,2) circle (.5ex);
\draw[black,fill=black] (4,4) circle (.5ex);
\draw[black,fill=black] (0,2) circle (.5ex);
\draw[black,fill=black] (0,4) circle (.5ex);
% Arrows
\draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
\draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
\draw[black,thick,->] (0,0.2)--(0,1.8);
\draw[black,thick,->] (0,2.2)--(0,3.8);
\draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
% Node (parameter) labels
\draw[] (0,0) node[anchor=east] {$(0,0,0)$};
\draw[] (0,4) node[anchor=east] {$(0,0,2)$};
\draw[] (4,0) node[anchor=west] {$(2,2,0)$};
\draw[] (4,4) node[anchor=west] {$(2,2,2)$};
% Arrow labels
\draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
\draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
% The key and headings
\draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
\draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
\draw[] (-2.3,3.2) node[anchor=west] {$(a,b)$};
\draw[] (-2.7,3.6) node[anchor=south] {$c$};
\draw[] (2,4.6) node[anchor=south] {basis conversion};
\end{tikzpicture} 
\caption{The Laplace operator acting on vectors of $\smash{\hdopnk^{(0,0,0)}}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{\hdopnk^{(2,2,2)}}$ coefficients. Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{\hdopnkabc}$ coefficients, where $(a,b,c)$ is at the tail of the arrow, and the range is $\smash{\hdopnk^{(\tilde{a},\tilde{b},\tilde{c})}}$ coefficients, where $(\tilde{a},\tilde{b},\tilde{c})$ is at the head of the arrow. %For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{\hdop_{n,k}^{(0,0,0)}}$ (resp.~$\smash{\hdop_{n,k}^{(1,0,1)}}$) and the range in $\smash{\hdop_{n,k}^{(1,0,1)}}$ (resp.~$\smash{\hdop_{n,k}^{(2,0,2)}}$).
}
\label{fig:Laplace} 
\end{figure}

In this section, we will derive the entries that form spherical partial differential operators, demonstrating their sparsity. To this end, we will use a different ordering of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal. Let $N \in \N$ and define: 
\begin{align}
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
\end{align}

We further denote the weighted set of OPs on $\Omega$ by 
\begin{align*}
	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z),
\end{align*}
and recall that a function $f(x,y,z)$ defined on $\Omega$ is approximated by its expansion $f(x,y,z) = \bigscopa(x,y,z)^\top \mathbf{f}$.

\begin{definition}\label{def:differentialoperators}
Define the operator matrices $D_\phi^{(a)}, \: W_\phi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ according to:
\begin{align*}
	\rho {\partial f \over \partial \phi}(x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\phi^{(a)} \: \mathbf{f}, \\
	\rho {\partial \over \partial \phi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\phi^{(a)}\: \mathbf{f}, \\
	{\partial f \over \partial \theta}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
	\nabla^2_s \big( \genjacw^{(a,0)}(z) \: f(x,y,z) \big) &= \bigscopNa(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ for } a = 1 \text{ only}.
\end{align*}
\end{definition}

The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, on the triangle \cite{olver2018recurrence}, and on the disk-slice \cite{snowball2019sparse}. The operators we define here are for partial derivatives with respect to the spherical coordinates $(\phi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).

\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
The operator matrices $D_\phi^{(a)}, \: W_\phi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure. More specifically:
\begin{itemize}
	\item $D_\phi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 4)$
  	\item $W_\phi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 2)$
	\item $D_\theta$ has  block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
	\item $\Delta^{(1)}_W$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
\end{itemize}
\end{theorem}

In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short Lemma.

\begin{lemma}\label{lemma:Rsecondderivative}
For any general parameter $a$ and any $n = 0,1,\dots$, $k = 0,\dots,n$ we have that
\begin{align*}
	&\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
	&\quad \quad = \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
	&\quad \quad = \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
\end{align*}
where 
\begin{align*}
	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
\end{align*}
\end{lemma}
\begin{proof}[Proof of Lemma \ref{lemma:Rsecondderivative}]
Since $\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] = \genjacw^{(a, 2k)} \: r_{n-k+1}$ where $r_{n-k+1}$ is a degree $n - k + 1$ polynomial, we have that 
\begin{align*}
	\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)}
\end{align*}
for some coefficients $\tilde c_{\{n,k\},m}$. These coefficients are given by
\begin{align*}
	\tilde c_{\{n,k\},m} &= \frac{1}{\normgenjac^{(a,2k)}} \: \ip<\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
	&= - \frac{1}{\normgenjac^{(a,2k)}}  \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
\end{align*}
We show that these are zero for $m < n - k - 1$ by integrating twice by parts:
\begin{align*}
	&\ip<\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
	&\quad \quad \quad = - \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z \\
	&\quad \quad \quad = \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: [(a+1) \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(0, 2)} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - 2(k+1) z \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(1, 0)} + \genjac_{m}^{(a,2k) \: \prime \prime} \: \genjacw^{(1, 2)}] \: \genjacw^{(a, 2k)} \: \D z
\end{align*}
which is indeed zero for $m < n - k - 1$ by orthogonality.
\end{proof}

\begin{proof}[Proof of Theorem \ref{theorem:sparsityofdifferentialoperators}]
First, note that:
\begin{align}
	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \label{eqn:derivativeofweightgenjac} \\
	\rho(z) \: \rho'(z) &= -z \label{eqn:rhoderivative} \\
	{\partial \over \partial \phi} &= -\rho \: {\partial \over \partial z} \label{eqn:partialphitoz} 
\end{align}

For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that
\begin{align*}
	{\partial \over \partial \theta}\scopnkia(x,y,z) &= \genjacnmk^{(a,2k)}(z) \: \rho(z)^{k} \: {\partial \over \partial \theta} \chki(\theta) \\
	&= 
	\begin{cases}
		(-1)^{i+1} \: k \: \scopa_{n,k,|i-1|}(x,y,z) & k > 0 \\
		0 & k = 0
	\end{cases}
\end{align*}

We now proceed with the case for the operator $D_\phi^{(a)}$ for partial differentiation by $\phi$. The entries of the operator are given by the coefficients in the expansion $\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)}$, where the coefficients are
\begin{align*}
	c_{m,j,h} = \norm{\scopmjh^{(a+1)}}^{-2}_{\scop^{(a+1)}} \ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}}.
\end{align*}
Now,
\begin{align*}
	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
	&\quad = - \int_\Omega \: \rho(z)^2 \ppz \:  [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \genjacw^{(a+1, 0)} \: \D \Omega \\
	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+1,2k)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, 2k)} \: \D z \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
\end{align*}
which is zero for $j \ne k$, $h \ne i$, and $m < n - 2$ by orthogonality.

Similarly for the operator $W_\phi^{(a)}$ for partial differentiation by $\phi$ on the weighted space, the entries of the operator are given by the coefficients in the expansion $\rhoppphi (\genjacw^{(a,0)} \: \scopnkia) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-1,0)} \: \scopmjh^{(a-1)}$, where the coefficients are
\begin{align*}
	c_{m,j,h} = \norm{\scopmjh^{(a-1)}}^{-2}_{\scop^{(a-1)}} \ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}}.
\end{align*}
Now,
\begin{align*}
	&\ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}} \\
	&\quad = - \int_\Omega \: \rho(z)^2 \ppz \:  [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \genjacw^{(a, 0)}(z)] \: \genjacmmj^{(a-1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \Omega \\
	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a-1,2j)} \: [k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2] \: \genjacw^{(a-1, k + j)} \: \D z \Big) \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \:  \int_\alpha^1 \: \genjac_{m-k}^{(a-1,2k)} \: [k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2] \: \genjacw^{(a-1, 2k)} \: \D z \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} - a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \genjacw^{(1, 2)} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a-1, 2k)} \: \D z \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
\end{align*}
which is zero for $j \ne k$, $h \ne i$, and $m < n - 1$ by orthogonality.

Finally, fix $a = 1$. For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are
\begin{align*}
	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
\end{align*}
Now, note that the Laplacian acting on the weighted spherical cap OP $\scopnki^{(a)}$ yields
\begin{align*}
	\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) &= 
\end{align*}
Hence,
\begin{align*}
	&\ip<\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(a, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(a, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(a, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(a, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(1,2k)} \: \genjacnmk^{(a, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\ 
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \Big\} \: \D z
\end{align*}
where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in Lemma \ref{lemma:Rsecondderivative}.
\end{proof}

There exist conversion matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.

\begin{definition}\label{def:parametertransformationoperators}
Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
\begin{align*}
	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z) \\
	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z)
\end{align*}
\end{definition}

\begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
The operator matrices in Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
\begin{itemize}
	\item $T^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block bandwidths $(0,2\tilde a)$
	\item $T_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block bandwidths $(2\tilde a, 0)$
\end{itemize}
\end{lemma}

\begin{proof}
We proceed with the case for the non-weighted operators $T^{(a)\to(a+\tilde a)}$. Since $\{\scopmjh^{(a+\tilde{a})}\}$ for $m = 0,\dots,n$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n$ polynomial, we can expand $\scopnkia = \sum_{m=0}^{n} \sum_{j=0}^m t_{m,j} \: \scopmjh^{(a+\tilde{a})}$. The coefficients of the expansion are then the entries of the operator matrix. We will show that the only non-zero coefficients are for $k = j$, $i = h$ and $m \ge  n - \tilde a$. Note that
\begin{align*}
	t_{m,j} = \norm{\scopmjh^{(a+\tilde{a})}}^{-2}_{\scop^{(a+\tilde a)}} \: \ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}}.
\end{align*}
where
\begin{align*}
	\ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}} &= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \cdot \: \Big( \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjacmmj^{(a+\tilde a,2j)} \: \rho^{k+j} \: \genjacw^{(a + \tilde a, 0)} \: \D z \Big) \\
	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a + \tilde a, 2k)} \: \D z
\end{align*}
which is zero for $n > m + \tilde a \iff m < n - \tilde a$. The sparsity argument for the weighted parameter transformation operator follows similarly.
\end{proof}

\begin{figure}[t]
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
        \centering
	\end{subfigure}
    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity. Left: the Laplace operator $\laplacewiii$. Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. Right: the biharmonic operator $\biharmonictwo$.}
        \label{fig:sparsity}
        \centering
\end{figure}

General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators. As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored surface Laplacian $\rho(z)^2 \: \nabla^2_s$, that will take us from coefficients for expansion in the weighted space
$$
\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)
$$
to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$. Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in Definition \ref{def:differentialoperators}. The matrix operator for this $\rho^2$-factored Laplacian we denote $\mathcal{L}_W^{(1)}$ acting on the coefficients vector is then given by
\begin{align*}
    \mathcal{L}_W^{(1)} := D_\phi^{(0)} \: W_\phi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
\end{align*}
Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.

Another important example is the Biharmonic operator $\Delta^2_s$ (defined by $\Delta^2_s f = \nabla^2_s (\nabla^2_s f)$), where we assume zero Dirichlet and Neumann conditions. To construct an operator for the Biharmonic function, we first note that we can obtain the matrix operator for the $\rho^2$-factored Laplacian $\rho(z)^2 \: \nabla^2_s$ that will take us from coefficients for expansion in the space $\bigscopN^{(0)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$. We denote this matrix operator that acts on the coefficients vector as $\mathcal{L}^{(0) \to (2)}$, and is given by
\begin{align*}
    \mathcal{L}^{(0) \to (2)} := D_\phi^{(2)} \: D_\phi^{(0)} + T^{(0)\to(2)} \: D_\theta.
\end{align*}
Further, we can represent this same Laplacian as a map from coefficients in the space $\bigW_N^{(2)}$ to coefficients in the space $\bigW_N^{(0)} \equiv \bigscopN^{(0)}$. Note that a function expanded in the $\bigW_N^{(2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$. We denote this matrix operator as $\mathcal{L}_W^{(2) \to (0)}$, and is given by
\begin{align*}
	\mathcal{L}_W^{(2) \to (0)} := W_\phi^{(1)} \: W_\phi^{(2)} + T_W^{(2)\to(0,0,0)} \: D_\theta.
\end{align*}
We can then construct a matrix operator for $\Delta^2_s$ that will take coefficients in the space $\bigW_N^{(2)}$ to coefficients in the space $\bigscopN^{(2)}$. Note that any function expanded in the $\bigW_N^{(2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$. Since, for any scalar function $f:\Omega \to \R$,
\begin{align*}
	\rho^2(z) \: \Delta^2 \: f(x,y,z) = \rho^2 \: \nabla^2_s (\rho^2 \: \nabla^2_s f) - z \: \rhoppphi (\rho^2 \: \nabla^2_s f) - [2z^2 - 2\rho^2 + 3z] \:\rho^2 \: \nabla^2_s f
\end{align*}
the matrix operator for the $\rho^2$-factored Biharmonic operator $\rho(z)^2 \: \Delta_s^2$ is then given by
\begin{align*}
	\mathcal{B}_W^{(2)} := \mathcal{L}^{(0) \to (2)} \: \mathcal{L}_W^{(2) \to (0)} - (J_z^{(2)})^\top \: T^{(1)\to(2)} \: D_\phi^{(0)} \: \mathcal{L}_W^{(2) \to (0)} - V^{(2)} \:  T^{(0)\to(2)} \: \mathcal{L}_W^{(2) \to (0)},
\end{align*}
where $V^{(a)}$ is an operator corresponding to multiplication by the scalar function $v(z) := 2z^2 - 2\rho(z)^2 + 3z = 4z^2 + 3z - 2$ acting on coefficients in the $\bigscopN^{(a)}$ space.
The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.

%
\section{Computational aspects}\label{Section:Computation}

In this section we discuss how to take advantage of the proposed basis and sparsity structure in partial differential operators in practical computational applications.

\subsection{Constructing $\genjac_n^{(a,b)}(x)$}

It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.

\subsection{Quadrature rule on the spherical cap}\label{subsection:quadrule}

In this section we construct a quadrature rule exact for polynomials on the spherical cap $\Omega$ that can be used to expand functions in the OPs $\scopnkia(x,y,z)$ for a given parameter $a$.

\begin{theorem}\label{Theorem:quadrule}
Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$. Further, we denote the $M_2$ Gauss quadrature nodes and weights for the unit circle \cite{???}\bstodo{good reference?} by $(\mathbf{s}_j, w_j^{(s)})$ where for $j = 1,\dots,M_2$:
\begin{align*}
	\mathbf{s}_j &= \Big( \cos\fpr(\frac{2 \pi j}{M_2}), \: \sin\fpr(\frac{2 \pi j}{M_2}) \Big), \\
	w_j^{(s)} &= \frac{2\pi}{M_2}.
\end{align*}
Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
\begin{align*}
	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
	z_{l+(j-1)M_2} &:= t_j, \\
	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
\end{align*}
Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
\begin{align*}
	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
\end{align*}
where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
\end{theorem}

\begin{proof}
Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
\begin{align*}
	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
\end{align*}
so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function. Note that if $f$ is a polynomial, then $f_e(\rho(t)\xvec, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $\xvec \in \R^2$.

Now, we have that
\begin{align*}
	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\mathbf{\xi}} = 1} f_e\big(\rho(z)\mathbf{\xi}, z)\big) \: \D \sigma(\mathbf{\xi}) \Big) \: \D z \\
	&\quad \quad \approx\int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) \mathbf{s}_l, z\big) \Big) \: \D z \quad (\star) \\
	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) \mathbf{s}_l, t_j\big) \quad (\star \star) \\
	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
\end{align*}
Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial. We therefore achieve equality at $(\star)$ if $M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$. \bstodo{For the circle rule, is this bound on N true?}

Next, note that 
\begin{align*}
	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\mathbf{\xi}} = 1} f_o\big(\rho(z)\mathbf{\xi}, z)\big) \: \D \sigma(\mathbf{\xi}) \Big) \: \D z \quad (\dagger) \\
	&\quad \quad = 0
\end{align*}
since the inner integral at $(\dagger)$ over $\mathbf{\xi}$ is zero, due to the symmetry over the domain.

Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
\begin{align*}
	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z  \\
	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
\end{align*}
where $M = M_1 M_2$ and $2M_1 - 1 \ge N, M_2 - 1 \ge N$.
\end{proof}

\subsection{Obtaining the coefficients for expansion of a function on the spherical cap}\label{subsection:expandingfunctions}

Fix $a \in \R$. Then for any function $f : \Omega \to \R$ we can express $f$ by
\begin{align*}
	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \mathbf{f}_k
\end{align*}
for N sufficiently large, where $\bigscopNka$ is defined in equation \label{eqn:OPdefNka} and where
\begin{align*}
	\mathbf{f}_k &:= 
		\begin{pmatrix}
			f_{k,k,0} \\
			f_{k,k,1} \\
			\vdots \\
			f_{N,k,0} \\
			f_{N,k,1}
		\end{pmatrix} \in \R^{2(N-k+1)} \quad \text{for } n = 1,2,\dots,N, \quad
	\mathbf{f}_0 := 
		\begin{pmatrix}
			f_{0,0,0} \\
			\vdots \\
			f_{N,0,0}
		\end{pmatrix} \in \R^{2(N+1)},
\end{align*}
\begin{align*}
	f_{n,k,i} &:= \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
\end{align*}
Recall from (\ref{eqn:OPnorm}) that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$. Using the quadrature rule detailed in Section \ref{subsection:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$: 
\begin{align*}
	f_{n,k,i} &= \frac{1}{2 \: \normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
\end{align*}
where the quadrature nodes and weights are those from Theorem \ref{Theorem:quadrule}, and $M = M_1 M_2$ with $2M_1 - 1 \ge N, M_2 - 1 \ge N$ (i.e. we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).

\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}

The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}. We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$. This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodo{correct the O(nd) complexity}.

%
\section{Examples on the disk-slice with zero Dirichlet conditions}\label{Section:Examples}

We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions on the spherical cap defined by $\Omega$. We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, as well as a time dependent heat equation, demonstrating the versatility of the approach.

\subsection{Poisson}

% FIGURES
\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
        	%\label{fig:solutionblocknorms-poisson}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions. This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
	\centering
	\label{fig:poisson}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
	\centering
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
	\centering
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
	\centering
	\end{subfigure}
	\centering
	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$. Left: Computed. Centre: Exact. Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
	\centering
	\label{fig:poissonexact}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
	%\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
	\centering
        	%\label{fig:solutionblocknorms-helmholtz}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
	\centering
	\label{fig:helmholtz}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
	%\label{fig:solution-biharmonic}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic-diskslice-alpha=0p2-beta=0p8-N=200}
	\centering
        	%\label{fig:solutionblocknorms-biharmonic}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
	\centering
	\label{fig:biharmonic}
\end{figure}
% END FIGURES

The Poisson equation is the classic problem of finding \(u(x,y)\) given a function \(f(x,y)\) such that:
\begin{align}
	\begin{cases}
    		\nabla^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
		u(x,y,z) = 0& \quad \text{on } \partial \Omega
	\end{cases}.
	\label{eqn:poisson}
\end{align}
noting the imposition of zero Dirichlet boundary conditions on $u$.

We can tackle the problem as follows. Choose an $N \in \N$ large enough for the problem, and denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}. In matrix-vector notation, our system hence becomes:
\begin{align*}
    \Delta_W^{(1)} \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$.
In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$. In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns. The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay. We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence. \bstodo{get figs and update the comments here}

In Figure \ref{fig:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error. The example was chosen so that the exact solution was $u(x,y,z) = \genjacw^{(1,0)}(z) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y,z) = \Delta \big[\genjacw^{(1,0)}(z) y^3 \exp(x)\big]$. We see that the computed solution is almost exact.

\subsection{Inhomogeneous variable-coefficient Helmholtz}

Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
\begin{align}
	\begin{cases}
    		\nabla^2 u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
		u(x,y,z) = 0 &\quad \text{on } \partial \Omega
	\end{cases}
	\label{eqn:helmholtz}
\end{align}
where $k \in \R$, noting the imposition of zero Dirichlet boundary conditions on $u$.

We can tackle the problem as follows. Denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}. We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}. We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top)$. In matrix-vector notation, our system hence becomes:
\begin{align*}
    (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example. In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y,z) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y,z) = x(1-x^2-y^2)e^x$. In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y,z) = 1 - (3(x-1)^2 + 5y^2)$ using N = 200, that is, 20,301 unknowns. The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution. We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence. \bstodo{get figs and update the comments here}

% For the final Gaussian bump example, we see that we achieve spectral convergence.

We can extend this to constant non-zero boundary conditions by simply noting that the problem 
\begin{align*}
	\begin{cases}
    		\nabla^2 u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
		u(x,y,z) = c \in \R &\quad \text{on } \partial \Omega
	\end{cases}
\end{align*}
is equivalent to letting $u = \tilde{u} + c$ and solving
\begin{align*}
	\begin{cases}
    		\nabla^2 \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - c \: k^2 \: v(x,y,z) \; =: g(x,y,z)  \quad \text{in } \Omega \\
		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
	\end{cases}
\end{align*}
and solving this zero boundary condition Helmholtz problem.

\subsection{Biharmonic equation}

Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
\begin{align}
	\begin{cases}
    		\rho(z)^2 \: \Delta^2 u(x,y,z) = \rho(z)^2 \: f(x,y,z) &\quad \text{in } \Omega \\
		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
	\end{cases}
	\label{eqn:biharmonic}
\end{align}
where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$. In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$. In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.  We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence. \bstodo{get figs and update the comments here}

% For the final Gaussian bump example, we see that we achieve spectral convergence.

%
\section{Tangent space (vector-valued functions)}\label{Section:tangentspace}

We can extend the methodology detailed here to the tangent space of the spherical cap $\Omega$. By creating a basis of orthogonal vector-valued functions (OVFs) for the spherical cap $\Omega$, we can expand vector-valued functions that lie in tangent space we call $\tangentspace$ in this basis. Such functions that are useful for the sphere are gradients and perpendicular-gradients of scalar functions. In this section we will define the OVFs, outline the Clenshaw matrices and how to build the OVFs, function evaluation as well as differential operators for gradient, curl and divergence that take us to and from coefficients in either of the OVF or scalar OP expansions.

Any function in the tangent space of $\Omega$ can be written as $\nabla f + \rvec \times \nabla g$ for some scalar functions $f, g$ on $\Omega$. Let $f$ be a scalar function on $\Omega$ that satisfies zero boundary conditions. Then there exist coefficients $\bm{f}$ such that $f$ can be expanded in the OP basis $\bigW_N^{(1)}$ for large enough N, i.e. $f = \genjacw^{(1,0)} \: \sum_{n,k,i} f_{n,k,i} \: \scopnki^{(1)}$. The gradient and perpendicular gradient of $f$ are then
\begin{align*}
	\nabla f &= \sum_{n,k,i} f_{n,k,i} \: \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \\
	\rvec \times \nabla f &= \sum_{n,k,i} f_{n,k,i} \: \rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big).
\end{align*}
The game is to find an orthogonal basis such that we can sparsely expand $\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ and $\rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ for each $n,k,i$ trio.

Define the OVFs as follows:
\begin{align}
	\tsopinki(x,y,z) &:= \phivec \: \frac{1}{\rho(z)} \: \scopnki^{(0)}(x,y,z) \\
	\tsopiinki(x,y,z) &:= \thetavec \: \frac{1}{\rho(z)} \: \scopnki^{(0)}(x,y,z)
\end{align}
where
\begin{align}
	\phivec &:= 
		\begin{pmatrix}
			\cos(\theta) \: \cos(\phi) \\
			\sin(\theta) \: \cos(\phi) \\
			- \sin(\phi)
		\end{pmatrix}
		=
		\begin{pmatrix}
			x z / \rho(z) \\
			y z / \rho(z) \\
			- \rho(z)
		\end{pmatrix} \\
	\thetavec &:= 
		\begin{pmatrix}
			-\sin(\theta) \\
			\cos(\theta) \\
			0
		\end{pmatrix}
		=
		\begin{pmatrix}
			-y / \rho(z) \\
			x / \rho(z) \\
			0
		\end{pmatrix}
\end{align}
These are orthogonal with respect to the inner product $\ip<\bm{A}, \bm{B}>_\tangentspace := \int_\Omega \bm{A} \cdot \bm{B} \: \rho(z)^2 \:\D \Omega$:
\begin{align*}
	\ip<\tsopinki, \tsopi_{m,j,h}>_\tangentspace &= \int_\Omega \tsopinki \cdot \tsopi_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
	&= \int_\Omega \scopnki^{(0)} \: \scopmjh^{(0)} \: \D \Omega \\
	&= \pi \: \normgenjac^{(0, 2k)} \: \delta_{n,m} \: \delta_{k,j} \: \delta_{i,h}, \\
	\ip<\tsopiinki, \tsopii_{m,j,h}>_\tangentspace &= \int_\Omega \tsopiinki \cdot \tsopii_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
	&= \int_\Omega \scopnki^{(0)} \: \scopmjh^{(0)} \: \D \Omega \\
	&= \pi \: \normgenjac^{(0, 2k)} \: \delta_{n,m} \: \delta_{k,j} \: \delta_{i,h}, \\
	\ip<\tsopinki, \tsopii_{m,j,h}>_\tangentspace &\equiv 0.
\end{align*}
To show that these span the tangent space $\tangentspace$, we merely require that we can sparsely expand $\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \: \rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ with the vectors $\tsopi_{m,j,h}, \tsopii_{m,j,h}$.  First, note that
\begin{align*}
	\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) &= \phivec \: \ppphi \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) + \thetavec \: \frac{1}{\rho(z)} \: \pptheta \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \\
	&= \phivec \: \Big[k \: z \: \genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) -  \big(\genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) \big)' \rho(z)^2 \Big] \: \rho(z)^{k-1} \: \chki(\theta) \\
	&\quad \quad + \thetavec \: (-1)^{i+1} \: k \: \genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) \: \rho(z)^{k-1} \: \ch_{k,|i-1|}(\theta).
\end{align*}
Then,
\begin{align*}
	&\ip<\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \tsopi_{m,j,h}>_\tangentspace \\
	&\quad \quad = \int_\Omega \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \cdot \tsopi_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big[k \: z \: \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} -  (\genjacw^{(1,0)} \: \genjacnmk^{(1,2k)})' \rho^2 \Big] \: \rho^{k-1} \: \genjac_{m-k}^{(0,2k)} \: \rho^{k-1} \: \rho^2 \:\D z \\
	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big[k \: z \: \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} -  (\genjacw^{(1,0)} \: \genjacnmk^{(1,2k)})' \rho^2 \Big] \: \genjac_{m-k}^{(0,2k)} \: \rho^{2k} \:\D z \quad (\star) \\
	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big( kz \: \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(0,2k)} \: \genjacw^{(1,2k)} \\
	&\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad+ \genjacw^{(1,2k)} \: \genjacnmk^{(1,2k)} \: \Big[ \genjac_{m-k}^{(0,2k) \; \prime} \: \rho^2 - (2k+2) \: z \: \genjac_{m-k}^{(0,2k)} \Big] \Big) \: \D z \\
	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \genjacnmk^{(1,2k)} \: \Big[ - (k + 2) \: z \: \genjac_{m-k}^{(0,2k)} + \genjac_{m-k}^{(0,2k) \; \prime} \: \rho^2 \Big] \: \genjacw^{(1,2k)} \: \D z \quad (\star \star)
\end{align*}
where by orthogonality of the $\{\genjac_n^{(a,2b)}\}$ OPs we have that the above is zero for $m - k > n - k + 2 \iff m > n + 2$ at $(\star)$ and also for $n - k > m - k + 1 \iff m < n - 1$ at $(\star \star)$. Further, 
\begin{align*}
	&\ip<\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \tsopii_{m,j,h}>_\tangentspace \\
	&\quad \quad = \int_\Omega \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \cdot \tsopii_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
	&\quad \quad = (-1)^{i+1} \: k \: \pi \: \delta_{k,j} \: \delta_{i,|h-1|} \: \int_\alpha^1 \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} \: \rho^{k-1} \: \genjac_{m-k}^{(0,2k)} \: \rho^{k-1} \: \rho^2 \:\D z \\
	&\quad \quad = (-1)^{i+1} \: k \: \pi \: \delta_{k,j} \: \delta_{i,|h-1|} \: \int_\alpha^1 \genjacnmk^{(1,2k)}  \: \genjac_{m-k}^{(0,2k)} \: \genjacw^{(1,2k)} \: \D z
\end{align*}
which again by orthogonality is zero for $m - k > n - k + 1 \iff m > n + 1$ and $n - k > m - k \iff m < n$. We of course have similar results for $\rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ since
\begin{align*}
	\rvec \times \phivec = \thetavec, \quad \rvec \times \thetavec = - \phivec.
\end{align*}
We now have a set of orthogonal vector valued functions that span the tangent space $\tangentspace$. By construction, we have the same recurrence coefficients for multiplication by $x,y,z$ for each of the OVFs as for the OPs $\scopnki^{(0)}$.

Define
\begin{align*}
	\bigtsoptn := 
		\begin{pmatrix}
			\tsopit_{n,0,0}(x,y,z) \\
			\tsopiit_{n,0,0}(x,y,z) \\
			\tsopit_{n,1,0}(x,y,z) \\
			\tsopiit_{n,1,0}(x,y,z) \\
			\tsopit_{n,1,1}(x,y,z) \\
			\tsopiit_{n,1,1}(x,y,z) \\
			\vdots \\
			\tsopit_{n,n,0}(x,y,z) \\
			\tsopiit_{n,n,0}(x,y,z) \\
			\tsopit_{n,n,1}(x,y,z) \\
			\tsopiit_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2(2n+1) \times 3}, 
	\quad \quad 
	\bigtsopt := 
		\begin{pmatrix}
			\bigtsopt_0 \\
			\bigtsopt_1 \\
			\bigtsopt_2 \\
			\vdots \\
		\end{pmatrix}
\end{align*}
and set $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ as the Jacobi matrices corresponding to
\begin{align}
	\jacobimattangentx \: \bigtsopt(x,y,z) = x \: \bigtsopt(x,y,z), \nonumber \\
	\jacobimattangenty \: \bigtsopt(x,y,z) = y \: \bigtsopt(x,y,z), \label{eqn:jacobimatricestangentdefinition} \\
	\jacobimattangentz \: \bigtsopt(x,y,z) = z \: \bigtsopt(x,y,z), \nonumber
\end{align}
The matrices $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ act on the coefficients vector of a function's expansion in the $\bigtsopt$ basis. For example, let a function $\bm{F}(x,y,z) : \Omega \to \tangentspace$ be approximated by its expansion $\bm{F}(x,y,z)^\top = \bigtsopt(x,y,z)^\top \bm{F}_c$, for some coefficients vector $\bm{F}_c$. Then $x \: \bm{F}(x,y,z)$ is approximated by $\bigtsopt(x,y,z)^\top {(\jacobimattangentx)}^\top \bm{F}_c$. In other words, ${(\jacobimattangentx)}^\top \bm{F}_c$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: \bm{F}(x,y,z)$ in the  $\bigtsopt$ basis. Further, note that $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ are banded-block-banded matrices -- they are block-tridiagonal (block-bandwidths $(1,1)$):
\begin{align*}
	\jacobimattangent_{x/y/z} &= 
		\begin{pmatrix}
			B_{x/y/z, 0} & A_{x/y/z, 0} & & & & \\
			C_{x/y/z, 1} & B_{x/y/z, 1} & A_{x/y/z, 1} & & & \\
			& C_{x/y/z, 2} & B_{x/y/z, 2} & A_{x/y/z, 2} & & & \\
			& & C_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}
\end{align*}
For $\jacobimattangentx$, the sub-blocks have sub-block-bandwidths $(?,?)$:
\begin{align*}
	A_{x,n} &:= 
		\begin{pmatrix}
			0 & A_{n,0,6} & 0 & & \\
			A_{n,1,5} & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & A_{n,n,5} & 0 & A_{n,n,6} \\
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
	B_{x,n} &:= 
		\begin{pmatrix}
			0 & A_{n,0,4} & & \\
			A_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & A_{n,n-1,4} \\
			& & A_{n,n,3} & 0
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
	C_{x,n} &:= 
		\begin{pmatrix}
			0 & A_{n,0,2} & & \\
			A_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots & A_{n,n-2,2} \\
			& & \ddots & 0 \\
			& & & A_{n,n,1} \\
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
\end{align*}
where
\begin{align*}
	A_{n,k,j} &:= \alphaonkj \: I_{4} \in \R^{4\times4}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
	A_{n,0,j} &:=
		\begin{pmatrix}
			\alphao_{n,0,j} & 0 & 0 & 0 \\
			0 & \alphao_{n,0,j} & 0 & 0
		\end{pmatrix} \in \R^{2\times4}, \quad j \text{ even} \\
	A_{n,1,j} &:=
		\begin{pmatrix}
			\alphao_{n,1,j} & 0 \\
			0 & \alphao_{n,1,j} \\
			0 & 0 \\
			0 & 0
		\end{pmatrix} \in \R^{4\times2}, \quad j \text{ odd}
\end{align*}
For $\jacobimattangenty$, the sub-blocks have sub-block-bandwidths $(?,?)$:
\begin{align*}
	A_{y,n} &:= 
		\begin{pmatrix}
			0 & B_{n,0,6} & 0 & & \\
			B_{n,1,5} & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & B_{n,n,5} & 0 & B_{n,n,6} \\
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
	B_{y,n} &:= 
		\begin{pmatrix}
			0 & B_{n,0,4} & & \\
			B_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & B_{n,n-1,4} \\
			& & B_{n,n,3} & 0
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
	C_{y,n} &:= 
		\begin{pmatrix}
			0 & B_{n,0,2} & & \\
			B_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots & B_{n,n-2,2} \\
			& & \ddots & 0 \\
			& & & B_{n,n,1} \\
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
\end{align*}
where
\begin{align*}
	B_{n,k,j} &:= 
		\begin{pmatrix}
			0 & 0 & \betaonkj & 0 \\
			0 & 0 & 0 & \betaonkj \\
			\betaonkj & 0 & 0 & 0 \\
			0 & \betaonkj & 0 & 0
		\end{pmatrix} \in \R^{4\times4}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
	B_{n,0,j} &:=
		\begin{pmatrix}
			0 & 0 & \betao_{n,0,j} & 0 \\
			0 & 0 & 0 & \betao_{n,0,j}
		\end{pmatrix} \in \R^{2\times4}, \quad j \text{ even} \\
	B_{n,1,j} &:=
		\begin{pmatrix}
			0 & 0 \\
			0 & 0 \\
			\betao_{n,1,j} & 0 \\
			0 & \betao_{n,1,j}
		\end{pmatrix} \in \R^{4\times2}, \quad j \text{ odd}
\end{align*}
For $\jacobimattangentz$, the sub-blocks have sub-block-bandwidths $(?,?)$:
\begin{align*}
	A_{z,n} &:= 
		\begin{pmatrix}
			\Gamma_{n,0,3} & 0 & 0 & & \\
			0 & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & 0 & \Gamma_{n,n,3} & 0 \\
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
	B_{z,n} &:= 
		\begin{pmatrix}
			\Gamma_{n,0,2} & 0 & & \\
			0 & \ddots & \ddots & \\
			& \ddots & \ddots & 0 \\
			& & 0 & \Gamma_{n,n,2}
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
	C_{z,n} &:= 
		\begin{pmatrix}
			\Gamma_{n,0,1} & 0 & & \\
			0 & \ddots & \ddots & \\
			& \ddots & \ddots & 0 \\
			& & \ddots &  \Gamma_{n,n-1,1} \\
			& & & 0 \\
		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
\end{align*}
where
\begin{align*}
	\Gamma_{n,k,j} &:= \gammaonkj \: I_{4} \in \R^{4\times4}, \quad k = 1,\dots,n \\
	\Gamma_{n,0,j} &:=
		\begin{pmatrix}
			\gammao_{n,0,j} & 0 \\
			0 & \gammao_{n,0,j}
		\end{pmatrix} \in \R^{2\times2}.
\end{align*}
Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.

\subsection{Building the OPs}

We can combine each system in (\ref{eqn:jacobimatricestangentdefinition}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:

\begin{align*}
\renewcommand\arraystretch{1.3}
\begin{pmatrix}
		I_2 & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
		& & \ddots &\ddots
\end{pmatrix}
\bigtsopt(x,y,z)
=
\begin{pmatrix}
	 \tsopi_0^\top \\ \tsopii_0^\top \\ 0 \\ 0 \\ \vdots  \\
\end{pmatrix},
\end{align*}
where we note $\scopa_{0,0,0}(x,y,z) =: \scopa_0 \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
\begin{align*}
A_n &:= \begin{pmatrix}
		A_{x,n} \\
		A_{y,n} \\
		A_{z,n}
	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n+3)}, \quad
C_n := \begin{pmatrix}
		C_{x,n} \\
		C_{y,n} \\
		C_{z,n}
	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n-1)} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{pmatrix}
		B_{x,n} \\
		B_{y,n} \\
		B_{z,n}
	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n+1)}, \quad
G_n(x,y,z) := \begin{pmatrix}
		xI_{2(2n+1)} \\
		yI_{2(2n+1)} \\
		zI_{2(2n+1)}
	    \end{pmatrix} \in \R^{6(2n+1)\times2(n+1)}.
\end{align*}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{2(2n+3)}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\begin{align*}
	\begin{cases}
		\bigtsopt_{-1}(x,y,z) := 0 \\
		\bigtsopt_{0}(x,y,z) = 
			\begin{pmatrix} 
				\tsopi_0^\top \\ 
				\tsopii_0^\top
			\end{pmatrix} :=
			\begin{pmatrix} 
				\tsopi_{0,0,0}^\top \\
				\tsopi_{0,0,0}^\top
			\end{pmatrix} \\
		\bigtsopt_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigtsoptn(x,y,z) - \Dnt C_n  \, \bigtsopt_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
	\end{cases}
\end{align*}

Note that we can define an explicit \(\Dnt\) as follows: \bstodo{type up DnT for tangent space}
\begin{align*}
\Dnt := \begin{pmatrix}
		0 & & & 0 & & & 1 / \gammaa_{n,0,3} & & \\
		& \ddots & & & \ddots & & & \ddots \\
		& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
		& & & & & \underline{\eta}^\top_{1} & & & \\
		& & & & & \underline{\eta}^\top_{0} & & &
	    \end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
\end{align*}
for $n = 2, 3, \dots$ where $\underline{\eta}_{0}, \underline{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
\begin{align*}
	\big(\underline{\eta}_{0}\big)_j &= 
		\begin{cases}
			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2 \\
			0 & o/w
		\end{cases} \\
	\big(\underline{\eta}_{1}\big)_j &= 
		\begin{cases}
			\big(\underline{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
			0 & o/w
		\end{cases}
\end{align*}
For $n=0, 1$, we can simply take
\begin{align*}
D^\top_0 &:= \begin{pmatrix}
		0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
		\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
		0 & \frac{1}{\betaa_{0,0,6}} & 0
	    \end{pmatrix} \in \R^{3\times3}, \\
D^\top_1 &:= \begin{pmatrix}
		0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,0,3} & & \\
		0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3} \\
		0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3}  \\
		0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & \eta & 0 & 0 \\
		0 & 0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & 0 & 0
	    \end{pmatrix} \in \R^{5\times9},
\end{align*}
where $\eta = \frac{- \: \alphaa_{1,1,5}}{\alphaa_{1,1,6} \: \gammaa_{1, 0, 3}}$.

It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigtsopt_0(x,y,z)$  through $\bigtsoptn(x,y,z)$ in optimal $O(n^2)$ complexity. \bstodo{correct for this case}

\subsection{Operators}

Gradient operator on a scalar function is pretty much outlined above. Divergence would either be worked out by knowing what $f,g$ are for a function $\bm{F} = \nabla f + \rvec \times \nabla g$, so that $\nabla \cdot \bm{F} = \nabla^2 f$. Alternatively, we have expressions for $\rhoppphi \scopnkia$ and $\pptheta \scopnkia$ so that we can calculate an operator for $\rho^2 \nabla \cdot$, i.e. 
\begin{align*}
	\rho^2 \nabla \cdot \bm{F} &= \rho^2 \: \frac{1}{\rho} \ppphi (\rho F^\phi) + \rho^2 \: \frac{1}{\rho} \pptheta F^\theta \\
	&= \sum_{n,k,i} F^\phi_{n,k,i} \: \rhoppphi \scopnki^{(0)} + F^\theta_{n,k,i} \: \pptheta \scopnki^{(0)}.
\end{align*}

\bstodoinline{My main concern is whether this vector function basis is correct/valid. In my mind I would rather something like. I have found it difficult to come up with something that gives orthogonal vectors while still maintaining a sparse expansion for the gradient of a function. I am also confused about what the rho factors "mean".}

%
\section{Conclusions}

We have shown that tri-variate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary. This work provides a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: \sqrt{(1-z^2)}^b$ defined for the disk-slice case \cite{snowball2019sparse}. This work also forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.

This work serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere - t is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere. The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}). Constructing these sparse spectral methods for surface PDEs on spherical bands and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.

%\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.  The second author was supported in part by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces.

%
\appendix
%
\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}

\begin{figure}[t]
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
        	%\label{fig:solutionblocknorms-poisson}
	\centering
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-poisson}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
	%\label{fig:solution-poisson}
	\centering
	\end{subfigure}
	\centering
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element. Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk. Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. in the trapezium.}
	\centering
	\label{fig:appendixfigs}
\end{figure}

We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric. Consider the 3D Dirichlet problem on a domain $\Omega$:
\begin{align*}
	\begin{cases}
         - \rho(z)^2 \: \nabla^2 u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
         u = 0 \quad \text{on } \partial \Omega
         \end{cases}
\end{align*}
This has the weak formulation
\begin{align*}
	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(x,y) \D z = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(x,y) \D z =: a(u,v)
\end{align*}
 for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$.

In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$. While spherical bands are an extension to this work, we note that we can still apply a $p$-FEM using two half-spheres ($\alpha = 0$). In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain. We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.

We seek $u_p \in V_p$ s.t.
\begin{align}
	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
	\label{eqn:FEMweakform}
\end{align}

Define $\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa}$ for any $a$. Note that due to orthogonality this is a diagonal matrix. By choosing suitable bases for our FE spaces, we can rewrite (\ref{eqn:FEMweakform}) in matrix-vector form. We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.

Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
\begin{align*}
	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(x,y) \D z \\
	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(x,y) \D z \\
	&= \int_\element \Big( \mathbf{v}^\top {W_\phi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\phi^{(1)} \mathbf{u} \nonumber \\
					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(x,y) \D z \\
	&= \mathbf{v}^\top \: \Big( {W_\phi^{(1)}}^\top \Lambda^{(0)}  W_\phi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
\end{align*}
Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p = \rho(z) \: \bigscopi_p$. Then,
\begin{align*}
	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(x,y) \D z \\
	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(x,y) \D z \\	
	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
\end{align*}

Since (\ref{eqn:FEMweakform}) is equivalent to stating that
\begin{align*}
	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
\end{align*}
(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
\begin{align*}
	A\mathbf{u} = \tilde{\mathbf{f}}.
\end{align*}
where the (element) stiffness matrix $A$ is defined by 
\begin{align*}
	A =  {W_\phi^{(1)}}^\top \Lambda^{(0)}  W_\phi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
\end{align*}
and the load vector $\tilde{\mathbf{f}}$ is given by 
\begin{align*}
	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
\end{align*}
Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).

\bibliography{spherical-caps}

\end{document}

### 2020-08-06 papers/spherical-caps/spherical-caps.tex
Further edits to paper. Need to update images and ensure prose is correct.

@@
 \newcommand{\gammaonkj}{\gammao_{n,k,j}}
+\newcommand{\uvec}{\bm{u}}
+\newcommand{\hvec}{\bm{h}}
+\newcommand{\uvecc}{\uvec^c}
+\newcommand{\hvecc}{\hvec^c}
+\newcommand{\uvecn}{\uvec_n}
+\newcommand{\hvecn}{\hvec_n}
+\newcommand{\uvecnpi}{\uvec_{n+1}}
+\newcommand{\hvecnpi}{\hvec_{n+1}}
+\newcommand{\uvecnc}{\uvecc_n}
+\newcommand{\hvecnc}{\hvecc_n}
+\newcommand{\uvecnpic}{\uvecnpi^c}
+\newcommand{\hvecnpic}{\hvecnpi^c}
+\newcommand{\href}{\mathcal{H}}
+\newcommand{\divergence}{\nabla \cdot}
+\newcommand{\grad}{\nabla}
 
@@
 \newtheorem{corollary}{Corollary}
+\newtheorem*{remark}{Remark}
 
@@
 
-\section{Introduction}
-
-We can use the similar idea of projecting up from the 2D orthogonal polynomials on the half disk to 3D OPs on the sphere. Define the disk-slice, denoted by $\omega$, by
-\begin{align}
-	\omega := \{(x,y) \in \R^2 \quad | \quad \alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)\}.
-\end{align}
-where:
-\begin{align*}
-	\begin{cases}
-		(\alpha,\beta) &:= (\alpha,1), \quad \alpha \ge 0 \\
-		(\gamma,\delta) &:= (-1,1) \\
-		\rho(x) &:= (1-x^2)^{\half}, \quad x \in (\alpha,\beta) \\
-	\end{cases}
-\end{align*}
-
-Further, define the spherical cap, denoted by $\Omega$, by
-\begin{align}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad (x,y) \in \omega, \: x^2 + y^2 + z^2 = 1\}.
-\end{align}
-We can define the 3D OP family on $\Omega$ as the set of polynomials $\bigQab(x,y,z)$ where
-\begin{align}
-	\bigQab_n(x,y,z) := \begin{pmatrix}
-		\bighdopab_n(x,y) \\
-		z \: \bighdop^{(a,b+1)}_{n-1}(x,y)
-	\end{pmatrix} \in \R^{2n+1}, 
-	\quad \quad 
-	\bigQab := \begin{pmatrix}
-		\bigQab_0 \\
-		\bigQab_1 \\
-		\vdots \\
-		\bigQab_N
-	\end{pmatrix}.
-\end{align}
-The relations for multiplication by $x, y$ are then trivial.
-For multiplication by $z$, note that $z^2 = W^{(0,1)}(x,y)$ for $(x,y,z) \in \Omega$ and consider that (as the 2D OPs for any parameters form a 2D basis of $\omega$):
-\begin{align}
-	z \: \hdopnkab(x,y) &= \sum_{m,j} c_{n,k,m,j}^{(a,b)\to(a,b+1)} \: z \: \hdop^{(a,b+1)}_{m,j}(x,y), \\
-	z^2 \: \hdopnk^{(a,b+1)}(x,y) &= \sum_{m,j} d_{n,k,m,j}^{(a,b+1)\to(a,b)} \: \hdopmjab(x,y).
-\end{align}
-The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ and $d_{n,k,m,j}^{(a,b+1)\to(a,b)}$ are given by
-\begin{align}
-	c_{n,k,m,j}^{(a,b)\to(a,b+1)} &= \frac{\ip< \hdopnkab, \: \hdop^{(a,b+1)}_{m,j} >_{W^{(a,b+1)}}}{\norm{\hdop^{(a,b+1)}_{m,j}}^2_{W^{(a,b+1)}}} \\
-	d_{n,k,m,j}^{(a,b+1)\to(a,b)} &= \frac{\ip< W^{(0,1)} \hdopnk^{(a,b+1)}, \: \hdopmjab >_\Wab}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
-        	&= \frac{\ip< \hdopnk^{(a,b+1)}, \: \hdopmjab >_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \nonumber \\
-        	&= \frac{\norm{\hdopnk^{(a,b+1)}}^2_{W^{(a,b+1)}}}{\norm{\hdopmjab}^2_\Wab} \: c_{m,j,n,k}^{(a,b)\to(a,b+1)}.
-\end{align}
-The coefficients $c_{n,k,m,j}^{(a,b)\to(a,b+1)}$ are just the entries to the parameter transformation operator matrix $T^{(a,b)\to(a,b+1)}$, defined by
-\begin{align}
-	\bighdop^{(a,b)}(x,y) &= \Big(T^{(a,b)\to(a,b+1)} \Big)^T \: \bighdop^{(a,b+1)}(x,y),
-\end{align}
-which can be shown to be sparse with banded-block-banded structure, meaning the resulting operator matrix for the Jacobi operator for multiplication by $z$ is sparse.
-
 %
@@
 
-This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
+This paper develops sparse spectral methods for solving linear partial differential equations on spherical subdomains such as spherical bands and spherical caps.
 More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
@@
 \end{align*}
-Of course, there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$, however for simplicity we focus on the spherical cap.
-
-We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
-On the disk-slice the family of weights we consider are of the form
-$$
-\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \qfor \alpha \leq x \leq \beta,\quad -\rho(x) \leq y \leq \rho(x).
-$$
-The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
-$$
-\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
-$$
-where $C_{n,k} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
+Of course, there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$. However for simplicity we focus on the spherical cap in this paper.
+
+As in the work in two dimensions on the show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that are defined in such a way that mirrors the approach on quadratic curves \cite{olver2020orthogonal}.
+On the spherical cap, the family of weights we consider are of the form
+\begin{align*}
+	\Wa(x,y,z) := (z - \alpha)^a, \qfor (x,y,z) \in \Omega,
+\end{align*}
+noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \; \Omega$.
+The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
+We define these to be orthogonalised lexicographically, that is,
+\begin{align*}
+	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k - \tilde k_i} \: y^{\tilde k_i} \: z^{n-k} + (\hbox{lower order terms})
+\end{align*} \bstodo{Ordering correct?}
+where $C_{n,k,i} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{j - \tilde j_i} \: y^{\tilde j_i} \: z^{n-j}$ where $j < k$.
 The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
 
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$.
+In this work, as for the disk-slice case in 2D \cite{snowball2019sparse}, we use a more general integration-by-parts argument to deduce the sparsity structure.
+The 3D OPs defined here involve the same non-classical (in fact, semi-classical) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
-
-Here is an overview of the paper:
-
-\noindent \secref{OPs}: We present our procedure to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the disk-slice domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.
-
-\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
-We present a method involving use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.
-
-\noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice.
-We present a method for explicitly deriving the recurrence coefficients of the non-classical 1D OPs we detail in \secref{OPs}.
-We derive a quadrature rule on the disk-slice that can be used to expand a function in the OP basis up to a given order.
-Further, we implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
-
-\noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the disk-slice.
-
-\noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods.
-This lays the groundwork for a future $hp$-finite element method in a disk, where the elements capture the circular geometry precisely.
-
-\noindent\appref{HalfDisk}: We discuss extension to the special case of end-disk-slices (e.g., half disks).
-
-\noindent\appref{Trapezium}: We discuss extension to trapezia.
+\bstodo{reword to avoid plagiarising self}
+
+An overview of the paper is as follows:
+
+\noindent \secref{OPs}: We present our definition of a (one-parameter) family of 3D orthogonal polynomials (OPs) on the spherical cap domain $\Omega$, by combining 1D OPs on the interval $(\alpha, 1)$ with Chebyshev polynomials, to form 3D OPs on the spherical cap surface.
+We show that these families will lead to sparse Jacobi operators for multiplication by $x, y, z$ and demonstrate how to obtain the 3D OPs.
+
+\noindent\secref{PDOs}: We define useful partial differential operators and further show that these will also be sparse.
+We can exactly calculate the non-zero entries of these sparse operators using the quadrature rule associated with the non-classical 1D OPs.
+
+\noindent\secref{Computation}: We derive a quadrature rule on the spherical cap that can be used to expand a function in the OP basis up to a given order $N$, and demonstrate how to evaluate a function using the Clenshaw algorithm using the coefficients of its expansion.
+
+\noindent\secref{Examples}: We demonstrate the proposed technique for solving differential equations on the spherical cap such as the Poisson equation, variable coefficient Helmholtz equation, and Biharmonic equation.
+
+\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
+We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
 
@@
 
-We can construct the 3D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from the 2D circular harmonics on the unit circle.
+We can construct the 3D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Chebyshev polynomials.
 
@@
 Define $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ as the 2D OPs on $(x,y)\in\R$ s.t.
-$x^2 + y^2 = 1$ (known as the circular harmonics on the unit circle); that is, if we define $\theta$ by $x = \cos(\theta)$, $y = \sin(\theta)$ for $\xvec := (x, y)$ where $||\xvec|| = 1$,
+$x^2 + y^2 = 1$; that is, if we define $\theta$ by $x = \cos(\theta)$, $y = \sin(\theta)$ for $\xvec := (x, y)$ where $||\xvec|| = 1$,
 \begin{align}
@@
 \begin{align}
-	\ip< p, \: q >_{\ch} &:= \frac{1}{\sqrt{\pi}} \: \int_{||\xvec|| = 1} p(\xvec) \: q(\xvec) \: \D \sigma(\xvec) = \frac{1}{\sqrt{\pi}} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
+	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_{||\xvec|| = 1} p(\xvec) \: q(\xvec) \: \D \sigma(\xvec) = \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
 \label{eqn:ipcircularharmonics}
@@
 \end{definition}
+Of course, these are simply Chebyshev polynomials $\ch_{k,0}(x, y) = T_k(x)$ and $\ch_{k,1}(x, y) = y \: U_{k-1}(x)$ for $k > 0$, where we define $\ch_0$ so as to ensure orthonormality.
 
@@
 
-For the spherical cap, we formalise Proposition \ref{prop:construction} in the following definitions. First we introduce notation for two families of univariate OPs.
-\begin{definition}\label{def:OPconstruction}
+For the spherical cap, we can use Proposition \ref{prop:construction} to create our one-parameter family of OPs.
+We first introduce notation for our family of non-classical univariate OPs that will be used as the $r_n$ polynomials above.
+\begin{definition}[\cite{snowball2019sparse}]\label{def:OPconstruction}
 Let $\genjacw^{(a,b)}(x)$ be a weight function on the interval $(\alpha, \beta)$ given by:
@@
 is a normalising constant.
-Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}).
+Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}).
 \end{definition}
 
+We can now define the 3D OPs for the spherical cap.
 \begin{definition}\label{def:constuction}
@@
 
-$\{\scopnkia\}$ are orthogonal with respect to the inner product
+By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
 \begin{align}
 	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D \sigma(\xvec, z) \\
-	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z.
+	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
 \end{align}
-
-We note that the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}), and has been used in the construction of 2D orthogonal polynomials on disk-slices and trapeziums \cite{snowball2019sparse}.
+with
+\begin{align}
+	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} \nonumber \\
+	&= \pi \: \normgenjac^{(a, 2k)} \label{eqn:normscop}
+\end{align}
+We note that the weight $\genjacw^{(a,b)}(z)$ has been used in the construction of 2D orthogonal polynomials on disk-slices and trapeziums \cite{snowball2019sparse}, where a method for obtaining recurrence coefficients and evaluating integrals was established (the weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
 
@@
 \begin{lemma}\label{lemma:Yrecurrence} 
-The following identities hold for $k = 2,3,\dots$:
+The following identities hold for $k = 2,3,\dots$, $j = 0,1,\dots$ and $i, h \in \{0,1\}$:
 \begin{align*}
@@
 \begin{align*}
-	2 \: \cos(k \theta) \cos(\theta) &= \cos((k-1)\theta) + \cos((k+1)\theta) \\
-	2 \: \sin(k \theta) \cos(\theta) &= \sin((k-1)\theta) + \sin((k+1)\theta) \\
-	2 \: \cos(k \theta) \sin(\theta) &= - \sin((k-1)\theta) + \sin((k+1)\theta) \\
-	2 \: \sin(k \theta) \sin(\theta) &= \cos((k-1)\theta) - \cos((k+1)\theta).
+	2 \cos(k \theta) \cos(\theta) &= \cos((k-1)\theta) + \cos((k+1)\theta) \\
+	2 \sin(k \theta) \cos(\theta) &= \sin((k-1)\theta) + \sin((k+1)\theta) \\
+	2 \cos(k \theta) \sin(\theta) &= - \sin((k-1)\theta) + \sin((k+1)\theta) \\
+	2 \sin(k \theta) \sin(\theta) &= \cos((k-1)\theta) - \cos((k+1)\theta).
 \end{align*}
@@
 		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
-z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z), \\
+	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
 \end{align*}
@@
 		\end{cases} \\	
-	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)} \\
+	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}
 \end{align*}
@@
 where $\delta_{k, j}$ is the standard Kronecker delta function, using Lemma \ref{lemma:Yrecurrence}. Similarly, for the recurrence for multiplication by $y$, we can expand 
-$$
+\begin{align*}
 	y \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 d_{m,j} \: \scopmjha(x,y,z).
-$$ 
+\end{align*}
 These coefficients are given by
@@
 \begin{align*}
-\bigscoptna := \begin{pmatrix}
+	\bigscoptna := 
+		\begin{pmatrix}
 			\scopa_{n,0,0}(x,y,z) \\
@@
 	\quad \quad 
-\bigscopta := \begin{pmatrix}
+	\bigscopta := 
+		\begin{pmatrix}
 			\bigscopta_0 \\
@@
 	J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinition} \\
-	J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z), \nonumber
+	J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
 \end{align}
-The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis.
+%The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis.
 For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated by its expansion $f(x,y,z) = \bigscopta(x,y,z)^\top \mathbf{f}$.
@@
 In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis.
-Further, note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
-
+We will note here that we present these definitions of $\bigscoptna$, $\bigscopta$ and the Jacobi matrices in order to explain how we can derive the relationship for building the OPs themselves, i.e.
+gather a relationship \bstodoinline{is Q the right way to go or shall i derive the clenshaw matrices in Fourier mode k order?}
+Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
 \begin{definition}
@@
 \end{definition}
-
 For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
 \begin{align*}
-J_{x/y/z}^{(a)} &= \begin{pmatrix}
+	J_{x/y/z}^{(a)} &= 
+		\begin{pmatrix}
 			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
@@
 \begin{align*}
-A^{(a)}_{x,n} &:= \begin{pmatrix}
+	A^{(a)}_{x,n} &:= 
+		\begin{pmatrix}
 			0 & A^{(a)}_{n,0,6} & 0 & & \\
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-B^{(a)}_{x,n} &:= \begin{pmatrix}
+	B^{(a)}_{x,n} &:= 
+		\begin{pmatrix}
 			0 & A^{(a)}_{n,0,4} & & \\
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-C^{(a)}_{x,n} &:= \begin{pmatrix}
+	C^{(a)}_{x,n} &:= 
+		\begin{pmatrix}
 			0 & A^{(a)}_{n,0,2} & & \\
@@
 \begin{align*}
-A^{(a)}_{y,n} &:= \begin{pmatrix}
+	A^{(a)}_{y,n} &:= 
+		\begin{pmatrix}
 			0 & B^{(a)}_{n,0,6} & 0 & & \\
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-B^{(a)}_{y,n} &:= \begin{pmatrix}
+	B^{(a)}_{y,n} &:= 
+		\begin{pmatrix}
 			0 & B^{(a)}_{n,0,4} & & \\
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-C^{(a)}_{y,n} &:= \begin{pmatrix}
+	C^{(a)}_{y,n} &:= 
+		\begin{pmatrix}
 			0 & B^{(a)}_{n,0,2} & & \\
@@
 \begin{align*}
-A^{(a)}_{z,n} &:= \begin{pmatrix}
+	A^{(a)}_{z,n} &:= 
+		\begin{pmatrix}
 			\Gamma^{(a)}_{n,0,3} & 0 & \\
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-B^{(a)}_{z,n} &:= \begin{pmatrix}
+	B^{(a)}_{z,n} &:= 
+		\begin{pmatrix}
 			\Gamma^{(a)}_{n,0,2} & \\
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-C^{(a)}_{z,n} &:= \begin{pmatrix}
+	C^{(a)}_{z,n} &:= 
+		\begin{pmatrix}
 			\Gamma^{(a)}_{n,0,1} & 0 & & \\
@@
 	\end{pmatrix}
-\bigscopta(x,y,z)
-=
+	\bigscopta(x,y,z) =
 		\begin{pmatrix}
@@
 \begin{align*}
-A_n &:= \begin{pmatrix}
+	A_n &:= 
+		\begin{pmatrix}
 			A^{(a)}_{x,n} \\
@@
 		\end{pmatrix} \in \R^{3(2n+1)\times(2n+3)}, \quad
-C_n := \begin{pmatrix}
+	C_n := 
+		\begin{pmatrix}
 			C^{(a)}_{x,n} \\
@@
 		\end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \nonumber \\
-B_n &:= \begin{pmatrix}
+	B_n &:= 
+		\begin{pmatrix}
 			B^{(a)}_{x,n} \\
@@
 		\end{pmatrix} \in \R^{3(2n+1)\times(2n+1)}, \quad
-G_n(x,y,z) := \begin{pmatrix}
+	G_n(x,y,z) := 
+		\begin{pmatrix}
 			xI_{2n+1} \\
@@
 \begin{align*}
-\Dnt := \begin{pmatrix}
+	\Dnt := 
+		\begin{pmatrix}
 			0 & & & 0 & & & 1 / \gammaa_{n,0,3} & & \\
@@
 			& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
-		& & & & & \underline{\eta}^\top_{1} & & & \\
-		& & & & & \underline{\eta}^\top_{0} & & &
+			& & & & & \bm{\eta}^\top_{1} & & & \\
+			& & & & & \bm{\eta}^\top_{0} & & &
 		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
 \end{align*}
-for $n = 2, 3, \dots$ where $\underline{\eta}_{0}, \underline{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
-\begin{align*}
-	\big(\underline{\eta}_{0}\big)_j &= 
+for $n = 2, 3, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
+\begin{align*}
+	\big(\bm{\eta}_{0}\big)_j &= 
 		\begin{cases}
@@
 		\end{cases} \\
-	\big(\underline{\eta}_{1}\big)_j &= 
+	\big(\bm{\eta}_{1}\big)_j &= 
 		\begin{cases}
-			\big(\underline{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
+			\big(\bm{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
 			0 & o/w
@@
 \begin{align*}
-D^\top_0 &:= \begin{pmatrix}
+	D^\top_0 &:= 
+		\begin{pmatrix}
 			0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
@@
 		\end{pmatrix} \in \R^{3\times3}, \\
-D^\top_1 &:= \begin{pmatrix}
+	D^\top_1 &:= 
+		\begin{pmatrix}
 			0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,0,3} & & \\
@@
 
+These operators are be rotationally invariant, and hence with the ordering we have for the OPs in $\bigscopNa$, obtaining and applying the operators can be done in parallel since each block need not talk to each other (they are block diagonal).
 In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short Lemma.
@@
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
-As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored surface Laplacian $\rho(z)^2 \: \nabla^2_s$, that will take us from coefficients for expansion in the weighted space
-$$
-\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)
-$$
-to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
+As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored surface Laplacian $\rho(z)^2 \: \nabla^2_s$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in Definition \ref{def:differentialoperators}.
@@
 
-Another important example is the Biharmonic operator $\Delta^2_s$ (defined by $\Delta^2_s f = \nabla^2_s (\nabla^2_s f)$), where we assume zero Dirichlet and Neumann conditions.
-To construct an operator for the Biharmonic function, we first note that we can obtain the matrix operator for the $\rho^2$-factored Laplacian $\rho(z)^2 \: \nabla^2_s$ that will take us from coefficients for expansion in the space $\bigscopN^{(0)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
-We denote this matrix operator that acts on the coefficients vector as $\mathcal{L}^{(0) \to (2)}$, and is given by
+Another desirable operator is the Biharmonic operator $\Delta^2_s$ (defined by $\Delta^2_s f = \nabla^2_s (\nabla^2_s f)$), for which we assume zero Dirichlet and Neumann conditions.
+To construct a matrix operator, we first note that we can obtain the matrix operator for the $\rho^2$-factored Laplacian $\rho(z)^2 \: \nabla^2_s$ that will take us from coefficients for expansion in the space $\bigscopN^{(0)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
+We denote this matrix operator that acts on the coefficients vector as $\mathcal{L}^{(0) \to (2)}$, and it is given by
 \begin{align*}
@@
 \end{align*}
-We can then construct a matrix operator for $\Delta^2_s$ that will take coefficients in the space $\bigW_N^{(2)}$ to coefficients in the space $\bigscopN^{(2)}$.
-Note that any function expanded in the $\bigW_N^{(2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$.
 Since, for any scalar function $f:\Omega \to \R$,
@@
 \end{align*}
-where $V^{(a)}$ is an operator corresponding to multiplication by the scalar function $v(z) := 2z^2 - 2\rho(z)^2 + 3z = 4z^2 + 3z - 2$ acting on coefficients in the $\bigscopN^{(a)}$ space.
+where $V^{(a)}$ is an operator corresponding to multiplication by the scalar polynomial $v(z) := 2z^2 - 2\rho(z)^2 + 3z = 4z^2 + 3z - 2$ acting on coefficients in the $\bigscopN^{(a)}$ space (see Section \ref{subsection:operatorclenshaw} for how to computationally obtain operators representing functional coefficients).
 The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
@@
 Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
-Further, we denote the $M_2$ Gauss quadrature nodes and weights for the unit circle \cite{???}\bstodo{good reference?} by $(\mathbf{s}_j, w_j^{(s)})$ where for $j = 1,\dots,M_2$:
-\begin{align*}
-	\mathbf{s}_j &= \Big( \cos\fpr(\frac{2 \pi j}{M_2}), \: \sin\fpr(\frac{2 \pi j}{M_2}) \Big), \\
-	w_j^{(s)} &= \frac{2\pi}{M_2}.
-\end{align*}
+Further, denote the $M_2$ Gauss quadrature nodes and weights $[-1,1]$ with weight $(1 - x^2)^{-\half}$ as $(s_j, w_j^{(s)})$.
 Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
@@
 \begin{align*}
-	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
+	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
 \end{align*}
@@
 
+\begin{remark} 
+	Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[3.5.23]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
+\end{remark}
+
 \begin{proof}
@@
 so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
-Note that if $f$ is a polynomial, then $f_e(\rho(t)\xvec, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $\xvec \in \R^2$.
-
-Now, we have that
+Note that if $f$ is a polynomial, then $f_e(\rho(t)x, \rho(t)y, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $(x,y) \in \R^2$.
+
+Firstly, we note that
+\begin{align*}
+	\int_0^{2\pi} g\big(\cos(\theta), \sin(\theta)\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
+\end{align*}
+for some function $g$. Then, integrating the even function $f_e$ we have
 \begin{align*}
 	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\mathbf{\xi}} = 1} f_e\big(\rho(z)\mathbf{\xi}, z)\big) \: \D \sigma(\mathbf{\xi}) \Big) \: \D z \\
-	&\quad \quad \approx\int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) \mathbf{s}_l, z\big) \Big) \: \D z \quad (\star) \\
-	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) \mathbf{s}_l, t_j\big) \quad (\star \star) \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z\big) \: \D \theta \Big) \: \D z \\
+	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \D x \Big) \: \D z \\
+	&\quad \quad \approx \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
+	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) s_l, \rho(t_j) \sqrt{1 - s_l^2}, t_j\big) \quad (\star \star) \\
 	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
@@
 Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
-We therefore achieve equality at $(\star)$ if $M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-\bstodo{For the circle rule, is this bound on N true?}
-
-Next, note that 
+It follows that $s \mapsto f_e\big(\rho(z)s,\rho(z)\sqrt{1-s^2}, z\big)$ for fixed $z$ is then a polynomial of degree $\le N$.
+We therefore achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
+
+Integrating the odd function $f_o$ results in
 \begin{align*}
 	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\mathbf{\xi}} = 1} f_o\big(\rho(z)\mathbf{\xi}, z)\big) \: \D \sigma(\mathbf{\xi}) \Big) \: \D z \quad (\dagger) \\
-	&\quad \quad = 0
-\end{align*}
-since the inner integral at $(\dagger)$ over $\mathbf{\xi}$ is zero, due to the symmetry over the domain.
-
-Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) + f_o\big(\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) - f_o\big(\rho(z)x, \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
+	&\quad \quad = 0.
+\end{align*}
+since $f_o(x,y,z) = - f_o(-x, -y, z)$. Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
 \begin{align*}
@@
 \end{align*}
-where $M = M_1 M_2$ and $2M_1 - 1 \ge N, M_2 - 1 \ge N$.
+where $M = M_1 M_2$ and $2M_1 - 1 \ge N, 2M_2 - 1 \ge N$.
 \end{proof}
+
+%\begin{theorem}\label{Theorem:quadrule}
+%Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
+Further, we denote the $M_2$ Gauss quadrature nodes and weights for the unit circle \cite{???}\bstodo{good reference?
+Or derive myself using Chebychev polys...} by $(\mathbf{s}_j, w_j^{(s)})$ where for $j = 1,\dots,M_2$:
+%\begin{align*}
+%	\mathbf{s}_j &= \Big( \cos\fpr(\frac{2 \pi j}{M_2}), \: \sin\fpr(\frac{2 \pi j}{M_2}) \Big), \\
+%	w_j^{(s)} &= \frac{2\pi}{M_2}.
+%\end{align*}
+%Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
+%\begin{align*}
+%	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
+%	z_{l+(j-1)M_2} &:= t_j, \\
+%	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
+%\end{align*}
+%Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
+%\begin{align*}
+%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
+%\end{align*}
+%where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
+%\end{theorem}
+%
+%\begin{proof}
+%Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
+%\begin{align*}
+%	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
+%	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
+%\end{align*}
+%so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
+Note that if $f$ is a polynomial, then $f_e(\rho(t)\xvec, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $\xvec \in \R^2$.
+%
+%Now, we have that
+%\begin{align*}
+%	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
+%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_e\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \\
+%	&\quad \quad \approx\int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) \mathbf{s}_l, z\big) \Big) \: \D z \quad (\star) \\
+%	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) \mathbf{s}_l, t_j\big) \quad (\star \star) \\
+%	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
+%\end{align*}
+%Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
+We therefore achieve equality at $(\star)$ if $M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
+\bstodo{For the circle rule, is this bound on N true?}
+%
+%Next, note that 
+%\begin{align*}
+%	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
+%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_o\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \quad (\dagger) \\
+%	&\quad \quad = 0
+%\end{align*}
+%since the inner integral at $(\dagger)$ over $\bm{\xi}$ is zero, due to the symmetry over the domain.
+%
+%Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
+%\begin{align*}
+%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z  \\
+%	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
+%	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
+%\end{align*}
+%where $M = M_1 M_2$ and $2M_1 - 1 \ge N, M_2 - 1 \ge N$.
+%\end{proof}
 
@@
 This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodo{correct the O(nd) complexity}.
+
+\subsection{Obtaining operator matrices for variable coefficients}\label{subsection:operatorclenshaw}
+
+Hey
 
@@
 		& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
-		& & & & & \underline{\eta}^\top_{1} & & & \\
-		& & & & & \underline{\eta}^\top_{0} & & &
+		& & & & & \bm{\eta}^\top_{1} & & & \\
+		& & & & & \bm{\eta}^\top_{0} & & &
 	    \end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
 \end{align*}
-for $n = 2, 3, \dots$ where $\underline{\eta}_{0}, \underline{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
-\begin{align*}
-	\big(\underline{\eta}_{0}\big)_j &= 
+for $n = 2, 3, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
+\begin{align*}
+	\big(\bm{\eta}_{0}\big)_j &= 
 		\begin{cases}
@@
 \bstodoinline{My main concern is whether this vector function basis is correct/valid.
-In my mind I would rather something like.
 I have found it difficult to come up with something that gives orthogonal vectors while still maintaining a sparse expansion for the gradient of a function.
 I am also confused about what the rho factors "mean".}
+
+\subsection{Linear SWE}
+
+\begin{align*}
+	\uvec_t &= - f \rvec \times \uvec + \grad h \\
+	h_t &= -\href \divergence \uvec
+\end{align*}
+$\implies$
+\begin{align*}
+	\uvecnpi &= \uvecn - \Delta t \: f \rvec \times \uvecnpi +\Delta t \: \grad h_{n+1} \\
+	h_{n+1} &= h_n - \Delta t \: \href \divergence \uvecnpi
+\end{align*}
+$\implies$
+\begin{align*}
+	\uvecnpic &= \uvecnc - \Delta t \: F \uvecnpic +\Delta t \: G \hvecnpic \\
+	P \hvecnpic &= P \hvecnc - \Delta t \: \href D \uvecnpic
+\end{align*}
+where $F : \bigtsop \to \bigtsop$ is the operator for $f \rvec \times$, $D : \bigtsop \to \bigscop^{(1)}$ is the operator for $\rho^2 \divergence$, $G : \bigW^{(1)} \to \bigtsop$ is the operator for $\grad$, and $P : \bigW^{(1)} \to \bigscop^{(1)}$ is the operator for multiplication by $\rho(z)^2$.
 
@@
 
-This work serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere - t is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
+This work serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere -- it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
 Constructing these sparse spectral methods for surface PDEs on spherical bands and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
+
+The next stage of this work is to develop an orthogonal basis for the tangent space of the spherical cap/band, and obtain sparse differential operators for gradient, divergence etc.
+On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
 

### 2020-08-06 papers/spherical-caps/spherical-caps.tex
Further edits to paper

@@
 \newcommand{\rhoppphi}{\rho \ppphi}
+\newcommand{\rhozppphi}{\rho(z) \ppphi}
 \newcommand{\pptheta}{\frac{\partial}{\partial \theta}}
+\newcommand{\ppthetatwo}{\frac{\partial^2}{\partial \theta^2}}
 \newcommand{\bigWNi}{\mathbb{W}_N^{(1)}}
@@
 \begin{align}
-	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(x) \: \rho(x)^k \: \chki\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega, 
 \end{align}
@@
 
-\subsection{Jacobi matrices}
+For clarity we state here that throughout we will regularly use coordinate transformations, mainly in any integrals over the surface $\Omega$, between the cartesian $(x,y,z)$ and the spherical coordinates $(\phi, \theta)$ -- that is, for any point $(x,y,z) \in \Omega$, let $(\phi, \theta)$ denote the angles such that:
+\begin{align*}
+	z &= \cos(\phi) \\
+	x &= \sin(\phi) \: \cos(\theta) = \rho(z) \: \cos(\theta) \\
+	y &= \sin(\phi) \: \sin(\theta) = \rho(z) \: \sin(\theta)
+\end{align*}
+
+\subsection{Jacobi matrices}\label{subsection:jacobimats}
 
@@
 \end{align}
-$\scopnkia(x,y,z)$ satisfy the following 3-term recurrences:
+$\scopnkia(x,y,z)$ satisfy the following recurrences:
 \begin{align*}
@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Define, for $n=0,1,2,\dots$: 
+The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define: 
+\begin{align}
+	\bigscopNka &:= 
+		\begin{pmatrix}
+			\scopa_{k,k,0}(x,y,z) \\
+			\scopa_{k,k,1}(x,y,z) \\
+			\vdots \\
+			\scopa_{N,k,0}(x,y,z) \\
+			\scopa_{N,k,1}(x,y,z)
+		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
+	\bigscopa_{N,0} &:= 
+		\begin{pmatrix}
+			\scopa_{0,0,0}(x,y,z) \\
+			\vdots \\
+			\scopa_{N,0,0}(x,y,z) \\
+		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
+	\bigscopNa &:= 
+		\begin{pmatrix}
+			\bigscopa_{N,0} \\
+			\vdots \\
+			\bigscopa_{N,N}
+		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
+\end{align}
+and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
+\begin{align}
+	J_x^{(a,b,c,d)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
+	J_y^{(a,b,c,d)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
+	J_z^{(a,b,c,d)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
+\end{align}
+The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis.
+For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$.
+Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$.
+In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis.
+Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
+
+\begin{definition}
+A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
+A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
+\end{definition}
+
+For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
+\begin{align*}
+	J_{x/y/z}^{(a)} &= 
+		\begin{pmatrix}
+			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
+			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
+			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
+			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
+			& & & \ddots & \ddots & \ddots \\
+		\end{pmatrix}
+\end{align*}
+
+For $J_x^{(a)}$, the sub-blocks $A^{(a)}_{x, k}, C^{(a)}_{x, k}$ have sub-block-bandwidths $(2,2)$ while the sub-blocks $B^{(a)}_{x,k}$ are zero matrices: \bstodo{fix bandwidths}
+\begin{align*}
+	B^{(a)}_{x,0} &= 0 \in \R^{(N+1)\times2N}, \quad B^{(a)}_{x,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	A^{(a)}_{x,k} &:= 
+		\begin{pmatrix}
+			A^{(a)}_{k,k,6} & & & \\
+			A^{(a)}_{k+1,k,4} & \ddots & & \\
+			A^{(a)}_{k+2,k,2} & \ddots & \ddots & \\
+			& \ddots & \ddots & A^{(a)}_{N-1,k,6} \\
+			& & A^{(a)}_{N,k,2} & A^{(a)}_{N,k,4}
+		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
+	A^{(a)}_{x,0} &\in \R^{(N+1)\times2N}, \quad A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	C^{(a)}_{x,k} &:= 
+		\begin{pmatrix}
+			A^{(a)}_{k,k,1} & A^{(a)}_{k,k,3} & A^{(a)}_{k,k,5} & & \\
+			& \ddots & \ddots & \ddots \\
+			& & \ddots & \ddots & A^{(a)}_{N-1,k,5} \\
+			& & & A^{(a)}_{N,k,1} & A^{(a)}_{N,k,3}
+		\end{pmatrix} \in \R^{2(N-k+1)\times2(N-k+2)}, \quad k = 1,2,\dots \\
+	C^{(a)}_{x,1} &\in \R^{2N\times(N+1)}, \quad C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N
+\end{align*}
+where for $k = 1,\dots,N, \: n = k,\dots,N$
+\begin{align*}
+	A^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			\alphaa_{n,k,j} & 0 \\
+			0 & \alphaa_{n,k,j}
+		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
+	A^{(a)}_{n,0,j} &:=
+		\begin{pmatrix}
+			\alphaa_{n,0,j} & 0
+		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \\
+	A^{(a)}_{n,1,j} &:=
+		\begin{pmatrix}
+			\alphaa_{n,1,j} \\
+			0
+		\end{pmatrix} \in \R^{2\times1}, j \text{ odd} \\
+\end{align*}
+
+For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$ while the sub-blocks $B^{(a)}_{y,k}$ are zero matrices:
+\begin{align*}
+	B^{(a)}_{y,0} &= 0 \in \R^{(N+1)\times2N}, \quad B^{(a)}_{y,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	A^{(a)}_{y,k} &:= 
+		\begin{pmatrix}
+			B^{(a)}_{k,k,6} & & & \\
+			B^{(a)}_{k+1,k,4} & \ddots & & \\
+			B^{(a)}_{k+2,k,2} & \ddots & \ddots & \\
+			& \ddots & \ddots & B^{(a)}_{N-1,k,6} \\
+			& & B^{(a)}_{N,k,2} & B^{(a)}_{N,k,4}
+		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
+	A^{(a)}_{y,0} &\in \R^{(N+1)\times2N}, \quad A^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	C^{(a)}_{y,k} &:= 
+		\begin{pmatrix}
+			B^{(a)}_{k,k,1} & B^{(a)}_{k,k,3} & B^{(a)}_{k,k,5} & & \\
+			& \ddots & \ddots & \ddots \\
+			& & \ddots & \ddots & B^{(a)}_{N-1,k,5} \\
+			& & & B^{(a)}_{N,k,1} & B^{(a)}_{N,k,3}
+		\end{pmatrix} \in \R^{2(N-k+1)\times2(N-k+2)}, \quad k = 1,2,\dots \\
+	C^{(a)}_{y,1} &\in \R^{2N\times(N+1)}, \quad C^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N
+\end{align*}
+where for $k = 1,\dots,N, \: n = k,\dots,N$
+\begin{align*}
+	B^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			0 & \betaa_{n,k,0,j} \\
+			\betaa_{n,k,1,j} & 0
+		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
+	B^{(a)}_{n,0,j} &:=
+		\begin{pmatrix}
+			0 & \betaa_{n,0,0,j}
+		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \\
+	B^{(a)}_{n,1,j} &:=
+		\begin{pmatrix}
+			0 \\
+			\betaa_{n,1,1,j}
+		\end{pmatrix}\in \R^{2\times1}, j \text{ odd} \\
+\end{align*}
+
+For $J_z^{(a)}$, the sub-blocks $B^{(a)}_{z,k}$ have sub-block-bandwidths $(3,3)$, while the sub-blocks $A^{(a)}_{z,k}, \: C^{(a)}_{z,k}$ are zero matrices:
+\begin{align*}
+	A^{(a)}_{z,0} &= 0 \in \R^{(N+1)\times2N}, \quad A^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	C^{(a)}_{z,1} &= 0 \in \R^{2N\times(N+1)}, \quad C^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N \\
+	B^{(a)}_{z,k} &:= 
+		\begin{pmatrix}
+			\Gamma^{(a)}_{k,k,2} & \Gamma^{(a)}_{k,k,3} & \\
+			\Gamma^{(a)}_{k+1,k,1} & \ddots & \ddots & \\
+			& \ddots & \ddots & \Gamma^{(a)}_{N-1,k,3} \\
+			& & \Gamma^{(a)}_{N,k,1} & \Gamma^{(a)}_{N,k,2}
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
+	B^{(a)}_{z,0} &\in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{z,k} \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N
+\end{align*}
+where for $k = 1,\dots,N, \: n = k,\dots,N$
+\begin{align*}
+	\Gamma^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			\gamma_{n,k,j} & 0 \\
+			0 & \gamma_{n,k,j}
+		\end{pmatrix} \in \R^{2\times2}, \\
+	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}.
+\end{align*}
+
+Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
+
+\subsection{Building the OPs}
+
+In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated 'Clenshaw matrices'.
+While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the 'Clenshaw matrices' necessary for efficient function evaluation using the Clenshaw algorithm.
+In practice, it is simply a matter of converting coefficients between the two orderings.
+
+Define, for $n=0,1,2,\dots$: 
 \begin{align*}
@@
 \end{align*}
-and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
+and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to
 \begin{align}
-	J_x^{(a,b,c,d)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
-	J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinition} \\
-	J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
+	\tilde J_x^{(a,b,c,d)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
+	\tilde J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
+	\tilde J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
 \end{align}
-%The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis.
-For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated by its expansion $f(x,y,z) = \bigscopta(x,y,z)^\top \mathbf{f}$.
-Then $x \: f(x,y,z)$ is approximated by $\bigscopta(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis.
-We will note here that we present these definitions of $\bigscoptna$, $\bigscopta$ and the Jacobi matrices in order to explain how we can derive the relationship for building the OPs themselves, i.e.
-gather a relationship \bstodoinline{is Q the right way to go or shall i derive the clenshaw matrices in Fourier mode k order?}
-Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
-\begin{definition}
-A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
-A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
-\end{definition}
-For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
-\begin{align*}
-	J_{x/y/z}^{(a)} &= 
-		\begin{pmatrix}
-			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
-			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
-			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
-			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
+where
+\begin{align*}
+	\tilde J_{x/y/z}^{(a)} &= 
+		\begin{pmatrix}
+			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
+			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
+			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
+			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
 			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}
-\end{align*}
-For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
-\begin{align*}
-	A^{(a)}_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A^{(a)}_{n,0,6} & 0 & & \\
-			A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
+		\end{pmatrix}.
+\end{align*}
+These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).
+
+For $\tilde J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
+\begin{align*}
+	\tilde A^{(a)}_{x,n} &:= 
+		\begin{pmatrix}
+			0 & \tilde A^{(a)}_{n,0,6} & 0 & & \\
+			\tilde A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
+			& & \tilde A^{(a)}_{n,n,5} & 0 & \tilde A^{(a)}_{n,n,6} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	B^{(a)}_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A^{(a)}_{n,0,4} & & \\
-			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
-			& & A^{(a)}_{n,n,3} & 0
+	\tilde B^{(a)}_{x,n} &:= 
+		\begin{pmatrix}
+			0 & \tilde A^{(a)}_{n,0,4} & & \\
+			\tilde A^{(a)}_{n,1,3} & \ddots & \ddots & \\
+			& \ddots & \ddots & \tilde A^{(a)}_{n,n-1,4} \\
+			& & \tilde A^{(a)}_{n,n,3} & 0
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	C^{(a)}_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A^{(a)}_{n,0,2} & & \\
-			A^{(a)}_{n,1,1} & \ddots & \ddots & \\
+	\tilde C^{(a)}_{x,n} &:= 
+		\begin{pmatrix}
+			0 & \tilde A^{(a)}_{n,0,2} & & \\
+			\tilde A^{(a)}_{n,1,1} & \ddots & \ddots & \\
 			& \ddots & \ddots &A^{(a)}_{n,n-2,2} \\
 			& & \ddots & 0 \\
-			& & & A^{(a)}_{n,n,1} \\
+			& & & \tilde A^{(a)}_{n,n,1} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
@@
 \begin{align*}
-	A^{(a)}_{n,k,j} &:= 
+	\tilde A^{(a)}_{n,k,j} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
-	A^{(a)}_{n,0,j} &:=
+	\tilde A^{(a)}_{n,0,j} &:=
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{1\times2}, \quad j \text{ even} \\
-	A^{(a)}_{n,1,j} &:=
+	\tilde A^{(a)}_{n,1,j} &:=
 		\begin{pmatrix}
@@
 \end{align*}
-For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
-\begin{align*}
-	A^{(a)}_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B^{(a)}_{n,0,6} & 0 & & \\
-			B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
+For $\tilde J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
+\begin{align*}
+	\tilde A^{(a)}_{y,n} &:= 
+		\begin{pmatrix}
+			0 & \tilde B^{(a)}_{n,0,6} & 0 & & \\
+			\tilde B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
+			& & \tilde B^{(a)}_{n,n,5} & 0 & \tilde B^{(a)}_{n,n,6} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	B^{(a)}_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B^{(a)}_{n,0,4} & & \\
-			B^{(a)}_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & B^{(a)}_{n,n-1,4} \\
-			& & B^{(a)}_{n,n,3} & 0
+	\tilde B^{(a)}_{y,n} &:= 
+		\begin{pmatrix}
+			0 & \tilde B^{(a)}_{n,0,4} & & \\
+			\tilde B^{(a)}_{n,1,3} & \ddots & \ddots & \\
+			& \ddots & \ddots & \tilde B^{(a)}_{n,n-1,4} \\
+			& & \tilde B^{(a)}_{n,n,3} & 0
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	C^{(a)}_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B^{(a)}_{n,0,2} & & \\
-			B^{(a)}_{n,1,1} & \ddots & \ddots & \\
-			& \ddots & \ddots & B^{(a)}_{n,n-2,2} \\
+	\tilde C^{(a)}_{y,n} &:= 
+		\begin{pmatrix}
+			0 & \tilde B^{(a)}_{n,0,2} & & \\
+			\tilde B^{(a)}_{n,1,1} & \ddots & \ddots & \\
+			& \ddots & \ddots & \tilde B^{(a)}_{n,n-2,2} \\
 			& & \ddots & 0 \\
-			& & & B^{(a)}_{n,n,1} \\
+			& & & \tilde B^{(a)}_{n,n,1} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
@@
 \begin{align*}
-	B^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			0 & \betaa_{n,k,j} \\
-			\betaa_{n,k,j} & 0
+	\tilde B^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			0 & \betaa_{n,k,0,j} \\
+			\betaa_{n,k,1,j} & 0
 		\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
-	B^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			0 & \betaa_{n,0,j}
+	\tilde B^{(a)}_{n,0,j} &:=
+		\begin{pmatrix}
+			0 & \betaa_{n,0,0,j}
 		\end{pmatrix} \in \R^{1\times2}, \quad j \text{ even} \\
-	B^{(a)}_{n,1,j} &:=
+	\tilde B^{(a)}_{n,1,j} &:=
 		\begin{pmatrix}
 			0 \\
-			\betaa_{n,1,j}
+			\betaa_{n,1,1,j}
 		\end{pmatrix} \in \R^{2\times1}, \quad j \text{ odd} \\
 \end{align*}
-For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
-\begin{align*}
-	A^{(a)}_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{n,0,3} & 0 & \\
+For $\tilde J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
+\begin{align*}
+	\tilde A^{(a)}_{z,n} &:= 
+		\begin{pmatrix}
+			\tilde \Gamma^{(a)}_{n,0,3} & 0 & \\
 			0 & \ddots & \ddots & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & 0 & \Gamma^{(a)}_{n,0,3} & 0 \\
+			& & 0 & \tilde \Gamma^{(a)}_{n,0,3} & 0 \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	B^{(a)}_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{n,0,2} & \\
+	\tilde B^{(a)}_{z,n} &:= 
+		\begin{pmatrix}
+			\tilde \Gamma^{(a)}_{n,0,2} & \\
 			& \ddots & & \\
 			& & \ddots & \\
-			& & & \Gamma^{(a)}_{n,0,2}
+			& & & \tilde \Gamma^{(a)}_{n,0,2}
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	C^{(a)}_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{n,0,1} & 0 & & \\
+	\tilde C^{(a)}_{z,n} &:= 
+		\begin{pmatrix}
+			\tilde \Gamma^{(a)}_{n,0,1} & 0 & & \\
 			0 & \ddots & \ddots & \\
 			& \ddots & \ddots & 0 \\
-			& & \ddots & \Gamma^{(a)}_{n,n,1} \\
+			& & \ddots & \tilde \Gamma^{(a)}_{n,n,1} \\
 			& & & 0 \\
@@
 \begin{align*}
-	\Gamma^{(a)}_{n,k,j} &:= 
+	\tilde \Gamma^{(a)}_{n,k,j} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \\
-	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}
-\end{align*}
-
-Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
-
-\subsection{Building the OPs}
-
-We can combine each system in (\ref{eqn:jacobimatricesdefinition}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
-
+	\tilde \Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}
+\end{align*}
+
+We can combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
 \begin{align*}
@@
 \end{align*}
-where we note $\scopa_{0,0,0}(x,y,z) =: \scopa_0 \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
+where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
 \begin{align*}
@@
 
+\begin{definition}\label{def:clenshawmats}
+The Clenshaw matrices associated with the OPs $\{\scopnkia\}$ given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
+\end{definition}
+
 %
@@
 
-\begin{figure} 
-\center
-\begin{tikzpicture} 
-% Box/square (4x4)
-\draw[black,solid,ultra thin] (0,0)--(0,4);
-\draw[black,solid,ultra thin] (0,0)--(4,0);
-\draw[black,solid,ultra thin] (4,0)--(4,4);
-\draw[black,solid,ultra thin] (0,4)--(4,4);
-% Dots
-\draw[black,fill=black] (0,0) circle (.5ex);
-\draw[black,fill=black] (2,2) circle (.5ex);
-\draw[black,fill=black] (4,4) circle (.5ex);
-\draw[black,fill=black] (0,2) circle (.5ex);
-\draw[black,fill=black] (0,4) circle (.5ex);
-% Arrows
-\draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
-\draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
-\draw[black,thick,->] (0,0.2)--(0,1.8);
-\draw[black,thick,->] (0,2.2)--(0,3.8);
-\draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
-% Node (parameter) labels
-\draw[] (0,0) node[anchor=east] {$(0,0,0)$};
-\draw[] (0,4) node[anchor=east] {$(0,0,2)$};
-\draw[] (4,0) node[anchor=west] {$(2,2,0)$};
-\draw[] (4,4) node[anchor=west] {$(2,2,2)$};
-% Arrow labels
-\draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
-\draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
-\draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
-\draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
-% The key and headings
-\draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
-\draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
-\draw[] (-2.3,3.2) node[anchor=west] {$(a,b)$};
-\draw[] (-2.7,3.6) node[anchor=south] {$c$};
-\draw[] (2,4.6) node[anchor=south] {basis conversion};
-\end{tikzpicture} 
-\caption{The Laplace operator acting on vectors of $\smash{\hdopnk^{(0,0,0)}}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{\hdopnk^{(2,2,2)}}$ coefficients.
-Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{\hdopnkabc}$ coefficients, where $(a,b,c)$ is at the tail of the arrow, and the range is $\smash{\hdopnk^{(\tilde{a},\tilde{b},\tilde{c})}}$ coefficients, where $(\tilde{a},\tilde{b},\tilde{c})$ is at the head of the arrow.
-%For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{\hdop_{n,k}^{(0,0,0)}}$ (resp.~$\smash{\hdop_{n,k}^{(1,0,1)}}$) and the range in $\smash{\hdop_{n,k}^{(1,0,1)}}$ (resp.~$\smash{\hdop_{n,k}^{(2,0,2)}}$).
-}
-\label{fig:Laplace} 
-\end{figure}
-
 In this section, we will derive the entries that form spherical partial differential operators, demonstrating their sparsity.
 To this end, we will use a different ordering of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
-Let $N \in \N$ and define:
-\begin{align}
+Let $N \in \N$ and recall from equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) that:
+\begin{align*}
 	\bigscopNka &:= 
@@
 			\scopa_{N,k,1}(x,y,z)
-		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
+		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \\ 
 	\bigscopa_{N,0} &:= 
@@
 			\scopa_{N,0,0}(x,y,z) \\
-		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
+		\end{pmatrix} \in \R^{N+1}, \\
 	\bigscopNa &:= 
@@
 			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
-\end{align}
+		\end{pmatrix} \in \R^{(N+1)^2}
+\end{align*}
 
@@
 	{\partial f \over \partial \theta}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
-	\nabla^2_s \big( \genjacw^{(a,0)}(z) \: f(x,y,z) \big) &= \bigscopNa(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ for } a = 1 \text{ only}.
-\end{align*}
+	\nabla^2_s \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
+\end{align*}
+where $\nabla^2_s$ is the surface Laplacian, given by $\nabla^2_s = \frac{1}{\rho(z)} \ppphi \big( \rhozppphi \big) + \frac{1}{\rho(z)^2} \ppthetatwo$.
 \end{definition}
@@
 The operators we define here are for partial derivatives with respect to the spherical coordinates $(\phi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).
+With the OP ordering by Fourier mode $k$ defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) these rotationally invariant operators are block-diagonal, meaning simple and parallelisable practical application.
 
@@
 
-These operators are be rotationally invariant, and hence with the ordering we have for the OPs in $\bigscopNa$, obtaining and applying the operators can be done in parallel since each block need not talk to each other (they are block diagonal).
 In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short Lemma.
@@
 
-In this section we discuss how to take advantage of the proposed basis and sparsity structure in partial differential operators in practical computational applications.
+In this section we discuss how to expand and evaluate functions in our proposed basis, and take advantage of the sparsity structure in partial differential operators in practical computational applications.
 
@@
 \begin{align*}
-	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \mathbf{f}_k
-\end{align*}
-for N sufficiently large, where $\bigscopNka$ is defined in equation \label{eqn:OPdefNka} and where
-\begin{align*}
-	\mathbf{f}_k &:= 
+	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \bm{f}_k = \bigscopNa(x,y,z)^\top \: \bm{f}
+\end{align*}
+for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) and where
+
+\begin{align*}
+	\bm{f}_k &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2(N-k+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	\mathbf{f}_0 := 
+	\bm{f}_0 := 
 		\begin{pmatrix}
@@
 			f_{N,0,0}
-		\end{pmatrix} \in \R^{2(N+1)},
-\end{align*}
-\begin{align*}
-	f_{n,k,i} &:= \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
-\end{align*}
-Recall from (\ref{eqn:OPnorm}) that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$.
+		\end{pmatrix} \in \R^{N+1}, \\
+	\bm{f} &:= 
+		\begin{pmatrix}
+			\bm{f}_0 \\
+			\vdots \\
+			\bm{f}_N
+		\end{pmatrix} \in \R^{2(N+1)^2}, \quad \quad
+	f_{n,k,i} := \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
+\end{align*}
+Recall from equation (\ref{eqn:normscop}) that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$.
 Using the quadrature rule detailed in Section \ref{subsection:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$:
@@
 we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).
+
+\subsection{Function evaluation}\label{subsection:functionevaluation}
+
+For a function $f$, with coefficients vector $\bm{f}$ for expansion in the $\{\scopnki\}$ basis as determined via the method in Section \ref{subsection:expandingfunctions} up to order $N$, we can use the Clenshaw algorithm to evaluate the function at a point $(x,y,z) \in \Omega$ as follows.
+Let $A_n, B_n, \Dnt, C_n$ be the Clenshaw matrices from Definition \ref{def:clenshawmats}, and define the rearranged coefficients vector $\tilde{\bm{f}}$ via
+\begin{align*}
+	\tilde{\bm{f}}_n &:= 
+		\begin{pmatrix}
+			f_{n,0,0} \\
+			f_{n,1,0} \\
+			f_{n,1,1} \\
+			\vdots \\
+			f_{n,n,0} \\
+			f_{n,n,1}
+		\end{pmatrix} \in \R^{2(N+1)} \quad \text{for } n = 1,2,\dots,N, \quad
+	\tilde{\bm{f}}_0 = f_{0,0,0} \in \R, \\
+	\tilde{\bm{f}} &:= 
+		\begin{pmatrix}
+			\tilde{\bm{f}}_0 \\
+			\vdots \\
+			\tilde{\bm{f}}_N
+		\end{pmatrix} \in \R^{(N+1)^2}.
+\end{align*}
+The Clenshaw algorithm is then \cite{}\bstodo{Reference for Clenshaw in >1 dimensions???} 
+\begin{align*}
+	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
+	\quad &\text{2) } \text{For } n = N:-1:0 \\
+	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{f}}_n}^T - \bm{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+\end{align*}
 
@@
 
-Hey
+The Clenshaw algorithm outlined in Section \ref{subsection:functionevaluation} can also be used with Jacobi matrices $J^{(a)}_x, J^{(a)}_y, J^{(a)}_z$ replacing the point $(x,y,z)$.
+Let $v : \Omega \to \R$ be the function that we wish to obtain an operator matrix $V$ for, so that
+\begin{align*}
+	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \bm{f}^\top \: \bigscopNa(x,y,z) = (V \bm{f})^\top \: \bigscopNa(x,y,z),
+\end{align*}
+i.e. $V \bm{f}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$.
+
+To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in Section \ref{subsection:functionevaluation} so that $v(x,y,z) = \tilde{\bm{v}}^\top \: \bigscopta(x,y,z)$).
+The operator $V$ is then the result of the following:
+\begin{align*}
+	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
+	\quad &\text{2) } \text{For } n = N:-1:0 \\
+	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(J^{(a)}_x, J^{(a)}_y, J^{(a)}_z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad &\text{3) } \text{Output: } V(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+\end{align*}
+where at each iteration, $\bm{\xi}_n$ is a vector of matrices.
 

### 2020-08-07 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
Edits to code, mainly typos

@@ -415,7 +415,7 @@ function pointswithweights(S::SphericalCapSpace{<:Any, B, T, <:Any}, n;
     # Need to maunally call the method to get R coeffs here
     m = isodd(M1) ? Int((M1 + 1) / 2) : Int((M1 + 2) / 2); m -= Int(S.params[end])
     getreccoeffsR!(S, m; maxk=0)
-    t, wt = pointswithweights(B, getRspace(S, 0), M1) # Quad for w_R^{(a,2b+1)}
+    t, wt = pointswithweights(B, getRspace(S, 0), M1) # Quad for w_R^{(a,2b)}
     s = [(cospi(B(2 * (it - 1)) / M2), sinpi(B(2 * (it - 1)) / M2)) for it=1:M2]
     ws = ones(B, M2) / M2  # Quad for circumference of unit circle
     if !nofactor
@@ -992,10 +992,15 @@ function operatorclenshaw(cfs::AbstractVector{T}, S::SphericalCapSpace, M) where
     f = PseudoBlockArray(f, [2n+1 for n=0:N])
 
     @show "getting Jacobi matrices"
+    # TODO Cange the Jacobi operators and id to BandedBlockBandedMatrix objects
+    # (not Sparse matrices!)
     Jx = jacobix(S, M)
     Jy = jacobiy(S, M)
     Jz = jacobiz(S, M)
     id = sparse(I, size(Jx))
+    # id = BandedBlockBandedMatrix(I, size(Jx),
+    #                                 ([M+1; 2M:-2:1], [M+1; 2M:-2:1]),
+    #                                 (0,0), (0,0))
 
     P0 = getdegzeropteval(T, S)
     if N == 0
@@ -1300,7 +1305,7 @@ function rho2operator(S::SphericalCapSpace{<:Any, B, T, <:Any},
     resizedataonedimops!(S, N+band2)
 
     # Get the operator for mult by ρ²
-    ptsr, wr = pointswithweights(B, getRspace(S, 0), N+2)
+    ptsr, wr = pointswithweights(B, getRspace(S, 0), N+3)
     rhoptsr2 = S.family.ρ.(ptsr).^2
     getopnorms(St, N+band2+1)
     for k = 0:N
@@ -1385,10 +1390,10 @@ function getlaplacianval(S::SphericalCapSpace{<:Any, B, T, <:Any},
 end
 # The operator for Δ
 function laplacianoperator(S::SphericalCapSpace{<:Any, B, T, <:Any}, N::Int;
[…]

### 2020-08-07 OrthogonalPolynomialFamilies/src/SphericalCapsTangent.jl

@@ -38,7 +38,7 @@ spacescompatible(A::SphericalCapTangentSpace, B::SphericalCapTangentSpace) = (A.
 function gettangentspace(D::SphericalCapFamily{B,T,N,<:Any,<:Any}) where {B,T,N}
     length(D.tangentspace) == 1 && return D.tangentspace[1]
     resize!(D.tangentspace, 1)
-    params = (B(0.0), B(0.0))
+    params = (B(0), B(0))
     D.tangentspace[1] = SphericalCapTangentSpace(D, params)
 end
 
@@ -295,7 +295,7 @@ function rhodivminusgradrhodotoperator(S::SphericalCapTangentSpace{<:Any, B, T, 
     resizedataonedimops!(S1, N+1)
     resizedataonedimops!(S0, N+1)
 
-    ptsr, wr = pointswithweights(B, R1, N+3)
+    ptsr, wr = pointswithweights(B, R1, 2N+3)
     rhoptsr2 = S.family.ρ.(ptsr).^2
     rhodxrhoptsr = S.family.ρ.(ptsr) .* differentiate(S.family.ρ).(ptsr)
 
@@ -333,7 +333,7 @@ function coriolisoperator(S::SphericalCapTangentSpace{<:Any, B, T, <:Any}, N;
     # We output the operator that results in coefficients {u\^Φ, u\^Φ}, for
     # expansion in the 𝕋 basis
 
-    Ω = 1.0 # B(72921) / 1e9 # TODO make this a global definition somehow (maybe a
+    Ω = B(72921) / 1e9 # TODO make this a global definition somehow (maybe a
                        # member of the class struct)
 
     # resizedataonedimops! for the SCSpace
@@ -369,58 +369,6 @@ function coriolisoperator(S::SphericalCapTangentSpace{<:Any, B, T, <:Any}, N;
 end
 
 
-#===#
-# "Coeffs degree increaser" operator matrix
-
-function increasedegreeoperator(S::SphericalCapTangentSpace{<:Any, B, T, <:Any},
-                                N, Nto; weighted=false) where {B,T}
-    # This operator acts on the coeffs vector of a Fun in the space S to just
-    # reorganise the coeffs so that the length is increased from deg N to
-    # deg Nto, with all coeffs representing the extra degrees being zero.
-
-    # TODO weighted=true
-    C = BandedBlockBandedMatrix(Zeros{B}(2(Nto+1)^2, 2(N+1)^2),
[…]

### 2020-08-07 experiments/sphericalcaptests-plots.jl
Experiment julia files

@@ -1,0 +1,227 @@
+using Revise
+using ApproxFun
+using Test, Profile, StaticArrays, LinearAlgebra, BlockArrays,
+        BlockBandedMatrices, SparseArrays
+using OrthogonalPolynomialFamilies
+import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
+                    opevalatpts, inner2, getopindex, getnk, resizecoeffs!,
+                    transformparamsoperator, weightedpartialoperatorx,
+                    weightedpartialoperatory, partialoperatorx, partialoperatory,
+                    derivopevalatpts, getderivopptseval, getopnorms,
+                    getopnorm, operatorclenshaw, weight, biharmonicoperator,
+                    getPspace, getRspace, differentiatespacex, differentiatespacey,
+                    differentiateweightedspacex, differentiateweightedspacey,
+                    resizedata!,
+                    resizedataonedimops!, getnki, convertcoeffsvecorder,
+                    diffoperatorphi, diffoperatortheta, diffoperatortheta2,
+                    differentiatespacephi, increasedegreeoperator,
+                    getptsevalforop, getderivptsevalforop, laplacianoperator,
+                    rho2laplacianoperator, gettangentspace,
+                    rhogradoperator, coriolisoperator, rho2operator,
+                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator
+using JLD
+using Makie
+
+
+isindomain(pt, D::SphericalCapFamily) = D.α ≤ pt[3] ≤ D.β && norm(pt) == 1.0
+isindomain(pt, S::SphericalCapSpace) = isindomain(pt, S.family)
+rhoval(z) = sqrt(1 - z^2)
+
+
+# Setup
+T = Float64; B = T#BigFloat
+α = 0.2
+SCF = SphericalCapFamily(B, T, B(α * 1000) / 1000)
+a = 1.0
[…]

### 2020-08-07 experiments/sphericalcaptests-swe.jl

@@ -18,10 +18,11 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
                     getptsevalforop, getderivptsevalforop, laplacianoperator,
                     rho2laplacianoperator, gettangentspace,
                     rhogradoperator, coriolisoperator, rho2operator,
-                    rhodivminusgradrhodotoperator, gradrhooperator
+                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator
 using Makie
 using GeometryTypes
 # using PyPlot
+
 
 
 isindomain(pt, D::SphericalCapFamily) = D.α ≤ pt[3] ≤ D.β && norm(pt) == 1.0
@@ -40,6 +41,23 @@ function gettangentspacecoeffsvec(f::Fun, g::Fun)
     ret
 end
 
+function squareoperator(S::SphericalCapSpace, A::BandedBlockBandedMatrix, N::Int)
+    """ this is to square the given operator, by reassigning the non-zero
+        entries correctly for given N value
+    """
+    @assert (N < nblocks(A)[1] - 1 && N == nblocks(A)[2] - 1) "Invalid size of A or incorrect N val"
+    C = BandedBlockBandedMatrix(Zeros{B}((N+1)^2, (N+1)^2),
+                                ([N+1; 2N:-2:1], [N+1; 2N:-2:1]),
+                                (A.l, A.u), (A.λ, A.μ))
+    for k = 1:N
+        inds = 1:2(N-k+1)
+        view(C, Block(k+1, k+1)) .= view(A, Block(k+1, k+1))[inds, inds]
+    end
+    k = 0
+    view(C, Block(k+1, k+1)) .= view(A, Block(k+1, k+1))[1:N+1, 1:N+1]
+    C
+end
+
 
 #======#
 # SWE
@@ -51,31 +69,70 @@ SCF = SphericalCapFamily(B, T, B(α * 1000) / 1000)
 S = SCF(1.0, 0.0); S0 = SCF(0.0, 0.0)
 ST = gettangentspace(SCF)
 
[…]

### 2020-08-07 experiments/sphericalcaptests-tangentspace.jl

@@ -1,0 +1,211 @@
+using Revise
+using ApproxFun
+using Test, Profile, StaticArrays, LinearAlgebra, BlockArrays,
+        BlockBandedMatrices, SparseArrays
+using OrthogonalPolynomialFamilies
+import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
+                    opevalatpts, inner2, getopindex, getnk, resizecoeffs!,
+                    transformparamsoperator, weightedpartialoperatorx,
+                    weightedpartialoperatory, partialoperatorx, partialoperatory,
+                    derivopevalatpts, getderivopptseval, getopnorms,
+                    getopnorm, operatorclenshaw, weight, biharmonicoperator,
+                    getPspace, getRspace, differentiatespacex, differentiatespacey,
+                    differentiateweightedspacex, differentiateweightedspacey,
+                    resizedata!,
+                    resizedataonedimops!, getnki, convertcoeffsvecorder,
+                    diffoperatorphi, diffoperatortheta, diffoperatortheta2,
+                    differentiatespacephi, increasedegreeoperator,
+                    getptsevalforop, getderivptsevalforop, laplacianoperator,
+                    rho2laplacianoperator, gettangentspace,
+                    rhogradoperator, coriolisoperator, rho2operator,
+                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator
+using JLD
+
+
+isindomain(pt, D::SphericalCapFamily) = D.α ≤ pt[3] ≤ D.β && norm(pt) == 1.0
+isindomain(pt, S::SphericalCapSpace) = isindomain(pt, S.family)
+rhoval(z) = sqrt(1 - z^2)
+
+
+# Setup
[…]

### 2020-08-07 experiments/sphericalcaptests.jl

@@ -18,7 +18,7 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
                     getptsevalforop, getderivptsevalforop, laplacianoperator,
                     rho2laplacianoperator, gettangentspace,
                     rhogradoperator, coriolisoperator, rho2operator,
-                    rhodivminusgradrhodotoperator
+                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator
 using JLD
 
 
@@ -67,14 +67,50 @@ function getscspacecoeffsvecs(ST::SphericalCapTangentSpace, Fc::AbstractVector{T
 end
 getscspacecoeffsvecs(F::Fun) = getscspacecoeffsvecs(F.space, F.coefficients)
 
+# Useful functions for testing
+function converttopseudo(S::SphericalCapSpace, cfs; converttobydegree=true)
+    N = getnki(S, length(cfs))[1]
+    @assert (length(cfs) == (N+1)^2) "Invalid coeffs length"
+    if converttobydegree
+        PseudoBlockArray(convertcoeffsvecorder(S, cfs), [2n+1 for n=0:N])
+    else
+        PseudoBlockArray(cfs, [2n+1 for n=0:N])
+    end
+end
+
+# Solve `A x = b` for `x` using iterative improvement
+# (for BigFloat sparse matrix and vector)
+function iterimprove(A::SparseMatrixCSC{T}, b::Vector{T};
+                        iters=5, verbose=true) where T
+    if eps(T) ≥ eps(Float64)
+        # throw(ArgumentError("wrong implementation"))
+        return A \ b
+    end
+    A0 = SparseMatrixCSC{Float64}(A)
+    F = factorize(A0)
+    x = zeros(T, size(A)[2]) # x = zeros(T, length(b))
+    r = copy(b)
+    for iter = 1:iters
[…]

### 2020-08-19 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
Rework of clenshaw and storing of coeffs

@@ -49,6 +49,10 @@ struct SphericalCapSpace{DF, B, T, N} <: Space{SphericalCap{B,T}, T}
     opnorms::Vector{B} # squared norms
     opptseval::Vector{Vector{B}} # Store the ops evaluated at the transform pts
 
+    # Stores the α recurrence coeffs (for mult by x). No need to store β's, γ's
+    # here too.
+    reccoeffsa::Dict{NTuple{2,Int}, Vector{B}}
+
     # These store the "Clenshaw" matrices. Each subvector contains the x,y,z
     # submatrices
     A::Vector{Vector{BandedBlockBandedMatrix{B}}}
@@ -64,7 +68,7 @@ end
 
 function SphericalCapSpace(fam::SphericalFamily{B,T,N}, params::NTuple{N,B}) where {B,T,N}
     SphericalCapSpace{typeof(fam), B, T, N}(
-        fam, params, Vector{B}(), Vector{Vector{B}}(),
+        fam, params, Vector{B}(), Vector{Vector{B}}(), Dict{NTuple{2,Int}, Vector{B}}(),
         Vector{Vector{BandedBlockBandedMatrix{B}}}(), Vector{Vector{BandedBlockBandedMatrix{B}}}(),
         Vector{Vector{BandedBlockBandedMatrix{B}}}(), Vector{Vector{BandedBlockBandedMatrix{B}}}())
 end
@@ -260,53 +264,71 @@ end
 # Recurrence coeffs retrieval methods
 
 # α are the mult by x coeffs, β the mult by y coeffs, γ the mult by z coeffs
-function recα(::Type{T}, S::SphericalCapSpace, n, k, j) where T
+function recα(::Type{T}, S::SphericalCapSpace, n::Int, k::Int, j::Int) where T
+
+    # Check if we already have this stored
+    haskey(S.reccoeffsa, (n, k)) && return S.reccoeffsa[(n, k)][j]
+
+    # Else, we calculate each (valid) coeff j=1:6 for the (n,k) pair
+    resizedataonedimops!(S, n+3)
+    S.reccoeffsa[(n, k)] = zeros(T, 6)
+
     getopnorms(S, k+1)
     R1 = getRspace(S, k-1)
     R2 = getRspace(S, k)
     R3 = getRspace(S, k+1)
 
-    # TODO check that the number of pts we are getting here is what is required!
-    if isodd(j)
-        pts, w = pointswithweights(T, R2, Int(ceil(n-k+1.5)))
-    else
[…]

### 2020-08-19 experiments/sphericalcaptests-plots.jl

@@ -18,7 +18,8 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
                     getptsevalforop, getderivptsevalforop, laplacianoperator,
                     rho2laplacianoperator, gettangentspace,
                     rhogradoperator, coriolisoperator, rho2operator,
-                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator
+                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator,
+                    squareoperator
 using JLD
 using Makie
 
@@ -38,8 +39,7 @@ ST = gettangentspace(SCF)
 
 y, z = B(-234)/1000, B(643)/1000; x = sqrt(1 - z^2 - y^2); p = [x; y; z]; isindomain(p, SCF)
 θ = atan(y / x)
-resizedata!(S, 20)
-resizedata!(S0, 20)
+
 
 function gettangentspacecoeffsvec(S::SphericalCapSpace, fc::AbstractVector{T},
                                     gc::AbstractVector{T}) where T
@@ -103,22 +103,7 @@ function iterimprove(A::SparseMatrixCSC{T}, b::Vector{T};
     x
 end
 
-function squareoperator(S::SphericalCapSpace, A::BandedBlockBandedMatrix, N::Int)
-    """ this is to square the given operator, by reassigning the non-zero
-        entries correctly for given N value
-    """
-    @assert (N < nblocks(A)[1] - 1 && N == nblocks(A)[2] - 1) "Invalid size of A or incorrect N val"
-    C = BandedBlockBandedMatrix(Zeros{B}((N+1)^2, (N+1)^2),
-                                ([N+1; 2N:-2:1], [N+1; 2N:-2:1]),
-                                (A.l, A.u), (A.λ, A.μ))
-    for k = 1:N
-        inds = 1:2(N-k+1)
-        view(C, Block(k+1, k+1)) .= view(A, Block(k+1, k+1))[inds, inds]
-    end
-    k = 0
[…]

### 2020-08-19 experiments/sphericalcaptests.jl

@@ -259,6 +259,7 @@ f = (x,y,z)->cos(y)
 pts, w = pointswithweights(S, n)
 vals = [f(pt...) for pt in pts]
 cfs = transform(S, vals)
+S
 N = getnki(S, length(cfs))[1]
 cfs2 = PseudoBlockArray(convertcoeffsvecorder(S, cfs), [2n+1 for n=0:N])
 F = Fun(S, cfs)
@@ -271,6 +272,7 @@ vals2 = itransform(S, cfs); pts2 = points(S, length(vals2)/2)
 F = Fun(f, S, 500); F.coefficients
 F(p) - f(p...)
 @test F(p) ≈ f(p...)
+S
 
 # ∂/∂θ operator
 f = (x,y,z)->x^2 + y^4 # (cos2t + sin4t*ρ2(z))ρ2(z)
@@ -352,8 +354,8 @@ N = 10
 Δ = laplacianoperator(S, N)
 c = 2.0
 u = (x,y,z)->c
-rho2f = (x,y,z)->-c * 2 * z * (1-z^2) # 2 * ρ * ρ' * ρ^2
-F = Fun(rho2f, S, 2*(N+4)^2); F.coefficients
+rho2f = (x,y,z)->-c * 2 * z #* (1-z^2) # 2 * ρ * ρ' * ρ^2
+F = Fun(rho2f, S, 2*(N+2)^2); F.coefficients
 ucfs = iterimprove(sparse(Δ), F.coefficients)
 U = Fun(S, ucfs)
 @test U(p) ≈ u(p...)
@@ -361,13 +363,14 @@ U = Fun(S, ucfs)
 # 2) u = monomial
 inds = [2, 3, 3]; N = 50
 u = (x,y,z)->x^inds[1] * y^inds[2] * z^inds[3]
-Δ = laplaceoperator(S, S, N)
-rho2f = (x,y,z)->rholaplacian(S, u, inds, [x;y;z]) # NOTE methods at bottom of script for this
-F = Fun(rho2f, S, 2*(sum(inds)+4)^2); resizecoeffs!(S, F, N+3)
-F(p) - rho2f(p...)
+Δ = laplacianoperator(S, N)
+fff = (x,y,z)->laplaciantest(S, u, inds, [x;y;z]) # NOTE methods at bottom of script for this
+F = Fun(fff, S, 2*(sum(inds)+2)^2); resizecoeffs!(S, F, N+1)
[…]

### 2020-08-19 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
Updated Julia, Atom etc and so needed to update these files. Includes some minor fixes too.

@@ -264,70 +264,76 @@ end
 # Recurrence coeffs retrieval methods
 
 # α are the mult by x coeffs, β the mult by y coeffs, γ the mult by z coeffs
-function recα(::Type{T}, S::SphericalCapSpace, n::Int, k::Int, j::Int) where T
 
-    # Check if we already have this stored
-    haskey(S.reccoeffsa, (n, k)) && return S.reccoeffsa[(n, k)][j]
+function getrecα(::Type{T}, S::SphericalCapSpace, n::Int, k::Int, jj::Int) where T
+    # This function returns the jjth α rec coeff for mult by x for the (n,k) pair
 
-    # Else, we calculate each (valid) coeff j=1:6 for the (n,k) pair
     resizedataonedimops!(S, n+3)
-    S.reccoeffsa[(n, k)] = zeros(T, 6)
 
     getopnorms(S, k+1)
     R1 = getRspace(S, k-1)
     R2 = getRspace(S, k)
     R3 = getRspace(S, k+1)
 
-    for jj = 1:6
+    failval = T(0)
+    if isodd(jj)
+        k == 0 && return failval
+        pts, w = pointswithweights(T, R2, Int(ceil(n-k+1.5)))
+    else
+        (n == k && jj != 6) && return failval
+        (n == 0 && jj == 2) && return failval
+        (n == k + 1 && jj == 2) && return failval
+        pts, w = pointswithweights(T, R3, Int(ceil(n-k+1.5)))
+    end
+    getopptseval(R2, n-k+1, pts)
 
-        if isodd(jj)
-            k == 0 && continue
-            pts, w = pointswithweights(T, R2, Int(ceil(n-k+1.5)))
-        else
-            (n == k && jj != 6) && continue
-            (n == 0 && jj == 2) && continue
-            (n == k + 1 && jj == 2) && continue
-            pts, w = pointswithweights(T, R3, Int(ceil(n-k+1.5)))
-        end
-        getopptseval(R2, n-k+1, pts)
-
-        if jj == 1
-            getopptseval(R1, n-k+1, pts)
-            ret = (inner2(R2, opevalatpts(R2, n-k+1, pts), opevalatpts(R1, n-k+1, pts), w)
-                    / getopnorm(R1))
-        elseif jj == 2
-            getopptseval(R3, n-k-1, pts)
[…]

### 2020-08-19 OrthogonalPolynomialFamilies/src/SphericalCapsTangent.jl

@@ -303,7 +303,7 @@ function rhodivminusgradrhodotoperator(S::SphericalCapTangentSpace{<:Any, B, T, 
     band2 = 1
     # TODO sort out bandwidths!
     A = BandedBlockBandedMatrix(Zeros{B}((N+band2+1)^2, 2 * (N+1)^2),
-                                ([N+band2+1; 2(N+band2):-2:1], 2 * [N+1; 2N:-2:1]),
+                                [N+band2+1; 2(N+band2):-2:1], 2 * [N+1; 2N:-2:1],
                                 (0, 0), (2band2, 2N+band1+1))
     for k = 0:N
         if k % 100 == 0
@@ -343,7 +343,7 @@ function coriolisoperator(S::SphericalCapTangentSpace{<:Any, B, T, <:Any}, N;
     band1 = 1
     band2 = 1
     A = BandedBlockBandedMatrix(Zeros{B}(2 * (N+band2+1)^2, 2 * (N+1)^2),
-                                (2 * [N+band2+1; 2(N+band2):-2:1], 2 * [N+1; 2N:-2:1]),
+                                2 * [N+band2+1; 2(N+band2):-2:1], 2 * [N+1; 2N:-2:1],
                                 (0, 0), (4band2+1, 4band1+1))
     for k = 0:N
         if k % 100 == 0
@@ -361,7 +361,7 @@ function coriolisoperator(S::SphericalCapTangentSpace{<:Any, B, T, <:Any}, N;
     A
     if square
         F = BandedBlockBandedMatrix(A[1:getopindex(S, N, N, 1, 1), 1:getopindex(S, N, N, 1, 1)],
-                                      (2 * [N+1; 2N:-2:1], 2 * [N+1; 2N:-2:1]),
+                                      2 * [N+1; 2N:-2:1], 2 * [N+1; 2N:-2:1],
                                       (0, 0), (4band2+1, 4band1+1))
     else
         A
@@ -388,7 +388,7 @@ function transformparamsoperator(S::SphericalCapTangentSpace{<:Any, B, <:Any, <:
     #                             (0, 0), (2band2, 2band1))
     band1 = 2; band2 = 0
     C = BandedBlockBandedMatrix(Zeros{B}(2(N+1)^2, 2(N+1)^2),
-                                (2 * [N+1; 2N:-2:1], 2 * [N+1; 2N:-2:1]),
[…]

### 2020-08-19 experiments/sphericalcaptests-plots.jl
Plotting experiments updated

@@ -104,7 +104,6 @@ function iterimprove(A::SparseMatrixCSC{T}, b::Vector{T};
 end
 
 
-
 #===#
 # Plotting functions
 function plot_on_domain_scalar!(scene, S, u, ϕ, θ; weighted=true)
@@ -129,10 +128,13 @@ Sparsity of Operators
 =#
 using Plots
 function getspymat(A)
-    B = sparse(A)
+    B = sparse(zeros(size(A)))
     m1, m2 = size(A)
     for i = 1:m1, j = 1:m2
-        B[i, j] = A[m1-i+1, m2-j+1]
+        aa = A[m1-i+1, m2-j+1]
+        if abs(aa) > 1e-17
+            B[i, j] = aa
+        end
     end
     B
 end
@@ -141,21 +143,13 @@ N = 30
 
 # Laplacian
 Δw = squareoperator(S, laplacianoperator(S, N), N)
-    Plots.spy(sparse(getspymat(Δw)), marker = (:square, 1), legend=nothing)
-    Plots.savefig("experiments/saved/sphericalcap/images/sparsity-of-laplacian.png")
+Plots.spy(sparse(getspymat(Δw)), marker = (:square, 1), legend=nothing)
+Plots.savefig("experiments/saved/sphericalcap/images/sparsity-of-laplacian.png")
 
 # Biharmonic
-Lw = squareoperator(S, rho2laplacianoperator(S2, S0, N; weighted=true), N)
-    L = squareoperator(S, rho2laplacianoperator(S0, S2, N; weighted=false), N)
-    Jz = OrthogonalPolynomialFamilies.jacobiz(S2, N)
-    t = transformparamsoperator(S, S2, N; weighted=false)
-    Dϕ = squareoperator(S, diffoperatorphi(S0, N), N)
-    t2 = transformparamsoperator(S0, S2, N; weighted=false)
-    v = Fun((x,y,z)->2z^2 - 2S.family.ρ(z)^2 + 3z, S2, 20)
-    V = operatorclenshaw(v.coefficients, S2, N)
-    Bw = sparse(L * Lw - Jz * t * Dϕ * Lw) - V * sparse(t2 * Lw)
+Bw = biharmonicoperator(S2, N; square=true)
[…]

### 2020-08-19 papers/spherical-caps/spherical-caps.tex
Further edits to paper, including new figs (not pushed)

@@
 \usepackage{todonotes}
+\usepackage{csquotes}
 
@@
 \begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| = \gamma \rho(z)\}
+	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| := \sqrt{x^2 + y^2} = \gamma \rho(z)\}
 \end{align*}
@@
 \end{align*} \bstodo{Ordering correct?}
-where $C_{n,k,i} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{j - \tilde j_i} \: y^{\tilde j_i} \: z^{n-j}$ where $j < k$.
+where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - \tilde j_i} \: y^{\tilde j_i} \: z^{n-j}$ where $j < k$.
 The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
@@
 	x &= \sin(\phi) \: \cos(\theta) = \rho(z) \: \cos(\theta) \\
-	y &= \sin(\phi) \: \sin(\theta) = \rho(z) \: \sin(\theta)
+	y &= \sin(\phi) \: \sin(\theta) = \rho(z) \: \sin(\theta).
 \end{align*}
@@
 \begin{align}
-	\gamma_{k} :=
+	\eta_{k} :=
 		\begin{cases}
@@
 \begin{align*}
-	\alphaa_{n,k,1} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
-	\alphaa_{n,k,2} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
-	\alphaa_{n,k,3} &:=\gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
-	\alphaa_{n,k,4} &:= \gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
-	\alphaa_{n,k,5} &:= \gamma_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
-	\alphaa_{n,k,6} &:=\gamma_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
+	\alphaa_{n,k,1} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
+	\alphaa_{n,k,2} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
+	\alphaa_{n,k,3} &:=\eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
+	\alphaa_{n,k,4} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
+	\alphaa_{n,k,5} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
+	\alphaa_{n,k,6} &:=\eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
 	\betaa_{n,k,i,j} &:= 
@@
 For the recurrence for multiplication by $x$, since $\{\scopmjha\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n+1$ polynomial on $\Omega$, we can expand
-$$
+\begin{align*}
 	x \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j} \: \scopmjha(x,y,z).
-$$
+\end{align*}
 These coefficients are given by
@@
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{i,h} \: (\gamma_{k-1} \: \delta_{k-1, j} + \gamma_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
+	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
 \end{align*}
@@
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \gamma_{k-1} \: \delta_{k-1, j} + (-1)^i \: \gamma_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
+	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
 \end{align*}
@@
 
-For example, $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
+For example, $J_x^{(a)}, J_y^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$) while $J_z^{(a)}$ is block-diagonal (block-bandwidths $(0,0)$):
 \begin{align*}
@@
 
-For $J_x^{(a)}$, the sub-blocks $A^{(a)}_{x, k}, C^{(a)}_{x, k}$ have sub-block-bandwidths $(2,2)$ while the sub-blocks $B^{(a)}_{x,k}$ are zero matrices: \bstodo{fix bandwidths}
+For $J_x^{(a)}$, the sub-blocks $A^{(a)}_{x, k}, C^{(a)}_{x, k}$ have sub-block-bandwidths $(4,4)$ while the sub-blocks $B^{(a)}_{x,k}$ are zero matrices:
 \begin{align*}
@@
 where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align*}
+\begin{align}
 	A^{(a)}_{n,k,j} &:= 
@@
 			0 & \alphaa_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
+		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \label{eqn:jacobisubblocksA1} \\
 	A^{(a)}_{n,0,j} &:=
@@
 			\alphaa_{n,0,j} & 0
-		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \\
+		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \label{eqn:jacobisubblocksA2} \\
 	A^{(a)}_{n,1,j} &:=
@@
 			0
-		\end{pmatrix} \in \R^{2\times1}, j \text{ odd} \\
-\end{align*}
-
-For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$ while the sub-blocks $B^{(a)}_{y,k}$ are zero matrices:
+		\end{pmatrix} \in \R^{2\times1}, j \text{ odd} \label{eqn:jacobisubblocksA3} \\
+\end{align}
+
+For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(5,5)$ while the sub-blocks $B^{(a)}_{y,k}$ are zero matrices:
 \begin{align*}
@@
 where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align*}
+\begin{align}
 	B^{(a)}_{n,k,j} &:= 
@@
 			\betaa_{n,k,1,j} & 0
-		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
+		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \label{eqn:jacobisubblocksB1} \\
 	B^{(a)}_{n,0,j} &:=
@@
 			0 & \betaa_{n,0,0,j}
-		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \\
+		\end{pmatrix} \in \R^{1\times2}, j \text{ odd}  \label{eqn:jacobisubblocksB2} \\
 	B^{(a)}_{n,1,j} &:=
@@
 			\betaa_{n,1,1,j}
-		\end{pmatrix}\in \R^{2\times1}, j \text{ odd} \\
-\end{align*}
-
-For $J_z^{(a)}$, the sub-blocks $B^{(a)}_{z,k}$ have sub-block-bandwidths $(3,3)$, while the sub-blocks $A^{(a)}_{z,k}, \: C^{(a)}_{z,k}$ are zero matrices:
+		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}  \label{eqn:jacobisubblocksB3} \\
+\end{align}
+
+For $J_z^{(a)}$, the sub-blocks $B^{(a)}_{z,k}$ have sub-block-bandwidths $(2,2)$, while the sub-blocks $A^{(a)}_{z,k}, \: C^{(a)}_{z,k}$ are zero matrices:
 \begin{align*}
@@
 where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align*}
+\begin{align}
 	\Gamma^{(a)}_{n,k,j} &:= 
@@
 			0 & \gamma_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, \\
-	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}.
-\end{align*}
+		\end{pmatrix} \in \R^{2\times2}, \label{eqn:jacobisubblocksGamma1} \\
+	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:jacobisubblocksGamma2}
+\end{align}
 
@@
 
-In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated 'Clenshaw matrices'.
-While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the 'Clenshaw matrices' necessary for efficient function evaluation using the Clenshaw algorithm.
+In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices}.
+While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
@@
 		\begin{pmatrix}
-			0 & \tilde A^{(a)}_{n,0,6} & 0 & & \\
-			\tilde A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
+			0 & A^{(a)}_{n,0,6} & 0 & & \\
+			A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & \tilde A^{(a)}_{n,n,5} & 0 & \tilde A^{(a)}_{n,n,6} \\
+			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
@@
 		\begin{pmatrix}
-			0 & \tilde A^{(a)}_{n,0,4} & & \\
-			\tilde A^{(a)}_{n,1,3} & \ddots & \ddots & \\
+			0 & A^{(a)}_{n,0,4} & & \\
+			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
 			& \ddots & \ddots & \tilde A^{(a)}_{n,n-1,4} \\
-			& & \tilde A^{(a)}_{n,n,3} & 0
+			& & A^{(a)}_{n,n,3} & 0
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
@@
 		\begin{pmatrix}
-			0 & \tilde A^{(a)}_{n,0,2} & & \\
-			\tilde A^{(a)}_{n,1,1} & \ddots & \ddots & \\
+			0 & A^{(a)}_{n,0,2} & & \\
+			A^{(a)}_{n,1,1} & \ddots & \ddots & \\
 			& \ddots & \ddots &A^{(a)}_{n,n-2,2} \\
 			& & \ddots & 0 \\
-			& & & \tilde A^{(a)}_{n,n,1} \\
+			& & & A^{(a)}_{n,n,1} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
 \end{align*}
-where
-\begin{align*}
-	\tilde A^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			\alphaa_{n,k,j} & 0 \\
-			0 & \alphaa_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
-	\tilde A^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			\alphaa_{n,0,j} & 0
-		\end{pmatrix} \in \R^{1\times2}, \quad j \text{ even} \\
-	\tilde A^{(a)}_{n,1,j} &:=
-		\begin{pmatrix}
-			\alphaa_{n,1,j} \\
-			0
-		\end{pmatrix} \in \R^{2\times1}, \quad j \text{ odd} \\
-\end{align*}
+where $A^{(a)}_{n,k,j}$ are defined in equations (\ref{eqn:jacobisubblocksA1}, \ref{eqn:jacobisubblocksA2}, \ref{eqn:jacobisubblocksA3}).
 For $\tilde J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
@@
 		\begin{pmatrix}
-			0 & \tilde B^{(a)}_{n,0,6} & 0 & & \\
-			\tilde B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
+			0 & B^{(a)}_{n,0,6} & 0 & & \\
+			B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & \tilde B^{(a)}_{n,n,5} & 0 & \tilde B^{(a)}_{n,n,6} \\
+			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
@@
 		\begin{pmatrix}
-			0 & \tilde B^{(a)}_{n,0,4} & & \\
-			\tilde B^{(a)}_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & \tilde B^{(a)}_{n,n-1,4} \\
-			& & \tilde B^{(a)}_{n,n,3} & 0
+			0 & B^{(a)}_{n,0,4} & & \\
+			B^{(a)}_{n,1,3} & \ddots & \ddots & \\
+			& \ddots & \ddots & B^{(a)}_{n,n-1,4} \\
+			& & B^{(a)}_{n,n,3} & 0
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
@@
 		\begin{pmatrix}
-			0 & \tilde B^{(a)}_{n,0,2} & & \\
-			\tilde B^{(a)}_{n,1,1} & \ddots & \ddots & \\
-			& \ddots & \ddots & \tilde B^{(a)}_{n,n-2,2} \\
+			0 & B^{(a)}_{n,0,2} & & \\
+			B^{(a)}_{n,1,1} & \ddots & \ddots & \\
+			& \ddots & \ddots & B^{(a)}_{n,n-2,2} \\
 			& & \ddots & 0 \\
-			& & & \tilde B^{(a)}_{n,n,1} \\
+			& & & B^{(a)}_{n,n,1} \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
 \end{align*}
-where
-\begin{align*}
-	\tilde B^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			0 & \betaa_{n,k,0,j} \\
-			\betaa_{n,k,1,j} & 0
-		\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
-	\tilde B^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			0 & \betaa_{n,0,0,j}
-		\end{pmatrix} \in \R^{1\times2}, \quad j \text{ even} \\
-	\tilde B^{(a)}_{n,1,j} &:=
-		\begin{pmatrix}
-			0 \\
-			\betaa_{n,1,1,j}
-		\end{pmatrix} \in \R^{2\times1}, \quad j \text{ odd} \\
-\end{align*}
-For $\tilde J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
+where $B^{(a)}_{n,k,j}$ are defined equations (\ref{eqn:jacobisubblocksB1}, \ref{eqn:jacobisubblocksB2}, \ref{eqn:jacobisubblocksB3}).
+For $\tilde J_z^{(a)}$, the sub-blocks are diagonal, i.e.
+have sub-block-bandwidths $(0,0)$:
 \begin{align*}
@@
 		\begin{pmatrix}
-			\tilde \Gamma^{(a)}_{n,0,3} & 0 & \\
+			\Gamma^{(a)}_{n,0,3} & 0 & \\
 			0 & \ddots & \ddots & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & 0 & \tilde \Gamma^{(a)}_{n,0,3} & 0 \\
+			& & 0 & \Gamma^{(a)}_{n,n,3} & 0 \\
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
@@
 		\begin{pmatrix}
-			\tilde \Gamma^{(a)}_{n,0,2} & \\
+			\Gamma^{(a)}_{n,0,2} & \\
 			& \ddots & & \\
 			& & \ddots & \\
-			& & & \tilde \Gamma^{(a)}_{n,0,2}
+			& & & \Gamma^{(a)}_{n,n,2}
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
@@
 		\begin{pmatrix}
-			\tilde \Gamma^{(a)}_{n,0,1} & 0 & & \\
+			\Gamma^{(a)}_{n,0,1} & 0 & & \\
 			0 & \ddots & \ddots & \\
 			& \ddots & \ddots & 0 \\
-			& & \ddots & \tilde \Gamma^{(a)}_{n,n,1} \\
+			& & \ddots & \Gamma^{(a)}_{n,n-1,1} \\
 			& & & 0 \\
@@
 \end{align*}
-where
-\begin{align*}
-	\tilde \Gamma^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			\gamma_{n,k,j} & 0 \\
-			0 & \gamma_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, \quad k = 1,\dots,n \\
-	\tilde \Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}
-\end{align*}
+where $\Gamma^{(a)}_{n,k,j}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}).
 
@@
 			& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
-			& & & & & \bm{\eta}^\top_{1} & & & \\
-			& & & & & \bm{\eta}^\top_{0} & & &
+			& & & & & \bm{\eta}^\top_{0} & & & \\
+			& & & & & \bm{\eta}^\top_{1} & & &
 		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
 \end{align*}
-for $n = 2, 3, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
+for $n = 1, 2, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
 \begin{align*}
@@
 		\begin{cases}
-			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
-			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2 \\
+			\frac{1}{\betaa_{n,n,1,6}} & j = 2(2n+1) \\
+			\frac{- \: \betaa_{n,n,1,5}}{\betaa_{n,n,1,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 3 \\
 			0 & o/w
@@
 		\begin{cases}
-			\big(\bm{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
+			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
+			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2  \text{ and } n > 1 \\
 			0 & o/w
@@
 \end{align*}
-For $n=0, 1$, we can simply take
+For $n=0$, we can simply take
 \begin{align*}
@@
 			0 & \frac{1}{\betaa_{0,0,6}} & 0
-		\end{pmatrix} \in \R^{3\times3}, \\
-	D^\top_1 &:= 
-		\begin{pmatrix}
-			0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,0,3} & & \\
-			0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3} \\
-			0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3}  \\
-			0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & \eta & 0 & 0 \\
-			0 & 0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & 0 & 0
-		\end{pmatrix} \in \R^{5\times9},
-\end{align*}
-where $\eta = \frac{- \: \alphaa_{1,1,5}}{\alphaa_{1,1,6} \: \gammaa_{1, 0, 3}}$.
+		\end{pmatrix} \in \R^{3\times3}.
+\end{align*}
 
@@
 			\vdots \\
-			\scopa_{N,0,0}(x,y,z) \\
+			\scopa_{N,0,0}(x,y,z)
 		\end{pmatrix} \in \R^{N+1}, \\
@@
 			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2}
+		\end{pmatrix} \in \R^{(N+1)^2}.
 \end{align*}
@@
 \begin{figure}[t]
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
+	\centering % <-- added
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[scale=0.3]{sparsity-of-laplacian}
 		\centering
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
+		\caption{The Laplace operator $\Delta^{(1)}_W$}
+	\end{subfigure}\hfil % <-- added
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[scale=0.3]{sparsity-of-rho2laplacian}
 		\centering
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
+		\caption{The $\rho^2$-factored Laplacian $L_W^{(1)}$}
+	\end{subfigure}\hfil % <-- added
+
+	\medskip
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[scale=0.3]{sparsity-of-helmholtz}
 		\centering
-	\end{subfigure}
-    	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
-Left: the Laplace operator $\laplacewiii$.
-Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
-Right: the biharmonic operator $\biharmonictwo$.}
-        \label{fig:sparsity}
+		\caption{A variable coefficient Helmholtz operator}
+	\end{subfigure}\hfil % <-- added
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[scale=0.3]{sparsity-of-biharmonic}
 		\centering
+		\caption{The Biharmonic operator $B_W^{(2)}$}
+	\end{subfigure}\hfil % <-- added
+	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
+For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 200$.}
+	\label{fig:images}
 \end{figure}
@@
 \end{align*}
-Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}. \bstodo{Add this fig}
 
@@
 The Clenshaw algorithm outlined in Section \ref{subsection:functionevaluation} can also be used with Jacobi matrices $J^{(a)}_x, J^{(a)}_y, J^{(a)}_z$ replacing the point $(x,y,z)$.
-Let $v : \Omega \to \R$ be the function that we wish to obtain an operator matrix $V$ for, so that
+Let $v : \Omega \to \R$ be the function that we wish to obtain an operator matrix $V$ for $v$, so that
 \begin{align*}
@@
 \begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
+	\begin{subfigure}{0.5\textwidth}
+	\includegraphics[scale=0.17]{poisson-u=wyexpx-N=60}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
+	\includegraphics[scale=0.3]{solutionblocknorms-poisson-N=100}
         	%\label{fig:solutionblocknorms-poisson}
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions.
-This demonstrates algebraic convergence with the rate dictated by the decay at the corners, with spectral convergence observed when the right-hand side vanishes to all orders.}
+	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
+Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions (here, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
+This demonstrates spectral convergence observed.}
 	\centering
@@
 \end{figure}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
-	\centering
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
-	\centering
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
-	\centering
-	\end{subfigure}
-	\centering
-	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$.
+%
+%\begin{figure}[t]
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
+%	\centering
+%	\end{subfigure}
+%	\begin{subfigure}{0.3\textwidth}
+%	\centering
+%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
+%	\centering
+%	\end{subfigure}
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
+%	\centering
+%	\end{subfigure}
+%	\centering
+%	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$.
 Left: Computed.
@@
 Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
+%	\centering
+%	\label{fig:poissonexact}
+%\end{figure}
+
+\begin{figure}[t]
+	\begin{subfigure}{0.5\textwidth}
+	\includegraphics[scale=0.17]{helmholtz-f=wyexpx-N=60-n=300}
 	\centering
-	\label{fig:poissonexact}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
-	%\label{fig:solution-helmholtz}
+	%\label{fig:solution-poisson}
 	\end{subfigure}
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
 	\centering
-        	%\label{fig:solutionblocknorms-helmholtz}
+	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-N=100}
+        	%\label{fig:solutionblocknorms-poisson}
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = x(1-x^2-y^2)e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$.}
+	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
+Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and the same function $v$ (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
+This demonstrates spectral convergence observed.}
 	\centering
@@
 \begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
+	\begin{subfigure}{0.5\textwidth}
+	\includegraphics[scale=0.17]{biharmonic-f=erf-N=80-n=50}
 	\centering
-	\includegraphics[scale=0.3]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
 	%\label{fig:solution-biharmonic}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic-diskslice-alpha=0p2-beta=0p8-N=200}
+	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-N=100}
 	\centering
@@
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions.}
+	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
+Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
+This demonstrates spectral convergence observed.}
 	\centering
@@
 
-The Poisson equation is the classic problem of finding \(u(x,y)\) given a function \(f(x,y)\) such that:
+The Poisson equation is the classic problem of finding $u(x,y)$ given a function $f(x,y)$ such that:
 \begin{align}
@@
 		u(x,y,z) = 0& \quad \text{on } \partial \Omega
-	\end{cases}.
+	\end{cases}
 	\label{eqn:poisson}
@@
 In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
-\bstodo{get figs and update the comments here}
-
-In Figure \ref{fig:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error.
-The example was chosen so that the exact solution was $u(x,y,z) = \genjacw^{(1,0)}(z) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y,z) = \Delta \big[\genjacw^{(1,0)}(z) y^3 \exp(x)\big]$.
-We see that the computed solution is almost exact.
+In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+In the plot, a "block" is simply the group of coefficients corresponding to OPs of the same degree, and so the plot shows how the norms of these blocks decay as the degree of the expansion increases.
+Thus, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
+We see that we achieve spectral convergence \bstodo{Add to the comments here}
+
+\bstodoinline{Are there more suitable RHSs for the block norms plot that will demonstrate further properties? Or as there are no corners, we are pretty limited?}
 
@@
 
-Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
+Find $u(x,y)$ given functions $v$, $f : \Omega \to \R$ such that:
 \begin{align}
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y,z) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y,z) = x(1-x^2-y^2)e^x$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y,z) = 1 - (3(x-1)^2 + 5y^2)$ using N = 200, that is, 20,301 unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
-\bstodo{get figs and update the comments here}
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
+In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ using $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.
+\bstodo{Add to the comments here}
 
@@
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
-We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
-\bstodo{get figs and update the comments here}
-
-% For the final Gaussian bump example, we see that we achieve spectral convergence.
-
-%
-\section{Tangent space (vector-valued functions)}\label{Section:tangentspace}
-
-We can extend the methodology detailed here to the tangent space of the spherical cap $\Omega$.
-By creating a basis of orthogonal vector-valued functions (OVFs) for the spherical cap $\Omega$, we can expand vector-valued functions that lie in tangent space we call $\tangentspace$ in this basis.
-Such functions that are useful for the sphere are gradients and perpendicular-gradients of scalar functions.
-In this section we will define the OVFs, outline the Clenshaw matrices and how to build the OVFs, function evaluation as well as differential operators for gradient, curl and divergence that take us to and from coefficients in either of the OVF or scalar OP expansions.
-
-Any function in the tangent space of $\Omega$ can be written as $\nabla f + \rvec \times \nabla g$ for some scalar functions $f, g$ on $\Omega$.
-Let $f$ be a scalar function on $\Omega$ that satisfies zero boundary conditions.
-Then there exist coefficients $\bm{f}$ such that $f$ can be expanded in the OP basis $\bigW_N^{(1)}$ for large enough N, i.e.
-$f = \genjacw^{(1,0)} \: \sum_{n,k,i} f_{n,k,i} \: \scopnki^{(1)}$.
-The gradient and perpendicular gradient of $f$ are then
-\begin{align*}
-	\nabla f &= \sum_{n,k,i} f_{n,k,i} \: \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \\
-	\rvec \times \nabla f &= \sum_{n,k,i} f_{n,k,i} \: \rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big).
-\end{align*}
-The game is to find an orthogonal basis such that we can sparsely expand $\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ and $\rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ for each $n,k,i$ trio.
-
-Define the OVFs as follows:
-\begin{align}
-	\tsopinki(x,y,z) &:= \phivec \: \frac{1}{\rho(z)} \: \scopnki^{(0)}(x,y,z) \\
-	\tsopiinki(x,y,z) &:= \thetavec \: \frac{1}{\rho(z)} \: \scopnki^{(0)}(x,y,z)
-\end{align}
-where
-\begin{align}
-	\phivec &:= 
-		\begin{pmatrix}
-			\cos(\theta) \: \cos(\phi) \\
-			\sin(\theta) \: \cos(\phi) \\
-			- \sin(\phi)
-		\end{pmatrix}
-		=
-		\begin{pmatrix}
-			x z / \rho(z) \\
-			y z / \rho(z) \\
-			- \rho(z)
-		\end{pmatrix} \\
-	\thetavec &:= 
-		\begin{pmatrix}
-			-\sin(\theta) \\
-			\cos(\theta) \\
-			0
-		\end{pmatrix}
-		=
-		\begin{pmatrix}
-			-y / \rho(z) \\
-			x / \rho(z) \\
-			0
-		\end{pmatrix}
-\end{align}
-These are orthogonal with respect to the inner product $\ip<\bm{A}, \bm{B}>_\tangentspace := \int_\Omega \bm{A} \cdot \bm{B} \: \rho(z)^2 \:\D \Omega$:
-\begin{align*}
-	\ip<\tsopinki, \tsopi_{m,j,h}>_\tangentspace &= \int_\Omega \tsopinki \cdot \tsopi_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
-	&= \int_\Omega \scopnki^{(0)} \: \scopmjh^{(0)} \: \D \Omega \\
-	&= \pi \: \normgenjac^{(0, 2k)} \: \delta_{n,m} \: \delta_{k,j} \: \delta_{i,h}, \\
-	\ip<\tsopiinki, \tsopii_{m,j,h}>_\tangentspace &= \int_\Omega \tsopiinki \cdot \tsopii_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
-	&= \int_\Omega \scopnki^{(0)} \: \scopmjh^{(0)} \: \D \Omega \\
-	&= \pi \: \normgenjac^{(0, 2k)} \: \delta_{n,m} \: \delta_{k,j} \: \delta_{i,h}, \\
-	\ip<\tsopinki, \tsopii_{m,j,h}>_\tangentspace &\equiv 0.
-\end{align*}
-To show that these span the tangent space $\tangentspace$, we merely require that we can sparsely expand $\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \: \rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ with the vectors $\tsopi_{m,j,h}, \tsopii_{m,j,h}$.
-First, note that
-\begin{align*}
-	\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) &= \phivec \: \ppphi \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) + \thetavec \: \frac{1}{\rho(z)} \: \pptheta \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \\
-	&= \phivec \: \Big[k \: z \: \genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) -  \big(\genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) \big)' \rho(z)^2 \Big] \: \rho(z)^{k-1} \: \chki(\theta) \\
-	&\quad \quad + \thetavec \: (-1)^{i+1} \: k \: \genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) \: \rho(z)^{k-1} \: \ch_{k,|i-1|}(\theta).
-\end{align*}
-Then,
-\begin{align*}
-	&\ip<\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \tsopi_{m,j,h}>_\tangentspace \\
-	&\quad \quad = \int_\Omega \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \cdot \tsopi_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
-	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big[k \: z \: \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} -  (\genjacw^{(1,0)} \: \genjacnmk^{(1,2k)})' \rho^2 \Big] \: \rho^{k-1} \: \genjac_{m-k}^{(0,2k)} \: \rho^{k-1} \: \rho^2 \:\D z \\
-	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big[k \: z \: \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} -  (\genjacw^{(1,0)} \: \genjacnmk^{(1,2k)})' \rho^2 \Big] \: \genjac_{m-k}^{(0,2k)} \: \rho^{2k} \:\D z \quad (\star) \\
-	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big( kz \: \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(0,2k)} \: \genjacw^{(1,2k)} \\
-	&\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad+ \genjacw^{(1,2k)} \: \genjacnmk^{(1,2k)} \: \Big[ \genjac_{m-k}^{(0,2k) \; \prime} \: \rho^2 - (2k+2) \: z \: \genjac_{m-k}^{(0,2k)} \Big] \Big) \: \D z \\
-	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \genjacnmk^{(1,2k)} \: \Big[ - (k + 2) \: z \: \genjac_{m-k}^{(0,2k)} + \genjac_{m-k}^{(0,2k) \; \prime} \: \rho^2 \Big] \: \genjacw^{(1,2k)} \: \D z \quad (\star \star)
-\end{align*}
-where by orthogonality of the $\{\genjac_n^{(a,2b)}\}$ OPs we have that the above is zero for $m - k > n - k + 2 \iff m > n + 2$ at $(\star)$ and also for $n - k > m - k + 1 \iff m < n - 1$ at $(\star \star)$.
-Further,
-\begin{align*}
-	&\ip<\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \tsopii_{m,j,h}>_\tangentspace \\
-	&\quad \quad = \int_\Omega \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \cdot \tsopii_{m,j,h} \: \rho(z)^2 \:\D \Omega \\
-	&\quad \quad = (-1)^{i+1} \: k \: \pi \: \delta_{k,j} \: \delta_{i,|h-1|} \: \int_\alpha^1 \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} \: \rho^{k-1} \: \genjac_{m-k}^{(0,2k)} \: \rho^{k-1} \: \rho^2 \:\D z \\
-	&\quad \quad = (-1)^{i+1} \: k \: \pi \: \delta_{k,j} \: \delta_{i,|h-1|} \: \int_\alpha^1 \genjacnmk^{(1,2k)}  \: \genjac_{m-k}^{(0,2k)} \: \genjacw^{(1,2k)} \: \D z
-\end{align*}
-which again by orthogonality is zero for $m - k > n - k + 1 \iff m > n + 1$ and $n - k > m - k \iff m < n$.
-We of course have similar results for $\rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ since
-\begin{align*}
-	\rvec \times \phivec = \thetavec, \quad \rvec \times \thetavec = - \phivec.
-\end{align*}
-We now have a set of orthogonal vector valued functions that span the tangent space $\tangentspace$.
-By construction, we have the same recurrence coefficients for multiplication by $x,y,z$ for each of the OVFs as for the OPs $\scopnki^{(0)}$.
-
-Define
-\begin{align*}
-	\bigtsoptn := 
-		\begin{pmatrix}
-			\tsopit_{n,0,0}(x,y,z) \\
-			\tsopiit_{n,0,0}(x,y,z) \\
-			\tsopit_{n,1,0}(x,y,z) \\
-			\tsopiit_{n,1,0}(x,y,z) \\
-			\tsopit_{n,1,1}(x,y,z) \\
-			\tsopiit_{n,1,1}(x,y,z) \\
-			\vdots \\
-			\tsopit_{n,n,0}(x,y,z) \\
-			\tsopiit_{n,n,0}(x,y,z) \\
-			\tsopit_{n,n,1}(x,y,z) \\
-			\tsopiit_{n,n,1}(x,y,z)
-		\end{pmatrix} \in \R^{2(2n+1) \times 3}, 
-	\quad \quad 
-	\bigtsopt := 
-		\begin{pmatrix}
-			\bigtsopt_0 \\
-			\bigtsopt_1 \\
-			\bigtsopt_2 \\
-			\vdots \\
-		\end{pmatrix}
-\end{align*}
-and set $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ as the Jacobi matrices corresponding to
-\begin{align}
-	\jacobimattangentx \: \bigtsopt(x,y,z) = x \: \bigtsopt(x,y,z), \nonumber \\
-	\jacobimattangenty \: \bigtsopt(x,y,z) = y \: \bigtsopt(x,y,z), \label{eqn:jacobimatricestangentdefinition} \\
-	\jacobimattangentz \: \bigtsopt(x,y,z) = z \: \bigtsopt(x,y,z), \nonumber
-\end{align}
-The matrices $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ act on the coefficients vector of a function's expansion in the $\bigtsopt$ basis.
-For example, let a function $\bm{F}(x,y,z) : \Omega \to \tangentspace$ be approximated by its expansion $\bm{F}(x,y,z)^\top = \bigtsopt(x,y,z)^\top \bm{F}_c$, for some coefficients vector $\bm{F}_c$.
-Then $x \: \bm{F}(x,y,z)$ is approximated by $\bigtsopt(x,y,z)^\top {(\jacobimattangentx)}^\top \bm{F}_c$.
-In other words, ${(\jacobimattangentx)}^\top \bm{F}_c$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: \bm{F}(x,y,z)$ in the  $\bigtsopt$ basis.
-Further, note that $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ are banded-block-banded matrices -- they are block-tridiagonal (block-bandwidths $(1,1)$):
-\begin{align*}
-	\jacobimattangent_{x/y/z} &= 
-		\begin{pmatrix}
-			B_{x/y/z, 0} & A_{x/y/z, 0} & & & & \\
-			C_{x/y/z, 1} & B_{x/y/z, 1} & A_{x/y/z, 1} & & & \\
-			& C_{x/y/z, 2} & B_{x/y/z, 2} & A_{x/y/z, 2} & & & \\
-			& & C_{x/y/z, 3} & \ddots & \ddots & \\
-			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}
-\end{align*}
-For $\jacobimattangentx$, the sub-blocks have sub-block-bandwidths $(?,?)$:
-\begin{align*}
-	A_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A_{n,0,6} & 0 & & \\
-			A_{n,1,5} & \ddots & \ddots & & \\
-			& \ddots & \ddots & \ddots & \\
-			& & A_{n,n,5} & 0 & A_{n,n,6} \\
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
-	B_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A_{n,0,4} & & \\
-			A_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & A_{n,n-1,4} \\
-			& & A_{n,n,3} & 0
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
-	C_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A_{n,0,2} & & \\
-			A_{n,1,1} & \ddots & \ddots & \\
-			& \ddots & \ddots & A_{n,n-2,2} \\
-			& & \ddots & 0 \\
-			& & & A_{n,n,1} \\
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
-where
-\begin{align*}
-	A_{n,k,j} &:= \alphaonkj \: I_{4} \in \R^{4\times4}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
-	A_{n,0,j} &:=
-		\begin{pmatrix}
-			\alphao_{n,0,j} & 0 & 0 & 0 \\
-			0 & \alphao_{n,0,j} & 0 & 0
-		\end{pmatrix} \in \R^{2\times4}, \quad j \text{ even} \\
-	A_{n,1,j} &:=
-		\begin{pmatrix}
-			\alphao_{n,1,j} & 0 \\
-			0 & \alphao_{n,1,j} \\
-			0 & 0 \\
-			0 & 0
-		\end{pmatrix} \in \R^{4\times2}, \quad j \text{ odd}
-\end{align*}
-For $\jacobimattangenty$, the sub-blocks have sub-block-bandwidths $(?,?)$:
-\begin{align*}
-	A_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B_{n,0,6} & 0 & & \\
-			B_{n,1,5} & \ddots & \ddots & & \\
-			& \ddots & \ddots & \ddots & \\
-			& & B_{n,n,5} & 0 & B_{n,n,6} \\
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
-	B_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B_{n,0,4} & & \\
-			B_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & B_{n,n-1,4} \\
-			& & B_{n,n,3} & 0
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
-	C_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B_{n,0,2} & & \\
-			B_{n,1,1} & \ddots & \ddots & \\
-			& \ddots & \ddots & B_{n,n-2,2} \\
-			& & \ddots & 0 \\
-			& & & B_{n,n,1} \\
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
-where
-\begin{align*}
-	B_{n,k,j} &:= 
-		\begin{pmatrix}
-			0 & 0 & \betaonkj & 0 \\
-			0 & 0 & 0 & \betaonkj \\
-			\betaonkj & 0 & 0 & 0 \\
-			0 & \betaonkj & 0 & 0
-		\end{pmatrix} \in \R^{4\times4}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
-	B_{n,0,j} &:=
-		\begin{pmatrix}
-			0 & 0 & \betao_{n,0,j} & 0 \\
-			0 & 0 & 0 & \betao_{n,0,j}
-		\end{pmatrix} \in \R^{2\times4}, \quad j \text{ even} \\
-	B_{n,1,j} &:=
-		\begin{pmatrix}
-			0 & 0 \\
-			0 & 0 \\
-			\betao_{n,1,j} & 0 \\
-			0 & \betao_{n,1,j}
-		\end{pmatrix} \in \R^{4\times2}, \quad j \text{ odd}
-\end{align*}
-For $\jacobimattangentz$, the sub-blocks have sub-block-bandwidths $(?,?)$:
-\begin{align*}
-	A_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma_{n,0,3} & 0 & 0 & & \\
-			0 & \ddots & \ddots & & \\
-			& \ddots & \ddots & \ddots & \\
-			& & 0 & \Gamma_{n,n,3} & 0 \\
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
-	B_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma_{n,0,2} & 0 & & \\
-			0 & \ddots & \ddots & \\
-			& \ddots & \ddots & 0 \\
-			& & 0 & \Gamma_{n,n,2}
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
-	C_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma_{n,0,1} & 0 & & \\
-			0 & \ddots & \ddots & \\
-			& \ddots & \ddots & 0 \\
-			& & \ddots &  \Gamma_{n,n-1,1} \\
-			& & & 0 \\
-		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
-where
-\begin{align*}
-	\Gamma_{n,k,j} &:= \gammaonkj \: I_{4} \in \R^{4\times4}, \quad k = 1,\dots,n \\
-	\Gamma_{n,0,j} &:=
-		\begin{pmatrix}
-			\gammao_{n,0,j} & 0 \\
-			0 & \gammao_{n,0,j}
-		\end{pmatrix} \in \R^{2\times2}.
-\end{align*}
-Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
-
-\subsection{Building the OPs}
-
-We can combine each system in (\ref{eqn:jacobimatricestangentdefinition}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
-
-\begin{align*}
-\renewcommand\arraystretch{1.3}
-\begin{pmatrix}
-		I_2 & & & \\
-		B_0-G_0(x,y,z) & A_0 & & \\
-		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
-		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
-		& & \ddots &\ddots
-\end{pmatrix}
-\bigtsopt(x,y,z)
-=
-\begin{pmatrix}
-	 \tsopi_0^\top \\ \tsopii_0^\top \\ 0 \\ 0 \\ \vdots  \\
-\end{pmatrix},
-\end{align*}
-where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
-\begin{align*}
-A_n &:= \begin{pmatrix}
-		A_{x,n} \\
-		A_{y,n} \\
-		A_{z,n}
-	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n+3)}, \quad
-C_n := \begin{pmatrix}
-		C_{x,n} \\
-		C_{y,n} \\
-		C_{z,n}
-	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n-1)} \quad (n \ne 0), \nonumber \\
-B_n &:= \begin{pmatrix}
-		B_{x,n} \\
-		B_{y,n} \\
-		B_{z,n}
-	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n+1)}, \quad
-G_n(x,y,z) := \begin{pmatrix}
-		xI_{2(2n+1)} \\
-		yI_{2(2n+1)} \\
-		zI_{2(2n+1)}
-	    \end{pmatrix} \in \R^{6(2n+1)\times2(n+1)}.
-\end{align*}
-
-For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e.
-such that $\Dnt A_n = I_{2(2n+3)}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
-\begin{align*}
-	\begin{cases}
-		\bigtsopt_{-1}(x,y,z) := 0 \\
-		\bigtsopt_{0}(x,y,z) = 
-			\begin{pmatrix} 
-				\tsopi_0^\top \\ 
-				\tsopii_0^\top
-			\end{pmatrix} :=
-			\begin{pmatrix} 
-				\tsopi_{0,0,0}^\top \\
-				\tsopi_{0,0,0}^\top
-			\end{pmatrix} \\
-		\bigtsopt_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigtsoptn(x,y,z) - \Dnt C_n  \, \bigtsopt_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
-	\end{cases}
-\end{align*}
-
-Note that we can define an explicit \(\Dnt\) as follows: \bstodo{type up DnT for tangent space}
-\begin{align*}
-\Dnt := \begin{pmatrix}
-		0 & & & 0 & & & 1 / \gammaa_{n,0,3} & & \\
-		& \ddots & & & \ddots & & & \ddots \\
-		& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
-		& & & & & \bm{\eta}^\top_{1} & & & \\
-		& & & & & \bm{\eta}^\top_{0} & & &
-	    \end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
-\end{align*}
-for $n = 2, 3, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
-\begin{align*}
-	\big(\bm{\eta}_{0}\big)_j &= 
-		\begin{cases}
-			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
-			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2 \\
-			0 & o/w
-		\end{cases} \\
-	\big(\bm{\eta}_{1}\big)_j &= 
-		\begin{cases}
-			\big(\bm{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
-			0 & o/w
-		\end{cases}
-\end{align*}
-For $n=0, 1$, we can simply take
-\begin{align*}
-D^\top_0 &:= \begin{pmatrix}
-		0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
-		\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
-		0 & \frac{1}{\betaa_{0,0,6}} & 0
-	    \end{pmatrix} \in \R^{3\times3}, \\
-D^\top_1 &:= \begin{pmatrix}
-		0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,0,3} & & \\
-		0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3} \\
-		0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3}  \\
-		0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & \eta & 0 & 0 \\
-		0 & 0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & 0 & 0
-	    \end{pmatrix} \in \R^{5\times9},
-\end{align*}
-where $\eta = \frac{- \: \alphaa_{1,1,5}}{\alphaa_{1,1,6} \: \gammaa_{1, 0, 3}}$.
-
-It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigtsopt_0(x,y,z)$  through $\bigtsoptn(x,y,z)$ in optimal $O(n^2)$ complexity. \bstodo{correct for this case}
-
-\subsection{Operators}
-
-Gradient operator on a scalar function is pretty much outlined above.
-Divergence would either be worked out by knowing what $f,g$ are for a function $\bm{F} = \nabla f + \rvec \times \nabla g$, so that $\nabla \cdot \bm{F} = \nabla^2 f$.
-Alternatively, we have expressions for $\rhoppphi \scopnkia$ and $\pptheta \scopnkia$ so that we can calculate an operator for $\rho^2 \nabla \cdot$, i.e.
-\begin{align*}
-	\rho^2 \nabla \cdot \bm{F} &= \rho^2 \: \frac{1}{\rho} \ppphi (\rho F^\phi) + \rho^2 \: \frac{1}{\rho} \pptheta F^\theta \\
-	&= \sum_{n,k,i} F^\phi_{n,k,i} \: \rhoppphi \scopnki^{(0)} + F^\theta_{n,k,i} \: \pptheta \scopnki^{(0)}.
-\end{align*}
-
-\bstodoinline{My main concern is whether this vector function basis is correct/valid.
-I have found it difficult to come up with something that gives orthogonal vectors while still maintaining a sparse expansion for the gradient of a function.
-I am also confused about what the rho factors "mean".}
-
-\subsection{Linear SWE}
-
-\begin{align*}
-	\uvec_t &= - f \rvec \times \uvec + \grad h \\
-	h_t &= -\href \divergence \uvec
-\end{align*}
-$\implies$
-\begin{align*}
-	\uvecnpi &= \uvecn - \Delta t \: f \rvec \times \uvecnpi +\Delta t \: \grad h_{n+1} \\
-	h_{n+1} &= h_n - \Delta t \: \href \divergence \uvecnpi
-\end{align*}
-$\implies$
-\begin{align*}
-	\uvecnpic &= \uvecnc - \Delta t \: F \uvecnpic +\Delta t \: G \hvecnpic \\
-	P \hvecnpic &= P \hvecnc - \Delta t \: \href D \uvecnpic
-\end{align*}
-where $F : \bigtsop \to \bigtsop$ is the operator for $f \rvec \times$, $D : \bigtsop \to \bigscop^{(1)}$ is the operator for $\rho^2 \divergence$, $G : \bigW^{(1)} \to \bigtsop$ is the operator for $\grad$, and $P : \bigW^{(1)} \to \bigscop^{(1)}$ is the operator for multiplication by $\rho(z)^2$.
+In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
+In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+\bstodo{Add to the comments here}
 
@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-        	%\label{fig:solutionblocknorms-poisson}
-	\centering
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-	%\label{fig:solution-poisson}
-	\centering
-	\end{subfigure}
-	\centering
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
+\bstodoinline{Is this worth including?}
+
+%\begin{figure}[t]
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
+%        	%\label{fig:solutionblocknorms-poisson}
+%	\centering
+%	\end{subfigure}
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.3]{solution-poisson}
+%	\centering
+%	%\label{fig:solution-poisson}
+%	\end{subfigure}
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
+%	%\label{fig:solution-poisson}
+%	\centering
+%	\end{subfigure}
+%	\centering
+%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
 Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
@@
 in the trapezium.}
-	\centering
-	\label{fig:appendixfigs}
-\end{figure}
+%	\centering
+%	\label{fig:appendixfigs}
+%\end{figure}
 

### 2020-08-20 papers/spherical-caps/spherical-caps.tex
Draft of paper, with pdf figs included. Should be able to compile.

@@
 	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k - \tilde k_i} \: y^{\tilde k_i} \: z^{n-k} + (\hbox{lower order terms})
-\end{align*} \bstodo{Ordering correct?}
+\end{align*} \bstodoinline{How can I write down what the ordering is? Is this necessary?}
 where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - \tilde j_i} \: y^{\tilde j_i} \: z^{n-j}$ where $j < k$.
@@
 In this work, as for the disk-slice case in 2D \cite{snowball2019sparse}, we use a more general integration-by-parts argument to deduce the sparsity structure.
-The 3D OPs defined here involve the same non-classical (in fact, semi-classical) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
-In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
-This compares favourably to $O(p^6)$ operations if one proceeds na\"ively.
-Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
-\bstodo{reword to avoid plagiarising self}
+The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
+In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
+This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations \bstodoinline{Presumably this is correct to say?}.
+Furthermore, we can use this framework to derive sparse $p$-finite element methods, which we introduce as an Appendix here.
+The methods are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
 
@@
 \begin{definition}
-Define $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ as the 2D OPs on $(x,y)\in\R$ s.t.
-$x^2 + y^2 = 1$; that is, if we define $\theta$ by $x = \cos(\theta)$, $y = \sin(\theta)$ for $\xvec := (x, y)$ where $||\xvec|| = 1$,
-\begin{align}
-	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) \equiv \ch_{0,0}(\theta) &:= \frac{\sqrt{2}}{2} =: \ch_{0} \nonumber \\
-	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \equiv \ch_{k,0}(\theta) &:= \cos(k \theta), \quad k = 1,2,3,\dots\label{eqn:circularharmonics} \\
-	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) \equiv \ch_{k,0}(\theta)& := \sin(k \theta), \quad k = 1,2,3,\dots \nonumber
-\end{align}
-orthonormal with respect to the inner product
+Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos(\theta)$, $y = \sin(\theta)$.
+Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
+\begin{align}
+	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \nonumber \\
+	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos(k \theta) =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots\label{eqn:chebyshev} \\
+	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin(k \theta) =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots \nonumber
+\end{align}
+where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
 \begin{align}
@@
 \end{definition}
-Of course, these are simply Chebyshev polynomials $\ch_{k,0}(x, y) = T_k(x)$ and $\ch_{k,1}(x, y) = y \: U_{k-1}(x)$ for $k > 0$, where we define $\ch_0$ so as to ensure orthonormality.
+Note that we have defined $\ch_0$ so as to ensure orthonormality.
 
@@
 	\ip< p, \: q > &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: w(z) \: \D \sigma(\xvec, z) \\
-	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: w(z) \: \D z.
+	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: w(z) \: \D z
 \end{align*}
@@
 \begin{definition}[\cite{snowball2019sparse}]\label{def:OPconstruction}
-Let $\genjacw^{(a,b)}(x)$ be a weight function on the interval $(\alpha, \beta)$ given by:
-\begin{align*}
-	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{b} \: \rho(x)^{b}
+Let $\genjacw^{(a,b)}(x)$ be a weight function on the interval $(\alpha, 1)$ given by:
+\begin{align*}
+	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b}
 \end{align*}
@@
 \begin{align}
-	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipgenjac}
+	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipgenjac}
 \end{align}
@@
 \begin{align}
-	\normgenjac^{(a,b)} := \int_\alpha^\beta \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipnormalisation}
+	\normgenjac^{(a,b)} := \int_\alpha^1 \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipnormalisation}
 \end{align}
@@
 \begin{align}
-	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} \nonumber \\
-	&= \pi \: \normgenjac^{(a, 2k)} \label{eqn:normscop}
+	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)} \label{eqn:normscop}
 \end{align}
@@
 	z &= \cos(\phi) \\
-	x &= \sin(\phi) \: \cos(\theta) = \rho(z) \: \cos(\theta) \\
-	y &= \sin(\phi) \: \sin(\theta) = \rho(z) \: \sin(\theta).
+	x &= \sin(\phi) \cos(\theta) = \rho(z) \cos(\theta) \\
+	y &= \sin(\phi) \sin(\theta) = \rho(z) \sin(\theta).
 \end{align*}
@@
 
-It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigscopta_{0}(x,y)$  through $\bigscopta_{n}(x,y)$ in optimal $O(n^2)$ complexity. \bstodo{correct for this case}
+It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigscopta_{0}(x,y,z)$  through $\bigscopta_{n}(x,y,z)$ in optimal $O(n^2)$ complexity.
+\bstodoinline{I think this is correct for this case}
 
 \begin{definition}\label{def:clenshawmats}
-The Clenshaw matrices associated with the OPs $\{\scopnkia\}$ given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
+The Clenshaw matrices associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
 \end{definition}
@@
 For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 200$.}
-	\label{fig:images}
+	\label{fig:sparsity}
 \end{figure}
@@
 \end{align*}
-Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}. \bstodo{Add this fig}
+Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
 
@@
 \end{align*}
-The Clenshaw algorithm is then \cite{}\bstodo{Reference for Clenshaw in >1 dimensions???} 
+The Clenshaw algorithm is then \cite{}\bstodoinline{Reference for Clenshaw in >1 dimensions???} 
 \begin{align*}
@@
 We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$.
-This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodo{correct the O(nd) complexity}.
+This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodoinline{correct the O(nd) complexity}.
 
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.17]{poisson-u=wyexpx-N=60}
+	\includegraphics[scale=0.25]{poisson-u=wyexpx-N=60}
 	\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.17]{helmholtz-f=wyexpx-N=60-n=300}
+	\includegraphics[scale=0.25]{helmholtz-f=wyexpx-N=60-n=300}
 	\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.17]{biharmonic-f=erf-N=80-n=50}
+	\includegraphics[scale=0.25]{biharmonic-f=erf-N=80-n=50}
 	\centering
@@
 Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
-This demonstrates spectral convergence observed.}
+This demonstrates algebraic convergence observed.}
 	\centering
@@
 \begin{align*}
-	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(x,y) \D z = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(x,y) \D z =: a(u,v)
-\end{align*}
- for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$.
+	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
+\end{align*}
+ for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
 
@@
 In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain.
-We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.
+We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p \rvert_\element) \le p\}$ for some $p \in \N$.
 
@@
 
-Define $\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa}$ for any $a$.
+Define 
+\begin{align*}
+	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
+\end{align*}
+for any $a$.
 Note that due to orthogonality this is a diagonal matrix.
@@
 \begin{align*}
-	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(x,y) \D z \\
-	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(x,y) \D z \\
+	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
 	&= \int_\element \Big( \mathbf{v}^\top {W_\phi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\phi^{(1)} \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(x,y) \D z \\
+					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
 	&= \mathbf{v}^\top \: \Big( {W_\phi^{(1)}}^\top \Lambda^{(0)}  W_\phi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
 \end{align*}
-Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p = \rho(z) \: \bigscopi_p$.
+Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
 Then,
 \begin{align*}
-	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(x,y) \D z \\
-	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(x,y) \D z \\	
+	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
 	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.

### 2020-09-02 papers/spherical-caps/spherical-caps.tex
Readthrough

@@
 \newcommand{\bigWNa}{\mathbb{W}_N^{(a)}}
-\newcommand{\ppphi}{\frac{\partial}{\partial \phi}}
+\newcommand{\ppphi}{\frac{\partial}{\partial \varphi}}
 \newcommand{\ppz}{\frac{\partial}{\partial z}}
@@
 \newcommand{\tangentspace}{{\Omega_T}}
-\newcommand{\phivec}{\hat{\bm{\phi}}}
+\newcommand{\phivec}{\hat{\bm{\varphi}}}
 \newcommand{\thetavec}{\hat{\bm{\theta}}}
@@
 
+\def\DeltaS{\Delta_{\rm S}}
+
 \usepackage{amsthm}
@@
 
-\title{Sparse spectral and $p$-finite element methods for partial differential equations on spherical caps}
+\title{Sparse spectral methods for partial differential equations on spherical caps and bands}
 \author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
@@
 In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
-In this work we extend the methodology on disk-slices developed in \cite{snowball2019sparse} to a hierarchy of non-classical orthogonal polynomials on spherical caps.
-This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
+In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps and bands.
+The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
+We demonstrate the results on partial differential equations involving the spherical Laplacian and Biharmonic operators, showing spectral convergence.
 \end{abstract}
@@
 
-This paper develops sparse spectral methods for solving linear partial differential equations on spherical subdomains such as spherical bands and spherical caps.
+This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere: spherical bands and spherical caps.
 More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
 \begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| := \sqrt{x^2 + y^2} = \gamma \rho(z)\}
-\end{align*}
-and consider the case where $\alpha \in (0,1)$ and where
-\begin{align*}
-	 \rho(z) &:= \sqrt{1-z^2} \\
-	 \beta &:= 1 \\
-	 \gamma &:= 1.
-\end{align*}
+	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| := \sqrt{x^2 + y^2} =  \rho(z)\}
+\end{align*}
+and consider the case where \sotodo{Why is $\alpha$ restricted to be positive?} $\alpha \in (0,1)$ and where
+$ \rho(z) := \sqrt{1-z^2}$,  $\beta := 1$ \sotodo{Don't we need $\beta \neq 1$ to get a band?}.
 This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
@@
 
-As in the work in two dimensions on the show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that are defined in such a way that mirrors the approach on quadratic curves \cite{olver2020orthogonal}.
+We advocate using a basis that is polynomial in cartesian coordinates, that is, polynomial in $x$, $y$, and $z$, and orthogonal with respect to a prescribed weight: that is, multivariate orthogonal polynomials, whose construction was considered in \cite{olver2020orthogonal}.
+Equivalently, we can think of these as polynomials modulo the vanishing ideal $\{ x^2 + y^2 + z^2 = 1 \}$, or simply as a linear recombination of spherical harmonics that are orthogonalised on a subset of the sphere.
+This is in contrast to more standard approaches based on mapping the geometry to a simpler one (e.g., a rectangle or disk) and using orthogonal polynomials in the mapped coordinates (e.g., a basis that is polynomial in the spherical coordinates $\varphi$ and $\theta$).
+The benefit of the new approach is that we do not need to resolve Jacobians, and thereby we can achieve sparse discretisations for partial differential operators, including those with polynomial variable coefficients.
 On the spherical cap, the family of weights we consider are of the form
@@
 The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
-We define these to be orthogonalised lexicographically, that is,
+We define these to be orthogonalised lexicographically, that is, \sotodo{I don't understand this ordering.
+Shouldn't it either be $x^k y^j z^{n-k-j}$ or $x^k y^{j-k} z^{n-j}$?
+PS I prefer $j$ to $i$ }
 \begin{align*}
@@
 
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$.
-In this work, as for the disk-slice case in 2D \cite{snowball2019sparse}, we use a more general integration-by-parts argument to deduce the sparsity structure.
-The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
+We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
+\begin{align*}
+	z &= \cos\varphi, \\
+	x &= \sin\varphi \cos\theta = \rho(z) \cos\theta, \\
+	y &= \sin\varphi \sin\theta = \rho(z) \sin\theta.
+\end{align*}
+we have \sotodo{Double check}
+\begin{align*}
+\DeltaS &= {1 \over \sin \varphi} {\partial \over \partial \varphi} \sin \varphi {\partial \over \partial \varphi} + {1 \over \sin^2 \varphi} {\partial^2 \over \partial \theta^2}  = {1 \over \rho} {\partial \over \partial \varphi} \rho {\partial \over \partial \varphi} + {1 \over \rho^2} {\partial^2 \over \partial \theta^2} \\
+\end{align*}
+We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
+As in the  disk-slice case in 2D \cite{snowball2019sparse}, we use an integration-by-parts argument to deduce the sparsity structure.
+
+ The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical\sotodo{Add citation}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
 In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
-This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations \bstodoinline{Presumably this is correct to say?}.
+This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations.
 Furthermore, we can use this framework to derive sparse $p$-finite element methods, which we introduce as an Appendix here.
@@
 
+Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + V(x,y,z)$ where $V$ is first approximated by a polynomial.
+A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axes (e.g., a Schr\"odinger operator with potential $V(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
+This improves the complexity further to an optimal $O(p^2$).
+\sotodo{Adding a complexity plot showing this would be great}
+
 An overview of the paper is as follows:
@@
 
-\section{Orthogonal polynomials on the disk-slice and the trapezium}\label{Section:OPs}
-
-In this section we outline the construction and some basic properties of $\scopnkia(x,y)$.
+\section{Orthogonal polynomials on spherical caps}\label{Section:OPs}
+
+In this section we outline the construction and some basic properties of $\scopnkia(x,y,z)$.
 
@@
 We can construct the 3D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Chebyshev polynomials.
+We do so in terms of Fourier series, which, following \cite{olver2020orthogonal}, we write here as orthogonal polynomials in $x$ and $y$:
 
 \begin{definition}
-Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos(\theta)$, $y = \sin(\theta)$.
+Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
 Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
@@
 	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \nonumber \\
-	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos(k \theta) =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots\label{eqn:chebyshev} \\
-	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin(k \theta) =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots \nonumber
-\end{align}
-where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
+	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots\label{eqn:chebyshev} \\
+	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots \nonumber
+\end{align}
+where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$.
+The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product\sotodo{The $\sigma$ notation needs explanation}
 \begin{align}
@@
 Note that we have defined $\ch_0$ so as to ensure orthonormality.
+
+\sotodoinline{Isn't the following from \cite{olver2020orthogonal}? We then don't need to repeat the proof}
 
@@
 	\ip< p, \: q > &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: w(z) \: \D \sigma(\xvec, z) \\
-	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: w(z) \: \D z
+	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: w(z) \: \D z
 \end{align*}
@@
 	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D \sigma(\xvec, z) \\
-	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: q(\rho(z) \cos(\theta), \rho(z) \sin(\theta), z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
+	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
 \end{align}
@@
 We note that the weight $\genjacw^{(a,b)}(z)$ has been used in the construction of 2D orthogonal polynomials on disk-slices and trapeziums \cite{snowball2019sparse}, where a method for obtaining recurrence coefficients and evaluating integrals was established (the weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
-
-For clarity we state here that throughout we will regularly use coordinate transformations, mainly in any integrals over the surface $\Omega$, between the cartesian $(x,y,z)$ and the spherical coordinates $(\phi, \theta)$ -- that is, for any point $(x,y,z) \in \Omega$, let $(\phi, \theta)$ denote the angles such that:
-\begin{align*}
-	z &= \cos(\phi) \\
-	x &= \sin(\phi) \cos(\theta) = \rho(z) \cos(\theta) \\
-	y &= \sin(\phi) \sin(\theta) = \rho(z) \sin(\theta).
-\end{align*}
 
@@
 \begin{align*}
-	&1) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \cos(\theta) \: \D \theta = \ch_0 \: \pi \: \delta_{0,h} \: \delta_{1,j} \\
-	&2) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \sin(\theta) \: \D \theta = \ch_0 \: \pi \: \delta_{1,h} \: \delta_{1,j} \\
-	&3) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta = \pi \: \delta_{i,h} \: (\ch_0 \: \delta_{0,j} + \half \delta_{2,j}) \\
-	&4) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta = \pi \:  \delta_{|i-1|,h} \: ((-1)^{i+1} \: \ch_0 \: \delta_{0,j} + (-1)^i \: \half \: \delta_{2,j}) \\
-	&5) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta = \half \: \pi \: \delta_{i,h} \: (\delta_{k-1,j} + \delta_{k+1,j}) \\
-	&6) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta = \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \delta_{k-1,j} + (-1)^i \: \delta_{k+1,j}).
+	&1) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \cos\theta \: \D \theta = \ch_0 \: \pi \: \delta_{0,h} \: \delta_{1,j} \\
+	&2) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \sin\theta \: \D \theta = \ch_0 \: \pi \: \delta_{1,h} \: \delta_{1,j} \\
+	&3) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta = \pi \: \delta_{i,h} \: (\ch_0 \: \delta_{0,j} + \half \delta_{2,j}) \\
+	&4) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta = \pi \:  \delta_{|i-1|,h} \: ((-1)^{i+1} \: \ch_0 \: \delta_{0,j} + (-1)^i \: \half \: \delta_{2,j}) \\
+	&5) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta = \half \: \pi \: \delta_{i,h} \: (\delta_{k-1,j} + \delta_{k+1,j}) \\
+	&6) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta = \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \delta_{k-1,j} + (-1)^i \: \delta_{k+1,j}).
 \end{align*}
@@
 \begin{align*}
-	2 \cos(k \theta) \cos(\theta) &= \cos((k-1)\theta) + \cos((k+1)\theta) \\
-	2 \sin(k \theta) \cos(\theta) &= \sin((k-1)\theta) + \sin((k+1)\theta) \\
-	2 \cos(k \theta) \sin(\theta) &= - \sin((k-1)\theta) + \sin((k+1)\theta) \\
-	2 \sin(k \theta) \sin(\theta) &= \cos((k-1)\theta) - \cos((k+1)\theta).
+	2 \cos k \theta \cos\theta &= \cos (k-1)\theta + \cos(k+1)\theta \\
+	2 \sin k \theta \cos\theta &= \sin (k-1)\theta + \sin(k+1)\theta \\
+	2 \cos k \theta \sin\theta &= - \sin (k-1)\theta + \sin(k+1)\theta \\
+	2 \sin k \theta \sin\theta &= \cos (k-1)\theta - \cos(k+1)\theta.
 \end{align*}
@@
 Recall from equation (\ref{eqn:normscop}) that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$.
-Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos(\theta) \sin(\phi), \: \sin(\theta)\sin(\phi), \: \cos(\phi)) = (x, y, z)$:
+Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos(\theta) \sin(\varphi), \: \sin\theta\sin(\varphi), \: \cos(\varphi)) = (x, y, z)$:
 \begin{align*}
@@
 	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: x \: \genjacw^{(a,0)}(z) \: \D\sigma(\xvec, z) \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta \Big) \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos(\theta) \: \D \theta \Big) \\
+	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
+	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
 	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
@@
 	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: y \: \genjacw^{(a,0)}(z) \: \D\sigma(\xvec, z) \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta \Big) \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin(\theta) \: \D \theta \Big) \\
+	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
+	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
 	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
@@
 
-In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices}.
+In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated\sotodo{Explain what a Clenshaw matrix is} \enquote{Clenshaw matrices}.
 While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
@@
 
-Note that we can define an explicit \(\Dnt\) as follows:
+Note that we can define an explicit \(\Dnt\) as follows:\sotodo{Improve following notation, say, by adding notation to the top part}
 \begin{align*}
@@
 
-In this section, we will derive the entries that form spherical partial differential operators, demonstrating their sparsity.
-To this end, we will use a different ordering of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
+In this section, we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process.
+To this end, we will use a different ordering\sotodo{I thought we would discuss this separately as a permutation of rows/columns?} of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
 Let $N \in \N$ and recall from equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) that:
@@
 \begin{definition}\label{def:differentialoperators}
-Define the operator matrices $D_\phi^{(a)}, \: W_\phi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ according to:
-\begin{align*}
-	\rho {\partial f \over \partial \phi}(x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\phi^{(a)} \: \mathbf{f}, \\
-	\rho {\partial \over \partial \phi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\phi^{(a)}\: \mathbf{f}, \\
+Define the operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ according to:
+\begin{align*}
+	\rho {\partial f \over \partial \varphi}(x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \mathbf{f}, \\
+	\rho {\partial \over \partial \varphi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\varphi^{(a)}\: \mathbf{f}, \\
 	{\partial f \over \partial \theta}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
-	\nabla^2_s \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
-\end{align*}
-where $\nabla^2_s$ is the surface Laplacian, given by $\nabla^2_s = \frac{1}{\rho(z)} \ppphi \big( \rhozppphi \big) + \frac{1}{\rho(z)^2} \ppthetatwo$.
+	\DeltaS \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
+\end{align*}
 \end{definition}
@@
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, on the triangle \cite{olver2018recurrence}, and on the disk-slice \cite{snowball2019sparse}.
-The operators we define here are for partial derivatives with respect to the spherical coordinates $(\phi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).
+The operators we define here are for partial derivatives with respect to the spherical coordinates $(\varphi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).
 With the OP ordering by Fourier mode $k$ defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) these rotationally invariant operators are block-diagonal, meaning simple and parallelisable practical application.
@@
 \begin{theorem}\label{theorem:sparsityofdifferentialoperators}
-The operator matrices $D_\phi^{(a)}, \: W_\phi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure. More specifically:
+The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
+More specifically:
 \begin{itemize}
-	\item $D_\phi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 4)$
-  	\item $W_\phi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 2)$
+	\item $D_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 4)$
+  	\item $W_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 2)$
 	\item $D_\theta$ has  block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
@@
 \begin{proof}[Proof of Theorem \ref{theorem:sparsityofdifferentialoperators}]
-First, note that:
+First, note that \sotodo{Here and elsewhere: ${\partial \over \partial z}$ is not well-defined as we are on the sphere.
+Though perhaps ${\rm d} \over {\rm d} z$ works as we are differentiating the formula with respect to $z$?}:
 \begin{align}
@@
 	\rho(z) \: \rho'(z) &= -z \label{eqn:rhoderivative} \\
-	{\partial \over \partial \phi} &= -\rho \: {\partial \over \partial z} \label{eqn:partialphitoz} 
-\end{align}
-
-For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that
+	{\partial \over \partial \varphi} &= -\rho \: {\partial \over \partial z} \label{eqn:partialphitoz} 
+\end{align}
+
+For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that \sotodo{Would it be better as ${\rm d} \over {\rm d} \theta$?}
 \begin{align*}
@@
 
-We now proceed with the case for the operator $D_\phi^{(a)}$ for partial differentiation by $\phi$.
+We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$.
 The entries of the operator are given by the coefficients in the expansion $\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)}$, where the coefficients are
@@
 
-Similarly for the operator $W_\phi^{(a)}$ for partial differentiation by $\phi$ on the weighted space, the entries of the operator are given by the coefficients in the expansion $\rhoppphi (\genjacw^{(a,0)} \: \scopnkia) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-1,0)} \: \scopmjh^{(a-1)}$, where the coefficients are
+Similarly for the operator $W_\varphi^{(a)}$ for partial differentiation by $\varphi$ on the weighted space, the entries of the operator are given by the coefficients in the expansion $\rhoppphi (\genjacw^{(a,0)} \: \scopnkia) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-1,0)} \: \scopmjh^{(a-1)}$, where the coefficients are
 \begin{align*}
@@
 Finally, fix $a = 1$.
-For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are
-\begin{align*}
-	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
+For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are
+\begin{align*}
+	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
 \end{align*}
@@
 \begin{align*}
-	\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) &= 
+	\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) &= 
 \end{align*}
@@
 \begin{align*}
-	&\ip<\nabla^2_s \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
+	&\ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
 	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
@@
 		\centering
-		\caption{The Biharmonic operator $B_W^{(2)}$}
+		\caption{The Biharmonic operator $B_W^{(2)}$ \sotodoinline{This one is surprising as I was expecting rotational invariance to give us block-diagonal operators}}
 	\end{subfigure}\hfil % <-- added
@@
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
-As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored surface Laplacian $\rho(z)^2 \: \nabla^2_s$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
+As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored surface Laplacian $\rho(z)^2 \: \DeltaS$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in Definition \ref{def:differentialoperators}.
@@
 \begin{align*}
-    \mathcal{L}_W^{(1)} := D_\phi^{(0)} \: W_\phi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
+    \mathcal{L}_W^{(1)} := D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
 \end{align*}
@@
 
-Another desirable operator is the Biharmonic operator $\Delta^2_s$ (defined by $\Delta^2_s f = \nabla^2_s (\nabla^2_s f)$), for which we assume zero Dirichlet and Neumann conditions.
-To construct a matrix operator, we first note that we can obtain the matrix operator for the $\rho^2$-factored Laplacian $\rho(z)^2 \: \nabla^2_s$ that will take us from coefficients for expansion in the space $\bigscopN^{(0)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
+Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions \sotodo{We need to explain Neumann conditions precisely, as the normal derivative is not so obvious}.
+To construct a matrix operator, we first note that we can obtain the matrix operator for the $\rho^2$-factored Laplacian $\rho(z)^2 \: \DeltaS$ that will take us from coefficients for expansion in the space $\bigscopN^{(0)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
 We denote this matrix operator that acts on the coefficients vector as $\mathcal{L}^{(0) \to (2)}$, and it is given by
 \begin{align*}
-    \mathcal{L}^{(0) \to (2)} := D_\phi^{(2)} \: D_\phi^{(0)} + T^{(0)\to(2)} \: D_\theta.
+    \mathcal{L}^{(0) \to (2)} := D_\varphi^{(2)} \: D_\varphi^{(0)} + T^{(0)\to(2)} \: D_\theta.
 \end{align*}
@@
 \begin{align*}
-	\mathcal{L}_W^{(2) \to (0)} := W_\phi^{(1)} \: W_\phi^{(2)} + T_W^{(2)\to(0,0,0)} \: D_\theta.
-\end{align*}
-Since, for any scalar function $f:\Omega \to \R$,
-\begin{align*}
-	\rho^2(z) \: \Delta^2 \: f(x,y,z) = \rho^2 \: \nabla^2_s (\rho^2 \: \nabla^2_s f) - z \: \rhoppphi (\rho^2 \: \nabla^2_s f) - [2z^2 - 2\rho^2 + 3z] \:\rho^2 \: \nabla^2_s f
-\end{align*}
-the matrix operator for the $\rho^2$-factored Biharmonic operator $\rho(z)^2 \: \Delta_s^2$ is then given by
-\begin{align*}
-	\mathcal{B}_W^{(2)} := \mathcal{L}^{(0) \to (2)} \: \mathcal{L}_W^{(2) \to (0)} - (J_z^{(2)})^\top \: T^{(1)\to(2)} \: D_\phi^{(0)} \: \mathcal{L}_W^{(2) \to (0)} - V^{(2)} \:  T^{(0)\to(2)} \: \mathcal{L}_W^{(2) \to (0)},
+	\mathcal{L}_W^{(2) \to (0)} := W_\varphi^{(1)} \: W_\varphi^{(2)} + T_W^{(2)\to(0,0,0)} \: D_\theta.
+\end{align*}
+Since, for any scalar function $f:\Omega \to \R$, \sotodo{The extra $\rho^2$ in front is weird to me}
+\begin{align*}
+	\rho^2(z) \: \Delta^2 \: f(x,y,z) = \rho^2 \: \DeltaS (\rho^2 \: \DeltaS f) - z \: \rhoppphi (\rho^2 \: \DeltaS f) - [2z^2 - 2\rho^2 + 3z] \:\rho^2 \: \DeltaS f
+\end{align*}
+the matrix operator for the $\rho^2$-factored Biharmonic operator $\rho(z)^2 \: \DeltaS^2$ is then given by
+\begin{align*}
+	\mathcal{B}_W^{(2)} := \mathcal{L}^{(0) \to (2)} \: \mathcal{L}_W^{(2) \to (0)} - (J_z^{(2)})^\top \: T^{(1)\to(2)} \: D_\varphi^{(0)} \: \mathcal{L}_W^{(2) \to (0)} - V^{(2)} \:  T^{(0)\to(2)} \: \mathcal{L}_W^{(2) \to (0)},
 \end{align*}
@@
 \begin{align*}
-	\int_0^{2\pi} g\big(\cos(\theta), \sin(\theta)\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
+	\int_0^{2\pi} g\big(\cos(\theta), \sin\theta\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
 \end{align*}
@@
 	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z\big) \: \D \theta \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
 	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \D x \Big) \: \D z \\
@@
 	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) + f_o\big(\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
@@
 %	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
+%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos(\theta)\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
 %	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_e\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \\
@@
 %	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos(\theta)\sin(\phi), \sin(\theta)\sin(\phi), \cos(\phi)\big) \: \genjacw^{(a,0)}\big(\cos(\phi)\big) \: \sin(\phi) \: \D \theta \: \D \phi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin(\theta), z)\big) \: \D \theta \Big) \: \D z \\
+%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos(\theta)\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
 %	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_o\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \quad (\dagger) \\
@@
 \end{align*}
-The Clenshaw algorithm is then \cite{}\bstodoinline{Reference for Clenshaw in >1 dimensions???} 
+The trivariate Clenshaw algorithm works similar to the bivariate Clenshaw algorithm introduced in \cite{olver2019triangle}:
 \begin{align*}
@@
 %
-\section{Examples on the disk-slice with zero Dirichlet conditions}\label{Section:Examples}
+\section{Examples on spherical caps with zero Dirichlet conditions}\label{Section:Examples}
 
@@
 Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and the same function $v$ (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
-This demonstrates spectral convergence observed.}
+This demonstrates spectral convergence observed.
+	\sotodoinline{Change $x$-axis to use normal (not logarithmic) scaling: this will change exponential convergence to more readable straight lines.
+For examples, having  nearby singularities (e.g.
+something like$f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$) should give different exponential convergence rates.}
+	\sotodoinline{Just write out $W$ in terms of $z$}
+	}
 	\centering
@@
 
-The Poisson equation is the classic problem of finding $u(x,y)$ given a function $f(x,y)$ such that:
+The Poisson equation is the classic problem of finding $u(x,y,z)$ given a function $f(x,y,z)$ such that:
 \begin{align}
 	\begin{cases}
-    		\nabla^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
+\DeltaS u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
 		u(x,y,z) = 0& \quad \text{on } \partial \Omega
@@
 In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
-In the plot, a "block" is simply the group of coefficients corresponding to OPs of the same degree, and so the plot shows how the norms of these blocks decay as the degree of the expansion increases.
+In the plot, a ``block'' is simply the group of coefficients corresponding to OPs of the same degree, and so the plot shows how the norms of these blocks decay as the degree of the expansion increases.
 Thus, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
@@
 	\begin{cases}
-    		\nabla^2 u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
+    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
 		u(x,y,z) = 0 &\quad \text{on } \partial \Omega
@@
 Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices\sotodo{It would be cleaner defining $X := (J_x^{(0)})^\top$, $Y := ...$, $Z := ...$} ${J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
 We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top)$.
@@
 	\begin{cases}
-    		\nabla^2 u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
+    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
 		u(x,y,z) = c \in \R &\quad \text{on } \partial \Omega
@@
 	\begin{cases}
-    		\nabla^2 \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - c \: k^2 \: v(x,y,z) \; =: g(x,y,z)  \quad \text{in } \Omega \\
+    		\DeltaS \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - c \: k^2 \: v(x,y,z) \; =: g(x,y,z)  \quad \text{in } \Omega \\
 		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
@@
 
-We have shown that tri-variate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary.
-This work provides a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: \sqrt{(1-z^2)}^b$ defined for the disk-slice case \cite{snowball2019sparse}.
-This work also forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.
+We have shown that trivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary.
+We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: \sqrt{(1-z^2)}^b$ defined for the disk-slice case \cite{snowball2019sparse}.
+This  also forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.
 
@@
 
-The next stage of this work is to develop an orthogonal basis for the tangent space of the spherical cap/band, and obtain sparse differential operators for gradient, divergence etc.
-On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
+The next stage is to develop an orthogonal basis for the tangent space of the spherical cap/band, and obtain sparse differential operators for gradient, divergence etc.
+On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
 
@@
 \bstodoinline{Is this worth including?}
+
+\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis}
 
@@
 	\begin{cases}
-         - \rho(z)^2 \: \nabla^2 u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
+         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
          u = 0 \quad \text{on } \partial \Omega
@@
 	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-	&= \int_\element \Big( \mathbf{v}^\top {W_\phi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\phi^{(1)} \mathbf{u} \nonumber \\
+	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
 					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-	&= \mathbf{v}^\top \: \Big( {W_\phi^{(1)}}^\top \Lambda^{(0)}  W_\phi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
+	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
 \end{align*}
@@
 \begin{align*}
-	A =  {W_\phi^{(1)}}^\top \Lambda^{(0)}  W_\phi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
+	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
 \end{align*}

### 2020-09-23 papers/spherical-caps/spherical-caps.tex
Updates to paper

@@
 								% TeX will automatically convert eps --> pdf in pdflatex
-\usepackage[margin=10pt,font=footnotesize,labelfont=bf]{caption}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{bbm}
-\usepackage{mleftright}
-\usepackage{todonotes}
-\usepackage{csquotes}
-
-%\bibliographystyle{apalike}
 
@@
 
-\newcommand{\sotodo}{\todo[color=green]}
-\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
-\newcommand{\bstodo}{\todo[color=pink]}
-\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}
-
-%SetFonts
-
-%SetFonts
-
-\usepackage[square,numbers]{natbib}
-\usepackage{url}
-%\bibliographystyle{elsarticle-harv}
-\bibliographystyle{plain}
-
-\newcommand{\bigO}{\mathcal{O}}
-\newcommand{\half}{\frac{1}{2}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\C}{\mathbb{C}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\No}{\mathbb{N}_0}
-\newcommand{\Ylm}{Y^m_l}
-\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
-\newcommand{\Plm}{\jac^m_l}
-\newcommand{\costheta}{\cos\theta}
-\newcommand{\sintheta}{\sin\theta}
-\newcommand{\cosphi}{\cos\varphi}
-\newcommand{\sinphi}{\sin\varphi}
-\newcommand{\eimphi}{e^{im\varphi}}
-\newcommand{\alphalm}{\alpha^m_l}
-\newcommand{\clm}{c^m_l}
-\newcommand{\ctilde}{\tilde{c}^m_l}
-\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
-\newcommand{\chat}{\hat{c}^m_l}
-\newcommand{\chatmod}{\hat{c}^{|m|}_l}
-\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
-\newcommand{\ddy}{\frac{\mathrm{d}}{\mathrm{d}y}}
-\newcommand{\pddx}{\frac{\partial}{\partial x}}
-\newcommand{\pddy}{\frac{\partial}{\partial y}}
-\newcommand{\pddn}{\frac{\partial}{\partial n}}
-\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}
-
-\newcommand{\Atilde}{\tilde{A}_{l,m}}
-\newcommand{\Btilde}{\tilde{B}_{l,m}}
-\newcommand{\Dtilde}{\tilde{D}_{l,m}}
-\newcommand{\Etilde}{\tilde{E}_{l,m}}
-\newcommand{\Ftilde}{\tilde{F}_{l,m}}
-\newcommand{\Gtilde}{\tilde{G}_{l,m}}
-\newcommand{\Alm}{A_{l,m}}
-\newcommand{\Blm}{B_{l,m}}
-\newcommand{\Dlm}{D_{l,m}}
-\newcommand{\Elm}{E_{l,m}}
-\newcommand{\Flm}{F_{l,m}}
-\newcommand{\Glm}{G_{l,m}}
-
-\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
-\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
-\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
-\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}
-
-\newcommand{\bigP}{\mathbb{P}}
-\newcommand{\Pl}{\mathbb{P}_l}
-\newcommand{\gradP}{T\mathbb{P}}
-\newcommand{\gradPl}{T\mathbb{P}_l}
-\newcommand{\gradY}{\nabla Y}
-\newcommand{\gradYlm}{\nabla Y^m_l}
-\newcommand{\gradpY}{\nabla^\perp Y}
-\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}
-
-\newcommand{\Dlt}{D^\top_l}
-
-\newcommand{\curlyy}{\bm{\mathcal{Y}}}
-\newcommand{\blone}{\beta_{l, 1}}
-\newcommand{\blzero}{\beta_{l, 0}}
-\newcommand{\blmone}{\beta_{l, -1}}
-\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
-\newcommand{\cgcoeff}{\mathcal{C}}
-
-\newcommand{\alm}{a_{l,m}}
-\newcommand{\blm}{b_{l,m}}
-\newcommand{\dlm}{d_{l,m}}
-\newcommand{\elm}{e_{l,m}}
-\newcommand{\flm}{f_{l,m}}
-\newcommand{\glm}{g_{l,m}}
-\newcommand{\hlm}{h_{l,m}}
-\newcommand{\jlm}{j_{l,m}}
-\newcommand{\klm}{k_{l,m}}
-\newcommand{\almperp}{a_{l,m}^\perp}
-\newcommand{\blmperp}{b_{l,m}^\perp}
-\newcommand{\dlmperp}{d_{l,m}^\perp}
-\newcommand{\elmperp}{e_{l,m}^\perp}
-\newcommand{\flmperp}{f_{l,m}^\perp}
-\newcommand{\glmperp}{g_{l,m}^\perp}
-\newcommand{\hlmperp}{h_{l,m}^\perp}
-\newcommand{\jlmperp}{j_{l,m}^\perp}
-\newcommand{\klmperp}{k_{l,m}^\perp}
-
-\newcommand{\unitvec}{\hat{\bm{k}}}
-
-\newcommand{\hdop}{H}
-\newcommand{\bighdop}{\mathbb{\hdop}}
-\newcommand{\hdopnk}{\hdop_{n,k}}
-\newcommand{\hdopnkab}{\hdop_{n,k}^{(a,b)}}
-\newcommand{\hdopab}{\hdop^{(a,b)}}
-\newcommand{\Wab}{{W^{(a,b)}}}
-\newcommand{\hdopmj}{\hdop_{m,j}}
-\newcommand{\hdopmjab}{\hdop_{m,j}^{(a,b)}}
-\newcommand{\alphaab}{\alpha^{(a,b)}}
-\newcommand{\betaab}{\beta^{(a,b)}}
-\newcommand{\bighdopab}{\bighdop^{(a,b)}}
-\newcommand{\Dnt}{D^\top_n}
-\newcommand{\Wii}{W^{(1,1)}}
-\newcommand{\hdopii}{\hdop^{(1,1)}}
-\newcommand{\bighdopii}{{\mathbb{\hdop}^{(1,1)}}}
-\newcommand{\hdopoo}{\hdop^{(0,0)}}
-\newcommand{\bighdopoo}{{\mathbb{\hdop}^{(0,0)}}}
-\newcommand{\hdopoi}{\hdop^{(0,1)}}
-\newcommand{\hdopio}{\hdop^{(1,0)}}
-%\newcommand{\jac}{{\tilde P}}
-\newcommand{\jac}{{P}}
-\newcommand{\genjac}{R}
-\newcommand{\genjacnmk}{\genjac_{n-k}}
-\newcommand{\genjacmmj}{\genjac_{m-j}}
-\newcommand{\genjacw}{w_\genjac}
-\newcommand{\jacw}{w_P}
-\newcommand{\normgenjac}{\omega_\genjac}
-\newcommand{\normjac}{\omega_P}
-
-\newcommand{\dx}{\frac{\partial}{\partial x}}
-\newcommand{\dy}{\frac{\partial}{\partial y}}
-\newcommand{\laplacewii}{\Delta_W^{(1,1)\to(1,1)}}
-\newcommand{\laplacewtt}{\Delta_W^{(2,2)\to(0,0)}}
-\newcommand{\laplaceoo}{\Delta^{(0,0)\to(2,2)}}
-\newcommand{\biharmonic}{_2\Delta_W^{(2,2)\to(2,2)}}
-
-\newcommand{\element}{\tau}
-\newcommand{\refelement}{\hat{\tau}}
-\newcommand{\FEset}{\mathcal{T}}
-\newcommand{\bigW}{\mathbb{W}}
-\newcommand{\bigWab}{\mathbb{W}^{(a,b)}}
-\newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
-\newcommand{\bigQ}{\mathbb{Q}}
-\newcommand{\bigQab}{\bigQ^{(a,b)}}
-
-\newcommand{\hdopnkabc}{\hdop_{n,k}^{(a,b,c)}}
-\newcommand{\hdopabc}{\hdop^{(a,b,c)}}
-\newcommand{\Wabc}{{W^{(a,b,c)}}}
-\newcommand{\hdopmjabc}{\hdop_{m,j}^{(a,b,c)}}
-\newcommand{\alphaabc}{\alpha^{(a,b,c)}}
-\newcommand{\betaabc}{\beta^{(a,b,c)}}
-\newcommand{\bighdopabc}{\bighdop^{(a,b,c)}}
-\newcommand{\Wiii}{W^{(1,1,1)}}
-\newcommand{\hdopiii}{\hdop^{(1,1,1)}}
-\newcommand{\bighdopiii}{{\mathbb{\hdop}^{(1,1,1)}}}
-\newcommand{\hdopooo}{\hdop^{(0,0,0)}}
-\newcommand{\bighdopooo}{{\mathbb{\hdop}^{(0,0,0)}}}
-\newcommand{\hdopooi}{\hdop^{(0,0,1)}}
-\newcommand{\hdopiio}{\hdop^{(1,1,0)}}
-\newcommand{\laplacewiii}{\Delta_W^{(1,1,1)\to(1,1,1)}}
-\newcommand{\laplacewttt}{\Delta_W^{(2,2,2)\to(0,0,0)}}
-\newcommand{\laplaceooo}{\Delta^{(0,0,0)\to(2,2,2)}}
-\newcommand{\biharmonictwo}{_2\Delta_W^{(2,2,2)\to(2,2,2)}}
-\newcommand{\bigWiii}{{\mathbb{W}^{(1,1,1)}}}
-\newcommand{\bigWabc}{\mathbb{W}^{(a,b,c)}}
-
-\newcommand{\alphaabcd}{\alpha^{(a,b,c,d)}}
-\newcommand{\betaabcd}{\beta^{(a,b,c,d)}}
-\newcommand{\hdopnkabcd}{\hdop_{n,k}^{(a,b,c,d)}}
-\newcommand{\hdopabcd}{\hdop^{(a,b,c,d)}}
-\newcommand{\Wabcd}{{W^{(a,b,c,d)}}}
-\newcommand{\hdopmjabcd}{\hdop_{m,j}^{(a,b,c,d)}}
-\newcommand{\bighdopabcd}{\bighdop^{(a,b,c,d)}}
-\newcommand{\bigWabcd}{\mathbb{W}^{(a,b,c,d)}}
-
-\newcommand{\genjact}{\tilde{\genjac}}
-\newcommand{\genjactnmk}{\genjact_{n-k}}
-\newcommand{\genjactmmj}{\genjact_{m-j}}
-\newcommand{\genjactw}{w_{\genjact}}
-\newcommand{\normgenjact}{\omega_{\genjact}}
-
-% Spherical Cap commands
-\newcommand{\scop}{Q}
-\newcommand{\scopnki}{\scop_{n,k,i}}
-\newcommand{\scopmjh}{\scop_{m,j,h}}
-\newcommand{\scopa}{\scop^{(a)}}
-\newcommand{\scopnkia}{\scopnki^{(a)}}
-\newcommand{\scopmjha}{\scopmjh^{(a)}}
-\newcommand{\bigscop}{{\mathbb{Q}}}
-\newcommand{\bigscopa}{\bigscop^{(a)}}
-\newcommand{\bigscopN}{\bigscop_{N}}
-\newcommand{\bigscopNa}{\bigscopa_{N}}
-\newcommand{\bigscopNka}{\bigscopa_{N,k}}
-\newcommand{\bigscopt}{\mathbb{\tilde{Q}}}
-\newcommand{\bigscopta}{\bigscopt^{(a)}}
-\newcommand{\bigscoptn}{\bigscopt_{n}}
-\newcommand{\bigscoptna}{\bigscopta_{n}}
-\newcommand{\xvec}{\mathbf{x}}
-\newcommand{\ch}{Y}
-\newcommand{\chki}{\ch_{k,i}}
-\newcommand{\chjh}{\ch_{j,h}}
-\newcommand{\chkitheta}{\ch_{k,i}(\theta)}
-\newcommand{\alphaa}{\alpha^{(a)}}
-\newcommand{\betaa}{\beta^{(a)}}
-\newcommand{\gammaa}{\gamma^{(a)}}
-\newcommand{\Wa}{W^{(a)}}
-\newcommand{\bigWNa}{\mathbb{W}_N^{(a)}}
-\newcommand{\ppphi}{\frac{\partial}{\partial \varphi}}
-\newcommand{\ppz}{\frac{\partial}{\partial z}}
-\newcommand{\rhoppphi}{\rho \ppphi}
-\newcommand{\rhozppphi}{\rho(z) \ppphi}
-\newcommand{\pptheta}{\frac{\partial}{\partial \theta}}
-\newcommand{\ppthetatwo}{\frac{\partial^2}{\partial \theta^2}}
-\newcommand{\bigWNi}{\mathbb{W}_N^{(1)}}
-\newcommand{\scopi}{\scop^{(1)}}
-\newcommand{\scopnkii}{\scopnki^{(1)}}
-\newcommand{\bigWNo}{\mathbb{W}_N^{(0)}}
-\newcommand{\scopo}{\scop^{(0)}}
-\newcommand{\scopnkio}{\scopnki^{(0)}}
-\newcommand{\bigscopi}{\bigscop^{(1)}}
-\newcommand{\bigscopNi}{\bigscopi_{N}}
-\newcommand{\bigscopNki}{\bigscopi_{N,k}}
-\newcommand{\bigscopo}{\bigscop^{(0)}}
-\newcommand{\bigscopNo}{\bigscopo_{N}}
-\newcommand{\bigscopNko}{\bigscopo_{N,k}}
-
-% Tangent space
-\newcommand{\tangentspace}{{\Omega_T}}
-\newcommand{\phivec}{\hat{\bm{\varphi}}}
-\newcommand{\thetavec}{\hat{\bm{\theta}}}
-\newcommand{\bigtsop}{\mathbb{T}}
-\newcommand{\bigtsopt}{\tilde{\bigtsop}}
-\newcommand{\bigtsoptn}{\bigtsopt_n}
-\newcommand{\bigtsopN}{\bigtsop_N}
-\newcommand{\bigtsopNk}{\bigtsop_{N,k}}
-\newcommand{\tsopi}{\bm{\Phi}}
-\newcommand{\tsopii}{\bm{\Psi}}
-\newcommand{\tsopiN}{\tsopi_N}
-\newcommand{\tsopiiN}{\tsopii_N}
-\newcommand{\tsopinki}{\tsopi_{n,k,i}}
-\newcommand{\tsopiinki}{\tsopii_{n,k,i}}
-\newcommand{\tsopinkit}{\tsopi_{n,k,i}^\top}
-\newcommand{\tsopiinkit}{\tsopii_{n,k,i}\top}
-\newcommand{\tsopit}{\tsopi^\top}
-\newcommand{\tsopiit}{\tsopii^\top}
-\newcommand{\rvec}{\hat{\bm{r}}}
-\newcommand{\jacobimattangent}{{}^T\!J}
-\newcommand{\jacobimattangentx}{\jacobimattangent_x}
-\newcommand{\jacobimattangenty}{\jacobimattangent_y}
-\newcommand{\jacobimattangentz}{\jacobimattangent_z}
-\newcommand{\alphao}{\alpha^{(0)}}
-\newcommand{\alphaonkj}{\alphao_{n,k,j}}
-\newcommand{\betao}{\beta^{(0)}}
-\newcommand{\betaonkj}{\betao_{n,k,j}}
-\newcommand{\gammao}{\gamma^{(0)}}
-\newcommand{\gammaonkj}{\gammao_{n,k,j}}
-\newcommand{\uvec}{\bm{u}}
-\newcommand{\hvec}{\bm{h}}
-\newcommand{\uvecc}{\uvec^c}
-\newcommand{\hvecc}{\hvec^c}
-\newcommand{\uvecn}{\uvec_n}
-\newcommand{\hvecn}{\hvec_n}
-\newcommand{\uvecnpi}{\uvec_{n+1}}
-\newcommand{\hvecnpi}{\hvec_{n+1}}
-\newcommand{\uvecnc}{\uvecc_n}
-\newcommand{\hvecnc}{\hvecc_n}
-\newcommand{\uvecnpic}{\uvecnpi^c}
-\newcommand{\hvecnpic}{\hvecnpi^c}
-\newcommand{\href}{\mathcal{H}}
-\newcommand{\divergence}{\nabla \cdot}
-\newcommand{\grad}{\nabla}
-
-\def\DeltaS{\Delta_{\rm S}}
-
-\usepackage{amsthm}
-
-\newtheorem{proposition}{Proposition}
-\newtheorem{lemma}{Lemma} 
-\newtheorem{theorem}{Theorem} 
-\newtheorem{definition}{Definition}
-\newtheorem{condition}{Condition}
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{corollary}{Corollary}
-\newtheorem*{remark}{Remark}
-
-\def\condref#1{Condition~\ref{cond:#1}}
-
+\include{bsthesisnewcommands}
+\include{bsmacros}
 \include{somacros}
-
-%\usepackage{media9}
-%\graphicspath{ {../sphere/plots} }
 
@@
 \end{align*}
-and consider the case where \sotodo{Why is $\alpha$ restricted to be positive?} $\alpha \in (0,1)$ and where
-$ \rho(z) := \sqrt{1-z^2}$,  $\beta := 1$ \sotodo{Don't we need $\beta \neq 1$ to get a band?}.
+and consider the case where $\alpha \in (-1,1)$ and where $ \rho(z) := \sqrt{1-z^2}$,  $\beta := 1$ \sotodoinline{Don't we need $\beta \neq 1$ to get a band?} \bstodoinline{yes, but to keep it simple this paper focuses on the cap (otherwise the maths would become more complex/longer.
+I state that the band is an extension below.}.
 This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
@@
 \end{align*}
-Of course, there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$. However for simplicity we focus on the spherical cap in this paper.
+Of course, there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$. However, for simplicity, we focus on the spherical cap in this paper.
 
@@
 The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
-We define these to be orthogonalised lexicographically, that is, \sotodo{I don't understand this ordering.
+We define these to be orthogonalised lexicographically, that is, \sotodoinline{I don't understand this ordering.
 Shouldn't it either be $x^k y^j z^{n-k-j}$ or $x^k y^{j-k} z^{n-j}$?
-PS I prefer $j$ to $i$ }
-\begin{align*}
-	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k - \tilde k_i} \: y^{\tilde k_i} \: z^{n-k} + (\hbox{lower order terms})
-\end{align*} \bstodoinline{How can I write down what the ordering is? Is this necessary?}
-where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - \tilde j_i} \: y^{\tilde j_i} \: z^{n-j}$ where $j < k$.
+PS I prefer $j$ to $i$ } \bstodoinline{I have changed to the below.
+I will change i -> j throughout later if we prefer this.}
+\begin{align*}
+	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms})
+\end{align*}
+where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - i} \: y^{i} \: z^{n-j}$ where $j < k$.
 The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
@@
 \end{align*}
-we have \sotodo{Double check}
-\begin{align*}
-\DeltaS &= {1 \over \sin \varphi} {\partial \over \partial \varphi} \sin \varphi {\partial \over \partial \varphi} + {1 \over \sin^2 \varphi} {\partial^2 \over \partial \theta^2}  = {1 \over \rho} {\partial \over \partial \varphi} \rho {\partial \over \partial \varphi} + {1 \over \rho^2} {\partial^2 \over \partial \theta^2} \\
-\end{align*}
+we have
+\begin{align*}
+	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
+\end{align*}
+i.e.
+$\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$.
 We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.
@@
 
- The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical\sotodo{Add citation}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
+ The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
 In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 This improves the complexity further to an optimal $O(p^2$).
-\sotodo{Adding a complexity plot showing this would be great}
+\sotodoinline{Adding a complexity plot showing this would be great} \bstodoinline{I am not sure how to do a complexity plot}
 
@@
 
-\begin{definition}
+\begin{definition}\label{def:Ydefinition}
 	Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
 Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
-\begin{align}
-	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \nonumber \\
-	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots\label{eqn:chebyshev} \\
-	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots \nonumber
-\end{align}
-where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$.
-The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product\sotodo{The $\sigma$ notation needs explanation}
-\begin{align}
-	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_{||\xvec|| = 1} p(\xvec) \: q(\xvec) \: \D \sigma(\xvec) = \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
-\label{eqn:ipcircularharmonics}
-\end{align}
+\begin{align*}
+	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
+	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots \\
+	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots
+\end{align*}
+where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
+\begin{align*}
+	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta
+\end{align*}
 \end{definition}
@@
 
-\sotodoinline{Isn't the following from \cite{olver2020orthogonal}? We then don't need to repeat the proof}
-
-\begin{proposition}\label{prop:construction}
+\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
 	Let $w : (\alpha,\beta) \: \to \R$ be a weight function.
@@
 \end{proposition}
-
-\begin{proof}
-\begin{align*}
-	\ip< \scopnki, \: \scopmjh > &= \int_\alpha^1 \int_0^{2\pi} r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}(\theta) \: r_{m-j,j}(z) \: \rho(z)^j \: \ch_{j,h}(\theta) \: \D \theta \: w(z) \: \D z \\
-	&= \Big(\int_0^{2\pi}  \: \ch_{k,i}(\theta) \: \: \ch_{j,h}(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: r_{n-k,k}(z) \: \rho(z)^k \: r_{m-j,j}(z) \: \rho(z)^j \: w(z) \: \D z \Big) \\
-	&= \pi \: \delta_{k,j} \: \delta_{i,h} \:  \int_\alpha^1 \: r_{n-k,k}(z) \: r_{m-k,k}(z) \: \rho(z)^{2k} \: w(z) \: \D z \\
-	&= C \: \pi \: \delta_{k,j} \: \delta_{i,h}
-\end{align*}
-for some constant $C$.
-\end{proof}
 
@@
 \begin{align}
-	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega, 
+	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega.
 \end{align}
@@
 By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
-\begin{align}
+\begin{align*}
 	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D \sigma(\xvec, z) \\
 	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
-\end{align}
-with
+\end{align*}
+where $\D \sigma(\xvec, z)$ is the Lebesgue measure on $\Omega$, with
 \begin{align}
-	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)} \label{eqn:normscop}
+	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:normscop}
 \end{align}
@@
 
-In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated\sotodo{Explain what a Clenshaw matrix is} \enquote{Clenshaw matrices}.
-While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
+In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices} -- we define the Clenshaw matrices as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the multidimensional three-term recurrences \bstodoinline{added explanation of clenshaw matrix in this context} While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
@@
 
-Note that we can define an explicit \(\Dnt\) as follows:\sotodo{Improve following notation, say, by adding notation to the top part}
+Note that we can define an explicit \(\Dnt\) as follows:\sotodoinline{Improve following notation, say, by adding notation to the top part} \bstodoinline{is this any better?}
 \begin{align*}
@@
 		\begin{pmatrix}
-			0 & & & 0 & & & 1 / \gammaa_{n,0,3} & & \\
+			0 & & & 0 & & & (\Gamma^{(a)}_{n,0,3})^{-1} & & \\
 			& \ddots & & & \ddots & & & \ddots \\
-			& & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
+			& & 0 & & & 0 & & & (\Gamma^{(a)}_{n,n,3})^{-1}  \\
 			& & & & & \bm{\eta}^\top_{0} & & & \\
@@
 \end{align*}
-for $n = 1, 2, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
+for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}) for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by
 \begin{align*}
@@
 In this section, we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process.
-To this end, we will use a different ordering\sotodo{I thought we would discuss this separately as a permutation of rows/columns?} of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
+To this end, we will use a different ordering\sotodoinline{I thought we would discuss this separately as a permutation of rows/columns?} \bstodoinline{Not sure what you mean by this comment - maybe I misunderstood before?} of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
 Let $N \in \N$ and recall from equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) that:
@@
 
+\bstodoinline{I have added defns (and sparsity proofs) for two other Laplacian operators.
+In practice, I cannot get the non-weighted one to work coded up, which I assume is a bug on my end and not the maths (which to me is correct).}
 \begin{definition}\label{def:differentialoperators}
-Define the operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ according to:
-\begin{align*}
-	\rho {\partial f \over \partial \varphi}(x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \mathbf{f}, \\
-	\rho {\partial \over \partial \varphi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\varphi^{(a)}\: \mathbf{f}, \\
-	{\partial f \over \partial \theta}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
+	Let $a$ be a nonnegative parameter, and $\tilde a \ge 2$ be a positive integer.
+Define the operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ according to:
+\begin{align*}
+	\rho \pfpx{\varphi}{f} (x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \mathbf{f}, \\
+	\rho \ppx{\varphi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\varphi^{(a)}\: \mathbf{f}, \\
+	\pfpx{\theta}{f}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
+	\DeltaS f(x,y,z) &= \bigscopN^{(a+\tilde a)}(x,y,z)^\top \: \calL^{(a)\to(a+\tilde a)} \: \mathbf{f}, \\
+	\DeltaS \big( \genjacw^{(a,0)}(z) \: f(x,y,z) \big) &= \bigW_N^{(a-\tilde a)}(x,y,z)^\top \: \calL_W^{(a)\to(a-\tilde a)} \: \mathbf{f},  \quad \text{ (for } a \ge 2 \text{ only)} \\
 	\DeltaS \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
@@
 	\item $D_\theta$ has  block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
-	\item $\Delta^{(1)}_W$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
+	\item $\calL^{(a)\to(a+\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(0, 4)$
+	\item $\calL_W^{(a)\to(a-\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 0)$
+	\item $\Delta^{(1)}_W$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 2)$
 \end{itemize}
@@
 \begin{align*}
-	&\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
+	&\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
 	&\quad \quad = \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
@@
 \end{lemma}
+
 \begin{proof}[Proof of Lemma \ref{lemma:Rsecondderivative}]
-Since $\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] = \genjacw^{(a, 2k)} \: r_{n-k+1}$ where $r_{n-k+1}$ is a degree $n - k + 1$ polynomial, we have that 
-\begin{align*}
-	\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)}
+	Since $\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] = \genjacw^{(a, 2k)} \: r_{n-k+1}$ where $r_{n-k+1}$ is a degree $n - k + 1$ polynomial, we have that 
+\begin{align*}
+	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)}
 \end{align*}
@@
 \begin{align*}
-	\tilde c_{\{n,k\},m} &= \frac{1}{\normgenjac^{(a,2k)}} \: \ip<\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
+	\tilde c_{\{n,k\},m} &= \frac{1}{\normgenjac^{(a,2k)}} \: \ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
 	&= - \frac{1}{\normgenjac^{(a,2k)}}  \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
@@
 \begin{align*}
-	&\ip<\ppz[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
+	&\ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
 	&\quad \quad \quad = - \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z \\
@@
 \begin{proof}[Proof of Theorem \ref{theorem:sparsityofdifferentialoperators}]
-First, note that \sotodo{Here and elsewhere: ${\partial \over \partial z}$ is not well-defined as we are on the sphere.
-Though perhaps ${\rm d} \over {\rm d} z$ works as we are differentiating the formula with respect to $z$?}:
-\begin{align}
-	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \label{eqn:derivativeofweightgenjac} \\
-	\rho(z) \: \rho'(z) &= -z \label{eqn:rhoderivative} \\
-	{\partial \over \partial \varphi} &= -\rho \: {\partial \over \partial z} \label{eqn:partialphitoz} 
-\end{align}
-
-For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that \sotodo{Would it be better as ${\rm d} \over {\rm d} \theta$?}
-\begin{align*}
-	{\partial \over \partial \theta}\scopnkia(x,y,z) &= \genjacnmk^{(a,2k)}(z) \: \rho(z)^{k} \: {\partial \over \partial \theta} \chki(\theta) \\
+
+For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that \sotodoinline{Would it be better as ${\rm d} \over {\rm d} \theta$?} \bstodoinline{fixed}
+\begin{align*}
+	\pptheta \scopnkia(x,y,z) &= \genjacnmk^{(a,2k)}(z) \: \rho(z)^{k} \: \ddx{\theta} \chki(\theta) \\
 	&= 
@@
 \end{align*}
-Now,
+Now, note that \sotodoinline{Here and elsewhere: ${\partial \over \partial z}$ is not well-defined as we are on the sphere.
+Though perhaps ${\rm d} \over {\rm d} z$ works as we are differentiating the formula with respect to $z$?} \bstodoinline{Is the following clearer?}:
+\begin{align*}
+	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \\
+	\rho(z) \: \rho'(z) &= -z \\
+	\ppphi \scopnkia(x,y,z) &= -\rho(z) \: \ddz \Big[ \rho(z)^k \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta), \\
+	\ppphi \Big[ \genjacw^{(a,0)}(z) \: \scopnkia(x,y,z) \Big] &= -\rho(z) \: \ddz \Big[ \genjacw^{(a,k)}(z) \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta).
+\end{align*}
+Then, 
 \begin{align*}
 	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
-	&\quad = - \int_\Omega \: \rho(z)^2 \ppz \:  [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \genjacw^{(a+1, 0)} \: \D \Omega \\
+	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \genjacw^{(a+1, 0)} \: \D z \\
 	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
@@
 	&\ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}} \\
-	&\quad = - \int_\Omega \: \rho(z)^2 \ppz \:  [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \genjacw^{(a, 0)}(z)] \: \genjacmmj^{(a-1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \Omega \\
+	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \genjacw^{(a, k)}(z)] \: \genjacmmj^{(a-1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \D z \\
 	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
@@
 
+We move onto the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yields
+\begin{align}
+	\DeltaS \: \scopnki^{(a)} &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
+	&\quad \quad \quad \quad + \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
+	&= \chki(\theta) \rho(z)^k \Big\{ -k(k+1) \genjacnmk^{(a,2k)}(z) - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime}(z) \nonumber \\
+	&\quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime}(z) \Big\}, \label{eqn:oplaplacian} \\
+	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
+	&\quad \quad \quad \quad + \genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
+	&= \chki(\theta) \Big\{\genjacnmk^{(a,2k)}(z) \big[ -k(k+1) \genjacw^{(a,k)}(z) - 2a(k+1) z \: \genjacw^{(a-1,k)}(z) \big] \nonumber \\
+	&\quad \quad \quad \quad \quad + a(a-1) \genjacnmk^{(a,2k)}(z) \: \genjacw^{(a-2,k+1)}(z) \nonumber \\
+	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime}(z) \big[ -2(k+1) z \: \genjacw^{(a,k)}(z) + 2a \: \genjacw^{(a-1,k+2)}(z) \big] \nonumber \\
+	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime}(z) \: \genjacw^{(a,k+2)}(z) \Big\}. \label{eqn:woplaplacian}
+\end{align}
+
+For the operator $\calL^{(a)\to(a+\tilde a)}$ for the surface Laplacian on a non-weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \scopnki^{(a)} = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a + \tilde a)}$, where the coefficients are
+\begin{align*}
+	c_{m,j,h} = \norm{\scopmjh^{(a + \tilde a)}}^{-2}_{\scop^{(a + \tilde a)}} \ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}}.
+\end{align*}
+Using \bsrefeqn{eqn:oplaplacian}, and integrating by parts twice, we then have that
+\begin{align*}
+	&\ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}} \\
+	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a + \tilde a, 2j)} \: \genjacw^{(a + \tilde a), k+j)} \: \Big\{-k(k+1) \genjacnmk^{(a,2k)} - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime} \Big\} \: \D z \Big) \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a+\tilde a, 0)} \Big( -k(k+1) \genjacnmk^{(a,2k)} \: \rho^{2k} + \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \rho^{2(k+1)}] \Big) \: \D z \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ -k(k+1) \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a+\tilde a, 2k)} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - \genjacnmk^{(a,2k) \: \prime} \genjacw^{(a+\tilde a-1, 2k)} \: [\genjac_{m-k}^{(a+\tilde a,2k) \: \prime} \: \genjacw^{(1,0)} + (a + \tilde a) \genjac_{m-k}^{(a+\tilde a,2k)}] \Big\} \: \D z \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z
+\end{align*}
+where $r_{m-k+\tilde a}$ is a degree $m - k + \tilde a$ polynomial in $z$, and so the above is zero for $n  - k > m - k + \tilde a \iff m < n - \tilde a$.
+
+For the operator $\calL_W^{(a)\to(a-\tilde a)}$ for the surface Laplacian on a weighted space, the entries of the operator are given by the coefficients in the expansion 
+\begin{align*}
+	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-\tilde a,0)} \: \scopmjh^{(a - \tilde a)}, 
+\end{align*}	
+where the coefficients are
+\begin{align*}
+	c_{m,j,h} = \norm{\scopmjh^{(a - \tilde a)}}^{-2}_{\scop^{(a - \tilde a)}} \ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}}.
+\end{align*}
+Using \bsrefeqn{eqn:woplaplacian}, and integrating by parts thrice, we then have that
+\begin{align*}
+	&\ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}} \\
+	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a - \tilde a, 2j)} \: \genjacw^{(a - 2, k+j)} \: \Big\{\genjacnmk^{(a,2k)} [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} [-2(k+1) z \: \genjacw^{(2,0)} + 2a \genjacw^{(1,2)}] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime} \: \genjacw^{(2,2)} \Big\} \: \D z \Big) \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \genjacw^{(a, 2k+2)}] \Big\} \: \D z \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} ] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a [\genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a-1, 2k+2)} \Big\} \: \D z \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} ] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - a \genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a-2, 2k)} \: [(a-1) \rho^2 - 2(k+1) z \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z
+\end{align*}
+where $r_{m-k}$ is a degree $m - k$ polynomial in $z$, and so the above is zero for $n - k > m - k \iff m < n$.
+
 Finally, fix $a = 1$.
@@
 \end{align*}
-Now, note that the Laplacian acting on the weighted spherical cap OP $\scopnki^{(a)}$ yields
-\begin{align*}
-	\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) &= 
-\end{align*}
-Hence,
+Using \bsrefeqn{eqn:woplaplacian} with $a = 1$, and \bsreflemma{lemma:Rsecondderivative}, we then have that
 \begin{align*}
@@
 \end{align*}
-where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in Lemma \ref{lemma:Rsecondderivative}.
+where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in \bsreflemma{lemma:Rsecondderivative}.
 \end{proof}
@@
 	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-biharmonic}
+		\includegraphics[scale=0.23]{sparsity-of-biharmonic}
 		\centering
-		\caption{The Biharmonic operator $B_W^{(2)}$ \sotodoinline{This one is surprising as I was expecting rotational invariance to give us block-diagonal operators}}
+		\caption{The Biharmonic operator $B_W^{(2)}$} 
 	\end{subfigure}\hfil % <-- added
@@
 
-Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions \sotodo{We need to explain Neumann conditions precisely, as the normal derivative is not so obvious}.
-To construct a matrix operator, we first note that we can obtain the matrix operator for the $\rho^2$-factored Laplacian $\rho(z)^2 \: \DeltaS$ that will take us from coefficients for expansion in the space $\bigscopN^{(0)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
-We denote this matrix operator that acts on the coefficients vector as $\mathcal{L}^{(0) \to (2)}$, and it is given by
-\begin{align*}
-    \mathcal{L}^{(0) \to (2)} := D_\varphi^{(2)} \: D_\varphi^{(0)} + T^{(0)\to(2)} \: D_\theta.
-\end{align*}
-Further, we can represent this same Laplacian as a map from coefficients in the space $\bigW_N^{(2)}$ to coefficients in the space $\bigW_N^{(0)} \equiv \bigscopN^{(0)}$.
-Note that a function expanded in the $\bigW_N^{(2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
-We denote this matrix operator as $\mathcal{L}_W^{(2) \to (0)}$, and is given by
-\begin{align*}
-	\mathcal{L}_W^{(2) \to (0)} := W_\varphi^{(1)} \: W_\varphi^{(2)} + T_W^{(2)\to(0,0,0)} \: D_\theta.
-\end{align*}
-Since, for any scalar function $f:\Omega \to \R$, \sotodo{The extra $\rho^2$ in front is weird to me}
-\begin{align*}
-	\rho^2(z) \: \Delta^2 \: f(x,y,z) = \rho^2 \: \DeltaS (\rho^2 \: \DeltaS f) - z \: \rhoppphi (\rho^2 \: \DeltaS f) - [2z^2 - 2\rho^2 + 3z] \:\rho^2 \: \DeltaS f
-\end{align*}
-the matrix operator for the $\rho^2$-factored Biharmonic operator $\rho(z)^2 \: \DeltaS^2$ is then given by
-\begin{align*}
-	\mathcal{B}_W^{(2)} := \mathcal{L}^{(0) \to (2)} \: \mathcal{L}_W^{(2) \to (0)} - (J_z^{(2)})^\top \: T^{(1)\to(2)} \: D_\varphi^{(0)} \: \mathcal{L}_W^{(2) \to (0)} - V^{(2)} \:  T^{(0)\to(2)} \: \mathcal{L}_W^{(2) \to (0)},
-\end{align*}
-where $V^{(a)}$ is an operator corresponding to multiplication by the scalar polynomial $v(z) := 2z^2 - 2\rho(z)^2 + 3z = 4z^2 + 3z - 2$ acting on coefficients in the $\bigscopN^{(a)}$ space (see Section \ref{subsection:operatorclenshaw} for how to computationally obtain operators representing functional coefficients).
+Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions \sotodoinline{We need to explain Neumann conditions precisely, as the normal derivative is not so obvious}\bstodoinline{Is the following correct?
+As in, isnt the normal at the boundary still just the unit outward normal on the sphere?}.
+That is,
+\begin{align*}
+	u(x,y,z) &= 0 \quad \text{for } (x,y,z) \in \Omega, \\
+	\pfpx{n}{u}(x,y,z) &= \nabla_S u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
+\end{align*}
+where $\partial\Omega$ is the $z=\alpha$ boundary, and $\unitvec{n}(x,y,z)$ is the outward unit normal vector at the point $(x,y,z)$ on the boundary, i.e.
+$\unitvec{n}(\vec{x}) = {\vec{x} \over \norm{\vec{x}}}$.
+The matrix operator for the Biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
+To construct this, we can simply multiply together two of the spherical Laplacian operators defined in \bsrefdef{def:differentialoperators}, namely $\calL^{(0)\to(2)}$ and $\calL_W^{(2)\to(0)}$:
+\begin{align*}
+	\calB^{(2)}_W := \calL^{(0)\to(2)} \: \calL_W^{(2)\to(0)}.
+\end{align*}
+Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions.
 The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
@@
 We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$.
-This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodoinline{correct the O(nd) complexity}.
+This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodo{correct the O(nd) complexity}.
 
@@
 To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in Section \ref{subsection:functionevaluation} so that $v(x,y,z) = \tilde{\bm{v}}^\top \: \bigscopta(x,y,z)$).
+Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, $Z := (J_z^{(a)})^\top$.
 The operator $V$ is then the result of the following:
@@
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(J^{(a)}_x, J^{(a)}_y, J^{(a)}_z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad &\text{3) } \text{Output: } V(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
 \end{align*}
@@
 	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-poisson-N=100}
+	\includegraphics[scale=0.3]{solutionblocknorms-poisson-N=100-xnolog}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-N=100}
+	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-N=100-xnolog}
         	%\label{fig:solutionblocknorms-poisson}
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
+	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
 Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and the same function $v$ (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
@@
 something like$f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$) should give different exponential convergence rates.}
-	\sotodoinline{Just write out $W$ in terms of $z$}
+	\bstodoinline{to do}
 	}
@@
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
+	\caption{\bstodoinline{update when biharmonic is fixed} Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
 Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
@@
 Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices\sotodo{It would be cleaner defining $X := (J_x^{(0)})^\top$, $Y := ...$, $Z := ...$} ${J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
-We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top)$.
+
+Define $X := (J_x^{(0)})^\top$, $Y := (J_y^{(0)})^\top$, $Z := (J_z^{(0)})^\top$.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
+We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $V(X, Y, Z)$.
 In matrix-vector notation, our system hence becomes:
@@
 %
-\appendix
-%
-\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
-
-\bstodoinline{Is this worth including?}
-
-\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis}
-
-%\begin{figure}[t]
-%	\begin{subfigure}{0.3\textwidth}
-%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-%        	%\label{fig:solutionblocknorms-poisson}
-%	\centering
-%	\end{subfigure}
-%	\begin{subfigure}{0.3\textwidth}
-%	\includegraphics[scale=0.3]{solution-poisson}
-%	\centering
-%	%\label{fig:solution-poisson}
-%	\end{subfigure}
-%	\begin{subfigure}{0.3\textwidth}
-%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-%	%\label{fig:solution-poisson}
-%	\centering
-%	\end{subfigure}
-%	\centering
-%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
+%%
+%\appendix
+%%
+%\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
+%
+%\bstodoinline{Is this worth including?}
+%
+%\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis} \bstodoinline{Yes thought so too}
+%
+%%\begin{figure}[t]
+%%	\begin{subfigure}{0.3\textwidth}
+%%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
+%%        	%\label{fig:solutionblocknorms-poisson}
+%%	\centering
+%%	\end{subfigure}
+%%	\begin{subfigure}{0.3\textwidth}
+%%	\includegraphics[scale=0.3]{solution-poisson}
+%%	\centering
+%%	%\label{fig:solution-poisson}
+%%	\end{subfigure}
+%%	\begin{subfigure}{0.3\textwidth}
+%%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
+%%	%\label{fig:solution-poisson}
+%%	\centering
+%%	\end{subfigure}
+%%	\centering
+%%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
 Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
@@
 in the trapezium.}
-%	\centering
-%	\label{fig:appendixfigs}
-%\end{figure}
-
-We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
+%%	\centering
+%%	\label{fig:appendixfigs}
+%%\end{figure}
+%
+%We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
 Consider the 3D Dirichlet problem on a domain $\Omega$:
-\begin{align*}
-	\begin{cases}
-         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
-         u = 0 \quad \text{on } \partial \Omega
-         \end{cases}
-\end{align*}
-This has the weak formulation
-\begin{align*}
-	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
-\end{align*}
- for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
-
-In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$.
+%\begin{align*}
+%	\begin{cases}
+%         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
+%         u = 0 \quad \text{on } \partial \Omega
+%         \end{cases}
+%\end{align*}
+%This has the weak formulation
+%\begin{align*}
+%	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
+%\end{align*}
+% for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
+%
+%In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$.
 While spherical bands are an extension to this work, we note that we can still apply a $p$-FEM using two half-spheres ($\alpha = 0$).
@@
 We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p \rvert_\element) \le p\}$ for some $p \in \N$.
-
-We seek $u_p \in V_p$ s.t.
-\begin{align}
-	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-	\label{eqn:FEMweakform}
-\end{align}
-
-Define 
-\begin{align*}
-	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
-\end{align*}
-for any $a$.
+%
+%We seek $u_p \in V_p$ s.t.
+%\begin{align}
+%	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
+%	\label{eqn:FEMweakform}
+%\end{align}
+%
+%Define 
+%\begin{align*}
+%	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
+%\end{align*}
+%for any $a$.
 Note that due to orthogonality this is a diagonal matrix.
@@
 We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.
-
-Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
-\begin{align*}
-	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
-	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
-\end{align*}
-Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
+%
+%Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
+%\begin{align*}
+%	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
+%	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
+%	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
+%					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
+%	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
+%\end{align*}
+%Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
 Then,
-\begin{align*}
-	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
-	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
-	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
-\end{align*}
-
-Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-\begin{align*}
-	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
-\end{align*}
-(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-\begin{align*}
-	A\mathbf{u} = \tilde{\mathbf{f}}.
-\end{align*}
-where the (element) stiffness matrix $A$ is defined by 
-\begin{align*}
-	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
-\end{align*}
-and the load vector $\tilde{\mathbf{f}}$ is given by 
-\begin{align*}
-	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
-\end{align*}
-Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+%\begin{align*}
+%	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
+%	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
+%	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
+%\end{align*}
+%
+%Since (\ref{eqn:FEMweakform}) is equivalent to stating that
+%\begin{align*}
+%	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
+%\end{align*}
+%(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
+%\begin{align*}
+%	A\mathbf{u} = \tilde{\mathbf{f}}.
+%\end{align*}
+%where the (element) stiffness matrix $A$ is defined by 
+%\begin{align*}
+%	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
+%\end{align*}
+%and the load vector $\tilde{\mathbf{f}}$ is given by 
+%\begin{align*}
+%	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
+%\end{align*}
+%Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+%
 

### 2020-09-23 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
Further edits to laplacian operator method. Need to fix L0to2 case.

@@ -1665,35 +1665,49 @@ function getlaplacianval(S::SphericalCapSpace{<:Any, B, T, <:Any},
                             ptsr::Vector{B}, wr::Vector{B}, rhoptsr2::Vector{B},
                             w10ptsr::Vector{B}, n::Int, m::Int, k::Int;
                             weighted::Bool=false, specialcase::Bool=false) where {B,T}
+    # specialcase == true if S.a == 1 == St.a and S is a weighted space
+
     a = Int(S.params[1])
-    if specialcase && weighted
+    rhoptsr2k = rhoptsr2.^k
+    rhoptsr2kp2 = rhoptsr2k .* rhoptsr2
+    if specialcase
+        if !weighted
+            error("specialcase true only if weighted true")
+        end
         R = getRspace(S, k)
-        ret = - inner2(R, getptsevalforop(R, n-k) .* getptsevalforop(R, m-k) .* rhoptsr2.^k,
+        ret = - inner2(R, getptsevalforop(R, n-k) .* getptsevalforop(R, m-k) .* rhoptsr2k,
                         (2(k+1) * ptsr .+ k*(k+1) * w10ptsr),
                         wr)
         ret -= inner2(R, w10ptsr .* getderivptsevalforop(R, n-k),
-                        getderivptsevalforop(R, m-k) .* rhoptsr2.^(k+1), wr)
+                        getderivptsevalforop(R, m-k) .* rhoptsr2kp2,
+                        wr)
         ret / getopnorm(R)
     elseif weighted
+        ã = a - Int(St.params[1])
         R = getRspace(S, k); Rt = getRspace(St, k)
         ret = inner2(Rt, getptsevalforop(R, n-k) .* getptsevalforop(Rt, m-k),
-                        (-k^2 * w10ptsr.^2
-                            .- 2 * a * (k+1) * ptsr .* w10ptsr
-                            .+ a * (a-1) * rhoptsr2) .* rhoptsr2.^k,
+                        (-k * (k+1) * w10ptsr.^ã
+                            - 2 * a * (k+1) * ptsr .* w10ptsr.^(ã-1)
+                            + a * (a-1) * rhoptsr2 .* w10ptsr.^(ã-2)) .* rhoptsr2k,
                         wr)
-        ret += inner2(Rt, a * getderivptsevalforop(R, n-k) .* w10ptsr,
[…]

### 2020-09-25 papers/spherical-caps/spherical-caps.tex
New figs
updated comments

@@
 \bstodoinline{I have added defns (and sparsity proofs) for two other Laplacian operators.
-In practice, I cannot get the non-weighted one to work coded up, which I assume is a bug on my end and not the maths (which to me is correct).}
+In practice, I have found that the non-weighted one is singular - is this correct/expected?
+The weighted one and the resulting Biharmonic operators both work (as far as I have tested).}
 \begin{definition}\label{def:differentialoperators}
@@
 This demonstrates spectral convergence observed.
-	\sotodoinline{Change $x$-axis to use normal (not logarithmic) scaling: this will change exponential convergence to more readable straight lines.
-For examples, having  nearby singularities (e.g.
-something like$f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$) should give different exponential convergence rates.}
-	\bstodoinline{to do}
+	\sotodoinline{For examples, having  nearby singularities (e.g. something like$f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$) should give different exponential convergence rates.}
+	\bstodoinline{New examples, such as the suggested $f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$.
+Though these look like I should remove the trivial examples entirely, as it ruins the scaling.}
 	}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{biharmonic-f=erf-N=80-n=50}
+	\includegraphics[scale=0.15]{biharmonic-f=erf-N=80-n=50}
 	\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-N=100}
+	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-N=100-xnolog}
 	\centering
@@
 	\end{subfigure}
-	\caption{\bstodoinline{update when biharmonic is fixed} Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
+	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
 Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
@@
 We see that we achieve spectral convergence \bstodo{Add to the comments here}
-
-\bstodoinline{Are there more suitable RHSs for the block norms plot that will demonstrate further properties? Or as there are no corners, we are pretty limited?}
 

### 2020-09-29 papers/spherical-caps/spherical-caps.tex
Remove completed todos

@@
 \end{align*}
-and consider the case where $\alpha \in (-1,1)$ and where $ \rho(z) := \sqrt{1-z^2}$,  $\beta := 1$ \sotodoinline{Don't we need $\beta \neq 1$ to get a band?} \bstodoinline{yes, but to keep it simple this paper focuses on the cap (otherwise the maths would become more complex/longer.
-I state that the band is an extension below.}.
+and consider the case where $\alpha \in (-1,1)$ and where $ \rho(z) := \sqrt{1-z^2}$.
 This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
 \begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < 1, \: x^2 + y^2 + z^2 = 1\}
-\end{align*}
-Of course, there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$. However, for simplicity, we focus on the spherical cap in this paper.
+	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
+\end{align*}
+For simplicity we focus on a spherical cap with   $\beta := 1$, though there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$.
 
@@
 The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
-We define these to be orthogonalised lexicographically, that is, \sotodoinline{I don't understand this ordering.
-Shouldn't it either be $x^k y^j z^{n-k-j}$ or $x^k y^{j-k} z^{n-j}$?
-PS I prefer $j$ to $i$ } \bstodoinline{I have changed to the below.
-I will change i -> j throughout later if we prefer this.}
+We define these to be orthogonalised lexicographically, that is,
 \begin{align*}
@@
 
-Note that we can define an explicit \(\Dnt\) as follows:\sotodoinline{Improve following notation, say, by adding notation to the top part} \bstodoinline{is this any better?}
+Note that we can define an explicit \(\Dnt\) as follows:
 \begin{align*}
@@
 It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigscopta_{0}(x,y,z)$  through $\bigscopta_{n}(x,y,z)$ in optimal $O(n^2)$ complexity.
-\bstodoinline{I think this is correct for this case}
 
@@
 
-In this section, we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process.
-To this end, we will use a different ordering\sotodoinline{I thought we would discuss this separately as a permutation of rows/columns?} \bstodoinline{Not sure what you mean by this comment - maybe I misunderstood before?} of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
+In this section, we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process. To this end, we will use a different ordering
+ of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
 Let $N \in \N$ and recall from equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) that:
@@
 
-For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that \sotodoinline{Would it be better as ${\rm d} \over {\rm d} \theta$?} \bstodoinline{fixed}
+For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that
 \begin{align*}
@@
 \end{align*}
-Now, note that \sotodoinline{Here and elsewhere: ${\partial \over \partial z}$ is not well-defined as we are on the sphere.
-Though perhaps ${\rm d} \over {\rm d} z$ works as we are differentiating the formula with respect to $z$?} \bstodoinline{Is the following clearer?}:
+Now, note that:
 \begin{align*}
@@
 
-Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions \sotodoinline{We need to explain Neumann conditions precisely, as the normal derivative is not so obvious}\bstodoinline{Is the following correct?
-As in, isnt the normal at the boundary still just the unit outward normal on the sphere?}.
-That is,
+Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions. That is, 
 \begin{align*}

### 2020-10-01 timing.jl
add timing

@@ -1,0 +1,17 @@
+using BenchmarkTools, Plots
+
+function foo(n)
+    A = randn(n,n)
+    b = randn(n)
+    A \ b
+end
+
+Ns = 1:100:1000
+tms = Array{Float64}(undef, length(Ns))
+for k = 1:length(Ns)
+    tms[k] = @belapsed foo($(Ns[k]))
+end
+
+plot(Ns, tms)
+
+plot(Ns, tms; yscale=:log10, xscale=:log10)

### 2020-10-08 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
Quickened up jacobiz and operatorclenshaw methods

@@ -1272,11 +1272,21 @@ end
 #     end
 #     B
 # end
-function operatorclenshaw(cfs::AbstractVector{T}, S::SphericalCapSpace, M) where T
+function operatorclenshaw(cfs::AbstractVector{T},
+                            S::SphericalCapSpace,
+                            M::Int,
+                            Jx::BandedBlockBandedMatrix,
+                            Jy::BandedBlockBandedMatrix,
+                            Jz::BandedBlockBandedMatrix;
+                            rotationalinvariant::Bool=false) where T
     # Outputs the operator MxM-blocked matrix operator corresponding to the
     # function f given by its coefficients of its expansion in the space S
 
-    @show "Operator Clenshaw"
+    # rotationalinvariance flag: true ⟹ function is only dependent on z, and
+    # so we can reduce the arithmatic significantly
+
+    # @show "Operator Clenshaw"
+
     # Convert the cfs vector from ordered by Fourier mode k, to by degree n
     f = convertcoeffsvecorder(S, cfs)
 
@@ -1285,32 +1295,59 @@ function operatorclenshaw(cfs::AbstractVector{T}, S::SphericalCapSpace, M) where
     resizedata!(S, N+1)
     f = PseudoBlockArray(f, [2n+1 for n=0:N])
 
-    @show "getting Jacobi matrices"
-    Jx = jacobix(S, M); @show "Jx done"
-    Jy = jacobiy(S, M); @show "Jy done"
-    Jz = jacobiz(S, M); @show "Jz done"
+    # identity matrix, requiredfor converting the coeffs vector into mat coeffs
     id = BandedBlockBandedMatrix(I, [M+1; 2M:-2:1], [M+1; 2M:-2:1], (0, 0), (0, 0))
-    # Begin alg
     P0 = getdegzeropteval(T, S)
+
+    # Special case
     if N == 0
         return f[1] * P0 * id
     end
-    ξ2 = operatorclenshawvector(S, view(f, Block(N+1)), id)
-    ξ1 = (- clenshawDTBmG(S, N-1, ξ2, Jx, Jy, Jz; operator=true)
-            + operatorclenshawvector(S, view(f, Block(N)), id))
-    for n = N-2:-1:0
[…]

### 2020-10-08 experiments/sphericalcaptests-plots.jl

@@ -34,8 +34,8 @@ T = Float64; B = T#BigFloat
 α = 0.2
 SCF = SphericalCapFamily(B, T, B(α * 1000) / 1000)
 a = 1.0
-S = SCF(a, 0.0); S2 = SCF(2.0, a); S0 = SCF(0.0, 0.0)
-ST = gettangentspace(SCF)
+S = SCF(a, 0.0); S2 = SCF(2.0, 0.0); S0 = SCF(0.0, 0.0)
+# ST = gettangentspace(SCF)
 
 y, z = B(-234)/1000, B(643)/1000; x = sqrt(1 - z^2 - y^2); p = [x; y; z]; isindomain(p, SCF)
 θ = atan(y / x)
@@ -142,9 +142,17 @@ maxm = sum(1:20+1)
 N = 30
 
 # Laplacian
-Δw = squareoperator(S, laplacianoperator(S, N), N)
+Δw = squareoperator(S, laplacianoperator(S, S, N), N)
 Plots.spy(sparse(getspymat(Δw)), marker = (:square, 1), legend=nothing)
 Plots.savefig("experiments/saved/sphericalcap/images/sparsity-of-laplacian.png")
+
+Δ2w = squareoperator(S, laplacianoperator(S2, S0, N; weighted=true), N)
+Plots.spy(sparse(getspymat(Δw)), marker = (:square, 1), legend=nothing)
+Plots.savefig("experiments/saved/sphericalcap/images/sparsity-of-laplacian-w2.png")
+
+Δ2 = laplacianoperator(S0, S2, N; weighted=false)
+Plots.spy(sparse(getspymat(Δw)), marker = (:square, 1), legend=nothing)
+Plots.savefig("experiments/saved/sphericalcap/images/sparsity-of-laplacian-2.png")
 
 # Biharmonic
 Bw = biharmonicoperator(S2, N; square=true)
@@ -168,6 +176,11 @@ Lw = squareoperator(S, rho2laplacianoperator(S, S, N), N)
 Plots.spy(sparse(getspymat(Lw)), marker = (:square, 1), legend=nothing)
 Plots.savefig("experiments/saved/sphericalcap/images/sparsity-of-rho2laplacian.png")
 
+a = randn(10)
+b = randn(10)
+c = randn(10)
+d = randn(10)
+(a' * b) * (c' * d) - a' * (b * d') * c
 
[…]

### 2020-10-08 papers/spherical-caps/spherical-caps.tex
edits to sc paper

@@
 \begin{abstract}
-In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, disk-slices and triangles.
 In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps and bands.
@@
 
-This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere: spherical bands and spherical caps.
+This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
 More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
@@
 \end{align*}
-and consider the case where $\alpha \in (-1,1)$ and where $ \rho(z) := \sqrt{1-z^2}$.
+and consider the case where $\alpha \in (-1,1)$, $\beta := 1$ and where $ \rho(z) := \sqrt{1-z^2}$.
 This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
@@
 \end{align*}
-For simplicity we focus on a spherical cap with   $\beta := 1$, though there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$.
+For simplicity we focus on the case of a spherical cap, though there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$.
+The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper (these adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same).
+For clarity we note that, for all examples and plots in this paper, we take $\alpha = 0.2$.
+\bstodo{removed "band" from title, and added a bit more here about the extension}
 
@@
 The benefit of the new approach is that we do not need to resolve Jacobians, and thereby we can achieve sparse discretisations for partial differential operators, including those with polynomial variable coefficients.
+Further, we remove the singular nature as $\alpha$ approaches $0$ that such a projection would give, since our new approach is applicable for all $\alpha \in (-1,1)$.
 On the spherical cap, the family of weights we consider are of the form
@@
 This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations.
-Furthermore, we can use this framework to derive sparse $p$-finite element methods, which we introduce as an Appendix here.
-The methods are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{beuchler2006new}, see also work by Li and Shen \cite{li2010optimal}.
+
+ \begin{figure}[t]
+	\centering % <-- added
+	\includegraphics[scale=0.3]{complexity-Nend=1000}
+	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos(z)$.}
+	\centering
+	\label{fig:complexity}
+\end{figure}
 
@@
 A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axes (e.g., a Schr\"odinger operator with potential $V(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
-This improves the complexity further to an optimal $O(p^2$).
-\sotodoinline{Adding a complexity plot showing this would be great} \bstodoinline{I am not sure how to do a complexity plot}
+This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity}.
+This plot presents the time taken (measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia) to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos(z)$, as we increase the orthogonal polynomial degree $p$.
+\bstodoinline{Added plot}
 
@@
 
-\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
+%\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
 We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
@@
 \begin{align*}
-	B^{(a)}_{x,0} &= 0 \in \R^{(N+1)\times2N}, \quad B^{(a)}_{x,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	B^{(a)}_{x,0} &= 0 \in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{x,k} = 0 \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N \\
 	A^{(a)}_{x,k} &:= 
@@
 		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
-	A^{(a)}_{x,0} &\in \R^{(N+1)\times2N}, \quad A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	&\big( \text{so that } A^{(a)}_{x,0} \in \R^{(N+1)\times2N}, \quad A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \big) \\
 	C^{(a)}_{x,k} &:= 
@@
 			& & & A^{(a)}_{N,k,1} & A^{(a)}_{N,k,3}
-		\end{pmatrix} \in \R^{2(N-k+1)\times2(N-k+2)}, \quad k = 1,2,\dots \\
-	C^{(a)}_{x,1} &\in \R^{2N\times(N+1)}, \quad C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N
+		\end{pmatrix}, \quad k = 1,2,\dots,N \\
+	&\big( \text{so that } C^{(a)}_{x,1} \in \R^{2N\times(N+1)}, \quad C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+2)} \text{ for } k = 2,3,\dots,N \big)
 \end{align*}
@@
 \begin{align*}
-	B^{(a)}_{y,0} &= 0 \in \R^{(N+1)\times2N}, \quad B^{(a)}_{y,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	B^{(a)}_{y,0} &= 0 \in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{y,k} = 0 \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N \\
 	A^{(a)}_{y,k} &:= 
@@
 		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
-	A^{(a)}_{y,0} &\in \R^{(N+1)\times2N}, \quad A^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
+	&\big( \text{so that } A^{(a)}_{y,0} \in \R^{(N+1)\times2N}, \quad A^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \big) \\
 	C^{(a)}_{y,k} &:= 
@@
 			& & & B^{(a)}_{N,k,1} & B^{(a)}_{N,k,3}
-		\end{pmatrix} \in \R^{2(N-k+1)\times2(N-k+2)}, \quad k = 1,2,\dots \\
-	C^{(a)}_{y,1} &\in \R^{2N\times(N+1)}, \quad C^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N
+		\end{pmatrix}, \quad k = 1,2,\dots,N \\
+	&\big( \text{so that } C^{(a)}_{y,1} \in \R^{2N\times(N+1)}, \quad C^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k+2)} \text{ for } k = 2,3,\dots,N \big)
 \end{align*}
@@
 	A^{(a)}_{z,0} &= 0 \in \R^{(N+1)\times2N}, \quad A^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
-	C^{(a)}_{z,1} &= 0 \in \R^{2N\times(N+1)}, \quad C^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k+3)} \text{ for } k = 2,3,\dots,N \\
+	C^{(a)}_{z,1} &= 0 \in \R^{2N\times(N+1)}, \quad C^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k+2)} \text{ for } k = 2,3,\dots,N \\
 	B^{(a)}_{z,k} &:= 
@@
 			& & \Gamma^{(a)}_{N,k,1} & \Gamma^{(a)}_{N,k,2}
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	B^{(a)}_{z,0} &\in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{z,k} \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N
+		\end{pmatrix}  \quad k = 0,1,2,\dots,N \\
+	&\big( \text{so that } B^{(a)}_{z,0} \in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{z,k} \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N \big)
 \end{align*}
@@
 
-In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices} -- we define the Clenshaw matrices as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the multidimensional three-term recurrences \bstodoinline{added explanation of clenshaw matrix in this context} While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
+In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices} -- we define the Clenshaw matrices as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the multidimensional three-term recurrences.
+While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
@@
 
-\bstodoinline{I have added defns (and sparsity proofs) for two other Laplacian operators.
-In practice, I have found that the non-weighted one is singular - is this correct/expected?
-The weighted one and the resulting Biharmonic operators both work (as far as I have tested).}
 \begin{definition}\label{def:differentialoperators}
@@
 Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions.
+This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
 The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
@@
 The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
-We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$.
-This proves much cheaper computationally than using the 3D quadrature rule to calculate the 3D inner products, which has a complexity of $\bigO(N^4)$ \bstodo{correct the O(nd) complexity}.
+We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^2)$.
+This proves much cheaper computationally than using the 3D quadrature rule to calculate the surface inner products, which has a complexity of $\bigO(N^3)$.
 
@@
 	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-poisson-N=100-xnolog}
+	\includegraphics[scale=0.3]{solutionblocknorms-poisson-epsilonfun-N=200}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
-Right: The norms of each block of the computed solution of the Poisson equation with the given right hand side functions (here, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
-This demonstrates spectral convergence observed.}
+Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
+This demonstrates spectral convergence observed.
+	\bstodoinline{New solution block norms plots.
+I have only gone up to N=100 (200 for Poisson) for now, mainly because expanding a RHS function f for larger N takes a looong time, which is where we would see the benefit of a higher N plot (for the simpler $f$'s, we would just see more super convergence).}}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-N=100-xnolog}
+	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-varyingk-N=100}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
-Right: The norms of each block of the computed solution of the Helmholtz equation with the given right hand side functions, with $k=20$ and the same function $v$ (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
-This demonstrates spectral convergence observed.
-	\sotodoinline{For examples, having  nearby singularities (e.g. something like$f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$) should give different exponential convergence rates.}
-	\bstodoinline{New examples, such as the suggested $f(\vc x) = \|\vc x - (1,1,1) * (1 /\sqrt 3 + \epsilon)\|$.
-Though these look like I should remove the trivial examples entirely, as it ruins the scaling.}
-	}
+Right: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
+This demonstrates spectral convergence observed.}
 	\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-N=100-xnolog}
+	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-expfun-N=100}
 	\centering
@@
 	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the given right hand side functions (again, $\bf{p}$ is simply a point on the spherical cap surface away from the pole and boundary).
+Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$, for various $\epsilon$ values.
 This demonstrates algebraic convergence observed.}
@@
 In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
-In the plot, a ``block'' is simply the group of coefficients corresponding to OPs of the same degree, and so the plot shows how the norms of these blocks decay as the degree of the expansion increases.
+In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
+The right hand sides we choose here are given by
+\begin{align*}
+	f(x,y,z) &= \norm{\Big(x - (\epsilon + 1 / \sqrt{3}), \: y - (\epsilon + 1 / \sqrt{3}), \: z - (\epsilon + 1 / \sqrt{3})\Big)^\top}
+\end{align*}
+for differing choices of $\epsilon$ -- this parameter serves to alter the distance from which we would have a singularity.
+In the plot, a \enquote{block} is simply the group of coefficients corresponding to OPs of the same degree, and so the plot shows how the norms of these blocks decay as the degree of the expansion increases.
 Thus, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
-We see that we achieve spectral convergence \bstodo{Add to the comments here}
+We see that we achieve spectral convergence for these examples.
 
@@
 In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ using $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
+Here, we use $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
 Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.
-\bstodo{Add to the comments here}
 
@@
 In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
-\bstodo{Add to the comments here}
+In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve algebraic convergence for these more complex functions.
 

### 2020-10-16 papers/thesis/ch2-spherical-harmonics/spherical-harmonics.tex
Added thesis tex files

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex

\input{../includes/bsmacros}
\input{../includes/bsthesisnewcommands}
\input{../includes/somacros}

\title{Spherical harmonics as orthogonal polynomials in three variables}
\author{Ben Snowball}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\section{Spherical harmonics as orthogonal polynomials in three variables}

\subsection{Introduction}

To introduce ourselves to the world of multidimensional orthogonal polynomials for solving PDEs on the sphere, we can naturally choose to look at the famous spherical harmonics. Our aim here is to express the spherical harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere. More precisely, we desire the solution to partial differential equations on the domain
\bseqn{
	\Omega := \bsset{(x,y,z) \in \R^3}{x^2 + y^2 = \rho(z)^2}
}
where 
\bseqn{
	\rho(z) := \sqrt{1-z^2}.
}
While it may seem somewhat odd, it should hopefully become apparent why this is a useful way to define our domain.

\subsection{Defining the spherical harmonics in three variables}

Before we proceed further, let's build up to our definition of the spherical harmonics. We first need to introduce a few classical orthogonal polynomials.

On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
\bseqn{
	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
}
where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$. Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
\bseqn{
	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
}
Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and are given by \bstodo{Add citation for associated legendre polys}:
\bseqn{
	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = \chat (1-x^2)^\frac{m}{2} P^{(m,m)}_{l-m}(x) \\
	P^{-m}_l(x) &:= \ctilde \Plm(x),
}
for $m = 0, 1, 2, \dots, l$ where
\bseqn{
	\chat &:= {(l+m)! \over (-2)^m \; l!} \\
	\ctilde &:= {(-1)^m \: (l-m)! \over (l+m)!}
}
and $n! := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N$ is the standard factorial. Further, define
\bseqnnumber{
	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
		\begin{cases} 
			\chat \quad \quad \quad \text{if } m \ge 0 \\
			\chatmod \ctildemod \quad \text{if } m < 0
		\end{cases}. \label{eqn:clmdef}
}
This allows us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.

Let $(x,y,z) \in \Omega$. It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$). On this note, throughout we will use the convention that the spherical coordinate angles be defined by
\bseqn{
	x &= \sinphi \costheta = \rho(z) \costheta \\
	y &= \sinphi \sintheta = \rho(z) \sintheta \\
	z &= \cosphi.
}
We can now write down the spherical harmonics. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
\bseqnnumber{
	\Ylmfull &:= \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \: \eimtheta \Plm (\cosphi) \nonumber \\
	&= \clm (1 - (\cosphi)^2)^\frac{|m|}{2} \eimtheta P^{(|m|,|m|)}_{l-|m|}(\cosphi) \nonumber \\
	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta \label{eqn:shdef}
}
for $0 \le |m| \le l, \, l \in \No$ where $\clm$ is defined in \bsrefeqn{eqn:clmdef}, and
\bseqn{
	&\int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
	& \quad = \delta_{l, l'} \: \delta_{m, m'}.
}
where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$. Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m\in\mathbb{Z}$. Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$. They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.

\subsection{Jacobi matrices}

Jacobi operators that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$. We start by expressing $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$ , and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.

\begin{lemma}
	For $l \in \No$, $m \in \Z$ s.t. $0 \le |m| \le l$, the spherical harmonics satisfy the relationships:
\bseqnnumber{
	x\,\Ylm(x,y,z) &= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \label{eqn:shrecx} \\
	y\,\Ylm(x,y,z) &= \beta_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \beta_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \beta_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \beta_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \label{eqn:shrecy} \\
	z\,\Ylm(x,y,z) &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z), \label{eqn:shrecz}
}
where
\bseqn{
	\alpha_{l,m,1} &:= 
		\begin{cases}
			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \tilde \alpha_{l,m,1} \quad \text{if } m>0 \\
			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \tilde \alpha_{l,|m|,2} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
		\end{cases} \\ 
	\alpha_{l,m,2} &:= 
		\begin{cases}
			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \tilde \alpha_{l,m,2} \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \tilde \alpha_{l,|m|,1} \quad \text{if } m<0 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
		\end{cases} \\
	\alpha_{l,m,3} &:= 
		\begin{cases}
			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \tilde \alpha_{l,m,3} \quad \text{if } m>0 \\
			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \tilde \alpha_{l,|m|,4} \quad \text{if } m\le0 
		\end{cases} \\
	\alpha_{l,m,4} &:=
		\begin{cases}
			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \tilde \alpha_{l,|m|,3} \quad \text{if } m<0 
		\end{cases} \\
	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4 \\
	\gamma_{l,m,1} &:= \frac{\clm}{c^{m}_{l-1}} \: \tilde \gamma_{l,m,1} \\
	\gamma_{l,m,2} &:= \frac{\clm}{c^{m}_{l+1}} \: \tilde \gamma_{l,m,2}.
}
and
\bseqn{
	\tilde \alpha_{l,m,1} &:= \frac{2l}{2l+1} \\
	\tilde \alpha_{l,m,2} &:= 
		\begin{cases}
			- \frac{l}{2(2l+1)} \quad \text{if } l - m \ge 2 \\
			\quad \quad 0 \quad \quad \text{otherwise}
		\end{cases} \\
	\tilde \alpha_{l,m,3} &:= - \frac{2(l-m+2)(l-m+1)}{(2l+1)(l+1)} \\
	\tilde \alpha_{l,m,4} &:= \frac{(l+m+2)(l+m+1)}{2(2l+1)(l+1)} \\
	\tilde \gamma_{l,m,1} &:= \begin{cases}
			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
			\quad 0 \quad \quad \text{otherwise}
		\end{cases} \\
	\tilde \gamma_{l,m,2} &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
}
\end{lemma}

\begin{proof}
	The recurrence relationship for the Jacobi polynomials satisfies \cite[18.9.1, 18.9.2]{DLMF} 
\bseqn{
	z P^{(m,m)}_{l-m}(z) = \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
}
for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylmfull$ by $z$ is
\bseqn{
	z\,\Ylm(x,y,z) &= \clm \eimphi \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
		     &= \clm \eimphi \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
		     &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z).
}

For multiplication by $x$ and $y$, we require some further relations for the complex exponential and the Jacobi polynomials. First, recall that
\bseqn{
	\cosphi \, \eimphi &= \frac{1}{2} (e^{i\varphi} + e^{-i\varphi}) \eimphi =  \frac{1}{2} (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \\
	\sinphi \, \eimphi &= \frac{1}{2i} (e^{i\varphi} - e^{-i\varphi}) \eimphi =  \frac{-i}{2} (e^{i(m+1)\varphi} - e^{i(m-1)\varphi})
}
Finally, we can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
\bseqn{
	{l! \: (2m)! \over (l+m)! \: m!} \: C^{(m+1/2)}_{l-m}(z) = P^{(m,m)}_{l-m}(z),
}
where the ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
\bseqn{
	C^{(\lambda)}_n(z) &= {\lambda \over n + \lambda} \: \big[ C^{(\lambda+1)}_n(z) - C^{(\lambda+1)}_{n-2}(z) \big]  \\
	(1-z^2) \, C^{(\lambda)}_n(z) &= {1 \over 4(\lambda-1)(n+\lambda)} \: \big[ (n+2\lambda-2)(n+2\lambda-1) \: C^{(\lambda-1)}_n(z) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (n+1)(n+2) \: C^{(\lambda-1)}_{n+2}(z) \big] ,
}
for $m \in \No$ where $\lambda > -\half$ is some parameter. Thus, combining these, we can write three-term recurrences for the Jacobi polynomials as:
\bseqn{
	P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z) + \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) \\
	(1-z^2) \, P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) + \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z),
}
for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylmfull$ by $x$ and $y$ is then:
\bseqn{
	x\,\Ylm(x,y,z) &= \clm \cosphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\varphi} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quad \quad + \half \clm e^{i(m-1)\varphi} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
	y\,\Ylm(x,y,z) &= \clm \sinphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\varphi} - e^{i(m-1)\varphi}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= - i \: \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \\
	&\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}

\end{proof}

These recurrences lead to Jacobi operators that correspond to multiplication by $x,y,z$. Define for $l \in \No$:
\bseqn{
	\SHvecl := 
		\begin{pmatrix}
			Y^{-l}_l \\
			\vdots \\
			Y^l_l
		\end{pmatrix} \in \C^{2l+1}, 
	\quad \quad 
	\SHvec := 
		\begin{pmatrix}
			\SHvec_0 \\
			\hline
			\SHvec_1 \\
			\hline
			\SHvec_2 \\
			\vdots \\
		\end{pmatrix}.
}
and let the Jacobi matrices $J^x, J^y, J^z$ be given by 
\bseqnnumber{
	J^x \SHvec = x \SHvec, \quad J^y \SHvec = y \SHvec, \quad J^z \SHvec = z \SHvec \label{eqn:jacobimatsdef}.
}
The Jacobi matrices act on the coefficients vector of a function's expansion in the spherical harmonic basis. For example, let the function $f(x,y,z) : \Omega \to \C$ be approximated by its expansion $f(x,y,z) = \SHvec(x,y,z)^\top \: \fvec$ up to some degree order $N$. Then $x f(x,y,z)$ is approximated by $\SHvec(x,y,z)^\top \: J^x \: \fvec$, i.e. $ J^x \: \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the spherical harmonics basis.

An important property of the Jacobi matrices is that they are sparse -- specifically they are \textit{banded-block-banded matrices}.

\begin{definition}\label{def:bandedblockbanded}
	A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L, U)$ if $A_{i,j} = 0$ for $-L \le j - i \le U$, and subblock-bandwidths $(\lambda, \mu)$ if all blocks are banded with bandwidths $(\lambda, \mu)$. A matrix where the block-bandwidths and subblock-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
\end{definition}

Using \bsrefeqns{eqn:shrecx}{eqn:shrecz}, we have that the Jacobi matrices take the following block-tridiagonal form (i.e. have block-bandwidths $(1,1)$):
\bseqn{
	J_x &= 
		\begin{pmatrix}
			B_{x,0} & A_{x,0} & & & & \\
			C_{x,1} & B_{x,1} & A_{x,1} & & & \\
			& C_{x,2} & B_{x,2} & A_{x,2}  & & & \\
			& & C_{x,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}, \\
	J_y &= 
		\begin{pmatrix}
			B_{y,0} & A_{y,0} & & & & \\
			C_{y,1} & B_{y,1} & A_{y,1} & & & \\
			& C_{y,2} & B_{y,2} & A_{y,2}  & & & \\
			& & C_{y,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}, \\
	J_z &= 
		\begin{pmatrix}
			B_{z,0} & A_{z,0} & & & & \\
			C_{z,1} & B_{z,1} & A_{z,1} & & & \\
			& C_{z,2} & B_{z,2} & A_{z,2}  & & & \\
			& & C_{z,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
$J_x$ has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{x,l} &:= 
		\begin{pmatrix}
			\alpha_{l,-l,3} & 0 & \alpha_{l,-l,4} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \alpha_{l,l,3} & 0 & \alpha_{l,l,4} \\
		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
	B_{x,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{x,l} &:= 
		\begin{pmatrix}
			\alpha_{l,-l,2} & & \\
			0 & \ddots & \\
			\alpha_{l,-l+2,1} & \ddots & \alpha_{l,l-2,2} \\
			& \ddots & 0 \\
			& & \alpha_{l,l,1}
		\end{pmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}
$J_y$ also has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{y,l} &:= 
		\begin{pmatrix}
			\beta_{l,-l,3} & 0 & \beta_{l,-l,4} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \beta_{l,l,3} & 0 & \beta_{l,l,4} \\
		\end{pmatrix} \in \C^{(2l+1)\times(2l+3)}, \\
	B_{y,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{y,l} &:= 
		\begin{pmatrix}
			\beta_{l,-l,2} & & \\
			0 & \ddots & \\
			\beta_{l,-l+2,1} & \ddots & \beta_{l,l-2,2} \\
			& \ddots & 0 \\
			& & \beta_{l,l,1}
		\end{pmatrix} \in \C^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}
Finally, $J_z$ has subblock-bandwidths $(1,1)$, where the blocks for $l \in \No$ are given by:
\bseqn{
	A_{z,l} &:= 
		\begin{pmatrix}
			0 & \gamma_{l,-l,2} & 0 & & \\
			& \ddots & \ddots & \ddots & \\
			& & 0 &  \gamma_{l,-l,2} & 0 \\
		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
	B_{z,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
	C_{z,l} &:= 
		\begin{pmatrix}
			0 & & \\
			\gamma_{l,-l+1,1} & \ddots & \\
			0 & \ddots & 0 \\
			& \ddots & \gamma_{l,l-1,1} \\
			& & 0
		\end{pmatrix} \in \R^{(2l+1)\times(2l-1)} \quad (l \ne 0).
}

\subsection{Three-term recurrence relation for $\SHvec$}

Three-term recurrence relations for orthogonal polynomials are well established (e.g. \cite[18.9]{DLMF}, \cite{dunkl2014orthogonal}). In a similar vein, we can obtain a recurrence relation for the spherical harmonics by combining each system in (\ref{eqn:jacobimatsdef}).
\bseqn{
	\renewcommand\arraystretch{1.3}
	\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\SHvec(x,y,z)
	=
	\begin{pmatrix}
		Y_0 \\ 0 \\ 0 \\ 0 \\ \vdots  \\
	\end{pmatrix},
}
where we note $Y_0^0(x,y,z) \equiv Y_0 := P_0^{(0,0)} \equiv 1$, and for each $l = 0,1,2\dots$,
\begin{align}
	A_l &:= 
		\begin{pmatrix}
			A_{x,l} \\
			A_{y,l} \\
			A_{z,l}
		\end{pmatrix} \in \C^{3(2l+1)\times(2l+3)}, \quad
	C_l := 
		\begin{pmatrix}
			C_{x,l} \\
			C_{y,l} \\
			C_{z,l}
		\end{pmatrix} \in \C^{3(2l+1)\times(2l-1)} \quad (n \ne 0), \label{eqn:clenshawmats1} \\
	B_l &:= 
		\begin{pmatrix}
			B_{x,l} \\
			B_{y,l} \\
			B_{z,l}
		\end{pmatrix} \in \C^{3(2l+1)\times(2l+1)}, \quad
	G_n(x,y) := 
		\begin{pmatrix}
			x\idmat{2l+1} \\
			y\idmat{2l+1} \\
			z\idmat{2l+1}
		\end{pmatrix}  \in \C^{3(2l+1)\times(2l+1)}. \label{eqn:clenshawmats2}
\end{align}

For each $l = 0,1,2\dots$ let $\Dlt$ be any matrix that is a left inverse of $A_l$, i.e. such that $\Dlt A_l = \idmat{2l+3}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\bseqn{
	\begin{cases}
		\SHvec_{-1}(x,y,z) := 0 \\
		\SHvec_{0}(x,y,z) := Y_0 \\
		\SHvec_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z), \quad l = 0,1,2,\dots.
	\end{cases}
}

Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices in the following way. For $l \in \N$, we set
\bseqnnumber{
	\Dlt = 
		\begin{pmatrix}
			\hat{A}^{x,y}_l & \zeromat{(2l+3)\times(2l+1)}
		\end{pmatrix} \in \R^{(2l+3)\times3(2l+1)} \label{eqn:Dltdef}
}
where $\zeromat{(2l+3)\times(2l+1)}$ the zero matrix in $\R^{(2l+3)\times(2l+1)}$, and $\hat{A}^{x,y}_l \in \R^{(2l+3)\times2(2l+1)}$ is the left inverse of the matrix $\begin{pmatrix} A^x_l \\ A^y_l \end{pmatrix}$, given by
\bseqn{
	\hat{A}^{x,y}_l = 
		\begin{pmatrix}
			\frac{1}{2\alpha_{l,-l,3}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,-l,3}} & 0 & \hdots & 0 \\
			& \ddots & & & & \ddots & & \\
			& & \ddots & & & & \ddots & \\
			& & & \frac{1}{2\alpha_{l,l,3}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,-l,3}} \\
			0 & \hdots & \frac{1}{2\alpha_{l,l-1,4}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,l-1,4}} & 0 \\
			0 & \hdots & 0 & \frac{1}{2\alpha_{l,l,4}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,l,4}}
		\end{pmatrix}.
}
For $l = 0$ we set
\bseqnnumber{
	D^T_0 = 
		\begin{pmatrix}
			\frac{1}{2\alpha_{0,0,3}} & \frac{1}{2\beta_{0,0,3}} & 0 \\
			0 & 0 & \frac{1}{\gamma_{0,0,2}} \\
			\frac{1}{2\alpha_{0,0,4}}&  \frac{1}{2\beta_{0,0,4}} & 0 
		\end{pmatrix}. \label{eqn:Dzerotdef}
}

It will be useful for us to give a formal name for these coefficient matrices above for a family of multidimensional orthogonal polynomials.
\begin{definition}\label{def:clenshawmats}
	The matrices $-\Dlt (B_l-G_l(x,y,z))$, $\Dlt C_l$ for $l \in \No$ defined via \bsrefeqns{eqn:clenshawmats1}{eqn:Dzerotdef} are called the \textbf{recurrence coefficient matrices} for a given family of multidimensional orthogonal polynomials.
\end{definition}

%
%\subsection{Laplacian operator}
%
%The spherical harmonics are named as such because they are indeed harmonic functions -- that is, they satisfy $\nabla^2 \Ylm = \lambda \Ylm$ for some eigenvalue $\lambda$. It can be shown that in fact, for any $l \in \No$, $-l \le m \le l$, \bstodo{cite}
%\bseqn{
%	\nabla^2 \Ylm = - l (l+1) \Ylm.
%}
%This allows us to write down a sparse (in fact, diagonal) matrix operator for the Laplacian. Recall that a function $f(x,y,z)$ defined on the unit sphere $\Omega$ is approximated by its expansion $\SHvec^\top \fvec$ for some coefficients vector $\fvec$ up to degree order $N \in \N$. Then the function $(x,y,z) \mapsto \nabla^2 f(x,y,z)$ is approximated by $\SHvec^\top \Delta \fvec$ where $\Delta$ is the matrix operator is then simply given by
%\bseqn{
%	\Delta := 
%		\begin{pmatrix}
%			\Delta_0 & & & \\
%			& \Delta_1 & & \\
%			& & \ddots & \\
%			& & & \Delta_N
%		\end{pmatrix}, \quad
%	\Delta_l := - l (l+1) \idmat{2l+1} \in \R^{(2l+1)\times(2l+1)}.
%}

\subsection{Computational aspects}

\bstodoinline{quad rule on sphere?}

\subsubsection{Function evaluation}\label{subsubsection:clenshaw}

Once again, let $f(x,y,z)$ be a function on the unit sphere $\Omega$ be approximated by its expansion 
\bseqn{
	f(x,y,z) \approx \SHvec(x,y,z)^\top \fvec = \sum^N_{l=0} \SHvecl(x,y,z)^\top \fvec_l,
}
where $\SHvecl (x,y,z), \fvec_l \in \C^{2l+1}$ for each  $l \in \{0,\dots,N\}$, for some coefficients vector $\fvec$ up to degree order $N \in \N$. We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}. The \textit{Clenshaw algorithm} is then as follows:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:0 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \fvec_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
}
where $\SHvec_0 \equiv \alpha^0_0$.

\subsubsection{Function evaluation}

The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used with the Jacobi matrices replacing the point $(x,y,z)$, to yield an operator matrix.

Let's explain what we mean. Suppose $v : \Omega \to \C$ is a function, and we encounter the problem of finding the coefficients of the expansion for $v(x,y,z) f(x,y,z)$. We wish to therefore find an operator $V$ for $v$ so that
\bseqn{
	v(x,y,z) f(x,y,z) = v(x,y,z) \fvec^\top \SHvec(x,y,z) = (V \; \fvec)^\top \SHvec(x,y,z),
}
i.e. $V \; \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto v(x,y,z) f(x,y,z)$ in the spherical harmonic basis.

Let $\vec{v}$ be the coefficients of the expansion of $v$ up to order $N$. The operator $V$ would then be the result of the following \textit{operator Clenshaw algorithm}:
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:0 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \vec{v}_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(J_x,J_y,J_z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
}
where at each iteration, $\vec{\xi}_n$ is a vector of matrices (note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).

%%%
\section{Tangent space -- vector spherical harmonics as orthogonal vectors in three variables}

Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics 
\bseqnnumber{
	\gradYlm, \quad \gradpYlm := \rvec \times \gradYlm, \label{eqn:VSHs}
}
span the tangent space. Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$. Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}. Other definitions for VSHs are used, however it is convenient for deriving explicit sparse relations and operators to use the one described here.

A simple calculation shows that such orthogonal vectors must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal. We will determine the coefficients of these Jacobi operators, as well as other important operators needed -- however first we note a useful property of the VSHs that is a result of the definition used.

\subsection{Vector spherical harmonics as a vector-weighted sum of scalar spherical harmonics} %{Spin-1 tensor spherical harmonics}

The vector spherical harmonics in \bsrefeqn{eqn:VSHs} can be expressed in terms of \textit{spin-1 tensor spherical harmonics}, which in turn can each be expressed as a vector-weighted sum of spherical harmonics \cite{khersonskii1988quantum}.

We start by defining what we mean by a spin-1 tensor spherical harmonic. In general, the tensor spherical harmonic is given by, for $2l, \, j, \, 2s \in \No$,
\bseqn{
	\curlyy^{j,s}_{l,m}(x,y,z) = \sum_{m_s=-s}^s \langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle \, Y^{m-m_s}_j (x,y,z) \, \vec{\chi}_{s, m_s},
}
where $\vec{\chi}_{s, m_s}$ are the simultaneous eigenstates of the spin operators $\bold{S}^2$ and $S_z$, and where $\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle$ is a Clebsch-Gordan coefficient \cite{} \bstodo{cite C-G coeffs}. We note that a property of the Clebsch--Gordan coefficients means that they vanish unless $|j - s| \le l \le j + s$. We further note that there are simple calculable expressions for the Clebsch--Gordan coefficients when the spin $s=1$.

To simplify the notation, for spin $s=1$, we define
\bseqnnumber{
	\cgcoeff^{L, m_s}_{l, m} := \, \langle L \quad m-m_s \; ; \; 1 \quad m_s \; | \; l \quad m \rangle. \label{eqn:cgcoeffdef}
}
Thus we have that the three spin-1 tensor spherical harmonics are given by
\bseqn{
	\curlyy^{l \pm1,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
	\curlyy^{l,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
}

Here, the vectors $\chivec$ are the orthonormal eigenvectors of the matrix
\bseqn{
	S_3 = \begin{pmatrix} 0 & -i & 0 \\ i & 0 & 0 \\ 0 & 0 & 0 \end{pmatrix},
}
and so are given as
\bseqn{
	\vec{\chi}_{1, \pm 1} = \frac{1}{\sqrt{2}} \begin{pmatrix} \mp 1 \\ -i \\ 0 \end{pmatrix}, \quad \vec{\chi}_{1,0} = \begin{pmatrix} 0 \\ 0 \\ 1 \end{pmatrix}.
}
Then, for any $l \in \No$, $m \in \Z$ s.t. $0 \le |m| \le l$ we have that
\bseqnnumber{
	\gradYlm &= \mu_{l,-1} \, \curlyy^{l-1,1}_{l,m} + \mu_{l,1} \curlyy^{l+1,1}_{l,m}, \quad \gradpYlm = \mu_{l,0} \, \curlyy^{l,1}_{l,m},\label{eqn:VSHsalt}
}
where \bstodoinline{get better reference than that Winter 2017 physics document}
\bseqnnumber{
	\mu_{l,-1} &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
	\mu_{l,0} := i \, \big(l(l+1)\big)^\half, \quad 
	\mu_{l,1} := l \, \Big(\frac{l+1}{2l+1}\Big)^\half. \label{eqn:spintensorvshcoeffsdef}
}

\subsection{Jacobi matrices}

We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc. in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$. 
\begin{lemma}\label{lemma:VSHrecurrences}
	For $l \in \No$, $-l \le m \le l$ the vector spherical harmonics satisfy the relationships:
\bseqn{
	x \gradYlm &= A_{l,m,1} \gradY^{m-1}_{l-1} + A_{l,m,2} \gradY^{m+1}_{l-1} + A_{l,m,5} \gradY^{m-1}_{l+1} + A_{l,m,6} \gradY^{m+1}_{l+1} \\
& \quad \quad \quad + A_{l,m,3}^{*} \gradpY^{m-1}_{l} + A_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
	x \gradpYlm &= A_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + A_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + A_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + A_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
	& \quad \quad \quad + A_{l,m,3} \gradY^{m+1}_{l} + A_{l,m,4} \gradY^{m-1}_{l}, \\
	y \gradYlm &= B_{l,m,1} \gradY^{m-1}_{l-1} + B_{l,m,2} \gradY^{m+1}_{l-1} + B_{l,m,5} \gradY^{m-1}_{l+1} + B_{l,m,6} \gradY^{m+1}_{l+1} \\
& \quad \quad \quad + B_{l,m,3}^{*} \gradpY^{m-1}_{l} + B_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
	y \gradpYlm &= B_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + B_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + B_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + B_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
	& \quad \quad \quad + B_{l,m,3} \gradY^{m+1}_{l} + B_{l,m,4} \gradY^{m-1}_{l}, \\
	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l}, \\
	z \gradpYlm &= \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l}
}
where $*$ denotes the complex conjugate, and where, for any valid $m_s$ value,
\bseqn{
	A_{l,m,1} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m-1}} \alpha_{l-1,m-m_s,1}, \\
	A_{l,m,2} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} \alpha_{l-1,m-m_s,2}, \\
	A_{l,m,3} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,3} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,1} \\
	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,5} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - A_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big], \\	
	A_{l,m,4} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,4} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,2} \\
	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,6} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - A_{l,m,2} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
	A_{l,m,5} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} \alpha_{l+1,m-m_s,3}, \\
	A_{l,m,6} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} \alpha_{l+1,m-m_s,4}, \\
	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6, \\
	\Gamma_{l,m,1} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} \gamma_{l-1,m-m_s,1}, \\
	\Gamma_{l,m,2} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \gamma_{l-1,m-m_s,2} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \gamma_{l+1,m-m_s,1} \\
	& \quad \quad \quad \quad \quad \quad \quad - \Gamma_{l,m,3} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \Gamma_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big], \\
	\Gamma_{l,m,3} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} \gamma_{l+1,m-m_s,2},
}
and where $\cgcoeff^{l,m_s}_{l',m'}$, $\mu_{l,k}$ are defined in \bsrefeqn{eqn:cgcoeffdef} and \bsrefeqn{eqn:spintensorvshcoeffsdef} respectively.
\end{lemma}

\remark It is the case that these coefficients are indeed the same for each $l,m$ pair despite appearing to depend on the value of $m_s$; we need only use any valid $m_s$ value for each expression. By \enquote{valid}, we mean the Clebsch--Gordan coefficients do not vanish for that $m_s$ value when used.

\begin{proof}[Proof of \bsreflemma{lemma:VSHrecurrences}]

\bstodoinline{Do the derivation of these rec coeffs, to have as the proof (and check they are correct!)}

\end{proof}

%\begin{align}
%x \,\gradYlm 
%&= x \, \blmone \, \curlyy^{l-1,1}_{l,m} + x \, \blone \curlyy^{l+1,1}_{l,m} \nonumber
%\\
%&= \sum_{m_s=-1}^{1} \chivec \Big[ \blmone \, \cgcoeff^{l-1,m_s}_{l,m} \big\{ A_{l-1,m-m_s} \, Y^{m-m_s+1}_{l} + B_{l-1,m-m_s} \, Y^{m-m_s+1}_{l-2} \nonumber \\ 
%& \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + D_{l-1,m-m_s} \, Y^{m-m_s-1}_{l} + E_{l-1,m-m_s} \, Y^{m-m_s-1}_{l-2} \big\} \nonumber \\
%& \quad \quad \quad \quad \quad \quad \quad + \blone \, \cgcoeff^{l+1,m_s}_{l,m} \big\{ A_{l+1,m-m_s} \, Y^{m-m_s+1}_{l+2} + B_{l+1,m-m_s} \, Y^{m-m_s+1}_{l} \nonumber \\ 
%& \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + D_{l+1,m-m_s} \, Y^{m-m_s-1}_{l+2} + E_{l+1,m-m_s} \, Y^{m-m_s-1}_{l} \big\} \Big]
%\end{align}
%\begin{align}
%\implies x \,\gradYlm &= \alm \gradY^{m+1}_{l+1} + \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
%& \quad \quad \quad + \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l},
%\end{align}
%where, for any valid $m_s$ value,
%\begin{align}
%\alm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} A_{l+1,m-m_s}, \\
%\blm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} B_{l-1,m-m_s}, \\
%\dlm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} D_{l+1,m-m_s}, \\
%\elm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m-1}} E_{l-1,m-m_s}, \\
%\hlm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, A_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, B_{l+1,m-m_s} \nonumber \\
%& \quad \quad \quad \quad \quad \quad \quad - \alm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - \blm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
%\jlm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, D_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, E_{l+1,m-m_s} \nonumber \\
%& \quad \quad \quad \quad \quad \quad \quad - \dlm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - \elm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big].
%\end{align}
%
%Note that (it can be shown) these are constants for each $l,m$ pair despite appearing to depend on the value of $m_s$; we need only use any valid $m_s$ value for each expression. By ``valid'', we mean the Clebsch--Gordan coefficients do not vanish for that $m_s$ value when used. 
%
%Similarly, we have that
%\begin{align}
%y \,\gradYlm &= i \, \Big[ -\alm \gradY^{m+1}_{l+1} - \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
%& \quad \quad \quad - \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l} \Big].
%\end{align}
%
%Further, we also have that
%\begin{align}
%z \,\gradYlm 
%&= z \, \blzero \, \curlyy^{l,1}_{l,m} \nonumber
%\\
%&= \sum_{m_s=-1}^{1} \chivec \Big[ \blmone \, \cgcoeff^{l-1,m_s}_{l,m} \big\{ F_{l-1,m-m_s} \, Y^{m-m_s}_{l} + G_{l-1,m-m_s} \, Y^{m-m_s}_{l-2} \big\} \nonumber \\ 
%& \quad \quad \quad \quad \quad \quad \quad + \blone \, \cgcoeff^{l+1,m_s}_{l,m} \big\{ F_{l+1,m-m_s} \, Y^{m-m_s}_{l+2} + G_{l+1,m-m_s} \, Y^{m-m_s}_{l} \big\} \Big]
%\end{align}
%\begin{align}
%\implies z \,\gradYlm &= \flm \gradY^{m}_{l+1} + \glm \gradY^{m}_{l-1} + \klm \gradpY^{m}_{l},
%\end{align}
%where, for any valid $m_s$ value,
%\begin{align}
%\flm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} F_{l+1,m-m_s}, \\
%\glm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} G_{l-1,m-m_s}, \\
%\klm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, F_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, G_{l+1,m-m_s}  \nonumber \\
%& \quad \quad \quad \quad \quad \quad \quad - \flm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \glm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big].
%\end{align}
%
%We can similarly yield the relations for the perpendicular gradients:
%\begin{align}
%x \,\gradpYlm &= \almperp \gradpY^{m+1}_{l+1} + \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
%& \quad \quad \quad + \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
%y \,\gradpYlm &= i \, \Big[ -\almperp \gradpY^{m+1}_{l+1} - \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
%& \quad \quad \quad - \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
%z \,\gradpYlm &= \flmperp \gradpY^{m}_{l+1} + \glmperp \gradpY^{m}_{l-1} + \klmperp \gradY^{m}_{l},
%\end{align}
%where it can be shown that
%\begin{align}
%\almperp &= \alm^*, \quad \blmperp = \blm^*, \quad \dlmperp = \dlm^*, \quad \elmperp = \elm^*, \nonumber \\
%\flmperp = \flm^*,& \quad \glmperp = \glm^*, \quad \hlmperp = \hlm^*, \quad \jlmperp = \jlm^*, \quad \klmperp = \klm^*,
%\end{align}
%where $^*$ denotes the complex conjugate.

As for the scalar case, these recurrences above lead us to Jacobi matrices for the tangent space basis. Define $\VSHvecfull$ by
\bseqnnumber{
	\VSHvecfull = 
		\begin{pmatrix} 
			\VSHvecfull_0 \\ \VSHvecfull_1 \\ \vdots 
		\end{pmatrix}, \quad \text{where } \quad 
	\VSHvecfulll = 
		\begin{pmatrix} 
			(\gradY^{-l}_l)^\top \\ (\gradpY^{-l}_l)^\top \\ \vdots \\ (\gradY^{l}_l)^\top \\ (\gradpY^{l}_l)^\top
		\end{pmatrix} \in \C^{2(2l+1) \times 3} \quad \forall \: l \in \No. \label{eqn:VSHfulldef}
}

The Jacobi matrices are then defined according to the matrices $\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$ satisfying
\bseqnnumber{
	\tangentjacobi_x \VSHvecfull(x,y,z) &= x \VSHvecfull(x,y,z) \nonumber \\
	\tangentjacobi_y \VSHvecfull(x,y,z) &= y \VSHvecfull(x,y,z) \label{eqn:tangentjacobimatdef} \\ 
	\tangentjacobi_z \VSHvecfull(x,y,z) &= z \VSHvecfull(x,y,z) \nonumber
}
for each $(x,y,z) \in \Omega$. The tangent space Jacobi matrices are banded-block-banded due to the sparse relationships we obtain from \bsreflemma{lemma:VSHrecurrences}. Specifically, they each have block-bandwidths $(1,1)$:
\bseqn{
	\tangentjacobi_{x/y/z} &= 
		\begin{pmatrix}
			B_{x/y/z,0} & A_{x/y/z,0} & & & & \\
			C_{x/y/z,1} & B_{x/y/z,1} & A_{x/y/z,1} & & & \\
			& C_{x/y/z,2} & B_{x/y/z,2} & A_{x/y/z,2}  & & & \\
			& & C_{x/y/z,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
Let $0_2 \in \R^{2\times2}$ be the 2x2 zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
\bseqn{
	A_{x,l} &:= 
		\begin{pmatrix}
			\mathcal{A}_{l,-l,5} & 0_2 & \mathcal{A}_{l,-l,6} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \mathcal{A}_{l,l,5} & 0_2 & \mathcal{A}_{l,l,6} \\
		\end{pmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
	B_{x,l} &:= 
		\begin{pmatrix}
			0_2 & \mathcal{A}_{l,-l,4} & \\
			\mathcal{A}_{l,-l+1,3} & \ddots & \ddots \\
			& \ddots & \ddots & \mathcal{A}_{l,l-1,4} \\
			& & \mathcal{A}_{l,l,3} & 0_2
		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
	C_{x,l} &:= 
		\begin{pmatrix}
			\mathcal{A}_{l,-l,2} & & & \\
			0_2 & \ddots & & \\
			\mathcal{A}_{l,-l+2,1} & \ddots & \ddots & \\
			& \ddots & \ddots & \mathcal{A}_{l,l-2,2} \\
			& & \ddots & 0_2 \\
			& & & \mathcal{A}_{l,l,1} \\
		\end{pmatrix} \in \R^{2(2l+1)\times2(2l-1)} \quad (l \ne 0),
}
where
\bseqn{
	\mathcal{A}_{l,m,j} := 
		\begin{pmatrix}
			A_{l,m,j} & 0 \\
			0 & A_{l,m,j}^*
		\end{pmatrix} \: \text{for } j = 1, 2, 5, 6, \quad
	\mathcal{A}_{l,m,j} := 
		\begin{pmatrix}
			0 & A_{l,m,j} \\
			A_{l,m,j}^* & 0
		\end{pmatrix} \: \text{for } j = 3,4.
}
Similarly, $\tangentjacobi_y$ also has subblock-bandwidths $(4,4)$ since for $l \in \No$:
\bseqn{
	A_{y,l} &:= 
		\begin{pmatrix}
			\mathcal{B}_{l,-l,5} & 0_2 & \mathcal{B}_{l,-l,6} & & \\
			& \ddots & \ddots & \ddots & \\
			& & \mathcal{B}_{l,l,5} & 0_2 & \mathcal{B}_{l,l,6} \\
		\end{pmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
	B_{y,l} &:= 
		\begin{pmatrix}
			0_2 & \mathcal{B}_{l,-l,4} & \\
			\mathcal{B}_{l,-l+1,3} & \ddots & \ddots \\
			& \ddots & \ddots & \mathcal{B}_{l,l-1,4} \\
			& & \mathcal{B}_{l,l,3} & 0_2
		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
	C_{y,l} &:= 
		\begin{pmatrix}
			\mathcal{B}_{l,-l,2} & & & \\
			0_2 & \ddots & & \\
			\mathcal{B}_{l,-l+2,1} & \ddots & \ddots & \\
			& \ddots & \ddots & \mathcal{B}_{l,l-2,2} \\
			& & \ddots & 0_2 \\
			& & & \mathcal{B}_{l,l,1} \\
		\end{pmatrix} \in \R^{2(2l+1)\times2(2l-1)} \quad (l \ne 0),
}
where
\bseqn{
	\mathcal{B}_{l,m,j} := 
		\begin{pmatrix}
			B_{l,m,j} & 0 \\
			0 & B_{l,m,j}^*
		\end{pmatrix} \: \text{for } j = 1, 2, 5, 6, \quad
	\mathcal{B}_{l,m,j} := 
		\begin{pmatrix}
			0 & B_{l,m,j} \\
			B_{l,m,j}^* & 0
		\end{pmatrix} \: \text{for } j = 3,4.
}
Finally, $\tangentjacobi_z$ has subblock-bandwidths $(2,2)$, as for $l \in \No$: 
\bseqn{
	A^z_l &:= 
		\begin{pmatrix}
			0 & 0 & \Gamma_{l,-l,3} \\
			& & & \Gamma_{l,-l,3}^* \\
			& & & & \ddots \\
			& & & & & \Gamma_{l,l,3} \\
			& & & & & & \Gamma_{l,l,3}^*& 0 & 0 \\
		\end{pmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
	B^z_l &:= 
		\begin{pmatrix}
			0 & \Gamma_{l,-l,2} \\
			\Gamma_{l,-l,2}^* & 0 \\
			& & \ddots \\
			& & & 0 & \Gamma_{l,l,2} \\
			& & & \Gamma_{l,l,2}^* & 0 \\
		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
	C^z_l &:= 
		\begin{pmatrix}
			0 \\
			0 \\
			\Gamma_{l,-l+1,1} \\
			& \Gamma_{l,-l+1,1}^* \\
			& & \ddots \\
			& & & \Gamma_{l,l-1,1} \\
			& & & & \Gamma_{l,l-1,1}^* \\
			& & & & 0 \\
			& & & & 0 \\
		\end{pmatrix} \in \R^{2(2l+1)\times2(2l-1)} \quad (l \ne 0).
}

\subsection{Three-term recurrence relation for $\VSHvecfull(x,y,z)$}

We can combine each system in \bsrefeqn{eqn:tangentjacobimatdef} into a block-tridiagonal system for any $(x,y,z) \in \Omega$:
\bseqn{
	\renewcommand\arraystretch{1.3}
	\begin{pmatrix}
		I_6 & & & \\
		B_1-G_1(x,y,z) & A_1 & & \\
		C_2 & B_2-G_2(x,y,z) & \quad A_3 \quad & \\
		& C_3 & B_3 - G_3(x,y,z)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\VSHvecfull(x,y,z)
	=
	\begin{pmatrix}
		\VSHvecfull_1(x,y,z) \\ \vec{0} \\ \vec{0} \\ \vdots
	\end{pmatrix},
}
where we note that 
\bseqn{
	\VSHvecfull_{1}(x,y,z) := 
		\begin{pmatrix}
			\gradY_1^{-1}(x,y,z) \\ \gradpY_1^{-1}(x,y,z) \\ \gradY_1^{0}(x,y,z) \\ \gradpY_1^{0}(x,y,z) \\ \gradY_1^{1}(x,y,z) \\ \gradpY_1^{1}(x,y,z) \\
			\end{pmatrix}, 
}
can be calculated explicitly, and for each $l \in \N$,
\bseqnnumber{
	A_l &:= 
		\begin{pmatrix}
			A_{x,l} \\
			A_{y,l} \\
			A_{z,l}
		\end{pmatrix} \in \C^{6(2l+1)\times2(2l+3)}, \quad
	C_l := 
		\begin{pmatrix}
			C_{x,l} \\
			C_{y,l} \\
			C_{z,l}
		\end{pmatrix} \in \C^{6(2l+1)\times2(2l-1)} \quad (n \ne 0), \label{eqn:tangentclenshawmats1} \\
	B_l &:= 
		\begin{pmatrix}
			B_{x,l} \\
			B_{y,l} \\
			B_{z,l}
		\end{pmatrix} \in \C^{6(2l+1)\times2(2l+1)}, \quad
	G_n(x,y) := 
		\begin{pmatrix}
			x\idmat{2l+1} \\
			y\idmat{2l+1} \\
			z\idmat{2l+1}
		\end{pmatrix}  \in \C^{6(2l+1)\times2(2l+1)}. \label{eqn:tangentclenshawmats1}
}
For each $l \in \N$ let $\Dlt$ be any matrix that is a left inverse of $A_l$, i.e. such that $\Dlt A_l = \idmat{2(2l+3)}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal} which can be expanded to obtain the three-term recurrence:
\bseqn{
	\VSHvecfull_{l+1}(x,y,z) = -\Dlt [B_l-G_l(x,y,z)] \VSHvecfull_l(x,y,z) - \Dlt C_l  \, \VSHvecfull_{l-1}(x,y,z), \quad l \in \N
}
where of course we have that $\VSHvec(x,y,z) \equiv \vec{0}_{2\times3}$.

We note that we can choose the matrices $\Dlt$ in the following way. For $l \in \N$, we set
\begin{align}
\Dlt = \begin{pmatrix}
		\hat{A}^{x,y}_l & \zeromat{2(2l+3)\times2(2l+1)}
	  \end{pmatrix} \in \R^{2(2l+3)\times6(2l+1)}
\end{align}
where $\zeromat{2(2l+3)\times2(2l+1)}$ the zero matrix in $\R^{2(2l+3)\times2(2l+1)}$, and $\hat{A}^{x,y}_l \in \R^{2(2l+3)\times4(2l+1)}$ is the left inverse of the matrix $\begin{pmatrix} A^x_l \\ A^y_l \end{pmatrix}$, given by
\begin{align}
\hat{A}^{x,y}_l = \begin{pmatrix}
		\frac{1}{2d_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2d_{l,-l}} & 0 & \hdots &  \\
		0 & \frac{1}{2d^\perp_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2d^\perp_{l,-l}} & 0 & \hdots &  \\
		& & \ddots & & & & \ddots & & \\
		\vdots & & & \ddots & & & & \ddots & \\
		& & & & \frac{1}{2d_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2d_{l,l}} & \\
		& & & & & \frac{1}{2d^\perp_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2d^\perp_{l,l}} \\
		0 & \hdots & \frac{1}{2a_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2a_{l,l-1}} & 0 & 0 & 0 \\
		0 & \hdots & 0 & \frac{1}{2a^\perp_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2a^\perp_{l,l-1}} & 0 & 0\\
		0 & \hdots & \hdots & 0 & \frac{1}{2a_{l,l}} & 0 & \hdots & 0 & \frac{i}{2a_{l,l}} & 0 \\
		0 & \hdots & \hdots & & 0 & \frac{1}{2a^\perp_{l,l}} & 0 & \hdots & 0 & \frac{i}{2a^\perp_{l,l}} \\
	  \end{pmatrix}.
\end{align}
\bstodoinline{Derive Dlts again}

\subsection{Function evaluation}

The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$. Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
\bseqn{
	\uvec(x,y,z) = u_\varphi(x,y,z) \: \phivec + u_\theta(x,y,z) \: \thetavec, \quad (x,y,z) \in \Omega
}
for some scalar functions $u^\varphi, u^\theta$ where $\phivec, \thetavec$ are the standard non-radial unit basis vectors on the sphere defined by $\phivec := (\cos \theta \sin \varphi, \sin \theta \sin \varphi, \cos \varphi)^\top$, $\thetavec := (-\sin \theta, \cos \theta, 0)^\top$. Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent space of $\Omega$}. We can then approximate this function by its expansion in the VSHs basis, i.e. for some large enough $N \in \N$:
\bseqn{
	\uvec(x,y,z) \approx \VSHvecfull(x,y,z)^\top \uvec^c := \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big].
}
The clenshaw algorithm to evaluate $\uvec$ at a point $(x,y,z) \in \Omega$ is then as follows
\bseqn{
	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
	\quad &\text{2) } \text{For } n = N:-1:1 \\
	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^\top = (\mathbf{u}^c_n)^\top - \vec{\xi}_{n+1}^\top \Dnt (B_n - G_n) -  \vec{\xi}_{n+2}^\top D^\top_{n+1} C_{n+1} \\
	\quad &\text{3) } \text{Output: } \mathbf{u}(x,y,z) \approx \VSHvecfull_1(x,y,z)^\top \vec{\xi}_{1} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad = (\vec{\xi}_{1})_1 \gradY_1^{-1} + (\vec{\xi}_{1})_2 \gradpY_1^{-1} + (\vec{\xi}_{1})_3 \gradY_1^{0} \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (\vec{\xi}_{1})_4 \gradpY_1^{0} + (\vec{\xi}_{1})_5 \gradY_1^{1} + (\vec{\xi}_{1})_6 \gradpY_1^{1}.
}

\subsection{Sparse partial differential operators}

The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplacian, as well as other key operators useful in examples such as the linearised shallow water equations. These operators will take coefficients of a function's expansion in either the SSH or VSH basis, to coefficients in either basis.

In order to ensure we fully capitalise on the natural sparsity of the relationships so that we yield banded-block-banded operators, we will \enquote{split} coefficient vectors of a vector-valued function on the tangent space of $\Omega$ into two. On that note, define $\VSHvec, \VSHvecperp$ by
\bseqnnumber{
	\VSHvec &:= 
		\begin{pmatrix} 
			\VSHvec_0 \\ \VSHvec_1 \\ \vdots 
		\end{pmatrix}, \quad \text{where } \quad 
	\VSHvecl := 
		\begin{pmatrix} 
			(\gradY^{-l}_l)^\top \\ \vdots \\ (\gradY^{l}_l)^\top
		\end{pmatrix} \in \C^{(2l+1) \times 3} \quad \forall \: l \in \No, \label{eqn:VSHdef} \\
	\VSHvecperp &:= 
		\begin{pmatrix} 
			\VSHvecperp_0 \\ \VSHvecperp_1 \\ \vdots 
		\end{pmatrix}, \quad \text{where } \quad 
	\VSHvecperpl := 
		\begin{pmatrix} 
			(\gradpY^{-l}_l)^\top \\ \vdots \\ (\gradpY^{l}_l)^\top
		\end{pmatrix} \in \C^{(2l+1) \times 3} \quad \forall \: l \in \No. \label{eqn:VSHperpdef}
}
Let $\uvec$ be a vector-valued function on the tangent space of $\Omega$ with coefficients vectors $\uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, and let $f$ be a scalar function on $\Omega$ with coefficients vector $\fvecc$ for its expansion in the SSH basis also up to order $N$, i.e.
\bseqn{
	\uvec(x,y,z) &\approx \VSHvec(x,y,z)^\top \: \uvecc + \VSHvecperp(x,y,z)^\top \: \uveccperp \\
	&:=  \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big], \\
	f(x,y,z) &\approx \SHvec(x,y,z)^\top \: \fvecc :=  \sum^N_{l=0} \sum^l_{m=-l} f_{l,m} \, \Ylm(x,y,z)
}
We are now in a position to derive the sparse differential operators that can be applied to the functions $\uvec, f$.

\begin{definition}\label{def:SHoperators}
	We define the operator matrices $\cal{D}, \calG, \calL$ according to:
\bseqn{
	\divergence \uvec(x,y,z) &= \SHvec(x,y,z)^\top \: \calD \: \uvecc, \\
	\gradS f(x,y,z) &= \VSHvec(x,y,z)^\top \: \calG \: \fvecc, \\
	\DeltaS f(x,y,z) &= \SHvec(x,y,z)^\top \: \calL \: \fvecc.
}
\end{definition}

\begin{theorem}
	The operator matrices defined in \bsrefdef{def:SHoperators} are diagonal, and are given by:
\bseqn{
	\calG &= I_{(N+1)^2}, \quad
	\calD \equiv \calL := 
		\begin{pmatrix}
			\calL_0 & & \\
			& \ddots & \\
			& & \calL_N
		\end{pmatrix},
}
where
\bseqn{
	\calL_l &:= 
		\begin{pmatrix}
			-l(l+1) & & \\
			& \ddots & \\
			& & -l(l+1)
		\end{pmatrix} \in \R^{(2l+1) \times (2l+1)}.
}
\end{theorem}

\begin{proof}
	The argument for $\calG$ is trivial by definition. Now, the spherical harmonics, as the name eludes to, satisfy a harmonic relationship $\DeltaS \Ylm = \: l(l+1) \Ylm$ \bstodo{cite}. Thus, for the proof for $\calD$, consider
\bseqn{
	\divergence \uvec &= \divergence \Big( \sum_{l=0}^N \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ] \Big) \\
	&= \sum_{l=0}^N \sum_{m=-l}^{l} u_{l,m} \: \DeltaS \Ylm \\
	&= \sum_{l=0}^N \sum_{m=-l}^{l} - u_{l,m} \: l(l+1) \Ylm,
}
using the fact that $\divergence (\grad^\perp f )\equiv 0$ for any function $f$. We similarly have for $\calL$:
\bseqn{
	\DeltaS f &= \DeltaS \Big( \sum_{l=0}^N \sum_{m=-l}^{l} f_{l,m} \Ylm \Big) \\
	&= \sum_{l=0}^N \sum_{m=-l}^{l} f_{l,m} \: \DeltaS \Ylm \\
	&= \sum_{l=0}^N \sum_{m=-l}^{l} - f_{l,m} \: l(l+1) \Ylm.
}
\end{proof}

Now that we have our differential operators, let's apply them in an example.

\subsection{Example: linear shallow water equations}

The linearised shallow water equations allow us to showcase the differential operators that we defined in \bsrefdef{def:SHoperators} (namely the divergence and gradient operators) and demonstrate how the natural sparsity that our framework brings to the problem leads to a simple sparse linear system to solve.

Let $\uvec(x,y,z)$ be the tangential velocity of a flow and $h(x,y,z)$ be the height deviation of the flow from some constant reference height $\calH$. Define $\rvec$ as the unit outward normal vector at the point on the sphere $(x,y,z)$, so that
\bseqn{
	\rvec = \begin{pmatrix} x \\ y \\ z \end{pmatrix}.
}
The linear SWEs are
\bseqn{
	\begin{cases}
		\pfpx{t}{\uvec} + f \rvec \times \uvec - \nabla h = \mathbf{0} \\
		\pfpx{t}{h} + \calH \nabla \cdot \uvec = 0
	\end{cases}
}
where $f = 2 \rotrate \cos(\varphi) =  2 \rotrate z$ is the Coriolis parameter, and $\rotrate = 7.2921 × 10^{-5}$ rad s$^{-1}$ is the rotation rate of the Earth (note that $\varphi = {\pi \over 2} - \text{latitude}$).

A keen eye may notice that we require one more operator not yet defined -- one for the Coriolis parameter cross product (i.e. $f \: \rvec \times \uvec$). 
\begin{definition}\label{def:coriolisoperator}
	Define the operator matrices $\calF, \calF^\perp$ according to:
\bseqn{
	f \: \rvec \times \uvec(x,y,z) &= \VSHvec(x,y,z)^\top \: \calF \: \uvec^{\perp c} + \Big(\VSHvecperp(x,y,z)\Big)^\top \: \calF^\perp \: \uvec^{c}
}
\end{definition}
\begin{lemma}
	The operators $\calF, \calF^\perp$ defined in \bsrefdef{def:coriolisoperator} are sparse with banded-block-banded structure. More specifically,
\begin{itemize}
	\item $\calF$ has block-bandwidths $(1,1)$, and sub-block-bandwidths $(1,1)$
	\item $\calF^\perp$ has block-bandwidths $(1,1)$, and sub-block-bandwidths $(1,1)$
\end{itemize}
\end{lemma}
\begin{proof}
	Using \bsreflemma{lemma:VSHrecurrences}, we have that:
\bseqn{
	f \: \rvec \times \bold{u} &= f \: \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ] \\
	&= 2 \: \rotrate \: z \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} \rvec \times \gradYlm + u^{\perp}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} z \: \gradpYlm - u^{\perp}_{l,m} z \: \gradYlm ) ] \\
	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} \Big[ u_{l,m} \: \Big( \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l} \Big) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad - u^{\perp}_{l,m} \: \Big( \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l} \Big) \Big].
}
We can then write down the operators $\calF, \calF^\perp$ as follows:
\bseqn{
	\calF &= \bar{\calF^\perp} = 2 \: \rotrate \: 
		\begin{pmatrix}
			\calF_{0,2} & \calF_{1,1} & & & & \\
			\calF_{0,3} & \calF_{1,2} & \calF_{2,1} & & & \\
			& \calF_{1,3} & \calF_{2,2} & \calF_{3,1}  & & & \\
			& & \calF_{2,3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}
}
where $\bar{A}$ denotes the conjugate matrix of $A$ (i.e. where the entries of $\bar{A}$ are simply the complex conjugates of the entries of $A$) \cite[p.10]{courant1953methods}, and for $l = 0,\dots,N$
\bseqn{
	\calF_{l,1} &:= 
		\begin{pmatrix}
			0 & \Gamma_{l,-l+1,1} & & \\
			& & \ddots & & \\
			& & & \Gamma_{l,l-1,1} & 0 \\
		\end{pmatrix} \in \R^{(2l-1)\times(2l+1)}, \quad (l \ne 0), \\
	\calF_{l,2} &:= 
		\begin{pmatrix}
			\Gamma_{l,-l,2} & & \\
			& \ddots & \\
			& & \Gamma_{l,l,2} \\
		\end{pmatrix}  \in \R^{(2l+1)\times(2l+1)}, \\
	\calF_{l,3} &:= 
		\begin{pmatrix}
			0 \\
			\Gamma_{l,-l,3} \\
			& & \ddots \\
			& & & & \Gamma_{l,l,3} \\
			& & & & 0 \\
		\end{pmatrix} \in \R^{(2l+3)\times(2l+1)}.
}
\end{proof}

For simplicity, we implement a backward Euler timestepping method to solve the linear SWEs with timestep $\Delta t$:
\bseqn{
	\uvec_{n+1} &= \uvec_{n} + \Delta t \: (\nabla h_{n+1} - f \rvec \times \uvec_{n+1}) \\
	h_{n+1} &= h_n - \Delta t \: \calH \nabla \cdot \uvec_{n+1}
}
In coefficient space, this system becomes:
\bseqn{
	\uvec^c_{n+1} &= \uvec^c_{n} + \Delta t \: (\calG \vec{h}^c_{n+1} - \calF \uvec^{\perp c}_{n+1}) \\
	\uvec^{\perp c}_{n+1} &=  - \Delta t \: \calF^\perp \uvec^{c}_{n+1} \\
	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvec^c_{n+1}
}
$\iff$
\bseqn{
	\uvec^c_{n+1} &= \uvec^c_{n} + \Delta t \: (\calG \vec{h}^c_n - \Delta t \: \calH \calG \calD \uvec^c_{n+1} + \Delta t \calF \calF^\perp \uvec^{c}_{n+1}) \\
	\uvec^{\perp c}_{n+1} &=  - \Delta t \: \calF^\perp \uvec^{c}_{n+1} \\
	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvec^c_{n+1}
}
$\iff$
\bseqn{
	\uvec^c_{n+1} &= (I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp)^{-1} \: (\uvec^c_{n} + \Delta t \: \calG \vec{h}^c_n) \\
	\uvec^{\perp c}_{n+1} &=  - \Delta t \: \calF^\perp \uvec^{c}_{n+1} \\
	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH \calD \uvec^c_{n+1}.
}
Here, it is clear that the matrix $I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve. Doing so at each iteration provides us with the coefficients vectors $\uvec^{c}_{n+1}, \uvec^{\perp c}_{n+1}, \vec{h}^{c}_{n+1}$ if we know $\uvec^{c}_{n}, \uvec^{\perp c}_{n}, \vec{h}^{c}_{n}$.

\bstodo{Image}

\section{Appendix}

\bibliography{../includes/thesis-ben-snowball}

\end{document}

### 2020-10-22 papers/thesis/full/ch1-introduction/introduction.tex
New thesis structure - commiting important files

\chapter{Introduction}

Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs}. While there are many famous examples -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few -- the area of multivariate orthogonal polynomials has a smaller array of research. However, this branch of mathematics has an encouraging future \cite{dunkl2014orthogonal}, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains, such as on the triangle \cite{olver2019triangle} with applications including solving Volterra integral equations \cite{gutleb2020sparse}.

One famous group of multidimensional orthogonal polynomials is of course the Spherical Harmonics. The European Centre for Medium-range Weather Forecasts (ECMWF) use a Spherical Harmonic spectral method in their weather and climate model \cite{cheong2006dynamical}. Chapter 2 of this thesis provides an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface. We present a framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion. Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.

While the whole sphere spectral method has been successful for numerous years \cite{wedi2013fast}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the ECMWF model \cite{ecmwf2020scalability}. This thesis serves to lay a foundation for addressing this while still utilising a spectral approach. More precisely, we develop a sparse spectral method for solving partial differential equations (PDEs) on the spherical cap as a surface in 3D, with a simple extension to a spherical band. Together, these frameworks can be pieced together to create a spectral element method for the whole sphere, or further developed to investigate spectral methods on other spherical subdomains.

To this end, in Chapter 3 we move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles. Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials, which we define. Triangles and the full disk are then special cases of our framework, which we formalise for the disk-slice and trapezium cases.

With a greater knowledge base in our quiver, we can adapt the techniques learnt from the founding of the disk-slice formulation to a surface in 3D in Chapter 4 -- in particular the spherical cap, a subdomain of the surface of a unit sphere. Using the same family of (non-classical) 1D OPs, we present a suitable orthogonal polynomial basis once more, this time as polynomials in $x,y,z$ complete with sparse differential operators.

### 2020-10-22 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
New thesis structure - commiting important files

--- 2020-10-16 papers/thesis/ch2-spherical-harmonics/spherical-harmonics.tex
@@
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
-%\geometry{landscape}                		% Activate for rotated page geometry
-\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
-								% TeX will automatically convert eps --> pdf in pdflatex
-
-\input{../includes/bsmacros}
-\input{../includes/bsthesisnewcommands}
-\input{../includes/somacros}
-
-\title{Spherical harmonics as orthogonal polynomials in three variables}
-\author{Ben Snowball}
-%\date{}							% Activate to display a given date or no date
-
-\begin{document}
-
-\maketitle
-
-\section{Spherical harmonics as orthogonal polynomials in three variables}
-
-\subsection{Introduction}
+\chapter{Spherical harmonics as orthogonal polynomials in three variables}
 
@@
 
-\subsection{Defining the spherical harmonics in three variables}
+\section{Defining the spherical harmonics in three variables}
 
@@
 
-\subsection{Jacobi matrices}
+\section{Jacobi matrices}
 
@@
 }
-for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylmfull$ by $z$ is
-\bseqn{
-	z\,\Ylm(x,y,z) &= \clm \eimphi \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
+for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is
+\bseqn{
+	z\,\Ylmxyz &= \clm \eimphi \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
 		     &= \clm \eimphi \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
@@
 }
-for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylmfull$ by $x$ and $y$ is then:
-\bseqn{
-	x\,\Ylm(x,y,z) &= \clm \cosphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
+for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ is then:
+\bseqn{
+	x\,\Ylmxyz &= \clm \cosphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
 	&= \half \clm (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
@@
 	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
-	y\,\Ylm(x,y,z) &= \clm \sinphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
+	y\,\Ylmxyz &= \clm \sinphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
 	&= -\half i \: \clm (e^{i(m+1)\varphi} - e^{i(m-1)\varphi}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
@@
 
-\subsection{Three-term recurrence relation for $\SHvec$}
+\section{Three-term recurrence relation for $\SHvec$}
 
@@
 	\begin{pmatrix}
-		Y_0 \\ 0 \\ 0 \\ 0 \\ \vdots  \\
+		Y_0 \\ 0 \\ 0 \\ 0 \\ \vdots
 	\end{pmatrix},
 }
-where we note $Y_0^0(x,y,z) \equiv Y_0 := P_0^{(0,0)} \equiv 1$, and for each $l = 0,1,2\dots$,
+where we note $Y_0^0(x,y,z) \equiv Y_0 := c_0^0 \: P_0^{(0,0)} \equiv \half {1 \over \sqrt{\pi}}$, and for each $l = 0,1,2\dots$,
 \begin{align}
@@
 
-\subsection{Computational aspects}
-
-\bstodoinline{quad rule on sphere?}
-
-\subsubsection{Function evaluation}\label{subsubsection:clenshaw}
+\section{Computational aspects}
 
@@
 \bseqn{
-	f(x,y,z) \approx \SHvec(x,y,z)^\top \fvec = \sum^N_{l=0} \SHvecl(x,y,z)^\top \fvec_l,
-}
-where $\SHvecl (x,y,z), \fvec_l \in \C^{2l+1}$ for each  $l \in \{0,\dots,N\}$, for some coefficients vector $\fvec$ up to degree order $N \in \N$.
-We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}.
-The \textit{Clenshaw algorithm} is then as follows:
+	f(x,y,z) \approx \SHvec(x,y,z)^\top \fvec = \sum^N_{l=0} \SHvecl(x,y,z)^\top \fvec_l =  \sum^N_{l=0}  \sum^l_{m=-l} f_{l,m} \: \Ylmxyz,
+}
+where $\SHvecl (x,y,z), \fvec_l \in \C^{2l+1}$ for each $l \in \{0,\dots,N\}$, for some coefficients vector $\fvec = (f_{l,m})$ up to degree order $N \in \N$.
+
+\subsection{Obtaining coefficients}
+
+\bstodoinline{obtaining coeffs/ transform}
+In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
+Naturally, we of course need a way to obtain the coefficients $f_{l,m}$.
+The coefficients can be calculated via the integral
+\bseqn{
+	f_{l,m} = \int_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta
+}
+using the orthonormality of the spherical harmonics.
+Methods to calculate these coefficients exist, known as spectral transforms, and are well established for the spherical harmonics (see e.g.
+\cite{mohlenkamp1999fast, suda2002fast}).
+
+\subsection{Function evaluation}\label{subsubsection:clenshaw}
+
+We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}. The \textit{Clenshaw algorithm} is then as follows:
 \bseqn{
@@
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \fvec_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
-}
-where $\SHvec_0 \equiv \alpha^0_0$.
-
-\subsubsection{Function evaluation}
+	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0.
+}
+
+\subsection{Variable coefficients as operators}
 
@@
 Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}.
-Other definitions for VSHs are used, however it is convenient for deriving explicit sparse relations and operators to use the one described here.
+Vector spherical harmonics have been widely used in electrostatics (e.g.
+\cite{barrera1985vector}), electrodynamics (e.g.
+\cite{carrascal1991vector}), and of course fluid dynamics including weather and climate modelling (e.g.
+\cite{moses1974use, fengler2005nonlinear, swarztrauber1996spectral}).
+Other definitions for VSHs are used (e.g.
+\cite{hill1954theory}), however it is convenient for deriving explicit sparse relations and operators to use the one described here.
 
@@
 
-\subsection{Function evaluation}
-
-The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$.
+\subsection{Computational aspects}
+
 Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
@@
 Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent space of $\Omega$}.
-We can then approximate this function by its expansion in the VSHs basis, i.e.
+We can then approximate this function by its expansion in the VSH basis, i.e.
 for some large enough $N \in \N$:
@@
 }
-The clenshaw algorithm to evaluate $\uvec$ at a point $(x,y,z) \in \Omega$ is then as follows
+
+\subsubsection{Obtaining coefficients}
+The development of vector spherical harmonic transforms is well established \cite{swarztrauber1993vector}, and continues today (see \cite{gia2019favest}). \bstodoinline{VSH Transform}
+
+\subsection{Function evaluation}
+The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$:
 \bseqn{
@@
 
-\subsection{Example: linear shallow water equations}
+\section{Example: linear shallow water equations}
 
@@
 \bstodo{Image}
-
-\section{Appendix}
-
-\bibliography{../includes/thesis-ben-snowball}
-
-\end{document}

### 2020-10-22 papers/thesis/full/conclusion/summary.tex
New thesis structure - commiting important files

\chapter{Summary and future directions}

\section{Summary}

In this thesis, we have developed sparse spectral methods for solving partial differential equations (PDEs) on disk-slices and trapeziums in 2D, and spherical caps as a surface 3D. The work can also be used as a template for developing similar methods on other such similar domains, in particular other subdomains of the sphere.

We began with an introduction to multidimensional sparse spectral methods by looking at the spherical harmonics on the whole sphere, explaining how they can be written as orthogonal polynomials in $x,y,z$, and how Jacobi and differential operators that apply to coefficient vectors will hence be banded-block-banded.

For the disk-slice in 2D, we defined the OPs that allow similar sparse and banded-block-banded operators required for solving PDEs in the domain, deriving their structure and providing a method to efficiently calculate numerically their entries. The reason they need to be calculated numerically is due to the non-classical univariate OPs that are involved in the 2D OP definitions. Finally, we moved on to use the same arguments for the spherical cap, a 3D surface that is a subdomain of the unit sphere. We once again defined the 3D OPs as orthogonal polynomials in in $x,y,z$, and showed how the differential operator matrices continue to elicit similar banded-block-banded structure.

\section{Future directions}

We hope that this thesis can serve as somewhat of a blueprint for formalising sparse spectral methods, including how one should define the OPs and derive the accompanying Jacobi and differential operators. The motivation behind this work was to develop sparse spectral methods for subdomains of the unit sphere in 3D space. Thus, the natural direction from this point would be to formalise the framework for spherical bands, which would simply involve slightly more complex arithmetic and the sub-block bands would be slightly larger (this due to the $\genjac$ polynomials we defined now having no zero coefficients in their three-term recurrences, meaning the expressions for multiplication by $x,y,z$ would gain some extra terms).

Ideally, we would like to pair this with a spectral element method for the sphere, where the elements would be the spherical subdomains of spherical bands and spherical caps. The goal here would be to test this method against the full sphere spectral method the is in place in the ECMWF model, with the hypothesis being that by reducing the size of the transforms, we can improve the overall efficiency while still maintaining the accuracy that we achieve from a spectral approach.

### 2020-10-25 papers/spherical-caps/spherical-caps.tex
Comments to Page 20

@@
 
-\title{Sparse spectral methods for partial differential equations on spherical caps and bands}
+\title{Sparse spectral methods for partial differential equations on spherical caps}
 \author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
@@
 In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, disk-slices and triangles.
-In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps and bands.
+In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps.
 The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
@@
 For clarity we note that, for all examples and plots in this paper, we take $\alpha = 0.2$.
-\bstodo{removed "band" from title, and added a bit more here about the extension}
 
@@
 	\includegraphics[scale=0.3]{complexity-Nend=1000}
-	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos(z)$.}
+	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$.
+\sotodoinline{Move to experiments section} \sotodoinline{Add solve time}}
 	\centering
@@
 This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity}.
-This plot presents the time taken (measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia) to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos(z)$, as we increase the orthogonal polynomial degree $p$.
-\bstodoinline{Added plot}
+This plot presents the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos z$, as we increase the orthogonal polynomial degree $p$.
 
@@
 
-\noindent\secref{PDOs}: We define useful partial differential operators and further show that these will also be sparse.
+\noindent\secref{PDOs}: We define several partial differential operators such as spherical Laplacians and  show that these will  be sparse when applied to a suitable choice of expansions in bases built from OPs on the spherical cap.
 We can exactly calculate the non-zero entries of these sparse operators using the quadrature rule associated with the non-classical 1D OPs.
@@
 	Let $w : (\alpha,\beta) \: \to \R$ be a weight function.
-$\forall$, $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$.
+For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$.
 Then the 3D polynomials defined on $\Omega$
@@
 \end{align*}
-on $\Omega$. 
+on $\Omega$, where $\D \sigma(\xvec, z)$ is the Lebesgue measure\sotodo{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} on $\Omega$. 
 \end{proposition}
@@
 \end{align*}
-where $\D \sigma(\xvec, z)$ is the Lebesgue measure on $\Omega$, with
+with
 \begin{align}
@@
 Importantly, we can deduce sparsity in the recurrence relationships.
-We first require the following Lemma.
+We first require the following lemma.
 
@@
 \end{lemma}
+
 \begin{proof}
@@
 \begin{align*}
-	\alphaa_{n,k,1} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
-	\alphaa_{n,k,2} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
-	\alphaa_{n,k,3} &:=\eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
-	\alphaa_{n,k,4} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
-	\alphaa_{n,k,5} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}} \\
-	\alphaa_{n,k,6} &:=\eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}} \\
+	\alphaa_{n,k,1} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
+	\alphaa_{n,k,2} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
+	\alphaa_{n,k,3} &:=\eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
+	\alphaa_{n,k,4} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
+	\alphaa_{n,k,5} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
+	\alphaa_{n,k,6} &:=\eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
 	\betaa_{n,k,i,j} &:= 
@@
 			\alphaa_{n,k,j} \quad &\text{otherwise}
-		\end{cases} \\	
-	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}
+		\end{cases}, \\	
+	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}.
 \end{align*}
 \end{lemma}
-
+\begin{remark}
+For $z$ multiplication, note that different Fourier modes do not interact. This is because $z$ is rotationally invariant. 
+\end{remark}
 \begin{proof}
@@
 \end{align*}
-where we show the non-zero coefficients that result are the $\alphaa_{n,k,1},\dots,\alphaa_{n,k,6}$ in the Lemma.
+where we show the non-zero coefficients that result are the $\alphaa_{n,k,1},\dots,\alphaa_{n,k,6}$ in the lemma.
 Recall from equation (\ref{eqn:normscop}) that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$.
-Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos(\theta) \sin(\varphi), \: \sin\theta\sin(\varphi), \: \cos(\varphi)) = (x, y, z)$:
+Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)$:
 \begin{align*}
@@
 \end{align*}
-where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the Lemma:
+where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the lemma:
 \begin{align*}
@@
 
-The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define: 
+The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$.
+Let $N \in \N$ and define notation that groups degree $N$ polynomials together[:
 \begin{align}
@@
 \end{align}
-and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
+and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to\sotodo{The following is wrong as it only applies if $N = \infty$}
 \begin{align}
-	J_x^{(a,b,c,d)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
-	J_y^{(a,b,c,d)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
-	J_z^{(a,b,c,d)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
+	J_x^{(a)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
+	J_y^{(a)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
+	J_z^{(a)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
 \end{align}
@@
 Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis.
+In other words, ${J_x^{(a)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis.
 Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
@@
 			& & A^{(a)}_{N,k,2} & A^{(a)}_{N,k,4}
-		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
-	&\big( \text{so that } A^{(a)}_{x,0} \in \R^{(N+1)\times2N}, \quad A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \big) \\
+		\end{pmatrix}, \quad k = 0,1,2,\dots,N 
+\end{align*}
+so that  $A^{(a)}_{x,0} \in \R^{(N+1)\times2N}$,  $A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)}$ for $k = 1,2,\dots,N$,
+\begin{align*}
 	C^{(a)}_{x,k} &:= 
@@
 			& & & A^{(a)}_{N,k,1} & A^{(a)}_{N,k,3}
-		\end{pmatrix}, \quad k = 1,2,\dots,N \\
-	&\big( \text{so that } C^{(a)}_{x,1} \in \R^{2N\times(N+1)}, \quad C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+2)} \text{ for } k = 2,3,\dots,N \big)
-\end{align*}
-where for $k = 1,\dots,N, \: n = k,\dots,N$
+		\end{pmatrix}, \quad k = 1,2,\dots,N 
+\end{align*}
+so that $C^{(a)}_{x,1} \in \R^{2N\times(N+1)}$,  $C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+2)}$  for  $k = 2,3,\dots,N$, here for $k = 1,\dots,N, \: n = k,\dots,N$,
 \begin{align}
@@
 
-Define, for $n=0,1,2,\dots$: 
+Define, for $n=0,1,2,\dots$:  \sotodo{Explain why these are needed}
 \begin{align*}
@@
 \end{align*}
-and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to
+and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to \sotodo{Can the following be simplified? Seems like it repeats}
 \begin{align}
-	\tilde J_x^{(a,b,c,d)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
-	\tilde J_y^{(a,b,c,d)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
-	\tilde J_z^{(a,b,c,d)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
+	\tilde J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
+	\tilde J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
+	\tilde J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
 \end{align}
@@
 
-We can combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
+Following the triangle case \cite{olver2019triangle}, we can combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
 \begin{align*}
@@
 
-In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short Lemma.
+In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short lemma.
 
@@
 \begin{align*}
-	\int_0^{2\pi} g\big(\cos(\theta), \sin\theta\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
+	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
 \end{align*}
@@
 	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
 	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \D x \Big) \: \D z \\
@@
 	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) + f_o\big(\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
@@
 %	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos(\theta)\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
+%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
 %	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_e\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \\
@@
 %	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos(\theta)\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos(\theta),\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
+%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
+%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
 %	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_o\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \quad (\dagger) \\

### 2020-10-26 papers/spherical-caps/spherical-caps.tex
Finish proof read

@@
 		0 & k = 0
-	\end{cases}
-\end{align*}
-
-We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$.
-The entries of the operator are given by the coefficients in the expansion $\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)}$, where the coefficients are
+	\end{cases}.
+\end{align*}
+
+We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$. The entries of the operator are given by the coefficients in the expansion 
+$$
+\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)},
+$$ 
+where the coefficients are 
 \begin{align*}
@@
 
-We move onto the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yields
+We move on to the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yield
 \begin{align}
@@
 
-For the operator $\calL^{(a)\to(a+\tilde a)}$ for the surface Laplacian on a non-weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \scopnki^{(a)} = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a + \tilde a)}$, where the coefficients are
+For the operator $\calL^{(a)\to(a+\tilde a)}$ for the surface Laplacian on a non-weighted space, the entries of the operator are given by the coefficients in the expansion 
+$$\DeltaS \scopnki^{(a)} = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a + \tilde a)},$$
+ where the coefficients are
 \begin{align*}
@@
 \begin{definition}\label{def:parametertransformationoperators}
-Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
+Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T_W^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
 \begin{align*}
@@
 	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
-For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 200$.}
+For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Shouldn't this be $\sqrt{1-x_0^2-y_0^2}$?} and $k = 200$.}
 	\label{fig:sparsity}
@@
 \end{align*}
-The trivariate Clenshaw algorithm works similar to the bivariate Clenshaw algorithm introduced in \cite{olver2019triangle}:
+The trivariate Clenshaw algorithm works similar to the bivariate Clenshaw algorithm introduced in \cite{olver2019triangle} for expansions in the triangle:
 \begin{align*}
@@
 Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
-This demonstrates spectral convergence observed.
-	\bstodoinline{New solution block norms plots.
-I have only gone up to N=100 (200 for Poisson) for now, mainly because expanding a RHS function f for larger N takes a looong time, which is where we would see the benefit of a higher N plot (for the simpler $f$'s, we would just see more super convergence).}}
+This indicates spectral convergence.}
 	\centering
@@
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
+	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Here too} and $k = 100$.
 Right: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
-This demonstrates spectral convergence observed.}
+This indicates spectral convergence.
+\sotodoinline{Since this example has simple right-hand side, wouldn't it be easy to extend to $N = 200$?}}
 	\centering
@@
 	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$, for various $\epsilon$ values.
+Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Here too}, for various $\epsilon$ values.
 This demonstrates algebraic convergence observed.}
@@
 We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
-We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $V(X, Y, Z)$.
+We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $v(X, Y, Z)$.
 In matrix-vector notation, our system hence becomes:
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ and $k = 100$.
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodo{Here too} and $k = 100$.
 In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
@@
 and solving this zero boundary condition Helmholtz problem.
+
+\sotodoinline{
+Wouldn't this trick work for any boundary conditions $c(x,y)$?
+That is, expand in Fourier series (a polynomial in $x$ and $y$) and use that to extend off the $z=0$ axis.
+If you agree we should explain that here.
+}
 
@@
 \end{align}
+\sotodoinline{I think we may need to explain what ${\partial u \over\partial n}$ means here: is the normal derivative well defined for 2D surfaces embedded in 3D?}
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
@@
 We have shown that trivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary.
-We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: \sqrt{(1-z^2)}^b$ defined for the disk-slice case \cite{snowball2019sparse}.
-This  also forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.
+We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: \sqrt{(1-z^2)}^b$ \sotodo{Should this be written as $(1-z^2)^{b/2}$?} defined for the disk-slice case \cite{snowball2019sparse}.
+Generalisation to spherical bands ($\alpha \leq z \leq \beta$) is straightforward.
+This work thus forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.
 
@@
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on spherical bands and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
-
-The next stage is to develop an orthogonal basis for the tangent space of the spherical cap/band, and obtain sparse differential operators for gradient, divergence etc.
+Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
+
+The next stage is to develop an orthogonal basis for the tangent space of the spherical cap (or band), and obtain sparse differential operators for gradient, divergence etc.
 On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.

### 2020-10-28 OrthogonalPolynomialFamilies/src/SphericalCaps.jl
Housekeeping with files. Edits to experiments file.

@@ -1801,9 +1801,9 @@ function laplacianoperator(S::SphericalCapSpace{<:Any, B, T, <:Any},
 
     # Begin assigning entries
     for k = 0:N
-        if k % 20 == 0
-            @show "Δ", weighted, k
-        end
+        # if k % 20 == 0
+        #     @show "Δ", weighted, k
+        # end
         R = getRspace(S, k)
         getopptseval(R, N-k+band2, ptsr)
         getderivopptseval(R, N-k+band2, ptsr)

### 2020-10-28 experiments/sphericalcaptests-plots.jl

@@ -296,11 +296,11 @@ function getsolutionblocknorms(N::Int, S::SphericalCapSpace, A::BandedBlockBande
 end
 
 # solutionblocknorms - Poisson
-N = 200
+N = 100
 Δw = squareoperator(S, laplacianoperator(S, S, N), N)
 
-ϵ = 0.5; f5anon = (x,y,z)->norm([x; y; z] .- (1.0 * (1 / sqrt(3) + ϵ)))
-f5 = Fun(f5anon, S, 2115*2); f5.coefficients
+ϵ = 0.1; f5anon = (x,y,z)->norm([x; y; z] .- (1.0 * (1 / sqrt(3) + ϵ)))
+f5 = Fun(f5anon, S, 2(N+1)^2); f5.coefficients
 f5(p) - f5anon(p...)
 ϵ = 1; f1anon = (x,y,z)->norm([x; y; z] .- (1.0 * (1 / sqrt(3) + ϵ)))
 f1 = Fun(f1anon, S, 500*2); f1.coefficients
@@ -325,7 +325,7 @@ unorms = getsolutionblocknorms(N, S, Δw,
                                     resizecoeffs!(S, f2, N), resizecoeffs!(S, f3, N),
                                     resizecoeffs!(S, f4, N)))
 using Plots
-Plots.plot(unorms[1], line=(3, :solid), label="ϵ = 0.5", yscale=:log10, legend=:bottomleft)
+Plots.plot(unorms[1], line=(3, :solid), label="ϵ = 0.1", yscale=:log10, legend=:bottomleft)
 Plots.plot!(unorms[2], line=(3, :dash), label="ϵ = 1")
 Plots.plot!(unorms[3], line=(3, :dashdot), label="ϵ = 2")
 Plots.plot!(unorms[4], line=(3, :dot), label = "ϵ = 3")
@@ -337,7 +337,7 @@ Plots.savefig("experiments/saved/sphericalcap/images/solutionblocknorms-poisson-
 # Helmholtz
 f1 = Fun((x,y,z)->1.0, S, 2(1+1)^2); f1.coefficients
 f2 = Fun((x,y,z)->weight(S, x, y, z) * S.family.ρ(z)^2, S, 2(3+1)^2); f2.coefficients
-N = 100 # TODO Run with N larger
+N = 200 # TODO Run with N larger
 Jx = OrthogonalPolynomialFamilies.jacobix(S, N)
[…]

### 2020-10-28 papers/spherical-caps/spherical-caps.tex
Edits to spherical cap paper. Added comments and new figs

@@
 
- \begin{figure}[t]
-	\centering % <-- added
-	\includegraphics[scale=0.3]{complexity-Nend=1000}
-	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$.
-\sotodoinline{Move to experiments section} \sotodoinline{Add solve time}}
-	\centering
-	\label{fig:complexity}
-\end{figure}
-
 Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + V(x,y,z)$ where $V$ is first approximated by a polynomial.
@@
 \begin{align*}
-	\ip< p, \: q > &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: w(z) \: \D \sigma(\xvec, z) \\
-	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: w(z) \: \D z
-\end{align*}
-on $\Omega$, where $\D \sigma(\xvec, z)$ is the Lebesgue measure\sotodo{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} on $\Omega$. 
+	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
+	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
+	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
+\end{align*}
+on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the Lebesgue measure on $\Omega$. 
+\sotodoinline{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} 
+\bstodoinline{I have changed it to dA and elaborated a bit more on it, but I am not sure on this so you could be correct that another term needs to be used.
+Your quadratic surfaces paper with Yuan says Lebesgue I think when discussing this.
+I have tried to keep it general here by using a measure definition, but we could omit this if its just going to cause confusion?}
 \end{proposition}
@@
 \end{align}
-and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to\sotodo{The following is wrong as it only applies if $N = \infty$}
+and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
+\sotodoinline{The following is wrong as it only applies if $N = \infty$} 
+\bstodoinline{What is better?
+To define them in this way for $N \to \infty$?
+Or to include the extra terms in this defn below?
+Or just state that the Jacobi operators approximate multiplication by x,y,z?
+This also relates to the next comments}
 \begin{align}
@@
 
-In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices} -- we define the Clenshaw matrices as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the multidimensional three-term recurrences.
-While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
+In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \textit{recurrence coefficient matrices} -- we define these as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the coefficients for the multidimensional three-term recurrences.
+
+While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the recurrence coefficient matrices necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
-
-Define, for $n=0,1,2,\dots$:  \sotodo{Explain why these are needed}
+To this end, we define our OP-building ordering as follows.
+For $n=0,1,2,\dots$:
 \begin{align*}
@@
 \end{align*}
-and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to \sotodo{Can the following be simplified? Seems like it repeats}
+and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to 
+\sotodoinline{Can the following be simplified? Seems like it repeats} 
+\bstodoinline{Is it worth removing the previous?
+And just explaining that we can reorder the Jacobi matrices according to the OP/coefficient vector ordering?
+This section is important for deriving how to build the OPs, whereas before I simply need to say that the Jacobi operators that apply to coefficients vectors are reorderings of the ones in this section.}
 \begin{align}
@@
 
-Note that we can define an explicit \(\Dnt\) as follows:
+Note that we can define an explicit $\Dnt$ as follows:
 \begin{align*}
@@
 \begin{definition}\label{def:clenshawmats}
-The Clenshaw matrices associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
+	The \textit{recurrence coefficient matrices} associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
 \end{definition}
@@
 \begin{theorem}\label{theorem:sparsityofdifferentialoperators}
-	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
+	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
 More specifically:
@@
 Finally, fix $a = 1$.
-For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are
+For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are given by
 \begin{align*}
@@
 	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(a, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(a, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(a, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(a, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
-	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(1,2k)} \: \genjacnmk^{(a, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \Big\} \: \D z
+	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(1, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
+	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(1, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(1, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(1, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
+	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(1, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z
 \end{align*}
@@
 
-There exist conversion matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.
-
+By applying these differential operators, we are (in some cases) incrementing or decrementing the parameter value $a$.
+It is therefore necessary to also be able to raise or lower the parameter by way of an independent operator.
+There exist conversion matrix operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 \begin{definition}\label{def:parametertransformationoperators}
@@
 		\centering
-		\caption{The Laplace operator $\Delta^{(1)}_W$}
+		\caption{The Laplace-Beltrami operator $\Delta^{(1)}_W$}
 	\end{subfigure}\hfil % <-- added
@@
 		\centering
-		\caption{The $\rho^2$-factored Laplacian $L_W^{(1)}$}
+		\caption{The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$}
 	\end{subfigure}\hfil % <-- added
@@
 		\centering
-		\caption{The Biharmonic operator $B_W^{(2)}$} 
+		\caption{The Biharmonic operator $\calB_W^{(2)}$} 
 	\end{subfigure}\hfil % <-- added
 	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
-For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Shouldn't this be $\sqrt{1-x_0^2-y_0^2}$?} and $k = 200$.}
+For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
 	\label{fig:sparsity}
@@
 
+\subsection{Further partial differential operators}\label{subsection:furtherdiffoperators}
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
-As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored surface Laplacian $\rho(z)^2 \: \DeltaS$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
+As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored spherical Laplacian $\rho(z)^2 \: \DeltaS$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
 Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in Definition \ref{def:differentialoperators}.
-The matrix operator for this $\rho^2$-factored Laplacian we denote $\mathcal{L}_W^{(1)}$ acting on the coefficients vector is then given by
-\begin{align*}
-    \mathcal{L}_W^{(1)} := D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
+The matrix operator for this $\rho^2$-factored spherical Laplacian acting on the coefficients vector is then given by
+\begin{align*}
+	D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
 \end{align*}
@@
 \begin{align*}
-	u(x,y,z) &= 0 \quad \text{for } (x,y,z) \in \Omega, \\
-	\pfpx{n}{u}(x,y,z) &= \nabla_S u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
+	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \nabla_S u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
 \end{align*}
 where $\partial\Omega$ is the $z=\alpha$ boundary, and $\unitvec{n}(x,y,z)$ is the outward unit normal vector at the point $(x,y,z)$ on the boundary, i.e.
-$\unitvec{n}(\vec{x}) = {\vec{x} \over \norm{\vec{x}}}$.
+$\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$.
 The matrix operator for the Biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
@@
 \end{align*}
-Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions.
+Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,u,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$ and thus its spherical gradient would be zero on the boundary $z = \alpha$.
 This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
@@
 % FIGURES
+
+\begin{figure}[t]
+	\centering % <-- added
+	\includegraphics[scale=0.35]{complexity-Nend=1000}
+	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$.
+This shows that the time to build and solve the system $\big[\DeltaS + V(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
+Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.
+\bstodoinline{Redo with belapsed}}
+	\centering
+	\label{fig:complexity}
+\end{figure}
+
 \begin{figure}[t]
@@
 Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
-This indicates spectral convergence.}
+This indicates spectral convergence.
+\bstodoinline{Plots increased to N=200 for each}}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-varyingk-N=100}
+	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-varyingk-N=200}
         	%\label{fig:solutionblocknorms-poisson}
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Here too} and $k = 100$.
+	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
 Right: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
-This indicates spectral convergence.
-\sotodoinline{Since this example has simple right-hand side, wouldn't it be easy to extend to $N = 200$?}}
+This indicates spectral convergence.}
 	\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-expfun-N=100}
+	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-expfun-N=200}
 	\centering
@@
 	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Here too}, for various $\epsilon$ values.
+Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
 This demonstrates algebraic convergence observed.}
@@
 We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodo{Here too} and $k = 100$.
+In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
 In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
@@
 Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.
-
-% For the final Gaussian bump example, we see that we achieve spectral convergence.
-
-We can extend this to constant non-zero boundary conditions by simply noting that the problem 
-\begin{align*}
-	\begin{cases}
-    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = c \in \R &\quad \text{on } \partial \Omega
-	\end{cases}
-\end{align*}
-is equivalent to letting $u = \tilde{u} + c$ and solving
-\begin{align*}
-	\begin{cases}
-    		\DeltaS \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - c \: k^2 \: v(x,y,z) \; =: g(x,y,z)  \quad \text{in } \Omega \\
-		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
-	\end{cases}
-\end{align*}
-and solving this zero boundary condition Helmholtz problem.
 
@@
 }
+\bstodoinline{Is the following is what you were thinking?}
+We can extend this to other boundary conditions that depend only on $x$ and $y$, i.e. $c = c(x,y)$. In this case, the problem
+\begin{align*}
+	\begin{cases}
+    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
+		u(x,y,z) = c(x,y) &\quad \text{on } \partial \Omega
+	\end{cases}
+\end{align*}
+is equivalent to letting $u(x,y,z) = \tilde{u}(x,y,z)+ c(x,y)$ and solving
+\begin{align*}
+	\begin{cases}
+    		\DeltaS \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)  \quad \text{in } \Omega \\
+		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
+	\end{cases}
+\end{align*}
+for $\tilde u$. This new problem is then a zero boundary condition Helmholtz problem with right hand side 
+\begin{align*}
+	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y) \\
+	&= f(x,y,z) + k^2 \: [{1 \over \rho(z)^2} - v(x,y,z)] \: c(x,y)
+\end{align*}
+for $(x,y,z) \in \Omega$. Notice that the spherical Laplacian applied to $c(x,y)$ is just 
+\begin{align*}
+	\DeltaS \: c(x,y) = {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) = -{k^2 \over \rho(z)^2} c(x,y)
+\end{align*}
+since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only.
 
@@
 
+\sotodoinline{I think we may need to explain what ${\partial u \over\partial n}$ means here: is the normal derivative well defined for 2D surfaces embedded in 3D?}
+\bstodoinline{Hopefully expanded on what the normal direction is here, and referenced back to where I defined the Biharmonic operator that also mentions it.
+Unless I am mistaken, the unit normal is simply the (normalised) positional vector $\vec x$ here.
+So I dont see a reason why normal derivative wont be well defined (though I very well could be wrong here).}
+\bstodoinline{The normal direction (including at the boundary) on a surface is the cross product $\ppx{\varphi}\vec r \times \ppx{\theta}\vec r$ where $\vec r = (x(\varphi,\theta),y(\varphi,\theta),z(\varphi,\theta))^\top$.
+For the spherical cap its the same as the sphere, i.e.
+is simply $\rho(z)^2 \: \vec r$ and so the unit normal would be $\vec r (= \vec x)$.}
 Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
@@
 	\begin{cases}
-    		\rho(z)^2 \: \Delta^2 u(x,y,z) = \rho(z)^2 \: f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
+    		\DeltaS^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
+		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = \grad_S \: u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
 	\end{cases}
@@
 \end{align}
-\sotodoinline{I think we may need to explain what ${\partial u \over\partial n}$ means here: is the normal derivative well defined for 2D surfaces embedded in 3D?}
-where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
+where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
+For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see Section \ref{subsection:furtherdiffoperators}).
 In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
@@
 We have shown that trivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary.
-We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: \sqrt{(1-z^2)}^b$ \sotodo{Should this be written as $(1-z^2)^{b/2}$?} defined for the disk-slice case \cite{snowball2019sparse}.
+We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: (1-z^2)^{b/2}$ defined for the disk-slice case \cite{snowball2019sparse}.
 Generalisation to spherical bands ($\alpha \leq z \leq \beta$) is straightforward.

### 2020-11-04 papers/spherical-caps/spherical-caps.tex
added some more comments

@@
 \begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| := \sqrt{x^2 + y^2} = \rho(z)\}
-\end{align*}
-and consider the case where $\alpha \in (-1,1)$, $\beta := 1$ and where $ \rho(z) := \sqrt{1-z^2}$.
+	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
+\end{align*}
+where $\alpha \in (-1,1)$ and $\beta := 1$.
 This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
-\begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
-\end{align*}
-For simplicity we focus on the case of a spherical cap, though there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$.
-The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper (these adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same).
-For clarity we note that, for all examples and plots in this paper, we take $\alpha = 0.2$.
+
+\begin{remark}
+For simplicity we focus on the case of a spherical cap, though there is an extension to a spherical band by taking $\beta \in (\alpha,1)$.
+The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper.
+These adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same.
+\end{remark}
 
@@
 \end{align*}
-noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \; \Omega$.
+noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \; \Omega$ when $a > 0$.
 The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
@@
 \end{align*}
-we have
+where $ \rho(z) := \sqrt{1-z^2}$, we have
 \begin{align*}
@@
 A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axes (e.g., a Schr\"odinger operator with potential $V(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
-This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity}.
-This plot presents the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos z$, as we increase the orthogonal polynomial degree $p$.
+This improves the complexity further to an optimal $O(p^2$), which is demonstrated\sotodo{This isn't actually shown in the referenced figure, which shows a non-radially symmetric potential Also, we need to make $V$ vs $v$ consistent with the figures and example} in Figure \ref{fig:complexity} with $V(x,y,z) = \cos z$.
 
@@
 For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$.
-Then the 3D polynomials defined on $\Omega$
+Then the 3D polynomials defined on $\Omega$\sotodo{Make notation consistent with Def 3, i.e., whether you use $\xvec$ or not}
 \begin{align*}
@@
 \end{align*}
-on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the Lebesgue measure on $\Omega$. 
-\sotodoinline{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} 
-\bstodoinline{I have changed it to dA and elaborated a bit more on it, but I am not sure on this so you could be correct that another term needs to be used.
-Your quadratic surfaces paper with Yuan says Lebesgue I think when discussing this.
-I have tried to keep it general here by using a measure definition, but we could omit this if its just going to cause confusion?}
+on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the uniform spherical measure on $\Omega$. 
 \end{proposition}
@@
 The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$.
-Let $N \in \N$ and define notation that groups degree $N$ polynomials together[:
+Let $N \in \N$ and define notation that groups degree $N$ polynomials together:
 \begin{align}
@@
 This also relates to the next comments}
+\sotodoinline{If we get rid of them as suggested below this is no longer an issue}
 \begin{align}
@@
 This section is important for deriving how to build the OPs, whereas before I simply need to say that the Jacobi operators that apply to coefficients vectors are reorderings of the ones in this section.}
+\sotodoinline{I agree: get rid of the previous definition and remove the tilde from these polynomials, as they are the "default" ones}
 \begin{align}
@@
 	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$.
-This shows that the time to build and solve the system $\big[\DeltaS + V(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
+This shows that the time to build and solve the system $\big[\DeltaS + V(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
 Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.
@@
 Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
-This indicates spectral convergence.
-\bstodoinline{Plots increased to N=200 for each}}
+This indicates spectral convergence.}
 	\centering
@@
 
-\sotodoinline{
-Wouldn't this trick work for any boundary conditions $c(x,y)$?
-That is, expand in Fourier series (a polynomial in $x$ and $y$) and use that to extend off the $z=0$ axis.
-If you agree we should explain that here.
-}
-\bstodoinline{Is the following is what you were thinking?}
+In Figure~\ref{fig:complexity} we plot presents the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos z$, as we increase the orthogonal polynomial degree $p$.
+\sotodo{The example is not the radially symmetric case}
+
 We can extend this to other boundary conditions that depend only on $x$ and $y$, i.e. $c = c(x,y)$. In this case, the problem
@@
 
-\sotodoinline{I think we may need to explain what ${\partial u \over\partial n}$ means here: is the normal derivative well defined for 2D surfaces embedded in 3D?}
-\bstodoinline{Hopefully expanded on what the normal direction is here, and referenced back to where I defined the Biharmonic operator that also mentions it.
-Unless I am mistaken, the unit normal is simply the (normalised) positional vector $\vec x$ here.
-So I dont see a reason why normal derivative wont be well defined (though I very well could be wrong here).}
-\bstodoinline{The normal direction (including at the boundary) on a surface is the cross product $\ppx{\varphi}\vec r \times \ppx{\theta}\vec r$ where $\vec r = (x(\varphi,\theta),y(\varphi,\theta),z(\varphi,\theta))^\top$.
-For the spherical cap its the same as the sphere, i.e.
-is simply $\rho(z)^2 \: \vec r$ and so the unit normal would be $\vec r (= \vec x)$.}
 Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:

### 2020-11-06 papers/spherical-caps/spherical-caps.tex
Addressed comments.

@@
 
+\bstodoinline{You mentioned before that you didnt like the use of $i$ being an index thoughout (as in $n,k,i$}. Is this still the case?}
+
 This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
@@
 
-Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + V(x,y,z)$ where $V$ is first approximated by a polynomial.
-A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axes (e.g., a Schr\"odinger operator with potential $V(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
-This improves the complexity further to an optimal $O(p^2$), which is demonstrated\sotodo{This isn't actually shown in the referenced figure, which shows a non-radially symmetric potential Also, we need to make $V$ vs $v$ consistent with the figures and example} in Figure \ref{fig:complexity} with $V(x,y,z) = \cos z$.
+Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + v(x,y,z)$ where $v$ is first approximated by a polynomial.
+A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
+This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity} with $v(x,y,z) = \cos z$.
+\sotodoinline{This isn't actually shown in the referenced figure, which shows a non-radially symmetric potential} 
+\bstodoinline{I dont really understand - isnt $V(x,y,z) =  V(z) = \cos z$ valid?
+I could be wrong but are you confusing this with the helmholtz example, with a different function $v$?
+If so, maybe I should make this clearer.}
 
@@
 For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$.
-Then the 3D polynomials defined on $\Omega$\sotodo{Make notation consistent with Def 3, i.e., whether you use $\xvec$ or not}
-\begin{align*}
-	\scopnki(x,y,z) \equiv \scopnki(\xvec,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr(\frac{\xvec}{\rho(z)})
+Then the 3D polynomials defined on $\Omega$
+\begin{align*}
+	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)})
 \end{align*}
@@
 \begin{align*}
-	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D \sigma(\xvec, z) \\
+	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
@@
 	&\ip<x \: \scopnkia, \scopmjha>_{\scopa} \\
-	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: x \: \genjacw^{(a,0)}(z) \: \D\sigma(\xvec, z) \\
+	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: x \: \genjacw^{(a,0)}(z) \: \D A \\
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
@@
 	&\ip<y \: \scopnkia, \scopmjha>_{\scopa} \\
-	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: y \: \genjacw^{(a,0)}(z) \: \D\sigma(\xvec, z) \\
+	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: y \: \genjacw^{(a,0)}(z) \: \D A \\
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
@@
 The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$.
-Let $N \in \N$ and define notation that groups degree $N$ polynomials together:
-\begin{align}
-	\bigscopNka &:= 
-		\begin{pmatrix}
-			\scopa_{k,k,0}(x,y,z) \\
-			\scopa_{k,k,1}(x,y,z) \\
-			\vdots \\
-			\scopa_{N,k,0}(x,y,z) \\
-			\scopa_{N,k,1}(x,y,z)
-		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
-	\bigscopa_{N,0} &:= 
-		\begin{pmatrix}
-			\scopa_{0,0,0}(x,y,z) \\
-			\vdots \\
-			\scopa_{N,0,0}(x,y,z) \\
-		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
-	\bigscopNa &:= 
-		\begin{pmatrix}
-			\bigscopa_{N,0} \\
-			\vdots \\
-			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
-\end{align}
-and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
-\sotodoinline{The following is wrong as it only applies if $N = \infty$} 
-\bstodoinline{What is better?
-To define them in this way for $N \to \infty$?
-Or to include the extra terms in this defn below?
-Or just state that the Jacobi operators approximate multiplication by x,y,z?
-This also relates to the next comments}
-\sotodoinline{If we get rid of them as suggested below this is no longer an issue}
-\begin{align}
-	J_x^{(a)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
-	J_y^{(a)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
-	J_z^{(a)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
-\end{align}
-The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis.
-For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$.
-Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis.
-Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:
-
-\begin{definition}
-A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
-A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
-\end{definition}
-
-For example, $J_x^{(a)}, J_y^{(a)}$ are block-tridiagonal (block-bandwidths $(1,1)$) while $J_z^{(a)}$ is block-diagonal (block-bandwidths $(0,0)$):
-\begin{align*}
-	J_{x/y/z}^{(a)} &= 
-		\begin{pmatrix}
-			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
-			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
-			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
-			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
-			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}
-\end{align*}
-
-For $J_x^{(a)}$, the sub-blocks $A^{(a)}_{x, k}, C^{(a)}_{x, k}$ have sub-block-bandwidths $(4,4)$ while the sub-blocks $B^{(a)}_{x,k}$ are zero matrices:
-\begin{align*}
-	B^{(a)}_{x,0} &= 0 \in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{x,k} = 0 \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N \\
-	A^{(a)}_{x,k} &:= 
-		\begin{pmatrix}
-			A^{(a)}_{k,k,6} & & & \\
-			A^{(a)}_{k+1,k,4} & \ddots & & \\
-			A^{(a)}_{k+2,k,2} & \ddots & \ddots & \\
-			& \ddots & \ddots & A^{(a)}_{N-1,k,6} \\
-			& & A^{(a)}_{N,k,2} & A^{(a)}_{N,k,4}
-		\end{pmatrix}, \quad k = 0,1,2,\dots,N 
-\end{align*}
-so that  $A^{(a)}_{x,0} \in \R^{(N+1)\times2N}$,  $A^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k)}$ for $k = 1,2,\dots,N$,
-\begin{align*}
-	C^{(a)}_{x,k} &:= 
-		\begin{pmatrix}
-			A^{(a)}_{k,k,1} & A^{(a)}_{k,k,3} & A^{(a)}_{k,k,5} & & \\
-			& \ddots & \ddots & \ddots \\
-			& & \ddots & \ddots & A^{(a)}_{N-1,k,5} \\
-			& & & A^{(a)}_{N,k,1} & A^{(a)}_{N,k,3}
-		\end{pmatrix}, \quad k = 1,2,\dots,N 
-\end{align*}
-so that $C^{(a)}_{x,1} \in \R^{2N\times(N+1)}$,  $C^{(a)}_{x,k} \in \R^{2(N-k+1)\times2(N-k+2)}$  for  $k = 2,3,\dots,N$, here for $k = 1,\dots,N, \: n = k,\dots,N$,
-\begin{align}
-	A^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			\alphaa_{n,k,j} & 0 \\
-			0 & \alphaa_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \label{eqn:jacobisubblocksA1} \\
-	A^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			\alphaa_{n,0,j} & 0
-		\end{pmatrix} \in \R^{1\times2}, j \text{ odd} \label{eqn:jacobisubblocksA2} \\
-	A^{(a)}_{n,1,j} &:=
-		\begin{pmatrix}
-			\alphaa_{n,1,j} \\
-			0
-		\end{pmatrix} \in \R^{2\times1}, j \text{ odd} \label{eqn:jacobisubblocksA3} \\
-\end{align}
-
-For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(5,5)$ while the sub-blocks $B^{(a)}_{y,k}$ are zero matrices:
-\begin{align*}
-	B^{(a)}_{y,0} &= 0 \in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{y,k} = 0 \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N \\
-	A^{(a)}_{y,k} &:= 
-		\begin{pmatrix}
-			B^{(a)}_{k,k,6} & & & \\
-			B^{(a)}_{k+1,k,4} & \ddots & & \\
-			B^{(a)}_{k+2,k,2} & \ddots & \ddots & \\
-			& \ddots & \ddots & B^{(a)}_{N-1,k,6} \\
-			& & B^{(a)}_{N,k,2} & B^{(a)}_{N,k,4}
-		\end{pmatrix}, \quad k = 0,1,2,\dots,N \\
-	&\big( \text{so that } A^{(a)}_{y,0} \in \R^{(N+1)\times2N}, \quad A^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \big) \\
-	C^{(a)}_{y,k} &:= 
-		\begin{pmatrix}
-			B^{(a)}_{k,k,1} & B^{(a)}_{k,k,3} & B^{(a)}_{k,k,5} & & \\
-			& \ddots & \ddots & \ddots \\
-			& & \ddots & \ddots & B^{(a)}_{N-1,k,5} \\
-			& & & B^{(a)}_{N,k,1} & B^{(a)}_{N,k,3}
-		\end{pmatrix}, \quad k = 1,2,\dots,N \\
-	&\big( \text{so that } C^{(a)}_{y,1} \in \R^{2N\times(N+1)}, \quad C^{(a)}_{y,k} \in \R^{2(N-k+1)\times2(N-k+2)} \text{ for } k = 2,3,\dots,N \big)
-\end{align*}
-where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align}
-	B^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			0 & \betaa_{n,k,0,j} \\
-			\betaa_{n,k,1,j} & 0
-		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \label{eqn:jacobisubblocksB1} \\
-	B^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			0 & \betaa_{n,0,0,j}
-		\end{pmatrix} \in \R^{1\times2}, j \text{ odd}  \label{eqn:jacobisubblocksB2} \\
-	B^{(a)}_{n,1,j} &:=
-		\begin{pmatrix}
-			0 \\
-			\betaa_{n,1,1,j}
-		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}  \label{eqn:jacobisubblocksB3} \\
-\end{align}
-
-For $J_z^{(a)}$, the sub-blocks $B^{(a)}_{z,k}$ have sub-block-bandwidths $(2,2)$, while the sub-blocks $A^{(a)}_{z,k}, \: C^{(a)}_{z,k}$ are zero matrices:
-\begin{align*}
-	A^{(a)}_{z,0} &= 0 \in \R^{(N+1)\times2N}, \quad A^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k)} \text{ for } k = 1,2,\dots,N \\
-	C^{(a)}_{z,1} &= 0 \in \R^{2N\times(N+1)}, \quad C^{(a)}_{z,k} = 0 \in \R^{2(N-k+1)\times2(N-k+2)} \text{ for } k = 2,3,\dots,N \\
-	B^{(a)}_{z,k} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{k,k,2} & \Gamma^{(a)}_{k,k,3} & \\
-			\Gamma^{(a)}_{k+1,k,1} & \ddots & \ddots & \\
-			& \ddots & \ddots & \Gamma^{(a)}_{N-1,k,3} \\
-			& & \Gamma^{(a)}_{N,k,1} & \Gamma^{(a)}_{N,k,2}
-		\end{pmatrix}  \quad k = 0,1,2,\dots,N \\
-	&\big( \text{so that } B^{(a)}_{z,0} \in \R^{(N+1)\times(N+1)}, \quad B^{(a)}_{z,k} \in \R^{2(N-k+1)\times2(N-k+1)} \text{ for } k = 1,2,\dots,N \big)
-\end{align*}
-where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align}
-	\Gamma^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			\gamma_{n,k,j} & 0 \\
-			0 & \gamma_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, \label{eqn:jacobisubblocksGamma1} \\
-	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:jacobisubblocksGamma2}
-\end{align}
-
-Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
-
-\subsection{Building the OPs}
-
-In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \textit{recurrence coefficient matrices} -- we define these as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the coefficients for the multidimensional three-term recurrences.
-
-While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the recurrence coefficient matrices necessary for efficient function evaluation using the Clenshaw algorithm.
+In later sections we will use an ordering of the OPs so that they are grouped by Fourier mode $k$, which is convenient for the application of differential and other operators to the vector of coefficients of a given function's expansion (some operators will exploit this ordering for operators where Fourier modes do not interact, and thus will be block-diagonal).
+Before that though, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
@@
 and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to
-\sotodoinline{Can the following be simplified? Seems like it repeats} 
-\bstodoinline{Is it worth removing the previous?
-And just explaining that we can reorder the Jacobi matrices according to the OP/coefficient vector ordering?
-This section is important for deriving how to build the OPs, whereas before I simply need to say that the Jacobi operators that apply to coefficients vectors are reorderings of the ones in this section.}
-\sotodoinline{I agree: get rid of the previous definition and remove the tilde from these polynomials, as they are the "default" ones}
 \begin{align}
@@
 \end{align*}
-These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).
-
-For $\tilde J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
+Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are \textit{banded-block-banded matrices}:
+
+\begin{definition}
+	A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
+A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
+\end{definition}
+
+Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $\tilde J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
 \begin{align*}
@@
 where $\Gamma^{(a)}_{n,k,j}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}).
+
+Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
+
+\subsection{Building the OPs}
 
@@
 
-In this section, we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process. To this end, we will use a different ordering
- of the OPs and coefficients of a function's expansion, in order to exploit the orthogonality the circular harmonics will bring and thus the operators will be block-diagonal.
-Let $N \in \N$ and recall from equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) that:
-\begin{align*}
+In this section we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process.
+To this end, as alluded to in Section~\ref{subsection:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal.
+Let $N \in \N$ and define:
+\begin{align}
 	\bigscopNka &:= 
@@
 			\scopa_{N,k,1}(x,y,z)
-		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \\ 
+		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
 	\bigscopa_{N,0} &:= 
@@
 			\vdots \\
-			\scopa_{N,0,0}(x,y,z)
-		\end{pmatrix} \in \R^{N+1}, \\
+			\scopa_{N,0,0}(x,y,z) \\
+		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
 	\bigscopNa &:= 
@@
 			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2}.
-\end{align*}
+		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
+\end{align}
 
@@
 \end{align*}
-and recall that a function $f(x,y,z)$ defined on $\Omega$ is approximated by its expansion $f(x,y,z) = \bigscopa(x,y,z)^\top \mathbf{f}$.
+
+The operator matrices we derive here act on coefficient vectors, that represent a function $f(x,y,z)$ defined on $\Omega$ in spectral space -- such a function is approximated by its expansion up to degree $N$:
+\begin{align*}
+	f(x,y,z) = \bigscopNa(x,y,z)^\top \vec f = \sum_{n=0}^{N} \sum_{k=0}^{n} \sum_{i=0}^{1} f_{n,k,i} \: \scopnkia(x,y,z),
+\end{align*}
+where $\vec f = (f_{n,k,i})$ is the coefficients vector for the function $f$.
 
@@
 \begin{align*}
-	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
+	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \approx \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
 \end{align*}
@@
 \begin{align*}
-	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
+	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
@@
 \begin{align*}
-	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
+	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
@@
 \begin{align*}
-	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z  \\
-	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
+	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D A  \\
+	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
@@
 % FIGURES
-
-\begin{figure}[t]
-	\centering % <-- added
-	\includegraphics[scale=0.35]{complexity-Nend=1000}
-	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$.
-This shows that the time to build and solve the system $\big[\DeltaS + V(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
-Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.
-\bstodoinline{Redo with belapsed}}
-	\centering
-	\label{fig:complexity}
-\end{figure}
 
@@
 \end{figure}
-%
-%\begin{figure}[t]
-%	\begin{subfigure}{0.3\textwidth}
-%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
-%	\centering
-%	\end{subfigure}
-%	\begin{subfigure}{0.3\textwidth}
-%	\centering
-%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
-%	\centering
-%	\end{subfigure}
-%	\begin{subfigure}{0.3\textwidth}
-%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
-%	\centering
-%	\end{subfigure}
-%	\centering
-%	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$.
-Left: Computed.
-Centre: Exact.
-Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
-%	\centering
-%	\label{fig:poissonexact}
-%\end{figure}
 
@@
 	\label{fig:helmholtz}
+\end{figure}
+
+\begin{figure}[t]
+	\centering % <-- added
+	\includegraphics[scale=0.35]{complexity-Nend=1000}
+	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + v(x,y,z)$, with rotationally invariant potential $v(x,y,z) = \cos z$.
+This shows that the time to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
+Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.
+\bstodoinline{Redo with belapsed}}
+	\centering
+	\label{fig:complexity}
 \end{figure}
@@
 
-In Figure~\ref{fig:complexity} we plot presents the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos z$, as we increase the orthogonal polynomial degree $p$.
-\sotodo{The example is not the radially symmetric case}
-
-We can extend this to other boundary conditions that depend only on $x$ and $y$, i.e. $c = c(x,y)$. In this case, the problem
+In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} \bstodo{cite benchmarktools github?} to construct the operator for $\DeltaS + v(x,y,z)$, this time with a rotationally invariant $v(x,y,z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
+The plot demonstrates that as we increase the degree of approximation $N$, we achieve a complexity of an optimal $\bigO(N^2)$.
+\sotodoinline{The example is not the radially symmetric case} \bstodoinline{see other comment on this}
+
+What about other boundary conditions?
+One simple extension is the case where the value on the boundary takes that of a function depending only on $x$ and $y$, i.e.
+$c = c(x,y)$.
+In this case, the problem
 \begin{align*}
@@
 \begin{align*}
-	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y) \\
-	&= f(x,y,z) + k^2 \: [{1 \over \rho(z)^2} - v(x,y,z)] \: c(x,y)
-\end{align*}
-for $(x,y,z) \in \Omega$. Notice that the spherical Laplacian applied to $c(x,y)$ is just 
-\begin{align*}
-	\DeltaS \: c(x,y) = {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) = -{k^2 \over \rho(z)^2} c(x,y)
+	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)
+\end{align*}
+for $(x,y,z) \in \Omega$. Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
+\begin{align*}
+	\DeltaS \: c(x,y) = {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) = - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
 \end{align*}
 since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only.
+Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manor.
 

### 2020-11-09 papers/spherical-caps/spherical-caps.tex
Added belapsed image and comment
Typo fixes

@@
 
-\bstodoinline{You mentioned before that you didnt like the use of $i$ being an index thoughout (as in $n,k,i$}. Is this still the case?}
+\bstodoinline{You mentioned before that you didnt like the use of $i$ being an index throughout (as in $n,k,i$). Is this still the case?}
 
@@
 In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
-Here, we use $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+Here, we use $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
 Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.
@@
 In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 100$, that is, $(N+1)^2 = 10,201$ unknowns.
+In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
 Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve algebraic convergence for these more complex functions.

### 2020-11-09 papers/spherical-caps/spherical-caps.tex
removed comment
add ref for BenchmarkTools.jl

@@
 
-\bstodoinline{You mentioned before that you didnt like the use of $i$ being an index throughout (as in $n,k,i$). Is this still the case?}
-
 This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
@@
 This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity} with $v(x,y,z) = \cos z$.
-\sotodoinline{This isn't actually shown in the referenced figure, which shows a non-radially symmetric potential} 
-\bstodoinline{I dont really understand - isnt $V(x,y,z) =  V(z) = \cos z$ valid?
-I could be wrong but are you confusing this with the helmholtz example, with a different function $v$?
-If so, maybe I should make this clearer.}
 
@@
 This shows that the time to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
-Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.
-\bstodoinline{Redo with belapsed}}
+Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.}
 	\centering
@@
 	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
-This demonstrates algebraic convergence observed.}
+Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
+This demonstrates algebraic convergence.}
 	\centering
@@
 
-In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} \bstodo{cite benchmarktools github?} to construct the operator for $\DeltaS + v(x,y,z)$, this time with a rotationally invariant $v(x,y,z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
+In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools.jl package \cite{BenchmarkTools.jl-2016} in Julia.}  to construct the operator for $\DeltaS + v(x,y,z)$, this time with a rotationally invariant $v(x,y,z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
 The plot demonstrates that as we increase the degree of approximation $N$, we achieve a complexity of an optimal $\bigO(N^2)$.
-\sotodoinline{The example is not the radially symmetric case} \bstodoinline{see other comment on this}
 

### 2020-11-18 papers/spherical-caps/spherical-caps.tex
Minor edits (I think)

@@
 \end{align*}
-where $A^{(a)}_{n,k,j}$ are defined in equations (\ref{eqn:jacobisubblocksA1}, \ref{eqn:jacobisubblocksA2}, \ref{eqn:jacobisubblocksA3}).
+where for $k = 1,\dots,N, \: n = k,\dots,N$
+\begin{align*}
+	A^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			\alphaa_{n,k,j} & 0 \\
+			0 & \alphaa_{n,k,j}
+		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
+	A^{(a)}_{n,0,j} &:=
+		\begin{pmatrix}
+			\alphaa_{n,0,j} & 0
+		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
+	A^{(a)}_{n,1,j} &:=
+		\begin{pmatrix}
+			\alphaa_{n,1,j} \\
+			0
+		\end{pmatrix} \in \R^{2\times1}, j \text{ odd}.
+\end{align*}
 For $\tilde J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
@@
 \end{align*}
-where $B^{(a)}_{n,k,j}$ are defined equations (\ref{eqn:jacobisubblocksB1}, \ref{eqn:jacobisubblocksB2}, \ref{eqn:jacobisubblocksB3}).
-For $\tilde J_z^{(a)}$, the sub-blocks are diagonal, i.e.
-have sub-block-bandwidths $(0,0)$:
+where for $k = 1,\dots,N, \: n = k,\dots,N$  
+\begin{align*}
+	B^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			0 & \betaa_{n,k,0,j} \\
+			\betaa_{n,k,1,j} & 0
+		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
+	B^{(a)}_{n,0,j} &:=
+		\begin{pmatrix}
+			0 & \betaa_{n,0,0,j}
+		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
+	B^{(a)}_{n,1,j} &:=
+		\begin{pmatrix}
+			0 \\
+			\betaa_{n,1,1,j}
+		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}.
+\end{align*}
+For $\tilde J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
 \begin{align*}
@@
 \end{align*}
-where $\Gamma^{(a)}_{n,k,j}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}).
-
+where for $k = 1,\dots,N, \: n = k,\dots,N$
+\begin{align}
+	\Gamma^{(a)}_{n,k,j} &:= 
+		\begin{pmatrix}
+			\gamma_{n,k,j} & 0 \\
+			0 & \gamma_{n,k,j}
+		\end{pmatrix} \in \R^{2\times2}, \label{eqn:jacobisubblocksGamma1} \\
+	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:jacobisubblocksGamma2}
+\end{align}
 Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.

### 2020-12-21 papers/spherical-caps/spherical-caps.tex
Fixed image sizing
final(?)

@@
 The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
-We demonstrate the results on partial differential equations involving the spherical Laplacian and Biharmonic operators, showing spectral convergence.
+We demonstrate the results on partial differential equations involving the spherical Laplacian and biharmonic operators, showing spectral convergence.
 \end{abstract}
@@
 
-This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
-More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
+This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere---specifically spherical caps.
+More precisely, we consider the solution of partial differential equations on the spherical cap $\Omega$ defined by
 \begin{align*}
@@
 where $\alpha \in (-1,1)$ and $\beta := 1$.
-This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
 
@@
 The benefit of the new approach is that we do not need to resolve Jacobians, and thereby we can achieve sparse discretisations for partial differential operators, including those with polynomial variable coefficients.
-Further, we remove the singular nature as $\alpha$ approaches $0$ that such a projection would give, since our new approach is applicable for all $\alpha \in (-1,1)$.
+Further, we avoid the singular nature at the poles or  as $\alpha$ approaches $0$ that  such a projection may give, since our new approach yields a smooth polynomial basis for all $\alpha \in [-1,1)$.
+
 On the spherical cap, the family of weights we consider are of the form
@@
 \end{align*}
-noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \; \Omega$ when $a > 0$.
+noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \Omega$ when $a > 0$.
 The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
@@
 
- The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals has already been outlined \cite{snowball2019sparse}.
+ The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined \cite{snowball2019sparse}.
 In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
-	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
+	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: \times \\
+	&\qquad w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
 	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
@@
 \end{align*}
-and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to
+and set $ J_x^{(a)},  J_y^{(a)},  J_z^{(a)}$ as the Jacobi matrices corresponding to
 \begin{align}
-	\tilde J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
-	\tilde J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
-	\tilde J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
+	J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
+	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
+	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
 \end{align}
@@
 \begin{align*}
-	\tilde J_{x/y/z}^{(a)} &= 
-		\begin{pmatrix}
-			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
-			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
-			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
-			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
+	J_{x/y/z}^{(a)} &= 
+		\begin{pmatrix}
+			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
+			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
+			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
+			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
 			& & & \ddots & \ddots & \ddots \\
@@
 
-Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $\tilde J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
-\begin{align*}
-	\tilde A^{(a)}_{x,n} &:= 
+Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
+\begin{align*}
+	A^{(a)}_{x,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	\tilde B^{(a)}_{x,n} &:= 
+	B^{(a)}_{x,n} &:= 
 		\begin{pmatrix}
@@
 			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & \tilde A^{(a)}_{n,n-1,4} \\
+			& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
 			& & A^{(a)}_{n,n,3} & 0
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	\tilde C^{(a)}_{x,n} &:= 
+	C^{(a)}_{x,n} &:= 
 		\begin{pmatrix}
@@
 \end{align*}
-For $\tilde J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
-\begin{align*}
-	\tilde A^{(a)}_{y,n} &:= 
+For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
+\begin{align*}
+	A^{(a)}_{y,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	\tilde B^{(a)}_{y,n} &:= 
+	B^{(a)}_{y,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	\tilde C^{(a)}_{y,n} &:= 
+	C^{(a)}_{y,n} &:= 
 		\begin{pmatrix}
@@
 \end{align*}
-For $\tilde J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
-\begin{align*}
-	\tilde A^{(a)}_{z,n} &:= 
+For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
+\begin{align*}
+	A^{(a)}_{z,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	\tilde B^{(a)}_{z,n} &:= 
+	B^{(a)}_{z,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	\tilde C^{(a)}_{z,n} &:= 
+	C^{(a)}_{z,n} &:= 
 		\begin{pmatrix}
@@
 
-Following the triangle case \cite{olver2019triangle}, we can combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
+Following the triangle case \cite{olver2019triangle}, we use a multivariate analogue of Clenshaw's algorithm for evaluation, where we  combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
 \begin{align*}
@@
 \begin{itemize}
-	\item $D_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 4)$
-  	\item $W_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 2)$
-	\item $D_\theta$ has  block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
-	\item $\calL^{(a)\to(a+\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(0, 4)$
-	\item $\calL_W^{(a)\to(a-\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 0)$
-	\item $\Delta^{(1)}_W$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 2)$
+	\item $D_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(2, 4)$
+  	\item $W_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(4, 2)$
+	\item $D_\theta$ is block-diagonal with sub-block-bandwidths $(1, 1)$
+	\item $\calL^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block-bandwidths $(0, 4)$
+	\item $\calL_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block-bandwidths $(4, 0)$
+	\item $\Delta^{(1)}_W$ is block-diagonal with sub-block-bandwidths $(2, 2)$
 \end{itemize}
@@
 	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a - \tilde a, 2j)} \: \genjacw^{(a - 2, k+j)} \: \Big\{\genjacnmk^{(a,2k)} [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quad \quad  \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a - \tilde a, 2j)} \: \genjacw^{(a - 2, k+j)} \: \Big\{\genjacnmk^{(a,2k)} [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
 	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} [-2(k+1) z \: \genjacw^{(2,0)} + 2a \genjacw^{(1,2)}] \\
@@
 	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
+	&\quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
+	&\quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
 	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
@@
 	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.23]{sparsity-of-biharmonic}
+		\includegraphics[scale=0.3]{sparsity-of-biharmonic}
 		\centering
@@
 \end{align*}
-Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,u,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$ and thus its spherical gradient would be zero on the boundary $z = \alpha$.
+Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,y,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$ and thus its spherical gradient would be zero on the boundary $z = \alpha$.
 This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
@@
 \begin{align*}
-	\tilde{\bm{f}}_n &:= 
+	{\bm{f}}_n &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2(N+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	\tilde{\bm{f}}_0 = f_{0,0,0} \in \R, \\
-	\tilde{\bm{f}} &:= 
-		\begin{pmatrix}
-			\tilde{\bm{f}}_0 \\
+	{\bm{f}}_0 = f_{0,0,0} \in \R, \\
+	{\bm{f}} &:= 
+		\begin{pmatrix}
+			{\bm{f}}_0 \\
 			\vdots \\
-			\tilde{\bm{f}}_N
+			{\bm{f}}_N
 		\end{pmatrix} \in \R^{(N+1)^2}.
@@
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{f}}_n}^T - \bm{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {{\bm{f}}_n}^T - \bm{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
 	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
@@
 \begin{align*}
-	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \bm{f}^\top \: \bigscopNa(x,y,z) = (V \bm{f})^\top \: \bigscopNa(x,y,z),
+	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \bm{f} \: \bigscopta(x,y,z) = (V \bm{f})^\top \: \bigscopta(x,y,z),
 \end{align*}
@@
 
-To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in Section \ref{subsection:functionevaluation} so that $v(x,y,z) = \tilde{\bm{v}}^\top \: \bigscopta(x,y,z)$).
+To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in Section \ref{subsection:functionevaluation} so that $v(x,y,z) = {\bm{v}}^\top \: \bigscopta(x,y,z)$).
 Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, $Z := (J_z^{(a)})^\top$.
@@
 	\includegraphics[scale=0.35]{complexity-Nend=1000}
-	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + v(x,y,z)$, with rotationally invariant potential $v(x,y,z) = \cos z$.
-This shows that the time to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
+	\caption{Time in seconds to  to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$.
+This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
 Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.}
@@
 \end{align*}
-since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only.
+since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only, which are precisely the Fourier coefficients of $c(\cos \theta, \sin \theta)$.
 Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manor.
@@
 
-Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
+Our last erxample is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
 \begin{align}
@@
 In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
-Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve algebraic convergence for these more complex functions.
+Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve exponential convergence for these more complex functions.
 
@@
 
-This work serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere -- it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
+This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere---it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
+Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct the necessary orthogonal polynomials.
 

### 2020-12-21 papers/spherical-caps/spherical-caps.tex
add grant support
Added EPSRC acknowledgement. Minor edits to paper.

@@
 We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
+
+\noindent{\bf Acknowledgments}: The first author was supported by the Engineering and Physical Sciences Research Council Mathematics of Planet Earth Centre for Doctoral Training at Imperial College London and the University of Reading, with grant number EP/L016613/1.The second author was supported by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces''.
 
@@
 	\includegraphics[scale=0.35]{complexity-Nend=1000}
-	\caption{Time in seconds to  to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$.
+	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.
 This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
-Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.}
+Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
 	\centering
@@
 
-In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools.jl package \cite{BenchmarkTools.jl-2016} in Julia.}  to construct the operator for $\DeltaS + v(x,y,z)$, this time with a rotationally invariant $v(x,y,z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
+In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools.jl package \cite{BenchmarkTools.jl-2016} in Julia.} to construct the operator for $\DeltaS + v(x,y,z)$, with a rotationally invariant $v(x,y,z) = v(z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
 The plot demonstrates that as we increase the degree of approximation $N$, we achieve a complexity of an optimal $\bigO(N^2)$.

### 2020-12-22 papers/spherical-caps/highlights.tex
Add statement on significance and highlights

--- 2020-12-21 papers/spherical-caps/spherical-caps.tex
@@
 
-\title{Sparse spectral methods for partial differential equations on spherical caps}
+\title{Sparse spectral methods for partial differential equations on spherical caps, highlights}
 \author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
@@
 
-\begin{abstract}
-In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, disk-slices and triangles.
-In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps.
-The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
-We demonstrate the results on partial differential equations involving the spherical Laplacian and biharmonic operators, showing spectral convergence.
-\end{abstract}
-
-%
-\section{Introduction}
-
-This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere---specifically spherical caps.
-More precisely, we consider the solution of partial differential equations on the spherical cap $\Omega$ defined by
-\begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
-\end{align*}
-where $\alpha \in (-1,1)$ and $\beta := 1$.
-
-\begin{remark}
-For simplicity we focus on the case of a spherical cap, though there is an extension to a spherical band by taking $\beta \in (\alpha,1)$.
-The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper.
-These adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same.
-\end{remark}
-
-We advocate using a basis that is polynomial in cartesian coordinates, that is, polynomial in $x$, $y$, and $z$, and orthogonal with respect to a prescribed weight: that is, multivariate orthogonal polynomials, whose construction was considered in \cite{olver2020orthogonal}.
-Equivalently, we can think of these as polynomials modulo the vanishing ideal $\{ x^2 + y^2 + z^2 = 1 \}$, or simply as a linear recombination of spherical harmonics that are orthogonalised on a subset of the sphere.
-This is in contrast to more standard approaches based on mapping the geometry to a simpler one (e.g., a rectangle or disk) and using orthogonal polynomials in the mapped coordinates (e.g., a basis that is polynomial in the spherical coordinates $\varphi$ and $\theta$).
-The benefit of the new approach is that we do not need to resolve Jacobians, and thereby we can achieve sparse discretisations for partial differential operators, including those with polynomial variable coefficients.
-Further, we avoid the singular nature at the poles or  as $\alpha$ approaches $0$ that  such a projection may give, since our new approach yields a smooth polynomial basis for all $\alpha \in [-1,1)$.
-
-On the spherical cap, the family of weights we consider are of the form
-\begin{align*}
-	\Wa(x,y,z) := (z - \alpha)^a, \qfor (x,y,z) \in \Omega,
-\end{align*}
-noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \Omega$ when $a > 0$.
-The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$.
-We define these to be orthogonalised lexicographically, that is,
-\begin{align*}
-	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms})
-\end{align*}
-where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - i} \: y^{i} \: z^{n-j}$ where $j < k$.
-The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
-
-We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
-\begin{align*}
-	z &= \cos\varphi, \\
-	x &= \sin\varphi \cos\theta = \rho(z) \cos\theta, \\
-	y &= \sin\varphi \sin\theta = \rho(z) \sin\theta.
-\end{align*}
-where $ \rho(z) := \sqrt{1-z^2}$, we have
-\begin{align*}
-	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
-\end{align*}
-i.e.
-$\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$.
-We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
-As in the  disk-slice case in 2D \cite{snowball2019sparse}, we use an integration-by-parts argument to deduce the sparsity structure.
-
- The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined \cite{snowball2019sparse}.
-In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
-This clearly compares favourably to proceeding in a na\"ive approach where one would require $O(p^6)$ operations.
-
-Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + v(x,y,z)$ where $v$ is first approximated by a polynomial.
-A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
-This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity} with $v(x,y,z) = \cos z$.
-
-An overview of the paper is as follows:
-
-\noindent \secref{OPs}: We present our definition of a (one-parameter) family of 3D orthogonal polynomials (OPs) on the spherical cap domain $\Omega$, by combining 1D OPs on the interval $(\alpha, 1)$ with Chebyshev polynomials, to form 3D OPs on the spherical cap surface.
-We show that these families will lead to sparse Jacobi operators for multiplication by $x, y, z$ and demonstrate how to obtain the 3D OPs.
-
-\noindent\secref{PDOs}: We define several partial differential operators such as spherical Laplacians and  show that these will  be sparse when applied to a suitable choice of expansions in bases built from OPs on the spherical cap.
-We can exactly calculate the non-zero entries of these sparse operators using the quadrature rule associated with the non-classical 1D OPs.
-
-\noindent\secref{Computation}: We derive a quadrature rule on the spherical cap that can be used to expand a function in the OP basis up to a given order $N$, and demonstrate how to evaluate a function using the Clenshaw algorithm using the coefficients of its expansion.
-
-\noindent\secref{Examples}: We demonstrate the proposed technique for solving differential equations on the spherical cap such as the Poisson equation, variable coefficient Helmholtz equation, and Biharmonic equation.
-
-%\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
-We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
-
-\section{Orthogonal polynomials on spherical caps}\label{Section:OPs}
-
-In this section we outline the construction and some basic properties of $\scopnkia(x,y,z)$.
-
-\subsection{Explicit construction}
-
-We can construct the 3D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Chebyshev polynomials.
-We do so in terms of Fourier series, which, following \cite{olver2020orthogonal}, we write here as orthogonal polynomials in $x$ and $y$:
-
-\begin{definition}\label{def:Ydefinition}
-	Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
-Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
-\begin{align*}
-	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
-	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots \\
-	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots
-\end{align*}
-where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
-\begin{align*}
-	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta
-\end{align*}
-\end{definition}
-Note that we have defined $\ch_0$ so as to ensure orthonormality.
-
-\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
-	Let $w : (\alpha,\beta) \: \to \R$ be a weight function.
-For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$.
-Then the 3D polynomials defined on $\Omega$
-\begin{align*}
-	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)})
-\end{align*}
-for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
-\begin{align*}
-	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
-	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: \times \\
-	&\qquad w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
-	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
-\end{align*}
-on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the uniform spherical measure on $\Omega$. 
-\end{proposition}
-
-For the spherical cap, we can use Proposition \ref{prop:construction} to create our one-parameter family of OPs.
-We first introduce notation for our family of non-classical univariate OPs that will be used as the $r_n$ polynomials above.
-\begin{definition}[\cite{snowball2019sparse}]\label{def:OPconstruction}
-	Let $\genjacw^{(a,b)}(x)$ be a weight function on the interval $(\alpha, 1)$ given by:
-\begin{align*}
-	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b}
-\end{align*}
-and define the associated inner product by:
-\begin{align}
-	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipgenjac}
-\end{align}
-where
-\begin{align}
-	\normgenjac^{(a,b)} := \int_\alpha^1 \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipnormalisation}
-\end{align}
-is a normalising constant.
-Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}).
-\end{definition}
-
-We can now define the 3D OPs for the spherical cap.
-\begin{definition}\label{def:constuction}
-	Define the one-parameter 3D orthogonal polynomials via:
-\begin{align}
-	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega.
-\end{align}
-\end{definition}
-
-By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
-\begin{align*}
-	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
-	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
-\end{align*}
-with
-\begin{align}
-	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:normscop}
-\end{align}
-We note that the weight $\genjacw^{(a,b)}(z)$ has been used in the construction of 2D orthogonal polynomials on disk-slices and trapeziums \cite{snowball2019sparse}, where a method for obtaining recurrence coefficients and evaluating integrals was established (the weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
-
-\subsection{Jacobi matrices}\label{subsection:jacobimats}
-
-We can express the three-term recurrences associated with $\genjac_n^{(a,b)}$ as
-\begin{align}
-	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Rrecurrence}
-\end{align}
-where the coefficients are calculatable (see \cite{snowball2019sparse}).
-We can use (\ref{eqn:Rrecurrence}) to determine the 3D recurrences for $\scopnkia(x,y,z)$.
-Importantly, we can deduce sparsity in the recurrence relationships.
-We first require the following lemma.
-
-\begin{lemma}\label{lemma:Yrecurrence} 
-The following identities hold for $k = 2,3,\dots$, $j = 0,1,\dots$ and $i, h \in \{0,1\}$:
-\begin{align*}
-	&1) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \cos\theta \: \D \theta = \ch_0 \: \pi \: \delta_{0,h} \: \delta_{1,j} \\
-	&2) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \sin\theta \: \D \theta = \ch_0 \: \pi \: \delta_{1,h} \: \delta_{1,j} \\
-	&3) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta = \pi \: \delta_{i,h} \: (\ch_0 \: \delta_{0,j} + \half \delta_{2,j}) \\
-	&4) \quad \int_0^{2\pi} \ch_{1,i}(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta = \pi \:  \delta_{|i-1|,h} \: ((-1)^{i+1} \: \ch_0 \: \delta_{0,j} + (-1)^i \: \half \: \delta_{2,j}) \\
-	&5) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta = \half \: \pi \: \delta_{i,h} \: (\delta_{k-1,j} + \delta_{k+1,j}) \\
-	&6) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta = \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \delta_{k-1,j} + (-1)^i \: \delta_{k+1,j}).
-\end{align*}
-\end{lemma}
-
-\begin{proof}
-Each follows from the definitions of $\ch_{k,i}$ and $\ch_0$, as well as the relationships:
-\begin{align*}
-	2 \cos k \theta \cos\theta &= \cos (k-1)\theta + \cos(k+1)\theta \\
-	2 \sin k \theta \cos\theta &= \sin (k-1)\theta + \sin(k+1)\theta \\
-	2 \cos k \theta \sin\theta &= - \sin (k-1)\theta + \sin(k+1)\theta \\
-	2 \sin k \theta \sin\theta &= \cos (k-1)\theta - \cos(k+1)\theta.
-\end{align*}
-\end{proof}
-
-\begin{lemma}\label{lemma:Qrecurrence} 
-Define
-\begin{align}
-	\eta_{k} :=
-		\begin{cases}
-			0 &\text{if } k < 0 \\
-			\ch_0 &\text{if } k = 0 \\
-			\half &\text{otherwise}
-		\end{cases}
-\end{align}
-$\scopnkia(x,y,z)$ satisfy the following recurrences:
-\begin{align*}
-	x \: \scopnkia(x,y,z) &= \alphaa_{n,k,1} \:  \scopa_{n-1, k-1, i}(x, y, z) + \alphaa_{n,k,2} \:  \scopa_{n-1, k+1, i}(x, y, z) \nonumber \\
-		& \quad \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
-		& \quad \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\
-	y \: \scopnkia(x,y,z) &= \betaa_{n,k,i,1} \:  \scopa_{n-1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,2} \:  \scopa_{n-1, k+1, |i-1|}(x, y, z) \nonumber \\
-		& \quad \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
-		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
-	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
-\end{align*}
-for $(x,y,z) \in \Omega$, where
-\begin{align*}
-	\alphaa_{n,k,1} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
-	\alphaa_{n,k,2} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-2}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
-	\alphaa_{n,k,3} &:=\eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+1}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
-	\alphaa_{n,k,4} &:= \eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjac_{n-k-1}^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
-	\alphaa_{n,k,5} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
-	\alphaa_{n,k,6} &:=\eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
-	\betaa_{n,k,i,j} &:= 
-		\begin{cases}
-			- \alphaa_{n,k,j} \quad &\text{if }(i = 0 \text{ and } j \text{ is odd}) \text{ or }(i = 1 \text{ and } j \text{ is even}) \\
-			\alphaa_{n,k,j} \quad &\text{otherwise}
-		\end{cases}, \\	
-	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}.
-\end{align*}
-\end{lemma}
-\begin{remark}
-For $z$ multiplication, note that different Fourier modes do not interact. This is because $z$ is rotationally invariant. 
-\end{remark}
-\begin{proof}
-The 3-term recurrence for multiplication by $z$ follows from (\ref{eqn:Rrecurrence}).
-For the recurrence for multiplication by $x$, since $\{\scopmjha\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n+1$ polynomial on $\Omega$, we can expand
-\begin{align*}
-	x \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j} \: \scopmjha(x,y,z).
-\end{align*}
-These coefficients are given by
-\begin{align*}
-	c_{m,j} = {\ip< x \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
-\end{align*}
-where we show the non-zero coefficients that result are the $\alphaa_{n,k,1},\dots,\alphaa_{n,k,6}$ in the lemma.
-Recall from equation (\ref{eqn:normscop}) that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$.
-Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)$:
-\begin{align*}
-	&\ip<x \: \scopnkia, \scopmjha>_{\scopa} \\
-	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: x \: \genjacw^{(a,0)}(z) \: \D A \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
-\end{align*}
-where $\delta_{k, j}$ is the standard Kronecker delta function, using Lemma \ref{lemma:Yrecurrence}. Similarly, for the recurrence for multiplication by $y$, we can expand 
-\begin{align*}
-	y \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 d_{m,j} \: \scopmjha(x,y,z).
-\end{align*}
-These coefficients are given by
-\begin{align*}
-	d_{m,j} = {\ip< y \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
-\end{align*}
-where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the lemma:
-\begin{align*}
-	&\ip<y \: \scopnkia, \scopmjha>_{\scopa} \\
-	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: y \: \genjacw^{(a,0)}(z) \: \D A \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
-\end{align*}
-where again $\delta_{k, j}$ is the standard Kronecker delta function, and we have used Lemma \ref{lemma:Yrecurrence}.
-
-\end{proof}
-
-The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$.
-In later sections we will use an ordering of the OPs so that they are grouped by Fourier mode $k$, which is convenient for the application of differential and other operators to the vector of coefficients of a given function's expansion (some operators will exploit this ordering for operators where Fourier modes do not interact, and thus will be block-diagonal).
-Before that though, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
-In practice, it is simply a matter of converting coefficients between the two orderings.
-To this end, we define our OP-building ordering as follows.
-For $n=0,1,2,\dots$:
-\begin{align*}
-	\bigscoptna := 
-		\begin{pmatrix}
-			\scopa_{n,0,0}(x,y,z) \\
-			\scopa_{n,1,0}(x,y,z) \\
-			\scopa_{n,1,1}(x,y,z) \\
-			\vdots \\
-			\scopa_{n,n,0}(x,y,z) \\
-			\scopa_{n,n,1}(x,y,z)
-		\end{pmatrix} \in \R^{2n+1}, 
-	\quad \quad 
-	\bigscopta := 
-		\begin{pmatrix}
-			\bigscopta_0 \\
-			\bigscopta_1 \\
-			\bigscopta_2 \\
-			\vdots \\
-		\end{pmatrix}
-\end{align*}
-and set $ J_x^{(a)},  J_y^{(a)},  J_z^{(a)}$ as the Jacobi matrices corresponding to
-\begin{align}
-	J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
-	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
-	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
-\end{align}
-where
-\begin{align*}
-	J_{x/y/z}^{(a)} &= 
-		\begin{pmatrix}
-			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
-			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
-			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
-			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
-			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}.
-\end{align*}
-Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are \textit{banded-block-banded matrices}:
-
-\begin{definition}
-	A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
-A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
-\end{definition}
-
-Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
-\begin{align*}
-	A^{(a)}_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A^{(a)}_{n,0,6} & 0 & & \\
-			A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
-			& \ddots & \ddots & \ddots & \\
-			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	B^{(a)}_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A^{(a)}_{n,0,4} & & \\
-			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
-			& & A^{(a)}_{n,n,3} & 0
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	C^{(a)}_{x,n} &:= 
-		\begin{pmatrix}
-			0 & A^{(a)}_{n,0,2} & & \\
-			A^{(a)}_{n,1,1} & \ddots & \ddots & \\
-			& \ddots & \ddots &A^{(a)}_{n,n-2,2} \\
-			& & \ddots & 0 \\
-			& & & A^{(a)}_{n,n,1} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
-where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align*}
-	A^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			\alphaa_{n,k,j} & 0 \\
-			0 & \alphaa_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
-	A^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			\alphaa_{n,0,j} & 0
-		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
-	A^{(a)}_{n,1,j} &:=
-		\begin{pmatrix}
-			\alphaa_{n,1,j} \\
-			0
-		\end{pmatrix} \in \R^{2\times1}, j \text{ odd}.
-\end{align*}
-For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
-\begin{align*}
-	A^{(a)}_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B^{(a)}_{n,0,6} & 0 & & \\
-			B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
-			& \ddots & \ddots & \ddots & \\
-			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	B^{(a)}_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B^{(a)}_{n,0,4} & & \\
-			B^{(a)}_{n,1,3} & \ddots & \ddots & \\
-			& \ddots & \ddots & B^{(a)}_{n,n-1,4} \\
-			& & B^{(a)}_{n,n,3} & 0
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	C^{(a)}_{y,n} &:= 
-		\begin{pmatrix}
-			0 & B^{(a)}_{n,0,2} & & \\
-			B^{(a)}_{n,1,1} & \ddots & \ddots & \\
-			& \ddots & \ddots & B^{(a)}_{n,n-2,2} \\
-			& & \ddots & 0 \\
-			& & & B^{(a)}_{n,n,1} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
-where for $k = 1,\dots,N, \: n = k,\dots,N$  
-\begin{align*}
-	B^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			0 & \betaa_{n,k,0,j} \\
-			\betaa_{n,k,1,j} & 0
-		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
-	B^{(a)}_{n,0,j} &:=
-		\begin{pmatrix}
-			0 & \betaa_{n,0,0,j}
-		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
-	B^{(a)}_{n,1,j} &:=
-		\begin{pmatrix}
-			0 \\
-			\betaa_{n,1,1,j}
-		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}.
-\end{align*}
-For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
-\begin{align*}
-	A^{(a)}_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{n,0,3} & 0 & \\
-			0 & \ddots & \ddots & & \\
-			& \ddots & \ddots & \ddots & \\
-			& & 0 & \Gamma^{(a)}_{n,n,3} & 0 \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
-	B^{(a)}_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{n,0,2} & \\
-			& \ddots & & \\
-			& & \ddots & \\
-			& & & \Gamma^{(a)}_{n,n,2}
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
-	C^{(a)}_{z,n} &:= 
-		\begin{pmatrix}
-			\Gamma^{(a)}_{n,0,1} & 0 & & \\
-			0 & \ddots & \ddots & \\
-			& \ddots & \ddots & 0 \\
-			& & \ddots & \Gamma^{(a)}_{n,n-1,1} \\
-			& & & 0 \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
-where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align}
-	\Gamma^{(a)}_{n,k,j} &:= 
-		\begin{pmatrix}
-			\gamma_{n,k,j} & 0 \\
-			0 & \gamma_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, \label{eqn:jacobisubblocksGamma1} \\
-	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:jacobisubblocksGamma2}
-\end{align}
-Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
-
-\subsection{Building the OPs}
-
-Following the triangle case \cite{olver2019triangle}, we use a multivariate analogue of Clenshaw's algorithm for evaluation, where we  combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
-\begin{align*}
-	\renewcommand\arraystretch{1.3}
-	\begin{pmatrix}
-		1 & & & \\
-		B_0-G_0(x,y,z) & A_0 & & \\
-		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
-		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
-		& & \ddots &\ddots
-	\end{pmatrix}
-	\bigscopta(x,y,z) =
-		\begin{pmatrix}
-	 		\scopa_0 \\ 0 \\ 0 \\ 0 \\ \vdots  \\
-		\end{pmatrix},
-\end{align*}
-where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
-\begin{align*}
-	A_n &:= 
-		\begin{pmatrix}
-			A^{(a)}_{x,n} \\
-			A^{(a)}_{y,n} \\
-			A^{(a)}_{z,n}
-		\end{pmatrix} \in \R^{3(2n+1)\times(2n+3)}, \quad
-	C_n := 
-		\begin{pmatrix}
-			C^{(a)}_{x,n} \\
-			C^{(a)}_{y,n} \\
-			C^{(a)}_{z,n}
-		\end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \nonumber \\
-	B_n &:= 
-		\begin{pmatrix}
-			B^{(a)}_{x,n} \\
-			B^{(a)}_{y,n} \\
-			B^{(a)}_{z,n}
-		\end{pmatrix} \in \R^{3(2n+1)\times(2n+1)}, \quad
-	G_n(x,y,z) := 
-		\begin{pmatrix}
-			xI_{2n+1} \\
-			yI_{2n+1} \\
-			zI_{2n+1}
-		\end{pmatrix} \in \R^{3(2n+1)\times(n+1)}.
-\end{align*}
-
-For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e.
-such that $\Dnt A_n = I_{2n+3}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
-\begin{align*}
-	\begin{cases}
-		\bigscopta_{-1}(x,y,z) := 0 \\
-		\bigscopta_{0}(x,y) := \scopa_0 \\
-		\bigscopta_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
-	\end{cases}
-\end{align*}
-
-Note that we can define an explicit $\Dnt$ as follows:
-\begin{align*}
-	\Dnt := 
-		\begin{pmatrix}
-			0 & & & 0 & & & (\Gamma^{(a)}_{n,0,3})^{-1} & & \\
-			& \ddots & & & \ddots & & & \ddots \\
-			& & 0 & & & 0 & & & (\Gamma^{(a)}_{n,n,3})^{-1}  \\
-			& & & & & \bm{\eta}^\top_{0} & & & \\
-			& & & & & \bm{\eta}^\top_{1} & & &
-		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
-\end{align*}
-for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}) for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by
-\begin{align*}
-	\big(\bm{\eta}_{0}\big)_j &= 
-		\begin{cases}
-			\frac{1}{\betaa_{n,n,1,6}} & j = 2(2n+1) \\
-			\frac{- \: \betaa_{n,n,1,5}}{\betaa_{n,n,1,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 3 \\
-			0 & o/w
-		\end{cases} \\
-	\big(\bm{\eta}_{1}\big)_j &= 
-		\begin{cases}
-			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
-			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2  \text{ and } n > 1 \\
-			0 & o/w
-		\end{cases}
-\end{align*}
-For $n=0$, we can simply take
-\begin{align*}
-	D^\top_0 &:= 
-		\begin{pmatrix}
-			0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
-			\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
-			0 & \frac{1}{\betaa_{0,0,6}} & 0
-		\end{pmatrix} \in \R^{3\times3}.
-\end{align*}
-
-It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigscopta_{0}(x,y,z)$  through $\bigscopta_{n}(x,y,z)$ in optimal $O(n^2)$ complexity.
-
-\begin{definition}\label{def:clenshawmats}
-	The \textit{recurrence coefficient matrices} associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
-\end{definition}
-
-%
-\section{Sparse partial differential operators}\label{Section:PDOs}
-
-In this section we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process.
-To this end, as alluded to in Section~\ref{subsection:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal.
-Let $N \in \N$ and define:
-\begin{align}
-	\bigscopNka &:= 
-		\begin{pmatrix}
-			\scopa_{k,k,0}(x,y,z) \\
-			\scopa_{k,k,1}(x,y,z) \\
-			\vdots \\
-			\scopa_{N,k,0}(x,y,z) \\
-			\scopa_{N,k,1}(x,y,z)
-		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
-	\bigscopa_{N,0} &:= 
-		\begin{pmatrix}
-			\scopa_{0,0,0}(x,y,z) \\
-			\vdots \\
-			\scopa_{N,0,0}(x,y,z) \\
-		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
-	\bigscopNa &:= 
-		\begin{pmatrix}
-			\bigscopa_{N,0} \\
-			\vdots \\
-			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
-\end{align}
-
-We further denote the weighted set of OPs on $\Omega$ by 
-\begin{align*}
-	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z),
-\end{align*}
-
-The operator matrices we derive here act on coefficient vectors, that represent a function $f(x,y,z)$ defined on $\Omega$ in spectral space -- such a function is approximated by its expansion up to degree $N$:
-\begin{align*}
-	f(x,y,z) = \bigscopNa(x,y,z)^\top \vec f = \sum_{n=0}^{N} \sum_{k=0}^{n} \sum_{i=0}^{1} f_{n,k,i} \: \scopnkia(x,y,z),
-\end{align*}
-where $\vec f = (f_{n,k,i})$ is the coefficients vector for the function $f$.
-
-\begin{definition}\label{def:differentialoperators}
-	Let $a$ be a nonnegative parameter, and $\tilde a \ge 2$ be a positive integer.
-Define the operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ according to:
-\begin{align*}
-	\rho \pfpx{\varphi}{f} (x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \mathbf{f}, \\
-	\rho \ppx{\varphi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\varphi^{(a)}\: \mathbf{f}, \\
-	\pfpx{\theta}{f}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
-	\DeltaS f(x,y,z) &= \bigscopN^{(a+\tilde a)}(x,y,z)^\top \: \calL^{(a)\to(a+\tilde a)} \: \mathbf{f}, \\
-	\DeltaS \big( \genjacw^{(a,0)}(z) \: f(x,y,z) \big) &= \bigW_N^{(a-\tilde a)}(x,y,z)^\top \: \calL_W^{(a)\to(a-\tilde a)} \: \mathbf{f},  \quad \text{ (for } a \ge 2 \text{ only)} \\
-	\DeltaS \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
-\end{align*}
-\end{definition}
-
-The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, on the triangle \cite{olver2018recurrence}, and on the disk-slice \cite{snowball2019sparse}.
-The operators we define here are for partial derivatives with respect to the spherical coordinates $(\varphi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).
-With the OP ordering by Fourier mode $k$ defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) these rotationally invariant operators are block-diagonal, meaning simple and parallelisable practical application.
-
-\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
-	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
-More specifically:
-\begin{itemize}
-	\item $D_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(2, 4)$
-  	\item $W_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(4, 2)$
-	\item $D_\theta$ is block-diagonal with sub-block-bandwidths $(1, 1)$
-	\item $\calL^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block-bandwidths $(0, 4)$
-	\item $\calL_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block-bandwidths $(4, 0)$
-	\item $\Delta^{(1)}_W$ is block-diagonal with sub-block-bandwidths $(2, 2)$
-\end{itemize}
-\end{theorem}
-
-In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short lemma.
-
-\begin{lemma}\label{lemma:Rsecondderivative}
-	For any general parameter $a$ and any $n = 0,1,\dots$, $k = 0,\dots,n$ we have that
-\begin{align*}
-	&\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
-	&\quad \quad = \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
-	&\quad \quad = \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
-\end{align*}
-where 
-\begin{align*}
-	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
-\end{align*}
-\end{lemma}
-
-\begin{proof}[Proof of Lemma \ref{lemma:Rsecondderivative}]
-	Since $\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] = \genjacw^{(a, 2k)} \: r_{n-k+1}$ where $r_{n-k+1}$ is a degree $n - k + 1$ polynomial, we have that 
-\begin{align*}
-	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)}
-\end{align*}
-for some coefficients $\tilde c_{\{n,k\},m}$. These coefficients are given by
-\begin{align*}
-	\tilde c_{\{n,k\},m} &= \frac{1}{\normgenjac^{(a,2k)}} \: \ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
-	&= - \frac{1}{\normgenjac^{(a,2k)}}  \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
-\end{align*}
-We show that these are zero for $m < n - k - 1$ by integrating twice by parts:
-\begin{align*}
-	&\ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
-	&\quad \quad \quad = - \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z \\
-	&\quad \quad \quad = \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: [(a+1) \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(0, 2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - 2(k+1) z \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(1, 0)} + \genjac_{m}^{(a,2k) \: \prime \prime} \: \genjacw^{(1, 2)}] \: \genjacw^{(a, 2k)} \: \D z
-\end{align*}
-which is indeed zero for $m < n - k - 1$ by orthogonality.
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{theorem:sparsityofdifferentialoperators}]
-
-For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that
-\begin{align*}
-	\pptheta \scopnkia(x,y,z) &= \genjacnmk^{(a,2k)}(z) \: \rho(z)^{k} \: \ddx{\theta} \chki(\theta) \\
-	&= 
-	\begin{cases}
-		(-1)^{i+1} \: k \: \scopa_{n,k,|i-1|}(x,y,z) & k > 0 \\
-		0 & k = 0
-	\end{cases}.
-\end{align*}
-
-We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$. The entries of the operator are given by the coefficients in the expansion 
-$$
-\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)},
-$$ 
-where the coefficients are 
-\begin{align*}
-	c_{m,j,h} = \norm{\scopmjh^{(a+1)}}^{-2}_{\scop^{(a+1)}} \ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}}.
-\end{align*}
-Now, note that:
-\begin{align*}
-	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \\
-	\rho(z) \: \rho'(z) &= -z \\
-	\ppphi \scopnkia(x,y,z) &= -\rho(z) \: \ddz \Big[ \rho(z)^k \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta), \\
-	\ppphi \Big[ \genjacw^{(a,0)}(z) \: \scopnkia(x,y,z) \Big] &= -\rho(z) \: \ddz \Big[ \genjacw^{(a,k)}(z) \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta).
-\end{align*}
-Then, 
-\begin{align*}
-	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
-	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \genjacw^{(a+1, 0)} \: \D z \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+1,2k)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, 2k)} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
-\end{align*}
-which is zero for $j \ne k$, $h \ne i$, and $m < n - 2$ by orthogonality.
-
-Similarly for the operator $W_\varphi^{(a)}$ for partial differentiation by $\varphi$ on the weighted space, the entries of the operator are given by the coefficients in the expansion $\rhoppphi (\genjacw^{(a,0)} \: \scopnkia) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-1,0)} \: \scopmjh^{(a-1)}$, where the coefficients are
-\begin{align*}
-	c_{m,j,h} = \norm{\scopmjh^{(a-1)}}^{-2}_{\scop^{(a-1)}} \ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}}.
-\end{align*}
-Now,
-\begin{align*}
-	&\ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}} \\
-	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \genjacw^{(a, k)}(z)] \: \genjacmmj^{(a-1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \D z \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a-1,2j)} \: [k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2] \: \genjacw^{(a-1, k + j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \:  \int_\alpha^1 \: \genjac_{m-k}^{(a-1,2k)} \: [k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2] \: \genjacw^{(a-1, 2k)} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} - a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \genjacw^{(1, 2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a-1, 2k)} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
-\end{align*}
-which is zero for $j \ne k$, $h \ne i$, and $m < n - 1$ by orthogonality.
-
-We move on to the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yield
-\begin{align}
-	\DeltaS \: \scopnki^{(a)} &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
-	&\quad \quad \quad \quad + \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
-	&= \chki(\theta) \rho(z)^k \Big\{ -k(k+1) \genjacnmk^{(a,2k)}(z) - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime}(z) \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime}(z) \Big\}, \label{eqn:oplaplacian} \\
-	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
-	&\quad \quad \quad \quad + \genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
-	&= \chki(\theta) \Big\{\genjacnmk^{(a,2k)}(z) \big[ -k(k+1) \genjacw^{(a,k)}(z) - 2a(k+1) z \: \genjacw^{(a-1,k)}(z) \big] \nonumber \\
-	&\quad \quad \quad \quad \quad + a(a-1) \genjacnmk^{(a,2k)}(z) \: \genjacw^{(a-2,k+1)}(z) \nonumber \\
-	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime}(z) \big[ -2(k+1) z \: \genjacw^{(a,k)}(z) + 2a \: \genjacw^{(a-1,k+2)}(z) \big] \nonumber \\
-	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime}(z) \: \genjacw^{(a,k+2)}(z) \Big\}. \label{eqn:woplaplacian}
-\end{align}
-
-For the operator $\calL^{(a)\to(a+\tilde a)}$ for the surface Laplacian on a non-weighted space, the entries of the operator are given by the coefficients in the expansion 
-$$\DeltaS \scopnki^{(a)} = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a + \tilde a)},$$
- where the coefficients are
-\begin{align*}
-	c_{m,j,h} = \norm{\scopmjh^{(a + \tilde a)}}^{-2}_{\scop^{(a + \tilde a)}} \ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}}.
-\end{align*}
-Using \bsrefeqn{eqn:oplaplacian}, and integrating by parts twice, we then have that
-\begin{align*}
-	&\ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}} \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a + \tilde a, 2j)} \: \genjacw^{(a + \tilde a), k+j)} \: \Big\{-k(k+1) \genjacnmk^{(a,2k)} - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime} \Big\} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a+\tilde a, 0)} \Big( -k(k+1) \genjacnmk^{(a,2k)} \: \rho^{2k} + \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \rho^{2(k+1)}] \Big) \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ -k(k+1) \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a+\tilde a, 2k)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - \genjacnmk^{(a,2k) \: \prime} \genjacw^{(a+\tilde a-1, 2k)} \: [\genjac_{m-k}^{(a+\tilde a,2k) \: \prime} \: \genjacw^{(1,0)} + (a + \tilde a) \genjac_{m-k}^{(a+\tilde a,2k)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z
-\end{align*}
-where $r_{m-k+\tilde a}$ is a degree $m - k + \tilde a$ polynomial in $z$, and so the above is zero for $n  - k > m - k + \tilde a \iff m < n - \tilde a$.
-
-For the operator $\calL_W^{(a)\to(a-\tilde a)}$ for the surface Laplacian on a weighted space, the entries of the operator are given by the coefficients in the expansion 
-\begin{align*}
-	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-\tilde a,0)} \: \scopmjh^{(a - \tilde a)}, 
-\end{align*}	
-where the coefficients are
-\begin{align*}
-	c_{m,j,h} = \norm{\scopmjh^{(a - \tilde a)}}^{-2}_{\scop^{(a - \tilde a)}} \ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}}.
-\end{align*}
-Using \bsrefeqn{eqn:woplaplacian}, and integrating by parts thrice, we then have that
-\begin{align*}
-	&\ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}} \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad  \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a - \tilde a, 2j)} \: \genjacw^{(a - 2, k+j)} \: \Big\{\genjacnmk^{(a,2k)} [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} [-2(k+1) z \: \genjacw^{(2,0)} + 2a \genjacw^{(1,2)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime} \: \genjacw^{(2,2)} \Big\} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \genjacw^{(a, 2k+2)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} ] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a [\genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a-1, 2k+2)} \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} ] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - a \genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a-2, 2k)} \: [(a-1) \rho^2 - 2(k+1) z \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z
-\end{align*}
-where $r_{m-k}$ is a degree $m - k$ polynomial in $z$, and so the above is zero for $n - k > m - k \iff m < n$.
-
-Finally, fix $a = 1$.
-For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are given by
-\begin{align*}
-	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
-\end{align*}
-Using \bsrefeqn{eqn:woplaplacian} with $a = 1$, and \bsreflemma{lemma:Rsecondderivative}, we then have that
-\begin{align*}
-	&\ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(1, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(1, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(1, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(1, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
-	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(1, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z
-\end{align*}
-where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in \bsreflemma{lemma:Rsecondderivative}.
-\end{proof}
-
-By applying these differential operators, we are (in some cases) incrementing or decrementing the parameter value $a$.
-It is therefore necessary to also be able to raise or lower the parameter by way of an independent operator.
-There exist conversion matrix operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
-\begin{definition}\label{def:parametertransformationoperators}
-Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T_W^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
-\begin{align*}
-	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z) \\
-	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z)
-\end{align*}
-\end{definition}
-
-\begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
-The operator matrices in Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
-\begin{itemize}
-	\item $T^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block bandwidths $(0,2\tilde a)$
-	\item $T_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block bandwidths $(2\tilde a, 0)$
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-We proceed with the case for the non-weighted operators $T^{(a)\to(a+\tilde a)}$.
-Since $\{\scopmjh^{(a+\tilde{a})}\}$ for $m = 0,\dots,n$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n$ polynomial, we can expand $\scopnkia = \sum_{m=0}^{n} \sum_{j=0}^m t_{m,j} \: \scopmjh^{(a+\tilde{a})}$.
-The coefficients of the expansion are then the entries of the operator matrix.
-We will show that the only non-zero coefficients are for $k = j$, $i = h$ and $m \ge  n - \tilde a$.
-Note that
-\begin{align*}
-	t_{m,j} = \norm{\scopmjh^{(a+\tilde{a})}}^{-2}_{\scop^{(a+\tilde a)}} \: \ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}}.
-\end{align*}
-where
-\begin{align*}
-	\ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}} &= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \cdot \: \Big( \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjacmmj^{(a+\tilde a,2j)} \: \rho^{k+j} \: \genjacw^{(a + \tilde a, 0)} \: \D z \Big) \\
-	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a + \tilde a, 2k)} \: \D z
-\end{align*}
-which is zero for $n > m + \tilde a \iff m < n - \tilde a$. The sparsity argument for the weighted parameter transformation operator follows similarly.
-\end{proof}
-
-\begin{figure}[t]
-	\centering % <-- added
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-laplacian}
-		\centering
-		\caption{The Laplace-Beltrami operator $\Delta^{(1)}_W$}
-	\end{subfigure}\hfil % <-- added
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-rho2laplacian}
-		\centering
-		\caption{The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$}
-	\end{subfigure}\hfil % <-- added
-
-	\medskip
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-helmholtz}
-		\centering
-		\caption{A variable coefficient Helmholtz operator}
-	\end{subfigure}\hfil % <-- added
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-biharmonic}
-		\centering
-		\caption{The Biharmonic operator $\calB_W^{(2)}$} 
-	\end{subfigure}\hfil % <-- added
-	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
-For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
-	\label{fig:sparsity}
-\end{figure}
-
-\subsection{Further partial differential operators}\label{subsection:furtherdiffoperators}
-General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
-As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored spherical Laplacian $\rho(z)^2 \: \DeltaS$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
-Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in Definition \ref{def:differentialoperators}.
-The matrix operator for this $\rho^2$-factored spherical Laplacian acting on the coefficients vector is then given by
-\begin{align*}
-	D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
-\end{align*}
-Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
-
-Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions. That is, 
-\begin{align*}
-	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \nabla_S u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
-\end{align*}
-where $\partial\Omega$ is the $z=\alpha$ boundary, and $\unitvec{n}(x,y,z)$ is the outward unit normal vector at the point $(x,y,z)$ on the boundary, i.e.
-$\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$.
-The matrix operator for the Biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
-To construct this, we can simply multiply together two of the spherical Laplacian operators defined in \bsrefdef{def:differentialoperators}, namely $\calL^{(0)\to(2)}$ and $\calL_W^{(2)\to(0)}$:
-\begin{align*}
-	\calB^{(2)}_W := \calL^{(0)\to(2)} \: \calL_W^{(2)\to(0)}.
-\end{align*}
-Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,y,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$ and thus its spherical gradient would be zero on the boundary $z = \alpha$.
-This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
-The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
-
-%
-\section{Computational aspects}\label{Section:Computation}
-
-In this section we discuss how to expand and evaluate functions in our proposed basis, and take advantage of the sparsity structure in partial differential operators in practical computational applications.
-
-\subsection{Constructing $\genjac_n^{(a,b)}(x)$}
-
-It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
-
-\subsection{Quadrature rule on the spherical cap}\label{subsection:quadrule}
-
-In this section we construct a quadrature rule exact for polynomials on the spherical cap $\Omega$ that can be used to expand functions in the OPs $\scopnkia(x,y,z)$ for a given parameter $a$.
-
-\begin{theorem}\label{Theorem:quadrule}
-Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
-Further, denote the $M_2$ Gauss quadrature nodes and weights $[-1,1]$ with weight $(1 - x^2)^{-\half}$ as $(s_j, w_j^{(s)})$.
-Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
-\begin{align*}
-	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
-	z_{l+(j-1)M_2} &:= t_j, \\
-	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
-\end{align*}
-Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
-\begin{align*}
-	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \approx \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
-\end{align*}
-where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
-\end{theorem}
-
-\begin{remark} 
-	Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[3.5.23]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
-\end{remark}
-
-\begin{proof}
-Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-\begin{align*}
-	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
-	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
-\end{align*}
-so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
-Note that if $f$ is a polynomial, then $f_e(\rho(t)x, \rho(t)y, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $(x,y) \in \R^2$.
-
-Firstly, we note that
-\begin{align*}
-	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
-\end{align*}
-for some function $g$. Then, integrating the even function $f_e$ we have
-\begin{align*}
-	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \D x \Big) \: \D z \\
-	&\quad \quad \approx \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
-	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) s_l, \rho(t_j) \sqrt{1 - s_l^2}, t_j\big) \quad (\star \star) \\
-	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
-\end{align*}
-Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
-It follows that $s \mapsto f_e\big(\rho(z)s,\rho(z)\sqrt{1-s^2}, z\big)$ for fixed $z$ is then a polynomial of degree $\le N$.
-We therefore achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-
-Integrating the odd function $f_o$ results in
-\begin{align*}
-	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) + f_o\big(\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) - f_o\big(\rho(z)x, \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
-	&\quad \quad = 0.
-\end{align*}
-since $f_o(x,y,z) = - f_o(-x, -y, z)$. Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
-\begin{align*}
-	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D A  \\
-	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
-	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
-\end{align*}
-where $M = M_1 M_2$ and $2M_1 - 1 \ge N, 2M_2 - 1 \ge N$.
-\end{proof}
-
-%\begin{theorem}\label{Theorem:quadrule}
-%Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
-Further, we denote the $M_2$ Gauss quadrature nodes and weights for the unit circle \cite{???}\bstodo{good reference?
-Or derive myself using Chebychev polys...} by $(\mathbf{s}_j, w_j^{(s)})$ where for $j = 1,\dots,M_2$:
-%\begin{align*}
-%	\mathbf{s}_j &= \Big( \cos\fpr(\frac{2 \pi j}{M_2}), \: \sin\fpr(\frac{2 \pi j}{M_2}) \Big), \\
-%	w_j^{(s)} &= \frac{2\pi}{M_2}.
-%\end{align*}
-%Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
-%\begin{align*}
-%	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
-%	z_{l+(j-1)M_2} &:= t_j, \\
-%	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
-%\end{align*}
-%Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
-%\begin{align*}
-%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
-%\end{align*}
-%where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
-%\end{theorem}
-%
-%\begin{proof}
-%Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-%\begin{align*}
-%	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
-%	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
-%\end{align*}
-%so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
-Note that if $f$ is a polynomial, then $f_e(\rho(t)\xvec, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $\xvec \in \R^2$.
-%
-%Now, we have that
-%\begin{align*}
-%	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_e\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \\
-%	&\quad \quad \approx\int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) \mathbf{s}_l, z\big) \Big) \: \D z \quad (\star) \\
-%	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) \mathbf{s}_l, t_j\big) \quad (\star \star) \\
-%	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
-%\end{align*}
-%Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
-We therefore achieve equality at $(\star)$ if $M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-\bstodo{For the circle rule, is this bound on N true?}
-%
-%Next, note that 
-%\begin{align*}
-%	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_o\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \quad (\dagger) \\
-%	&\quad \quad = 0
-%\end{align*}
-%since the inner integral at $(\dagger)$ over $\bm{\xi}$ is zero, due to the symmetry over the domain.
-%
-%Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
-%\begin{align*}
-%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z  \\
-%	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
-%\end{align*}
-%where $M = M_1 M_2$ and $2M_1 - 1 \ge N, M_2 - 1 \ge N$.
-%\end{proof}
-
-\subsection{Obtaining the coefficients for expansion of a function on the spherical cap}\label{subsection:expandingfunctions}
-
-Fix $a \in \R$. Then for any function $f : \Omega \to \R$ we can express $f$ by
-\begin{align*}
-	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \bm{f}_k = \bigscopNa(x,y,z)^\top \: \bm{f}
-\end{align*}
-for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) and where
-
-\begin{align*}
-	\bm{f}_k &:= 
-		\begin{pmatrix}
-			f_{k,k,0} \\
-			f_{k,k,1} \\
-			\vdots \\
-			f_{N,k,0} \\
-			f_{N,k,1}
-		\end{pmatrix} \in \R^{2(N-k+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	\bm{f}_0 := 
-		\begin{pmatrix}
-			f_{0,0,0} \\
-			\vdots \\
-			f_{N,0,0}
-		\end{pmatrix} \in \R^{N+1}, \\
-	\bm{f} &:= 
-		\begin{pmatrix}
-			\bm{f}_0 \\
-			\vdots \\
-			\bm{f}_N
-		\end{pmatrix} \in \R^{2(N+1)^2}, \quad \quad
-	f_{n,k,i} := \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
-\end{align*}
-Recall from equation (\ref{eqn:normscop}) that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$.
-Using the quadrature rule detailed in Section \ref{subsection:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$:
-\begin{align*}
-	f_{n,k,i} &= \frac{1}{2 \: \normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
-	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
-\end{align*}
-where the quadrature nodes and weights are those from Theorem \ref{Theorem:quadrule}, and $M = M_1 M_2$ with $2M_1 - 1 \ge N, M_2 - 1 \ge N$ (i.e.
-we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).
-
-\subsection{Function evaluation}\label{subsection:functionevaluation}
-
-For a function $f$, with coefficients vector $\bm{f}$ for expansion in the $\{\scopnki\}$ basis as determined via the method in Section \ref{subsection:expandingfunctions} up to order $N$, we can use the Clenshaw algorithm to evaluate the function at a point $(x,y,z) \in \Omega$ as follows.
-Let $A_n, B_n, \Dnt, C_n$ be the Clenshaw matrices from Definition \ref{def:clenshawmats}, and define the rearranged coefficients vector $\tilde{\bm{f}}$ via
-\begin{align*}
-	{\bm{f}}_n &:= 
-		\begin{pmatrix}
-			f_{n,0,0} \\
-			f_{n,1,0} \\
-			f_{n,1,1} \\
-			\vdots \\
-			f_{n,n,0} \\
-			f_{n,n,1}
-		\end{pmatrix} \in \R^{2(N+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	{\bm{f}}_0 = f_{0,0,0} \in \R, \\
-	{\bm{f}} &:= 
-		\begin{pmatrix}
-			{\bm{f}}_0 \\
-			\vdots \\
-			{\bm{f}}_N
-		\end{pmatrix} \in \R^{(N+1)^2}.
-\end{align*}
-The trivariate Clenshaw algorithm works similar to the bivariate Clenshaw algorithm introduced in \cite{olver2019triangle} for expansions in the triangle:
-\begin{align*}
-	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
-	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {{\bm{f}}_n}^T - \bm{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
-\end{align*}
-
-\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
-
-The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
-We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^2)$.
-This proves much cheaper computationally than using the 3D quadrature rule to calculate the surface inner products, which has a complexity of $\bigO(N^3)$.
-
-\subsection{Obtaining operator matrices for variable coefficients}\label{subsection:operatorclenshaw}
-
-The Clenshaw algorithm outlined in Section \ref{subsection:functionevaluation} can also be used with Jacobi matrices $J^{(a)}_x, J^{(a)}_y, J^{(a)}_z$ replacing the point $(x,y,z)$.
-Let $v : \Omega \to \R$ be the function that we wish to obtain an operator matrix $V$ for $v$, so that
-\begin{align*}
-	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \bm{f} \: \bigscopta(x,y,z) = (V \bm{f})^\top \: \bigscopta(x,y,z),
-\end{align*}
-i.e. $V \bm{f}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$.
-
-To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in Section \ref{subsection:functionevaluation} so that $v(x,y,z) = {\bm{v}}^\top \: \bigscopta(x,y,z)$).
-Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, $Z := (J_z^{(a)})^\top$.
-The operator $V$ is then the result of the following:
-\begin{align*}
-	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
-	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
-\end{align*}
-where at each iteration, $\bm{\xi}_n$ is a vector of matrices.
-
-%
-\section{Examples on spherical caps with zero Dirichlet conditions}\label{Section:Examples}
-
-We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions on the spherical cap defined by $\Omega$.
-We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, as well as a time dependent heat equation, demonstrating the versatility of the approach.
-
-\subsection{Poisson}
-
-% FIGURES
-
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{poisson-u=wyexpx-N=60}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-poisson-epsilonfun-N=200}
-        	%\label{fig:solutionblocknorms-poisson}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
-Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
-This indicates spectral convergence.}
-	\centering
-	\label{fig:poisson}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{helmholtz-f=wyexpx-N=60-n=300}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-varyingk-N=200}
-        	%\label{fig:solutionblocknorms-poisson}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
-Right: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
-This indicates spectral convergence.}
-	\centering
-	\label{fig:helmholtz}
-\end{figure}
-
-\begin{figure}[t]
-	\centering % <-- added
-	\includegraphics[scale=0.35]{complexity-Nend=1000}
-	\caption{Time in seconds to  to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$.
-This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
-Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$.}
-	\centering
-	\label{fig:complexity}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.15]{biharmonic-f=erf-N=80-n=50}
-	\centering
-	%\label{fig:solution-biharmonic}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-expfun-N=200}
-	\centering
-        	%\label{fig:solutionblocknorms-biharmonic}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
-This demonstrates algebraic convergence.}
-	\centering
-	\label{fig:biharmonic}
-\end{figure}
-% END FIGURES
-
-The Poisson equation is the classic problem of finding $u(x,y,z)$ given a function $f(x,y,z)$ such that:
-\begin{align}
-	\begin{cases}
-\DeltaS u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0& \quad \text{on } \partial \Omega
-	\end{cases}
-	\label{eqn:poisson}
-\end{align}
-noting the imposition of zero Dirichlet boundary conditions on $u$.
-
-We can tackle the problem as follows.
-Choose an $N \in \N$ large enough for the problem, and denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}.
-In matrix-vector notation, our system hence becomes:
-\begin{align*}
-    \Delta_W^{(1)} \mathbf{u} = \mathbf{f}
-\end{align*}
-which can be solved to find $\mathbf{u}$.
-In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
-The right hand sides we choose here are given by
-\begin{align*}
-	f(x,y,z) &= \norm{\Big(x - (\epsilon + 1 / \sqrt{3}), \: y - (\epsilon + 1 / \sqrt{3}), \: z - (\epsilon + 1 / \sqrt{3})\Big)^\top}
-\end{align*}
-for differing choices of $\epsilon$ -- this parameter serves to alter the distance from which we would have a singularity.
-In the plot, a \enquote{block} is simply the group of coefficients corresponding to OPs of the same degree, and so the plot shows how the norms of these blocks decay as the degree of the expansion increases.
-Thus, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
-We see that we achieve spectral convergence for these examples.
-
-\subsection{Inhomogeneous variable-coefficient Helmholtz}
-
-Find $u(x,y)$ given functions $v$, $f : \Omega \to \R$ such that:
-\begin{align}
-	\begin{cases}
-    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0 &\quad \text{on } \partial \Omega
-	\end{cases}
-	\label{eqn:helmholtz}
-\end{align}
-where $k \in \R$, noting the imposition of zero Dirichlet boundary conditions on $u$.
-
-We can tackle the problem as follows.
-Denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}.
-
-Define $X := (J_x^{(0)})^\top$, $Y := (J_y^{(0)})^\top$, $Z := (J_z^{(0)})^\top$.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
-We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $v(X, Y, Z)$.
-In matrix-vector notation, our system hence becomes:
-\begin{align*}
-    (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \mathbf{u} = \mathbf{f}
-\end{align*}
-which can be solved to find $\mathbf{u}$.
-We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
-Here, we use $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
-Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve spectral convergence.
-
-In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools.jl package \cite{BenchmarkTools.jl-2016} in Julia.}  to construct the operator for $\DeltaS + v(x,y,z)$, this time with a rotationally invariant $v(x,y,z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
-The plot demonstrates that as we increase the degree of approximation $N$, we achieve a complexity of an optimal $\bigO(N^2)$.
-
-What about other boundary conditions?
-One simple extension is the case where the value on the boundary takes that of a function depending only on $x$ and $y$, i.e.
-$c = c(x,y)$.
-In this case, the problem
-\begin{align*}
-	\begin{cases}
-    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = c(x,y) &\quad \text{on } \partial \Omega
-	\end{cases}
-\end{align*}
-is equivalent to letting $u(x,y,z) = \tilde{u}(x,y,z)+ c(x,y)$ and solving
-\begin{align*}
-	\begin{cases}
-    		\DeltaS \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)  \quad \text{in } \Omega \\
-		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
-	\end{cases}
-\end{align*}
-for $\tilde u$. This new problem is then a zero boundary condition Helmholtz problem with right hand side 
-\begin{align*}
-	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)
-\end{align*}
-for $(x,y,z) \in \Omega$. Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
-\begin{align*}
-	\DeltaS \: c(x,y) = {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) = - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
-\end{align*}
-since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only, which are precisely the Fourier coefficients of $c(\cos \theta, \sin \theta)$.
-Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manor.
-
-\subsection{Biharmonic equation}
-
-Our last erxample is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
-\begin{align}
-	\begin{cases}
-    		\DeltaS^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = \grad_S \: u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
-	\end{cases}
-	\label{eqn:biharmonic}
-\end{align}
-where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see Section \ref{subsection:furtherdiffoperators}).
-In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
-Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve exponential convergence for these more complex functions.
-
-%
-\section{Conclusions}
-
-We have shown that trivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary.
-We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: (1-z^2)^{b/2}$ defined for the disk-slice case \cite{snowball2019sparse}.
-Generalisation to spherical bands ($\alpha \leq z \leq \beta$) is straightforward.
-This work thus forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.
-
-This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere---it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
-The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct the necessary orthogonal polynomials.
-
-The next stage is to develop an orthogonal basis for the tangent space of the spherical cap (or band), and obtain sparse differential operators for gradient, divergence etc.
-On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
-
-%\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.
-The second author was supported in part by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces.
-
-%
-%%
-%\appendix
-%%
-%\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
-%
-%\bstodoinline{Is this worth including?}
-%
-%\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis} \bstodoinline{Yes thought so too}
-%
-%%\begin{figure}[t]
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-%%        	%\label{fig:solutionblocknorms-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-poisson}
-%%	\centering
-%%	%\label{fig:solution-poisson}
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-%%	%\label{fig:solution-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\centering
-%%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
-Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
-Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-in the trapezium.}
-%%	\centering
-%%	\label{fig:appendixfigs}
-%%\end{figure}
-%
-%We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
-Consider the 3D Dirichlet problem on a domain $\Omega$:
-%\begin{align*}
-%	\begin{cases}
-%         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
-%         u = 0 \quad \text{on } \partial \Omega
-%         \end{cases}
-%\end{align*}
-%This has the weak formulation
-%\begin{align*}
-%	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
-%\end{align*}
-% for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
-%
-%In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$.
-While spherical bands are an extension to this work, we note that we can still apply a $p$-FEM using two half-spheres ($\alpha = 0$).
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain.
-We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p \rvert_\element) \le p\}$ for some $p \in \N$.
-%
-%We seek $u_p \in V_p$ s.t.
-%\begin{align}
-%	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-%	\label{eqn:FEMweakform}
-%\end{align}
-%
-%Define 
-%\begin{align*}
-%	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
-%\end{align*}
-%for any $a$.
-Note that due to orthogonality this is a diagonal matrix.
-By choosing suitable bases for our FE spaces, we can rewrite (\ref{eqn:FEMweakform}) in matrix-vector form.
-We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.
-%
-%Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
-%\begin{align*}
-%	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
-%					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
-%\end{align*}
-%Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
-Then,
-%\begin{align*}
-%	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
-%	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%
-%Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-%\begin{align*}
-%	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
-%\end{align*}
-%(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-%\begin{align*}
-%	A\mathbf{u} = \tilde{\mathbf{f}}.
-%\end{align*}
-%where the (element) stiffness matrix $A$ is defined by 
-%\begin{align*}
-%	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
-%\end{align*}
-%and the load vector $\tilde{\mathbf{f}}$ is given by 
-%\begin{align*}
-%	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-%
-
-\bibliography{spherical-caps}
+\begin{enumerate}
+\item Constructs sparse discretisations for partial differential operators on spherical caps using orthogonal polynomials. 
+\item Approximations are observed to converge spectrally fast.
+\item Optimal complexity is achieved for operators that are invariant to rotation around the pole.
+\item Examples include (surface) Poisson equation, Helmholtz equation with variable coefficients, and the biharmonic equation.
+\end{enumerate}
 

### 2021-02-11 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Edits to SH chapter

@@
-\chapter{Spherical harmonics as orthogonal polynomials in three variables}
+\chapter{Spherical harmonics as orthogonal polynomials}
 
@@
 		\begin{cases}
-			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \tilde \alpha_{l,m,1} \quad \text{if } m>0 \\
-			\frac{\alphalm}{2\alpha^{m-1}_{l-1}} \tilde \alpha_{l,|m|,2} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
+			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,m,1} \quad \text{if } m>0 \\
+			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,|m|,2} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
 			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
@@
 		\begin{cases}
-			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \tilde \alpha_{l,m,2} \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
-			\frac{\alphalm}{2\alpha^{m+1}_{l-1}} \tilde \alpha_{l,|m|,1} \quad \text{if } m<0 \\
+			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,m,2} \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
+			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,|m|,1} \quad \text{if } m<0 \\
 			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
@@
 		\begin{cases}
-			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \tilde \alpha_{l,m,3} \quad \text{if } m>0 \\
-			\frac{\alphalm}{2\alpha^{m-1}_{l+1}} \tilde \alpha_{l,|m|,4} \quad \text{if } m\le0 
+			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,m,3} \quad \text{if } m>0 \\
+			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,|m|,4} \quad \text{if } m\le0 
 		\end{cases} \\
@@
 		\begin{cases}
-			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
-			\frac{\alphalm}{2\alpha^{m+1}_{l+1}} \tilde \alpha_{l,|m|,3} \quad \text{if } m<0 
+			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
+			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,|m|,3} \quad \text{if } m<0 
 		\end{cases} \\
@@
 
-\subsection{Variable coefficients as operators}
+\subsection{Obtaining operator matrices for variable coefficients}
 
@@
 %%%
-\section{Tangent space -- vector spherical harmonics as orthogonal vectors in three variables}
+\section{Vector spherical harmonics as orthogonal vectors in three variables}
 
@@
 
-\subsection{Vector spherical harmonics as a vector-weighted sum of scalar spherical harmonics} %{Spin-1 tensor spherical harmonics}
-
-The vector spherical harmonics in \bsrefeqn{eqn:VSHs} can be expressed in terms of \textit{spin-1 tensor spherical harmonics}, which in turn can each be expressed as a vector-weighted sum of spherical harmonics \cite{khersonskii1988quantum}.
-
-We start by defining what we mean by a spin-1 tensor spherical harmonic. In general, the tensor spherical harmonic is given by, for $2l, \, j, \, 2s \in \No$,
+\subsection{Jacobi matrices} %{Spin-1 tensor spherical harmonics}
+
+\bstodoinline{cite C-G coeffs, spin-1 tensor harmonics, and the combination expression in \label{eqn:VSHsalt}}
+\bstodoinline{get better reference than that Winter 2017 physics document}
+
+We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc.
+in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$.
+To this end, we use the helpful property that the vector spherical harmonics as defined in \bsrefeqn{eqn:VSHs} can be expressed in terms of \textit{spin-1 tensor spherical harmonics}, which in turn can each be expressed as a vector-weighted sum of spherical harmonics \cite{khersonskii1988quantum}.
+
+We start by defining what we mean by a spin-1 tensor spherical harmonic. In general, the tensor spherical harmonic is given by, for $2l, \, j, \, 2s \in \No$, \cite{}
 \bseqn{
@@
 }
-where $\vec{\chi}_{s, m_s}$ are the simultaneous eigenstates of the spin operators $\bold{S}^2$ and $S_z$, and where $\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle$ is a Clebsch-Gordan coefficient \cite{} \bstodo{cite C-G coeffs}.
-We note that a property of the Clebsch--Gordan coefficients means that they vanish unless $|j - s| \le l \le j + s$.
-We further note that there are simple calculable expressions for the Clebsch--Gordan coefficients when the spin $s=1$.
+where $\vec{\chi}_{s, m_s}$ are the simultaneous eigenstates of the spin operators $\bold{S}^2$ and $S_z$, and where $\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle$ is a Clebsch--Gordan coefficient \cite{} .
+We note that a property of the Clebsch--Gordan coefficients means that they vanish unless $|j - s| \le l \le j + s$ \cite{}.
+We further note that there are simple calculable expressions for the Clebsch--Gordan coefficients when the spin $s=1$ \cite{}.
 
@@
 \bseqn{
-	\curlyy^{l \pm1,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
-	\curlyy^{l,s}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
+	\curlyy^{l \pm1,1}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
+	\curlyy^{l,1}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
 }
@@
 \bseqnnumber{
-	\gradYlm &= \mu_{l,-1} \, \curlyy^{l-1,1}_{l,m} + \mu_{l,1} \curlyy^{l+1,1}_{l,m}, \quad \gradpYlm = \mu_{l,0} \, \curlyy^{l,1}_{l,m},\label{eqn:VSHsalt}
-}
-where \bstodoinline{get better reference than that Winter 2017 physics document}
+	\gradYlm &= \mu_{l,-1} \, \curlyy^{l-1,1}_{l,m} + \mu_{l,1} \curlyy^{l+1,1}_{l,m}, \quad \gradpYlm = \mu_{l,0} \, \curlyy^{l,1}_{l,m}, \label{eqn:VSHsalt}
+}
+where
 \bseqnnumber{
@@
 
-\subsection{Jacobi matrices}
-
-We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc. in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$. 
+The expressions for the VSHs as given in \bsrefeqn{eqn:VSHsalt} allow us to be able to find the coefficients for our desired expressions for $x \gradYlm$, $y \gradYlm$ and $z \gradYlm$. 
 \begin{lemma}\label{lemma:VSHrecurrences}
@@
 \bseqn{
-	x \gradYlm &= A_{l,m,1} \gradY^{m-1}_{l-1} + A_{l,m,2} \gradY^{m+1}_{l-1} + A_{l,m,5} \gradY^{m-1}_{l+1} + A_{l,m,6} \gradY^{m+1}_{l+1} \\
-& \quad \quad \quad + A_{l,m,3}^{*} \gradpY^{m-1}_{l} + A_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
-	x \gradpYlm &= A_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + A_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + A_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + A_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
-	& \quad \quad \quad + A_{l,m,3} \gradY^{m+1}_{l} + A_{l,m,4} \gradY^{m-1}_{l}, \\
-	y \gradYlm &= B_{l,m,1} \gradY^{m-1}_{l-1} + B_{l,m,2} \gradY^{m+1}_{l-1} + B_{l,m,5} \gradY^{m-1}_{l+1} + B_{l,m,6} \gradY^{m+1}_{l+1} \\
-& \quad \quad \quad + B_{l,m,3}^{*} \gradpY^{m-1}_{l} + B_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
-	y \gradpYlm &= B_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + B_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + B_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + B_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
-	& \quad \quad \quad + B_{l,m,3} \gradY^{m+1}_{l} + B_{l,m,4} \gradY^{m-1}_{l}, \\
-	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l}, \\
-	z \gradpYlm &= \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l}
-}
-where $*$ denotes the complex conjugate, and where, for any valid $m_s$ value,
+	x \gradYlm &= A_{l,m,1} \gradY^{m-1}_{l-1} + A_{l,m,2} \gradY^{m+1}_{l-1} + A_{l,m,3} \gradY^{m-1}_{l+1} + A_{l,m,4} \gradY^{m+1}_{l+1} \\
+& \quad \quad \quad + A_{l,m,5} \gradpY^{m-1}_{l} + A_{l,m,6} \gradpY^{m+1}_{l}, \\
+	x \gradpYlm &= A_{l,m,1} \gradpY^{m-1}_{l-1} + A_{l,m,2} \gradpY^{m+1}_{l-1} + A_{l,m,3} \gradpY^{m-1}_{l+1} + A_{l,m,4} \gradpY^{m+1}_{l+1} \\
+& \quad \quad \quad - A_{l,m,5} \gradY^{m-1}_{l} - A_{l,m,6} \gradY^{m+1}_{l}, \\
+	y \gradYlm &= B_{l,m,1} \gradY^{m-1}_{l-1} + B_{l,m,2} \gradY^{m+1}_{l-1} + B_{l,m,3} \gradY^{m-1}_{l+1} + B_{l,m,4} \gradY^{m+1}_{l+1} \\
+& \quad \quad \quad + B_{l,m,5} \gradpY^{m-1}_{l} + B_{l,m,6} \gradpY^{m+1}_{l}, \\
+	y \gradpYlm &= B_{l,m,1} \gradpY^{m-1}_{l-1} + B_{l,m,2} \gradpY^{m+1}_{l-1} + B_{l,m,3} \gradpY^{m-1}_{l+1} + B_{l,m,4} \gradpY^{m+1}_{l+1} \\
+& \quad \quad \quad - B_{l,m,5} \gradY^{m-1}_{l} - B_{l,m,6} \gradY^{m+1}_{l}, \\
+	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,2} \gradY^{m}_{l+1} + \Gamma_{l,m,3} \gradpY^{m}_{l}, \\
+	z \gradpYlm &= \Gamma_{l,m,1} \gradpY^{m}_{l-1} + \Gamma_{l,m,2} \gradpY^{m}_{l+1} - \Gamma_{l,m,3} \gradY^{m}_{l}
+}
+%\bseqn{
+%	x \gradYlm &= A_{l,m,1} \gradY^{m-1}_{l-1} + A_{l,m,2} \gradY^{m+1}_{l-1} + A_{l,m,5} \gradY^{m-1}_{l+1} + A_{l,m,6} \gradY^{m+1}_{l+1} \\
+%& \quad \quad \quad + A_{l,m,3}^{*} \gradpY^{m-1}_{l} + A_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
+%	x \gradpYlm &= A_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + A_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + A_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + A_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
+%	& \quad \quad \quad + A_{l,m,3} \gradY^{m+1}_{l} + A_{l,m,4} \gradY^{m-1}_{l}, \\
+%	y \gradYlm &= B_{l,m,1} \gradY^{m-1}_{l-1} + B_{l,m,2} \gradY^{m+1}_{l-1} + B_{l,m,5} \gradY^{m-1}_{l+1} + B_{l,m,6} \gradY^{m+1}_{l+1} \\
+%& \quad \quad \quad + B_{l,m,3}^{*} \gradpY^{m-1}_{l} + B_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
+%	y \gradpYlm &= B_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + B_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + B_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + B_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
+%	& \quad \quad \quad + B_{l,m,3} \gradY^{m+1}_{l} + B_{l,m,4} \gradY^{m-1}_{l}, \\
+%	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l}, \\
+%	z \gradpYlm &= \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l}
+%}
+%where $*$ denotes the complex conjugate, and where, for any valid $m_s$ value,
+where, for any valid $m_s$ value,
 \bseqn{
@@
 	A_{l,m,2} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} \alpha_{l-1,m-m_s,2}, \\
-	A_{l,m,3} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,3} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,1} \\
-	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,5} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - A_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big], \\	
-	A_{l,m,4} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,4} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,2} \\
-	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,6} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - A_{l,m,2} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
-	A_{l,m,5} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} \alpha_{l+1,m-m_s,3}, \\
-	A_{l,m,6} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} \alpha_{l+1,m-m_s,4}, \\
+	A_{l,m,3} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} \alpha_{l+1,m-m_s,3}, \\
+	A_{l,m,4} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} \alpha_{l+1,m-m_s,4}, \\
+	A_{l,m,5} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,3} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,1} \\
+	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,3} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - A_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big], \\	
+	A_{l,m,6} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,4} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,2} \\
+	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,4} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - A_{l,m,2} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
 	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6, \\
-	\Gamma_{l,m,1} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} \gamma_{l-1,m-m_s,1}, \\
-	\Gamma_{l,m,2} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \gamma_{l-1,m-m_s,2} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \gamma_{l+1,m-m_s,1} \\
-	& \quad \quad \quad \quad \quad \quad \quad - \Gamma_{l,m,3} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \Gamma_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big], \\
-	\Gamma_{l,m,3} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} \gamma_{l+1,m-m_s,2},
-}
-and where $\cgcoeff^{l,m_s}_{l',m'}$, $\mu_{l,k}$ are defined in \bsrefeqn{eqn:cgcoeffdef} and \bsrefeqn{eqn:spintensorvshcoeffsdef} respectively.
+	\Gamma_{l,m,1} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m}} \gamma_{l-1,m-m_s,1}, \\
+	\Gamma_{l,m,2} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m}} \gamma_{l+1,m-m_s,2}, \\
+	\Gamma_{l,m,3} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \gamma_{l-1,m-m_s,2} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \gamma_{l+1,m-m_s,1} \\
+	& \quad \quad \quad \quad \quad \quad \quad - \Gamma_{l,m,3} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \Gamma_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big],
+}
+and where $\cgcoeff^{L,m_s}_{l,m}$, $\mu_{l,k}$ are defined in \bsrefeqn{eqn:cgcoeffdef} and \bsrefeqn{eqn:spintensorvshcoeffsdef} respectively.
 \end{lemma}
@@
 
-\bstodoinline{Do the derivation of these rec coeffs, to have as the proof (and check they are correct!)}
+\bstodoinline{Do the derivation of these rec coeffs, to have as the proof (and check they are correct!). Need to account for l=1 case.}
+
+Fix $l \in \N, m\in\{-l,\dots,l\}$, and define $Y_{-1}^m := 0$ for any $m \in \Z$.
+We proceed first with the case for multiplication by $z$.
+Using the definitions given for $\Gamma_{l,m,1}, \Gamma_{l,m,2}, \Gamma_{l,m,3}$ in the Lemma, we have that
+\bseqn{
+	&\Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l} \\
+	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: \Big\{ \Gamma_{l,m,1} \mu_{l-1,-1} \cgcoeff^{l-2, m_s}_{l-1, m} Y^{m-m_s}_{l-2} + \Gamma_{l,m,1} \mu_{l-1,1} \cgcoeff^{l, m_s}_{l-1, m} Y^{m-m_s}_{l} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad + \Gamma_{l,m,3} \mu_{l+1,-1} \cgcoeff^{l, m_s}_{l+1, m} Y^{m-m_s}_{l} + \Gamma_{l,m,3} \mu_{l+1,1} \cgcoeff^{l+2, m_s}_{l+1, m} Y^{m-m_s}_{l+2} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad + \Gamma_{l,m,2} \mu_{l,0} \cgcoeff^{l, m_s}_{l, m} Y^{m-m_s}_{l} \Big\} \\
+	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: \Big\{ Y^{m-m_s}_{l-2} \Gamma_{l,m,1} \mu_{l-1,-1} \cgcoeff^{l-2, m_s}_{l-1, m} + Y^{m-m_s}_{l+2} \Gamma_{l,m,3} \mu_{l+1,1} \cgcoeff^{l+2, m_s}_{l+1, m} \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad + Y^{m-m_s}_{l} \Big[  \Gamma_{l,m,1} \mu_{l-1,1} \cgcoeff^{l, m_s}_{l-1, m} + \Gamma_{l,m,2} \mu_{l,0} \cgcoeff^{l, m_s}_{l, m} +  \Gamma_{l,m,3} \mu_{l+1,-1} \cgcoeff^{l, m_s}_{l+1, m} \Big] \Big\} \\
+	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: \Big\{ \cgcoeff^{l-1, m_s}_{l, m}  \mu_{l,-1} \Big[ \gamma_{l-1,m-m_s,1} \: Y^{m-m_s}_{l-2} + \gamma_{l-1,m-m_s,2} \: Y^{m-m_s}_{l} \Big] \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad + \cgcoeff^{l+1, m_s}_{l, m}  \mu_{l,1} \Big[ \gamma_{l+1,m-m_s,1} \: Y^{m-m_s}_{l} + \gamma_{l+1,m-m_s,2} \: Y^{m-m_s}_{l+2} \Big] \Big\} \\
+	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: z \: \Big\{ \cgcoeff^{l, m_s}_{l-1, m} \mu_{l,-1} Y^{m-m_s}_{l-1} + \cgcoeff^{l, m_s}_{l+1, m} \mu_{l,1} Y^{m-m_s}_{l+1} \Big\} \\
+	&\quad = z \gradYlm,
+}
+demonstrating that the expression for $z \gradYlm(x,y,z)$ holds.
+
+Firstly, reiterating the previous remark, the ratios of the CG coeffs are equal for each valid $m_s$ value.
+
+Now, we compare coeffs and we are done.
 
@@
 
-\subsection{Function evaluation}
+\subsubsection{Function evaluation}
 The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$:

### 2021-02-11 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
VSH reccurences changed - need to finish

@@
 Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$.
-Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}.
+Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}, orthogonal with respect to the inner product $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$.
 Vector spherical harmonics have been widely used in electrostatics (e.g.
@@
 A simple calculation shows that such orthogonal vectors must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal.
-We will determine the coefficients of these Jacobi operators, as well as other important operators needed -- however first we note a useful property of the VSHs that is a result of the definition used.
-
-\subsection{Jacobi matrices} %{Spin-1 tensor spherical harmonics}
-
-\bstodoinline{cite C-G coeffs, spin-1 tensor harmonics, and the combination expression in \label{eqn:VSHsalt}}
-\bstodoinline{get better reference than that Winter 2017 physics document}
+
+\subsection{Jacobi matrices}
+
+% REMOVED SPIN harmonics and Clebsch Gorden coeff stuff as its much neater to use an inner product argument I have found.
 
@@
 in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$.
-To this end, we use the helpful property that the vector spherical harmonics as defined in \bsrefeqn{eqn:VSHs} can be expressed in terms of \textit{spin-1 tensor spherical harmonics}, which in turn can each be expressed as a vector-weighted sum of spherical harmonics \cite{khersonskii1988quantum}.
-
-We start by defining what we mean by a spin-1 tensor spherical harmonic. In general, the tensor spherical harmonic is given by, for $2l, \, j, \, 2s \in \No$, \cite{}
-\bseqn{
-	\curlyy^{j,s}_{l,m}(x,y,z) = \sum_{m_s=-s}^s \langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle \, Y^{m-m_s}_j (x,y,z) \, \vec{\chi}_{s, m_s},
-}
-where $\vec{\chi}_{s, m_s}$ are the simultaneous eigenstates of the spin operators $\bold{S}^2$ and $S_z$, and where $\langle j \quad m-m_s \; ; \; s \quad m_s \; | \; l \quad m \rangle$ is a Clebsch--Gordan coefficient \cite{} .
-We note that a property of the Clebsch--Gordan coefficients means that they vanish unless $|j - s| \le l \le j + s$ \cite{}.
-We further note that there are simple calculable expressions for the Clebsch--Gordan coefficients when the spin $s=1$ \cite{}.
-
-To simplify the notation, for spin $s=1$, we define
-\bseqnnumber{
-	\cgcoeff^{L, m_s}_{l, m} := \, \langle L \quad m-m_s \; ; \; 1 \quad m_s \; | \; l \quad m \rangle. \label{eqn:cgcoeffdef}
-}
-Thus we have that the three spin-1 tensor spherical harmonics are given by
-\bseqn{
-	\curlyy^{l \pm1,1}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l\pm1, m_s}_{l, m} \, Y^{m-m_s}_{l \pm 1} (x,y,z) \, \chivec, \\
-	\curlyy^{l,1}_{l,m}(x,y,z) &= \sum_{m_s=-1}^1 \cgcoeff^{l, m_s}_{l, m} \, Y^{m-m_s}_l (x,y,z) \, \chivec.
-}
-
-Here, the vectors $\chivec$ are the orthonormal eigenvectors of the matrix
-\bseqn{
-	S_3 = \begin{pmatrix} 0 & -i & 0 \\ i & 0 & 0 \\ 0 & 0 & 0 \end{pmatrix},
-}
-and so are given as
-\bseqn{
-	\vec{\chi}_{1, \pm 1} = \frac{1}{\sqrt{2}} \begin{pmatrix} \mp 1 \\ -i \\ 0 \end{pmatrix}, \quad \vec{\chi}_{1,0} = \begin{pmatrix} 0 \\ 0 \\ 1 \end{pmatrix}.
-}
-Then, for any $l \in \No$, $m \in \Z$ s.t. $0 \le |m| \le l$ we have that
-\bseqnnumber{
-	\gradYlm &= \mu_{l,-1} \, \curlyy^{l-1,1}_{l,m} + \mu_{l,1} \curlyy^{l+1,1}_{l,m}, \quad \gradpYlm = \mu_{l,0} \, \curlyy^{l,1}_{l,m}, \label{eqn:VSHsalt}
+A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations.
+To this end, we recall and derive some important relations of the complex exponential and Jacobi polynomials that are a part of the spherical harmonics definitions.
+In particular we first note that:
+\bseqnnumber{
+	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}) \label{eqn:intcomplexexpcos} \\
+	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}) \label{eqn:intcomplexexpsin}
+}
+and
+\bseqnnumber{
+	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
+	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } d_{l,m} := \half (l + \abs{m} + 1), \label{eqn:diffjac},
+}
+where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:normjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}.
+One final relation we will need is that for incrementing the Jacobi parameters, which we will present in the following Lemma.
+\begin{lemma}
+	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
+\bseqn{
+	P_{n}^{(\alpha, \beta)}(z) =
+		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
+									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
+									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
+									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}.
+}
+\end{lemma}
+\begin{proof}
+	This is simply as consequence of the relationships established for incrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
+\end{proof}
+We thus can present a useful corollary specific for our needs.
+\begin{corollary}\label{corollary:incrementjacparams}
+	Let $m \in \N_0$ and $z \in [-1,1]$. Then, for any $n \in \N_0$:
+\bseqn{
+	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z) \Big],
 }
 where
-\bseqnnumber{
-	\mu_{l,-1} &:= (l+1) \, \Big(\frac{l}{2l+1}\Big)^\half, \quad
-	\mu_{l,0} := i \, \big(l(l+1)\big)^\half, \quad 
-	\mu_{l,1} := l \, \Big(\frac{l+1}{2l+1}\Big)^\half. \label{eqn:spintensorvshcoeffsdef}
-}
-
-The expressions for the VSHs as given in \bsrefeqn{eqn:VSHsalt} allow us to be able to find the coefficients for our desired expressions for $x \gradYlm$, $y \gradYlm$ and $z \gradYlm$. 
+\bseqn{
+	\mu_{l,m,1} &:= {(l+m+\beta+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
+	\mu_{l,m,2} &:= {l \over 2(2l + 1)}.
+}
+\end{corollary}
+
+The expressions for the VSHs as given in \bsrefeqn{eqn:VSHs} together with the above allow us to be able to find the coefficients for our desired expressions for $x \gradYlm$, $y \gradYlm$, $z \gradYlm$, $x \gradpYlm$, $y \gradpYlm$ and $z \gradpYlm$.
 \begin{lemma}\label{lemma:VSHrecurrences}
@@
 }
-%\bseqn{
-%	x \gradYlm &= A_{l,m,1} \gradY^{m-1}_{l-1} + A_{l,m,2} \gradY^{m+1}_{l-1} + A_{l,m,5} \gradY^{m-1}_{l+1} + A_{l,m,6} \gradY^{m+1}_{l+1} \\
-%& \quad \quad \quad + A_{l,m,3}^{*} \gradpY^{m-1}_{l} + A_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
-%	x \gradpYlm &= A_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + A_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + A_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + A_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
-%	& \quad \quad \quad + A_{l,m,3} \gradY^{m+1}_{l} + A_{l,m,4} \gradY^{m-1}_{l}, \\
-%	y \gradYlm &= B_{l,m,1} \gradY^{m-1}_{l-1} + B_{l,m,2} \gradY^{m+1}_{l-1} + B_{l,m,5} \gradY^{m-1}_{l+1} + B_{l,m,6} \gradY^{m+1}_{l+1} \\
-%& \quad \quad \quad + B_{l,m,3}^{*} \gradpY^{m-1}_{l} + B_{l,m,4}^{*} \gradpY^{m+1}_{l}, \\
-%	y \gradpYlm &= B_{l,m,1}^{*} \gradpY^{m-1}_{l-1} + B_{l,m,2}^{*} \gradpY^{m+1}_{l-1} + B_{l,m,5}^{*} \gradpY^{m-1}_{l+1} + B_{l,m,6}^{*} \gradpY^{m+1}_{l+1} \\
-%	& \quad \quad \quad + B_{l,m,3} \gradY^{m+1}_{l} + B_{l,m,4} \gradY^{m-1}_{l}, \\
-%	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l}, \\
-%	z \gradpYlm &= \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l}
-%}
-%where $*$ denotes the complex conjugate, and where, for any valid $m_s$ value,
-where, for any valid $m_s$ value,
-\bseqn{
-	A_{l,m,1} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m-1}} \alpha_{l-1,m-m_s,1}, \\
-	A_{l,m,2} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} \alpha_{l-1,m-m_s,2}, \\
-	A_{l,m,3} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} \alpha_{l+1,m-m_s,3}, \\
-	A_{l,m,4} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} \alpha_{l+1,m-m_s,4}, \\
+where, \bstodoinline{finish typing up (will take a while as is quite fiddly to set out)}
+\bseqn{
+	A_{l,m,1} &:= \clm
+		\begin{cases}
+			\tilde A_{l,m,1} \quad \text{if } m > 1 \\
+			something \quad \text{if } m = 1 \\
+			something2 \quad \text{if } m = 0 \\
+			\tilde A_{l,m,2} \quad \text{if } m < 0
+		\end{cases} \\
+	A_{l,m,2} &:= \clm
+		\begin{cases}
+			\tilde A_{l,m,2} \quad \text{if } m > 0 \\
+			something \quad \text{if } m = 0 \\
+			something2 \quad \text{if } m = -1 \\
+			\tilde A_{l,m,1} \quad \text{if } m < -1
+		\end{cases} \\
+	A_{l,m,3} &:= \clm
+		\begin{cases}
+			\tilde A_{l,m,3} \quad \text{if } m > 1 \\
+			something \quad \text{if } m = 1 \\
+			something2 \quad \text{if } m = 0 \\
+			\tilde A_{l,m,4} \quad \text{if } m < 0
+		\end{cases} \\
+	A_{l,m,4} &:= \clm
+		\begin{cases}
+			\tilde A_{l,m,4} \quad \text{if } m > 0 \\
+			something \quad \text{if } m = 0 \\
+			something2 \quad \text{if } m = -1 \\
+			\tilde A_{l,m,3} \quad \text{if } m < -1
+		\end{cases} \\
 	A_{l,m,5} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,3} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,1} \\
@@
 }
-and where $\cgcoeff^{L,m_s}_{l,m}$, $\mu_{l,k}$ are defined in \bsrefeqn{eqn:cgcoeffdef} and \bsrefeqn{eqn:spintensorvshcoeffsdef} respectively.
 \end{lemma}
 
-\remark It is the case that these coefficients are indeed the same for each $l,m$ pair despite appearing to depend on the value of $m_s$; we need only use any valid $m_s$ value for each expression.
-By \enquote{valid}, we mean the Clebsch--Gordan coefficients do not vanish for that $m_s$ value when used.
+\remark We emphasise again that it is due to the unique relationships that the Jacobi polynomials possess that we are able to explicitly write down these coefficients.
 
 \begin{proof}[Proof of \bsreflemma{lemma:VSHrecurrences}]
-
-\bstodoinline{Do the derivation of these rec coeffs, to have as the proof (and check they are correct!). Need to account for l=1 case.}
 
@@
 \end{proof}
-
-%\begin{align}
-%x \,\gradYlm 
-%&= x \, \blmone \, \curlyy^{l-1,1}_{l,m} + x \, \blone \curlyy^{l+1,1}_{l,m} \nonumber
-%\\
-%&= \sum_{m_s=-1}^{1} \chivec \Big[ \blmone \, \cgcoeff^{l-1,m_s}_{l,m} \big\{ A_{l-1,m-m_s} \, Y^{m-m_s+1}_{l} + B_{l-1,m-m_s} \, Y^{m-m_s+1}_{l-2} \nonumber \\ 
-%& \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + D_{l-1,m-m_s} \, Y^{m-m_s-1}_{l} + E_{l-1,m-m_s} \, Y^{m-m_s-1}_{l-2} \big\} \nonumber \\
-%& \quad \quad \quad \quad \quad \quad \quad + \blone \, \cgcoeff^{l+1,m_s}_{l,m} \big\{ A_{l+1,m-m_s} \, Y^{m-m_s+1}_{l+2} + B_{l+1,m-m_s} \, Y^{m-m_s+1}_{l} \nonumber \\ 
-%& \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + D_{l+1,m-m_s} \, Y^{m-m_s-1}_{l+2} + E_{l+1,m-m_s} \, Y^{m-m_s-1}_{l} \big\} \Big]
-%\end{align}
-%\begin{align}
-%\implies x \,\gradYlm &= \alm \gradY^{m+1}_{l+1} + \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
-%& \quad \quad \quad + \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l},
-%\end{align}
-%where, for any valid $m_s$ value,
-%\begin{align}
-%\alm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} A_{l+1,m-m_s}, \\
-%\blm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} B_{l-1,m-m_s}, \\
-%\dlm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m-1}} D_{l+1,m-m_s}, \\
-%\elm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m-1}} E_{l-1,m-m_s}, \\
-%\hlm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, A_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, B_{l+1,m-m_s} \nonumber \\
-%& \quad \quad \quad \quad \quad \quad \quad - \alm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - \blm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
-%\jlm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, D_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, E_{l+1,m-m_s} \nonumber \\
-%& \quad \quad \quad \quad \quad \quad \quad - \dlm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - \elm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big].
-%\end{align}
-%
-%Note that (it can be shown) these are constants for each $l,m$ pair despite appearing to depend on the value of $m_s$; we need only use any valid $m_s$ value for each expression.
-By ``valid'', we mean the Clebsch--Gordan coefficients do not vanish for that $m_s$ value when used.
-%
-%Similarly, we have that
-%\begin{align}
-%y \,\gradYlm &= i \, \Big[ -\alm \gradY^{m+1}_{l+1} - \blm \gradY^{m+1}_{l-1} + \dlm \gradY^{m-1}_{l+1} + \elm \gradY^{m-1}_{l-1} \nonumber \\
-%& \quad \quad \quad - \hlm \gradpY^{m+1}_{l} + \jlm \gradpY^{m-1}_{l} \Big].
-%\end{align}
-%
-%Further, we also have that
-%\begin{align}
-%z \,\gradYlm 
-%&= z \, \blzero \, \curlyy^{l,1}_{l,m} \nonumber
-%\\
-%&= \sum_{m_s=-1}^{1} \chivec \Big[ \blmone \, \cgcoeff^{l-1,m_s}_{l,m} \big\{ F_{l-1,m-m_s} \, Y^{m-m_s}_{l} + G_{l-1,m-m_s} \, Y^{m-m_s}_{l-2} \big\} \nonumber \\ 
-%& \quad \quad \quad \quad \quad \quad \quad + \blone \, \cgcoeff^{l+1,m_s}_{l,m} \big\{ F_{l+1,m-m_s} \, Y^{m-m_s}_{l+2} + G_{l+1,m-m_s} \, Y^{m-m_s}_{l} \big\} \Big]
-%\end{align}
-%\begin{align}
-%\implies z \,\gradYlm &= \flm \gradY^{m}_{l+1} + \glm \gradY^{m}_{l-1} + \klm \gradpY^{m}_{l},
-%\end{align}
-%where, for any valid $m_s$ value,
-%\begin{align}
-%\flm &:= \frac{\blone}{\beta_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m+1}} F_{l+1,m-m_s}, \\
-%\glm &:= \frac{\blmone}{\beta_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m+1}} G_{l-1,m-m_s}, \\
-%\klm &:= \frac{1}{\blzero \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \blmone \, F_{l-1,m-m_s} + \cgcoeff^{l+1,m_s}_{l,m} \, \blone \, G_{l+1,m-m_s}  \nonumber \\
-%& \quad \quad \quad \quad \quad \quad \quad - \flm \, \beta_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \glm \, \beta_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big].
-%\end{align}
-%
-%We can similarly yield the relations for the perpendicular gradients:
-%\begin{align}
-%x \,\gradpYlm &= \almperp \gradpY^{m+1}_{l+1} + \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
-%& \quad \quad \quad + \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
-%y \,\gradpYlm &= i \, \Big[ -\almperp \gradpY^{m+1}_{l+1} - \blmperp \gradpY^{m+1}_{l-1} + \dlmperp \gradpY^{m-1}_{l+1} + \elmperp \gradpY^{m-1}_{l-1} \nonumber \\
-%& \quad \quad \quad - \hlmperp \gradY^{m+1}_{l} + \jlmperp \gradY^{m-1}_{l}, \\
-%z \,\gradpYlm &= \flmperp \gradpY^{m}_{l+1} + \glmperp \gradpY^{m}_{l-1} + \klmperp \gradY^{m}_{l},
-%\end{align}
-%where it can be shown that
-%\begin{align}
-%\almperp &= \alm^*, \quad \blmperp = \blm^*, \quad \dlmperp = \dlm^*, \quad \elmperp = \elm^*, \nonumber \\
-%\flmperp = \flm^*,& \quad \glmperp = \glm^*, \quad \hlmperp = \hlm^*, \quad \jlmperp = \jlm^*, \quad \klmperp = \klm^*,
-%\end{align}
-%where $^*$ denotes the complex conjugate.
 

### 2021-02-17 papers/thesis/full/abstract/abstract.tex
Necessary files for thesis

\addcontentsline{toc}{chapter}{Abstract}

\begin{abstract}

This thesis develops sparse spectral methods for solving partial differential equations (PDEs) on various multidimensional domains, with a specific focus on the disk-slice and trapezium in 2D, and the spherical cap as a surface in 3D.

We begin with an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface. We present a framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion. Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.

Throughout this thesis we wish to utilise our understanding of the 3D spherical harmonics to build up to a new orthogonal polynomial basis for subdomains of the sphere surface. To this end, we move on working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles. Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials. Triangles and the full disk are then special cases of our framework, which we formalise for the disk-slice and trapezium cases.

Finally, we adapt the tools developed in two dimensions to derive a similar OP basis complete with sparse differential operators.

\end{abstract}

### 2021-02-17 papers/thesis/full/acknowledgements/acknowledgements.tex
Necessary files for thesis

\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

Firstly, I would like to express my thanks to ESPRC for their financial support, and to the Mathematics of Planet Earth Centre for Doctoral Training (MPE CDT) for providing me with the incredible opportunities that the course has brought.

I would like to thank the MPE CDT staff for any knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for his guidance, understanding and general enthusiasm (it meant a lot). And to all my fellow 'Charlies' -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.

I thank my family and friends for all their support and words of encouragement -- it all helped a lot. An extra thanks to all my undergrad mates for many well needed fun times too.

A special thanks goes to my partner Jessica for putting up with me and helping me when things got tough during my 4 years at Imperial College London.

Finally, to Dr. Sheehan Olver -- I cannot thank you enough for all your help, wisdom, advice, kindness, understanding and patience you have given me. I may not have expressed my gratitude enough over the past few years, but you have truly been a great supervisor to me.

I would like to express (whatever feelings I have) to:

\begin{itemize}
 \item My supervisor
 \vspace*{3mm}
 \item My second supervisor
 \vspace*{3mm}
 \item Other researchers
 \vspace*{3mm}
 \item My family and friends
\end{itemize}

\end{acknowledgements}

### 2021-02-17 papers/thesis/full/declaration/declaration.tex
Necessary files for thesis

\addcontentsline{toc}{chapter}{Declaration}

\begin{declaration}
	Declaration here.
\end{declaration}

### 2021-02-17 papers/thesis/full/foreword/foreword.tex
Necessary files for thesis

\addcontentsline{toc}{chapter}{Foreword}

\begin{foreword}

Text of the foreword.

\end{foreword}

### 2021-02-17 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
All files for thesis committed

@@
 \bseqn{
-	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = \chat (1-x^2)^\frac{m}{2} P^{(m,m)}_{l-m}(x) \\
-	P^{-m}_l(x) &:= \ctilde \Plm(x),
+	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; \chat \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x) \\
+	P^{-m}_l(x) &:= (-1)^m \; \ctilde \; \Plm(x),
 }
@@
 \bseqn{
-	\chat &:= {(l+m)! \over (-2)^m \; l!} \\
-	\ctilde &:= {(-1)^m \: (l-m)! \over (l+m)!}
-}
-and $n! := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N$ is the standard factorial. Further, define
-\bseqnnumber{
-	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
-		\begin{cases} 
-			\chat \quad \quad \quad \text{if } m \ge 0 \\
-			\chatmod \ctildemod \quad \text{if } m < 0
-		\end{cases}. \label{eqn:clmdef}
-}
-This allows us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.
-
+	\chat &:= {(l+m)! \over 2^m \; l!} \\
+	\ctilde &:= {(l-m)! \over (l+m)!}
+}
+and $n!
+:= n \: (n-1) \: (n-2) \dots 1$ for $n \in \N_0$ is the standard factorial, with $0!
+= 1$.
+These relationships for the Jacobi polynomials and associated Legendre polynomials allow us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.
+
+On that note, we can now write down the spherical harmonics.
 Let $(x,y,z) \in \Omega$.
 It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$).
-On this note, throughout we will use the convention that the spherical coordinate angles be defined by
+Throughout this document we will use the convention that the spherical coordinate angles be defined by
 \bseqn{
@@
 }
-We can now write down the spherical harmonics. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
+where $\rho(z) := \sqrt{1-z^2}$. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
 \bseqnnumber{
@@
 }
-for $0 \le |m| \le l, \, l \in \No$ where $\clm$ is defined in \bsrefeqn{eqn:clmdef}, and
+for $0 \le |m| \le l, \, l \in \No$, where
+\bseqnnumber{
+	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
+		\begin{cases} 
+			(-1)^m \: \chat &\text{if } m \ge 0 \\
+			\chatmod &\text{if } m < 0
+		\end{cases}. \label{eqn:clmdef}
+} 
+The spherical harmonics as defined are then orthonormal with respect to the inner product
 \bseqn{
@@
 	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
-	& \quad = \delta_{l, l'} \: \delta_{m, m'}.
+	& \quad = \delta_{l, l'} \: \delta_{m, m'}
 }
 where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$.
-Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m\in\mathbb{Z}$.
+Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m \in \Z$.
 Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$.
@@
 \bseqn{
-	z\,\Ylmxyz &= \clm \eimphi \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
-		     &= \clm \eimphi \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
+	z\,\Ylmxyz &= \clm \eimtheta \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
+		     &= \clm \eimtheta \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
 		     &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z).
@@
 \bseqn{
-	\cosphi \, \eimphi &= \frac{1}{2} (e^{i\varphi} + e^{-i\varphi}) \eimphi =  \frac{1}{2} (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \\
-	\sinphi \, \eimphi &= \frac{1}{2i} (e^{i\varphi} - e^{-i\varphi}) \eimphi =  \frac{-i}{2} (e^{i(m+1)\varphi} - e^{i(m-1)\varphi})
-}
-Finally, we can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
+	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
+	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
+}
+We can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an Ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
 \bseqn{
@@
 }
-where the ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
+where the Ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
 \bseqn{
@@
 }
-for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ is then:
-\bseqn{
-	x\,\Ylmxyz &= \clm \cosphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
-	&= \half \clm (e^{i(m+1)\varphi} + e^{i(m-1)\varphi}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
-	&= \half \clm e^{i(m+1)\varphi} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
-	&\quad \quad + \half \clm e^{i(m-1)\varphi} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
+for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
+\bseqn{
+	x \,\Ylmxyz &= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
+	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
+	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
+	&\quad \quad + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
 	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
 	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
-	y\,\Ylmxyz &= \clm \sinphi \: \eimphi \: \sintheta \: \rho(z)^{|m|} P^{(|m|,|m|)}_{l-|m|}(z) \\
-	&= -\half i \: \clm (e^{i(m+1)\varphi} - e^{i(m-1)\varphi}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
-	&= - i \: \big[\Alm Y^{m+1}_{l+1}(x,y,z) +  \Blm Y^{m+1}_{l-1}(x,y,z) \big] \\
-	&\quad \quad \quad + i \, \big[ \Dlm Y^{m-1}_{l+1}(x,y,z) + \Elm Y^{m-1}_{l-1}(x,y,z) \big], \\
+	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
+	&= -\half i \: \clm (e^{i(m+1)\theta} - e^{i(m-1)\theta}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
 	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
@@
 \begin{corollary}\label{corollary:incrementjacparams}
-	Let $m \in \N_0$ and $z \in [-1,1]$. Then, for any $n \in \N_0$:
-\bseqn{
-	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z) \Big],
+	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
+\bseqn{
+	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z),
 }
@@
 \bseqn{
-	\mu_{l,m,1} &:= {(l+m+\beta+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
+	\mu_{l,m,1} &:= {(l+m+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
 	\mu_{l,m,2} &:= {l \over 2(2l + 1)}.
@@
 }
-where, \bstodoinline{finish typing up (will take a while as is quite fiddly to set out)}
-\bseqn{
-	A_{l,m,1} &:= \clm
-		\begin{cases}
-			\tilde A_{l,m,1} \quad \text{if } m > 1 \\
-			something \quad \text{if } m = 1 \\
-			something2 \quad \text{if } m = 0 \\
-			\tilde A_{l,m,2} \quad \text{if } m < 0
-		\end{cases} \\
-	A_{l,m,2} &:= \clm
-		\begin{cases}
-			\tilde A_{l,m,2} \quad \text{if } m > 0 \\
-			something \quad \text{if } m = 0 \\
-			something2 \quad \text{if } m = -1 \\
-			\tilde A_{l,m,1} \quad \text{if } m < -1
-		\end{cases} \\
-	A_{l,m,3} &:= \clm
-		\begin{cases}
-			\tilde A_{l,m,3} \quad \text{if } m > 1 \\
-			something \quad \text{if } m = 1 \\
-			something2 \quad \text{if } m = 0 \\
-			\tilde A_{l,m,4} \quad \text{if } m < 0
-		\end{cases} \\
-	A_{l,m,4} &:= \clm
-		\begin{cases}
-			\tilde A_{l,m,4} \quad \text{if } m > 0 \\
-			something \quad \text{if } m = 0 \\
-			something2 \quad \text{if } m = -1 \\
-			\tilde A_{l,m,3} \quad \text{if } m < -1
-		\end{cases} \\
-	A_{l,m,5} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m-1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,3} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,1} \\
-	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,3} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m-1} - A_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m-1} \Big], \\	
-	A_{l,m,6} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m+1}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \alpha_{l-1,m-m_s,4} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \alpha_{l-1,m-m_s,2} \\
-	& \quad \quad \quad \quad \quad \quad \quad - A_{l,m,4} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m+1} - A_{l,m,2} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,m+1} \Big], \\
+where,
+\bseqn{
+	A_{l,m,1} &:= {\clm c_{l-1}^{m-1} \over 2}
+		\begin{cases}
+			\tilde A_{l,m,1} & \text{if } m > 0 \\
+			\tilde A_{l,\abs{m},2} & \text{if } m \le 0
+		\end{cases} \\
+	A_{l,m,2} &:= {\clm c_{l-1}^{m+1} \over 2}
+		\begin{cases}
+			\tilde A_{l,m,2} & \text{if } m \ge 0 \\
+			\tilde A_{l,\abs{m},1} & \text{if } m < 0
+		\end{cases} \\
+	A_{l,m,3} &:= {\clm c_{l-1}^{m-1} \over 2}
+		\begin{cases}
+			\tilde A_{l,m,3} &\text{if } m > 0 \\
+			\tilde A_{l,\abs{m},4} &\text{if } m \le 0
+		\end{cases} \\
+	A_{l,m,4} &:= {\clm c_{l-1}^{m+1} \over 2}
+		\begin{cases}
+			\tilde A_{l,m,4} &\text{if } m \ge 0 \\
+			\tilde A_{l,\abs{m},3} &\text{if } m < 0
+		\end{cases} \\
+	A_{l,m,5} &:=
+		\begin{cases}
+			- i \: d_{l,m - 1} \: {c_l^{m-1} \over 2 \clm} &\text{if } m > 0 \\
+			(-1)^m \: i \: d_{l,\abs{m}} \: {\clm \over 2 c_l^{m-1}} &\text{if } m \le 0
+		\end{cases} \\
+	A_{l,m,6} &:=
+		\begin{cases}
+			- i \: d_{l,m} \: {\clm \over 2 c_l^{m+1}} &\text{if } m \ge 0 \\
+			(-1)^{m+1} \: i \: d_{l,\abs{m} - 1} \: {c_l^{m+1} \over 2 \clm} &\text{if } m < 0
+		\end{cases} \\
 	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6, \\
-	\Gamma_{l,m,1} &:= \frac{\mu_{l,-1}}{\mu_{l-1,-1}} \frac{\cgcoeff^{l-1,m_s}_{l,m}}{\cgcoeff^{l-2,m_s}_{l-1,m}} \gamma_{l-1,m-m_s,1}, \\
-	\Gamma_{l,m,2} &:= \frac{\mu_{l,1}}{\mu_{l+1,1}} \frac{\cgcoeff^{l+1,m_s}_{l,m}}{\cgcoeff^{l+2,m_s}_{l+1,m}} \gamma_{l+1,m-m_s,2}, \\
-	\Gamma_{l,m,3} &:= \frac{1}{\mu_{l,0} \, \cgcoeff^{l,m_s}_{l, m}} \, \Big[ \cgcoeff^{l-1,m_s}_{l,m} \, \mu_{l,-1} \, \gamma_{l-1,m-m_s,2} + \cgcoeff^{l+1,m_s}_{l,m} \, \mu_{l,1} \, \gamma_{l+1,m-m_s,1} \\
-	& \quad \quad \quad \quad \quad \quad \quad - \Gamma_{l,m,3} \, \mu_{l+1,-1} \, \cgcoeff^{l,m_s}_{l+1,m} - \Gamma_{l,m,1} \, \mu_{l-1,1} \, \cgcoeff^{l,m_s}_{l-1,1} \Big],
+	\Gamma_{l,m,1} &:= \clm c_{l-1}^{m} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \Big] \\
+	\Gamma_{l,m,2} &:= \clm c_{l+1}^{m} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \Big] \\
+	\Gamma_{l,m,3} &:= (-1)^m \: i \: m
+}
+and
+\bseqn{
+	\tilde A_{l,m,1} &:= \Big( (m^2 - 1) \mu_{l-1,m-1,2} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l-1,m-1} {\mu_{l-1,m,2} \over (c_{l}^{m+1})^2} \\
+	\tilde A_{l,m,2} &:= \Big( m(m+2) \mu_{l,m,1} - d_{l,m} \tilde \gamma_{l,m+1,1} \Big) {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\mu_{l,m+1,1} \over (c_{l-1}^{m+2})^2} \\
+	\tilde A_{l,m,3} &:= \Big( (m^2 - 1) \mu_{l+1,m-1,1} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l+1,m-1} {\mu_{l+1,m,1} \over (c_{l}^{m+1})^2} \\
+	\tilde A_{l,m,4} &:= \Big( m(m+2) \mu_{l,m,2} - d_{l,m} \tilde \gamma_{l,m+1,2} \Big) {1 \over (c_{l+1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\mu_{l,m+1,2} \over (c_{l+1}^{m+2})^2}.
 }
@@
 
-Fix $l \in \N, m\in\{-l,\dots,l\}$, and define $Y_{-1}^m := 0$ for any $m \in \Z$.
-We proceed first with the case for multiplication by $z$.
-Using the definitions given for $\Gamma_{l,m,1}, \Gamma_{l,m,2}, \Gamma_{l,m,3}$ in the Lemma, we have that
-\bseqn{
-	&\Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l} \\
-	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: \Big\{ \Gamma_{l,m,1} \mu_{l-1,-1} \cgcoeff^{l-2, m_s}_{l-1, m} Y^{m-m_s}_{l-2} + \Gamma_{l,m,1} \mu_{l-1,1} \cgcoeff^{l, m_s}_{l-1, m} Y^{m-m_s}_{l} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad + \Gamma_{l,m,3} \mu_{l+1,-1} \cgcoeff^{l, m_s}_{l+1, m} Y^{m-m_s}_{l} + \Gamma_{l,m,3} \mu_{l+1,1} \cgcoeff^{l+2, m_s}_{l+1, m} Y^{m-m_s}_{l+2} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad + \Gamma_{l,m,2} \mu_{l,0} \cgcoeff^{l, m_s}_{l, m} Y^{m-m_s}_{l} \Big\} \\
-	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: \Big\{ Y^{m-m_s}_{l-2} \Gamma_{l,m,1} \mu_{l-1,-1} \cgcoeff^{l-2, m_s}_{l-1, m} + Y^{m-m_s}_{l+2} \Gamma_{l,m,3} \mu_{l+1,1} \cgcoeff^{l+2, m_s}_{l+1, m} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad + Y^{m-m_s}_{l} \Big[  \Gamma_{l,m,1} \mu_{l-1,1} \cgcoeff^{l, m_s}_{l-1, m} + \Gamma_{l,m,2} \mu_{l,0} \cgcoeff^{l, m_s}_{l, m} +  \Gamma_{l,m,3} \mu_{l+1,-1} \cgcoeff^{l, m_s}_{l+1, m} \Big] \Big\} \\
-	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: \Big\{ \cgcoeff^{l-1, m_s}_{l, m}  \mu_{l,-1} \Big[ \gamma_{l-1,m-m_s,1} \: Y^{m-m_s}_{l-2} + \gamma_{l-1,m-m_s,2} \: Y^{m-m_s}_{l} \Big] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad + \cgcoeff^{l+1, m_s}_{l, m}  \mu_{l,1} \Big[ \gamma_{l+1,m-m_s,1} \: Y^{m-m_s}_{l} + \gamma_{l+1,m-m_s,2} \: Y^{m-m_s}_{l+2} \Big] \Big\} \\
-	&\quad = \sum_{m_s=-1}^1 \vec{\chi}_{1, m_s} \: z \: \Big\{ \cgcoeff^{l, m_s}_{l-1, m} \mu_{l,-1} Y^{m-m_s}_{l-1} + \cgcoeff^{l, m_s}_{l+1, m} \mu_{l,1} Y^{m-m_s}_{l+1} \Big\} \\
-	&\quad = z \gradYlm,
-}
-demonstrating that the expression for $z \gradYlm(x,y,z)$ holds.
-
-Firstly, reiterating the previous remark, the ratios of the CG coeffs are equal for each valid $m_s$ value.
-
-Now, we compare coeffs and we are done.
+Fix $l \in \N, m\in\{-l,\dots,l\}$.
 

### 2021-02-17 papers/thesis/full/ch3-disk-slices/disk-slices.tex
All files for thesis committed

--- 2020-01-23 papers/disk-slices-and-trapeziums/revised/halfdisk.tex
@@
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
-%\geometry{landscape}                		% Activate for rotated page geometry
-\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
-								% TeX will automatically convert eps --> pdf in pdflatex		
-\usepackage[margin=10pt,font=footnotesize,labelfont=bf]{caption}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{bbm}
-\usepackage{mleftright}
-\usepackage{todonotes}
-
-%\bibliographystyle{apalike}
-
-\graphicspath{ {images/} }
-
-%SetFonts
-
-%SetFonts
-
-\usepackage[square,numbers]{natbib}
-\usepackage{url}
-%\bibliographystyle{elsarticle-harv}
-\bibliographystyle{plain}
-
-\newcommand{\bigO}{\mathcal{O}}
-\newcommand{\half}{\frac{1}{2}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\C}{\mathbb{C}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\No}{\mathbb{N}_0}
-\newcommand{\Ylm}{Y^m_l}
-\newcommand{\Ylmfull}{Y^m_l(\theta,\varphi)}
-\newcommand{\Plm}{\jac^m_l}
-\newcommand{\costheta}{\cos\theta}
-\newcommand{\sintheta}{\sin\theta}
-\newcommand{\cosphi}{\cos\varphi}
-\newcommand{\sinphi}{\sin\varphi}
-\newcommand{\eimphi}{e^{im\varphi}}
-\newcommand{\alphalm}{\alpha^m_l}
-\newcommand{\clm}{c^m_l}
-\newcommand{\ctilde}{\tilde{c}^m_l}
-\newcommand{\ctildemod}{\tilde{c}^{|m|}_l}
-\newcommand{\chat}{\hat{c}^m_l}
-\newcommand{\chatmod}{\hat{c}^{|m|}_l}
-\newcommand{\ddx}{\frac{\mathrm{d}}{\mathrm{d}x}}
-\newcommand{\ddy}{\frac{\mathrm{d}}{\mathrm{d}y}}
-\newcommand{\pddx}{\frac{\partial}{\partial x}}
-\newcommand{\pddy}{\frac{\partial}{\partial y}}
-\newcommand{\pddn}{\frac{\partial}{\partial n}}
-\newcommand{\dmdxm}{\frac{\mathrm{d}^m}{\mathrm{d}x^m}}
-
-\newcommand{\Atilde}{\tilde{A}_{l,m}}
-\newcommand{\Btilde}{\tilde{B}_{l,m}}
-\newcommand{\Dtilde}{\tilde{D}_{l,m}}
-\newcommand{\Etilde}{\tilde{E}_{l,m}}
-\newcommand{\Ftilde}{\tilde{F}_{l,m}}
-\newcommand{\Gtilde}{\tilde{G}_{l,m}}
-\newcommand{\Alm}{A_{l,m}}
-\newcommand{\Blm}{B_{l,m}}
-\newcommand{\Dlm}{D_{l,m}}
-\newcommand{\Elm}{E_{l,m}}
-\newcommand{\Flm}{F_{l,m}}
-\newcommand{\Glm}{G_{l,m}}
-
-\newcommand{\xione}{\xi^{(1)}_{n, \lambda}}
-\newcommand{\xitwo}{\xi^{(2)}_{n, \lambda}}
-\newcommand{\xithree}{\xi^{(3)}_{n, \lambda}}
-\newcommand{\xifour}{\xi^{(4)}_{n, \lambda}}
-
-\newcommand{\bigP}{\mathbb{P}}
-\newcommand{\Pl}{\mathbb{P}_l}
-\newcommand{\gradP}{T\mathbb{P}}
-\newcommand{\gradPl}{T\mathbb{P}_l}
-\newcommand{\gradY}{\nabla Y}
-\newcommand{\gradYlm}{\nabla Y^m_l}
-\newcommand{\gradpY}{\nabla^\perp Y}
-\newcommand{\gradpYlm}{\nabla^\perp Y^m_l}
-
-\newcommand{\Dlt}{D^\top_l}
-
-\newcommand{\curlyy}{\bm{\mathcal{Y}}}
-\newcommand{\blone}{\beta_{l, 1}}
-\newcommand{\blzero}{\beta_{l, 0}}
-\newcommand{\blmone}{\beta_{l, -1}}
-\newcommand{\chivec}{\bm{\chi}_{1,m_s}}
-\newcommand{\cgcoeff}{\mathcal{C}}
-
-\newcommand{\alm}{a_{l,m}}
-\newcommand{\blm}{b_{l,m}}
-\newcommand{\dlm}{d_{l,m}}
-\newcommand{\elm}{e_{l,m}}
-\newcommand{\flm}{f_{l,m}}
-\newcommand{\glm}{g_{l,m}}
-\newcommand{\hlm}{h_{l,m}}
-\newcommand{\jlm}{j_{l,m}}
-\newcommand{\klm}{k_{l,m}}
-\newcommand{\almperp}{a_{l,m}^\perp}
-\newcommand{\blmperp}{b_{l,m}^\perp}
-\newcommand{\dlmperp}{d_{l,m}^\perp}
-\newcommand{\elmperp}{e_{l,m}^\perp}
-\newcommand{\flmperp}{f_{l,m}^\perp}
-\newcommand{\glmperp}{g_{l,m}^\perp}
-\newcommand{\hlmperp}{h_{l,m}^\perp}
-\newcommand{\jlmperp}{j_{l,m}^\perp}
-\newcommand{\klmperp}{k_{l,m}^\perp}
-
-\newcommand{\unitvec}{\hat{\bm{k}}}
-
-\newcommand{\hdop}{H}
-\newcommand{\bighdop}{\mathbb{\hdop}}
-\newcommand{\hdopnk}{\hdop_{n,k}}
-\newcommand{\hdopnkab}{\hdop_{n,k}^{(a,b)}}
-\newcommand{\hdopab}{\hdop^{(a,b)}}
-\newcommand{\Wab}{{W^{(a,b)}}}
-\newcommand{\hdopmj}{\hdop_{m,j}}
-\newcommand{\hdopmjab}{\hdop_{m,j}^{(a,b)}}
-\newcommand{\alphaab}{\alpha^{(a,b)}}
-\newcommand{\betaab}{\beta^{(a,b)}}
-\newcommand{\bighdopab}{\bighdop^{(a,b)}}
-\newcommand{\Dnt}{D^\top_n}
-\newcommand{\Wii}{W^{(1,1)}}
-\newcommand{\hdopii}{\hdop^{(1,1)}}
-\newcommand{\bighdopii}{{\mathbb{\hdop}^{(1,1)}}}
-\newcommand{\hdopoo}{\hdop^{(0,0)}}
-\newcommand{\bighdopoo}{{\mathbb{\hdop}^{(0,0)}}}
-\newcommand{\hdopoi}{\hdop^{(0,1)}}
-\newcommand{\hdopio}{\hdop^{(1,0)}}
-%\newcommand{\jac}{{\tilde P}}
-\newcommand{\jac}{{P}}
-\newcommand{\genjac}{R}
-\newcommand{\genjacnmk}{\genjac_{n-k}}
-\newcommand{\genjacmmj}{\genjac_{m-j}}
-\newcommand{\genjacw}{w_\genjac}
-\newcommand{\jacw}{w_P}
-\newcommand{\normgenjac}{\omega_\genjac}
-\newcommand{\normjac}{\omega_P}
-
-\newcommand{\dx}{\frac{\partial}{\partial x}}
-\newcommand{\dy}{\frac{\partial}{\partial y}}
-\newcommand{\laplacewii}{\Delta_W^{(1,1)\to(1,1)}}
-\newcommand{\laplacewtt}{\Delta_W^{(2,2)\to(0,0)}}
-\newcommand{\laplaceoo}{\Delta^{(0,0)\to(2,2)}}
-\newcommand{\biharmonic}{_2\Delta_W^{(2,2)\to(2,2)}}
-
-\newcommand{\element}{\tau}
-\newcommand{\refelement}{\hat{\tau}}
-\newcommand{\FEset}{\mathcal{T}}
-\newcommand{\bigW}{\mathbb{W}}
-\newcommand{\bigWab}{\mathbb{W}^{(a,b)}}
-\newcommand{\bigWii}{{\mathbb{W}^{(1,1)}}}
-\newcommand{\bigQ}{\mathbb{Q}}
-\newcommand{\bigQab}{\bigQ^{(a,b)}}
-
-\newcommand{\hdopnkabc}{\hdop_{n,k}^{(a,b,c)}}
-\newcommand{\hdopabc}{\hdop^{(a,b,c)}}
-\newcommand{\Wabc}{{W^{(a,b,c)}}}
-\newcommand{\hdopmjabc}{\hdop_{m,j}^{(a,b,c)}}
-\newcommand{\alphaabc}{\alpha^{(a,b,c)}}
-\newcommand{\betaabc}{\beta^{(a,b,c)}}
-\newcommand{\bighdopabc}{\bighdop^{(a,b,c)}}
-\newcommand{\Wiii}{W^{(1,1,1)}}
-\newcommand{\hdopiii}{\hdop^{(1,1,1)}}
-\newcommand{\bighdopiii}{{\mathbb{\hdop}^{(1,1,1)}}}
-\newcommand{\hdopooo}{\hdop^{(0,0,0)}}
-\newcommand{\bighdopooo}{{\mathbb{\hdop}^{(0,0,0)}}}
-\newcommand{\hdopooi}{\hdop^{(0,0,1)}}
-\newcommand{\hdopiio}{\hdop^{(1,1,0)}}
-\newcommand{\laplacewiii}{\Delta_W^{(1,1,1)\to(1,1,1)}}
-\newcommand{\laplacewttt}{\Delta_W^{(2,2,2)\to(0,0,0)}}
-\newcommand{\laplaceooo}{\Delta^{(0,0,0)\to(2,2,2)}}
-\newcommand{\biharmonictwo}{_2\Delta_W^{(2,2,2)\to(2,2,2)}}
-\newcommand{\bigWiii}{{\mathbb{W}^{(1,1,1)}}}
-\newcommand{\bigWabc}{\mathbb{W}^{(a,b,c)}}
-
-\newcommand{\alphaabcd}{\alpha^{(a,b,c,d)}}
-\newcommand{\betaabcd}{\beta^{(a,b,c,d)}}
-\newcommand{\hdopnkabcd}{\hdop_{n,k}^{(a,b,c,d)}}
-\newcommand{\hdopabcd}{\hdop^{(a,b,c,d)}}
-\newcommand{\Wabcd}{{W^{(a,b,c,d)}}}
-\newcommand{\hdopmjabcd}{\hdop_{m,j}^{(a,b,c,d)}}
-\newcommand{\bighdopabcd}{\bighdop^{(a,b,c,d)}}
-\newcommand{\bigWabcd}{\mathbb{W}^{(a,b,c,d)}}
-
-\newcommand{\genjact}{\tilde{\genjac}}
-\newcommand{\genjactnmk}{\genjact_{n-k}}
-\newcommand{\genjactmmj}{\genjact_{m-j}}
-\newcommand{\genjactw}{w_{\genjact}}
-\newcommand{\normgenjact}{\omega_{\genjact}}
-
-\usepackage{amsthm}
-
-\newtheorem{proposition}{Proposition}
-\newtheorem{lemma}{Lemma} 
-\newtheorem{theorem}{Theorem} 
-\newtheorem{definition}{Definition}
-\newtheorem{condition}{Condition}
-\newtheorem{algorithm}{Algorithm}
-\newtheorem{corollary}{Corollary}
-
-\def\condref#1{Condition~\ref{cond:#1}}
-
-\include{somacros}
-
-%\usepackage{media9}
-%\graphicspath{ {../sphere/plots} }
-
-\title{Sparse spectral and $p$-finite element methods for partial differential equations on disk slices and trapeziums}
-\author{Sheehan Olver, Ben Snowball}
-%\date{}							% Activate to display a given date or no date
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-Sparse spectral methods for solving partial differential equations have been derived in recent years using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
-In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices and trapeziums.
-This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
-We apply the framework to solving the Poisson, variable coefficient Helmholtz, and Biharmonic equations.
-In this paper we focus on constant Dirichlet boundary conditions, as well as zero Dirichlet and Neumann boundary conditions, with other types of boundary conditions requiring future work.
- \end{abstract}
-
-%
-\section{Introduction}
-
-This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and trapeziums.  
-More precisely, we consider the solution of partial differential equations on the domain
-\begin{align*}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
+\chapter{Disk slices and trapeziums}
+
+The contents of the previous chapter on spherical harmonics can allow us to use similar techniques for developing sparse spectral methods for solving linear partial differential equations on other domains.
+Before we can return to the 3D surface realm, let's first achieve a foundation in two-dimensions, by looking at a special class of geometries that includes disk slices and trapeziums.
+
+More precisely, we will consider the solution of partial differential equations on the domain
+\bseqn{
+	\Omega := \bsset{(x,y) \in \R^2}{\alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)} 
+}
 where  either of the following conditions hold:
@@
 \begin{condition}\label{cond:trap}
- \(\rho\) is a degree 1 polynomial.
+	$\rho$ is a degree 1 polynomial.
 \end{condition}
-
 \begin{condition}\label{cond:disk}
-\(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
+	$\rho$ is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
 \end{condition}
 
-For simplicity of presentation we focus on the disk-slice, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$, and discuss an extension to other geometries in the appendix (including the half-disk and trapeziums).
+For simplicity of presentation we focus on the disk-slice in this chapter, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$.
+However, we will discuss an extension to other geometries, including the half-disk and trapeziums.
 
@@
 On the disk-slice the family of weights we consider are of the form
-$$
+\bseqn{
 	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \qfor \alpha \leq x \leq \beta,\quad -\rho(x) \leq y \leq \rho(x).
-$$
+}
 The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
-$$
+\bseqn{
 	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
-$$
+}
 where $C_{n,k} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
@@
 
-Here is an overview of the paper:
-
-\noindent \secref{OPs}: We present our procedure to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the disk-slice domain, by combining 1D OPs on the interval, to form 2D OPs on the disk.
-
-\noindent\secref{PDOs}: We demonstrate that these families will lead to sparse operators, including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators.
-We present a method involving use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} to obtain the recurrence coefficients for the non-classical 1D OPs, allowing us to exactly use the 1D quadrature rules we present to calculate the non-zero entries to the sparse operators.
-
-\noindent\secref{Computation}: We discuss computational issues, in particular, how to realise the results of the preceding sections in practice.
-We present a method for explicitly deriving the recurrence coefficients of the non-classical 1D OPs we detail in \secref{OPs}.
-We derive a quadrature rule on the disk-slice that can be used to expand a function in the OP basis up to a given order.
-Further, we implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm.
-
-\noindent\secref{Examples}: We demonstrate the proposed technique for solving Poisson, Helmholtz, and Biharmonic equations on the disk-slice.
-
-\noindent\appref{PFEM}: We use the procedure to construct sparse $p$-finite element methods.
-This lays the groundwork for a future $hp$-finite element method in a disk, where the elements capture the circular geometry precisely.
-
-\noindent\appref{HalfDisk}: We discuss extension to the special case of end-disk-slices (e.g., half disks).
-
-\noindent\appref{Trapezium}: We discuss extension to trapezia.
-
 \section{Orthogonal polynomials on the disk-slice and the trapezium}\label{Section:OPs}
 
-In this section we outline the construction and some basic properties of $\hdopnkabcd(x,y)$.
+We can mirror the approach for the spherical harmonics, in that we will outline the construction and some basic properties of our 2D OPs on $\Omega$ that we denote by $\hdopnkabcd(x,y)$.
 The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.
@@
 \begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
-Let \(w_1 : (\alpha,\beta) \: \to \R\), \(w_2 : (\gamma,\delta) \: \to \R\) be weight functions with \(\alpha,\beta,\gamma,\delta \in \R\), and let \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either \condref{trap} or \condref{disk} with \(w_2\) being an even function hold.
+	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \condref{trap} or \condref{disk} with $w_2$ being an even function hold.
 $\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$.
 Then the 2D polynomials defined on $\Omega$
-$$
+\bseqn{
 	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
-$$
+}
 are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)}) \) on $\Omega$. 
@@
 
-For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definition. First we introduce notation for two families of univariate OPs.
+For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definitions. First we introduce notation for two families of univariate OPs.
 \begin{definition}\label{def:OPconstruction}
 	Let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
-\begin{align*}
+\bseqn{
 	\begin{cases}
@@
 	\end{cases}
-\end{align*}
+}
 and define the associated inner products by:
-\begin{align}
+\bseqnnumber{
 	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ipgenjac} \\
 	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ipjac}
-\end{align}
+}
 where
-\begin{align}
+\bseqnnumber{
 	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
-\end{align}
+}
 Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
@@
 
+We can now write down our 2D OP family.
 \begin{definition}\label{def:constuction}
 	Define the four-parameter 2D orthogonal polynomials via:
-\begin{align*}
+\bseqn{
 	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align*}
+}
 \end{definition}
 
+The comparison with the construction of the 3D spherical harmonics is evident here (the $\genjac$ OPs are in place of the Jacobi polynomials, and the $\jac$ OPs are in place of the complex exponentials, which are just Chebyshev polynomials).
+By defining our OPs in this way, we will be able to yield sparse and banded relations for the necessary operators involved in PDEs, just as we did before.
+
 $\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
-\begin{align*}
+\bseqn{
 	W^{(a,b,c,d)}(x,y) := \genjacw^{(a,b,c+d)}(x) \: \jacw^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega,
-\end{align*}
+}
 assuming that either \condref{trap} or \condref{disk} with  \(\jacw^{(a,b)}\) being an even function (i.e.
@@
 That is,
-\begin{align*}
+\bseqn{
 	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
-\end{align*}
+}
 where for $f, g : \Omega \to \R$ the inner product is defined as 
-\begin{align*}
+\bseqn{
 	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
-\end{align*}
+}
 
 We can see that they are indeed orthogonal using the change of variable $t = \frac{y}{\rho(x)}$, for the following normalisation:
-\begin{align}
+\bseqnnumber{
 	&\ip< \hdopnkabcd, \: \hdopmjabcd >_\Wabcd \\
@@
 	&\quad \quad \quad = \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}. \label{eqn:normhdop}
-\end{align}
+}
 
 For the disk-slice, the weight $\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c$ results from setting:
-\begin{align*}
+\bseqn{
 \begin{cases}
@@
 \end{cases}
-\end{align*}
+}
 so that
-\begin{align*}
+\bseqn{
 \begin{cases}
@@
 \end{cases}
-\end{align*}
+}
 Note here we can simply remove the need for including a fourth parameter $d$. The 2D OPs orthogonal with respect to the weight above on the disk-slice $\Omega$ are then given by:
-\begin{align}\label{eq:diskpolys}
+\bseqnnumber{\label{eq:diskpolys}
 	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
-\end{align}
+}
 In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
@@
 
-We can express the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$ as 
-\begin{align}
+Recall that Jacobi matrices are an important piece of the puzzle for using these orthogonal polynomials to practically solve PDEs in a sparse spectral method.
+To obtain the entries for them, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$.
+These can be expressed as:
+\bseqnnumber{
 	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:Hrecurrence} \\
 	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:Precurrence}
-\end{align}
-Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0 \quad \forall n=0,1,2,\dots$.
+}
+Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0$ $\forall n=0,1,2,\dots$.
 We can use (\ref{eqn:Hrecurrence}) and (\ref{eqn:Precurrence}) to determine the 2D recurrences for $\hdopnkabcd(x,y)$.
@@
 	$\hdopnkabcd(x,y)$ satisfy the following 3-term recurrences:
-\begin{align*}
+\bseqn{
 x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
@@
 		& \quad \quad + \betaabcd_{n,k,7} \: \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \: \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,9} \: \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
-\end{align*}
-for \((x,y) \in \Omega\), where
-\begin{align*}
-	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b+k+\half)} \\
+}
+for $(x,y) \in \Omega$, where
+\bseqn{
+	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)} \\
 	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
@@
 	\betaabcd_{n,k,9} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjacnmk^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}. 
-\end{align*}
+}
 
@@
 These coefficients are given by
-\begin{align*}
+\bseqn{
 	c_{m,j} = {\ip< y \: \hdopnkabcd, \hdopmjabcd >_{\Wabcd}}{\norm{\hdopmjabcd}^{-2}_{\Wabcd}}.
-\end{align*}
+}
 Recall from equation (\ref{eqn:normhdop}) that $\norm{\hdopmjabcd}_{\Wabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$.
 Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
-\begin{align*}
+\bseqn{
 	&\ip<y \hdopnkabcd, \hdopmjabcd>_\Wabcd \\
@@
 	&\quad \quad \quad \quad \quad\cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big)
-\end{align*}
-\begin{align*}
+}
+\bseqn{
 	=
@@
       	\end{cases}
-\end{align*}
+}
 where, by orthogonality,
-\begin{align*}
+\bseqn{
 	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} &= 0 \qfor m < n-1, \\
 	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} &= 0 \qfor m < n-1.
-\end{align*}
+}
 
@@
 Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
-\begin{align*}
+\bseqn{
 \bighdopabcd_n := \begin{pmatrix}
@@
 	\end{pmatrix}
-\end{align*}
+}
 and set $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ as the Jacobi matrices corresponding to
-\begin{align}
+\bseqnnumber{
 J_x^{(a,b,c,d)} \: \bighdopabcd(x,y) = x \: \bighdopabcd(x,y), \quad J_y^{(a,b,c,d)} \: \bighdopabcd(x,y) = y \: \bighdopabcd(x,y).
 \label{eqn:jacobimatricesdefinition}
-\end{align}
+}
 The matrices $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ act on the coefficients vector of a function's expansion in the $\{\hdopnkabcd\}$ basis.
@@
 In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkabcd\}$ basis.
-Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices:
-
-\begin{definition}
-A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
-A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
-\end{definition}
-
-For example, $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are block-tridiagonal (block-bandwidths $(1,1)$):
-\begin{align*}
+
+Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
+\bseqn{
 J_{x/y}^{(a,b,c,d)} &= \begin{pmatrix}
@@
 	\end{pmatrix}
-\end{align*}
+}
 where the blocks themselves are diagonal for $J_x^{(a,b,c,d)}$ (sub-block-bandwidths $(0,0)$),
-\begin{align*}
+\bseqn{
 A^x_n &:= \begin{pmatrix}
@@
 \nonumber
-\end{align*}
+}
 and tridiagonal for $J_y^{(a,b,c,d)}$ (sub-block-bandwidths $(1,1)$),
-\begin{align*}
+\bseqn{
 A^y_n &:= \begin{pmatrix}
@@
 	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
-\end{align*}
+}
 
@@
 We can combine each system in (\ref{eqn:jacobimatricesdefinition}) into a block-tridiagonal system:
-\begin{align*}
+\bseqn{
 \renewcommand\arraystretch{1.3}
@@
 \end{pmatrix},
-\end{align*}
+}
 where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
-\begin{align*}
+\bseqn{
 A_n &:= \begin{pmatrix}
@@
 	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
-\end{align*}
+}
 
@@
 Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
-\begin{align*}
+\bseqn{
 \begin{cases}
@@
 \end{cases}
-\end{align*}
+}
 
 Note that we can define an explicit \(\Dnt\) as follows:
-\begin{align*}
+\bseqn{
 \Dnt := \begin{pmatrix}
@@
 	    \end{pmatrix},
-\end{align*}
+}
  where
- \begin{align*}
+ \bseqn{
 	\eta_{n+1} &= \frac{1}{\betaabcd_{n,n,9}}, \\ 
 	\eta_{n} &= -\frac{1}{\betaabcd_{n,n-1,9}} \big( \betaabcd_{n,n,8} \: \eta_{n+1} \big), \\
-	\eta_j &= \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
-	\eta_0 &= \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
-\end{align*}
+	\eta_j &= - \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
+	\eta_0 &= - \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
+}
 It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopabcd_{0}(x,y)$  through $\bighdopabcd_{n}(x,y)$ in optimal $O(n^2)$ complexity.
@@
 }
-\label{fig:Laplace} 
+\label{fig:laplaceparameterconversion} 
 \end{figure}
@@
 In this section, we concentrate on the disk-slice case, and simply note that similar arguments apply for the trapezium case. Recall that, for the disk-slice,
-\begin{align*}
+\bseqn{
 	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
+}
 where
-\begin{align*}
+\bseqn{
 \begin{cases}
@@
 \end{cases}.
-\end{align*}
+}
 The 2D OPs on the disk-slice $\Omega$, orthogonal with respect to the weight
-\begin{align*}
+\bseqn{
 	\Wabc(x,y) &:= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
 	&= (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \quad (x,y) \in \Omega,
-\end{align*}
+}
 are then given by:
-\begin{align*}
+\bseqn{
 	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
-\end{align*}
+}
 where the 1D OPs $\{\genjac^{(a, b, c)}_n\}$ are orthogonal on the interval $(\alpha, \beta)$ with respect to the weight
@@
 Denote the weighted OPs by
-\begin{align*}
+\bseqn{
 \bigWabc(x,y) := \Wabc(x,y) \: \bighdopabc(x,y),
-\end{align*}
+}
 and recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopabc(x,y)^\top \mathbf{f}$.
@@
 Define the operator matrices $D_x^{(a,b,c)}, \: D_y^{(a,b,c)}, \: W_x^{(a,b,c)}, \: W_y^{(a,b,c)}$ according to:
-\begin{align*}
+\bseqn{
 {\partial f \over \partial x}&= \bighdop^{(a+1,b+1,c+1)}(x,y)^\top \: D_x^{(a,b,c)} \: \mathbf{f}, \\
@@
 {\partial \over \partial y}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a,b,c-1)}(x,y)^\top \: W_y^{(a,b,c)} \: \mathbf{f}.
-\end{align*}
+}
 \end{definition}
@@
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
+An illustration of how the non-weighted differential operators increment the parameters (a,b,c) is seen in Figure \ref{fig:laplaceparameterconversion}.
 
@@
 First, note that:
-\begin{align}
+\bseqnnumber{
 	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
@@
 	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
-\end{align}
+}
 
 We proceed with the case for the operator $D_y^{(a,b,c)}$ for partial differentiation by $y$.
-Since $\{\hdop^{(a,b,c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddy \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^y \: \hdop^{(a, b, c+1)}_{m,j}$.
+Since $\{\hdop^{(a,b,c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\ppx{y} \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^y \: \hdop^{(a, b, c+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 First, note that
-\begin{align*}
-	c_{m,j}^y = {\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
-\end{align*}
+\bseqn{
+	c_{m,j}^y = {\ip<\ppx{y} \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
+}
 Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
-\begin{align*}
-	&\ip<\pddy \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
+\bseqn{
+	&\ip<\ppx{y} \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
 	&= \iint_\Omega \: \Big[ \genjacnmk^{(a,b,2c+2k+1)}(x) \: \rho(x)^{k-1} \: \jac_k^{(c,c) \: \prime}\fpr(\frac{y}{\rho(x)}) \\
@@
 	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}
-\end{align*}
+}
 Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
-\begin{align*}
+\bseqn{
 	\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} &= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
-	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddy [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
+	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddx{y} [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
 	&= - \int_{-1}^1 \: \jac_k^{(c,c)} \: [ \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: y \: \jac_{j}^{(c+1,c+1)} \:\jacw^{(c)} ] \: \D y \\
 	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}}
-\end{align*}
+}
 which is zero for $j < k-1$ by orthogonality. Further, when $j = k-1$, we have that
-\begin{align*}
+\bseqn{
 	&\normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
@@
 	&= \normgenjac^{(a, b, 2c+2k+1)} \: \delta_{n,m+1},
-\end{align*}
+}
 showing that the only possible non-zero coefficient is when $m=n-1, j=k-1$. Finally,
-\begin{align*}
+\bseqn{
 	c_{n-1,k-1}^y &= \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{k-1}^{(c+1,c+1)} >_{\jacw^{(c+1)}}.
-\end{align*}
+}
 
 We next consider the case for the operator $D_x^{(a,b,c)}$ for partial differentiation by $x$.
-Since $\{\hdop^{(a+1, b+1, c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\pddx \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^x \: \hdop^{(a+1, b+1, c+1)}_{m,j}$.
+Since $\{\hdop^{(a+1, b+1, c+1)}_{m,j}\}$ for $m = 0,\dots,n-1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n-1$ polynomial, we can expand $\ppx{x} \hdopnkabc = \sum_{m=0}^{n-1} \sum_{j=0}^m c_{m,j}^x \: \hdop^{(a+1, b+1, c+1)}_{m,j}$.
 The coefficients of the expansion are then the entries of the relevant operator matrix.
@@
 First, note that
-\begin{align*}
-	c_{m,j}^x = {\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
-\end{align*}
+\bseqn{
+	c_{m,j}^x = {\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
+}
 Now, again using the change of variable $t= \frac{y}{\rho(x)}$, we have that
-\begin{align}
-	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
+\bseqnnumber{
+	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
 	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
@@
 	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:dxsparsity}
-\end{align}
+}
 We will first show that the second factor of each term in (\ref{eqn:dxsparsity}) are zero for $j < k - 2$ and also for $j = k - 1$.
@@
 Hence
-\begin{align*}
+\bseqn{
 	\int_\gamma^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t = \int_{-\delta}^\delta \jac_k^{(c,c)} \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} \: \jacw^{(c)} \: \D t 
-\end{align*}
+}
 is zero for $j < k - 2$ by orthogonality, and is zero for $j = k-1$ due to symmetry over the domain.
 Moreover, $t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t)$ is also an odd polynomial for any $k$ and so
-\begin{align*}
+\bseqn{
 	\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime}(t) \: \jac_j^{(c+1,c+1)}(t) \: \jacw^{(c+1)}(t) \: \D t
-\end{align*}
+}
 is zero for $j = k-1$ due to symmetry over the domain, and
-\begin{align*}
+\bseqn{
 	&\int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \\
@@
 	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}}
-\end{align*}
+}
 which is zero for $j < k - 2$ by orthogonality. Thus, (\ref{eqn:dxsparsity}) is zero for $j \notin \{k-2, k\}$.
@@
 Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
-\begin{align}
+\bseqnnumber{
 	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \nonumber  \\
 	&= \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \: \D x \nonumber \\
-	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \nonumber \\
+	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \ddx{x} \Big[ \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \Big] \: \D x \nonumber \\
 	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \nonumber \\
@@
 		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:dxsparsityR}
-\end{align}
+}
 By recalling (\ref{eqn:rhoderivative}) and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
@@
 Finally,
-\begin{align*}
+\bseqn{
 	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \\
 	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}}
-\end{align*}
+}
 which is also zero for $m < n - 3$. Thus
-\begin{align*}
-	&\ip<\pddx \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
-\end{align*}
-showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,\dots,n$ and $j = k-2,k$ $(j \le m)$.
+\bseqn{
+	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
+}
+showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,n-2,n-1$ and $j = k-2,k$ $(j \le m)$.
 
 We can gain the non-zero entries of the weighted differential operators similarly, by noting that for the disk-slice
-\begin{align}
-	\pddx \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
-	\pddy \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
-\end{align}
+\bseqnnumber{
+	\ppx{x} \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
+	\ppx{y} \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
+}
 and also that
-\begin{align*}
+\bseqn{
 	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{W^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wabc.
-\end{align*}
+}
 
@@
 for conversion between weighted spaces, according to:
-\begin{align*}
+\bseqn{
 	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c)} \Big)^\top \: \bighdop^{(a+1,b+1,c)}(x,y) \\
@@
 	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a-1,b-1,c-1)} \Big)^\top \: \bigW^{(a-1,b-1,c-1)}(x,y).
-\end{align*}
+}
 \end{definition}
@@
 First, note that
-\begin{align*}
+\bseqn{
 	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
-\end{align*}
+}
 Then, using the change of variable $t = \frac{y}{\rho(x)}$, we have that
-\begin{align*}
+\bseqn{
 	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
@@
 	&\quad \quad \quad \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
-\end{align*}
+}
 Since $\jacw^{(\tilde{c})}$ is a polynomial degree $2\tilde{c}$, we have that the above is then zero for $j < k-2\tilde{c}$.
@@
 As a canonical example, we can obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
-$$
+\bseqn{
 	\bigWiii(x,y) = \Wiii(x,y) \: \bighdopiii(x,y)
-$$
+}
 to coefficients in the non-weighted space $\bighdopiii(x,y)$.
@@
 The matrix operator for the Laplacian we denote $\laplacewiii$ acting on the coefficients vector is then given by
-\begin{align*}
+\bseqn{
     \laplacewiii := D_x^{(0,0,0)} \: W_x^{(1,1,1)} + T^{(0,0,1)\to(1,1)} \: D_y^{(0,0,0)} \: T_W^{(1,1,0)\to(0,0,0)} \: W_y^{(1,1,1)}.
-\end{align*}
+}
 Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
@@
 Another important example is the Biharmonic operator $\Delta^2$, where we assume zero Dirichlet and Neumann conditions.
-To construct this operator, we first note that we can obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopooo(x,y)$ to coefficients in the space $\bighdop^{(2,2)}(x,y)$.
+To construct this operator, we first note that we can obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopooo(x,y)$ to coefficients in the space $\bighdop^{(2,2,2)}(x,y)$.
 We denote this matrix operator that acts on the coefficients vector as $\laplaceooo$, and is given by
-\begin{align*}
+\bseqn{
     \laplaceooo := D_x^{(1,1,1)} \: D_x^{(0,0,0)} + T^{(1,1,2)\to(2,2,2)} \: D_y^{(1,1,1)} \: T^{(0,0,1)\to(1,1,1)} \: D_y^{(0,0,0)}.
-\end{align*}
+}
 Further, we can represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopooo$.
@@
 We denote this matrix operator as $\laplacewttt$, and is given by
-\begin{align*}
+\bseqn{
 	\laplacewttt := W_x^{(1,1,1)} \: W_x^{(2,2,2)} + T_W^{(1,1,0)\to(0,0,0)} \: W_y^{(1,1,1)} \: T_W^{(2,2,1)\to(1,1,1)} \: W_y^{(2,2,2)}.
-\end{align*}
+}
 We can then construct a matrix operator for $\Delta^2$ that will take coefficients in the space $\bigW^{(2,2,2)}$ to coefficients in the space $\bighdop^{(2,2,2)}$.
@@
 The matrix operator for the Biharmonic operator is then given by
-\begin{align*}
+\bseqn{
 	\biharmonictwo = \laplaceooo \: \laplacewttt.
-\end{align*}
+}
 The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
@@
 Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
-\begin{align}
+\bseqnnumber{
 	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ipgenjact}
-\end{align}
+}
 where
-\begin{align}
+\bseqnnumber{
 	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
-\end{align}
+}
 Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjact}).
@@
 Note that the OPs $\{\genjac_n^{(a,b,2c)}\}$ are then equivalent to the OPs $\{\genjact_n^{(a,b,c,c)}\}$. Let the recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d)}\}$ be given by:
-\begin{align}
+\bseqnnumber{
 	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
-\end{align}
+}
 \begin{proposition}
 There exist constants $\mathcal{C}_n^{(a,b,c,d)}$, $\mathcal{D}_n^{(a,b,c,d)}$ such that
-\begin{align}
+\bseqnnumber{
 	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
 	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
-\end{align}
+}
 \end{proposition}
@@
 Fix $n,m \in \{0,1,\dots\}$ and without loss of generality, assume $m \le n$. First recall that
-\begin{align*}
+\bseqn{
 	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
-\end{align*}
+}
 and define
-\begin{align}
+\bseqnnumber{
 	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
@@
 \label{eqn:genjactnormalisationii}
-\end{align}
+}
 Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
-\begin{align}
+\bseqnnumber{
 	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}.
 \label{eqn:christoffeldarboux}
-\end{align}
+}
 Then,
-\begin{align*}
+\bseqn{
 	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
@@
 	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
-\end{align*}
+}
 using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
-\begin{align*}
+\bseqn{
 	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
@@
 	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
-\end{align*}
+}
 using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
@@
 The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c+1,d)}\}$ are given by:
-\begin{align}
+\bseqnnumber{
 	\tilde{\alpha}_n^{(a,b,c+1,d)} &= \frac{\genjact_{n+2}^{(a,b,c,d)}(1)}{\genjact_{n+1}^{(a,b,c,d)}(1)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} - \frac{\genjact_{n+1}^{(a,b,c,d)}(1)}{\genjact_{n}^{(a,b,c,d)}(1)} \: \tilde{\beta}_{n}^{(a,b,c,d)} + \tilde{\alpha}_{n+1}^{(a,b,c,d)}, \\
 	\tilde{\beta}_n^{(a,b,c+1,d)} &= \frac{\mathcal{C}_n^{(a,b,c,d)}}{\mathcal{C}_{n+1}^{(a,b,c,d)}} \: \frac{\genjact_{n}^{(a,b,c,d)}(1)}{\genjact_{n+1}^{(a,b,c,d)}(1)} \: \tilde{\beta}_n^{(a,b,c,d)}.
-\end{align}
+}
 The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d+1)}\}$ are given by:
-\begin{align}
+\bseqnnumber{
 	\tilde{\alpha}_n^{(a,b,c,d+1)} &= \frac{\genjact_{n+2}^{(a,b,c,d)}(-1)}{\genjact_{n+1}^{(a,b,c,d)}(-1)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} - \frac{\genjact_{n+1}^{(a,b,c,d)}(-1)}{\genjact_{n}^{(a,b,c,d)}(-1)} \: \tilde{\beta}_{n}^{(a,b,c,d)} + \tilde{\alpha}_{n+1}^{(a,b,c,d)}, \\
 	\tilde{\beta}_n^{(a,b,c,d+1)} &= \frac{\mathcal{D}_n^{(a,b,c,d)}}{\mathcal{D}_{n+1}^{(a,b,c,d)}} \: \frac{\genjact_{n}^{(a,b,c,d)}(-1)}{\genjact_{n+1}^{(a,b,c,d)}(-1)} \: \tilde{\beta}_n^{(a,b,c,d)}.
-\end{align}
+}
 \end{proposition}
@@
 First, using (\ref{eqn:genjacexpansioni}) and (\ref{eqn:christoffeldarboux}) we have that
-\begin{align}
+\bseqnnumber{
 	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
@@
 	& \quad \quad \quad - \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \: \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x)\Big) \: \genjact_{n}^{(a,b,c,d)}(1) \Big] \label{eqn:genjactproofa}
-\end{align}
+}
 Next, note that the recurrence coefficients for $\genjact_{n}^{(a,b,c+1,d)}(x)$ satisfy
-\begin{align}
+\bseqnnumber{
 	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
@@
 	& \quad \quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \: \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:genjactproofb}
-\end{align}
+}
 We can set $\tilde{\beta}_{-1}^{(a,b,c+1,d)} = 0$.
@@
 The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c+1,d)}\}$ can be written as:
-\begin{align}
+\bseqnnumber{
 	\tilde{\alpha}_n^{(a,b,c+1,d)} &= \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(1)} - \frac{\tilde{\beta}_{n}^{(a,b,c,d)}}{\chi_{n}^{(a,b,c,d)}(1)} + \tilde{\alpha}_{n}^{(a,b,c,d)}, \\
 	\tilde{\beta}_n^{(a,b,c+1,d)} &= \fpr(\frac{1 - \tilde{\alpha}_{n+1}^{(a,b,c,d)} - \frac{\tilde{\beta}_n^{(a,b,c,d)}}{\chi_n^{(a,b,c,d)}(1)}}{1 - \tilde{\alpha}_{n}^{(a,b,c,d)} - \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(1)}})^\half \tilde{\beta}_n^{(a,b,c,d)}.
-\end{align}
+}
 The recurrence coefficients for the OPs $\{\genjact_n^{(a,b,c,d+1)}\}$ can be written as:
-\begin{align}
+\bseqnnumber{
 	\tilde{\alpha}_n^{(a,b,c,d+1)} &= \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(-1)} - \frac{\tilde{\beta}_{n}^{(a,b,c,d)}}{\chi_{n}^{(a,b,c,d)}(-1)} + \tilde{\alpha}_{n}^{(a,b,c,d)}, \\
 	\tilde{\beta}_n^{(a,b,c,d+1)} &= \fpr(\frac{- 1 + \tilde{\alpha}_{n+1}^{(a,b,c,d)} + \frac{\tilde{\beta}_n^{(a,b,c,d)}}{\chi_n^{(a,b,c,d)}(-1)}}{- 1 + \tilde{\alpha}_{n}^{(a,b,c,d)} + \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(-1)}})^\half \tilde{\beta}_n^{(a,b,c,d)}.
-\end{align}
+}
 where 
-\begin{align}
+\bseqnnumber{
 	\chi_{n}^{(a,b,c,d)}(y) &:= \frac{\genjact_{n+1}^{(a,b,c,d)}(y)}{\genjact_{n}^{(a,b,c,d)}(y)} \\
 	&= \frac{1}{\tilde{\beta}_{n}^{(a,b,c,d)}} \: \fpr(y - \tilde{\alpha}_{n}^{(a,b,c,d)} - \frac{\tilde{\beta}_{n-1}^{(a,b,c,d)}}{\chi_{n-1}^{(a,b,c,d)}(y)}), \quad y \in \{-1,1\}.
-\end{align}
+}
 \end{corollary}
@@
  on \([-1,1]\) with weight \((1-t^2)^c\) as $(t_k,w_k^{(t)})$. Define
-\begin{align*}
+\bseqn{
 	x_{i+(j-1)N} &:= s_j, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
@@
 	w_{i+(j-1)N} &:= w_j^{(s)} w_i^{(t)}, \quad  i,j = 1,\dots, \ceil{\frac{N+1}{2}}.
-\end{align*}
+}
 Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
@@
 We will use the substitution that
-\begin{align*}
+\bseqn{
 	x &= s, \quad y = \rho(s) \: t.
-\end{align*}
+}
 First, note that, for $(x,y) \in \Omega$,
-\begin{align*}
+\bseqn{
 	\Wabc(x,y) &= \genjacw^{(a,b,2c)}(x) \: \jacw^{(c)}\fpr(\frac{y}{\rho(x)}) \\
@@
 	&=: V^{(a,b,c)}(s,t), \quad \text{for } (s,t) \in [\alpha,\beta] \times [-1,1].
-\end{align*}
+}
 
 Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-\begin{align*}
+\bseqn{
 	f_e(x,y) &:= \half \Big(f(x, y) + f(x, -y)\Big), \quad \forall (x,y) \in \Omega\\
 	f_o(x,y) &:= \half \Big(f(x, y) - f(x, -y)\Big), \quad \forall (x,y) \in \Omega
-\end{align*}
+}
 so that $y \mapsto f_e(x,y)$ for fixed $x$ is an even function, and $y \mapsto f_o(x,y)$ for fixed $x$ is an odd function.
@@
 Now, we have that
-\begin{align*}
+\bseqn{
 	\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
@@
 	&= \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
-\end{align*}
+}
 Suppose $f$ is a polynomial in $x$ and $y$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
@@
 Next, note that 
-\begin{align*}
+\bseqn{
 	\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
@@
 	&= 0
-\end{align*}
+}
 since the inner integral at $(\dagger)$ over $t$ is zero, due to the symmetry over the domain.
@@
 Hence, for a polynomial $f$ in $x$ and $y$ of degree $N$,
-\begin{align*}
+\bseqn{
 	\iint_\Omega f(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \iint_\Omega \Big(f_e(x,y) + f_o(x,y)\Big) \: \Wabc(x,y) \: \D y \: \D x \\
@@
 	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j),
-\end{align*}
+}
 where $M = \ceil{\half(N+1)}^2$.
@@
 Fix \(a,b,c \in \R\). Then for any function \(f : \Omega \to \R\) we can express \(f\) by
-\begin{align*}
+\bseqn{
 	f(x,y) \approx \sum_{n=0}^N \bighdop_n^{(a,b,c)}(x,y)^\top \: \mathbf{f}_n
-\end{align*}
+}
 for N sufficiently large, where
-\begin{align*}
+\bseqn{
 	\bighdop^{(a,b,c)}_n(x,y) &:= \begin{pmatrix}
@@
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
-\end{align*}
+}
 and where
-\begin{align*}
+\bseqn{
 	\mathbf{f}_n &:= \begin{pmatrix}
@@
 	f_{n,k} := \frac{\ip< f, \: \hdopnk^{(a,b,c)}>_{\Wabc}}{\norm{\hdopnk^{(a,b,c)}}_{\Wabc}}
-\end{align*}
+}
 Recall from (\ref{eqn:normhdop}) that $\norm{\hdopnkabc}^2_\Wabc = \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}$.
 Using the quadrature rule detailed in Section 4.2 for the inner product, we can calculate the coefficients $f_{n,k}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$:
-\begin{align*}
+\bseqn{
 	f_{n,k} &= \frac{1}{2 \: \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}} \: \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) \: \hdopnkabc(x_j, y_j) +f(x_j, -y_j) \: \hdopnkabc(x_j, -y_j) \big]
-\end{align*}
+}
 where $M = \ceil{\half(N+1)}^2$.
@@
 
-The Poisson equation is the classic problem of finding \(u(x,y)\) given a function \(f(x,y)\) such that:
-\begin{align}
+The Poisson equation is the classic problem of finding $u(x,y)$ given a function $f(x,y)$ such that:
+\bseqnnumber{
 	\begin{cases}
@@
 	\label{eqn:poisson}
-\end{align}
+}
 noting the imposition of zero Dirichlet boundary conditions on $u$.
@@
 In matrix-vector notation, our system hence becomes:
-\begin{align*}
+\bseqn{
     \laplacewiii \mathbf{u} = \mathbf{f}
-\end{align*}
+}
 which can be solved to find $\mathbf{u}$.
 In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 990$, that is, $491,536$ unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
@@
 
-Find \(u(x,y)\) given functions $v$, $f : \Omega \to \R$ such that:
-\begin{align}
+Find $u(x,y)$ given functions $v$, $f : \Omega \to \R$ such that:
+\bseqnnumber{
 	\begin{cases}
@@
 	\label{eqn:helmholtz}
-\end{align}
+}
 where $k \in \R$, noting the imposition of zero Dirichlet boundary conditions on $u$.
@@
 In matrix-vector notation, our system hence becomes:
-\begin{align*}
+\bseqn{
     (\laplacewiii + k^2 T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}) \mathbf{u} = \mathbf{f}
-\end{align*}
+}
 which can be solved to find $\mathbf{u}$.
@@
 In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using $N = 500$, that is, $125,751$ unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
 We can extend this to constant non-zero boundary conditions by simply noting that the problem 
-\begin{align*}
+\bseqn{
 	\begin{cases}
@@
 	\end{cases}
-\end{align*}
+}
 is equivalent to letting $u = \tilde{u} + c$ and solving
-\begin{align*}
+\bseqn{
 	\begin{cases}
@@
 	\end{cases}.
-\end{align*}
+}
 
@@
 
-Find \(u(x,y)\) given a function \(f(x,y)\) such that:
-\begin{align}
+Find $u(x,y)$ given a function $f(x,y)$ such that:
+\bseqnnumber{
 	\begin{cases}
@@
 	\label{eqn:biharmonic}
-\end{align}
+}
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
 In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with N = 200, that is, 20,301 unknowns.
+In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with $N = 500$, that is, $125,751$ unknowns.
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
 
-% For the final Gaussian bump example, we see that we achieve spectral convergence.
-
+%%%%%
+\section{Other domains}
+
+\subsection{End-Disk-Slice}\label{Appendix:HalfDisk}
+
+The work in this paper on the disk-slice can be easily transferred to the special-case domain of the end-disk-slice, such as half disks, by which we mean
+\bseqn{
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+}
+with
+\bseqn{
+\begin{cases}
+\alpha &\in (0,1) \\
+\beta &:= 1 \\
+(\gamma, \delta) &:= (-1,1) \\
+\rho(x) &:= (1-x^2)^{\half}.
+\end{cases}
+}
+Our 1D weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively are then given by:
+\bseqn{
+\begin{cases}
+\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b} \\
+\jacw^{(a)}(x) &:= (1-x^2)^b.
+\end{cases}
+}
+
+Note here how we can remove the need for third parameter, which is why we consider this a special case.
+This will make some calculations easier, and the operator matrices more sparse.
+The weight $\jacw^{(b)}(x)$ is a still the same ultraspherical weight (and the corresponding OPs are the Jacobi polynomials $\{\jac_n^{(b, b)}\}$).
+$\genjacw^{(a,b)}(x)$ is the (non-classical) weight for the OPs denoted $\{\genjac_n^{(a,b)}\}$.
+Thus we arrive at the two-parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\bseqn{
+	\hdopnkab(x,y) := \genjacnmk^{(a, 2b+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+}
+orthogonal with respect to the weight
+\bseqn{
+	\Wab(x,y) &:= \genjacw^{(a,2b)}(x) w^{(b)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
+	&= (x - \alpha)^{a} \: (\rho(x)^2 -y^2)^b \nonumber \\
+	&= (x - \alpha)^{a} \: (1 - x^2 -y^2)^b , \quad (x,y) \in \Omega.
+}
+
+The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such end-disk-slice domains.
+For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for an end-disk-slice.
+
+In Figure \ref{fig:appendixfigs} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
+
+\subsection{Trapeziums}\label{Appendix:Trapezium}
+
+We can further extend this work to trapezium shaped domains. Note that for any trapezium there exists an affine map to the canonical trapezium domain that we consider here, given by
+\bseqn{
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+}
+with 
+\bseqn{
+\begin{cases}
+(\alpha, \beta) &:= (0,1) \\
+(\gamma, \delta) &:= (0,1) \\
+\rho(x) &:= 1- \xi x, \quad \xi \in (0,1) \\
+\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\xi x)^{c} \\
+\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
+\end{cases}
+}
+The weight $\jacw^{(a,b)}(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
+We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = \jac_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
+$\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\).
+Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
+\bseqn{
+	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+}
+orthogonal with respect to the weight
+\bseqn{
+	W^{(a,b,c,d)}(x,y) &:= \genjacw^{(a,b,c+d)}(x) \: w^{(d,c)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
+	&= (1-x)^a \: x^b \: y^c \: (1- \xi x - y)^d, \quad (x,y) \in \Omega.
+}
+
+In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
+
+%\subsection{Quadrature rule on the trapezium}
 %
-\section{Conclusions}
-
-We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
+%In this section we construct a quadrature rule exact for polynomials in the trapezium $\Omega$ that can be used to expand functions in $\hdopnkabcd(x,y)$.
+%
+%\begin{theorem}
+%Denote the Gauss quadrature nodes and weight on $[\alpha,\beta]$ with weight $\genjacw^{(a,b,c+d)}(s) = (\beta - s)^a \: (s - \alpha)^b \: \rho(s)^{c+d+1}$ as $(s_k,w_k^{(s)})$ , and
+% on $[0,1]$ with weight $\jacw^{(d,c)}(t) = t^c \: (1-t)^d$ as $(t_k,w_k^{(t)})$. Define
+%\bseqn{
+%	x_{i+(j-1)N} &:= s_j, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
+%	y_{i+(j-1)N} &:= \rho(s_j) \: t_i, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
+%	w_{i+(j-1)N} &:= w_j^{(s)} w_i^{(t)}, \quad  i,j = 1,\dots, \ceil{\frac{N+1}{2}}.
+%}
+%Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
+%$$
+%\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D A \approx \sum_{j=1}^{M} w_j \: f(x_j, y_j)
+%$$
+%where $M = \ceil{\half(N+1)}^2$, and the quadrature rule is exact if $f(x,y)$ is a polynomial of degree $\le N$.
+%\end{theorem}
+%
+%\begin{proof}
+%\bseqn{
+%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
+%	&\approx \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
+%	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
+%	&= \sum_{k=1}^{M_1 M_2}  w_k \: f(x_k, y_k).
+%}
+%Suppose f is a polynomial in $x$ and $y$ of degree $N$.
+First, note that the degree of the polynomial given by $x \mapsto f(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f(x,y)$ for fixed $x$ is $\le N$.
+Also note that $s \mapsto f\big(s,\rho(s) t\big)$ for fixed $t$ is then a degree $N$ polynomial (since $\rho$ is a degree $1$ polynomial).
+Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
+%
+%Thus, for a polynomial $f$ in $x$ and $y$ of degree $N$, 
+%\bseqn{
+%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
+%	&= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad \\
+%	&= \sum_{j=1}^{\ceil{\half(N+1)}} \Bigg( w_j^{(s)} \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \\
+%	&= \sum_{k=1}^{M}  w_k \: f(x_k, y_k)
+%}
+%where $M = \ceil{\half(N+1)}^2$.
+%
+%\end{proof}
+
+%%
+%\section{Conclusions}
+%
+%
+%We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
 This work extends  the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements, for example, by dividing the disk into disk slice elements.
-%A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
+%%A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
 With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
-This work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles.
+%This work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles.
 In particular, orthogonal polynomials (OPs) in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere.
@@
 Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
-
-\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.  The second author was supported in part by a Leverhulme Trust Research Grant.
-
 %
-\appendix
-%
+%\noindent{\bf Acknowledgements}: We would like to thank the anonymous referees for their helpful comments.
+The author SO was supported in part by the Leverhulme Trust Research Project Grant RPG?2019?144 \enquote{Constructive approximation theory on and inside algebraic curves and surfaces}.
+The author BS was supported by the Engineering and Physical Sciences Research Council Mathematics of Planet Earth Centre for Doctoral Training at Imperial College London and the University of Reading, with grant number EP/L016613/1.
+
 \section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
@@
 
-We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
+It is possible for our framework to be applied to a $p$-finite element method -- that is, one where we can vary the polynomial degree $p$ of the basis functions used in each element (compare this to a normal $h$-FEM, where we can tune the element size $h$).
+For example, one could discretise the disk into disk-slice elements, and apply a $p$-finite element method to solve PDEs on the disk.
+As a precursor to this, in this section we limit our discretisation to a single element.
+Specifically, we follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
 Consider the 2D Dirichlet problem on a domain $\Omega$:
-\begin{align*}
+\bseqn{
 	\begin{cases}
@@
          \end{cases}
-\end{align*}
+}
 This has the weak formulation for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v|_{\partial \Omega} = 0 \}$,
-\begin{align*}
+\bseqn{
 	L(v) := \int_\Omega f \: v \: d\mathbf{x} = \int_\Omega \nabla u \cdot \nabla v \: d\mathbf{x} =: a(u,v).
-\end{align*}
-
-In general, we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a trapezium or disk slice for example.
-
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a disk-slice domain.
+}
+
+As eluded to, in general we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a trapezium or disk slice for example.
+However, here we simply consider our domain to be a disk-slice and our discretisation to be a single element -- that is we let $\element = \Omega$ for a disk-slice domain.
 We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.
@@
 We seek $u_p \in V_p$ s.t.
-\begin{align}
+\bseqnnumber{
 	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
 	\label{eqn:FEMweakform}
-\end{align}
-
-Define $\Lambda^{(a,b,c)} :=  \ip< \bighdop^{(a,b,c)}, \: {\bighdop^{(a,b,c)}}^\top >_{\Wabc}$ where $\Wabc$ is the weight with which the OPs in $\bighdop^{(a,b,c)}$ are orthogonal with respect to.
+}
+
+Recall that the OPs $\bighdop^{(a,b,c)}$ are orthogonal with respect to the weight $\Wabc$ on $\Omega$, and define the matrix $\Lambda^{(a,b,c)} :=  \ip< \bighdop^{(a,b,c)}, \: {\bighdop^{(a,b,c)}}^\top >_{\Wabc}$.
 Note that due to orthogonality this is a diagonal matrix.
 We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
-\begin{align*}
+\bseqn{
 \bigWiii(x,y) &:= \begin{pmatrix}
@@
 					\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
-\end{align*}
+}
 and rewrite (\ref{eqn:FEMweakform}) in matrix form:
-\begin{align*}
+\bseqn{
 	a(u_p,v_p) &= \int_\element \nabla u_p \cdot \nabla v_p \: \D \mathbf{x} \\
@@
 	& \quad \quad \quad \quad + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \Big) \: \mathbf{u}
-\end{align*}
+}
 where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs), and
-\begin{align*}
+\bseqn{
 	L(v_p) &= \int_\element \: v_p \: f \: \D \mathbf{x}\\
@@
 	&= \mathbf{v}^\top \Lambda^{(1,1,1)} \: \mathbf{f},
-\end{align*}
+}
 where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bighdopiii$ OP basis.
@@
 Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-\begin{align*}
+\bseqn{
 	L(\Wiii \hdopiii_{n,k}) = a(u_p,\Wiii \hdopiii_{n,k}) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n,
-\end{align*}
+}
 (i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-\begin{align*}
+\bseqn{
 	A\mathbf{u} = \tilde{\mathbf{f}}.
-\end{align*}
+}
 where the (element) stiffness matrix $A$ is defined by 
-\begin{align*}
+\bseqn{
 	A = {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y
-\end{align*}
+}
 and the load vector $\tilde{\mathbf{f}}$ is given by 
-\begin{align*}
+\bseqn{
 	\tilde{\mathbf{f}} = \Lambda^{(1,1,1)} \: \mathbf{f}.
-\end{align*}
+}
 
 Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-
-\section{End-Disk-Slice}\label{Appendix:HalfDisk}
-
-The work in this paper on the disk-slice can be easily transferred to the special-case domain of the end-disk-slice , such as half disks, by which we mean
-\begin{align*}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
-with
-\begin{align*}
-\begin{cases}
-\alpha &\in (0,1) \\
-\beta &:= 1 \\
-(\gamma, \delta) &:= (-1,1) \\
-\rho(x) &:= (1-x^2)^{\half}.
-\end{cases}
-\end{align*}
-Our 1D weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively are then given by:
-\begin{align*}
-\begin{cases}
-\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b} \\
-\jacw^{(a)}(x) &:= (1-x^2)^b.
-\end{cases}
-\end{align*}
-
-Note here how we can remove the need for third parameter, which is why we consider this a special case.
-This will make some calculations easier, and the operator matrices more sparse.
-The weight $\jacw^{(b)}(x)$ is a still the same ultraspherical weight (and the corresponding OPs are the Jacobi polynomials $\{\jac_n^{(b, b)}\}$).
-$\genjacw^{(a,b)}(x)$ is the (non-classical) weight for the OPs denoted $\{\genjac_n^{(a,b)}\}$.
-Thus we arrive at the two-parameter family of 2D orthogonal polynomials $\{\hdopnkab\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align*}
-	\hdopnkab(x,y) := \genjacnmk^{(a, 2b+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(b,b)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align*}
-orthogonal with respect to the weight
-\begin{align*}
-	\Wab(x,y) &:= \genjacw^{(a,2b)}(x) w^{(b)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
-	&= (x - \alpha)^{a} \: (\rho(x)^2 -y^2)^b \nonumber \\
-	&= (x - \alpha)^{a} \: (1 - x^2 -y^2)^b , \quad (x,y) \in \Omega.
-\end{align*}
-
-The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such end-disk-slice domains.
-For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for an end-disk-slice.
-
-In Figure \ref{fig:appendixfigs} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
-
-\section{Trapeziums}\label{Appendix:Trapezium}
-
-We can further extend this work to trapezium shaped domains. Note that for any trapezium there exists an affine map to the canonical trapezium domain that we consider here, given by
-\begin{align*}
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
-\end{align*}
-with 
-\begin{align*}
-\begin{cases}
-(\alpha, \beta) &:= (0,1) \\
-(\gamma, \delta) &:= (0,1) \\
-\rho(x) &:= 1- \xi x, \quad \xi \in (0,1) \\
-\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} = (1-x)^{a} \: x^b \: (1-\xi x)^{c} \\
-\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b = (1-x)^a \: x^b.
-\end{cases}
-\end{align*}
-The weight $\jacw^{(a,b)}(x)$ is the weight for the shifted Jacobi polynomials on the interval $[0,1]$, and hence the corresponding OPs are the shifted Jacobi polynomials $\{\tilde{P}_n^{(a, b)}\}$.
-We note that the shifted Jacobi polynomials relate to the normal Jacobi polynomials by the relationship $\tilde{P}_n^{(a,b)}(x) = \jac_n^{(a,b)}(2x-1)$ for any degree $n = 0,1,2,\dots$ and $x \in [0,1]$.
-$\genjacw^{(a,b,c)}(x)$ is the (non-classical) weight for the OPs we dentote $\{\genjac_n^{(a,b,c)}\}\).
-Thus we arrive at the four-parameter family of 2D orthogonal polynomials $\{\hdopnk^{(a,b,c,d)}\}$ on $\Omega$ given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
-\begin{align*}
-	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \tilde{P}_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
-\end{align*}
-orthogonal with respect to the weight
-\begin{align*}
-	W^{(a,b,c,d)}(x,y) &:= \genjacw^{(a,b,c+d)}(x) \: w^{(d,c)}_P\fpr(\frac{y}{\rho(x)}) \nonumber \\
-	&= (1-x)^a \: x^b \: y^c \: (1- \xi x - y)^d, \quad (x,y) \in \Omega.
-\end{align*}
-
-In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
-
-%\subsection{Quadrature rule on the trapezium}
-%
-%In this section we construct a quadrature rule exact for polynomials in the trapezium $\Omega$ that can be used to expand functions in $\hdopnkabcd(x,y)$.
-%
-%\begin{theorem}
-%Denote the Gauss quadrature nodes and weight on $[\alpha,\beta]$ with weight $\genjacw^{(a,b,c+d)}(s) = (\beta - s)^a \: (s - \alpha)^b \: \rho(s)^{c+d+1}$ as $(s_k,w_k^{(s)})$ , and
-% on $[0,1]$ with weight $\jacw^{(d,c)}(t) = t^c \: (1-t)^d$ as $(t_k,w_k^{(t)})$. Define
-%\begin{align*}
-%	x_{i+(j-1)N} &:= s_j, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
-%	y_{i+(j-1)N} &:= \rho(s_j) \: t_i, \quad i,j = 1,\dots, \ceil{\frac{N+1}{2}}, \\
-%	w_{i+(j-1)N} &:= w_j^{(s)} w_i^{(t)}, \quad  i,j = 1,\dots, \ceil{\frac{N+1}{2}}.
-%\end{align*}
-%Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
-%$$
-%\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D A \approx \sum_{j=1}^{M} w_j \: f(x_j, y_j)
-%$$
-%where $M = \ceil{\half(N+1)}^2$, and the quadrature rule is exact if $f(x,y)$ is a polynomial of degree $\le N$.
-%\end{theorem}
-%
-%\begin{proof}
-%\begin{align*}
-%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
-%	&\approx \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
-%	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
-%	&= \sum_{k=1}^{M_1 M_2}  w_k \: f(x_k, y_k).
-%\end{align*}
-%Suppose f is a polynomial in $x$ and $y$ of degree $N$.
-First, note that the degree of the polynomial given by $x \mapsto f(x,y)$ for fixed $y$ is $\le N$ and the degree of the polynomial given by $y \mapsto f(x,y)$ for fixed $x$ is $\le N$.
-Also note that $s \mapsto f\big(s,\rho(s) t\big)$ for fixed $t$ is then a degree $N$ polynomial (since $\rho$ is a degree $1$ polynomial).
-Hence, we achieve equality at $(\star)$ if $2M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-%
-%Thus, for a polynomial $f$ in $x$ and $y$ of degree $N$, 
-%\begin{align*}
-%	\iint_\Omega f(x,y) \: \Wabcd(x,y) \: \D y \: \D x &= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \Big( \int_{0}^{1} f\big(s,\rho(s) t\big) \: \jacw^{(d,c)}(t) \: \D t \Big) \: \D s \\
-%	&= \int_\alpha^\beta \genjacw^{(a,b,c+d+1)}(s) \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s,\rho(s) t_i\big) \Big) \: \D s \quad \\
-%	&= \sum_{j=1}^{\ceil{\half(N+1)}} \Bigg( w_j^{(s)} \: \sum_{i=1}^{\ceil{\half(N+1)}} \Big( w_i^{(t)} f\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \\
-%	&= \sum_{k=1}^{M}  w_k \: f(x_k, y_k)
-%\end{align*}
-%where $M = \ceil{\half(N+1)}^2$.
-%
-%\end{proof}
-
-\bibliography{halfdisk}
-
-\end{document}

### 2021-02-17 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
All files for thesis committed

--- 2020-11-18 papers/spherical-caps/spherical-caps.tex
@@
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
-%\geometry{landscape}                		% Activate for rotated page geometry
-\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
-								% TeX will automatically convert eps --> pdf in pdflatex
-
-\graphicspath{ {images/} }
-
-\include{bsthesisnewcommands}
-\include{bsmacros}
-\include{somacros}
-
-\title{Sparse spectral methods for partial differential equations on spherical caps}
-\author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
-%\date{}							% Activate to display a given date or no date
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, disk-slices and triangles.
-In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps.
+\chapter{Spherical Caps}
+
+While the work in the previous chapter looked at developing a sparse spectral method inside a two dimensional domain, we move on to investigating the realm of a surface in three dimensional space.
+Specifically, we look to extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps, which we will formally define.
 The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
 We demonstrate the results on partial differential equations involving the spherical Laplacian and Biharmonic operators, showing spectral convergence.
-\end{abstract}
 
@@
 
-This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
+Our aim in this chapter is to develop sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
 More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
@@
 This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity} with $v(x,y,z) = \cos z$.
-
-An overview of the paper is as follows:
-
-\noindent \secref{OPs}: We present our definition of a (one-parameter) family of 3D orthogonal polynomials (OPs) on the spherical cap domain $\Omega$, by combining 1D OPs on the interval $(\alpha, 1)$ with Chebyshev polynomials, to form 3D OPs on the spherical cap surface.
-We show that these families will lead to sparse Jacobi operators for multiplication by $x, y, z$ and demonstrate how to obtain the 3D OPs.
-
-\noindent\secref{PDOs}: We define several partial differential operators such as spherical Laplacians and  show that these will  be sparse when applied to a suitable choice of expansions in bases built from OPs on the spherical cap.
-We can exactly calculate the non-zero entries of these sparse operators using the quadrature rule associated with the non-classical 1D OPs.
-
-\noindent\secref{Computation}: We derive a quadrature rule on the spherical cap that can be used to expand a function in the OP basis up to a given order $N$, and demonstrate how to evaluate a function using the Clenshaw algorithm using the coefficients of its expansion.
-
-\noindent\secref{Examples}: We demonstrate the proposed technique for solving differential equations on the spherical cap such as the Poisson equation, variable coefficient Helmholtz equation, and Biharmonic equation.
-
-%\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
-We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
 
@@
 
-%
-\section{Conclusions}
-
-We have shown that trivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on spherical cap domains, with Dirichlet boundary conditions on the $z = \alpha \in (0,1)$ boundary.
-We have provided a detailed practical framework for the application of the methods described for quadratic surfaces of revolution \cite{olver2020orthogonal}, by utilising the non-classical 1D OPs on the interval $[\alpha, 1]$ with the weight $(z - \alpha)^a \: (1-z^2)^{b/2}$ defined for the disk-slice case \cite{snowball2019sparse}.
-Generalisation to spherical bands ($\alpha \leq z \leq \beta$) is straightforward.
-This work thus forms a building block in developing an $hp-$finite element method to solve PDEs on the sphere by using spherical band and spherical cap shaped elements.
-
-This work serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere -- it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
-The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
-
-The next stage is to develop an orthogonal basis for the tangent space of the spherical cap (or band), and obtain sparse differential operators for gradient, divergence etc.
-On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
-
-%\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.
-The second author was supported in part by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces.
-
-%
-%%
-%\appendix
-%%
-%\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
-%
-%\bstodoinline{Is this worth including?}
-%
-%\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis} \bstodoinline{Yes thought so too}
-%
-%%\begin{figure}[t]
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-%%        	%\label{fig:solutionblocknorms-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-poisson}
-%%	\centering
-%%	%\label{fig:solution-poisson}
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-%%	%\label{fig:solution-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\centering
-%%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
+\section{$p$-finite element methods using sparse operators}
+
+%\begin{figure}[t]
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
+%        	%\label{fig:solutionblocknorms-poisson}
+%	\centering
+%	\end{subfigure}
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.3]{solution-poisson}
+%	\centering
+%	%\label{fig:solution-poisson}
+%	\end{subfigure}
+%	\begin{subfigure}{0.3\textwidth}
+%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
+%	%\label{fig:solution-poisson}
+%	\centering
+%	\end{subfigure}
+%	\centering
+%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
 Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
@@
 in the trapezium.}
-%%	\centering
-%%	\label{fig:appendixfigs}
-%%\end{figure}
-%
-%We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
-Consider the 3D Dirichlet problem on a domain $\Omega$:
-%\begin{align*}
-%	\begin{cases}
-%         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
-%         u = 0 \quad \text{on } \partial \Omega
-%         \end{cases}
-%\end{align*}
-%This has the weak formulation
-%\begin{align*}
-%	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
-%\end{align*}
-% for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
-%
-%In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$.
-While spherical bands are an extension to this work, we note that we can still apply a $p$-FEM using two half-spheres ($\alpha = 0$).
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain.
+%	\centering
+%	\label{fig:appendixfigs}
+%\end{figure}
+
+As was the case in Section ???, it is possible for our framework in this chapter to be applied to a $p$-finite element method.
+The ultimate aim here would be to discretise the sphere into spherical cap and spherical band elements, and apply a $p$-finite element method to solve PDEs on the sphere.
+As a precursor to this, in this section we limit our discretisation to a single spherical cap element, and follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, once again with the benefit of ensuring that the resulting discretisation is symmetric.
+Consider the Dirichlet problem on a surface in 3D $\Omega$:
+\begin{align*}
+	\begin{cases}
+         	- \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
+		u = 0 \quad \text{on } \partial \Omega
+	\end{cases}
+\end{align*}
+This has the weak formulation
+\begin{align*}
+	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
+\end{align*}
+ for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
+
+As mentioned, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain.
 We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p \rvert_\element) \le p\}$ for some $p \in \N$.
-%
-%We seek $u_p \in V_p$ s.t.
-%\begin{align}
-%	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-%	\label{eqn:FEMweakform}
-%\end{align}
-%
-%Define 
-%\begin{align*}
-%	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
-%\end{align*}
-%for any $a$.
+
+We seek $u_p \in V_p$ s.t.
+\begin{align}
+	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p. \label{eqn:FEMweakform}
+\end{align}
+
+Define 
+\begin{align*}
+	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D A = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D A
+\end{align*}
+for any $a$.
 Note that due to orthogonality this is a diagonal matrix.
@@
 We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.
-%
-%Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
-%\begin{align*}
-%	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
-%					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
-%\end{align*}
-%Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
+
+Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
+\begin{align*}
+	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
+					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
+	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
+\end{align*}
+Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
 Then,
-%\begin{align*}
-%	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
-%	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%
-%Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-%\begin{align*}
-%	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
-%\end{align*}
-%(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-%\begin{align*}
-%	A\mathbf{u} = \tilde{\mathbf{f}}.
-%\end{align*}
-%where the (element) stiffness matrix $A$ is defined by 
-%\begin{align*}
-%	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
-%\end{align*}
-%and the load vector $\tilde{\mathbf{f}}$ is given by 
-%\begin{align*}
-%	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-%
-
-\bibliography{spherical-caps}
-
-\end{document}
+\begin{align*}
+	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
+	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
+\end{align*}
+
+Since (\ref{eqn:FEMweakform}) is equivalent to stating that
+\begin{align*}
+	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
+\end{align*}
+(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
+\begin{align*}
+	A\mathbf{u} = \tilde{\mathbf{f}}.
+\end{align*}
+where the (element) stiffness matrix $A$ is defined by 
+\begin{align*}
+	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
+\end{align*}
+and the load vector $\tilde{\mathbf{f}}$ is given by 
+\begin{align*}
+	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
+\end{align*}
+Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+
+%%%%%%%%%%%%%%%%
+%%% TANGENT SPACE %%%
+
+\section{Vector-valued functions in the tangent space}
+
+We can extend the methodology detailed here to the tangent space of the spherical cap $\Omega$.
+By creating a basis of orthogonal vector-valued functions (OVFs) for the spherical cap $\Omega$, we can expand vector-valued functions that lie in tangent space we call $\tangentspace$ in this basis.
+Such functions that are useful for the sphere are gradients and perpendicular-gradients of scalar functions.
+In this section we will define the OVFs, outline the Clenshaw matrices and how to build the OVFs, function evaluation as well as differential operators for gradient, curl and divergence that take us to and from coefficients in either of the OVF or scalar OP expansions.
+
+Any function in the tangent space of $\Omega$ can be written as $\nabla f + \rvec \times \nabla g$ for some scalar functions $f, g$ on $\Omega$.
+Let $f$ be a scalar function on $\Omega$ that satisfies zero boundary conditions.
+Then there exist coefficients $\bm{f}$ such that $f$ can be expanded in the OP basis $\bigW_N^{(1)}$ for large enough N, i.e.
+$f = \genjacw^{(1,0)} \: \sum_{n,k,i} f_{n,k,i} \: \scopnki^{(1)}$.
+The gradient and perpendicular gradient of $f$ are then
+\begin{align*}
+	\nabla f &= \sum_{n,k,i} f_{n,k,i} \: \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \\
+	\rvec \times \nabla f &= \sum_{n,k,i} f_{n,k,i} \: \rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big).
+\end{align*}
+For the whole sphere, recall from Section ???
+that we simply chose the vector spherical harmonics as our tangent space basis, defined simply as the gradient and perpendicular gradients of the scalar spherical harmonics.
+We could do that because the vector spherical harmonics as defined were naturally orthogonal.
+Unfortunately, we do not have the same orthogonality here with the set $\{ \nabla \Big( \genjacw^{(a,0)} \: \scopnki^{(a)} \Big), \rvec \times \nabla \Big( \genjacw^{(a,0)} \: \scopnki^{(a)} \Big) \}$ for any parameter $a$.
+
+The game is to find an orthogonal basis that spans the tangent space of the spherical cap $\Omega$ such that we can sparsely expand $\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ and $\rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ for each $n,k,i$ trio.
+
+Define the OVFs as follows:
+\begin{align}
+	\tsopinki(x,y,z) &:= \phivec \: \frac{1}{\rho(z)} \: \scopnki^{(0)}(x,y,z) \\
+	\tsopiinki(x,y,z) &:= \thetavec \: \frac{1}{\rho(z)} \: \scopnki^{(0)}(x,y,z)
+\end{align}
+where
+\begin{align}
+	\phivec &:= 
+		\begin{pmatrix}
+			\cos(\theta) \: \cos(\phi) \\
+			\sin(\theta) \: \cos(\phi) \\
+			- \sin(\phi)
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+			x z / \rho(z) \\
+			y z / \rho(z) \\
+			- \rho(z)
+		\end{pmatrix} \\
+	\thetavec &:= 
+		\begin{pmatrix}
+			-\sin(\theta) \\
+			\cos(\theta) \\
+			0
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+			-y / \rho(z) \\
+			x / \rho(z) \\
+			0
+		\end{pmatrix}
+\end{align}
+These are orthogonal with respect to the inner product $\ip<\bm{A}, \bm{B}>_\tangentspace := \int_\Omega \bm{A} \cdot \bm{B} \: \rho(z)^2 \: \D A$:
+\begin{align*}
+	\ip<\tsopinki, \tsopi_{m,j,h}>_\tangentspace &= \int_\Omega \tsopinki \cdot \tsopi_{m,j,h} \: \rho(z)^2 \:\D A \\
+	&= \int_\Omega \scopnki^{(0)} \: \scopmjh^{(0)} \: \D \Omega \\
+	&= \pi \: \normgenjac^{(0, 2k)} \: \delta_{n,m} \: \delta_{k,j} \: \delta_{i,h}, \\
+	\ip<\tsopiinki, \tsopii_{m,j,h}>_\tangentspace &= \int_\Omega \tsopiinki \cdot \tsopii_{m,j,h} \: \rho(z)^2 \:\D A \\
+	&= \int_\Omega \scopnki^{(0)} \: \scopmjh^{(0)} \: \D \Omega \\
+	&= \pi \: \normgenjac^{(0, 2k)} \: \delta_{n,m} \: \delta_{k,j} \: \delta_{i,h}, \\
+	\ip<\tsopinki, \tsopii_{m,j,h}>_\tangentspace &\equiv 0.
+\end{align*}
+To show that these span the tangent space $\tangentspace$, we merely require that we can sparsely expand $\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \: \rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ with the vectors $\tsopi_{m,j,h}, \tsopii_{m,j,h}$.
+First, note that
+\begin{align*}
+	\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) &= \phivec \: \ppphi \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) + \thetavec \: \frac{1}{\rho(z)} \: \pptheta \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \\
+	&= \phivec \: \Big[k \: z \: \genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) -  \big(\genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) \big)' \rho(z)^2 \Big] \: \rho(z)^{k-1} \: \chki(\theta) \\
+	&\quad \quad + \thetavec \: (-1)^{i+1} \: k \: \genjacw^{(1,0)}(z) \: \genjacnmk^{(1,2k)}(z) \: \rho(z)^{k-1} \: \ch_{k,|i-1|}(\theta).
+\end{align*}
+Then,
+\begin{align*}
+	&\ip<\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \tsopi_{m,j,h}>_\tangentspace \\
+	&\quad \quad = \int_\Omega \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \cdot \tsopi_{m,j,h} \: \rho(z)^2 \:\D A \\
+	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big[k \: z \: \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} -  (\genjacw^{(1,0)} \: \genjacnmk^{(1,2k)})' \rho^2 \Big] \: \rho^{k-1} \: \genjac_{m-k}^{(0,2k)} \: \rho^{k-1} \: \rho^2 \:\D z \\
+	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big[k \: z \: \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} -  (\genjacw^{(1,0)} \: \genjacnmk^{(1,2k)})' \rho^2 \Big] \: \genjac_{m-k}^{(0,2k)} \: \rho^{2k} \:\D z \quad (\star) \\
+	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \Big( kz \: \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(0,2k)} \: \genjacw^{(1,2k)} \\
+	&\quad \quad\quad \quad\quad \quad\quad \quad\quad \quad+ \genjacw^{(1,2k)} \: \genjacnmk^{(1,2k)} \: \Big[ \genjac_{m-k}^{(0,2k) \; \prime} \: \rho^2 - (2k+2) \: z \: \genjac_{m-k}^{(0,2k)} \Big] \Big) \: \D z \\
+	&\quad \quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \genjacnmk^{(1,2k)} \: \Big[ - (k + 2) \: z \: \genjac_{m-k}^{(0,2k)} + \genjac_{m-k}^{(0,2k) \; \prime} \: \rho^2 \Big] \: \genjacw^{(1,2k)} \: \D z \quad (\star \star)
+\end{align*}
+where by orthogonality of the $\{\genjac_n^{(a,2b)}\}$ OPs we have that the above is zero for $m - k > n - k + 2 \iff m > n + 2$ at $(\star)$ and also for $n - k > m - k + 1 \iff m < n - 1$ at $(\star \star)$.
+Further,
+\begin{align*}
+	&\ip<\nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big), \tsopii_{m,j,h}>_\tangentspace \\
+	&\quad \quad = \int_\Omega \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big) \cdot \tsopii_{m,j,h} \: \rho(z)^2 \:\D A \\
+	&\quad \quad = (-1)^{i+1} \: k \: \pi \: \delta_{k,j} \: \delta_{i,|h-1|} \: \int_\alpha^1 \genjacw^{(1,0)} \: \genjacnmk^{(1,2k)} \: \rho^{k-1} \: \genjac_{m-k}^{(0,2k)} \: \rho^{k-1} \: \rho^2 \:\D z \\
+	&\quad \quad = (-1)^{i+1} \: k \: \pi \: \delta_{k,j} \: \delta_{i,|h-1|} \: \int_\alpha^1 \genjacnmk^{(1,2k)}  \: \genjac_{m-k}^{(0,2k)} \: \genjacw^{(1,2k)} \: \D z
+\end{align*}
+which again by orthogonality is zero for $m - k > n - k + 1 \iff m > n + 1$ and $n - k > m - k \iff m < n$.
+We of course have similar results for $\rvec \times \nabla \Big( \genjacw^{(1,0)} \: \scopnki^{(1)} \Big)$ since
+\begin{align*}
+	\rvec \times \phivec = \thetavec, \quad \rvec \times \thetavec = - \phivec.
+\end{align*}
+We now have a set of orthogonal vector valued functions that span the tangent space $\tangentspace$.
+By construction, we have the same recurrence coefficients for multiplication by $x,y,z$ for each of the OVFs as for the OPs $\scopnki^{(0)}$.
+
+Define
+\begin{align*}
+	\bigtsoptn := 
+		\begin{pmatrix}
+			\tsopit_{n,0,0}(x,y,z) \\
+			\tsopiit_{n,0,0}(x,y,z) \\
+			\tsopit_{n,1,0}(x,y,z) \\
+			\tsopiit_{n,1,0}(x,y,z) \\
+			\tsopit_{n,1,1}(x,y,z) \\
+			\tsopiit_{n,1,1}(x,y,z) \\
+			\vdots \\
+			\tsopit_{n,n,0}(x,y,z) \\
+			\tsopiit_{n,n,0}(x,y,z) \\
+			\tsopit_{n,n,1}(x,y,z) \\
+			\tsopiit_{n,n,1}(x,y,z)
+		\end{pmatrix} \in \R^{2(2n+1) \times 3}, 
+	\quad \quad 
+	\bigtsopt := 
+		\begin{pmatrix}
+			\bigtsopt_0 \\
+			\bigtsopt_1 \\
+			\bigtsopt_2 \\
+			\vdots \\
+		\end{pmatrix}
+\end{align*}
+and set $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ as the Jacobi matrices corresponding to
+\begin{align}
+	\jacobimattangentx \: \bigtsopt(x,y,z) = x \: \bigtsopt(x,y,z), \nonumber \\
+	\jacobimattangenty \: \bigtsopt(x,y,z) = y \: \bigtsopt(x,y,z), \label{eqn:jacobimatricestangentdefinition} \\
+	\jacobimattangentz \: \bigtsopt(x,y,z) = z \: \bigtsopt(x,y,z), \nonumber
+\end{align}
+The matrices $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ act on the coefficients vector of a function's expansion in the $\bigtsopt$ basis.
+For example, let a function $\bm{F}(x,y,z) : \Omega \to \tangentspace$ be approximated by its expansion $\bm{F}(x,y,z)^\top = \bigtsopt(x,y,z)^\top \bm{F}_c$, for some coefficients vector $\bm{F}_c$.
+Then $x \: \bm{F}(x,y,z)$ is approximated by $\bigtsopt(x,y,z)^\top {(\jacobimattangentx)}^\top \bm{F}_c$.
+In other words, ${(\jacobimattangentx)}^\top \bm{F}_c$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: \bm{F}(x,y,z)$ in the  $\bigtsopt$ basis.
+Further, note that $\jacobimattangentx, \jacobimattangenty, \jacobimattangentz$ are banded-block-banded matrices -- they are block-tridiagonal (block-bandwidths $(1,1)$):
+\begin{align*}
+	\jacobimattangent_{x/y/z} &= 
+		\begin{pmatrix}
+			B_{x/y/z, 0} & A_{x/y/z, 0} & & & & \\
+			C_{x/y/z, 1} & B_{x/y/z, 1} & A_{x/y/z, 1} & & & \\
+			& C_{x/y/z, 2} & B_{x/y/z, 2} & A_{x/y/z, 2} & & & \\
+			& & C_{x/y/z, 3} & \ddots & \ddots & \\
+			& & & \ddots & \ddots & \ddots \\
+		\end{pmatrix}
+\end{align*}
+For $\jacobimattangentx$, the sub-blocks have sub-block-bandwidths $(?,?)$:
+\begin{align*}
+	A_{x,n} &:= 
+		\begin{pmatrix}
+			0 & A_{n,0,6} & 0 & & \\
+			A_{n,1,5} & \ddots & \ddots & & \\
+			& \ddots & \ddots & \ddots & \\
+			& & A_{n,n,5} & 0 & A_{n,n,6} \\
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
+	B_{x,n} &:= 
+		\begin{pmatrix}
+			0 & A_{n,0,4} & & \\
+			A_{n,1,3} & \ddots & \ddots & \\
+			& \ddots & \ddots & A_{n,n-1,4} \\
+			& & A_{n,n,3} & 0
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
+	C_{x,n} &:= 
+		\begin{pmatrix}
+			0 & A_{n,0,2} & & \\
+			A_{n,1,1} & \ddots & \ddots & \\
+			& \ddots & \ddots & A_{n,n-2,2} \\
+			& & \ddots & 0 \\
+			& & & A_{n,n,1} \\
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
+\end{align*}
+where
+\begin{align*}
+	A_{n,k,j} &:= \alphaonkj \: I_{4} \in \R^{4\times4}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
+	A_{n,0,j} &:=
+		\begin{pmatrix}
+			\alphao_{n,0,j} & 0 & 0 & 0 \\
+			0 & \alphao_{n,0,j} & 0 & 0
+		\end{pmatrix} \in \R^{2\times4}, \quad j \text{ even} \\
+	A_{n,1,j} &:=
+		\begin{pmatrix}
+			\alphao_{n,1,j} & 0 \\
+			0 & \alphao_{n,1,j} \\
+			0 & 0 \\
+			0 & 0
+		\end{pmatrix} \in \R^{4\times2}, \quad j \text{ odd}
+\end{align*}
+For $\jacobimattangenty$, the sub-blocks have sub-block-bandwidths $(?,?)$:
+\begin{align*}
+	A_{y,n} &:= 
+		\begin{pmatrix}
+			0 & B_{n,0,6} & 0 & & \\
+			B_{n,1,5} & \ddots & \ddots & & \\
+			& \ddots & \ddots & \ddots & \\
+			& & B_{n,n,5} & 0 & B_{n,n,6} \\
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
+	B_{y,n} &:= 
+		\begin{pmatrix}
+			0 & B_{n,0,4} & & \\
+			B_{n,1,3} & \ddots & \ddots & \\
+			& \ddots & \ddots & B_{n,n-1,4} \\
+			& & B_{n,n,3} & 0
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
+	C_{y,n} &:= 
+		\begin{pmatrix}
+			0 & B_{n,0,2} & & \\
+			B_{n,1,1} & \ddots & \ddots & \\
+			& \ddots & \ddots & B_{n,n-2,2} \\
+			& & \ddots & 0 \\
+			& & & B_{n,n,1} \\
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
+\end{align*}
+where
+\begin{align*}
+	B_{n,k,j} &:= 
+		\begin{pmatrix}
+			0 & 0 & \betaonkj & 0 \\
+			0 & 0 & 0 & \betaonkj \\
+			\betaonkj & 0 & 0 & 0 \\
+			0 & \betaonkj & 0 & 0
+		\end{pmatrix} \in \R^{4\times4}, \quad k = 1,\dots,n \: (j \text{ even}), \quad k = 2,\dots,n \: (j \text{ odd}) \\
+	B_{n,0,j} &:=
+		\begin{pmatrix}
+			0 & 0 & \betao_{n,0,j} & 0 \\
+			0 & 0 & 0 & \betao_{n,0,j}
+		\end{pmatrix} \in \R^{2\times4}, \quad j \text{ even} \\
+	B_{n,1,j} &:=
+		\begin{pmatrix}
+			0 & 0 \\
+			0 & 0 \\
+			\betao_{n,1,j} & 0 \\
+			0 & \betao_{n,1,j}
+		\end{pmatrix} \in \R^{4\times2}, \quad j \text{ odd}
+\end{align*}
+For $\jacobimattangentz$, the sub-blocks have sub-block-bandwidths $(?,?)$:
+\begin{align*}
+	A_{z,n} &:= 
+		\begin{pmatrix}
+			\Gamma_{n,0,3} & 0 & 0 & & \\
+			0 & \ddots & \ddots & & \\
+			& \ddots & \ddots & \ddots & \\
+			& & 0 & \Gamma_{n,n,3} & 0 \\
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+3)}, \quad n = 0,1,2,\dots \\
+	B_{z,n} &:= 
+		\begin{pmatrix}
+			\Gamma_{n,0,2} & 0 & & \\
+			0 & \ddots & \ddots & \\
+			& \ddots & \ddots & 0 \\
+			& & 0 & \Gamma_{n,n,2}
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n+1)}  \quad n = 0,1,2,\dots \\
+	C_{z,n} &:= 
+		\begin{pmatrix}
+			\Gamma_{n,0,1} & 0 & & \\
+			0 & \ddots & \ddots & \\
+			& \ddots & \ddots & 0 \\
+			& & \ddots &  \Gamma_{n,n-1,1} \\
+			& & & 0 \\
+		\end{pmatrix} \in \R^{2(2n+1)\times2(2n-1)}, \quad n = 1,2,\dots
+\end{align*}
+where
+\begin{align*}
+	\Gamma_{n,k,j} &:= \gammaonkj \: I_{4} \in \R^{4\times4}, \quad k = 1,\dots,n \\
+	\Gamma_{n,0,j} &:=
+		\begin{pmatrix}
+			\gammao_{n,0,j} & 0 \\
+			0 & \gammao_{n,0,j}
+		\end{pmatrix} \in \R^{2\times2}.
+\end{align*}
+Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
+
+\subsection{Building the OPs}
+
+We can combine each system in (\ref{eqn:jacobimatricestangentdefinition}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
+
+\begin{align*}
+\renewcommand\arraystretch{1.3}
+\begin{pmatrix}
+		I_2 & & & \\
+		B_0-G_0(x,y,z) & A_0 & & \\
+		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
+		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
+		& & \ddots &\ddots
+\end{pmatrix}
+\bigtsopt(x,y,z)
+=
+\begin{pmatrix}
+	 \tsopi_0^\top \\ \tsopii_0^\top \\ 0 \\ 0 \\ \vdots  \\
+\end{pmatrix},
+\end{align*}
+where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
+\begin{align*}
+A_n &:= \begin{pmatrix}
+		A_{x,n} \\
+		A_{y,n} \\
+		A_{z,n}
+	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n+3)}, \quad
+C_n := \begin{pmatrix}
+		C_{x,n} \\
+		C_{y,n} \\
+		C_{z,n}
+	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n-1)} \quad (n \ne 0), \nonumber \\
+B_n &:= \begin{pmatrix}
+		B_{x,n} \\
+		B_{y,n} \\
+		B_{z,n}
+	    \end{pmatrix} \in \R^{6(2n+1)\times2(2n+1)}, \quad
+G_n(x,y,z) := \begin{pmatrix}
+		xI_{2(2n+1)} \\
+		yI_{2(2n+1)} \\
+		zI_{2(2n+1)}
+	    \end{pmatrix} \in \R^{6(2n+1)\times2(n+1)}.
+\end{align*}
+
+For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e.
+such that $\Dnt A_n = I_{2(2n+3)}$.
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+\begin{align*}
+	\begin{cases}
+		\bigtsopt_{-1}(x,y,z) := 0 \\
+		\bigtsopt_{0}(x,y,z) = 
+			\begin{pmatrix} 
+				\tsopi_0^\top \\ 
+				\tsopii_0^\top
+			\end{pmatrix} :=
+			\begin{pmatrix} 
+				\tsopi_{0,0,0}^\top \\
+				\tsopi_{0,0,0}^\top
+			\end{pmatrix} \\
+		\bigtsopt_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigtsoptn(x,y,z) - \Dnt C_n  \, \bigtsopt_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
+	\end{cases}
+\end{align*}
+
+Note that we can define an explicit \(\Dnt\) as follows: \bstodo{type up DnT for tangent space}
+\begin{align*}
+\Dnt := \begin{pmatrix}
+		0 & & & 0 & & & 1 / \gammaa_{n,0,3} & & \\
+		& \ddots & & & \ddots & & & \ddots \\
+		& & \ddots & & & \ddots & & & \ddots \\
+		& & & 0 & & & 0 & & & 1 / \gammaa_{n,n,3}  \\
+		& & & & & \bm{\eta}^\top_{1} & & & \\
+		& & & & & \bm{\eta}^\top_{0} & & &
+	    \end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
+\end{align*}
+for $n = 2, 3, \dots$ where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
+\begin{align*}
+	\big(\bm{\eta}_{0}\big)_j &= 
+		\begin{cases}
+			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
+			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2 \\
+			0 & o/w
+		\end{cases} \\
+	\big(\bm{\eta}_{1}\big)_j &= 
+		\begin{cases}
+			\big(\bm{\eta}_{0}\big)_{j+1} & j = 2n, 3(2n+1) - 3 \\
+			0 & o/w
+		\end{cases}
+\end{align*}
+For $n=0, 1$, we can simply take
+\begin{align*}
+D^\top_0 &:= \begin{pmatrix}
+		0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
+		\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
+		0 & \frac{1}{\betaa_{0,0,6}} & 0
+	    \end{pmatrix} \in \R^{3\times3}, \\
+D^\top_1 &:= \begin{pmatrix}
+		0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,0,3} & & \\
+		0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3} \\
+		0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 / \gammaa_{1,1,3}  \\
+		0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & \eta & 0 & 0 \\
+		0 & 0 & 1 / \alphaa_{1,1,6} & 0 & 0 & 0 & 0 & 0 & 0
+	    \end{pmatrix} \in \R^{5\times9},
+\end{align*}
+where $\eta = \frac{- \: \alphaa_{1,1,5}}{\alphaa_{1,1,6} \: \gammaa_{1, 0, 3}}$.
+
+It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigtsopt_0(x,y,z)$  through $\bigtsoptn(x,y,z)$ in optimal $O(n^2)$ complexity. \bstodo{correct for this case}
+
+\subsection{Operators}
+
+Gradient operator on a scalar function is pretty much outlined above.
+Divergence would either be worked out by knowing what $f,g$ are for a function $\bm{F} = \nabla f + \rvec \times \nabla g$, so that $\nabla \cdot \bm{F} = \nabla^2 f$.
+Alternatively, we have expressions for $\rhoppphi \scopnkia$ and $\pptheta \scopnkia$ so that we can calculate an operator for $\rho^2 \nabla \cdot$, i.e.
+\begin{align*}
+	\rho^2 \nabla \cdot \bm{F} &= \rho^2 \: \frac{1}{\rho} \ppphi (\rho F^\phi) + \rho^2 \: \frac{1}{\rho} \pptheta F^\theta \\
+	&= \sum_{n,k,i} F^\phi_{n,k,i} \: \rhoppphi \scopnki^{(0)} + F^\theta_{n,k,i} \: \pptheta \scopnki^{(0)}.
+\end{align*}
+
+\subsection{Linear SWE}
+
+\begin{align*}
+	\uvec_t &= - f \rvec \times \uvec + \grad h \\
+	h_t &= -\href \divergence \uvec
+\end{align*}
+$\implies$
+\begin{align*}
+	\uvecnpi &= \uvecn - \Delta t \: f \rvec \times \uvecnpi +\Delta t \: \grad h_{n+1} \\
+	h_{n+1} &= h_n - \Delta t \: \href \divergence \uvecnpi
+\end{align*}
+$\implies$
+\begin{align*}
+	\uvecnpic &= \uvecnc - \Delta t \: F \uvecnpic +\Delta t \: G \hvecnpic \\
+	P \hvecnpic &= P \hvecnc - \Delta t \: \href D \uvecnpic
+\end{align*}
+where $F : \bigtsop \to \bigtsop$ is the operator for $f \rvec \times$, $D : \bigtsop \to \bigscop^{(1)}$ is the operator for $\rho^2 \divergence$, $G : \bigW^{(1)} \to \bigtsop$ is the operator for $\grad$, and $P : \bigW^{(1)} \to \bigscop^{(1)}$ is the operator for multiplication by $\rho(z)^2$.

### 2021-02-18 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Edits to SH chapter

@@
 Jacobi operators that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$.
-We start by expressing $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$ , and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.
-
+We can obtain the entries to these operator matrices by determining the coefficients of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.
+
+First though, we require a few results concerning the incrementing and decrementing of the parameters for the Jacobi polynomials, and the three-term recurrence for the Jacobi polynomials, that will also prove useful later on in this chapter.
 \begin{lemma}
+	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
+\bseqn{
+	P_{n}^{(\alpha, \beta)}(z) =
+		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
+									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
+									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
+									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
+	P_{n}^{(\alpha, \beta)}(z) =
+		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
+									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
+									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
+									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
+	z \, P_{n}^{(\alpha, \beta)}(z) = 
+		\Big\{ & {2(n+\alpha)(n+\beta) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 1)} \: P_{n-1}^{(\alpha, \beta)}(z) \\
+			 & {(\beta^2 - \alpha^2) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 2)} \: P_{n}^{(\alpha, \beta)}(z) \\
+			 & {2(n+1)(n + \alpha + \beta + 1) \over 2(2n + \alpha + \beta + 1)(2n + \alpha + \beta + 2)} \: P_{n+1}^{(\alpha, \beta)}(z).
+}
+\end{lemma}
+\begin{proof}
+	\bstodoinline{Do the second result (i.e. $(1-z^2)P_n$ relationship).}
+	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5, 18.9.6]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
+The final one is simply a rearranging of the established three-term recurrence detailed in \cite[18.9.1, 18.9.2]{DLMF}
+\end{proof}
+We thus can present a useful corollary specific for our needs.
+\begin{corollary}\label{corollary:incrementjacparams}
+	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
+\bseqn{
+	(1-z^2) \, \jaclm{l}{m}(z) &=  \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z) + \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) \\
+	\jaclm{l}{m}(z) &= \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) + \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z), \\
+	z \, \jaclm{l}{m}(z) &= \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
+}
+where
+\bseqn{
+	\tilde \alpha_{l,m,1} &:= {2l \over 2l+1} \\
+	\tilde \alpha_{l,m,2} &:= 
+		\begin{cases}
+			- {l \over 2(2l+1)} \quad \text{if } l - m \ge 2 \\
+			\quad \quad 0 \quad \quad \text{otherwise}
+		\end{cases} \\
+	\tilde \alpha_{l,m,3} &:= - {2(l-m+2)(l-m+1) \over (2l+1)(l+1)} \\
+	\tilde \alpha_{l,m,4} &:= {(l+m+2)(l+m+1) \over 2(2l+1)(l+1)} \\
+	\tilde \gamma_{l,m,1} &:= \begin{cases}
+			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
+			\quad 0 \quad \quad \text{otherwise}
+		\end{cases} \\
+	\tilde \gamma_{l,m,2} &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
+}
+\end{corollary}
+
+\bsrefcor{corollary:incrementjacparams} allows us to easily determine the aforementioned coefficients, that we will formalise in the following Lemma.
+
+\begin{lemma}\label{lemma:shxyzrelations}
 	For $l \in \No$, $m \in \Z$ s.t. $0 \le |m| \le l$, the spherical harmonics satisfy the relationships:
@@
 }
-and
-\bseqn{
-	\tilde \alpha_{l,m,1} &:= \frac{2l}{2l+1} \\
-	\tilde \alpha_{l,m,2} &:= 
-		\begin{cases}
-			- \frac{l}{2(2l+1)} \quad \text{if } l - m \ge 2 \\
-			\quad \quad 0 \quad \quad \text{otherwise}
-		\end{cases} \\
-	\tilde \alpha_{l,m,3} &:= - \frac{2(l-m+2)(l-m+1)}{(2l+1)(l+1)} \\
-	\tilde \alpha_{l,m,4} &:= \frac{(l+m+2)(l+m+1)}{2(2l+1)(l+1)} \\
-	\tilde \gamma_{l,m,1} &:= \begin{cases}
-			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
-			\quad 0 \quad \quad \text{otherwise}
-		\end{cases} \\
-	\tilde \gamma_{l,m,2} &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
-}
 \end{lemma}
@@
 \begin{proof}
-	The recurrence relationship for the Jacobi polynomials satisfies \cite[18.9.1, 18.9.2]{DLMF} 
-\bseqn{
-	z P^{(m,m)}_{l-m}(z) = \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
-}
-for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is
+	The expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is simply a consequence of the third result in \bsrefcor{corollary:incrementjacparams}:
 \bseqn{
@@
 
-For multiplication by $x$ and $y$, we require some further relations for the complex exponential and the Jacobi polynomials. First, recall that
+For multiplication by $x$ and $y$, note that the complex exponential satisfies
 \bseqn{
@@
 }
-We can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an Ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
-\bseqn{
-	{l! \: (2m)! \over (l+m)! \: m!} \: C^{(m+1/2)}_{l-m}(z) = P^{(m,m)}_{l-m}(z),
-}
-where the Ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
-\bseqn{
-	C^{(\lambda)}_n(z) &= {\lambda \over n + \lambda} \: \big[ C^{(\lambda+1)}_n(z) - C^{(\lambda+1)}_{n-2}(z) \big]  \\
-	(1-z^2) \, C^{(\lambda)}_n(z) &= {1 \over 4(\lambda-1)(n+\lambda)} \: \big[ (n+2\lambda-2)(n+2\lambda-1) \: C^{(\lambda-1)}_n(z) \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (n+1)(n+2) \: C^{(\lambda-1)}_{n+2}(z) \big] ,
-}
-for $m \in \No$ where $\lambda > -\half$ is some parameter. Thus, combining these, we can write three-term recurrences for the Jacobi polynomials as:
-\bseqn{
-	P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z) + \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) \\
-	(1-z^2) \, P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) + \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z),
-}
-for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
+Combining this with the results in \bsrefcor{corollary:incrementjacparams}, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
 \bseqn{
@@
 	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
-	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
+	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z),
+}
+and
+\bseqn{
 	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
@@
 }
-
 \end{proof}
@@
 }
-where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:normjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}.
-One final relation we will need is that for incrementing the Jacobi parameters, which we will present in the following Lemma.
-\begin{lemma}
-	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
-\bseqn{
-	P_{n}^{(\alpha, \beta)}(z) =
-		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
-									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
-									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
-									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}.
-}
-\end{lemma}
-\begin{proof}
-	This is simply as consequence of the relationships established for incrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
-\end{proof}
-We thus can present a useful corollary specific for our needs.
-\begin{corollary}\label{corollary:incrementjacparams}
-	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
-\bseqn{
-	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z),
-}
-where
-\bseqn{
-	\mu_{l,m,1} &:= {(l+m+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
-	\mu_{l,m,2} &:= {l \over 2(2l + 1)}.
-}
-\end{corollary}
+where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:diffjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}.
 
 The expressions for the VSHs as given in \bsrefeqn{eqn:VSHs} together with the above allow us to be able to find the coefficients for our desired expressions for $x \gradYlm$, $y \gradYlm$, $z \gradYlm$, $x \gradpYlm$, $y \gradpYlm$ and $z \gradpYlm$.
+
 \begin{lemma}\label{lemma:VSHrecurrences}
-	For $l \in \No$, $-l \le m \le l$ the vector spherical harmonics satisfy the relationships:
+	For $l \in \N$ and $m \in \Z$ s.t. $-l \le m \le l$, the vector spherical harmonics satisfy the relationships:
 \bseqn{
@@
 }
-and
+where the $c_{l'}^{m'}$, $d_{l',m'}$ are defined in \bsrefeqnslist{eqn:clmdef}{eqn:diffjac}{} respectively, and
 \bseqn{
@@
 
-Fix $l \in \N, m\in\{-l,\dots,l\}$.
+Fix $l \in \N$ and $m \in \Z$ s.t. $-l \le m \le l$.
 

### 2021-02-22 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Further edits. Need to type up proof of VSH x,y,z mults

@@
 	P_{n}^{(\alpha, \beta)}(z) =
-		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
-									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
-									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
-									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
-	P_{n}^{(\alpha, \beta)}(z) =
-		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
-									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
-									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
-									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
+		{1 \over 2n + \alpha + \beta + 1} \Big\{ & - {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \\
+									  & + (n+\alpha+\beta+1) \Big[{n+\alpha \over 2n+\alpha+\beta} - {n+\beta+1 \over 2n+\alpha+\beta+2}\Big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
+									  & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
+	(1-z^2) \, P_{n}^{(\alpha, \beta)}(z) =
+		{4 \over 2n + \alpha + \beta + 1} \Big\{ & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n}^{(\alpha-1, \beta-1)}(z) \\
+									  & + (n+1) \Big[{n+\alpha+1 \over 2n+\alpha+\beta+2} - {n+\beta \over 2n+\alpha+\beta}\Big] \: P_{n+1}^{(\alpha-1, \beta-1)}(z) \\
+									  & - {(n+1)(n+2) \over 2n+\alpha+\beta+2} \: P_{n+2}^{(\alpha-1, \beta-1)}(z) \Big\}, \\
 	z \, P_{n}^{(\alpha, \beta)}(z) = 
-		\Big\{ & {2(n+\alpha)(n+\beta) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 1)} \: P_{n-1}^{(\alpha, \beta)}(z) \\
+		& {2(n+\alpha)(n+\beta) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 1)} \: P_{n-1}^{(\alpha, \beta)}(z) \\
 		& {(\beta^2 - \alpha^2) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 2)} \: P_{n}^{(\alpha, \beta)}(z) \\
@@
 \begin{proof}
-	\bstodoinline{Do the second result (i.e. $(1-z^2)P_n$ relationship).}
 	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5, 18.9.6]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
@@
 We thus can present a useful corollary specific for our needs.
-\begin{corollary}\label{corollary:incrementjacparams}
+\begin{corollary}\label{corollary:jacobirelations}
 	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
@@
 
-\bsrefcor{corollary:incrementjacparams} allows us to easily determine the aforementioned coefficients, that we will formalise in the following Lemma.
+\bsrefcor{corollary:jacobirelations} allows us to easily determine the aforementioned coefficients, that we will formalise in the following Lemma.
 
@@
 \begin{proof}
-	The expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is simply a consequence of the third result in \bsrefcor{corollary:incrementjacparams}:
+	The expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is simply a consequence of the third result in \bsrefcor{corollary:jacobirelations}:
 \bseqn{
@@
 }
-Combining this with the results in \bsrefcor{corollary:incrementjacparams}, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
+Combining this with the results in \bsrefcor{corollary:jacobirelations}, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
 \bseqn{
@@
 
-%%%
+%%% TANGENT SPACE ----- VECTOR SHs
 \section{Vector spherical harmonics as orthogonal vectors in three variables}
@@
 \subsection{Jacobi matrices}
-
-% REMOVED SPIN harmonics and Clebsch Gorden coeff stuff as its much neater to use an inner product argument I have found.
 
@@
 }
-where the $c_{l'}^{m'}$, $d_{l',m'}$ are defined in \bsrefeqnslist{eqn:clmdef}{eqn:diffjac}{} respectively, and
-\bseqn{
-	\tilde A_{l,m,1} &:= \Big( (m^2 - 1) \mu_{l-1,m-1,2} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l-1,m-1} {\mu_{l-1,m,2} \over (c_{l}^{m+1})^2} \\
-	\tilde A_{l,m,2} &:= \Big( m(m+2) \mu_{l,m,1} - d_{l,m} \tilde \gamma_{l,m+1,1} \Big) {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\mu_{l,m+1,1} \over (c_{l-1}^{m+2})^2} \\
-	\tilde A_{l,m,3} &:= \Big( (m^2 - 1) \mu_{l+1,m-1,1} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l+1,m-1} {\mu_{l+1,m,1} \over (c_{l}^{m+1})^2} \\
-	\tilde A_{l,m,4} &:= \Big( m(m+2) \mu_{l,m,2} - d_{l,m} \tilde \gamma_{l,m+1,2} \Big) {1 \over (c_{l+1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\mu_{l,m+1,2} \over (c_{l+1}^{m+2})^2}.
-}
+where the $c_{l'}^{m'}$, $d_{l',m'}$, $\tilde \gamma_{l',m',j}$ are defined in \bsrefeqn{eqn:clmdef}, \bsrefeqn{eqn:diffjac} and \bsrefcor{corollary:jacobirelations} respectively, and
+\bseqn{
+	\tilde A_{l,m,1} &:= \Big( (m^2 - 1) \tilde \alpha_{l-1,m-1,4} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l-1,m-1} {\tilde \alpha_{l,m+1,1} \over (c_{l-1}^{m})^2} \\
+	\tilde A_{l,m,2} &:= \Big( m(m+2) \tilde \alpha_{l,m,2} - d_{l,m} \tilde \gamma_{l,m+1,1} \Big) {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\tilde \alpha_{l-1,m+2,3} \over (c_{l}^{m+1})^2} \\
+	\tilde A_{l,m,3} &:= \Big( (m^2 - 1) \tilde \alpha_{l+1,m-1,2} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l+1,m-1} {\tilde \alpha_{l,m+1,3} \over (c_{l+1}^{m})^2} \\
+	\tilde A_{l,m,4} &:= \Big( m(m+2) \tilde \alpha_{l,m,4} - d_{l,m} \tilde \gamma_{l,m+1,2} \Big) {1 \over (c_{l+1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\tilde \alpha_{l+1,m+2,1} \over (c_{l}^{m+1})^2}.
+}
+where again the $\tilde \alpha_{l',m',j}$ are defined in \bsrefcor{corollary:jacobirelations} too.
 \end{lemma}

### 2021-02-24 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Nearly completed proof of VSH mult. New bscommands additions

@@
 \bseqnnumber{
-	\gradYlm, \quad \gradpYlm := \rvec \times \gradYlm, \label{eqn:VSHs}
+	\gradYlm &:= \nabla \Ylm = \ppx{\phi} \Ylm \phivec + {1 \over \sinphi} \ppx{\theta} \Ylm \thetavec, \nonumber \\
+	\gradpYlm &:= \nabla^\perp \Ylm \equiv \rvec \times \nabla \Ylm = \ppx{\phi} \Ylm \thetavec - {1 \over \sinphi} \ppx{\theta} \Ylm \phivec, \label{eqn:VSHs}
 }
 span the tangent space.
-Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$.
-Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}, orthogonal with respect to the inner product $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$.
+Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$, while $\phivec, \thetavec$ are the standard spherical coordinate unit vectors, i.e.
+$\phivec := [\costheta \cosphi, \sintheta \cosphi, -\sinphi]^\top$, $\thetavec := [-\sintheta, \costheta, 0]^\top$.
+Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}, orthogonal with respect to the inner product $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$ -- in particular, for any $l, l' \in \N$, $m, m' \in \Z$ s.t.
+$-l \le m \le l$, $-l' \le m' \le l'$,
+\bseqn{
+	\ip<\gradYlm, \gradYlmprime> &= \iint_\Omega \gradYlm \cdot {\gradYlmprime}^* \: \sinphi \: \D\varphi \: \D\theta = \delta_{l,l'} \delta_{m,m'} \: l (l+1), \\
+	\ip<\gradpYlm, \gradpYlmprime> &= \iint_\Omega \gradpYlm \cdot {\gradpYlmprime}^* \: \sinphi \: \D\varphi \: \D\theta = \delta_{l,l'} \delta_{m,m'} \: l (l+1), \\
+	\ip<\gradYlm, \gradpYlmprime> &= \iint_\Omega \gradpYlm \cdot {\gradpYlmprime}^* \: \sinphi \: \D\varphi \: \D\theta = 0.
+}
 Vector spherical harmonics have been widely used in electrostatics (e.g.
@@
 
-We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc.
-in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$.
+We start as before for the scalar case by finding $x \, \gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc.
+in terms of $\gradYlmprime (x,y,z)$, $\gradpYlmprime (x,y,z)$.
 A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations.
@@
 \bseqn{
-	A_{l,m,1} &:= {\clm c_{l-1}^{m-1} \over 2}
+	A_{l,m,1} &:= {\clm c_{l-1}^{m-1} \over 2 l (l-1)}
 		\begin{cases}
@@
 		\end{cases} \\
-	A_{l,m,2} &:= {\clm c_{l-1}^{m+1} \over 2}
+	A_{l,m,2} &:= {\clm c_{l-1}^{m+1} \over 2 l (l-1)}
 		\begin{cases}
@@
 		\end{cases} \\
-	A_{l,m,3} &:= {\clm c_{l-1}^{m-1} \over 2}
+	A_{l,m,3} &:= {\clm c_{l-1}^{m-1} \over 2 (l+1)(l+2)}
 		\begin{cases}
@@
 		\end{cases} \\
-	A_{l,m,4} &:= {\clm c_{l-1}^{m+1} \over 2}
+	A_{l,m,4} &:= {\clm c_{l-1}^{m+1} \over 2 (l+1)(l+2)}
 		\begin{cases}
@@
 		\end{cases} \\
-	A_{l,m,5} &:=
+	A_{l,m,5} &:= {1 \over l(l+1)}
 		\begin{cases}
@@
 		\end{cases} \\
-	A_{l,m,6} &:=
+	A_{l,m,6} &:= {1 \over l(l+1)}
 		\begin{cases}
@@
 	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6, \\
-	\Gamma_{l,m,1} &:= \clm c_{l-1}^{m} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \Big] \\
-	\Gamma_{l,m,2} &:= \clm c_{l+1}^{m} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \Big] \\
-	\Gamma_{l,m,3} &:= (-1)^m \: i \: m
+	\Gamma_{l,m,1} &:= {\clm c_{l-1}^{m} \over l(l-1)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \Big] \\
+	\Gamma_{l,m,2} &:= {\clm c_{l+1}^{m} \over (l+1)(l+2)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \Big] \\
+	\Gamma_{l,m,3} &:= {(-1)^m \: i \: m \over l(l+1)}
 }
@@
 
-Fix $l \in \N$ and $m \in \Z$ s.t. $-l \le m \le l$.
+	Fix $l \in \N$ and $m \in \Z$ s.t. $-l \le m \le l$. The proof we will use is to directly calculate the nonzero coefficients in each expansion using inner products. To this end, note that
+\bseqn{
+	x \gradYlm = \sum_{l'=1}^{l+1} \sum_{m'=-l'}^{l'} \calA_{l',m'} \gradYlmprime + \calA^\perp_{l',m'} \gradpYlmprime
+}
+where the coefficients are give by
+\bseqn{
+	\calA_{l',m'} = {\ip<x \gradYlm, \gradYlmprime> \over \norm{\gradYlmprime}^2} = {1 \over l'(l'+1)} \ip<x \gradYlm, \gradYlmprime> \\
+	\calA^\perp_{l',m'} = {\ip<x \gradYlm, \gradpYlmprime> \over \norm{\gradpYlmprime}^2} = {1 \over l'(l'+1)} \ip<x \gradYlm, \gradpYlmprime>.
+}
+The aim will be to show that the only nonzero coefficients $\calA_{l',m'}$, $\calA^\perp_{l',m'}$ match the $A_{l,m,j}$ for $j=1,\dots,6$ in the Lemma. Now, using a change of variable, we have that
+\bseqnnumber{
+	&\ip<x \gradYlm, \gradYlmprime> \nonumber \\
+	&= \iint_\Omega \costheta \sinphi \: \gradYlm \cdot {\gradYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \rho(z)^{\abs{m}+\abs{m'}-1} \nonumber \\
+	&\quadeight + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: \D z \Big) \nonumber \\
+	&= \clm \clmprime \pi (\delta_{m',m-1} + \delta_{m',m+1}) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m}+\abs{m'}-1} \big(mm' + \abs{m}\abs{m'} z^2 \big) \nonumber \\
+	&\quadeight - \abs{m} z \jaclmabs{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m}+\abs{m'}+1} \nonumber \\
+	&\quadeight - \abs{m'} z \jaclmabsdz{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m}+\abs{m'}+1} \nonumber  \\
+	&\quadeight + \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m}+\abs{m'}+3} \Big\} \: \D z \Big). \label{eqn:lemmaVSHrecurrencesproof}
+}
+We can conclude that the only possible nonzero coefficients are when $m' = m \pm 1$, and so it is left to evaluate the integral that is left at $(\star)$ when $m'$ takes either value.
+Since we need to take account of the fact that $m$ can be negative, to simplify our argument we note that the case $m' = m - 1$ for $m < 0$ is analogous to the case $m' = m+1$ for $m > 0$, and vice versa.
+
+On that note, we will first consider the case that $m > 0$ and $m' = m - 1$, and calculate the integral in \bsrefeqn{eqn:lemmaVSHrecurrencesproof}:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ m(m-1) \jaclm{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m-2} (1 + z^2) - m z \jaclm{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m} \nonumber \\
+	&\quadtwo \quad - (m-1) z \jaclmdz{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m} + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m(m-1) \jaclm{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m-2} (1 + z^2) - m z \jaclm{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m} \nonumber \\
+	&\quadfour + (m-1) \jaclm{l}{m} \rho^{2m-2} \big( - 2m z^2 \jaclmbracket{l'}{m-1} + \jaclmbracket{l'}{m-1} \rho^{2} - z \jaclmbracketdz{l'}{m-1} \rho^{2} \big) \nonumber \\
+	&\quadfour + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ (m-1)(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m} - z \jaclm{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m} \nonumber \\
+	&\quadfour + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ (m-1)(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m} - d_{l',m-1} z \jaclm{l}{m} \jaclm{l'}{m} \rho^{2m} \nonumber \\
+	&\quadfour + d_{l,m} d_{l',m-1} \jaclmbracket{l}{m+1} \jaclm{l'}{m} \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ (m^2 - 1) \jaclm{l}{m} \big( \tilde \alpha_{l', m-1, 2} \jaclm{l'-1}{m} + \tilde \alpha_{l', m-1, 4} \jaclm{l'+1}{m} \big) \rho^{2m} \nonumber \\
+	&\quadfour - d_{l',m-1} \jaclm{l}{m} \big( \tilde \gamma_{l',m,1} \jaclm{l'-1}{m} + \tilde \gamma_{l',m,2} \jaclm{l'+1}{m} \big) \rho^{2m} \nonumber \\
+	&\quadfour + d_{l,m} d_{l',m-1} \jaclm{l'}{m} \big( \tilde \alpha_{l, m+1, 1} \jaclm{l-1}{m} + \tilde \alpha_{l, m+1, 3} \jaclm{l+1}{m} \big) \rho^{2m} \Big\} \: \D z \nonumber \\
+	&= \delta_{l',l-1} \Big( \big[ (m^2 - 1) \tilde \alpha_{l-1, m-1, 4} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \big] {1\over2\pi(\clm)^2} + d_{l,m} d_{l-1,m-1} {\tilde \alpha_{l-1, m-1, 1} \over 2\pi (c_{l-1}^m)^2} \Big) \nonumber \\
+	&\quadtwo + \delta_{l',l+1} \Big( \big[ (m^2 - 1) \tilde \alpha_{l+1, m-1, 2} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \big] {1\over2\pi(\clm)^2} + d_{l,m} d_{l+1,m-1} {\tilde \alpha_{l-1, m-1, 3} \over 2\pi (c_{l+1}^m)^2} \Big).
+\label{eqn:lemmaVSHrecurrencesproof2}
+}
+As we can see, the coefficients are then also only possibly nonzero for $l' = l \pm 1$.
+Combining the above with the \bsrefeqn{eqn:lemmaVSHrecurrencesproof}, we have that our two coefficients for $\calA_{l-1,m-1}, \calA_{l+1,m-1}$ are those stated in the Lemma as $A_{l,m,1}$ and $A_{l,m,3}$, for $m > 0$.
+
+Next, assume $m \ge 0$ and consider the case $m' = m + 1$. The integral in \bsrefeqn{eqn:lemmaVSHrecurrencesproof} is then:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ m(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m} (1 + z^2) - m z \jaclm{l}{m} \jaclmbracketdz{l'}{m+1} \rho^{2m+2} \nonumber \\
+	&\quadtwo \quad - (m+1) z \jaclmdz{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m+1} \rho^{2m+4} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m} (1 + z^2) - (m+1) z \jaclmdz{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} \nonumber \\
+	&\quadfour + m \jaclmbracket{l'}{m+1} \rho^{2m} \big( - 2(m+1) z^2 \jaclm{l}{m} + \jaclm{l}{m} \rho^{2} - z \jaclmdz{l}{m} \rho^{2} \big) \nonumber \\
+	&\quadfour + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m+1} \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m(m+2) \jaclm{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} - z \jaclmdz{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} \nonumber \\
+	&\quadfour + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m+1} \rho^{2m+4} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m(m+2) \jaclm{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} - d_{l,m} z \jaclmbracket{l}{m+1} \jaclmbracket{l'}{m+1} \rho^{2m+2} \nonumber \\
+	&\quadfour + d_{l,m} d_{l',m+1} \jaclmbracket{l}{m+1} \jaclmbracket{l'}{m+2} \rho^{2m+4} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m(m+2) \jaclmbracket{l'}{m+1} \big( \tilde \alpha_{l, m, 2} \jaclmbracket{l-1}{m+1} + \tilde \alpha_{l, m, 4} \jaclmbracket{l+1}{m+1} \big) \rho^{2m+2} \nonumber \\
+	&\quadfour - d_{l,m} \jaclmbracket{l'}{m+1} \big( \tilde \gamma_{l,m+1,1} \jaclmbracket{l-1}{m+1} + \tilde \gamma_{l,m+1,2} \jaclmbracket{l+1}{m+1} \big) \rho^{2m+2} \nonumber \\
+	&\quadfour + d_{l,m} d_{l',m+1} \jaclmbracket{l}{m+1} \big( \tilde \alpha_{l', m+2, 1} \jaclmbracket{l'-1}{m+1} + \tilde \alpha_{l', m+2, 3} \jaclmbracket{l'+1}{m+1}  \big) \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= \delta_{l',l-1} \Big( \big[ m(m+2) \tilde \alpha_{l, m, 2} - d_{l,m} \tilde \gamma_{l,m+1,1} \big] {1 \over 2 \pi (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\tilde \alpha_{l-1, m+2, 3} \over 2\pi (c_{l}^{m+1})^2} \Big) \nonumber \\
+	&\quadtwo + \delta_{l',l+1} \Big( \big[ m(m+2) \tilde \alpha_{l, m, 4} - d_{l,m} \tilde \gamma_{l,m+1,2} \big] {1 \over 2 \pi (c_{l-1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\tilde \alpha_{l+1, m+2, 1} \over 2\pi (c_{l}^{m+1})^2} \Big).
+\label{eqn:lemmaVSHrecurrencesproof3}
+}
+Once again, we can see that the coefficients are then also only possibly nonzero for $l' = l \pm 1$.
+Combining the above with the \bsrefeqn{eqn:lemmaVSHrecurrencesproof}, we have that our two coefficients for $\calA_{l-1,m-1}, \calA_{l+1,m-1}$ are those stated in the Lemma as $A_{l,m,2}$ and $A_{l,m,4}$, for $m \ge 0$.
+
+We still have to account for the remaining $m = 0$ cases and $m$ negative cases.
+Thankfully, as noted earlier, the case of $m \le 0$, $m' = m-1$ is the same as replacing $m$ with $\abs{m}$ in \bsrefeqn{eqn:lemmaVSHrecurrencesproof3}.
+Thus, our two coefficients for $\calA_{l-1,m-1}, \calA_{l+1,m-1}$ are those stated in the Lemma as $A_{l,m,1}$ and $A_{l,m,3}$, for $m \le 0$.
+Similarly the case of $m < 0$, $m' = m+1$ is the same as replacing $m$ with $\abs{m}$ in \bsrefeqn{eqn:lemmaVSHrecurrencesproof2}, and thus our two coefficients for $\calA_{l-1,m+1}, \calA_{l+1,m+1}$ are those stated in the Lemma as $A_{l,m,2}$ and $A_{l,m,4}$ for $m < 0$.
+
+The remaining coefficients we need to determine are $\calA^\perp_{l',m'}$. To this end, we have that %make this lemmaVSHrecurrencesproof4
+\bseqnnumber{
+	&\ip<x \gradYlm, \gradpYlmprime> \nonumber \\
+}
+
+An almost identical argument holds for multiplication by $y$, since
+\bseqn{
+	 \int_0^{2pi} \eimtheta \eimthetaprime \sintheta \: \D\theta = i \pi (\delta_{m',m-1} - \delta_{m',m+1})
+}
+replaces the integrals over $\theta$ in \bsrefeqn{eqn:lemmaVSHrecurrencesproof} and \bsrefeqn{eqn:lemmaVSHrecurrencesproof4}.
+
+Finally, mult by z...
+
+Finally, for multiplication by $x,y,z$ of $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
+noting that $\rvec \times \gradYlm = \gradpYlm$ and $\rvec \times \gradpYlm = - \norm{\rvec} \gradYlm = - \gradYlm$ by the triple vector product and orthogonality.
 

### 2021-03-01 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Further edits to VSH xyz mult proof

@@
 }
-We can conclude that the only possible nonzero coefficients are when $m' = m \pm 1$, and so it is left to evaluate the integral that is left at $(\star)$ when $m'$ takes either value.
-Since we need to take account of the fact that $m$ can be negative, to simplify our argument we note that the case $m' = m - 1$ for $m < 0$ is analogous to the case $m' = m+1$ for $m > 0$, and vice versa.
+We can conclude that the only possible nonzero coefficients are when $m' = m \pm 1$, and so it is left to evaluate the integral that is left at \bsrefeqn{eqn:lemmaVSHrecurrencesproof} when $m'$ takes either value.
+Since we need to take account of the fact that $m$ can be negative, to simplify our argument we make use of the fact that $m' = m - 1$ for $m < 0$ is analogous to the case $m' = m+1$ for $m > 0$, and vice versa.
 
@@
 }
-Once again, we can see that the coefficients are then also only possibly nonzero for $l' = l \pm 1$.
+Again, we can see that the coefficients are then also only possibly nonzero for $l' = l \pm 1$.
 Combining the above with the \bsrefeqn{eqn:lemmaVSHrecurrencesproof}, we have that our two coefficients for $\calA_{l-1,m-1}, \calA_{l+1,m-1}$ are those stated in the Lemma as $A_{l,m,2}$ and $A_{l,m,4}$, for $m \ge 0$.
@@
 
-The remaining coefficients we need to determine are $\calA^\perp_{l',m'}$. To this end, we have that %make this lemmaVSHrecurrencesproof4
+The remaining coefficients we need to determine are $\calA^\perp_{l',m'}$. To this end, we have that
 \bseqnnumber{
 	&\ip<x \gradYlm, \gradpYlmprime> \nonumber \\
-}
+	&= \iint_\Omega \costheta \sinphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ i m' \jaclmabs{l'}{m'}(z) \rho(z)^{\abs{m} + \abs{m'} - 1} \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
+	&\quadeight + i m \jaclmabs{l}{m}(z) \rho(z)^{\abs{m} + \abs{m'} - 1} \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
+	&= \clm \clmprime \pi (\delta_{m',m-1} + \delta_{m',m+1}) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ 2 i \, m \, m' \, z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} - 1} \nonumber \\
+	&\quadeight - i m \jaclmabsdz{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} + 1} \nonumber \\
+	&\quadeight - i m' \jaclmabs{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m} + \abs{m'} +1} \Big\} \: \D z \Big). \label{eqn:lemmaVSHrecurrencesproof4}
+}
+Once again, we can conclude that $\calA^\perp_{l',m'}$ are zero for $m' \notin \{m-1, m+1\}$.
+It is left to evaluate the integral that is left at \bsrefeqn{eqn:lemmaVSHrecurrencesproof4} when $m'$ takes either $m\pm 1$ value.
+First, consider the case that $m > 0$ and $m' = m - 1$.
+The integral is then, using integration by parts:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m-1} + \abs{m} (m-1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} - 1} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \nonumber \\
+	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour + i (m-1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^2 - 2\abs{m} z \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= - i \; {d_{l,\abs{m}-1} \over 2 \pi (c_l^{\abs{m}})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof5}
+}
+Combining \bsrefeqn{eqn:lemmaVSHrecurrencesproof5} with \bsrefeqn{eqn:lemmaVSHrecurrencesproof4} means we retrieve the correct values for our coefficients $\calA^\perp_{l,m-1}$ when $m > 0$, i.e.
+they are $A_{l,m,5}$ as stated in the Lemma for $m > 0$.
+The case of $m < 0$ and $m' = m + 1$ is very similar with the integral becoming:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m+1} + \abs{m} (m+1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} - 1} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \nonumber \\
+	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour + i (m+1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^2 - 2\abs{m} z \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= - i \; {d_{l,\abs{m}-1} \over 2 \pi (c_l^{\abs{m}})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof6}
+}
+Combining \bsrefeqn{eqn:lemmaVSHrecurrencesproof6} with \bsrefeqn{eqn:lemmaVSHrecurrencesproof4} means we retrieve the correct values for our coefficients $\calA^\perp_{l,m+1}$ when $m < 0$, i.e.
+they are $A_{l,m,6}$ as stated in the Lemma for $m < 0$.
+
+Next, assume $m \ge 0$ and $m' = m + 1$. Then the integral in \bsrefeqn{eqn:lemmaVSHrecurrencesproof4} becomes:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m+1} + \abs{m} (m+1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} - 1} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \nonumber \\
+	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&\quadfour + i m \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m}} \big[\jaclmabsdz{l}{m} \rho^2 - 2 (\abs{m} + 1) z \jaclmabs{l}{m} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l,\abs{m}} \; \jaclmbracketdz{l}{\abs{m} + 1} \jaclmbracketdz{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \: \D z \nonumber \\
+	&= - i \; {d_{l,\abs{m}} \over 2 \pi (c_l^{\abs{m}+1})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof7}
+}
+Again, combining \bsrefeqn{eqn:lemmaVSHrecurrencesproof7} with \bsrefeqn{eqn:lemmaVSHrecurrencesproof4} means we retrieve the correct values for our coefficients $\calA^\perp_{l,m+1}$ when $m \ge 0$, i.e.
+they are $A_{l,m,6}$ as stated in the Lemma for $m \ge 0$.
+In a similar vein, the integral for the case $m \le 0$ and $m' = m - 1$ is:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m-1} + \abs{m} (m-1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} - 1} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \nonumber \\
+	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&\quadfour + i m \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m}} \big[\jaclmabsdz{l}{m} \rho^2 - 2 (\abs{m} + 1) z \jaclmabs{l}{m} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l,\abs{m}} \; \jaclmbracketdz{l}{\abs{m} + 1} \jaclmbracketdz{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \: \D z \nonumber \\
+	&= - i \; {d_{l,\abs{m}} \over 2 \pi (c_l^{\abs{m}+1})^2} \; \delta_{l',l}, \label{eqn:lemmaVSHrecurrencesproof8}
+}
+and so combining \bsrefeqn{eqn:lemmaVSHrecurrencesproof8} with \bsrefeqn{eqn:lemmaVSHrecurrencesproof4}, we have that our coefficients for $\calA^\perp_{l,m-1}$ are those stated in the Lemma as $A_{l,m,5}$ for $m \le 0$, concluding the argument for the expansion of $x \gradYlm$.
 
@@
 
-Finally, mult by z...
-
-Finally, for multiplication by $x,y,z$ of $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
+For mult by z... \bstodoinline{do z}
+
+Finally, for multiplication by $x,y,z$ of the other VSHs $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
 noting that $\rvec \times \gradYlm = \gradpYlm$ and $\rvec \times \gradpYlm = - \norm{\rvec} \gradYlm = - \gradYlm$ by the triple vector product and orthogonality.

### 2021-03-02 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Finished proof. Need to finish chapter and tidy spacing.

@@
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \rho(z)^{\abs{m}+\abs{m'}-1} \nonumber \\
-	&\quadeight + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: \D z \Big) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
+	&\quadeight + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: \rho(z)^{\abs{m}+\abs{m'}-1} \: \D z \Big) \nonumber \\
 	&= \clm \clmprime \pi (\delta_{m',m-1} + \delta_{m',m+1}) \nonumber \\
@@
 	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \nonumber \\
-	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
-	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
-	&\quadfour + i (m-1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^2 - 2\abs{m} z \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \big] \Big\} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \: \D z \nonumber \\
-	&= - i \; {d_{l,\abs{m}-1} \over 2 \pi (c_l^{\abs{m}})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof5}
+	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m-1} \rho^{m + \abs{m-1} + 1} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclm{l}{m} \jaclmbracket{l'}{m - 1} \rho^{2m - 2} - i m \jaclm{l}{m} \jaclmbracketdz{l'}{m - 1} \rho^{2m} \nonumber \\
+	&\quadfour - i (m-1) \jaclmdz{l}{m} \jaclmbracket{l'}{m - 1} \rho^{2m} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclm{l}{m}\jaclmbracket{l'}{m - 1} \rho^{2m - 2} - i m \jaclm{l}{m} \jaclmbracketdz{l'}{m - 1} \rho^{2m} \nonumber \\
+	&\quadfour + i (m-1) \jaclm{l}{m} \rho^{2m - 2} \big[ \jaclmbracketdz{l'}{m - 1} \rho^2 - 2m z \jaclmbracket{l'}{m - 1} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclm{l}{m} \jaclmbracketdz{l'}{m - 1} \rho^{2m} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclm{l}{m} \jaclm{l'}{m} \rho^{2m} \: \D z \nonumber \\
+	&= - i \; {d_{l,m-1} \over 2 \pi (c_l^{m})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof5}
 }
@@
 	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
-	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \nonumber \\
-	&\quadfour + i (m+1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^2 - 2\abs{m} z \jaclmuser{l' - \abs{m} + 1}{\abs{m}-1} \big] \Big\} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmuserdz{l' - \abs{m} + 1}{\abs{m}-1} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour + i (m+1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmbracketdz{l'}{\abs{m} - 1} \rho^2 - 2\abs{m} z \jaclmbracket{l'}{\abs{m} - 1} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \: \D z \nonumber \\
 	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \: \D z \nonumber \\
@@
 	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
-	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
-	&\quadfour + i m \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m}} \big[\jaclmabsdz{l}{m} \rho^2 - 2 (\abs{m} + 1) z \jaclmabs{l}{m} \big] \Big\} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \; d_{l,\abs{m}} \; \jaclmbracketdz{l}{\abs{m} + 1} \jaclmbracketdz{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \: \D z \nonumber \\
-	&= - i \; {d_{l,\abs{m}} \over 2 \pi (c_l^{\abs{m}+1})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof7}
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclm{l}{m} \jaclmbracket{l'}{m + 1} \rho^{2 m} - i m \jaclm{l}{m} \jaclmbracketdz{l'}{m + 1}  \rho^{2m + 2} \nonumber \\
+	&\quadfour - i (m+1) \jaclmdz{l}{m} \jaclmbracket{l'}{m + 1} \rho^{2m + 2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclm{l}{m} \jaclmbracket{l'}{m + 1} \rho^{2 m} - i (m+1) \jaclmdz{l}{m} \jaclmbracket{l'}{m + 1}  \rho^{2m + 2} \nonumber \\
+	&\quadfour + i m \jaclmbracket{l'}{m + 1} \rho^{2m} \big[ \jaclmdz{l}{m} \rho^2 - 2 (m + 1) z \jaclm{l}{m} \big] \Big\} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclm{l}{m} \jaclmbracketdz{l'}{m + 1}  \rho^{2m + 2} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l,m} \; \jaclmbracketdz{l}{m + 1} \jaclmbracketdz{l'}{m + 1} \rho^{2m + 2} \: \D z \nonumber \\
+	&= - i \; {d_{l,m} \over 2 \pi (c_l^{m+1})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof7}
 }
@@
 
-For mult by z... \bstodoinline{do z}
-
-Finally, for multiplication by $x,y,z$ of the other VSHs $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
+For multiplication by z, we will use the same method to directly calculate the nonzero coefficients in each expansion using inner products. To this end, note that
+\bseqn{
+	z \gradYlm = \sum_{l'=1}^{l+1} \sum_{m'=-l'}^{l'} \calC_{l',m'} \gradYlmprime + \calC^\perp_{l',m'} \gradpYlmprime
+}
+where the coefficients are give by
+\bseqn{
+	\calC_{l',m'} = {\ip<z \gradYlm, \gradYlmprime> \over \norm{\gradYlmprime}^2} = {1 \over l'(l'+1)} \ip<z \gradYlm, \gradYlmprime> \\
+	\calC^\perp_{l',m'} = {\ip<z \gradYlm, \gradpYlmprime> \over \norm{\gradpYlmprime}^2} = {1 \over l'(l'+1)} \ip<z \gradYlm, \gradpYlmprime>.
+}
+The aim once again will be to show that the only nonzero coefficients $\calC_{l',m'}$, $\calC^\perp_{l',m'}$ match the $\Gamma_{l,m,j}$ for $j=1,2,3$ in the Lemma.
+Now, using a change of variable and integration by parts, we have that
+\bseqnnumber{
+	&\ip<z \gradYlm, \gradYlmprime> \nonumber \\
+	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
+	&\quadeight + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: z \; \rho(z)^{\abs{m}+\abs{m'}-2} \: \D z \Big) \nonumber \\
+	&= \clm \clmprime 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m}+\abs{m'}-2} \big(mm' + \abs{m}\abs{m'} z^2 \big) \nonumber \\
+	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabs{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m}+\abs{m'}} \nonumber \\
+	&\quadeight \quadtwo - \abs{m'} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m}+\abs{m'}} \nonumber  \\
+	&\quadeight \quadtwo + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m}+\abs{m'}+2} \Big\} \: \D z \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ m^2 z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} (1 + z^2) \nonumber \\
+	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabs{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}} \nonumber \\
+	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber  \\
+	&\quadeight \quadtwo + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ m^2 z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} (1 + z^2) \nonumber \\
+	&\quadeight \quadtwo + \abs{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} \big[ z^2 \jaclmabsdz{l}{m} \rho^2 + 2z \jaclmabs{l'}{m} - 2\abs{m}z^3 \jaclmabs{l'}{m} \rho^2 \big] \nonumber \\
+	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber  \\
+	&\quadeight \quadtwo + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber \\
+	&\quadeight \quadtwo + z d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l}{\abs{m}+1} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) \jaclmabs{l'}{m} \rho^{2\abs{m}} \big[ \tilde\gamma_{l,\abs{m},1} \jaclmabs{l-1}{m} + \tilde\gamma_{l,\abs{m},2} \jaclmabs{l+1}{m} \big] \nonumber \\
+	&\quadeight \quadtwo + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\
+	&= \clm 2 \pi \delta_{m',m} \Big\{ c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over 2\pi (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over 2\pi (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
+	&\quadfour c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over 2\pi (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over 2\pi (c_{l+1}^{\abs{m}+1})^2} \big] \Big\} \nonumber \\
+	&= \clm \delta_{m',m} \Big\{ c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
+	&\quadfour c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}.
+\label{eqn:lemmaVSHrecurrencesproofz}
+}
+We can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz} that the coefficients $\calC_{l',m'}$ are zero when $l' \ne l \pm 1$ or $m' \ne m \pm 1$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,1}$ and $\Gamma_{l,m,3}$.
+
+Next, we also have that
+\bseqnnumber{
+	&\ip<z \gradYlm, \gradpYlmprime> \nonumber \\
+	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ i m' z \jaclmabs{l'}{m'}(z) \rho(z)^{\abs{m} + \abs{m'} - 2} \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
+	&\quadeight + i m z \jaclmabs{l}{m}(z) \rho(z)^{\abs{m} + \abs{m'} - 2} \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ 2i m\abs{m} z^2 \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m} - 2} \nonumber \\
+	&\quadeight \quadtwo - i m z \jaclmabs{l}{m}(z) \jaclmabsdz{l'}{m}(z) \rho(z)^{2\abs{m}} \nonumber \\
+	&\quadeight \quadtwo - i m z \jaclmabsdz{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \Big\} \: \D z \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} i m z^2 \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
+	&= \delta_{l',l} \delta_{m',m} i m {(\clm)^2 \over (\clmprime)^2} \nonumber \\
+	&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz2}
+}
+Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz2} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,2}$.
+
+Finally, to complete our proof, for multiplication by $x,y,z$ of the other VSHs $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
 noting that $\rvec \times \gradYlm = \gradpYlm$ and $\rvec \times \gradpYlm = - \norm{\rvec} \gradYlm = - \gradYlm$ by the triple vector product and orthogonality.

### 2021-03-12 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Finished proofs for SH work. Removed Tangent space stuff from CH4, made appendix. Need to address comments.

@@
 }
-Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and are given by \bstodo{Add citation for associated legendre polys}:
-\bseqn{
-	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; \chat \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x) \\
-	P^{-m}_l(x) &:= (-1)^m \; \ctilde \; \Plm(x),
-}
-for $m = 0, 1, 2, \dots, l$ where
-\bseqn{
-	\chat &:= {(l+m)! \over 2^m \; l!} \\
-	\ctilde &:= {(l-m)! \over (l+m)!}
-}
-and $n!
+Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Jacobi polynomial: \bstodo{Add citation for associated legendre polys}
+\bseqn{
+	\Plm(x) &:= (-1)^m (1-x^2)^\frac{m}{2} \dmdxm{x}{m} P_l(x) = (-1)^m \; {(l+m)! \over 2^m \; l!} \; (1-x^2)^\frac{m}{2} \; P^{(m,m)}_{l-m}(x),
+}
+for $m = 0,1,2,\dots,l$. It is also standard to include a further definition for the case of $m$ negative -- that is we also define:
+\bseqn{
+	\Plm(x) &:= (-1)^m \; {(l+m)! \over (l-m)!} \; P_l^{\abs{m}}(x) = {(l+m)! \over 2^{\abs{m}} \; l!} \; (1-x^2)^{\abs{m} \over 2} \; \jaclmabs{l}{m}(x)
+}
+for $m = -l, \dots, -1$.
+Here, $n!
 := n \: (n-1) \: (n-2) \dots 1$ for $n \in \N_0$ is the standard factorial, with $0!
 = 1$.
+Although referred to as the associated Legendre \enquote{polynomials}, they are not strictly polynomials when $m$ is odd, and as such some authors refer to them as associated Legendre functions instead.
+Moreover, it should be noted too that the inclusion of the $(-1)^m$ factor in the definition for $m > 0$, known as the \textit{Condon-Shortley phase} in physics, is sometimes omitted.
+
 These relationships for the Jacobi polynomials and associated Legendre polynomials allow us to explicitly see where the normalising constants that we shall be using for our definition of the spherical harmonics come from.
@@
 \bseqnnumber{
-	\clm := \Bigg({(2l+1) \: (l-m)! \over 4 \pi \: (l+m)!}\Bigg)^\frac{1}{2} \:
-		\begin{cases} 
-			(-1)^m \: \chat &\text{if } m \ge 0 \\
-			\chatmod &\text{if } m < 0
+	\clm := \Bigg({(2l+1) \: (l-m)! \over 4\pi \: (l+m)!}\Bigg)^\half \: {(l+m)! \over 2^{\abs{m}} \; l!} \:
+		\begin{cases} 
+			(-1)^m &\text{if } m \ge 0 \\
+			1 &\text{if } m < 0
 		\end{cases}. \label{eqn:clmdef}
@@
 			& \ddots & \ddots & \ddots & \\
-			& & 0 &  \gamma_{l,-l,2} & 0 \\
+			& & 0 &  \gamma_{l,l,2} & 0 \\
 		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
@@
 
-Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices in the following way. For $l \in \N$, we set
+Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices in the following way. For $l \in \N$, we can set
 \bseqnnumber{
@@
 		\begin{pmatrix}
-			\hat{A}^{x,y}_l & \zeromat{(2l+3)\times(2l+1)}
-		\end{pmatrix} \in \R^{(2l+3)\times3(2l+1)} \label{eqn:Dltdef}
-}
-where $\zeromat{(2l+3)\times(2l+1)}$ the zero matrix in $\R^{(2l+3)\times(2l+1)}$, and $\hat{A}^{x,y}_l \in \R^{(2l+3)\times2(2l+1)}$ is the left inverse of the matrix $\begin{pmatrix} A^x_l \\ A^y_l \end{pmatrix}$, given by
-\bseqn{
-	\hat{A}^{x,y}_l = 
-		\begin{pmatrix}
-			\frac{1}{2\alpha_{l,-l,3}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,-l,3}} & 0 & \hdots & 0 \\
-			& \ddots & & & & \ddots & & \\
-			& & \ddots & & & & \ddots & \\
-			& & & \frac{1}{2\alpha_{l,l,3}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,-l,3}} \\
-			0 & \hdots & \frac{1}{2\alpha_{l,l-1,4}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,l-1,4}} & 0 \\
-			0 & \hdots & 0 & \frac{1}{2\alpha_{l,l,4}} & 0 & \hdots & 0 & \frac{1}{2\beta_{l,l,4}}
-		\end{pmatrix}.
-}
-For $l = 0$ we set
+			& & & & & \vec{\eta}^\top_{0} \\
+			0 & & & 0 & & & {1 \over \gamma_{l,-l,2}} & \\
+			& \ddots & & & \ddots & & & \ddots \\
+			& & 0 & & & 0 & & & {1 \over \gamma_{l,l,2}} \\
+			& & & & & \vec{\eta}^\top_{1} 
+		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)}, \label{eqn:Dltdef}
+}
+where $\vec{\eta}_{0}, \vec{\eta}_{1} \in \R^{3(2n+1)}$ are vectors with entries given by 
+\bseqn{
+	\big(\vec{\eta}_{0}\big)_j &= 
+		\begin{cases}
+			{1 \over \alpha_{l,-l,3}} & j = 1 \\
+			{- \; \alpha_{l,-l,4} \over \alpha_{l,-l,3} \: \gamma_{l,1-l,2}} & j = 2(2n+1) + 2 \\
+			0 & o/w
+		\end{cases}, \\
+	\big(\vec{\eta}_{1}\big)_j &= 
+		\begin{cases}
+			{1 \over \alpha_{l,l,4}} & j = 2l+1 \\
+			{- \; \alpha_{l,l,3} \over \alpha_{l,l,4} \: \gamma_{l,l-1,2}} & j = 3(2n+1) - 1 \\
+			0 & o/w
+		\end{cases}.
+}
+For $l = 0$ we can simply set
 \bseqnnumber{
@@
 		\begin{pmatrix}
-			\frac{1}{2\alpha_{0,0,3}} & \frac{1}{2\beta_{0,0,3}} & 0 \\
-			0 & 0 & \frac{1}{\gamma_{0,0,2}} \\
-			\frac{1}{2\alpha_{0,0,4}}&  \frac{1}{2\beta_{0,0,4}} & 0 
+			{\beta_{0,0,4} \over \alpha_{0,0,3} \beta_{0,0,4} - \alpha_{0,0,4} \beta_{0,0,3}} & {- \alpha_{0,0,4} \over \alpha_{0,0,3} \beta_{0,0,4} - \alpha_{0,0,4} \beta_{0,0,3}} & 0 \\
+			0 & 0 & {1 \over \gamma_{0,0,2}} \\
+			{\beta_{0,0,3} \over \alpha_{0,0,4} \beta_{0,0,3} - \alpha_{0,0,3} \beta_{0,0,4}} & {- \alpha_{0,0,3} \over \alpha_{0,0,4} \beta_{0,0,3} - \alpha_{0,0,3} \beta_{0,0,4}} & 0
 		\end{pmatrix}. \label{eqn:Dzerotdef}
 }
-
 It will be useful for us to give a formal name for these coefficient matrices above for a family of multidimensional orthogonal polynomials.
@@
 
-\bstodoinline{obtaining coeffs/ transform}
+\bstodoinline{Is this para enough? coeffs/ transform}
 In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
@@
 \bseqn{
-	v(x,y,z) f(x,y,z) = v(x,y,z) \fvec^\top \SHvec(x,y,z) = (V \; \fvec)^\top \SHvec(x,y,z),
-}
-i.e. $V \; \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto v(x,y,z) f(x,y,z)$ in the spherical harmonic basis.
+	v(x,y,z) f(x,y,z) = v(x,y,z) \fvec^\top \SHvec(x,y,z) = (V \fvec)^\top \SHvec(x,y,z),
+}
+i.e. $V \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto v(x,y,z) f(x,y,z)$ in the spherical harmonic basis.
 
@@
 %%% TANGENT SPACE ----- VECTOR SHs
-\section{Vector spherical harmonics as orthogonal vectors in three variables}
+\section{Vector spherical harmonics as orthogonal vectors in three variables}\label{Section:VSHs}
 
@@
 
-As for the scalar case, these recurrences above lead us to Jacobi matrices for the tangent space basis. Define $\VSHvecfull$ by
+As for the scalar case, these coefficients detailed above are then the nonzero entries of the Jacobi matrices for the tangent space basis. Define $\VSHvecfull$ by
 \bseqnnumber{
@@
 		\begin{pmatrix}
-			\mathcal{A}_{l,-l,5} & 0_2 & \mathcal{A}_{l,-l,6} & & \\
+			\mathcal{A}_{l,-l,3} & 0_2 & \mathcal{A}_{l,-l,4} & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & \mathcal{A}_{l,l,5} & 0_2 & \mathcal{A}_{l,l,6} \\
+			& & \mathcal{A}_{l,l,3} & 0_2 & \mathcal{A}_{l,l,4} \\
 		\end{pmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
@@
 		\begin{pmatrix}
-			0_2 & \mathcal{A}_{l,-l,4} & \\
-			\mathcal{A}_{l,-l+1,3} & \ddots & \ddots \\
-			& \ddots & \ddots & \mathcal{A}_{l,l-1,4} \\
-			& & \mathcal{A}_{l,l,3} & 0_2
+			0_2 & \mathcal{A}_{l,-l,6} & \\
+			\mathcal{A}_{l,-l+1,5} & \ddots & \ddots \\
+			& \ddots & \ddots & \mathcal{A}_{l,l-1,6} \\
+			& & \mathcal{A}_{l,l,5} & 0_2
 		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
@@
 			A_{l,m,j} & 0 \\
-			0 & A_{l,m,j}^*
-		\end{pmatrix} \: \text{for } j = 1, 2, 5, 6, \quad
+			0 & A_{l,m,j}
+		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad
 	\mathcal{A}_{l,m,j} := 
@@
 			0 & A_{l,m,j} \\
-			A_{l,m,j}^* & 0
-		\end{pmatrix} \: \text{for } j = 3,4.
+			- A_{l,m,j} & 0
+		\end{pmatrix} \: \text{for } j = 5, 6.
 }
@@
 		\begin{pmatrix}
-			\mathcal{B}_{l,-l,5} & 0_2 & \mathcal{B}_{l,-l,6} & & \\
+			\mathcal{B}_{l,-l,3} & 0_2 & \mathcal{B}_{l,-l,4} & & \\
 			& \ddots & \ddots & \ddots & \\
-			& & \mathcal{B}_{l,l,5} & 0_2 & \mathcal{B}_{l,l,6} \\
+			& & \mathcal{B}_{l,l,3} & 0_2 & \mathcal{B}_{l,l,4} \\
 		\end{pmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
@@
 		\begin{pmatrix}
-			0_2 & \mathcal{B}_{l,-l,4} & \\
-			\mathcal{B}_{l,-l+1,3} & \ddots & \ddots \\
-			& \ddots & \ddots & \mathcal{B}_{l,l-1,4} \\
-			& & \mathcal{B}_{l,l,3} & 0_2
+			0_2 & \mathcal{B}_{l,-l,6} & \\
+			\mathcal{B}_{l,-l+1,5} & \ddots & \ddots \\
+			& \ddots & \ddots & \mathcal{B}_{l,l-1,6} \\
+			& & \mathcal{B}_{l,l,5} & 0_2
 		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
@@
 			B_{l,m,j} & 0 \\
-			0 & B_{l,m,j}^*
-		\end{pmatrix} \: \text{for } j = 1, 2, 5, 6, \quad
+			0 & B_{l,m,j}
+		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad
 	\mathcal{B}_{l,m,j} := 
@@
 			0 & B_{l,m,j} \\
-			B_{l,m,j}^* & 0
-		\end{pmatrix} \: \text{for } j = 3,4.
+			- B_{l,m,j} & 0
+		\end{pmatrix} \: \text{for } j = 5, 6.
 }
@@
 		\begin{pmatrix}
-			0 & 0 & \Gamma_{l,-l,3} \\
-			& & & \Gamma_{l,-l,3}^* \\
+			0 & 0 & \Gamma_{l,-l,2} \\
+			& & & \Gamma_{l,-l,2} \\
 			& & & & \ddots \\
-			& & & & & \Gamma_{l,l,3} \\
-			& & & & & & \Gamma_{l,l,3}^*& 0 & 0 \\
+			& & & & & \Gamma_{l,l,2} \\
+			& & & & & & \Gamma_{l,l,2}& 0 & 0 \\
 		\end{pmatrix} \in \R^{2(2l+1)\times2(2l+3)}, \\
@@
 		\begin{pmatrix}
-			0 & \Gamma_{l,-l,2} \\
-			\Gamma_{l,-l,2}^* & 0 \\
+			0 & \Gamma_{l,-l,3} \\
+			- \Gamma_{l,-l,3} & 0 \\
 			& & \ddots \\
-			& & & 0 & \Gamma_{l,l,2} \\
-			& & & \Gamma_{l,l,2}^* & 0 \\
+			& & & 0 & \Gamma_{l,l,3} \\
+			& & & - \Gamma_{l,l,3} & 0 \\
 		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
@@
 			\Gamma_{l,-l+1,1} \\
-			& \Gamma_{l,-l+1,1}^* \\
+			& \Gamma_{l,-l+1,1} \\
 			& & \ddots \\
 			& & & \Gamma_{l,l-1,1} \\
-			& & & & \Gamma_{l,l-1,1}^* \\
+			& & & & \Gamma_{l,l-1,1} \\
 			& & & & 0 \\
@@
 		\begin{pmatrix}
-			\gradY_1^{-1}(x,y,z) \\ \gradpY_1^{-1}(x,y,z) \\ \gradY_1^{0}(x,y,z) \\ \gradpY_1^{0}(x,y,z) \\ \gradY_1^{1}(x,y,z) \\ \gradpY_1^{1}(x,y,z) \\
+			\gradY_1^{-1}(x,y,z)^\top \\ \gradpY_1^{-1}(x,y,z)^\top \\ \gradY_1^{0}(x,y,z)^\top \\ \gradpY_1^{0}(x,y,z)^\top \\ \gradY_1^{1}(x,y,z)^\top \\ \gradpY_1^{1}(x,y,z)^\top \\
 			\end{pmatrix}, 
@@
 }
-For each $l \in \N$ let $\Dlt$ be any matrix that is a left inverse of $A_l$, i.e.
+Just like for the scalar case, for each $l = 0,1,2\dots$ let $\Dlt$ be any matrix that is a left inverse of $A_l$, i.e.
 such that $\Dlt A_l = \idmat{2(2l+3)}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal} which can be expanded to obtain the three-term recurrence:
-\bseqn{
-	\VSHvecfull_{l+1}(x,y,z) = -\Dlt [B_l-G_l(x,y,z)] \VSHvecfull_l(x,y,z) - \Dlt C_l  \, \VSHvecfull_{l-1}(x,y,z), \quad l \in \N
-}
-where of course we have that $\VSHvec(x,y,z) \equiv \vec{0}_{2\times3}$.
-
-We note that we can choose the matrices $\Dlt$ in the following way. For $l \in \N$, we set
-\begin{align}
-\Dlt = \begin{pmatrix}
-		\hat{A}^{x,y}_l & \zeromat{2(2l+3)\times2(2l+1)}
-	  \end{pmatrix} \in \R^{2(2l+3)\times6(2l+1)}
-\end{align}
-where $\zeromat{2(2l+3)\times2(2l+1)}$ the zero matrix in $\R^{2(2l+3)\times2(2l+1)}$, and $\hat{A}^{x,y}_l \in \R^{2(2l+3)\times4(2l+1)}$ is the left inverse of the matrix $\begin{pmatrix} A^x_l \\ A^y_l \end{pmatrix}$, given by
-\begin{align}
-\hat{A}^{x,y}_l = \begin{pmatrix}
-		\frac{1}{2d_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2d_{l,-l}} & 0 & \hdots &  \\
-		0 & \frac{1}{2d^\perp_{l,-l}} & 0 & \hdots & 0 & -\frac{i}{2d^\perp_{l,-l}} & 0 & \hdots &  \\
-		& & \ddots & & & & \ddots & & \\
-		\vdots & & & \ddots & & & & \ddots & \\
-		& & & & \frac{1}{2d_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2d_{l,l}} & \\
-		& & & & & \frac{1}{2d^\perp_{l,l}} & 0 & \hdots & 0 & -\frac{i}{2d^\perp_{l,l}} \\
-		0 & \hdots & \frac{1}{2a_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2a_{l,l-1}} & 0 & 0 & 0 \\
-		0 & \hdots & 0 & \frac{1}{2a^\perp_{l,l-1}} & 0 & \hdots & 0 & \frac{i}{2a^\perp_{l,l-1}} & 0 & 0\\
-		0 & \hdots & \hdots & 0 & \frac{1}{2a_{l,l}} & 0 & \hdots & 0 & \frac{i}{2a_{l,l}} & 0 \\
-		0 & \hdots & \hdots & & 0 & \frac{1}{2a^\perp_{l,l}} & 0 & \hdots & 0 & \frac{i}{2a^\perp_{l,l}} \\
-	  \end{pmatrix}.
-\end{align}
-\bstodoinline{Derive Dlts again}
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+\bseqn{
+	\begin{cases}
+		\VSHvecfull_{0}(x,y,z) := 0_{2,3} \\
+		\VSHvecfull_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z), \quad l \in \N.
+	\end{cases}
+}
+For $l \in \N$, we can set
+\bseqnnumber{
+	\Dlt = 
+		\begin{pmatrix}
+			& \vec{\eta}^\top_{0} & \\
+			& \vec{\eta}^\top_{1} & \\
+			0_{2(2l+1)} & 0_{2(2l+1)} & \quad \calD_l \quad \\
+			& \vec{\eta}^\top_{2} & \\
+			& \vec{\eta}^\top_{3} &
+		\end{pmatrix} \in \R^{2(2l+3)\times6(2l+1)}, \label{eqn:VSHDltdef}
+}
+where $\calD_l \in \R^{2(2l+1) \times 2(2l+1)}$ is the matrix
+\bseqn{
+	\calD_l := 
+		\begin{pmatrix}
+			{1 \over \Gamma_{l,-l,2}} \\
+			& {1 \over \Gamma_{l,-l,2}} \\
+			& & \ddots \\
+			& & & {1 \over \Gamma_{l,l,2}} \\
+			& & & & {1 \over \Gamma_{l,l,2}}
+		\end{pmatrix}
+}
+and where $\vec{\eta}_{0}, \vec{\eta}_{1}, \vec{\eta}_{2}, \vec{\eta}_{3} \in \R^{6(2l+1)}$ are vectors with entries given by 
+\bseqn{
+	\big(\vec{\eta}_{0}\big)_j &= 
+		\begin{cases}
+			{1 \over A_{l,-l,3}} & j = 1 \\
+			{- \; A_{l,-l,4} \over A_{l,-l,3} \: \Gamma_{l,1-l,2}} & j = 4(2n+1) + 3 \\
+			0 & o/w
+		\end{cases}, \\
+	\big(\vec{\eta}_{1}\big)_j &= 
+		\begin{cases}
+			\big(\vec{\eta}_{0}\big)_j & j = 2, 4(2l+1) + 4 \\
+			0 & o/w
+		\end{cases}, \\
+	\big(\vec{\eta}_{2}\big)_j &= 
+		\begin{cases}
+			{1 \over A_{l,l,4}} & j = 2(2l+1) - 1 \\
+			{- \; A_{l,l,3} \over A_{l,l,4} \: \Gamma_{l,l-1,2}} & j = 6(2l+1) - 5 \\
+			0 & o/w
+		\end{cases}, \\
+	\big(\vec{\eta}_{3}\big)_j &= 
+		\begin{cases}
+			\big(\vec{\eta}_{2}\big)_j & j = 2(2l+1), 6(2l+1) - 4 \\
+			0 & o/w
+		\end{cases}.
+}
+For $l = 0$ we can simply set
+\bseqnnumber{
+	D^T_0 = 
+		\begin{pmatrix}
+			\delta_1 & 0 & \delta_2 & 0 & 0 & 0 \\
+			0 & \delta_1 & 0 & \delta_2 & 0 & 0 \\
+			0 & 0 & 0 & 0 & {1 \over \Gamma_{0,0,2}} & 0 \\
+			0 & 0 & 0 & 0 & 0 & {1 \over \Gamma_{0,0,2}} \\
+			\delta_3 & 0 & \delta_4 & 0 & 0 & 0 \\
+			0 & \delta_3 & 0 & \delta_4 & 0 & 0
+		\end{pmatrix}. \label{eqn:VSHDzerotdef}
+}
+where
+\bseqn{
+	\delta_1 = {B_{0,0,4} \over A_{0,0,3} B_{0,0,4} - A_{0,0,4} B_{0,0,3}}, \quad \delta_2 = {- A_{0,0,4} \over A_{0,0,3} B_{0,0,4} - A_{0,0,4} B_{0,0,3}}, \\
+	\delta_3 = {B_{0,0,3} \over A_{0,0,4} B_{0,0,3} - A_{0,0,3} B_{0,0,4}}, \quad \delta_4 = {- A_{0,0,3} \over A_{0,0,4} B_{0,0,3} - A_{0,0,3} B_{0,0,4}}.
+}
 
@@
 \bseqn{
-	\uvec(x,y,z) \approx \VSHvecfull(x,y,z)^\top \uvec^c := \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big].
+	\uvec(x,y,z) \approx \VSHvecfull(x,y,z)^\top \uvec^c := \sum^N_{l=0} \sum^l_{m=-l} \big[u^{\gradY}_{l,m} \, \gradYlm(x,y,z)  + u^{\gradpY}_{l,m} \, \gradpYlm(x,y,z) \big].
 }
@@
 \subsubsection{Obtaining coefficients}
-The development of vector spherical harmonic transforms is well established \cite{swarztrauber1993vector}, and continues today (see \cite{gia2019favest}). \bstodoinline{VSH Transform}
+\bstodoinline{Do I need to say more here? VSH Transform}
+Just like for the scalar case, we would like to be able to obtain the coefficients for the expansion of a vector valued function in the VSH basis up to a given degree.
+To do this, we can use a vector spherical harmonic transform.
+The development of vector spherical harmonic transforms is well established \cite{swarztrauber1993vector}, and continues today (see \cite{gia2019favest}).
 
@@
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^\top = (\mathbf{u}^c_n)^\top - \vec{\xi}_{n+1}^\top \Dnt (B_n - G_n) -  \vec{\xi}_{n+2}^\top D^\top_{n+1} C_{n+1} \\
-	\quad &\text{3) } \text{Output: } \mathbf{u}(x,y,z) \approx \VSHvecfull_1(x,y,z)^\top \vec{\xi}_{1} \\
+	\quad &\text{3) } \text{Output: } \uvec(x,y,z) \approx \VSHvecfull_1(x,y,z)^\top \vec{\xi}_{1} \\
 	&\quad \quad \quad \quad \quad \quad \quad \quad \quad = (\vec{\xi}_{1})_1 \gradY_1^{-1} + (\vec{\xi}_{1})_2 \gradpY_1^{-1} + (\vec{\xi}_{1})_3 \gradY_1^{0} \\
@@
 	\uvec(x,y,z) &\approx \VSHvec(x,y,z)^\top \: \uvecc + \VSHvecperp(x,y,z)^\top \: \uveccperp \\
-	&:=  \sum^N_{l=0} \sum^l_{m=-l} \big[u_{l,m} \, \gradYlm(x,y,z)  + u^\perp_{l,m} \, \gradpYlm(x,y,z) \big], \\
+	&:=  \sum^N_{l=0} \sum^l_{m=-l} \big[u^{\gradY}_{l,m} \, \gradYlm(x,y,z)  + u^{\gradpY}_{l,m} \, \gradpYlm(x,y,z) \big], \\
 	f(x,y,z) &\approx \SHvec(x,y,z)^\top \: \fvecc :=  \sum^N_{l=0} \sum^l_{m=-l} f_{l,m} \, \Ylm(x,y,z)
@@
 	The argument for $\calG$ is trivial by definition.
-Now, the spherical harmonics, as the name eludes to, satisfy a harmonic relationship $\DeltaS \Ylm = \: l(l+1) \Ylm$ \bstodo{cite}.
+Now, the spherical harmonics, as the name eludes to, satisfy a harmonic relationship $\DeltaS \Ylm = \: l(l+1) \Ylm$ (see e.g.
+\cite{humi1987factorisation}).
 Thus, for the proof for $\calD$, consider
 \bseqn{
-	\divergence \uvec &= \divergence \Big( \sum_{l=0}^N \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ] \Big) \\
-	&= \sum_{l=0}^N \sum_{m=-l}^{l} u_{l,m} \: \DeltaS \Ylm \\
-	&= \sum_{l=0}^N \sum_{m=-l}^{l} - u_{l,m} \: l(l+1) \Ylm,
+	\divergence \uvec &= \divergence \Big( \sum_{l=0}^N \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \Big) \\
+	&= \sum_{l=0}^N \sum_{m=-l}^{l} u^{\gradY}_{l,m} \: \DeltaS \Ylm \\
+	&= \sum_{l=0}^N \sum_{m=-l}^{l} - u^{\gradY}_{l,m} \: l(l+1) \Ylm,
 }
@@
 \bseqn{
-	f \: \rvec \times \uvec(x,y,z) &= \VSHvec(x,y,z)^\top \: \calF \: \uvec^{\perp c} + \Big(\VSHvecperp(x,y,z)\Big)^\top \: \calF^\perp \: \uvec^{c}
+	f \: \rvec \times \uvec(x,y,z) &= \VSHvec(x,y,z)^\top \: \calF \: \uveccperp + \Big(\VSHvecperp(x,y,z)\Big)^\top \: \calF^\perp \: \uvecc
 }
@@
 \bseqn{
-	f \: \rvec \times \bold{u} &= f \: \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} \gradYlm + u^{\perp}_{l,m} \gradpYlm ] \\
-	&= 2 \: \rotrate \: z \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} \rvec \times \gradYlm + u^{\perp}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
-	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u_{l,m} z \: \gradpYlm - u^{\perp}_{l,m} z \: \gradYlm ) ] \\
-	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} \Big[ u_{l,m} \: \Big( \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l} \Big) \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad - u^{\perp}_{l,m} \: \Big( \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l} \Big) \Big].
+	f \: \rvec \times \bold{u} &= f \: \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
+	&= 2 \: \rotrate \: z \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \rvec \times \gradYlm + u^{\gradpY}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
+	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} z \: \gradpYlm - u^{\gradpY}_{l,m} z \: \gradYlm ) ] \\
+	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} \Big[ u^{\gradY}_{l,m} \: \Big( \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l} \Big) \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad - u^{\gradpY}_{l,m} \: \Big( \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l} \Big) \Big].
 }
@@
 \bseqn{
-	\uvec^c_{n+1} &= \uvec^c_{n} + \Delta t \: (\calG \vec{h}^c_{n+1} - \calF \uvec^{\perp c}_{n+1}) \\
-	\uvec^{\perp c}_{n+1} &=  - \Delta t \: \calF^\perp \uvec^{c}_{n+1} \\
-	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvec^c_{n+1}
+	\uvecc_{n+1} &= \uvecnc + \Delta t \: (\calG \vec{h}^c_{n+1} - \calF \uveccperp_{n+1}) \\
+	\uveccperp_{n+1} &=  - \Delta t \: \calF^\perp \uvecc_{n+1} \\
+	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvecc_{n+1}
 }
@@
 \bseqn{
-	\uvec^c_{n+1} &= \uvec^c_{n} + \Delta t \: (\calG \vec{h}^c_n - \Delta t \: \calH \calG \calD \uvec^c_{n+1} + \Delta t \calF \calF^\perp \uvec^{c}_{n+1}) \\
-	\uvec^{\perp c}_{n+1} &=  - \Delta t \: \calF^\perp \uvec^{c}_{n+1} \\
-	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvec^c_{n+1}
+	\uvecc_{n+1} &= \uvecc_{n} + \Delta t \: (\calG \vec{h}^c_n - \Delta t \: \calH \calG \calD \uvecc_{n+1} + \Delta t \calF \calF^\perp \uvecc_{n+1}) \\
+	\uveccperp_{n+1} &=  - \Delta t \: \calF^\perp \uvecc_{n+1} \\
+	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvecc_{n+1}
 }
@@
 \bseqn{
-	\uvec^c_{n+1} &= (I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp)^{-1} \: (\uvec^c_{n} + \Delta t \: \calG \vec{h}^c_n) \\
-	\uvec^{\perp c}_{n+1} &=  - \Delta t \: \calF^\perp \uvec^{c}_{n+1} \\
-	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH \calD \uvec^c_{n+1}.
+	\uvecc_{n+1} &= (I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp)^{-1} \: (\uvecc_{n} + \Delta t \: \calG \vec{h}^c_n) \\
+	\uveccperp_{n+1} &=  - \Delta t \: \calF^\perp \uvecc_{n+1} \\
+	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH \calD \uvecc_{n+1}.
 }
 Here, it is clear that the matrix $I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
-Doing so at each iteration provides us with the coefficients vectors $\uvec^{c}_{n+1}, \uvec^{\perp c}_{n+1}, \vec{h}^{c}_{n+1}$ if we know $\uvec^{c}_{n}, \uvec^{\perp c}_{n}, \vec{h}^{c}_{n}$.
+Doing so at each iteration provides us with the coefficients vectors $\uvecc_{n+1}, \uveccperp_{n+1}, \vec{h}^{c}_{n+1}$ if we know $\uvecc_{n}, \uveccperp_{n}, \vec{h}^{c}_{n}$.
 

### 2021-03-17 papers/thesis/full/ch1-introduction/introduction.tex
Edit to intro, and comments

@@
 
-Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs}.
+Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs} \bstodo{citation here}.
 While there are many famous examples -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few -- the area of multivariate orthogonal polynomials has a smaller array of research.

### 2021-03-17 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Edit to intro, and comments

@@
 }
-Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Jacobi polynomial: \bstodo{Add citation for associated legendre polys}
+Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Jacobi polynomial: \bstodo{Sheehan - do you know what citation I can use for the definition of the Associated Legendre Polynomials (as the mth derivative of the ordinary Legendre Polynomials)?}
 \bseqn{
@@
 
-\bstodoinline{Is this para enough? coeffs/ transform}
+\bstodoinline{Is this paragraph enough? Or do I need to put more detail into explaining these methods, or derive my own?}
 In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
@@
 \subsubsection{Obtaining coefficients}
-\bstodoinline{Do I need to say more here? VSH Transform}
+\bstodoinline{As for the scalar case, do I need to say more here about the transforms?}
 Just like for the scalar case, we would like to be able to obtain the coefficients for the expansion of a vector valued function in the VSH basis up to a given degree.
@@
 
-\bstodo{Image}
+\bstodo{It would be nice to have some images here. However I need to update/fix/tidy my SH code as its probably got a few bugs in it having now gone back over the derivations that I have done.}

### 2021-04-09 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
edits to SH

@@
 
-\subsection{Three-term recurrence relation for $\VSHvecfull(x,y,z)$}
+\subsection{Three-term recurrence relation for $\VSHvecfull(x,y,z)$}\label{subsection:VSHrecurrence}
 
@@
 		\VSHvecfull_{0}(x,y,z) := 0_{2,3} \\
-		\VSHvecfull_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z), \quad l \in \N.
+		\VSHvecfull_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \VSHvecfull_l(x,y,z) - \Dlt C_l  \, \VSHvecfull_{l-1}(x,y,z), \quad l \in \N.
 	\end{cases}
@@
 
+It will be useful to be able to describe the coefficients of a vector-valued function on the tangent space of $\Omega$ in multiple ways, and as such we will introduce some more notation for the \enquote{splitting} of the basis vectors into the two groups stemming from the gradient and perpendicular gradient directions.
+Let's explain what we mean by this.
 Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
@@
 Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent space of $\Omega$}.
-We can then approximate this function by its expansion in the VSH basis, i.e.
-for some large enough $N \in \N$:
-\bseqn{
-	\uvec(x,y,z) \approx \VSHvecfull(x,y,z)^\top \uvec^c := \sum^N_{l=0} \sum^l_{m=-l} \big[u^{\gradY}_{l,m} \, \gradYlm(x,y,z)  + u^{\gradpY}_{l,m} \, \gradpYlm(x,y,z) \big].
-}
+
+Define $\VSHvec, \VSHvecperp$ by
+\bseqnnumber{
+	\VSHvec &:= 
+		\begin{pmatrix} 
+			\VSHvec_0 \\ \VSHvec_1 \\ \vdots 
+		\end{pmatrix}, \quad \text{where } \quad 
+	\VSHvecl := 
+		\begin{pmatrix} 
+			(\gradY^{-l}_l)^\top \\ \vdots \\ (\gradY^{l}_l)^\top
+		\end{pmatrix} \in \C^{(2l+1) \times 3} \quad \forall \: l \in \No, \label{eqn:VSHdef} \\
+	\VSHvecperp &:= 
+		\begin{pmatrix} 
+			\VSHvecperp_0 \\ \VSHvecperp_1 \\ \vdots 
+		\end{pmatrix}, \quad \text{where } \quad 
+	\VSHvecperpl := 
+		\begin{pmatrix} 
+			(\gradpY^{-l}_l)^\top \\ \vdots \\ (\gradpY^{l}_l)^\top
+		\end{pmatrix} \in \C^{(2l+1) \times 3} \quad \forall \: l \in \No. \label{eqn:VSHperpdef}
+}
+If $\uvec$ is a vector-valued function on the tangent space of $\Omega$, it will have the coefficients vectors $\uvec^c, \uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, defined such that
+\bseqn{
+	\uvec(x,y,z) &\approx \VSHvecfull(x,y,z)^\top \: \uvec^c \\
+	&\equiv \VSHvec(x,y,z)^\top \: \uvecc + \VSHvecperp(x,y,z)^\top \: \uveccperp \\
+	&:=  \sum^N_{l=0} \sum^l_{m=-l} \big[u^{\gradY}_{l,m} \, \gradYlm(x,y,z)  + u^{\gradpY}_{l,m} \, \gradpYlm(x,y,z) \big].
+}
+Of course, just like for the scalar case, we can use transforms to compute the coefficients of such a known function, and further evaluate, and apply differential operators to, a function knowing just its coefficients.
 
@@
 \bstodoinline{As for the scalar case, do I need to say more here about the transforms?}
-Just like for the scalar case, we would like to be able to obtain the coefficients for the expansion of a vector valued function in the VSH basis up to a given degree.
+As eluded to, we would like to be able to obtain the coefficients for the expansion of a vector valued function in the VSH basis up to a given degree, similar to how we approach the scalar case.
 To do this, we can use a vector spherical harmonic transform.
@@
 \subsubsection{Function evaluation}
-The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$:
+The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$, where the matrices $C_n, B_n, \Dnt, G_n(x,y,z)$ are all defined in \bsrefsection{subsection:VSHrecurrence}:
 \bseqn{
@@
 	\quad &\text{2) } \text{For } n = N:-1:1 \\
-	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^\top = (\mathbf{u}^c_n)^\top - \vec{\xi}_{n+1}^\top \Dnt (B_n - G_n) -  \vec{\xi}_{n+2}^\top D^\top_{n+1} C_{n+1} \\
+	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^\top = (\uvec^c_n)^\top - \vec{\xi}_{n+1}^\top \Dnt \big(B_n - G_n(x,y,z)\big) -  \vec{\xi}_{n+2}^\top D^\top_{n+1} C_{n+1} \\
 	\quad &\text{3) } \text{Output: } \uvec(x,y,z) \approx \VSHvecfull_1(x,y,z)^\top \vec{\xi}_{1} \\
@@
 }
-
-\subsection{Sparse partial differential operators}
+Here, at each iteration, the vector $\uvec^c_n$ is given by the merging of the two order $n$ subvector of $\uvec^c$:
+\bseqn{
+	\uvec^c_n := 
+		\begin{pmatrix} 
+			u^{\gradY}_{n,-n} \\
+			u^{\gradpY}_{n,-n} \\
+			\vdots \\
+			u^{\gradY}_{n,n} \\
+			u^{\gradpY}_{n,n} \\
+		\end{pmatrix}.
+}
+Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs in our tangent space basis (and hence coefficients) for degree $n=0$ are zero.
+
+\subsection{Sparse partial differential operators}\label{section:SHdiffoperators}
 
@@
 
-In order to ensure we fully capitalise on the natural sparsity of the relationships so that we yield banded-block-banded operators, we will \enquote{split} coefficient vectors of a vector-valued function on the tangent space of $\Omega$ into two.
-On that note, define $\VSHvec, \VSHvecperp$ by
-\bseqnnumber{
-	\VSHvec &:= 
-		\begin{pmatrix} 
-			\VSHvec_0 \\ \VSHvec_1 \\ \vdots 
-		\end{pmatrix}, \quad \text{where } \quad 
-	\VSHvecl := 
-		\begin{pmatrix} 
-			(\gradY^{-l}_l)^\top \\ \vdots \\ (\gradY^{l}_l)^\top
-		\end{pmatrix} \in \C^{(2l+1) \times 3} \quad \forall \: l \in \No, \label{eqn:VSHdef} \\
-	\VSHvecperp &:= 
-		\begin{pmatrix} 
-			\VSHvecperp_0 \\ \VSHvecperp_1 \\ \vdots 
-		\end{pmatrix}, \quad \text{where } \quad 
-	\VSHvecperpl := 
-		\begin{pmatrix} 
-			(\gradpY^{-l}_l)^\top \\ \vdots \\ (\gradpY^{l}_l)^\top
-		\end{pmatrix} \in \C^{(2l+1) \times 3} \quad \forall \: l \in \No. \label{eqn:VSHperpdef}
-}
 Let $\uvec$ be a vector-valued function on the tangent space of $\Omega$ with coefficients vectors $\uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, and let $f$ be a scalar function on $\Omega$ with coefficients vector $\fvecc$ for its expansion in the SSH basis also up to order $N$, i.e.
 \bseqn{
-	\uvec(x,y,z) &\approx \VSHvec(x,y,z)^\top \: \uvecc + \VSHvecperp(x,y,z)^\top \: \uveccperp \\
+	\uvec(x,y,z) &\approx \VSHvecfull(x,y,z)^\top \: \uvec^c \\
+	&\equiv \VSHvec(x,y,z)^\top \: \uvecc + \VSHvecperp(x,y,z)^\top \: \uveccperp \\
 	&:=  \sum^N_{l=0} \sum^l_{m=-l} \big[u^{\gradY}_{l,m} \, \gradYlm(x,y,z)  + u^{\gradpY}_{l,m} \, \gradpYlm(x,y,z) \big], \\
-	f(x,y,z) &\approx \SHvec(x,y,z)^\top \: \fvecc :=  \sum^N_{l=0} \sum^l_{m=-l} f_{l,m} \, \Ylm(x,y,z)
-}
+	f(x,y,z) &\approx \SHvec(x,y,z)^\top \: \fvecc :=  \sum^N_{l=0} \sum^l_{m=-l} f_{l,m} \, \Ylm(x,y,z).
+}
+In order to fully exploit the sparsity of the relationships and ensure we are using banded-block-banded operators, in problems where we have an interdependency of scalar and vector-valued functions it can be useful to \enquote{expand} the scalar function's coefficients vector $\fvecc$ to be the length of the vector-valued one $\uvec^c$, with zeros being inserted after each entry, i.e.
+define the \textit{expanded coefficients for a scalar function on $\Omega$ in the SH basis} by
+\bseqn{
+	\fveccexp :=
+		\begin{pmatrix}
+			\fveccexp_0 \\
+			\vdots \\
+			\fveccexp_N
+		\end{pmatrix} \in \R^{2(N+1)^2}, \quad
+	\fveccexp_l := 
+		\begin{pmatrix}
+			f_{l,-l} \\
+			0 \\
+			f_{l,-l+1} \\
+			0 \\
+			\vdots \\
+		 	f_{l,l} \\
+			0
+		\end{pmatrix} \in \R^{4(l+1)}.
+}
+To help with the definitions of the operators we will discuss, let $\calE_s, \calE_v$ be (severely) rectangular matrices defined such that $\calE_s \fveccexp = \fvecc$ and $\calE_v \fvecc = \fveccexp$.
 We are now in a position to derive the sparse differential operators that can be applied to the functions $\uvec, f$.
@@
 \begin{definition}\label{def:SHoperators}
-	We define the operator matrices $\cal{D}, \calG, \calL$ according to:
+	We define the operator matrices $\calD, \tilde \calD, \calG, \tilde \calG \calL$ according to:
 \bseqn{
 	\divergence \uvec(x,y,z) &= \SHvec(x,y,z)^\top \: \calD \: \uvecc, \\
+	\divergence \uvec(x,y,z) &= \SHvec(x,y,z)^\top \: \calE_s \: \big(\tilde \calD \: \uvec^c), \\
 	\gradS f(x,y,z) &= \VSHvec(x,y,z)^\top \: \calG \: \fvecc, \\
+	\gradS f(x,y,z) &= \VSHvecfull(x,y,z)^\top \: \tilde \calG \: \fveccexp, \\
 	\DeltaS f(x,y,z) &= \SHvec(x,y,z)^\top \: \calL \: \fvecc.
@@
 \end{definition}
+\begin{remark}
+	The operators $\tilde \calD, \tilde \calG$ would be used when vectors of the resulting length are required for the problem you are wishing to solve.
+\end{remark}
 
@@
 \bseqn{
-	\calG &= I_{(N+1)^2}, \quad
-	\calD \equiv \calL := 
+	\calD &\equiv \calL := 
 		\begin{pmatrix}
@@
 			& & \calL_N
-		\end{pmatrix},
-}
-where
-\bseqn{
-	\calL_l &:= 
-		\begin{pmatrix}
-			-l(l+1) & & \\
-			& \ddots & \\
-			& & -l(l+1)
-		\end{pmatrix} \in \R^{(2l+1) \times (2l+1)}.
-}
+		\end{pmatrix} \in \R^{(N+1)^2}, \quad
+	\tilde \calD = 
+		\begin{pmatrix}
+			\calL_0 & & & & \\
+			& 0 & & & \\
+			& & \ddots & & \\
+			& & & \calL_N & \\
+			& & & & 0
+		\end{pmatrix} \in \R^{2(N+1)^2}, \\
+	\calG &= I_{(N+1)^2}, \quad
+	\tilde \calG = 
+		\begin{pmatrix}
+			1 & & & & \\
+			& 0 & & & \\
+			& & \ddots & & \\
+			& & & 1 & \\
+			& & & & 0
+		\end{pmatrix}  \in \R^{2(N+1)^2},
+}
+where $\calL_l := -l(l+1) \: I_{2l+1}$.
 \end{theorem}
@@
 \begin{proof}
-	The argument for $\calG$ is trivial by definition.
-Now, the spherical harmonics, as the name eludes to, satisfy a harmonic relationship $\DeltaS \Ylm = \: l(l+1) \Ylm$ (see e.g.
+	The arguments for $\calG$ and $\tilde \calG$ are trivial by definition.
+The spherical harmonics, as the name eludes to, satisfy a harmonic relationship $\DeltaS \Ylm = \: l(l+1) \Ylm$ (see e.g.
 \cite{humi1987factorisation}).
-Thus, for the proof for $\calD$, consider
+Thus, for the proof for $\calD$ and $\tilde \calD$, consider
 \bseqn{
@@
 	\begin{cases}
-		\pfpx{t}{\uvec} + f \rvec \times \uvec - \nabla h = \mathbf{0} \\
-		\pfpx{t}{h} + \calH \nabla \cdot \uvec = 0
+		\pfpx{t}{\uvec} + f \rvec \times \uvec - \gradS h = \mathbf{0} \\
+		\pfpx{t}{h} + \calH \gradS \cdot \uvec = 0
 	\end{cases}
 }
-where $f = 2 \rotrate \cos(\varphi) =  2 \rotrate z$ is the Coriolis parameter, and $\rotrate = 7.2921 × 10^{-5}$ rad s$^{-1}$ is the rotation rate of the Earth (note that $\varphi = {\pi \over 2} - \text{latitude}$).
-
-A keen eye may notice that we require one more operator not yet defined -- one for the Coriolis parameter cross product (i.e. $f \: \rvec \times \uvec$). 
-\begin{definition}\label{def:coriolisoperator}
-	Define the operator matrices $\calF, \calF^\perp$ according to:
-\bseqn{
-	f \: \rvec \times \uvec(x,y,z) &= \VSHvec(x,y,z)^\top \: \calF \: \uveccperp + \Big(\VSHvecperp(x,y,z)\Big)^\top \: \calF^\perp \: \uvecc
+where $f = 2 \rotrate \cos(\varphi) =  2 \rotrate z$ is the Coriolis parameter, and $\rotrate$ is the rotation rate of the Earth (note that $\varphi = {\pi \over 2} - \text{latitude}$ \bstodo{make a footnote}).
+On a full scale Earth, the rotation rate is $\rotrate = 7.2921 × 10^{-5}$ rad s$^{-1}$ \bstodo{make a footnote}.
+
+A keen eye may notice that -- while we have operators for the multiplication by z, the spherical gradient and divergence -- we require one more operator not yet defined for the unit vector cross product $\rvec \times \uvec$.
+\begin{definition}\label{def:unitveccrossproduct}
+	Define the operator matrix $\calR$ according to:
+\bseqn{
+	\rvec(x,y,z) \times \uvec(x,y,z) &= \VSHvecfull(x,y,z)^\top \: \calR \uvec^c.
 }
@@
 \begin{lemma}
-	The operators $\calF, \calF^\perp$ defined in \bsrefdef{def:coriolisoperator} are sparse with banded-block-banded structure. More specifically,
-\begin{itemize}
-	\item $\calF$ has block-bandwidths $(1,1)$, and sub-block-bandwidths $(1,1)$
-	\item $\calF^\perp$ has block-bandwidths $(1,1)$, and sub-block-bandwidths $(1,1)$
-\end{itemize}
+	The operator $\calR$ defined in \bsrefdef{def:unitveccrossproduct} is sparse with banded-block-banded structure.
+More specifically, $\calR$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(1,1)$ (i.e.
+is tridiagonal).
 \end{lemma}
 \begin{proof}
-	Using \bsreflemma{lemma:VSHrecurrences}, we have that:
-\bseqn{
-	f \: \rvec \times \bold{u} &= f \: \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
-	&= 2 \: \rotrate \: z \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \rvec \times \gradYlm + u^{\gradpY}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
-	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} z \: \gradpYlm - u^{\gradpY}_{l,m} z \: \gradYlm ) ] \\
-	&= 2 \: \rotrate \: \sum_{l=0}^{N} \sum_{m=-l}^{l} \Big[ u^{\gradY}_{l,m} \: \Big( \Gamma_{l,m,1}^* \gradpY^{m}_{l-1} + \Gamma_{l,m,3}^* \gradpY^{m}_{l+1} + \Gamma_{l,m,2} \gradY^{m}_{l} \Big) \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad - u^{\gradpY}_{l,m} \: \Big( \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,3} \gradY^{m}_{l+1} + \Gamma_{l,m,2}^* \gradpY^{m}_{l} \Big) \Big].
-}
-We can then write down the operators $\calF, \calF^\perp$ as follows:
-\bseqn{
-	\calF &= \bar{\calF^\perp} = 2 \: \rotrate \: 
-		\begin{pmatrix}
-			\calF_{0,2} & \calF_{1,1} & & & & \\
-			\calF_{0,3} & \calF_{1,2} & \calF_{2,1} & & & \\
-			& \calF_{1,3} & \calF_{2,2} & \calF_{3,1}  & & & \\
-			& & \calF_{2,3} & \ddots & \ddots & \\
-			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}
-}
-where $\bar{A}$ denotes the conjugate matrix of $A$ (i.e.
-where the entries of $\bar{A}$ are simply the complex conjugates of the entries of $A$) \cite[p.10]{courant1953methods}, and for $l = 0,\dots,N$
-\bseqn{
-	\calF_{l,1} &:= 
-		\begin{pmatrix}
-			0 & \Gamma_{l,-l+1,1} & & \\
-			& & \ddots & & \\
-			& & & \Gamma_{l,l-1,1} & 0 \\
-		\end{pmatrix} \in \R^{(2l-1)\times(2l+1)}, \quad (l \ne 0), \\
-	\calF_{l,2} &:= 
-		\begin{pmatrix}
-			\Gamma_{l,-l,2} & & \\
-			& \ddots & \\
-			& & \Gamma_{l,l,2} \\
-		\end{pmatrix}  \in \R^{(2l+1)\times(2l+1)}, \\
-	\calF_{l,3} &:= 
-		\begin{pmatrix}
-			0 \\
-			\Gamma_{l,-l,3} \\
-			& & \ddots \\
-			& & & & \Gamma_{l,l,3} \\
-			& & & & 0 \\
-		\end{pmatrix} \in \R^{(2l+3)\times(2l+1)}.
+	Using the relations that $\gradYlm = \gradS \Ylm$ and $\gradpYlm = \rvec \times \gradS \Ylm$ for each $l \in \N$, $m \in \Z$ s.t. $-l \le m \le l$, we have that:
+\bseqn{
+	\rvec \times \bold{u} &= \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
+	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \rvec \times \gradYlm + u^{\gradpY}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\
+	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \: \gradpYlm - u^{\gradpY}_{l,m} \: \gradYlm ) ].
+}
+We can then write down the operator $\calR$ as follows:
+\bseqn{
+	\calR &=
+		\begin{pmatrix}
+			R & & & \\
+			& R & & \\
+			& & & \ddots
+		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2},
+}
+where the blocks $R$ are simply given by
+\bseqn{
+	R &:= 
+		\begin{pmatrix}
+			0 & 1 \\
+			-1 & 0
+		\end{pmatrix} \in \R^{2\times2}.
 }
@@
 
-For simplicity, we implement a backward Euler timestepping method to solve the linear SWEs with timestep $\Delta t$:
-\bseqn{
-	\uvec_{n+1} &= \uvec_{n} + \Delta t \: (\nabla h_{n+1} - f \rvec \times \uvec_{n+1}) \\
-	h_{n+1} &= h_n - \Delta t \: \calH \nabla \cdot \uvec_{n+1}
-}
-In coefficient space, this system becomes:
-\bseqn{
-	\uvecc_{n+1} &= \uvecnc + \Delta t \: (\calG \vec{h}^c_{n+1} - \calF \uveccperp_{n+1}) \\
-	\uveccperp_{n+1} &=  - \Delta t \: \calF^\perp \uvecc_{n+1} \\
-	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvecc_{n+1}
+For simplicity, we will implement a backward Euler timestepping method to solve the linear SWEs with timestep $\Delta t$:
+\bseqn{
+	\uvec_{n+1} &= \uvec_{n} + \Delta t \: (\gradS h_{n+1} - f \rvec \times \uvec_{n+1}) \\
+	h_{n+1} &= h_n - \Delta t \: \calH \gradS \cdot \uvec_{n+1}
+}
+Let's write this in coefficient space. Firstly, our coefficients vectors for the functions $\uvec, h$ we will denote $\uvec^c, \vec{h}^c$ -- that is:
+\bseqn{
+	\uvec(x,y,z) &\approx \VSHvecfull(x,y,z)^\top \: \uvec^c \\
+	h(x,y,z) &\approx \SHvec(x,y,z)^\top \: \vec{h}^c.
+}
+As eluded to earlier, we will \enquote{extend} the scalar function's coefficients vector, writing $\tilde{\vec{h}}^c = E_v \vec{h}^c$, as described in \bsrefsection{section:SHdiffoperators}.
+Thus, our system can be written as a matrix-vector system:
+\bseqn{
+	\uvec^c_{n+1} &= \uvec^c_n + \Delta t \: (\tilde \calG \tilde{\vec{h}}^c_{n+1} - 2 \rotrate J_z \calR \uvec^c_{n+1}) \\
+	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}
 }
@@
 \bseqn{
-	\uvecc_{n+1} &= \uvecc_{n} + \Delta t \: (\calG \vec{h}^c_n - \Delta t \: \calH \calG \calD \uvecc_{n+1} + \Delta t \calF \calF^\perp \uvecc_{n+1}) \\
-	\uveccperp_{n+1} &=  - \Delta t \: \calF^\perp \uvecc_{n+1} \\
-	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH D \uvecc_{n+1}
+	\uvec^c_{n+1} &= \uvec^c_n + \Delta t \: (\tilde \calG \tilde{\vec{h}}^c_{n} - \Delta t \: \calH \tilde \calG \tilde \calD \uvecc_{n+1} - 2 \rotrate J_z \calR \uvec^c_{n+1}) \\
+	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}
 }
@@
 \bseqn{
-	\uvecc_{n+1} &= (I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp)^{-1} \: (\uvecc_{n} + \Delta t \: \calG \vec{h}^c_n) \\
-	\uveccperp_{n+1} &=  - \Delta t \: \calF^\perp \uvecc_{n+1} \\
-	\vec{h}^c_{n+1} &= \vec{h}^c_n - \Delta t \: \calH \calD \uvecc_{n+1}.
-}
-Here, it is clear that the matrix $I + \Delta t^2 \calH \calG \calD - \Delta t^2 \calF \calF^\perp$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
-Doing so at each iteration provides us with the coefficients vectors $\uvecc_{n+1}, \uveccperp_{n+1}, \vec{h}^{c}_{n+1}$ if we know $\uvecc_{n}, \uveccperp_{n}, \vec{h}^{c}_{n}$.
-
-\bstodo{It would be nice to have some images here. However I need to update/fix/tidy my SH code as its probably got a few bugs in it having now gone back over the derivations that I have done.}
+	\Big[ I + \Delta t^2 \calH \tilde \calG \tilde \calD - 2 \rotrate \Delta t \: J_z \calR \Big] \: \uvec^c_{n+1} &= \uvec^c_n + \Delta t \tilde \calG \tilde{\vec{h}}^c_{n} \\
+	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}.
+}
+Here, it is clear that the matrix $I + \Delta t^2 \calH \tilde \calG \tilde \calD - 2 \rotrate \Delta t \: J_z \calR$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
+Doing so at each iteration provides us with the coefficients vectors $\uvec^c_{n+1}, \tilde{\vec{h}}^c_{n+1}$ if we know $\uvec^c_{n}, \tilde{\vec{h}}^c_{n}$.
+Of course, when evaluating the function $h$, we can simply gather back the coefficients vector $\vec{h}^c$ by $\vec{h}^c = E_s \tilde{\vec{h}}^c$.
+
+\bstodo{Images}

### 2021-04-15 papers/thesis/full/ch1-introduction/introduction.tex
Improved intro and background. Edits to SH chapter.

@@
 Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs} \bstodo{citation here}.
-While there are many famous examples -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few -- the area of multivariate orthogonal polynomials has a smaller array of research.
+For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations \cite{boyd2001chebyshev, olver2013fast}.
+There are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few \cite[Section 18.3]{DLMF} -- but the area of multivariate orthogonal polynomials has a smaller array of research.
 However, this branch of mathematics has an encouraging future \cite{dunkl2014orthogonal}, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains, such as on the triangle \cite{olver2019triangle} with applications including solving Volterra integral equations \cite{gutleb2020sparse}.
 
-One famous group of multidimensional orthogonal polynomials is of course the Spherical Harmonics.
-The European Centre for Medium-range Weather Forecasts (ECMWF) use a Spherical Harmonic spectral method in their weather and climate model \cite{cheong2006dynamical}.
-Chapter 2 of this thesis provides an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface.
-We present a framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
+For the uninitiated, a spectral method is one that, given a domain, numerically approximates the solution of a (differential) equation as sum of basis functions for said domain, with the coefficients to be found.
+The basis functions are non-zero over the whole domain, meaning the solution is found globally and thus helps such methods to be inherently accurate (providing the solution is smooth).
+By utilising OPs as basis functions, we can develop sparse spectral methods (SSMs), meaning that the naturally sparse relationships between the basis OPs leads to sparse operator matrices that represent the differential operations in the equation to solve.
+Sparse spectral methods for one dimensional problems have been shown to lead to \enquote{almost banded} matrices \cite{olver2013fast}.
+
+In this thesis we expand upon this knowledge by providing frameworks for similar sparse spectral methods for solving PDEs on other multidimensional domains (notably including the disk-slice and trapezium in 2D, and the spherical cap surface in 3D) that also yield matrices that are similarly \enquote{almost banded}.
+The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
+
+Spherical Harmonics are of course one long-established and famous group of multidimensional orthogonal polynomials.
+They are used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
+While the whole sphere spectral method using the Spherical Harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast, diamantakisecmwf}.
+
+This spherical harmonic transform in fact uses two transforms -- a Fourier transform (using the well established Fast Fourier Transform \cite{cooley1965algorithm}) in the longitudinal direction and a Legendre transform in the latitudinal direction -- and it is the Legendre transform that has been identified as inhibiting future performance due to its computational cost.
+While a Fast Legendre Transform (FLT) \cite{wedi2013fast} has been incorporated into the model, along with new grid types \cite{malardel2016new}, to help to extend the lifespan of the spectral method for numerical weather prediction (NWP), it may not be sufficient for certain desired cases and resolutions \cite{diamantakisecmwf}.
+
+The motivation for this project was to help address this problem while still utilising a spectral approach.
+More precisely, we aim to develop a sparse spectral method for solving PDEs on the spherical cap as a surface in 3D, with a simple extension to a spherical band.
+Together, these frameworks can be pieced together to create a spectral element method for the whole sphere, or further developed to investigate spectral methods on other spherical subdomains.
+By spectral element method, we mean a finite element method (FEM) that uses high degree basis polynomials for its elements (this could also be referred to as a $p$-FEM with large $p$).
+In other words, we can use our spectral methods developed for the spherical band and cap elements as part of a finite element framework.
+By using this approach, one would be able to avoid having to complete the global spectral transform (in particular, the global Legendre transform) and instead be able to simply apply the local element transforms in parallel.
+Moreover, by still using a spectral approach, one can maintain the high accuracy and excellent error properties that such methods bring.
+
+Chapter 2 of this thesis provides an introduction to sparse spectral methods via the Spherical Harmonics on the whole sphere surface.
+We present a framework for how the Spherical Harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
 Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.
 
-While the whole sphere spectral method has been successful for numerous years \cite{wedi2013fast}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the ECMWF model \cite{ecmwf2020scalability}.
-This thesis serves to lay a foundation for addressing this while still utilising a spectral approach.
-More precisely, we develop a sparse spectral method for solving partial differential equations (PDEs) on the spherical cap as a surface in 3D, with a simple extension to a spherical band.
-Together, these frameworks can be pieced together to create a spectral element method for the whole sphere, or further developed to investigate spectral methods on other spherical subdomains.
-
-To this end, in Chapter 3 we move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+In Chapter 3 we move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials, which we define.
@@
 
-With a greater knowledge base in our quiver, we can adapt the techniques learnt from the founding of the disk-slice formulation to a surface in 3D in Chapter 4 -- in particular the spherical cap, a subdomain of the surface of a unit sphere.
-Using the same family of (non-classical) 1D OPs, we present a suitable orthogonal polynomial basis once more, this time as polynomials in $x,y,z$ complete with sparse differential operators.
+With a greater knowledge base in our quiver, we can adapt the techniques learnt from the founding of the disk-slice formulation to surfaces in 3D in Chapter 4.
+Using the same family of (non-classical) 1D OPs, we present a suitable orthogonal polynomial basis for the spherical cap, a subdomain of the surface of a unit sphere, complete with sparse differential operators.
+A relatively simple adaption permits this framework to be extended to a spherical band.
+From here, a spectral element method could be devised for the whole sphere using the aforementioned as elements.

### 2021-04-15 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Improved intro and background. Edits to SH chapter.

@@
 \chapter{Spherical harmonics as orthogonal polynomials}
+
+\bstodoinline{Summarise the ECMWF approach, and how this work is then a stepping stone to understanding more about multi dim OPs on complex domains.}
 
@@
 }
-and
+and, for $l \in \No, m \in \Z$ s.t. $-l \le m \le l$,
 \bseqnnumber{
 	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
-	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } d_{l,m} := \half (l + \abs{m} + 1), \label{eqn:diffjac},
-}
-where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:diffjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}.
+	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } 
+	d_{l,m} := 
+		\begin{cases}
+			\half (l + \abs{m} + 1) & l > \abs{m} \\
+			0 & l = \abs{m}
+		\end{cases}. \label{eqn:diffjac}
+}
+\bsrefeqncapital{eqn:normjac} and \bsrefeqn{eqn:diffjac} are simply consequences of the definition of the Jacobi polynomials -- see \cite[Table 18.3.1]{DLMF} and \cite[18.9.15]{DLMF}.
 
@@
 		\end{cases} \\
-	A_{l,m,3} &:= {\clm c_{l-1}^{m-1} \over 2 (l+1)(l+2)}
+	A_{l,m,3} &:= {\clm c_{l+1}^{m-1} \over 2 (l+1)(l+2)}
 		\begin{cases}
@@
 		\end{cases} \\
-	A_{l,m,4} &:= {\clm c_{l-1}^{m+1} \over 2 (l+1)(l+2)}
+	A_{l,m,4} &:= {\clm c_{l+1}^{m+1} \over 2 (l+1)(l+2)}
 		\begin{cases}
@@
 		\begin{cases}
-			\big(\vec{\eta}_{0}\big)_j & j = 2, 4(2l+1) + 4 \\
+			\big(\vec{\eta}_{0}\big)_{j-1} & j = 2, 4(2l+1) + 4 \\
 			0 & o/w
@@
 		\begin{cases}
-			\big(\vec{\eta}_{2}\big)_j & j = 2(2l+1), 6(2l+1) - 4 \\
+			\big(\vec{\eta}_{2}\big)_{j-1} & j = 2(2l+1), 6(2l+1) - 4 \\
 			0 & o/w
@@
 			& & \calL_N
-		\end{pmatrix} \in \R^{(N+1)^2}, \quad
+		\end{pmatrix} \in \R^{(N+1)^2 \times (N+1)^2}, \quad
 	\tilde \calD = 
 		\begin{pmatrix}
-			\calL_0 & & & & \\
-			& 0 & & & \\
-			& & \ddots & & \\
-			& & & \calL_N & \\
-			& & & & 0
-		\end{pmatrix} \in \R^{2(N+1)^2}, \\
+			\tilde \calD_0 & & \\
+			& \ddots & \\
+			& & \tilde \calD_N
+		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2}, \\
 	\calG &= I_{(N+1)^2}, \quad
@@
 			& & & & 0
-		\end{pmatrix}  \in \R^{2(N+1)^2},
-}
-where $\calL_l := -l(l+1) \: I_{2l+1}$.
+		\end{pmatrix}  \in \R^{2(N+1)^2 \times 2(N+1)^2},
+}
+where
+\bseqn{
+	\calL_l := -l(l+1) \: I_{2l+1}, \quad
+	\tilde \calD_l := \begin{pmatrix}
+			-l(l+1) & & & & \\
+			& 0 & & & \\
+			& & \ddots & & \\
+			& & & -l(l+1) & \\
+			& & & & 0
+		\end{pmatrix} \in \R^{2(2l+1) \times 2(2l+1)}.
+}.
 \end{theorem}
@@
 \bseqn{
-	\Big[ I + \Delta t^2 \calH \tilde \calG \tilde \calD - 2 \rotrate \Delta t \: J_z \calR \Big] \: \uvec^c_{n+1} &= \uvec^c_n + \Delta t \tilde \calG \tilde{\vec{h}}^c_{n} \\
+	\Big[ I + \Delta t^2 \calH \tilde \calG \tilde \calD + 2 \rotrate \Delta t \: J_z \calR \Big] \: \uvec^c_{n+1} &= \uvec^c_n + \Delta t \tilde \calG \tilde{\vec{h}}^c_{n} \\
 	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}.
 }
-Here, it is clear that the matrix $I + \Delta t^2 \calH \tilde \calG \tilde \calD - 2 \rotrate \Delta t \: J_z \calR$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
+Here, it is clear that the matrix $I + \Delta t^2 \calH \tilde \calG \tilde \calD + 2 \rotrate \Delta t \: J_z \calR$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
 Doing so at each iteration provides us with the coefficients vectors $\uvec^c_{n+1}, \tilde{\vec{h}}^c_{n+1}$ if we know $\uvec^c_{n}, \tilde{\vec{h}}^c_{n}$.

### 2021-04-15 papers/spherical-caps/submitted/significance.tex
Moved submitted SC paper into new directory. Revised version to be added.

--- 2020-12-22 papers/spherical-caps/highlights.tex
@@
 
-\title{Sparse spectral methods for partial differential equations on spherical caps, highlights}
+\title{Sparse spectral methods for partial differential equations on spherical caps, significant and novelty}
 \author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
@@
 
-\begin{enumerate}
-\item Constructs sparse discretisations for partial differential operators on spherical caps using orthogonal polynomials. 
-\item Approximations are observed to converge spectrally fast.
-\item Optimal complexity is achieved for operators that are invariant to rotation around the pole.
-\item Examples include (surface) Poisson equation, Helmholtz equation with variable coefficients, and the biharmonic equation.
-\end{enumerate}
+This paper introduces a novel approach for solving partial differential equations on an important subset of the full sphere: spherical caps.
+The approach leads to sparse discretisations of general partial differential operators that can be efficiently solved with spectral convergence, and in the case of operators that are invariant to rotation around the pole the resulting linear systems can be constructed and solved in optimal complexity.
+There is a big desire for more localised approaches to solving PDEs on subsections in weather prediction and other applications, and this manuscript lays the groundwork for high order, high accuracy new numerical methods.
 

### 2021-04-15 papers/spherical-caps/revised/response-to-referees.tex
Addressed most comments except the ones requiring major additions. Added tex file for the responses/changes logging

\documentclass[10pt]{letter}

\usepackage[papersize={8.5in,11in},margin=1in]{geometry}
\usepackage[final,stretch=10,shrink=10]{microtype} % default of stretch=shrink=20 is a little too much
\usepackage{amsmath,amsfonts,amsxtra,amssymb}
\usepackage{url}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}

\definecolor{bluey}{rgb}{0.0,0.2,0.5}
\newcommand{\comment}[1]{\textit{\color{bluey}#1}}

%\definecolor{rev1}{HTML}{cb270f}
%\definecolor{rev2}{HTML}{1c8235}

\usepackage{todonotes}
\newcommand{\sotodo}{\todo[color=green]}
\newcommand{\sotodoinline}{\todo[color=green,inline=true]}
\newcommand{\bstodo}{\todo[color=pink]}
\newcommand{\bstodoinline}{\todo[color=pink,inline=true]}

\newcommand{\half}{\frac{1}{2}}
\newcommand{\genjac}{R}

\begin{document}

\thispagestyle{empty}

\hfill
\begin{flushright}
Correspondence to: \\
Sheehan Olver \\
s.olver@imperial.ac.uk \\
\end{flushright}

\vspace{1em}
\hfill\today

Dear Jose Bautista Francisco III

Thank you for forwarding the reviews of our article. We greatly appreciate your time as well as that of the referees.  We have made changes to reflect the comments and suggestions of the referees. Below we detail our response to the referees' reports.

Ben Snowball and Sheehan Olver

\bigskip

\centerline{\textbf{Response to Referee 1}}

\begin{itemize}[parsep=1em,leftmargin=1em]

\item \comment{The authors spend an inadaquate amount of time discussing other approaches for solving this problem and comparing their methodology (e.g., closest point method, radial basis functions). What is the complexity convergence behavior of other approaches that can handle spherical caps?}

\bstodoinline{write comment}

\item \comment{Definition 2: It should be mentioned somewhere that OPs should defined by the condition that they are orthogonal to all lower degree polynomials.}

Added a comment before Definition 2 to explain this, as well as an extra line in Definition 2.

\item \comment{Above eq 7: These are not Jacobi matrices in the classical sense, but they are block Jacobi matrices.}

Addressed.

\item \comment{Beginning of Section 2.3: What is the "triangle case"?}

Explained via a comment in the relevant place. Typo has also been corrected in one equation in the section (missing z from $Q(x,y,z)$).

\item \comment{Remark at the bottom of page 7: I do not know what it means for z to be rotionally invariant. But the z coordinate is invariant under any rotation involving only the $\theta$ variable.}

Addressed.

\item \comment{Definition 6: Since $\tilde a \ge 2$, it only needs to be stated that $\tilde a$ is an integer.}

Fixed.

\item \comment{Figure 1: Extra the in the the}

Fixed.

\item \comment{All figures: They should be larger with larger font sizes.}

\bstodoinline{Fix this}

\item \comment{About Section 4.3: Is it possible to leverage the DCT to speed up the expansion of functions in the basis?}

\bstodoinline{Reference Michael Slavinsky paper on Transforms, say it's an open problem.}

\item \comment{Figure 4: 'N is the the degree' is ambiguous. How does N relate to the number of unknowns?}

This is mentioned in the body of the text, but now explained in the caption.

\end{itemize}

\bigskip

\centerline{\textbf{Response to Referee 2}}

\begin{itemize}[parsep=1em,leftmargin=1em]

\item \comment{One shortcoming of this paper is that it has no discussion of the condition number of the linear systems to be solved. I suspect that the condition numbers can get large, but if so, can the matrices be preconditioned in a similar way the Olver?Townsend ultraspherical spectral method (so that it has bounded condition number)?}

\bstodoinline{Plot condition number of each m-block of Laplacian. Also do plot of preconditioned Laplacian (where preconditioner P is the inverse of the diagonal of the Laplacian) which should mean bounded condition numbers.}

\item \comment{If the condition numbers do not get large, can you prove it, indicate why, or give numerical evidence?}

The evidence we can provide is that which addresses the first question.

\item \comment{The paper only includes numerical experiments involving zero Dirichlet boundary conditions. Please could the authors clarify in the manuscript if there are any roadblocks to boundary conditions beyond the zero Dirichlet ones demonstrated in Section 5?}

Added comment a small subsection to Section 5 to address this. \bstodoinline{Is this enough?}

\item \comment{Is the code used to produce the numerical experiments going to be made publicly available? For example in the ApproxFun Julia package?}

The package that makes up the framework for expanding functions, obtaining the operator matrices etc. is purely experimental at this stage, but is publicly available. This fact and where it can be found has been added to the paper.

\end{itemize}

\end{document}

### 2021-04-15 papers/spherical-caps/revised/spherical-caps.tex
Submitted version of paper, to track changes.
Addressed most comments except the ones requiring major additions. Added tex file for the responses/changes logging

--- 2020-12-21 papers/spherical-caps/spherical-caps.tex
@@
 
-%\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
-We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
+The code that allows one to produce the numerical examples in this paper is publicly available as a Julia package\footnote{https://github.com/snowball13/OrthogonalPolynomialFamilies} to partner the ApproxFun package \cite{ApproxFun} -- however, it is strongly warned that the package is purely experimental at this stage.
 
@@
 \end{proposition}
+Of course, \enquote{orthogonal polynomials} are defined as such by the condition that each polynomial is orthogonal to each other, with respect to the given inner product.
 
@@
 is a normalising constant.
-Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}).
+Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}). That is,
+\begin{align}
+	\ip< \genjac_n^{(a,b)}, \: \genjac_m^{(a,b)} >_{\genjacw^{(a,b)}} = \delta_{n,m}.
+\end{align}
 \end{definition}
@@
 \begin{remark}
-For $z$ multiplication, note that different Fourier modes do not interact. This is because $z$ is rotationally invariant. 
+	For $z$ multiplication, note that different Fourier modes do not interact. This is because multiplication by $z$ is invariant with respect to rotation around the $z$-axis. 
 \end{remark}
@@
 \end{align*}
-Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are \textit{banded-block-banded matrices}:
+While $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are not Jacobi matrices in the classical sense, they are block Jacobi matrices.
+However, it is useful to simply label them as such.
+In fact, note that they are in fact \textit{banded-block-banded matrices}:
 
@@
 
-Following the triangle case \cite{olver2019triangle}, we use a multivariate analogue of Clenshaw's algorithm for evaluation, where we  combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
+A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here.
+Combining each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yeilds:
 \begin{align*}
@@
 		\bigscopta_{-1}(x,y,z) := 0 \\
-		\bigscopta_{0}(x,y) := \scopa_0 \\
-		\bigscopta_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
+		\bigscopta_{0}(x,y,z) := \scopa_0 \\
+		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
 	\end{cases}
@@
 \begin{definition}\label{def:differentialoperators}
-	Let $a$ be a nonnegative parameter, and $\tilde a \ge 2$ be a positive integer.
+	Let $a$ be a nonnegative parameter, and $\tilde a \ge 2$ be an integer.
 Define the operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ according to:
@@
 	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
-For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
+For (c), the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
 	\label{fig:sparsity}
@@
 \end{proof}
-
-%\begin{theorem}\label{Theorem:quadrule}
-%Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
-Further, we denote the $M_2$ Gauss quadrature nodes and weights for the unit circle \cite{???}\bstodo{good reference?
-Or derive myself using Chebychev polys...} by $(\mathbf{s}_j, w_j^{(s)})$ where for $j = 1,\dots,M_2$:
-%\begin{align*}
-%	\mathbf{s}_j &= \Big( \cos\fpr(\frac{2 \pi j}{M_2}), \: \sin\fpr(\frac{2 \pi j}{M_2}) \Big), \\
-%	w_j^{(s)} &= \frac{2\pi}{M_2}.
-%\end{align*}
-%Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
-%\begin{align*}
-%	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
-%	z_{l+(j-1)M_2} &:= t_j, \\
-%	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
-%\end{align*}
-%Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
-%\begin{align*}
-%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
-%\end{align*}
-%where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
-%\end{theorem}
-%
-%\begin{proof}
-%Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-%\begin{align*}
-%	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
-%	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
-%\end{align*}
-%so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
-Note that if $f$ is a polynomial, then $f_e(\rho(t)\xvec, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $\xvec \in \R^2$.
-%
-%Now, we have that
-%\begin{align*}
-%	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_e\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \\
-%	&\quad \quad \approx\int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) \mathbf{s}_l, z\big) \Big) \: \D z \quad (\star) \\
-%	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) \mathbf{s}_l, t_j\big) \quad (\star \star) \\
-%	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
-%\end{align*}
-%Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
-We therefore achieve equality at $(\star)$ if $M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-\bstodo{For the circle rule, is this bound on N true?}
-%
-%Next, note that 
-%\begin{align*}
-%	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_o\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \quad (\dagger) \\
-%	&\quad \quad = 0
-%\end{align*}
-%since the inner integral at $(\dagger)$ over $\bm{\xi}$ is zero, due to the symmetry over the domain.
-%
-%Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
-%\begin{align*}
-%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z  \\
-%	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
-%\end{align*}
-%where $M = M_1 M_2$ and $2M_1 - 1 \ge N, M_2 - 1 \ge N$.
-%\end{proof}
 
@@
 	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.
-This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
+This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution (the number of unknowns is then $(N+1)^2$).
 Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
@@
 
+\subsection{Other boundary conditions}
+
+Other boundary conditions beyond those already presented here are beyond the scope of this paper.
+However, the approach established for the case of a triangle domain in 2D shows how Neumann conditions can be done by writing these as a system of equations in partial derivatives \cite{olver2019triangle}.
+This approach may also work here for the spherical cap.
+Ideally, one would properly tackle the tangent space (which again is beyond the scope of this paper) which would lower the degrees of freedom.
+
 %
@@
 
-%\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.
-The second author was supported in part by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces.
-
-%
-%%
-%\appendix
-%%
-%\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
-%
-%\bstodoinline{Is this worth including?}
-%
-%\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis} \bstodoinline{Yes thought so too}
-%
-%%\begin{figure}[t]
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-%%        	%\label{fig:solutionblocknorms-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-poisson}
-%%	\centering
-%%	%\label{fig:solution-poisson}
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-%%	%\label{fig:solution-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\centering
-%%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
-Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
-Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-in the trapezium.}
-%%	\centering
-%%	\label{fig:appendixfigs}
-%%\end{figure}
-%
-%We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
-Consider the 3D Dirichlet problem on a domain $\Omega$:
-%\begin{align*}
-%	\begin{cases}
-%         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
-%         u = 0 \quad \text{on } \partial \Omega
-%         \end{cases}
-%\end{align*}
-%This has the weak formulation
-%\begin{align*}
-%	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
-%\end{align*}
-% for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
-%
-%In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$.
-While spherical bands are an extension to this work, we note that we can still apply a $p$-FEM using two half-spheres ($\alpha = 0$).
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain.
-We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p \rvert_\element) \le p\}$ for some $p \in \N$.
-%
-%We seek $u_p \in V_p$ s.t.
-%\begin{align}
-%	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-%	\label{eqn:FEMweakform}
-%\end{align}
-%
-%Define 
-%\begin{align*}
-%	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
-%\end{align*}
-%for any $a$.
-Note that due to orthogonality this is a diagonal matrix.
-By choosing suitable bases for our FE spaces, we can rewrite (\ref{eqn:FEMweakform}) in matrix-vector form.
-We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.
-%
-%Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
-%\begin{align*}
-%	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
-%					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
-%\end{align*}
-%Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
-Then,
-%\begin{align*}
-%	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
-%	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%
-%Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-%\begin{align*}
-%	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
-%\end{align*}
-%(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-%\begin{align*}
-%	A\mathbf{u} = \tilde{\mathbf{f}}.
-%\end{align*}
-%where the (element) stiffness matrix $A$ is defined by 
-%\begin{align*}
-%	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
-%\end{align*}
-%and the load vector $\tilde{\mathbf{f}}$ is given by 
-%\begin{align*}
-%	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-%
-
 \bibliography{spherical-caps}

### 2021-04-15 papers/spherical-caps/revised/response-to-referees.tex
Changed to quad rule proof to fix some typos and make clearer

@@
 
-Explained via a comment in the relevant place. Typo has also been corrected in one equation in the section (missing z from $Q(x,y,z)$).
+Explained via a comment in the relevant place.
 
@@
 
+\bigskip
+
+\centerline{\textbf{Other Changes}}
+
+\begin{itemize}[parsep=1em,leftmargin=1em]
+	\item Typo has been corrected in one equation in Section 2.3 (missing z from $Q(x,y,z)$).
+	\item Proof for quadrature rule has been fixed - reference to $\bm{s}_l$ removed; $(1-x^2)^{-1/2}$ weight needed adding to the integrals over $dx$; clarified the maths that shows the symmetry that is referenced.
+\end{itemize}
+
 \end{document}

### 2021-04-15 papers/spherical-caps/revised/response-to-referees.tex
Addressed comments on CPM/RBF and DCT

@@
 
-\bstodoinline{write comment}
+We have added in a discussion of the closest point method.
+We have chosen to omit discussion on radial basis functions as we do not feel there is particularly suitable to reference and appropriate to discuss.
 
@@
 
-\bstodoinline{Reference Michael Slavinsky paper on Transforms, say it's an open problem.}
+Added a remark to state its an open problem.
 

### 2021-04-15 papers/spherical-caps/revised/spherical-caps.tex
Changed to quad rule proof to fix some typos and make clearer
Addressed comments on CPM/RBF and DCT

@@
 Further, we avoid the singular nature at the poles or  as $\alpha$ approaches $0$ that such a projection may give, since our new approach yields a smooth polynomial basis for all $\alpha \in [-1,1)$.
+
+There are of course other approaches for solving PDEs on surfaces, such as the closest point method \cite{macdonald2010implicit, macdonald2008level}.
+This involves solving the PDE inside a 3D volume that acts as a \enquote{shell} for the surface, and using a \enquote{closest point} to as the solution on the surface.
+While such an approach is useful for other geometries, it is of low order and does not achieve spectral convergence with sparse discretisations.
+Further, for our interest of the sphere, it would not take advantage of any rotational symmetry.
 
@@
 \begin{align*}
-	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
+	x_{l+(j-1)M_2} &:= \rho(t_j) \: s_l, \\
+	y_{l+(j-1)M_2} &:= \rho(t_j) \: \sqrt{1-s_l^2}, \\
 	z_{l+(j-1)M_2} &:= t_j, \\
@@
 \begin{align*}
-	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
-\end{align*}
-for some function $g$. Then, integrating the even function $f_e$ we have
+	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta &= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}} \\
+	&= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(-x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}}
+\end{align*}
+for any function $g$, using a change of variables $x \to -x$ for the second term in the integral. Then, integrating the even function $f_e$ we have
 \begin{align*}
@@
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \D x \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_e\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
+	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
 	&\quad \quad \approx \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
@@
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) + f_o\big(\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) - f_o\big(\rho(z)x, \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_o\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
 	&\quad \quad = 0.
@@
 we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).
+
+\remark{While it may be possible to leverage a discrete cosine transform (DCT) to speed up the calculation of the coefficients  \cite{slevinsky2018use}, this is an open question.}
 

### 2021-04-16 papers/spherical-caps/revised/response-to-referees.tex
Finished making the changes to paper, with new figs for condition numbers.

@@
 
-\bstodoinline{Fix this}
+Fixed.
 
@@
 
-\bstodoinline{Plot condition number of each m-block of Laplacian.
-Also do plot of preconditioned Laplacian (where preconditioner P is the inverse of the diagonal of the Laplacian) which should mean bounded condition numbers.}
+Added small subsection and plots to show the condition numbers of each m-block of Laplacian operator, along with a preconditioned Laplacian (where the preconditioner $P$ is the inverse of the diagonal of the Laplacian).
+While the Laplacian does not have bounded condition number, the trivial preconditioning keeps it bounded.
 

### 2021-04-16 papers/spherical-caps/revised/spherical-caps.tex
Finished making the changes to paper, with new figs for condition numbers.

@@
 
-\begin{figure}[t]
+\begin{figure}[tp]
 	\centering % <-- added
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-laplacian}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsity-of-laplacian}
 		\centering
@@
 	\end{subfigure}\hfil % <-- added
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-rho2laplacian}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsity-of-rho2laplacian}
 		\centering
@@
 	\medskip
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-helmholtz}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsity-of-helmholtz}
 		\centering
@@
 	\end{subfigure}\hfil % <-- added
-	\begin{subfigure}{0.45\textwidth}
-		\includegraphics[scale=0.3]{sparsity-of-biharmonic}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsity-of-biharmonic}
 		\centering
@@
 The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
+
+\subsection{Stability of the Laplacian operator}
+
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}[b]{0.55\textwidth}
+		\includegraphics[scale=0.4]{condition-numbers-laplacian-N=200}
+	\end{subfigure}
+	\hfill%
+
+	\begin{subfigure}[b]{0.55\textwidth}
+		\includegraphics[scale=0.4]{condition-numbers-max-laplacian}
+	\end{subfigure}
+	\hfill%
+	\caption{Plots of the condition number for the Laplacian operator matrix $\Delta^{(1)}_W$ and the preconditioned matrix $P^{-1} \; \Delta^{(1)}_W$ where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
+Left: The condition numbers of each diagonal block of $\Delta^{(1)}_W$ for $N = 200$, as well as those for $P^{-1} \; \Delta^{(1)}_W$.
+Right: The maximum condition number of all diagonal blocks of the Laplacian and the preconditioned Laplacian operators, as the degree $N$ increases.}
+	\label{fig:conditionnumbers}
+\end{figure}
+
+For a given problem $A\vec{x} = \vec{b}$, the condition number is a measure of how accurate solving the system will be (ignoring any numerical rounding errors), and is given as the condition number of the matrix $A$:
+\begin{align*}
+	\kappa(A) := \norm{A}_2 \norm{A^{-1}}_2.
+\end{align*}
+Recall that the Laplacian operator is block-diagonal.
+One can therefore look at the condition numbers of each block on the diagonal as proxy for the condition number of the Laplacian (and hence the Poisson equation), where graphical representations are seen in \ref{fig:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
+As the degree $N$ increases, the condition numbers of these blocks becomes increasingly large (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
+Fortunately however, by applying a trivial preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound the condition numbers -- this is also seen in \ref{fig:conditionnumbers}.
+The preconditioning matrix chosen is $P^{-1}$, where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
 
@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{poisson-u=wyexpx-N=60}
+\begin{figure}[tp]
 	\centering
-	%\label{fig:solution-poisson}
+	\begin{subfigure}{0.85\textwidth}
+		\includegraphics[scale=0.4]{poisson-u=wyexpx-N=60}
 	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-poisson-epsilonfun-N=200}
-        	%\label{fig:solutionblocknorms-poisson}
+	\hfill%
+
+	\begin{subfigure}{0.55\textwidth}
+		\includegraphics[scale=0.4]{solutionblocknorms-poisson-epsilonfun-N=200}
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
-Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
+	\hfill%
+	\caption{Top: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
+Bottom: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
 This indicates spectral convergence.}
-	\centering
 	\label{fig:poisson}
@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{helmholtz-f=wyexpx-N=60-n=300}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-varyingk-N=200}
-        	%\label{fig:solutionblocknorms-poisson}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
-Right: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
-This indicates spectral convergence.}
-	\centering
-	\label{fig:helmholtz}
-\end{figure}
-
-\begin{figure}[t]
-	\centering % <-- added
-	\includegraphics[scale=0.35]{complexity-Nend=1000}
-	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.
-This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution (the number of unknowns is then $(N+1)^2$).
-Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
-	\centering
-	\label{fig:complexity}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.15]{biharmonic-f=erf-N=80-n=50}
-	\centering
-	%\label{fig:solution-biharmonic}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-expfun-N=200}
-	\centering
-        	%\label{fig:solutionblocknorms-biharmonic}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
-This demonstrates algebraic convergence.}
-	\centering
-	\label{fig:biharmonic}
-\end{figure}
 % END FIGURES
@@
 
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}{0.85\textwidth}
+		\includegraphics[scale=0.4]{helmholtz-f=wyexpx-N=60-n=300}
+	\end{subfigure}
+
+	\begin{subfigure}{0.55\textwidth}
+		\includegraphics[scale=0.4]{solutionblocknorms-helmholtz-varyingk-N=200}
+	\end{subfigure}
+	\caption{Top: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
+Bottom: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
+This indicates spectral convergence.}
+	\label{fig:helmholtz}
+\end{figure}
+
+\begin{figure}[tp]
+	\centering % <-- added
+	\includegraphics[scale=0.4]{complexity-Nend=1000}
+	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.
+This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution (the number of unknowns is then $(N+1)^2$).
+Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
+	\label{fig:complexity}
+\end{figure}
+
 Find $u(x,y)$ given functions $v$, $f : \Omega \to \R$ such that:
@@
 \subsection{Biharmonic equation}
+
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}{0.85\textwidth}
+		\includegraphics[scale=0.4]{biharmonic-f=erf-N=80-n=50}
+	\end{subfigure}
+	\hfill%
+
+	\begin{subfigure}{0.55\textwidth}
+		\includegraphics[scale=0.4]{solutionblocknorms-biharmonic-expfun-N=200}
+	\end{subfigure}
+	\hfill%
+	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
+Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
+This demonstrates algebraic convergence.}
+	\label{fig:biharmonic}
+\end{figure}
 

### 2021-04-19 experiments/sphericalcaptests-plots.jl
Code to get condition number plots

@@ -1,3 +1,4 @@
+using Plots
 using BandedMatrices
 using Revise
 using ApproxFun
@@ -18,11 +19,11 @@ import OrthogonalPolynomialFamilies: points, pointswithweights, getopptseval,
                     differentiatespacephi, increasedegreeoperator,
                     getptsevalforop, getderivptsevalforop, laplacianoperator,
                     rho2laplacianoperator, gettangentspace,
-                    rhogradoperator, coriolisoperator, rho2operator,
-                    rhodivminusgradrhodotoperator, gradrhooperator, weightoperator,
-                    squareoperator
+                    coriolisoperator, rho2operator, gradrhooperator,
+                    weightoperator, squareoperator
+using LinearAlgebra
 using JLD
-using Makie
+# using Makie
 
 
 isindomain(pt, D::SphericalCapFamily) = D.α ≤ pt[3] ≤ D.β && norm(pt) == 1.0
@@ -183,7 +184,10 @@ c = randn(10)
 d = randn(10)
 (a' * b) * (c' * d) - a' * (b * d') * c
 
+
 #=======#
+# Examples
+
 # Poisson
 using SpecialFunctions
 # f = (x,y,z) -> 1 + erf(5(1 - 10((x - p[1])^2 + (y - p[2])^2 + (z - p[3])^2)))
@@ -215,12 +219,12 @@ vs = vbox(scene, cl)
 Makie.save("experiments/saved/sphericalcap/images/poisson-f=erf-N=$N-n=$n.png", vs)
 
 # Helmholtz
-N = 60
+N = 100
 k = 100
 z1 = S.family.α; x1 = 0.7; vpt = [x1; sqrt(1 - x1^2 - z1^2); z1]
 v = (x,y,z) -> 1 - (3(x - vpt[1])^2 + 5(y - vpt[2])^2 + 2(z - vpt[3])^2)
 V = operatorclenshaw(Fun(v, S, 100), S, N)
-Δw = laplacianoperator(S, N; square=true)
+Δw = squareoperator(S, laplacianoperator(S, S, N), N)
 tw = transformparamsoperator(S, S0, N; weighted=true)
 t = transformparamsoperator(S0, S, N+1; weighted=false)
 A = Δw + k^2 * V * squareoperator(S, t * tw, N)
[…]

### 2021-04-19 papers/spherical-caps/revised/response-to-referees.tex
Minor edit to response tex

@@
 \item \comment{One shortcoming of this paper is that it has no discussion of the condition number of the linear systems to be solved.
-I suspect that the condition numbers can get large, but if so, can the matrices be preconditioned in a similar way the Olver?Townsend ultraspherical spectral method (so that it has bounded condition number)?}
+I suspect that the condition numbers can get large, but if so, can the matrices be preconditioned in a similar way the Olver-Townsend ultraspherical spectral method (so that it has bounded condition number)?}
 

### 2021-04-23 papers/thesis/full/ch1-introduction/introduction.tex
Made some changes to address points in comments pdf

@@
 Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs} \bstodo{citation here}.
-For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations \cite{boyd2001chebyshev, olver2013fast}.
-There are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few \cite[Section 18.3]{DLMF} -- but the area of multivariate orthogonal polynomials has a smaller array of research.
-However, this branch of mathematics has an encouraging future \cite{dunkl2014orthogonal}, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains, such as on the triangle \cite{olver2019triangle} with applications including solving Volterra integral equations \cite{gutleb2020sparse}.
+For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations (see \cite{boyd2001chebyshev, olver2013fast} and references therein).
+While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[Section 18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
 
-For the uninitiated, a spectral method is one that, given a domain, numerically approximates the solution of a (differential) equation as sum of basis functions for said domain, with the coefficients to be found.
-The basis functions are non-zero over the whole domain, meaning the solution is found globally and thus helps such methods to be inherently accurate (providing the solution is smooth).
-By utilising OPs as basis functions, we can develop sparse spectral methods (SSMs), meaning that the naturally sparse relationships between the basis OPs leads to sparse operator matrices that represent the differential operations in the equation to solve.
+One could say this is surprising, given that there is a long history of around 150 years associated with multivariate OPs beginning with Hermite first presenting the multivariate Hermite polynomials in 1865 \cite{appel1926fonctions, ismail2017review}.
+Zernike polynomials \cite{zernike1934diffraction} were first introduced in 1934, for example, that are a group of bivariate polynomials orthogonal on a unit circle.
+Koornwinder in 1975 described a method for constructing two-variable OPs from univariate OPs \cite{koornwinder1975two}.
+However, few books have been published over the years on the topic of multivariate orthogonal polynomials \cite{dunkl2014orthogonal}.
+
+This branch of mathematics has an encouraging future though, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains.
+Spectral methods have been established on the triangle \cite{olver2019triangle} using OPs inspired by the Koornwinder approach for defining them, and on the disk \cite{vasil2016tensor} using Zernike polynomials.
+Applications include solving Volterra integral equations \cite{gutleb2020sparse}.
+
+There are various interpretations of what a \enquote{spectral method} is.
+It could be described as one that achieves spectral convergence of its solutions, or one that uses eigenfunctions as basis functions, or one that uses OPs as basis polynomials \cite{}\bstodo{cite these as interpretations}.
+It is the latter that we refer to for our purposes as a spectral method.
+By utilising OPs as basis functions, we can develop sparse spectral methods (SSMs), meaning that the naturally sparse relationships between the basis OPs lead to sparse operator matrices that represent the differential operations in the equation to solve.
 Sparse spectral methods for one dimensional problems have been shown to lead to \enquote{almost banded} matrices \cite{olver2013fast}.
 
-In this thesis we expand upon this knowledge by providing frameworks for similar sparse spectral methods for solving PDEs on other multidimensional domains (notably including the disk-slice and trapezium in 2D, and the spherical cap surface in 3D) that also yield matrices that are similarly \enquote{almost banded}.
-The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
+In this thesis we expand upon this knowledge by providing frameworks for similar sparse spectral methods for solving PDEs on other multidimensional domains (notably including the disk-slice and trapezium in 2D, and the spherical cap surface in 3D) that also yield matrices that are what is defined as \enquote{banded-block-banded}.
+We take inspiration from the work established for the triangle \cite{olver2019triangle} and the unit disk \cite{vasil2016tensor}, which can  be seen as special cases of the disk-slice and trapezium in the framework we present here.
+The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
 
-Spherical Harmonics are of course one long-established and famous group of multidimensional orthogonal polynomials.
-They are used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
+Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used for many applications involving PDEs \cite{}\bstodo{cite applications of SHs}.
+Notably, they are used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
 While the whole sphere spectral method using the Spherical Harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast, diamantakisecmwf}.
 
-This spherical harmonic transform in fact uses two transforms -- a Fourier transform (using the well established Fast Fourier Transform \cite{cooley1965algorithm}) in the longitudinal direction and a Legendre transform in the latitudinal direction -- and it is the Legendre transform that has been identified as inhibiting future performance due to its computational cost.
+Many implementations of an algorithm to compute the spectral transform (or spherical harmonic transform) exist (see e.g.
+\cite{slevinsky2019fast, suda2002fast}).
+For the IFS, the spherical harmonic transform in fact uses two transforms -- a Fourier transform (using the well-established Fast Fourier Transform (FFT) \cite{cooley1965algorithm}) in the longitudinal direction and a Legendre transform in the latitudinal direction -- and it is the Legendre transform that has been identified as inhibiting future performance due to its computational cost.
 While a Fast Legendre Transform (FLT) \cite{wedi2013fast} has been incorporated into the model, along with new grid types \cite{malardel2016new}, to help to extend the lifespan of the spectral method for numerical weather prediction (NWP), it may not be sufficient for certain desired cases and resolutions \cite{diamantakisecmwf}.
@@
 Together, these frameworks can be pieced together to create a spectral element method for the whole sphere, or further developed to investigate spectral methods on other spherical subdomains.
+This, however, is future work beyond the scope of this thesis.
 By spectral element method, we mean a finite element method (FEM) that uses high degree basis polynomials for its elements (this could also be referred to as a $p$-FEM with large $p$).
@@
 Chapter 2 of this thesis provides an introduction to sparse spectral methods via the Spherical Harmonics on the whole sphere surface.
-We present a framework for how the Spherical Harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
+Here, we think of the Spherical Harmonics in a non-traditional way and write them as a group of multidimensional orthogonal polynomials in $(x,y,z)$ as opposed to functions of spherical coordinates.
+We present an OP framework for how the Spherical Harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for a function's expansion.
 Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.

### 2021-04-23 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Made some changes to address points in comments pdf

@@
 }
-Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Jacobi polynomial: \bstodo{Sheehan - do you know what citation I can use for the definition of the Associated Legendre Polynomials (as the mth derivative of the ordinary Legendre Polynomials)?}
+Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[14.21.1]{DLMF}:
 \bseqn{
@@
 for $m = -l, \dots, -1$.
-Here, $n!
-:= n \: (n-1) \: (n-2) \dots 1$ for $n \in \N_0$ is the standard factorial, with $0!
-= 1$.
-Although referred to as the associated Legendre \enquote{polynomials}, they are not strictly polynomials when $m$ is odd, and as such some authors refer to them as associated Legendre functions instead.
+Although often referred to as the associated Legendre \enquote{polynomials} as we do here, they are not strictly polynomials when $m$ is odd, and as such some authors refer to them as associated Legendre functions instead.
 Moreover, it should be noted too that the inclusion of the $(-1)^m$ factor in the definition for $m > 0$, known as the \textit{Condon-Shortley phase} in physics, is sometimes omitted.
+\bstodoinline{I have cited DLMF on ALFs equation that defines the ALFs.
+Now need to explain how that relates to the definition in terms of mth derivative of normal Legendre poly when l,m are integers}
 
@@
 Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$.
-They span all polynomials modulo the ideal generated by $x^2 + y^2 + z^2 - 1$.
+They span all polynomials modulo the vanishing ideal associated by the zero set of $x^2 + y^2 + z^2 - 1$ in $\R^3$.
 
@@
 
-\begin{proof}
+The results in \bsreflemma{lemma:shxyzrelations} are fairly involved, and can be derived via \textit{Clebsch-Gordan coefficients}.
+
+\begin{proof}[Proof of \bsreflemma{lemma:shxyzrelations}]
 	The expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is simply a consequence of the third result in \bsrefcor{corollary:jacobirelations}:
@@
 
-These recurrences lead to Jacobi operators that correspond to multiplication by $x,y,z$. Define for $l \in \No$:
+These recurrences lead to Jacobi operators that correspond to multiplication by $x,y,z$ of the OPs. Define for $l \in \No$:
 \bseqn{
@@
 }
-and let the Jacobi matrices $J^x, J^y, J^z$ be given by 
-\bseqnnumber{
-	J^x \SHvec = x \SHvec, \quad J^y \SHvec = y \SHvec, \quad J^z \SHvec = z \SHvec \label{eqn:jacobimatsdef}.
-}
-The Jacobi matrices act on the coefficients vector of a function's expansion in the spherical harmonic basis.
-For example, let the function $f(x,y,z) : \Omega \to \C$ be approximated by its expansion $f(x,y,z) = \SHvec(x,y,z)^\top \: \fvec$ up to some degree order $N$.
-Then $x f(x,y,z)$ is approximated by $\SHvec(x,y,z)^\top \: J^x \: \fvec$, i.e.
-$ J^x \: \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the spherical harmonics basis.
+Note that these OP vectors we have defined are grouped and ordered according to the degree $l$, and within each degree ordered by the Fourier mode $m$.
+Let the Jacobi matrices $J_x, J_y, J_z$ hence be given by
+\bseqnnumber{
+	J_x \SHvec(x,y,z) = x \SHvec(x,y,z), \quad J_y \SHvec(x,y,z) = y \SHvec(x,y,z), \quad J_z \SHvec(x,y,z) = z \SHvec(x,y,z) \label{eqn:jacobimatsdef}
+}
+for $(x,y,z) \in \Omega$.
+These Jacobi matrices can act on the coefficients vector of a function's expansion in the spherical harmonic basis to implement the operation for multiplication by $x$, $y$ or $z$.
+For example, given sufficiently large $n \in \N$ let the function $f(x,y,z) : \Omega \to \C$ be given by its expansion
+\bseqn{
+	f(x,y,z) = \sum_{l=0}^{N} \sum_{m=-l}^{l} f_{l,m} \; \Ylm(x,y,z) = \SHvec(x,y,z)^\top \: \fvec
+}
+where 
+\bseqn{
+	\fvec_l := 
+		\begin{pmatrix}
+			f_{l,-l} \\
+			\vdots \\
+			f_{l,-l}
+		\end{pmatrix} \in \C^{2l+1}, 
+	\quad \quad 
+	\fvec := 
+		\begin{pmatrix}
+			\fvec_0 \\
+			\vdots \\
+			\fvec_N
+		\end{pmatrix}.
+}
+We say that $\fvec$ is the \textit{coefficients vector} of the function $f$.
+Then $x f(x,y,z)$ is approximated by $\SHvec(x,y,z)^\top \: J_x^\top \: \fvec$.
+In other words, $J_x^\top \: \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the spherical harmonics basis.
 
@@
 
-%
-%\subsection{Laplacian operator}
-%
-%The spherical harmonics are named as such because they are indeed harmonic functions -- that is, they satisfy $\nabla^2 \Ylm = \lambda \Ylm$ for some eigenvalue $\lambda$.
-It can be shown that in fact, for any $l \in \No$, $-l \le m \le l$, \bstodo{cite}
-%\bseqn{
-%	\nabla^2 \Ylm = - l (l+1) \Ylm.
-%}
-%This allows us to write down a sparse (in fact, diagonal) matrix operator for the Laplacian.
-Recall that a function $f(x,y,z)$ defined on the unit sphere $\Omega$ is approximated by its expansion $\SHvec^\top \fvec$ for some coefficients vector $\fvec$ up to degree order $N \in \N$.
-Then the function $(x,y,z) \mapsto \nabla^2 f(x,y,z)$ is approximated by $\SHvec^\top \Delta \fvec$ where $\Delta$ is the matrix operator is then simply given by
-%\bseqn{
-%	\Delta := 
-%		\begin{pmatrix}
-%			\Delta_0 & & & \\
-%			& \Delta_1 & & \\
-%			& & \ddots & \\
-%			& & & \Delta_N
-%		\end{pmatrix}, \quad
-%	\Delta_l := - l (l+1) \idmat{2l+1} \in \R^{(2l+1)\times(2l+1)}.
-%}
-
 \section{Computational aspects}
@@
 
-\bstodoinline{Is this paragraph enough? Or do I need to put more detail into explaining these methods, or derive my own?}
 In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
@@
 using the orthonormality of the spherical harmonics.
-Methods to calculate these coefficients exist, known as spectral transforms, and are well established for the spherical harmonics (see e.g.
+As discussed in \bsrefsection{section:introduction}, methods to calculate these coefficients exist known as spectral transforms, and are well established for the spherical harmonics (see e.g.
 \cite{mohlenkamp1999fast, suda2002fast}).
+Additionally, packages are available for public use that implement efficient spherical harmonics transforms, for example the FastTransforms package \cite{slevinsky2019fast}.
 
@@
 
-We start as before for the scalar case by finding $x \, \gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc.
-in terms of $\gradYlmprime (x,y,z)$, $\gradpYlmprime (x,y,z)$.
+We start as before for the scalar case by finding $x \, \gradYlm (x,y,z)$, $y \,\gradYlm (x,y,z)$, $z \, \gradYlm (x,y,z)$, and $x \,\gradpYlm (x,y,z)$, $y \, \gradpYlm (x,y,z)$, $z \,\gradpYlm (x,y,z)$ in terms of $\gradYlmprime (x,y,z)$, $\gradpYlmprime (x,y,z)$.
 A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations.
@@
 \subsubsection{Obtaining coefficients}
-\bstodoinline{As for the scalar case, do I need to say more here about the transforms?}
+
 As eluded to, we would like to be able to obtain the coefficients for the expansion of a vector valued function in the VSH basis up to a given degree, similar to how we approach the scalar case.
@@
 }
-where $f = 2 \rotrate \cos(\varphi) =  2 \rotrate z$ is the Coriolis parameter, and $\rotrate$ is the rotation rate of the Earth (note that $\varphi = {\pi \over 2} - \text{latitude}$ \bstodo{make a footnote}).
-On a full scale Earth, the rotation rate is $\rotrate = 7.2921 × 10^{-5}$ rad s$^{-1}$ \bstodo{make a footnote}.
-
-A keen eye may notice that -- while we have operators for the multiplication by z, the spherical gradient and divergence -- we require one more operator not yet defined for the unit vector cross product $\rvec \times \uvec$.
+where $f = 2 \rotrate \cos(\varphi) =  2 \rotrate z$ is the Coriolis parameter\footnote{Normally, the Coriolis parameter is expressed in terms of the latitude: $\text{latitude} = {\pi \over 2} - \varphi$}, and $\rotrate$ is the rotation rate of the sphere surface (that is, the amount the body rotates per unit time)\footnote{On a full scale Earth, the rotation rate is $\rotrate = 7.2921 × 10^{-5}$ rad s$^{-1}$.}.
+
+A keen eye may notice that -- while we have operators for the multiplication by $z$, the spherical gradient and divergence -- we require one more operator not yet defined for the unit vector cross product $\rvec \times \uvec$.
 \begin{definition}\label{def:unitveccrossproduct}

### 2021-04-27 papers/thesis/full/ch1-introduction/introduction.tex
Addressed more of SO comments

@@
 The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
+There are also potential applications in physics, where
 
-Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used for many applications involving PDEs \cite{}\bstodo{cite applications of SHs}.
-Notably, they are used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
+Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used for many applications including in computer graphics (e.g.
+\cite{moon2008efficient, sloan2013efficient}), physics (e.g.
+\cite{vasil2019tensor}), quantum theory \cite{varshalovich1988quantum} and geosciences \bstodo{cite applications of SHs}.
+Notably, they are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
 While the whole sphere spectral method using the Spherical Harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast, diamantakisecmwf}.

### 2021-04-27 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Addressed more of SO comments

@@
 
-Jacobi operators that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$.
+Jacobi operators are matrices that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$.
 We can obtain the entries to these operator matrices by determining the coefficients of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.
-
-First though, we require a few results concerning the incrementing and decrementing of the parameters for the Jacobi polynomials, and the three-term recurrence for the Jacobi polynomials, that will also prove useful later on in this chapter.
+Relations concerning the products of Spherical Harmonics have already been established, and so the expressions we desire here could just be seen as special cases of these.
+More specifically, the coefficients of the expansion in the SH basis of the product of two SHs are composed of Clebsch-Gordan coefficients \cite[p231]{sakurai1995modern}.
+Clebsch-Gordan coefficients are important in quantum mechanics and angular momentum, with tables and algorithms existing to calculate them \cite{mcnamee1964tables}.
+By simply choosing appropriate combinations, one could gain the desired coefficients for the expansion of each of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$.
+
+However we reiterate our goal here is to use techniques that can generalise to spectral methods on other domains using multidimensional OPs, and thus we proceed by exploiting the three-term recurrences and other relations of the univariate OPs that make up the Spherical Harmonics as we have defined them.
+With that being said, we require a few results concerning the incrementing and decrementing of the parameters for the Jacobi polynomials, and the three-term recurrence for the Jacobi polynomials, that will also prove useful later on in this chapter.
 \begin{lemma}
-	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
-\bseqn{
-	P_{n}^{(\alpha, \beta)}(z) =
-		{1 \over 2n + \alpha + \beta + 1} \Big\{ & - {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \\
-									  & + (n+\alpha+\beta+1) \Big[{n+\alpha \over 2n+\alpha+\beta} - {n+\beta+1 \over 2n+\alpha+\beta+2}\Big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
-									  & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
-	(1-z^2) \, P_{n}^{(\alpha, \beta)}(z) =
-		{4 \over 2n + \alpha + \beta + 1} \Big\{ & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n}^{(\alpha-1, \beta-1)}(z) \\
-									  & + (n+1) \Big[{n+\alpha+1 \over 2n+\alpha+\beta+2} - {n+\beta \over 2n+\alpha+\beta}\Big] \: P_{n+1}^{(\alpha-1, \beta-1)}(z) \\
-									  & - {(n+1)(n+2) \over 2n+\alpha+\beta+2} \: P_{n+2}^{(\alpha-1, \beta-1)}(z) \Big\}, \\
-	z \, P_{n}^{(\alpha, \beta)}(z) = 
-		& {2(n+\alpha)(n+\beta) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 1)} \: P_{n-1}^{(\alpha, \beta)}(z) \\
-		& {(\beta^2 - \alpha^2) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 2)} \: P_{n}^{(\alpha, \beta)}(z) \\
-		& {2(n+1)(n + \alpha + \beta + 1) \over 2(2n + \alpha + \beta + 1)(2n + \alpha + \beta + 2)} \: P_{n+1}^{(\alpha, \beta)}(z).
+	Let $\alpha, \beta \in \R$ be general parameters.
+Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$.
+Then, for any $n \in \N_0$ the following identities hold:
+\bseqn{
+	&1)\quad P_{n}^{(\alpha, \beta)}(z) = {1 \over 2n + \alpha + \beta + 1} \Big\{- {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \\
+									  					& \quadfour \quadtwo \quad + (n+\alpha+\beta+1) \Big[{n+\alpha \over 2n+\alpha+\beta} - {n+\beta+1 \over 2n+\alpha+\beta+2}\Big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
+									 					& \quadfour \quadtwo \quad + {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \Big\}, \\ \\
+	&2)\quad (1-z^2) \, P_{n}^{(\alpha, \beta)}(z) = {4 \over 2n + \alpha + \beta + 1} \Big\{{(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n}^{(\alpha-1, \beta-1)}(z) \\
+									 							& \quadeight \quadtwo \: + (n+1) \Big[{n+\alpha+1 \over 2n+\alpha+\beta+2} - {n+\beta \over 2n+\alpha+\beta}\Big] \: P_{n+1}^{(\alpha-1, \beta-1)}(z) \\
+									 							& \quadeight \quadtwo \: - {(n+1)(n+2) \over 2n+\alpha+\beta+2} \: P_{n+2}^{(\alpha-1, \beta-1)}(z) \Big\}, \\ \\
+	&3)\quad z \, P_{n}^{(\alpha, \beta)}(z) = {2(n+\alpha)(n+\beta) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 1)} \: P_{n-1}^{(\alpha, \beta)}(z) \\
+								& \quadeight + {(\beta^2 - \alpha^2) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 2)} \: P_{n}^{(\alpha, \beta)}(z) \\
+								& \quadeight + {2(n+1)(n + \alpha + \beta + 1) \over 2(2n + \alpha + \beta + 1)(2n + \alpha + \beta + 2)} \: P_{n+1}^{(\alpha, \beta)}(z).
 }
@@
 \bseqn{
-	\tilde \alpha_{l,m,1} &:= {2l \over 2l+1} \\
-	\tilde \alpha_{l,m,2} &:= 
+	\tilde \alpha_{l,m,1} &:= {2l \over 2l+1}, \quad
+	&&\tilde \alpha_{l,m,2} := 
 		\begin{cases}
@@
 			\quad \quad 0 \quad \quad \text{otherwise}
-		\end{cases} \\
-	\tilde \alpha_{l,m,3} &:= - {2(l-m+2)(l-m+1) \over (2l+1)(l+1)} \\
-	\tilde \alpha_{l,m,4} &:= {(l+m+2)(l+m+1) \over 2(2l+1)(l+1)} \\
+		\end{cases}, \\
+	\tilde \alpha_{l,m,3} &:= - {2(l-m+2)(l-m+1) \over (2l+1)(l+1)}, \quad
+	&&\tilde \alpha_{l,m,4} := {(l+m+2)(l+m+1) \over 2(2l+1)(l+1)}, \\
 	\tilde \gamma_{l,m,1} &:= \begin{cases}
@@
 			\quad 0 \quad \quad \text{otherwise}
-		\end{cases} \\
-	\tilde \gamma_{l,m,2} &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
+		\end{cases}, \quad
+	&&\tilde \gamma_{l,m,2} := \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
 }
 \end{corollary}
-
 \bsrefcor{corollary:jacobirelations} allows us to easily determine the aforementioned coefficients, that we will formalise in the following Lemma.
@@
 \bseqn{
-	\alpha_{l,m,1} &:= 
-		\begin{cases}
-			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,m,1} \quad \text{if } m>0 \\
-			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,|m|,2} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
-			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
-		\end{cases} \\ 
-	\alpha_{l,m,2} &:= 
-		\begin{cases}
-			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,m,2} \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
-			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,|m|,1} \quad \text{if } m<0 \\
-			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
-		\end{cases} \\
-	\alpha_{l,m,3} &:= 
-		\begin{cases}
-			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,m,3} \quad \text{if } m>0 \\
-			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,|m|,4} \quad \text{if } m\le0 
-		\end{cases} \\
-	\alpha_{l,m,4} &:=
-		\begin{cases}
-			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
-			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,|m|,3} \quad \text{if } m<0 
-		\end{cases} \\
-	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4 \\
-	\gamma_{l,m,1} &:= \frac{\clm}{c^{m}_{l-1}} \: \tilde \gamma_{l,m,1} \\
-	\gamma_{l,m,2} &:= \frac{\clm}{c^{m}_{l+1}} \: \tilde \gamma_{l,m,2}.
+	\alpha_{l,m,1} &:= {\clm \over 2c^{m-1}_{l-1}}
+		\begin{cases}
+			\tilde \alpha_{l,m,1} &\text{if } m>0 \\
+			\tilde \alpha_{l,\abs{m},2} &\text{if } m\le0 \\
+			0 &\text{otherwise} 
+		\end{cases}, \quad
+	&&\alpha_{l,m,2} := {\clm \over 2c^{m+1}_{l-1}}
+		\begin{cases}
+			\tilde \alpha_{l,m,2} &\text{if } m\ge0 \\
+			\tilde \alpha_{l,\abs{m},1} &\text{if } m<0 \\
+			0 &\text{otherwise} 
+		\end{cases}, \\
+	\alpha_{l,m,3} &:= {\clm \over 2c^{m-1}_{l+1}}
+		\begin{cases}
+			\tilde \alpha_{l,m,3}  &\text{if } m>0 \\
+			\tilde \alpha_{l,\abs{m},4} &\text{if } m\le0 
+		\end{cases}, \quad
+	&&\alpha_{l,m,4} := {\clm \over 2c^{m+1}_{l+1}}
+		\begin{cases}
+			\tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
+			\tilde \alpha_{l,\abs{m},3} \quad \text{if } m<0 
+		\end{cases}, \\
+	\gamma_{l,m,1} &:= \frac{\clm}{c^{m}_{l-1}} \: \tilde \gamma_{l,m,1}, \quad
+	&&\gamma_{l,m,2} := \frac{\clm}{c^{m}_{l+1}} \: \tilde \gamma_{l,m,2}, \\
+	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4.
 }
@@
 
-The results in \bsreflemma{lemma:shxyzrelations} are fairly involved, and can be derived via \textit{Clebsch-Gordan coefficients}.
+\remark The results in \bsreflemma{lemma:shxyzrelations} are fairly involved and can be derived via the relationships established for products of Spherical Harmonics and Clebsch-Gordan coefficients as mentioned earlier, but we include this proof so as to generalise to other OPs later on in later chapters.
 
@@
 
-Three-term recurrence relations for orthogonal polynomials are well established (e.g.
-\cite[18.9]{DLMF}, \cite{dunkl2014orthogonal}).
-In a similar vein, we can obtain a recurrence relation for the spherical harmonics by combining each system in (\ref{eqn:jacobimatsdef}).
+Three-term recurrence relations for orthogonal polynomials are well established (see for example \cite[Section 18.9]{DLMF}), including for multidimensional OPs \cite[Section 3.2]{dunkl2014orthogonal}.
+In a similar vein, we can obtain a recurrence relation for the Spherical Harmonics by combining each system in \bsrefeqn{eqn:jacobimatsdef} in a specific way.
+It should be emphasised at this point that this is not necessarily the optimal way of computing the Spherical Harmonics; various efficient methods and algorithms have been proposed (see e.g.
+\cite{guy1998efficient, sloan2013efficient, fukushima2012numerical}).
+We present this method in order to be able to generalise it to other multidimensional OP families later on in this thesis.
+
+Rewriting \bsrefeqn{eqn:jacobimatsdef}, we have that:
+\bseqn{
+	(J_x - xI)\SHvec(x,y,z) = (J_y - yI)\SHvec(x,y,z) = (J_z - zI)\SHvec(x,y,z) = \vec{0}.
+}
+Combining these relationships into a single system, complete with initial (degree 0) value, results in:
 \bseqn{
@@
 where we note $Y_0^0(x,y,z) \equiv Y_0 := c_0^0 \: P_0^{(0,0)} \equiv \half {1 \over \sqrt{\pi}}$, and for each $l = 0,1,2\dots$,
-\begin{align}
+\bseqnnumber{
 	A_l &:= 
@@
 		\end{pmatrix}  \in \C^{3(2l+1)\times(2l+1)}. \label{eqn:clenshawmats2}
-\end{align}
+}
 
@@
 such that $\Dlt A_l = \idmat{2l+3}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[Section 3.2]{dunkl2014orthogonal}, which can be expanded to obtain the three-term recurrence:
 \bseqn{
@@
 
-%%% TANGENT SPACE ----- VECTOR SHs
+%%% TANGENT BUNDLE ----- VECTOR SHs
 \section{Vector spherical harmonics as orthogonal vectors in three variables}\label{Section:VSHs}
 
-Since the spherical harmonics are a basis for the surface of the sphere, and the tangent space of the sphere is spanned by the gradient and perpendicular gradient of a scalar function, we have that the gradients and perpendicular gradients of the spherical harmonics
+We have established that the Spherical Harmonics can be viewed as a basis of orthonormal polynomials on the sphere.
+Specifically, given a function $f : \Omega \to \C$, we can write $f = \sum_{l = 0}^{N} \sum_{m = -l}^{l} f_{l,m} \; \Ylm$ where $\{f_{l,m}\}$ are coefficients and $N \in \N$ is sufficiently large.
+But what about vector-valued functions?
+\begin{definition}\label{def:tangentbundle}
+	Let $(x,y,z) \in \Omega$.
+We refer to the \textbf{tangent space} of the point $(x,y,z)$ as the space of all vectors that are orthogonal to the vector $\rvec := [x,y,x]^\top$.
+We further define the \textbf{tangent bundle} of $\Omega$ as the set of all such spaces for every point in $\Omega$.
+\end{definition}
+For the purposes of this discussion, we will only consider the tangent bundle, and thus wish to write down a basis for vector-valued functions with values in the tangent bundle.
+Fortunately, there is a simple solution known as the \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}:
 \bseqnnumber{
 	\gradYlm &:= \nabla \Ylm = \ppx{\phi} \Ylm \phivec + {1 \over \sinphi} \ppx{\theta} \Ylm \thetavec, \nonumber \\
-	\gradpYlm &:= \nabla^\perp \Ylm \equiv \rvec \times \nabla \Ylm = \ppx{\phi} \Ylm \thetavec - {1 \over \sinphi} \ppx{\theta} \Ylm \phivec, \label{eqn:VSHs}
-}
-span the tangent space.
+	\gradpYlm &:= \nabla^\perp \Ylm \equiv \rvec \times \nabla \Ylm = \ppx{\phi} \Ylm \thetavec - {1 \over \sinphi} \ppx{\theta} \Ylm \phivec. \label{eqn:VSHs}
+}
 Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$, while $\phivec, \thetavec$ are the standard spherical coordinate unit vectors, i.e.
 $\phivec := [\costheta \cosphi, \sintheta \cosphi, -\sinphi]^\top$, $\thetavec := [-\sintheta, \costheta, 0]^\top$.
-Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for vector valued functions in the tangent space of the sphere and are known as \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}, orthogonal with respect to the inner product $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$ -- in particular, for any $l, l' \in \N$, $m, m' \in \Z$ s.t.
+They are orthogonal with respect to the inner product $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$ -- in particular, for any $l, l' \in \N$, $m, m' \in \Z$ s.t.
 $-l \le m \le l$, $-l' \le m' \le l'$,
@@
 	\ip<\gradYlm, \gradpYlmprime> &= \iint_\Omega \gradpYlm \cdot {\gradpYlmprime}^* \: \sinphi \: \D\varphi \: \D\theta = 0.
+} 
+Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
+\bseqn{
+	\uvec(x,y,z) = u_\varphi(x,y,z) \: \phivec + u_\theta(x,y,z) \: \thetavec, \quad (x,y,z) \in \Omega
+}
+for some scalar functions $u^\varphi, u^\theta$.
+Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent bundle} of the sphere.
+Importantly, the set $\{\gradYlm, \gradpYlm\}$ form a complete and orthogonal basis for such functions \cite{barrera1985vector}.
+That is, given $\uvec$, there exist coefficients $\{u_{l,m}^{\gradYlm}, u_{l,m}^{\gradpYlm} \}$ such that for large enough $N$:
+\bseqn{
+	\uvec = \sum_{l = 0}^{N} \sum_{m = -l}^{l} u_{l,m}^{\gradYlm} \; \gradYlm + u_{l,m}^{\gradpYlm} \; \gradpYlm.
 }
@@
 Other definitions for VSHs are used (e.g.
-\cite{hill1954theory}), however it is convenient for deriving explicit sparse relations and operators to use the one described here.
+\cite{hill1954theory}), however it is convenient for deriving explicit sparse relations and operators to use the ones described here.
+In particular, we can note that computing the gradient or divergence of a function is merely a matter of conversion between the SH and the VSH bases.
 
 A simple calculation shows that such orthogonal vectors must still have block-tridiagonal Jacobi operators, as multiplication by $x, y$, or $z$ remains inside the ideal.
+Thus, we we can use the same techniques that we used for the scalar SHs to derive Jacobi matrices, function evaluation and transforms.
 
@@
 	&\ip<z \gradYlm, \gradYlmprime> \nonumber \\
-	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
+	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradYlmprime}^* \sinphi \: \D\varphi \: \D\theta \nonumber \\
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
-	&\quadeight + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: z \; \rho(z)^{\abs{m}+\abs{m'}-2} \: \D z \Big) \nonumber \\
-	&= \clm \clmprime 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m}+\abs{m'}-2} \big(mm' + \abs{m}\abs{m'} z^2 \big) \nonumber \\
-	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabs{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m}+\abs{m'}} \nonumber \\
-	&\quadeight \quadtwo - \abs{m'} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m}+\abs{m'}} \nonumber  \\
-	&\quadeight \quadtwo + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m}+\abs{m'}+2} \Big\} \: \D z \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ m^2 z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} (1 + z^2) \nonumber \\
-	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabs{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}} \nonumber \\
-	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber  \\
-	&\quadeight \quadtwo + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ m^2 z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} (1 + z^2) \nonumber \\
-	&\quadeight \quadtwo + \abs{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} \big[ z^2 \jaclmabsdz{l}{m} \rho^2 + 2z \jaclmabs{l'}{m} - 2\abs{m}z^3 \jaclmabs{l'}{m} \rho^2 \big] \nonumber \\
-	&\quadeight \quadtwo - \abs{m} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber  \\
-	&\quadeight \quadtwo + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber \\
-	&\quadeight \quadtwo + z d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l}{\abs{m}+1} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) \jaclmabs{l'}{m} \rho^{2\abs{m}} \big[ \tilde\gamma_{l,\abs{m},1} \jaclmabs{l-1}{m} + \tilde\gamma_{l,\abs{m},2} \jaclmabs{l+1}{m} \big] \nonumber \\
-	&\quadeight \quadtwo + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\
-	&= \clm 2 \pi \delta_{m',m} \Big\{ c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over 2\pi (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over 2\pi (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
-	&\quadfour c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over 2\pi (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over 2\pi (c_{l+1}^{\abs{m}+1})^2} \big] \Big\} \nonumber \\
-	&= \clm \delta_{m',m} \Big\{ c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
-	&\quadfour c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}.
-\label{eqn:lemmaVSHrecurrencesproofz}
-}
-We can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz} that the coefficients $\calC_{l',m'}$ are zero when $l' \ne l \pm 1$ or $m' \ne m \pm 1$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,1}$ and $\Gamma_{l,m,3}$.
+	&\quad \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
+	&\quadfour \quadtwo + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: z \; \rho(z)^{\abs{m}+\abs{m'}-2} \: \D z \Big) \label{eqn:lemmaVSHrecurrencesproofz1}
+}
+The integral over $\theta$ in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz1} is simply $2 \; \pi \; \delta_{m',m}$ showing the above is zero when $m' \ne m$, and thus so will the coefficients $\calC_{l',m'}$ be when this is the case.
+Focusing on the second integral (over $z$) when $m' = m$, this becomes:
+\bseqnnumber{
+	&\int_{-1}^{1} \Big\{ m^2 z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} (1 + z^2) \nonumber \\
+	&\quadfour - \abs{m} z^2 \jaclmabs{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour - \abs{m} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber  \\
+	&\quadfour + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m^2 z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} (1 + z^2) \nonumber \\
+	&\quadfour + \abs{m} \jaclmabs{l'}{m} \rho^{2\abs{m}-2} \big[ z^2 \jaclmabsdz{l}{m} \rho^2 + 2z \jaclmabs{l'}{m} - 2\abs{m}z^3 \jaclmabs{l'}{m} \rho^2 \big] \nonumber \\
+	&\quadfour - \abs{m} z^2 \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber  \\
+	&\quadfour + z \jaclmabsdz{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) z \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \nonumber \\
+	&\quadfour + z d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l}{\abs{m}+1} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) \jaclmabs{l'}{m} \rho^{2\abs{m}} \big[ \tilde\gamma_{l,\abs{m},1} \jaclmabs{l-1}{m} + \tilde\gamma_{l,\abs{m},2} \jaclmabs{l+1}{m} \big] \nonumber \\
+	&\quadtwo + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\
+	&= c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over 2\pi (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over 2\pi (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
+	&\quadtwo + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over 2\pi (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over 2\pi (c_{l+1}^{\abs{m}+1})^2} \big] \nonumber \\
+	&= c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
+	&\quadtwo + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big].
+\label{eqn:lemmaVSHrecurrencesproofz2}
+}
+Combining \bsrefeqn{eqn:lemmaVSHrecurrencesproofz2} with \bsrefeqn{eqn:lemmaVSHrecurrencesproofz1}, we obtain that
+\bseqnnumber{
+	\ip<z \gradYlm, \gradYlmprime> &= \clm \delta_{m',m} \Big\{ c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
+	&\quadfour + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}.
+\label{eqn:lemmaVSHrecurrencesproofz3}
+}
+We can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz3} that the nonzero coefficients $\calC_{l',m'}$ are those described in the Lemma as $\Gamma_{l,m,1}$ and $\Gamma_{l,m,3}$.
 
@@
 	&\ip<z \gradYlm, \gradpYlmprime> \nonumber \\
-	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
+	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\varphi \: \D\theta \nonumber \\
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
@@
 	&= \delta_{l',l} \delta_{m',m} i m {(\clm)^2 \over (\clmprime)^2} \nonumber \\
-	&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz2}
-}
-Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz2} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,2}$.
+	&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz4}
+}
+Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz4} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,2}$.
 
@@
 noting that $\rvec \times \gradYlm = \gradpYlm$ and $\rvec \times \gradpYlm = - \norm{\rvec} \gradYlm = - \gradYlm$ by the triple vector product and orthogonality.
-
 \end{proof}
 
-As for the scalar case, these coefficients detailed above are then the nonzero entries of the Jacobi matrices for the tangent space basis. Define $\VSHvecfull$ by
+As for the scalar case, these coefficients detailed above are then the nonzero entries of the Jacobi matrices for the VSHs. Define $\VSHvecfull$ by
 \bseqnnumber{
@@
 for each $(x,y,z) \in \Omega$.
-The tangent space Jacobi matrices are banded-block-banded due to the sparse relationships we obtain from \bsreflemma{lemma:VSHrecurrences}.
+The VSH Jacobi matrices are banded-block-banded due to the sparse relationships we obtain from \bsreflemma{lemma:VSHrecurrences}.
 Specifically, they each have block-bandwidths $(1,1)$:
@@
 
-It will be useful to be able to describe the coefficients of a vector-valued function on the tangent space of $\Omega$ in multiple ways, and as such we will introduce some more notation for the \enquote{splitting} of the basis vectors into the two groups stemming from the gradient and perpendicular gradient directions.
-Let's explain what we mean by this.
-Let $\uvec$ be a vector-valued function on $\Omega$ whose values are of the form
-\bseqn{
-	\uvec(x,y,z) = u_\varphi(x,y,z) \: \phivec + u_\theta(x,y,z) \: \thetavec, \quad (x,y,z) \in \Omega
-}
-for some scalar functions $u^\varphi, u^\theta$ where $\phivec, \thetavec$ are the standard non-radial unit basis vectors on the sphere defined by $\phivec := (\cos \theta \sin \varphi, \sin \theta \sin \varphi, \cos \varphi)^\top$, $\thetavec := (-\sin \theta, \cos \theta, 0)^\top$.
-Such a function $\uvec$ we refer to as a \textit{vector-valued function in the tangent space of $\Omega$}.
+It will be useful to be able to describe the coefficients of a vector-valued function in the tangent bundle of $\Omega$ in multiple ways, and as such we will introduce some more notation for the \enquote{splitting} of the basis vectors into the two groups stemming from the gradient and perpendicular gradient directions.
+Let us explain what we mean by this.
 
@@
 }
-If $\uvec$ is a vector-valued function on the tangent space of $\Omega$, it will have the coefficients vectors $\uvec^c, \uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, defined such that
+If $\uvec$ is a vector-valued function on the tangent bundle of $\Omega$, it will have the coefficients vectors $\uvec^c, \uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, defined such that
 \bseqn{
@@
 \subsubsection{Function evaluation}
-The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent space for a point $(x,y,z) \in \Omega$, where the matrices $C_n, B_n, \Dnt, G_n(x,y,z)$ are all defined in \bsrefsection{subsection:VSHrecurrence}:
+The Clenshaw algorithm presented in \bsrefsection{subsubsection:clenshaw} can also be used to evaluate a vector-valued function in the tangent bundle of $\Omega$, where the matrices $C_n, B_n, \Dnt, G_n(x,y,z)$ are all defined in \bsrefsection{subsection:VSHrecurrence}:
 \bseqn{
@@
 }
-Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs in our tangent space basis (and hence coefficients) for degree $n=0$ are zero.
+Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs (and hence the coefficients) for order $n=0$ are zero.
 
@@
 
-Let $\uvec$ be a vector-valued function on the tangent space of $\Omega$ with coefficients vectors $\uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, and let $f$ be a scalar function on $\Omega$ with coefficients vector $\fvecc$ for its expansion in the SSH basis also up to order $N$, i.e.
+Let $\uvec$ be a vector-valued function on the tangent bundle of $\Omega$ with coefficients vectors $\uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, and let $f$ be a scalar function on $\Omega$ with coefficients vector $\fvecc$ for its expansion in the SSH basis also up to order $N$, i.e.
 \bseqn{
@@
 
-For simplicity, we will implement a backward Euler timestepping method to solve the linear SWEs with timestep $\Delta t$:
+For simplicity, we will implement a backward Euler timestepping method to solve the linear SWEs with timestep $\Delta t$.
+Of course, other timestepping methods are equally implementable (for example, the method used in the IFS model that ECMWF uses is a semi-implicit, semi-Lagrangian (SISL) timestepping method \cite{diamantakisecmwf}), but we simplify things here to show proof of concept.
+Using $\uvec_n, h_n$ to represent the solution at time $n\Delta t$:
 \bseqn{
@@
 \bstodo{Images}
+
+\section{Conclusion}
+
+Spherical Harmonics have been widely used to solve PDEs, in particular a Spherical Harmonics approach is used as part of the Integrated Forecasting System (IFS) that ECMWF uses for their forecasts.
+In this chapter, we have chosen to view the Spherical Harmonics as orthogonal (in fact orthonormal) polynomials in $x$, $y$ and $z$ in order to illustrate and understand more about how the framework presented here for using multivariate OPs in a spectral method translate to the sphere.
+We have derived the entries of \textit{Jacobi matrices} that represent multiplication by the coordinates, showing how they are \textit{banded-block-banded} in structure.
+We have set out how one can find coefficients for the expansion of a function in the Spherical Harmonics basis using established transforms, and how one can evaluate a function given its coefficients vector using a multidimensional version of the Clenshaw algorithm.
+We extended our framework to include the \textit{tangent bundle} of the sphere -- that is, the space of vectors orthogonal to the unit outward normal for each point on the sphere -- where we can use the \textit{Vector Spherical Harmonics} as vector-valued orthogonal polynomials to be a basis, demonstrating how similar such Jacobi matrices, function expansion and evaluation can be derived.
+Further, by exploiting the natural sparsity of the relationships between the OPs in the bases, we have shown how matrices for differential and other operators will be sparse and banded-block-banded in structure too.
+In fact, for the Spherical Harmonics, these are simply diagonal.
+Finally, we presented a simple example of how one can put this into practice for solving the linearised shallow water equations.

### 2021-04-28 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Minor edit

@@
 
-\remark The results in \bsreflemma{lemma:shxyzrelations} are fairly involved and can be derived via the relationships established for products of Spherical Harmonics and Clebsch-Gordan coefficients as mentioned earlier, but we include this proof so as to generalise to other OPs later on in later chapters.
+\remark The results in \bsreflemma{lemma:shxyzrelations} are fairly involved and can be derived via the relationships established for products of Spherical Harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
 

### 2021-04-29 papers/thesis/full/abstract/abstract.tex
Edit to abstract

@@
 This thesis develops sparse spectral methods for solving partial differential equations (PDEs) on various multidimensional domains, with a specific focus on the disk-slice and trapezium in 2D, and the spherical cap as a surface in 3D.
+For the latter scenario, the PDEs are surface PDEs involving Laplace-Beltrami operators, spherical gradients and other spherical operators.
 
-We begin with an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface.
-We present a framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
+Throughout this thesis we wish to utilise our understanding of the orthogonal polynomials and sparse spectral methods to build up to new methods for subdomains of the sphere surface.
+To this end, we begin with an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface.
+We present a framework for how the spherical harmonics can be thought of as a set of multidimensional orthogonal polynomials in $x$, $y$ and $z$, and can be used to expand functions defined on the sphere.
+We explain how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
+This way of thinking about the spherical harmonics differs from the traditional sense (where they are considered as orthogonal functions in spherical coordinates) but we do so as to be able to generalise the techniques to spectral methods on other domains using multidimensional OPs.
 Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.
 
-Throughout this thesis we wish to utilise our understanding of the 3D spherical harmonics to build up to a new orthogonal polynomial basis for subdomains of the sphere surface.
-To this end, we move on working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+We move on working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
@@
 
-Finally, we adapt the tools developed in two dimensions to derive a similar OP basis complete with sparse differential operators.
+Finally, we return the the surface of the sphere, namely the subdomain we call the spherical cap.
+Piecing together techniques and OPs from both the Spherical Harmonics and disk-slice work, we once again present a new orthogonal polynomial basis and sparse spectral method for the spherical cap, complete with the same observation about the guaranteed sparsity of differential and other operators.
+The motivation is for one to use spherical caps and spherical bands (which are a simple extension) as elements in a spectral element method for the sphere, with many applications in meteorology and astrophysics -- in particular, as a potential replacement of the Spherical Harmonics approach that is currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) which is predicted in the future to be too costly in the due to a parallel scalability bottleneck arising from the global spectral transform.
 

### 2021-04-29 papers/thesis/full/ch1-introduction/introduction.tex
Some more edits to chapters one and two

@@
 
-Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs} \bstodo{citation here}.
-For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations (see \cite{boyd2001chebyshev, olver2013fast} and references therein).
+Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs} \bstodo{citations here}.
+For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations (see e.g.
+\cite{boyd2001chebyshev, olver2013fast} \bstodo{add more from the references in these papers}).
 While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[Section 18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
 
-One could say this is surprising, given that there is a long history of around 150 years associated with multivariate OPs beginning with Hermite first presenting the multivariate Hermite polynomials in 1865 \cite{appel1926fonctions, ismail2017review}.
-Zernike polynomials \cite{zernike1934diffraction} were first introduced in 1934, for example, that are a group of bivariate polynomials orthogonal on a unit circle.
+One could say this is surprising, given that there is a long history of around 150 years associated with multivariate OPs, beginning with Hermite first presenting the multivariate Hermite polynomials in 1865 \cite{appel1926fonctions, ismail2017review}.
+Zernike polynomials \cite{zernike1934diffraction} were first introduced in 1934, as another example, that are a group of bivariate polynomials orthogonal on a unit circle.
 Koornwinder in 1975 described a method for constructing two-variable OPs from univariate OPs \cite{koornwinder1975two}.
@@
 The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
-There are also potential applications in physics, where
 
-Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used for many applications including in computer graphics (e.g.
-\cite{moon2008efficient, sloan2013efficient}), physics (e.g.
+Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used in many fields including computer graphics (e.g.
+\cite{moon2008efficient, sloan2013efficient}), astrophysics (e.g.
 \cite{vasil2019tensor}), quantum theory \cite{varshalovich1988quantum} and geosciences \bstodo{cite applications of SHs}.
@@
 
+There are also potential applications in physics too, particularly in astrophysics, where solving equations on the sphere and working in spherical geometries is also desirable \cite{vasil2019tensor}\bstodo{cite more, see p2 of Vasil and the references}.
+Recently, a method for computing tensor fields in spherical coordinates using Jacobi polynomials has been proposed \cite{vasil2019tensor}.
+In this work the authors present the method for both the surface of the unit sphere and the three-dimensional generalisation of the unit ball.
+\bstodoinline{Describe/summarise the Vasil 2019 Tensor paper approach. Compare to our approach (spectral method with OPs on spherical sub-domains).}
+
+The structure of this thesis is as follows.
+
 Chapter 2 of this thesis provides an introduction to sparse spectral methods via the Spherical Harmonics on the whole sphere surface.

### 2021-04-29 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Some more edits to chapters one and two

@@
 }
-Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[14.21.1]{DLMF}:
+Legendre polynomials are special cases of \textit{Legendre functions} when $l$ is a non-negative integer.
+Legendre functions are a class of univariate functions that are solutions to the differential equation
+\bseqn{
+	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x} y + l(l+1)y = 0
+}
+which is helpfully known as the \textit{Legendre equation} \cite[14.2.1]{DLMF}.
+Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p5]{belousov1962tables}:
 \bseqn{
@@
 for $m = -l, \dots, -1$.
-Although often referred to as the associated Legendre \enquote{polynomials} as we do here, they are not strictly polynomials when $m$ is odd, and as such some authors refer to them as associated Legendre functions instead.
+
+Once again, associated Legendre polynomials are special cases of \textit{associated Legendre functions} when $l$ is a non-negative integer and $m \in \{0,\dots,l\}$, which are also a class of univariate functions that are solutions to a modified version of the Legendre equation
+\bseqn{
+	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x}y + \Big[l(l+1) - {m^2 \over 1-x^2} \Big] y = 0
+}
+which is also helpfully known as the \textit{associated Legendre equation} (\cite[p2]{belousov1962tables}, \cite[14.2.2]{DLMF}).
+We can see that Legendre functions (and polynomials) are just special cases of the associated versions when $m = 0$.
+
+Although often referred to as the associated Legendre \enquote{polynomials} as we do here when $l$ is a non-negative integer and $m \in \{0,\dots,l\}$, $\Plm$ are not strictly polynomials when $m$ is odd, and as such some authors refer to them as associated Legendre functions regardless instead.
 Moreover, it should be noted too that the inclusion of the $(-1)^m$ factor in the definition for $m > 0$, known as the \textit{Condon-Shortley phase} in physics, is sometimes omitted.
-\bstodoinline{I have cited DLMF on ALFs equation that defines the ALFs.
-Now need to explain how that relates to the definition in terms of mth derivative of normal Legendre poly when l,m are integers}
 

### 2021-05-04 papers/spherical-caps/revised/response-to-referees.tex
response edits

@@
 
-Dear Jose Bautista Francisco III
-
-Thank you for forwarding the reviews of our article.
-We greatly appreciate your time as well as that of the referees.
-We have made changes to reflect the comments and suggestions of the referees.
-Below we detail our response to the referees' reports.
+We greatly appreciate the thorough comments from the editor and referees, and have made changes as requested. Below we detail our response to the referees' reports.
 
@@
 
-We have added in a discussion of the closest point method.
-We have chosen to omit discussion on radial basis functions as we do not feel there is particularly suitable to reference and appropriate to discuss.
+We have added in a discussion of the closest point method. We have chosen to omit discussion on radial basis functions as we are not aware of references that consider their usage on spherical caps.
 
@@
 
-Added comment a small subsection to Section 5 to address this. \bstodoinline{Is this enough?}
+Added comment a small subsection to Section 5 to address this.
 

### 2021-05-04 papers/spherical-caps/revised/spherical-caps.tex
response edits

@@
 Other boundary conditions beyond those already presented here are beyond the scope of this paper.
-However, the approach established for the case of a triangle domain in 2D shows how Neumann conditions can be done by writing these as a system of equations in partial derivatives \cite{olver2019triangle}.
+However, the approach established for the case of a triangle domain in 2D shows how Neumann conditions can be considered  as a system of equations in partial derivatives involving Dirchlet conditions \cite{olver2019triangle}.
 This approach may also work here for the spherical cap.

### 2021-05-05 papers/spherical-caps/revised/spherical-caps.tex
Minor edits and typos, and new figs

@@
 A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
-This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity} with $v(x,y,z) = \cos z$.
+This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure~\ref{fig:complexity} with $v(x,y,z) = \cos z$.
 
@@
 is a normalising constant.
-Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}). That is,
+Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in equation (\ref{eqn:ipgenjac}).
+That is,
 \begin{align}
@@
 where the coefficients are calculatable (see \cite{snowball2019sparse}).
-We can use (\ref{eqn:Rrecurrence}) to determine the 3D recurrences for $\scopnkia(x,y,z)$.
+We can use equation (\ref{eqn:Rrecurrence}) to determine the 3D recurrences for $\scopnkia(x,y,z)$.
 Importantly, we can deduce sparsity in the recurrence relationships.
@@
 \begin{proof}
-	The 3-term recurrence for multiplication by $z$ follows from (\ref{eqn:Rrecurrence}).
+	The 3-term recurrence for multiplication by $z$ follows from equation (\ref{eqn:Rrecurrence}).
 For the recurrence for multiplication by $x$, since $\{\scopmjha\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n+1$ polynomial on $\Omega$, we can expand
@@
 A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here.
-Combining each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yeilds:
+Combining each system in equation (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yeilds:
 \begin{align*}
@@
 \end{align*}
-Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure~\ref{fig:sparsity}.
 
@@
 This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
-The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
+The sparsity and structure of this biharmonic operator are seen in Figure~\ref{fig:sparsity}.
 
@@
 	\begin{subfigure}[b]{0.55\textwidth}
-		\includegraphics[scale=0.4]{condition-numbers-laplacian-N=200}
+		\includegraphics[scale=0.4]{condition-numbers-laplacian-log-N=200}
 	\end{subfigure}
@@
 	\begin{subfigure}[b]{0.55\textwidth}
-		\includegraphics[scale=0.4]{condition-numbers-max-laplacian}
+		\includegraphics[scale=0.4]{condition-numbers-max-laplacian-log}
 	\end{subfigure}
 	\hfill%
-	\caption{Plots of the condition number for the Laplacian operator matrix $\Delta^{(1)}_W$ and the preconditioned matrix $P^{-1} \; \Delta^{(1)}_W$ where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
-Left: The condition numbers of each diagonal block of $\Delta^{(1)}_W$ for $N = 200$, as well as those for $P^{-1} \; \Delta^{(1)}_W$.
-Right: The maximum condition number of all diagonal blocks of the Laplacian and the preconditioned Laplacian operators, as the degree $N$ increases.}
+	\caption{Plots of the condition number for the Laplacian operator matrix $\Delta^{(1)}_W$ and the preconditioned matrix $P^{-1} \Delta^{(1)}_W$ where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
+Top: The condition numbers of each diagonal block of $\Delta^{(1)}_W$ for $N = 200$, as well as those for $P^{-1} \Delta^{(1)}_W$.
+Bottom: The maximum condition number of all diagonal blocks of the Laplacian and the preconditioned Laplacian operators, as the degree $N$ increases.}
 	\label{fig:conditionnumbers}
@@
 Recall that the Laplacian operator is block-diagonal.
-One can therefore look at the condition numbers of each block on the diagonal as proxy for the condition number of the Laplacian (and hence the Poisson equation), where graphical representations are seen in \ref{fig:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
-As the degree $N$ increases, the condition numbers of these blocks becomes increasingly large (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
-Fortunately however, by applying a trivial preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound the condition numbers -- this is also seen in \ref{fig:conditionnumbers}.
+One can therefore look at the condition numbers of each block on the diagonal as proxy for the condition number of the Laplacian (and hence the Poisson equation), where graphical representations are seen in Figure~\ref{fig:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
+As the degree $N$ increases, the condition numbers of these blocks becomes grows fast algebraically (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
+Fortunately however, by applying a trivial preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in Figure Figure~\ref{fig:conditionnumbers}.
 The preconditioning matrix chosen is $P^{-1}$, where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
@@
 which can be solved to find $\mathbf{u}$.
-In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
+In Figure~\ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in equation (\ref{eqn:poisson}) in the disk-slice $\Omega$.
+In Figure~\ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
 The right hand sides we choose here are given by
@@
 which can be solved to find $\mathbf{u}$.
-We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
+We can see the sparsity and structure of this matrix system in Figure~\ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
+In Figure~\ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in equation (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
+In Figure~\ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
 Here, we use $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
@@
 For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see Section \ref{subsection:furtherdiffoperators}).
-In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
+In Figure~\ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
+In Figure~\ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
 Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve exponential convergence for these more complex functions.

### 2021-05-05 papers/spherical-caps/revised/spherical-caps.tex
revision changes

@@
 The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
-We demonstrate the results on partial differential equations involving the spherical Laplacian and biharmonic operators, showing spectral convergence.
+We demonstrate the results on partial differential equations involving the spherical Laplacian and biharmonic operators, showing spectral convergence with  discretisations that can be made well-conditioned using a simple preconditioner.
 \end{abstract}
@@
 The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper.
-These adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same.
+These adjustments make the construction more involved, which is why they are omitted here, but the approach is essentially the same.
 \end{remark}
@@
 There are of course other approaches for solving PDEs on surfaces, such as the closest point method \cite{macdonald2010implicit, macdonald2008level}.
-This involves solving the PDE inside a 3D volume that acts as a \enquote{shell} for the surface, and using a \enquote{closest point} to as the solution on the surface.
+This involves recasting the PDE as one involving a \enquote{closest point} operator in a 3D volume that acts as a \enquote{shell} for the surface.
 While such an approach is useful for other geometries, it is of low order and does not achieve spectral convergence with sparse discretisations.
-Further, for our interest of the sphere, it would not take advantage of any rotational symmetry.
+Further, for our interest of the sphere, it would not take advantage of any rotational symmetry to achieve optimal complexity with a direct solver.
 
@@
 We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
 As in the  disk-slice case in 2D \cite{snowball2019sparse}, we use an integration-by-parts argument to deduce the sparsity structure.
@@
 Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + v(x,y,z)$ where $v$ is first approximated by a polynomial.
-A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
+A nice feature  is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
 This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure~\ref{fig:complexity} with $v(x,y,z) = \cos z$.
@@
 
-The code that allows one to produce the numerical examples in this paper is publicly available as a Julia package\footnote{https://github.com/snowball13/OrthogonalPolynomialFamilies} to partner the ApproxFun package \cite{ApproxFun} -- however, it is strongly warned that the package is purely experimental at this stage.
+The code that allows one to produce the numerical examples in this paper is publicly available as a Julia package\footnote{https://github.com/snowball13/OrthogonalPolynomialFamilies.jl} to partner the ApproxFun package \cite{ApproxFun} -- however, this package is purely experimental at this stage.
 
@@
 \end{proposition}
-Of course, \enquote{orthogonal polynomials} are defined as such by the condition that each polynomial is orthogonal to each other, with respect to the given inner product.
+\enquote{Orthogonal polynomials} are defined as such by the condition that each polynomial is orthogonal to each other, with respect to the given inner product.
+In particular, they are orthogonal with respect to all lower degree polynomials.
 
@@
 
-The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$.
+The recurrences in Lemma \ref{lemma:Qrecurrence} lead to (block) Jacobi matrices that correspond to multiplication by $x$, $y$ and $z$.
 In later sections we will use an ordering of the OPs so that they are grouped by Fourier mode $k$, which is convenient for the application of differential and other operators to the vector of coefficients of a given function's expansion (some operators will exploit this ordering for operators where Fourier modes do not interact, and thus will be block-diagonal).
-Before that though, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
+Before that, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
@@
 A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here.
-Combining each system in equation (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yeilds:
+Combining each system in equation (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yields:
 \begin{align*}
@@
 	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \\
-	\rho(z) \: \rho'(z) &= -z \\
+	\rho(z) \: \rho'(z) &= -z, \\
 	\ppphi \scopnkia(x,y,z) &= -\rho(z) \: \ddz \Big[ \rho(z)^k \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta), \\
@@
 It is therefore necessary to also be able to raise or lower the parameter by way of an independent operator.
-There exist conversion matrix operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+There exist conversion  operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 \begin{definition}\label{def:parametertransformationoperators}
@@
 
-For a given problem $A\vec{x} = \vec{b}$, the condition number is a measure of how accurate solving the system will be (ignoring any numerical rounding errors), and is given as the condition number of the matrix $A$:
-\begin{align*}
-	\kappa(A) := \norm{A}_2 \norm{A^{-1}}_2.
-\end{align*}
-Recall that the Laplacian operator is block-diagonal.
-One can therefore look at the condition numbers of each block on the diagonal as proxy for the condition number of the Laplacian (and hence the Poisson equation), where graphical representations are seen in Figure~\ref{fig:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
-As the degree $N$ increases, the condition numbers of these blocks becomes grows fast algebraically (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
-Fortunately however, by applying a trivial preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in Figure Figure~\ref{fig:conditionnumbers}.
+Denote the condition number of the matrix $A$: 
+\begin{align*}
+	\kappa(A) := \norm{A}_2 \norm{A^{-1}}_2,
+\end{align*}
+which encodes how accurate the solution to $A \vc x = \vc b$ is under perturbation.
+Since the Laplacian operator is block-diagonal one can determine the condition number from the condition numbers of each block on the diagonal,  and hence deduce the accuracy in solving the Poisson equation, where graphical representations are seen in Figure~\ref{fig:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
+As the degree $N$ increases, the condition numbers of these blocks becomes grows  algebraically fast (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
+Fortunately however, by applying a trivial diagonal preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in Figure Figure~\ref{fig:conditionnumbers}.
 The preconditioning matrix chosen is $P^{-1}$, where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
@@
 
-In this section we construct a quadrature rule exact for polynomials on the spherical cap $\Omega$ that can be used to expand functions in the OPs $\scopnkia(x,y,z)$ for a given parameter $a$.
+In this section we construct a quadrature rule that is exact for polynomials on the spherical cap $\Omega$ that can be used to expand functions in the OPs $\scopnkia(x,y,z)$ for a given parameter $a$.
 
@@
 
-\remark{While it may be possible to leverage a discrete cosine transform (DCT) to speed up the calculation of the coefficients  \cite{slevinsky2018use}, this is an open question.}
+\remark{While it may be possible to leverage fast transforms, as in \cite{slevinsky2018use,slevinsky2019fast}, to speed up the calculation of the coefficients, this remains an open question.}
 
@@
 	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+	\quad &\text{3) } \text{Output: } V = v(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
 \end{align*}
@@
 since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only, which are precisely the Fourier coefficients of $c(\cos \theta, \sin \theta)$.
-Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manor.
+Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manner.
 
@@
 
-Our last erxample is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
+Our last example is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
 \begin{align}
@@
 
-This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere---it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
+This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere such as spherical triangles.
+It is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct the necessary orthogonal polynomials.
+Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct, or perhaps {\it compute}, the necessary orthogonal polynomials.
 
 The next stage is to develop an orthogonal basis for the tangent space of the spherical cap (or band), and obtain sparse differential operators for gradient, divergence etc.
-On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
+On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the construction of a basis will not be as straightforward.
 

### 2021-05-09 papers/spherical-caps/revised/response-to-referees.tex
resubmitted version

@@
 
+\item\comment{If there is a key technical issue that the authors overcome to allow this construction to proceed it is not obvious from the paper, meaning that the reader is left with the impression that, of course, the previously developed sparse spectral methods on 2D planar domains should extend}
+
+Handling partial differential operators on the sphere is significantly more challenging than partial differential operators in Euclidean space.
+This is similar to the significantly increased challenge faced in solving PDEs on spheres (e.g.
+[22]) compared with disks (e.g.
+[21]).
+Note further that the construction for 2D planar domains did not take advantage of rotational symmetry.
+
 \item \comment{The authors spend an inadaquate amount of time discussing other approaches for solving this problem and comparing their methodology (e.g., closest point method, radial basis functions).
@@
 
-We have added in a discussion of the closest point method. We have chosen to omit discussion on radial basis functions as we are not aware of references that consider their usage on spherical caps.
+We have added  a brief discussion of the closest point method.
+Note that the closest point method does not achieve spectral accuracy so is not really comparable to our approach.
+Further it does not take advantage of rotational symmetry which means that direct solvers will not have optimal complexity.
+We have chosen to omit discussion on radial basis functions: while they have been used with great success on a full sphere, we are not aware of references that consider their usage on spherical caps.
+If the referee can point us to relevant references that would be very much  appreciated.
 
@@
 Added a comment before Definition 2 to explain this, as well as an extra line in Definition 2.
+Please note that in more than 1D this condition on its  own is only sufficient to define the span of the degree $n$ polynomials: the choice of basis for the degree $n$ polynomials is not unique.
+Definition 2 therefore determines our choice of basis.
 
@@
 Added small subsection and plots to show the condition numbers of each m-block of Laplacian operator, along with a preconditioned Laplacian (where the preconditioner $P$ is the inverse of the diagonal of the Laplacian).
-While the Laplacian does not have bounded condition number, the trivial preconditioning keeps it bounded.
+While the Laplacian itself has a growing condition number, the trivial preconditioning makes it well-conditioned.
 
@@
 
-The evidence we can provide is that which addresses the first question.
+We have added numerical evidence demonstrating that the discretisation is well-conditioned under a simple preconditioner.
+To prove this rigorously is likely to be quite involved as we believe it would involve determining the asymptotics of the operators.
+Even in the case of the full disk [21], where the entries of operators are known explicitly, full condition number analysis is still open, as one needs to handle singularly perturbed operators for the large $m$ modes.
 
@@
 
-Added comment a small subsection to Section 5 to address this.
+Added comment a small subsection to Section 5 to address this. Neumann conditions should be doable by writing the equation as a system of PDEs, as shown in [13] on the triangle.
 
@@
 The package that makes up the framework for expanding functions, obtaining the operator matrices etc.
-is purely experimental at this stage, but is publicly available.
-This fact and where it can be found has been added to the paper.
+is purely experimental at this stage, but is publicly available, and a link has been added to the paper.
 

### 2021-05-20 papers/thesis/full/ch1-introduction/introduction.tex
Further edits to ch1 and 2. Added images for examples and spy plots of jacobi mats.

@@
 \cite{moon2008efficient, sloan2013efficient}), astrophysics (e.g.
-\cite{vasil2019tensor}), quantum theory \cite{varshalovich1988quantum} and geosciences \bstodo{cite applications of SHs}.
+\cite{vasil2019tensor}), quantum theory (e.g.
+\cite{varshalovich1988quantum}), geosciences (e.g.
+\cite{hollerbach2013parity}) and meteorology (e.g.
+\cite{evans1998spherical, rubinstein2015scalar}) \bstodo{cite applications of SHs}.
 Notably, they are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
@@
 
-There are also potential applications in physics too, particularly in astrophysics, where solving equations on the sphere and working in spherical geometries is also desirable \cite{vasil2019tensor}\bstodo{cite more, see p2 of Vasil and the references}.
+There are also potential applications in physics too where solving equations on the sphere and working in spherical geometries is also desirable (e.g.
+\cite{beyer2014numerical, varshalovich1988quantum, slevinsky2018spectral})\bstodo{cite more, see p2 of Vasil and the references}, particularly in astrophysics \cite{vasil2019tensor, reinecke2013libsharp}.
+
 Recently, a method for computing tensor fields in spherical coordinates using Jacobi polynomials has been proposed \cite{vasil2019tensor}.
 In this work the authors present the method for both the surface of the unit sphere and the three-dimensional generalisation of the unit ball.
-\bstodoinline{Describe/summarise the Vasil 2019 Tensor paper approach. Compare to our approach (spectral method with OPs on spherical sub-domains).}
+Their method involves using a spectral basis to represent functions too, choosing for the angular part of the basis to be the spin-weighted spherical harmonics (of which the spherical harmonics are a special case, with spin $0$) in spherical coordinates.
+By using these as their basis functions, they are able to derive sparse relations for the Laplacian operator, as well as for multiplication of the basis by $\cosphi$ and $\sinphi$ (where $\varphi$ is the polar angle from the $z$ axis in cartesian coordinates) which lead to operators for operations with angular dependencies involving these.
+
+On the other hand, we aim to propose a strictly orthogonal \textit{polynomial} basis, and derive sparse operators for multiplication by the cartesian coordinate axes, which can in turn lead to operators for multiplication by trigonometric functions too.
+By utilising orthogonal polynomials, we can develop a sparse spectral method on subdomains of the sphere surface (e.g.
+the spherical cap) using our knowledge of other geometries and methods involving OPs developed for them.
 
@@
 From here, a spectral element method could be devised for the whole sphere using the aforementioned as elements.
+
+Finally, we summarise the work we have presented and detail avenues for future directions that one could take it in Chapter 5.

### 2021-05-20 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Further edits to ch1 and 2. Added images for examples and spy plots of jacobi mats.

@@
 \chapter{Spherical harmonics as orthogonal polynomials}
-
-\bstodoinline{Summarise the ECMWF approach, and how this work is then a stepping stone to understanding more about multi dim OPs on complex domains.}
 
@@
 			1 &\text{if } m < 0
-		\end{cases}. \label{eqn:clmdef}
+		\end{cases} \label{eqn:clmdef}
+}
+which can be simplified to
+\bseqn{
+	\clm = \Bigg({(2l+1) \: (l+m)! \: (l-m)! \over \pi }\Bigg)^\half \: {1 \over 2^{\abs{m} + 1} \; l!} \:
+		\begin{cases} 
+			(-1)^m &\text{if } m \ge 0 \\
+			1 &\text{if } m < 0
+		\end{cases}.
 }
@@
 
+\begin{figure}[tp]
+	\centering % <-- added
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{spy-of-Jx}
+		\centering
+		\caption{$J_x$}
+	\end{subfigure}\hfil % <-- added
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{spy-of-Jy}
+		\centering
+		\caption{$J_y$}
+	\end{subfigure}\hfil % <-- added
+
+	\medskip
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{spy-of-Jz}
+		\centering
+		\caption{$J_z$}
+	\end{subfigure}\hfil % <-- added
+	\caption{\enquote{Spy} plots of the Jacobi operator matrices, showing their sparsity and structure, up to order $N=10$.}
+	\label{fig:sh-spy-jacobi}
+\end{figure}
+
 Using \bsrefeqns{eqn:shrecx}{eqn:shrecz}, we have that the Jacobi matrices take the following block-tridiagonal form (i.e. have block-bandwidths $(1,1)$):
@@
 }
+The sparsity and structure of the Jacobi matrices as we have defined them can be seen visually in \bsreffig{fig:sh-spy-jacobi}.
 
@@
 		\end{cases} \\
-	A_{l,m,5} &:= {1 \over l(l+1)}
+	A_{l,m,5} &:= {i \over 2l(l+1)}
 		\begin{cases}
 			- i \: d_{l,m - 1} \: {c_l^{m-1} \over 2 \clm} &\text{if } m > 0 \\
-			(-1)^m \: i \: d_{l,\abs{m}} \: {\clm \over 2 c_l^{m-1}} &\text{if } m \le 0
+			i \: d_{l,\abs{m}} \: {\clm \over 2 c_l^{m-1}} &\text{if } m \le 0
 		\end{cases} \\
-	A_{l,m,6} &:= {1 \over l(l+1)}
-		\begin{cases}
-			- i \: d_{l,m} \: {\clm \over 2 c_l^{m+1}} &\text{if } m \ge 0 \\
-			(-1)^{m+1} \: i \: d_{l,\abs{m} - 1} \: {c_l^{m+1} \over 2 \clm} &\text{if } m < 0
+	A_{l,m,6} &:= {i \over 2l(l+1)}
+		\begin{cases}
+			- d_{l,m} \: {\clm \over c_l^{m+1}} &\text{if } m \ge 0 \\
+			d_{l,\abs{m} - 1} \: {c_l^{m+1} \over \clm} &\text{if } m < 0
 		\end{cases} \\
-	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6, \\
+	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6 \\
 	\Gamma_{l,m,1} &:= {\clm c_{l-1}^{m} \over l(l-1)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \Big] \\
 	\Gamma_{l,m,2} &:= {\clm c_{l+1}^{m} \over (l+1)(l+2)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \Big] \\
-	\Gamma_{l,m,3} &:= {(-1)^m \: i \: m \over l(l+1)}
+	\Gamma_{l,m,3} &:= {i \: m \over l(l+1)}
 }
@@
 }
-where the coefficients are give by
+where the coefficients are given by
 \bseqn{
@@
 	&= \clm \clmprime \pi (\delta_{m',m-1} + \delta_{m',m+1}) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ 2 i \, m \, m' \, z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} - 1} \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \sgn{m} 2 i \, m \, m' \, z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} - 1} \nonumber \\
 	&\quadeight - i m \jaclmabsdz{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} + 1} \nonumber \\
-	&\quadeight - i m' \jaclmabs{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m} + \abs{m'} +1} \Big\} \: \D z \Big). \label{eqn:lemmaVSHrecurrencesproof4}
-}
+	&\quadeight - i m' \jaclmabs{l}{m} \jaclmabsdz{l'}{m'} \rho^{\abs{m} + \abs{m'} +1} \Big\} \: \D z \Big), \label{eqn:lemmaVSHrecurrencesproof4}
+}
+since when the above is nonzero, $m, m'$ will have the same sign (or one is zero) and so $\abs{m}m' + \abs{m'}m = \sgn{m}mm'$.
 Once again, we can conclude that $\calA^\perp_{l',m'}$ are zero for $m' \notin \{m-1, m+1\}$.
@@
 \bseqnnumber{
-	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m-1} + \abs{m} (m-1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} - 1} \nonumber \\
-	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \nonumber \\
-	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m-1} \rho^{m + \abs{m-1} + 1} \Big\} \: \D z \nonumber \\
 	&= \int_{-1}^{1} \Big\{ 2 i \, m(m-1) \, z \jaclm{l}{m} \jaclmbracket{l'}{m - 1} \rho^{2m - 2} - i m \jaclm{l}{m} \jaclmbracketdz{l'}{m - 1} \rho^{2m} \nonumber \\
@@
 	&= - \int_{-1}^{1} i \jaclm{l}{m} \jaclmbracketdz{l'}{m - 1} \rho^{2m} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclm{l}{m} \jaclm{l'}{m} \rho^{2m} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l',m-1} \; \jaclm{l}{m} \jaclm{l'}{m} \rho^{2m} \: \D z \nonumber \\
 	&= - i \; {d_{l,m-1} \over 2 \pi (c_l^{m})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof5}
@@
 \bseqnnumber{
-	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m+1} + \abs{m} (m+1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} - 1} \nonumber \\
-	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \nonumber \\
-	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \Big\} \: \D z \nonumber \\
 	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
@@
 	&\quadfour + i (m+1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmbracketdz{l'}{\abs{m} - 1} \rho^2 - 2\abs{m} z \jaclmbracket{l'}{\abs{m} - 1} \big] \Big\} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \: \D z \nonumber \\
-	&= - i \; {d_{l,\abs{m}-1} \over 2 \pi (c_l^{\abs{m}})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof6}
+	&= \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= \int_{-1}^{1} i \; d_{l',\abs{m}-1} \; \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \: \D z \nonumber \\
+	&= i \; {d_{l,\abs{m}-1} \over 2 \pi (c_l^{\abs{m}})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof6}
 }
@@
 \bseqnnumber{
-	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m+1} + \abs{m} (m+1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} - 1} \nonumber \\
-	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \nonumber \\
-	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m+1} \rho^{\abs{m} + \abs{m+1} + 1} \Big\} \: \D z \nonumber \\
 	&= \int_{-1}^{1} \Big\{ 2 i \, m(m+1) \, z \jaclm{l}{m} \jaclmbracket{l'}{m + 1} \rho^{2 m} - i m \jaclm{l}{m} \jaclmbracketdz{l'}{m + 1}  \rho^{2m + 2} \nonumber \\
@@
 	&\quadfour + i m \jaclmbracket{l'}{m + 1} \rho^{2m} \big[ \jaclmdz{l}{m} \rho^2 - 2 (m + 1) z \jaclm{l}{m} \big] \Big\} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \jaclm{l}{m} \jaclmbracketdz{l'}{m + 1}  \rho^{2m + 2} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \; d_{l,m} \; \jaclmbracketdz{l}{m + 1} \jaclmbracketdz{l'}{m + 1} \rho^{2m + 2} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \jaclmdz{l}{m} \jaclmbracket{l'}{m + 1}  \rho^{2m + 2} \: \D z \nonumber \\
+	&= - \int_{-1}^{1} i \; d_{l,m} \; \jaclmbracket{l}{m + 1} \jaclmbracket{l'}{m + 1} \rho^{2m + 2} \: \D z \nonumber \\
 	&= - i \; {d_{l,m} \over 2 \pi (c_l^{m+1})^2} \; \delta_{l',l}. \label{eqn:lemmaVSHrecurrencesproof7}
@@
 \bseqnnumber{
-	&\int_{-1}^{1} \Big\{ i \, \big[ m \abs{m-1} + \abs{m} (m-1) \big] \, z \jaclmabs{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} - 1} \nonumber \\
-	&\quadfour - i m \jaclmabs{l}{m} \jaclmabsdz{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \nonumber \\
-	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmabs{l'}{m-1} \rho^{\abs{m} + \abs{m-1} + 1} \Big\} \: \D z \nonumber \\
 	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
-	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \Big\} \: \D z \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i (m-1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
 	&\quadfour + i m \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m}} \big[\jaclmabsdz{l}{m} \rho^2 - 2 (\abs{m} + 1) z \jaclmabs{l}{m} \big] \Big\} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \: \D z \nonumber \\
-	&= - \int_{-1}^{1} i \; d_{l,\abs{m}} \; \jaclmbracketdz{l}{\abs{m} + 1} \jaclmbracketdz{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \: \D z \nonumber \\
-	&= - i \; {d_{l,\abs{m}} \over 2 \pi (c_l^{\abs{m}+1})^2} \; \delta_{l',l}, \label{eqn:lemmaVSHrecurrencesproof8}
+	&= \int_{-1}^{1} i \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \: \D z \nonumber \\
+	&= \int_{-1}^{1} i \; d_{l,\abs{m}} \; \jaclmbracket{l}{\abs{m} + 1} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \: \D z \nonumber \\
+	&= i \; {d_{l,\abs{m}} \over 2 \pi (c_l^{\abs{m}+1})^2} \; \delta_{l',l}, \label{eqn:lemmaVSHrecurrencesproof8}
 }
@@
 }
-The integral over $\theta$ in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz1} is simply $2 \; \pi \; \delta_{m',m}$ showing the above is zero when $m' \ne m$, and thus so will the coefficients $\calC_{l',m'}$ be when this is the case.
+The integral over $\theta$ in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz1} is simply $2 \pi \delta_{m',m}$ showing the above is zero when $m' \ne m$, and thus so will the coefficients $\calC_{l',m'}$ be when this is the case.
 Focusing on the second integral (over $z$) when $m' = m$, this becomes:
@@
 }
-We can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz3} that the nonzero coefficients $\calC_{l',m'}$ are those described in the Lemma as $\Gamma_{l,m,1}$ and $\Gamma_{l,m,3}$.
+We can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz3} that the nonzero coefficients $\calC_{l',m'}$ are those described in the Lemma as $\Gamma_{l,m,1}$ and $\Gamma_{l,m,2}$.
 
@@
 	&\quadeight \quadtwo - i m z \jaclmabsdz{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \Big\} \: \D z \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} i m z^2 \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
-	&= \delta_{l',l} \delta_{m',m} i m {(\clm)^2 \over (\clmprime)^2} \nonumber \\
-	&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz4}
-}
-Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz4} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,2}$.
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} i m \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
+	&= \delta_{l',l} \delta_{m',m} i m. \label{eqn:lemmaVSHrecurrencesproofz4}
+}
+%&= \delta_{l',l} \delta_{m',m} i m (-1)^m. \label{eqn:lemmaVSHrecurrencesproofz4}
+Again, we can see here in \bsrefeqn{eqn:lemmaVSHrecurrencesproofz4} that the coefficients $\calC^\perp_{l',m'}$ are zero when $l' \ne l$ or $m' \ne m$, and that the nonzero coefficients are those described in the Lemma as $\Gamma_{l,m,3}$.
 
@@
 }
+
+\begin{figure}[tp]
+	\centering % <-- added
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{spy-of-Jx-tangent}
+		\centering
+		\caption{$\tangentjacobi_x$}
+	\end{subfigure}\hfil % <-- added
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{spy-of-Jy-tangent}
+		\centering
+		\caption{$\tangentjacobi_y$}
+	\end{subfigure}\hfil % <-- added
+
+	\medskip
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{spy-of-Jz-tangent}
+		\centering
+		\caption{$\tangentjacobi_z$}
+	\end{subfigure}\hfil % <-- added
+	\caption{\enquote{Spy} plots of the Jacobi operator matrices for the Vector Spherical Harmonic OP basis, showing their sparsity and structure, up to order $N=10$.}
+	\label{fig:sh-spy-jacobi}
+\end{figure}
 
@@
 }
+The sparsity and structure of the Jacobi matrices for the VSHs as we have defined them can be seen visually in \bsreffig{fig:vsh-spy-jacobi}.
 
@@
 			{1 \over A_{l,l,4}} & j = 2(2l+1) - 1 \\
-			{- \; A_{l,l,3} \over A_{l,l,4} \: \Gamma_{l,l-1,2}} & j = 6(2l+1) - 5 \\
+			{- \; A_{l,l,3} \over A_{l,l,4} \: \Gamma_{l,l-1,2}} & j = 6(2l+1) - 3 \\
 			0 & o/w
@@
 		\begin{cases}
-			\big(\vec{\eta}_{2}\big)_{j-1} & j = 2(2l+1), 6(2l+1) - 4 \\
+			\big(\vec{\eta}_{2}\big)_{j-1} & j = 2(2l+1), 6(2l+1) - 2 \\
 			0 & o/w
@@
 }
-For $l = 0$ we can simply set
-\bseqnnumber{
-	D^T_0 = 
-		\begin{pmatrix}
-			\delta_1 & 0 & \delta_2 & 0 & 0 & 0 \\
-			0 & \delta_1 & 0 & \delta_2 & 0 & 0 \\
-			0 & 0 & 0 & 0 & {1 \over \Gamma_{0,0,2}} & 0 \\
-			0 & 0 & 0 & 0 & 0 & {1 \over \Gamma_{0,0,2}} \\
-			\delta_3 & 0 & \delta_4 & 0 & 0 & 0 \\
-			0 & \delta_3 & 0 & \delta_4 & 0 & 0
-		\end{pmatrix}. \label{eqn:VSHDzerotdef}
-}
-where
-\bseqn{
-	\delta_1 = {B_{0,0,4} \over A_{0,0,3} B_{0,0,4} - A_{0,0,4} B_{0,0,3}}, \quad \delta_2 = {- A_{0,0,4} \over A_{0,0,3} B_{0,0,4} - A_{0,0,4} B_{0,0,3}}, \\
-	\delta_3 = {B_{0,0,3} \over A_{0,0,4} B_{0,0,3} - A_{0,0,3} B_{0,0,4}}, \quad \delta_4 = {- A_{0,0,3} \over A_{0,0,4} B_{0,0,3} - A_{0,0,3} B_{0,0,4}}.
-}
+%For $l = 0$ we can simply set
+%\bseqnnumber{
+%	D^T_0 = 
+%		\begin{pmatrix}
+%			\delta_1 & 0 & \delta_2 & 0 & 0 & 0 \\
+%			0 & \delta_1 & 0 & \delta_2 & 0 & 0 \\
+%			0 & 0 & 0 & 0 & {1 \over \Gamma_{0,0,2}} & 0 \\
+%			0 & 0 & 0 & 0 & 0 & {1 \over \Gamma_{0,0,2}} \\
+%			\delta_3 & 0 & \delta_4 & 0 & 0 & 0 \\
+%			0 & \delta_3 & 0 & \delta_4 & 0 & 0
+%		\end{pmatrix}. \label{eqn:VSHDzerotdef}
+%}
+%where
+%\bseqn{
+%	\delta_1 = {B_{0,0,4} \over A_{0,0,3} B_{0,0,4} - A_{0,0,4} B_{0,0,3}}, \quad \delta_2 = {- A_{0,0,4} \over A_{0,0,3} B_{0,0,4} - A_{0,0,4} B_{0,0,3}}, \\
+%	\delta_3 = {B_{0,0,3} \over A_{0,0,4} B_{0,0,3} - A_{0,0,3} B_{0,0,4}}, \quad \delta_4 = {- A_{0,0,3} \over A_{0,0,4} B_{0,0,3} - A_{0,0,3} B_{0,0,4}}.
+%}
 
@@
 
-Now that we have our differential operators, let's apply them in an example.
-
-\section{Example: linear shallow water equations}
-
-The linearised shallow water equations allow us to showcase the differential operators that we defined in \bsrefdef{def:SHoperators} (namely the divergence and gradient operators) and demonstrate how the natural sparsity that our framework brings to the problem leads to a simple sparse linear system to solve.
+Now that we have our differential operators, let's apply them in some examples.
+
+\section{Examples}
+
+\subsection{Heat equation}
+
+\begin{figure}[tp]
+	\centering % <-- added
+	\includegraphics[scale=0.5]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100}
+	\caption{Four snapshots of the solution $u$ of the heat equation with $\alpha = 1/42$ and a step size of $\Delta t = 0.01$, solved using the BDF2 timestepping method.
+Top Left: Initial conditions (0 timesteps).
+Top Right: After 25 timesteps.
+Bottom Left: After 50 timesteps.
+Bottom Right: After 75 timesteps.}
+	\label{fig:heateqn}
+\end{figure}
+
+To demonstrate proof of concept, we can apply the framework we have proposed to modelling the heat equation on the unit sphere.
+In doing so, we can demonstrate how our method allows us to expand and evaluate functions, and turn the differential equation into a matrix-vector problem at each timestep.
+
+Let $u : \Omega \to \C$ represent the temperature at a point $(x,y,z)$ on the sphere and suppose that it is modelled by the equation
+\bseqnnumber{
+	\ppx{t}{u} = \alpha \DeltaS u \label{eqn:heateqn}
+}
+for some $\alpha \in \R$.
+A simple timestepping method we can use to iterate with is the BDF2 method \cite[§2.3]{iserles2009first}.
+Let $u_n$ for $n = 0, 1, 2, \dots$ be the solution at time $n \Delta t$ where $\Delta t$ is the timestep.
+The are timestepping method is
+\bseqn{
+	{u_1 - u_0 \over \Delta t} &= \alpha \DeltaS u_1 \\
+	{3u_{n+1} - 4u_n + u_{n-1} \over 2\Delta t} &= \alpha \DeltaS u_{n+1}, \quad n = 1,2,\dots
+}
+$\iff$
+\bseqn{
+	[1 - \Delta t \alpha \DeltaS] u_1 -  &= u_0 \\
+	[3 - 2\Delta t \alpha \DeltaS] u_{n+1} &= 4u_n - u_{n-1}, \quad n = 1,2,\dots.
+}
+Let us write this in coefficient space. Firstly, our coefficients vectors for the function $u$ we will denote $\uvec^c$ -- that is:
+\bseqn{
+	u(x,y,z) &\approx \SHvec(x,y,z)^\top \: \uvec^c.
+}
+Our system then becomes
+\bseqn{
+	[I - \Delta t \alpha \calL] \uvec^c_1 -  &= u_0 \\
+	[3I - 2\Delta t \alpha \calL] \uvec^c_{n+1} &= 4\uvec^c_n - \uvec^c_{n-1}, \quad n = 1,2,\dots.
+}
+From here, it is clear that eat each timestep we simply have a fairly trivial linear system to solve, with the matrices $I - \Delta t \alpha \calL$, $3I - 2\Delta t \alpha \calL$ being banded-block-banded (in fact, simply diagonal here).
+We demonstrate this theory in \bsreffig{fig:heateqn}, where we see four snapshots of the solution to the heat equation in \bsrefeqn{eqn:heateqn} on the unit sphere with our order $N$ set large enough (here, $N=20$).
+We set $\alpha = {1 \over 42}$ and a step size of $\Delta t = 0.01$, and take the initial condition for the heat $u$ to be a \enquote{football function}\footnote{These initial conditions are taken from an example found at https://www.chebfun.org/examples/sphere/SphereHeatConduction.html} -- that is we take
+\bseqn{
+	u_0 := Y_6^0 + \sqrt{14 \over 11} Y_6^6.
+}
+The snapshots of the solution $u$ in bsreffig{fig:heateqn} are taken of the initial conditions, plus after $25$, $50$ and $75$ timesteps respectively.
+
+\subsection{Linearised shallow water equations}
+
+\begin{figure}[tp]
+	\centering % <-- added
+	\includegraphics[scale=0.5]{linear-swe-H=1-dt=0p01-nsteps=100}
+	\caption{Four snapshots of the solution for the height $h$ of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method.
+Top Left: Initial conditions (0 timesteps).
+Top Right: After 50 timesteps.
+Bottom Left: After 75 timesteps.
+Bottom Right: After 100 timesteps.}
+	\label{fig:swe}
+\end{figure}
+
+The linearised shallow water equations allow us to showcase more of the differential operators that we defined in \bsrefdef{def:SHoperators} (namely the divergence and gradient operators) and demonstrate how the natural sparsity that our framework brings to the problem leads to a simple sparse linear system to solve.
 
@@
 The linear SWEs are
-\bseqn{
+\bseqnnumber{
 	\begin{cases}
@@
 		\pfpx{t}{h} + \calH \gradS \cdot \uvec = 0
-	\end{cases}
+	\end{cases} \label{eqn:swe}
 }
@@
 		\begin{pmatrix}
-			0 & 1 \\
-			-1 & 0
+			0 & -1 \\
+			1 & 0
 		\end{pmatrix} \in \R^{2\times2}.
@@
 }
-Let's write this in coefficient space. Firstly, our coefficients vectors for the functions $\uvec, h$ we will denote $\uvec^c, \vec{h}^c$ -- that is:
+Let us write this in coefficient space. Firstly, our coefficients vectors for the functions $\uvec, h$ we will denote $\uvec^c, \vec{h}^c$ -- that is:
 \bseqn{
@@
 \bseqn{
-	\uvec^c_{n+1} &= \uvec^c_n + \Delta t \: (\tilde \calG \tilde{\vec{h}}^c_{n+1} - 2 \rotrate J_z \calR \uvec^c_{n+1}) \\
+	\uvec^c_{n+1} &= \uvec^c_n + \Delta t \: (\tilde \calG \tilde{\vec{h}}^c_{n+1} - 2 \rotrate \tangentjacobi_z \calR \uvec^c_{n+1}) \\
 	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}
@@
 \bseqn{
-	\uvec^c_{n+1} &= \uvec^c_n + \Delta t \: (\tilde \calG \tilde{\vec{h}}^c_{n} - \Delta t \: \calH \tilde \calG \tilde \calD \uvecc_{n+1} - 2 \rotrate J_z \calR \uvec^c_{n+1}) \\
+	\uvec^c_{n+1} &= \uvec^c_n + \Delta t \: (\tilde \calG \tilde{\vec{h}}^c_{n} - \Delta t \: \calH \tilde \calG \tilde \calD \uvecc_{n+1} - 2 \rotrate \tangentjacobi_z \calR \uvec^c_{n+1}) \\
 	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}
@@
 \bseqn{
-	\Big[ I + \Delta t^2 \calH \tilde \calG \tilde \calD + 2 \rotrate \Delta t \: J_z \calR \Big] \: \uvec^c_{n+1} &= \uvec^c_n + \Delta t \tilde \calG \tilde{\vec{h}}^c_{n} \\
+	\Big[ I + \Delta t^2 \calH \tilde \calG \tilde \calD + 2 \rotrate \Delta t \: \tangentjacobi_z \calR \Big] \: \uvec^c_{n+1} &= \uvec^c_n + \Delta t \tilde \calG \tilde{\vec{h}}^c_{n} \\
 	\tilde{\vec{h}}^c_{n+1} &= \tilde{\vec{h}}^c_n - \Delta t \: \calH \tilde \calD \uvec^c_{n+1}.
 }
-Here, it is clear that the matrix $I + \Delta t^2 \calH \tilde \calG \tilde \calD + 2 \rotrate \Delta t \: J_z \calR$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
+Here, it is clear that the matrix $I + \Delta t^2 \calH \tilde \calG \tilde \calD + 2 \rotrate \Delta t \: \tangentjacobi_z \calR$ is sparse with banded-block-banded structure, and hence this system will be efficient to solve.
 Doing so at each iteration provides us with the coefficients vectors $\uvec^c_{n+1}, \tilde{\vec{h}}^c_{n+1}$ if we know $\uvec^c_{n}, \tilde{\vec{h}}^c_{n}$.
@@
 
-\bstodo{Images}
+As a demonstration, in \bsreffig{fig:swe} we see four snapshots of the solution to the system in \bsrefeqn{eqn:swe} on the unit sphere, with $\rotrate = \calH = 1$ and initial conditions given by
+\bseqn{
+	\uvec_0 := \phivec \mu \sinphi, \quad h_0 := \mu \sin^2\varphi,
+}
+where $\mu = {\pi \over 6}$, with the solution order $N$ set large enough (here, $N=20$).
+This example is very similar to \enquote{test case 2} in the well-known standard test set for shallow water equations in spherical geometries \cite{williamson1992standard}.
+The snapshots of the wave height $h$ are taken of the initial conditions, plus after $50$, $75$ and $100$ timesteps respectively, using a step size of $\Delta t = 0.01$.
 
@@
 In fact, for the Spherical Harmonics, these are simply diagonal.
-Finally, we presented a simple example of how one can put this into practice for solving the linearised shallow water equations.
+Finally, we presented a couple of simple examples of how one can put this all into practice, namely solving the heat equation and the linearised shallow water equations on the unit sphere.

### 2021-05-20 papers/spherical-caps/revised/spherical-caps.tex

@@
 While such an approach is useful for other geometries, it is of low order and does not achieve spectral convergence with sparse discretisations.
-Further, for our interest of the sphere, it would not take advantage of any rotational symmetry to achieve optimal complexity with a direct solver.
+Further, for our domain of interest, the closest point method does not take advantage of any rotational symmetry, and so is unable to achieve optimal complexity with a direct solver.
 

### 2021-05-20 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Edit to intro of ch2. Added fig files.

@@
 To introduce ourselves to the world of multidimensional orthogonal polynomials for solving PDEs on the sphere, we can naturally choose to look at the famous spherical harmonics.
-Our aim here is to express the spherical harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere.
+Our aim here is to express the Spherical Harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere.
 More precisely, we desire the solution to partial differential equations on the domain
@@
 While it may seem somewhat odd, it should hopefully become apparent why this is a useful way to define our domain.
+
+In this chapter, we demonstrate how one can write the Spherical Harmonics as polynomials in $(x, y, z)$, and how the relationships between them can be used to obtain sparse \enquote{Jacobi operators} for multiplication by $x,y,z$.
+We demonstrate how these in turn lead to an algorithm for evaluating the Spherical Harmonics, and further an efficient algorithm for evaluating functions when expanded in the SH OP basis.
+The techniques used here are done so with the aim of applying them to other OP families of other domains (and we do so in later chapters).
+We will finally demonstrate a proof-of-concept via the examples of the heat equation and the linearised shallow water equations on the unit sphere.
+
+The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners the ApproxFun package \cite{ApproxFun} ?
+however, this package is purely experimental at this stage.
 

### 2021-05-20 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
minor edit

@@
 In this chapter, we demonstrate how one can write the Spherical Harmonics as polynomials in $(x, y, z)$, and how the relationships between them can be used to obtain sparse \enquote{Jacobi operators} for multiplication by $x,y,z$.
-We demonstrate how these in turn lead to an algorithm for evaluating the Spherical Harmonics, and further an efficient algorithm for evaluating functions when expanded in the SH OP basis.
+We demonstrate how these in turn lead to an algorithm for evaluating the Spherical Harmonics, an efficient algorithm for evaluating functions when expanded in the SH OP basis, and sparse \enquote{banded-block-banded} differential operator matrices.
 The techniques used here are done so with the aim of applying them to other OP families of other domains (and we do so in later chapters).

### 2021-05-24 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Added background to intro for ch3. needs finishing.

@@
 The contents of the previous chapter on spherical harmonics can allow us to use similar techniques for developing sparse spectral methods for solving linear partial differential equations on other domains.
-Before we can return to the 3D surface realm, let's first achieve a foundation in two-dimensions, by looking at a special class of geometries that includes disk slices and trapeziums.
+Before we can return to the 3D surface realm, let us first achieve a foundation in two-dimensions by looking at a special class of geometries that includes disk slices and trapeziums.
+
+\bstodoinline{Background on multivariate OPs on the disk and triangle.
+Give whole disk OP formula and triangle OP formula as reference points (and refer back to it when defining the disk slice OPs and trapezium OPs respectively).
+Mention there are other choices for OPs on these domains too.}
+
+There has been some really interesting work in the area of mutlivariate orthogonal polynomials in 2D domains \bstodo{list some domains and cite references for them}, in particular spectral methods have been established on the disk \cite{} and the triangle \cite{}.
+For the triangle, the OPs have been chosen as a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values:
+\bseqnnumber{
+	P^{(a,b,c)}(x,y) := define, \quad (x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x} \label{eqn:triangleops}
+}
+while for the disk, so-and-so choose the OPs to be:
+\bseqnnumber{
+	define \label{eqn:diskops}.
+}
+We use the word \enquote{choose} here deliberately, as naturally there is more than one way to define an OP basis on these domains, including the generalisations that we will discuss in this chapter.
+\bstodo{Give one?}
+
+We will see that the full unit disk and triangle cases are special cases of the spectral methods we investigate here.
 
@@
 \bseqn{
-	\Omega := \bsset{(x,y) \in \R^2}{\alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)} 
+	\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)} 
 }
@@
 \begin{condition}\label{cond:disk}
-	$\rho$ is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\).
+	$\rho$ is the square root of a non-negative degree $\le 2$ polynomial, $-\gamma = \delta > 0$.
 \end{condition}
@@
 For simplicity of presentation we focus on the disk-slice in this chapter, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$.
-However, we will discuss an extension to other geometries, including the half-disk and trapeziums.
-
-We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
+However, we will discuss an extension to other geometries, including the half-disk and trapeziums, by choosing the function $\rho$ and constants $\alpha, \beta, \gamma, \delta$ appropriately.
+Moreover, one should be able to see that the full disk and triangle can also be obtained in such a way.
+
+We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
 On the disk-slice the family of weights we consider are of the form
 \bseqn{
-	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c, \qfor \alpha \leq x \leq \beta,\quad -\rho(x) \leq y \leq \rho(x).
-}
-The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$. We define these to be orthogonalised lexicographically, that is,
+	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c
+}
+for $\alpha \leq x \leq \beta,$ $-\rho(x) \leq y \leq \rho(x)$.
+The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$.
+We define these to be orthogonalised lexicographically, that is,
 \bseqn{
@@
 }
-where $C_{n,k} \neq 0$ and ``lower order terms''   includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
-The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
+where $C_{n,k} \neq 0$ and \enquote{lower order terms'} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
+The precise normalization arises from their definition in terms of one-dimensional OPs in \bsrefdef{def:OPconstruction}.
 

### 2021-05-24 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
moved SC paper into thesis as CH4 (replacing old version). Other edits.

@@
 
-The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners the ApproxFun package \cite{ApproxFun} ?
-however, this package is purely experimental at this stage.
+The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners the ApproxFun package \cite{ApproxFun} -- however, this package is purely experimental at this stage.
 
@@
 
-Before we proceed further, let's build up to our definition of the spherical harmonics. We first need to introduce a few classical orthogonal polynomials.
-
-On the unit interval, $[-1,1]$, we note that there is a hierarchy of orthogonal polynomials (OPs) in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
+Before we proceed further, let's build up to our definition of the spherical harmonics.
+We first need to introduce a few classical orthogonal polynomials.
+\enquote{Orthogonal polynomials} (OPs) are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product.
+In particular, they are orthogonal with respect to all lower degree polynomials.
+
+On the unit interval, $[-1,1]$, we note that there is a hierarchy of OPs in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
 \bseqn{

### 2021-05-24 papers/thesis/full/ch3-disk-slices/disk-slices.tex
moved SC paper into thesis as CH4 (replacing old version). Other edits.

@@
 \begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
-	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \condref{trap} or \condref{disk} with $w_2$ being an even function hold.
+	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
 $\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$.
@@
 }
-assuming that either \condref{trap} or \condref{disk} with  \(\jacw^{(a,b)}\) being an even function (i.e.
+assuming that either \bsrefcondition{cond:disk} or \bsrefcondition{cond:disk} with  \(\jacw^{(a,b)}\) being an even function (i.e.
 $a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$) hold.
@@
 
-Finally, if \condref{trap} holds we have that 
+Finally, if \bsrefcondition{cond:trap} holds we have that 
 $$
@@
 $$
-If \condref{disk} holds we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.
+If \bsrefcondition{cond:disk} holds we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.
 

### 2021-05-24 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
moved SC paper into thesis as CH4 (replacing old version). Other edits.

--- 2020-12-21 papers/spherical-caps/spherical-caps.tex
@@
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
-%\geometry{landscape}                		% Activate for rotated page geometry
-\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
-								% TeX will automatically convert eps --> pdf in pdflatex
-
-\graphicspath{ {images/} }
-
-\include{bsthesisnewcommands}
-\include{bsmacros}
-\include{somacros}
-
-\title{Sparse spectral methods for partial differential equations on spherical caps}
-\author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
-%\date{}							% Activate to display a given date or no date
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, disk-slices and triangles.
-In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps.
+\chapter{Spherical caps}
+
+While the work in the previous chapter looked at developing a sparse spectral method inside a two dimensional domain, we move on to investigating the realm of a surface in three dimensional space.
+Specifically, we look to extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps.
 The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
-We demonstrate the results on partial differential equations involving the spherical Laplacian and biharmonic operators, showing spectral convergence.
-\end{abstract}
-
-%
-\section{Introduction}
-
-This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere---specifically spherical caps.
-More precisely, we consider the solution of partial differential equations on the spherical cap $\Omega$ defined by
-\begin{align*}
-	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
-\end{align*}
+We demonstrate the results on partial differential equations involving the spherical Laplacian and biharmonic operators, showing spectral convergence with discretisations that can be made well-conditioned using a simple preconditioner.
+
+Our aim in this chapter is to develop a sparse spectral method for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps.
+More precisely, we consider the solution of partial differential equations on the \textit{spherical cap} $\Omega$ defined by
+\bseqn{
+	\Omega := \bsset{(x,y,z) \in \R^3}{\alpha < z < \beta, \: x^2 + y^2 + z^2 = 1}
+}
 where $\alpha \in (-1,1)$ and $\beta := 1$.
+Simply put, the region of the surface of the $2$-sphere where the $z$-coordinate range is limited to a sub-interval of $[-1,1]$ is what we refer to as the spherical cap.
 
@@
 The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper.
-These adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same.
+These adjustments make the construction more involved, which is why they are omitted here, but the approach is essentially the same.
 \end{remark}
 
-We advocate using a basis that is polynomial in cartesian coordinates, that is, polynomial in $x$, $y$, and $z$, and orthogonal with respect to a prescribed weight: that is, multivariate orthogonal polynomials, whose construction was considered in \cite{olver2020orthogonal}.
-Equivalently, we can think of these as polynomials modulo the vanishing ideal $\{ x^2 + y^2 + z^2 = 1 \}$, or simply as a linear recombination of spherical harmonics that are orthogonalised on a subset of the sphere.
+For the spherical cap, we advocate using a basis that is polynomial in cartesian coordinates, that is, polynomial in $x$, $y$, and $z$, and orthogonal with respect to a prescribed weight: that is, multivariate orthogonal polynomials, whose construction was considered in \cite{olver2020orthogonal}.
+Equivalently, we can think of these as polynomials modulo the vanishing ideal $\bsset{(x,y,z) \in \R^3}{x^2 + y^2 + z^2 = 1}$, or simply as a linear recombination of spherical harmonics that are orthogonalised on a subset of the sphere.
 This is in contrast to more standard approaches based on mapping the geometry to a simpler one (e.g., a rectangle or disk) and using orthogonal polynomials in the mapped coordinates (e.g., a basis that is polynomial in the spherical coordinates $\varphi$ and $\theta$).
@@
 
-On the spherical cap, the family of weights we consider are of the form
-\begin{align*}
+There are of course other approaches for solving PDEs on surfaces, such as the closest point method \cite{macdonald2010implicit, macdonald2008level}.
+This involves recasting the PDE as one involving a \enquote{closest point} operator in a 3D volume that acts as a \enquote{shell} for the surface.
+While such an approach is useful for other geometries, it is of low order and does not achieve spectral convergence with sparse discretisations.
+Further, for our domain of interest, the closest point method does not take advantage of any rotational symmetry, and so is unable to achieve optimal complexity with a direct solver.
+
+On the spherical cap, the family of weights we will consider are of the form
+\bseqn{
 	\Wa(x,y,z) := (z - \alpha)^a, \qfor (x,y,z) \in \Omega,
-\end{align*}
+}
 noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \Omega$ when $a > 0$.
@@
 We define these to be orthogonalised lexicographically, that is,
-\begin{align*}
+\bseqn{
 	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms})
-\end{align*}
+}
 where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - i} \: y^{i} \: z^{n-j}$ where $j < k$.
-The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.
+The precise normalization arises from their definition in terms of one-dimensional OPs that we will see in \bsrefdef{def:sc:OPconstruction}.
 
 We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
-\begin{align*}
-	z &= \cos\varphi, \\
-	x &= \sin\varphi \cos\theta = \rho(z) \cos\theta, \\
-	y &= \sin\varphi \sin\theta = \rho(z) \sin\theta.
-\end{align*}
+\bseqn{
+	z &= \cosphi, \\
+	x &= \sinphi \costheta = \rho(z) \costheta, \\
+	y &= \sinphi \sintheta = \rho(z) \sintheta.
+}
 where $ \rho(z) := \sqrt{1-z^2}$, we have
-\begin{align*}
+\bseqn{
 	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
-\end{align*}
+}
 i.e.
 $\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$.
-We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
-As in the  disk-slice case in 2D \cite{snowball2019sparse}, we use an integration-by-parts argument to deduce the sparsity structure.
-
- The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined \cite{snowball2019sparse}.
+We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
+Just as we proceeded in \bsrefchapter{chapter:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.
+
+The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined in \bsrefchapter{chapter:diskslice} (see also \cite{snowball2019sparse}).
 In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for Schr\"odinger operators $\DeltaS + v(x,y,z)$ where $v$ is first approximated by a polynomial.
-A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
-This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity} with $v(x,y,z) = \cos z$.
-
-An overview of the paper is as follows:
-
-\noindent \secref{OPs}: We present our definition of a (one-parameter) family of 3D orthogonal polynomials (OPs) on the spherical cap domain $\Omega$, by combining 1D OPs on the interval $(\alpha, 1)$ with Chebyshev polynomials, to form 3D OPs on the spherical cap surface.
-We show that these families will lead to sparse Jacobi operators for multiplication by $x, y, z$ and demonstrate how to obtain the 3D OPs.
-
-\noindent\secref{PDOs}: We define several partial differential operators such as spherical Laplacians and  show that these will  be sparse when applied to a suitable choice of expansions in bases built from OPs on the spherical cap.
-We can exactly calculate the non-zero entries of these sparse operators using the quadrature rule associated with the non-classical 1D OPs.
-
-\noindent\secref{Computation}: We derive a quadrature rule on the spherical cap that can be used to expand a function in the OP basis up to a given order $N$, and demonstrate how to evaluate a function using the Clenshaw algorithm using the coefficients of its expansion.
-
-\noindent\secref{Examples}: We demonstrate the proposed technique for solving differential equations on the spherical cap such as the Poisson equation, variable coefficient Helmholtz equation, and Biharmonic equation.
-
-%\noindent\appref{PFEM}: We outline how our framework can allow us to construct sparse $hp$-finite element method on the sphere, using spherical band and cap elements that would capture spherical geometry precisely.
-We lay the groundwork by simply presenting a sparse $p$-finite element method using a single spherical cap element.
-
-\noindent{\bf Acknowledgments}: The first author was supported by the Engineering and Physical Sciences Research Council Mathematics of Planet Earth Centre for Doctoral Training at Imperial College London and the University of Reading, with grant number EP/L016613/1.The second author was supported by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces''.
-
-\section{Orthogonal polynomials on spherical caps}\label{Section:OPs}
-
-In this section we outline the construction and some basic properties of $\scopnkia(x,y,z)$.
-
-\subsection{Explicit construction}
-
-We can construct the 3D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Chebyshev polynomials.
-We do so in terms of Fourier series, which, following \cite{olver2020orthogonal}, we write here as orthogonal polynomials in $x$ and $y$:
-
-\begin{definition}\label{def:Ydefinition}
+A nice feature is that if the partial differential operator is invariant with respect to rotation around the $z$ axis (e.g., a Schr\"odinger operator with potential $v(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix.
+This improves the complexity further to an optimal $O(p^2$), which is demonstrated in \bsreffig{fig:sc:complexity} with $v(x,y,z) = \cos z$.
+
+The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/OrthogonalPolynomialFamilies.jl} to partner the ApproxFun package \cite{ApproxFun} -- however, this package is purely experimental at this stage.
+
+\section{The circle arc}\label{section:sc:arc}
+
+The spherical cap can be thought of as higher dimensional version of the circle arc (a one-dimensional \enquote{surface} in two-dimensional space).
+One can define OPs on the circle in terms of Fourier series, which we write here as orthogonal polynomials in $x$ and $y$:
+\begin{definition}\label{def:sc:Ydefinition}
 	Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
 Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
-\begin{align*}
+\bseqn{
 	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
@@
 	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots
-\end{align*}
+}
 where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
-\begin{align*}
-	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta
-\end{align*}
+\bseqn{
+	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
+}
+Note that $\ch_0$ is defined so as to ensure orthonormality. 
 \end{definition}
-Note that we have defined $\ch_0$ so as to ensure orthonormality.
-
-\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
-	Let $w : (\alpha,\beta) \: \to \R$ be a weight function.
+
+However, for an arc of the unit circle $\{ define arc \}$ (note we can orientate the arc as we please without loss of generality), we would need to make a modification so that the polynomials are orthogonal with respect to an inner product integral with the truncated domain of the arc:
+\bseqn{
+	give arc ops (T^h, yU^h) and the inner product.
+}
+\bstodoinline{Expand on this. Read the Olver Xu "OPs on quadratic curves paper." where the arc might apply??? Either way, cite this paper in intro (CH1).}
+
+\section{Orthogonal polynomials on algebraic curves}
+
+This idea of using 1D OPs on a shortened interval to create the modified OPs for an arc has be generalised to higher dimensional domains, where a way of constructing orthogonal polynomials on algebraic curves has been presented \cite{olver2020orthogonal}.
+\bstodoinline{Expand on this.
+Summarise the "on" surfaces approach.
+Use references in the paper.}
+
+\section{Orthogonal polynomials on spherical caps}\label{section:sc:OPs}
+
+\subsection{Explicit construction}
+
+Using the approach presented for constructing OPs on algebraic curves, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Fourier series written as orthogonal polynomials in $x$ and $y$:
+\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:sc:construction}
+	Let $w : (\alpha,\beta) \to \R$ be a weight function.
 For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$.
-Then the 3D polynomials defined on $\Omega$
-\begin{align*}
+Then the 3D polynomials defined on $\Omega$ given by
+\bseqn{
 	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)})
-\end{align*}
+}
 for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
-\begin{align*}
+\bseqn{
 	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
@@
 	&\qquad w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
-	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
-\end{align*}
+	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: q\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: w(z) \: \D \theta \: \D z
+}
 on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the uniform spherical measure on $\Omega$. 
@@
 
-For the spherical cap, we can use Proposition \ref{prop:construction} to create our one-parameter family of OPs.
-We first introduce notation for our family of non-classical univariate OPs that will be used as the $r_n$ polynomials above.
-\begin{definition}[\cite{snowball2019sparse}]\label{def:OPconstruction}
-	Let $\genjacw^{(a,b)}(x)$ be a weight function on the interval $(\alpha, 1)$ given by:
-\begin{align*}
+For the spherical cap, we can use \bsrefprop{prop:sc:construction} to create our one-parameter family of OPs.
+The univariate OPs that we will choose for the $r_{n,k}$ polynomials above will be the non-classical $\genjac^{(a,b)}$ OPs that we defined in \bsrefdef{def:ds:OPconstruction}.
+Since there is only one boundary for the spherical cap, we will only need to use a two-parameter version\footnote{For a spherical band with two boundaries, we would need the three-parameter version}).
+For reference, the family of orthonormal polynomials on $[\alpha,\beta]$ denoted $\{\genjac_n^{(a,b)}\}$ are defined such that they are orthonormal with respect to the inner product
+\bseqnnumber{
+	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:sc:ipgenjac}
+}
+where
+\bseqn{
 	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b}
-\end{align*}
-and define the associated inner product by:
-\begin{align}
-	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipgenjac}
-\end{align}
-where
-\begin{align}
-	\normgenjac^{(a,b)} := \int_\alpha^1 \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:ipnormalisation}
-\end{align}
+} is the weight function and 
+\bseqnnumber{
+	\normgenjac^{(a,b)} := \int_\alpha^1 \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:sc:ipnormalisation}
+}
 is a normalising constant.
-Denote the two-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}).
+
+We can now define the 3D OPs for the spherical cap.
+\begin{definition}\label{def:sc:constuction}
+	Define the one-parameter 3D orthogonal polynomials via:
+\bseqnnumber{
+	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega.
+}
 \end{definition}
 
-We can now define the 3D OPs for the spherical cap.
-\begin{definition}\label{def:constuction}
-	Define the one-parameter 3D orthogonal polynomials via:
-\begin{align}
-	\scopnkia(x,y,z) := \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \: \chki\fpr(\frac{x}{\rho(z)}, \frac{y}{\rho(z)}), \quad (x,y,z) \in \Omega.
-\end{align}
-\end{definition}
-
 By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
-\begin{align*}
+\bseqn{
 	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
-\end{align*}
+}
 with
-\begin{align}
-	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:normscop}
-\end{align}
-We note that the weight $\genjacw^{(a,b)}(z)$ has been used in the construction of 2D orthogonal polynomials on disk-slices and trapeziums \cite{snowball2019sparse}, where a method for obtaining recurrence coefficients and evaluating integrals was established (the weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
-
-\subsection{Jacobi matrices}\label{subsection:jacobimats}
-
-We can express the three-term recurrences associated with $\genjac_n^{(a,b)}$ as
-\begin{align}
-	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:Rrecurrence}
-\end{align}
-where the coefficients are calculatable (see \cite{snowball2019sparse}).
-We can use (\ref{eqn:Rrecurrence}) to determine the 3D recurrences for $\scopnkia(x,y,z)$.
+\bseqnnumber{
+	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:sc:normscop}
+}
+A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{chapter:diskslice} (see also \cite{snowball2019sparse}).
+The weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}.
+
+\subsection{Jacobi matrices}\label{section:sc:jacobimats}
+
+Recall that we can express the three-term recurrences associated with $\genjac_n^{(a,b)}$ as
+\bseqnnumber{
+	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:sc:Rrecurrence}
+}
+where the coefficients are calculable (see \bsrefchapter{chapter:diskslice}).
+We can use \bsrefeqn{eqn:sc:Rrecurrence} to determine the 3D recurrences for $\scopnkia(x,y,z)$.
 Importantly, we can deduce sparsity in the recurrence relationships.
@@
 
-\begin{lemma}\label{lemma:Yrecurrence} 
+\begin{lemma}\label{lemma:sc:Yrecurrence} 
 The following identities hold for $k = 2,3,\dots$, $j = 0,1,\dots$ and $i, h \in \{0,1\}$:
-\begin{align*}
+\bseqn{
 	&1) \quad \int_0^{2\pi} \ch_0 \: \chjh(\theta) \: \cos\theta \: \D \theta = \ch_0 \: \pi \: \delta_{0,h} \: \delta_{1,j} \\
@@
 	&6) \quad \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta = \half \: \pi \: \delta_{|i-1|,h} \: ((-1)^{i+1} \: \delta_{k-1,j} + (-1)^i \: \delta_{k+1,j}).
-\end{align*}
+}
 \end{lemma}
@@
 \begin{proof}
-Each follows from the definitions of $\ch_{k,i}$ and $\ch_0$, as well as the relationships:
-\begin{align*}
+Each follows from the definitions of $\ch_{k,i}$ and $\ch_0$, as well as the trigonometric relationships:
+\bseqn{
 	2 \cos k \theta \cos\theta &= \cos (k-1)\theta + \cos(k+1)\theta \\
@@
 	2 \sin k \theta \sin\theta &= \cos (k-1)\theta - \cos(k+1)\theta.
-\end{align*}
+}
 \end{proof}
 
-\begin{lemma}\label{lemma:Qrecurrence} 
+\begin{lemma}\label{lemma:sc:Qrecurrence} 
 Define
-\begin{align}
+\bseqnnumber{
 	\eta_{k} :=
@@
 		\end{cases}
-\end{align}
+}
 $\scopnkia(x,y,z)$ satisfy the following recurrences:
-\begin{align*}
+\bseqn{
 	x \: \scopnkia(x,y,z) &= \alphaa_{n,k,1} \:  \scopa_{n-1, k-1, i}(x, y, z) + \alphaa_{n,k,2} \:  \scopa_{n-1, k+1, i}(x, y, z) \nonumber \\
@@
 	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
-\end{align*}
+}
 for $(x,y,z) \in \Omega$, where
-\begin{align*}
+\bseqn{
 	\alphaa_{n,k,1} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjacnmk^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
@@
 	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}.
-\end{align*}
+}
 \end{lemma}
 \begin{remark}
-For $z$ multiplication, note that different Fourier modes do not interact. This is because $z$ is rotationally invariant. 
+	For $z$ multiplication, note that different Fourier modes do not interact. This is because multiplication by $z$ is invariant with respect to rotation around the $z$-axis. 
 \end{remark}
 \begin{proof}
-The 3-term recurrence for multiplication by $z$ follows from (\ref{eqn:Rrecurrence}).
+	The 3-term recurrence for multiplication by $z$ follows from \bsrefeqn{eqn:sc:Rrecurrence}.
 For the recurrence for multiplication by $x$, since $\{\scopmjha\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$, $h = 0,1$ is an orthogonal basis for any degree $n+1$ polynomial on $\Omega$, we can expand
-\begin{align*}
+\bseqn{
 	x \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j} \: \scopmjha(x,y,z).
-\end{align*}
+}
 These coefficients are given by
-\begin{align*}
+\bseqn{
 	c_{m,j} = {\ip< x \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
-\end{align*}
+}
 where we show the non-zero coefficients that result are the $\alphaa_{n,k,1},\dots,\alphaa_{n,k,6}$ in the lemma.
-Recall from equation (\ref{eqn:normscop}) that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$.
+Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$.
 Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)$:
-\begin{align*}
+\bseqn{
 	&\ip<x \: \scopnkia, \scopmjha>_{\scopa} \\
@@
 	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
-\end{align*}
-where $\delta_{k, j}$ is the standard Kronecker delta function, using Lemma \ref{lemma:Yrecurrence}. Similarly, for the recurrence for multiplication by $y$, we can expand 
-\begin{align*}
+}
+where $\delta_{k, j}$ is the standard Kronecker delta function, using \bsreflemma{lemma:sc:Yrecurrence}. Similarly, for the recurrence for multiplication by $y$, we can expand 
+\bseqn{
 	y \: \scopnkia(x,y,z) = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 d_{m,j} \: \scopmjha(x,y,z).
-\end{align*}
+}
 These coefficients are given by
-\begin{align*}
+\bseqn{
 	d_{m,j} = {\ip< y \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
-\end{align*}
-where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the lemma:
-\begin{align*}
+}
+where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the lemma:sc:
+\bseqn{
 	&\ip<y \: \scopnkia, \scopmjha>_{\scopa} \\
@@
 	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
-\end{align*}
-where again $\delta_{k, j}$ is the standard Kronecker delta function, and we have used Lemma \ref{lemma:Yrecurrence}.
+}
+where again $\delta_{k, j}$ is the standard Kronecker delta function, and we have used \bsreflemma{lemma:sc:Yrecurrence}.
 
@@
 
-The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$.
+The recurrences in \bsreflemma{lemma:sc:Qrecurrence} lead to (block) Jacobi matrices that correspond to multiplication by $x$, $y$ and $z$.
 In later sections we will use an ordering of the OPs so that they are grouped by Fourier mode $k$, which is convenient for the application of differential and other operators to the vector of coefficients of a given function's expansion (some operators will exploit this ordering for operators where Fourier modes do not interact, and thus will be block-diagonal).
-Before that though, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
+Before that, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm.
 In practice, it is simply a matter of converting coefficients between the two orderings.
@@
 For $n=0,1,2,\dots$:
-\begin{align*}
+\bseqn{
 	\bigscoptna := 
@@
 		\end{pmatrix}
-\end{align*}
+}
 and set $ J_x^{(a)},  J_y^{(a)},  J_z^{(a)}$ as the Jacobi matrices corresponding to
-\begin{align}
+\bseqnnumber{
 	J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
-	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
+	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:sc:jacobimatricesdefinitionalt} \\
 	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
-\end{align}
+}
 where
-\begin{align*}
+\bseqn{
 	J_{x/y/z}^{(a)} &= 
@@
 		\end{pmatrix}.
-\end{align*}
-Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are \textit{banded-block-banded matrices}:
+}
+While $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are not Jacobi matrices in the classical sense, they are block Jacobi matrices.
+However, it is useful to simply label them as such.
+In fact, note that they are in fact \textit{banded-block-banded matrices}:
 
@@
 Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
-\begin{align*}
+\bseqn{
 	A^{(a)}_{x,n} &:= 
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
+}
 where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align*}
+\bseqn{
 	A^{(a)}_{n,k,j} &:= 
@@
 		\end{pmatrix} \in \R^{2\times1}, j \text{ odd}.
-\end{align*}
+}
 For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
-\begin{align*}
+\bseqn{
 	A^{(a)}_{y,n} &:= 
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
+}
 where for $k = 1,\dots,N, \: n = k,\dots,N$  
-\begin{align*}
+\bseqn{
 	B^{(a)}_{n,k,j} &:= 
@@
 		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}.
-\end{align*}
+}
 For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
-\begin{align*}
+\bseqn{
 	A^{(a)}_{z,n} &:= 
@@
 		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
-\end{align*}
+}
 where for $k = 1,\dots,N, \: n = k,\dots,N$
-\begin{align}
+\bseqnnumber{
 	\Gamma^{(a)}_{n,k,j} &:= 
@@
 			0 & \gamma_{n,k,j}
-		\end{pmatrix} \in \R^{2\times2}, \label{eqn:jacobisubblocksGamma1} \\
-	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:jacobisubblocksGamma2}
-\end{align}
+		\end{pmatrix} \in \R^{2\times2}, \label{eqn:sc:jacobisubblocksGamma1} \\
+	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:sc:jacobisubblocksGamma2}
+}
 Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.
@@
 
-Following the triangle case \cite{olver2019triangle}, we use a multivariate analogue of Clenshaw's algorithm for evaluation, where we  combine each system in (\ref{eqn:jacobimatricesdefinitionalt}) into a block-tridiagonal system, for any $(x,y,z) \in \Omega$:
-\begin{align*}
+A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here.
+Combining each system in \bsrefeqn{eqn:sc:jacobimatricesdefinitionalt} into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yields:
+\bseqn{
 	\renewcommand\arraystretch{1.3}
@@
 		\end{pmatrix},
-\end{align*}
+}
 where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
-\begin{align*}
+\bseqn{
 	A_n &:= 
@@
 		\end{pmatrix} \in \R^{3(2n+1)\times(n+1)}.
-\end{align*}
+}
 
@@
 Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
-\begin{align*}
+\bseqn{
 	\begin{cases}
 		\bigscopta_{-1}(x,y,z) := 0 \\
-		\bigscopta_{0}(x,y) := \scopa_0 \\
-		\bigscopta_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
+		\bigscopta_{0}(x,y,z) := \scopa_0 \\
+		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
 	\end{cases}
-\end{align*}
+}
 
 Note that we can define an explicit $\Dnt$ as follows:
-\begin{align*}
+\bseqn{
 	\Dnt := 
@@
 		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
-\end{align*}
-for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}) for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by
-\begin{align*}
+}
+for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in \bsrefeqn{eqn:sc:jacobisubblocksGamma1} and \bsrefeqn{eqn:sc:jacobisubblocksGamma2} for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by
+\bseqn{
 	\big(\bm{\eta}_{0}\big)_j &= 
@@
 		\end{cases}
-\end{align*}
+}
 For $n=0$, we can simply take
-\begin{align*}
+\bseqn{
 	D^\top_0 &:= 
@@
 		\end{pmatrix} \in \R^{3\times3}.
-\end{align*}
+}
 
@@
 
-\begin{definition}\label{def:clenshawmats}
+\begin{definition}\label{def:sc:clenshawmats}
 	The \textit{recurrence coefficient matrices} associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
@@
 %
-\section{Sparse partial differential operators}\label{Section:PDOs}
+\section{Sparse partial differential operators}\label{Section:sc:PDOs}
 
 In this section we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process.
-To this end, as alluded to in Section~\ref{subsection:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal.
+To this end, as alluded to in \bsrefsection{section:sc:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal.
 Let $N \in \N$ and define:
-\begin{align}
+\bseqnnumber{
 	\bigscopNka &:= 
@@
 			\scopa_{N,k,1}(x,y,z)
-		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
+		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:sc:OPdefNka} \\ 
 	\bigscopa_{N,0} &:= 
@@
 			\scopa_{N,0,0}(x,y,z) \\
-		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
+		\end{pmatrix} \in \R^{N+1}, \label{eqn:sc:OPdefN0a} \\
 	\bigscopNa &:= 
@@
 			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
-\end{align}
+		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:sc:OPdefNa}
+}
 
 We further denote the weighted set of OPs on $\Omega$ by 
-\begin{align*}
+\bseqn{
 	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z),
-\end{align*}
+}
 
 The operator matrices we derive here act on coefficient vectors, that represent a function $f(x,y,z)$ defined on $\Omega$ in spectral space -- such a function is approximated by its expansion up to degree $N$:
-\begin{align*}
+\bseqn{
 	f(x,y,z) = \bigscopNa(x,y,z)^\top \vec f = \sum_{n=0}^{N} \sum_{k=0}^{n} \sum_{i=0}^{1} f_{n,k,i} \: \scopnkia(x,y,z),
-\end{align*}
+}
 where $\vec f = (f_{n,k,i})$ is the coefficients vector for the function $f$.
 
-\begin{definition}\label{def:differentialoperators}
-	Let $a$ be a nonnegative parameter, and $\tilde a \ge 2$ be a positive integer.
+\begin{definition}\label{def:sc:differentialoperators}
+	Let $a$ be a nonnegative parameter, and $\tilde a \ge 2$ be an integer.
 Define the operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ according to:
-\begin{align*}
+\bseqn{
 	\rho \pfpx{\varphi}{f} (x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \mathbf{f}, \\
@@
 	\DeltaS \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
-\end{align*}
+}
 \end{definition}
@@
 The operators we define here are for partial derivatives with respect to the spherical coordinates $(\varphi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).
-With the OP ordering by Fourier mode $k$ defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) these rotationally invariant operators are block-diagonal, meaning simple and parallelisable practical application.
-
-\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
-	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
+With the OP ordering by Fourier mode $k$ defined in \bsrefeqns{eqn:sc:OPdefNka}{eqn:sc:OPdefNa} these rotationally invariant operators are block-diagonal, meaning simple and parallelisable practical application.
+
+\begin{theorem}\label{theorem:sc:sparsityofdifferentialoperators}
+	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ from \bsrefdef{def:sc:differentialoperators} are sparse, with banded-block-banded structure.
 More specifically:
@@
 
-In order to show the last part of Theorem \ref{theorem:sparsityofdifferentialoperators}, we require the following short lemma.
-
-\begin{lemma}\label{lemma:Rsecondderivative}
+In order to show the last part of \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}, we require the following short lemma.
+
+\begin{lemma}\label{lemma:sc:Rsecondderivative}
 	For any general parameter $a$ and any $n = 0,1,\dots$, $k = 0,\dots,n$ we have that
-\begin{align*}
+\bseqn{
 	&\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
@@
 	&\quad \quad = \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
-\end{align*}
+}
 where 
-\begin{align*}
+\bseqn{
 	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
-\end{align*}
+}
 \end{lemma}
 
-\begin{proof}[Proof of Lemma \ref{lemma:Rsecondderivative}]
+\begin{proof}[Proof of \bsreflemma{lemma:sc:Rsecondderivative}]
 	Since $\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] = \genjacw^{(a, 2k)} \: r_{n-k+1}$ where $r_{n-k+1}$ is a degree $n - k + 1$ polynomial, we have that 
-\begin{align*}
+\bseqn{
 	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)}
-\end{align*}
+}
 for some coefficients $\tilde c_{\{n,k\},m}$. These coefficients are given by
-\begin{align*}
+\bseqn{
 	\tilde c_{\{n,k\},m} &= \frac{1}{\normgenjac^{(a,2k)}} \: \ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
 	&= - \frac{1}{\normgenjac^{(a,2k)}}  \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
-\end{align*}
+}
 We show that these are zero for $m < n - k - 1$ by integrating twice by parts:
-\begin{align*}
+\bseqn{
 	&\ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
@@
 	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - 2(k+1) z \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(1, 0)} + \genjac_{m}^{(a,2k) \: \prime \prime} \: \genjacw^{(1, 2)}] \: \genjacw^{(a, 2k)} \: \D z
-\end{align*}
+}
 which is indeed zero for $m < n - k - 1$ by orthogonality.
@@
 
-\begin{proof}[Proof of Theorem \ref{theorem:sparsityofdifferentialoperators}]
+\begin{proof}[Proof of \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}]
 
 For the operator $D_\theta$ for partial differentiation by $\theta$, we simply have that
-\begin{align*}
+\bseqn{
 	\pptheta \scopnkia(x,y,z) &= \genjacnmk^{(a,2k)}(z) \: \rho(z)^{k} \: \ddx{\theta} \chki(\theta) \\
@@
 where the coefficients are 
-\begin{align*}
+\bseqn{
 	c_{m,j,h} = \norm{\scopmjh^{(a+1)}}^{-2}_{\scop^{(a+1)}} \ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}}.
-\end{align*}
+}
 Now, note that:
-\begin{align*}
+\bseqn{
 	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \\
-	\rho(z) \: \rho'(z) &= -z \\
+	\rho(z) \: \rho'(z) &= -z, \\
 	\ppphi \scopnkia(x,y,z) &= -\rho(z) \: \ddz \Big[ \rho(z)^k \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta), \\
 	\ppphi \Big[ \genjacw^{(a,0)}(z) \: \scopnkia(x,y,z) \Big] &= -\rho(z) \: \ddz \Big[ \genjacw^{(a,k)}(z) \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta).
-\end{align*}
+}
 Then, 
-\begin{align*}
+\bseqn{
 	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
@@
 	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
-\end{align*}
+}
 which is zero for $j \ne k$, $h \ne i$, and $m < n - 2$ by orthogonality.
@@
 Similarly for the operator $W_\varphi^{(a)}$ for partial differentiation by $\varphi$ on the weighted space, the entries of the operator are given by the coefficients in the expansion $\rhoppphi (\genjacw^{(a,0)} \: \scopnkia) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-1,0)} \: \scopmjh^{(a-1)}$, where the coefficients are
-\begin{align*}
+\bseqn{
 	c_{m,j,h} = \norm{\scopmjh^{(a-1)}}^{-2}_{\scop^{(a-1)}} \ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}}.
-\end{align*}
+}
 Now,
-\begin{align*}
+\bseqn{
 	&\ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}} \\
@@
 	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
-\end{align*}
+}
 which is zero for $j \ne k$, $h \ne i$, and $m < n - 1$ by orthogonality.
@@
 We move on to the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yield
-\begin{align}
+\bseqnnumber{
 	\DeltaS \: \scopnki^{(a)} &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
@@
 	&= \chki(\theta) \rho(z)^k \Big\{ -k(k+1) \genjacnmk^{(a,2k)}(z) - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime}(z) \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime}(z) \Big\}, \label{eqn:oplaplacian} \\
+	&\quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime}(z) \Big\}, \label{eqn:sc:oplaplacian} \\
 	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
@@
 	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime}(z) \big[ -2(k+1) z \: \genjacw^{(a,k)}(z) + 2a \: \genjacw^{(a-1,k+2)}(z) \big] \nonumber \\
-	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime}(z) \: \genjacw^{(a,k+2)}(z) \Big\}. \label{eqn:woplaplacian}
-\end{align}
+	&\quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime}(z) \: \genjacw^{(a,k+2)}(z) \Big\}. \label{eqn:sc:woplaplacian}
+}
 
@@
  where the coefficients are
-\begin{align*}
+\bseqn{
 	c_{m,j,h} = \norm{\scopmjh^{(a + \tilde a)}}^{-2}_{\scop^{(a + \tilde a)}} \ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}}.
-\end{align*}
-Using \bsrefeqn{eqn:oplaplacian}, and integrating by parts twice, we then have that
-\begin{align*}
+}
+Using \bsrefeqn{eqn:sc:oplaplacian}, and integrating by parts twice, we then have that
+\bseqn{
 	&\ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}} \\
@@
 	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z
-\end{align*}
+}
 where $r_{m-k+\tilde a}$ is a degree $m - k + \tilde a$ polynomial in $z$, and so the above is zero for $n  - k > m - k + \tilde a \iff m < n - \tilde a$.
@@
 For the operator $\calL_W^{(a)\to(a-\tilde a)}$ for the surface Laplacian on a weighted space, the entries of the operator are given by the coefficients in the expansion 
-\begin{align*}
+\bseqn{
 	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \genjacw^{(a-\tilde a,0)} \: \scopmjh^{(a - \tilde a)}, 
-\end{align*}	
+}	
 where the coefficients are
-\begin{align*}
+\bseqn{
 	c_{m,j,h} = \norm{\scopmjh^{(a - \tilde a)}}^{-2}_{\scop^{(a - \tilde a)}} \ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}}.
-\end{align*}
-Using \bsrefeqn{eqn:woplaplacian}, and integrating by parts thrice, we then have that
-\begin{align*}
+}
+Using \bsrefeqn{eqn:sc:woplaplacian}, and integrating by parts thrice, we then have that
+\bseqn{
 	&\ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}} \\
@@
 	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z
-\end{align*}
+}
 where $r_{m-k}$ is a degree $m - k$ polynomial in $z$, and so the above is zero for $n - k > m - k \iff m < n$.
@@
 For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are given by
-\begin{align*}
+\bseqn{
 	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
-\end{align*}
-Using \bsrefeqn{eqn:woplaplacian} with $a = 1$, and \bsreflemma{lemma:Rsecondderivative}, we then have that
-\begin{align*}
+}
+Using \bsrefeqn{eqn:sc:woplaplacian} with $a = 1$, and \bsreflemma{lemma:sc:Rsecondderivative}, we then have that
+\bseqn{
 	&\ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
@@
 	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z
-\end{align*}
-where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in \bsreflemma{lemma:Rsecondderivative}.
+}
+where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in \bsreflemma{lemma:sc:Rsecondderivative}.
 \end{proof}
@@
 It is therefore necessary to also be able to raise or lower the parameter by way of an independent operator.
-There exist conversion matrix operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
-\begin{definition}\label{def:parametertransformationoperators}
+There exist conversion  operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+\begin{definition}\label{def:sc:parametertransformationoperators}
 Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T_W^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
-\begin{align*}
+\bseqn{
 	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z) \\
 	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z)
-\end{align*}
+}
 \end{definition}
 
-\begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
-The operator matrices in Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
+\begin{lemma}\label{lemma:sc:sparsityofparametertransformationoperators}
+The operator matrices in \bsrefdef{def:sc:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
 \begin{itemize}
@@
 Note that
-\begin{align*}
+\bseqn{
 	t_{m,j} = \norm{\scopmjh^{(a+\tilde{a})}}^{-2}_{\scop^{(a+\tilde a)}} \: \ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}}.
-\end{align*}
+}
 where
-\begin{align*}
+\bseqn{
 	\ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}} &= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \cdot \: \Big( \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjacmmj^{(a+\tilde a,2j)} \: \rho^{k+j} \: \genjacw^{(a + \tilde a, 0)} \: \D z \Big) \\
 	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a + \tilde a, 2k)} \: \D z
-\end{align*}
+}
 which is zero for $n > m + \tilde a \iff m < n - \tilde a$. The sparsity argument for the weighted parameter transformation operator follows similarly.
@@
 	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
-For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
-	\label{fig:sparsity}
+For (c), the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
+	\label{fig:sc:sparsity}
 \end{figure}
 
-\subsection{Further partial differential operators}\label{subsection:furtherdiffoperators}
+\subsection{Further partial differential operators}\label{section:sc:furtherdiffoperators}
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
 As a canonical example, we can obtain the matrix operator for the $\rho^2$-factored spherical Laplacian $\rho(z)^2 \: \DeltaS$, that will take us from coefficients for expansion in the weighted space $\bigW_N^{(1)}(x,y,z) = \genjacw^{(1,0)}(z) \: \bigscopN^{(1)}(x,y,z)$ to coefficients in the non-weighted space $\bigscopN^{(1)}(x,y,z)$.
-Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in Definition \ref{def:differentialoperators}.
+Note that this construction will ensure the imposition of the Dirichlet zero boundary conditions on $\Omega$, similar to how the Dirichlet zero boundary conditions would be imposed for the operator $\Delta^{(1)}_W$ in \bsrefdef{def:sc:differentialoperators}.
 The matrix operator for this $\rho^2$-factored spherical Laplacian acting on the coefficients vector is then given by
-\begin{align*}
+\bseqn{
 	D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2.
-\end{align*}
-Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+}
+Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in \bsreffig{fig:sc:sparsity}.
 
 Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions. That is, 
-\begin{align*}
+\bseqn{
 	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \nabla_S u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
-\end{align*}
+}
 where $\partial\Omega$ is the $z=\alpha$ boundary, and $\unitvec{n}(x,y,z)$ is the outward unit normal vector at the point $(x,y,z)$ on the boundary, i.e.
@@
 The matrix operator for the Biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
-To construct this, we can simply multiply together two of the spherical Laplacian operators defined in \bsrefdef{def:differentialoperators}, namely $\calL^{(0)\to(2)}$ and $\calL_W^{(2)\to(0)}$:
-\begin{align*}
+To construct this, we can simply multiply together two of the spherical Laplacian operators defined in \bsrefdef{def:sc:differentialoperators}, namely $\calL^{(0)\to(2)}$ and $\calL_W^{(2)\to(0)}$:
+\bseqn{
 	\calB^{(2)}_W := \calL^{(0)\to(2)} \: \calL_W^{(2)\to(0)}.
-\end{align*}
+}
 Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,y,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$ and thus its spherical gradient would be zero on the boundary $z = \alpha$.
 This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
-The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
+The sparsity and structure of this biharmonic operator are seen in \bsreffig{fig:sc:sparsity}.
+
+\subsection{Stability of the Laplacian operator}
+
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}[b]{0.55\textwidth}
+		\includegraphics[scale=0.4]{condition-numbers-laplacian-log-N=200}
+	\end{subfigure}
+	\hfill%
+
+	\begin{subfigure}[b]{0.55\textwidth}
+		\includegraphics[scale=0.4]{condition-numbers-max-laplacian-log}
+	\end{subfigure}
+	\hfill%
+	\caption{Plots of the condition number for the Laplacian operator matrix $\Delta^{(1)}_W$ and the preconditioned matrix $P^{-1} \Delta^{(1)}_W$ where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
+Top: The condition numbers of each diagonal block of $\Delta^{(1)}_W$ for $N = 200$, as well as those for $P^{-1} \Delta^{(1)}_W$.
+Bottom: The maximum condition number of all diagonal blocks of the Laplacian and the preconditioned Laplacian operators, as the degree $N$ increases.}
+	\label{fig:sc:conditionnumbers}
+\end{figure}
+
+Denote the condition number of the matrix $A$: 
+\bseqn{
+	\kappa(A) := \norm{A}_2 \norm{A^{-1}}_2,
+}
+which encodes how accurate the solution to $A \vc x = \vc b$ is under perturbation.
+Since the Laplacian operator is block-diagonal one can determine the condition number from the condition numbers of each block on the diagonal,  and hence deduce the accuracy in solving the Poisson equation, where graphical representations are seen in \bsreffig{fig:sc:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
+As the degree $N$ increases, the condition numbers of these blocks becomes grows  algebraically fast (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
+Fortunately however, by applying a trivial diagonal preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in Figure \bsreffig{fig:sc:conditionnumbers}.
+The preconditioning matrix chosen is $P^{-1}$, where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
 
 %
-\section{Computational aspects}\label{Section:Computation}
+\section{Computational aspects}\label{Section:sc:Computation}
 
@@
 
-It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
-
-\subsection{Quadrature rule on the spherical cap}\label{subsection:quadrule}
-
-In this section we construct a quadrature rule exact for polynomials on the spherical cap $\Omega$ that can be used to expand functions in the OPs $\scopnkia(x,y,z)$ for a given parameter $a$.
-
-\begin{theorem}\label{Theorem:quadrule}
+It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:sc:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
+
+\subsection{Quadrature rule on the spherical cap}\label{section:sc:quadrule}
+
+In this section we construct a quadrature rule that is exact for polynomials on the spherical cap $\Omega$ that can be used to expand functions in the OPs $\scopnkia(x,y,z)$ for a given parameter $a$.
+
+\begin{theorem}\label{theorem:sc:quadrule}
 	Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
@@
 Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
-\begin{align*}
-	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
+\bseqn{
+	x_{l+(j-1)M_2} &:= \rho(t_j) \: s_l, \\
+	y_{l+(j-1)M_2} &:= \rho(t_j) \: \sqrt{1-s_l^2}, \\
 	z_{l+(j-1)M_2} &:= t_j, \\
 	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
-\end{align*}
+}
 Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
-\begin{align*}
+\bseqn{
 	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \approx \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
-\end{align*}
+}
 where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
@@
 Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-\begin{align*}
+\bseqn{
 	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
 	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
-\end{align*}
+}
 so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
@@
 Firstly, we note that
-\begin{align*}
-	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta = \int_{-1}^1 \Big( g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big) \frac{\D x}{\sqrt{1-x^2}}
-\end{align*}
-for some function $g$. Then, integrating the even function $f_e$ we have
-\begin{align*}
+\bseqn{
+	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta &= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}} \\
+	&= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(-x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}}
+}
+for any function $g$, using a change of variables $x \to -x$ for the second term in the integral. Then, integrating the even function $f_e$ we have
+\bseqn{
 	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \D x \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_e\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
+	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
 	&\quad \quad \approx \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
@@
 	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
-\end{align*}
+}
 Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
@@
 Integrating the odd function $f_o$ results in
-\begin{align*}
+\bseqn{
 	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) + f_o\big(\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
-	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) - f_o\big(\rho(z)x, \rho(z)\sqrt{1-x^2}, z\big) \big] \: \D x \Big) \: \D z \\
+	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 \Big[ f_o\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \\
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_o\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
 	&\quad \quad = 0.
-\end{align*}
+}
 since $f_o(x,y,z) = - f_o(-x, -y, z)$. Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
-\begin{align*}
+\bseqn{
 	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D A &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D A  \\
@@
 	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
-\end{align*}
+}
 where $M = M_1 M_2$ and $2M_1 - 1 \ge N, 2M_2 - 1 \ge N$.
@@
 
-%\begin{theorem}\label{Theorem:quadrule}
-%Let $M_1, M_2 \in \N$ and denote the $M_1$ Gauss quadrature nodes and weights on $[\alpha,1]$ with weight $(t - \alpha)^a$ as $(t_j, w_j^{(t)})$.
-Further, we denote the $M_2$ Gauss quadrature nodes and weights for the unit circle \cite{???}\bstodo{good reference?
-Or derive myself using Chebychev polys...} by $(\mathbf{s}_j, w_j^{(s)})$ where for $j = 1,\dots,M_2$:
-%\begin{align*}
-%	\mathbf{s}_j &= \Big( \cos\fpr(\frac{2 \pi j}{M_2}), \: \sin\fpr(\frac{2 \pi j}{M_2}) \Big), \\
-%	w_j^{(s)} &= \frac{2\pi}{M_2}.
-%\end{align*}
-%Define for $j = 1,\dots,M_1, \: l=1,\dots,M_2$:
-%\begin{align*}
-%	\big(x_{l+(j-1)M_2}, \: y_{l+(j-1)M_2} \big) &:= \rho(t_j) \: \mathbf{s}_l, \\
-%	z_{l+(j-1)M_2} &:= t_j, \\
-%	w_{l+(j-1)M_2} &:= w_j^{(t)} w_l^{(s)}.
-%\end{align*}
-%Let $f(x,y,z)$ be a function on $\Omega$, and $N \in \N$. The quadrature rule is then
-%\begin{align*}
-%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j, z_j) + f(-x_j, -y_j, z_j) \big],
-%\end{align*}
-%where $M = M_1 \: M_2$, and the quadrature rule is exact if $f(x,y,z)$ is a polynomial of degree $\le N$ with $M_1 \ge \half (N+1), M_2 \ge N+1$.
-%\end{theorem}
-%
-%\begin{proof}
-%Let $f : \Omega \to \R$. Define the functions $f_e, f_o : \Omega \to \R$ by 
-%\begin{align*}
-%	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
-%	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
-%\end{align*}
-%so that $\xvec \mapsto f_e(\xvec, z)$ for fixed $z$ is an even function, and $\xvec \mapsto f_o(\xvec, z)$ for fixed $z$ is an odd function.
-Note that if $f$ is a polynomial, then $f_e(\rho(t)\xvec, t)$ is a polynomial in $t \in [\alpha,1]$ for fixed $\xvec \in \R^2$.
-%
-%Now, we have that
-%\begin{align*}
-%	&\int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_e\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_e\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_e\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \\
-%	&\quad \quad \approx\int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) \mathbf{s}_l, z\big) \Big) \: \D z \quad (\star) \\
-%	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) \mathbf{s}_l, t_j\big) \quad (\star \star) \\
-%	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
-%\end{align*}
-%Suppose $f$ is a polynomial in $x,y,z$ of degree $N$, and hence that $f_e$ is a degree $\le N$ polynomial.
-We therefore achieve equality at $(\star)$ if $M_2 - 1 \ge N$ and we achieve equality at $(\star \star)$ if also $2M_1 - 1 \ge N$.
-\bstodo{For the circle rule, is this bound on N true?}
-%
-%Next, note that 
-%\begin{align*}
-%	&\int_\Omega f_o(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&\quad \quad = \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} f_o\big(\cos\theta\sin(\varphi), \sin\theta\sin(\varphi), \cos(\varphi)\big) \: \genjacw^{(a,0)}\big(\cos(\varphi)\big) \: \sin(\varphi) \: \D \theta \: \D \varphi \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_0^{2\pi} f_o\big(\rho(z)\cos\theta,\rho(z)\sin\theta, z)\big) \: \D \theta \Big) \: \D z \\
-%	&\quad \quad = \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{\norm{\bm{\xi}} = 1} f_o\big(\rho(z)\bm{\xi}, z)\big) \: \D \sigma(\bm{\xi}) \Big) \: \D z \quad (\dagger) \\
-%	&\quad \quad = 0
-%\end{align*}
-%since the inner integral at $(\dagger)$ over $\bm{\xi}$ is zero, due to the symmetry over the domain.
-%
-%Hence, for a polynomial $f$ in $x,y,z$ of degree $N$,
-%\begin{align*}
-%	\int_\Omega f(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z &= \int_\Omega \Big(f_e(x,y,z) + f_o(x,y,z)\Big) \:  \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z  \\
-%	&= \int_\Omega f_e(x,y,z) \: \genjacw^{(a,0)}(z) \: \D \sigma(x,y) \: \D z \\
-%	&= \sum_{j=1}^{M}  w_j \: f_e(x_j, y_j, z_j),
-%\end{align*}
-%where $M = M_1 M_2$ and $2M_1 - 1 \ge N, M_2 - 1 \ge N$.
-%\end{proof}
-
-\subsection{Obtaining the coefficients for expansion of a function on the spherical cap}\label{subsection:expandingfunctions}
+\subsection{Obtaining the coefficients for expansion of a function on the spherical cap}\label{section:sc:expandingfunctions}
 
 Fix $a \in \R$. Then for any function $f : \Omega \to \R$ we can express $f$ by
-\begin{align*}
+\bseqn{
 	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \bm{f}_k = \bigscopNa(x,y,z)^\top \: \bm{f}
-\end{align*}
-for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in equations (\ref{eqn:OPdefNka}, \ref{eqn:OPdefN0a}, \ref{eqn:OPdefNa}) and where
-
-\begin{align*}
+}
+for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in \bsrefeqns{eqn:sc:OPdefNka}{eqn:sc:OPdefNa} and where
+
+\bseqn{
 	\bm{f}_k &:= 
@@
 	f_{n,k,i} := \ip< f, \: \scopnkia>_{\scopa} \: \norm{\scopnkia}^{-2}_{\scopa}.
-\end{align*}
-Recall from equation (\ref{eqn:normscop}) that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$.
-Using the quadrature rule detailed in Section \ref{subsection:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$:
-\begin{align*}
+}
+Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopnkia}^2_{\scopa} = \normgenjac^{(a,2k)} \: \pi$.
+Using the quadrature rule detailed in \bsrefsection{section:sc:quadrule} for the inner product, we can calculate the coefficients $f_{n,k,i}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$, $i = 0,1$:
+\bseqn{
 	f_{n,k,i} &= \frac{1}{2 \: \normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
 	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
-\end{align*}
-where the quadrature nodes and weights are those from Theorem \ref{Theorem:quadrule}, and $M = M_1 M_2$ with $2M_1 - 1 \ge N, M_2 - 1 \ge N$ (i.e.
+}
+where the quadrature nodes and weights are those from \bsreftheorem{theorem:sc:quadrule}, and $M = M_1 M_2$ with $2M_1 - 1 \ge N, M_2 - 1 \ge N$ (i.e.
 we can choose $M_2 := N + 1$ and $M_1 := \ceil{\frac{N+1}{2}}$).
 
-\subsection{Function evaluation}\label{subsection:functionevaluation}
-
-For a function $f$, with coefficients vector $\bm{f}$ for expansion in the $\{\scopnki\}$ basis as determined via the method in Section \ref{subsection:expandingfunctions} up to order $N$, we can use the Clenshaw algorithm to evaluate the function at a point $(x,y,z) \in \Omega$ as follows.
-Let $A_n, B_n, \Dnt, C_n$ be the Clenshaw matrices from Definition \ref{def:clenshawmats}, and define the rearranged coefficients vector $\tilde{\bm{f}}$ via
-\begin{align*}
+\remark{While it may be possible to leverage fast transforms, as in \cite{slevinsky2018use,slevinsky2019fast}, to speed up the calculation of the coefficients, this remains an open question.}
+
+\subsection{Function evaluation}\label{section:sc:functionevaluation}
+
+For a function $f$, with coefficients vector $\bm{f}$ for expansion in the $\{\scopnki\}$ basis as determined via the method in \bsrefsection{section:sc:expandingfunctions} up to order $N$, we can use the Clenshaw algorithm to evaluate the function at a point $(x,y,z) \in \Omega$ as follows.
+Let $A_n, B_n, \Dnt, C_n$ be the Clenshaw matrices from \bsrefdef{def:sc:clenshawmats}, and define the rearranged coefficients vector $\tilde{\bm{f}}$ via
+\bseqn{
 	{\bm{f}}_n &:= 
@@
 		\end{pmatrix} \in \R^{(N+1)^2}.
-\end{align*}
+}
 The trivariate Clenshaw algorithm works similar to the bivariate Clenshaw algorithm introduced in \cite{olver2019triangle} for expansions in the triangle:
-\begin{align*}
+\bseqn{
 	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
@@
 	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
-\end{align*}
-
-\subsection{Calculating non-zero entries of the operator matrices}\label{subsection:Computation-operatormatrices}
-
-The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
+}
+
+\subsection{Calculating non-zero entries of the operator matrices}\label{section:sc:Computation-operatormatrices}
+
+The proofs of \bsreftheorem{theorem:sc:sparsityofdifferentialoperators} and \bsreflemma{lemma:sc:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in \bsrefdef{def:sc:differentialoperators} and \bsrefdef{def:sc:parametertransformationoperators}.
 We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^2)$.
@@
 
-\subsection{Obtaining operator matrices for variable coefficients}\label{subsection:operatorclenshaw}
-
-The Clenshaw algorithm outlined in Section \ref{subsection:functionevaluation} can also be used with Jacobi matrices $J^{(a)}_x, J^{(a)}_y, J^{(a)}_z$ replacing the point $(x,y,z)$.
+\subsection{Obtaining operator matrices for variable coefficients}\label{section:sc:operatorclenshaw}
+
+The Clenshaw algorithm outlined in \bsrefsection{section:sc:functionevaluation} can also be used with Jacobi matrices $J^{(a)}_x, J^{(a)}_y, J^{(a)}_z$ replacing the point $(x,y,z)$.
 Let $v : \Omega \to \R$ be the function that we wish to obtain an operator matrix $V$ for $v$, so that
-\begin{align*}
+\bseqn{
 	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \bm{f} \: \bigscopta(x,y,z) = (V \bm{f})^\top \: \bigscopta(x,y,z),
-\end{align*}
+}
 i.e. $V \bm{f}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$.
 
-To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in Section \ref{subsection:functionevaluation} so that $v(x,y,z) = {\bm{v}}^\top \: \bigscopta(x,y,z)$).
+To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in \bsrefsection{section:sc:functionevaluation} so that $v(x,y,z) = {\bm{v}}^\top \: \bigscopta(x,y,z)$).
 Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, $Z := (J_z^{(a)})^\top$.
 The operator $V$ is then the result of the following:
-\begin{align*}
+\bseqn{
 	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
@@
 	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
-\end{align*}
+	\quad &\text{3) } \text{Output: } V = v(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+}
 where at each iteration, $\bm{\xi}_n$ is a vector of matrices.
@@
 %
-\section{Examples on spherical caps with zero Dirichlet conditions}\label{Section:Examples}
+\section{Examples on spherical caps with zero Dirichlet conditions}\label{Section:sc:Examples}
 
@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{poisson-u=wyexpx-N=60}
+\begin{figure}[tp]
 	\centering
-	%\label{fig:solution-poisson}
+	\begin{subfigure}{0.85\textwidth}
+		\includegraphics[scale=0.4]{poisson-u=wyexpx-N=60}
 	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-poisson-epsilonfun-N=200}
-        	%\label{fig:solutionblocknorms-poisson}
+	\hfill%
+
+	\begin{subfigure}{0.55\textwidth}
+		\includegraphics[scale=0.4]{solutionblocknorms-poisson-epsilonfun-N=200}
 	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
-Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
+	\hfill%
+	\caption{Top: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$.
+Bottom: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values.
 This indicates spectral convergence.}
-	\centering
-	\label{fig:poisson}
+	\label{fig:sc:poisson}
 \end{figure}
 
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.25]{helmholtz-f=wyexpx-N=60-n=300}
-	\centering
-	%\label{fig:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\centering
-	\includegraphics[scale=0.3]{solutionblocknorms-helmholtz-varyingk-N=200}
-        	%\label{fig:solutionblocknorms-poisson}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
-Right: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
-This indicates spectral convergence.}
-	\centering
-	\label{fig:helmholtz}
-\end{figure}
-
-\begin{figure}[t]
-	\centering % <-- added
-	\includegraphics[scale=0.35]{complexity-Nend=1000}
-	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.
-This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution.
-Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
-	\centering
-	\label{fig:complexity}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.15]{biharmonic-f=erf-N=80-n=50}
-	\centering
-	%\label{fig:solution-biharmonic}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.3]{solutionblocknorms-biharmonic-expfun-N=200}
-	\centering
-        	%\label{fig:solutionblocknorms-biharmonic}
-	\end{subfigure}
-	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
-Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
-This demonstrates algebraic convergence.}
-	\centering
-	\label{fig:biharmonic}
-\end{figure}
 % END FIGURES
@@
 The Poisson equation is the classic problem of finding $u(x,y,z)$ given a function $f(x,y,z)$ such that:
-\begin{align}
+\bseqnnumber{
 	\begin{cases}
@@
 	\end{cases}
-	\label{eqn:poisson}
-\end{align}
+	\label{eqn:sc:poisson}
+}
 noting the imposition of zero Dirichlet boundary conditions on $u$.
@@
 Choose an $N \in \N$ large enough for the problem, and denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}.
+Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule in \bsrefsection{section:sc:expandingfunctions}.
 In matrix-vector notation, our system hence becomes:
-\begin{align*}
+\bseqn{
     \Delta_W^{(1)} \mathbf{u} = \mathbf{f}
-\end{align*}
+}
 which can be solved to find $\mathbf{u}$.
-In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
+In \bsreffig{fig:sc:poisson} we see the solution to the Poisson equation with zero boundary conditions given in \bsrefeqn{eqn:sc:poisson} in the disk-slice $\Omega$.
+In \bsreffig{fig:sc:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
 The right hand sides we choose here are given by
-\begin{align*}
+\bseqn{
 	f(x,y,z) &= \norm{\Big(x - (\epsilon + 1 / \sqrt{3}), \: y - (\epsilon + 1 / \sqrt{3}), \: z - (\epsilon + 1 / \sqrt{3})\Big)^\top}
-\end{align*}
+}
 for differing choices of $\epsilon$ -- this parameter serves to alter the distance from which we would have a singularity.
@@
 
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}{0.85\textwidth}
+		\includegraphics[scale=0.4]{helmholtz-f=wyexpx-N=60-n=300}
+	\end{subfigure}
+
+	\begin{subfigure}{0.55\textwidth}
+		\includegraphics[scale=0.4]{solutionblocknorms-helmholtz-varyingk-N=200}
+	\end{subfigure}
+	\caption{Top: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y,z) = y e^x (z - \alpha) $, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
+Bottom: The norms of each block of the computed solution of the Helmholtz equation with the right hand side function $f(x,y,z) = 1$ and the same function $v(x,y,z)$, for various $k$ values.
+This indicates spectral convergence.}
+	\label{fig:sc:helmholtz}
+\end{figure}
+
+\begin{figure}[tp]
+	\centering % <-- added
+	\includegraphics[scale=0.4]{complexity-Nend=1000}
+	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.
+This demonstrates that the approach is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution (the number of unknowns is then $(N+1)^2$).
+Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$ and $v(x,y,z) = v(z) = \cos(z)$.}
+	\label{fig:sc:complexity}
+\end{figure}
+
 Find $u(x,y)$ given functions $v$, $f : \Omega \to \R$ such that:
-\begin{align}
+\bseqnnumber{
 	\begin{cases}
@@
 	\end{cases}
-	\label{eqn:helmholtz}
-\end{align}
+	\label{eqn:sc:helmholtz}
+}
 where $k \in \R$, noting the imposition of zero Dirichlet boundary conditions on $u$.
@@
 Denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in Section \ref{subsection:expandingfunctions}.
+Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in \bsrefsection{section:sc:expandingfunctions}.
 
@@
 In matrix-vector notation, our system hence becomes:
-\begin{align*}
+\bseqn{
     (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \mathbf{u} = \mathbf{f}
-\end{align*}
+}
 which can be solved to find $\mathbf{u}$.
-We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y,z) = zxy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
+We can see the sparsity and structure of this matrix system in \bsreffig{fig:sc:sparsity} with $v(x,y,z) = zxy^2$ as an example.
+In \bsreffig{fig:sc:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in \bsrefeqn{eqn:sc:helmholtz} in the spherical cap $\Omega$, with $f(x,y,z) = y e^x \genjacw^{(1,0)}(z)$, $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 100$.
+In \bsreffig{fig:sc:helmholtz} we also show the norms of each block of calculated coefficients for the approximation of the solution to the inhomogeneous variable-coefficient Helmholtz equation with various $k$ values.
 Here, we use $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
@@
 
-In Figure~\ref{fig:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools.jl package \cite{BenchmarkTools.jl-2016} in Julia.} to construct the operator for $\DeltaS + v(x,y,z)$, with a rotationally invariant $v(x,y,z) = v(z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
+In \bsreffig{fig:sc:complexity} we plot the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools.jl package \cite{BenchmarkTools.jl-2016} in Julia.} to construct the operator for $\DeltaS + v(x,y,z)$, with a rotationally invariant $v(x,y,z) = v(z) = \cos z$, and solve a zero boundary condition Helmholtz problem.
 The plot demonstrates that as we increase the degree of approximation $N$, we achieve a complexity of an optimal $\bigO(N^2)$.
@@
 In this case, the problem
-\begin{align*}
+\bseqn{
 	\begin{cases}
@@
 	\end{cases}
-\end{align*}
+}
 is equivalent to letting $u(x,y,z) = \tilde{u}(x,y,z)+ c(x,y)$ and solving
-\begin{align*}
+\bseqn{
 	\begin{cases}
@@
 	\end{cases}
-\end{align*}
+}
 for $\tilde u$. This new problem is then a zero boundary condition Helmholtz problem with right hand side 
-\begin{align*}
+\bseqn{
 	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)
-\end{align*}
+}
 for $(x,y,z) \in \Omega$. Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
-\begin{align*}
+\bseqn{
 	\DeltaS \: c(x,y) = {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) = - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
-\end{align*}
+}
 since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only, which are precisely the Fourier coefficients of $c(\cos \theta, \sin \theta)$.
-Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manor.
+Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manner.
 
@@
 
-Our last erxample is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
-\begin{align}
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}{0.85\textwidth}
+		\includegraphics[scale=0.4]{biharmonic-f=erf-N=80-n=50}
+	\end{subfigure}
+	\hfill%
+
+	\begin{subfigure}{0.55\textwidth}
+		\includegraphics[scale=0.4]{solutionblocknorms-biharmonic-expfun-N=200}
+	\end{subfigure}
+	\hfill%
+	\caption{Left: The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y,z) = (1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))) \rho(z)^2$.
+Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
+This demonstrates algebraic convergence.}
+	\label{fig:sc:biharmonic}
+\end{figure}
+
+Our last example is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
+\bseqnnumber{
 	\begin{cases}
@@
 	\end{cases}
-	\label{eqn:biharmonic}
-\end{align}
+	\label{eqn:sc:biharmonic}
+}
 where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see Section \ref{subsection:furtherdiffoperators}).
-In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the spherical cap $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
+For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see \bsrefsection{Section:sc:furtherdiffoperators}).
+In \bsreffig{fig:sc:biharmonic} we see the solution to the Biharmonic \bsrefeqn{eqn:sc:biharmonic} in the spherical cap $\Omega$.
+In \bsreffig{fig:sc:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.
 Once again, the rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution, and we see that we achieve exponential convergence for these more complex functions.
+
+\subsection{Other boundary conditions}
+
+Other boundary conditions beyond those already presented here are beyond the scope of this paper.
+However, the approach established for the case of a triangle domain in 2D shows how Neumann conditions can be considered  as a system of equations in partial derivatives involving Dirchlet conditions \cite{olver2019triangle}.
+This approach may also work here for the spherical cap.
+Ideally, one would properly tackle the tangent space (which again is beyond the scope of this paper) which would lower the degrees of freedom.
 
@@
 
-This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere---it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
+This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere such as spherical triangles.
+It is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct the necessary orthogonal polynomials.
+Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct, or perhaps {\it compute}, the necessary orthogonal polynomials.
 
 The next stage is to develop an orthogonal basis for the tangent space of the spherical cap (or band), and obtain sparse differential operators for gradient, divergence etc.
-On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the choice of basis will not be as straightforward.
-
-%\noindent{\bf Acknowledgements}:  We would like to thank the anonymous referees for their helpful comments.
-The second author was supported in part by the Leverhulme Trust Research Project Grant RPG-2019-144 ``Constructive approximation theory on and inside algebraic curves and surfaces.
-
-%
-%%
-%\appendix
-%%
-%\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
-%
-%\bstodoinline{Is this worth including?}
-%
-%\sotodoinline{It's already a very long paper so probably better to skip this. But include it in your thesis} \bstodoinline{Yes thought so too}
-%
-%%\begin{figure}[t]
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-%%        	%\label{fig:solutionblocknorms-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-poisson}
-%%	\centering
-%%	%\label{fig:solution-poisson}
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.3\textwidth}
-%%	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
-%%	%\label{fig:solution-poisson}
-%%	\centering
-%%	\end{subfigure}
-%%	\centering
-%%	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
-Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
-Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-in the trapezium.}
-%%	\centering
-%%	\label{fig:appendixfigs}
-%%\end{figure}
-%
-%We follow the method of \cite{beuchler2006new} to construct a sparse $p$-finite element method in terms of the operators constructed above, with the benefit of ensuring that the resulting discretisation is symmetric.
-Consider the 3D Dirichlet problem on a domain $\Omega$:
-%\begin{align*}
-%	\begin{cases}
-%         - \rho(z)^2 \: \DeltaS u(x,y,z) = \rho(z)^2 \: f(x,y,z) \quad \text{in } \Omega \\
-%         u = 0 \quad \text{on } \partial \Omega
-%         \end{cases}
-%\end{align*}
-%This has the weak formulation
-%\begin{align*}
-%	L(v) := \int_\Omega \rho^2 \: f \: v \: \D \sigma(\bm{x}, z) = \int_\Omega \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) =: a(u,v)
-%\end{align*}
-% for any test function $v \in V := H_0^1(\Omega) = \{v \in H^1(\Omega) \quad | \quad v \rvert_{\partial \Omega} = 0 \}$.
-%
-%In general, we would split a sphere domain into spherical band elements with spherical caps over the poles, and use a $p$-finite element method to solve PDEs on the sphere using these elements $\{\element_n\}$.
-While spherical bands are an extension to this work, we note that we can still apply a $p$-FEM using two half-spheres ($\alpha = 0$).
-In this section, we limit our discretisation to a single element, that is we let $\element = \Omega$ for a spherical cap domain.
-We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p \rvert_\element) \le p\}$ for some $p \in \N$.
-%
-%We seek $u_p \in V_p$ s.t.
-%\begin{align}
-%	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-%	\label{eqn:FEMweakform}
-%\end{align}
-%
-%Define 
-%\begin{align*}
-%	\Lambda^{(a)} :=  \ip< \bigscopa_p, \: {\bigscopa_p}^\top >_{\scopa} := \int_\element \bigscopa_p \: {\bigscopa_p}^\top \: \genjacw^{(a,0)} \: \D \sigma(\bm{x}, z) = \int_\element \bigW_p^{(a)} \: {\bigscopa_p}^\top \: \D \sigma(\bm{x}, z)
-%\end{align*}
-%for any $a$.
-Note that due to orthogonality this is a diagonal matrix.
-By choosing suitable bases for our FE spaces, we can rewrite (\ref{eqn:FEMweakform}) in matrix-vector form.
-We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.
-%
-%Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
-%\begin{align*}
-%	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
-%					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
-%\end{align*}
-%Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
-Then,
-%\begin{align*}
-%	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
-%	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%
-%Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-%\begin{align*}
-%	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
-%\end{align*}
-%(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-%\begin{align*}
-%	A\mathbf{u} = \tilde{\mathbf{f}}.
-%\end{align*}
-%where the (element) stiffness matrix $A$ is defined by 
-%\begin{align*}
-%	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
-%\end{align*}
-%and the load vector $\tilde{\mathbf{f}}$ is given by 
-%\begin{align*}
-%	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-%
-
-\bibliography{spherical-caps}
-
-\end{document}
+On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the construction of a basis will not be as straightforward.

### 2021-05-26 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Fig sizes changed to be consistent and bigger.

@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
-        \centering
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
-        \centering
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.35]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
-        \centering
-	\end{subfigure}
+\begin{figure}[tp]
+	\centering % <-- added
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
+		\centering
+		\caption{The Laplacian operator $\laplacewiii$}
+	\end{subfigure}\hfil % <-- added
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
+		\centering
+		\caption{A variable coefficient Helmholtz operator}
+	\end{subfigure}\hfil % <-- added
+
+	\medskip
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.5]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
+		\centering
+		\caption{The biharmonic operator $\biharmonictwo$}
+	\end{subfigure}\hfil % <-- added
     	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
-Left: the Laplace operator $\laplacewiii$.
-Centre: the weighted variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
-Right: the biharmonic operator $\biharmonictwo$.}
+For (b), the weighted variable coefficient Helmholtz operator is $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.}
 	\label{fig:sparsity}
-        \centering
 \end{figure}
@@
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
-As a canonical example, we can obtain the matrix operator for the Laplacian \(\Delta\), that will take us from coefficients for expansion in the weighted space
+As a canonical example, we can obtain the matrix operator for the Laplacian $\Delta$, that will take us from coefficients for expansion in the weighted space
 \bseqn{
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
+		\includegraphics[scale=0.35]{solution-poisson-diskslice-alpha=0p2-beta=0p8}
 		\centering
@@
 		\centering
-	\includegraphics[scale=0.5]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
+		\includegraphics[scale=0.45]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
         		%\label{fig:solutionblocknorms-poisson}
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
+	\includegraphics[scale=0.35]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
+	\includegraphics[scale=0.35]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
 	\centering
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
+	\includegraphics[scale=0.35]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.3]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
+	\includegraphics[scale=0.35]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
 	%\label{fig:solution-helmholtz}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
+	\includegraphics[scale=0.45]{solutionblocknorms-helmholtz-diskslice-alpha=0p2-beta=0p8-N=196-k=20}
 	\centering
@@
 	\centering
-	\includegraphics[scale=0.3]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
+	\includegraphics[scale=0.35]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
 	%\label{fig:solution-biharmonic}
@@
 	\begin{subfigure}{0.5\textwidth}
-	\includegraphics[scale=0.5]{solutionblocknorms-biharmonic-diskslice-alpha=0p2-beta=0p8-N=200}
+	\includegraphics[scale=0.45]{solutionblocknorms-biharmonic-diskslice-alpha=0p2-beta=0p8-N=200}
 	\centering
@@
 
-\subsection{End-Disk-Slice}\label{Appendix:HalfDisk}
+\subsection{End-Disk-Slice}\label{section:ds:enddiskslice}
 
@@
 
-In Figure \ref{fig:appendixfigs} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
-
-\subsection{Trapeziums}\label{Appendix:Trapezium}
+In Figure \ref{fig:ds:trappfem} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
+
+\subsection{Trapeziums}\label{section:ds:trapezium}
 
@@
 
-In Figure \ref{fig:appendixfigs} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
+In Figure \ref{fig:ds:trappfem} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
 
@@
 
-%%
-%\section{Conclusions}
-%
-%
-%We have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve---notably here the disk-slice---with Dirichlet boundary conditions.
-This work extends  the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements, for example, by dividing the disk into disk slice elements.
-%%A further extension is to develop similar relationships for 3D domains such as the half-sphere surface.
-With this, we can obtain a family of OPs on the half-sphere, with sparse relations, and construct sparse spectral methods for surface PDEs on half-spheres.
-%This work serves as a stepping stone to constructing similar methods to solve partial differential equations on 3D sub-domains of the sphere, such as spherical caps and spherical triangles.
-In particular, orthogonal polynomials (OPs) in cartesian coordinates ($x$, $y$, and $z$) on a half-sphere can be represented using two families of OPs on the half-disk, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere.
-The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
-Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.
-%
-%\noindent{\bf Acknowledgements}: We would like to thank the anonymous referees for their helpful comments.
-The author SO was supported in part by the Leverhulme Trust Research Project Grant RPG?2019?144 \enquote{Constructive approximation theory on and inside algebraic curves and surfaces}.
-The author BS was supported by the Engineering and Physical Sciences Research Council Mathematics of Planet Earth Centre for Doctoral Training at Imperial College London and the University of Reading, with grant number EP/L016613/1.
-
-\section{P-finite element methods using sparse operators}\label{Appendix:PFEM}
+\section{P-finite element methods using sparse operators}\label{section:ds:PFEM}
 
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.28]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
+	\includegraphics[scale=0.35]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
         	%\label{fig:solutionblocknorms-poisson}
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-poisson}
+	\includegraphics[scale=0.38]{solution-poisson}
 	\centering
@@
 	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.3]{solution-trapezium-helmholtz-k=100-n=1500}
+	\includegraphics[scale=0.38]{solution-trapezium-helmholtz-k=100-n=1500}
 	%\label{fig:solution-poisson}
@@
 	\centering
-	\label{fig:appendixfigs}
+	\label{fig:ds:trappfem}
 \end{figure}
@@
 Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+
+\section{Conclusions}
+
+In this chapter we have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve -- notably here the disk-slice -- with Dirichlet boundary conditions.
+This work extends the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} and whole disk case \cite{} \bstodo{cite} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{}\bstodo{cite}).
+
+We have demonstrated how one can construct the three-or-four-parameter OP families (depending on the domain) that form the basis for our sparse spectral methods, and presented a procedure, utilising the Christoffel-Darboux formula \cite[18.2.2]{DLMF}, of explicitly calculating the recurrence coefficients for the univariate OPs that are part of the construction.
+These coefficients contribute to calculations of the entries in the Jacobi matrices and other differential operators.
+Moreover, we have defined a quadrature rule that can be used for expanding functions in our OP basis and for calculating said entries of important operators.
+We have looked at a few mathematical examples including the Poisson, variable coefficient Helmholtz, and biharmonic equations and shown that our method performs well.
+Importantly, all operator matrices used are shown to be sparse, and in fact banded-block-banded in structure.
+
+Looking ahead, this work serves as a stepping stone to constructing similar methods to solve partial differential equations on sub-domains of the $2$-sphere surface, such as spherical caps that we will discuss in the next chapter.
+
+%, see \cite[Theorem 3.1]{olver2018orthogonal} for a similar construction of OPs on an arc in 2D, and it is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on half-spheres and other suitable sub-components of the sphere.
+The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
+Constructing these sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}.
+Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in turbulent pipe flow.

### 2021-05-26 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Fig sizes changed to be consistent and bigger.

@@
 	\begin{cases}
-    		\DeltaS \tilde{u}(x,y,z) + k^2 \: v(x,y,z) \; \tilde{u}(x,y,z) = f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)  \quad \text{in } \Omega \\
+    		\DeltaS \tilde{u}(x,y,z) + k^2 \; v(x,y,z) \tilde{u}(x,y,z) = g(x,y,z) \quad \text{in } \Omega \\
 		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
@@
 }
-for $\tilde u$. This new problem is then a zero boundary condition Helmholtz problem with right hand side 
+for $\tilde u$, where 
 \bseqn{
@@
 }
-for $(x,y,z) \in \Omega$. Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
-\bseqn{
-	\DeltaS \: c(x,y) = {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) = - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
+for $(x,y,z) \in \Omega$.
+This new problem is then a zero boundary condition Helmholtz problem with right hand side $g$.
+Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
+\bseqn{
+	\DeltaS \: c(x,y) &= {1 \over \rho(z)^2} \pmpxm{\theta}{2} c(x,y) \\
+	&= {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) \\
+	&= - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
 }

### 2021-05-29 papers/thesis/full/ch1-introduction/introduction.tex
Further edits. Included more references.

@@
 
-Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics \cite{onedimopsrefs} \bstodo{citations here}.
+Univariate orthogonal polynomials (hereon also referred to as OPs) have been extensively involved in the development of multiple fields of computational and applied mathematics.
 For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations (see e.g.
-\cite{boyd2001chebyshev, olver2013fast} \bstodo{add more from the references in these papers}).
+\cite{trefethen2000spectral, boyd2001chebyshev, mason2002chebyshev, shen2011spectral, olver2013fast} \bstodo{add more from the references in these papers}).
 While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[Section 18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.

### 2021-05-29 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Further edits. Included more references.

@@
 
-\bstodoinline{Background on multivariate OPs on the disk and triangle.
-Give whole disk OP formula and triangle OP formula as reference points (and refer back to it when defining the disk slice OPs and trapezium OPs respectively).
-Mention there are other choices for OPs on these domains too.}
-
-There has been some really interesting work in the area of mutlivariate orthogonal polynomials in 2D domains \bstodo{list some domains and cite references for them}, in particular spectral methods have been established on the disk \cite{} and the triangle \cite{}.
-For the triangle, the OPs have been chosen as a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values:
-\bseqnnumber{
-	P^{(a,b,c)}(x,y) := define, \quad (x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x} \label{eqn:triangleops}
-}
-while for the disk, so-and-so choose the OPs to be:
-\bseqnnumber{
-	define \label{eqn:diskops}.
-}
-We use the word \enquote{choose} here deliberately, as naturally there is more than one way to define an OP basis on these domains, including the generalisations that we will discuss in this chapter.
-\bstodo{Give one?}
-
-We will see that the full unit disk and triangle cases are special cases of the spectral methods we investigate here.
+In recent years, there has been some interesting work in the area of multivariate orthogonal polynomials in two-dimensional domains (see \cite{dunkl2014orthogonal}), such as for a Chebychev-based spectral method on the rectangle \cite{julien2009efficient} \bstodo{list some domains and cite references for them}.
+In particular, spectral methods have been established on the disk \cite{boyd2011comparing, vasil2016tensor} and the triangle \cite{olver2019triangle}.
+Jacobi polynomials are extremely useful for solving a number of problems in not just 1D but 2D also, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.
+In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, for the triangle a choice of OPs are the Jacobi polynomials on the triangle, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
+\bseqnnumber{
+	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:triangleops}
+}
+for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values.
+
+For the disk, a review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials.
+It is concluded that using Chebyshev polynomials yields mixed results, but that there is also no single advised basis to use.
+While Zernike polynomials yield accurate results, there is a lack of radial fast transform, which is a disadvantage.
+However, ideas for how to overcome this have progressed \cite{slevinsky2018use, sakai2009application, wilber2017computing}.
+One solution proposed is a radial basis defined in terms of Jacobi polynomials \cite{vasil2016tensor}:
+\bseqnnumber{
+	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:diskops}.
+}
+where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to $\int_0^1 Q_n^{k,m}(r) \; Q_{n'}^{k,m}(r) \; (1-r^2)^k \; r \; \D r = \delta_{n,n'}$.
+$\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials.
+Again, the idea here is that differential and other operators, as well as basis (parameter) conversion operators, can be represented as banded matrices.
+
+In this chapter, we aim to present a basis and develop a sparse spectral method on generalisations of these domains, namely the disk-slice and trapezium.
+The full unit disk and triangle cases are special cases of the spectral methods we investigate here too.
+We will show that the operators for multiplication by $x$ and $y$, as well as differential and parameter conversion, will be banded-block-banded.
 
@@
 For simplicity of presentation we focus on the disk-slice in this chapter, where $\rho(x) = \sqrt{1-x^2}$, $(\alpha,\beta) \subset (0, 1)$, and $(\gamma, \delta)  = (-1,1)$.
-However, we will discuss an extension to other geometries, including the half-disk and trapeziums, by choosing the function $\rho$ and constants $\alpha, \beta, \gamma, \delta$ appropriately.
+However, we will discuss an extension to other geometries, including the half-disk and trapeziums, by choosing the function $\rho$ and constants $\alpha, \beta, \gamma, \delta$ appropriately (for example, choosing $\rho(x) = 1-\xi x$ and $(\alpha,\beta) = (\gamma, \delta) = (0, 1)$ for some constant $\xi > 0$ returns the trapezium case).
 Moreover, one should be able to see that the full disk and triangle can also be obtained in such a way.
 
-We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{olver2013fast} and its analogue on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
+We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
 On the disk-slice the family of weights we consider are of the form
@@
 where $C_{n,k} \neq 0$ and \enquote{lower order terms'} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
-The precise normalization arises from their definition in terms of one-dimensional OPs in \bsrefdef{def:OPconstruction}.
+The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:OPconstruction}.
 
 Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$.
-Whereas the sparsity pattern and entries derived in \cite{olver2018recurrence,olver2019triangle} for equations on the triangle and \cite{vasil2016tensor} for equations on the disk results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel--Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
+Whereas the sparsity pattern and entries derived for equations on the triangle \cite{olver2018recurrence, olver2019triangle} and for equations on the disk \cite{vasil2016tensor} with the frameworks discussed earlier both result from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 $\forall$, $n = 0,1,2,\dots, $ let $\{p_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k+1} w_1(x)$ where $0 \le k \le n$, and $\{q_{n}\}$ be polynomials orthogonal with respect to the weight $w_2(x)$.
-Then the 2D polynomials defined on $\Omega$
-\bseqn{
-	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}) \qquad\hbox{for} \qquad 0 \le k \le n, \: n = 0,1,2,\dots
-}
-are orthogonal polynomials with respect to the weight \(W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)}) \) on $\Omega$. 
+Then, for $n = 0,1,2,\dots$, $k = 0,\dots,n$, the 2D polynomials defined on $\Omega$ given by
+\bseqn{
+	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)})
+}
+are orthogonal polynomials with respect to the weight $W(x,y) := w_1(x) \: w_2\fpr(\frac{y}{\rho(x)})$ on $\Omega$. 
 \end{proposition}
@@
 \begin{definition}\label{def:constuction}
-	Define the four-parameter 2D orthogonal polynomials via:
+	For parameters $a,b,c,d \in \R$, define the four-parameter 2D orthogonal polynomials as the set $\{\hdopnk^{(a,b,c,d)}\}$ where:
 \bseqn{
@@
 }
-assuming that either \bsrefcondition{cond:disk} or \bsrefcondition{cond:disk} with  \(\jacw^{(a,b)}\) being an even function (i.e.
-$a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) =  (\delta-x^2)^a$) hold.
+assuming that either \bsrefcondition{cond:trap} holds, or \bsrefcondition{cond:disk} holds with $\jacw^{(a,b)}$ being an even function (i.e.
+$a = b$, and we can hence denote the weight as $\jacw^{(a)}(x) = w^{(a,a)}_P(x) = (\delta-x^2)^a$).
 That is,
 \bseqn{
-	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
+	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}
 }
@@
 We can use (\ref{eqn:Hrecurrence}) and (\ref{eqn:Precurrence}) to determine the 2D recurrences for $\hdopnkabcd(x,y)$.
-Importantly, we can deduce sparsity in the recurrence relationships:
+Importantly, we can deduce sparsity in the recurrence relationships.
 
@@
 \bseqn{
-x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n+1,k,1} \: \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
-y \hdopnkabcd(x,y) &= \betaabcd_{n,k,1} \: \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \: \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,3} \: \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
-		& \quad \quad + \betaabcd_{n,k,4} \: \hdop_{n, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,5} \: \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \: \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \nonumber \\
-		& \quad \quad + \betaabcd_{n,k,7} \: \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \: \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,9} \: \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
+	&x \hdopnkabcd(x,y) \\
+	&\quadtwo = \alphaabcd_{n,k,1} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n+1,k,1} \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
+	&y \hdopnkabcd(x,y) \\
+	&\quadtwo = \betaabcd_{n,k,1} \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,3} \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) \\
+		&\quadfour + \betaabcd_{n,k,4} \hdop_{n, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,5} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \\
+		&\quadfour + \betaabcd_{n,k,7} \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,9} \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
 }

### 2021-05-29 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Further edits. Included more references.

@@
 \begin{definition}\label{def:sc:Ydefinition}
-	Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
+	Define the unit circle $\omega := \bsset{\vec{x} = (x,y) \in \R^2}{x^2 + y^2 = 1}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
 Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
@@
 	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
-	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots \\
-	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots
-}
-where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
+	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta) \\
+	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta)
+}
+for $k = 1,2,3,\dots$ where $\ch_{0} := {\sqrt{2} \over 2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$.
+The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
 \bseqn{
@@
 
-However, for an arc of the unit circle $\{ define arc \}$ (note we can orientate the arc as we please without loss of generality), we would need to make a modification so that the polynomials are orthogonal with respect to an inner product integral with the truncated domain of the arc:
-\bseqn{
-	give arc ops (T^h, yU^h) and the inner product.
-}
+However, for an arc of the unit circle $\bsset{\vec{x} = (x,y) \in \R^2}{x \ge h, \: x^2 + y^2 = 1}$ for some $h \in (-1, 1)$ (note we can orientate the arc as we please without loss of generality), we would need to make a modification so that the polynomials are orthonormal with respect to a different inner product accounting for the truncated domain of the arc:
+\bseqn{
+	\ip< p, \: q >_{\ch^h} &:= c^h \: \int_{-\arccos{h}}^{\arccos{h}} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta \\
+	&= c^h \: \int_h^1 \Big\{ p(x,\sqrt{1-x^2}) \; q(x,\sqrt{1-x^2}) \\
+	&\quadfour \quadtwo + p(x,-\sqrt{1-x^2}) \; q(x,-\sqrt{1-x^2}) \Big\} {\D x \over \sqrt{1-x^2}}
+}
+for some normalising constant $c^h$.
+In the same vein, we can choose to construct the arc OPs in the same way as the whole circle OPs -- that is, define $\{T_k^h\}$, $\{U_k^h\}$ as two sets of univariate orthogonal polynomials on the interval $[h,1]$, orthogonal with respect to the weight functions $(1-x^2)^{-\half}$ and $(1-x^2)^{\half}$ respectively, and then define
+\bseqn{
+	\ch^h_{0,0}(\xvec) \equiv \ch^h_{0,0}(x, y) &:= \ch^h_{0} =: \ch^h_{0,0}(\theta) \\
+	\ch^h_{k,0}(\xvec) \equiv \ch^h_{k,0}(x, y) &:= T^h_k(x) =: \ch^h_{k,0}(\theta) \\
+	\ch^h_{k,1}(\xvec) \equiv \ch^h_{k,1}(x, y) &:= y \: U^h_{k-1}(x) =: \ch^h_{k,1}(\theta)
+}
+where $Y_0^h$ is chosen so that $\ip< Y_0^h, \: Y_0^h >_{\ch^h} = 1$.
+
 \bstodoinline{Expand on this. Read the Olver Xu "OPs on quadratic curves paper." where the arc might apply??? Either way, cite this paper in intro (CH1).}
@@
 
-This idea of using 1D OPs on a shortened interval to create the modified OPs for an arc has be generalised to higher dimensional domains, where a way of constructing orthogonal polynomials on algebraic curves has been presented \cite{olver2020orthogonal}.
+This idea of using 1D OPs on a shortened interval to create the modified OPs for a circle arc has be generalised to higher dimensional domains, where a way of constructing orthogonal polynomials on algebraic curves has been presented \cite{olver2020orthogonal}.
 \bstodoinline{Expand on this.
@@
 
-Using the approach presented for constructing OPs on algebraic curves, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Fourier series written as orthogonal polynomials in $x$ and $y$:
+Using the approach presented for constructing OPs on algebraic curves, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval $[\alpha,\beta]$, and from Fourier series written as orthogonal polynomials in $x$ and $y$.
+Recall that the spherical cap is defined by
+\bseqn{
+	\Omega := \bsset{(x,y,z) \in \R^3}{\alpha < z < \beta, \: x^2 + y^2 + z^2 = 1}.
+}
 \begin{proposition}[\cite{olver2020orthogonal}]\label{prop:sc:construction}
@@
 \bseqn{
-	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
-	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: \times \\
-	&\qquad w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
+	&\ip< p, \: q > := \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
+	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} \Big\{ p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \\
+	&\quadeight \quad \cdot w(\cosphi) \: \sinphi \Big\} \: \D \theta \: \D \varphi \\
 	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: q\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: w(z) \: \D \theta \: \D z

### 2021-05-31 papers/thesis/full/ch1-introduction/introduction.tex
Further edits, mainly including additional references.

@@
 However, few books have been published over the years on the topic of multivariate orthogonal polynomials \cite{dunkl2014orthogonal}.
-
 This branch of mathematics has an encouraging future though, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains.
 Spectral methods have been established on the triangle \cite{olver2019triangle} using OPs inspired by the Koornwinder approach for defining them, and on the disk \cite{vasil2016tensor} using Zernike polynomials.
-Applications include solving Volterra integral equations \cite{gutleb2020sparse}.
 
@@
 
-Spherical Harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used in many fields including computer graphics (e.g.
+Spherical harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
 \cite{moon2008efficient, sloan2013efficient}), astrophysics (e.g.
 \cite{vasil2019tensor}), quantum theory (e.g.
-\cite{varshalovich1988quantum}), geosciences (e.g.
+\cite{varshalovich1988quantum}), biochemistry (e.g.\cite{parimal2014application, basko1998application}), geosciences (e.g.
 \cite{hollerbach2013parity}) and meteorology (e.g.
-\cite{evans1998spherical, rubinstein2015scalar}) \bstodo{cite applications of SHs}.
+\cite{evans1998spherical, rubinstein2015scalar, wedi2013fast, ecmwf2020scalability, diamantakisecmwf, courtier1998ecmwf}).
 Notably, they are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
@@
 On the other hand, we aim to propose a strictly orthogonal \textit{polynomial} basis, and derive sparse operators for multiplication by the cartesian coordinate axes, which can in turn lead to operators for multiplication by trigonometric functions too.
-By utilising orthogonal polynomials, we can develop a sparse spectral method on subdomains of the sphere surface (e.g.
+By utilising orthogonal polynomials in cartesian coordinates, we can develop a sparse spectral method on subdomains of the sphere surface (e.g.
 the spherical cap) using our knowledge of other geometries and methods involving OPs developed for them.

### 2021-05-31 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Further edits, mainly including additional references.

@@
 Jacobi polynomials are extremely useful for solving a number of problems in not just 1D but 2D also, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.
+Applications for the disk include \bstodo{give applications and references} and for the triangle include solving Volterra integral equations \cite{gutleb2020sparse}, \bstodo{give more}.
+
 In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, for the triangle a choice of OPs are the Jacobi polynomials on the triangle, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
@@
 In this chapter we have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve -- notably here the disk-slice -- with Dirichlet boundary conditions.
-This work extends the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} and whole disk case \cite{} \bstodo{cite} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{}\bstodo{cite}).
+This work extends the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} and whole disk case \cite{vasil2016tensor} \bstodo{cite more here} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, vasil2016tensor}).
 

### 2021-06-01 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Fig sizes changed to be consistent and bigger.
Spacing and margins of SH chapter eqns fixed.

@@
 such that $\Dlt A_l = \idmat{2l+3}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[Section 3.2]{dunkl2014orthogonal}, which can be expanded to obtain the three-term recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[Section 3.2]{dunkl2014orthogonal}, which can be expanded to obtain the three-term recurrence, for $l = 0,1,2,\dots$:
 \bseqn{
@@
 		\SHvec_{0}(x,y,z) := Y_0 \\
-		\SHvec_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z), \quad l = 0,1,2,\dots.
+		\SHvec_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z)
 	\end{cases}
@@
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
+	&\quadeight \quadtwo \cdot \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
 	&\quadeight + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: \rho(z)^{\abs{m}+\abs{m'}-1} \: \D z \Big) \nonumber \\
@@
 	&= \int_{-1}^{1} \Big\{ m(m-1) \jaclm{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m-2} (1 + z^2) - m z \jaclm{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m} \nonumber \\
-	&\quadfour + (m-1) \jaclm{l}{m} \rho^{2m-2} \big( - 2m z^2 \jaclmbracket{l'}{m-1} + \jaclmbracket{l'}{m-1} \rho^{2} - z \jaclmbracketdz{l'}{m-1} \rho^{2} \big) \nonumber \\
-	&\quadfour + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&\quadtwo + (m-1) \jaclm{l}{m} \rho^{2m-2} \big(\jaclmbracket{l'}{m-1} \rho^{2} - z \jaclmbracketdz{l'}{m-1} \rho^{2} - 2m z^2 \jaclmbracket{l'}{m-1} \big) \nonumber \\
+	&\quadtwo + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m+2} \Big\} \: \D z \nonumber \\
 	&= \int_{-1}^{1} \Big\{ (m-1)(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m-1} \rho^{2m} - z \jaclm{l}{m} \jaclmbracketdz{l'}{m-1} \rho^{2m} \nonumber \\
@@
 	&\quadfour + d_{l,m} d_{l',m-1} \jaclm{l'}{m} \big( \tilde \alpha_{l, m+1, 1} \jaclm{l-1}{m} + \tilde \alpha_{l, m+1, 3} \jaclm{l+1}{m} \big) \rho^{2m} \Big\} \: \D z \nonumber \\
-	&= \delta_{l',l-1} \Big( \big[ (m^2 - 1) \tilde \alpha_{l-1, m-1, 4} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \big] {1\over2\pi(\clm)^2} + d_{l,m} d_{l-1,m-1} {\tilde \alpha_{l-1, m-1, 1} \over 2\pi (c_{l-1}^m)^2} \Big) \nonumber \\
-	&\quadtwo + \delta_{l',l+1} \Big( \big[ (m^2 - 1) \tilde \alpha_{l+1, m-1, 2} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \big] {1\over2\pi(\clm)^2} + d_{l,m} d_{l+1,m-1} {\tilde \alpha_{l-1, m-1, 3} \over 2\pi (c_{l+1}^m)^2} \Big).
-\label{eqn:lemmaVSHrecurrencesproof2}
+	&= {\delta_{l',l-1} \over 2\pi} \Big\{ \big[ (m^2 - 1) \tilde \alpha_{l-1, m-1, 4} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \big] {1 \over (\clm)^2} + d_{l,m} d_{l-1,m-1} {\tilde \alpha_{l-1, m-1, 1} \over (c_{l-1}^m)^2} \Big\} \nonumber \\
+	&\quadtwo + {\delta_{l',l+1} \over 2\pi} \Big\{ \big[ (m^2 - 1) \tilde \alpha_{l+1, m-1, 2} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \big] {1 \over (\clm)^2} \nonumber \\
+	&\quadeight + d_{l,m} d_{l+1,m-1} {\tilde \alpha_{l-1, m-1, 3} \over (c_{l+1}^m)^2} \Big\}. \label{eqn:lemmaVSHrecurrencesproof2}
 }
@@
 	&\quadtwo \quad - (m+1) z \jaclmdz{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} + \jaclmdz{l}{m} \jaclmbracketdz{l'}{m+1} \rho^{2m+4} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ m(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m} (1 + z^2) - (m+1) z \jaclmdz{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} \nonumber \\
+	&= \int_{-1}^{1} \Big\{ m(m+1) \jaclm{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m} (1 + z^2) \nonumber \\
+	&\quadfour - (m+1) z \jaclmdz{l}{m} \jaclmbracket{l'}{m+1} \rho^{2m+2} \nonumber \\
 	&\quadfour + m \jaclmbracket{l'}{m+1} \rho^{2m} \big( - 2(m+1) z^2 \jaclm{l}{m} + \jaclm{l}{m} \rho^{2} - z \jaclmdz{l}{m} \rho^{2} \big) \nonumber \\
@@
 	&= \int_{-1}^{1} \Big\{ m(m+2) \jaclmbracket{l'}{m+1} \big( \tilde \alpha_{l, m, 2} \jaclmbracket{l-1}{m+1} + \tilde \alpha_{l, m, 4} \jaclmbracket{l+1}{m+1} \big) \rho^{2m+2} \nonumber \\
-	&\quadfour - d_{l,m} \jaclmbracket{l'}{m+1} \big( \tilde \gamma_{l,m+1,1} \jaclmbracket{l-1}{m+1} + \tilde \gamma_{l,m+1,2} \jaclmbracket{l+1}{m+1} \big) \rho^{2m+2} \nonumber \\
-	&\quadfour + d_{l,m} d_{l',m+1} \jaclmbracket{l}{m+1} \big( \tilde \alpha_{l', m+2, 1} \jaclmbracket{l'-1}{m+1} + \tilde \alpha_{l', m+2, 3} \jaclmbracket{l'+1}{m+1}  \big) \rho^{2m+2} \Big\} \: \D z \nonumber \\
-	&= \delta_{l',l-1} \Big( \big[ m(m+2) \tilde \alpha_{l, m, 2} - d_{l,m} \tilde \gamma_{l,m+1,1} \big] {1 \over 2 \pi (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\tilde \alpha_{l-1, m+2, 3} \over 2\pi (c_{l}^{m+1})^2} \Big) \nonumber \\
-	&\quadtwo + \delta_{l',l+1} \Big( \big[ m(m+2) \tilde \alpha_{l, m, 4} - d_{l,m} \tilde \gamma_{l,m+1,2} \big] {1 \over 2 \pi (c_{l-1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\tilde \alpha_{l+1, m+2, 1} \over 2\pi (c_{l}^{m+1})^2} \Big).
+	&\quadtwo - d_{l,m} \jaclmbracket{l'}{m+1} \big( \tilde \gamma_{l,m+1,1} \jaclmbracket{l-1}{m+1} + \tilde \gamma_{l,m+1,2} \jaclmbracket{l+1}{m+1} \big) \rho^{2m+2} \nonumber \\
+	&\quadtwo + d_{l,m} d_{l',m+1} \jaclmbracket{l}{m+1} \big( \tilde \alpha_{l', m+2, 1} \jaclmbracket{l'-1}{m+1} + \tilde \alpha_{l', m+2, 3} \jaclmbracket{l'+1}{m+1}  \big) \rho^{2m+2} \Big\} \: \D z \nonumber \\
+	&= {\delta_{l',l-1} \over 2\pi} \Big\{ \big[ m(m+2) \tilde \alpha_{l, m, 2} - d_{l,m} \tilde \gamma_{l,m+1,1} \big] {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\tilde \alpha_{l-1, m+2, 3} \over (c_{l}^{m+1})^2} \Big\} \nonumber \\
+	&\quadtwo + {\delta_{l',l+1} \over 2\pi} \Big\{ \big[ m(m+2) \tilde \alpha_{l, m, 4} - d_{l,m} \tilde \gamma_{l,m+1,2} \big] {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\tilde \alpha_{l+1, m+2, 1} \over (c_{l}^{m+1})^2} \Big\}.
 \label{eqn:lemmaVSHrecurrencesproof3}
@@
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ i m' \jaclmabs{l'}{m'}(z) \rho(z)^{\abs{m} + \abs{m'} - 1} \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
-	&\quadeight + i m \jaclmabs{l}{m}(z) \rho(z)^{\abs{m} + \abs{m'} - 1} \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
+	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho(z)^{\abs{m} + \abs{m'} - 1} \Big\{ i m' \jaclmabs{l'}{m'}(z) \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
+	&\quadfour + i m \jaclmabs{l}{m}(z) \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
 	&= \clm \clmprime \pi (\delta_{m',m-1} + \delta_{m',m+1}) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \sgn{m} 2 i \, m \, m' \, z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} - 1} \nonumber \\
+	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \signfunc{m} 2 i \, m \, m' \, z \jaclmabs{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} - 1} \nonumber \\
 	&\quadeight - i m \jaclmabsdz{l}{m} \jaclmabs{l'}{m'} \rho^{\abs{m} + \abs{m'} + 1} \nonumber \\
@@
 }
-since when the above is nonzero, $m, m'$ will have the same sign (or one is zero) and so $\abs{m}m' + \abs{m'}m = \sgn{m}mm'$.
+since when the above is nonzero, $m, m'$ will have the same sign (or one is zero) and so $\abs{m}m' + \abs{m'}m = \signfunc{m}m\:m'$.
 Once again, we can conclude that $\calA^\perp_{l',m'}$ are zero for $m' \notin \{m-1, m+1\}$.
@@
 \bseqnnumber{
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
 	&\quadfour - i (m+1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m+1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} - 1} \rho^{2 \abs{m} - 2} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} - 1} \rho^{2\abs{m}} \nonumber \\
 	&\quadfour + i (m+1) \jaclmabs{l}{m} \rho^{2\abs{m} - 2} \big[ \jaclmbracketdz{l'}{\abs{m} - 1} \rho^2 - 2\abs{m} z \jaclmbracket{l'}{\abs{m} - 1} \big] \Big\} \: \D z \nonumber \\
@@
 \bseqnnumber{
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} \nonumber \\
+	&\quadfour - i m \jaclmabs{l}{m} \jaclmbracketdz{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
 	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m} + 2} \Big\} \: \D z \nonumber \\
-	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} - i (m-1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
+	&= \int_{-1}^{1} \Big\{ - 2 i \, m(m-1) \, z \jaclmabs{l}{m} \jaclmbracket{l'}{\abs{m} + 1} \rho^{2 \abs{m}} \nonumber \\
+	&\quadfour - i (m-1) \jaclmabsdz{l}{m} \jaclmbracket{l'}{\abs{m} + 1}  \rho^{2\abs{m} + 2} \nonumber \\
 	&\quadfour + i m \jaclmbracket{l'}{\abs{m} + 1} \rho^{2\abs{m}} \big[\jaclmabsdz{l}{m} \rho^2 - 2 (\abs{m} + 1) z \jaclmabs{l}{m} \big] \Big\} \: \D z \nonumber \\
@@
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
-	&\quad \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
+	&\quadtwo \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
+	&\quadeight \cdot \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
 	&\quadfour \quadtwo + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: z \; \rho(z)^{\abs{m}+\abs{m'}-2} \: \D z \Big) \label{eqn:lemmaVSHrecurrencesproofz1}
@@
 	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
-	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ i m' z \jaclmabs{l'}{m'}(z) \rho(z)^{\abs{m} + \abs{m'} - 2} \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
-	&\quadeight + i m z \jaclmabs{l}{m}(z) \rho(z)^{\abs{m} + \abs{m'} - 2} \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
+	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho(z)^{\abs{m} + \abs{m'} - 2} \Big\{ i m' z \jaclmabs{l'}{m'}(z) \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
+	&\quadfour + i m z \jaclmabs{l}{m}(z) \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
 	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ 2i m\abs{m} z^2 \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m} - 2} \nonumber \\
@@
 	\centering % <-- added
-	\includegraphics[scale=0.5]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100}
+	\centerline{\includegraphics[scale=0.6]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100}}
 	\caption{Four snapshots of the solution $u$ of the heat equation with $\alpha = 1/42$ and a step size of $\Delta t = 0.01$, solved using the BDF2 timestepping method.
@@
 	\centering % <-- added
-	\includegraphics[scale=0.5]{linear-swe-H=1-dt=0p01-nsteps=100}
+	\centerline{\includegraphics[scale=0.6]{linear-swe-H=1-dt=0p01-nsteps=100}}
 	\caption{Four snapshots of the solution for the height $h$ of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method.

### 2021-06-01 OrthogonalPolynomialFamilies/src/DiskSliceFamilies.jl
Edits to SC and DS code

@@ -757,7 +757,7 @@ function jacobix(S::DiskSliceSpace, N)
     rows = cols = 1:N+1
     l, u = 1, 1
     λ, μ = 0, 0
-    J = BandedBlockBandedMatrix(0.0I, (rows, cols), (l, u), (λ, μ))
+    J = BandedBlockBandedMatrix(0.0I, rows, cols, (l, u), (λ, μ))
     J[1, 1] = S.B[1][1, 1]
     view(J, Block(1, 2)) .= S.A[1][1, :]'
     for n = 2:N
@@ -775,7 +775,7 @@ function jacobiy(S::DiskSliceSpace, N)
     rows = cols = 1:N+1
     l, u = 1, 1
     λ, μ = 1, 1
-    J = BandedBlockBandedMatrix(0.0I, (rows, cols), (l, u), (λ, μ))
+    J = BandedBlockBandedMatrix(0.0I, rows, cols, (l, u), (λ, μ))
     n = 1
     J[1, 1] = S.B[1][2, 1]
     view(J, Block(n, n+1)) .= S.C[n+1][Int(end/2)+1:end, :]'
@@ -1144,10 +1144,11 @@ function partialoperatorx(S::DiskSliceSpace{<:Any, B, T, <:Any}, N;
     band = S.family.nparams
     if transposed
         A = BandedBlockBandedMatrix(
-            Zeros{B}(sum(1:(N+1)), sum(1:N)), (1:N+1, 1:N), (band, -1), (2, 0))
+            Zeros{B}(sum(1:(N+1)), sum(1:N)), 1:N+1, 1:N, (band, -1), (2, 0))
     else
+        @show N, band
         A = BandedBlockBandedMatrix(
-            Zeros{B}(sum(1:N), sum(1:(N+1))), (1:N, 1:N+1), (-1, band), (0, 2))
+            Zeros{B}(sum(1:N), sum(1:(N+1))), 1:N, 1:N+1, (-1, band), (0, 2))
     end
 
     n, k = 1, 0
@@ -1201,10 +1202,10 @@ function partialoperatory(S::DiskSliceSpace{<:Any, B, T, <:Any}, N;
     # Takes the space H^{a,b,c} -> H^{a,b,c+1}
     if transposed
         A = BandedBlockBandedMatrix(
-            Zeros{B}(sum(1:(N+1)),sum(1:N)), (1:N+1, 1:N), (1,-1), (1,-1))
[…]

### 2021-06-01 OrthogonalPolynomialFamilies/src/SphericalCaps.jl

@@ -2121,7 +2121,8 @@ end
 # Uselful for SWE
 function convertcoeffssize(S::SphericalCapSpace{<:Any, B, T, <:Any},
                             cfs::AbstractArray{B};
-                            totangentspace::Bool=true) where {B,T}
+                            totangentspace::Bool=true,
+                            asfun::Bool=false) where {B,T}
     if totangentspace
         M = length(cfs)
         ret = zeros(B, 2M)
@@ -2131,6 +2132,7 @@ function convertcoeffssize(S::SphericalCapSpace{<:Any, B, T, <:Any},
             ret[it+1] = cfs[j]
             it += 2
         end
+        ret
     else
         M = Int(length(cfs) / 2)
         ret = zeros(B, M)
@@ -2139,8 +2141,12 @@ function convertcoeffssize(S::SphericalCapSpace{<:Any, B, T, <:Any},
             ret[j] = cfs[it]
             it += 2
         end
+        if asfun
+            Fun(S, ret)
+        else
+            ret
+        end
     end
-    ret
 end
 
 
@@ -2317,12 +2323,16 @@ end
 #===#
 # Resizing coeffs vectors
 
-function resizecoeffs!(S::SphericalCapSpace, f::Fun, N::Int)
+function resizecoeffs!(S::SphericalCapSpace, f::Fun, N::Int; force::Bool=false)
     ncfs = length(f.coefficients)
     m = (N+1)^2
     cfs = convertcoeffsvecorder(S, f.coefficients) # creates new cfs vec
     if ncfs > m && any(i -> abs(i) > 1e-16, cfs[m+1:end])
-        error("Trying to decrease degree of f")
+        if force
+            @show "Trying to decrease degree of f but force=true passed so continuing"
+        else
+            error("Trying to decrease degree of f")
+        end
     end
     cfs = convertcoeffsvecorder(S, f.coefficients) # creates new cfs vec
     resize!(cfs, m)

### 2021-06-01 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Edits to fig layouts mainly

@@
-\chapter{Disk slices and trapeziums}
+\chapter{Disk slices and trapeziums}\label{CHAPTER:diskslice}
 
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}
-		\centering
-		\caption{The Laplacian operator $\laplacewiii$}
+		\centerline{\includegraphics[scale=0.5]{sparsityoflaplacian-w11-diskslice-alpha=0p2-beta=0p8}}
+		\centering
+		\caption{}
 	\end{subfigure}\hfil % <-- added
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}
-		\centering
-		\caption{A variable coefficient Helmholtz operator}
+		\centerline{\includegraphics[scale=0.5]{sparsityofhelmholtz-diskslice-alpha=0p2-beta=0p8}}
+		\centering
+		\caption{}
 	\end{subfigure}\hfil % <-- added
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}
-		\centering
-		\caption{The biharmonic operator $\biharmonictwo$}
+		\centerline{\includegraphics[scale=0.5]{sparsityofbiharmonic-diskslice-alpha=0p2-beta=0p8}}
+		\centering
+		\caption{}
 	\end{subfigure}\hfil % <-- added
     	\caption{"Spy" plots of (differential) operator matrices, showing their sparsity.
-For (b), the weighted variable coefficient Helmholtz operator is $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.}
+(a) The Laplacian operator $\laplacewiii$.
+(b) The variable coefficient Helmholtz operator $\laplacewiii + k^2 \: T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}$ for $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$.
+(c) The biharmonic operator $\biharmonictwo$}
 	\label{fig:sparsity}
@@
 In this chapter we have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve -- notably here the disk-slice -- with Dirichlet boundary conditions.
-This work extends the triangle case \cite{beuchler2006new,li2010optimal,olver2019triangle} and whole disk case \cite{vasil2016tensor} \bstodo{cite more here} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, vasil2016tensor}).
+This work extends the triangle case \cite{beuchler2006new, li2010optimal, olver2019triangle} and whole disk case \cite{boyd2011comparing, vasil2016tensor} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, vasil2016tensor}).
 

### 2021-06-01 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Fixed appendix numbering
Edits to fig layouts mainly

@@
-\chapter{Spherical caps}
+\chapter{Spherical caps}\label{CHAPTER:sphericalcaps}
 
@@
 Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
-Just as we proceeded in \bsrefchapter{chapter:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.
-
-The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined in \bsrefchapter{chapter:diskslice} (see also \cite{snowball2019sparse}).
+Just as we proceeded in \bsrefchapter{CHAPTER:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.
+
+The three-dimensional orthogonal polynomials defined here involve the same non-classical (in fact, semi-classical \cite[\S5]{magnus1995painleve}) 1D OPs as those outlined for the disk-slice, and so methods for calculating these 1D OP recurrence coefficients and integrals have already been outlined in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}).
 In particular, by exploiting the connection with these 1D OPs we can construct discretizations of general partial differential operators of size $(p+1)^2 \times (p+1)^2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 }
-A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{chapter:diskslice} (see also \cite{snowball2019sparse}).
+A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}).
 The weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}.
@@
 }
-where the coefficients are calculable (see \bsrefchapter{chapter:diskslice}).
+where the coefficients are calculable (see \bsrefchapter{CHAPTER:diskslice}).
 We can use \bsrefeqn{eqn:sc:Rrecurrence} to determine the 3D recurrences for $\scopnkia(x,y,z)$.
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsity-of-laplacian}
+		\centerline{\includegraphics[scale=0.5]{sparsity-of-laplacian}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsity-of-rho2laplacian}
+		\centerline{\includegraphics[scale=0.5]{sparsity-of-rho2laplacian}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsity-of-helmholtz}
+		\centerline{\includegraphics[scale=0.5]{sparsity-of-helmholtz}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{sparsity-of-biharmonic}
+		\centerline{\includegraphics[scale=0.5]{sparsity-of-biharmonic}}
 		\centering
@@
 	\begin{subfigure}[b]{0.55\textwidth}
-		\includegraphics[scale=0.4]{condition-numbers-laplacian-log-N=200}
+		\centerline{\includegraphics[scale=0.4]{condition-numbers-laplacian-log-N=200}}
 	\end{subfigure}
@@
 	\begin{subfigure}[b]{0.55\textwidth}
-		\includegraphics[scale=0.4]{condition-numbers-max-laplacian-log}
+		\centerline{\includegraphics[scale=0.4]{condition-numbers-max-laplacian-log}}
 	\end{subfigure}
@@
 	&\quad \quad = 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \int_{-1}^1 f_e\big(\rho(z)x,\rho(z)\sqrt{1-x^2}, z\big) \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
-	&\quad \quad \approx \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
-	&\quad \quad \approx \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) s_l, \rho(t_j) \sqrt{1 - s_l^2}, t_j\big) \quad (\star \star) \\
-	&\quad \quad = \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
+	&\quad \quad \approx 2 \int_\alpha^1 \genjacw^{(a,0)}(z) \: \Big( \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(z) s_l, \rho(z) \sqrt{1 - s_l^2}, z\big) \Big) \: \D z \quad (\star) \\
+	&\quad \quad \approx 2 \sum_{j=1}^{M_1} w_j^{(t)} \sum_{l=1}^{M_2} w_l^{(s)} f_e\big(\rho(t_j) s_l, \rho(t_j) \sqrt{1 - s_l^2}, t_j\big) \quad (\star \star) \\
+	&\quad \quad = 2 \sum_{k=1}^{M_1 M_2}  w_j \: f_e(x_j, y_j, z_j).
 }
@@
 for N sufficiently large, where $\bigscopNka, \bigscopNa$ is defined in \bsrefeqns{eqn:sc:OPdefNka}{eqn:sc:OPdefNa} and where
-
 \bseqn{
@@
 \bseqn{
-	f_{n,k,i} &= \frac{1}{2 \: \normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
-	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) +f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
+	f_{n,k,i} &= \frac{1}{\normgenjac^{(a,2k)} \: \pi} \sum_{j=1}^{M} w_j \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) \\
+	&\quadeight \quad + f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big] \\
+	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M_1} w^{(t)}_j \sum_{l=1}^{M_2} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) \\
+	&\quadeight \quadfour + f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
 }
@@
 	\centering % <-- added
-	\includegraphics[scale=0.4]{complexity-Nend=1000}
+	\centerline{\includegraphics[scale=0.4]{complexity-Nend=1000}}
 	\caption{Time in seconds to build and solve the system $\big[\DeltaS + v(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$, for a rotationally invariant $v(x,y,z) = v(z)$.

### 2021-06-03 papers/thesis/full/ch1-introduction/introduction.tex
Further edits. New spy plots to make uniform across thesis. New references and fixing of capitals etc.

@@
 For example, univariate OPs have been used to derive spectral methods to numerically solve one-dimensional differential equations (see e.g.
-\cite{trefethen2000spectral, boyd2001chebyshev, mason2002chebyshev, shen2011spectral, olver2013fast} \bstodo{add more from the references in these papers}).
-While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[Section 18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
+\cite{trefethen2000spectral, canuto2007spectral, gottlieb1977numerical, boyd2001chebyshev, mason2002chebyshev, shen2011spectral, olver2013fast}).
+While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[\S18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
 
@@
 This branch of mathematics has an encouraging future though, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains.
-Spectral methods have been established on the triangle \cite{olver2019triangle} using OPs inspired by the Koornwinder approach for defining them, and on the disk \cite{vasil2016tensor} using Zernike polynomials.
+Spectral methods have been developed for solving PDEs as an alternative to finite difference and finite element methods.
+For example, in recent years spectral methods have been established on the triangle \cite{olver2019triangle} using OPs inspired by the Koornwinder approach for defining them, and on the disk \cite{vasil2016tensor} using Zernike polynomials.
 
 There are various interpretations of what a \enquote{spectral method} is.
-It could be described as one that achieves spectral convergence of its solutions, or one that uses eigenfunctions as basis functions, or one that uses OPs as basis polynomials \cite{}\bstodo{cite these as interpretations}.
-It is the latter that we refer to for our purposes as a spectral method.
+It could be described as one that achieves spectral convergence of its solutions \cite{gottlieb1977numerical}, or one that uses Laplacian eigenfunctions as basis functions \cite{zhong2007numerical}, or one that uses OPs as basis functions for the approximation of a solution \cite{olver2019triangle}.
+It is the latter that we refer to for our purposes as a spectral method in the body of the thesis.
 By utilising OPs as basis functions, we can develop sparse spectral methods (SSMs), meaning that the naturally sparse relationships between the basis OPs lead to sparse operator matrices that represent the differential operations in the equation to solve.
@@
 We take inspiration from the work established for the triangle \cite{olver2019triangle} and the unit disk \cite{vasil2016tensor}, which can  be seen as special cases of the disk-slice and trapezium in the framework we present here.
-The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the Spherical Harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
+The spherical cap work serves to lay a foundation for using spherical caps and spherical bands as elements in a spectral element method for solving PDEs on the whole sphere, as an alternative to the spherical harmonic transform approach that the European Centre for Medium-range Weather Forecasts (ECMWF) use in their weather and climate model \cite{cheong2006dynamical}.
 
@@
 \cite{varshalovich1988quantum}), biochemistry (e.g.\cite{parimal2014application, basko1998application}), geosciences (e.g.
-\cite{hollerbach2013parity}) and meteorology (e.g.
-\cite{evans1998spherical, rubinstein2015scalar, wedi2013fast, ecmwf2020scalability, diamantakisecmwf, courtier1998ecmwf}).
-Notably, they are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
-While the whole sphere spectral method using the Spherical Harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast, diamantakisecmwf}.
+\cite{fletcher2017data, hollerbach2013parity}) and meteorology (e.g.
+\cite{evans1998spherical, rubinstein2015scalar, wedi2013fast, ecmwf2020scalability, courtier1998ecmwf, silberman1954planetary, gottlieb1977numerical}).
+Spectral methods on the sphere involving spherical harmonics have been used for over 60 years \cite{silberman1954planetary}.
+Notably, spherical harmonics are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
+While the whole sphere spectral method using the spherical harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast}.
 
@@
 For the IFS, the spherical harmonic transform in fact uses two transforms -- a Fourier transform (using the well-established Fast Fourier Transform (FFT) \cite{cooley1965algorithm}) in the longitudinal direction and a Legendre transform in the latitudinal direction -- and it is the Legendre transform that has been identified as inhibiting future performance due to its computational cost.
-While a Fast Legendre Transform (FLT) \cite{wedi2013fast} has been incorporated into the model, along with new grid types \cite{malardel2016new}, to help to extend the lifespan of the spectral method for numerical weather prediction (NWP), it may not be sufficient for certain desired cases and resolutions \cite{diamantakisecmwf}.
+While a Fast Legendre Transform (FLT) \cite{wedi2013fast} has been incorporated into the model, along with new grid types \cite{malardel2016new}, to help to extend the lifespan of the spectral method for numerical weather prediction (NWP), it may not be sufficient for certain desired cases and resolutions \cite{wedi2014increasing}.
 
@@
 Moreover, by still using a spectral approach, one can maintain the high accuracy and excellent error properties that such methods bring.
-
-There are also potential applications in physics too where solving equations on the sphere and working in spherical geometries is also desirable (e.g.
-\cite{beyer2014numerical, varshalovich1988quantum, slevinsky2018spectral})\bstodo{cite more, see p2 of Vasil and the references}, particularly in astrophysics \cite{vasil2019tensor, reinecke2013libsharp}.
+There are also potential applications in physics too, particularly in astrophysics, where solving PDEs on the sphere surface and working in spherical geometries is also desirable (e.g.
+\cite{vasil2019tensor, reinecke2013libsharp, beyer2014numerical, varshalovich1988quantum, slevinsky2018spectral, rubinstein2015scalar}).
 
@@
 
-Chapter 2 of this thesis provides an introduction to sparse spectral methods via the Spherical Harmonics on the whole sphere surface.
-Here, we think of the Spherical Harmonics in a non-traditional way and write them as a group of multidimensional orthogonal polynomials in $(x,y,z)$ as opposed to functions of spherical coordinates.
-We present an OP framework for how the Spherical Harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for a function's expansion.
-Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.
+Chapter 2 of this thesis provides an introduction to sparse spectral methods via the spherical harmonics on the whole sphere surface.
+Here, we think of the spherical harmonics in a non-traditional way and write them as a group of multidimensional orthogonal polynomials in $(x,y,z)$ as opposed to functions of spherical coordinates.
+We present an OP framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for a function's expansion.
+Further, we demonstrate how the vector spherical harmonics can be used as an orthogonal basis for vector valued functions lying in the \enquote{tangent bundle} of the sphere, and thus how one can additionally derive gradient and divergence operators.
 

### 2021-06-03 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Edits to fig layouts mainly
Further edits. New spy plots to make uniform across thesis. New references and fixing of capitals etc.

@@
 To introduce ourselves to the world of multidimensional orthogonal polynomials for solving PDEs on the sphere, we can naturally choose to look at the famous spherical harmonics.
-Our aim here is to express the Spherical Harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere.
+Our aim here is to express the spherical harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere.
 More precisely, we desire the solution to partial differential equations on the domain
@@
 
-In this chapter, we demonstrate how one can write the Spherical Harmonics as polynomials in $(x, y, z)$, and how the relationships between them can be used to obtain sparse \enquote{Jacobi operators} for multiplication by $x,y,z$.
-We demonstrate how these in turn lead to an algorithm for evaluating the Spherical Harmonics, an efficient algorithm for evaluating functions when expanded in the SH OP basis, and sparse \enquote{banded-block-banded} differential operator matrices.
+In this chapter, we demonstrate how one can write the spherical harmonics as polynomials in $(x, y, z)$, and how the relationships between them can be used to obtain sparse \enquote{Jacobi operators} for multiplication by $x,y,z$.
+We demonstrate how these in turn lead to an algorithm for evaluating the spherical harmonics, an efficient algorithm for evaluating functions when expanded in the SH OP basis, and sparse \enquote{banded-block-banded} differential operator matrices.
 The techniques used here are done so with the aim of applying them to other OP families of other domains (and we do so in later chapters).
@@
 We can obtain the entries to these operator matrices by determining the coefficients of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.
-Relations concerning the products of Spherical Harmonics have already been established, and so the expressions we desire here could just be seen as special cases of these.
+Relations concerning the products of spherical harmonics have already been established, and so the expressions we desire here could just be seen as special cases of these.
 More specifically, the coefficients of the expansion in the SH basis of the product of two SHs are composed of Clebsch-Gordan coefficients \cite[p231]{sakurai1995modern}.
@@
 
-However we reiterate our goal here is to use techniques that can generalise to spectral methods on other domains using multidimensional OPs, and thus we proceed by exploiting the three-term recurrences and other relations of the univariate OPs that make up the Spherical Harmonics as we have defined them.
+However we reiterate our goal here is to use techniques that can generalise to spectral methods on other domains using multidimensional OPs, and thus we proceed by exploiting the three-term recurrences and other relations of the univariate OPs that make up the spherical harmonics as we have defined them.
 With that being said, we require a few results concerning the incrementing and decrementing of the parameters for the Jacobi polynomials, and the three-term recurrence for the Jacobi polynomials, that will also prove useful later on in this chapter.
@@
 
-\remark The results in \bsreflemma{lemma:shxyzrelations} are fairly involved and can be derived via the relationships established for products of Spherical Harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
+\remark The results in \bsreflemma{lemma:shxyzrelations} are fairly involved and can be derived via the relationships established for products of spherical harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
 
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{spy-of-Jx}
+		\centerline{\includegraphics[scale=0.5]{spy-of-Jx}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{spy-of-Jy}
+		\centerline{\includegraphics[scale=0.5]{spy-of-Jy}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{spy-of-Jz}
+		\centerline{\includegraphics[scale=0.5]{spy-of-Jz}}
 		\centering
@@
 Three-term recurrence relations for orthogonal polynomials are well established (see for example \cite[Section 18.9]{DLMF}), including for multidimensional OPs \cite[Section 3.2]{dunkl2014orthogonal}.
-In a similar vein, we can obtain a recurrence relation for the Spherical Harmonics by combining each system in \bsrefeqn{eqn:jacobimatsdef} in a specific way.
-It should be emphasised at this point that this is not necessarily the optimal way of computing the Spherical Harmonics; various efficient methods and algorithms have been proposed (see e.g.
+In a similar vein, we can obtain a recurrence relation for the spherical harmonics by combining each system in \bsrefeqn{eqn:jacobimatsdef} in a specific way.
+It should be emphasised at this point that this is not necessarily the optimal way of computing the spherical harmonics; various efficient methods and algorithms have been proposed (see e.g.
 \cite{guy1998efficient, sloan2013efficient, fukushima2012numerical}).
@@
 
-We have established that the Spherical Harmonics can be viewed as a basis of orthonormal polynomials on the sphere.
+We have established that the spherical harmonics can be viewed as a basis of orthonormal polynomials on the sphere.
 Specifically, given a function $f : \Omega \to \C$, we can write $f = \sum_{l = 0}^{N} \sum_{m = -l}^{l} f_{l,m} \; \Ylm$ where $\{f_{l,m}\}$ are coefficients and $N \in \N$ is sufficiently large.
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{spy-of-Jx-tangent}
+		\centerline{\includegraphics[scale=0.5]{spy-of-Jx-tangent}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{spy-of-Jy-tangent}
+		\centerline{\includegraphics[scale=0.5]{spy-of-Jy-tangent}}
 		\centering
@@
 	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.5]{spy-of-Jz-tangent}
+		\centerline{\includegraphics[scale=0.5]{spy-of-Jz-tangent}}
 		\centering
@@
 	\end{subfigure}\hfil % <-- added
-	\caption{\enquote{Spy} plots of the Jacobi operator matrices for the Vector Spherical Harmonic OP basis, showing their sparsity and structure, up to order $N=10$.}
+	\caption{\enquote{Spy} plots of the Jacobi operator matrices for the vector spherical harmonic OP basis, showing their sparsity and structure, up to order $N=10$.}
 	\label{fig:sh-spy-jacobi}
@@
 
-\begin{figure}[tp]
+\begin{figure}[t]
 	\centering % <-- added
-	\centerline{\includegraphics[scale=0.6]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100}}
+	\centerline{\includegraphics[scale=0.5]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100}}
 	\caption{Four snapshots of the solution $u$ of the heat equation with $\alpha = 1/42$ and a step size of $\Delta t = 0.01$, solved using the BDF2 timestepping method.
@@
 
-\begin{figure}[tp]
+\begin{figure}[t]
 	\centering % <-- added
-	\centerline{\includegraphics[scale=0.6]{linear-swe-H=1-dt=0p01-nsteps=100}}
-	\caption{Four snapshots of the solution for the height $h$ of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method.
+	\centerline{\includegraphics[scale=0.5]{linear-swe-H=1-dt=0p01-nsteps=100-arrows}}
+	\caption{Four snapshots of the solution for the height $h$ (surface colour) and the wave velocity $u$ (red arrows) of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method (BDF1).
+The initial conditions are given in \bsrefeqn{eqn:sweics}.
 Top Left: Initial conditions (0 timesteps).
@@
 As a demonstration, in \bsreffig{fig:swe} we see four snapshots of the solution to the system in \bsrefeqn{eqn:swe} on the unit sphere, with $\rotrate = \calH = 1$ and initial conditions given by
-\bseqn{
-	\uvec_0 := \phivec \mu \sinphi, \quad h_0 := \mu \sin^2\varphi,
+\bseqnnumber{
+	\uvec_0 := \phivec \mu \sinphi, \quad h_0 := \mu \sin^2\varphi, \label{eqn:sweics}
 }
@@
 
-Spherical Harmonics have been widely used to solve PDEs, in particular a Spherical Harmonics approach is used as part of the Integrated Forecasting System (IFS) that ECMWF uses for their forecasts.
-In this chapter, we have chosen to view the Spherical Harmonics as orthogonal (in fact orthonormal) polynomials in $x$, $y$ and $z$ in order to illustrate and understand more about how the framework presented here for using multivariate OPs in a spectral method translate to the sphere.
+Spherical harmonics have been widely used to solve PDEs, in particular a spherical harmonics approach is used as part of the Integrated Forecasting System (IFS) that ECMWF uses for their forecasts.
+In this chapter, we have chosen to view the spherical harmonics as orthogonal (in fact orthonormal) polynomials in $x$, $y$ and $z$ in order to illustrate and understand more about how the framework presented here for using multivariate OPs in a spectral method translate to the sphere.
 We have derived the entries of \textit{Jacobi matrices} that represent multiplication by the coordinates, showing how they are \textit{banded-block-banded} in structure.
-We have set out how one can find coefficients for the expansion of a function in the Spherical Harmonics basis using established transforms, and how one can evaluate a function given its coefficients vector using a multidimensional version of the Clenshaw algorithm.
-We extended our framework to include the \textit{tangent bundle} of the sphere -- that is, the space of vectors orthogonal to the unit outward normal for each point on the sphere -- where we can use the \textit{Vector Spherical Harmonics} as vector-valued orthogonal polynomials to be a basis, demonstrating how similar such Jacobi matrices, function expansion and evaluation can be derived.
+We have set out how one can find coefficients for the expansion of a function in the spherical harmonics basis using established transforms, and how one can evaluate a function given its coefficients vector using a multidimensional version of the Clenshaw algorithm.
+We extended our framework to include the \textit{tangent bundle} of the sphere -- that is, the space of vectors orthogonal to the unit outward normal for each point on the sphere -- where we can use the \textit{vector spherical harmonics} as vector-valued orthogonal polynomials to be a basis, demonstrating how similar such Jacobi matrices, function expansion and evaluation can be derived.
 Further, by exploiting the natural sparsity of the relationships between the OPs in the bases, we have shown how matrices for differential and other operators will be sparse and banded-block-banded in structure too.
-In fact, for the Spherical Harmonics, these are simply diagonal.
+In fact, for the spherical harmonics, these are simply diagonal.
 Finally, we presented a couple of simple examples of how one can put this all into practice, namely solving the heat equation and the linearised shallow water equations on the unit sphere.

### 2021-06-03 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Further edits. New spy plots to make uniform across thesis. New references and fixing of capitals etc.

@@
 
-In recent years, there has been some interesting work in the area of multivariate orthogonal polynomials in two-dimensional domains (see \cite{dunkl2014orthogonal}), such as for a Chebychev-based spectral method on the rectangle \cite{julien2009efficient} \bstodo{list some domains and cite references for them}.
+In recent years, there has been some interesting work in the area of multivariate orthogonal polynomials in two-dimensional domains (see \cite{dunkl2014orthogonal}), such as for a Chebychev-based spectral method on the rectangle \cite{julien2009efficient}.
 In particular, spectral methods have been established on the disk \cite{boyd2011comparing, vasil2016tensor} and the triangle \cite{olver2019triangle}.
-Jacobi polynomials are extremely useful for solving a number of problems in not just 1D but 2D also, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.
-Applications for the disk include \bstodo{give applications and references} and for the triangle include solving Volterra integral equations \cite{gutleb2020sparse}, \bstodo{give more}.
-
-In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, for the triangle a choice of OPs are the Jacobi polynomials on the triangle, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
-\bseqnnumber{
-	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:triangleops}
-}
-for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values.
+Examples of applications for the disk include use in optics \cite{mahajan2007orthonormal}, astrophysics \cite{pringle1981accretion}, and fluid dynamics \cite{eggels1994fully, noll1976zernike, kerswell2005recent}, and for the triangle include solving Volterra integral equations \cite{gutleb2020sparse}.
 
@@
 However, ideas for how to overcome this have progressed \cite{slevinsky2018use, sakai2009application, wilber2017computing}.
-One solution proposed is a radial basis defined in terms of Jacobi polynomials \cite{vasil2016tensor}:
+One solution proposed is pairing the Zernike disk OPs with a radial basis defined in terms of Jacobi polynomials \cite{vasil2016tensor}:
 \bseqnnumber{
@@
 $\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials.
-Again, the idea here is that differential and other operators, as well as basis (parameter) conversion operators, can be represented as banded matrices.
+The use of polar coordinates here then allows for radially symmetric partial differential (and other) operators, as well as basis (parameter) conversion operators, to be represented as banded matrices by exploiting the relations of the Jacobi polynomials -- they are extremely useful for solving a number of problems in not just 1D but 2D also, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.
+
+Conversely, in a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
+\bseqnnumber{
+	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:triangleops}
+}
+for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values.
+Since there is no radial symmetry, differential operators for the triangle (as will be the case for the domains we investigate in this chapter) will be more complex than those for the disk.
 
 In this chapter, we aim to present a basis and develop a sparse spectral method on generalisations of these domains, namely the disk-slice and trapezium.
-The full unit disk and triangle cases are special cases of the spectral methods we investigate here too.
+The full unit disk and triangle cases can be viewed as special cases of the spectral methods we investigate here too.
 We will show that the operators for multiplication by $x$ and $y$, as well as differential and parameter conversion, will be banded-block-banded.
@@
 where $C_{n,k} \neq 0$ and \enquote{lower order terms'} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
-The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:OPconstruction}.
+The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:ds:OPconstruction}.
 
@@
 
-\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:construction}
+\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
 	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
@@
 
-For disk slices and trapeziums, we specialise Proposition \ref{prop:construction} in the following definitions. First we introduce notation for two families of univariate OPs.
-\begin{definition}\label{def:OPconstruction}
+For disk slices and trapeziums, we specialise \bsrefprop{prop:ds:construction} in the following definitions. First we introduce notation for two families of univariate OPs.
+\begin{definition}\label{def:ds:OPconstruction}
 	Let $\genjacw^{(a,b,c)}(x)$ and $\jacw^{(a,b)}(x)$ be two weight functions on the intervals $(\alpha, \beta)$ and $(\gamma, \delta)$ respectively, given by:
@@
 We can now write down our 2D OP family.
-\begin{definition}\label{def:constuction}
+\begin{definition}\label{def:ds:constuction}
 	For parameters $a,b,c,d \in \R$, define the four-parameter 2D orthogonal polynomials as the set $\{\hdopnk^{(a,b,c,d)}\}$ where:
@@
 
-Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
+Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:sh:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
 \bseqn{
@@
 }
-\label{fig:laplaceparameterconversion} 
+\label{fig:ds:laplaceparameterconversion} 
 \end{figure}
@@
 
-\begin{definition}\label{def:differentialoperators}
+\begin{definition}\label{def:ds:differentialoperators}
 Define the operator matrices $D_x^{(a,b,c)}, \: D_y^{(a,b,c)}, \: W_x^{(a,b,c)}, \: W_y^{(a,b,c)}$ according to:
@@
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
-An illustration of how the non-weighted differential operators increment the parameters (a,b,c) is seen in Figure \ref{fig:laplaceparameterconversion}.
-
-\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
-The operator matrices $D_x^{(a,b,c)}, \: D_y^{(a,b,c)}, \: W_x^{(a,b,c)}, \: W_y^{(a,b,c)}$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure.
+An illustration of how the non-weighted differential operators increment the parameters (a,b,c) is seen in \bsreffig{fig:ds:laplaceparameterconversion}.
+
+\begin{theorem}\label{theorem:ds:sparsityofdifferentialoperators}
+The operator matrices $D_x^{(a,b,c)}, \: D_y^{(a,b,c)}, \: W_x^{(a,b,c)}, \: W_y^{(a,b,c)}$ from \bsrefdef{def:ds:differentialoperators} are sparse, with banded-block-banded structure.
 More specifically:
@@
 
-\begin{definition}\label{def:parametertransformationoperators}
+\begin{definition}\label{def:ds:parametertransformationoperators}
 Define the operator matrices 
@@
 
-\begin{lemma}\label{lemma:sparsityofparametertransformationoperators}
-The operator matrices in Definition \ref{def:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
+\begin{lemma}\label{lemma:ds:sparsityofparametertransformationoperators}
+The operator matrices in \bsrefdef{def:ds:parametertransformationoperators} are sparse, with banded-block-banded structure. More specifically:
 \begin{itemize}
@@
 (c) The biharmonic operator $\biharmonictwo$}
-	\label{fig:sparsity}
+	\label{fig:ds:sparsity}
 \end{figure}
@@
 }
-Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in Figure \ref{fig:sparsity}.
+Importantly, this operator will have banded-block-banded structure, and hence will be sparse, as seen in \bsreffig{fig:ds:sparsity}.
 
@@
 }
-The sparsity and structure of this biharmonic operator are seen in Figure \ref{fig:sparsity}.
+The sparsity and structure of this biharmonic operator are seen in \bsreffig{fig:ds:sparsity}.
 
@@
 We thus first need to define a new set of `interim' 1D OPs.
-\begin{definition}\label{def:InterimOPconstruction}
+\begin{definition}\label{def:ds:InterimOPconstruction}
 Let $\genjactw^{(a,b,c,d)}(x) := (\beta - x)^a \: (x - \alpha)^{b} \: (1-x)^{c} \: (1+x)^{d} $ be a weight function on the interval $(\alpha, \beta)$, and define the associated inner product by:
@@
 
-The proofs of Theorem \ref{theorem:sparsityofdifferentialoperators} and Lemma \ref{lemma:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in Definition \ref{def:differentialoperators} and Definition \ref{def:parametertransformationoperators}.
+The proofs of \bsreftheorem{theorem:ds:sparsityofdifferentialoperators} and \bsreflemma{lemma:ds:sparsityofparametertransformationoperators} provide a way to calculate the non-zero entries of the operator matrices given in \bsrefdef{def:ds:differentialoperators} and \bsrefdef{def:ds:parametertransformationoperators}.
 We can simply use quadrature to calculate the 1D inner products, which has a complexity of $\bigO(N^3)$.
@@
 		\centering
-		%\label{fig:solution-poisson}
+		%\label{fig:ds:solution-poisson}
 	\end{subfigure}
@@
 		\includegraphics[scale=0.45]{solutionblocknorms-poisson-diskslice-alpha=0p2-beta=0p8-N=200}
-        		%\label{fig:solutionblocknorms-poisson}
+        		%\label{fig:ds:solutionblocknorms-poisson}
 	\end{subfigure}
@@
 	\centering
-	\label{fig:poisson}
+	\label{fig:ds:poisson}
 \end{figure}
@@
 	\centering
-	\label{fig:poissonexact}
+	\label{fig:ds:poissonexact}
 \end{figure}
@@
 	\includegraphics[scale=0.35]{solution-helmholtz-diskslice-alpha=0p2-beta=0p8-k=100-n=300}
-	%\label{fig:solution-helmholtz}
+	%\label{fig:ds:solution-helmholtz}
 	\end{subfigure}
@@
 	\centering
-        	%\label{fig:solutionblocknorms-helmholtz}
+        	%\label{fig:ds:solutionblocknorms-helmholtz}
 	\end{subfigure}
@@
 	\centering
-	\label{fig:helmholtz}
+	\label{fig:ds:helmholtz}
 \end{figure}
@@
 	\includegraphics[scale=0.35]{solution-biharmonic-diskslice-alpha=0p2-beta=0p8}
-	%\label{fig:solution-biharmonic}
+	%\label{fig:ds:solution-biharmonic}
 	\end{subfigure}
@@
 	\centering
-        	%\label{fig:solutionblocknorms-biharmonic}
+        	%\label{fig:ds:solutionblocknorms-biharmonic}
 	\end{subfigure}
@@
 	\centering
-	\label{fig:biharmonic}
+	\label{fig:ds:biharmonic}
 \end{figure}
@@
 which can be solved to find $\mathbf{u}$.
-In Figure \ref{fig:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
-In Figure \ref{fig:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 990$, that is, $491,536$ unknowns.
+In \bsreffig{fig:ds:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
+In \bsreffig{fig:ds:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 990$, that is, $491,536$ unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
@@
 
-In Figure \ref{fig:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error.
+In \bsreffig{fig:ds:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error.
 The example was chosen so that the exact solution was $u(x,y) = \Wiii (x,y) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y) = \Delta \big[\Wiii (x,y) y^3 \exp(x)\big]$.
@@
 which can be solved to find $\mathbf{u}$.
-We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example.
-In Figure \ref{fig:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
-In Figure \ref{fig:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using $N = 500$, that is, $125,751$ unknowns.
+We can see the sparsity and structure of this matrix system in \bsreffig{fig:ds:sparsity} with $v(x,y) = xy^2$ as an example.
+In \bsreffig{fig:ds:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
+In \bsreffig{fig:ds:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using $N = 500$, that is, $125,751$ unknowns.
 The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution.
@@
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-In Figure \ref{fig:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
-In Figure \ref{fig:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with $N = 500$, that is, $125,751$ unknowns.
+In \bsreffig{fig:ds:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
+In \bsreffig{fig:ds:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with $N = 500$, that is, $125,751$ unknowns.
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we see faster convergence.
@@
 The sparsity of operator matrices for partial differentiation by $x, y$ as well as for parameter transformations generalise to such end-disk-slice domains.
-For instance, if we inspect the proof of Lemma \ref{theorem:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for an end-disk-slice.
-
-In Figure \ref{fig:ds:trappfem} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
+For instance, if we inspect the proof of \bsreftheorem{theorem:ds:sparsityofdifferentialoperators}, we see that it can easily generalise to the weights and domain $\Omega$ for an end-disk-slice.
+
+In \bsreffig{fig:ds:trappfem} we see the solution to the Poisson equation with zero boundary conditions in the half-disk $\Omega$ with $(\alpha,\beta) := (0,1)$.
 
@@
 
-In Figure \ref{fig:ds:trappfem} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
+In \bsreffig{fig:ds:trappfem} we see the solution to the Helmholtz equation with zero boundary conditions in the trapezium $\Omega$ with $\xi := \half$.
 
@@
 	\includegraphics[scale=0.35]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-        	%\label{fig:solutionblocknorms-poisson}
+        	%\label{fig:ds:solutionblocknorms-poisson}
 	\centering
@@
 	\centering
-	%\label{fig:solution-poisson}
+	%\label{fig:ds:solution-poisson}
 	\end{subfigure}
@@
 	\includegraphics[scale=0.38]{solution-trapezium-helmholtz-k=100-n=1500}
-	%\label{fig:solution-poisson}
+	%\label{fig:ds:solution-poisson}
 	\centering
@@
 In this chapter we have shown that bivariate orthogonal polynomials can lead to sparse discretizations of general linear PDEs on specific domains whose boundary is specified by an algebraic curve -- notably here the disk-slice -- with Dirichlet boundary conditions.
-This work extends the triangle case \cite{beuchler2006new, li2010optimal, olver2019triangle} and whole disk case \cite{boyd2011comparing, vasil2016tensor} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, vasil2016tensor}).
+This work extends the triangle case \cite{beuchler2006new, li2010optimal, olver2019triangle} and whole disk case \cite{boyd2011comparing, vasil2016tensor} to non-classical geometries, and forms a building block in developing an $hp-$finite element method to solve PDEs on other polygonal domains by using suitably shaped elements (for example, by dividing the disk or a section of the disk into disk-slice elements, which has applications in turbulent pipe flow \cite{eggels1994fully, kerswell2005recent, vasil2016tensor}).
 

### 2021-06-03 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Further edits. New spy plots to make uniform across thesis. New references and fixing of capitals etc.

@@
 
-The spherical cap can be thought of as higher dimensional version of the circle arc (a one-dimensional \enquote{surface} in two-dimensional space).
+The spherical cap can be thought of as a higher dimensional version of the circle arc (a one-dimensional \enquote{surface} in two-dimensional space).
+Recently, a procedure for defining OPs in two variables on quadratic curves, such as ellipses and parabolas, has been presented \cite{olver2021orthogonal}, where the unit circle is one of the five classes of curve that can be generalised to give any curve in the plane $\R^2$.
 One can define OPs on the circle in terms of Fourier series, which we write here as orthogonal polynomials in $x$ and $y$:
-\begin{definition}\label{def:sc:Ydefinition}
+\begin{definition}[\cite{olver2021orthogonal}]\label{def:sc:Ydefinition}
 	Define the unit circle $\omega := \bsset{\vec{x} = (x,y) \in \R^2}{x^2 + y^2 = 1}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$.
@@
 
-\bstodoinline{Expand on this. Read the Olver Xu "OPs on quadratic curves paper." where the arc might apply??? Either way, cite this paper in intro (CH1).}
-
-\section{Orthogonal polynomials on algebraic curves}
-
-This idea of using 1D OPs on a shortened interval to create the modified OPs for a circle arc has be generalised to higher dimensional domains, where a way of constructing orthogonal polynomials on algebraic curves has been presented \cite{olver2020orthogonal}.
-\bstodoinline{Expand on this.
-Summarise the "on" surfaces approach.
-Use references in the paper.}
-
 \section{Orthogonal polynomials on spherical caps}\label{section:sc:OPs}
@@
 
-Using the approach presented for constructing OPs on algebraic curves, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval $[\alpha,\beta]$, and from Fourier series written as orthogonal polynomials in $x$ and $y$.
+A continuation to the procedure for defining OPs on quadratic curves (see \cite{olver2021orthogonal}) has also been presented \cite{olver2020orthogonal}, where OPs are defined on quadratic surfaces of revolution in higher dimensions (specifically, surfaces in $\R^{d+1} for d \ge 2$).
+The authors describe their technique as modelled on the structure of the unit sphere, by approaching the construction as a generalisation of spherical harmonics.
+
+Using this approach, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval $[\alpha,\beta]$, and from Fourier series written as orthogonal polynomials in $x$ and $y$.
 Recall that the spherical cap is defined by
@@
 		\centering
-		\caption{The Laplace-Beltrami operator $\Delta^{(1)}_W$}
+		\caption{}
 	\end{subfigure}\hfil % <-- added
@@
 		\centering
-		\caption{The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$}
+		\caption{}
 	\end{subfigure}\hfil % <-- added
@@
 		\centering
-		\caption{A variable coefficient Helmholtz operator}
+		\caption{}
 	\end{subfigure}\hfil % <-- added
@@
 		\centering
-		\caption{The Biharmonic operator $\calB_W^{(2)}$} 
+		\caption{} 
 	\end{subfigure}\hfil % <-- added
-	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity.
-For (c), the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
+	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity, plotted for $N = 20$.
+(a) The Laplace-Beltrami operator $\Delta^{(1)}_W$.
+(b) The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$.
+(c) The variable coefficient Helmholtz operator $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.
+(d) The Biharmonic operator $\calB_W^{(2)}$.}
 	\label{fig:sc:sparsity}
@@
 where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see \bsrefsection{Section:sc:furtherdiffoperators}).
+For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see \bsrefsection{section:sc:furtherdiffoperators}).
 In \bsreffig{fig:sc:biharmonic} we see the solution to the Biharmonic \bsrefeqn{eqn:sc:biharmonic} in the spherical cap $\Omega$.

### 2021-06-03 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Fixed spacing and margins for eqns in ch4. Need to do same for ch3.

@@
 	x \: \scopnkia(x,y,z) &= \alphaa_{n,k,1} \:  \scopa_{n-1, k-1, i}(x, y, z) + \alphaa_{n,k,2} \:  \scopa_{n-1, k+1, i}(x, y, z) \nonumber \\
-		& \quad \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
-		& \quad \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\
+		& \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
+		& \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\
 	y \: \scopnkia(x,y,z) &= \betaa_{n,k,i,1} \:  \scopa_{n-1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,2} \:  \scopa_{n-1, k+1, |i-1|}(x, y, z) \nonumber \\
-		& \quad \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
-		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
-	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
+		& \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
+		& \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
+	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) \nonumber \\
+		& \quad + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),
 }
@@
 where we show the non-zero coefficients that result are the $\alphaa_{n,k,1},\dots,\alphaa_{n,k,6}$ in the lemma.
-Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$.
-Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)$:
+Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$. Using a change of variables 
+\bseqn{
+	(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)
+}
+we have that, for $m = 0,\dots,n+1$, $j = 0,\dots,m$, 
 \bseqn{
@@
 	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: x \: \genjacw^{(a,0)}(z) \: \D A \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
+	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \\
+	&\quadfour \cdot \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
+	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z
 }
@@
 	&= \int_\Omega \scopnkia(\xvec,z) \: \scopmjha(\xvec,z) \: y \: \genjacw^{(a,0)}(z) \: \D A \\
-	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
+	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \\
+	&\quadfour \cdot \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.
+	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \\
+	&\quadfour \cdot \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big)
 }
@@
 			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
 	B^{(a)}_{x,n} &:= 
@@
 			& & A^{(a)}_{n,n,3} & 0
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
 	C^{(a)}_{x,n} &:= 
@@
 			& & & A^{(a)}_{n,n,1} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
+		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad (n \ne 0)
 }
@@
 			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
 	B^{(a)}_{y,n} &:= 
@@
 			& & B^{(a)}_{n,n,3} & 0
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
 	C^{(a)}_{y,n} &:= 
@@
 			& & & B^{(a)}_{n,n,1} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
+		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad (n \ne 0)
 }
@@
 			& & 0 & \Gamma^{(a)}_{n,n,3} & 0 \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
 	B^{(a)}_{z,n} &:= 
@@
 			& & & \Gamma^{(a)}_{n,n,2}
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
 	C^{(a)}_{z,n} &:= 
@@
 			& & & 0 \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
+		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad (n \ne 0)
 }
@@
 	\renewcommand\arraystretch{1.3}
-	\begin{pmatrix}
+	&\begin{pmatrix}
 		1 & & & \\
@@
 	\end{pmatrix}
-	\bigscopta(x,y,z) =
-		\begin{pmatrix}
-	 		\scopa_0 \\ 0 \\ 0 \\ 0 \\ \vdots  \\
-		\end{pmatrix},
+	\bigscopta(x,y,z) \\
+	&=
+		\begin{pmatrix}
+	 		\scopa_0 & 0 & 0 & 0 & \hdots
+		\end{pmatrix}^\top,
 }
@@
 such that $\Dnt A_n = I_{2n+3}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
 \bseqn{
@@
 		\bigscopta_{0}(x,y,z) := \scopa_0 \\
-		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
-	\end{cases}
+		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z)
+	\end{cases}.
 }
@@
 \bseqn{
-	&\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
-	&\quad \quad = \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
-	&\quad \quad = \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
+	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] &= \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} \\
+	&\quadfour + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
+	&= \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
 }
@@
 	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
-	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \genjacw^{(a+1, 0)} \: \D z \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+1,2k)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, 2k)} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
+	&= - \int_\alpha^1 \Big\{ \int_0^{2\pi} \rho(z)^2 \ddz \:  \big[ \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \big] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \\
+	&\quadfour \quadtwo \cdot \genjacw^{(a+1, 0)}(z) \Big\} \: \D z \\
+	&= \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quadfour \cdot \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: \big[ k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime} \big] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
+	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjac_{m-k}^{(a+1,2k)} \: \big[ k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime} \big] \: \genjacw^{(a+1, 2k)} \: \D z \\
+	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
+	&\quadeight + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
 }
@@
 	&\ip<\rhoppphi (\genjacw^{(a,0)} \: \scopnkia), \: \scopmjh^{(a-1)}>_{\scop^{(0)}} \\
-	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \genjacw^{(a, k)}(z)] \: \genjacmmj^{(a-1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \D z \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a-1,2j)} \: [k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2] \: \genjacw^{(a-1, k + j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \:  \int_\alpha^1 \: \genjac_{m-k}^{(a-1,2k)} \: [k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2] \: \genjacw^{(a-1, 2k)} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} - a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \genjacw^{(1, 2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a-1, 2k)} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
+	&= - \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quadfour \cdot \Big( \int_\alpha^1 \rho(z)^2 \ddz \: \big[ \genjacnmk^{(a,2k)}(z) \: \genjacw^{(a, k)}(z) \big] \: \genjacmmj^{(a-1,2j)}(z) \: \rho(z)^j \: \D z \Big) \\
+	&= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quad \cdot \: \Big( \int_\alpha^1 \genjacmmj^{(a-1,2j)} \:  \big[k z \genjacnmk^{(a, 2k)} \: \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \: \genjacw^{(1, 2)} - a \: \genjacnmk^{(a, 2k)} \: \rho^2 \big] \genjacw^{(a-1, k + j)} \: \D z \Big) \\
+	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjac_{m-k}^{(a-1,2k)} \big[ k z \genjacnmk^{(a, 2k)} \genjacw^{(1, 0)} - \genjacnmk^{(a, 2k) \: \prime} \genjacw^{(1, 2)} - a \genjacnmk^{(a, 2k)} \rho^2 \big] \genjacw^{(a-1, 2k)} \: \D z \\
+	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} - a \: \rho^2 \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \genjacw^{(1, 2)} \\
+	&\quadeight \quadtwo + a \: \rho^2 \genjac_{m-k}^{(a-1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} \Big\} \genjacw^{(a-1, 2k)} \: \D z \\
+	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjacnmk^{(a, 2k)} \big[ k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \big] \genjacw^{(a, 2k)} \: \D z
 }
@@
 \bseqnnumber{
-	\DeltaS \: \scopnki^{(a)} &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
-	&\quad \quad \quad \quad + \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
-	&= \chki(\theta) \rho(z)^k \Big\{ -k(k+1) \genjacnmk^{(a,2k)}(z) - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime}(z) \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime}(z) \Big\}, \label{eqn:sc:oplaplacian} \\
-	\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) &= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
-	&\quad \quad \quad \quad + \genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
+	&\DeltaS \: \scopnki^{(a)} \nonumber \\
+	&= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) + \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
+	&= \chki(\theta) \: \rho(z)^k \: \Big\{ -k(k+1) \genjacnmk^{(a,2k)}(z) - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime}(z) \nonumber \\
+	&\quadeight + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime}(z) \Big\}, \label{eqn:sc:oplaplacian}
+}
+and
+\bseqnnumber{
+	&\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big) \nonumber \\
+	&= {1 \over \rho} \ppx{\varphi} \Big( \rho \ppx{\varphi} [\genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^k\varphi] \Big) \chki(\theta) \nonumber \\
+	&\quadfour + \genjacw^{(a,0)}(\cosphi) \: \genjacnmk^{(a,2k)}(\cosphi) \: \sin^{k-2}\varphi \pmpxm{\theta}{2} \chki(\theta) \nonumber \\
 	&= \chki(\theta) \Big\{\genjacnmk^{(a,2k)}(z) \big[ -k(k+1) \genjacw^{(a,k)}(z) - 2a(k+1) z \: \genjacw^{(a-1,k)}(z) \big] \nonumber \\
@@
 	&\ip<\DeltaS \scopnki^{(a)}, \: \scopmjh^{(a + \tilde a)}>_{\scop^{(a + \tilde a)}} \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a + \tilde a, 2j)} \: \genjacw^{(a + \tilde a), k+j)} \: \Big\{-k(k+1) \genjacnmk^{(a,2k)} - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime} \Big\} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a+\tilde a, 0)} \Big( -k(k+1) \genjacnmk^{(a,2k)} \: \rho^{2k} + \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \rho^{2(k+1)}] \Big) \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ -k(k+1) \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a+\tilde a, 2k)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - \genjacnmk^{(a,2k) \: \prime} \genjacw^{(a+\tilde a-1, 2k)} \: [\genjac_{m-k}^{(a+\tilde a,2k) \: \prime} \: \genjacw^{(1,0)} + (a + \tilde a) \genjac_{m-k}^{(a+\tilde a,2k)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z
+	&= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quadfour\cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a + \tilde a, 2j)} \: \genjacw^{(a + \tilde a), k+j)} \: \Big\{-k(k+1) \genjacnmk^{(a,2k)} - 2(k+1) z \: \genjacnmk^{(a,2k) \: \prime} \\
+	&\quadeight \quadeight \quad + \rho(z)^2 \genjacnmk^{(a,2k) \: \prime \prime} \Big\} \: \D z \Big) \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a+\tilde a, 0)} \Big\{ -k(k+1) \genjacnmk^{(a,2k)} \: \rho^{2k} \\
+	&\quadeight \quadeight + \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \rho^{2(k+1)}] \Big\} \: \D z \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ -k(k+1) \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a+\tilde a, 2k)} \\
+	&\quadfour \quadtwo \quad - \genjacnmk^{(a,2k) \: \prime} \genjacw^{(a+\tilde a-1, 2k)} \: [\genjac_{m-k}^{(a+\tilde a,2k) \: \prime} \: \genjacw^{(1,0)} + (a + \tilde a) \genjac_{m-k}^{(a+\tilde a,2k)}] \Big\} \: \D z \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z
 }
@@
 	&\ip<\DeltaS \big(\genjacw^{(a,0)} \: \scopnki^{(a)} \big), \: \scopmjh^{(a - \tilde a)}>_{\scop^{(0)}} \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad  \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(a - \tilde a, 2j)} \: \genjacw^{(a - 2, k+j)} \: \Big\{\genjacnmk^{(a,2k)} [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} [-2(k+1) z \: \genjacw^{(2,0)} + 2a \genjacw^{(1,2)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime \prime} \: \genjacw^{(2,2)} \Big\} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \genjacw^{(a, 2k+2)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} ] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a [\genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a-1, 2k+2)} \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} [\genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} ] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - a \genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a-2, 2k)} \: [(a-1) \rho^2 - 2(k+1) z \: \genjacw^{(1, 0)}] \Big\} \: \D z \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z
+	&= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quadfour \cdot \: \Big( \int_\alpha^1 \Big\{\genjacnmk^{(a,2k)} [-k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} + a (a-1) \rho^2] \\
+	&\quadeight + \genjacnmk^{(a,2k) \: \prime} [-2(k+1) z \: \genjacw^{(2,0)} + 2a \genjacw^{(1,2)}] \\
+	&\quadeight + \genjacnmk^{(a,2k) \: \prime \prime} \: \genjacw^{(2,2)} \Big\} \: \genjacmmj^{(a - \tilde a, 2j)} \: \genjacw^{(a - 2, k+j)} \:  \D z \Big) \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: \big[ -k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} \\
+	&\quadeight \quadeight \quadtwo \quad + a (a-1) \rho^2 \big] \\
+	&\quadeight + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
+	&\quadeight + \genjac_{m-k}^{(a - \tilde a, 2k)} \: \ddx{z} [\genjacnmk^{(a,2k) \: \prime} \: \genjacw^{(a, 2k+2)}] \Big\} \: \D z \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: \big[ -k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} \\
+	&\quadeight \quadeight \quadtwo \quad + a (a-1) \rho^2 \big] \\
+	&\quadeight + a \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a - 1, 2k+2)} \\
+	&\quadeight + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} \big[ \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 \\
+	&\quadeight \quadeight \quadtwo \quad - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)} \big] \Big\} \: \D z \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: \big[ -k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} \\
+	&\quadeight \quadeight \quadtwo \quad + a (a-1) \rho^2 \big] \\
+	&\quadfour \quadtwo + \genjacnmk^{(a,2k)} \: \genjacw^{(a - 1, 2k+2)} \big[ \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \rho^2 - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \big] \\
+	&\quadfour \quadtwo - a \genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a-2, 2k)} \: \big[ (a-1) \rho^2 - 2(k+1) z \: \genjacw^{(1, 0)} \big] \Big\} \: \D z \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z
 }
@@
 	&\ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
-	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
-	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(1, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
-	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(1, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(1, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(1, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
-	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(1, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z
+	&= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
+	&\quadfour \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(1, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
+	&\quadeight \quadfour + \genjacnmk^{(1, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
+	&\quadeight \quadfour + \genjacnmk^{(1, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(1, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
+	&\quadeight \quadfour + c_{n-1,k} \genjac_{n-k-1}^{(1, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(1, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
+	&= - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \\
+	&\quadfour \cdot \Big( \int_\alpha^1 \: \Big\{ \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(1, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\
+	&\quadeight + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z \Big)
 }
@@
 \bseqn{
-	\ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}} &= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \cdot \: \Big( \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjacmmj^{(a+\tilde a,2j)} \: \rho^{k+j} \: \genjacw^{(a + \tilde a, 0)} \: \D z \Big) \\
+	&\ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}} \\
+	&= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjacmmj^{(a+\tilde a,2j)} \: \rho^{k+j} \: \genjacw^{(a + \tilde a, 0)} \: \D z \Big) \\
 	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a + \tilde a, 2k)} \: \D z
@@
 \subsection{Further partial differential operators}\label{section:sc:furtherdiffoperators}
+
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.

### 2021-06-04 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Fixed margins for eqns in CH3

@@
 \bseqnnumber{
-	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:diskops}.
+	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}.
 }
@@
 \bseqnnumber{
-	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:triangleops}
+	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:ds:triangleops}
 }
@@
 \bseqnnumber{
-	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ipgenjac} \\
-	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ipjac}
+	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ds:ipgenjac} \\
+	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ds:ipjac}
 }
@@
 \bseqnnumber{
-	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ipnormalisations}
-}
-Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjac}), and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipjac}).
+	\normgenjac^{(a,b,c)} := \int_\alpha^\beta \: \genjacw^{(a,b,c)}(x) \: \D x, \quad \normjac^{(a,b)} := \int_\gamma^\delta \: \jacw^{(a,b)}(y) \: \D y. \label{eqn:ds:ipnormalisations}
+}
+Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjac}, and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipjac}.
 \end{definition}
@@
 	&\ip< \hdopnkabcd, \: \hdopmjabcd >_\Wabcd \\
-	&\quad \quad \quad =  \iint_\Omega \: \Big[ \genjacnmk^{(a,b,c+d+2k+1)}(x) \: \genjacmmj^{(a,b,c+d+2j+1)}(x) \: \rho(x)^{k+j}  \nonumber \\
-		&\quad \quad \quad \quad \quad \quad \cdot \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}) \: \jac_j^{(d,c)}\fpr(\frac{y}{\rho(x)})  \: \Wabcd(x,y) \Big] \: \D y \: \D x \nonumber \\
-	&\quad \quad \quad = \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b,c+d+2k+1)}}(x) \: {\genjacmmj^{(a,b,c+d+2j+1)}}(x) \: \genjacw^{(a,b,c+d+k+j+1)}(x) \: \D x \Big) \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta \: {\jac_k^{(d,c)}}(t) \:  {\jac_j^{(d,c)}}(t) \: \jacw^{(d,)}(t) \: \D t \Big) \nonumber \\
-	&\quad \quad \quad = \normjac^{(d,c)} \: \delta_{k,j} \: \int_\alpha^\beta \: {\genjacnmk^{(a,b,c+d+2k+1)}}(x) \: {\genjac_{m-k}^{(a,b,c+d+2k+1)}}(x) \: \genjacw^{(a,b,c+d+2k+1)}(x) \: \D x \nonumber \\
-	&\quad \quad \quad = \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}. \label{eqn:normhdop}
+	&=  \iint_\Omega \: \Big[ \genjacnmk^{(a,b,c+d+2k+1)}(x) \: \genjacmmj^{(a,b,c+d+2j+1)}(x) \: \rho(x)^{k+j}  \nonumber \\
+		&\quadfour \cdot \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}) \: \jac_j^{(d,c)}\fpr(\frac{y}{\rho(x)})  \: \Wabcd(x,y) \Big] \: \D y \: \D x \nonumber \\
+	&= \Big( \int_\alpha^\beta \: {\genjacnmk^{(a,b,c+d+2k+1)}}(x) \: {\genjacmmj^{(a,b,c+d+2j+1)}}(x) \: \genjacw^{(a,b,c+d+k+j+1)}(x) \: \D x \Big) \nonumber \\
+		&\quadfour \cdot \: \Big( \int_\gamma^\delta \: {\jac_k^{(d,c)}}(t) \:  {\jac_j^{(d,c)}}(t) \: \jacw^{(d,)}(t) \: \D t \Big) \nonumber \\
+	&= \normjac^{(d,c)} \: \delta_{k,j} \: \int_\alpha^\beta \: {\genjacnmk^{(a,b,c+d+2k+1)}}(x) \: {\genjac_{m-k}^{(a,b,c+d+2k+1)}}(x) \: \genjacw^{(a,b,c+d+2k+1)}(x) \: \D x \nonumber \\ 
+	&= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}. \label{eqn:ds:normhdop}
 }
@@
 \bseqnnumber{
-	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:Hrecurrence} \\
-	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:Precurrence}
+	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:ds:Hrecurrence} \\
+	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:ds:Precurrence}
 }
 Of course, for the disk-slice case, we have that $c = d$ and $\gamma_{n}^{(c,c)} = 0$ $\forall n=0,1,2,\dots$.
-We can use (\ref{eqn:Hrecurrence}) and (\ref{eqn:Precurrence}) to determine the 2D recurrences for $\hdopnkabcd(x,y)$.
+We can use \bsrefeqns{eqn:ds:Hrecurrence}{eqn:ds:Precurrence} to determine the 2D recurrences for $\hdopnkabcd(x,y)$.
 Importantly, we can deduce sparsity in the recurrence relationships.
@@
 \bseqn{
-	&x \hdopnkabcd(x,y) \\
-	&\quadtwo = \alphaabcd_{n,k,1} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n+1,k,1} \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
-	&y \hdopnkabcd(x,y) \\
-	&\quadtwo = \betaabcd_{n,k,1} \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,3} \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) \\
-		&\quadfour + \betaabcd_{n,k,4} \hdop_{n, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,5} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \\
-		&\quadfour + \betaabcd_{n,k,7} \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \hdop_{n+1, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,9} \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
+	x \hdopnkabcd(x,y) &= \alphaabcd_{n,k,1} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) + \alphaabcd_{n,k,2} \hdop_{n, k}^{(a,b,c,d)}(x, y) \\
+		&\quad + \alphaabcd_{n+1,k,1} \hdop_{n+1, k}^{(a,b,c)}(x, y), \\
+	y \hdopnkabcd(x,y) &= \betaabcd_{n,k,1} \hdop_{n-1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,2} \hdop_{n-1, k}^{(a,b,c,d)}(x, y) \\
+		&\quad + \betaabcd_{n,k,3} \hdop_{n-1, k+1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,4} \hdop_{n, k-1}^{(a,b,c,d)}(x, y) \\
+		&\quad + \betaabcd_{n,k,5} \hdop_{n, k}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,6} \hdop_{n, k+1}^{(a,b,c,d)}(x, y) \\
+		&\quad + \betaabcd_{n,k,7} \hdop_{n+1, k-1}^{(a,b,c,d)}(x, y) + \betaabcd_{n,k,8} \hdop_{n+1, k}^{(a,b,c,d)}(x, y) \\
+		&\quad + \betaabcd_{n,k,9} \hdop_{n+1, k+1}^{(a,b,c,d)}(x, y),
 }
@@
 \begin{proof}
-The 3-term recurrence for multiplication by $x$ follows from equation (\ref{eqn:Hrecurrence}).
+The 3-term recurrence for multiplication by $x$ follows from \bsrefeqn{eqn:ds:Hrecurrence}.
 For the recurrence for multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$.
@@
 }
-Recall from equation (\ref{eqn:normhdop}) that $\norm{\hdopmjabcd}_{\Wabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$.
+Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopmjabcd}_{\Wabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$.
 Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
@@
 	&\ip<y \hdopnkabcd, \hdopmjabcd>_\Wabcd \\
-	&\quad \quad =  \iint_\Omega \hdopnkabcd(x,y) \: \hdopmjabcd(x,y) \: y \: \Wabcd(x,y) \: dy \: dx \\
-	&\quad \quad = \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b,c+d)}(x) \: \D x \Big) \nonumber \\
-	&\quad \quad \quad \quad \quad\cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
-	&\quad \quad = \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \genjacw^{(a, b,c+d+k+j+2)}(x) \: \D x \Big) \nonumber \\
-	&\quad \quad \quad \quad \quad\cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big)
-}
-\bseqn{
-	=
+	&=  \iint_\Omega \hdopnkabcd(x,y) \: \hdopmjabcd(x,y) \: y \: \Wabcd(x,y) \: dy \: dx \\
+	&= \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \rho(x)^{k+j+2} \: \genjacw^{(a, b,c+d)}(x) \: \D x \Big) \\
+		&\quadfour \cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
+	&= \Big( \int^\beta_\alpha \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \genjacmmj^{(a, b, c+d+2j+1)}(x) \: \genjacw^{(a, b,c+d+k+j+2)}(x) \: \D x \Big) \\
+		&\quadfour \cdot \: \Big( \int^\delta_{\gamma} \jac_k^{(d,c)}(t) \: \jac_j^{(d,c)}(t) \: t \: \jacw^{(d,c)}(t) \: \D t \Big) \\
+	&=
 	\begin{cases}
-    		\delta_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+3)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \quad& \text{if } j = k+1 \\
-		\gamma_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+1)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \quad& \text{if } j = k \\
-		\delta_{k-1}^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k-1)} \: \ip<\genjacnmk^{(a, b, c+d+2k-1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} \quad& \text{if } j = k-1 \\
-		0 & \text{otherwise}
+    		\delta_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+3)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{m-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
+			\quadeight \quadeight \quadeight \: \: \text{if } j = k+1 \\
+		\gamma_k^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k+1)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
+			\quadeight \quadeight \quadeight \: \: \text{if } j = k \\
+		\delta_{k-1}^{(d,c)} \: \normjac^{(d,c)} \: \normgenjac^{(a, b, c+d+2k-1)} \: \ip<\genjacnmk^{(a, b, c+d+2k-1)}, \: \rho(x)^2 \genjac_{m-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k-1)}} \\
+			\quadeight \quadeight \quadeight \: \: \text{if } j = k-1 \\
+		\quadeight \quadfour 0 \quadfour \quadeight \text{otherwise}
       	\end{cases}
@@
 Finally, if \bsrefcondition{cond:trap} holds we have that 
-$$
+\bseqn{
 	\ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{m-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} = 0 \qfor m < n-1.
-$$
+}
 If \bsrefcondition{cond:disk} holds we have that $\gamma^{(d,c)}_{k} = \gamma^{(c,c)}_{k} \equiv 0$ for any $k$.
@@
 \bseqnnumber{
-J_x^{(a,b,c,d)} \: \bighdopabcd(x,y) = x \: \bighdopabcd(x,y), \quad J_y^{(a,b,c,d)} \: \bighdopabcd(x,y) = y \: \bighdopabcd(x,y).
-\label{eqn:jacobimatricesdefinition}
+	J_x^{(a,b,c,d)} \: \bighdopabcd(x,y) &= x \: \bighdopabcd(x,y), \label{eqn:ds:jacobimatricesdefinitionx} \\
+	J_y^{(a,b,c,d)} \: \bighdopabcd(x,y) &= y \: \bighdopabcd(x,y). \label{eqn:ds:jacobimatricesdefinitiony}
 }
 The matrices $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ act on the coefficients vector of a function's expansion in the $\{\hdopnkabcd\}$ basis.
-For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopabcd(x,y)^\top \mathbf{f}$.
-Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b,c,d)\top}} \mathbf{f}$.
-In other words, ${J_x^{(a,b,c,d)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the  $\{\hdopnkabcd\}$ basis.
+For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bighdopabcd(x,y)^\top \vec{f}$.
+Then $x \: f(x,y)$ is approximated by $\bighdopabcd(x,y)^\top {J_x^{(a,b,c,d)\top}} \vec{f}$.
+In other words, ${J_x^{(a,b,c,d)\top}} \vec{f}$ is the coefficients vector for the expansion of the function $(x,y) \mapsto x \: f(x,y)$ in the $\{\hdopnkabcd\}$ basis.
 
@@
 		& & & \ddots & \ddots & \ddots \\
-	\end{pmatrix}
-}
-where the blocks themselves are diagonal for $J_x^{(a,b,c,d)}$ (sub-block-bandwidths $(0,0)$),
-\bseqn{
-A^x_n &:= \begin{pmatrix}
+	\end{pmatrix}.
+}
+For $J_x^{(a,b,c,d)}$, the blocks are diagonal (sub-block-bandwidths $(0,0)$) for $n = 0,1,2,\dots$:
+\bseqn{
+	A^x_n &:= 
+		\begin{pmatrix}
 			\alphaabcd_{n+1,0,1} & 0 & \hdots & 0 \\
@@
 			& & \alphaabcd_{n+1,n,1} & 0 \\
-	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
-B^x_n &:= \begin{pmatrix}
+		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
+	B^x_n &:= 
+		\begin{pmatrix}
 			\alphaabcd_{n,0,2} & & \\
@@
 			& & \alphaabcd_{n,n,2} \\
-	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
-C^x_n &:= \big( A^x_n \big)^\top \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
-\nonumber
-}
-and tridiagonal for $J_y^{(a,b,c,d)}$ (sub-block-bandwidths $(1,1)$),
-\bseqn{
-A^y_n &:= \begin{pmatrix}
+		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
+	C^x_n &:= \big( A^x_{n-1} \big)^\top \in \R^{(n+1)\times n}, \quad (n \ne 0)
+}
+and for $J_y^{(a,b,c,d)}$ the blocks are tridiagonal (sub-block-bandwidths $(1,1)$):
+\bseqn{
+	A^y_n &:= 
+		\begin{pmatrix}
 			\betaabcd_{n,0,8} & \betaabcd_{n,0,9} & & & \\
@@
 			& & \betaabcd_{n,n,7} & \betaabcd_{n,n,8} & \betaabcd_{n,n,9} \\
-	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
-B^y_n &:= \begin{pmatrix}
+		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
+	B^y_n &:= 
+		\begin{pmatrix}
 			\betaabcd_{n,0,5} & \betaabcd_{n,0,6} & & \\
@@
 			& & \betaabcd_{n,n,4} & \betaabcd_{n,n,5}
-	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
-C^y_n &:= \begin{pmatrix}
+		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
+	C^y_n &:= 
+		\begin{pmatrix}
 			\betaabcd_{n,0,2} & \betaabcd_{n,0,3} & & \\
@@
 			& & & \betaabcd_{n,n,1} \\
-	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
-}
-
+		\end{pmatrix} \in \R^{(n+1)\times n}, \quad (n \ne 0).
+}
 Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs, meaning that the sparsity is not limited to the specific disk-slice case.
@@
 
-We can combine each system in (\ref{eqn:jacobimatricesdefinition}) into a block-tridiagonal system:
+We can combine each system in \bsrefeqns{eqn:ds:jacobimatricesdefinitionx}{eqn:ds:jacobimatricesdefinitiony} into a block-tridiagonal system:
 \bseqn{
 	\renewcommand\arraystretch{1.3}
-\begin{pmatrix}
+	&\begin{pmatrix}
 		1 & & & \\
@@
 	\end{pmatrix}
-\bighdopabcd(x,y)
-=
+	\bighdopabcd(x,y) \\
+	&=
 	\begin{pmatrix}
-	\hdopabcd_{0,0} \\ 0 \\ 0 \\ 0 \\ \vdots  \\
-\end{pmatrix},
+		\hdopabcd_{0,0} & 0 & 0 & 0 & \hdots  \\
+	\end{pmatrix}^\top,
 }
@@
 such that $\Dnt A_n = I_{n+2}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
 \bseqn{
@@
 		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
-\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y), \quad n = 0,1,2,\dots.
-\end{cases}
-}
-
-Note that we can define an explicit \(\Dnt\) as follows:
+		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y)
+	\end{cases}.
+}
+Note that we can define an explicit $\Dnt$ as follows:
 \bseqn{
@@
 where the 1D OPs $\{\genjac^{(a, b, c)}_n\}$ are orthogonal on the interval $(\alpha, \beta)$ with respect to the weight
-$$
+\bseqn{
 	\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c}
-$$
+}
 and the 1D OPs $\{\jac^{(c, c)}_n\}$ are orthogonal on the interval $(\gamma, \delta) = (-1, 1)$ with respect to the weight
-$$
+\bseqn{
 	\jacw^{(c)}(x) := (1-x)^c \: (1+x)^c = (1-x^2)^c.
-$$
-
+}
 Denote the weighted OPs by
@@
 }
-and recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopabc(x,y)^\top \mathbf{f}$.
+and recall that a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bighdopabc(x,y)^\top \vec{f}$.
 
@@
 \bseqn{
-{\partial f \over \partial x}&= \bighdop^{(a+1,b+1,c+1)}(x,y)^\top \: D_x^{(a,b,c)} \: \mathbf{f}, \\
-{\partial f \over \partial y} &= \bighdop^{(a,b,c+1)}(x,y)^\top \: D_y^{(a,b,c)} \: \mathbf{f}, \\
-{\partial \over \partial x}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1,c-1)}(x,y)^\top \: W_x^{(a,b,c)} \: \mathbf{f}, \\
-{\partial \over \partial y}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a,b,c-1)}(x,y)^\top \: W_y^{(a,b,c)} \: \mathbf{f}.
+	{\partial f \over \partial x}&= \bighdop^{(a+1,b+1,c+1)}(x,y)^\top \: D_x^{(a,b,c)} \: \vec{f}, \\
+	{\partial f \over \partial y} &= \bighdop^{(a,b,c+1)}(x,y)^\top \: D_y^{(a,b,c)} \: \vec{f}, \\
+	{\partial \over \partial x}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a-1,b-1,c-1)}(x,y)^\top \: W_x^{(a,b,c)} \: \vec{f}, \\
+	{\partial \over \partial y}[\Wabc(x,y) \: f(x,y)] &= \bigW^{(a,b,c-1)}(x,y)^\top \: W_y^{(a,b,c)} \: \vec{f}.
 }
@@
 \bseqnnumber{
-	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:derivativeofweightgenjac} \\
-	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(c-1)}(y), \label{eqn:derivativeofweightjac} \\
-	\rho(x) \: \rho'(x) &= -x. \label{eqn:rhoderivative}
-}
-
+	\genjacw^{(a,b,c) \: \prime}(x) &= - a \: \genjacw^{(a-1,b,c)}(x) + b \: \genjacw^{(a,b-1,c)}(x) + c \: \rho(x) \: \rho'(x) \:\genjacw^{(a,b,c-2)}(x), \label{eqn:ds:derivativeofweightgenjac} \\
+	\jacw^{(c) \: \prime}(y) &= - 2c \: y \: \jacw^{(c-1)}(y), \label{eqn:ds:derivativeofweightjac} \\
+	\rho(x) \: \rho'(x) &= -x. \label{eqn:ds:rhoderivative}
+}
 We proceed with the case for the operator $D_y^{(a,b,c)}$ for partial differentiation by $y$.
@@
 }
-Now, using (\ref{eqn:derivativeofweightjac}), integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
-\bseqn{
-	\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} &= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
+Now, using \bsrefeqn{eqn:ds:derivativeofweightjac}, integration-by-parts, and noting that the weight $\jacw^{(c)}$ is a polynomial of degree $2c$ and vanishes at the limits of the integral for positive parameter $c$, we have that
+\bseqn{
+	&\normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}} \\
+	&= \int_{\gamma}^\delta \: \jac_k^{(c,c) \: \prime}(y) \: \jac_{j}^{(c+1,c+1)}(y) \: \jacw^{(c+1)}(y) \: \D y \\
 	&= - \int_{-1}^1 \: \jac_k^{(c,c)}(y) \: \ddx{y} [\jacw^{(c+1)}(y) \: \jac_{j}^{(c+1,c+1)}(y)] \: \D y \\
@@
 	& \quad - \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
-	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:dxsparsity}
-}
-We will first show that the second factor of each term in (\ref{eqn:dxsparsity}) are zero for $j < k - 2$ and also for $j = k - 1$.
+	& \quad \quad \quad \quad \quad \cdot \: \Big( \int_\gamma^\delta t \: \jac_k^{(c,c) \: \prime} \: \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} \: \D t \Big). \label{eqn:ds:dxsparsity}
+}
+We will first show that the second factor of each term in \bsrefeqn{eqn:ds:dxsparsity} are zero for $j < k - 2$ and also for $j = k - 1$.
 To this end, observe that, for any integer $c$, $\jac^{(c,c)}(-t) = (-1)^k \: \jac^{(c,c)}(t)$ and so $\jac_k^{(c,c)}$ is an even polynomial for even $k$, and an odd polynomial for odd $k$.
@@
 }
-which is zero for $j < k - 2$ by orthogonality. Thus, (\ref{eqn:dxsparsity}) is zero for $j \notin \{k-2, k\}$.
-
-Now, using (\ref{eqn:derivativeofweightgenjac}), integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
+which is zero for $j < k - 2$ by orthogonality. Thus, \bsrefeqn{eqn:ds:dxsparsity} is zero for $j \notin \{k-2, k\}$.
+
+Now, using \bsrefeqn{eqn:ds:derivativeofweightgenjac}, integration-by-parts, and noting that the weight $\genjacw^{(a,b,2c)}$ is a polynomial degree $a+b+2c$ and vanishes at the limits of the integral for positive parameters $a,b,c$, we have that
 \bseqnnumber{
@@
 	&= - \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \Big\{ \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime} \: \genjacw^{(a+1, b+1, 2c+k+j+3)} \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \nonumber \\
+		&\quadeight + (2c+k+j+3) \: \rho \: \rho' \:  \genjacw^{(a+1, b+1, 2c+k+j+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
+		&\quadeight + (b+1) \: \genjacw^{(a+1, b, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \nonumber \\
+		&\quadeight - (a+1) \: \genjacw^{(a, b+1, 2c+k+j+3)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \Big\} \: \D x \nonumber \\
 	&= - \: \normgenjac^{(a, b, 2c+2k+1)} \:  \Big\{ \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1, 1, j-k+2)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3) \: \prime}>_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-		&\quad \quad \quad + (2c+k+j+3) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-		&\quad \quad \quad + (b+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
-		&\quad \quad \quad - (a+1) \: \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:dxsparsityR}
-}
-By recalling (\ref{eqn:rhoderivative}) and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
-$$\text{deg}(\rho \: \rho' \: \genjacw^{(1,1, j-k)}) = \text{deg} (\genjacw^{(1,0, j-k+2)}) = \text{deg} (\genjacw^{(0,1, j-k+2)}) = 3 + j-k.$$ 
-Hence, by orthogonality, each term in (\ref{eqn:dxsparsityR}) is is zero for $m-j+3+j-k < n-k \iff m < n - 3$.
+		&\quad + (2c+k+j+3) \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \rho' \: \genjacw^{(1,1, j-k)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+		&\quad + (b+1) \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(1,0, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \nonumber \\
+		&\quad - (a+1) \ip<\genjacnmk^{(a, b, 2c+2k+1)}, \: \genjacw^{(0,1, j-k+2)} \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}} \Big\}. \label{eqn:ds:dxsparsityR}
+}
+By recalling \bsrefeqn{eqn:ds:rhoderivative} and noting that $j-k$ is even by the earlier argument, we can see $\rho \: \rho' \: \genjacw^{(1,1, j-k)}$, $\genjacw^{(1,0, j-k+2)}$ and $\genjacw^{(1,0, j-k+2)}$ are all polynomials, and further that
+\bseqn{
+	\deg(\rho \rho' \: \genjacw^{(1,1, j-k)}) = \deg(\genjacw^{(1,0, j-k+2)}) = \deg(\genjacw^{(0,1, j-k+2)}) = 3 + j-k.
+}
+Hence, by orthogonality, each term in \bsrefeqn{eqn:ds:dxsparsityR} is is zero for $m-j+3+j-k < n-k \iff m < n - 3$.
 
@@
 \bseqn{
-	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0 \qfor m < n - 3, \: j \notin \{k-2, k\},
-}
-showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,n-2,n-1$ and $j = k-2,k$ $(j \le m)$.
+	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0
+}
+for $m < n - 3$, $j \notin \{k-2, k\}$, showing that the only possible non-zero coefficients $c_{m,j}^x$ are when $m = n-3,n-2,n-1$ and $j = k-2,k$ $(j \le m)$.
 
@@
 \bseqnnumber{
-	\ppx{x} \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:weightderivativex} \\
-	\ppx{y} \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:weightderivativey}
+	\ppx{x} \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:ds:weightderivativex} \\
+	\ppx{y} \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:ds:weightderivativey}
 }
@@
 	Define the operator matrices 
-$$
+\bseqn{
 	T^{(a,b,c)\to(a+1,b+1,c)}, \quad T^{(a,b,c)\to(a,b,c+1)}\qqand T^{(a,b,c)\to(a+1,b+1,c+1)}
-$$ 
+}
 for conversion between non-weighted spaces, and 
-$$
+\bseqn{
 	T_W^{(a,b,c)\to(a-1,b-1,c)}, \quad T_W^{(a,b,c)\to(a,b,c-1)} \qqand T_W^{(a,b,c)\to(a-1,b-1,c-1)}
-$$ 
+}
 for conversion between weighted spaces, according to:
@@
 	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
-	&\quad = \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\
-	&\quad \quad \quad \cdot \: \normjac^{(c+\tilde{c})} \: \ip< \jac_k^{(c,c)}, \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c+\tilde{c})}} \\
-	&\quad= \normgenjac^{(a,b,2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-	&\quad \quad \quad \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
+	&= \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\
+		&\quadfour \cdot \: \normjac^{(c+\tilde{c})} \: \ip< \jac_k^{(c,c)}, \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c+\tilde{c})}} \\
+	&= \normgenjac^{(a,b,2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \genjacw^{(\tilde{a}, \tilde{b}, \: 2\tilde{c}+j-k)} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
+		&\quadfour \cdot \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(\tilde{c})} \: \jac_{j}^{(c+\tilde{c},c+\tilde{c})} >_{\jacw^{(c)}}.
 }
@@
 
-It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in (\ref{eqn:Hrecurrence}), by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
+It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in \bsrefeqn{eqn:ds:Hrecurrence}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
 We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case.
@@
 \bseqnnumber{
-	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ipgenjact}
+	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ds:ipgenjact}
 }
@@
 }
-Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in (\ref{eqn:ipgenjact}).
+Denote the four-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjact_n^{(a,b,c,d)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjact}.
 \end{definition}
@@
 \bseqnnumber{
-	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansioni} \\
-	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:genjacexpansionii}
+	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansioni} \\
+	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansionii}
 }
@@
 \bseqnnumber{
-	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:genjactnormalisationi} \\
+	\mathcal{C}_n^{(a,b,c,d)} &= \fpr(\frac{\normgenjact^{(a,b,c+1,d)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half, \label{eqn:ds:genjactnormalisationi} \\
 	\mathcal{D}_n^{(a,b,c,d)} &= (-1)^n \: \fpr(\frac{- \normgenjact^{(a,b,c,d+1)}}{\normgenjact^{(a,b,c,d)} \: \genjact_n^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \: \tilde{\beta}_n^{(a,b,c,d)}})^\half.
-\label{eqn:genjactnormalisationii}
+\label{eqn:ds:genjactnormalisationii}
 }
@@
 \bseqnnumber{
-	\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) &= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}.
-\label{eqn:christoffeldarboux}
+	&\sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(y) \: \genjact_{k}^{(a,b,c,d)}(x) \nonumber \\
+	&= \tilde{\beta}_n^{(a,b,c,d)} \: \frac{\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(y) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(y)}{y - x}.
+\label{eqn:ds:christoffeldarboux}
 }
@@
 	&\int_\alpha^\beta \: \Big(\big[\mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
-	&\quad \quad \quad \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
+		&\quadfour \cdot \big[\mathcal{C}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c+1,d)}(x) \Big) \: \D x \\
 	&= \mathcal{C}_n^{(a,b,c,d)} \: \mathcal{C}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
-	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
-	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
+		&\quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \genjactw^{(a,b,c,d)}(x) \Big\{ \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
+		&\quadeight \quadtwo \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1)\big] \Big\} \: \D x \\
 	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
@@
 }
-using (\ref{eqn:genjactnormalisationi}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansioni}) are equivalent. Further,
-\bseqn{
-	&\int_\alpha^\beta \: \Big(\big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
-	&\quad \quad \quad \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big) \: \D x \\
+using \bsrefeqn{eqn:ds:genjactnormalisationi} and \bsrefeqn{eqn:ds:christoffeldarboux}, showing that the RHS and LHS of \bsrefeqn{eqn:ds:genjacexpansioni} are equivalent. Further,
+\bseqn{
+	&\int_\alpha^\beta \Big\{ \big[\mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \\
+		&\quadfour \cdot \big[\mathcal{D}_m^{(a,b,c,d)} \: \sum_{k=0}^m \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x)\big] \: \genjactw^{(a,b,c,d+1)}(x) \Big\} \: \D x \\
 	&= - \mathcal{D}_n^{(a,b,c,d)} \: \mathcal{D}_m^{(a,b,c,d)} \: \tilde{\beta}_n^{(a,b,c,d)} \\
-	&\quad \quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \: \Big(\genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
-	&\quad \quad \quad \quad \quad \quad \quad \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \: \genjactw^{(a,b,c,d)}(x)\Big) \: \D x \\
+		&\quad \cdot \sum_{k=0}^m \: \int_\alpha^\beta \genjactw^{(a,b,c,d)}(x) \Big\{ \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \\
+		&\quadeight \cdot \big[\genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(-1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(-1)\big] \Big\} \: \D x \\
 	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
@@
 }
-using (\ref{eqn:genjactnormalisationii}) and (\ref{eqn:christoffeldarboux}), showing that the RHS and LHS of (\ref{eqn:genjacexpansionii}) are also equivalent. 
+using \bsrefeqn{eqn:ds:genjactnormalisationii} and \bsrefeqn{eqn:ds:christoffeldarboux}, showing that the RHS and LHS of \bsrefeqn{eqn:ds:genjacexpansionii} are also equivalent. 
 \end{proof}
@@
 \begin{proof}
-First, using (\ref{eqn:genjacexpansioni}) and (\ref{eqn:christoffeldarboux}) we have that
+First, using \bsrefeqn{eqn:ds:genjacexpansioni} and \bsrefeqn{eqn:ds:christoffeldarboux} we have that
 \bseqnnumber{
@@
 	&= \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \nonumber \\
-	& \quad \quad \cdot \Big[\Big(\tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)\Big) \: \genjact_{n+1}^{(a,b,c,d)}(1) \nonumber \\
-	& \quad \quad \quad - \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \: \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\beta}_{n}^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(x)\Big) \: \genjact_{n}^{(a,b,c,d)}(1) \Big] \label{eqn:genjactproofa}
+		&\quad \cdot \Big\{  \genjact_{n+1}^{(a,b,c,d)}(1) \: \Big(\tilde{\beta}_{n}^{(a,b,c,d)} \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c,d)} \genjact_{n}^{(a,b,c,d)}(x) \nonumber \\
+			&\quadeight \quad + \tilde{\beta}_{n-1}^{(a,b,c,d)} \genjact_{n-1}^{(a,b,c,d)}(x)\Big) \nonumber \\
+		&\quadfour - \genjact_{n}^{(a,b,c,d)}(1) \: \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \genjact_{n+1}^{(a,b,c,d)}(x) \nonumber \\
+			&\quadeight \quadfour + \tilde{\beta}_{n}^{(a,b,c,d)} \genjact_{n}^{(a,b,c,d)}(x)\Big) \Big\} \label{eqn:ds:genjactproofa}
 }
@@
 	&(1-x) \: x \: \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
-	&= (1-x) \: \Big[\tilde{\beta}_{n}^{(a,b,c+1,d)} \: \genjact_{n+1}^{(a,b,c+1,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \genjact_{n}^{(a,b,c+1,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \genjact_{n-1}^{(a,b,c+1,d)}(x)\Big] \nonumber \\
-	&= \mathcal{C}_{n+1}^{(a,b,c,d)} \: \tilde{\beta}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n+1}^{(a,b,c,d)} \: \Big( \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n+2}^{(a,b,c,d)}(1) - \genjact_{n+2}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) \Big) \nonumber \\
-	& \quad \quad + \mathcal{C}_n^{(a,b,c,d)} \: \tilde{\alpha}_{n}^{(a,b,c+1,d)} \: \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
-	& \quad \quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \: \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \: \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \: \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \: \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:genjactproofb}
+	&= (1-x) \: \Big\{ \tilde{\beta}_{n}^{(a,b,c+1,d)} \genjact_{n+1}^{(a,b,c+1,d)}(x) + \tilde{\alpha}_{n}^{(a,b,c+1,d)} \genjact_{n}^{(a,b,c+1,d)}(x) \nonumber \\
+		&\quadeight + \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \genjact_{n-1}^{(a,b,c+1,d)}(x) \Big\} \nonumber \\
+	&= \mathcal{C}_{n+1}^{(a,b,c,d)} \tilde{\beta}_{n}^{(a,b,c+1,d)} \tilde{\beta}_{n+1}^{(a,b,c,d)} \: \Big( \genjact_{n+1}^{(a,b,c,d)}(x) \genjact_{n+2}^{(a,b,c,d)}(1) - \genjact_{n+2}^{(a,b,c,d)}(x) \genjact_{n+1}^{(a,b,c,d)}(1) \Big) \nonumber \\
+		&\quad + \mathcal{C}_n^{(a,b,c,d)} \tilde{\alpha}_{n}^{(a,b,c+1,d)} \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
+		&\quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:ds:genjactproofb}
 }
 We can set $\tilde{\beta}_{-1}^{(a,b,c+1,d)} = 0$.
-By comparing coefficients of $\genjact_{n+2}^{(a,b,c,d)}(x)$ and $\genjact_{n+1}^{(a,b,c,d)}(x)$ in both (\ref{eqn:genjactproofa}) and (\ref{eqn:genjactproofb}) we obtain the desired recurrence coefficients for the OP $\genjact_{n}^{(a,b,c+1,d)}(x)$.
+By comparing coefficients of $\genjact_{n+2}^{(a,b,c,d)}(x)$ and $\genjact_{n+1}^{(a,b,c,d)}(x)$ in both \bsrefeqn{eqn:ds:genjactproofa} and \bsrefeqn{eqn:ds:genjactproofb} we obtain the desired recurrence coefficients for the OP $\genjact_{n}^{(a,b,c+1,d)}(x)$.
 The recurrence coefficients for the OPs $\genjact_{n}^{(a,b,c,d+1)}(x)$ are found similarly.
@@
 \bseqn{
-	\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
-	&= \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \\
-	&\approx \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
-	&\approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
-	&= \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
+	&\iint_\Omega f_e(x,y) \: \Wabc(x,y) \: \D y \: \D x \\
+	&\quadfour = \int_\alpha^\beta \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
+	&\quadfour = \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_e\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \\
+	&\quadfour \approx \int_\alpha^\beta \genjacw^{(a,b,2c+1)}(s) \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s,\rho(s) t_i\big) \Big) \: \D s \quad (\star) \\
+	&\quadfour \approx \sum_{j=1}^{M_1} \Bigg( w_j^{(s)} \: \sum_{i=1}^{M_2} \Big( w_i^{(t)} f_e\big(s_j,\rho(s_j) t_i\big) \Big) \Bigg) \quad (\star \star) \\
+	&\quadfour = \sum_{k=1}^{M_1 M_2}  w_k \: f_e(x_k, y_k).
 }
@@
 \bseqn{
-	\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x &= \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
-	&= \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
-	&= 0
+	&\iint_\Omega f_o(x,y) \: \Wabc(x,y) \: \D y \: \D x \\
+	&\quadfour = \int_\alpha^\beta \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: V^{(a,b,c)}(s,t) \: \rho(s) \: \D t \: \D s \\
+	&\quadfour = \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
+	&\quadfour = 0
 }
@@
 
-Fix \(a,b,c \in \R\). Then for any function \(f : \Omega \to \R\) we can express \(f\) by
-\bseqn{
-	f(x,y) \approx \sum_{n=0}^N \bighdop_n^{(a,b,c)}(x,y)^\top \: \mathbf{f}_n
+Fix $a,b,c \in \R$. Then for any function $f : \Omega \to \R$ we can express $f$ by
+\bseqn{
+	f(x,y) \approx \sum_{n=0}^N \bighdop_n^{(a,b,c)}(x,y)^\top \: \vec{f}_n
 }
@@
 		\hdop^{(a,b,c)}_{n,n}(x,y)
-	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N,
-}
-and where
-\bseqn{
-	\mathbf{f}_n &:= \begin{pmatrix}
+	\end{pmatrix} \in \R^{n+1}
+}
+for all $n = 0,1,2,\dots,N$, and where
+\bseqn{
+	\vec{f}_n &:= \begin{pmatrix}
 		f_{n,0} \\
@@
 }
-Recall from (\ref{eqn:normhdop}) that $\norm{\hdopnkabc}^2_\Wabc = \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}$.
+Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopnkabc}^2_\Wabc = \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}$.
 Using the quadrature rule detailed in Section 4.2 for the inner product, we can calculate the coefficients $f_{n,k}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$:
 \bseqn{
-	f_{n,k} &= \frac{1}{2 \: \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}} \: \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) \: \hdopnkabc(x_j, y_j) +f(x_j, -y_j) \: \hdopnkabc(x_j, -y_j) \big]
-}
-where $M = \ceil{\half(N+1)}^2$.
+	f_{n,k} &= \lambda^{(a,b,c)}_k \: \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) \: \hdopnkabc(x_j, y_j) +f(x_j, -y_j) \: \hdopnkabc(x_j, -y_j) \big]
+}
+where $\lambda^{(a,b,c)}_k := {1 \over 2 \: \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}}$ and $M = \ceil{\half(N+1)}^2$.
 
@@
 	\end{cases}.
-	\label{eqn:poisson}
+	\label{eqn:ds:poisson}
 }
@@
 We can tackle the problem as follows.
-Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above.
+Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\vec{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\vec{f}$.
+Since $f$ is known, we can obtain $\vec{f}$ using the quadrature rule above.
 In matrix-vector notation, our system hence becomes:
 \bseqn{
-    \laplacewiii \mathbf{u} = \mathbf{f}
-}
-which can be solved to find $\mathbf{u}$.
-In \bsreffig{fig:ds:poisson} we see the solution to the Poisson equation with zero boundary conditions given in (\ref{eqn:poisson}) in the disk-slice $\Omega$.
+    \laplacewiii \vec{u} = \vec{f}
+}
+which can be solved to find $\vec{u}$.
+In \bsreffig{fig:ds:poisson} we see the solution to the Poisson equation with zero boundary conditions given in \bsrefeqn{eqn:ds:poisson} in the disk-slice $\Omega$.
 In \bsreffig{fig:ds:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 990$, that is, $491,536$ unknowns.
@@
 In \bsreffig{fig:ds:poissonexact} we see an example where the solution calculated to the Poisson equation is shown together with a plot of the exact solution and the error.
-The example was chosen so that the exact solution was $u(x,y) = \Wiii (x,y) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y) = \Delta \big[\Wiii (x,y) y^3 \exp(x)\big]$.
 We see that the computed solution is almost exact.
+The example was chosen so that the exact solution was $u(x,y) = \Wiii (x,y) y^3 \exp(x)$, and thus the RHS function $f$ would be $f(x,y) = \Delta \big[ \Wiii(x,y) \: y^3 \: \exp(x) \big]$.
 
@@
 	\end{cases}.
-	\label{eqn:helmholtz}
+	\label{eqn:ds:helmholtz}
 }
@@
 We can tackle the problem as follows.
-Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule above.
+Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\vec{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\vec{f}$.
+Since $f$ is known, we can obtain  the coefficients $\vec{f}$ using the quadrature rule above.
 We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
@@
 \bseqn{
-    (\laplacewiii + k^2 T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}) \mathbf{u} = \mathbf{f}
-}
-which can be solved to find $\mathbf{u}$.
+    (\laplacewiii + k^2 T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}) \vec{u} = \vec{f}
+}
+which can be solved to find $\vec{u}$.
 We can see the sparsity and structure of this matrix system in \bsreffig{fig:ds:sparsity} with $v(x,y) = xy^2$ as an example.
-In \bsreffig{fig:ds:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
+In \bsreffig{fig:ds:helmholtz} we see the solution to the inhomogeneous variable-coefficient Helmholtz equation with zero boundary conditions given in \bsrefeqn{eqn:ds:helmholtz} in the half-disk $\Omega$, with $k=100$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $f(x,y) = x(1-x^2-y^2)e^x$.
 In \bsreffig{fig:ds:helmholtz} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogeneous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ using $N = 500$, that is, $125,751$ unknowns.
@@
 	\end{cases}.
-	\label{eqn:biharmonic}
+	\label{eqn:ds:biharmonic}
 }
 where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-In \bsreffig{fig:ds:biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the disk-slice $\Omega$.
+In \bsreffig{fig:ds:biharmonic} we see the solution to the Biharmonic \bsrefeqn{eqn:ds:biharmonic} in the disk-slice $\Omega$.
 In \bsreffig{fig:ds:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the biharmonic equation with $N = 500$, that is, $125,751$ unknowns.
@@
 \bseqn{
-	L(v) := \int_\Omega f \: v \: d\mathbf{x} = \int_\Omega \nabla u \cdot \nabla v \: d\mathbf{x} =: a(u,v).
+	L(v) := \int_\Omega f \: v \: \D \vec{x} = \int_\Omega \nabla u \cdot \nabla v \: \D \vec{x} =: a(u,v).
 }
@@
 However, here we simply consider our domain to be a disk-slice and our discretisation to be a single element -- that is we let $\element = \Omega$ for a disk-slice domain.
-We can choose our finite dimensional space $V_p = \{v_p \in V \quad | \quad {\rm deg}\,(v_p|_\element) \le p\}$ for some $p \in \N$.
+We can choose our finite dimensional space $V_p = \bsset{v_p \in V}{\deg(v_p|_\element) \le p}$ for some $p \in \N$.
 
@@
 	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-	\label{eqn:FEMweakform}
+	\label{eqn:ds:FEMweakform}
 }
@@
 Note that due to orthogonality this is a diagonal matrix.
-We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = 1$:
-\bseqn{
-\bigWiii(x,y) &:= \begin{pmatrix}
-		\bigWiii_0(x,y) \\
-		\bigWiii_1(x,y) \\
-		\bigWiii_2(x,y) \\
-		\vdots \\
-		\bigWiii_p(x,y)
-	\end{pmatrix}, \\
-\bigWiii_n(x,y) &:=  \Wiii(x,y) \begin{pmatrix}
-						\hdopiii_{n,0}(x,y) \\
-						\vdots \\
-						\hdopiii_{n,n}(x,y)
-					\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,p,
-}
-and rewrite (\ref{eqn:FEMweakform}) in matrix form:
-\bseqn{
-	a(u_p,v_p) &= \int_\element \nabla u_p \cdot \nabla v_p \: \D \mathbf{x} \\
+We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameters $a = b = c = 1$
+\bseqn{
+	\bigWiii(x,y) = \Wiii(x,y) \: \bighdopiii(x,y)
+}
+and rewrite \bsrefeqn{eqn:ds:FEMweakform} in matrix form:
+\bseqn{
+	&a(u_p,v_p) \\
 	&= \int_\element \begin{pmatrix}
@@
 					\partial_y u_p
-				\end{pmatrix} \: \D \mathbf{x}
+				\end{pmatrix} \: \D \vec{x}
 				\\
 	&= \int_\element \begin{pmatrix}
-					\bighdopooo^\top \Wiii_x \mathbf{v} \\
-					\bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \mathbf{v}
+					\bighdopooo^\top \Wiii_x \vec{v} \\
+					\bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \vec{v}
 				\end{pmatrix}^\top 
 				\begin{pmatrix}
-					\bighdopooo^\top \Wiii_x \mathbf{u} \\
-					\bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \mathbf{u}
-				\end{pmatrix} \: \D \mathbf{x}
+					\bighdopooo^\top \Wiii_x \vec{u} \\
+					\bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \vec{u}
+				\end{pmatrix} \: \D \vec{x}
 				\\
-	&= \int_\element \Big( \mathbf{v}^\top {\Wiii_x}^\top \bighdopooo \bighdopooo^\top \Wiii_x \mathbf{u} \nonumber \\
-					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \bighdopooo \bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \mathbf{u}  \Big) \: \D \mathbf{x} \\
-	&= \mathbf{v}^\top \: \Big( {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x \\
-	& \quad \quad \quad \quad + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \Big) \: \mathbf{u}
-}
-where $\mathbf{u}, \mathbf{v}$ are the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs), and
-\bseqn{
-	L(v_p) &= \int_\element \: v_p \: f \: \D \mathbf{x}\\
-	&= \int_\element \: \mathbf{v}^\top \: \bigWiii \: \bighdopiii^\top \: \mathbf{f} \: \D \mathbf{x} \\
-	&= \mathbf{v}^\top \: \ip< \bighdopiii, {\bighdopiii}^\top >_{\Wiii} \: \D \mathbf{x} \\
-	&= \mathbf{v}^\top \Lambda^{(1,1,1)} \: \mathbf{f},
-}
-where $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bighdopiii$ OP basis.
-
-Since (\ref{eqn:FEMweakform}) is equivalent to stating that
+	&= \int_\element \Big( \vec{v}^\top {\Wiii_x}^\top \bighdopooo \bighdopooo^\top \Wiii_x \vec{u} \nonumber \\
+					& \quad \quad \quad + \vec{v}^\top ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \bighdopooo \bighdopooo^\top T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \vec{u}  \Big) \: \D \vec{x} \\
+	&= \vec{v}^\top \: \Big( {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x \\
+	& \quad \quad \quad \quad + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \Big) \: \vec{u}
+}
+where $\vec{u}, \vec{v}$ are the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs), and
+\bseqn{
+	L(v_p) &= \int_\element \: v_p \: f \: \D \vec{x}\\
+	&= \int_\element \: \vec{v}^\top \: \bigWiii \: \bighdopiii^\top \: \vec{f} \: \D \vec{x} \\
+	&= \vec{v}^\top \: \ip< \bighdopiii, {\bighdopiii}^\top >_{\Wiii} \: \D \vec{x} \\
+	&= \vec{v}^\top \Lambda^{(1,1,1)} \: \vec{f},
+}
+where $\vec{f}$ is the coefficient vector for the expansion of the function $f(x,y)$ in the $\bighdopiii$ OP basis.
+
+Since \bsrefeqn{eqn:ds:FEMweakform} is equivalent to stating that
 \bseqn{
@@
 \bseqn{
-	A\mathbf{u} = \tilde{\mathbf{f}}.
+	A\vec{u} = \tilde{\vec{f}}.
 }
@@
 }
-and the load vector $\tilde{\mathbf{f}}$ is given by 
-\bseqn{
-	\tilde{\mathbf{f}} = \Lambda^{(1,1,1)} \: \mathbf{f}.
+and the load vector $\tilde{\vec{f}}$ is given by 
+\bseqn{
+	\tilde{\vec{f}} = \Lambda^{(1,1,1)} \: \vec{f}.
 }

### 2021-06-04 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Further edits to help make sure the thesis flows ok and legacy paper specific language is removed from ch3 and ch4.

@@
-\chapter{Spherical harmonics as orthogonal polynomials}
+\chapter{Spherical harmonics as orthogonal polynomials}\label{CHAPTER:sphericalharmonics}
 
@@
 While it may seem somewhat odd, it should hopefully become apparent why this is a useful way to define our domain.
+
+It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$).
+Throughout this thesis we will use the convention that the spherical coordinate angles be defined by
+\bseqn{
+	x &= \sinphi \costheta = \rho(z) \costheta \\
+	y &= \sinphi \sintheta = \rho(z) \sintheta \\
+	z &= \cosphi
+}
+where $\rho(z) := \sqrt{1-z^2}$.
 
@@
 We will finally demonstrate a proof-of-concept via the examples of the heat equation and the linearised shallow water equations on the unit sphere.
+These examples consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): written in spherical coordinates this is
+\bseqnnumber{
+	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo \label{eqn:laplacebeltrami}
+}
+i.e. $\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for some function $f$ where $\xvec := (x,y,z) \in \R^3$.
 
@@
 Let $(x,y,z) \in \Omega$.
-It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$).
-Throughout this document we will use the convention that the spherical coordinate angles be defined by
-\bseqn{
-	x &= \sinphi \costheta = \rho(z) \costheta \\
-	y &= \sinphi \sintheta = \rho(z) \sintheta \\
-	z &= \cosphi.
-}
-where $\rho(z) := \sqrt{1-z^2}$. We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
+We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
 \bseqnnumber{
@@
 
-We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}. The \textit{Clenshaw algorithm} is then as follows:
+A multivariate analogue of Clenshaw's recursive algorithm for evaluation of a function expanded in an OP basis has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle}, and we proceed similarly here.
+We can use the Clenshaw algorithm to evaluate this function at a given point $(x,y,z)$ on $\Omega$ \cite{press2007numerical}.
+The multidimensional \textit{Clenshaw algorithm} is as follows:
 \bseqn{
@@
 %%% TANGENT BUNDLE ----- VECTOR SHs
-\section{Vector spherical harmonics as orthogonal vectors in three variables}\label{Section:VSHs}
+\section{Vector spherical harmonics as orthogonal vectors in three variables}\label{section:VSHs}
 
@@
 
-The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplacian, as well as other key operators useful in examples such as the linearised shallow water equations.
+The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other key operators useful in examples such as the linearised shallow water equations.
 These operators will take coefficients of a function's expansion in either the SSH or VSH basis, to coefficients in either basis.
@@
 \begin{definition}\label{def:SHoperators}
-	We define the operator matrices $\calD, \tilde \calD, \calG, \tilde \calG \calL$ according to:
+	We define the operator matrices $\calD, \tilde \calD, \calG, \tilde \calG, \calL$ according to:
 \bseqn{
@@
 \end{remark}
+
+For clarity, we note that $\DeltaS$ is defined in \bsrefeqn{eqn:laplacebeltrami}, and the spherical gradient and divergence operators are given by
+\bseqn{
+	 \gradS &= \ppx{\varphi} + {1 \over \sinphi} \ppx{\theta}, \\
+	 \divergence \uvec &= {1 \over \sinphi} \ppx{\varphi}\big( \sinphi u_\varphi \big) + {1 \over \sinphi} \ppx{\theta} u_\theta,
+}
+where $\uvec \equiv \phivec \: u_\varphi + \thetavec \: u_\theta$.
 

### 2021-06-04 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Further edits to help make sure the thesis flows ok and legacy paper specific language is removed from ch3 and ch4.

@@
 The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper.
-These adjustments make the construction more involved, which is why they are omitted here, but the approach is essentially the same.
+These adjustments make the construction more involved, but the approach is essentially the same which is why they are omitted here.
 \end{remark}
@@
 
-We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
-\bseqn{
-	z &= \cosphi, \\
-	x &= \sinphi \costheta = \rho(z) \costheta, \\
-	y &= \sinphi \sintheta = \rho(z) \sintheta.
-}
-where $ \rho(z) := \sqrt{1-z^2}$, we have
+Just as for the spherical harmonics framework for the whole sphere in \bsrefchapter{CHAPTER:sphericalharmonics}, we consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator).
+Recall that we write the spherical coordinates as
+\bseqn{
+	x &= \sinphi \costheta = \rho(z) \costheta \\
+	y &= \sinphi \sintheta = \rho(z) \sintheta \\
+	z &= \cosphi
+}
+where $ \rho(z) := \sqrt{1-z^2}$, and that the Laplace--Beltrami operator is then
 \bseqn{
@@
 We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}.
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method we investigated in \bsrefchapter{CHAPTER:diskslice} for disk-slices and trapeziums (see also \cite{snowball2019sparse}), as well as for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
 Just as we proceeded in \bsrefchapter{CHAPTER:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.
@@
 }
-\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:sc:construction}
+\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:sc:OPconstruction}
 	Let $w : (\alpha,\beta) \to \R$ be a weight function.
@@
 
-For the spherical cap, we can use \bsrefprop{prop:sc:construction} to create our one-parameter family of OPs.
+For the spherical cap, we can use \bsrefprop{prop:sc:OPconstruction} to create our one-parameter family of OPs.
 The univariate OPs that we will choose for the $r_{n,k}$ polynomials above will be the non-classical $\genjac^{(a,b)}$ OPs that we defined in \bsrefdef{def:ds:OPconstruction}.
@@
 We can now define the 3D OPs for the spherical cap.
-\begin{definition}\label{def:sc:constuction}
+\begin{definition}\label{def:sc:OPconstruction}
 	Define the one-parameter 3D orthogonal polynomials via:
@@
 However, it is useful to simply label them as such.
-In fact, note that they are in fact \textit{banded-block-banded matrices}:
-
-\begin{definition}
-	A block matrix $A$ with blocks $A_{i,j}$ has block-bandwidths $(L,U)$ if $A_{i,j} = 0$ for $- L \leq j-i \leq U$, and sub-block-bandwidths $(\lambda, \mu)$ if all blocks $A_{i,j}$ are banded with bandwidths $(\lambda,\mu)$.
-A matrix where the block-bandwidths and sub-block-bandwidths are small compared to the dimensions is referred to as a banded-block-banded matrix.
-\end{definition}
-
-Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
+They are once again in fact banded-block-banded matrices -- see \bsrefdef{def:bandedblockbanded}.
+More specifically, each Jacobi matrix is block-tridiagonal (block-bandwidths $(1,1)$).
+
+For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
 \bseqn{
@@
 }
+
 For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
@@
 }
+
 For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
@@
 
-A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here.
+Just as we saw for the spherical harmonics and the disk-slice, one can obtain a recursive method of calculating point evaluations for the OPs by gaining a multidimensional three-term recurrence relation.
 Combining each system in \bsrefeqn{eqn:sc:jacobimatricesdefinitionalt} into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yields:
@@
 			& & 0 & & & 0 & & & (\Gamma^{(a)}_{n,n,3})^{-1}  \\
-			& & & & & \bm{\eta}^\top_{0} & & & \\
-			& & & & & \bm{\eta}^\top_{1} & & &
+			& & & & & \vec{\eta}^\top_{0} & & & \\
+			& & & & & \vec{\eta}^\top_{1} & & &
 		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
 }
-for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in \bsrefeqn{eqn:sc:jacobisubblocksGamma1} and \bsrefeqn{eqn:sc:jacobisubblocksGamma2} for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by
-\bseqn{
-	\big(\bm{\eta}_{0}\big)_j &= 
+for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in \bsrefeqn{eqn:sc:jacobisubblocksGamma1} and \bsrefeqn{eqn:sc:jacobisubblocksGamma2} for $k=0,\dots,n$, and where $\vec{\eta}_{0}, \vec{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by
+\bseqn{
+	\big(\vec{\eta}_{0}\big)_j &= 
 		\begin{cases}
@@
 		\end{cases} \\
-	\big(\bm{\eta}_{1}\big)_j &= 
+	\big(\vec{\eta}_{1}\big)_j &= 
 		\begin{cases}
@@
 \bseqn{
-	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \bm{f}_k = \bigscopNa(x,y,z)^\top \: \bm{f}
+	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \vec{f}_k = \bigscopNa(x,y,z)^\top \: \vec{f}
 }
@@
 \bseqn{
-	\bm{f}_k &:= 
+	\vec{f}_k &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2(N-k+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	\bm{f}_0 := 
+	\vec{f}_0 := 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{N+1}, \\
-	\bm{f} &:= 
-		\begin{pmatrix}
-			\bm{f}_0 \\
+	\vec{f} &:= 
+		\begin{pmatrix}
+			\vec{f}_0 \\
 			\vdots \\
-			\bm{f}_N
+			\vec{f}_N
 		\end{pmatrix} \in \R^{2(N+1)^2}, \quad \quad
@@
 
-For a function $f$, with coefficients vector $\bm{f}$ for expansion in the $\{\scopnki\}$ basis as determined via the method in \bsrefsection{section:sc:expandingfunctions} up to order $N$, we can use the Clenshaw algorithm to evaluate the function at a point $(x,y,z) \in \Omega$ as follows.
-Let $A_n, B_n, \Dnt, C_n$ be the Clenshaw matrices from \bsrefdef{def:sc:clenshawmats}, and define the rearranged coefficients vector $\tilde{\bm{f}}$ via
-\bseqn{
-	{\bm{f}}_n &:= 
+For a function $f$, with coefficients vector $\vec{f}$ for expansion in the $\{\scopnki\}$ basis as determined via the method in \bsrefsection{section:sc:expandingfunctions} up to order $N$, we can use the Clenshaw algorithm to evaluate the function at a point $(x,y,z) \in \Omega$ as follows, as we have seen for the spherical harmonics and disk-slice in \bsrefchapter{CHAPTER:sphericalharmonics} and \bsrefchapter{CHAPTER:diskslice} respectively.
+Let $A_n, B_n, \Dnt, C_n$ be the Clenshaw matrices from \bsrefdef{def:sc:clenshawmats}, and define the rearranged coefficients vector $\tilde{\vec{f}}$ via
+\bseqn{
+	{\vec{f}}_n &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2(N+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	{\bm{f}}_0 = f_{0,0,0} \in \R, \\
-	{\bm{f}} &:= 
-		\begin{pmatrix}
-			{\bm{f}}_0 \\
+	{\vec{f}}_0 = f_{0,0,0} \in \R, \\
+	{\vec{f}} &:= 
+		\begin{pmatrix}
+			{\vec{f}}_0 \\
 			\vdots \\
-			{\bm{f}}_N
+			{\vec{f}}_N
 		\end{pmatrix} \in \R^{(N+1)^2}.
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {{\bm{f}}_n}^T - \bm{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {{\vec{f}}_n}^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
 }
@@
 \bseqn{
-	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \bm{f} \: \bigscopta(x,y,z) = (V \bm{f})^\top \: \bigscopta(x,y,z),
-}
-i.e. $V \bm{f}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$.
-
-To this end, let $\tilde{\bm{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in \bsrefsection{section:sc:functionevaluation} so that $v(x,y,z) = {\bm{v}}^\top \: \bigscopta(x,y,z)$).
+	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \vec{f} \: \bigscopta(x,y,z) = (V \vec{f})^\top \: \bigscopta(x,y,z),
+}
+i.e. $V \vec{f}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$.
+
+To this end, let $\tilde{\vec{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in \bsrefsection{section:sc:functionevaluation} so that $v(x,y,z) = {\vec{v}}^\top \: \bigscopta(x,y,z)$).
 Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, $Z := (J_z^{(a)})^\top$.
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \bm{\xi}_{N+2} = \bold{0}, \: \bm{\xi}_{N+2} = \bold{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \bm{\xi}_{n}^T = {\tilde{\bm{v}}_n}^T - \bm{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \bm{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V = v(X, Y, Z) \approx \bm{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
-}
-where at each iteration, $\bm{\xi}_n$ is a vector of matrices.
+	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {\tilde{\vec{v}}_n}^T - \vec{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad &\text{3) } \text{Output: } V = v(X, Y, Z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+}
+where at each iteration, $\vec{\xi}_n$ is a vector of matrices.
 
@@
 What about other boundary conditions?
-One simple extension is the case where the value on the boundary takes that of a function depending only on $x$ and $y$, i.e.
-$c = c(x,y)$.
-In this case, the problem
-\bseqn{
-	\begin{cases}
-    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = c(x,y) &\quad \text{on } \partial \Omega
-	\end{cases}
-}
-is equivalent to letting $u(x,y,z) = \tilde{u}(x,y,z)+ c(x,y)$ and solving
-\bseqn{
-	\begin{cases}
-    		\DeltaS \tilde{u}(x,y,z) + k^2 \; v(x,y,z) \tilde{u}(x,y,z) = g(x,y,z) \quad \text{in } \Omega \\
-		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
-	\end{cases}
-}
-for $\tilde u$, where 
-\bseqn{
-	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)
-}
-for $(x,y,z) \in \Omega$.
-This new problem is then a zero boundary condition Helmholtz problem with right hand side $g$.
-Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
-\bseqn{
-	\DeltaS \: c(x,y) &= {1 \over \rho(z)^2} \pmpxm{\theta}{2} c(x,y) \\
-	&= {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) \\
-	&= - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
-}
-since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only, which are precisely the Fourier coefficients of $c(\cos \theta, \sin \theta)$.
-Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manner.
 
@@
 
-Other boundary conditions beyond those already presented here are beyond the scope of this paper.
+One simple extension is the case where the value on the boundary takes that of a function depending only on $x$ and $y$, i.e. $c = c(x,y)$. In this case, the Helmholtz problem
+\bseqn{
+	\begin{cases}
+    		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
+		u(x,y,z) = c(x,y) &\quad \text{on } \partial \Omega
+	\end{cases}
+}
+is equivalent to letting $u(x,y,z) = \tilde{u}(x,y,z)+ c(x,y)$ and solving
+\bseqn{
+	\begin{cases}
+    		\DeltaS \tilde{u}(x,y,z) + k^2 \; v(x,y,z) \tilde{u}(x,y,z) = g(x,y,z) \quad \text{in } \Omega \\
+		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
+	\end{cases}
+}
+for $\tilde u$, where 
+\bseqn{
+	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)
+}
+for $(x,y,z) \in \Omega$.
+This new problem is then a zero boundary condition Helmholtz problem with right hand side $g$.
+Notice that the spherical Laplacian applied to $c(x,y)$, expanded in the $\bigscopN^{(1)}$ basis with coefficients vector $\vec c = (c_{n,k,i})$, is just
+\bseqn{
+	\DeltaS \: c(x,y) &= {1 \over \rho(z)^2} \pmpxm{\theta}{2} c(x,y) \\
+	&= {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) \\
+	&= - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
+}
+since the coefficients $\{c_{n,k,i}\}$ for such a function are zero for $k < n$ due to the dependence on $x$ and $y$ only, which are precisely the Fourier coefficients of $c(\cos \theta, \sin \theta)$.
+Thus, since the function $c(x,y)$ is known, it is simple to evaluate $\pmpxm{\theta}{2} c(x,y)$ and hence one can obtain the coefficients for the expansion of $g(x,y,z)$ in the $\bigscopN^{(1)}$ basis in the usual manner.
+
+Other boundary conditions beyond those already presented here are beyond the scope of this thesis.
 However, the approach established for the case of a triangle domain in 2D shows how Neumann conditions can be considered  as a system of equations in partial derivatives involving Dirchlet conditions \cite{olver2019triangle}.
 This approach may also work here for the spherical cap.
-Ideally, one would properly tackle the tangent space (which again is beyond the scope of this paper) which would lower the degrees of freedom.
+Ideally, one would properly tackle the tangent bundle (that is defined in \bsrefdef{def:tangentbundle}) which would lower the degrees of freedom, but this is future work -- see \bsrefappendix{appendix:tangent}.
 
@@
 This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere such as spherical triangles.
-It is clear from the construction in this paper that discretizations of spherical gradients and Laplacian's are sparse on other suitable sub-components of the sphere.
+It is clear from the construction in this paper that discretizations spherical Laplacian's  and other partial differential operators are sparse on other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).
-Constructing these sparse spectral methods for surface PDEs on  spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct, or perhaps {\it compute}, the necessary orthogonal polynomials.
-
-The next stage is to develop an orthogonal basis for the tangent space of the spherical cap (or band), and obtain sparse differential operators for gradient, divergence etc.
+Constructing these sparse spectral methods for surface PDEs on spherical triangles is future work, and has applications in weather prediction \cite{staniforth2012horizontal}, though it is not yet clear how to directly construct, or perhaps \textit{compute}, the necessary orthogonal polynomials.
+
+The next stage is to develop an orthogonal basis for the tangent bundle of the spherical cap (or band), and obtain sparse differential operators for the spherical gradient, divergence etc.
 On the complete sphere, the vector spherical harmonics that form the orthogonal basis are simply the gradients and perpendicular gradients of the scalar spherical harmonics \cite{barrera1985vector} which has been used effectively for solving PDEs on the sphere \cite{vasil2019tensor,lecoanet2019tensor} -- however, we do not have that luxury for the spherical cap or band, and hence the construction of a basis will not be as straightforward.

### 2021-06-07 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
New images for SH plots (removing shading/shadowing)

@@
 	\centering % <-- added
-	\centerline{\includegraphics[scale=0.5]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100}}
+	\centerline{\includegraphics[scale=0.5]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100-2}}
 	\caption{Four snapshots of the solution $u$ of the heat equation with $\alpha = 1/42$ and a step size of $\Delta t = 0.01$, solved using the BDF2 timestepping method.
@@
 	\centering % <-- added
-	\centerline{\includegraphics[scale=0.5]{linear-swe-H=1-dt=0p01-nsteps=100-arrows}}
+	\centerline{\includegraphics[scale=0.5]{linear-swe-H=1-dt=0p01-nsteps=100-arrows-2}}
 	\caption{Four snapshots of the solution for the height $h$ (surface colour) and the wave velocity $u$ (red arrows) of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method (BDF1).

### 2021-06-08 papers/thesis/full/abstract/abstract.tex
Addressed most comments so far.

@@
 
-We move on working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+We move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.

### 2021-06-08 papers/thesis/full/acknowledgements/acknowledgements.tex
Pushing the pdf of thesis
Addressed most comments so far.

@@
 I would like to thank the MPE CDT staff for any knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for his guidance, understanding and general enthusiasm (it meant a lot).
-And to all my fellow 'Charlies' -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.
+And to all my fellow \enquote{Charlies} -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.
 
-I thank my family and friends for all their support and words of encouragement -- it all helped a lot. An extra thanks to all my undergrad mates for many well needed fun times too.
+I thank my family and friends for all their support and words of encouragement -- it all helped a lot.
 
-A special thanks goes to my partner Jessica for putting up with me and helping me when things got tough during my 4 years at Imperial College London.
+A special thanks goes to my partner Jess for putting up with me and helping me when things got tough during my 4 years at Imperial College London.
 
@@
 
-I would like to express (whatever feelings I have) to:
-
-\begin{itemize}
- \item My supervisor
- \vspace*{3mm}
- \item My second supervisor
- \vspace*{3mm}
- \item Other researchers
- \vspace*{3mm}
- \item My family and friends
-\end{itemize}
+%I would like to express (whatever feelings I have) to:
+%
+%\begin{itemize}
+% \item My supervisor
+% \vspace*{3mm}
+% \item My second supervisor
+% \vspace*{3mm}
+% \item Other researchers
+% \vspace*{3mm}
+% \item My family and friends
+%\end{itemize}
 

### 2021-06-08 papers/thesis/full/ch1-introduction/introduction.tex
Addressed most comments so far.

@@
-\chapter{Introduction}
+\chapter{Introduction}\label{CHAPTER:intro}
 
@@
 Koornwinder in 1975 described a method for constructing two-variable OPs from univariate OPs \cite{koornwinder1975two}.
-However, few books have been published over the years on the topic of multivariate orthogonal polynomials \cite{dunkl2014orthogonal}.
+However, few books have been published over the years on the topic of multivariate orthogonal polynomials, with the notable exception of the book by Dunkl \& Xu \cite{dunkl2014orthogonal}.
+As Dumitriu, Edelman and Shuman said in 2007, \enquote{\textit{[multivariate orthogonal polynomials] are understudied, underapplied, and important applications may be being missed}} \cite{dumitriu2007mops}, while in 2016, Olver opined that \enquote{\textit{there is still a long way to go before multivariate orthogonal polynomials can reach their full potential in applications}}.
+He hypothesised that \enquote{\textit{[p]art of this neglect may be due to the fact that mutually orthogonal polynomials are not uniquely defined: there is no canonical ordering}} and continued \enquote{\textit{[e]ven when explicit constructions are known, these are often unwieldy}} \cite{olver2016review}.
+
 This branch of mathematics has an encouraging future though, not least as a basis for sparse spectral methods for solving partial differential equations (PDEs) on multidimensional domains.
@@
 Moreover, by still using a spectral approach, one can maintain the high accuracy and excellent error properties that such methods bring.
-There are also potential applications in physics too, particularly in astrophysics, where solving PDEs on the sphere surface and working in spherical geometries is also desirable (e.g.
+Solving PDEs on a section of the sphere surface is still useful for weather prediction in its own right too.
+Moreover, there are also prospective applications in physics, particularly in astrophysics, where solving PDEs on the sphere surface and working in spherical geometries is also desirable (e.g.
 \cite{vasil2019tensor, reinecke2013libsharp, beyer2014numerical, varshalovich1988quantum, slevinsky2018spectral, rubinstein2015scalar}).
@@
 
-Chapter 2 of this thesis provides an introduction to sparse spectral methods via the spherical harmonics on the whole sphere surface.
+\bsrefchapter{CHAPTER:sphericalharmonics} of this thesis provides an introduction to sparse spectral methods via the spherical harmonics on the whole sphere surface.
 Here, we think of the spherical harmonics in a non-traditional way and write them as a group of multidimensional orthogonal polynomials in $(x,y,z)$ as opposed to functions of spherical coordinates.
@@
 
-In Chapter 3 we move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+In \bsrefchapter{CHAPTER:diskslice} we move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials, which we define.
 Triangles and the full disk are then special cases of our framework, which we formalise for the disk-slice and trapezium cases.
+The work in this chapter has been previously published in \cite{snowball2019sparse}
 
-With a greater knowledge base in our quiver, we can adapt the techniques learnt from the founding of the disk-slice formulation to surfaces in 3D in Chapter 4.
+With a greater knowledge base in our quiver, we can adapt the techniques learnt from the founding of the disk-slice formulation to surfaces in 3D in \bsrefchapter{CHAPTER:sphericalcaps}.
 Using the same family of (non-classical) 1D OPs, we present a suitable orthogonal polynomial basis for the spherical cap, a subdomain of the surface of a unit sphere, complete with sparse differential operators.
@@
 From here, a spectral element method could be devised for the whole sphere using the aforementioned as elements.
+The work in this chapter has been accepted for publication \cite{snowball2020sparse}.
 
-Finally, we summarise the work we have presented and detail avenues for future directions that one could take it in Chapter 5.
+Chapter 5 gives a summary of the work we have presented and details avenues for future directions that one could take it.
+
+Finally, as an appendix, we outline how one could approach a vector OP basis for the \enquote{tangent bundle} of the spherical cap, analogous to the vector spherical harmonics on the whole sphere.
+This constitutes \bsrefappendix{appendix:tangent}.

### 2021-06-08 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Addressed most comments so far.

@@
 }
-While it may seem somewhat odd, it should hopefully become apparent why this is a useful way to define our domain.
+While it may seem somewhat odd, it should hopefully become apparent why this is a practical way to define our domain.
 
@@
 
-The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners the ApproxFun package \cite{ApproxFun} -- however, this package is purely experimental at this stage.
+The code that allows one to produce the numerical examples in this chapter is publicly available as a Julia package\footnote{https://github.com/snowball13/SphericalHarmonics.jl} that partners with the ApproxFun package \cite{ApproxFun} -- however, this package is purely experimental at this stage.
 
@@
 
-Before we proceed further, let's build up to our definition of the spherical harmonics.
+Before we proceed further, let us build up to our definition of the spherical harmonics.
 We first need to introduce a few classical orthogonal polynomials.
@@
 }
-The spherical harmonics as defined are then orthonormal with respect to the inner product
+The spherical harmonics as defined are then orthonormal with respect to the inner product with uniform measure
 \bseqn{
@@
 \end{proof}
-We thus can present a useful corollary specific for our needs.
+We thus can present a helpful corollary specific for our needs.
 \begin{corollary}\label{corollary:jacobirelations}
@@
 	&&\gamma_{l,m,2} := \frac{\clm}{c^{m}_{l+1}} \: \tilde \gamma_{l,m,2}, \\
-	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4.
-}
+	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4,
+}
+where the $\clmprime$ are defined in \bsrefeqn{eqn:clmdef}.
 \end{lemma}
 
-\remark The results in \bsreflemma{lemma:shxyzrelations} are fairly involved and can be derived via the relationships established for products of spherical harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
+\remark The results in \bsreflemma{lemma:shxyzrelations} can be derived via the relationships established for products of spherical harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
 
@@
 	\end{subfigure}\hfil % <-- added
-	\caption{\enquote{Spy} plots of the Jacobi operator matrices, showing their sparsity and structure, up to order $N=10$.}
+	\caption{\enquote{Spy} plots of the Jacobi operator matrices, showing their sparsity and structure, up to order $N=10$.
+Each are block-tridiagonal, with the blocks corresponding to polynomial degree.}
 	\label{fig:sh-spy-jacobi}
@@
 \bseqn{
-	J_x &= 
-		\begin{pmatrix}
-			B_{x,0} & A_{x,0} & & & & \\
-			C_{x,1} & B_{x,1} & A_{x,1} & & & \\
-			& C_{x,2} & B_{x,2} & A_{x,2}  & & & \\
-			& & C_{x,3} & \ddots & \ddots & \\
-			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}, \\
-	J_y &= 
-		\begin{pmatrix}
-			B_{y,0} & A_{y,0} & & & & \\
-			C_{y,1} & B_{y,1} & A_{y,1} & & & \\
-			& C_{y,2} & B_{y,2} & A_{y,2}  & & & \\
-			& & C_{y,3} & \ddots & \ddots & \\
-			& & & \ddots & \ddots & \ddots \\
-		\end{pmatrix}, \\
-	J_z &= 
-		\begin{pmatrix}
-			B_{z,0} & A_{z,0} & & & & \\
-			C_{z,1} & B_{z,1} & A_{z,1} & & & \\
-			& C_{z,2} & B_{z,2} & A_{z,2}  & & & \\
-			& & C_{z,3} & \ddots & \ddots & \\
+	J_{x/y/z} &= 
+		\begin{pmatrix}
+			B_{x/y/z,0} & A_{x/y/z,0} & & & & \\
+			C_{x/y/z,1} & B_{x/y/z,1} & A_{x/y/z,1} & & & \\
+			& C_{x/y/z,2} & B_{x/y/z,2} & A_{x/y/z,2}  & & & \\
+			& & C_{x/y/z,3} & \ddots & \ddots & \\
 			& & & \ddots & \ddots & \ddots \\
@@
 }
-It will be useful for us to give a formal name for these coefficient matrices above for a family of multidimensional orthogonal polynomials.
+It will be convenient for us to give a formal name for these coefficient matrices above for a family of multidimensional orthogonal polynomials.
 \begin{definition}\label{def:clenshawmats}
@@
 using the orthonormality of the spherical harmonics.
-As discussed in \bsrefsection{section:introduction}, methods to calculate these coefficients exist known as spectral transforms, and are well established for the spherical harmonics (see e.g.
+As discussed in \bsrefchapter{CHAPTER:intro}, methods to calculate these coefficients exist known as spectral transforms, and are well established for the spherical harmonics (see e.g.
 \cite{mohlenkamp1999fast, suda2002fast}).
@@
 
-Let's explain what we mean.
+Let us explain what we mean by this.
 Suppose $v : \Omega \to \C$ is a function, and we encounter the problem of finding the coefficients of the expansion for $v(x,y,z) f(x,y,z)$.
@@
 	&= \iint_\Omega \costheta \sinphi \: \gradYlm \cdot {\gradYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
-	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
 	&\quadfour \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
@@
 	&= \iint_\Omega \costheta \sinphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\phi \: \D\theta \nonumber \\
-	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \costheta \: \D\theta \Big) \nonumber \\
 	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho(z)^{\abs{m} + \abs{m'} - 1} \Big\{ i m' \jaclmabs{l'}{m'}(z) \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
@@
 \bseqn{
-	 \int_0^{2pi} \eimtheta \eimthetaprime \sintheta \: \D\theta = i \pi (\delta_{m',m-1} - \delta_{m',m+1})
+	 \int_0^{2\pi} \eimtheta \eimthetaprime \sintheta \: \D\theta = i \pi (\delta_{m',m-1} - \delta_{m',m+1})
 }
@@
 
-For multiplication by z, we will use the same method to directly calculate the nonzero coefficients in each expansion using inner products. To this end, note that
+For multiplication by $z$, we will use the same method to directly calculate the nonzero coefficients in each expansion using inner products. To this end, note that
 \bseqn{
@@
 	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradYlmprime}^* \sinphi \: \D\varphi \: \D\theta \nonumber \\
-	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
 	&\quadtwo \cdot \Big( \int_{-1}^{1} \Big\{ \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
@@
 	&= \iint_\Omega \cosphi \: \gradYlm \cdot {\gradpYlmprime}^* \sinphi \: \D\varphi \: \D\theta \nonumber \\
-	&= \clm \clmprime \Big( \int_0^{2pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
+	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
 	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho(z)^{\abs{m} + \abs{m'} - 2} \Big\{ i m' z \jaclmabs{l'}{m'}(z) \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
@@
 	\end{subfigure}\hfil % <-- added
-	\caption{\enquote{Spy} plots of the Jacobi operator matrices for the vector spherical harmonic OP basis, showing their sparsity and structure, up to order $N=10$.}
-	\label{fig:sh-spy-jacobi}
+	\caption{\enquote{Spy} plots of the Jacobi operator matrices for the vector spherical harmonic OP basis, showing their sparsity and structure, up to order $N=10$.
+These are block-tridiagonal, with sub-block bandwidths: (a) (4,4), (b) (4,4), (c) (2,2).}
+	\label{fig:sh-spy-jacobi-tangent}
 \end{figure}
@@
 }
-Let $0_2 \in \R^{2\times2}$ be the 2x2 zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
+Let $0_2 \in \R^{2\times2}$ be the $2\times2$ zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
 \bseqn{
@@
 }
-The sparsity and structure of the Jacobi matrices for the VSHs as we have defined them can be seen visually in \bsreffig{fig:vsh-spy-jacobi}.
+The sparsity and structure of the Jacobi matrices for the VSHs as we have defined them can be seen visually in \bsreffig{fig:sh-spy-jacobi-tangent}.
 
@@
 }
-Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs (and hence the coefficients) for order $n=0$ are zero.
+Notice how the iteration also ends at $n=1$ (not $n=0$) here, since the VSHs for order $n=0$ are zero.
 
@@
 
-The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other key operators useful in examples such as the linearised shallow water equations.
+The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other key operators found in examples such as the linearised shallow water equations.
 These operators will take coefficients of a function's expansion in either the SSH or VSH basis, to coefficients in either basis.
@@
 \end{definition}
-\begin{remark}
-	The operators $\tilde \calD, \tilde \calG$ would be used when vectors of the resulting length are required for the problem you are wishing to solve.
-\end{remark}
+
+\remark The operators $\tilde \calD, \tilde \calG$ would be used when vectors of the resulting length are required for the problem you are wishing to solve.
 
@@
 
-Now that we have our differential operators, let's apply them in some examples.
-
 \section{Examples}
+
+Now that we have our differential operators, let us apply them in some examples.
 
@@
 }
-The snapshots of the solution $u$ in bsreffig{fig:heateqn} are taken of the initial conditions, plus after $25$, $50$ and $75$ timesteps respectively.
+The snapshots of the solution $u$ in \bsreffig{fig:heateqn} are taken of the initial conditions, plus after $25$, $50$ and $75$ timesteps respectively.
 
@@
 }
-As eluded to earlier, we will \enquote{extend} the scalar function's coefficients vector, writing $\tilde{\vec{h}}^c = E_v \vec{h}^c$, as described in \bsrefsection{section:SHdiffoperators}.
+As eluded to earlier, we will \enquote{extend} the scalar function's coefficients vector, writing $\tilde{\vec{h}}^c = \calE_v \vec{h}^c$, as described in \bsrefsection{section:SHdiffoperators}.
 Thus, our system can be written as a matrix-vector system:
@@
 Doing so at each iteration provides us with the coefficients vectors $\uvec^c_{n+1}, \tilde{\vec{h}}^c_{n+1}$ if we know $\uvec^c_{n}, \tilde{\vec{h}}^c_{n}$.
-Of course, when evaluating the function $h$, we can simply gather back the coefficients vector $\vec{h}^c$ by $\vec{h}^c = E_s \tilde{\vec{h}}^c$.
+Of course, when evaluating the function $h$, we can simply gather back the coefficients vector $\vec{h}^c$ by $\vec{h}^c = \calE_s \tilde{\vec{h}}^c$.
 

### 2021-06-08 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Addressed most comments so far.

@@
 
-In recent years, there has been some interesting work in the area of multivariate orthogonal polynomials in two-dimensional domains (see \cite{dunkl2014orthogonal}), such as for a Chebychev-based spectral method on the rectangle \cite{julien2009efficient}.
+In recent years, there has been some interesting work in the area of multivariate orthogonal polynomials in two-dimensional domains (see \cite{dunkl2014orthogonal}), such as for a Chebyshev-based spectral method on the rectangle \cite{julien2009efficient}.
 In particular, spectral methods have been established on the disk \cite{boyd2011comparing, vasil2016tensor} and the triangle \cite{olver2019triangle}.
@@
 It is concluded that using Chebyshev polynomials yields mixed results, but that there is also no single advised basis to use.
-While Zernike polynomials yield accurate results, there is a lack of radial fast transform, which is a disadvantage.
-However, ideas for how to overcome this have progressed \cite{slevinsky2018use, sakai2009application, wilber2017computing}.
+Zernike polynomials were found to yield accurate results, and ideas for how to implement a radial transform have progressed in recent years \cite{slevinsky2018use, sakai2009application, wilber2017computing} \bstodo{not sure what to say here?}.
 One solution proposed is pairing the Zernike disk OPs with a radial basis defined in terms of Jacobi polynomials \cite{vasil2016tensor}:
 \bseqnnumber{
-	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}.
+	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
 }
@@
 }
-where $C_{n,k} \neq 0$ and \enquote{lower order terms'} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
+where $C_{n,k} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{n-j} y^j$ where $j < k$.
 The precise normalization arises from their definition in terms of one-dimensional OPs we will see in \bsrefdef{def:ds:OPconstruction}.
@@
 In this case the weight $\jacw(x)$ is an ultraspherical weight, and the corresponding OPs are the normalized Jacobi polynomials $\{\jac_n^{(b, b)}\}$, while the weight $w_R(x)$ is non-classical (it is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}).
+
+\remark We should note at this stage that hereon we will continue to use the notation $\jac_n^{(a, b)}$ to refer to the univariate OPs orthonormal with respect to the weight $\jacw^{(a,b)}$.
+For the disk-slice case, this means they are the orthonormal Jacobi polynomials, which is why we choose this notation.
+We stress that this is in difference to the spherical harmonic framework in \bsrefchapter{CHAPTER:sphericalharmonics}, where conventionally they are defined using the un-normalized (i.e.
+simply orthogonal) classical Jacobi polynomials, and for which we also used the notation $\jac_n^{(a, b)}$.
 
@@
 
-Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:sh:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
-\bseqn{
-J_{x/y}^{(a,b,c,d)} &= \begin{pmatrix}
+Further, note that $J_x^{(a,b,c,d)}, J_y^{(a,b,c,d)}$ are banded-block-banded matrices (see \bsrefdef{def:bandedblockbanded}). They are block-tridiagonal (block-bandwidths $(1,1)$):
+\bseqn{
+	J_{x/y}^{(a,b,c,d)} &= 
+		\begin{pmatrix}
 			B^{x/y}_0 & A^{x/y}_0 & & & & \\
@@
 }
-where the 1D OPs $\{\genjac^{(a, b, c)}_n\}$ are orthogonal on the interval $(\alpha, \beta)$ with respect to the weight
+where the 1D OPs $\{\genjac^{(a, b, c)}_n\}$ are orthonormal on the interval $(\alpha, \beta)$ with respect to the weight
 \bseqn{
@@
 }
-and the 1D OPs $\{\jac^{(c, c)}_n\}$ are orthogonal on the interval $(\gamma, \delta) = (-1, 1)$ with respect to the weight
+and the 1D OPs $\{\jac^{(c, c)}_n\}$ are orthonormal on the interval $(\gamma, \delta) = (-1, 1)$ with respect to the weight
 \bseqn{
@@
 The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, and on the triangle \cite{olver2018recurrence}.
-An illustration of how the non-weighted differential operators increment the parameters (a,b,c) is seen in \bsreffig{fig:ds:laplaceparameterconversion}.
+An illustration of how the non-weighted differential operators increment the parameters $(a,b,c)$ is seen in \bsreffig{fig:ds:laplaceparameterconversion}.
 

### 2021-06-08 papers/thesis/full/declaration/declaration.tex
Added declaration, foreword, list of figs, shortened abstract, full name (things the checklist says to do)

@@
 \begin{declaration}
-	Declaration here.
+	I declare that all the work included in this thesis is, to the best of my knowledge, original unless otherwise attributed.
+
+\begin{flushright} Ben Snowball, June 2021. \end{flushright}
+
+\vfill
+
+The copyright of this thesis rests with the author. Unless otherwise indicated, its contents are licensed under a Creative Commons Attribution-Non Commercial 4.0 International Licence (CC BY-NC).
+
+Under this licence, you may copy and redistribute the material in any medium or format.
+You may also create and distribute modified versions of the work.
+This is on the condition that: you credit the author and do not use it, or any derivative works, for a commercial purpose.
+
+When reusing or sharing this work, ensure you make the licence terms clear to others by naming the licence and linking to the licence text.
+Where a work has been adapted, you should indicate that the work has been changed and describe those changes.
+
+Please seek permission from the copyright holder for uses of this work that are not included in this licence or permitted under UK Copyright Law.
+
 \end{declaration}

### 2021-06-08 papers/thesis/full/foreword/foreword.tex
Added declaration, foreword, list of figs, shortened abstract, full name (things the checklist says to do)

@@
 
-Text of the foreword.
+Much of the work in this thesis has already been published, or submitted for publication.
+
+\begin{itemize}
+
+\item The contents of \bsrefchapter{CHAPTER:diskslice} has been published in:
+
+Ben Snowball and Sheehan Olver.
+Sparse spectral and p-finite element methods for partial differential equations on disk slices and trapeziums.
+\textit{Studies in Applied Mathematics}, 145(1):3--35, 2020.
+
+DOI: https://doi.org/10.1111/sapm.12303
+
+\item The contents of \bsrefchapter{CHAPTER:sphericalcaps} has been accepted for publication in \textit{Transactions of Mathematics and Its Applications}:
+
+Ben Snowball and Sheehan Olver. Sparse spectral methods for partial differential equations on spherical caps. arXiv preprint arXiv:2012.11493, 2020.
+
+\end{itemize}
 

### 2021-06-09 papers/thesis/full/ch1-introduction/introduction.tex
Minor addition and typo to intro

@@
 Moreover, by still using a spectral approach, one can maintain the high accuracy and excellent error properties that such methods bring.
-Solving PDEs on a section of the sphere surface is still useful for weather prediction in its own right too.
+Solving PDEs on a section of the sphere surface is still useful for numerical weather prediction in its own right too, where one could for example perform a more localised simulation near a pole while still using a \enquote{global} (in the domain sense) spectral method approach.
 Moreover, there are also prospective applications in physics, particularly in astrophysics, where solving PDEs on the sphere surface and working in spherical geometries is also desirable (e.g.
@@
 Triangles and the full disk are then special cases of our framework, which we formalise for the disk-slice and trapezium cases.
-The work in this chapter has been previously published in \cite{snowball2019sparse}
+The work in this chapter has been previously published \cite{snowball2019sparse}.
 

### 2021-06-10 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Added index of notation. Uses makeindex.

@@
 Our aim here is to express the spherical harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere.
-More precisely, we desire the solution to partial differential equations on the domain
+More precisely, we desire the solution to partial differential equations on the domain \nomenclature[Omega]{$\Omega$}{The domain of interest}
 \bseqn{
@@
 
-It is useful to be able to transform between these cartesian coordinates and the spherical coordinates ($\varphi, \theta$).
-Throughout this thesis we will use the convention that the spherical coordinate angles be defined by
+It is useful to be able to transform between these cartesian coordinates and the spherical coordinates \nomenclature[phi]{$\varphi$}{The polar angle in spherical coordinates} \nomenclature[theta]{$\theta$}{The azimuthal angle in spherical coordinates}($\varphi, \theta$).
+Throughout this thesis we will use the convention that the spherical coordinate angles be defined by \nomenclature[xyz]{$x,y,z$}{Lower case non-bold $x,y,z$ represents cartesian coordinates}
 \bseqn{
@@
 We will finally demonstrate a proof-of-concept via the examples of the heat equation and the linearised shallow water equations on the unit sphere.
-These examples consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): written in spherical coordinates this is
+These examples consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): written in spherical coordinates this is \nomenclature[Deltas]{$\DeltaS$}{The Laplace--Beltrami operator}
 \bseqnnumber{
@@
 
-On the unit interval, $[-1,1]$, we note that there is a hierarchy of OPs in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}:
+On the unit interval, $[-1,1]$, we note that there is a hierarchy of OPs in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}: \nomenclature[Pnab]{$P^{(a,b)}_n$}{The degree $n$ Jacobi polynomial (context determines normalized or not) unless otherwise stated}
 \bseqn{
@@
 }
-where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$.
-Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$
+where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and \nomenclature[Pl]{$P_l$}{The degree $l$ Legendre polynomial} $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$.
+Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$ \nomenclature[Kronecker delta]{$\delta_{n, n'}$}{The Kronecker delta function; $1$ if $n=n'$, else $0$}
 \bseqn{
@@
 which is helpfully known as the \textit{Legendre equation} \cite[14.2.1]{DLMF}.
-Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p5]{belousov1962tables}:
+Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p5]{belousov1962tables}: \nomenclature[Plm]{$\Plm$}{The Associated Legendre polynomial of degree $l$ and order $m$}
 \bseqn{
@@
 Let $(x,y,z) \in \Omega$.
-We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}:
+We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}: \nomenclature[Ylm]{$\Ylm$}{The degree $l$ and order $m$ spherical harmonic}
 \bseqnnumber{
@@
 }
-for $0 \le |m| \le l, \, l \in \No$, where
+for $0 \le |m| \le l, \, l \in \No$, where \nomenclature[clm]{$\clm$}{The normalising constant for the spherical harmonic $\Ylm$}
 \bseqnnumber{
@@
 }
-The spherical harmonics as defined are then orthonormal with respect to the inner product with uniform measure
+The spherical harmonics as defined are then orthonormal with respect to the inner product with uniform measure \nomenclature[conjuagate]{$a^*$}{The complex conjugate of $a \in \C$}
 \bseqn{
@@
 }
-where $\alpha^*$ denotes the complex conjugate of $\alpha \in \C$.
-Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any $m \in \Z$.
+where $\alpha^*$ denotes the complex conjugate of \nomenclature[C]{$\C$}{The complex numbers} $\alpha \in \C$.
+Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any \nomenclature[Z]{$\Z$}{The integers} $m \in \Z$.
 Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$.
-They span all polynomials modulo the vanishing ideal associated by the zero set of $x^2 + y^2 + z^2 - 1$ in $\R^3$.
+They span all polynomials modulo the vanishing ideal associated by the zero set of $x^2 + y^2 + z^2 - 1$ in \nomenclature[R]{$\R$}{The real numbers} \nomenclature[Rd]{$\R^d$}{The d-dimensional coordinate space over the real numbers} $\R^3$.
 
@@
 We thus can present a helpful corollary specific for our needs.
+\nomenclature[alpha]{$\alpha_{l,m,j}$}{Lower case greek letters with subscripts refer to coefficients in scalar OP relations (unless otherwise stated)}
 \begin{corollary}\label{corollary:jacobirelations}
@@
 
-These recurrences lead to Jacobi operators that correspond to multiplication by $x,y,z$ of the OPs. Define for $l \in \No$:
+These recurrences lead to Jacobi operators that correspond to multiplication by $x,y,z$ of the OPs.
+Define for $l \in \No$: \nomenclature[Pvector]{$\SHvec$}{Vector of the SH OPs} \nomenclature[Pvectorl]{$\SHvec_l$}{(Sub-)Vector of the degree $l$ SH OPs}
 \bseqn{
@@
 Note that these OP vectors we have defined are grouped and ordered according to the degree $l$, and within each degree ordered by the Fourier mode $m$.
-Let the Jacobi matrices $J_x, J_y, J_z$ hence be given by
+Let the Jacobi matrices \nomenclature[J]{$J_x, J_y, J_z$}{Jacobi operator matrices} $J_x, J_y, J_z$ hence be given by
 \bseqnnumber{
@@
 These Jacobi matrices can act on the coefficients vector of a function's expansion in the spherical harmonic basis to implement the operation for multiplication by $x$, $y$ or $z$.
-For example, given sufficiently large $n \in \N$ let the function $f(x,y,z) : \Omega \to \C$ be given by its expansion
+For example, given sufficiently large $n \in \N$ let the function $f(x,y,z) : \Omega \to \C$ be given by its expansion \nomenclature[fvec]{$\vec{f}, \vec{\xi}$}{Lower case bold letters refer to vectors} \nomenclature[transpose]{$\vec{v}^\top, A^\top$}{Transpose of the vector $\vec{v}$ or matrix $A$ (non complex conjuagate)}
 \bseqn{
@@
 }
-$J_x$ has subblock-bandwidths $(2,2)$, where the blocks for $l \in \No$ are given by:
+$J_x$ has subblock-bandwidths $(2,2)$, where the blocks for \nomenclature[N]{$\N$}{The natural numbers (positive integers)} \nomenclature[N0]{$\No$}{The set $\N \cup \{0\}$} $l \in \No$ are given by:
 \bseqn{
@@
 
-Rewriting \bsrefeqn{eqn:jacobimatsdef}, we have that:
+Rewriting \bsrefeqn{eqn:jacobimatsdef}, we have that: \nomenclature[I]{$I$, $I_d$}{The identity matrix (in $\R^{d \times d}$)}
 \bseqn{
@@
 But what about vector-valued functions?
+\nomenclature[r]{$\rvec$}{The unit normal vector at the point $(x,y,z)$}
 \begin{definition}\label{def:tangentbundle}
@@
 For the purposes of this discussion, we will only consider the tangent bundle, and thus wish to write down a basis for vector-valued functions with values in the tangent bundle.
-Fortunately, there is a simple solution known as the \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}:
-\bseqnnumber{
-	\gradYlm &:= \nabla \Ylm = \ppx{\phi} \Ylm \phivec + {1 \over \sinphi} \ppx{\theta} \Ylm \thetavec, \nonumber \\
-	\gradpYlm &:= \nabla^\perp \Ylm \equiv \rvec \times \nabla \Ylm = \ppx{\phi} \Ylm \thetavec - {1 \over \sinphi} \ppx{\theta} \Ylm \phivec. \label{eqn:VSHs}
-}
-Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$, while $\phivec, \thetavec$ are the standard spherical coordinate unit vectors, i.e.
+Fortunately, there is a simple solution known as the \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}: \nomenclature[grad]{$\gradS$}{The spherical gradient operator} \nomenclature[Psilm]{$\gradYlm$}{The vector spherical harmonic $\gradS \Ylm$} \nomenclature[Philm]{$\gradpYlm$}{The vector spherical harmonic $\gradS^\perp \Ylm = \rvec \times \gradS \Ylm$}
+\bseqnnumber{
+	\gradYlm &:= \gradS \Ylm = \ppx{\phi} \Ylm \phivec + {1 \over \sinphi} \ppx{\theta} \Ylm \thetavec, \nonumber \\
+	\gradpYlm &:= \gradS^\perp \Ylm \equiv \rvec \times \gradS \Ylm = \ppx{\phi} \Ylm \thetavec - {1 \over \sinphi} \ppx{\theta} \Ylm \phivec. \label{eqn:VSHs}
+}
+Here, $\rvec$ is simply the outward unit normal vector to the surface of the sphere at the point $(x,y,z)$, while \nomenclature[phi theta vector]{$\phivec$, $\thetavec$}{The standard spherical coordinate angular basis vectors} $\phivec, \thetavec$ are the standard spherical coordinate unit vectors, i.e.
 $\phivec := [\costheta \cosphi, \sintheta \cosphi, -\sinphi]^\top$, $\thetavec := [-\sintheta, \costheta, 0]^\top$.
-They are orthogonal with respect to the inner product $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$ -- in particular, for any $l, l' \in \N$, $m, m' \in \Z$ s.t.
+They are orthogonal with respect to the inner product \nomenclature[inner]{$\ip<u, v>$}{An inner product (defined by the context)}  $\ip<\vec{A}, \vec{B}> = \iint_\Omega \vec{A} \cdot \vec{B}^* \: \sinphi \: \D\varphi \: \D\theta$ -- in particular, for any $l, l' \in \N$, $m, m' \in \Z$ s.t.
 $-l \le m \le l$, $-l' \le m' \le l'$,
@@
 \begin{lemma}\label{lemma:VSHrecurrences}
-	For $l \in \N$ and $m \in \Z$ s.t. $-l \le m \le l$, the vector spherical harmonics satisfy the relationships:
+	For $l \in \N$ and $m \in \Z$ s.t.
+$-l \le m \le l$, the vector spherical harmonics satisfy the relationships: \nomenclature[Alpha]{$A_{l,m,j}$}{Capital greek letters with subscripts refer to coefficients in vector OP (VSH) relations (unless otherwise stated)}
 \bseqn{
@@
 
-As for the scalar case, these coefficients detailed above are then the nonzero entries of the Jacobi matrices for the VSHs. Define $\VSHvecfull$ by
+As for the scalar case, these coefficients detailed above are then the nonzero entries of the Jacobi matrices for the VSHs.
+Define $\VSHvecfull$ by \nomenclature[T vector]{$\VSHvecfull$}{Vector of the VSHs} \nomenclature[Tl vector]{$\VSHvecfulll$}{(Sub)-Vector of the VSHs of order $l$}
 \bseqnnumber{
@@
 }
-Let $0_2 \in \R^{2\times2}$ be the $2\times2$ zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
+Let \nomenclature[zerod]{$0_d$}{The zero matrix in $\R^{d \times d}$} $0_2 \in \R^{2\times2}$ be the $2\times2$ zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
 \bseqn{
@@
 Let $u_n$ for $n = 0, 1, 2, \dots$ be the solution at time $n \Delta t$ where $\Delta t$ is the timestep.
-The are timestepping method is
+The are timestepping method is \nomenclature[Delta t]{$\Delta t$}{A timestep for a simulation}
 \bseqn{

### 2021-06-10 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Added index of notation. Uses makeindex.

@@
 We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
-On the disk-slice the family of weights we consider are of the form
+On the disk-slice the family of weights we consider are of the form \nomenclature[a]{$a,b,c,d$}{Lower case letters $a,b,c,d$ refer to OP family parameters (unless otherwise stated)}
 \bseqn{
@@
 
-We can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals \([\alpha,\beta]\) and \([\gamma,\delta]\).
+We can construct 2D orthogonal polynomials on $\Omega$ from 1D orthogonal polynomials on the intervals $[\alpha,\beta]$ and $[\gamma,\delta]$.
 
@@
 
-For disk slices and trapeziums, we specialise \bsrefprop{prop:ds:construction} in the following definitions. First we introduce notation for two families of univariate OPs.
+For disk slices and trapeziums, we specialise \bsrefprop{prop:ds:construction} in the following definitions.
+First we introduce notation for two families of univariate OPs.
+\nomenclature[weight]{$\genjacw^{(a,b,c)}$, $\jacw^{(a,b)}$}{Weight functions for the one-dimensional OPs $\{\genjac_n^{(a,b,c)}\}$ and $\{\jac_n^{(a,b)}\}$ respectively}
 \begin{definition}\label{def:ds:OPconstruction}
@@
 }
-and define the associated inner products by:
+and define the associated inner products by: \nomenclature[omega]{$\normgenjac$, $\normjac$}{Normalising constants for inner products} \nomenclature[innerw]{$\ip< p, \: q >_w$}{A one-dimensional inner product with weight $w$}
 \bseqnnumber{
@@
 }
-Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjac}, and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipjac}.
+Denote the three-parameter family of orthonormal polynomials on $[\alpha,\beta]$ by \nomenclature[R]{$\genjac_n^{(a,b,c)}$}{Non-classical three-parameter univariate OP of degree $n$} $\{\genjac_n^{(a,b,c)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipgenjac}, and the two-parameter family of orthonormal polynomials on $[\gamma,\delta]$ by $\{\jac_n^{(a,b)}\}$, orthonormal with respect to the inner product defined in \bsrefeqn{eqn:ds:ipjac}.
 \end{definition}
 
-We can now write down our 2D OP family.
+We can now write down our 2D OP family. \nomenclature[Hnk]{$\hdopnk^{(a,b,c,d)}$}{Four-parameter bivariate OP of degree $n$ and order $k$ defined on the given domain and by the given inner product}
 \begin{definition}\label{def:ds:constuction}
@@
 
-$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight
+$\{\hdopnk^{(a,b,c,d)}\}$ are orthogonal with respect to the weight \nomenclature[Wa]{$W^{(a)}, W^{(a,b,c)}, W^{(a,b,c,d)}$}{Weight function for relevant multidimensional OPs, with the given parameters}
 \bseqn{
@@
 \bseqn{
-	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{W^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}
-}
-where for $f, g : \Omega \to \R$ the inner product is defined as 
-\bseqn{
-	\ip< f, \: g >_{W^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
+	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{\hdop^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}
+}
+where for $f, g : \Omega \to \R$ the inner product is defined as \nomenclature[innerH]{$\ip< p, \: q >_{\hdopabcd}$}{A multidimensional inner product over $\Omega$ with the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
+\bseqn{
+	\ip< f, \: g >_{\hdop^{(a,b,c,d)}} := \iint_\Omega \: f(x,y) \: g(x,y) \: W^{(a,b,c,d)}(x,y) \: \D y \: \D x.
 }
@@
 \bseqnnumber{
-	&\ip< \hdopnkabcd, \: \hdopmjabcd >_\Wabcd \\
+	&\ip< \hdopnkabcd, \: \hdopmjabcd >_{\hdopabcd} \\
 	&=  \iint_\Omega \: \Big[ \genjacnmk^{(a,b,c+d+2k+1)}(x) \: \genjacmmj^{(a,b,c+d+2j+1)}(x) \: \rho(x)^{k+j}  \nonumber \\
@@
 simply orthogonal) classical Jacobi polynomials, and for which we also used the notation $\jac_n^{(a, b)}$.
-
-%Again, for clarity we will use the notation that for $p : \Omega \to \R$, $\norm{p}_{\Wab}^2 := \ip<p, p>_{\Wab}$.
 
@@
 For the recurrence for multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$.
-These coefficients are given by
-\bseqn{
-	c_{m,j} = {\ip< y \: \hdopnkabcd, \hdopmjabcd >_{\Wabcd}}{\norm{\hdopmjabcd}^{-2}_{\Wabcd}}.
-}
-Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopmjabcd}_{\Wabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$.
+These coefficients are given by \nomenclature[abs]{$\abs{m}$}{The absolute value of a scalar $m$} \nomenclature[norm]{$\norm{f}$}{The norm of a function f, defined by the square root of the relevant inner product} \nomenclature[normH]{$\norm{f}_{\hdopabcd}$}{The norm of a function f, defined by the square root of the inner product over $\Omega$ given by the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
+\bseqn{
+	c_{m,j} = {\ip< y \: \hdopnkabcd, \hdopmjabcd >_{\hdopabcd}}{\norm{\hdopmjabcd}^{-2}_{\hdopabcd}}.
+}
+Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopmjabcd}_{\hdopabcd}^2 = \normgenjac^{(a,b,c+d+2j+1)} \: \normjac^{(d,c)}$.
 Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using the change of variable $t = \frac{y}{\rho(x)}$:
 \bseqn{
-	&\ip<y \hdopnkabcd, \hdopmjabcd>_\Wabcd \\
+	&\ip<y \hdopnkabcd, \hdopmjabcd>_{\hdopabcd} \\
 	&=  \iint_\Omega \hdopnkabcd(x,y) \: \hdopmjabcd(x,y) \: y \: \Wabcd(x,y) \: dy \: dx \\
@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
-\bseqn{
-\bighdopabcd_n := \begin{pmatrix}
+Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
+Define, for $n=0,1,2,\dots$: \nomenclature[Hvectorabcd]{$\bighdopabcd$}{Vector of the disk-slice/trapezium OPs ordered by degree, with parameters $a,b,c,d$} \nomenclature[Hvectornabcd]{$\bighdopabcd_n$}{(Sub-)Vector of the disk-slice/trapezium OPs of degree $n$, with parameters $a,b,c,d$}
+\bseqn{
+	\bighdopabcd_n := 
+		\begin{pmatrix}
 			\hdopabcd_{n,0}(x,y) \\
@@
 	\quad \quad 
-\bighdopabcd := \begin{pmatrix}
+	\bighdopabcd := 
+		\begin{pmatrix}
 			\bighdopabcd_0 \\
@@
 }
-Denote the weighted OPs by
+Denote the weighted OPs by \nomenclature[Wvectorabcd]{$\bigWabcd$}{Vector of the weighted disk-slice/trapezium OPs ordered by degree, with parameters $a,b,c,d$}
 \bseqn{
@@
 \bseqn{
-	c_{m,j}^y = {\ip<\ppx{y} \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{W^{(a, b+1)}}}.
+	c_{m,j}^y = {\ip<\ppx{y} \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{\hdop^{(a, b, c+1)}}}{\norm{\hdop^{(a, b, c+1)}_{m,j}}^{-2}_{\hdop^{(a, b+1)}}}.
 }
@@
 \bseqn{
-	&\ip<\ppx{y} \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{W^{(a, b, c+1)}} \\
+	&\ip<\ppx{y} \hdopnkabc, \hdop^{(a, b, c+1)}_{m,j} >_{\hdop^{(a, b, c+1)}} \\
 	&= \iint_\Omega \: \Big[ \genjacnmk^{(a,b,2c+2k+1)}(x) \: \rho(x)^{k-1} \: \jac_k^{(c,c) \: \prime}\fpr(\frac{y}{\rho(x)}) \\
@@
 \bseqn{
-	c_{m,j}^x = {\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{W^{(a+1, b+1, c+1)}}}.
+	c_{m,j}^x = {\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{\hdop^{(a+1, b+1, c+1)}}}{\norm{\hdop^{(a+1, b+1, c+1)}_{m,j}}^{-2}_{\hdop^{(a+1, b+1, c+1)}}}.
 }
@@
 \bseqnnumber{
-	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} \nonumber \\ 
+	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{\hdop^{(a+1, b+1, c+1)}} \nonumber \\ 
 	&= \Big( \int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1) \: \prime} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j+1} \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \Big) \nonumber \\
@@
 \bseqn{
-	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{W^{(a+1, b+1, c+1)}} = 0
+	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{\hdop^{(a+1, b+1, c+1)}} = 0
 }
@@
 \bseqn{
-	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{W^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkab, \hdopmj^{(\tilde{a},\tilde{b})} >_\Wabc.
+	\ip< \Wabc \hdopnkabc, W^{(\tilde{a},\tilde{b},\tilde{c})} \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{\hdop^{(-\tilde{a},-\tilde{b},-\tilde{c})}} = \ip< \hdopnkabc, \hdopmj^{(\tilde{a},\tilde{b},\tilde{c})} >_{\hdopabc}.
 }
@@
 \bseqn{
-	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
+	c_{m,j} = {\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}{\norm{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j}}^{-2}_{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}}}.
 }
@@
 \bseqn{
-	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{W^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
+	&\ip< \hdopnkabc, \hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}_{m,j} >_{\hdop^{(a+\tilde{a},b+\tilde{b},c+\tilde{c})}} \\
 	&= \normgenjac^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{k+j+1} \: \genjacmmj^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c}+2j+1)} >_{\genjacw^{(a+\tilde{a},b+\tilde{b},2c+2\tilde{c})}} \\
@@
 General linear partial differential operators with polynomial variable coefficients can be constructed by composing the sparse representations for partial derivatives, conversion between bases, and Jacobi operators.
-As a canonical example, we can obtain the matrix operator for the Laplacian $\Delta$, that will take us from coefficients for expansion in the weighted space
+As a canonical example, we can obtain the matrix operator for the Laplacian \nomenclature[Delta]{$\Delta$}{The Laplacian operator in $\R^2$} $\Delta$, that will take us from coefficients for expansion in the weighted space
 \bseqn{
@@
 
-Another important example is the Biharmonic operator $\Delta^2$, where we assume zero Dirichlet and Neumann conditions.
+Another important example is the biharmonic operator \nomenclature[Delta2]{$\Delta^2$}{The biharmonic operator in $\R^2$} $\Delta^2$, where we assume zero Dirichlet and Neumann conditions.
 To construct this operator, we first note that we can obtain the matrix operator for the Laplacian $\Delta$ that will take us from coefficients for expansion in the space $\bighdopooo(x,y)$ to coefficients in the space $\bighdop^{(2,2,2)}(x,y)$.
@@
 
-\noindent{\bf Remark}:
-The Corollary demonstrates that in order to obtain the recurrence coefficients $\{\alpha_{m}^{(a,b,2c+2k+1)}\}$, $\{\beta_{m}^{(a,b,2c+2k+1)}\}$ for some $m$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{m+2}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{m+2}^{(a,b,2c+2(k-1)+1)}\}$.
+\remark The Corollary demonstrates that in order to obtain the recurrence coefficients $\{\alpha_{m}^{(a,b,2c+2k+1)}\}$, $\{\beta_{m}^{(a,b,2c+2k+1)}\}$ for some $m$ and $k$, we require that we obtain the recurrence coefficients $\{\alpha_{m+2}^{(a,b,2c+2(k-1)+1)}\}$, $\{\beta_{m+2}^{(a,b,2c+2(k-1)+1)}\}$.
 Thus, for large $N$, this recursive method of obtaining the recurrence coefficients requires a large initialisation (i.e.
@@
 Denote the  Gauss quadrature nodes and weight on $[\alpha,\beta]$ with weight $(\beta - s)^a \: (s - \alpha)^b \: \rho(s)^{2c+1}$ as $(s_k,w_k^{(s)})$ , and
- on \([-1,1]\) with weight \((1-t^2)^c\) as $(t_k,w_k^{(t)})$. Define
+ on $[-1,1]$ with weight $(1-t^2)^c$ as $(t_k,w_k^{(t)})$. Define
 \bseqn{
@@
 Let $f(x,y)$ be a polynomial on $\Omega$. The quadrature rule is then
-$$
+\bseqn{
 	\iint_\Omega f(x,y) \: \Wab(x,y) \: \D A \approx \half \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) + f(x_j, -y_j) \big],
-$$
+}
 where $M = \ceil{\half(N+1)}^2$, and the quadrature rule is exact if $f(x,y)$ is a polynomial of degree $\le N$.
@@
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
-	f_{n,k} := \frac{\ip< f, \: \hdopnk^{(a,b,c)}>_{\Wabc}}{\norm{\hdopnk^{(a,b,c)}}_{\Wabc}}
-}
-Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopnkabc}^2_\Wabc = \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}$.
+	f_{n,k} := \frac{\ip< f, \: \hdopnk^{(a,b,c)}>_{\hdopabc}}{\norm{\hdopnk^{(a,b,c)}}_{\hdopabc}}
+}
+Recall from \bsrefeqn{eqn:ds:normhdop} that $\norm{\hdopnkabc}^2_{\hdopabc} = \normgenjac^{(a,b,2c+2k+1)} \: \normjac^{(c)}$.
 Using the quadrature rule detailed in Section 4.2 for the inner product, we can calculate the coefficients $f_{n,k}$ for each $n = 0,\dots,N$, $k = 0,\dots,n$:
@@
 
-Recall that the OPs $\bighdop^{(a,b,c)}$ are orthogonal with respect to the weight $\Wabc$ on $\Omega$, and define the matrix $\Lambda^{(a,b,c)} :=  \ip< \bighdop^{(a,b,c)}, \: {\bighdop^{(a,b,c)}}^\top >_{\Wabc}$.
+Recall that the OPs $\bighdop^{(a,b,c)}$ are orthogonal with respect to the weight $\Wabc$ on $\Omega$, and define the matrix $\Lambda^{(a,b,c)} :=  \ip< \bighdop^{(a,b,c)}, \: {\bighdop^{(a,b,c)}}^\top >_{\hdopabc}$.
 Note that due to orthogonality this is a diagonal matrix.
@@
 	&= \int_\element \: \vec{v}^\top \: \bigWiii \: \bighdopiii^\top \: \vec{f} \: \D \vec{x} \\
-	&= \vec{v}^\top \: \ip< \bighdopiii, {\bighdopiii}^\top >_{\Wiii} \: \D \vec{x} \\
+	&= \vec{v}^\top \: \ip< \bighdopiii, {\bighdopiii}^\top >_{\hdopiii} \: \D \vec{x} \\
 	&= \vec{v}^\top \Lambda^{(1,1,1)} \: \vec{f},

### 2021-06-10 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Added index of notation. Uses makeindex.

@@
 
-By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
+By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product \nomenclature[innerQ]{$\ip< p, \: q >_{\scopa}$}{A multidimensional inner product over $\Omega$ with the weight $\Wa$ associated with the OPs $\{\scopnkia\}$}
 \bseqn{
@@
 }
-with
+with \nomenclature[normQ]{$\norm{f}_{\scopa}$}{The norm of a function f, defined by the square root of the inner product over $\Omega$ given by the weight $\Wa$ associated with the OPs $\{\scopnkia\}$}
 \bseqnnumber{
@@
 To this end, we define our OP-building ordering as follows.
-For $n=0,1,2,\dots$:
+For $n=0,1,2,\dots$: \nomenclature[Qtildevectora]{$\bigscopta$}{Vector of the spherical cap OPs ordered by degree, with parameter $a$} \nomenclature[Qtildevectorna]{$\bigscoptna$}{(Sub-)Vector of the spherical cap OPs of degree $n$, with parameter $a$}
 \bseqn{
@@
 To this end, as alluded to in \bsrefsection{section:sc:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal.
-Let $N \in \N$ and define:
+Let $N \in \N$ and define: \nomenclature[QvectorNa]{$\bigscopNa$}{Vector of the spherical cap OPs up to degree $N$, grouped by order $k$, with parameter $a$} \nomenclature[QvectorNka]{$\bigscopNka$}{(Sub-)Vector of the spherical cap OPs of order $k$ up to degree $N$, with parameter $a$}
 \bseqnnumber{
@@
 
-We further denote the weighted set of OPs on $\Omega$ by 
+We further denote the weighted set of OPs on $\Omega$ by \nomenclature[WvectorNa]{$\bigWNa$}{Vector of the weighted spherical cap OPs up to degree $N$, grouped by order $k$, with parameter $a$}
 \bseqn{

### 2021-06-10 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Addressed the new minor comments, though I have questions for the other points. Also changed the rec coeff mats defs for each chapter to make consistent.

@@
 for $(x,y,z) \in \Omega$.
+While $J_x, J_y, J_z$ are not Jacobi matrices in the classical sense, they are block Jacobi matrices.
+However, it is useful to simply label them as such.
 These Jacobi matrices can act on the coefficients vector of a function's expansion in the spherical harmonic basis to implement the operation for multiplication by $x$, $y$ or $z$.
@@
 \begin{definition}\label{def:clenshawmats}
-	The matrices $-\Dlt (B_l-G_l(x,y,z))$, $\Dlt C_l$ for $l \in \No$ defined via \bsrefeqns{eqn:clenshawmats1}{eqn:Dzerotdef} are called the \textbf{recurrence coefficient matrices} for a given family of multidimensional orthogonal polynomials.
+	The matrices $-\Dlt (B_l-G_l(x,y,z))$, $\Dlt C_l$ for $l \in \No$ defined via \bsrefeqns{eqn:clenshawmats1}{eqn:Dzerotdef} are called the recurrence coefficient matrices for a given family of multidimensional orthogonal polynomials.
 \end{definition}

### 2021-06-10 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Addressed the new minor comments, though I have questions for the other points. Also changed the rec coeff mats defs for each chapter to make consistent.

@@
 where $\alpha \in (-1,1)$ and $\beta := 1$.
-Simply put, the region of the surface of the $2$-sphere where the $z$-coordinate range is limited to a sub-interval of $[-1,1]$ is what we refer to as the spherical cap.
-
-\begin{remark}
-For simplicity we focus on the case of a spherical cap, though there is an extension to a spherical band by taking $\beta \in (\alpha,1)$.
+Simply put, the region of the surface of the $2$-sphere where the $z$-coordinate range is limited to the interval $[\alpha,1]$ is what we refer to as the spherical cap.
+
+\remark For simplicity we focus on the case of a spherical cap, though there is an extension to a spherical band by taking $\beta \in (\alpha,1)$.
 The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper.
 These adjustments make the construction more involved, but the approach is essentially the same which is why they are omitted here.
-\end{remark}
 
@@
 \end{lemma}
-\begin{remark}
-	For $z$ multiplication, note that different Fourier modes do not interact. This is because multiplication by $z$ is invariant with respect to rotation around the $z$-axis. 
-\end{remark}
+
+\remark For $z$ multiplication, note that different Fourier modes do not interact. This is because multiplication by $z$ is invariant with respect to rotation around the $z$-axis.
+
 \begin{proof}
@@
 }
-While $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are not Jacobi matrices in the classical sense, they are block Jacobi matrices.
-However, it is useful to simply label them as such.
-They are once again in fact banded-block-banded matrices -- see \bsrefdef{def:bandedblockbanded}.
-More specifically, each Jacobi matrix is block-tridiagonal (block-bandwidths $(1,1)$).
+The Jacobi matrices are once again in fact banded-block-banded matrices -- see \bsrefdef{def:bandedblockbanded}. More specifically, each Jacobi matrix is block-tridiagonal (block-bandwidths $(1,1)$).
 
@@
 where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
-\bseqn{
+\bseqnnumber{
 	A_n &:= 
@@
 			C^{(a)}_{z,n}
-		\end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \nonumber \\
+		\end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \label{eqn:sc:jacobisubmats1} \\
 	B_n &:= 
@@
 			zI_{2n+1}
-		\end{pmatrix} \in \R^{3(2n+1)\times(n+1)}.
+		\end{pmatrix} \in \R^{3(2n+1)\times(n+1)}. \label{eqn:sc:jacobisubmats2}
 }
@@
 Note that we can define an explicit $\Dnt$ as follows:
-\bseqn{
+\bseqnnumber{
 	\Dnt := 
@@
 			& & & & & \vec{\eta}^\top_{1} & & &
-		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
+		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)}, \label{eqn:sc:Dtdef}
 }
@@
 For $n=0$, we can simply take
-\bseqn{
+\bseqnnumber{
 	D^\top_0 &:= 
@@
 			0 & \frac{1}{\betaa_{0,0,6}} & 0
-		\end{pmatrix} \in \R^{3\times3}.
-}
-
+		\end{pmatrix} \in \R^{3\times3}. \label{eqn:sc:Dzerotdef}
+}
 It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bigscopta_{0}(x,y,z)$  through $\bigscopta_{n}(x,y,z)$ in optimal $O(n^2)$ complexity.
@@
 \begin{definition}\label{def:sc:clenshawmats}
-	The \textit{recurrence coefficient matrices} associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
+	The matrices $-\Dnt (B_n-G_n(x,y,z))$, $\Dnt C_n$ defined via \bsrefeqns{eqn:sc:jacobisubmats1}{eqn:sc:Dzerotdef} are the recurrence coefficient matrices associated with the OPs $\{\scopnkia\}$.
 \end{definition}
@@
 By applying these differential operators, we are (in some cases) incrementing or decrementing the parameter value $a$.
-It is therefore necessary to also be able to raise or lower the parameter by way of an independent operator.
-There exist conversion  operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
+It is therefore necessary to also be able to raise or lower the parameter by way of conversion operators, transforming the OPs from one (weighted or non-weighted) parameter space to another.
 \begin{definition}\label{def:sc:parametertransformationoperators}
@@
 which encodes how accurate the solution to $A \vc x = \vc b$ is under perturbation.
-Since the Laplacian operator is block-diagonal one can determine the condition number from the condition numbers of each block on the diagonal,  and hence deduce the accuracy in solving the Poisson equation, where graphical representations are seen in \bsreffig{fig:sc:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
+Since the Laplacian is block-diagonal one can determine the condition number from the condition numbers of each block on the diagonal,  and hence deduce the accuracy in solving the Poisson equation, where graphical representations are seen in \bsreffig{fig:sc:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
 As the degree $N$ increases, the condition numbers of these blocks becomes grows  algebraically fast (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
-Fortunately however, by applying a trivial diagonal preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in Figure \bsreffig{fig:sc:conditionnumbers}.
+Fortunately however, by applying a trivial diagonal preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in \bsreffig{fig:sc:conditionnumbers}.
 The preconditioning matrix chosen is $P^{-1}$, where $P$ is the matrix of the diagonal of $\Delta^{(1)}_W$.
@@
 
-\begin{remark} 
-	Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[3.5.23]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
-\end{remark}
+\remark Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[3.5.23]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
 
@@
 
-What about other boundary conditions?
-
 \subsection{Biharmonic equation}
@@
 Right: The norms of each block of the computed solution of the biharmonic equation with the right hand side function $f(x,y,z) = \exp(-\epsilon((x-x_0)^2 + (y-y_0)^2 + (z-z_0)^2))$ where  $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$, for various $\epsilon$ values.
-This demonstrates algebraic convergence.}
+This demonstrates spectral convergence.}
 	\label{fig:sc:biharmonic}
@@
 
-Other boundary conditions beyond those already presented here are beyond the scope of this thesis.
+Other boundary conditions aside from those already presented here are beyond the scope of this thesis.
 However, the approach established for the case of a triangle domain in 2D shows how Neumann conditions can be considered  as a system of equations in partial derivatives involving Dirchlet conditions \cite{olver2019triangle}.
@@
 This work also serves as a stepping stone to constructing similar methods to solve partial differential equations on other 3D sub-domains of the sphere such as spherical triangles.
-It is clear from the construction in this paper that discretizations spherical Laplacian's  and other partial differential operators are sparse on other suitable sub-components of the sphere.
+It is clear from the construction in this chapter that discretizations of the spherical Laplacian and other partial differential operators are sparse on other suitable sub-components of the sphere.
 The resulting sparsity in high-polynomial degree discretizations presents an attractive alternative to methods based on bijective mappings (e.g., \cite{DGShallowWater,FEMShallowWater,boyd2005sphere}).

### 2021-06-10 papers/thesis/full/ch1-introduction/introduction.tex
Added a sentence to explain there are other global spectral models in the intro

@@
 Notably, spherical harmonics are also used as basis functions for the spectral transform method that makes up part of the model in the Integrated Forecasting System (IFS), which is used by ECMWF for their forecasts \cite{wedi2013fast}.
+
+There are other modelling systems using global spectral models too.
+For example, the AGCM3 model developed by the Canadian Centre for Climate Modelling and Analysis (CCCma) and the Community Earth System Model (CESM) both use a spectral dynamical core\footnote{A \textit{dynamical core} is part of a model that deals with numerically solving the equations of motion on the underlying grid, as opposed to other physical processes.} \cite{hurrell2013community, scinocca2008cccma}, while the Geophysical Fluid Dynamics Laboratory (GFDL) has also developed a global spectral atmospheric model \cite{gordon1982description}.
+
 While the whole sphere spectral method using the spherical harmonics has been successful for numerous years \cite{williamson2007evolution}, there is a drawback in the parallel scalability bottleneck that arises from the global spectral transform, which is expected to inhibit future performance of the IFS \cite{ecmwf2020scalability, wedi2013fast}.

### 2021-06-11 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Further edits to address comments. Some questions remain.

@@
 
-A continuation to the procedure for defining OPs on quadratic curves (see \cite{olver2021orthogonal}) has also been presented \cite{olver2020orthogonal}, where OPs are defined on quadratic surfaces of revolution in higher dimensions (specifically, surfaces in $\R^{d+1} for d \ge 2$).
+A continuation to the procedure for defining OPs on quadratic curves (see \cite{olver2021orthogonal}) has also been presented \cite{olver2020orthogonal}, where OPs are defined on quadratic surfaces of revolution in higher dimensions (specifically, surfaces in $\R^{d+1}$ for $d \ge 2$).
 The authors describe their technique as modelled on the structure of the unit sphere, by approaching the construction as a generalisation of spherical harmonics.
@@
 
-Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions. That is, 
+Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions.
+That is, \bstodo{All functions not dependent on r (ie any we expand in this basis) satisfy the derivative BC automatically.
+I am fairly sure this is still the BC for the sphere though..?}
 \bseqn{

### 2021-06-11 papers/thesis/full/conclusion/summary.tex
Further edits to address comments. Some questions remain.

@@
 
-In this thesis, we have developed sparse spectral methods for solving partial differential equations (PDEs) on disk-slices and trapeziums in 2D, and spherical caps as a surface 3D.
+In this thesis, we have developed sparse spectral methods for solving partial differential equations (PDEs) on disk-slices and trapeziums in 2D, and spherical caps as a surface in 3D.
 The work can also be used as a template for developing similar methods on other such similar domains, in particular other subdomains of the sphere.
@@
 Finally, we moved on to use the same arguments for the spherical cap, a 3D surface that is a subdomain of the unit sphere.
-We once again defined the 3D OPs as orthogonal polynomials in in $x,y,z$, and showed how the differential operator matrices continue to elicit similar banded-block-banded structure.
+We once again defined the 3D OPs as orthogonal polynomials in $x,y,z$, and showed how the differential operator matrices continue to elicit similar banded-block-banded structure.
 
@@
 
-Ideally, we would like to pair this with a spectral element method for the sphere, where the elements would be the spherical subdomains of spherical bands and spherical caps.
+A further avenue to take the framework for scalar functions is to construct a vector polynomial basis for functions whose values lie on the tangent bundle of the sphere.
+In \bsrefappendix{appendix:tangent}, we provide a summary for how one needs to approach this.
+The motivation for such is so as to be able to derive similarly sparsely structured operators for the spherical gradient and divergence.
+The main observation however is that, while the vector spherical harmonics are already established as being such an orthogonal and complete basis for the whole sphere, a similar construction for the spherical cap would not yield the same result.
+
+Ideally, we would like to pair the framework presented in \bsrefchapter{CHAPTER:sphericalcaps} with a spectral element method for the sphere, where the elements would be the spherical subdomains of spherical bands and spherical caps.
 The goal here would be to test this method against the full sphere spectral method the is in place in the ECMWF model, with the hypothesis being that by reducing the size of the transforms, we can improve the overall efficiency while still maintaining the accuracy that we achieve from a spectral approach.
+
+Discretizations of partial differential operators are sparse on other suitable sub-domains of the sphere, as evidenced by the construction presented for the spherical cap.
+Hence, using the techniques learnt here, one could also look to constructing similar methods for solving PDEs on spherical triangles, with a view to using these sub-domains as elements for a spherical spectral element method.
+Moreover, one could use this framework to also develop sparse spectral methods on more general domains defined by an algebraic surface -- in particular those defined by non-uniform rational B-splines (NURBS) \bstodo{Not sure what to say exactly or how in depth to go here}.
+However, it is not obvious how one would derive or compute the OPs for such domains.

### 2021-06-11 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Addressed the new minor comments, though I have questions for the other points. Also changed the rec coeff mats defs for each chapter to make consistent.
Edits to Zernike stuff and other minor edits

@@
 For the disk, a review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials.
-It is concluded that using Chebyshev polynomials yields mixed results, but that there is also no single advised basis to use.
-Zernike polynomials were found to yield accurate results, and ideas for how to implement a radial transform have progressed in recent years \cite{slevinsky2018use, sakai2009application, wilber2017computing} \bstodo{not sure what to say here?}.
-One solution proposed is pairing the Zernike disk OPs with a radial basis defined in terms of Jacobi polynomials \cite{vasil2016tensor}:
+It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results.
+In the years since however, such radial fast transforms have been developed \cite{slevinsky2018use, sakai2009application, wilber2017computing}.
+The Zernike disk OPs can be written in terms of Jacobi polynomials \cite{boyd2011comparing}:
+\bseqn{
+	Z_n^m(r, \theta) := R_n^m(r) \sin(m\theta), \quad Z_n^{-m}(r, \theta) := R_n^m(r) \cos(m\theta),
+}
+where the radial part is
+\bseqn{
+	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1)
+}
+for $n - m$ even, while it is defined as $0$ for $n - m$ odd.
+A generalisation of the Zernike polynomials has been used for the development of a sparse spectral method on the disk \cite{vasil2016tensor}.
+These radial OPs are:
 \bseqnnumber{
@@
 }
-where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to $\int_0^1 Q_n^{k,m}(r) \; Q_{n'}^{k,m}(r) \; (1-r^2)^k \; r \; \D r = \delta_{n,n'}$.
-$\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials.
-The use of polar coordinates here then allows for radially symmetric partial differential (and other) operators, as well as basis (parameter) conversion operators, to be represented as banded matrices by exploiting the relations of the Jacobi polynomials -- they are extremely useful for solving a number of problems in not just 1D but 2D also, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.
-
-Conversely, in a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
+where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$.
+$\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials on the unit disk.
+The classical Zernike polynomials are retrieved via setting $k = 0$.
+The use of polar coordinates here then allows for radially symmetric partial differential and other operators, as well as basis (parameter) conversion operators, to be represented as banded matrices by exploiting the relations of the Jacobi polynomials.
+Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.
+
+In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
 \bseqnnumber{
@@
 where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
-\bseqn{
+\bseqnnumber{
 A_n &:= \begin{pmatrix}
@@
 		C^y_n
-	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
+	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \label{eqn:ds:jacobisubmats1} \\
 B_n &:= \begin{pmatrix}
@@
 		yI_{n+1}
-	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
+	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}. \label{eqn:ds:jacobisubmats2}
 }
@@
 Note that we can define an explicit $\Dnt$ as follows:
-\bseqn{
+\bseqnnumber{
 \Dnt := \begin{pmatrix}
@@
 		\eta_{0} & 0 & \hdots & 0 & \eta_{1} & \hdots & \hdots & \eta_{n+1}
-	    \end{pmatrix},
+	    \end{pmatrix}, \label{eqn:ds:Dtdef}
 }
@@
 For the disk-slice, $\betaabc_{n,k,2} = \betaabc_{n,k,5} = \betaabc_{n,k,8} \equiv 0$ for any $n, k$.
+
+\begin{definition}\label{def:sc:clenshawmats}
+	The matrices $-\Dnt (B_n-G_n(x,y,z))$, $\Dnt C_n$ defined via \bsrefeqns{eqn:ds:jacobisubmats1}{eqn:ds:Dtdef} are the recurrence coefficient matrices associated with the OPs $\{\hdopnkabcd\}$.
+\end{definition}
 

### 2021-06-14 papers/thesis/full/ch3-disk-slices/disk-slices.tex
(Hopefully almost) final edits made to thesis

@@
 where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$.
-$\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials on the unit disk.
-The classical Zernike polynomials are retrieved via setting $k = 0$.
-The use of polar coordinates here then allows for radially symmetric partial differential and other operators, as well as basis (parameter) conversion operators, to be represented as banded matrices by exploiting the relations of the Jacobi polynomials.
+These OPs constitute a one-parameter family of orthogonal polynomials on the unit disk, with the classical Zernike polynomials being retrieved via setting the parameter $k = 0$.
+The use of polar coordinates allows for radially symmetric partial differential operators, while the incorporation of a parameter means they can be represented as banded matrices by exploiting the relations of the Jacobi polynomials.
 Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

### 2021-06-14 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Edits to Zernike stuff and other minor edits
(Hopefully almost) final edits made to thesis

@@
 	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity, plotted for $N = 20$.
+Note that here each block corresponds to the order $k$ and not the degree $n$.
 (a) The Laplace-Beltrami operator $\Delta^{(1)}_W$.
@@
 (c) The variable coefficient Helmholtz operator $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.
-(d) The Biharmonic operator $\calB_W^{(2)}$.}
+(d) The biharmonic operator $\calB_W^{(2)}$.}
 	\label{fig:sc:sparsity}
@@
 
-Another desirable operator is the Biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions.
-That is, \bstodo{All functions not dependent on r (ie any we expand in this basis) satisfy the derivative BC automatically.
-I am fairly sure this is still the BC for the sphere though..?}
-\bseqn{
-	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \nabla_S u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
-}
-where $\partial\Omega$ is the $z=\alpha$ boundary, and $\unitvec{n}(x,y,z)$ is the outward unit normal vector at the point $(x,y,z)$ on the boundary, i.e.
-$\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$.
-The matrix operator for the Biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
+Another desirable operator is the biharmonic operator $\DeltaS^2$, for which we assume zero Dirichlet and Neumann conditions. That is, 
+\bseqn{
+	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \gradS u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
+}
+where $\partial\Omega$ is the $z=\alpha$ boundary, and $\unitvec{n}(x,y,z)$ is the outward unit normal vector at the point $(x,y,z)$ on the boundary.
+For the spherical cap $\Omega$, this is simply $\phivec(x,y,z)$.
+Hence,
+\bseqn{
+	\gradS u(x,y,z) \cdot \unitvec{n}(x,y,z) = \gradS u(x,y,z) \cdot \phivec(x,y,z) = - \rho(z) \pfpx{z}{u}(x,y,z)
+} 
+and the boundary conditions simplify to
+\bseqn{
+	u(x,y,z) &= 0, \quad \pfpx{z}{u}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega.
+}
+The matrix operator for the biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
 To construct this, we can simply multiply together two of the spherical Laplacian operators defined in \bsrefdef{def:sc:differentialoperators}, namely $\calL^{(0)\to(2)}$ and $\calL_W^{(2)\to(0)}$:
@@
 }
-Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,y,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$ and thus its spherical gradient would be zero on the boundary $z = \alpha$.
+Since the operator $\calL_W^{(2)\to(0)}$ acts on coefficients in the $\bigW^{(2)}(x,y,z)$ space, we ensure that we satisfy the zero Dirichlet and Neumann boundary conditions -- such a function could be written $u(x,y,z) = \genjacw^{(2,0)}(z) \: \tilde u(x,y,z)$.
 This allows us to apply the $\calL^{(0)\to(2)}$ operator after, safe in the knowledge that boundary conditions have been accounted for.
@@
 We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions on the spherical cap defined by $\Omega$.
-We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, as well as a time dependent heat equation, demonstrating the versatility of the approach.
+We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the biharmonic equation, as well as a time dependent heat equation, demonstrating the versatility of the approach.
 
@@
 
-Our last example is the biharmonic equation: find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
+Our last example is the biharmonic equation. Find $u(x,y,z)$ given a function $f(x,y,z)$ such that:
 \bseqnnumber{
@@
 }
-where $\DeltaS^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
-For clarity, we reiterate that the unit normal vector in this sense is simply $\unitvec{n}(x,y,z) = \unitvec{n}(\vec x) := {\vec x \over \norm{\vec x}} = \vec x$ (see \bsrefsection{section:sc:furtherdiffoperators}).
-In \bsreffig{fig:sc:biharmonic} we see the solution to the Biharmonic \bsrefeqn{eqn:sc:biharmonic} in the spherical cap $\Omega$.
+where $\DeltaS^2$ is the biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$.
+Recall from \bsrefsection{section:sc:furtherdiffoperators} that for the spherical cap, the boundary conditions here simplify to $u = \pfpx{z}{u} = 0$ on $\partial \Omega$ (since the normal vector of the boundary is simply $\phivec(x,y,z)$).
+In \bsreffig{fig:sc:biharmonic} we see the solution to the biharmonic \bsrefeqn{eqn:sc:biharmonic} in the spherical cap $\Omega$.
 In \bsreffig{fig:sc:biharmonic} we also show the norms of each block of calculated coefficients of the approximation for four more complex right-hand sides of the biharmonic equation with $N = 200$, that is, $(N+1)^2 = 40,401$ unknowns.

### 2021-06-14 papers/thesis/full/conclusion/summary.tex
(Hopefully almost) final edits made to thesis

@@
 Hence, using the techniques learnt here, one could also look to constructing similar methods for solving PDEs on spherical triangles, with a view to using these sub-domains as elements for a spherical spectral element method.
-Moreover, one could use this framework to also develop sparse spectral methods on more general domains defined by an algebraic surface -- in particular those defined by non-uniform rational B-splines (NURBS) \bstodo{Not sure what to say exactly or how in depth to go here}.
+Moreover, one could use this framework to also develop sparse spectral methods on more general domains defined by an algebraic surface -- in particular those defined by non-uniform rational B-splines (NURBS).
 However, it is not obvious how one would derive or compute the OPs for such domains.

### 2021-06-14 papers/thesis/full/abstract/abstract.tex
Added comments/questions

--- 2021-02-17 papers/thesis/full/abstract/abstract.tex
@@
 
-This thesis develops sparse spectral methods for solving partial differential equations (PDEs) on various multidimensional domains, with a specific focus on the disk-slice and trapezium in 2D, and the spherical cap as a surface in 3D.
-
-We begin with an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface.
-We present a framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
+%This thesis develops sparse spectral methods for solving partial differential equations (PDEs) on various multidimensional domains, with a specific focus on the disk-slice and trapezium in 2D, and the spherical cap as a surface in 3D.
+For the latter scenario, the PDEs are surface PDEs involving Laplace-Beltrami operators, spherical gradients and other spherical operators.
+%
+%Throughout this thesis we wish to utilise our understanding of the orthogonal polynomials and sparse spectral methods to build up to new methods for subdomains of the sphere surface.
+To this end, we begin with an introduction to sparse spectral methods via the long established Spherical Harmonics on the whole sphere surface.
+We present a framework for how the spherical harmonics can be thought of as a set of multidimensional orthogonal polynomials in $x$, $y$ and $z$, and can be used to expand functions defined on the sphere.
+We explain how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for the function's expansion.
+This way of thinking about the spherical harmonics differs from the traditional sense (where they are considered as orthogonal functions in spherical coordinates) but we do so as to be able to generalise the techniques to spectral methods on other domains using multidimensional OPs.
 Further, we demonstrate how the Vector Spherical Harmonics can be used as an orthogonal basis for vector valued functions lying in the tangent space of the sphere, and thus how one can additionally derive gradient and divergence operators.
-
-Throughout this thesis we wish to utilise our understanding of the 3D spherical harmonics to build up to a new orthogonal polynomial basis for subdomains of the sphere surface.
-To this end, we move on working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
+%
+%We move on to working in 2D, where in recent years sparse spectral methods for solving PDEs have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles.
 Presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary, this work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
 Triangles and the full disk are then special cases of our framework, which we formalise for the disk-slice and trapezium cases.
+%
+%Finally, we return the the surface of the sphere, namely the subdomain we call the spherical cap.
+Piecing together techniques and OPs from both the Spherical Harmonics and disk-slice work, we once again present a new orthogonal polynomial basis and sparse spectral method for the spherical cap, complete with the same observation about the guaranteed sparsity of differential and other operators.
+The motivation is for one to use spherical caps and spherical bands (which are a simple extension) as elements in a spectral element method for the sphere, with many applications in meteorology and astrophysics -- in particular, as a potential replacement of the Spherical Harmonics approach that is currently in use at the European Centre for Medium-range Weather Forecasts (ECMWF) which is predicted in the future to be too costly in the due to a parallel scalability bottleneck arising from the global spectral transform.
 
-Finally, we adapt the tools developed in two dimensions to derive a similar OP basis complete with sparse differential operators.
+\bstodoinline{New (shortened) abstract - any comments?}
+
+This thesis develops sparse spectral methods for solving partial differential equations (PDEs) on various multidimensional domains, with a specific focus on the disk-slice and trapezium in 2D, and the spherical cap in 3D.
+For the latter, the PDEs are surface PDEs involving Laplace-Beltrami operators, spherical gradients and other spherical operators.
+
+We begin with an introduction to sparse spectral methods via viewing spherical harmonics as multidimensional orthogonal polynomials in $x$, $y$, $z$.
+We explain how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for a function's expansion.
+Further, we demonstrate how vector spherical harmonics in $x$, $y$ and $z$ can be used as an orthogonal basis for vector-valued functions, yielding similar banded-block-banded gradient and divergence operators.
+
+We move on to presenting a new framework for choosing a suitable orthogonal polynomial basis for more general 2D domains defined via an algebraic curve as a boundary.
+This work builds on the observation that sparsity is guaranteed due to this definition of the boundary, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials.
+Triangles and the full disk are then special cases of our framework, which we formalise for the disk-slice and trapezium.
+
+Piecing together the techniques used thus far, we present a new orthogonal polynomial basis and sparse spectral method for the spherical cap, complete with the same observation of the guaranteed sparsity of operators.
+The motivation is for one to use spherical caps bands as in a spectral element method for the sphere, with many applications in meteorology and astrophysics -- in particular, as a potential replacement of the spherical harmonics approach currently in use at ECMWF (which is predicted to become too costly due to a parallel scalability bottleneck arising from the global spectral transform).
 

### 2021-06-14 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Added comments/questions

@@
 }
+
+\bstodoinline{The above I think is now correct. Does it make sense?}
+
 The matrix operator for the biharmonic operator will take us from coefficients in the space $\bigW^{(2)}(x,y,z)$ to coefficients in the space $\bigscopN^{(2)}(x,y,z)$.
@@
 Recall from \bsrefsection{section:sc:furtherdiffoperators} that for the spherical cap, the boundary conditions here simplify to $u = \pfpx{z}{u} = 0$ on $\partial \Omega$ (since the normal vector of the boundary is simply $\phivec(x,y,z)$).
+
+\bstodoinline{The above just references the BCs explained earlier.}
+
 In \bsreffig{fig:sc:biharmonic} we see the solution to the biharmonic \bsrefeqn{eqn:sc:biharmonic} in the spherical cap $\Omega$.

### 2021-06-14 papers/thesis/full/ch1-introduction/introduction.tex
Added comments/questions
Addressed final points. Removed comments.

@@
 However, few books have been published over the years on the topic of multivariate orthogonal polynomials, with the notable exception of the book by Dunkl \& Xu \cite{dunkl2014orthogonal}.
-As Dumitriu, Edelman and Shuman said in 2007, \enquote{\textit{[multivariate orthogonal polynomials] are understudied, underapplied, and important applications may be being missed}} \cite{dumitriu2007mops}, while in 2016, Olver opined that \enquote{\textit{there is still a long way to go before multivariate orthogonal polynomials can reach their full potential in applications}}.
-He hypothesised that \enquote{\textit{[p]art of this neglect may be due to the fact that mutually orthogonal polynomials are not uniquely defined: there is no canonical ordering}} and continued \enquote{\textit{[e]ven when explicit constructions are known, these are often unwieldy}} \cite{olver2016review}.
+As Dumitriu, Edelman and Shuman said in 2007, \enquote{[multivariate orthogonal polynomials] are understudied, underapplied, and important applications may be being missed} \cite{dumitriu2007mops}, while in 2016, Olver opined that \enquote{there is still a long way to go before multivariate orthogonal polynomials can reach their full potential in applications}.
+He hypothesised that \enquote{[p]art of this neglect may be due to the fact that mutually orthogonal polynomials are not uniquely defined: there is no canonical ordering} and continued \enquote{[e]ven when explicit constructions are known, these are often unwieldy} \cite{olver2016review}.
 

### 2021-06-14 papers/thesis/full/foreword/foreword.tex
Added comments/questions
Addressed final points. Removed comments.

@@
 
-\item The contents of \bsrefchapter{CHAPTER:sphericalcaps} has been accepted for publication in \textit{Transactions of Mathematics and Its Applications}:
+\item The contents of \bsrefchapter{CHAPTER:sphericalcaps} has been submitted for publication in \textit{Transactions of Mathematics and Its Applications}:
 

### 2021-06-15 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Further edits to address comments. Some questions remain.
Minor edits to preamble and help with spacing issues for large eqn environments

@@
 \bseqn{
-	x \,\Ylmxyz &= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
+	&x \,\Ylmxyz \\
+	&= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
 	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
 	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
-	&\quad \quad + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
+	&\quadtwo + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
 	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
-	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z),
+	&\quadtwo + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z),
 }
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
@@
 \bseqn{
-	\uvec = \sum_{l = 0}^{N} \sum_{m = -l}^{l} u_{l,m}^{\gradYlm} \; \gradYlm + u_{l,m}^{\gradpYlm} \; \gradpYlm.
+	\uvec = \sum_{l = 0}^{N} \sum_{m = -l}^{l} u_{l,m}^{\gradY} \; \gradYlm + u_{l,m}^{\gradpY} \; \gradpYlm.
 }
@@
 	&= \int_{-1}^{1} \Big\{ \abs{m}(\abs{m}+2) \jaclmabs{l'}{m} \rho^{2\abs{m}} \big[ \tilde\gamma_{l,\abs{m},1} \jaclmabs{l-1}{m} + \tilde\gamma_{l,\abs{m},2} \jaclmabs{l+1}{m} \big] \nonumber \\
-	&\quadtwo + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\
+	&\quadfour + d_{l,\abs{m}} d_{l',\abs{m}} \jaclmbracket{l'}{\abs{m}+1} \rho^{2\abs{m}+2} \nonumber \\
+	&\quadfour \quadtwo \cdot \big[ \tilde\gamma_{l,\abs{m}+1,1} \jaclmbracket{l-1}{\abs{m}+1} + \tilde\gamma_{l,\abs{m}+1,2} \jaclmbracket{l+1}{\abs{m}+1} \big]\Big\} \: \D z \nonumber \\
 	&= c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over 2\pi (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over 2\pi (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
@@
 	\ip<z \gradYlm, \gradYlmprime> &= \clm \delta_{m',m} \Big\{ c_{l-1}^{m} \delta_{l',l-1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \big] \nonumber \\
-	&\quadfour + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}.
+	&\quadtwo + c_{l+1}^{m} \delta_{l',l+1} \big[ \abs{m}(\abs{m}+2) {\tilde\gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde\gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \big] \Big\}.
 \label{eqn:lemmaVSHrecurrencesproofz3}
@@
 	&= \clm \clmprime \Big( \int_0^{2\pi} \eimtheta \eimthetaprime \: \D\theta \Big) \nonumber \\
-	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho(z)^{\abs{m} + \abs{m'} - 2} \Big\{ i m' z \jaclmabs{l'}{m'}(z) \big[ \abs{m} z \jaclmabs{l}{m}(z) - \rho(z)^2 \jaclmabsdz{l}{m}(z) \big] \nonumber \\
-	&\quadfour + i m z \jaclmabs{l}{m}(z) \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \Big\} \: \D z \Big) \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ 2i m\abs{m} z^2 \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m} - 2} \nonumber \\
-	&\quadeight \quadtwo - i m z \jaclmabs{l}{m}(z) \jaclmabsdz{l'}{m}(z) \rho(z)^{2\abs{m}} \nonumber \\
-	&\quadeight \quadtwo - i m z \jaclmabsdz{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \Big\} \: \D z \nonumber \\
-	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} i m \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
-	&= \delta_{l',l} \delta_{m',m} i m. \label{eqn:lemmaVSHrecurrencesproofz4}
+	&\quadtwo \cdot \Big( \int_{-1}^{1} \rho^{\abs{m} + \abs{m'} - 2} \Big\{ i m' z \jaclmabs{l'}{m'} \big[ \abs{m} z \jaclmabs{l}{m} - \rho^2 \jaclmabsdz{l}{m} \big] \nonumber \\
+	&\quadfour \quadfour \quadtwo + i m z \jaclmabs{l}{m} \big[ \abs{m'} z \jaclmabs{l'}{m'} - \rho^2 \jaclmabs{l'}{m'} \big] \Big\} \: \D z \Big) \nonumber \\
+	&= \clm c_{l'}^{m} 2 \pi \delta_{m',m} \: \int_{-1}^{1} \Big\{ 2i m\abs{m} z^2 \jaclmabs{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m} - 2} \nonumber \\
+	&\quadeight \quadtwo - i m z \jaclmabs{l}{m} \jaclmabsdz{l'}{m} \rho^{2\abs{m}} \nonumber \\
+	&\quadeight \quadtwo - i m z \jaclmabsdz{l}{m} \jaclmabs{l'}{m} \rho^{2\abs{m}} \Big\} \: \D z \nonumber \\
+	&= \clm c_{l'}^{m} \; 2 \pi \; \delta_{m',m} \: \int_{-1}^{1} i m \jaclmabs{l}{m}(z) \jaclmabs{l'}{m}(z) \rho(z)^{2\abs{m}} \: \D z \nonumber \\
+	&= \delta_{l',l} \; \delta_{m',m} \; i \; m. \label{eqn:lemmaVSHrecurrencesproofz4}
 }
@@
 }
-Let \nomenclature[zerod]{$0_d$}{The zero matrix in $\R^{d \times d}$} $0_2 \in \R^{2\times2}$ be the $2\times2$ zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
+Let \nomenclature[zerod]{$0_d$}{The zero matrix in $\R^{d \times d}$} $0_2 \in \R^{2\times2}$ be the zero matrix. $\tangentjacobi_x$ has subblock-bandwidths $(4,4)$ since for $l \in \No$:
 \bseqn{
@@
 			& & \mathcal{A}_{l,l,5} & 0_2
-		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
+		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)},
+}
+and
+\bseqn{
 	C_{x,l} &:= 
@@
 \bseqn{
-	\mathcal{A}_{l,m,j} := 
+	\mathcal{A}_{l,m,j} &:= 
 		\begin{pmatrix}
@@
 			0 & A_{l,m,j}
-		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad
-	\mathcal{A}_{l,m,j} := 
+		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \\
+	\mathcal{A}_{l,m,j} &:= 
 		\begin{pmatrix}
@@
 			& & \mathcal{B}_{l,l,5} & 0_2
-		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
+		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)},
+}
+and
+\bseqn{
 	C_{y,l} &:= 
@@
 \bseqn{
-	\mathcal{B}_{l,m,j} := 
+	\mathcal{B}_{l,m,j} &:= 
 		\begin{pmatrix}
@@
 			0 & B_{l,m,j}
-		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \quad
-	\mathcal{B}_{l,m,j} := 
+		\end{pmatrix} \: \text{for } j = 1, 2, 3, 4, \\
+	\mathcal{B}_{l,m,j} &:= 
 		\begin{pmatrix}
@@
 			& & & - \Gamma_{l,l,3} & 0 \\
-		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)}, \\
+		\end{pmatrix}  \in \R^{2(2l+1)\times2(2l+1)},
+}
+and
+\bseqn{
 	C^z_l &:= 
@@
 
-\subsection{Three-term recurrence relation for $\VSHvecfull(x,y,z)$}\label{subsection:VSHrecurrence}
+\subsection{Three-term recurrence relation for $\VSHvecfull$}\label{subsection:VSHrecurrence}
 
@@
 		I_6 & & & \\
-		B_1-G_1(x,y,z) & A_1 & & \\
-		C_2 & B_2-G_2(x,y,z) & \quad A_3 \quad & \\
-		& C_3 & B_3 - G_3(x,y,z)  & \ddots \\
+		B_1-G_1 & A_1 & & \\
+		C_2 & B_2-G_2 & \quad A_3 \quad & \\
+		& C_3 & B_3 - G_3  & \ddots \\
 		& & \ddots &\ddots
 	\end{pmatrix}
-	\VSHvecfull(x,y,z)
+	\VSHvecfull
 	=
 	\begin{pmatrix}
-		\VSHvecfull_1(x,y,z) \\ \vec{0} \\ \vec{0} \\ \vdots
+		\VSHvecfull_1 \\ \vec{0} \\ \vec{0} \\ \vdots
 	\end{pmatrix},
 }
-where we note that 
+where 
 \bseqn{
@@
 }
-can be calculated explicitly, and for each $l \in \N$,
+which we note can be calculated explicitly, and for each $l \in \N$,
 \bseqnnumber{
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \bold{0}, \: \vec{\xi}_{N+2} = \bold{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
 	\quad &\text{2) } \text{For } n = N:-1:1 \\
@@
 \bseqn{
-	\calD &\equiv \calL := 
-		\begin{pmatrix}
-			\calL_0 & & \\
-			& \ddots & \\
-			& & \calL_N
-		\end{pmatrix} \in \R^{(N+1)^2 \times (N+1)^2}, \quad
-	\tilde \calD = 
-		\begin{pmatrix}
-			\tilde \calD_0 & & \\
-			& \ddots & \\
-			& & \tilde \calD_N
-		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2}, \\
 	\calG &= I_{(N+1)^2}, \quad
@@
 			& & & & 0
+		\end{pmatrix}  \in \R^{2(N+1)^2 \times 2(N+1)^2}, \\
+	\calD &\equiv \calL := 
+		\begin{pmatrix}
+			\calL_0 & & \\
+			& \ddots & \\
+			& & \calL_N
+		\end{pmatrix} \in \R^{(N+1)^2 \times (N+1)^2}, \\
+	\tilde \calD &= 
+		\begin{pmatrix}
+			\tilde \calD_0 & & \\
+			& \ddots & \\
+			& & \tilde \calD_N
 		\end{pmatrix} \in \R^{2(N+1)^2 \times 2(N+1)^2},
@@
 
-\section{Examples}
+\section{Examples}\label{section:examples}
 
@@
 \bseqn{
-	\rvec \times \bold{u} &= \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
+	\rvec \times \vec{u} &= \rvec \times \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \gradYlm + u^{\gradpY}_{l,m} \gradpYlm ] \\
 	&= \sum_{l=0}^{N} \sum_{m=-l}^{l} [ u^{\gradY}_{l,m} \rvec \times \gradYlm + u^{\gradpY}_{l,m} \rvec \times ( \rvec \times \gradYlm ) ] \\

### 2021-06-15 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Minor edits to preamble and help with spacing issues for large eqn environments

--- 2021-06-14 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
@@
 \bseqn{
-	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
+	&\ip< p, \: q >_{\scopa} \\
+	&:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
 	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
@@
 	\alphaa_{n,k,5} &:= \eta_{k-1} \: \ip<\genjacnmk^{(a, 2k)}, \: \genjac_{n-k+2}^{(a, 2(k-1))}>_{\genjacw^{(a, 2k)}}, \\
-	\alphaa_{n,k,6} &:=\eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}}, \\
+	\alphaa_{n,k,6} &:=\eta_{k} \: \ip<\genjacnmk^{(a, 2k)}, \genjacnmk^{(a, 2(k+1))}>_{\genjacw^{(a, 2(k+1))}},
+}
+\bseqn{
 	\betaa_{n,k,i,j} &:= 
@@
 }
-where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ in the lemma:sc:
+where we show the non-zero coefficients that result are the $\betaa_{n,k,1},\dots,\betaa_{n,k,6}$ as stated in the Lemma:
 \bseqn{
@@
 	&\quadeight + \genjacnmk^{(a,2k)} \: \genjacw^{(a-1, 2k)} \big[ \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime \prime} \: \genjacw^{(1, 2)} + a \genjac_{m-k}^{(a - \tilde a, 2k) \: \prime} \: \rho^2 \\
-	&\quadeight \quadeight \quadtwo \quad - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)} \big] \Big\} \: \D z \\
+	&\quadeight \quadeight \quadtwo \quad - 2(k+1) z \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(1, 0)} \big] \Big\} \: \D z
+}
+\bseqn{
 	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacnmk^{(a,2k)} \: \genjacw^{(a - 2, 2k)} \: \big[ -k(k+1) \genjacw^{(2,0)} - 2a(k+1)z \: \genjacw^{(1,0)} \\

### 2021-06-18 papers/thesis/full/ch1-introduction/introduction.tex
minor edits after readthrough (typos, repetition of accronym defs, etc.)

@@
 
-Spherical harmonics are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
+Spherical harmonics (SHs) are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
 \cite{moon2008efficient, sloan2013efficient}), astrophysics (e.g.
@@
 We present an OP framework for how the spherical harmonics can be used to expand functions defined on the sphere as multidimensional polynomials in $x, y, z$, and how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for a function's expansion.
-Further, we demonstrate how the vector spherical harmonics can be used as an orthogonal basis for vector valued functions lying in the \enquote{tangent bundle} of the sphere, and thus how one can additionally derive gradient and divergence operators.
+Further, we demonstrate how the vector spherical harmonics (VSHs) can be used as an orthogonal basis for vector valued functions lying in the \enquote{tangent bundle} of the sphere, and thus how one can additionally derive gradient and divergence operators.
 

### 2021-06-18 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
minor edits after readthrough (typos, repetition of accronym defs, etc.)

@@
 We first need to introduce a few classical orthogonal polynomials.
-\enquote{Orthogonal polynomials} (OPs) are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product.
+Orthogonal polynomials are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product.
 In particular, they are orthogonal with respect to all lower degree polynomials.
@@
 
-An important property of the Jacobi matrices is that they are sparse -- specifically they are \textit{banded-block-banded matrices}.
+An important property of the Jacobi matrices is that they are sparse -- specifically they are banded-block-banded matrices.
 
@@
 
-Three-term recurrence relations for orthogonal polynomials are well established (see for example \cite[Section 18.9]{DLMF}), including for multidimensional OPs \cite[Section 3.2]{dunkl2014orthogonal}.
+Three-term recurrence relations for orthogonal polynomials are well established (see for example \cite[\S18.9]{DLMF}), including for multidimensional OPs \cite[\S3.2]{dunkl2014orthogonal}.
 In a similar vein, we can obtain a recurrence relation for the spherical harmonics by combining each system in \bsrefeqn{eqn:jacobimatsdef} in a specific way.
@@
 such that $\Dlt A_l = \idmat{2l+3}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[Section 3.2]{dunkl2014orthogonal}, which can be expanded to obtain the three-term recurrence, for $l = 0,1,2,\dots$:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[\S3.2]{dunkl2014orthogonal}, which can be expanded to obtain the three-term recurrence, for $l = 0,1,2,\dots$:
 \bseqn{
@@
 For the purposes of this discussion, we will only consider the tangent bundle, and thus wish to write down a basis for vector-valued functions with values in the tangent bundle.
-Fortunately, there is a simple solution known as the \textit{vector spherical harmonics (VSHs)} \cite{barrera1985vector}: \nomenclature[grad]{$\gradS$}{The spherical gradient operator} \nomenclature[Psilm]{$\gradYlm$}{The vector spherical harmonic $\gradS \Ylm$} \nomenclature[Philm]{$\gradpYlm$}{The vector spherical harmonic $\gradS^\perp \Ylm = \rvec \times \gradS \Ylm$}
+Fortunately, there is a simple solution known as the vector spherical harmonics (VSHs) \cite{barrera1985vector}: \nomenclature[grad]{$\gradS$}{The spherical gradient operator} \nomenclature[Psilm]{$\gradYlm$}{The vector spherical harmonic $\gradS \Ylm$} \nomenclature[Philm]{$\gradpYlm$}{The vector spherical harmonic $\gradS^\perp \Ylm = \rvec \times \gradS \Ylm$}
 \bseqnnumber{
@@
 }
-As we can see, the coefficients are then also only possibly nonzero for $l' = l \pm 1$.
+As we can see, the coefficients zero for $l' \ne l \pm 1$.
 Combining the above with the \bsrefeqn{eqn:lemmaVSHrecurrencesproof}, we have that our two coefficients for $\calA_{l-1,m-1}, \calA_{l+1,m-1}$ are those stated in the Lemma as $A_{l,m,1}$ and $A_{l,m,3}$, for $m > 0$.
@@
 }
-Again, we can see that the coefficients are then also only possibly nonzero for $l' = l \pm 1$.
+Again, we can see that the coefficients hence zero for $l' \ne l \pm 1$.
 Combining the above with the \bsrefeqn{eqn:lemmaVSHrecurrencesproof}, we have that our two coefficients for $\calA_{l-1,m-1}, \calA_{l+1,m-1}$ are those stated in the Lemma as $A_{l,m,2}$ and $A_{l,m,4}$, for $m \ge 0$.
@@
 The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other key operators found in examples such as the linearised shallow water equations.
-These operators will take coefficients of a function's expansion in either the SSH or VSH basis, to coefficients in either basis.
-
-Let $\uvec$ be a vector-valued function on the tangent bundle of $\Omega$ with coefficients vectors $\uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, and let $f$ be a scalar function on $\Omega$ with coefficients vector $\fvecc$ for its expansion in the SSH basis also up to order $N$, i.e.
+These operators will take coefficients of a function's expansion in either the SH or VSH basis, to coefficients in either basis.
+
+Let $\uvec$ be a vector-valued function on the tangent bundle of $\Omega$ with coefficients vectors $\uvecc, \uveccperp$ for its expansion in the VSH basis up to order $N \in \N$, and let $f$ be a scalar function on $\Omega$ with coefficients vector $\fvecc$ for its expansion in the SH basis also up to order $N$, i.e.
 \bseqn{
@@
 for some $\alpha \in \R$.
-A simple timestepping method we can use to iterate with is the BDF2 method \cite[§2.3]{iserles2009first}.
+A simple timestepping method we can use to iterate with is the BDF2 method \cite[\S2.3]{iserles2009first}.
 Let $u_n$ for $n = 0, 1, 2, \dots$ be the solution at time $n \Delta t$ where $\Delta t$ is the timestep.
-The are timestepping method is \nomenclature[Delta t]{$\Delta t$}{A timestep for a simulation}
+The timestepping method is \nomenclature[Delta t]{$\Delta t$}{A timestep for a simulation}
 \bseqn{
@@
 }
-From here, it is clear that eat each timestep we simply have a fairly trivial linear system to solve, with the matrices $I - \Delta t \alpha \calL$, $3I - 2\Delta t \alpha \calL$ being banded-block-banded (in fact, simply diagonal here).
+From here, it is clear that at each timestep we simply have a fairly trivial linear system to solve, with the matrices $I - \Delta t \alpha \calL$, $3I - 2\Delta t \alpha \calL$ being banded-block-banded (in fact, simply diagonal here).
 We demonstrate this theory in \bsreffig{fig:heateqn}, where we see four snapshots of the solution to the heat equation in \bsrefeqn{eqn:heateqn} on the unit sphere with our order $N$ set large enough (here, $N=20$).
@@
 
-The linearised shallow water equations allow us to showcase more of the differential operators that we defined in \bsrefdef{def:SHoperators} (namely the divergence and gradient operators) and demonstrate how the natural sparsity that our framework brings to the problem leads to a simple sparse linear system to solve.
+The linearised shallow water equations (SWEs) allow us to showcase more of the differential operators that we defined in \bsrefdef{def:SHoperators} (namely the divergence and gradient operators) and demonstrate how the natural sparsity that our framework brings to the problem leads to a simple sparse linear system to solve.
 
@@
 In this chapter, we have chosen to view the spherical harmonics as orthogonal (in fact orthonormal) polynomials in $x$, $y$ and $z$ in order to illustrate and understand more about how the framework presented here for using multivariate OPs in a spectral method translate to the sphere.
-We have derived the entries of \textit{Jacobi matrices} that represent multiplication by the coordinates, showing how they are \textit{banded-block-banded} in structure.
+We have derived the entries of Jacobi matrices that represent multiplication by the coordinates, showing how they are banded-block-banded in structure.
 We have set out how one can find coefficients for the expansion of a function in the spherical harmonics basis using established transforms, and how one can evaluate a function given its coefficients vector using a multidimensional version of the Clenshaw algorithm.
-We extended our framework to include the \textit{tangent bundle} of the sphere -- that is, the space of vectors orthogonal to the unit outward normal for each point on the sphere -- where we can use the \textit{vector spherical harmonics} as vector-valued orthogonal polynomials to be a basis, demonstrating how similar such Jacobi matrices, function expansion and evaluation can be derived.
+We extended our framework to include the tangent bundle of the sphere -- that is, the space of vectors orthogonal to the unit outward normal for each point on the sphere -- where we can use the vector spherical harmonics as vector-valued orthogonal polynomials to be a basis, demonstrating how similar such Jacobi matrices, function expansion and evaluation can be derived.
 Further, by exploiting the natural sparsity of the relationships between the OPs in the bases, we have shown how matrices for differential and other operators will be sparse and banded-block-banded in structure too.

### 2021-06-18 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Added comments/questions
minor edits after readthrough (typos, repetition of accronym defs, etc.)

@@
 
-For the disk, a review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials.
+A review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials.
 It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results.
@@
 
-We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials (OPs) that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
+We will show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations on the interval \cite{olver2013fast}, and its analogues on the disk \cite{vasil2016tensor} and triangle \cite{olver2018recurrence,olver2019triangle}.
 On the disk-slice the family of weights we consider are of the form \nomenclature[a]{$a,b,c,d$}{Lower case letters $a,b,c,d$ refer to OP family parameters (unless otherwise stated)}
@@
 for $\alpha \leq x \leq \beta,$ $-\rho(x) \leq y \leq \rho(x)$.
-The corresponding OPs denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$.
+The corresponding OPs are denoted $\hdopnkabc(x,y)$, where $n$ denotes the polynomial degree, and $0 \le k \le n$.
 We define these to be orthogonalised lexicographically, that is,
@@
 
-The comparison with the construction of the 3D spherical harmonics is evident here (the $\genjac$ OPs are in place of the Jacobi polynomials, and the $\jac$ OPs are in place of the complex exponentials, which are just Chebyshev polynomials).
+The comparison with the construction of the 3D spherical harmonics is evident here (the $\genjac_n$ OPs are in place of the Jacobi polynomials, and the $\jac_n$ OPs are in place of the complex exponentials, which are just Chebyshev polynomials).
 By defining our OPs in this way, we will be able to yield sparse and banded relations for the necessary operators involved in PDEs, just as we did before.
@@
 Recall that Jacobi matrices are an important piece of the puzzle for using these orthogonal polynomials to practically solve PDEs in a sparse spectral method.
-To obtain the entries for them, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$.
+To obtain their entries, we first need to establish the three-term recurrences associated with $\genjac_n^{(a,b,c)}$ and $\jac_n^{(d,c)}$.
 These can be expressed as:
@@
 \begin{lemma}
-	$\hdopnkabcd(x,y)$ satisfy the following 3-term recurrences:
+	$\hdopnkabcd(x,y)$ satisfy the following relations:
 \bseqn{
@@
 \begin{proof}
-The 3-term recurrence for multiplication by $x$ follows from \bsrefeqn{eqn:ds:Hrecurrence}.
-For the recurrence for multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$.
+The relation for multiplication by $x$ follows from \bsrefeqn{eqn:ds:Hrecurrence}.
+For multiplication by $y$, since $\{\hdopmjabcd\}$ for $m = 0,\dots,n+1$, $j = 0,\dots,m$ is an orthogonal basis for any degree $n+1$ polynomial, we can expand $y \: \hdopnkabcd(x,y) = \sum_{m=0}^{n+1} \sum_{j=0}^m c_{m,j} \: \hdopmjabcd(x,y)$.
 These coefficients are given by \nomenclature[abs]{$\abs{m}$}{The absolute value of a scalar $m$} \nomenclature[norm]{$\norm{f}$}{The norm of a function f, defined by the square root of the relevant inner product} \nomenclature[normH]{$\norm{f}_{\hdopabcd}$}{The norm of a function f, defined by the square root of the inner product over $\Omega$ given by the weight $\Wabcd$ associated with the OPs $\{\hdopnkabcd\}$}
@@
 
-Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
+These relations lead to Jacobi operators that correspond to multiplication by $x$ and $y$.
 Define, for $n=0,1,2,\dots$: \nomenclature[Hvectorabcd]{$\bighdopabcd$}{Vector of the disk-slice/trapezium OPs ordered by degree, with parameters $a,b,c,d$} \nomenclature[Hvectornabcd]{$\bighdopabcd_n$}{(Sub-)Vector of the disk-slice/trapezium OPs of degree $n$, with parameters $a,b,c,d$}
@@
 In \bsreffig{fig:ds:poisson} we also show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the Poisson equation with $N = 990$, that is, $491,536$ unknowns.
-The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution: as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
+The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution; as typical of spectral methods, we expect the numerical scheme to converge at the same rate as the coefficients decay.
 We see that we achieve algebraic convergence for the first three examples, noting that for right hand-sides that vanish at the corners of our disk-slice ($x\in\{\alpha,\beta\}, \: y = \pm \rho(x)$) we observe faster convergence.
@@
 Since $f$ is known, we can obtain  the coefficients $\vec{f}$ using the quadrature rule above.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices \`a la the procedure introduced in \cite{olver2019triangle}.
 We can denote the resulting operator acting on coefficients in the $\bighdopooo$ space by $V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top)$.
@@
 
+\begin{figure}[t]
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.35]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
+        	%\label{fig:ds:solutionblocknorms-poisson}
+	\centering
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.38]{solution-poisson}
+	\centering
+	%\label{fig:ds:solution-poisson}
+	\end{subfigure}
+	\begin{subfigure}{0.3\textwidth}
+	\includegraphics[scale=0.38]{solution-trapezium-helmholtz-k=100-n=1500}
+	%\label{fig:ds:solution-poisson}
+	\centering
+	\end{subfigure}
+	\centering
+	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
+Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
+Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
+in the trapezium.}
+	\centering
+	\label{fig:ds:trappfem}
+\end{figure}
+
 \subsection{End-Disk-Slice}\label{section:ds:enddiskslice}
@@
 
-Note here how we can remove the need for third parameter, which is why we consider this a special case.
+Note here how we can remove the need for a third parameter, which is why we consider this a special case.
 This will make some calculations easier, and the operator matrices more sparse.
@@
 
-\begin{figure}[t]
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.35]{solution-pfem-poisson-diskslice-alpha=0p2-beta=0p8-f=Wycosx}
-        	%\label{fig:ds:solutionblocknorms-poisson}
-	\centering
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.38]{solution-poisson}
-	\centering
-	%\label{fig:ds:solution-poisson}
-	\end{subfigure}
-	\begin{subfigure}{0.3\textwidth}
-	\includegraphics[scale=0.38]{solution-trapezium-helmholtz-k=100-n=1500}
-	%\label{fig:ds:solution-poisson}
-	\centering
-	\end{subfigure}
-	\centering
-	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = \Wiii(x,y) y \cos(x)$ in the disk-slice using the $p$-FEM approach with a single element.
-Centre: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$ in the half-disk.
-Right: The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = (1-x) \: x \: y \: (1- \half x - y) \: e^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$.
-in the trapezium.}
-	\centering
-	\label{fig:ds:trappfem}
-\end{figure}
-
 It is possible for our framework to be applied to a $p$-finite element method -- that is, one where we can vary the polynomial degree $p$ of the basis functions used in each element (compare this to a normal $h$-FEM, where we can tune the element size $h$).
@@
 
-As eluded to, in general we would let $\FEset$ be the set of elements $\element$ that make up our finite element discretisation of the domain, where each $\element$ is a trapezium or disk slice for example.
+As eluded to, in general we would discretise our domain $\Omega$ with a mesh $\FEset$ consisting of elements $\element$, where each $\element \in \FEset$ is a trapezium or disk slice for example.
 However, here we simply consider our domain to be a disk-slice and our discretisation to be a single element -- that is we let $\element = \Omega$ for a disk-slice domain.
@@
 }
-
 Recall that the OPs $\bighdop^{(a,b,c)}$ are orthogonal with respect to the weight $\Wabc$ on $\Omega$, and define the matrix $\Lambda^{(a,b,c)} :=  \ip< \bighdop^{(a,b,c)}, \: {\bighdop^{(a,b,c)}}^\top >_{\hdopabc}$.
@@
 				\end{pmatrix} \: \D \vec{x}
-				\\
+}
+\bseqn{
 	&= \int_\element \Big( \vec{v}^\top {\Wiii_x}^\top \bighdopooo \bighdopooo^\top \Wiii_x \vec{u} \nonumber \\
@@
 }
-
-Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
+Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector.
 

### 2021-06-18 papers/thesis/submitted/acknowledgements/acknowledgements.tex
Submitted version saved in new dir

--- 2021-06-08 papers/thesis/full/acknowledgements/acknowledgements.tex
@@
 
-I would like to thank the MPE CDT staff for any knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for his guidance, understanding and general enthusiasm (it meant a lot).
+I would like to thank the MPE CDT staff for the knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for his guidance, assistance and general enthusiasm (it meant a lot).
 And to all my fellow \enquote{Charlies} -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.
 
-I thank my family and friends for all their support and words of encouragement -- it all helped a lot.
+I thank my friends for their support and the fun times they have provided me with over these few years.
+I also thank my family for all their support and encouragement, not just during my PhD but throughout my life as well to help me get to this point.
 
-A special thanks goes to my partner Jess for putting up with me and helping me when things got tough during my 4 years at Imperial College London.
+A special thanks goes to my wonderful partner Jess for putting up with me and helping me when things got tough during my time at Imperial College London.
 
@@
 Sheehan Olver -- I cannot thank you enough for all your help, wisdom, advice, kindness, understanding and patience you have given me.
-I may not have expressed my gratitude enough over the past few years, but you have truly been a great supervisor to me.
+I may not have expressed my gratitude enough, but you have truly been a great supervisor to me.
 

### 2021-06-18 papers/thesis/submitted/ch4-spherical-caps/spherical-caps-old.tex
Submitted version saved in new dir

--- 2020-11-18 papers/spherical-caps/spherical-caps.tex
@@
-\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
-\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
-\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
-%\geometry{landscape}                		% Activate for rotated page geometry
-\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
-								% TeX will automatically convert eps --> pdf in pdflatex
-
-\graphicspath{ {images/} }
-
-\include{bsthesisnewcommands}
-\include{bsmacros}
-\include{somacros}
-
-\title{Sparse spectral methods for partial differential equations on spherical caps}
-\author{Ben Snowball, Sheehan Olver\thanks{Department of Mathematics, Imperial College, London, England}}
-%\date{}							% Activate to display a given date or no date
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-In recent years, sparse spectral methods for solving partial differential equations have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, disk-slices and triangles.
-In this work we extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps.
+\chapter{Spherical Caps}\label{CH:SCs}
+
+While the work in the previous chapter looked at developing a sparse spectral method inside a two dimensional domain, we move on to investigating the realm of a surface in three dimensional space.
+Specifically, we look to extend the methodology to a hierarchy of non-classical multivariate orthogonal polynomials on spherical caps, which we will formally define.
 The entries of discretisations of partial differential operators can be effectively computed using formulae in terms of (non-classical) univariate orthogonal polynomials.
 We demonstrate the results on partial differential equations involving the spherical Laplacian and Biharmonic operators, showing spectral convergence.
-\end{abstract}
 
@@
 We can choose a basis for $V_p$ by using the weighted orthogonal polynomials on $\element$ with parameter $a = 1$, the $\bigW_p^{(1)}$ OPs, and we can expand the function $f$ in the $\bigscopi_p$ basis.
-%
-%Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
-%\begin{align*}
-%	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
-%					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
-%	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
-%\end{align*}
-%Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
+
+Let $\mathbf{u}, \mathbf{v}$ be the coefficient vectors of the expansions of $u_p, v_p \in V_p$ respectively in the $V_p$ basis ($\bigWiii$ OPs). Then,
+\begin{align*}
+	a(u_p,v_p) &= \int_\element \big(\rho \: \nabla u\big) \cdot \big(\rho \: \nabla v\big) \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \Big( \rhoppphi u_p \: \rhoppphi v_p + \pptheta u_p \: \pptheta v_p \Big) \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \Big( \mathbf{v}^\top {W_\varphi^{(1)}}^\top \bigscopo_p {\bigscopo_p}^\top W_\varphi^{(1)} \mathbf{u} \nonumber \\
+					& \quad \quad \quad + \mathbf{v}^\top ({T_W^{(1)\to(0)} D_\theta})^\top \bigscopa_p {\bigscopa_p}^\top T_W^{(1)\to(0)} D_\theta \mathbf{u}  \Big) \: \D \sigma(\bm{x}, z) \\
+	&= \mathbf{v}^\top \: \Big( {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta \Big) \: \mathbf{u}.
+\end{align*}
+Further, let $\mathbf{f}$ is the coefficient vector for the expansion of the function $f(x,y,z)$ in the $\bigscopi_p$ OP basis and let $P$ be the operator for multiplication by $\rho(z)^2$, that is the operator matrix defined by $P^\top \bigscopi_p(x,y,z) = \rho(z)^2 \: \bigscopi_p(x,y,z)$.
 Then,
-%\begin{align*}
-%	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
-%	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
-%	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%
-%Since (\ref{eqn:FEMweakform}) is equivalent to stating that
-%\begin{align*}
-%	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
-%\end{align*}
-%(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
-%\begin{align*}
-%	A\mathbf{u} = \tilde{\mathbf{f}}.
-%\end{align*}
-%where the (element) stiffness matrix $A$ is defined by 
-%\begin{align*}
-%	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
-%\end{align*}
-%and the load vector $\tilde{\mathbf{f}}$ is given by 
-%\begin{align*}
-%	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
-%\end{align*}
-%Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).
-%
-
-\bibliography{spherical-caps}
-
-\end{document}
+\begin{align*}
+	L(v_p) &= \int_\element \: v_p \: \rho^2 \: f \: \D \sigma(\bm{x}, z) \\
+	&= \int_\element \: \mathbf{v}^\top \: \bigW_p^{(1)} \: {\bigscopi_p}^\top \: P \: \mathbf{f} \: \D \sigma(\bm{x}, z) \\	
+	&= \mathbf{v}^\top \Lambda^{(1)} \: P \: \mathbf{f}.
+\end{align*}
+
+Since (\ref{eqn:FEMweakform}) is equivalent to stating that
+\begin{align*}
+	L(\genjacw^{(1)} \scopnkii) = a(u_p, \genjacw^{(1)} \scopnkii) \quad \forall \: n = 0,\dots,p, \: k = 0,\dots,n, \: i = 0,1
+\end{align*}
+(i.e. holds for all basis functions of $V_p$) by choosing $v_p$ as each basis function, we can equivalently write the linear system for our finite element problem as:
+\begin{align*}
+	A\mathbf{u} = \tilde{\mathbf{f}}.
+\end{align*}
+where the (element) stiffness matrix $A$ is defined by 
+\begin{align*}
+	A =  {W_\varphi^{(1)}}^\top \Lambda^{(0)}  W_\varphi^{(1)} + ({T_W^{(1)\to(0)} D_\theta})^\top \Lambda^{(0)} T_W^{(1)\to(0)} D_\theta
+\end{align*}
+and the load vector $\tilde{\mathbf{f}}$ is given by 
+\begin{align*}
+	\tilde{\mathbf{f}} = \Lambda^{(1)} \: P \: \mathbf{f}.
+\end{align*}
+Note that since we have sparse operator matrices for partial derivatives and basis-transform, we obtain a symmetric sparse (element) stiffness matrix, as well as a sparse operator matrix for calculating the load vector (rhs).

### 2021-06-18 papers/thesis/submitted/conclusion/summary.tex
Submitted version saved in new dir

--- 2021-06-14 papers/thesis/full/conclusion/summary.tex
@@
 
-In this thesis, we have developed sparse spectral methods for solving partial differential equations (PDEs) on disk-slices and trapeziums in 2D, and spherical caps as a surface in 3D.
+In this thesis, we have developed sparse spectral methods for solving partial differential equations on disk-slices and trapeziums in 2D, and spherical caps as a surface in 3D.
 The work can also be used as a template for developing similar methods on other such similar domains, in particular other subdomains of the sphere.

### 2021-09-28 papers/thesis/resubmitted/abstract/abstract.tex
Resubmitted version of thesis to track the changes for the corrections. Fixed typo in SH clenshaws.

--- 2021-06-14 papers/thesis/full/abstract/abstract.tex
@@
 
-\bstodoinline{New (shortened) abstract - any comments?}
-
 This thesis develops sparse spectral methods for solving partial differential equations (PDEs) on various multidimensional domains, with a specific focus on the disk-slice and trapezium in 2D, and the spherical cap in 3D.
@@
 
-We begin with an introduction to sparse spectral methods via viewing spherical harmonics as multidimensional orthogonal polynomials in $x$, $y$, $z$.
+We begin with an introduction to sparse spectral methods via viewing spherical harmonics as multidimensional orthogonal polynomials in $x$, $y$, and $z$.
 We explain how differential operators can be applied as banded-block-banded matrix operators to coefficient vectors for a function's expansion.
-Further, we demonstrate how vector spherical harmonics in $x$, $y$ and $z$ can be used as an orthogonal basis for vector-valued functions, yielding similar banded-block-banded gradient and divergence operators.
+Further, we demonstrate how vector spherical harmonics in $x$, $y$, and $z$ can be used as an orthogonal basis for vector-valued functions, yielding similar banded-block-banded gradient and divergence operators.
 
@@
 Piecing together the techniques used thus far, we present a new orthogonal polynomial basis and sparse spectral method for the spherical cap, complete with the same observation of the guaranteed sparsity of operators.
-The motivation is for one to use spherical caps bands as in a spectral element method for the sphere, with many applications in meteorology and astrophysics -- in particular, as a potential replacement of the spherical harmonics approach currently in use at ECMWF (which is predicted to become too costly due to a parallel scalability bottleneck arising from the global spectral transform).
+The motivation is for one to use spherical caps bands as in a spectral element method for the sphere, with many applications in meteorology and astrophysics -- in particular, as a potential replacement of the spherical harmonics approach currently in use at ECMWF, which is predicted to become too costly due to a parallel scalability bottleneck arising from the global spectral transform.
 

### 2021-09-28 papers/thesis/resubmitted/acknowledgements/acknowledgements.tex
Resubmitted version of thesis to track the changes for the corrections. Fixed typo in SH clenshaws.

--- 2021-06-18 papers/thesis/submitted/acknowledgements/acknowledgements.tex
@@
 
-I would like to thank the MPE CDT staff for the knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for his guidance, assistance and general enthusiasm (it meant a lot).
+I would like to thank the MPE CDT staff for the knowledge and wisdom they have bestowed, and in particular say a huge thank you to Colin Cotter for your guidance, assistance and general enthusiasm.
 And to all my fellow \enquote{Charlies} -- I truly mean it when I say you are all such great, smart and kind people, and you all deserve the absolute best for the future.

### 2021-09-28 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
Resubmitted version of thesis to track the changes for the corrections. Fixed typo in SH clenshaws.

--- 2021-06-18 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+1} = \vec{0}. \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+1} = \vec{0}. \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \vec{v}_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(J_x,J_y,J_z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
+	\quad &\text{3) } \text{Output: } V = V(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
 }

### 2021-09-28 papers/thesis/resubmitted/ch4-spherical-caps/spherical-caps.tex
Resubmitted version of thesis to track the changes for the corrections. Fixed typo in SH clenshaws.

--- 2021-06-15 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
@@
 We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see \bsreftheorem{theorem:sc:sparsityofdifferentialoperators}.
-Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method we investigated in \bsrefchapter{CHAPTER:diskslice} for disk-slices and trapeziums (see also \cite{snowball2019sparse}), as well as for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
+Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, \`a la the ultraspherical spectral method we investigated in \bsrefchapter{CHAPTER:diskslice} for disk-slices and trapeziums (see also \cite{snowball2019sparse}), as well as for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle}, and the related work on sparse discretisations involving Jacobi polynomials on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.
 Just as we proceeded in \bsrefchapter{CHAPTER:diskslice} for the disk-slice case in 2D (see also \cite{snowball2019sparse}), we will use an integration-by-parts argument to deduce the sparsity structure.
@@
 Recently, a procedure for defining OPs in two variables on quadratic curves, such as ellipses and parabolas, has been presented \cite{olver2021orthogonal}, where the unit circle is one of the five classes of curve that can be generalised to give any curve in the plane $\R^2$.
-One can define OPs on the circle in terms of Fourier series, which we write here as orthogonal polynomials in $x$ and $y$:
+One can define OPs on the circle in terms of Fourier series, which we write here as orthogonal polynomials in $x$ and $y$.
 \begin{definition}[\cite{olver2021orthogonal}]\label{def:sc:Ydefinition}
@@
 
-The incrementing and decrementing of parameters as seen here is analogous to other well known orthogonal polynomial families' derivatives, for example the Jacobi polynomials on the interval, as seen in the DLMF \cite[(18.9.3)]{DLMF}, on the triangle \cite{olver2018recurrence}, and on the disk-slice \cite{snowball2019sparse}.
+The incrementing and decrementing of parameters as seen here is analogous to other to that seen in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}), and we further reiterate that it is also seen for other well known orthogonal polynomial families' derivatives -- for example the Jacobi polynomials on the interval (as seen in the NIST Digital Library of Mathematical Functions \cite[(18.9.3)]{DLMF}) and on the triangle \cite{olver2018recurrence}.
 The operators we define here are for partial derivatives with respect to the spherical coordinates $(\varphi, \theta)$, so that we can more easily apply the operators to PDEs on the surface of a sphere (for example, surface Laplacian operator in the Poisson equation).
@@
 Since the Laplacian is block-diagonal one can determine the condition number from the condition numbers of each block on the diagonal,  and hence deduce the accuracy in solving the Poisson equation, where graphical representations are seen in \bsreffig{fig:sc:conditionnumbers} for the $\Delta^{(1)}_W$ operator matrix.
-As the degree $N$ increases, the condition numbers of these blocks becomes grows  algebraically fast (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
+As the degree $N$ increases, the condition numbers of these blocks grows algebraically fast (at least, the condition number for the first block where $m=0$ does, which will be the largest for each $N$).
 Fortunately however, by applying a trivial diagonal preconditioning matrix similar to that seen for the ultraspherical method \cite{olver2013fast}, we can significantly bound this growth of the condition numbers -- this is also seen in \bsreffig{fig:sc:conditionnumbers}.
@@
 Define $X := (J_x^{(0)})^\top$, $Y := (J_y^{(0)})^\top$, $Z := (J_z^{(0)})^\top$.
-We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}.
+We can obtain the matrix operator for the variable-coefficient function $v(x,y,z)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices $X, Y, Z$, yielding an operator matrix of the same dimension as the input Jacobi matrices \`a la the procedure introduced in \cite{olver2019triangle}.
 We can denote the resulting operator acting on coefficients in the $\bigscopNo$ space by $v(X, Y, Z)$.

### 2021-09-28 papers/thesis/resubmitted/ch3-disk-slices/disk-slices.tex
Fixed typos and added more explanation of BCs in ch3.4

--- 2021-06-18 papers/thesis/full/ch3-disk-slices/disk-slices.tex
@@
 }
-Further, we can represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2)}$ to coefficients in the space $\bighdopooo$.
-Note that a function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
+Further, we can represent the Laplacian as a map from coefficients in the space $\bigW^{(2,2,2)}$ to coefficients in the space $\bighdopooo$.
+Note that a function expanded in the $\bigW^{(2,2,2)}$ basis will satisfy both zero Dirichlet and Neumann boundary conditions on $\Omega$.
 We denote this matrix operator as $\laplacewttt$, and is given by
@@
 We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions.
-We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, demonstrating the versatility of the approach.
+We consider the Poisson equation, an inhomogeneous variable coefficient Helmholtz equation and the Biharmonic equation, demonstrating the versatility of the approach.
+
+As mentioned earlier, we can handle these boundary conditions by incorporating them into the chosen basis OP family.
+The weight $\Wiii(x,y)$ is of course zero on the boundary of $\Omega$, and in fact any polynomial that vanishes on the boundary will have $\Wiii(x,y)$ as a factor.
+Hence, any function expanded in the $\bigWiii$ basis will satisfy zero Dirichlet conditions.
+The same argument can be used to see why any function expanded in the $\bigW^{(2,2,2)}$ basis will satisfy zero Neumann conditions too.
 
@@
 We can tackle the problem as follows.
-Denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\vec{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ by $\vec{f}$.
+Since $u$ vanishes on the boundary, any polynomial expansion will have $\Wiii(x,y)$ as a factor, and so we can denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\vec{u}$, while the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ we can denote by $\vec{f}$.
 Since $f$ is known, we can obtain $\vec{f}$ using the quadrature rule above.

### 2021-09-28 papers/thesis/resubmitted/ch1-introduction/introduction.tex
Fixed typos and (most) notes from AT corrections for intro

--- 2021-06-18 papers/thesis/full/ch1-introduction/introduction.tex
@@
 \cite{trefethen2000spectral, canuto2007spectral, gottlieb1977numerical, boyd2001chebyshev, mason2002chebyshev, shen2011spectral, olver2013fast}).
-While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[\S18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
+While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, with special cases of the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[\S18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
 
@@
 Koornwinder in 1975 described a method for constructing two-variable OPs from univariate OPs \cite{koornwinder1975two}.
-However, few books have been published over the years on the topic of multivariate orthogonal polynomials, with the notable exception of the book by Dunkl \& Xu \cite{dunkl2014orthogonal}.
+However, few books have been published over the years on the topic of multivariate orthogonal polynomials, with the notable exception of the book by Dunkl and Xu \cite{dunkl2014orthogonal}.
 As Dumitriu, Edelman and Shuman said in 2007, \enquote{[multivariate orthogonal polynomials] are understudied, underapplied, and important applications may be being missed} \cite{dumitriu2007mops}, while in 2016, Olver opined that \enquote{there is still a long way to go before multivariate orthogonal polynomials can reach their full potential in applications}.
@@
 
-Spherical harmonics (SHs) are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere and form a basis for expanding functions defined on the sphere) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
+Spherical harmonics (SHs) are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere, have sparse recurrences, and are eigenfunctions of the spherical Laplacian) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
 \cite{moon2008efficient, sloan2013efficient}), astrophysics (e.g.
@@
 In other words, we can use our spectral methods developed for the spherical band and cap elements as part of a finite element framework.
-By using this approach, one would be able to avoid having to complete the global spectral transform (in particular, the global Legendre transform) and instead be able to simply apply the local element transforms in parallel.
+By using this approach, one can avoid having to compute the global spectral transform (in particular, the global Legendre transform) and instead simply apply the local element transforms in parallel.
 Moreover, by still using a spectral approach, one can maintain the high accuracy and excellent error properties that such methods bring.
@@
 Their method involves using a spectral basis to represent functions too, choosing for the angular part of the basis to be the spin-weighted spherical harmonics (of which the spherical harmonics are a special case, with spin $0$) in spherical coordinates.
-By using these as their basis functions, they are able to derive sparse relations for the Laplacian operator, as well as for multiplication of the basis by $\cosphi$ and $\sinphi$ (where $\varphi$ is the polar angle from the $z$ axis in cartesian coordinates) which lead to operators for operations with angular dependencies involving these.
+By using these as their basis functions, they are able to derive sparse relations for the Laplacian operator, as well as for multiplication of the basis by $\cosphi$ and $\sinphi$ (where $\varphi$ is the polar angle from the $z$ axis in cartesian coordinates), which lead to operators for operations with angular dependencies involving these.
 

### 2021-09-28 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
Added TJx etc. to notation index

@@
 
-The Jacobi matrices are then defined according to the matrices $\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$ satisfying
+The Jacobi matrices are then defined according to the matrices \nomenclature[JT]{$\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$}{The Jacobi matrices for the tangent bundle VSH basis} $\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$ satisfying
 \bseqnnumber{

### 2021-09-28 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
Fixed further typos and corrections from AT notes

@@
 To introduce ourselves to the world of multidimensional orthogonal polynomials for solving PDEs on the sphere, we can naturally choose to look at the famous spherical harmonics.
-Our aim here is to express the spherical harmonics as polynomials in three variables $x, y, z$ to evaluate functions and solve PDEs on the whole sphere.
+Our aim here is to express the spherical harmonics as polynomials in three variables $x$, $y$, and $z$ to evaluate functions and solve PDEs on the whole sphere.
 More precisely, we desire the solution to partial differential equations on the domain \nomenclature[Omega]{$\Omega$}{The domain of interest}
 \bseqn{
-	\Omega := \bsset{(x,y,z) \in \R^3}{x^2 + y^2 = \rho(z)^2}
+	\Omega := \bigS^2 = \bsset{(x,y,z) \in \R^3}{x^2 + y^2 = \rho(z)^2}
 }
@@
 }
-While it may seem somewhat odd, it should hopefully become apparent why this is a practical way to define our domain.
+While it may seem somewhat odd, it should hopefully become apparent why using $\rho$ here is a practical way to define our domain.
 
 It is useful to be able to transform between these cartesian coordinates and the spherical coordinates \nomenclature[phi]{$\varphi$}{The polar angle in spherical coordinates} \nomenclature[theta]{$\theta$}{The azimuthal angle in spherical coordinates}($\varphi, \theta$).
-Throughout this thesis we will use the convention that the spherical coordinate angles be defined by \nomenclature[xyz]{$x,y,z$}{Lower case non-bold $x,y,z$ represents cartesian coordinates}
-\bseqn{
-	x &= \sinphi \costheta = \rho(z) \costheta \\
-	y &= \sinphi \sintheta = \rho(z) \sintheta \\
+Throughout this thesis we will use the convention that the spherical coordinate angles be defined by \nomenclature[xyz]{$x,y,z$}{Lower case non-bold $x$, $y$, and $z$ represent cartesian coordinates}
+\bseqn{
+	x &= \sinphi \costheta = \rho(z) \costheta, \\
+	y &= \sinphi \sintheta = \rho(z) \sintheta, \\
 	z &= \cosphi
@@
 
-In this chapter, we demonstrate how one can write the spherical harmonics as polynomials in $(x, y, z)$, and how the relationships between them can be used to obtain sparse \enquote{Jacobi operators} for multiplication by $x,y,z$.
+In this chapter, we demonstrate how one can write the spherical harmonics as polynomials in $(x, y, z)$, and how the relationships between them can be used to obtain sparse \enquote{Jacobi operators} for multiplication by $x$, $y$, and $z$.
 We demonstrate how these in turn lead to an algorithm for evaluating the spherical harmonics, an efficient algorithm for evaluating functions when expanded in the SH OP basis, and sparse \enquote{banded-block-banded} differential operator matrices.
-The techniques used here are done so with the aim of applying them to other OP families of other domains (and we do so in later chapters).
+The techniques used here are done so with the aim of applying them to other OP families on other domains (and we do so in later chapters).
 We will finally demonstrate a proof-of-concept via the examples of the heat equation and the linearised shallow water equations on the unit sphere.
@@
 We first need to introduce a few classical orthogonal polynomials.
-Orthogonal polynomials are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product.
+Orthogonal polynomials are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product, and ordered by degree.
 In particular, they are orthogonal with respect to all lower degree polynomials.
 
-On the unit interval, $[-1,1]$, we note that there is a hierarchy of OPs in the sense that \cite[table 18.3.1, eqn 18.9.15]{DLMF}: \nomenclature[Pnab]{$P^{(a,b)}_n$}{The degree $n$ Jacobi polynomial (context determines normalized or not) unless otherwise stated}
+On the unit interval, $[-1,1]$, we note that there is a hierarchy of OPs in the sense that \cite[Table 18.3.1, (18.9.15)]{DLMF}: \nomenclature[Pnab]{$P^{(a,b)}_n$}{The degree $n$ Jacobi polynomial (context determines normalized or not) unless otherwise stated}
 \bseqn{
@@
 }
-where $P^{(a,b)}_l (x)$ is the $l$ degree \textit{Jacobi polynomial}, and \nomenclature[Pl]{$P_l$}{The degree $l$ Legendre polynomial} $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$.
-Jacobi polynomials are orthogonal with weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$ \nomenclature[Kronecker delta]{$\delta_{n, n'}$}{The Kronecker delta function; $1$ if $n=n'$, else $0$}
-\bseqn{
-	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)} \\
-	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'}.
-}
+where $P^{(a,b)}_l (x)$ is the degree $l$ \textit{Jacobi polynomial} with $a, b > -1$, and \nomenclature[Pl]{$P_l$}{The degree $l$ Legendre polynomial} $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$.
+Jacobi polynomials are orthogonal with respect to the weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$ \nomenclature[Kronecker delta]{$\delta_{n, n'}$}{The Kronecker delta function; $1$ if $n=n'$, else $0$}
+\bseqn{
+	\int_{-1}^1 {P^{(a,b)}_l(x)}^2 \:(1-x)^a(1+x)^b \: \D x &=: \omega_{P,l}^{(a,b)}, \\
+	\int_{-1}^1 P^{(a,b)}_l(x) \: P^{(a,b)}_{l'}(x) \:(1-x)^a(1+x)^b \: \D x &= \omega_{P,l}^{(a,b)} \: \delta_{l, l'},
+}
+where $\delta_{l, l'}$ is the regular Kronecker delta function. For the Jacobi polynomials,
+\bseqn{
+	 \omega_{P,l}^{(a,b)} = {2^{a + b + 1} \; \Gamma(l + a + 1) \; \Gamma(l + b + 1) \over (2l + a + b + 1) \; \Gamma(l + a + b + 1) \; \Gamma(l + 1)}
+}
+where $\Gamma(x)$ is the regular Gamma function \cite[(5.2.1)]{DLMF} (with $\Gamma(n) = (n-1)!$ for $n \in \N$).
 Legendre polynomials are special cases of \textit{Legendre functions} when $l$ is a non-negative integer.
-Legendre functions are a class of univariate functions that are solutions to the differential equation
-\bseqn{
-	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x} y + l(l+1)y = 0
-}
-which is helpfully known as the \textit{Legendre equation} \cite[14.2.1]{DLMF}.
-Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p5]{belousov1962tables}: \nomenclature[Plm]{$\Plm$}{The Associated Legendre polynomial of degree $l$ and order $m$}
+Legendre functions are a class of univariate functions that are one set of solutions to the differential equation
+\bseqn{
+	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x} y + l(l+1)y = 0, \quad x \in [-1,1]
+}
+which is helpfully known as the \textit{Legendre equation} \cite[(14.2.1)]{DLMF}.
+Further, the \textit{associated Legendre polynomials} are a set of polynomials orthogonal with respect to unit weight on the unit interval, and can be defined as the $m^{\text{th}}$ derivative of a Legendre polynomial \cite[p.5]{belousov1962tables}: \nomenclature[Plm]{$\Plm$}{The Associated Legendre polynomial of degree $l$ and order $m$}
 \bseqn{
@@
 \bseqn{
-	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x}y + \Big[l(l+1) - {m^2 \over 1-x^2} \Big] y = 0
-}
-which is also helpfully known as the \textit{associated Legendre equation} (\cite[p2]{belousov1962tables}, \cite[14.2.2]{DLMF}).
+	(1 - x^2) \dmdxm{x}{2}y - 2x \ddx{x}y + \Big[l(l+1) - {m^2 \over 1-x^2} \Big] y = 0,
+}
+which is also helpfully known as the \textit{associated Legendre equation} (\cite[p.2]{belousov1962tables}, \cite[(14.2.2)]{DLMF}).
 We can see that Legendre functions (and polynomials) are just special cases of the associated versions when $m = 0$.
@@
 Let $(x,y,z) \in \Omega$.
-We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[14.30.1]{DLMF}: \nomenclature[Ylm]{$\Ylm$}{The degree $l$ and order $m$ spherical harmonic}
+We will use the standard definition -- that is, the spherical harmonics, orthonormal on the unit sphere, are \cite[(14.30.1)]{DLMF}: \nomenclature[Ylm]{$\Ylm$}{The degree $l$ and order $m$ spherical harmonic}
 \bseqnnumber{
@@
 			(-1)^m &\text{if } m \ge 0 \\
-			1 &\text{if } m < 0
+			1 &\text{if } m < 0,
 		\end{cases} \label{eqn:clmdef}
@@
 			(-1)^m &\text{if } m \ge 0 \\
-			1 &\text{if } m < 0
-		\end{cases}.
+			1 &\text{if } m < 0.
+		\end{cases}
 }
@@
 \bseqn{
-	&\int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
-	& \quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
-	& \quad = \delta_{l, l'} \: \delta_{m, m'}
+	&\iint_{\Omega} \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \D S \\
+	&\quad = \int_0^{2\pi} \int_0^\pi \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \sinphi \: \D \varphi \: \D \theta \\
+	&\quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
+	&\quad = \delta_{l, l'} \: \delta_{m, m'}
 }
 where $\alpha^*$ denotes the complex conjugate of \nomenclature[C]{$\C$}{The complex numbers} $\alpha \in \C$.
-Note how we can express the spherical harmonics $\Ylm$ in terms of $x,y,z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x,y,z$ for any \nomenclature[Z]{$\Z$}{The integers} $m \in \Z$.
-Indeed, they are polynomials in $x,y,z$ which we denote $\Ylm(x,y,z)$.
+Note how we can express the spherical harmonics $\Ylm$ in terms of $x$, $y$, and $z$ instead of $\varphi, \theta$ by noting that $\rho(z)^{|m|} \eimtheta$ can be expressed in terms of $x$, $y$, and $z$ for any \nomenclature[Z]{$\Z$}{The integers} $m \in \Z$.
+Indeed, they are polynomials in $x$, $y$, and $z$, which we denote $\Ylm(x,y,z)$.
 They span all polynomials modulo the vanishing ideal associated by the zero set of $x^2 + y^2 + z^2 - 1$ in \nomenclature[R]{$\R$}{The real numbers} \nomenclature[Rd]{$\R^d$}{The d-dimensional coordinate space over the real numbers} $\R^3$.
@@
 
-Jacobi operators are matrices that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$.
+Jacobi operators are matrices that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x$, $y$, and $z$.
 We can obtain the entries to these operator matrices by determining the coefficients of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.
 Relations concerning the products of spherical harmonics have already been established, and so the expressions we desire here could just be seen as special cases of these.
-More specifically, the coefficients of the expansion in the SH basis of the product of two SHs are composed of Clebsch-Gordan coefficients \cite[p231]{sakurai1995modern}.
-Clebsch-Gordan coefficients are important in quantum mechanics and angular momentum, with tables and algorithms existing to calculate them \cite{mcnamee1964tables}.
+More specifically, the coefficients of the expansion in the SH basis of the product of two SHs are composed of Clebsch--Gordan coefficients \cite[p.231]{sakurai1995modern}.
+Clebsch--Gordan coefficients are important in quantum mechanics and angular momentum, with tables and algorithms existing to calculate them \cite{mcnamee1964tables}.
 By simply choosing appropriate combinations, one could gain the desired coefficients for the expansion of each of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$.
@@
 \begin{proof}
-	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5, 18.9.6]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
-The final one is simply a rearranging of the established three-term recurrence detailed in \cite[18.9.1, 18.9.2]{DLMF}
+	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[(18.9.5), (18.9.6)]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
+The final one is obtained by rearranging the three-term recurrence in \cite[(18.9.1), (18.9.2)]{DLMF}
 \end{proof}
@@
 
-\remark The results in \bsreflemma{lemma:shxyzrelations} can be derived via the relationships established for products of spherical harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
+\remark The results in \bsreflemma{lemma:shxyzrelations} can be derived via the relationships established for products of spherical harmonics and Clebsch-Gordan coefficients as mentioned earlier \cite[p.231]{sakurai1995modern} but we include this proof so as to generalise to other OPs later on in later chapters.
 
@@
 
-These recurrences lead to Jacobi operators that correspond to multiplication by $x,y,z$ of the OPs.
+These recurrences lead to Jacobi operators that correspond to multiplication by $x$, $y$, and $z$ of the OPs.
 Define for $l \in \No$: \nomenclature[Pvector]{$\SHvec$}{Vector of the SH OPs} \nomenclature[Pvectorl]{$\SHvec_l$}{(Sub-)Vector of the degree $l$ SH OPs}
@@
 Then $x f(x,y,z)$ is approximated by $\SHvec(x,y,z)^\top \: J_x^\top \: \fvec$.
-In other words, $J_x^\top \: \fvec$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the spherical harmonics basis.
+In other words, $J_x^\top \: \fvec$ is the coefficient vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the spherical harmonics basis.
 
@@
 	\end{subfigure}\hfil % <-- added
-	\caption{\enquote{Spy} plots of the Jacobi operator matrices, showing their sparsity and structure, up to order $N=10$.
+	\caption{\enquote{Spy} plots of the Jacobi operator matrices, showing their sparsity and structure, up to order $N=10$ (i.e.
+$(N+1)^2$ unknowns).
+Dots represent non-zero entries of the operator.
 Each are block-tridiagonal, with the blocks corresponding to polynomial degree.}
@@
 		\end{pmatrix} \in \C^{3(2l+1)\times(2l+1)}, \quad
-	G_n(x,y) := 
+	G_l(x,y,z) := 
 		\begin{pmatrix}
@@
 
-For each $l = 0,1,2\dots$ let $\Dlt$ be any matrix that is a left inverse of $A_l$, i.e.
-such that $\Dlt A_l = \idmat{2l+3}$.
+For each $l = 0,1,2\dots$ let $\Dlt$ be any matrix that is a left inverse of $A_l$, i.e., such that $\Dlt A_l = \idmat{2l+3}$.
 Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[\S3.2]{dunkl2014orthogonal}, which can be expanded to obtain the three-term recurrence, for $l = 0,1,2,\dots$:
@@
 	\begin{cases}
-		\SHvec_{-1}(x,y,z) := 0 \\
-		\SHvec_{0}(x,y,z) := Y_0 \\
-		\SHvec_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z)
+		\SHvec_{-1}(x,y,z) := 0, \\
+		\SHvec_{0}(x,y,z) := Y_0, \\
+		\SHvec_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \SHvec_l(x,y,z) - \Dlt C_l  \,\SHvec_{l-1}(x,y,z).
 	\end{cases}
 }
-
-Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices in the following way. For $l \in \N$, we can set
+Since the above holds for any $\Dlt$ that is a left inverse of $A_l$, we are free to choose the $\Dlt$ matrices.
+In order to attempt to maximise sparsity, we define them in the following way\footnote{There may be a more computationally efficient way of defining these matrices, but in practice we only desire the output of the action of $\Dlt C_l$ and $\Dlt (B_l-G_l(x,y,z))$, and so minimising the resulting number of operations is the goal.
+The choices of the $\Dlt$ matrices we make in this thesis are designed to do this, but whether there are more computationally efficient ways of defining them is an open question.}.
+For $l \in \N$, we can set
 \bseqnnumber{
@@
 		\begin{pmatrix}
-			& & & & & \vec{\eta}^\top_{0} \\
+			& & & & \vec{\eta}^\top_{0} \\
 			0 & & & 0 & & & {1 \over \gamma_{l,-l,2}} & \\
@@
 			& & 0 & & & 0 & & & {1 \over \gamma_{l,l,2}} \\
-			& & & & & \vec{\eta}^\top_{1} 
-		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)}, \label{eqn:Dltdef}
-}
-where $\vec{\eta}_{0}, \vec{\eta}_{1} \in \R^{3(2n+1)}$ are vectors with entries given by 
+			& & & & \vec{\eta}^\top_{1} 
+		\end{pmatrix} \in \R^{(2l+3)\times3(2l+1)}, \label{eqn:Dltdef}
+}
+where $\vec{\eta}_{0}, \vec{\eta}_{1} \in \R^{3(2l+1)}$ are vectors with entries given by 
 \bseqn{
@@
 			{1 \over \alpha_{l,-l,3}} & j = 1 \\
-			{- \; \alpha_{l,-l,4} \over \alpha_{l,-l,3} \: \gamma_{l,1-l,2}} & j = 2(2n+1) + 2 \\
+			{- \; \alpha_{l,-l,4} \over \alpha_{l,-l,3} \: \gamma_{l,1-l,2}} & j = 2(2l+1) + 2 \\
 			0 & o/w
@@
 			{1 \over \alpha_{l,l,4}} & j = 2l+1 \\
-			{- \; \alpha_{l,l,3} \over \alpha_{l,l,4} \: \gamma_{l,l-1,2}} & j = 3(2n+1) - 1 \\
+			{- \; \alpha_{l,l,3} \over \alpha_{l,l,4} \: \gamma_{l,l-1,2}} & j = 3(2l+1) - 1 \\
 			0 & o/w
@@
 
-In spectral space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
+In coefficient space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
 Naturally, we of course need a way to obtain the coefficients $f_{l,m}$.
@@
 \bseqn{
-	f_{l,m} = \int_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta
-}
+	f_{l,m} = \iint_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta
+}\bstodo{the complex conj should be included here, shouldn't it?}
 using the orthonormality of the spherical harmonics.
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+1} = \vec{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+1} = \vec{0} \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = \vec{v}_n^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(J_x,J_y,J_z)) - \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V = V(x,y,z) \approx \vec{\xi}_{0}^T \SHvec_0(x,y,z) \equiv \xi_0 Y_0
-}
-where at each iteration, $\vec{\xi}_n$ is a vector of matrices (note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).
+	\quad &\text{3) } \text{Output: } V = V(x,y,z) \approx \xi_0 Y_0
+}
+where at each iteration, $\vec{\xi}_n$ is a vector of matrices\footnote{It may help to think of this symbolically.
+At each iteration, the entries of $\vec{\xi}_n$ are polynomials in $J_x$, $J_y$ and $J_z$.
+On the final step, $\vec{\xi}_{0} = \xi_{0}$ is hence some polynomial expression in $J_x$, $J_y$ and $J_z$, which we can then evaluate by inserting the Jacobi matrices.} (note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).
 
@@
 To this end, we recall and derive some important relations of the complex exponential and Jacobi polynomials that are a part of the spherical harmonics definitions.
-In particular we first note that:
+In particular, we first note that:
 \bseqnnumber{
@@
 }
-\bsrefeqncapital{eqn:normjac} and \bsrefeqn{eqn:diffjac} are simply consequences of the definition of the Jacobi polynomials -- see \cite[Table 18.3.1]{DLMF} and \cite[18.9.15]{DLMF}.
+\bsrefeqncapital{eqn:normjac} and \bsrefeqn{eqn:diffjac} are simply consequences of the definition of the Jacobi polynomials -- see \cite[Table 18.3.1]{DLMF} and \cite[(18.9.15)]{DLMF}.
 
@@
 
-Finally, to complete our proof, for multiplication by $x,y,z$ of the other VSHs $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
+Finally, to complete our proof, for multiplication by $x$, $y$, and $z$ of the other VSHs $\gradpYlm$, simply take the cross product by $\unitvec{r}$ of $x \gradYlm$, etc.
 noting that $\rvec \times \gradYlm = \gradpYlm$ and $\rvec \times \gradpYlm = - \norm{\rvec} \gradYlm = - \gradYlm$ by the triple vector product and orthogonality.
@@
 		\end{pmatrix} \in \C^{6(2l+1)\times2(2l+1)}, \quad
-	G_n(x,y) := 
+	G_l(x,y,z) := 
 		\begin{pmatrix}
@@
 such that $\Dlt A_l = \idmat{2(2l+3)}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dlt$'s, we obtain a lower triangular system \cite[p.78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
 \bseqn{
@@
 			& 0 & & & \\
-			& & \ddots & & \\
-			& & & 1 & \\
-			& & & & 0
+			& & 1 & & & & \\
+			& & & 0 & & & \\
+			& & & & \ddots & & \\
+			& & & & & 1 & \\
+			& & & & & & 0
 		\end{pmatrix}  \in \R^{2(N+1)^2 \times 2(N+1)^2}, \\
@@
 	\centerline{\includegraphics[scale=0.5]{heat-eqn-alpha-0p0238-dt=0p01-nsteps=100-2}}
-	\caption{Four snapshots of the solution $u$ of the heat equation with $\alpha = 1/42$ and a step size of $\Delta t = 0.01$, solved using the BDF2 timestepping method.
+	\caption{Four snapshots of the solution $u$ of the heat equation with $\alpha = 1/42$ and a step size of $\Delta t = 0.01$, solved using the BDF2 timestepping method and order $N = 20$.
 Top Left: Initial conditions (0 timesteps).
@@
 	\centerline{\includegraphics[scale=0.5]{linear-swe-H=1-dt=0p01-nsteps=100-arrows-2}}
-	\caption{Four snapshots of the solution for the height $h$ (surface colour) and the wave velocity $u$ (red arrows) of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method (BDF1).
+	\caption{Four snapshots of the solution for the height $h$ (surface colour) and the wave velocity $u$ (red arrows) of the linearised shallow water equations with $\rotrate = \calH = 1$ and a step size of $\Delta t = 0.01$, solved using the backwards Euler timestepping method (BDF1) and order $N = 20$.
 The initial conditions are given in \bsrefeqn{eqn:sweics}.

### 2021-09-30 papers/thesis/resubmitted/ch1-introduction/introduction.tex
Added general framework to the intro (draft)

@@
 
-The structure of this thesis is as follows.
+% GENERAL FRAMEWORK
+The construction of the spectral methods in this thesis follow a similar pattern that is worth detailing here as a general framework.
+\begin{itemize}
+	\item \textit{Define the domain of interest via a restriction of one coordinate in terms of the remainder, that uses a function $\rho$ that satisfies one of two conditions.}
+
+		For example, for a domain in 2D space, we would define this as $\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)}$.
+
+	\item \textit{Express the multivariate and multi-parameter OPs as polynomials in cartesian coordinates.}
+
+		The construction of the multivariate OPs takes its inspiration from the Koornwinder approach \cite{koornwinder1975two}.
+These OPs will be $d$-parameter, meaning that we can construct parameter lowering and raising operators for conversion between the OP bases.
+
+	\item \textit{Derive expressions for multiplication of the OPs by the coordinates.}
+
+	\item \textit{Use these relations to determine the entries to \enquote{Jacobi matrices} $J_x$, etc.
+that represent multiplication by the coordinates when applied to the degree-ordered vector of OPs $\bigP$.}
+
+		These Jacobi matrices will be banded-block-banded in structure. In fact, due to the construction, they will be block-tridiagonal:
+		\bseqn{
+			J_{x/y} &= 
+			\begin{pmatrix}
+				B_{x/y,0} & A_{x/y,0} & & & & \\
+				C_{x/y,1} & B_{x/y,1} & A_{x/y,1} & & & \\
+				& C_{x/y,2} & B_{x/y,2} & A_{x/y,2}  & & & \\
+				& & C_{x/y,3} & \ddots & \ddots & \\
+				& & & \ddots & \ddots & \ddots \\
+			\end{pmatrix}.
+		}
+		The degree-ordered vector of OPs, $\bigP$, can be thought of as being made up of ordered sub-vectors grouped by polynomial degree, $\bigP_n$ for $n \in \No$.
+
+	\item \textit{Create a three-term recurrence for the vector of degree $n$ OPs by combining the systems $J_x \bigP = x \bigP$, etc.}
+
+		The resulting three-term recurrence will have the form:
+		\bseqn{
+			\bigP_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigP_n(x,y) - \Dnt C_n  \, \bigP_{n-1}(x,y)
+		}
+		Here, the matrices $B_n$, $C_n$ and $G_n$ are defined via:
+		\bseqn{
+                    	C_n := 
+                    		\begin{pmatrix}
+                    			C_{x,n} \\
+	                    		C_{y,n}
+				\end{pmatrix} \: (n \ne 0), \quad
+			B_n &:= 
+				\begin{pmatrix}
+                    			B_{x,n} \\
+	                    		B_{y,n}
+				\end{pmatrix}, \quad
+			G_n(x,y) := 
+                    		\begin{pmatrix}
+					xI_{n+1} \\
+	                    		yI_{n+1}
+				\end{pmatrix}.
+		}
+		The matrices $\Dnt$ are left-inverses of the matrices $A_n$ respectively, which are defined similarly:
+		\bseqn{		
+			A_n &:=
+				\begin{pmatrix}
+                    			A_{x,n} \\
+                    			A_{y,n}
+				\end{pmatrix}, \quad
+		}
+
+	\item \textit{Use the multivariate version of the Clenshaw algorithm for evaluation of functions expanded in the OP basis.}
+
+	\item \textit{Derive (sparse) relationships for various differential operations, and relations for incrementing and decrementing the parameter values of the OPs.
+Hence construct the sparse operator matrices representing these operations.}
+
+		The differential operators may raise or lower parameter values, which is why the basis conversion operator matrices are required. Further operators may be constructed as combinations of others.
+
+	\item \textit{Use the \enquote{operator Clenshaw algorithm} to construct operator matrices for the action of multiplying by a general function.}
+
+		For instance, this would be useful for a variable coefficient in a Helmholtz example.
+\end{itemize}
+
+The structure of the remainder of thesis is as follows.
 

### 2021-10-11 papers/thesis/resubmitted/ch4-spherical-caps/spherical-caps.tex
Fixed typos. Added better explanation of operator clenshaw in CH4.

@@
 \bseqn{
-	{\vec{f}}_n &:= 
+	\tilde{\vec{f}}_n &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{2(N+1)} \quad \text{for } n = 1,2,\dots,N, \quad
-	{\vec{f}}_0 = f_{0,0,0} \in \R, \\
-	{\vec{f}} &:= 
-		\begin{pmatrix}
-			{\vec{f}}_0 \\
+	\tilde{\vec{f}}_0 = f_{0,0,0} \in \R, \\
+	\tilde{\vec{f}} &:= 
+		\begin{pmatrix}
+			\tilde{\vec{f}}_0 \\
 			\vdots \\
-			{\vec{f}}_N
+			\tilde{\vec{f}}_N
 		\end{pmatrix} \in \R^{(N+1)^2}.
@@
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
-	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {{\vec{f}}_n}^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
+	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {\tilde{\vec{f}}_n}^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
 	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
@@
 \bseqn{
-	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \vec{f} \: \bigscopta(x,y,z) = (V \vec{f})^\top \: \bigscopta(x,y,z),
-}
-i.e. $V \vec{f}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$.
-
-To this end, let $\tilde{\vec{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in \bsrefsection{section:sc:functionevaluation} so that $v(x,y,z) = {\vec{v}}^\top \: \bigscopta(x,y,z)$).
-Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, $Z := (J_z^{(a)})^\top$.
-The operator $V$ is then the result of the following:
+	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: \tilde{\vec{f}} \: \bigscopta(x,y,z) = (\tilde V \tilde{\vec{f}})^\top \: \bigscopta(x,y,z),
+}
+i.e. $\tilde V \tilde{\vec{f}}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$ in the $\bigscopta$ basis.
+
+To this end, let $\tilde{\vec{v}}$ be the coefficients for expansion up to order $N$ in the $\{\scopnki\}$ basis of $v$ (rearranged as in \bsrefsection{section:sc:functionevaluation} so that $v(x,y,z) = \tilde{\vec{v}}^\top \: \bigscopta(x,y,z)$).
+Denote $X := (J_x^{(a)})^\top$, $Y := (J_y^{(a)})^\top$, and $Z := (J_z^{(a)})^\top$.
+The operator $\tilde V$ is then the result of the following:
 \bseqn{
@@
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {\tilde{\vec{v}}_n}^T - \vec{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } V = v(X, Y, Z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
-}
-where at each iteration, $\vec{\xi}_n$ is a vector of matrices.
+	\quad &\text{3) } \text{Output: } \tilde V = v(X, Y, Z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+}
+where at each iteration, $\vec{\xi}_n$ is a vector of matrices\footnote{It may help to think of this symbolically.
+At each iteration, the entries of $\vec{\xi}_n$ are polynomials in $X$, $Y$, and $Z$.
+On the final step, $\vec{\xi}_0 = \xi_{0}$ is hence some polynomial expression in $X$, $Y$, and $Z$, which we can then evaluate by inserting the Jacobi matrices.}.
+Of course, by using rearranged versions of the Jacobi matrices (that act on coefficients $\vec{f}$ that expand the function $f$ in the $\bigscopa$ basis), the same algorithm would return the operator $V$ so that
+\bseqn{
+	v(x,y,z) \: f(x,y,z) = v(x,y,z) \: {\vec{f}} \: \bigscopa(x,y,z) = (V {\vec{f}})^\top \: \bigscopa(x,y,z),
+}
+i.e. $V {\vec{f}}$ is the coefficients vector for the function $v(x,y,z) \: f(x,y,z)$ in the $\bigscopa$ basis.
 
@@
 We now demonstrate how the sparse linear systems constructed as above can be used to efficiently solve PDEs with zero Dirichlet conditions on the spherical cap defined by $\Omega$.
-We consider Poisson, inhomogeneous variable coefficient Helmholtz equation and the biharmonic equation, as well as a time dependent heat equation, demonstrating the versatility of the approach.
+We consider Poisson, an inhomogeneous variable coefficient Helmholtz equation, and the biharmonic equation, as well as a time dependent heat equation, demonstrating the versatility of the approach.
 

### 2021-10-11 papers/thesis/resubmitted/ch3-disk-slices/disk-slices.tex
Made a portion more consistent with rest of thesis
Fixed typos for Jac mats defs. Added brackets to eqn numbers in citations

@@
 Sparsity comes from expanding the domain and range of an operator using different choices of the parameters $a$, $b$ and $c$.
-Whereas the sparsity pattern and entries derived for equations on the triangle \cite{olver2018recurrence, olver2019triangle} and for equations on the disk \cite{vasil2016tensor} with the frameworks discussed earlier both result from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[18.2.2]{DLMF} and quadrature rules to determine the entries.
+Whereas the sparsity pattern and entries derived for equations on the triangle \cite{olver2018recurrence, olver2019triangle} and for equations on the disk \cite{vasil2016tensor} with the frameworks discussed earlier both result from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside careful use of the Christoffel-Darboux formula \cite[(18.2.12)]{DLMF} and quadrature rules to determine the entries.
 In particular, by exploiting the connection with one-dimensional orthogonal polynomials we can construct discretizations of general partial differential operators of size $p(p-1)/2 \times p(p-1)/2$ in $O(p^3)$ operations, where $p$ is the total polynomial degree.
@@
 
-\begin{proposition}[{\cite[p55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
+\begin{proposition}[{\cite[p.55--56]{dunkl2014orthogonal}}]\label{prop:ds:construction}
 	Let $w_1 : (\alpha,\beta) \: \to \R$, $w_2 : (\gamma,\delta) \: \to \R$ be weight functions with $\alpha,\beta,\gamma,\delta \in \R$, and let $\rho \: : \: (\alpha,\beta) \: \to (0,\infty)$ be such that either \bsrefcondition{cond:trap} or \bsrefcondition{cond:disk} with $w_2$ being an even function hold.
@@
 		\begin{pmatrix}
-			B^{x/y}_0 & A^{x/y}_0 & & & & \\
-			C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
-			& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
-			& & C^{x/y}_3 & \ddots & \ddots & \\
+			B_{x/y,0} & A_{x/y,0} & & & & \\
+			C_{x/y,1} & B_{x/y,1} & A_{x/y,1} & & & \\
+			& C_{x/y,2} & B_{x/y,2} & A_{x/y,2}  & & & \\
+			& & C_{x/y,3} & \ddots & \ddots & \\
 			& & & \ddots & \ddots & \ddots \\
@@
 \bseqn{
-	A^x_n &:= 
+	A_{x,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
-	B^x_n &:= 
+	B_{x,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
-	C^x_n &:= \big( A^x_{n-1} \big)^\top \in \R^{(n+1)\times n}, \quad (n \ne 0)
+	C_{x,n} &:= \big( A^x_{n-1} \big)^\top \in \R^{(n+1)\times n}, \quad (n \ne 0)
 }
@@
 \bseqn{
-	A^y_n &:= 
+	A_{y,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(n+1)\times(n+2)}, \\
-	B^y_n &:= 
+	B_{y,n} &:= 
 		\begin{pmatrix}
@@
 		\end{pmatrix} \in \R^{(n+1)\times(n+1)}, \\
-	C^y_n &:= 
+	C_{y,n} &:= 
 		\begin{pmatrix}
@@
 \bseqnnumber{
-A_n &:= \begin{pmatrix}
-		A^x_n \\
-		A^y_n
-	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)}, \quad
-C_n := \begin{pmatrix}
-		C^x_n \\
-		C^y_n
+	A_n &:= 
+		\begin{pmatrix}
+                    	A_{x,n} \\
+			A_{y,n}
+		\end{pmatrix}, \quad \in \R^{2(n+1)\times(n+2)}, \quad
+	C_n &:= 
+		\begin{pmatrix}
+                    	C_{x,n} \\
+			C_{y,n}
 		\end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \label{eqn:ds:jacobisubmats1} \\
-B_n &:= \begin{pmatrix}
-		B^x_n \\
-		B^y_n
+	B_n &:= 
+		\begin{pmatrix}
+                    	B_{x,n} \\
+			B_{y,n}
 		\end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
-G_n(x,y) := \begin{pmatrix}
+	G_n(x,y) := 
+		\begin{pmatrix}
 			xI_{n+1} \\
@@
 such that $\Dnt A_n = I_{n+2}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p.78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
 \bseqn{
@@
 
-It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in \bsrefeqn{eqn:ds:Hrecurrence}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
+It is possible to obtain the recurrence coefficients for the $\{\genjac_n^{(a,b,c)}\}$ OPs in \bsrefeqn{eqn:ds:Hrecurrence}, by careful application of the Christoffel--Darboux formula \cite[(18.2.12)]{DLMF}.
 We explain the process here for the disk-slice case, however we note that a similar but simpler argument holds for the trapezium case.
@@
 }
-Now, by the Christoffel--Darboux formula \cite[18.2.12]{DLMF}, we have that for any $x, y \in \R$,
+Now, by the Christoffel--Darboux formula \cite[(18.2.12)]{DLMF}, we have that for any $x, y \in \R$,
 \bseqnnumber{
@@
 
-We have demonstrated how one can construct the three-or-four-parameter OP families (depending on the domain) that form the basis for our sparse spectral methods, and presented a procedure, utilising the Christoffel-Darboux formula \cite[18.2.2]{DLMF}, of explicitly calculating the recurrence coefficients for the univariate OPs that are part of the construction.
+We have demonstrated how one can construct the three-or-four-parameter OP families (depending on the domain) that form the basis for our sparse spectral methods, and presented a procedure, utilising the Christoffel-Darboux formula \cite[(18.2.12)]{DLMF}, of explicitly calculating the recurrence coefficients for the univariate OPs that are part of the construction.
 These coefficients contribute to calculations of the entries in the Jacobi matrices and other differential operators.

### 2021-10-11 papers/thesis/resubmitted/ch4-spherical-caps/spherical-caps.tex
Fixed typos for Jac mats defs. Added brackets to eqn numbers in citations

@@
 such that $\Dnt A_n = I_{2n+3}$.
-Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
+Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p.78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
 \bseqn{
@@
 
-It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:sc:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[18.2.12]{DLMF}.
+It is possible to recursively obtain the recurrence coefficients for the $\{\genjac_n^{(a,b)}\}$ OPs in (\ref{eqn:sc:Rrecurrence}), see \cite{snowball2019sparse}, by careful application of the Christoffel--Darboux formula \cite[(18.2.12)]{DLMF}.
 
@@
 
-\remark Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[3.5.23]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
+\remark Note that the Gauss quadrature nodes and weights $(t_j, w_j^{(t)})$ will have to be calculated, however the Gauss quadrature nodes and weights $(s_j, w_j^{(s)})$ are simply the Chebyshev--Gauss quadrature nodes and weights given explicitly \cite[(3.5.23)]{DLMF} as $s_j := \cos \fpr(\frac{2j - 1}{2M_2} \pi)$, $w_j^{(s)} := \frac{\pi}{M_2}$.
 

### 2021-10-11 papers/thesis/resubmitted/ch3-disk-slices/disk-slices.tex
Fixed commas, fullstops etc. that follow equations to make consistent.

@@
 \bseqn{
-	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1)
+	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1),
 }
@@
 \bseqnnumber{
-	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:ds:triangleops}
+	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}), \label{eqn:ds:triangleops}
 }
@@
 \bseqn{
-	\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)} 
+	\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)},
 }
@@
 \bseqn{
-	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c
+	\Wabc(x,y) = (\beta - x)^a \: (x - \alpha)^b \: (1-x^2-y^2)^c,
 }
@@
 \bseqn{
-	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms})
+	\hdopnkabc(x,y) = C_{n,k} x^{n-k} y^k + (\hbox{lower order terms}),
 }
@@
 \bseqn{
-	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)})
+	\hdopnk(x,y) := p_{n-k,k}(x) \: \rho(x)^k \: q_k\fpr(\frac{y}{\rho(x)}),
 }
@@
 	\begin{cases}
-		\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c} \\
-		\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b
+		\genjacw^{(a,b,c)}(x) &:= (\beta - x)^a \: (x - \alpha)^{b} \: \rho(x)^{c}, \\
+		\jacw^{(a,b)}(x) &:= (\delta-x)^a \: (x - \gamma)^b,
 	\end{cases}
@@
 \bseqnnumber{
-	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x \label{eqn:ds:ipgenjac} \\
-	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y \label{eqn:ds:ipjac}
+	\ip< p, \: q >_{\genjacw^{(a,b,c)}} &:= \frac{1}{\normgenjac^{(a,b,c)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjacw^{(a,b,c)}(x) \: \D x, \label{eqn:ds:ipgenjac} \\
+	\ip<p, q>_{\jacw^{(a,b)}} &:= \frac{1}{\normjac^{(a,b)}} \: \int_\gamma^\delta p(y) \: q(y) \: \jacw^{(a,b)}(y)\: \D y, \label{eqn:ds:ipjac}
 }
@@
 \bseqn{
-	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
+	\hdopnk^{(a,b,c,d)}(x,y) := \genjacnmk^{(a, b, c+d+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(d,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega.
 }
@@
 \bseqn{
-	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{\hdop^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j}
+	\ip< \hdopnk^{(a,b,c,d)}, \: \hdopmj^{(a,b,c,d)} >_{\hdop^{(a,b,c,d)}} &= \normgenjac^{(a,b,c+d+2k+1)} \: \normjac^{(d,c)} \: \delta_{n,m} \: \delta_{k,j},
 }
@@
 \bseqnnumber{\label{eq:diskpolys}
-	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega
+	\hdopnkabc(x,y) := \genjacnmk^{(a, b, 2c+2k+1)}(x) \: \rho(x)^k \: \jac_k^{(c,c)}\fpr(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega.
 }
@@
 \bseqnnumber{
-	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x) \label{eqn:ds:Hrecurrence} \\
+	x \genjac_n^{(a,b,c)}(x) &= \beta_n^{(a,b,c)} \genjac_{n+1}^{(a,b,c)}(x) + \alpha_n^{(a,b,c)} \genjac_n^{(a,b,c)}(x) + \beta_{n-1}^{(a,b,c)} \genjac_{n-1}^{(a,b,c)}(x), \label{eqn:ds:Hrecurrence} \\
 	y \jac_n^{(d,c)}(y) &= \delta_n^{(d,c)} \jac_{n+1}^{(d,c)}(y) + \gamma_{n}^{(d,c)} \jac_{n}^{(d,c)}(y) + \delta_{n-1}^{(d,c)} \jac_{n-1}^{(d,c)}(y). \label{eqn:ds:Precurrence}
@@
 \bseqn{
-	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)} \\
-	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
-	\betaabcd_{n,k,2} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k-1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
-	\betaabcd_{n,k,3} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-2}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
-	\betaabcd_{n,k,4} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
-	\betaabcd_{n,k,5} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
-	\betaabcd_{n,k,6} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}} \\
-	\betaabcd_{n,k,7} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+2}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
-	\betaabcd_{n,k,8} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k+1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}} \\
+	\alphaabcd_{n,k,1} &:= \beta_{n-k-1}^{(a, b, c+d+2k+1)}, \qquad \alphaabcd_{n,k,2} := \alpha_{n-k}^{(a, b, c+d+2k+1)}, \\
+	\betaabcd_{n,k,1} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjacnmk^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
+	\betaabcd_{n,k,2} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k-1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
+	\betaabcd_{n,k,3} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-2}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}, \\
+	\betaabcd_{n,k,4} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+1}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
+	\betaabcd_{n,k,5} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
+	\betaabcd_{n,k,6} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjac_{n-k-1}^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}, \\
+	\betaabcd_{n,k,7} &:= \delta_{k-1}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \genjac_{n-k+2}^{(a, b, c+d+2k-1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
+	\betaabcd_{n,k,8} &:= \gamma_k^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \: \rho(x) \genjac_{n-k+1}^{(a, b, c+d+2k+1)}>_{\genjacw^{(a, b, c+d+2k+1)}}, \\
 	\betaabcd_{n,k,9} &:= \delta_{k}^{(d,c)} \: \ip<\genjacnmk^{(a, b, c+d+2k+1)}, \genjacnmk^{(a, b, c+d+2k+3)}>_{\genjacw^{(a, b, c+d+2k+3)}}.
@@
 			\quadeight \quadeight \quadeight \: \: \text{if } j = k-1 \\
-		\quadeight \quadfour 0 \quadfour \quadeight \text{otherwise}
+		\quadeight \quadfour 0 \quadfour \quadeight \text{otherwise},
       	\end{cases}
@@
 			\vdots \\
-		\end{pmatrix}
+		\end{pmatrix},
 }
@@
 	\begin{cases}
-		\bighdopabcd_{-1}(x,y) := 0 \\
-		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
-		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y)
+		\bighdopabcd_{-1}(x,y) := 0, \\
+		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0}, \\
+		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y).
 	\end{cases}.
@@
 \bseqn{
-	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
+	\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\},
 }
@@
 		(\gamma,\delta) &:= (-1,1) \\
-		\rho(x) &:= (1-x^2)^{\half}
-	\end{cases}.
+		\rho(x) &:= (1-x^2)^{\half}.
+	\end{cases}
 }
@@
 \bseqn{
-	\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c}
+	\genjacw^{(a,b,c)}(x) :=  (\beta - x)^a \: (x - \alpha)^b \: \rho(x)^{c},
 }
@@
 \bseqn{
-	{\partial f \over \partial x}&= \bighdop^{(a+1,b+1,c+1)}(x,y)^\top \: D_x^{(a,b,c)} \: \vec{f}, \\
+	{\partial f \over \partial x} &= \bighdop^{(a+1,b+1,c+1)}(x,y)^\top \: D_x^{(a,b,c)} \: \vec{f}, \\
 	{\partial f \over \partial y} &= \bighdop^{(a,b,c+1)}(x,y)^\top \: D_y^{(a,b,c)} \: \vec{f}, \\
@@
 	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a,b,2c+2k+1)}, \rho(x)^{j-k+1} \: \genjacmmj^{(a,b,2c+2j+3)} >_{\genjacw^{(a,b,2c+2k+1)}} \\
-	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}
+	&\quad \quad \quad \quad \quad \quad \cdot \: \normjac^{(c+1)} \: \ip< \jac_k^{(c,c) \: \prime}, \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c+1)}}.
 }
@@
 	&= - \int_{-1}^1 \: \jac_k^{(c,c)} \: [ \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: y \: \jac_{j}^{(c+1,c+1)} \:\jacw^{(c)} ] \: \D y \\
-	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}}
+	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jacw^{(1)} \: \jac_{j}^{(c+1,c+1) \: \prime} -2c \:y \: \jac_{j}^{(c+1,c+1)} >_{\jacw^{(c)}},
 }
@@
 	&= - \int_{-\delta}^\delta \jac_k^{(c,c)} \: \big[ \jac_j^{(c+1,c+1)} \: \jacw^{(c+1)} + t \: \jac_j^{(c+1,c+1) \: \prime} \: \jacw^{(c+1)} - 2 c \: t^2 \: \jac_j^{(c+1,c+1)} \: \jacw^{(c)} \big] \: \D t \\
-	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}}
+	&= - \: \normjac^{(c)} \: \ip< \jac_k^{(c,c)}, \: \jac_j^{(c+1,c+1)} \: \jacw^{(1)} + t \: \jac_{j}^{(c+1,c+1) \: \prime} \: \jacw^{(1)} - 2c \: t^2 \: \jac_{j}^{(c+1,c+1)}  >_{\jacw^{(c)}},
 }
@@
 	&\int_\alpha^\beta \: \genjacnmk^{(a, b, 2c+2k+1)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} \: \rho^{k+j} \: \rho' \: \genjacw^{(a+1, b+1, 2c+2)} \: \D x \\
-	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}}
+	&= \normgenjac^{(a, b, 2c+2k+1)} \: \ip< \genjacnmk^{(a, b, 2c+2k+1)}, \: \rho \: \rho' \: \genjacw^{(1, 1, j-k)} \: \genjacmmj^{(a+1, b+1, 2c+2j+3)} >_{\genjacw^{(a, b, 2c+2k+1)}},
 }
@@
 \bseqn{
-	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{\hdop^{(a+1, b+1, c+1)}} = 0
+	&\ip<\ppx{x} \hdopnkabc, \hdop^{(a+1, b+1, c+1)}_{m,j} >_{\hdop^{(a+1, b+1, c+1)}} = 0,
 }
@@
 \bseqnnumber{
-	\ppx{x} \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)} \label{eqn:ds:weightderivativex} \\
-	\ppx{y} \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y) \label{eqn:ds:weightderivativey}
+	\ppx{x} \Wabc(x,y) &= -a W^{(a-1, b, c)}(x,y) + bW^{(a, b, c)}(x,y) + 2c \rho(x) \: \rho'(x) \: W^{(a, b, c-1)}, \label{eqn:ds:weightderivativex} \\
+	\ppx{y} \Wabc(x,y) &= -2c \: y \: W^{(a, b, c-1)}(x,y), \label{eqn:ds:weightderivativey}
 }
@@
 \bseqn{
-	T^{(a,b,c)\to(a+1,b+1,c)}, \quad T^{(a,b,c)\to(a,b,c+1)}\qqand T^{(a,b,c)\to(a+1,b+1,c+1)}
+	T^{(a,b,c)\to(a+1,b+1,c)}, \quad T^{(a,b,c)\to(a,b,c+1)}, \qqand T^{(a,b,c)\to(a+1,b+1,c+1)}
 }
@@
 \bseqn{
-	T_W^{(a,b,c)\to(a-1,b-1,c)}, \quad T_W^{(a,b,c)\to(a,b,c-1)} \qqand T_W^{(a,b,c)\to(a-1,b-1,c-1)}
+	T_W^{(a,b,c)\to(a-1,b-1,c)}, \quad T_W^{(a,b,c)\to(a,b,c-1)}, \qqand T_W^{(a,b,c)\to(a-1,b-1,c-1)}
 }
@@
 \bseqn{
-	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c)} \Big)^\top \: \bighdop^{(a+1,b+1,c)}(x,y) \\
-	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a,b,c+1)} \Big)^\top \: \bighdop^{(a,b,c+1)}(x,y) \\
-	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c+1)} \Big)^\top \: \bighdop^{(a+1,b+1,c+1)}(x,y) \\
-	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a-1,b-1,c)} \Big)^\top \: \bigW^{(a-1,b-1,c)}(x,y) \\
-	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a,b,c-1)} \Big)^\top \: \bigW^{(a,b,c-1)}(x,y) \\
+	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c)} \Big)^\top \: \bighdop^{(a+1,b+1,c)}(x,y), \\
+	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a,b,c+1)} \Big)^\top \: \bighdop^{(a,b,c+1)}(x,y), \\
+	\bighdop^{(a,b,c)}(x,y) &= \Big(T^{(a,b,c)\to(a+1,b+1,c+1)} \Big)^\top \: \bighdop^{(a+1,b+1,c+1)}(x,y), \\
+	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a-1,b-1,c)} \Big)^\top \: \bigW^{(a-1,b-1,c)}(x,y), \\
+	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a,b,c-1)} \Big)^\top \: \bigW^{(a,b,c-1)}(x,y), \\
 	\bigW^{(a,b,c)}(x,y) &= \Big(T_W^{(a,b,c)\to(a-1,b-1,c-1)} \Big)^\top \: \bigW^{(a-1,b-1,c-1)}(x,y).
@@
 \bseqn{
-	\bigWiii(x,y) = \Wiii(x,y) \: \bighdopiii(x,y)
+	\bigWiii(x,y) = \Wiii(x,y) \: \bighdopiii(x,y),
 }
@@
 \bseqnnumber{
-	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x \label{eqn:ds:ipgenjact}
+	\ip< p, \: q >_{\genjactw^{(a,b,c,d)}} &:= \frac{1}{\normgenjact^{(a,b,c,d)}} \: \int_\alpha^\beta p(x) \: q(x) \: \genjactw^{(a,b,c,d)}(x) \: \D x, \label{eqn:ds:ipgenjact}
 }
@@
 \bseqnnumber{
-	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x
+	\normgenjact^{(a,b,c,d)} := \int_\alpha^\beta \: \genjactw^{(a,b,c,d)}(x) \: \D x.
 }
@@
 \bseqnnumber{
-	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x)
+	x \: \genjact_n^{(a,b,c,d)}(x) = \tilde{\beta}_n^{(a,b,c,d)} \: \genjact_{n+1}^{(a,b,c,d)}(x) + \tilde{\alpha}_n^{(a,b,c,d)} \:\genjact_n^{(a,b,c,d)}(x) + \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \genjact_{n-1}^{(a,b,c,d)}(x).
 }
@@
 \bseqnnumber{
-	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansioni} \\
-	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x) \label{eqn:ds:genjacexpansionii}
+	\genjact_n^{(a,b,c+1,d)}(x) &= \mathcal{C}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(1) \: \genjact_{k}^{(a,b,c,d)}(x), \label{eqn:ds:genjacexpansioni} \\
+	\genjact_n^{(a,b,c,d+1)}(x) &= \mathcal{D}_n^{(a,b,c,d)} \: \sum_{k=0}^n \: \genjact_{k}^{(a,b,c,d)}(-1) \: \genjact_{k}^{(a,b,c,d)}(x). \label{eqn:ds:genjacexpansionii}
 }
@@
 \bseqn{
-	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)}
+	\int_\alpha^\beta \: \genjact_n^{(a,b,c+1,d)}(x) \: \genjact_m^{(a,b,c+1,d)}(x) \: \genjactw^{(a,b,c+1,d)}(x) \: \D x &= \delta_{n,m} \: \normgenjact^{(a,b,c+1,d)},
 }
@@
 	&= \delta_{m,n} \: {\mathcal{C}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(1) \: \genjact_{n+1}^{(a,b,c,d)}(1) \\
-	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)}
+	&= \delta_{m,n} \: \normgenjact^{(a,b,c+1,d)},
 }
@@
 	&= - \delta_{m,n} \: {\mathcal{D}_n^{(a,b,c,d)}}^2 \: \tilde{\beta}_n^{(a,b,c,d)} \: \normgenjact^{(a,b,c,d)} \: \genjact_{n}^{(a,b,c,d)}(-1) \: \genjact_{n+1}^{(a,b,c,d)}(-1) \\
-	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)}
+	&= \delta_{m,n} \: \normgenjact^{(a,b,c,d+1)},
 }
@@
 		&\quadfour - \genjact_{n}^{(a,b,c,d)}(1) \: \Big(\tilde{\beta}_{n+1}^{(a,b,c,d)} \genjact_{n+2}^{(a,b,c,d)}(x) + \tilde{\alpha}_{n+1}^{(a,b,c,d)} \genjact_{n+1}^{(a,b,c,d)}(x) \nonumber \\
-			&\quadeight \quadfour + \tilde{\beta}_{n}^{(a,b,c,d)} \genjact_{n}^{(a,b,c,d)}(x)\Big) \Big\} \label{eqn:ds:genjactproofa}
+			&\quadeight \quadfour + \tilde{\beta}_{n}^{(a,b,c,d)} \genjact_{n}^{(a,b,c,d)}(x)\Big) \Big\}. \label{eqn:ds:genjactproofa}
 }
@@
 		&\quad + \mathcal{C}_n^{(a,b,c,d)} \tilde{\alpha}_{n}^{(a,b,c+1,d)} \tilde{\beta}_{n}^{(a,b,c,d)} \: \Big( \genjact_{n}^{(a,b,c,d)}(x) \genjact_{n+1}^{(a,b,c,d)}(1) - \genjact_{n+1}^{(a,b,c,d)}(x) \genjact_{n}^{(a,b,c,d)}(1) \Big) \nonumber \\
-		&\quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \genjact_{n-1}^{(a,b,c,d)}(1) \Big) \label{eqn:ds:genjactproofb}
+		&\quad + \mathcal{C}_{n-1}^{(a,b,c,d)} \tilde{\beta}_{n-1}^{(a,b,c+1,d)} \tilde{\beta}_{n-1}^{(a,b,c,d)} \: \Big( \genjact_{n-1}^{(a,b,c,d)}(x) \genjact_{n}^{(a,b,c,d)}(1) - \genjact_{n}^{(a,b,c,d)}(x) \genjact_{n-1}^{(a,b,c,d)}(1) \Big).\label{eqn:ds:genjactproofb}
 }
@@
 	&\quadfour = \int_\alpha^\beta  \genjacw^{(a,b,2c+1)}(s) \: \Big( \int_{-1}^1 f_o\big(s,\rho(s) t\big) \: \jacw^{(c)}(t) \: \D t \Big) \: \D s \quad (\dagger) \\
-	&\quadfour = 0
+	&\quadfour = 0,
 }
@@
 		\hdop^{(a,b,c)}_{n,n}(x,y)
-	\end{pmatrix} \in \R^{n+1}
+	\end{pmatrix} \in \R^{n+1},
 }
@@
 	\end{pmatrix} \in \R^{n+1} \quad \forall n = 0,1,2,\dots,N, \quad
-	f_{n,k} := \frac{\ip< f, \: \hdopnk^{(a,b,c)}>_{\hdopabc}}{\norm{\hdopnk^{(a,b,c)}}_{\hdopabc}}
+	f_{n,k} := \frac{\ip< f, \: \hdopnk^{(a,b,c)}>_{\hdopabc}}{\norm{\hdopnk^{(a,b,c)}}_{\hdopabc}}.
 }
@@
 \bseqn{
-	f_{n,k} &= \lambda^{(a,b,c)}_k \: \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) \: \hdopnkabc(x_j, y_j) +f(x_j, -y_j) \: \hdopnkabc(x_j, -y_j) \big]
+	f_{n,k} &= \lambda^{(a,b,c)}_k \: \sum_{j=1}^{M} w_j \: \big[ f(x_j, y_j) \: \hdopnkabc(x_j, y_j) +f(x_j, -y_j) \: \hdopnkabc(x_j, -y_j) \big],
 }
@@
 	\begin{cases}
-    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
-		u(x,y) = 0 \quad \text{on } \partial \Omega
-	\end{cases}.
+    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega, \\
+		u(x,y) = 0 \quad \text{on } \partial \Omega.
+	\end{cases}
 	\label{eqn:ds:helmholtz}
@@
 \bseqn{
-    (\laplacewiii + k^2 T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}) \vec{u} = \vec{f}
+    (\laplacewiii + k^2 T^{(0,0,0)\to(1,1,1)} \: V({J_x^{(0,0,0)}}^\top, {J_y^{(0,0,0)}}^\top) \: T_W^{(1,1,1)\to(0,0,0)}) \vec{u} = \vec{f},
 }
@@
 	\begin{cases}
-    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega \\
-		u(x,y) = c \in \R \quad \text{on } \partial \Omega
+    		\Delta u(x,y) + k^2 \: v(x,y) \; u(x,y) = f(x,y) \quad \text{in } \Omega, \\
+		u(x,y) = c \in \R \quad \text{on } \partial \Omega,
 	\end{cases}
@@
 	\begin{cases}
-    		\Delta \tilde{u}(x,y) + k^2 \: v(x,y) \; \tilde{u}(x,y) = f(x,y) - c \: k^2 \: v(x,y) \; =: g(x,y)  \quad \text{in } \Omega \\
+    		\Delta \tilde{u}(x,y) + k^2 \: v(x,y) \; \tilde{u}(x,y) = f(x,y) - c \: k^2 \: v(x,y) \; =: g(x,y)  \quad \text{in } \Omega, \\
 		\tilde{u}(x,y) = 0 \quad \text{on } \partial \Omega.
-	\end{cases}.
+	\end{cases}
 }
@@
 	\begin{cases}
-    		\Delta^2 u(x,y) = f(x,y) \quad \text{in } \Omega \\
-		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
-	\end{cases}.
+    		\Delta^2 u(x,y) = f(x,y) \quad \text{in } \Omega, \\
+		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega.
+	\end{cases}
 	\label{eqn:ds:biharmonic}
@@
 	\begin{cases}
-         - \Delta u(x,y) = f(x,y) \quad \text{in } \Omega \\
-         u = 0 \quad \text{on } \partial \Omega
+		- \Delta u(x,y) = f(x,y) \quad \text{in } \Omega, \\
+		u = 0 \quad \text{on } \partial \Omega.
          \end{cases}
@@
 \bseqnnumber{
-	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p.
-	\label{eqn:ds:FEMweakform}
+	L(v_p) = a(u_p,v_p) \quad \forall \: v_p \in V_p. \label{eqn:ds:FEMweakform}
 }
@@
 	&= \vec{v}^\top \: \Big( {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x \\
-	& \quad \quad \quad \quad + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \Big) \: \vec{u}
+	& \quad \quad \quad \quad + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y \Big) \: \vec{u},
 }
@@
 \bseqn{
-	A\vec{u} = \tilde{\vec{f}}.
+	A\vec{u} = \tilde{\vec{f}},
 }
@@
 \bseqn{
-	A = {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y
+	A = {\Wiii_x}^\top \Lambda^{(0,0,0)} \Wiii_x + ({T_W^{(1,1,0)\to(0,0,0)} \Wiii_y})^\top \Lambda^{(0,0,0)} T_W^{(1,1,0)\to(0,0,0)} \Wiii_y,
 }

### 2021-10-11 papers/thesis/resubmitted/ch4-spherical-caps/spherical-caps.tex
Fixed commas, fullstops etc. that follow equations to make consistent in CH4.

@@
 \bseqn{
-	\Omega := \bsset{(x,y,z) \in \R^3}{\alpha < z < \beta, \: x^2 + y^2 + z^2 = 1}
+	\Omega := \bsset{(x,y,z) \in \R^3}{\alpha < z < \beta, \: x^2 + y^2 + z^2 = 1},
 }
@@
 \bseqn{
-	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms})
+	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms}),
 }
@@
 \bseqn{
-	x &= \sinphi \costheta = \rho(z) \costheta \\
-	y &= \sinphi \sintheta = \rho(z) \sintheta \\
-	z &= \cosphi
+	x &= \sinphi \costheta = \rho(z) \costheta, \\
+	y &= \sinphi \sintheta = \rho(z) \sintheta, \\
+	z &= \cosphi,
 }
@@
 \bseqn{
-	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
+	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo,
 }
@@
 \bseqn{
-	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
-	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta) \\
-	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta)
+	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta), \\
+	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \\
+	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta),
 }
@@
 	&= c^h \: \int_h^1 \Big\{ p(x,\sqrt{1-x^2}) \; q(x,\sqrt{1-x^2}) \\
-	&\quadfour \quadtwo + p(x,-\sqrt{1-x^2}) \; q(x,-\sqrt{1-x^2}) \Big\} {\D x \over \sqrt{1-x^2}}
+	&\quadfour \quadtwo + p(x,-\sqrt{1-x^2}) \; q(x,-\sqrt{1-x^2}) \Big\} {\D x \over \sqrt{1-x^2}},
 }
@@
 \bseqn{
-	\ch^h_{0,0}(\xvec) \equiv \ch^h_{0,0}(x, y) &:= \ch^h_{0} =: \ch^h_{0,0}(\theta) \\
-	\ch^h_{k,0}(\xvec) \equiv \ch^h_{k,0}(x, y) &:= T^h_k(x) =: \ch^h_{k,0}(\theta) \\
-	\ch^h_{k,1}(\xvec) \equiv \ch^h_{k,1}(x, y) &:= y \: U^h_{k-1}(x) =: \ch^h_{k,1}(\theta)
+	\ch^h_{0,0}(\xvec) \equiv \ch^h_{0,0}(x, y) &:= \ch^h_{0} =: \ch^h_{0,0}(\theta), \\
+	\ch^h_{k,0}(\xvec) \equiv \ch^h_{k,0}(x, y) &:= T^h_k(x) =: \ch^h_{k,0}(\theta), \\
+	\ch^h_{k,1}(\xvec) \equiv \ch^h_{k,1}(x, y) &:= y \: U^h_{k-1}(x) =: \ch^h_{k,1}(\theta),
 }
@@
 \bseqn{
-	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)})
+	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)}),
 }
@@
 	&\quadeight \quad \cdot w(\cosphi) \: \sinphi \Big\} \: \D \theta \: \D \varphi \\
-	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: q\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: w(z) \: \D \theta \: \D z
+	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: q\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: w(z) \: \D \theta \: \D z,
 }
@@
 \bseqnnumber{
-	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:sc:ipgenjac}
+	\ip< p, \: q >_{\genjacw^{(a,b)}} &:= \frac{1}{\normgenjac^{(a,b)}} \: \int_\alpha^1 p(x) \: q(x) \: \genjacw^{(a,b)}(x) \: \D x, \label{eqn:sc:ipgenjac}
 }
@@
 \bseqn{
-	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b}
-} is the weight function and 
-\bseqnnumber{
-	\normgenjac^{(a,b)} := \int_\alpha^1 \: \genjacw^{(a,b)}(x) \: \D x \label{eqn:sc:ipnormalisation}
+	\genjacw^{(a,b)}(x) &:= (x - \alpha)^{a} \: \rho(x)^{b},
+} 
+is the weight function and 
+\bseqnnumber{
+	\normgenjac^{(a,b)} := \int_\alpha^1 \: \genjacw^{(a,b)}(x) \: \D x, \label{eqn:sc:ipnormalisation}
 }
@@
 \bseqnnumber{
-	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x) \label{eqn:sc:Rrecurrence}
+	x \genjac_n^{(a,b)}(x) &= \beta_n^{(a,b)} \genjac_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} \genjac_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} \genjac_{n-1}^{(a,b)}(x), \label{eqn:sc:Rrecurrence}
 }
@@
 \bseqn{
-	2 \cos k \theta \cos\theta &= \cos (k-1)\theta + \cos(k+1)\theta \\
-	2 \sin k \theta \cos\theta &= \sin (k-1)\theta + \sin(k+1)\theta \\
-	2 \cos k \theta \sin\theta &= - \sin (k-1)\theta + \sin(k+1)\theta \\
+	2 \cos k \theta \cos\theta &= \cos (k-1)\theta + \cos(k+1)\theta, \\
+	2 \sin k \theta \cos\theta &= \sin (k-1)\theta + \sin(k+1)\theta, \\
+	2 \cos k \theta \sin\theta &= - \sin (k-1)\theta + \sin(k+1)\theta, \\
 	2 \sin k \theta \sin\theta &= \cos (k-1)\theta - \cos(k+1)\theta.
@@
 			\ch_0 &\text{if } k = 0 \\
-			\half &\text{otherwise}
+			\half &\text{otherwise}.
 		\end{cases}
@@
 			- \alphaa_{n,k,j} \quad &\text{if }(i = 0 \text{ and } j \text{ is odd}) \text{ or }(i = 1 \text{ and } j \text{ is even}) \\
-			\alphaa_{n,k,j} \quad &\text{otherwise}
-		\end{cases}, \\	
+			\alphaa_{n,k,j} \quad &\text{otherwise},
+		\end{cases} \\	
 	\gammaa_{n,k,1} &:= \beta_{n-k-1}^{(a, 2k)}, \qquad \gammaa_{n,k,2} := \alpha_{n-k}^{(a, 2k)}, \qquad \gammaa_{n,k,3} := \beta_{n-k}^{(a, 2k)}.
@@
 \bseqn{
-	c_{m,j} = {\ip< x \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
+	c_{m,j} = {\ip< x \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}},
 }
@@
 \bseqn{
-	(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)
+	(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z),
 }
@@
 	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\
-	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z
+	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z,
 }
@@
 \bseqn{
-	d_{m,j} = {\ip< y \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}}
+	d_{m,j} = {\ip< y \: \scopnkia, \scopmjha >_{\scopa}}{\norm{\scopmjha}^{-2}_{\scopa}},
 }
@@
 	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \\
-	&\quadfour \cdot \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big)
+	&\quadfour \cdot \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big),
 }
@@
 	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:sc:jacobimatricesdefinitionalt} \\
-	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
+	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z), \nonumber
 }
@@
 			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \\
 	B^{(a)}_{x,n} &:= 
@@
 			& & A^{(a)}_{n,n,3} & 0
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}, \\
 	C^{(a)}_{x,n} &:= 
@@
 			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \\
 	B^{(a)}_{y,n} &:= 
@@
 			& & B^{(a)}_{n,n,3} & 0
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}, \\
 	C^{(a)}_{y,n} &:= 
@@
 			& & 0 & \Gamma^{(a)}_{n,n,3} & 0 \\
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \\
 	B^{(a)}_{z,n} &:= 
@@
 			& & & \Gamma^{(a)}_{n,n,2}
-		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
+		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}, \\
 	C^{(a)}_{z,n} &:= 
@@
 	\begin{cases}
-		\bigscopta_{-1}(x,y,z) := 0 \\
-		\bigscopta_{0}(x,y,z) := \scopa_0 \\
-		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z)
-	\end{cases}.
+		\bigscopta_{-1}(x,y,z) := 0, \\
+		\bigscopta_{0}(x,y,z) := \scopa_0, \\
+		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z).
+	\end{cases}
 }
@@
 			\frac{- \: \betaa_{n,n,1,5}}{\betaa_{n,n,1,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 3 \\
-			0 & o/w
+			0 &\text{otherwise},
 		\end{cases} \\
@@
 			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2  \text{ and } n > 1 \\
-			0 & o/w
+			0 &\text{otherwise}.
 		\end{cases}
@@
 			\bigscopa_{N,N}
-		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:sc:OPdefNa}
+		\end{pmatrix} \in \R^{(N+1)^2}. \label{eqn:sc:OPdefNa}
 }
@@
 \bseqn{
-	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z),
+	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z).
 }
@@
 \bseqn{
-	\rho \pfpx{\varphi}{f} (x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \mathbf{f}, \\
-	\rho \ppx{\varphi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\varphi^{(a)}\: \mathbf{f}, \\
-	\pfpx{\theta}{f}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \mathbf{f}, \\
-	\DeltaS f(x,y,z) &= \bigscopN^{(a+\tilde a)}(x,y,z)^\top \: \calL^{(a)\to(a+\tilde a)} \: \mathbf{f}, \\
-	\DeltaS \big( \genjacw^{(a,0)}(z) \: f(x,y,z) \big) &= \bigW_N^{(a-\tilde a)}(x,y,z)^\top \: \calL_W^{(a)\to(a-\tilde a)} \: \mathbf{f},  \quad \text{ (for } a \ge 2 \text{ only)} \\
-	\DeltaS \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \mathbf{f}, \quad \text{ (for } a = 1 \text{ only)}
+	\rho \pfpx{\varphi}{f} (x,y,z)&= \bigscop_N^{(a+1)}(x,y,z)^\top \: D_\varphi^{(a)} \: \vec{f}, \\
+	\rho \ppx{\varphi}[\genjacw^{(a,0)}(z) \: f(x,y,z)] &= \bigW_N^{(a-1)}(x,y)^\top \: W_\varphi^{(a)}\: \vec{f}, \\
+	\pfpx{\theta}{f}(x,y,z) &= \bigscopNa(x,y,z)^\top \: D_\theta \: \vec{f}, \\
+	\DeltaS f(x,y,z) &= \bigscopN^{(a+\tilde a)}(x,y,z)^\top \: \calL^{(a)\to(a+\tilde a)} \: \vec{f}, \\
+	\DeltaS \big( \genjacw^{(a,0)}(z) \: f(x,y,z) \big) &= \bigW_N^{(a-\tilde a)}(x,y,z)^\top \: \calL_W^{(a)\to(a-\tilde a)} \: \vec{f},  \quad \text{ (for } a \ge 2 \text{ only)} \\
+	\DeltaS \big( \genjacw^{(1,0)}(z) \: f(x,y,z) \big) &= \bigscopN^{(1)}(x,y,z)^\top \: \Delta^{(1)}_W \: \vec{f}, \quad \text{ (for } a = 1 \text{ only)}.
 }
@@
 \begin{itemize}
-	\item $D_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(2, 4)$
-  	\item $W_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(4, 2)$
-	\item $D_\theta$ is block-diagonal with sub-block-bandwidths $(1, 1)$
-	\item $\calL^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block-bandwidths $(0, 4)$
-	\item $\calL_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block-bandwidths $(4, 0)$
-	\item $\Delta^{(1)}_W$ is block-diagonal with sub-block-bandwidths $(2, 2)$
+	\item $D_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(2, 4)$.
+  	\item $W_\varphi^{(a)}$ is block-diagonal with sub-block-bandwidths $(4, 2)$.
+	\item $D_\theta$ is block-diagonal with sub-block-bandwidths $(1, 1)$.
+	\item $\calL^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block-bandwidths $(0, 4)$.
+	\item $\calL_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block-bandwidths $(4, 0)$.
+	\item $\Delta^{(1)}_W$ is block-diagonal with sub-block-bandwidths $(2, 2)$.
 \end{itemize}
@@
 	&\quadfour + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
-	&= \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
+	&= \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)},
 }
@@
 \bseqn{
-	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
+	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z.
 }
@@
 \bseqn{
-	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)}
+	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] =\sum_{m=0}^{n-k+1} \: \tilde c_{\{n,k\},m} \: \genjacw^{(a, 2k)} \: \genjac_{m}^{(a,2k)},
 }
@@
 	\tilde c_{\{n,k\},m} &= \frac{1}{\normgenjac^{(a,2k)}} \: \ip<\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}], \genjac_{m}^{(a,2k)}>_{\genjacw^{(0,0)}} \\
-	&= - \frac{1}{\normgenjac^{(a,2k)}}  \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
+	&= - \frac{1}{\normgenjac^{(a,2k)}}  \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z.
 }
@@
 	&\quad \quad \quad = \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: [(a+1) \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(0, 2)} \\
-	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - 2(k+1) z \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(1, 0)} + \genjac_{m}^{(a,2k) \: \prime \prime} \: \genjacw^{(1, 2)}] \: \genjacw^{(a, 2k)} \: \D z
+	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - 2(k+1) z \genjac_{m}^{(a,2k) \: \prime} \: \genjacw^{(1, 0)} + \genjac_{m}^{(a,2k) \: \prime \prime} \: \genjacw^{(1, 2)}] \: \genjacw^{(a, 2k)} \: \D z,
 }
@@
 		(-1)^{i+1} \: k \: \scopa_{n,k,|i-1|}(x,y,z) & k > 0 \\
-		0 & k = 0
-	\end{cases}.
+		0 & k = 0.
+	\end{cases}
 }
@@
 We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$. The entries of the operator are given by the coefficients in the expansion 
-$$
+\bseqn{
 	\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)},
-$$ 
+}
 where the coefficients are 
@@
 	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
-	&\quadeight + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
+	&\quadeight + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z,
 }
@@
 	&\quadeight \quadtwo + a \: \rho^2 \genjac_{m-k}^{(a-1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \: \genjacw^{(1, 0)} \Big\} \genjacw^{(a-1, 2k)} \: \D z \\
-	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjacnmk^{(a, 2k)} \big[ k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \big] \genjacw^{(a, 2k)} \: \D z
+	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjacnmk^{(a, 2k)} \big[ k z \: \genjac_{m-k}^{(a-1,2k)} + \genjac_{m-k}^{(a-1,2k) \: \prime} \: \rho^2- (2k+2) z \: \genjac_{m-k}^{(a-1,2k)} \big] \genjacw^{(a, 2k)} \: \D z,
 }
@@
 
-We move on to the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yield
+We move on to the spherical Laplacian operators. Note that the Laplacian acting on the weighted and non-weighted spherical cap OP $\scopnki^{(a)}$ yields
 \bseqnnumber{
@@
 For the operator $\calL^{(a)\to(a+\tilde a)}$ for the surface Laplacian on a non-weighted space, the entries of the operator are given by the coefficients in the expansion 
-$$\DeltaS \scopnki^{(a)} = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a + \tilde a)},$$
+\bseqn{
+	\DeltaS \scopnki^{(a)} = \sum_{m=0}^{n} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a + \tilde a)},
+}
 where the coefficients are
@@
 	&\quadfour \quadtwo \quad - \genjacnmk^{(a,2k) \: \prime} \genjacw^{(a+\tilde a-1, 2k)} \: [\genjac_{m-k}^{(a+\tilde a,2k) \: \prime} \: \genjacw^{(1,0)} + (a + \tilde a) \genjac_{m-k}^{(a+\tilde a,2k)}] \Big\} \: \D z \\
-	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k+\tilde a} \: \D z,
 }
@@
 	&\quadfour \quadtwo - a \genjacnmk^{(a,2k)} \: \genjac_{m-k}^{(a - \tilde a, 2k)} \: \genjacw^{(a-2, 2k)} \: \big[ (a-1) \rho^2 - 2(k+1) z \: \genjacw^{(1, 0)} \big] \Big\} \: \D z \\
-	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a,2k)} \: \genjacw^{(a, 2k)} \: r_{m-k} \: \D z,
 }
@@
 	&\quadfour \cdot \Big( \int_\alpha^1 \: \Big\{ \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(1, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\
-	&\quadeight + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z \Big)
+	&\quadeight + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z \Big),
 }
@@
 \bseqn{
-	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z) \\
-	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z)
+	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z), \\
+	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z).
 }
@@
 \begin{itemize}
-	\item $T^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block bandwidths $(0,2\tilde a)$
-	\item $T_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block bandwidths $(2\tilde a, 0)$
+	\item $T^{(a)\to(a+\tilde a)}$ is block-diagonal with sub-block bandwidths $(0,2\tilde a)$.
+	\item $T_W^{(a)\to(a-\tilde a)}$ is block-diagonal with sub-block bandwidths $(2\tilde a, 0)$.
 \end{itemize}
@@
 \bseqn{
-	t_{m,j} = \norm{\scopmjh^{(a+\tilde{a})}}^{-2}_{\scop^{(a+\tilde a)}} \: \ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}}.
+	t_{m,j} = \norm{\scopmjh^{(a+\tilde{a})}}^{-2}_{\scop^{(a+\tilde a)}} \: \ip< \scopnkia, \scopmjh^{(a+\tilde{a})}>_{\scop^{(a+\tilde a)}},
 }
@@
 	&= \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjacmmj^{(a+\tilde a,2j)} \: \rho^{k+j} \: \genjacw^{(a + \tilde a, 0)} \: \D z \Big) \\
-	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a + \tilde a, 2k)} \: \D z
+	&= \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \genjac_{m-k}^{(a+\tilde a,2k)} \: \genjacw^{(a + \tilde a, 2k)} \: \D z,
 }
@@
 \bseqn{
-	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \gradS u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega
+	u(x,y,z) &= 0, \quad \pfpx{n}{u}(x,y,z) = \gradS u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 \quad \text{for } (x,y,z) \in \partial\Omega,
 }
@@
 \bseqn{
-	\gradS u(x,y,z) \cdot \unitvec{n}(x,y,z) = \gradS u(x,y,z) \cdot \phivec(x,y,z) = - \rho(z) \pfpx{z}{u}(x,y,z)
+	\gradS u(x,y,z) \cdot \unitvec{n}(x,y,z) = \gradS u(x,y,z) \cdot \phivec(x,y,z) = - \rho(z) \pfpx{z}{u}(x,y,z),
 } 
@@
 \bseqn{
-	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega\\
-	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega
+	f_e(x,y,z) &:= \half \Big(f(x, y,z) + f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega, \\
+	f_o(x,y,z) &:= \half \Big(f(x, y,z) - f(-x, -y,z)\Big), \quad \forall (x,y,z) \in \Omega,
 }
@@
 	\int_0^{2\pi} g\big(\cos \theta, \sin\theta\big) \: \D \theta &= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}} \\
-	&= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(-x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}}
+	&= \int_{-1}^1 \Big[ g\big(x, \sqrt{1-x^2}\big) + g\big(-x, -\sqrt{1-x^2}\big) \Big] \frac{\D x}{\sqrt{1-x^2}},
 }
@@
 	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + f_o\big(-\rho(z)x, - \rho(z)\sqrt{1-x^2}, z\big) \Big] \: \frac{\D x}{\sqrt{1-x^2}} \Big) \: \D z \\
-	&\quad \quad = 0.
+	&\quad \quad = 0,
 }
@@
 \bseqn{
-	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \vec{f}_k = \bigscopNa(x,y,z)^\top \: \vec{f}
+	f(x,y,z) \approx \sum_{k=0}^N \bigscopNka(x,y,z)^\top \: \vec{f}_k = \bigscopNa(x,y,z)^\top \: \vec{f},
 }
@@
 	&= \frac{1}{M_2 \: \normgenjac^{(a,2k)}} \sum_{j=1}^{M_1} w^{(t)}_j \sum_{l=1}^{M_2} \big[ f(x_j, y_j, z_j) \scopnkia(x_j, y_j, z_j) \\
-	&\quadeight \quadfour + f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big]
+	&\quadeight \quadfour + f(-x_j, -y_j, z_j) \scopnkia(-x_j, -y_j, z_j) \big],
 }
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0} \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {\tilde{\vec{f}}_n}^T - \vec{\xi}_{n+1}^T D^T_n (B_n - G_n(x,y,z)) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+	\quad &\text{3) } \text{Output: } f(x,y,z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0.
 }
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+2} = \vec{0} \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
 	\quad & \quad \quad \quad \text{set } \vec{\xi}_{n}^T = {\tilde{\vec{v}}_n}^T - \vec{\xi}_{n+1}^T D^T_n \big(B_n - G_n(X, Y, Z)\big) -  \vec{\xi}_{n+2}^T D^T_{n+1}C_{n+1} \\
-	\quad &\text{3) } \text{Output: } \tilde V = v(X, Y, Z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0
+	\quad &\text{3) } \text{Output: } \tilde V = v(X, Y, Z) \approx \vec{\xi}_{0} \: \bigscopta_{0}  = \xi_{0} \: \scopa_0,
 }
@@
 		\DeltaS u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0& \quad \text{on } \partial \Omega
+		u(x,y,z) = 0& \quad \text{on } \partial \Omega,
 	\end{cases}
@@
 We can tackle the problem as follows.
-Choose an $N \in \N$ large enough for the problem, and denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule in \bsrefsection{section:sc:expandingfunctions}.
+Choose an $N \in \N$ large enough for the problem, and denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\vec{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\vec{f}$.
+Since $f$ is known, we can obtain $\vec{f}$ using the quadrature rule in \bsrefsection{section:sc:expandingfunctions}.
 In matrix-vector notation, our system hence becomes:
 \bseqn{
-    \Delta_W^{(1)} \mathbf{u} = \mathbf{f}
-}
-which can be solved to find $\mathbf{u}$.
+    \Delta_W^{(1)} \vec{u} = \vec{f},
+}
+which can be solved to find $\vec{u}$.
 In \bsreffig{fig:sc:poisson} we see the solution to the Poisson equation with zero boundary conditions given in \bsrefeqn{eqn:sc:poisson} in the disk-slice $\Omega$.
@@
 \bseqn{
-	f(x,y,z) &= \norm{\Big(x - (\epsilon + 1 / \sqrt{3}), \: y - (\epsilon + 1 / \sqrt{3}), \: z - (\epsilon + 1 / \sqrt{3})\Big)^\top}
+	f(x,y,z) &= \norm{\Big(x - (\epsilon + 1 / \sqrt{3}), \: y - (\epsilon + 1 / \sqrt{3}), \: z - (\epsilon + 1 / \sqrt{3})\Big)^\top},
 }
@@
     		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0 &\quad \text{on } \partial \Omega
+		u(x,y,z) = 0 &\quad \text{on } \partial \Omega,
 	\end{cases}
@@
 We can tackle the problem as follows.
-Denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\mathbf{f}$.
-Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule in \bsrefsection{section:sc:expandingfunctions}.
+Denote the coefficient vector for expansion of $u$ in the $\bigWNi$ OP basis up to degree $N$ by $\vec{u}$, and the coefficient vector for expansion of $f$ in the $\bigscopNi$ OP basis up to degree $N$ by $\vec{f}$.
+Since $f$ is known, we can obtain  the coefficients $\vec{f}$ using the quadrature rule in \bsrefsection{section:sc:expandingfunctions}.
 
@@
 \bseqn{
-    (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \mathbf{u} = \mathbf{f}
-}
-which can be solved to find $\mathbf{u}$.
+    (\Delta_W^{(1)} + k^2 \:T^{(0)\to(1)} \: V \: T_W^{(1)\to(0)}) \: \vec{u} = \vec{f},
+}
+which can be solved to find $\vec{u}$.
 We can see the sparsity and structure of this matrix system in \bsreffig{fig:sc:sparsity} with $v(x,y,z) = zxy^2$ as an example.
@@
     		\DeltaS^2 u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = \grad_S \: u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 &\quad \text{on } \partial \Omega
+		u(x,y,z) = 0, \quad \frac{\partial u}{\partial n}(x,y,z) = \grad_S \: u(x,y,z) \cdot \unitvec{n}(x,y,z) = 0 &\quad \text{on } \partial \Omega,
 	\end{cases}
@@
     		\DeltaS u(x,y,z) + k^2 \: v(x,y,z) \; u(x,y,z) = f(x,y,z) &\quad \text{in } \Omega \\
-		u(x,y,z) = c(x,y) &\quad \text{on } \partial \Omega
+		u(x,y,z) = c(x,y) &\quad \text{on } \partial \Omega,
 	\end{cases}
@@
     		\DeltaS \tilde{u}(x,y,z) + k^2 \; v(x,y,z) \tilde{u}(x,y,z) = g(x,y,z) \quad \text{in } \Omega \\
-		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega
+		\tilde{u}(x,y,z) = 0 \quad \text{on } \partial \Omega,
 	\end{cases}
@@
 \bseqn{
-	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y)
+	g(x,y,z) &:= f(x,y,z) - k^2 \: v(x,y,z) \: c(x,y) - \DeltaS \: c(x,y),
 }
@@
 	&= {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 c_{n,n,i} \: \pmpxm{\theta}{2} \ch_{n,i}(\theta) \\
-	&= - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta)
+	&= - {1 \over \rho(z)^2} \sum_{n=0}^N \sum_{i=0}^1 n^2 \: c_{n,n,i} \: \ch_{n,i}(\theta),
 }

### 2021-10-19 papers/thesis/resubmitted/ch1-introduction/introduction.tex
Further edits to intro framework

--- 2021-09-28 papers/thesis/resubmitted/ch1-introduction/introduction.tex
@@
 
-The structure of this thesis is as follows.
+% GENERAL FRAMEWORK
+The construction of the spectral methods in this thesis follow a similar pattern that is worth detailing here as a general framework.
+\begin{itemize}
+	\item \textit{Define the domain of interest.}
+
+		It is best to do this via a restriction of one coordinate in terms of the others, using a function $\rho$ that satisfies one of two conditions we will detail later.
+As an example, for a domain in 2D space, we would define this as $\Omega := \bsset{(x,y) \in \R^2}{\alpha \le x \le \beta, \: \gamma \rho(x) \le y \le \delta \rho(x)}$ for given constants $\alpha, \beta, \gamma, \delta$.
+
+	\item \textit{Express the multivariate and multi-parameter OPs as polynomials in cartesian coordinates.}
+
+		The construction of the multivariate OPs takes its inspiration from the Koornwinder approach \cite{koornwinder1975two}.
+The construction will make use of the function $\rho$.
+These OPs will be $d$-parameter, meaning that we can construct operators that raise or lower the parameter values, and thus represent conversion between the OP bases.
+
+	\item \textit{Derive expressions for multiplication of the OPs by the coordinates.}
+
+	\item \textit{Use these relations to determine the entries to \enquote{Jacobi matrices}}
+
+		These \enquote{Jacobi matrices} $J_x$, $J_y$, etc.
+represent multiplication by the coordinates when applied to the degree-ordered vector of OPs, $\bigP$.
+They will be banded-block-banded in structure -- in fact, due to the construction, they will be block-tridiagonal.
+Continuing with the 2D example:
+		\bseqn{
+			J_{x/y} &= 
+			\begin{pmatrix}
+				B_{x/y,0} & A_{x/y,0} & & & & \\
+				C_{x/y,1} & B_{x/y,1} & A_{x/y,1} & & & \\
+				& C_{x/y,2} & B_{x/y,2} & A_{x/y,2}  & & & \\
+				& & C_{x/y,3} & \ddots & \ddots & \\
+				& & & \ddots & \ddots & \ddots \\
+			\end{pmatrix}.
+		}
+		The degree-ordered vector of OPs, $\bigP$, can be thought of as being made up of ordered sub-vectors grouped by polynomial degree, $\bigP_n$ for $n \in \No$.
+
+	\item \textit{Create a three-term recurrence for the vector of OPs of degree $n$.}
+
+		This is done by combining the systems $J_x \bigP = x \bigP$ and $J_y \bigP = y \bigP$ (and $J_z \bigP = z \bigP$ if in three-dimensional space).
+The resulting three-term recurrence will have the form:
+		\bseqn{
+			\bigP_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigP_n(x,y) - \Dnt C_n  \, \bigP_{n-1}(x,y).
+		}
+		Here, the matrices $B_n$, $C_n$ and $G_n$ are defined via:
+		\bseqn{
+                    	C_n := 
+                    		\begin{pmatrix}
+                    			C_{x,n} \\
+	                    		C_{y,n}
+				\end{pmatrix} \: (n \ne 0), \quad
+			B_n &:= 
+				\begin{pmatrix}
+                    			B_{x,n} \\
+	                    		B_{y,n}
+				\end{pmatrix}, \quad
+			G_n(x,y) := 
+                    		\begin{pmatrix}
+					xI_{n+1} \\
+	                    		yI_{n+1}
+				\end{pmatrix}.
+		}
+		The matrices $\Dnt$ are left-inverses of the matrices $A_n$ respectively, which are defined similarly:
+		\bseqn{		
+			A_n &:=
+				\begin{pmatrix}
+                    			A_{x,n} \\
+                    			A_{y,n}
+				\end{pmatrix}.
+		}
+
+	\item \textit{Use the multivariate version of the Clenshaw algorithm for evaluation of functions that are expanded in the OP basis.}
+
+	\item \textit{Derive (sparse) relationships for various differential operations, and relations for raising and lowering the parameter values of the OPs.}
+
+	\item \textit{Construct the sparse operator matrices representing these operations.}
+
+		The differential operators may raise or lower parameter values, which is why the basis conversion operators are required.
+Further operators can be constructed as combinations of others -- for example, a second derivative would be the multiplication of two first derivative operators.
+Notably, these operator matrices will be banded-block-banded in structure.
+
+	\item \textit{Use the \enquote{operator Clenshaw algorithm} to construct operator matrices for the action of multiplying by a general function.}
+
+		For instance, this would be useful for constructing an operator matrix that represents a variable coefficient in a Helmholtz example.
+\end{itemize}
+
+By following this framework, one will be able to write down a given PDE in spectral space (i.e.
+as a matrix-vector equation).
+The variables are given by their coefficients for their expansion in the OP basis, while the differential and other operations are given by their relevant operator matrices.
+The resulting system's sparsity is a result of the sparse relations of the OPs.
+
+The structure of the remainder of thesis is as follows.
 

### 2021-10-19 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
Edits to CH2

@@
 \bseqn{
-	\Omega := \bigS^2 = \bsset{(x,y,z) \in \R^3}{x^2 + y^2 = \rho(z)^2}
+	\Omega := \bigS^2 = \bsset{(x,y,z) \in \R^3}{x^2 + y^2 = \rho(z)^2},
 }
@@
 	y &= \sinphi \sintheta = \rho(z) \sintheta, \\
-	z &= \cosphi
+	z &= \cosphi,
 }
@@
 \bseqnnumber{
-	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo \label{eqn:laplacebeltrami}
+	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo, \label{eqn:laplacebeltrami}
 }
@@
 \bseqn{
-	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x) \\
-	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x)
+	& \ddx{x} P^{(a,b)}_l (x) = \half \: (l + a + b + 1) \: P^{(a+1,b+1)}_{l-1}(x), \\
+	\implies & \dmdxm{x}{m} P_l(x) = {(l+m)! \over 2^m \: l!} \: P^{(m,m)}_{l-m}(x),
 }
@@
 \bseqn{
-	 \omega_{P,l}^{(a,b)} = {2^{a + b + 1} \; \Gamma(l + a + 1) \; \Gamma(l + b + 1) \over (2l + a + b + 1) \; \Gamma(l + a + b + 1) \; \Gamma(l + 1)}
+	 \omega_{P,l}^{(a,b)} = {2^{a + b + 1} \; \Gamma(l + a + 1) \; \Gamma(l + b + 1) \over (2l + a + b + 1) \; \Gamma(l + a + b + 1) \; \Gamma(l + 1)},
 }
@@
 \bseqn{
-	\Plm(x) &:= (-1)^m \; {(l+m)! \over (l-m)!} \; P_l^{\abs{m}}(x) = {(l+m)! \over 2^{\abs{m}} \; l!} \; (1-x^2)^{\abs{m} \over 2} \; \jaclmabs{l}{m}(x)
+	\Plm(x) &:= (-1)^m \; {(l+m)! \over (l-m)!} \; P_l^{\abs{m}}(x) = {(l+m)! \over 2^{\abs{m}} \; l!} \; (1-x^2)^{\abs{m} \over 2} \; \jaclmabs{l}{m}(x),
 }
@@
 	&= \clm (1 - (\cosphi)^2)^\frac{|m|}{2} \eimtheta P^{(|m|,|m|)}_{l-|m|}(\cosphi) \nonumber \\
-	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta \label{eqn:shdef}
+	&= \clm P^{(|m|,|m|)}_{l-|m|}(z) \: \rho(z)^{|m|} \: \eimtheta, \label{eqn:shdef}
 }
@@
 	&\quad = 2\pi \: \delta_{m, m'} \: \clm \: c_{l'}^{m'} \: \int_{-1}^1 P^{(|m|,|m|)}_{l-|m|}(z) \: P^{(|m|,|m|)}_{l'-|m|}(z) \: \rho(z)^{2|m|} \: \D z \\
-	&\quad = \delta_{l, l'} \: \delta_{m, m'}
+	&\quad = \delta_{l, l'} \: \delta_{m, m'},
 }
@@
 	&1)\quad P_{n}^{(\alpha, \beta)}(z) = {1 \over 2n + \alpha + \beta + 1} \Big\{- {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \\
-									  					& \quadfour \quadtwo \quad + (n+\alpha+\beta+1) \Big[{n+\alpha \over 2n+\alpha+\beta} - {n+\beta+1 \over 2n+\alpha+\beta+2}\Big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
+									  					& \quadfour \quadtwo \quad + (n+\alpha+\beta+1) \Big[{n+\alpha \over 2n+\alpha+\beta} - {n+\beta+1 \over 2n+\alpha+\beta+2}\Big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z), \\
 									 					& \quadfour \quadtwo \quad + {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \Big\}, \\ \\
 	&2)\quad (1-z^2) \, P_{n}^{(\alpha, \beta)}(z) = {4 \over 2n + \alpha + \beta + 1} \Big\{{(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n}^{(\alpha-1, \beta-1)}(z) \\
-									 							& \quadeight \quadtwo \: + (n+1) \Big[{n+\alpha+1 \over 2n+\alpha+\beta+2} - {n+\beta \over 2n+\alpha+\beta}\Big] \: P_{n+1}^{(\alpha-1, \beta-1)}(z) \\
+									 							& \quadeight \quadtwo \: + (n+1) \Big[{n+\alpha+1 \over 2n+\alpha+\beta+2} - {n+\beta \over 2n+\alpha+\beta}\Big] \: P_{n+1}^{(\alpha-1, \beta-1)}(z)\\
 									 							& \quadeight \quadtwo \: - {(n+1)(n+2) \over 2n+\alpha+\beta+2} \: P_{n+2}^{(\alpha-1, \beta-1)}(z) \Big\}, \\ \\
@@
 \bseqn{
-	(1-z^2) \, \jaclm{l}{m}(z) &=  \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z) + \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) \\
+	(1-z^2) \, \jaclm{l}{m}(z) &=  \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z) + \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z), \\
 	\jaclm{l}{m}(z) &= \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) + \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z), \\
@@
 			- {l \over 2(2l+1)} \quad \text{if } l - m \ge 2 \\
-			\quad \quad 0 \quad \quad \text{otherwise}
-		\end{cases}, \\
+			\quad \quad 0 \quad \quad \text{otherwise},
+		\end{cases} \\
 	\tilde \alpha_{l,m,3} &:= - {2(l-m+2)(l-m+1) \over (2l+1)(l+1)}, \quad
@@
 			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
-			\quad 0 \quad \quad \text{otherwise}
-		\end{cases}, \quad
+			\quad 0 \quad \quad \text{otherwise},
+		\end{cases} \quad
 	&&\tilde \gamma_{l,m,2} := \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
@@
 			\tilde \alpha_{l,\abs{m},2} &\text{if } m\le0 \\
-			0 &\text{otherwise} 
-		\end{cases}, \quad
+			0 &\text{otherwise},
+		\end{cases} \quad
 	&&\alpha_{l,m,2} := {\clm \over 2c^{m+1}_{l-1}}
@@
 			\tilde \alpha_{l,\abs{m},1} &\text{if } m<0 \\
-			0 &\text{otherwise} 
-		\end{cases}, \\
+			0 &\text{otherwise},
+		\end{cases} \\
 	\alpha_{l,m,3} &:= {\clm \over 2c^{m-1}_{l+1}}
@@
 			\tilde \alpha_{l,m,3}  &\text{if } m>0 \\
-			\tilde \alpha_{l,\abs{m},4} &\text{if } m\le0 
-		\end{cases}, \quad
+			\tilde \alpha_{l,\abs{m},4} &\text{if } m\le0,
+		\end{cases} \quad
 	&&\alpha_{l,m,4} := {\clm \over 2c^{m+1}_{l+1}}
@@
 			\tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
-			\tilde \alpha_{l,\abs{m},3} \quad \text{if } m<0 
-		\end{cases}, \\
+			\tilde \alpha_{l,\abs{m},3} \quad \text{if } m<0,
+		\end{cases} \\
 	\gamma_{l,m,1} &:= \frac{\clm}{c^{m}_{l-1}} \: \tilde \gamma_{l,m,1}, \quad
@@
 \bseqn{
-	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
-	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
+	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}), \\
+	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta}).
 }
@@
 \bseqnnumber{
-	J_x \SHvec(x,y,z) = x \SHvec(x,y,z), \quad J_y \SHvec(x,y,z) = y \SHvec(x,y,z), \quad J_z \SHvec(x,y,z) = z \SHvec(x,y,z) \label{eqn:jacobimatsdef}
+	J_x \SHvec(x,y,z) &= x \SHvec(x,y,z), \nonumber \\
+	J_y \SHvec(x,y,z) &= y \SHvec(x,y,z), \label{eqn:jacobimatsdef} \\
+	J_z \SHvec(x,y,z) &= z \SHvec(x,y,z), \nonumber
 }
@@
 \bseqn{
-	f(x,y,z) = \sum_{l=0}^{N} \sum_{m=-l}^{l} f_{l,m} \; \Ylm(x,y,z) = \SHvec(x,y,z)^\top \: \fvec
+	f(x,y,z) = \sum_{l=0}^{N} \sum_{m=-l}^{l} f_{l,m} \; \Ylm(x,y,z) = \SHvec(x,y,z)^\top \: \fvec,
 }
@@
 		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
-	B_{x,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
+	B_{x,l} &:= 0 \in \R^{(2l+1)\times(2l+1)}, \\
 	C_{x,l} &:= 
@@
 		\end{pmatrix} \in \C^{(2l+1)\times(2l+3)}, \\
-	B_{y,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
+	B_{y,l} &:= 0 \in \R^{(2l+1)\times(2l+1)}, \\
 	C_{y,l} &:= 
@@
 		\end{pmatrix} \in \R^{(2l+1)\times(2l+3)}, \\
-	B_{z,l} &:= 0 \in \R^{(2l+1)\times(2l+1)} \\
+	B_{z,l} &:= 0 \in \R^{(2l+1)\times(2l+1)}, \\
 	C_{z,l} &:= 
@@
 			{- \; \alpha_{l,-l,4} \over \alpha_{l,-l,3} \: \gamma_{l,1-l,2}} & j = 2(2l+1) + 2 \\
-			0 & o/w
-		\end{cases}, \\
+			0 &\text{otherwise},
+		\end{cases} \\
 	\big(\vec{\eta}_{1}\big)_j &= 
@@
 			{- \; \alpha_{l,l,3} \over \alpha_{l,l,4} \: \gamma_{l,l-1,2}} & j = 3(2l+1) - 1 \\
-			0 & o/w
+			0 &\text{otherwise}.
 		\end{cases}.
@@
 \bseqn{
-	f_{l,m} = \iint_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta
-}\bstodo{the complex conj should be included here, shouldn't it?}
+	f_{l,m} = \iint_\Omega f(\costheta \sinphi, \sintheta \sinphi, \cosphi) \: \Ylmfull^* \: \sinphi \: \D \varphi \: \D \theta,
+}
 using the orthonormality of the spherical harmonics.
@@
 \bseqn{
-	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+1} = \vec{0}. \\
+	\quad &\text{1) } \text{Set } \vec{\xi}_{N+2} = \vec{0}, \: \vec{\xi}_{N+1} = \vec{0} \\
 	\quad &\text{2) } \text{For } n = N:-1:0 \\
@@
 \bseqnnumber{
-	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}) \label{eqn:intcomplexexpcos} \\
-	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}) \label{eqn:intcomplexexpsin}
+	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}), \label{eqn:intcomplexexpcos} \\
+	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}), \label{eqn:intcomplexexpsin}
 }
@@
 \bseqnnumber{
-	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
-	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } 
+	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2}, \label{eqn:normjac} \\
+	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1}, \quad \text{where } 
 	d_{l,m} := 
@@
 			\half (l + \abs{m} + 1) & l > \abs{m} \\
-			0 & l = \abs{m}
-		\end{cases}. \label{eqn:diffjac}
+			0 & l = \abs{m}.
+		\end{cases} \label{eqn:diffjac}
 }
@@
 	z \gradYlm &= \Gamma_{l,m,1} \gradY^{m}_{l-1} + \Gamma_{l,m,2} \gradY^{m}_{l+1} + \Gamma_{l,m,3} \gradpY^{m}_{l}, \\
-	z \gradpYlm &= \Gamma_{l,m,1} \gradpY^{m}_{l-1} + \Gamma_{l,m,2} \gradpY^{m}_{l+1} - \Gamma_{l,m,3} \gradY^{m}_{l}
+	z \gradpYlm &= \Gamma_{l,m,1} \gradpY^{m}_{l-1} + \Gamma_{l,m,2} \gradpY^{m}_{l+1} - \Gamma_{l,m,3} \gradY^{m}_{l},
 }
@@
 			\tilde A_{l,m,1} & \text{if } m > 0 \\
-			\tilde A_{l,\abs{m},2} & \text{if } m \le 0
+			\tilde A_{l,\abs{m},2} & \text{if } m \le 0,
 		\end{cases} \\
@@
 			\tilde A_{l,m,2} & \text{if } m \ge 0 \\
-			\tilde A_{l,\abs{m},1} & \text{if } m < 0
+			\tilde A_{l,\abs{m},1} & \text{if } m < 0,
 		\end{cases} \\
@@
 			\tilde A_{l,m,3} &\text{if } m > 0 \\
-			\tilde A_{l,\abs{m},4} &\text{if } m \le 0
+			\tilde A_{l,\abs{m},4} &\text{if } m \le 0,
 		\end{cases} \\
@@
 			\tilde A_{l,m,4} &\text{if } m \ge 0 \\
-			\tilde A_{l,\abs{m},3} &\text{if } m < 0
+			\tilde A_{l,\abs{m},3} &\text{if } m < 0,
 		\end{cases} \\
@@
 			- i \: d_{l,m - 1} \: {c_l^{m-1} \over 2 \clm} &\text{if } m > 0 \\
-			i \: d_{l,\abs{m}} \: {\clm \over 2 c_l^{m-1}} &\text{if } m \le 0
+			i \: d_{l,\abs{m}} \: {\clm \over 2 c_l^{m-1}} &\text{if } m \le 0,
 		\end{cases} \\
@@
 			- d_{l,m} \: {\clm \over c_l^{m+1}} &\text{if } m \ge 0 \\
-			d_{l,\abs{m} - 1} \: {c_l^{m+1} \over \clm} &\text{if } m < 0
+			d_{l,\abs{m} - 1} \: {c_l^{m+1} \over \clm} &\text{if } m < 0,
 		\end{cases} \\
-	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6 \\
-	\Gamma_{l,m,1} &:= {\clm c_{l-1}^{m} \over l(l-1)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \Big] \\
-	\Gamma_{l,m,2} &:= {\clm c_{l+1}^{m} \over (l+1)(l+2)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \Big] \\
-	\Gamma_{l,m,3} &:= {i \: m \over l(l+1)}
+	B_{l,m,j} &:= i (-1)^{j+1} A_{l,m,j} \quad \text{for } j = 1,\dots,6, \\
+	\Gamma_{l,m,1} &:= {\clm c_{l-1}^{m} \over l(l-1)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},1} \over (c_{l-1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l-1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,1} \over (c_{l-1}^{\abs{m}+1})^2} \Big], \\
+	\Gamma_{l,m,2} &:= {\clm c_{l+1}^{m} \over (l+1)(l+2)} \Big[ \abs{m}(\abs{m} + 2) {\tilde \gamma_{l,\abs{m},2} \over (c_{l+1}^{\abs{m}})^2} + d_{l,\abs{m}} d_{l+1,\abs{m}} {\tilde \gamma_{l,\abs{m}+1,2} \over (c_{l+1}^{\abs{m}+1})^2} \Big] ,\\
+	\Gamma_{l,m,3} &:= {i \: m \over l(l+1)},
 }
@@
 \bseqn{
-	\tilde A_{l,m,1} &:= \Big( (m^2 - 1) \tilde \alpha_{l-1,m-1,4} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l-1,m-1} {\tilde \alpha_{l,m+1,1} \over (c_{l-1}^{m})^2} \\
-	\tilde A_{l,m,2} &:= \Big( m(m+2) \tilde \alpha_{l,m,2} - d_{l,m} \tilde \gamma_{l,m+1,1} \Big) {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\tilde \alpha_{l-1,m+2,3} \over (c_{l}^{m+1})^2} \\
-	\tilde A_{l,m,3} &:= \Big( (m^2 - 1) \tilde \alpha_{l+1,m-1,2} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l+1,m-1} {\tilde \alpha_{l,m+1,3} \over (c_{l+1}^{m})^2} \\
-	\tilde A_{l,m,4} &:= \Big( m(m+2) \tilde \alpha_{l,m,4} - d_{l,m} \tilde \gamma_{l,m+1,2} \Big) {1 \over (c_{l+1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\tilde \alpha_{l+1,m+2,1} \over (c_{l}^{m+1})^2}.
+	\tilde A_{l,m,1} &:= \Big( (m^2 - 1) \tilde \alpha_{l-1,m-1,4} - d_{l-1,m-1} \tilde \gamma_{l-1,m,2} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l-1,m-1} {\tilde \alpha_{l,m+1,1} \over (c_{l-1}^{m})^2}, \\
+	\tilde A_{l,m,2} &:= \Big( m(m+2) \tilde \alpha_{l,m,2} - d_{l,m} \tilde \gamma_{l,m+1,1} \Big) {1 \over (c_{l-1}^{m+1})^2} + d_{l,m} d_{l-1,m+1} {\tilde \alpha_{l-1,m+2,3} \over (c_{l}^{m+1})^2}, \\
+	\tilde A_{l,m,3} &:= \Big( (m^2 - 1) \tilde \alpha_{l+1,m-1,2} - d_{l+1,m-1} \tilde \gamma_{l+1,m,1} \Big) {1 \over (\clm)^2} + d_{l,m} d_{l+1,m-1} {\tilde \alpha_{l,m+1,3} \over (c_{l+1}^{m})^2}, \\
+	\tilde A_{l,m,4} &:= \Big( m(m+2) \tilde \alpha_{l,m,4} - d_{l,m} \tilde \gamma_{l,m+1,2} \Big) {1 \over (c_{l+1}^{m+1})^2} + d_{l,m} d_{l+1,m+1} {\tilde \alpha_{l+1,m+2,1} \over (c_{l}^{m+1})^2},
 }
@@
 \bseqn{
-	x \gradYlm = \sum_{l'=1}^{l+1} \sum_{m'=-l'}^{l'} \calA_{l',m'} \gradYlmprime + \calA^\perp_{l',m'} \gradpYlmprime
+	x \gradYlm = \sum_{l'=1}^{l+1} \sum_{m'=-l'}^{l'} \calA_{l',m'} \gradYlmprime + \calA^\perp_{l',m'} \gradpYlmprime,
 }
@@
 \bseqn{
-	\calA_{l',m'} = {\ip<x \gradYlm, \gradYlmprime> \over \norm{\gradYlmprime}^2} = {1 \over l'(l'+1)} \ip<x \gradYlm, \gradYlmprime> \\
+	\calA_{l',m'} = {\ip<x \gradYlm, \gradYlmprime> \over \norm{\gradYlmprime}^2} = {1 \over l'(l'+1)} \ip<x \gradYlm, \gradYlmprime>, \\
 	\calA^\perp_{l',m'} = {\ip<x \gradYlm, \gradpYlmprime> \over \norm{\gradpYlmprime}^2} = {1 \over l'(l'+1)} \ip<x \gradYlm, \gradpYlmprime>.
@@
 \bseqn{
-	z \gradYlm = \sum_{l'=1}^{l+1} \sum_{m'=-l'}^{l'} \calC_{l',m'} \gradYlmprime + \calC^\perp_{l',m'} \gradpYlmprime
+	z \gradYlm = \sum_{l'=1}^{l+1} \sum_{m'=-l'}^{l'} \calC_{l',m'} \gradYlmprime + \calC^\perp_{l',m'} \gradpYlmprime,
 }
@@
 \bseqn{
-	\calC_{l',m'} = {\ip<z \gradYlm, \gradYlmprime> \over \norm{\gradYlmprime}^2} = {1 \over l'(l'+1)} \ip<z \gradYlm, \gradYlmprime> \\
+	\calC_{l',m'} = {\ip<z \gradYlm, \gradYlmprime> \over \norm{\gradYlmprime}^2} = {1 \over l'(l'+1)} \ip<z \gradYlm, \gradYlmprime>, \\
 	\calC^\perp_{l',m'} = {\ip<z \gradYlm, \gradpYlmprime> \over \norm{\gradpYlmprime}^2} = {1 \over l'(l'+1)} \ip<z \gradYlm, \gradpYlmprime>.
@@
 	&\quadeight \cdot \big[ \abs{m'} z \jaclmabs{l'}{m'}(z) - \rho(z)^2 \jaclmabs{l'}{m'}(z) \big] \nonumber \\
-	&\quadfour \quadtwo + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: z \; \rho(z)^{\abs{m}+\abs{m'}-2} \: \D z \Big) \label{eqn:lemmaVSHrecurrencesproofz1}
+	&\quadfour \quadtwo + mm' \jaclmabs{l}{m}(z) \jaclmabs{l'}{m'}(z) \Big\} \: z \; \rho(z)^{\abs{m}+\abs{m'}-2} \: \D z \Big). \label{eqn:lemmaVSHrecurrencesproofz1}
 }
@@
 \bseqnnumber{
-	\tangentjacobi_x \VSHvecfull(x,y,z) &= x \VSHvecfull(x,y,z) \nonumber \\
-	\tangentjacobi_y \VSHvecfull(x,y,z) &= y \VSHvecfull(x,y,z) \label{eqn:tangentjacobimatdef} \\ 
-	\tangentjacobi_z \VSHvecfull(x,y,z) &= z \VSHvecfull(x,y,z) \nonumber
+	\tangentjacobi_x \VSHvecfull(x,y,z) &= x \VSHvecfull(x,y,z), \nonumber \\
+	\tangentjacobi_y \VSHvecfull(x,y,z) &= y \VSHvecfull(x,y,z), \label{eqn:tangentjacobimatdef} \\ 
+	\tangentjacobi_z \VSHvecfull(x,y,z) &= z \VSHvecfull(x,y,z), \nonumber
 }
@@
 	\begin{cases}
-		\VSHvecfull_{0}(x,y,z) := 0_{2,3} \\
+		\VSHvecfull_{0}(x,y,z) := 0_{2,3}, \\
 		\VSHvecfull_{l+1}(x,y,z) = -\Dlt (B_l-G_l(x,y,z)) \VSHvecfull_l(x,y,z) - \Dlt C_l  \, \VSHvecfull_{l-1}(x,y,z), \quad l \in \N.
@@
 			{- \; A_{l,-l,4} \over A_{l,-l,3} \: \Gamma_{l,1-l,2}} & j = 4(2n+1) + 3 \\
-			0 & o/w
-		\end{cases}, \\
+			0 &\text{otherwise},
+		\end{cases} \\
 	\big(\vec{\eta}_{1}\big)_j &= 
@@
 			\big(\vec{\eta}_{0}\big)_{j-1} & j = 2, 4(2l+1) + 4 \\
-			0 & o/w
-		\end{cases}, \\
+			0 &\text{otherwise},
+		\end{cases} \\
 	\big(\vec{\eta}_{2}\big)_j &= 
@@
 			{- \; A_{l,l,3} \over A_{l,l,4} \: \Gamma_{l,l-1,2}} & j = 6(2l+1) - 3 \\
-			0 & o/w
-		\end{cases}, \\
+			0 &\text{otherwise},
+		\end{cases} \\
 	\big(\vec{\eta}_{3}\big)_j &= 
@@
 			\big(\vec{\eta}_{2}\big)_{j-1} & j = 2(2l+1), 6(2l+1) - 2 \\
-			0 & o/w
-		\end{cases}.
+			0 &\text{otherwise}.
+		\end{cases}
 }

### 2021-10-29 papers/thesis/resubmitted/ch1-introduction/introduction.tex
Intro roadmap is now numbered
Added bstodo notes to indicate changes

@@
 \chapter{Introduction}\label{CHAPTER:intro}
+
+\bstodoinline{Any comments in boxes like these indicate changes made.
+
+Aside from where I have marked, I have also:
+
+- Fixed typos pointed out and found myself
+
+- Added commas or full stops after equations
+
+- Changed how commas are presented in equations with \enquote{cases}
+
+- Added round brackets to citations of equation numbers in the DLMF
+
+- Added a dot where citing a page number (eg, p.76)
+
+- Replaced \& with and}
 
@@
 \cite{trefethen2000spectral, canuto2007spectral, gottlieb1977numerical, boyd2001chebyshev, mason2002chebyshev, shen2011spectral, olver2013fast}).
-While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, with special cases of the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[\S18.3]{DLMF} -- the area of multivariate orthogonal polynomials has a smaller array of research.
+While there are many famous examples of univariate OPs -- such as the Jacobi polynomials, with special cases of the Legendre polynomials and the Chebyshev polynomials but to name a few of the classical families \cite[\S18.3]{DLMF}\bstodo{Clarified that these families named are all Jacobi} -- the area of multivariate orthogonal polynomials has a smaller array of research.
 
@@
 
-Spherical harmonics (SHs) are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere, have sparse recurrences, and are eigenfunctions of the spherical Laplacian) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
+Spherical harmonics (SHs) are of course a long-established and famous group of functions defined on the surface of a sphere with certain useful properties (for example, they are orthogonal to each other on the unit sphere, have sparse recurrences, and are eigenfunctions of the spherical Laplacian \bstodo{Re-worded the properties}) and as a result are widely used in many scientific fields for solving PDEs including computer graphics (e.g.
 \cite{moon2008efficient, sloan2013efficient}), astrophysics (e.g.
@@
 % GENERAL FRAMEWORK
+\bstodoinline{The following is the added "roadmap"}
 The construction of the spectral methods in this thesis follow a similar pattern that is worth detailing here as a general framework.
-\begin{itemize}
+\begin{enumerate}
 	\item \textit{Define the domain of interest.}
@@
 		For instance, this would be useful for constructing an operator matrix that represents a variable coefficient in a Helmholtz example.
-\end{itemize}
+\end{enumerate}
 
@@
 The resulting system's sparsity is a result of the sparse relations of the OPs.
+\bstodoinline{End of added roadmap}
 

### 2021-10-29 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
Added bstodo notes to indicate changes

@@
 Our aim here is to express the spherical harmonics as polynomials in three variables $x$, $y$, and $z$ to evaluate functions and solve PDEs on the whole sphere.
-More precisely, we desire the solution to partial differential equations on the domain \nomenclature[Omega]{$\Omega$}{The domain of interest}
+More precisely, we desire the solution to partial differential equations on the domain \nomenclature[Omega]{$\Omega$}{The domain of interest} \bstodo{Added that this domain is $\bigS^2$}
 \bseqn{
@@
 We first need to introduce a few classical orthogonal polynomials.
-Orthogonal polynomials are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product, and ordered by degree.
-In particular, they are orthogonal with respect to all lower degree polynomials.
+Orthogonal polynomials are defined as such by the condition that each polynomial is orthogonal to each other, with respect to a given inner product, and ordered by degree.\bstodo{added that they are degree ordered} In particular, they are orthogonal with respect to all lower degree polynomials.
 
@@
 }
-where $P^{(a,b)}_l (x)$ is the degree $l$ \textit{Jacobi polynomial} with $a, b > -1$, and \nomenclature[Pl]{$P_l$}{The degree $l$ Legendre polynomial} $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$.
+where $P^{(a,b)}_l (x)$ is the degree $l$ \textit{Jacobi polynomial} with $a, b > -1$ \bstodo{added the a,b constraint}, and \nomenclature[Pl]{$P_l$}{The degree $l$ Legendre polynomial} $P_l(x) := P^{(0,0)}_l (x)$ is simply the \textit{Legendre polynomial} of degree $l$.
 Jacobi polynomials are orthogonal with respect to the weight $w(x) = (1-x)^a(1+x)^b$; that is for $l, l' \in \No,$ \nomenclature[Kronecker delta]{$\delta_{n, n'}$}{The Kronecker delta function; $1$ if $n=n'$, else $0$}
@@
 }
-where $\delta_{l, l'}$ is the regular Kronecker delta function. For the Jacobi polynomials,
+where $\delta_{l, l'}$ is the regular Kronecker delta function.\bstodo{clarified that this delta is the Kronecker} For the Jacobi polynomials,
 \bseqn{
@@
 Legendre polynomials are special cases of \textit{Legendre functions} when $l$ is a non-negative integer.
-Legendre functions are a class of univariate functions that are one set of solutions to the differential equation
+Legendre functions are a class of univariate functions that are one set of solutions to the differential equation \bstodo{clarified that they are one set of solutions to the following eqn.
+Added the x domain.}
 \bseqn{
@@
 }
-The spherical harmonics as defined are then orthonormal with respect to the inner product with uniform measure \nomenclature[conjuagate]{$a^*$}{The complex conjugate of $a \in \C$}
+The spherical harmonics as defined are then orthonormal with respect to the inner product with uniform measure \nomenclature[conjuagate]{$a^*$}{The complex conjugate of $a \in \C$} \bstodo{Added that the following is $\iint_{\Omega} \Ylmfull \: {Y_{l'}^{m'}(\varphi, \theta)}^{*} \: \D S$.}
 \bseqn{
@@
 	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[(18.9.5), (18.9.6)]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
-The final one is obtained by rearranging the three-term recurrence in \cite[(18.9.1), (18.9.2)]{DLMF}
+The final one is obtained by rearranging the three-term recurrence in \cite[(18.9.1), (18.9.2)]{DLMF}.
+\bstodo{Reworded this sentence.}
 \end{proof}
@@
 In order to attempt to maximise sparsity, we define them in the following way\footnote{There may be a more computationally efficient way of defining these matrices, but in practice we only desire the output of the action of $\Dlt C_l$ and $\Dlt (B_l-G_l(x,y,z))$, and so minimising the resulting number of operations is the goal.
-The choices of the $\Dlt$ matrices we make in this thesis are designed to do this, but whether there are more computationally efficient ways of defining them is an open question.}.
-For $l \in \N$, we can set
+The choices of the $\Dlt$ matrices we make in this thesis are designed to do this, but whether there are more computationally efficient ways of defining them is an open question.}.\bstodo{Added footnote to help explain choice of $\Dlt$'s} For $l \in \N$, we can set \bstodo{Tried to reformat the matrix defn to make it clearer.}
 \bseqnnumber{
@@
 
-In coefficient space, we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
+In coefficient space,\bstodo{Changed "spectral space" to "coefficient space".} we wish to work only with vectors of coefficients for the expansion of a function, to which we can apply operator matrices to that represent differential or other operations.
 Naturally, we of course need a way to obtain the coefficients $f_{l,m}$.
@@
 
-Let $\vec{v}$ be the coefficients of the expansion of $v$ up to order $N$. The operator $V$ would then be the result of the following \textit{operator Clenshaw algorithm}:
+Let $\vec{v}$ be the coefficients of the expansion of $v$ up to order $N$.
+The operator $V$ would then be the result of the following \textit{operator Clenshaw algorithm}:\bstodo{Corrected the typo on the last line of the algorithm.}
 \bseqn{
@@
 At each iteration, the entries of $\vec{\xi}_n$ are polynomials in $J_x$, $J_y$ and $J_z$.
-On the final step, $\vec{\xi}_{0} = \xi_{0}$ is hence some polynomial expression in $J_x$, $J_y$ and $J_z$, which we can then evaluate by inserting the Jacobi matrices.} (note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).
+On the final step, $\vec{\xi}_{0} = \xi_{0}$ is hence some polynomial expression in $J_x$, $J_y$ and $J_z$, which we can then evaluate by inserting the Jacobi matrices.} \bstodo{Added footnote to help explain better what the notation here is meaning.}(note that we are abusing notation here a bit, however it is the simplest way to present the algorithm without introducing yet more matrices!).
 
@@
 
-The Jacobi matrices are then defined according to the matrices \nomenclature[JT]{$\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$}{The Jacobi matrices for the tangent bundle VSH basis} $\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$ satisfying
+The Jacobi matrices are then defined according to the matrices \nomenclature[JT]{$\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$}{The Jacobi matrices for the tangent bundle VSH basis} $\tangentjacobi_x, \tangentjacobi_y, \tangentjacobi_z$ satisfying \bstodo{Added the notation for $\tangentjacobi_x$, etc.
+to the index.}
 \bseqnnumber{
@@
 
-The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other key operators found in examples such as the linearised shallow water equations.
-These operators will take coefficients of a function's expansion in either the SH or VSH basis, to coefficients in either basis.
+The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators.
+Of course, not all PDEs can be sparsely represented, however we can do this for key operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other operators found in examples such as the linearised shallow water equations.\bstodo{Added the comment that not all PDEs operators can be sparsely represented.} These operators will take coefficients of a function's expansion in either the SH or VSH basis, to coefficients in either basis.
 
@@
 \begin{theorem}
-	The operator matrices defined in \bsrefdef{def:SHoperators} are diagonal, and are given by:
+	The operator matrices defined in \bsrefdef{def:SHoperators} are diagonal, and are given by: \bstodo{Tried to make clearer what $\calG$ is here.}
 \bseqn{
@@
 
-\begin{figure}[t]
+\bstodo{added the order used $N=20$ in the figure caption}\begin{figure}[t]
 	\centering % <-- added

### 2021-10-29 papers/thesis/resubmitted/ch3-disk-slices/disk-slices.tex
Added bstodo notes to indicate changes

@@
 
-As mentioned earlier, we can handle these boundary conditions by incorporating them into the chosen basis OP family.
+\bstodoinline{Added the following paragraph to explain BCs better.}As mentioned earlier, we can handle these boundary conditions by incorporating them into the chosen basis OP family.
 The weight $\Wiii(x,y)$ is of course zero on the boundary of $\Omega$, and in fact any polynomial that vanishes on the boundary will have $\Wiii(x,y)$ as a factor.
@@
 
-We can tackle the problem as follows.
+\bstodo{Reworded this paragraph to help make the BC incorporation clearer.}We can tackle the problem as follows.
 Since $u$ vanishes on the boundary, any polynomial expansion will have $\Wiii(x,y)$ as a factor, and so we can denote the coefficient vector for expansion of $u$ in the $\bigWiii$ OP basis up to degree $N$ by $\vec{u}$, while the coefficient vector for expansion of $f$ in the $\bighdopiii$ OP basis up to degree $N$ we can denote by $\vec{f}$.
@@
 \bseqn{
-    \laplacewiii \vec{u} = \vec{f}
+    \laplacewiii \vec{u} = \vec{f},
 }

### 2021-10-29 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
Final version with bstodo notes removed

--- 2021-10-19 papers/thesis/resubmitted/ch2-spherical-harmonics/spherical-harmonics.tex
@@
 	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[(18.9.5), (18.9.6)]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
-The final one is obtained by rearranging the three-term recurrence in \cite[(18.9.1), (18.9.2)]{DLMF}
+The final one is obtained by rearranging the three-term recurrence in \cite[(18.9.1), (18.9.2)]{DLMF}.
 \end{proof}
@@
 
-The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other key operators found in examples such as the linearised shallow water equations.
+The framework we have outlined for the spherical harmonics and the vector spherical harmonics allows us to easily and explicitly derive sparse operator matrices for partial differential operators.
+Of course, not all PDEs can be sparsely represented, however we can do this for key operators such as the spherical gradient, divergence and Laplace--Beltrami, as well as other operators found in examples such as the linearised shallow water equations.
 These operators will take coefficients of a function's expansion in either the SH or VSH basis, to coefficients in either basis.

### 2021-11-09 papers/thesis/resubmitted/acknowledgements/acknowledgements.tex
Actually, spotted typo in EPSRC. Fixed.

@@
 
-Firstly, I would like to express my thanks to ESPRC for their financial support, and to the Mathematics of Planet Earth Centre for Doctoral Training (MPE CDT) for providing me with the incredible opportunities that the course has brought.
+Firstly, I would like to express my thanks to EPSRC for their financial support, and to the Mathematics of Planet Earth Centre for Doctoral Training (MPE CDT) for providing me with the incredible opportunities that the course has brought.
 
